### 2018-08-10 README.Rmd
Adding package infrastructure

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# forestecology

The goal of forestecology is to ...

## Installation

You can install the released version of forestecology from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("forestecology")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("rudeboybert/forestecology")
```
## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
## basic example code
```

What is special about using `README.Rmd` instead of just `README.md`? You can include R chunks like so:

```{r cars}
summary(cars)
```

You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date.

You can also embed plots, for example:

```{r pressure, echo = FALSE}
plot(pressure)
```

In that case, don't forget to commit and push the resulting figure files, so they display on GitHub!

### 2018-08-15 R/data.R
Fixed all CRAN checks

@@ -8,11 +8,11 @@
 #'   \item{species}{Linnaean species name}
 #'   \item{family_phylo}{Corresponding species phylogenetic family}
 #'   \item{trait_group}{Corresponding species trait-based cluster}
-#'   \item{x}{x-coordinate (meters from reference point)}
-#'   \item{y}{y-coordinate (meters from reference point)}
-#'   \item{growth}{Average annual growth in diameter at breast height from 2008-2014 (6 years)}
+#'   \item{x}{x-coordinate meters from reference point}
+#'   \item{y}{y-coordinate meters from reference point}
+#'   \item{growth}{Average annual growth in diameter at breast height from 2008-2014 6 years}
 #'   \item{dbh08}{Diameter at breast high in 2008}
-#'   \item{dhb14}{Diameter at breast high in 2014}
+#'   \item{dbh14}{Diameter at breast high in 2014}
 #'   \item{code14}{Code}
 #' }
 #' @seealso \code{\link{families}}
@@ -29,13 +29,12 @@
 #'   \item{y}{y-coordinate (meters from reference point)}
 #' }
 #' @seealso \code{\link{families}}
-#' @examples{
+#' @examples
 #' library(ggplot2)
 #' ggplot(data = bigwoods, aes(x,y)) +
 #'   geom_point() +
 #'   coord_fixed(ratio=1) +
-#'   geom_path(data = study_region_vertices, col = "red", size = 1)
-#' }
+#'   geom_path(data = bigwoods_study_region, col = "red", size = 1)
 "bigwoods_study_region"
 
 

### 2018-08-15 R/forestecology.R

@@ -1,0 +1,21 @@
+#' \code{forestecology} package
+#'
+#' Methods and data for forest ecology model selection and assessment
+#'
+#' See the README on
+#' \href{https://github.com/rudeboybert/forestecology#readme}{GitHub}
+#'
+#' @docType package
+#' @name forestecology
+NULL
+
+## quiets concerns of R CMD check re: the .'s that appear in pipelines
+if(getRversion() >= "2.15.1")
+  utils::globalVariables(
+    c("dbh08", "dbh14", "family_phylo", "trait_group", "buffer", "code14",
+      "species", "focal_ID", "ID", "spCode", "fold", "basal_area", "dbh",
+      "focal_ID_neighbor", "focal_bw_ID", "comp_species", "comp_biomass",
+      "inside", "growth", "biomass_total", "model_formula", "growth_hat",
+      "comp_ID", "x", "y", "x_fold", "y_fold", "neighbor", ".",
+      "notion_of_competitor_species"
+      ))

### 2018-08-15 R/modeling.R

@@ -4,7 +4,8 @@
 #' @param model_number Number of model out of 1-5 from paper
 #'
 #' @import dplyr
-#' @return A list ouf outputs
+#' @importFrom stats as.formula
+#' @return A list of outputs
 #' @export
 #' @examples
 #' 1+1
@@ -70,11 +71,13 @@ get_model_specs <- function(forest, model_number){
 
 #' Fit bayesian model
 #'
-#' @param focal_trees training set
 #' @param focal_vs_comp from \code{\link{create_focal_vs_comp}}
 #' @param run_shuffle boolean
 #'
 #' @import dplyr
+#' @importFrom stats model.matrix
+#' @importFrom tidyr unnest
+#' @importFrom tidyr spread
 #' @return Posterior parameter values
 #' @export
 #'
@@ -162,6 +165,8 @@ fit_bayesian_model <- function(focal_vs_comp, run_shuffle = FALSE){
 #' @param posterior_param Output of \code{\link{fit_bayesian_model}}
 #'
 #' @import dplyr
+#' @importFrom stats model.matrix
+#' @importFrom tidyr nest
 #' @return \code{focal_vs_comp} with new column of predicted \code{growth_hat}
 #' @export
 #'

### 2018-08-15 R/spatial.R

@@ -38,13 +38,17 @@ define_bigwoods_buffer <- function(bigwoods, max_dist){
 #'
 #' @inheritParams define_bigwoods_buffer
 #' @param forest data frame containing tree data
+#' @param cv_fold_size grid size for spatial crossvalidation fold
 #'
 #' @return The same \code{forest} data frame but with a new numerical variable
 #'   \code{fold} indicating the fold number.
 #' @export
 #' @import dplyr
+#' @importFrom stringr str_c
 #' @examples
 #' library(ggplot2)
+#' max_dist <- 7.5
+#' cv_fold_size <- 100
 #' bigwoods <- bigwoods %>%
 #'   define_bigwoods_buffer(max_dist) %>%
 #'   define_cv_grid(cv_fold_size)
@@ -75,6 +79,7 @@ define_cv_grid <- function(forest, cv_fold_size){
 #'   all its neighbors in a nested list.
 #' @export
 #' @import dplyr
+#' @importFrom tidyr nest
 #' @examples
 #' 1+1
 get_cv_fold_info <- function(forest, cv_fold_size){

### 2019-12-03 README.Rmd
Fixed all CRAN checks
Started README

@@
   fig.path = "man/figures/README-",
-  out.width = "100%"
+  out.width = "100%",
+  message = FALSE,
+  warning = FALSE
 )
 ```
+
 # forestecology
 
-The goal of forestecology is to ...
+[![Travis Build Status](https://travis-ci.org/rudeboybert/forestecology.svg?branch=master)](https://travis-ci.org/rudeboybert/forestecology)
+[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
+[![CRAN status](https://www.r-pkg.org/badges/version/forestecology)](https://cran.r-project.org/package=forestecology)
 
@@
 ```
-## Example
 
-This is a basic example which shows you how to solve a common problem:
+## Example analysis
 
-```{r example}
-## basic example code
+We present an example analysis using the [Michigan Big Woods research plot
+data](https://doi.org/10.7302/wx55-kt18) which is part of the Smithsonian
+Institution's [Forest Global Earth Observatory
+(ForestGEO)](https://forestgeo.si.edu/) global network of forest research sites.
+
+Porting over code from `SCBI.R`
+
+```{r, cache = TRUE}
+library(tidyverse)
+library(forestecology)
+library(snakecase)
+
+# Read in Big Woods census and species data
+bw_2008 <- read_delim("https://deepblue.lib.umich.edu/data/downloads/z603qx485", delim = "\t") %>% 
+  select(
+    treeID = treeid, stemID = stemtag, sp = spcode, quadrat, gx, gy, dbh, 
+    ExactDate = date, code_2008 = codes
+  )
+bw_2014 <- read_delim("https://deepblue.lib.umich.edu/data/downloads/1831ck00f", delim = "\t") %>% 
+  select(
+    treeID = treeid, stemID = stemtag, sp = spcode, quadrat, gx, gy, dbh, 
+    ExactDate = date, code_2014 = codes
+  )
+bw_species <- read_delim("https://deepblue.lib.umich.edu/data/downloads/000000086", delim = "\t") %>% 
+  mutate(spcode = to_any_case(spcode)) %>% 
+  left_join(families, by = "spcode") %>% 
+  mutate(
+    sp = str_sub(genus, 1, 2), 
+    sp = str_c(sp, str_sub(species, 1, 2)),
+    sp = tolower(sp),
+    latin = str_c(genus, species, sep = " ")
+  ) %>% 
+  select(sp, genus, species, family, latin, trait_group)
 ```
-
-What is special about using `README.Rmd` instead of just `README.md`? You can include R chunks like so:
-
-```{r cars}
-summary(cars)
-```
-
-You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date.
-
-You can also embed plots, for example:
-
-```{r pressure, echo = FALSE}
-plot(pressure)
-```
-
-In that case, don't forget to commit and push the resulting figure files, so they display on GitHub!

### 2020-01-23 README.Rmd
Added compute_growth() to README

--- 2018-08-10 README.Rmd
@@
   fig.path = "man/figures/README-",
-  out.width = "100%"
+  out.width = "100%",
+  fig.width = 16/2,
+  fig.height = 9/2,
+  message = FALSE,
+  warning = FALSE
 )
 ```
+
 # forestecology
 
-The goal of forestecology is to ...
+[![Travis Build Status](https://travis-ci.org/rudeboybert/forestecology.svg?branch=master)](https://travis-ci.org/rudeboybert/forestecology)
+[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
+[![CRAN status](https://www.r-pkg.org/badges/version/forestecology)](https://cran.r-project.org/package=forestecology)
 
@@
 
-``` r
+```{r, eval = FALSE}
 install.packages("forestecology")
@@
 
-``` r
-# install.packages("devtools")
-devtools::install_github("rudeboybert/forestecology")
-```
-## Example
-
-This is a basic example which shows you how to solve a common problem:
-
-```{r example}
-## basic example code
+```{r, eval = FALSE}
+# install.packages("remotes")
+remotes::install_github("rudeboybert/forestecology")
 ```
 
-What is special about using `README.Rmd` instead of just `README.md`? You can include R chunks like so:
+## Example analysis
 
-```{r cars}
-summary(cars)
+We present an example analysis using:
+
+* [Michigan Big Woods](https://doi.org/10.7302/wx55-kt18) research plot data
+* [Smithsonian Conservation Biology Institute (SCBI) ForestGEO](https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/tree/master/tree_main_census) research plot data
+
+```{r}
+library(tidyverse)
+library(forestecology)
+library(snakecase)
+library(skimr)
 ```
 
-You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date.
+### Load & preprocess data
 
-You can also embed plots, for example:
+**Big Woods**:
 
-```{r pressure, echo = FALSE}
-plot(pressure)
+```{r, cache=TRUE}
+# Read in Big Woods census data from 2008 & 2014
+bw_2008 <- 
+  "https://deepblue.lib.umich.edu/data/downloads/z603qx485" %>% 
+  read_delim(delim = "\t") %>% 
+  select(
+    treeID = treeid, stemID = stemtag, sp = spcode, quadrat, gx, gy, dbh, 
+    date, codes
+  )
+bw_2014 <- 
+  "https://deepblue.lib.umich.edu/data/downloads/1831ck00f" %>% 
+  read_delim(delim = "\t") %>% 
+  select(
+    treeID = treeid, stemID = stemtag, sp = spcode, quadrat, gx, gy, dbh, 
+    date, codes
+  )
+
+# Read in bigwoods grouping classification data
+bw_species <- 
+  "https://deepblue.lib.umich.edu/data/downloads/000000086" %>% 
+  read_delim(delim = "\t") %>% 
+  # convert all to snake case:
+  mutate_at(c("species", "genus", "family", "idlevel", "spcode"), to_any_case) %>% 
+  # join trait group
+  left_join(families, by = c("spcode", "family")) %>% 
+  mutate(
+    sp = str_sub(genus, 1, 2), 
+    sp = str_c(sp, str_sub(species, 1, 2)),
+    sp = tolower(sp),
+    latin = str_c(genus, species, sep = " "),
+    latin = to_any_case(latin)
+  ) %>% 
+  select(sp, spcode, genus, species, latin, family, trait_group)
 ```
 
-In that case, don't forget to commit and push the resulting figure files, so they display on GitHub!
+**SCBI**:
+
+```{r, cache=TRUE}
+scbi_2013 <- 
+  "https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/raw/master/tree_main_census/data/census-csv-files/scbi.stem2.csv" %>% 
+  read_csv() %>% 
+  select(
+    treeID, stemID, sp, quadrat, gx, gy, dbh, 
+    date = ExactDate, codes, status
+  )
+
+scbi_2018 <-
+  "https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/raw/master/tree_main_census/data/census-csv-files/scbi.stem3.csv" %>% 
+  read_csv() %>% 
+  select(
+    treeID, stemID, sp, quadrat, gx, gy, dbh, 
+    date = ExactDate, codes, status
+  ) %>% 
+  mutate(dbh = as.numeric(dbh))
+```
+
+### Compute growth
+
+**Big Woods**:
+
+```{r, cache=TRUE}
+census_df1 <- bw_2008
+census_df2 <- bw_2014
+id <- "treeID"
+
+bw_growth_df <- 
+  # Merge both censuses and compute growth:
+  compute_growth(census_df1, census_df2, id) %>% 
+  mutate(sp = to_any_case(sp)) %>% 
+  left_join(bw_species, by = c("sp" = "spcode"))
+```
+
+**SCBI**:
+
+```{r, cache=TRUE}
+census_df1 <- scbi_2013
+census_df2 <- scbi_2018
+id <- "stemID"
+
+scbi_growth_df <- 
+  # Merge both censuses and compute growth:
+  compute_growth(census_df1, census_df2, id)
+```
+
+**Joint**:
+
+```{r, cache = TRUE}
+growth_df <- bind_rows(
+  bw_growth_df %>% select(growth) %>% mutate(site = "bw"),
+  scbi_growth_df %>% select(growth) %>% mutate(site = "scbi")
+)
+ggplot(growth_df, aes(x = growth, y = ..density.., fill = site)) +
+  geom_histogram(alpha = 0.5, position = "identity", binwidth = 0.5) +
+  labs(x = "Average annual growth in dbh") +
+  coord_cartesian(xlim = c(-5, 10))
+```

### 2020-01-26 README.Rmd
Leaving breadcrumbs
Plugging along add_spatial_info()

@@
 library(skimr)
+# devtools::install_github("rvalavi/blockCV")
+library(blockCV)
 ```
@@
 ### Load & preprocess data
+
+We first load and preprocess the data corresponding to the two census reflecting the time period over which we consider growth in dbh.
+We load them into R as "tibble" data frames thereby ensuring a standardized input/output format that can be used across all `tidyverse` packages @tidyverse.
+Furthermore, we ensure that the different variables have the correct names, types (`dbl`, `data`, `factor`).
 
@@
 
+We then combine the two census data frames into a single data frame  that now includes a numerical variable `growth` reflecting the average annual growth of dbh in cm.
+Furthermore, variables that (in theory) remain unchanged between censuses appear only once, such as `gx`, `gy`, species-related variables.
+Variables that should change between censuses are tagged with `1/2` indicating earlier/later, such as `dbh1/dbh2`, `codes1/codes2`.
+
+The resulting data frames are named with some variation of `growth_df`.
+
 **Big Woods**:
@@
 
-**Joint**:
+**EDA of both BigWoods & SCBI*: Note the large variation in growths for the SCBI trees over the BigWoods trees.
 
@@
 ```
+
+### Add spatial information
+
+We now add spatial information to the `growth_df` data frame. Specifically:
+
+1.
+In order to control for study region "edge effects" (cite Waller?), we need a function that adds "buffers" of trees.
+In our case, since our model of interspecies competition relies on a spatial definition of who is a "neighboring competitor" and certain explanatory varibles such as biomass are cummulative, we need to ensure that all trees being modeled are not biased to have different neighbor structure, in particular the trees at the boundary of study regions.
+1.
+Assign each tree to a "fold" for cross-validation purposes.
+Conventional cross-validation schemes assign observations to folds by resampling individual observations at random.
+However, underlying this scheme is an assumption that the observations are independent.
+In the case of forest census data, observations exhibit spatial autocorrelation, and thus this dependence must be incorporated in our resampling scheme @roberts2017 @pohjankukka2017.
+Packages that have implemented spatial cross-validation include @valavi2019
+
+Thus, before we can incorporate the above information to `growth_df`, we need to define two constants:
+
+```{r}
+cv_fold_size <- 100
+max_dist <- 7.5
+```
+
+For a focal tree of interest, the `max_dist` of `r max_dist` meters acts as a radius defining a neighborhood within which all trees are considered competitors.
+In other words, all trees within `r max_dist` meters of the focal tree are considered competitor trees.
+Other studies have estimated this distance; we used `r max_dist` meters as an average of estimated values [@canham2004, @canham2006, @uriarte2004, @tatsumi2013].
+Following Tobler's first law of geography that "everything is related to everything else, but near things are more related than distant things." @tobler1970firstlawofgeo, we assume that the degree of spatial autocorrelation is inversely-related to distance.
+However, we further assume that once trees are more than `r max_dist` meters apart, this autocorrelation is negligeable.
+
+Using the [`blockCV`](https://github.com/rvalavi/blockCV) package, in particular the example code in the package [vignette](http://htmlpreview.github.io/?https://github.com/rvalavi/blockCV/blob/master/vignettes/BlockCV_for_SDM.html), we
+
+* Define the blocking structure
+* Determine the degree of spatial autocorrelation for various `max_dist` values
+
+```{r}
+# spatial blocking by specified range with random assignment
+
+```
+
+If the above doesn't work, the go back to what we used for PLOS paper.
+
+**Big Woods**:
+
+```{r, cache = TRUE}
+
+```
+
+**SCBI**:
+
+```{r, cache = TRUE}
+```

### 2020-01-30 README.Rmd
Added sf pkg to README
Added initial attempt at define_buffer()

@@
   message = FALSE,
-  warning = FALSE
+  warning = FALSE,
+  cache = TRUE
 )
@@
 library(skimr)
+library(sf)
+library(sfheaders)
 # devtools::install_github("rvalavi/blockCV")
-library(blockCV)
+# library(blockCV)
 ```
@@
 
-```{r, cache=TRUE}
+```{r}
 # Read in Big Woods census data from 2008 & 2014
@@
 
-```{r, cache=TRUE}
+```{r}
 scbi_2013 <- 
@@
 
-```{r, cache=TRUE}
+```{r}
 census_df1 <- bw_2008
@@
 
-```{r, cache=TRUE}
+```{r}
 census_df1 <- scbi_2013
@@
 
-If the above doesn't work, the go back to what we used for PLOS paper.
+* If the above doesn't work, the go back to what we used for PLOS paper.
+* Consider making `growth_df` an `sf` object
 
@@
 ```{r, cache = TRUE}
+square <- tibble(
+  x = c(0,0,1,1),
+  y = c(0,1,1,0)
+) %>% 
+  sf_polygon()
+buffer <- square %>% 
+  st_buffer(dist = -0.1)
+ggplot() +
+  geom_sf(data = square) +
+  geom_sf(data = buffer, col="red")
 
+# Bigwoods study region boundary polygon
+bw_boundary <- bigwoods_study_region %>% 
+  sf_polygon()
+
+# Buffer polygon
+bw_buffer <- bw_boundary %>%
+  st_buffer(dist = -25)
+
+# Sample of 5000 trees
+set.seed(76)
+bw_trees_sample <- bw_growth_df %>% 
+  sample_n(5000) %>% 
+  select(x = gx, y = gy) %>% 
+  st_as_sf(coords = c("x", "y"))
+
+index <- st_intersects(bw_trees_sample, bw_buffer, sparse = FALSE)
+bw_trees_sample <- bw_trees_sample %>% 
+  mutate(buffer = index)
+
+# Plot
+ggplot() +
+  geom_sf(data = bw_boundary) +
+  geom_sf(data = bw_buffer, col="red") +
+  geom_sf(data = bw_trees_sample, aes(col=buffer), size = 0.5)
 ```

### 2020-01-30 README.Rmd
Cleaned up README a bit

@@
 
-**EDA of both BigWoods & SCBI*: Note the large variation in growths for the SCBI trees over the BigWoods trees.
+**EDA of both BigWoods & SCBI**: Note the large variation in growths for the SCBI trees over the BigWoods trees.
 
@@
 
+#### Defining buffers
+
 For a focal tree of interest, the `max_dist` of `r max_dist` meters acts as a radius defining a neighborhood within which all trees are considered competitors.
@@
 However, we further assume that once trees are more than `r max_dist` meters apart, this autocorrelation is negligeable.
-
-Using the [`blockCV`](https://github.com/rvalavi/blockCV) package, in particular the example code in the package [vignette](http://htmlpreview.github.io/?https://github.com/rvalavi/blockCV/blob/master/vignettes/BlockCV_for_SDM.html), we
-
-* Define the blocking structure
-* Determine the degree of spatial autocorrelation for various `max_dist` values
-
-```{r}
-# spatial blocking by specified range with random assignment
-
-```
-
-* If the above doesn't work, the go back to what we used for PLOS paper.
-* Consider making `growth_df` an `sf` object
 
@@
 bw_buffer <- bw_boundary %>%
-  st_buffer(dist = -25)
+  st_buffer(dist = -max_dist)
 
@@
 ```
+
+#### Spatial cross-validation
+
+Using the [`blockCV`](https://github.com/rvalavi/blockCV) package, in particular the example code in the package [vignette](http://htmlpreview.github.io/?https://github.com/rvalavi/blockCV/blob/master/vignettes/BlockCV_for_SDM.html), we
+
+* Define the blocking structure
+* Determine the degree of spatial autocorrelation for various `max_dist` values
+* If the above doesn't work, the go back to what we used for PLOS paper.

### 2020-02-06 README.Rmd
found the problem for why the SCBI and BW histograms looked different

@@
 census_df1 <- bw_2008
-census_df2 <- bw_2014
+# we need to filter out the resprouts
+census_df2 <- bw_2014 %>% filter(codes != 'R')
 id <- "treeID"
@@
   # Merge both censuses and compute growth:
-  compute_growth(census_df1, census_df2, id)
+  compute_growth(census_df1, census_df2, id) %>%
+  # they are mesuaring in mm while we are measuring in cm!!!
+  mutate(growth = growth/10)
 ```
@@
 ggplot(growth_df, aes(x = growth, y = ..density.., fill = site)) +
-  geom_histogram(alpha = 0.5, position = "identity", binwidth = 0.5) +
+  geom_histogram(alpha = 0.5, position = "identity", binwidth = 0.05) +
   labs(x = "Average annual growth in dbh") +
-  coord_cartesian(xlim = c(-5, 10))
+  coord_cartesian(xlim = c(-0.5, 1))
 ```

### 2020-02-06 README.Rmd
small change to filter out resprouts
specify of growth

@@
 # we need to filter out the resprouts
-census_df2 <- bw_2014 %>% filter(codes != 'R')
+census_df2 <- bw_2014 %>% filter(!str_detect(codes,'R'))
 id <- "treeID"
@@
   geom_histogram(alpha = 0.5, position = "identity", binwidth = 0.05) +
-  labs(x = "Average annual growth in dbh") +
+  labs(x = "Average annual growth in dbh (cm per yr)") +
   coord_cartesian(xlim = c(-0.5, 1))

### 2020-02-13 README.Rmd
Created buffer for SCBI

@@
 
-**Big Woods**:
+**Example of buffer at work**:
 
 ```{r, cache = TRUE}
-square <- tibble(
+# Boundary polygon
+square_boundary <- tibble(
   x = c(0,0,1,1),
@@
   sf_polygon()
-buffer <- square %>% 
+
+# Buffer polygon
+square_buffer <- square_boundary %>% 
   st_buffer(dist = -0.1)
+
 ggplot() +
-  geom_sf(data = square) +
-  geom_sf(data = buffer, col="red")
+  geom_sf(data = square_boundary) +
+  geom_sf(data = square_buffer, col="red") 
+```
 
+**Big Woods**:
+
+```{r, cache = TRUE}
 # Bigwoods study region boundary polygon
@@
 
-# Sample of 5000 trees
-set.seed(76)
-bw_trees_sample <- bw_growth_df %>% 
-  sample_n(5000) %>% 
-  select(x = gx, y = gy) %>% 
-  st_as_sf(coords = c("x", "y"))
+# Convert data frame to sf object
+bw_growth_df <- bw_growth_df %>% 
+  st_as_sf(coords = c("gx", "gy"))
 
-index <- st_intersects(bw_trees_sample, bw_buffer, sparse = FALSE)
-bw_trees_sample <- bw_trees_sample %>% 
+# ID which points are in buffer and which are not
+index <- st_intersects(bw_growth_df, bw_buffer, sparse = FALSE)
+bw_growth_df <- bw_growth_df %>% 
   mutate(buffer = index)
@@
   geom_sf(data = bw_buffer, col="red") +
-  geom_sf(data = bw_trees_sample, aes(col=buffer), size = 0.5)
+  # Only random sample of 1000 trees:
+  geom_sf(data = bw_growth_df %>% sample_n(1000), aes(col=buffer), size = 0.5)
 ```
@@
 ```{r, cache = TRUE}
+# Bigwoods study region boundary polygon
+scbi_boundary <- tibble(
+  x = c(0, 0, 400, 400, 0),
+  y = c(0, 640, 640, 0, 0)
+) %>% 
+  sf_polygon()
+
+# Buffer polygon
+scbi_buffer <- scbi_boundary %>%
+  st_buffer(dist = -max_dist)
+
+# Convert data frame to sf object
+scbi_growth_df <- scbi_growth_df %>% 
+  st_as_sf(coords = c("gx", "gy"))
+
+# ID which points are in buffer and which are not
+index <- st_intersects(scbi_growth_df, scbi_buffer, sparse = FALSE)
+scbi_growth_df <- scbi_growth_df %>% 
+  mutate(buffer = index)
+
+# Plot
+ggplot() +
+  geom_sf(data = scbi_boundary) +
+  geom_sf(data = scbi_buffer, col="red") +
+  # Only random sample of 1000 trees:
+  geom_sf(data = scbi_growth_df %>% sample_n(1000), aes(col=buffer), size = 0.5)
 ```

### 2020-02-13 README.Rmd
Add buffer for SCBI

--- 2020-02-06 README.Rmd
@@
 
-**EDA of both BigWoods & SCBI**: Note the large variation in growths for the SCBI trees over the BigWoods trees.
-
 ```{r, cache = TRUE}
@@
 ```{r, cache = TRUE}
+scbi_boundary <- tibble(x = c(0,400,400,0,0), y = c(0,0,640,640,0)) %>% 
+  sf_polygon()
+
+# Buffer polygon
+scbi_buffer <- scbi_boundary %>%
+  st_buffer(dist = -max_dist)
+
+# Sample of 5000 trees
+set.seed(76)
+scbi_trees_sample <- scbi_growth_df %>% 
+  sample_n(5000) %>% 
+  select(x = gx, y = gy) %>% 
+  st_as_sf(coords = c("x", "y"))
+
+index <- st_intersects(scbi_trees_sample, scbi_buffer, sparse = FALSE)
+scbi_trees_sample <- scbi_trees_sample %>% 
+  mutate(buffer = index)
+
+# Plot
+ggplot() +
+  geom_sf(data = scbi_boundary) +
+  geom_sf(data = scbi_buffer, col="red") +
+  geom_sf(data = scbi_trees_sample, aes(col=buffer), size = 0.5)
 ```

### 2020-02-13 README.Rmd
Added CV grid

--- 2020-02-13 README.Rmd
@@
 # devtools::install_github("rvalavi/blockCV")
-# library(blockCV)
+library(blockCV)
 ```
@@
 
+**Big Woods**:
+
+```{r, cache = TRUE}
+bw_cv_grid <- spatialBlock(
+  speciesData = bw_growth_df, theRange = 100, k = 28, yOffset = 0.6
+)
+
+# Add foldID to data
+bw_growth_df <- bw_growth_df %>% 
+  mutate(
+    foldID = bw_cv_grid$foldID,
+    foldID = factor(foldID)
+  )
+
+# Visualize grid
+bw_cv_grid$plots +
+  geom_sf(data = bw_growth_df, aes(col=foldID), size = 0.1)
+```
+
+**SCBI**:
+
+```{r, cache = TRUE}
+scbi_cv_grid <- spatialBlock(
+  speciesData = scbi_growth_df, theRange = 100, k = 28, yOffset = 0.6
+)
+
+# Add foldID to data
+scbi_growth_df <- scbi_growth_df %>% 
+  mutate(
+    foldID = scbi_cv_grid$foldID,
+    foldID = factor(foldID)
+  )
+
+# Visualize grid
+scbi_cv_grid$plots +
+  geom_sf(data = scbi_growth_df, aes(col=foldID), size = 0.1)
+```
+
 Using the [`blockCV`](https://github.com/rvalavi/blockCV) package, in particular the example code in the package [vignette](http://htmlpreview.github.io/?https://github.com/rvalavi/blockCV/blob/master/vignettes/BlockCV_for_SDM.html), we

### 2020-02-13 README.Rmd

@@
 
-**EDA of both BigWoods & SCBI**: Note the large variation in growths for the SCBI trees over the BigWoods trees.
-
-```{r, cache = TRUE}
+```{r}
 growth_df <- bind_rows(
@@
 
-```{r, cache = TRUE}
+```{r}
 # Boundary polygon
@@
 
-```{r, cache = TRUE}
+```{r}
 # Bigwoods study region boundary polygon
@@
 
-```{r, cache = TRUE}
+```{r}
 # Bigwoods study region boundary polygon
-scbi_boundary <- tibble(
-  x = c(0, 0, 400, 400, 0),
-  y = c(0, 640, 640, 0, 0)
-) %>% 
+scbi_boundary <- tibble(x = c(0,400,400,0,0), y = c(0,0,640,640,0)) %>% 
   sf_polygon()
@@
 
-```{r, cache = TRUE}
+```{r}
 bw_cv_grid <- spatialBlock(
-  speciesData = bw_growth_df, theRange = 100, k = 28, yOffset = 0.6
+  speciesData = bw_growth_df, theRange = 100, k = 28, yOffset = 0.6, verbose = FALSE
 )
@@
 
-```{r, cache = TRUE}
+```{r}
 scbi_cv_grid <- spatialBlock(
-  speciesData = scbi_growth_df, theRange = 100, k = 28, yOffset = 0.6
+  speciesData = scbi_growth_df, theRange = 100, k = 28, yOffset = 0.6, verbose = FALSE
 )

### 2020-02-13 README.Rmd
Tweaking BigWoods CV grid
Cleaned up spatial CV

@@
 
+We use the [`blockCV`](https://github.com/rvalavi/blockCV) package to define the spatial grid, whose elements will act as the folds in our leave-one-out (by "one" we meen "one grid block") cross-validation scheme.
+
 **Big Woods**:
@@
 bw_cv_grid <- spatialBlock(
-  speciesData = bw_growth_df, theRange = 100, k = 28, yOffset = 0.6, verbose = FALSE
+  speciesData = bw_growth_df, theRange = 100, k = 28, 
+  xOffset = 0.5, yOffset = 0.99999,
+  verbose = FALSE
 )
@@
 ```
-
-Using the [`blockCV`](https://github.com/rvalavi/blockCV) package, in particular the example code in the package [vignette](http://htmlpreview.github.io/?https://github.com/rvalavi/blockCV/blob/master/vignettes/BlockCV_for_SDM.html), we
-
-* Define the blocking structure
-* Determine the degree of spatial autocorrelation for various `max_dist` values
-* If the above doesn't work, the go back to what we used for PLOS paper.

### 2020-02-13 README.Rmd
Added outline for remainder of code

@@
 ```
+
+### Model specification
+
+* Model formula: Growth as a function of
+
+* `get_model_specs()`. Inputs:
+    + define identification/grouping classification
+    + `growth_df` format
+* `create_focal_and_comp()`. Inputs:
+    + `growth_df`
+    + `max_dist`: who are your competitors
+    + `model_specs`: grouping of competitor biomasses
+
+### Model fit and prediction
+
+* `fit_bayesian_model()`. 
+    + Inputs: `focal_and_comp`, `model_specs`
+    + Output: `posterior_param`
+* `predict_bayesian_model()`. 
+    + Inputs: `focal_and_comp`, `model_specs`, `posterior_param`
+    + Output: `focal_and_comp` with updated `growth_hat` variable
+* Show observed vs predicted growth
+* CV: `run_cv()` is a wrapper to `fit_bayesian_model()` and `predict_bayesian_model()`. 
+* Permutation test:
+  + Add line to that shuffles competitor species within `focal_and_comp` just before fitting model
+  + permutation AND CV: add `run_shuffle` argument to `run_cv()`
+
+### Model output and performance
+
+* Make plot posterior parameter functions

### 2020-02-25 README.Rmd
Change to make get_model_specs more flexible

@@
   read_delim(delim = "\t") %>% 
+  mutate(spcode = to_any_case(spcode)) %>%
   select(
@@
   ) %>% 
-  select(sp, spcode, genus, species, latin, family, trait_group)
+  select(sp = spcode, genus, species, latin, family, trait_group)
+
+bw_2008 <- bw_2008 %>%
+  left_join(bw_species,by='sp')
+
 ```
@@
 
-* Model formula: Growth as a function of
+Next we specify the growth model we want.
+We will model growth as a function of tree identiy, size, and the size and identiy of its neighbors.
+The model can be run on different grouping of individuals (e.g., based on species or trait groupings).
+It can also be run with different measures of competition.
+
+* `model_number = 1`: No competition growth only depends on `dbh` and individual grouping.
+* `model_number = 2`: Competition but identity of competitor does not matter, just sum of competitior biomass.
+* `model_number = 3`: Full model with competition and competitor identity.
+
+```{r}
+bw_specs <- get_model_specs(bw_2008, 3, 'trait_group')
+bw_specs
+
+scbi_specs <- get_model_specs(scbi_2013, 3, 'sp')
+
+```
 

### 2020-02-25 R/modeling.R

@@ -9,45 +9,24 @@
 #' @export
 #' @examples
 #' 1+1
-get_model_specs <- function(forest, model_number){
+get_model_specs <- function(forest, model_number, species_notion){
   # Define model formula to use out of 5 possible choices defined in paper
   # TODO: Generalize this for any species/family list
 
   model_1_formula <-
-    "growth ~ species + dbh + dbh*species"
+    paste0("growth ~ ", species_notion, " + dbh + dbh*", species_notion)
 
   model_2_formula <-
-    "growth ~ species + dbh + dbh*species + biomass + biomass*species"
+    paste0("growth ~ ", species_notion, " + dbh + dbh*", species_notion, " + biomass + biomass*", species_notion)
 
-  model_3_formula <- forest$trait_group %>%
+
+  model_3_formula <- forest %>%
+    pull(species_notion) %>%
     unique() %>%
     sort() %>%
-    paste(., "*species", sep = "", collapse = " + ") %>%
-    paste("growth ~ species + dbh + dbh*species + ", .)
-
-  model_4_formula <-
-    "growth ~ species + dbh + dbh*species"
-
-  model_5_formula <-
-    "growth ~ species + dbh + dbh*species + biomass + biomass*species"
-
-  model_6_formula <- forest$family_phylo %>%
-    unique() %>%
-    sort() %>%
-    paste(., "*species", sep = "", collapse = " + ") %>%
-    paste("growth ~ species + dbh + dbh*species + ", .)
-
-  model_7_formula <-
-    "growth ~ species + dbh + dbh*species"
-
-  model_8_formula <-
-    "growth ~ species + dbh + dbh*species + biomass + biomass*species"
-
-  model_9_formula <- forest$species %>%
-    unique() %>%
-    sort() %>%
-    paste(., "*species", sep = "", collapse = " + ") %>%
-    paste("growth ~ species + dbh + dbh*species + ", .)
+    paste0('`', ., '`') %>%
+    paste(., "*", species_notion, sep = "", collapse = " + ") %>%
+    paste(model_2_formula,'+', .)
 
   # Convert to formula object:
   model_formula <- model_number %>%
@@ -55,26 +34,18 @@ get_model_specs <- function(forest, model_number){
     get() %>%
     as.formula()
 
-  # Notion of species for both focal and competitor trees. Either:
-  # + "trait_group" (6 levels) or
-  # + "family_phylo" (23 levels) or
-  # + "species" (46 levels)
[…]

### 2020-03-04 README.Rmd
Working on the readme and some changes to package.

@@
   compute_growth(census_df1, census_df2, id) %>% 
-  mutate(sp = to_any_case(sp)) %>% 
-  left_join(bw_species, by = c("sp" = "spcode"))
+  mutate(sp = to_any_case(sp))
 ```
@@
 bw_growth_df <- bw_growth_df %>% 
-  mutate(buffer = index)
+  mutate(buffer = index[,1])
 
@@
 scbi_growth_df <- scbi_growth_df %>% 
-  mutate(buffer = index)
+  mutate(buffer = index[,1])
 
@@
 
-* `model_number = 1`: No competition growth only depends on `dbh` and individual grouping.
-* `model_number = 2`: Competition but identity of competitor does not matter, just sum of competitior biomass.
-* `model_number = 3`: Full model with competition and competitor identity.
+* `model_number = 1`: No competition growth only depends on focal tree `dbh` and grouping.
+* `model_number = 2`: Competition but identity of competitor does not matter, just sum of competitior basal area.
+* `model_number = 3`: Full model with competition including competitor identity.
 
@@
 
-* `get_model_specs()`. Inputs:
-    + define identification/grouping classification
-    + `growth_df` format
-* `create_focal_and_comp()`. Inputs:
-    + `growth_df`
-    + `max_dist`: who are your competitors
-    + `model_specs`: grouping of competitor biomasses
+Next we create the `focal_vs_comp` object which connects each tree to the trees in its competitive neighborhood (with `max_dist`).
+This required `growth_df`, `max_dist`, `cv_fold_size`, and `model_specs` as inputs.
+
+```{r}
+bw_focal_vs_comp <- create_focal_vs_comp(bw_growth_df, max_dist, cv_fold_size, bw_specs)
+scbi_focal_vs_comp <- create_focal_vs_comp(scbi_growth_df, max_dist, cv_fold_size, scbi_specs)
+
+```
 

### 2020-03-04 R/modeling.R

@@ -17,7 +17,7 @@ get_model_specs <- function(forest, model_number, species_notion){
     paste0("growth ~ ", species_notion, " + dbh + dbh*", species_notion)
 
   model_2_formula <-
-    paste0("growth ~ ", species_notion, " + dbh + dbh*", species_notion, " + biomass + biomass*", species_notion)
+    paste0("growth ~ ", species_notion, " + dbh + dbh*", species_notion, " + comp_basal_area + comp_basal_area*", species_notion)
 
 
   model_3_formula <- forest %>%

### 2020-03-04 R/preprocess_data.R

@@ -63,7 +63,7 @@ compute_growth <- function(census_df1, census_df2, id) {
 create_focal_vs_comp <- function(forest, max_dist, cv_fold_size, model_specs){
 
   # get folds info
-  folds <- growth_df %>%
+  folds <- forest %>%
     filter(!buffer) %>%
     get_cv_fold_info(cv_fold_size) %>%
     mutate(

### 2020-03-04 R/preprocess_data.R
Made some changes but not running! SOrry. I left a note where the problem is.

@@ -90,7 +90,7 @@ create_focal_vs_comp <- function(forest, max_dist, cv_fold_size, model_specs){
     mutate(spCode = as.numeric(factor(species))) %>%
     # ID numbers to join focal trees with competitor trees
     mutate(focal_ID = 1:n()) %>%
-    select(focal_ID, ID, species, spCode, x, y, dbh, growth, fold)
+    select(focal_ID, ID, species, spCode, x, y, dbh, growth, foldID)
 
   # 2. Define competitor trees where notion of "species" depends on
   # notion_of_competitor_species
@@ -107,7 +107,7 @@ create_focal_vs_comp <- function(forest, max_dist, cv_fold_size, model_specs){
       # https://en.wikipedia.org/wiki/Basal_area
       basal_area = 0.0001 * pi * (dbh/2)^2
     ) %>%
-    select(comp_ID, ID, species, x, y, dbh, basal_area, fold)
+    select(comp_ID, ID, species, x, y, dbh, basal_area, foldID)
 
 
   # 2. Define distances of focal trees to other focal trees.
@@ -119,11 +119,11 @@ create_focal_vs_comp <- function(forest, max_dist, cv_fold_size, model_specs){
   # notion_of_competitor_species == "species"
   # species_of_interest == unique(bw$species)
   focal_vs_focal <- NULL
-  for(i in 1:max(forest$fold)){
+  for(i in 1:max(forest$foldID)){
     x <- focal_trees %>%
-      filter(fold == i)
+      filter(foldID == i)
     y <- focal_trees %>%
-      filter(fold %in% unlist(folds$data[folds$fold == i]))
+      filter(foldID %in% unlist(folds$data[folds$foldID == i]))
 
     D <- proxy::dist(x=x[, c("x", "y")], y=y[, c("x", "y")])
 
@@ -155,11 +155,11 @@ create_focal_vs_comp <- function(forest, max_dist, cv_fold_size, model_specs){
   # notion_of_competitor_species == "species"
   # species_of_interest == unique(bw$species)
   focal_vs_comp <- NULL
-  for(i in 1:max(forest$fold)){
+  for(i in 1:max(forest$foldID)){
     x <- focal_trees %>%
-      filter(fold == i)
+      filter(foldID == i)
     y <- comp_trees %>%
-      filter(fold %in% unlist(folds$data[i]))
[…]

### 2020-03-04 R/spatial.R

@@ -132,9 +132,9 @@ get_cv_fold_info <- function(forest, cv_fold_size){
   # Save output here
   fold_neighbors <- NULL
 
-  for(i in 1:n_distinct(forest$fold)){
+  for(i in 1:n_distinct(forest$foldID)){
     center_tree <- forest %>%
-      filter(fold == i) %>%
+      filter(foldID == i) %>%
       summarise(x = mean(x), y = mean(y))
 
     fold_neighbors <- forest %>%
@@ -142,7 +142,7 @@ get_cv_fold_info <- function(forest, cv_fold_size){
         between(x, center_tree$x - cv_fold_size, center_tree$x + cv_fold_size),
         between(y, center_tree$y - cv_fold_size, center_tree$y + cv_fold_size)
       ) %>%
-      pull(fold) %>%
+      pull(foldID) %>%
       unique() %>%
       sort() %>%
       data_frame(neighbor = .) %>%
@@ -151,13 +151,13 @@ get_cv_fold_info <- function(forest, cv_fold_size){
         x = center_tree$x,
         y = center_tree$y
       ) %>%
-      select(fold, x, y, neighbor) %>%
+      select(foldID, x, y, neighbor) %>%
       nest(data=c(neighbor)) %>%
       bind_rows(fold_neighbors)
   }
 
   fold_neighbors <- fold_neighbors %>%
-    arrange(fold)
+    arrange(foldID)
   return(fold_neighbors)
 }
 

### 2020-03-04 README.Rmd
Made some changes but not running! SOrry. I left a note where the problem is.
comment out the parts that don't work

@@
 
+**DA NOTE: THIS IS BROKEN HERE!!
+The problem is the new `sf` structure of `growth_df` breaks `get_cv_fold_info` which is called in `create_focal_vs_comp`.
+Also at some point `fold` in `growth_bf` became `foldID`.
+So I had to go change that in a bunch of the package code.**
+
 ```{r}
-bw_focal_vs_comp <- create_focal_vs_comp(bw_growth_df, max_dist, cv_fold_size, bw_specs)
-scbi_focal_vs_comp <- create_focal_vs_comp(scbi_growth_df, max_dist, cv_fold_size, scbi_specs)
+#bw_focal_vs_comp <- create_focal_vs_comp(bw_growth_df, max_dist, cv_fold_size, bw_specs)
+#scbi_focal_vs_comp <- create_focal_vs_comp(scbi_growth_df, max_dist, cv_fold_size, scbi_specs)
 

### 2020-03-06 README.Rmd
Done create_focal_vs_comp()

--- 2020-03-04 README.Rmd
@@
 # we need to filter out the resprouts
-census_df2 <- bw_2014 %>% filter(!str_detect(codes,'R'))
+census_df2 <- bw_2014 %>% 
+  filter(!str_detect(codes, 'R'))
 id <- "treeID"
@@
 # ID which points are in buffer and which are not
-index <- st_intersects(bw_growth_df, bw_buffer, sparse = FALSE)
+buffer_index <- !st_intersects(bw_growth_df, bw_buffer, sparse = FALSE)
 bw_growth_df <- bw_growth_df %>% 
-  mutate(buffer = index[,1])
+  mutate(buffer = as.vector(buffer_index))
 
@@
 # ID which points are in buffer and which are not
-index <- st_intersects(scbi_growth_df, scbi_buffer, sparse = FALSE)
+buffer_index <- st_intersects(scbi_growth_df, scbi_buffer, sparse = FALSE)
 scbi_growth_df <- scbi_growth_df %>% 
-  mutate(buffer = index[,1])
+  mutate(buffer = as.vector(buffer_index))
 
@@
 ```{r}
+set.seed(76)
 bw_cv_grid <- spatialBlock(
-  speciesData = bw_growth_df, theRange = 100, k = 28, 
-  xOffset = 0.5, yOffset = 0.99999,
-  verbose = FALSE
+  speciesData = bw_growth_df, theRange = 100, verbose = FALSE,
+  # Some guess work in figuring this out:
+  k = 28, xOffset = 0.5, yOffset = 0
 )
@@
   mutate(
-    foldID = bw_cv_grid$foldID,
-    foldID = factor(foldID)
+    foldID = bw_cv_grid$foldID
   )
 
-# Visualize grid
+# Visualize grid. Why does fold 19 repeat?
 bw_cv_grid$plots +
-  geom_sf(data = bw_growth_df, aes(col=foldID), size = 0.1)
+  geom_sf(data = bw_growth_df %>% sample_frac(0.2), aes(col=factor(foldID)), size = 0.1)
+
+# Remove weird folds with no trees in them from viz above
+bw_growth_df <- bw_growth_df %>%
+  filter(!foldID %in% c(19, 23, 21, 17, 8, 19))
 ```
@@
 scbi_cv_grid <- spatialBlock(
-  speciesData = scbi_growth_df, theRange = 100, k = 28, yOffset = 0.6, verbose = FALSE
+  speciesData = scbi_growth_df, theRange = 100, k = 28, yOffset = 0.9999, verbose = FALSE
 )
@@
   mutate(
-    foldID = scbi_cv_grid$foldID,
-    foldID = factor(foldID)
+    foldID = scbi_cv_grid$foldID
   )
@@
 scbi_cv_grid$plots +
-  geom_sf(data = scbi_growth_df, aes(col=foldID), size = 0.1)
+  geom_sf(data = scbi_growth_df, aes(col=factor(foldID)), size = 0.1)
 ```
@@
 Next we specify the growth model we want.
-We will model growth as a function of tree identiy, size, and the size and identiy of its neighbors.
+We will model growth as a function of tree identity, size (in dbh), and the identity and size (in dbh) of its neighbors.
 The model can be run on different grouping of individuals (e.g., based on species or trait groupings).
-It can also be run with different measures of competition.
+It can also be run with different notions of competition:
 
@@
 ```{r}
-bw_specs <- get_model_specs(bw_2008, 3, 'trait_group')
+bw_specs <- bw_growth_df %>% 
+  get_model_specs(model_number = 3, species_notion = 'trait_group')
 bw_specs
 
-scbi_specs <- get_model_specs(scbi_2013, 3, 'sp')
-
+scbi_specs <- scbi_growth_df %>% 
+  get_model_specs(model_number = 3, species_notion = 'sp')
 ```
 
-Next we create the `focal_vs_comp` object which connects each tree to the trees in its competitive neighborhood (with `max_dist`).
-This required `growth_df`, `max_dist`, `cv_fold_size`, and `model_specs` as inputs.
+Next we create the `focal_vs_comp` data frame which connects each tree in `growth_df` to the trees in its competitive neighborhood range (as defined by `max_dist`).
+So for example, if `growth_df` consisted of two focal trees with two and three neighbors respectively, `focal_vs_comp` would consist of 5 rows.
+
+This requires the `growth_df` data frame, `max_dist` scalar defining competitive range, `cv_fold_size` defining the size of the spatial cross-validation blocks, and `model_specs` as inputs.
 
 ```{r}
-bw_focal_vs_comp <- create_focal_vs_comp(bw_growth_df, max_dist, cv_fold_size, bw_specs)
-scbi_focal_vs_comp <- create_focal_vs_comp(scbi_growth_df, max_dist, cv_fold_size, scbi_specs)
+focal_vs_comp_bw <- bw_growth_df %>% 
+  create_focal_vs_comp(max_dist, model_specs = bw_specs, cv_grid = bw_cv_grid, id = "treeID")
+```
 
+```{r}
+glimpse(focal_vs_comp_bw)
 ```

### 2020-03-06 R/modeling.R

@@ -1,24 +1,23 @@
 #' Define notion of species based on model choice
 #'
 #' @inheritParams define_cv_grid
-#' @param model_number Number of model out of 1-5 from paper
+#' @param model_number Number of model out of 1-3 from paper
+#' @param species_notion Notion of grouping of individuals
 #'
 #' @import dplyr
 #' @importFrom stats as.formula
+#' @importFrom stringr str_c
 #' @return A list of outputs
 #' @export
 #' @examples
 #' 1+1
 get_model_specs <- function(forest, model_number, species_notion){
-  # Define model formula to use out of 5 possible choices defined in paper
-  # TODO: Generalize this for any species/family list
-
+  # Define 3 possible models for 3 notions of competition
   model_1_formula <-
-    paste0("growth ~ ", species_notion, " + dbh + dbh*", species_notion)
+    paste0("growth ~ ", species_notion, " + dbh + dbh * ", species_notion)
 
   model_2_formula <-
-    paste0("growth ~ ", species_notion, " + dbh + dbh*", species_notion, " + comp_basal_area + comp_basal_area*", species_notion)
-
+    paste0("growth ~ ", species_notion, " + dbh + dbh * ", species_notion, " + comp_basal_area + comp_basal_area * ", species_notion)
 
   model_3_formula <- forest %>%
     pull(species_notion) %>%
@@ -26,9 +25,9 @@ get_model_specs <- function(forest, model_number, species_notion){
     sort() %>%
     paste0('`', ., '`') %>%
     paste(., "*", species_notion, sep = "", collapse = " + ") %>%
-    paste(model_2_formula,'+', .)
+    paste(model_2_formula, '+', .)
 
-  # Convert to formula object:
+  # Convert desied model to formula object:
   model_formula <- model_number %>%
     paste("model_", ., "_formula", sep="") %>%
     get() %>%
@@ -37,9 +36,8 @@ get_model_specs <- function(forest, model_number, species_notion){
   # Species of interest to model. These should be a subset of the species
   # correspoding to notion_of_focal_species.
   species_of_interest <- forest %>%
-    select(species_notion) %>%
-    distinct() %>%
-    pull()
+    pull(species_notion) %>%
+    unique()
 
   # Return output list
   output <- list(

### 2020-03-06 R/preprocess_data.R

@@ -60,57 +60,50 @@ compute_growth <- function(census_df1, census_df2, id) {
 #' @seealso \code{\link{define_cv_grid}} and \code{\link{get_model_specs}}
 #' @examples
 #' 1+1
-create_focal_vs_comp <- function(forest, max_dist, cv_fold_size, model_specs){
-
-  # get folds info
-  folds <- forest %>%
-    filter(!buffer) %>%
-    get_cv_fold_info(cv_fold_size) %>%
-    mutate(
-      x = round(x/(cv_fold_size/2))*(cv_fold_size/2),
-      y = round(y/(cv_fold_size/2))*(cv_fold_size/2)
-    )
-
+create_focal_vs_comp <- function(growth_df, max_dist, model_specs, cv_grid, id, plot_folds = FALSE){
   # Extract model specifications
   notion_of_focal_species <- model_specs$notion_of_focal_species
   notion_of_competitor_species <- model_specs$notion_of_competitor_species
-  model_formula <- model_specs$model_formula
-  species_of_interest <- model_specs$species_of_interest
 
   # 1. Define focal trees where notion of "species" depends on
   # notion_of_focal_species
-  focal_trees <- forest %>%
-    # Define notion of speices
-    mutate_(species = notion_of_focal_species) %>%
-    select(-c(family_phylo, trait_group)) %>%
-    # Only alive both dates, not in buffer
-    filter(dbh1>0, dbh2>0, !buffer, code != 'R') %>%
+  growth_df_focal_trees <- growth_df %>%
+    # Define notion of species as factor
+    mutate(
+      focal_notion_of_species = .data[[notion_of_focal_species]],
+      focal_notion_of_species = factor(focal_notion_of_species)
+    ) %>%
+    # Only trees alive both dates, not in buffer, and not a resprout at 2nd
+    # census (OK to be resprout at first census):
+    filter(dbh1 > 0, dbh2 > 0, !buffer, codes2 != 'R') %>%
     rename(dbh = dbh1) %>%
     # Assign species numerical code.
-    mutate(spCode = as.numeric(factor(species))) %>%
+    # mutate(spCode = as.numeric(notion_of_species)) %>%
     # ID numbers to join focal trees with competitor trees
-    mutate(focal_ID = 1:n()) %>%
-    select(focal_ID, ID, species, spCode, x, y, dbh, growth, foldID)
+    mutate(focal_ID = .data[[id]]) %>%
[…]

### 2020-03-09 README.Rmd
Added fit_bayesain_model to readme and necessary changes to package

@@
 
+Now we are ready to fit the competition model with `fit_bayesian_model` this takes two inputs: `focal_vs_comp` the data frame with focal trees connected to their competitors and `model_specs` which specifies the notion of competition.
+
+```{r}
+bw_fit_model <- focal_vs_comp_bw %>% fit_bayesian_model(model_specs = bw_specs)
+```
+
 * `fit_bayesian_model()`. 

### 2020-03-09 R/modeling.R

@@ -75,12 +75,12 @@ fit_bayesian_model <- function(focal_vs_comp, model_specs, run_shuffle = FALSE,
 
   # Prepare data for regression Generate data frame of all focal trees
   focal_trees <- focal_vs_comp %>%
-    group_by(ID, focal_ID, species, spCode, x, y, dbh, growth, fold, comp_species) %>%
-    # Sum biomass & count of all neighbors; set to 0 for cases of no neighbors
+    group_by(focal_ID, focal_notion_of_species, dbh, growth, foldID, comp_notion_of_species) %>%
+    # Sum basal area & count of all neighbors; set to 0 for cases of no neighbors
     # within range.
     summarise(
-      biomass_total = sum(comp_biomass),
-      biomass = sum(comp_biomass),
+      basal_area_total = sum(comp_basal_area),
+      comp_basal_area = sum(comp_basal_area),
       n_comp = n()
     ) %>%
     arrange(focal_ID)
@@ -88,8 +88,8 @@ fit_bayesian_model <- function(focal_vs_comp, model_specs, run_shuffle = FALSE,
   # Shuffle group label only if flag is set
   if(run_shuffle){
     focal_trees <- focal_trees %>%
-      group_by(species) %>%
-      mutate(comp_species = sample(comp_species))
+      group_by(focal_notion_of_species) %>%
+      mutate(comp_notion_of_species = sample(comp_notion_of_species))
   }
 
   # Continue processing focal_trees
@@ -97,16 +97,17 @@ fit_bayesian_model <- function(focal_vs_comp, model_specs, run_shuffle = FALSE,
     ungroup() %>%
     # sum biomass and n_comp for competitors of same species. we need to do this
     # for the cases when we do permutation shuffle.
-    group_by(ID, focal_ID, species, spCode, x, y, dbh, growth, fold, comp_species) %>%
+    group_by(focal_ID, focal_notion_of_species, dbh, growth, foldID, comp_notion_of_species) %>%
     summarise_all(list(sum)) %>%
     ungroup() %>%
     # compute biomass for each tree type
-    spread(comp_species, biomass_total, fill = 0) %>%
-    group_by(ID, focal_ID, species, spCode, x, y, dbh, growth, fold) %>%
+    spread(comp_notion_of_species, basal_area_total, fill = 0) %>%
+    group_by(focal_ID, focal_notion_of_species, dbh, growth, foldID) %>%
     summarise_all(list(sum)) %>%
     ungroup() %>%
[…]

### 2020-03-09 README.Rmd
change from spread to pivot_wider

@@
 
-* `fit_bayesian_model()`. 
-    + Inputs: `focal_and_comp`, `model_specs`
-    + Output: `posterior_param`
+This output has the posterior parameters for the specified competition model.
+If `model_number = 1` this includes just the beta values for how focal tree DBH and identity affect growth.
+If `model_number = 2` it also includes a vector of lambdas for how total competitor basal area affect individuals of each group.
+If `model_number = 3` then lambda is a matrix which gives how individuals of each focal group respond to the competition of each competitor group.
+This `posterior_param` output can be used to get predicted growths for each individual (with `predict_bayesain_model`) to test how well the model performs.
+Or this `posterior_param` output can be plots (either the betas or lambdas) to understand what controls individual growth.
+
 * `predict_bayesian_model()`. 

### 2020-03-09 R/modeling.R

@@ -101,7 +101,7 @@ fit_bayesian_model <- function(focal_vs_comp, model_specs, run_shuffle = FALSE,
     summarise_all(list(sum)) %>%
     ungroup() %>%
     # compute biomass for each tree type
-    spread(comp_notion_of_species, basal_area_total, fill = 0) %>%
+    pivot_wider(names_from = comp_notion_of_species, values_from = basal_area_total, values_fill = list(basal_area_total = 0)) %>%
     group_by(focal_ID, focal_notion_of_species, dbh, growth, foldID) %>%
     summarise_all(list(sum)) %>%
     ungroup() %>%

### 2020-03-09 README.Rmd
Also wrote predict_bayesian_model. For some reason I can get fit_bayesain_model to run in console but not in markdown document.

@@
 # install.packages("remotes")
-remotes::install_github("rudeboybert/forestecology")
+# remotes::install_github("rudeboybert/forestecology")
 ```
@@
 
-```{r}
-bw_fit_model <- focal_vs_comp_bw %>% fit_bayesian_model(model_specs = bw_specs)
+```{r, eval = F}
+bw_fit_model <- focal_vs_comp_bw %>% 
+  fit_bayesian_model(model_specs = bw_specs)
 ```
@@
 Or this `posterior_param` output can be plots (either the betas or lambdas) to understand what controls individual growth.
+
+```{r, eval = F}
+bw_predict <- focal_vs_comp_bw %>% 
+  predict_bayesian_model(model_specs = bw_specs, posterior_param = bw_fit_model)
+
+bw_predict %>%
+  ggplot(aes(growth, growth_hat)) +
+  geom_point() +
+  stat_smooth(method = 'lm') +
+  geom_abline(slope = 1, intercept = 0)
+```
 

### 2020-03-09 R/modeling.R

@@ -192,17 +192,18 @@ fit_bayesian_model <- function(focal_vs_comp, model_specs, run_shuffle = FALSE,
 #' @examples
 #' 1+1
 predict_bayesian_model <- function(focal_vs_comp, model_specs, posterior_param){
+
   # Get model formula
   model_formula <- model_specs$model_formula
 
   # Prepare data for regression Generate data frame of all focal trees
   focal_trees <- focal_vs_comp %>%
-    group_by(ID, focal_ID, species, spCode, x, y, dbh, growth, fold, comp_species) %>%
-    # Sum biomass & count of all neighbors; set to 0 for cases of no neighbors
+    group_by(focal_ID, focal_notion_of_species, dbh, growth, foldID, comp_notion_of_species) %>%
+    # Sum basal area & count of all neighbors; set to 0 for cases of no neighbors
     # within range.
     summarise(
-      biomass_total = sum(comp_biomass),
-      biomass = sum(comp_biomass),
+      basal_area_total = sum(comp_basal_area),
+      comp_basal_area = sum(comp_basal_area),
       n_comp = n()
     ) %>%
     arrange(focal_ID)
@@ -212,16 +213,17 @@ predict_bayesian_model <- function(focal_vs_comp, model_specs, posterior_param){
     ungroup() %>%
     # sum biomass and n_comp for competitors of same species. we need to do this
     # for the cases when we do permutation shuffle.
-    group_by(ID, focal_ID, species, spCode, x, y, dbh, growth, fold, comp_species) %>%
+    group_by(focal_ID, focal_notion_of_species, dbh, growth, foldID, comp_notion_of_species) %>%
     summarise_all(list(sum)) %>%
     ungroup() %>%
     # compute biomass for each tree type
-    spread(comp_species, biomass_total, fill = 0) %>%
-    group_by(ID, focal_ID, species, spCode, x, y, dbh, growth, fold) %>%
+    pivot_wider(names_from = comp_notion_of_species, values_from = basal_area_total, values_fill = list(basal_area_total = 0)) %>%
+    group_by(focal_ID, focal_notion_of_species, dbh, growth, foldID) %>%
     summarise_all(list(sum)) %>%
     ungroup() %>%
     # sort by focal tree ID number
-    arrange(focal_ID)
+    arrange(focal_ID) %>%
+    rename(!!model_specs$notion_of_focal_species := focal_notion_of_species)
 
[…]

### 2020-03-12 README.Rmd
Problem with readme at chunk 19

@@
 bw_predict %>%
+  group_by(focal_ID) %>%
+  summarise(growth = mean(growth), growth_hat = mean(growth_hat)) %>%
   ggplot(aes(growth, growth_hat)) +
-  geom_point() +
+  geom_point(size = 0.5, color = rgb(0,0,0,0.25)) +
   stat_smooth(method = 'lm') +

### 2020-03-12 just_bw_code_blocks.R

@@ -1,0 +1,125 @@
+library(tidyverse)
+library(forestecology)
+library(snakecase)
+library(skimr)
+library(sf)
+library(sfheaders)
+library(blockCV)
+
+bw_2008 <-
+  "https://deepblue.lib.umich.edu/data/downloads/z603qx485" %>%
+  read_delim(delim = "\t") %>%
+  mutate(spcode = to_any_case(spcode)) %>%
+  select(
+    treeID = treeid, stemID = stemtag, sp = spcode, quadrat, gx, gy, dbh,
+    date, codes
+  )
+bw_2014 <-
+  "https://deepblue.lib.umich.edu/data/downloads/1831ck00f" %>%
+  read_delim(delim = "\t") %>%
+  mutate(spcode = to_any_case(spcode)) %>%
+  select(
+    treeID = treeid, stemID = stemtag, sp = spcode, quadrat, gx, gy, dbh,
+    date, codes
+  )
+
+# Read in bigwoods grouping classification data
+bw_species <-
+  "https://deepblue.lib.umich.edu/data/downloads/000000086" %>%
+  read_delim(delim = "\t") %>%
+  # convert all to snake case:
+  mutate_at(c("species", "genus", "family", "idlevel", "spcode"), to_any_case) %>%
+  # join trait group
+  left_join(families, by = c("spcode", "family")) %>%
+  mutate(
+    sp = str_sub(genus, 1, 2),
+    sp = str_c(sp, str_sub(species, 1, 2)),
+    sp = tolower(sp),
+    latin = str_c(genus, species, sep = " "),
+    latin = to_any_case(latin)
+  ) %>%
+  select(sp = spcode, genus, species, latin, family, trait_group)
+
+bw_2008 <- bw_2008 %>%
+  left_join(bw_species,by='sp')
+
+
+census_df1 <- bw_2008
+# we need to filter out the resprouts
+census_df2 <- bw_2014 %>%
+  filter(!str_detect(codes, 'R'))
+id <- "treeID"
+
+bw_growth_df <-
+  # Merge both censuses and compute growth:
+  compute_growth(census_df1, census_df2, id) %>%
+  mutate(sp = to_any_case(sp))
+
+cv_fold_size <- 100
+max_dist <- 7.5
+
+# Bigwoods study region boundary polygon
+bw_boundary <- bigwoods_study_region %>%
+  sf_polygon()
+
[…]

### 2020-03-16 README.Rmd
Bayesian model

--- 2020-03-09 README.Rmd
@@
 ```{r}
-bw_fit_model <- focal_vs_comp_bw %>% fit_bayesian_model(model_specs = bw_specs)
+bw_fit_model <- focal_vs_comp_bw %>% 
+  fit_bayesian_model(model_specs = bw_specs)
 ```
@@
 Or this `posterior_param` output can be plots (either the betas or lambdas) to understand what controls individual growth.
+
+```{r}
+bw_predict <- focal_vs_comp_bw %>% 
+  predict_bayesian_model(model_specs = bw_specs, posterior_param = bw_fit_model)
+
+bw_predict %>%
+  group_by(focal_ID) %>%
+  summarise(growth = mean(growth), growth_hat = mean(growth_hat)) %>%
+  ggplot(aes(growth, growth_hat)) +
+  geom_point(size = 0.5, color = rgb(0,0,0,0.25)) +
+  stat_smooth(method = 'lm') +
+  geom_abline(slope = 1, intercept = 0) +
+  labs(x = "Observed growth in dbh", y = "Predicted growth in dbh")
+```
 

### 2020-03-16 README.Rmd
final tweaks

@@
 bw_predict %>%
-  group_by(focal_ID) %>%
+  group_by(focal_ID, focal_notion_of_species) %>%
+  # Since for the same focal tree all growth & growth_hat values are the same:
   summarise(growth = mean(growth), growth_hat = mean(growth_hat)) %>%
-  ggplot(aes(growth, growth_hat)) +
+  ggplot(aes(x = growth, y = growth_hat)) +
   geom_point(size = 0.5, color = rgb(0,0,0,0.25)) +
@@
   geom_abline(slope = 1, intercept = 0) +
+  coord_fixed() + 
   labs(x = "Observed growth in dbh", y = "Predicted growth in dbh")

### 2020-03-18 README.Rmd
Added TODO's based on meeting
working on fit_bayesian_model

--- 2020-03-16 README.Rmd
@@
 
-#### Spatial cross-validation
+#### Defining spatial cross-validation folds
 
@@
 focal_vs_comp_bw <- bw_growth_df %>% 
-  create_focal_vs_comp(max_dist, model_specs = bw_specs, cv_grid = bw_cv_grid, id = "treeID")
+  create_focal_vs_comp(max_dist = max_dist, model_specs = bw_specs, cv_grid = bw_cv_grid, id = "treeID")
 ```
@@
 
+# TODO: Convert output of predict_bayesian_model() to be of focal_vs_focal type
+# TODO: Preserve geometry variable, that way we can easily create residual plots
+
 bw_predict %>%
-  group_by(focal_ID) %>%
+  group_by(focal_ID, focal_notion_of_species) %>%
+  # Since for the same focal tree all growth & growth_hat values are the same:
   summarise(growth = mean(growth), growth_hat = mean(growth_hat)) %>%
-  ggplot(aes(growth, growth_hat)) +
+  ggplot(aes(x = growth, y = growth_hat)) +
   geom_point(size = 0.5, color = rgb(0,0,0,0.25)) +
@@
   geom_abline(slope = 1, intercept = 0) +
-  labs(x = "Observed growth in dbh", y = "Predicted growth in dbh")
+  coord_fixed() + 
+  labs(
+    x = "Observed growth in dbh", y = "Predicted growth in dbh", 
+    title = "Predicted vs Observed Growth"
+  )
 ```
 
-* `predict_bayesian_model()`. 
-    + Inputs: `focal_and_comp`, `model_specs`, `posterior_param`
-    + Output: `focal_and_comp` with updated `growth_hat` variable
-* Show observed vs predicted growth
+### Run spatial cross-validation
+
 * CV: `run_cv()` is a wrapper to `fit_bayesian_model()` and `predict_bayesian_model()`.
+
+### Run permutations
+
 * Permutation test:
@@
 
-### Model output and performance
+### Visualize posterior distributions
 
-* Make plot posterior parameter functions
+* Plot posterior distributions of all parameters

### 2020-03-18 R/modeling.R

@@ -17,7 +17,8 @@ get_model_specs <- function(forest, model_number, species_notion){
     paste0("growth ~ ", species_notion, " + dbh + dbh * ", species_notion)
 
   model_2_formula <-
-    paste0("growth ~ ", species_notion, " + dbh + dbh * ", species_notion, " + comp_basal_area + comp_basal_area * ", species_notion)
+    paste0("growth ~ ", species_notion, " + dbh + dbh * ", species_notion,
+           " + comp_basal_area + comp_basal_area * ", species_notion)
 
   model_3_formula <- forest %>%
     pull(species_notion) %>%
@@ -70,12 +71,14 @@ get_model_specs <- function(forest, model_number, species_notion){
 fit_bayesian_model <- function(focal_vs_comp, model_specs, run_shuffle = FALSE,
                                prior_hyperparameters = NULL){
 
+  # Save original group_by variables
+  group_variables <- group_vars(focal_vs_comp)
+
   # Get model formula
   model_formula <- model_specs$model_formula
 
-  # Prepare data for regression Generate data frame of all focal trees
+  # Prepare data for regression generate data frame of all focal trees
   focal_trees <- focal_vs_comp %>%
-    group_by(focal_ID, focal_notion_of_species, dbh, growth, foldID, comp_notion_of_species) %>%
     # Sum basal area & count of all neighbors; set to 0 for cases of no neighbors
     # within range.
     summarise(
@@ -89,21 +92,26 @@ fit_bayesian_model <- function(focal_vs_comp, model_specs, run_shuffle = FALSE,
   if(run_shuffle){
     focal_trees <- focal_trees %>%
       group_by(focal_notion_of_species) %>%
-      mutate(comp_notion_of_species = sample(comp_notion_of_species))
+      mutate(comp_notion_of_species = sample(comp_notion_of_species)) %>%
+      # Revert group_by variables
+      group_by_at(vars(one_of(group_variables)))
   }
 
   # Continue processing focal_trees
-  focal_trees <- focal_trees %>%
-    ungroup() %>%
+  focal_trees_2 <- focal_trees %>%
+    filter(focal_ID %in% c(1:1500)) %>%
     # sum biomass and n_comp for competitors of same species. we need to do this
     # for the cases when we do permutation shuffle.
-    group_by(focal_ID, focal_notion_of_species, dbh, growth, foldID, comp_notion_of_species) %>%
-    summarise_all(list(sum)) %>%
[…]

### 2020-03-18 R/preprocess_data.R

@@ -140,7 +140,11 @@ create_focal_vs_comp <- function(growth_df, max_dist, model_specs, cv_grid, id, 
     # and cyan larger dots are the test set. orange ones separating test set
     # from training set (trees in all other folds)
     if(plot_folds){
-      plot_title <- str_c("fold ", all_folds[i], ": Small black dots = competitor, cyan dots = test set, orange dots = buffer")
+      plot_title <- str_c(
+        "fold ", all_folds[i],
+        ": Small black dots = competitor, cyan dots = test set, orange dots = buffer"
+      )
+
       ggplot() +
         geom_sf(data = bw_boundary, col = "black") +
         geom_sf(data = current_fold_boundary, col = "black") +
@@ -184,7 +188,10 @@ create_focal_vs_comp <- function(growth_df, max_dist, model_specs, cv_grid, id, 
   focal_vs_comp <- focal_vs_comp %>%
     arrange(focal_ID, comp_ID) %>%
     mutate(growth_hat = NA) %>%
-    select(everything(), growth, growth_hat)
+    select(everything(), growth, growth_hat) %>%
+    st_as_sf() %>%
+    group_by(focal_ID, focal_notion_of_species, dbh, growth, foldID,
+             comp_notion_of_species)
 
   return(focal_vs_comp)
 }

### 2020-03-18 R/modeling.R
Revert "working on fit_bayesian_model"
This reverts commit f0113776f76d6185b3d0fe029c93ab64062e6231.

@@ -17,8 +17,7 @@ get_model_specs <- function(forest, model_number, species_notion){
     paste0("growth ~ ", species_notion, " + dbh + dbh * ", species_notion)
 
   model_2_formula <-
-    paste0("growth ~ ", species_notion, " + dbh + dbh * ", species_notion,
-           " + comp_basal_area + comp_basal_area * ", species_notion)
+    paste0("growth ~ ", species_notion, " + dbh + dbh * ", species_notion, " + comp_basal_area + comp_basal_area * ", species_notion)
 
   model_3_formula <- forest %>%
     pull(species_notion) %>%
@@ -71,14 +70,12 @@ get_model_specs <- function(forest, model_number, species_notion){
 fit_bayesian_model <- function(focal_vs_comp, model_specs, run_shuffle = FALSE,
                                prior_hyperparameters = NULL){
 
-  # Save original group_by variables
-  group_variables <- group_vars(focal_vs_comp)
-
   # Get model formula
   model_formula <- model_specs$model_formula
 
-  # Prepare data for regression generate data frame of all focal trees
+  # Prepare data for regression Generate data frame of all focal trees
   focal_trees <- focal_vs_comp %>%
+    group_by(focal_ID, focal_notion_of_species, dbh, growth, foldID, comp_notion_of_species) %>%
     # Sum basal area & count of all neighbors; set to 0 for cases of no neighbors
     # within range.
     summarise(
@@ -92,26 +89,21 @@ fit_bayesian_model <- function(focal_vs_comp, model_specs, run_shuffle = FALSE,
   if(run_shuffle){
     focal_trees <- focal_trees %>%
       group_by(focal_notion_of_species) %>%
-      mutate(comp_notion_of_species = sample(comp_notion_of_species)) %>%
-      # Revert group_by variables
-      group_by_at(vars(one_of(group_variables)))
+      mutate(comp_notion_of_species = sample(comp_notion_of_species))
   }
 
   # Continue processing focal_trees
-  focal_trees_2 <- focal_trees %>%
-    filter(focal_ID %in% c(1:1500)) %>%
+  focal_trees <- focal_trees %>%
+    ungroup() %>%
     # sum biomass and n_comp for competitors of same species. we need to do this
     # for the cases when we do permutation shuffle.
-    group_by(focal_ID, focal_notion_of_species, dbh, growth, foldID,
-             comp_notion_of_species) %>%
[…]

### 2020-03-18 R/preprocess_data.R

@@ -140,11 +140,7 @@ create_focal_vs_comp <- function(growth_df, max_dist, model_specs, cv_grid, id, 
     # and cyan larger dots are the test set. orange ones separating test set
     # from training set (trees in all other folds)
     if(plot_folds){
-      plot_title <- str_c(
-        "fold ", all_folds[i],
-        ": Small black dots = competitor, cyan dots = test set, orange dots = buffer"
-      )
-
+      plot_title <- str_c("fold ", all_folds[i], ": Small black dots = competitor, cyan dots = test set, orange dots = buffer")
       ggplot() +
         geom_sf(data = bw_boundary, col = "black") +
         geom_sf(data = current_fold_boundary, col = "black") +
@@ -188,10 +184,7 @@ create_focal_vs_comp <- function(growth_df, max_dist, model_specs, cv_grid, id, 
   focal_vs_comp <- focal_vs_comp %>%
     arrange(focal_ID, comp_ID) %>%
     mutate(growth_hat = NA) %>%
-    select(everything(), growth, growth_hat) %>%
-    st_as_sf() %>%
-    group_by(focal_ID, focal_notion_of_species, dbh, growth, foldID,
-             comp_notion_of_species)
+    select(everything(), growth, growth_hat)
 
   return(focal_vs_comp)
 }

### 2020-03-18 README.Rmd
Revert "working on fit_bayesian_model"
This reverts commit f0113776f76d6185b3d0fe029c93ab64062e6231.
Done spatial residuals plot

@@
 focal_vs_comp_bw <- bw_growth_df %>% 
-  create_focal_vs_comp(max_dist = max_dist, model_specs = bw_specs, cv_grid = bw_cv_grid, id = "treeID")
+  create_focal_vs_comp(max_dist, model_specs = bw_specs, cv_grid = bw_cv_grid, id = "treeID")
 ```
@@
 ```{r}
-bw_predict <- focal_vs_comp_bw %>% 
-  predict_bayesian_model(model_specs = bw_specs, posterior_param = bw_fit_model)
+bw_growth_df <- focal_vs_comp_bw %>% 
+  predict_bayesian_model(model_specs = bw_specs, posterior_param = bw_fit_model) %>% 
+  right_join(bw_growth_df, by = c("focal_ID" = "treeID"))
 
-# TODO: Convert output of predict_bayesian_model() to be of focal_vs_focal type
-# TODO: Preserve geometry variable, that way we can easily create residual plots
+bw_growth_df <- bw_growth_df %>% 
+  st_as_sf()
 
-bw_predict %>%
-  group_by(focal_ID, focal_notion_of_species) %>%
-  # Since for the same focal tree all growth & growth_hat values are the same:
-  summarise(growth = mean(growth), growth_hat = mean(growth_hat)) %>%
-  ggplot(aes(x = growth, y = growth_hat)) +
-  geom_point(size = 0.5, color = rgb(0,0,0,0.25)) +
+# Observed vs predicted growth  
+ggplot(bw_growth_df, aes(x = growth, y = growth_hat)) +
+  geom_point(size = 0.5, color = rgb(0, 0, 0, 0.25)) +
   stat_smooth(method = 'lm') +
@@
   )
+
+reslab <- expression(paste('Residual (cm ',y^{-1},')'))
+bw_growth_df %>% 
+  # Need to investigate missingness
+  filter(!is.na(growth_hat)) %>% 
+  mutate(
+    error = growth - growth_hat,
+    error_bin = cut_number(error, n = 5), 
+    error_compress = ifelse(error < -0.75, -0.75, ifelse(error > 0.75, 0.75, error))
+  ) %>% 
+  ggplot() + 
+  geom_sf(aes(col = error_compress), size = 0.4) + 
+  theme_bw() + 
+  scale_color_gradient2(
+    low = "#ef8a62", mid = "#f7f7f7", high = "#67a9cf", 
+    name = reslab,
+    breaks = seq(from = -0.75, to = 0.75, by = 0.25),
+    labels = c('< -0.75', '-0.5', '0.25', '0', '0.25', '0.5', '> 0.75')) +
+  labs(x = "Meter", y = "Meter")
 ```

### 2020-03-18 R/modeling.R

@@ -251,11 +251,6 @@ predict_bayesian_model <- function(focal_vs_comp, model_specs, posterior_param){
     mutate(growth_hat = as.vector(X %*% mu_star)) %>%
     select(focal_ID, growth_hat)
 
-  focal_vs_comp <- focal_vs_comp %>%
-    select(-growth_hat) %>%
-    left_join(focal_trees, by = "focal_ID") %>%
-    select(focal_ID, focal_notion_of_species, dbh, growth, growth_hat, foldID, everything())
-
   # why do we return focal_vs_comp?? Shouldn't we return focal_trees (one row per focal tree not per interaction)
-  return(focal_vs_comp)
+  return(focal_trees)
 }

### 2020-03-18 R/preprocess_data.R

@@ -184,7 +184,16 @@ create_focal_vs_comp <- function(growth_df, max_dist, model_specs, cv_grid, id, 
   focal_vs_comp <- focal_vs_comp %>%
     arrange(focal_ID, comp_ID) %>%
     mutate(growth_hat = NA) %>%
-    select(everything(), growth, growth_hat)
+    select(
+      # Relating to focal tree:
+      focal_ID, focal_notion_of_species, dbh, foldID, geometry, growth,
+      # Relating to competitor tree:
+      comp_ID, dist, comp_notion_of_species, comp_basal_area
+    )
+    # Should we do grouping here?
+    # group_by(focal_ID, focal_notion_of_species, dbh, foldID, geometry, growth) %>%
+    # should we convert to sf object here?
+    # st_as_sf()
 
   return(focal_vs_comp)
 }

### 2020-03-19 README.Rmd
Done SCBI fit/predict model (sort of)

@@
 
+**Big Woods**:
+
 ```{r}
@@
 bw_specs
-
-scbi_specs <- scbi_growth_df %>% 
-  get_model_specs(model_number = 3, species_notion = 'sp')
 ```
@@
 
+**SCBI**:
+
+```{r}
+scbi_specs <- scbi_growth_df %>% 
+  get_model_specs(model_number = 3, species_notion = 'sp')
+scbi_specs
+```
+
+```{r}
+focal_vs_comp_scbi <- scbi_growth_df %>% 
+  create_focal_vs_comp(max_dist, model_specs = scbi_specs, cv_grid = scbi_cv_grid, id = "stemID")
+```
+
+```{r}
+glimpse(focal_vs_comp_scbi)
+```
+
 ### Model fit and prediction
+
+**Big Woods**:
 
@@
 
+**SCBI**:
+
+```{r}
+scbi_fit_model <- focal_vs_comp_scbi %>% 
+  fit_bayesian_model(model_specs = scbi_specs)
+```
+
+```{r}
+scbi_growth_df <- focal_vs_comp_scbi %>% 
+  predict_bayesian_model(model_specs = scbi_specs, posterior_param = scbi_fit_model) %>% 
+  right_join(scbi_growth_df, by = c("focal_ID" = "stemID"))
+
+scbi_growth_df <- scbi_growth_df %>% 
+  st_as_sf()
+
+# Observed vs predicted growth  
+ggplot(scbi_growth_df, aes(x = growth, y = growth_hat)) +
+  geom_point(size = 0.5, color = rgb(0, 0, 0, 0.25)) +
+  stat_smooth(method = 'lm') +
+  geom_abline(slope = 1, intercept = 0) +
+  coord_fixed() + 
+  labs(
+    x = "Observed growth in dbh", y = "Predicted growth in dbh", 
+    title = "Predicted vs Observed Growth"
+  )
+
+reslab <- expression(paste('Residual (cm ',y^{-1},')'))
+scbi_growth_df %>% 
+  # Need to investigate missingness
+  filter(!is.na(growth_hat)) %>% 
+  mutate(
+    error = growth - growth_hat,
+    error_bin = cut_number(error, n = 5), 
+    error_compress = ifelse(error < -0.75, -0.75, ifelse(error > 0.75, 0.75, error))
+  ) %>% 
+  ggplot() + 
+  geom_sf(aes(col = error_compress), size = 0.4) + 
+  theme_bw() + 
+  scale_color_gradient2(
+    low = "#ef8a62", mid = "#f7f7f7", high = "#67a9cf", 
+    name = reslab,
+    breaks = seq(from = -0.75, to = 0.75, by = 0.25),
+    labels = c('< -0.75', '-0.5', '0.25', '0', '0.25', '0.5', '> 0.75')) +
+  labs(x = "Meter", y = "Meter")
+```
+
 ### Run spatial cross-validation
@@
 * CV: `run_cv()` is a wrapper to `fit_bayesian_model()` and `predict_bayesian_model()`.
+
+```{r}
+
+```
 

### 2020-03-20 README.Rmd
trying to get run_cv to work

@@
 # ID which points are in buffer and which are not
-buffer_index <- st_intersects(scbi_growth_df, scbi_buffer, sparse = FALSE)
+buffer_index <- !st_intersects(scbi_growth_df, scbi_buffer, sparse = FALSE)
 scbi_growth_df <- scbi_growth_df %>% 
@@
 
-* CV: `run_cv()` is a wrapper to `fit_bayesian_model()` and `predict_bayesian_model()`.
+For the above results we fit the model to the entire data set, and then make predictions across the entire data set from that fit.
+This could lead to overfitting because we are using the training data to also test the model.
+If model error is spatially correlated this could be a large issue (cite important sources here!).
+We can use the spatial block structure we defined above to deal with with.
+The function `run_cv` goes through each fold in the `cv_grid` and fits the model on all the other folds.
+Then applies that fit to the focal fold.
+It is a wrapper for `fit_bayesain_model` and `predict_bayesain_model` but fits a seperate model for each fold.
 
 ```{r}
+scbi_cv_predict <- focal_vs_comp_scbi %>%
+  run_cv(model_specs = scbi_specs, max_dist = max_dist, cv_grid = scbi_cv_grid)
 

### 2020-03-20 R/modeling.R

@@ -254,3 +254,45 @@ predict_bayesian_model <- function(focal_vs_comp, model_specs, posterior_param){
   # why do we return focal_vs_comp?? Shouldn't we return focal_trees (one row per focal tree not per interaction)
   return(focal_trees)
 }
+
+
+#' Run the bayesain model with spatial cross validation
+#'
+#' @inheritParams fit_bayesian_model
+#' @param max_dist
+#' @param cv_grid
+#'
+#' @import dplyr
+#' @import sf
+#' @importFrom stats model.matrix
+#' @importFrom tidyr nest
+#' @return \code{focal_vs_comp} with new column of predicted \code{growth_hat}
+#' @export
+#'
+#' @examples
+#' 1+1
+run_cv <- function(focal_vs_comp, model_specs, max_dist, cv_grid,
+                        run_shuffle = FALSE, prior_hyperparameters = NULL){
+
+  folds <- focal_vs_comp %>% pull(foldID) %>% unique()
+  focal_trees <- tibble(focal_ID = NA, growth_hat  = NA)
+
+  for (i in folds)
+  {
+    ### Need to deal with the buffer issue!!! Buffer out max_dist
+    ### It looks like our new cv_grid is a spatialBlock object (from blockCV)
+    ## not sure how to pull the coordiants out of that fold
+
+    train <- focal_vs_comp %>% filter(foldID != i)
+    test <- focal_vs_comp %>% filter(foldID == i)
+
+    # test <- test %>% filter out buffer
+
+    fold_fit <- train %>% fit_bayesian_model(model_specs)
+    fold_predict <- test %>% predict_bayesian_model(model_specs, fold_fit)
+    focal_trees <- focal_trees %>% rbind(fold_predict) %>% filter(!is.na(focal_ID))
+
+  }
+
+  return(focal_trees)
+}

### 2020-03-20 R/modeling.R
run_cv changes and README changes

@@ -259,13 +259,11 @@ predict_bayesian_model <- function(focal_vs_comp, model_specs, posterior_param){
 #' Run the bayesain model with spatial cross validation
 #'
 #' @inheritParams fit_bayesian_model
-#' @param max_dist
-#' @param cv_grid
+#' @param max_dist distance of competitive neighborhood
+#' @param cv_grid length of the cross validation grid
 #'
 #' @import dplyr
 #' @import sf
-#' @importFrom stats model.matrix
-#' @importFrom tidyr nest
 #' @return \code{focal_vs_comp} with new column of predicted \code{growth_hat}
 #' @export
 #'

### 2020-03-21 README.Rmd
run_cv changes and README changes
sill pugging away at `run_cv`

@@
 library(blockCV)
+library(tictoc)
 ```
@@
 
+Here I cheated `run_cv` to run the cross valididation for just two training set. It runs on the whole data set up just for `test = fold 23` and `test = fold 2`. Just to see how long it takes.
 ```{r}
+tic()
+
 scbi_cv_predict <- focal_vs_comp_scbi %>%
   run_cv(model_specs = scbi_specs, max_dist = max_dist, cv_grid = scbi_cv_grid)
+toc()
+```
+
+Running just two folds took 5440 seconds. There are 28 folds. So running everything should take 5440 * 14 / 60 / 60 = 21 hours.
+
+Comapre results
+
+```{r}
+
+scbi_cv_predict %>%
+  inner_join(scbi_growth_df, by = 'focal_ID', suffix = c('_cv','')) %>%
+  summarise(rmse_cv = sqrt(mean((growth - growth_hat_cv)^2)),
+            rmse = sqrt(mean((growth - growth_hat)^2)) )
 

### 2020-03-23 README.Rmd
run_cv with test fold buffer

@@
   summarise(rmse_cv = sqrt(mean((growth - growth_hat_cv)^2)),
-            rmse = sqrt(mean((growth - growth_hat)^2)) )
+            rmse = sqrt(mean((growth - growth_hat)^2)), n() )
 

### 2020-03-23 just_bw_code_blocks.R

@@ -1,125 +1,0 @@
-library(tidyverse)
-library(forestecology)
-library(snakecase)
-library(skimr)
-library(sf)
-library(sfheaders)
-library(blockCV)
-
-bw_2008 <-
-  "https://deepblue.lib.umich.edu/data/downloads/z603qx485" %>%
-  read_delim(delim = "\t") %>%
-  mutate(spcode = to_any_case(spcode)) %>%
-  select(
-    treeID = treeid, stemID = stemtag, sp = spcode, quadrat, gx, gy, dbh,
-    date, codes
-  )
-bw_2014 <-
-  "https://deepblue.lib.umich.edu/data/downloads/1831ck00f" %>%
-  read_delim(delim = "\t") %>%
-  mutate(spcode = to_any_case(spcode)) %>%
-  select(
-    treeID = treeid, stemID = stemtag, sp = spcode, quadrat, gx, gy, dbh,
-    date, codes
-  )
-
-# Read in bigwoods grouping classification data
-bw_species <-
-  "https://deepblue.lib.umich.edu/data/downloads/000000086" %>%
-  read_delim(delim = "\t") %>%
-  # convert all to snake case:
-  mutate_at(c("species", "genus", "family", "idlevel", "spcode"), to_any_case) %>%
-  # join trait group
-  left_join(families, by = c("spcode", "family")) %>%
-  mutate(
-    sp = str_sub(genus, 1, 2),
-    sp = str_c(sp, str_sub(species, 1, 2)),
-    sp = tolower(sp),
-    latin = str_c(genus, species, sep = " "),
-    latin = to_any_case(latin)
-  ) %>%
-  select(sp = spcode, genus, species, latin, family, trait_group)
-
-bw_2008 <- bw_2008 %>%
-  left_join(bw_species,by='sp')
-
-
-census_df1 <- bw_2008
-# we need to filter out the resprouts
-census_df2 <- bw_2014 %>%
-  filter(!str_detect(codes, 'R'))
-id <- "treeID"
-
-bw_growth_df <-
-  # Merge both censuses and compute growth:
-  compute_growth(census_df1, census_df2, id) %>%
-  mutate(sp = to_any_case(sp))
-
-cv_fold_size <- 100
-max_dist <- 7.5
-
-# Bigwoods study region boundary polygon
-bw_boundary <- bigwoods_study_region %>%
-  sf_polygon()
-
[…]

### 2020-03-23 README.Rmd
run_cv working and example for SCBI and BW

@@
 
+**SCBI**
+
 For the above results we fit the model to the entire data set, and then make predictions across the entire data set from that fit.
@@
 
-Here I cheated `run_cv` to run the cross valididation for just two training set. It runs on the whole data set up just for `test = fold 23` and `test = fold 2`. Just to see how long it takes.
+This will fit the model for each fold.
+On each fold it fits the data for all trees outside of that fold.
+If you have N folds then `run_cv` will take N-times longer than `fit_bayesain_model`.
+Here I cheated `run_cv` to fit for just `test = fold 23` and `test = fold 2`.
+It still fits to all other folds for those two, but only does it twice (rather than 28 times for the SCBI data set).
+Do this with the arguement `all_folds = FALSE`.
+
 ```{r}
 tic()
-
 scbi_cv_predict <- focal_vs_comp_scbi %>%
-  run_cv(model_specs = scbi_specs, max_dist = max_dist, cv_grid = scbi_cv_grid)
+  run_cv(model_specs = scbi_specs, max_dist = max_dist, cv_grid = scbi_cv_grid, all_folds = FALSE)
 toc()
@@
 
-Running just two folds took 5440 seconds. There are 28 folds. So running everything should take 5440 * 14 / 60 / 60 = 21 hours.
+Running just two folds took 2119 seconds. There are 28 folds. So running everything should take 2119 * 14 / 60 / 60 = 8 hours.
 
-Comapre results
+Then we can compare the results to show that the RMSE for the cross-validated fit is larger than for the none CV fit above.
 
@@
   inner_join(scbi_growth_df, by = 'focal_ID', suffix = c('_cv','')) %>%
+  summarise(rmse_cv = sqrt(mean((growth - growth_hat_cv)^2)),
+            rmse = sqrt(mean((growth - growth_hat)^2)), n() )
+
+```
+
+**Big Woods**
+
+Big woods is faster because it has fewer trees, but also because we are fitting the model for trait groups (6 groups) rather than species for SCBI (40 species).
+The lambda matrix is much smaller (6 x 6 versus 40 x 40).
+Means it fits much faster.
+
+```{r}
+tic()
+bw_cv_predict <- focal_vs_comp_bw %>%
+  run_cv(model_specs = bw_specs, max_dist = max_dist, cv_grid = bw_cv_grid)
+toc()
+```
+
+Big Woods must faster because it is a smaller plot? Mabye also because we are fitting the trait group version (6 spp versus 40 for SCBI).
+
+```{r}
+
+bw_cv_predict %>%
+  inner_join(bw_growth_df, by = 'focal_ID', suffix = c('_cv','')) %>%
   summarise(rmse_cv = sqrt(mean((growth - growth_hat_cv)^2)),

### 2020-03-23 R/modeling.R

@@ -271,11 +271,17 @@ predict_bayesian_model <- function(focal_vs_comp, model_specs, posterior_param){
 #' @examples
 #' 1+1
 run_cv <- function(focal_vs_comp, model_specs, max_dist, cv_grid,
-                        run_shuffle = FALSE, prior_hyperparameters = NULL){
+                        run_shuffle = FALSE, prior_hyperparameters = NULL,
+                        all_folds = TRUE){
 
-  # cheat for a second to see how long it takes to do two folds
-  folds <- c(23, 2)
-  # folds <- focal_vs_comp %>% pull(foldID) %>% unique()
+  # if subset is true just two folds
+  if (all_folds)
+  {
+    folds <- focal_vs_comp %>% pull(foldID) %>% unique()
+  } else {
+    folds <- c(23, 2)
+  }
+
   focal_trees <- tibble(focal_ID = NA, growth_hat  = NA)
 
   for (i in folds)

### 2020-05-25 README.Rmd
trivial change

--- 2020-03-21 README.Rmd
@@
   fig.path = "man/figures/README-",
-  out.width = "100%",
-  fig.width = 16/2,
-  fig.height = 9/2,
-  message = FALSE,
-  warning = FALSE,
-  cache = TRUE
+  out.width = "100%"
 )
 ```
-
 # forestecology
@@
 [![Travis Build Status](https://travis-ci.org/rudeboybert/forestecology.svg?branch=master)](https://travis-ci.org/rudeboybert/forestecology)
-[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
 [![CRAN status](https://www.r-pkg.org/badges/version/forestecology)](https://cran.r-project.org/package=forestecology)
@@
 
-You can install the released version of forestecology from [CRAN](https://CRAN.R-project.org) with:
+You can install the released version of `forestecology` from [CRAN](https://CRAN.R-project.org) with:
 
-```{r, eval = FALSE}
+``` r
 install.packages("forestecology")
@@
 
-```{r, eval = FALSE}
-# install.packages("remotes")
-# remotes::install_github("rudeboybert/forestecology")
+``` r
+# install.packages("devtools")
+devtools::install_github("rudeboybert/forestecology")
+```
+## Example
+
+This is a basic example which shows you how to solve a common problem:
+
+```{r example}
+## basic example code
 ```
 
-## Example analysis
+What is special about using `README.Rmd` instead of just `README.md`? You can include R chunks like so:
 
-We present an example analysis using:
-
-* [Michigan Big Woods](https://doi.org/10.7302/wx55-kt18) research plot data
-* [Smithsonian Conservation Biology Institute (SCBI) ForestGEO](https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/tree/master/tree_main_census) research plot data
-
-```{r}
-library(tidyverse)
-library(forestecology)
-library(snakecase)
-library(skimr)
-library(sf)
-library(sfheaders)
-# devtools::install_github("rvalavi/blockCV")
-library(blockCV)
-library(tictoc)
+```{r cars}
+summary(cars)
 ```
 
-### Load & preprocess data
+You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date.
 
-We first load and preprocess the data corresponding to the two census reflecting the time period over which we consider growth in dbh.
-We load them into R as "tibble" data frames thereby ensuring a standardized input/output format that can be used across all `tidyverse` packages @tidyverse.
-Furthermore, we ensure that the different variables have the correct names, types (`dbl`, `data`, `factor`).
+You can also embed plots, for example:
 
-**Big Woods**:
-
-```{r}
-# Read in Big Woods census data from 2008 & 2014
-bw_2008 <- 
-  "https://deepblue.lib.umich.edu/data/downloads/z603qx485" %>% 
-  read_delim(delim = "\t") %>% 
-  mutate(spcode = to_any_case(spcode)) %>%
-  select(
-    treeID = treeid, stemID = stemtag, sp = spcode, quadrat, gx, gy, dbh, 
-    date, codes
-  )
-bw_2014 <- 
-  "https://deepblue.lib.umich.edu/data/downloads/1831ck00f" %>% 
-  read_delim(delim = "\t") %>% 
-  mutate(spcode = to_any_case(spcode)) %>%
-  select(
-    treeID = treeid, stemID = stemtag, sp = spcode, quadrat, gx, gy, dbh, 
-    date, codes
-  )
-
-# Read in bigwoods grouping classification data
-bw_species <- 
-  "https://deepblue.lib.umich.edu/data/downloads/000000086" %>% 
-  read_delim(delim = "\t") %>% 
-  # convert all to snake case:
-  mutate_at(c("species", "genus", "family", "idlevel", "spcode"), to_any_case) %>% 
-  # join trait group
-  left_join(families, by = c("spcode", "family")) %>% 
-  mutate(
-    sp = str_sub(genus, 1, 2), 
-    sp = str_c(sp, str_sub(species, 1, 2)),
-    sp = tolower(sp),
-    latin = str_c(genus, species, sep = " "),
-    latin = to_any_case(latin)
-  ) %>% 
-  select(sp = spcode, genus, species, latin, family, trait_group)
-
-bw_2008 <- bw_2008 %>%
-  left_join(bw_species,by='sp')
-
+```{r pressure, echo = FALSE}
+plot(pressure)
 ```
 
-**SCBI**:
-
-```{r}
-scbi_2013 <- 
-  "https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/raw/master/tree_main_census/data/census-csv-files/scbi.stem2.csv" %>% 
-  read_csv() %>% 
-  select(
-    treeID, stemID, sp, quadrat, gx, gy, dbh, 
-    date = ExactDate, codes, status
-  )
-
-scbi_2018 <-
-  "https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/raw/master/tree_main_census/data/census-csv-files/scbi.stem3.csv" %>% 
-  read_csv() %>% 
-  select(
-    treeID, stemID, sp, quadrat, gx, gy, dbh, 
-    date = ExactDate, codes, status
-  ) %>% 
-  mutate(dbh = as.numeric(dbh))
-```
-
-### Compute growth
-
-We then combine the two census data frames into a single data frame  that now includes a numerical variable `growth` reflecting the average annual growth of dbh in cm.
-Furthermore, variables that (in theory) remain unchanged between censuses appear only once, such as `gx`, `gy`, species-related variables.
-Variables that should change between censuses are tagged with `1/2` indicating earlier/later, such as `dbh1/dbh2`, `codes1/codes2`.
-
-The resulting data frames are named with some variation of `growth_df`.
-
-**Big Woods**:
-
-```{r}
-census_df1 <- bw_2008
-# we need to filter out the resprouts
-census_df2 <- bw_2014 %>% 
-  filter(!str_detect(codes, 'R'))
-id <- "treeID"
-
-bw_growth_df <- 
-  # Merge both censuses and compute growth:
-  compute_growth(census_df1, census_df2, id) %>% 
-  mutate(sp = to_any_case(sp))
-```
-
-**SCBI**:
-
-```{r}
-census_df1 <- scbi_2013
-census_df2 <- scbi_2018
-id <- "stemID"
-
-scbi_growth_df <- 
-  # Merge both censuses and compute growth:
-  compute_growth(census_df1, census_df2, id) %>%
-  # they are mesuaring in mm while we are measuring in cm!!!
-  mutate(growth = growth/10)
-```
-
-```{r}
-growth_df <- bind_rows(
-  bw_growth_df %>% select(growth) %>% mutate(site = "bw"),
-  scbi_growth_df %>% select(growth) %>% mutate(site = "scbi")
-)
-ggplot(growth_df, aes(x = growth, y = ..density.., fill = site)) +
-  geom_histogram(alpha = 0.5, position = "identity", binwidth = 0.05) +
-  labs(x = "Average annual growth in dbh (cm per yr)") +
-  coord_cartesian(xlim = c(-0.5, 1))
-```
-
-### Add spatial information
-
-We now add spatial information to the `growth_df` data frame. Specifically:
-
-1.
-In order to control for study region "edge effects" (cite Waller?), we need a function that adds "buffers" of trees.
-In our case, since our model of interspecies competition relies on a spatial definition of who is a "neighboring competitor" and certain explanatory varibles such as biomass are cummulative, we need to ensure that all trees being modeled are not biased to have different neighbor structure, in particular the trees at the boundary of study regions.
-1.
-Assign each tree to a "fold" for cross-validation purposes.
-Conventional cross-validation schemes assign observations to folds by resampling individual observations at random.
-However, underlying this scheme is an assumption that the observations are independent.
-In the case of forest census data, observations exhibit spatial autocorrelation, and thus this dependence must be incorporated in our resampling scheme @roberts2017 @pohjankukka2017.
-Packages that have implemented spatial cross-validation include @valavi2019
-
-Thus, before we can incorporate the above information to `growth_df`, we need to define two constants:
-
-```{r}
-cv_fold_size <- 100
-max_dist <- 7.5
-```
-
-#### Defining buffers
-
-For a focal tree of interest, the `max_dist` of `r max_dist` meters acts as a radius defining a neighborhood within which all trees are considered competitors.
-In other words, all trees within `r max_dist` meters of the focal tree are considered competitor trees.
-Other studies have estimated this distance; we used `r max_dist` meters as an average of estimated values [@canham2004, @canham2006, @uriarte2004, @tatsumi2013].
-Following Tobler's first law of geography that "everything is related to everything else, but near things are more related than distant things." @tobler1970firstlawofgeo, we assume that the degree of spatial autocorrelation is inversely-related to distance.
-However, we further assume that once trees are more than `r max_dist` meters apart, this autocorrelation is negligeable.
-
-**Example of buffer at work**:
-
-```{r}
-# Boundary polygon
-square_boundary <- tibble(
-  x = c(0,0,1,1),
-  y = c(0,1,1,0)
-) %>% 
-  sf_polygon()
-
-# Buffer polygon
-square_buffer <- square_boundary %>% 
-  st_buffer(dist = -0.1)
-
-ggplot() +
-  geom_sf(data = square_boundary) +
-  geom_sf(data = square_buffer, col="red") 
-```
-
-**Big Woods**:
-
-```{r}
-# Bigwoods study region boundary polygon
-bw_boundary <- bigwoods_study_region %>% 
-  sf_polygon()
-
-# Buffer polygon
-bw_buffer <- bw_boundary %>%
-  st_buffer(dist = -max_dist)
-
-# Convert data frame to sf object
-bw_growth_df <- bw_growth_df %>% 
-  st_as_sf(coords = c("gx", "gy"))
-
-# ID which points are in buffer and which are not
-buffer_index <- !st_intersects(bw_growth_df, bw_buffer, sparse = FALSE)
-bw_growth_df <- bw_growth_df %>% 
-  mutate(buffer = as.vector(buffer_index))
-
-# Plot
-ggplot() +  
-  geom_sf(data = bw_boundary) +
-  geom_sf(data = bw_buffer, col="red") +
-  # Only random sample of 1000 trees:
-  geom_sf(data = bw_growth_df %>% sample_n(1000), aes(col=buffer), size = 0.5)
-```
-
-**SCBI**:
-
-```{r}
-# Bigwoods study region boundary polygon
-scbi_boundary <- tibble(x = c(0,400,400,0,0), y = c(0,0,640,640,0)) %>% 
-  sf_polygon()
-
-# Buffer polygon
-scbi_buffer <- scbi_boundary %>%
-  st_buffer(dist = -max_dist)
-
-# Convert data frame to sf object
-scbi_growth_df <- scbi_growth_df %>% 
-  st_as_sf(coords = c("gx", "gy"))
-
-# ID which points are in buffer and which are not
-buffer_index <- !st_intersects(scbi_growth_df, scbi_buffer, sparse = FALSE)
-scbi_growth_df <- scbi_growth_df %>% 
-  mutate(buffer = as.vector(buffer_index))
-
-# Plot
-ggplot() +
-  geom_sf(data = scbi_boundary) +
-  geom_sf(data = scbi_buffer, col="red") +
-  # Only random sample of 1000 trees:
-  geom_sf(data = scbi_growth_df %>% sample_n(1000), aes(col=buffer), size = 0.5)
-```
-
-#### Defining spatial cross-validation folds
-
-We use the [`blockCV`](https://github.com/rvalavi/blockCV) package to define the spatial grid, whose elements will act as the folds in our leave-one-out (by "one" we meen "one grid block") cross-validation scheme.
-
-**Big Woods**:
-
-```{r}
-set.seed(76)
-bw_cv_grid <- spatialBlock(
-  speciesData = bw_growth_df, theRange = 100, verbose = FALSE,
-  # Some guess work in figuring this out:
-  k = 28, xOffset = 0.5, yOffset = 0
-)
-
-# Add foldID to data
-bw_growth_df <- bw_growth_df %>% 
-  mutate(
-    foldID = bw_cv_grid$foldID
-  )
-
-# Visualize grid. Why does fold 19 repeat?
-bw_cv_grid$plots +
-  geom_sf(data = bw_growth_df %>% sample_frac(0.2), aes(col=factor(foldID)), size = 0.1)
-
-# Remove weird folds with no trees in them from viz above
-bw_growth_df <- bw_growth_df %>%
-  filter(!foldID %in% c(19, 23, 21, 17, 8, 19))
-```
-
-**SCBI**:
-
-```{r}
-scbi_cv_grid <- spatialBlock(
-  speciesData = scbi_growth_df, theRange = 100, k = 28, yOffset = 0.9999, verbose = FALSE
-)
-
-# Add foldID to data
-scbi_growth_df <- scbi_growth_df %>% 
-  mutate(
-    foldID = scbi_cv_grid$foldID
-  )
-
-# Visualize grid
-scbi_cv_grid$plots +
-  geom_sf(data = scbi_growth_df, aes(col=factor(foldID)), size = 0.1)
-```
-
-### Model specification
-
-Next we specify the growth model we want.
-We will model growth as a function of tree identity, size (in dbh), and the identity and size (in dbh) of its neighbors.
-The model can be run on different grouping of individuals (e.g., based on species or trait groupings).
-It can also be run with different notions of competition:
-
-* `model_number = 1`: No competition growth only depends on focal tree `dbh` and grouping.
-* `model_number = 2`: Competition but identity of competitor does not matter, just sum of competitior basal area.
-* `model_number = 3`: Full model with competition including competitor identity.
-
-**Big Woods**:
-
-```{r}
-bw_specs <- bw_growth_df %>% 
-  get_model_specs(model_number = 3, species_notion = 'trait_group')
-bw_specs
-```
-
-Next we create the `focal_vs_comp` data frame which connects each tree in `growth_df` to the trees in its competitive neighborhood range (as defined by `max_dist`).
-So for example, if `growth_df` consisted of two focal trees with two and three neighbors respectively, `focal_vs_comp` would consist of 5 rows.
-
-This requires the `growth_df` data frame, `max_dist` scalar defining competitive range, `cv_fold_size` defining the size of the spatial cross-validation blocks, and `model_specs` as inputs.
-
-```{r}
-focal_vs_comp_bw <- bw_growth_df %>% 
-  create_focal_vs_comp(max_dist, model_specs = bw_specs, cv_grid = bw_cv_grid, id = "treeID")
-```
-
-```{r}
-glimpse(focal_vs_comp_bw)
-```
-
-**SCBI**:
-
-```{r}
-scbi_specs <- scbi_growth_df %>% 
-  get_model_specs(model_number = 3, species_notion = 'sp')
-scbi_specs
-```
-
-```{r}
-focal_vs_comp_scbi <- scbi_growth_df %>% 
-  create_focal_vs_comp(max_dist, model_specs = scbi_specs, cv_grid = scbi_cv_grid, id = "stemID")
-```
-
-```{r}
-glimpse(focal_vs_comp_scbi)
-```
-
-### Model fit and prediction
-
-**Big Woods**:
-
-Now we are ready to fit the competition model with `fit_bayesian_model` this takes two inputs: `focal_vs_comp` the data frame with focal trees connected to their competitors and `model_specs` which specifies the notion of competition.
-
-```{r}
-bw_fit_model <- focal_vs_comp_bw %>% 
-  fit_bayesian_model(model_specs = bw_specs)
-```
-
-This output has the posterior parameters for the specified competition model.
-If `model_number = 1` this includes just the beta values for how focal tree DBH and identity affect growth.
-If `model_number = 2` it also includes a vector of lambdas for how total competitor basal area affect individuals of each group.
-If `model_number = 3` then lambda is a matrix which gives how individuals of each focal group respond to the competition of each competitor group.
-This `posterior_param` output can be used to get predicted growths for each individual (with `predict_bayesain_model`) to test how well the model performs.
-Or this `posterior_param` output can be plots (either the betas or lambdas) to understand what controls individual growth.
-
-```{r}
-bw_growth_df <- focal_vs_comp_bw %>% 
-  predict_bayesian_model(model_specs = bw_specs, posterior_param = bw_fit_model) %>% 
-  right_join(bw_growth_df, by = c("focal_ID" = "treeID"))
-
-bw_growth_df <- bw_growth_df %>% 
-  st_as_sf()
-
-# Observed vs predicted growth  
-ggplot(bw_growth_df, aes(x = growth, y = growth_hat)) +
-  geom_point(size = 0.5, color = rgb(0, 0, 0, 0.25)) +
-  stat_smooth(method = 'lm') +
-  geom_abline(slope = 1, intercept = 0) +
-  coord_fixed() + 
-  labs(
-    x = "Observed growth in dbh", y = "Predicted growth in dbh", 
-    title = "Predicted vs Observed Growth"
-  )
-
-reslab <- expression(paste('Residual (cm ',y^{-1},')'))
-bw_growth_df %>% 
-  # Need to investigate missingness
-  filter(!is.na(growth_hat)) %>% 
-  mutate(
-    error = growth - growth_hat,
-    error_bin = cut_number(error, n = 5), 
-    error_compress = ifelse(error < -0.75, -0.75, ifelse(error > 0.75, 0.75, error))
-  ) %>% 
-  ggplot() + 
-  geom_sf(aes(col = error_compress), size = 0.4) + 
-  theme_bw() + 
-  scale_color_gradient2(
-    low = "#ef8a62", mid = "#f7f7f7", high = "#67a9cf", 
-    name = reslab,
-    breaks = seq(from = -0.75, to = 0.75, by = 0.25),
-    labels = c('< -0.75', '-0.5', '0.25', '0', '0.25', '0.5', '> 0.75')) +
-  labs(x = "Meter", y = "Meter")
-```
-
-**SCBI**:
-
-```{r}
-scbi_fit_model <- focal_vs_comp_scbi %>% 
-  fit_bayesian_model(model_specs = scbi_specs)
-```
-
-```{r}
-scbi_growth_df <- focal_vs_comp_scbi %>% 
-  predict_bayesian_model(model_specs = scbi_specs, posterior_param = scbi_fit_model) %>% 
-  right_join(scbi_growth_df, by = c("focal_ID" = "stemID"))
-
-scbi_growth_df <- scbi_growth_df %>% 
-  st_as_sf()
-
-# Observed vs predicted growth  
-ggplot(scbi_growth_df, aes(x = growth, y = growth_hat)) +
-  geom_point(size = 0.5, color = rgb(0, 0, 0, 0.25)) +
-  stat_smooth(method = 'lm') +
-  geom_abline(slope = 1, intercept = 0) +
-  coord_fixed() + 
-  labs(
-    x = "Observed growth in dbh", y = "Predicted growth in dbh", 
-    title = "Predicted vs Observed Growth"
-  )
-
-reslab <- expression(paste('Residual (cm ',y^{-1},')'))
-scbi_growth_df %>% 
-  # Need to investigate missingness
-  filter(!is.na(growth_hat)) %>% 
-  mutate(
-    error = growth - growth_hat,
-    error_bin = cut_number(error, n = 5), 
-    error_compress = ifelse(error < -0.75, -0.75, ifelse(error > 0.75, 0.75, error))
-  ) %>% 
-  ggplot() + 
-  geom_sf(aes(col = error_compress), size = 0.4) + 
-  theme_bw() + 
-  scale_color_gradient2(
-    low = "#ef8a62", mid = "#f7f7f7", high = "#67a9cf", 
-    name = reslab,
-    breaks = seq(from = -0.75, to = 0.75, by = 0.25),
-    labels = c('< -0.75', '-0.5', '0.25', '0', '0.25', '0.5', '> 0.75')) +
-  labs(x = "Meter", y = "Meter")
-```
-
-### Run spatial cross-validation
-
-For the above results we fit the model to the entire data set, and then make predictions across the entire data set from that fit.
-This could lead to overfitting because we are using the training data to also test the model.
-If model error is spatially correlated this could be a large issue (cite important sources here!).
-We can use the spatial block structure we defined above to deal with with.
-The function `run_cv` goes through each fold in the `cv_grid` and fits the model on all the other folds.
-Then applies that fit to the focal fold.
-It is a wrapper for `fit_bayesain_model` and `predict_bayesain_model` but fits a seperate model for each fold.
-
-Here I cheated `run_cv` to run the cross valididation for just two training set. It runs on the whole data set up just for `test = fold 23` and `test = fold 2`. Just to see how long it takes.
-```{r}
-tic()
-
-scbi_cv_predict <- focal_vs_comp_scbi %>%
-  run_cv(model_specs = scbi_specs, max_dist = max_dist, cv_grid = scbi_cv_grid)
-toc()
-```
-
-Running just two folds took 5440 seconds. There are 28 folds. So running everything should take 5440 * 14 / 60 / 60 = 21 hours.
-
-Comapre results
-
-```{r}
-
-scbi_cv_predict %>%
-  inner_join(scbi_growth_df, by = 'focal_ID', suffix = c('_cv','')) %>%
-  summarise(rmse_cv = sqrt(mean((growth - growth_hat_cv)^2)),
-            rmse = sqrt(mean((growth - growth_hat)^2)) )
-
-```
-
-### Run permutations
-
-* Permutation test:
-  + Add line to that shuffles competitor species within `focal_and_comp` just before fitting model
-  + permutation AND CV: add `run_shuffle` argument to `run_cv()`
-
-### Visualize posterior distributions
-
-* Plot posterior distributions of all parameters
+In that case, don't forget to commit and push the resulting figure files, so they display on GitHub!

### 2020-05-28 README.Rmd
another change
Reran README

--- 2020-03-23 README.Rmd
@@
 library(tictoc)
+
+run_scbi <- FALSE
+run_bw <- TRUE
 ```
@@
 
-```{r}
+```{r, eval = run_bw}
 # Read in Big Woods census data from 2008 & 2014
@@
 
-```{r}
+```{r, eval = run_scbi}
 scbi_2013 <- 
@@
 
-```{r}
+```{r, eval = run_bw}
 census_df1 <- bw_2008
@@
 
-```{r}
+```{r, eval = run_scbi}
 census_df1 <- scbi_2013
@@
 
-```{r}
+**Comparison**:
+
+```{r, eval = run_scbi & run_bw}
 growth_df <- bind_rows(
@@
 
-```{r}
+```{r, eval = run_bw}
 # Bigwoods study region boundary polygon
@@
 
-```{r}
+```{r, eval = run_scbi}
 # Bigwoods study region boundary polygon
@@
 
-```{r}
+```{r, eval = run_bw}
 set.seed(76)
@@
 
-```{r}
+```{r, eval = run_scbi}
 scbi_cv_grid <- spatialBlock(
@@
 
-```{r}
+```{r, eval = run_bw}
 bw_specs <- bw_growth_df %>% 
@@
 
-```{r}
+```{r, eval = run_bw}
+tic()
 focal_vs_comp_bw <- bw_growth_df %>% 
   create_focal_vs_comp(max_dist, model_specs = bw_specs, cv_grid = bw_cv_grid, id = "treeID")
+toc()
 ```
 
-```{r}
+```{r, eval = run_bw}
 glimpse(focal_vs_comp_bw)
@@
 
-```{r}
+```{r, eval = run_scbi}
 scbi_specs <- scbi_growth_df %>% 
@@
 
-```{r}
+```{r, eval = run_scbi}
+tic()
 focal_vs_comp_scbi <- scbi_growth_df %>% 
   create_focal_vs_comp(max_dist, model_specs = scbi_specs, cv_grid = scbi_cv_grid, id = "stemID")
+toc()
 ```
 
-```{r}
+```{r, eval = run_scbi}
 glimpse(focal_vs_comp_scbi)
@@
 
-```{r}
+```{r, eval = run_bw}
+tic()
 bw_fit_model <- focal_vs_comp_bw %>% 
   fit_bayesian_model(model_specs = bw_specs)
+toc()
 ```
@@
 
-```{r}
+```{r, eval = run_bw}
 bw_growth_df <- focal_vs_comp_bw %>% 
@@
 
-```{r}
+```{r, eval = run_scbi}
+tic()
 scbi_fit_model <- focal_vs_comp_scbi %>% 
   fit_bayesian_model(model_specs = scbi_specs)
+toc()
 ```
 
-```{r}
+```{r, eval = run_scbi}
 scbi_growth_df <- focal_vs_comp_scbi %>% 
@@
 
-```{r}
+```{r, eval = run_scbi}
 tic()
@@
 
-```{r}
-
+```{r, eval = run_scbi}
 scbi_cv_predict %>%
   inner_join(scbi_growth_df, by = 'focal_ID', suffix = c('_cv','')) %>%
-  summarise(rmse_cv = sqrt(mean((growth - growth_hat_cv)^2)),
-            rmse = sqrt(mean((growth - growth_hat)^2)), n() )
-
+  summarise(
+    rmse_cv = sqrt(mean((growth - growth_hat_cv)^2)),
+    rmse = sqrt(mean((growth - growth_hat)^2)), 
+    n = n() 
+  )
 ```
@@
 
-```{r}
+```{r, eval = run_bw}
 tic()
@@
 
-```{r}
+We have a problem: not all focal trees have a CV predicted value of growth_hat
 
-bw_cv_predict %>%
-  inner_join(bw_growth_df, by = 'focal_ID', suffix = c('_cv','')) %>%
-  summarise(rmse_cv = sqrt(mean((growth - growth_hat_cv)^2)),
-            rmse = sqrt(mean((growth - growth_hat)^2)), n() )
+```{r, eval = run_bw}
+bw_growth_df %>% 
+  left_join(bw_cv_predict, by = 'focal_ID', suffix = c('', '_cv')) %>% 
+  mutate(has_cv = !is.na(growth_hat_cv)) %>% 
+  st_as_sf() %>% 
+  ggplot() +
+  geom_sf(size = 0.1, aes(col = has_cv))
+```
 
+```{r, eval = run_bw}
+bw_growth_df %>% 
+  left_join(bw_cv_predict, by = 'focal_ID', suffix = c('', '_cv')) %>%
+  filter(!is.na(growth_hat_cv)) %>% 
+  as_tibble() %>% 
+  summarise(
+    rmse_cv = sqrt(mean((growth - growth_hat_cv)^2)),
+    rmse = sqrt(mean((growth - growth_hat)^2)), 
+    n = n() 
+  )
 ```

### 2020-06-04 README.Rmd
working on README

@@
 ```{r, eval = run_bw}
+if (!file.exists("focal_vs_comp_bw.Rdata")) {
   tic()
@@
   toc()
+} else {
+  load("focal_vs_comp_bw.Rdata)
+}
 ```

### 2020-06-04 README.Rmd
minor changes

--- 2020-05-28 README.Rmd
@@
   toc()
+} else {
+  load("focal_vs_comp_bw.Rdata")
+}
 ```

### 2020-06-04 README.Rmd

--- 2020-05-28 README.Rmd
@@
 ```{r, eval = run_bw}
-# Read in Big Woods census data from 2008 & 2014
+# Big Woods
+# Read in census data from 2008 & 2014
 bw_2008 <- 
@@
 
-# Read in bigwoods grouping classification data
+# Read in grouping classification data
 bw_species <- 
@@
 ```{r, eval = run_scbi}
+# SCBI
 scbi_2013 <- 
@@
 ```{r, eval = run_bw}
+# Big Woods
 census_df1 <- bw_2008
@@
 ```{r, eval = run_scbi}
+# SCBI
 census_df1 <- scbi_2013
@@
 ```{r, eval = run_scbi & run_bw}
+# Both Big Woods & SCBI
 growth_df <- bind_rows(
@@
 ```{r, eval = run_bw}
-# Bigwoods study region boundary polygon
+# Bigwoods
+# Study region boundary polygon
 bw_boundary <- bigwoods_study_region %>% 
@@
 ```{r, eval = run_scbi}
-# Bigwoods study region boundary polygon
+# SCBI
+# Study region boundary polygon
 scbi_boundary <- tibble(x = c(0,400,400,0,0), y = c(0,0,640,640,0)) %>% 
@@
 ```{r, eval = run_bw}
+# Big Woods
 set.seed(76)
@@
 ```{r, eval = run_scbi}
+# SCBI
 scbi_cv_grid <- spatialBlock(
@@
 ```{r, eval = run_bw}
+# Big Woods
 bw_specs <- bw_growth_df %>% 
@@
 ```{r, eval = run_bw}
+# Big Woods
+if (!file.exists("focal_vs_comp_bw.Rdata")) {
   tic()
@@
   toc()
+  save(focal_vs_comp_bw, file = "focal_vs_comp_bw.Rdata")
+} else {
+  load("focal_vs_comp_bw.Rdata")
+}
 ```
@@
 ```{r, eval = run_bw}
+# Big Woods
 glimpse(focal_vs_comp_bw)
@@
 ```{r, eval = run_scbi}
+# SCBI
 scbi_specs <- scbi_growth_df %>% 
@@
 ```{r, eval = run_scbi}
+# SCBI
 tic()
@@
 ```{r, eval = run_scbi}
+# SCBI
 glimpse(focal_vs_comp_scbi)
@@
 ```{r, eval = run_bw}
+# Big Woods
 tic()
@@
 ```{r, eval = run_bw}
+# Big Woods
 bw_growth_df <- focal_vs_comp_bw %>% 
@@
 ```{r, eval = run_scbi}
+# SCBI
 scbi_growth_df <- focal_vs_comp_scbi %>% 
@@
 ```{r, eval = run_scbi}
+# SCBI
 scbi_cv_predict %>%
@@
 ```{r, eval = run_bw}
+# Big Woods
 bw_growth_df %>% 

### 2020-06-08 R/modeling.R
starting to work on plotting hit error

@@ -369,3 +369,20 @@ run_cv <- function(focal_vs_comp, model_specs, max_dist, cv_grid,
 
   return(focal_trees)
 }
+
+#' Plot beta_0 parameters
+#'
+#' @inheritParams fit_bayesian_model
+#' @param posterior_param Output of \code{\link{fit_bayesian_model}}
+#'
+#' @import ggridges
+#' @return \code{focal_vs_comp} with new column of predicted \code{growth_hat}
+#' @export
+#'
+#' @examples
+#' 1+1
+plot_beta0 <- function(posterior_param){
+
+
+}
+

### 2020-06-11 README.Rmd
starting to work on plotting hit error
Done run_cv()

@@
 # Big Woods
+if (!file.exists("bw_cv_predict.Rdata")) {
 tic()
@@
 toc()
+  save(bw_cv_predict, file = "bw_cv_predict.Rdata")
+} else {
+  load("bw_cv_predict.Rdata")
+}
 ```
@@
 
-We have a problem: not all focal trees have a CV predicted value of growth_hat
+```{r, eval = run_bw}
+# Big Woods
+bw_growth_df <- bw_growth_df %>% 
+  left_join(bw_cv_predict, by = "focal_ID", suffix = c('', '_cv')) %>% 
+  mutate(has_cv = !is.na(growth_hat))
+
+ggplot(bw_growth_df) +
+  geom_sf(size = 0.1, aes(col = has_cv))
+```
+
+It appears not all trees have a CV predicted value of `growth_hat`. This is because either:
+
+* The tree is part of the buffer
+* The tree was dropped in the `create_focal_vs_comp()` phase
+
+```{r, eval = run_bw}
+# This code is a hideous sanity check of the two facts above. To delete.
+trees_dropped_during_create_focal_vs_comp <- 
+  setdiff(unique(bw_growth_df$focal_ID), unique(focal_vs_comp_bw$focal_ID)) %>% 
+  sort()
+trees_with_no_cv <- bw_growth_df %>% 
+  filter(!has_cv) %>% 
+  pull(focal_ID) %>% 
+  sort()
+mean(trees_dropped_during_create_focal_vs_comp == trees_with_no_cv)
+```
+
+Let's remove these trees that have no
+
+```{r, eval = run_bw}
+bw_growth_df <- bw_growth_df %>% 
+  filter(!is.na(growth_hat))
+```
+
+Let's compute summary statistics of our cross-validated errors:
 
@@
 bw_growth_df %>% 
-  left_join(bw_cv_predict, by = 'focal_ID', suffix = c('', '_cv')) %>% 
-  mutate(has_cv = !is.na(growth_hat_cv)) %>% 
-  st_as_sf() %>% 
-  ggplot() +
-  geom_sf(size = 0.1, aes(col = has_cv))
-```
-
-```{r, eval = run_bw}
-bw_growth_df %>% 
-  left_join(bw_cv_predict, by = 'focal_ID', suffix = c('', '_cv')) %>%
-  filter(!is.na(growth_hat_cv)) %>% 
   as_tibble() %>% 
   summarise(
-    rmse_cv = sqrt(mean((growth - growth_hat_cv)^2)),
+    rmse_cv = sqrt(mean((growth - growth_hat)^2)),
     rmse = sqrt(mean((growth - growth_hat)^2)), 
@@
   )
+```
+
+Let's visualize the spatial distribution of our cross-validated errors:
+
+```{r, eval = run_bw}
+# Big Woods
+bw_growth_df <- bw_growth_df %>%
+  mutate(
+    error = growth - growth_hat,
+    error_bin = cut_number(error, n = 5),
+    error_compress = ifelse(error < -0.75, -0.75, ifelse(error > 0.75, 0.75, error))
+  )
+
+ggplot(bw_growth_df) +
+  geom_sf(aes(col = error_compress), size = 0.4) +
+  scale_color_gradient2(
+    low = "#ef8a62", mid = "#f7f7f7", high = "#67a9cf", 
+    name = expression(paste("Residual (cm ", y^{-1}, ")")),
+    breaks = seq(from = -0.75, to = 0.75, by = 0.25),
+    labels = c("< -0.75", "-0.5", "0.25", "0", "0.25", "0.5", "> 0.75")
+  ) +
+  labs(x = "Meter", y = "Meter")
 ```
@@
 * Plot posterior distributions of all parameters
+
+```{r, eval = run_bw}
+
+# plot_beta0(bw_fit_model)
+
+```

### 2020-06-18 README.Rmd
Done permutations

@@
 library(tictoc)
+library(yardstick)
+library(viridis)
 
@@
   filter(!foldID %in% c(19, 23, 21, 17, 8, 19))
+
+# Save original
+bw_growth_df_orig <- bw_growth_df
 ```
@@
 # Big Woods
-bw_specs <- bw_growth_df %>% 
+bw_specs <- bw_growth_df_orig %>% 
   get_model_specs(model_number = 3, species_notion = 'trait_group')
@@
   tic()
-  focal_vs_comp_bw <- bw_growth_df %>% 
+  focal_vs_comp_bw <- bw_growth_df_orig %>% 
     create_focal_vs_comp(max_dist, model_specs = bw_specs, cv_grid = bw_cv_grid, id = "treeID")
@@
   predict_bayesian_model(model_specs = bw_specs, posterior_param = bw_fit_model) %>% 
-  right_join(bw_growth_df, by = c("focal_ID" = "treeID"))
+  right_join(bw_growth_df_orig, by = c("focal_ID" = "treeID"))
 
@@
 
+```{r, eval = FALSE}
+# Number of permutation shuffles:
+num_shuffle <- 4
+
+# Compute observed RMSE for all models, but only do permutation shuffling for
+# models 3,6,9
+model_numbers <- c(3)
+
+# Save results here
+run_time <- rep(0, length(model_numbers))
+observed_RMSE <- rep(0, length(model_numbers))
+observed_RMSE_CV <- rep(0, length(model_numbers))
+shuffle_RMSE <- list()
+shuffle_RMSE_CV <- list()
+
+for(i in 1:length(model_numbers)){
+  # Start clock
+  tic()
+
+  # Modeling and species stuff
+  model_number <- model_numbers[i]
+  bw_specs <- bw_growth_df %>% 
+    get_model_specs(model_number = model_number, species_notion = 'trait_group')
+
+  # Focal vs comp main dataframe for analysis
+  # Computed earlier
+  load("focal_vs_comp_bw.Rdata")
+
+  # 1. Compute observed test statistic: RMSE with no cross-validation ----
+  # Fit model (compute posterior parameters)
+  bw_fit_model <- focal_vs_comp_bw %>% 
+    fit_bayesian_model(model_specs = bw_specs)
+
+  # Make predictions, compute and save RMSE, and reset
+  observed_RMSE[i] <- focal_vs_comp_bw %>% 
+    predict_bayesian_model(model_specs = bw_specs, posterior_param = bw_fit_model) %>% 
+    right_join(bw_growth_df_orig, by = c("focal_ID" = "treeID")) %>%
+    rmse(truth = growth, estimate = growth_hat) %>%
+    pull(.estimate)
+
+  # 2. Compute observed test statistic: RMSE with cross-validation ----
+  observed_RMSE_CV[i] <- focal_vs_comp_bw %>%
+    run_cv(model_specs = bw_specs, max_dist = max_dist, cv_grid = bw_cv_grid) %>% 
+    right_join(bw_growth_df_orig, by = c("focal_ID" = "treeID")) %>%
+    rmse(truth = growth, estimate = growth_hat) %>%
+    pull(.estimate)
+
+  # 3. Permutation distribution: RMSE with no cross-validation ----
+  # Only do permutation shuffling for models 3, 6, 9
+  if(model_number %in% c(3, 6, 9)){
+    # Compute num_shuffle permutation test statistics
+    shuffle_RMSE[[i]] <- numeric(length = num_shuffle)
+
+    for(j in 1:num_shuffle){
+      # Fit model (compute posterior parameters) with shuffling
+      bw_fit_model_shuffle <- focal_vs_comp_bw %>% 
+        fit_bayesian_model(model_specs = bw_specs, run_shuffle = TRUE)
+
+      # Make predictions, compute and save RMSE, and reset
+      shuffle_RMSE[[i]][j] <- focal_vs_comp_bw %>% 
+        predict_bayesian_model(model_specs = bw_specs, posterior_param = bw_fit_model_shuffle) %>% 
+        right_join(bw_growth_df_orig, by = c("focal_ID" = "treeID")) %>%
+        rmse(truth = growth, estimate = growth_hat) %>%
+        pull(.estimate)
+    }
+  } else {
+    shuffle_RMSE[[i]] <- NA
+  }
+
+  # 4. Permutation distribution: RMSE with cross-validation ----
+  # Only do permutation shuffling for models 3, 6, 9
+  if(model_number %in% c(3, 6, 9)){
+    # Compute num_shuffle permutation test statistics
+    shuffle_RMSE_CV[[i]] <- numeric(length = num_shuffle)
+
+    # Compute num_shuffle permutation test statistics
+    for(j in 1:num_shuffle){
+      # Compute and save RMSE, and reset
+      shuffle_RMSE_CV[[i]][j] <- focal_vs_comp_bw %>%
+        run_cv(model_specs = bw_specs, max_dist = max_dist, cv_grid = bw_cv_grid, run_shuffle = TRUE) %>% 
+        right_join(bw_growth_df_orig, by = c("focal_ID" = "treeID")) %>%
+        rmse(truth = growth, estimate = growth_hat) %>%
+        pull(.estimate)
+
+      # Status update
+      str_c("Model ", model_numbers[i], ", shuffle with permutation ", j) %>% print()
+    }
+  } else {
+    shuffle_RMSE_CV[[i]] <- NA
+  }
+
+  # 5. Stop clock and save
+  clock <- toc(quiet = TRUE)
+  run_time[i] <- clock$toc - clock$tic
+}
+
+model_comp_tbl <- tibble(
+  model_number = model_numbers,
+  run_time = run_time,
+  observed_RMSE = observed_RMSE,
+  observed_RMSE_CV = observed_RMSE_CV,
+  shuffle_RMSE = shuffle_RMSE,
+  shuffle_RMSE_CV = shuffle_RMSE_CV,
+)
+
+save(model_comp_tbl, file = "model_comp_tbl_4_shuffles.RData")
+```
+
+```{r}
+load("model_comp_tbl_4_shuffles.RData")
+model_comp <- bind_rows(
+  model_comp_tbl %>% select(model_number, run_time, observed = observed_RMSE, shuffle = shuffle_RMSE) %>% mutate(CV = FALSE),
+  model_comp_tbl %>% select(model_number, run_time, observed = observed_RMSE_CV, shuffle = shuffle_RMSE_CV) %>% mutate(CV = TRUE)
+) %>%
+  gather(type, RMSE, -c(model_number, run_time, CV)) %>%
+  arrange(CV, model_number) %>%
+  mutate(
+    model_number = case_when(
+      model_number == 3 ~ "1. Trait-based (6): lambda = 6 x 6",
+      model_number == 6 ~ "2. Phylogenetic family (20): lambda = 20 x 20",
+      model_number == 9 ~ "3. Actual species (36): lambda = 36 x 36"
+    )
+  )
+
+model_comp_observed <- model_comp %>%
+  filter(type == "observed") %>%
+  unnest()
+model_comp_shuffle <- model_comp %>%
+  filter(type == "shuffle") %>%
+  unnest()
+
+xlab <- expression(paste('RMSE (cm ',y^{-1},')'))
+
+ggplot() +
+  geom_vline(data = model_comp_observed, aes(xintercept = RMSE, col = CV), linetype = "dashed", show.legend = F) +
+  geom_histogram(data = model_comp_shuffle, aes(x = RMSE, fill = CV), bins = 200) +
+  labs(fill = "Cross-validated?", x = xlab) +
+  facet_wrap(~model_number, ncol = 1) +
+  scale_color_viridis(discrete = TRUE, option = "D")+
+  scale_fill_viridis(discrete = TRUE)
+```
+
 ### Visualize posterior distributions

### 2020-06-18 R/modeling.R

@@ -357,7 +357,7 @@ run_cv <- function(focal_vs_comp, model_specs, max_dist, cv_grid,
 
     # now pretty easy to just call the two functions!
     fold_fit <- train %>%
-      fit_bayesian_model(model_specs)
+      fit_bayesian_model(model_specs, run_shuffle = run_shuffle)
     fold_predict <- test %>%
       predict_bayesian_model(model_specs, fold_fit)
 

### 2020-06-18 README.Rmd
Readied code for Dave inspection

@@
 
-* Permutation test:
-  + Add line to that shuffles competitor species within `focal_and_comp` just before fitting model
-  + permutation AND CV: add `run_shuffle` argument to `run_cv()`
+Hey Dave! Start here!
+
+**Repeat earlier code**:
+
+```{r, eval = FALSE}
+# Load packages ----------------------------------------------------------------
+library(tidyverse)
+library(forestecology)
+library(snakecase)
+library(skimr)
+library(sf)
+library(sfheaders)
+# devtools::install_github("rvalavi/blockCV")
+library(blockCV)
+library(tictoc)
+library(yardstick)
+library(viridis)
+
+# Load & preprocess data -------------------------------------------------------
+# Read in census data from 2008 & 2014
+bw_2008 <- 
+  "https://deepblue.lib.umich.edu/data/downloads/z603qx485" %>% 
+  read_delim(delim = "\t") %>% 
+  mutate(spcode = to_any_case(spcode)) %>%
+  select(
+    treeID = treeid, stemID = stemtag, sp = spcode, quadrat, gx, gy, dbh, 
+    date, codes
+  )
+bw_2014 <- 
+  "https://deepblue.lib.umich.edu/data/downloads/1831ck00f" %>% 
+  read_delim(delim = "\t") %>% 
+  mutate(spcode = to_any_case(spcode)) %>%
+  select(
+    treeID = treeid, stemID = stemtag, sp = spcode, quadrat, gx, gy, dbh, 
+    date, codes
+  )
+
+# Read in grouping classification data
+bw_species <- 
+  "https://deepblue.lib.umich.edu/data/downloads/000000086" %>% 
+  read_delim(delim = "\t") %>% 
+  # convert all to snake case:
+  mutate_at(c("species", "genus", "family", "idlevel", "spcode"), to_any_case) %>% 
+  # join trait group
+  left_join(families, by = c("spcode", "family")) %>% 
+  mutate(
+    sp = str_sub(genus, 1, 2), 
+    sp = str_c(sp, str_sub(species, 1, 2)),
+    sp = tolower(sp),
+    latin = str_c(genus, species, sep = " "),
+    latin = to_any_case(latin)
+  ) %>% 
+  select(sp = spcode, genus, species, latin, family, trait_group)
+
+bw_2008 <- bw_2008 %>%
+  left_join(bw_species,by='sp')
+
+# Compute growth ---------------------------------------------------------------
+census_df1 <- bw_2008
+# we need to filter out the resprouts
+census_df2 <- bw_2014 %>% 
+  filter(!str_detect(codes, 'R'))
+id <- "treeID"
+
+bw_growth_df <- 
+  # Merge both censuses and compute growth:
+  compute_growth(census_df1, census_df2, id) %>% 
+  mutate(sp = to_any_case(sp))
+
+# Define buffers ---------------------------------------------------------------
+cv_fold_size <- 100
+max_dist <- 7.5
+
+# Study region boundary polygon
+bw_boundary <- bigwoods_study_region %>% 
+  sf_polygon()
+
+# Buffer polygon
+bw_buffer <- bw_boundary %>%
+  st_buffer(dist = -max_dist)
+
+# Convert data frame to sf object
+bw_growth_df <- bw_growth_df %>% 
+  st_as_sf(coords = c("gx", "gy"))
+
+# ID which points are in buffer and which are not
+buffer_index <- !st_intersects(bw_growth_df, bw_buffer, sparse = FALSE)
+bw_growth_df <- bw_growth_df %>% 
+  mutate(buffer = as.vector(buffer_index))
+
+# Define spatial CV folds ------------------------------------------------------
+set.seed(76)
+bw_cv_grid <- spatialBlock(
+  speciesData = bw_growth_df, theRange = 100, verbose = FALSE,
+  # Some guess work in figuring this out:
+  k = 28, xOffset = 0.5, yOffset = 0
+)
+
+# Add foldID to data
+bw_growth_df <- bw_growth_df %>% 
+  mutate(
+    foldID = bw_cv_grid$foldID
+  )
+
+# Visualize grid. Why does fold 19 repeat?
+bw_cv_grid$plots +
+  geom_sf(data = bw_growth_df %>% sample_frac(0.2), aes(col=factor(foldID)), size = 0.1)
+
+# Remove weird folds with no trees in them from viz above
+bw_growth_df <- bw_growth_df %>%
+  filter(!foldID %in% c(19, 23, 21, 17, 8, 19))
+
+# Save original
+bw_growth_df_orig <- bw_growth_df
+```
+
+**Loop through 4 scenarios of model fitting**:
 
@@
   # Focal vs comp main dataframe for analysis
-  # Computed earlier
-  load("focal_vs_comp_bw.Rdata")
+  focal_vs_comp_bw <- bw_growth_df_orig %>% 
+    create_focal_vs_comp(max_dist, model_specs = bw_specs, cv_grid = bw_cv_grid, id = "treeID")
 
@@
 
-save(model_comp_tbl, file = "model_comp_tbl_4_shuffles.RData")
+save(model_comp_tbl, file = str_c(format(Sys.time(), "%Y-%m-%d"), "_model_comp_tbl_99_shuffles.RData"))
 ```
+
+**Plot results**:
 

### 2020-06-23 README.Rmd
tweak:
Working on the plotting code

@@
 
-save(model_comp_tbl, file = str_c(format(Sys.time(), "%Y-%m-%d"), "_model_comp_tbl_99_shuffles.RData"))
+save(model_comp_tbl, file = str_c(format(Sys.time(), "%Y-%m-%d"), "_model_comp_tbl_", num_shuffle, "_shuffles.RData"))
 ```
@@
 
+We might be interested in the posterior distributions of parameters.
+The betas tell us about how fast each species grows and how this depends on DBH.
+The lambdas, which are often of more interest, are the species-specific competition coefficents.
+The full lambda matrix gives competition strength between species.
+There is a rich literature how this matrxi (cite).
+
+Because of the strucutre of the `bw_fit_model` object we cannot simply draw these curves based on the posterior distribution.
+`bw_fit_model` gives the parameteres *compared* to a baseline.
+This is often not of interest.
+So to display these parameters as we care about them we have to sample from the baseline distrubiton and from the comparison one to get the posterior distribution of interest.
+
 ```{r, eval = run_bw}

### 2020-06-23 R/modeling.R

@@ -376,6 +376,8 @@ run_cv <- function(focal_vs_comp, model_specs, max_dist, cv_grid,
 #' @param posterior_param Output of \code{\link{fit_bayesian_model}}
 #'
 #' @import ggridges
+#' @importFrom mvnfast rmvt
+
 #' @return \code{focal_vs_comp} with new column of predicted \code{growth_hat}
 #' @export
 #'
@@ -451,42 +453,6 @@ plot_beta0 <- function(posterior_param, model_specs){
   }
 
 
-  # do we have to sample from the distribution to make the curves?
-  # since we know mu and sigma, can't we just plot the exact curves?
-  # would this be faster?
-  # I am having a hard time pulling them out from the posterior_param object
-  # but here is my thought
-
-  params <- tibble(species = c('Oak', 'Maple', 'Hickory', 'Cherry'),
-                   mu = c(0.1, 0.05, 0.15, 0.075),
-                   sigma = c(0.025, 0.015, 0.05, 0.03))
-
-  x_lims <- params %>%
-    mutate(max_val = mu + sigma*4,
-           min_val = mu - sigma*4) %>%
-    summarise(min_val = min(min_val), max_val = max(max_val)) %>%
-    as.numeric()
-
-  x_range_len <- 500
-  x_range <- seq(from = x_lims[1], to = x_lims[2], length = x_range_len)
-
-  posterior_curves <- tibble(species = rep(params$species[1], x_range_len),
-                             x = x_range,
-                             y = dnorm(x_range,params$mu[1],params$sigma[1]))
-
-  # can you do this will purrr??
-  for (sp in 2:dim(params)[1])
-  {
-    to_add <- tibble(species = rep(params$species[sp], x_range_len),
-                               x = x_range,
-                               y = dnorm(x_range,params$mu[sp],params$sigma[sp]))
-
-    posterior_curves <- rbind(posterior_curves, to_add)
-  }
-
-  ggplot(posterior_curves, aes(x = x, y = species, height = y)) +
-    geom_density_ridges(stat = 'identity') +
-    geom_vline(xintercept = 0, linetype = "dashed")
 
 
 

### 2020-06-25 README.Rmd
Fixed README to account for diff notion of species

--- 2020-06-18 README.Rmd
@@
 # Number of permutation shuffles:
-num_shuffle <- 4
+num_shuffle <- 2
 
@@
 model_numbers <- c(3)
+species_notion_vector <- c("trait_group", "family", "species")
 
 # Save results here
-run_time <- rep(0, length(model_numbers))
-observed_RMSE <- rep(0, length(model_numbers))
-observed_RMSE_CV <- rep(0, length(model_numbers))
+run_time <- rep(0, length(species_notion_vector))
+observed_RMSE <- rep(0, length(species_notion_vector))
+observed_RMSE_CV <- rep(0, length(species_notion_vector))
 shuffle_RMSE <- list()
@@
 
-for(i in 1:length(model_numbers)){
+for(i in 1:length(species_notion_vector)){
   # Start clock
@@
   # Modeling and species stuff
-  model_number <- model_numbers[i]
+  model_number <- 3
+  species_notion <- species_notion_vector[i]
   bw_specs <- bw_growth_df %>% 
-    get_model_specs(model_number = model_number, species_notion = 'trait_group')
+    get_model_specs(model_number = model_number, species_notion = species_notion)
 
@@
       # Status update
-      str_c("Model ", model_numbers[i], ", shuffle with permutation ", j) %>% print()
+      str_c("notion of species: ", species_notion_vector[i], ", shuffle with permutation ", j) %>% print()
     }

### 2020-06-25 README.Rmd
About to relaunch full simulation

@@
 # Number of permutation shuffles:
-num_shuffle <- 2
+num_shuffle <- 4
 
@@
 observed_RMSE_CV <- rep(0, length(species_notion_vector))
-shuffle_RMSE <- list()
-shuffle_RMSE_CV <- list()
+shuffle_RMSE <- vector("list", length(species_notion_vector))
+shuffle_RMSE_CV <- vector("list", length(species_notion_vector))
+filename <- str_c(format(Sys.time(), "%Y-%m-%d"), "_model_comp_tbl_", num_shuffle, "_shuffles.RData")
 
@@
   # Modeling and species stuff
-  model_number <- 3
   species_notion <- species_notion_vector[i]
   bw_specs <- bw_growth_df %>% 
-    get_model_specs(model_number = model_number, species_notion = species_notion)
+    get_model_specs(model_number = 3, species_notion = species_notion)
 
@@
   # Only do permutation shuffling for models 3, 6, 9
-  if(model_number %in% c(3, 6, 9)){
   # Compute num_shuffle permutation test statistics
@@
   }
-  } else {
-    shuffle_RMSE[[i]] <- NA
-  }
 
   # 4. Permutation distribution: RMSE with cross-validation ----
-  # Only do permutation shuffling for models 3, 6, 9
-  if(model_number %in% c(3, 6, 9)){
   # Compute num_shuffle permutation test statistics
@@
   }
-  } else {
-    shuffle_RMSE_CV[[i]] <- NA
-  }
 
-  # 5. Stop clock and save
+  # 5. Stop clock
   clock <- toc(quiet = TRUE)
   run_time[i] <- clock$toc - clock$tic
-}
 
+  # 6. Save
   model_comp_tbl <- tibble(
-  model_number = model_numbers,
+    species_notion = species_notion_vector,
     run_time = run_time,
@@
   )
-
-save(model_comp_tbl, file = str_c(format(Sys.time(), "%Y-%m-%d"), "_model_comp_tbl_", num_shuffle, "_shuffles.RData"))
+  save(model_comp_tbl, file = filename)
+}
 ```
@@
 
-```{r}
+```{r, eval=FALSE}
 load("model_comp_tbl_4_shuffles.RData")
 model_comp <- bind_rows(
-  model_comp_tbl %>% select(model_number, run_time, observed = observed_RMSE, shuffle = shuffle_RMSE) %>% mutate(CV = FALSE),
-  model_comp_tbl %>% select(model_number, run_time, observed = observed_RMSE_CV, shuffle = shuffle_RMSE_CV) %>% mutate(CV = TRUE)
+  model_comp_tbl %>% select(species_notion, run_time, observed = observed_RMSE, shuffle = shuffle_RMSE) %>% mutate(CV = FALSE),
+  model_comp_tbl %>% select(species_notion, run_time, observed = observed_RMSE_CV, shuffle = shuffle_RMSE_CV) %>% mutate(CV = TRUE)
 ) %>%
-  gather(type, RMSE, -c(model_number, run_time, CV)) %>%
+  gather(type, RMSE, -c(species_notion, run_time, CV)) %>%
   arrange(CV, model_number) %>%
   mutate(
-    model_number = case_when(
-      model_number == 3 ~ "1. Trait-based (6): lambda = 6 x 6",
-      model_number == 6 ~ "2. Phylogenetic family (20): lambda = 20 x 20",
-      model_number == 9 ~ "3. Actual species (36): lambda = 36 x 36"
+    species_notion = case_when(
+      species_notion == "trait_group" ~ "1. Trait-based (6): lambda = 6 x 6",
+      species_notion == "family" ~ "2. Phylogenetic family (20): lambda = 20 x 20",
+      species_notion == "species" ~ "3. Actual species (36): lambda = 36 x 36"
     )
@@
   filter(type == "observed") %>%
-  unnest()
+  unnest(cols = c(RMSE))
 model_comp_shuffle <- model_comp %>%
   filter(type == "shuffle") %>%
-  unnest()
+  unnest(cols = c(RMSE))
 
@@
   labs(fill = "Cross-validated?", x = xlab) +
-  facet_wrap(~model_number, ncol = 1) +
+  facet_wrap(~species_notion, ncol = 1) +
   scale_color_viridis(discrete = TRUE, option = "D")+

### 2020-07-06 README.Rmd
Build README

@@
 
-```{r, eval=FALSE}
-load("model_comp_tbl_4_shuffles.RData")
+```{r, eval=TRUE}
+load("2020-06-25_model_comp_tbl_49_shuffles.RData")
 model_comp <- bind_rows(
@@
   gather(type, RMSE, -c(species_notion, run_time, CV)) %>%
-  arrange(CV, model_number) %>%
   mutate(
@@
     )
-  )
+  ) %>% 
+  filter(species_notion != "3. Actual species (36): lambda = 36 x 36")
 

### 2020-08-21 README.Rmd
small change
Working on readme edit

@@
 
-run_scbi <- FALSE
-run_bw <- TRUE
+run_scbi <- TRUE
+run_bw <- FALSE
 ```
@@
 
-We first load and preprocess the data corresponding to the two census reflecting the time period over which we consider growth in dbh.
+The data for the Big Woods plor are included in the package.
+For SCBI, we load and preprocess the data corresponding to the two census reflecting the time period over which we consider growth in dbh.
 We load them into R as "tibble" data frames thereby ensuring a standardized input/output format that can be used across all `tidyverse` packages @tidyverse.
@@
 # Read in census data from 2008 & 2014
-bw_2008 <- 
-  "https://deepblue.lib.umich.edu/data/downloads/z603qx485" %>% 
-  read_delim(delim = "\t") %>% 
-  mutate(spcode = to_any_case(spcode)) %>%
-  select(
-    treeID = treeid, stemID = stemtag, sp = spcode, quadrat, gx, gy, dbh, 
-    date, codes
-  )
-bw_2014 <- 
-  "https://deepblue.lib.umich.edu/data/downloads/1831ck00f" %>% 
-  read_delim(delim = "\t") %>% 
-  mutate(spcode = to_any_case(spcode)) %>%
-  select(
-    treeID = treeid, stemID = stemtag, sp = spcode, quadrat, gx, gy, dbh, 
-    date, codes
-  )
-
-# Read in grouping classification data
-bw_species <- 
-  "https://deepblue.lib.umich.edu/data/downloads/000000086" %>% 
-  read_delim(delim = "\t") %>% 
-  # convert all to snake case:
-  mutate_at(c("species", "genus", "family", "idlevel", "spcode"), to_any_case) %>% 
-  # join trait group
-  left_join(families, by = c("spcode", "family")) %>% 
-  mutate(
-    sp = str_sub(genus, 1, 2), 
-    sp = str_c(sp, str_sub(species, 1, 2)),
-    sp = tolower(sp),
-    latin = str_c(genus, species, sep = " "),
-    latin = to_any_case(latin)
-  ) %>% 
-  select(sp = spcode, genus, species, latin, family, trait_group)
-
-bw_2008 <- bw_2008 %>%
-  left_join(bw_species,by='sp')
+data(bw_census_2008, bw_census_2014 )
 
@@
 # Big Woods
-census_df1 <- bw_2008
-# we need to filter out the resprouts
-census_df2 <- bw_2014 %>% 
-  filter(!str_detect(codes, 'R'))
 id <- "treeID"
@@
   # Merge both censuses and compute growth:
-  compute_growth(census_df1, census_df2, id) %>% 
-  mutate(sp = to_any_case(sp))
+  compute_growth(bw_census_2008, bw_census_2014, id) %>%
+  mutate(sp = to_any_case(sp)) %>%
+  # Convert data frame to sf object
+  st_as_sf(coords = c("gx", "gy")) %>%
+  # drop stemID
+  select(-stemID)
+
 ```
@@
 
+We use the function `add_buffer_variable` to identifiy which trees are inside of a buffer region of size `max_dist`.
+
 **Example of buffer at work**:
@@
 # Bigwoods
-# Study region boundary polygon
-bw_boundary <- bigwoods_study_region %>% 
-  sf_polygon()
+data(bw_study_region)
 
-# Buffer polygon
-bw_buffer <- bw_boundary %>%
-  st_buffer(dist = -max_dist)
+# Add buffer variable to data frame
+bw_growth_df <- bw_growth_df %>%
+  add_buffer_variable(direction = "in", size = max_dist, region = bw_study_region)
 
-# Convert data frame to sf object
-bw_growth_df <- bw_growth_df %>% 
-  st_as_sf(coords = c("gx", "gy"))
+ggplot() +
+  geom_sf(data = bw_growth_df, aes(col = buffer))
 
-# ID which points are in buffer and which are not
-buffer_index <- !st_intersects(bw_growth_df, bw_buffer, sparse = FALSE)
-bw_growth_df <- bw_growth_df %>% 
-  mutate(buffer = as.vector(buffer_index))
-
-# Plot
-ggplot() +  
-  geom_sf(data = bw_boundary) +
-  geom_sf(data = bw_buffer, col="red") +
-  # Only random sample of 1000 trees:
-  geom_sf(data = bw_growth_df %>% sample_n(1000), aes(col=buffer), size = 0.5)
 ```

### 2020-08-27 README.Rmd
working through changes to readme with new package functions

@@
 library(viridis)
+library(lubridate)
 
@@
     date = ExactDate, codes, status
-  )
+  ) %>%
+  mutate(date = mdy(date))
 
@@
   ) %>% 
-  mutate(dbh = as.numeric(dbh))
+  mutate(dbh = as.numeric(dbh),
+         date = mdy(date)) 
 ```
@@
   mutate(sp = to_any_case(sp)) %>%
-  # Convert data frame to sf object
+  # Convert data frame to sf object DA note: should compute_growth do this??
   st_as_sf(coords = c("gx", "gy")) %>%
@@
   # they are mesuaring in mm while we are measuring in cm!!!
-  mutate(growth = growth/10)
+  mutate(growth = growth/10) %>%
+  st_as_sf(coords = c("gx", "gy"))
+
 ```
@@
 growth_df <- bind_rows(
-  bw_growth_df %>% select(growth) %>% mutate(site = "bw"),
-  scbi_growth_df %>% select(growth) %>% mutate(site = "scbi")
+  bw_growth_df %>% st_drop_geometry %>% select(growth) %>% mutate(site = "bw"),
+  scbi_growth_df %>% st_drop_geometry %>% select(growth) %>% mutate(site = "scbi")
 )
@@
 
+# "Trees" in polygon
+trees_df <- tibble(
+    x = runif(100),
+    y = runif(100)
+  ) %>%
+  sf_point()
+
 # Buffer polygon
-square_buffer <- square_boundary %>% 
-  st_buffer(dist = -0.1)
+trees_df <- trees_df %>%
+  add_buffer_variable(direction = "in", size = 0.1, region = square_boundary)
 
 ggplot() +
-  geom_sf(data = square_boundary) +
-  geom_sf(data = square_buffer, col="red") 
+  geom_sf(data = trees_df, aes(col = buffer))
+
 ```
@@
 # Study region boundary polygon
-scbi_boundary <- tibble(x = c(0,400,400,0,0), y = c(0,0,640,640,0)) %>% 
+scbi_study_region <- tibble(x = c(0,400,400,0,0), y = c(0,0,640,640,0)) %>% 
   sf_polygon()
 
-# Buffer polygon
-scbi_buffer <- scbi_boundary %>%
-  st_buffer(dist = -max_dist)
+# Add buffer variable to data frame
+scbi_growth_df <- scbi_growth_df %>%
+  add_buffer_variable(direction = "in", size = max_dist, region = scbi_study_region)
 
-# Convert data frame to sf object
-scbi_growth_df <- scbi_growth_df %>% 
-  st_as_sf(coords = c("gx", "gy"))
-
-# ID which points are in buffer and which are not
-buffer_index <- !st_intersects(scbi_growth_df, scbi_buffer, sparse = FALSE)
-scbi_growth_df <- scbi_growth_df %>% 
-  mutate(buffer = as.vector(buffer_index))
-
-# Plot
 ggplot() +
-  geom_sf(data = scbi_boundary) +
-  geom_sf(data = scbi_buffer, col="red") +
-  # Only random sample of 1000 trees:
-  geom_sf(data = scbi_growth_df %>% sample_n(1000), aes(col=buffer), size = 0.5)
+  geom_sf(data = scbi_growth_df, aes(col = buffer))
 ```

### 2020-08-28 README.Rmd
remove st_as_sf since now in compute_growth

@@
   mutate(sp = to_any_case(sp)) %>%
-  # Convert data frame to sf object DA note: should compute_growth do this??
-  st_as_sf(coords = c("gx", "gy")) %>%
   # drop stemID
@@
   # they are mesuaring in mm while we are measuring in cm!!!
-  mutate(growth = growth/10) %>%
-  st_as_sf(coords = c("gx", "gy"))
+  mutate(growth = growth/10)
 

### 2020-09-16 README.Rmd
deal with resprout issue
Updated upto the blockCV stuff.
readme through shuffle RMSE comparison

@@
 
+# we need to filter out the resprouts
+bw_census_2014 <- bw_census_2014 %>% 
+  filter(!str_detect(codes, 'R'))
+
 bw_growth_df <-
@@
 
-We use the function `add_buffer_variable` to identifiy which trees are inside of a buffer region of size `max_dist`.
+We use the function `add_buffer_variable` to identifiy which trees are inside of a buffer region of size `max_dist`. The user will need to specify a study region and convert to a `sf_polygon`.
 
@@
 
-# Remove weird folds with no trees in them from viz above
 bw_growth_df <- bw_growth_df %>%
-  filter(!foldID %in% c(19, 23, 21, 17, 8, 19))
+  filter(!foldID %in% c(19, 23, 21, 17, 8, 19)) %>%
+  mutate(foldID = as.character(foldID))
 
-# Save original
-bw_growth_df_orig <- bw_growth_df
+# Deliverable
+ggplot() +
+  geom_sf(data = bw_growth_df, aes(col = foldID, alpha = buffer))
+
+bw_cv_grid_sf <- bw_cv_grid$blocks %>%
+  st_as_sf()
 ```
@@
 
-### Model specification
-
-Next we specify the growth model we want.
-We will model growth as a function of tree identity, size (in dbh), and the identity and size (in dbh) of its neighbors.
-The model can be run on different grouping of individuals (e.g., based on species or trait groupings).
-It can also be run with different notions of competition:
-
-* `model_number = 1`: No competition growth only depends on focal tree `dbh` and grouping.
-* `model_number = 2`: Competition but identity of competitor does not matter, just sum of competitior basal area.
-* `model_number = 3`: Full model with competition including competitor identity.
-
-**Big Woods**:
-
-```{r, eval = run_bw}
-# Big Woods
-bw_specs <- bw_growth_df_orig %>% 
-  get_model_specs(model_number = 3, species_notion = 'trait_group')
-bw_specs
-```
+### Focal versus competitor trees
 
@@
 
-This requires the `growth_df` data frame, `max_dist` scalar defining competitive range, `cv_fold_size` defining the size of the spatial cross-validation blocks, and `model_specs` as inputs.
+This requires the `growth_df` data frame; `max_dist`, the scalar defining competitive range; `cv_fold_size`, defining the size of the spatial cross-validation blocks; and the `id` variable as inputs.
 
@@
   tic()
-  focal_vs_comp_bw <- bw_growth_df_orig %>% 
-    create_focal_vs_comp(max_dist, model_specs = bw_specs, cv_grid = bw_cv_grid, id = "treeID")
+  focal_vs_comp_bw <- bw_growth_df %>% 
+    create_focal_vs_comp(max_dist, cv_grid_sf = bw_cv_grid_sf, id = "treeID")
   toc()
@@
 
-Now we are ready to fit the competition model with `fit_bayesian_model` this takes two inputs: `focal_vs_comp` the data frame with focal trees connected to their competitors and `model_specs` which specifies the notion of competition.
+Now we are ready to fit the competition model with `fit_bayesian_model`.
+This function needs only the `focal_vs_comp` as an input.
+Other options allow the user to specify prior parameters and run a species identity shuffle (see below).
 
@@
 tic()
-bw_fit_model <- focal_vs_comp_bw %>% 
-  fit_bayesian_model(model_specs = bw_specs)
+posterior_param_bw <- focal_vs_comp_bw %>% 
+  fit_bayesian_model(prior_param = NULL, run_shuffle = FALSE)
 toc()
@@
 This output has the posterior parameters for the specified competition model.
-If `model_number = 1` this includes just the beta values for how focal tree DBH and identity affect growth.
-If `model_number = 2` it also includes a vector of lambdas for how total competitor basal area affect individuals of each group.
-If `model_number = 3` then lambda is a matrix which gives how individuals of each focal group respond to the competition of each competitor group.
 This `posterior_param` output can be used to get predicted growths for each individual (with `predict_bayesain_model`) to test how well the model performs.
@@
 
+Here we calculate the RMSE
+
 ```{r, eval = run_bw}
 # Big Woods
-bw_growth_df <- focal_vs_comp_bw %>% 
-  predict_bayesian_model(model_specs = bw_specs, posterior_param = bw_fit_model) %>% 
-  right_join(bw_growth_df_orig, by = c("focal_ID" = "treeID"))
+predictions <- focal_vs_comp_bw %>%
+  predict_bayesian_model(posterior_param = posterior_param_bw) %>%
+  right_join(bw_growth_df, by = c("focal_ID" = "treeID"))
+predictions %>%
+  rmse(truth = growth, estimate = growth_hat) %>%
+  pull(.estimate)
+```
 
-bw_growth_df <- bw_growth_df %>% 
-  st_as_sf()
+Now we test whether the identity of the competitor matters.
+We do this by shuffling the identity of competitors (but not of focal trees or spatial locations or sizes) and fitting the model again.
+We then compare RMSEs to see whether competitor identity matters to competitive effects
 
-# Observed vs predicted growth  
-ggplot(bw_growth_df, aes(x = growth, y = growth_hat)) +
-  geom_point(size = 0.5, color = rgb(0, 0, 0, 0.25)) +
-  stat_smooth(method = 'lm') +
-  geom_abline(slope = 1, intercept = 0) +
-  coord_fixed() + 
-  labs(
-    x = "Observed growth in dbh", y = "Predicted growth in dbh", 
-    title = "Predicted vs Observed Growth"
-  )
+```{r, eval = run_bw}
+posterior_param_bw_shuffle <- focal_vs_comp_bw %>%
+  fit_bayesian_model(prior_param = NULL, run_shuffle = TRUE)
 
-reslab <- expression(paste('Residual (cm ',y^{-1},')'))
-bw_growth_df %>% 
-  # Need to investigate missingness
-  filter(!is.na(growth_hat)) %>% 
-  mutate(
-    error = growth - growth_hat,
-    error_bin = cut_number(error, n = 5), 
-    error_compress = ifelse(error < -0.75, -0.75, ifelse(error > 0.75, 0.75, error))
-  ) %>% 
-  ggplot() + 
-  geom_sf(aes(col = error_compress), size = 0.4) + 
-  theme_bw() + 
-  scale_color_gradient2(
-    low = "#ef8a62", mid = "#f7f7f7", high = "#67a9cf", 
-    name = reslab,
-    breaks = seq(from = -0.75, to = 0.75, by = 0.25),
-    labels = c('< -0.75', '-0.5', '0.25', '0', '0.25', '0.5', '> 0.75')) +
-  labs(x = "Meter", y = "Meter")
+# b) Make predictions and compute RMSE
+predictions_shuffle <- focal_vs_comp_bw %>%
+  predict_bayesian_model(posterior_param = posterior_param_bw_shuffle) %>%
+  right_join(bw_growth_df, by = c("focal_ID" = "treeID"))
+predictions_shuffle %>%
+  rmse(truth = growth, estimate = growth_hat) %>%
+  pull(.estimate)
 ```

### 2020-09-17 README.Rmd
Small change to SCBI section getting rid of get_model_specs
Some changes so it works for SCBI

@@
   geom_sf(data = scbi_growth_df, aes(col=factor(foldID)), size = 0.1)
+
+scbi_cv_grid_sf <- scbi_cv_grid$blocks %>%
+  st_as_sf()
 ```
@@
 # SCBI
-scbi_specs <- scbi_growth_df %>% 
-  get_model_specs(model_number = 3, species_notion = 'sp')
-scbi_specs
-```
-
-```{r, eval = run_scbi}
-# SCBI
 tic()
 focal_vs_comp_scbi <- scbi_growth_df %>% 
-  create_focal_vs_comp(max_dist, model_specs = scbi_specs, cv_grid = scbi_cv_grid, id = "stemID")
+  create_focal_vs_comp(max_dist, cv_grid_sf = scbi_cv_grid_sf, id = "stemID")
 toc()
@@
 tic()
-scbi_fit_model <- focal_vs_comp_scbi %>% 
-  fit_bayesian_model(model_specs = scbi_specs)
+posterior_param_scbi <- focal_vs_comp_scbi %>% 
+  fit_bayesian_model(prior_param = NULL, run_shuffle = TRUE)
 toc()

### 2020-09-18 README.Rmd
All done through CV

@@
 scbi_growth_df <- focal_vs_comp_scbi %>% 
-  predict_bayesian_model(model_specs = scbi_specs, posterior_param = scbi_fit_model) %>% 
+  predict_bayesian_model(posterior_param = posterior_param_scbi) %>% 
   right_join(scbi_growth_df, by = c("focal_ID" = "stemID"))
-
-scbi_growth_df <- scbi_growth_df %>% 
-  st_as_sf()
 
@@
 scbi_growth_df %>% 
+  st_as_sf() %>%
   # Need to investigate missingness

### 2020-09-20 README.Rmd
Working through the chagens for run_shuffle and run_cv

@@
 # Read in census data from 2008 & 2014
-data(bw_census_2008, bw_census_2014 )
+data(bw_census_2008, bw_census_2014, bw_species)
+
+# Append additional species data
+bw_census_2008 <- bw_census_2008 %>%
+  left_join(bw_species, by = "sp") %>%
+  select(-c(genus, species, latin))
 
@@
   compute_growth(bw_census_2008, bw_census_2014, id) %>%
-  mutate(sp = to_any_case(sp)) %>%
+  mutate(
+    sp = to_any_case(sp),
+    sp = as.factor(sp),
+    species = sp,
+    family = as.factor(family),
+    trait_group = as.factor(trait_group)) %>%
   # drop stemID

### 2020-09-25 README.Rmd
full run!

@@
 
-run_scbi <- TRUE
-run_bw <- FALSE
+run_scbi <- FALSE
+run_bw <- TRUE
 ```
@@
 
-**SCBI**
-
 For the above results we fit the model to the entire data set, and then make predictions across the entire data set from that fit.
@@
 Then applies that fit to the focal fold.
-It is a wrapper for `fit_bayesain_model` and `predict_bayesain_model` but fits a seperate model for each fold.
+It is a wrapper for `fit_bayesain_model` and `predict_bayesain_model` but fits a separate model for each fold.
 
-This will fit the model for each fold.
-On each fold it fits the data for all trees outside of that fold.
-If you have N folds then `run_cv` will take N-times longer than `fit_bayesain_model`.
-Here I cheated `run_cv` to fit for just `test = fold 23` and `test = fold 2`.
-It still fits to all other folds for those two, but only does it twice (rather than 28 times for the SCBI data set).
-Do this with the arguement `all_folds = FALSE`.
+```{r, eval = run_bw}
+# big woods
+tic()
+cv_bw <- focal_vs_comp_bw %>%
+  run_cv(max_dist = max_dist, cv_grid = bw_cv_grid) %>%
+  right_join(bw_growth_df, by = c("focal_ID" = "treeID"))
+toc()
+
+cv_bw %>%
+  rmse(truth = growth, estimate = growth_hat) %>%
+  pull(.estimate)
+```
 
@@
 tic()
-scbi_cv_predict <- focal_vs_comp_scbi %>%
-  run_cv(model_specs = scbi_specs, max_dist = max_dist, cv_grid = scbi_cv_grid, all_folds = FALSE)
+cv_scbi <- focal_vs_comp_scbi %>%
+  run_cv(max_dist = max_dist, cv_grid = scbi_cv_grid) %>%
+  right_join(scbi_growth_df, by = c("focal_ID" = "treeID"))
 toc()
-```
 
-Running just two folds took 2119 seconds. There are 28 folds. So running everything should take 2119 * 14 / 60 / 60 = 8 hours.
-
-Then we can compare the results to show that the RMSE for the cross-validated fit is larger than for the none CV fit above.
-
-```{r, eval = run_scbi}
-# SCBI
-scbi_cv_predict %>%
-  inner_join(scbi_growth_df, by = 'focal_ID', suffix = c('_cv','')) %>%
-  summarise(
-    rmse_cv = sqrt(mean((growth - growth_hat_cv)^2)),
-    rmse = sqrt(mean((growth - growth_hat)^2)), 
-    n = n() 
-  )
-```
-
-**Big Woods**
-
-Big woods is faster because it has fewer trees, but also because we are fitting the model for trait groups (6 groups) rather than species for SCBI (40 species).
-The lambda matrix is much smaller (6 x 6 versus 40 x 40).
-Means it fits much faster.
-
-```{r, eval = run_bw}
-# Big Woods
-if (!file.exists("bw_cv_predict.Rdata")) {
-tic()
-bw_cv_predict <- focal_vs_comp_bw %>%
-  run_cv(model_specs = bw_specs, max_dist = max_dist, cv_grid = bw_cv_grid)
-toc()
-  save(bw_cv_predict, file = "bw_cv_predict.Rdata")
-} else {
-  load("bw_cv_predict.Rdata")
-}
-```
-
-Big Woods must faster because it is a smaller plot? Mabye also because we are fitting the trait group version (6 spp versus 40 for SCBI).
-
-```{r, eval = run_bw}
-# Big Woods
-bw_growth_df <- bw_growth_df %>% 
-  left_join(bw_cv_predict, by = "focal_ID", suffix = c('', '_cv')) %>% 
-  mutate(has_cv = !is.na(growth_hat))
-
-ggplot(bw_growth_df) +
-  geom_sf(size = 0.1, aes(col = has_cv))
-```
-
-It appears not all trees have a CV predicted value of `growth_hat`. This is because either:
-
-* The tree is part of the buffer
-* The tree was dropped in the `create_focal_vs_comp()` phase
-
-```{r, eval = run_bw}
-# This code is a hideous sanity check of the two facts above. To delete.
-trees_dropped_during_create_focal_vs_comp <- 
-  setdiff(unique(bw_growth_df$focal_ID), unique(focal_vs_comp_bw$focal_ID)) %>% 
-  sort()
-trees_with_no_cv <- bw_growth_df %>% 
-  filter(!has_cv) %>% 
-  pull(focal_ID) %>% 
-  sort()
-mean(trees_dropped_during_create_focal_vs_comp == trees_with_no_cv)
-```
-
-Let's remove these trees that have no
-
-```{r, eval = run_bw}
-bw_growth_df <- bw_growth_df %>% 
-  filter(!is.na(growth_hat))
-```
-
-Let's compute summary statistics of our cross-validated errors:
-
-```{r, eval = run_bw}
-# Big Woods
-bw_growth_df %>% 
-  as_tibble() %>% 
-  summarise(
-    rmse_cv = sqrt(mean((growth - growth_hat)^2)),
-    rmse = sqrt(mean((growth - growth_hat)^2)), 
-    n = n() 
-  )
-```
-
-Let's visualize the spatial distribution of our cross-validated errors:
-
-```{r, eval = run_bw}
-# Big Woods
-bw_growth_df <- bw_growth_df %>%
-  mutate(
-    error = growth - growth_hat,
-    error_bin = cut_number(error, n = 5),
-    error_compress = ifelse(error < -0.75, -0.75, ifelse(error > 0.75, 0.75, error))
-  )
-
-ggplot(bw_growth_df) +
-  geom_sf(aes(col = error_compress), size = 0.4) +
-  scale_color_gradient2(
-    low = "#ef8a62", mid = "#f7f7f7", high = "#67a9cf", 
-    name = expression(paste("Residual (cm ", y^{-1}, ")")),
-    breaks = seq(from = -0.75, to = 0.75, by = 0.25),
-    labels = c("< -0.75", "-0.5", "0.25", "0", "0.25", "0.5", "> 0.75")
-  ) +
-  labs(x = "Meter", y = "Meter")
-```
-
-### Run permutations
-
-Hey Dave! Start here!
-
-**Repeat earlier code**:
-
-```{r, eval = FALSE}
-# Load packages ----------------------------------------------------------------
-library(tidyverse)
-library(forestecology)
-library(snakecase)
-library(skimr)
-library(sf)
-library(sfheaders)
-# devtools::install_github("rvalavi/blockCV")
-library(blockCV)
-library(tictoc)
-library(yardstick)
-library(viridis)
-
-# Load & preprocess data -------------------------------------------------------
-# Read in census data from 2008 & 2014
-bw_2008 <- 
-  "https://deepblue.lib.umich.edu/data/downloads/z603qx485" %>% 
-  read_delim(delim = "\t") %>% 
-  mutate(spcode = to_any_case(spcode)) %>%
-  select(
-    treeID = treeid, stemID = stemtag, sp = spcode, quadrat, gx, gy, dbh, 
-    date, codes
-  )
-bw_2014 <- 
-  "https://deepblue.lib.umich.edu/data/downloads/1831ck00f" %>% 
-  read_delim(delim = "\t") %>% 
-  mutate(spcode = to_any_case(spcode)) %>%
-  select(
-    treeID = treeid, stemID = stemtag, sp = spcode, quadrat, gx, gy, dbh, 
-    date, codes
-  )
-
-# Read in grouping classification data
-bw_species <- 
-  "https://deepblue.lib.umich.edu/data/downloads/000000086" %>% 
-  read_delim(delim = "\t") %>% 
-  # convert all to snake case:
-  mutate_at(c("species", "genus", "family", "idlevel", "spcode"), to_any_case) %>% 
-  # join trait group
-  left_join(families, by = c("spcode", "family")) %>% 
-  mutate(
-    sp = str_sub(genus, 1, 2), 
-    sp = str_c(sp, str_sub(species, 1, 2)),
-    sp = tolower(sp),
-    latin = str_c(genus, species, sep = " "),
-    latin = to_any_case(latin)
-  ) %>% 
-  select(sp = spcode, genus, species, latin, family, trait_group)
-
-bw_2008 <- bw_2008 %>%
-  left_join(bw_species,by='sp')
-
-# Compute growth ---------------------------------------------------------------
-census_df1 <- bw_2008
-# we need to filter out the resprouts
-census_df2 <- bw_2014 %>% 
-  filter(!str_detect(codes, 'R'))
-id <- "treeID"
-
-bw_growth_df <- 
-  # Merge both censuses and compute growth:
-  compute_growth(census_df1, census_df2, id) %>% 
-  mutate(sp = to_any_case(sp))
-
-# Define buffers ---------------------------------------------------------------
-cv_fold_size <- 100
-max_dist <- 7.5
-
-# Study region boundary polygon
-bw_boundary <- bigwoods_study_region %>% 
-  sf_polygon()
-
-# Buffer polygon
-bw_buffer <- bw_boundary %>%
-  st_buffer(dist = -max_dist)
-
-# Convert data frame to sf object
-bw_growth_df <- bw_growth_df %>% 
-  st_as_sf(coords = c("gx", "gy"))
-
-# ID which points are in buffer and which are not
-buffer_index <- !st_intersects(bw_growth_df, bw_buffer, sparse = FALSE)
-bw_growth_df <- bw_growth_df %>% 
-  mutate(buffer = as.vector(buffer_index))
-
-# Define spatial CV folds ------------------------------------------------------
-set.seed(76)
-bw_cv_grid <- spatialBlock(
-  speciesData = bw_growth_df, theRange = 100, verbose = FALSE,
-  # Some guess work in figuring this out:
-  k = 28, xOffset = 0.5, yOffset = 0
-)
-
-# Add foldID to data
-bw_growth_df <- bw_growth_df %>% 
-  mutate(
-    foldID = bw_cv_grid$foldID
-  )
-
-# Visualize grid. Why does fold 19 repeat?
-bw_cv_grid$plots +
-  geom_sf(data = bw_growth_df %>% sample_frac(0.2), aes(col=factor(foldID)), size = 0.1)
-
-# Remove weird folds with no trees in them from viz above
-bw_growth_df <- bw_growth_df %>%
-  filter(!foldID %in% c(19, 23, 21, 17, 8, 19))
-
-# Save original
-bw_growth_df_orig <- bw_growth_df
-```
-
-**Loop through 4 scenarios of model fitting**:
-
-```{r, eval = FALSE}
-# Number of permutation shuffles:
-num_shuffle <- 4
-
-# Compute observed RMSE for all models, but only do permutation shuffling for
-# models 3,6,9
-model_numbers <- c(3)
-species_notion_vector <- c("trait_group", "family", "species")
-
-# Save results here
-run_time <- rep(0, length(species_notion_vector))
-observed_RMSE <- rep(0, length(species_notion_vector))
-observed_RMSE_CV <- rep(0, length(species_notion_vector))
-shuffle_RMSE <- vector("list", length(species_notion_vector))
-shuffle_RMSE_CV <- vector("list", length(species_notion_vector))
-filename <- str_c(format(Sys.time(), "%Y-%m-%d"), "_model_comp_tbl_", num_shuffle, "_shuffles.RData")
-
-for(i in 1:length(species_notion_vector)){
-  # Start clock
-  tic()
-
-  # Modeling and species stuff
-  species_notion <- species_notion_vector[i]
-  bw_specs <- bw_growth_df %>% 
-    get_model_specs(model_number = 3, species_notion = species_notion)
-
-  # Focal vs comp main dataframe for analysis
-  focal_vs_comp_bw <- bw_growth_df_orig %>% 
-    create_focal_vs_comp(max_dist, model_specs = bw_specs, cv_grid = bw_cv_grid, id = "treeID")
-
-  # 1. Compute observed test statistic: RMSE with no cross-validation ----
-  # Fit model (compute posterior parameters)
-  bw_fit_model <- focal_vs_comp_bw %>% 
-    fit_bayesian_model(model_specs = bw_specs)
-
-  # Make predictions, compute and save RMSE, and reset
-  observed_RMSE[i] <- focal_vs_comp_bw %>% 
-    predict_bayesian_model(model_specs = bw_specs, posterior_param = bw_fit_model) %>% 
-    right_join(bw_growth_df_orig, by = c("focal_ID" = "treeID")) %>%
+cv_scbi %>%
   rmse(truth = growth, estimate = growth_hat) %>%
   pull(.estimate)
-
-  # 2. Compute observed test statistic: RMSE with cross-validation ----
-  observed_RMSE_CV[i] <- focal_vs_comp_bw %>%
-    run_cv(model_specs = bw_specs, max_dist = max_dist, cv_grid = bw_cv_grid) %>% 
-    right_join(bw_growth_df_orig, by = c("focal_ID" = "treeID")) %>%
-    rmse(truth = growth, estimate = growth_hat) %>%
-    pull(.estimate)
-
-  # 3. Permutation distribution: RMSE with no cross-validation ----
-  # Only do permutation shuffling for models 3, 6, 9
-  # Compute num_shuffle permutation test statistics
-  shuffle_RMSE[[i]] <- numeric(length = num_shuffle)
-
-  for(j in 1:num_shuffle){
-    # Fit model (compute posterior parameters) with shuffling
-    bw_fit_model_shuffle <- focal_vs_comp_bw %>% 
-      fit_bayesian_model(model_specs = bw_specs, run_shuffle = TRUE)
-
-    # Make predictions, compute and save RMSE, and reset
-    shuffle_RMSE[[i]][j] <- focal_vs_comp_bw %>% 
-      predict_bayesian_model(model_specs = bw_specs, posterior_param = bw_fit_model_shuffle) %>% 
-      right_join(bw_growth_df_orig, by = c("focal_ID" = "treeID")) %>%
-      rmse(truth = growth, estimate = growth_hat) %>%
-      pull(.estimate)
-  }
-
-  # 4. Permutation distribution: RMSE with cross-validation ----
-  # Compute num_shuffle permutation test statistics
-  shuffle_RMSE_CV[[i]] <- numeric(length = num_shuffle)
-
-  # Compute num_shuffle permutation test statistics
-  for(j in 1:num_shuffle){
-    # Compute and save RMSE, and reset
-    shuffle_RMSE_CV[[i]][j] <- focal_vs_comp_bw %>%
-      run_cv(model_specs = bw_specs, max_dist = max_dist, cv_grid = bw_cv_grid, run_shuffle = TRUE) %>% 
-      right_join(bw_growth_df_orig, by = c("focal_ID" = "treeID")) %>%
-      rmse(truth = growth, estimate = growth_hat) %>%
-      pull(.estimate)
-
-    # Status update
-    str_c("notion of species: ", species_notion_vector[i], ", shuffle with permutation ", j) %>% print()
-  }
-
-  # 5. Stop clock
-  clock <- toc(quiet = TRUE)
-  run_time[i] <- clock$toc - clock$tic
-
-  # 6. Save
-  model_comp_tbl <- tibble(
-    species_notion = species_notion_vector,
-    run_time = run_time,
-    observed_RMSE = observed_RMSE,
-    observed_RMSE_CV = observed_RMSE_CV,
-    shuffle_RMSE = shuffle_RMSE,
-    shuffle_RMSE_CV = shuffle_RMSE_CV,
-  )
-  save(model_comp_tbl, file = filename)
-}
-```
-
-**Plot results**:
-
-```{r, eval=TRUE}
-load("2020-06-25_model_comp_tbl_49_shuffles.RData")
-model_comp <- bind_rows(
-  model_comp_tbl %>% select(species_notion, run_time, observed = observed_RMSE, shuffle = shuffle_RMSE) %>% mutate(CV = FALSE),
-  model_comp_tbl %>% select(species_notion, run_time, observed = observed_RMSE_CV, shuffle = shuffle_RMSE_CV) %>% mutate(CV = TRUE)
-) %>%
-  gather(type, RMSE, -c(species_notion, run_time, CV)) %>%
-  mutate(
-    species_notion = case_when(
-      species_notion == "trait_group" ~ "1. Trait-based (6): lambda = 6 x 6",
-      species_notion == "family" ~ "2. Phylogenetic family (20): lambda = 20 x 20",
-      species_notion == "species" ~ "3. Actual species (36): lambda = 36 x 36"
-    )
-  ) %>% 
-  filter(species_notion != "3. Actual species (36): lambda = 36 x 36")
-
-model_comp_observed <- model_comp %>%
-  filter(type == "observed") %>%
-  unnest(cols = c(RMSE))
-model_comp_shuffle <- model_comp %>%
-  filter(type == "shuffle") %>%
-  unnest(cols = c(RMSE))
-
-xlab <- expression(paste('RMSE (cm ',y^{-1},')'))
-
-ggplot() +
-  geom_vline(data = model_comp_observed, aes(xintercept = RMSE, col = CV), linetype = "dashed", show.legend = F) +
-  geom_histogram(data = model_comp_shuffle, aes(x = RMSE, fill = CV), bins = 200) +
-  labs(fill = "Cross-validated?", x = xlab) +
-  facet_wrap(~species_notion, ncol = 1) +
-  scale_color_viridis(discrete = TRUE, option = "D")+
-  scale_fill_viridis(discrete = TRUE)
 ```

### 2020-09-30 README.Rmd
Some littel changes

@@
 
-run_scbi <- FALSE
+run_scbi <- TRUE
 run_bw <- TRUE
@@
 
+The RMSE is lower for the non-shuffled version. This gives support for the idea that competitor identiy does matter for competitive interactions.
+
 **SCBI**:
@@
 
+As expected this RMSE is higher than that when the model is fit without cross validation. See Allen and Kim (2020) for more discussion of this.
+
 ```{r, eval = run_scbi}

### 2020-10-02 README.Rmd
adding plot code

@@
 The betas tell us about how fast each species grows and how this depends on DBH.
-The lambdas, which are often of more interest, are the species-specific competition coefficents.
-The full lambda matrix gives competition strength between species.
-There is a rich literature how this matrxi (cite).
+The lambdas, which are often of more interest, are the species-specific competition coefficents The full lambda matrix gives competition strength between species.
+There is a rich literature how this matrix (cite).
 
-Because of the strucutre of the `bw_fit_model` object we cannot simply draw these curves based on the posterior distribution.
+Because of the structure of the `bw_fit_model` object we cannot simply draw these curves based on the posterior distribution.
 `bw_fit_model` gives the parameteres *compared* to a baseline.
@@
 
+Here we re-run the Big Woods model but using the family as the group for comparison.
+This makes the posterior distributions easier to follow.
+Also, surprisingly, grouping by family performed just as well as grouping by species (see Allen and Kim 2020).
+
 ```{r, eval = run_bw}
 
-# plot_beta0(bw_fit_model)
+focal_vs_comp_bw <- bw_growth_df %>%
+  # mutate(sp = trait_group) %>%
+  mutate(sp = family) %>%
+  create_focal_vs_comp(max_dist = max_dist, cv_grid_sf = bw_cv_grid_sf, id = "treeID")
 
+# a) Fit model (compute posterior parameters) with no permutation shuffling
+posterior_param_bw <- focal_vs_comp_bw %>%
+  fit_bayesian_model(prior_param = NULL, run_shuffle = FALSE)
+
+# b) Recreate Fig5 from Allen (2020): Posterior distributions of selected lambdas
+posterior_plots <- plot_posterior_parameters(
+  posterior_param = posterior_param_bw,
+  sp_to_plot = c("cornaceae", "fagaceae", "hamamelidaceae", "juglandaceae", "lauraceae", "rosaceae", "sapindaceae", "ulmaceae")
+)
+
+posterior_plots[["beta_0"]]
+posterior_plots[["beta_dbh"]]
+posterior_plots[["lambda"]]
 ```

### 2020-10-02 README.Rmd
send readme edits

@@
 
-run_scbi <- TRUE
+run_scbi <- FALSE
 run_bw <- TRUE
@@
 Also, surprisingly, grouping by family performed just as well as grouping by species (see Allen and Kim 2020).
+First we re-run `create_focal_vs_comp` and `fit_bayesian_model` with the grouping variable as family.
 
@@
 
+```
+
+Now the output of `fit_bayesian_model` is passed to `plot_posterior_parameters`.
+
+```{r, eval = run_bw}
 # b) Recreate Fig5 from Allen (2020): Posterior distributions of selected lambdas
@@
 )
+```
 
+The output is a list with three plots stored. The element `beta_0` gives the growth intercept, i.e., how fast an individual of each group grows independent of DBH).
+
+```{r, eval = run_bw}
 posterior_plots[["beta_0"]]
+```
+
+Next `beta_dbh` gives the DBH-growth slope for each group.
+
+```{r, eval = run_bw}
 posterior_plots[["beta_dbh"]]
+```
+
+Finally `lambda` gives the competition coeffiencts.
+
+```{r, eval = run_bw}
 posterior_plots[["lambda"]]

### 2020-10-02 README.Rmd
Works for SCBI!

@@
 
+For SCBI we will just look at a 16 ha plot to speed calculations.
+
 ```{r, eval = run_scbi}
@@
   ) %>%
-  mutate(date = mdy(date))
+  mutate(date = mdy(date)) %>%
+  filter(gy <= 400)
 
@@
   mutate(dbh = as.numeric(dbh),
-         date = mdy(date)) 
+         date = mdy(date)) %>%
+  filter(gy <= 400)
 ```
@@
   # they are mesuaring in mm while we are measuring in cm!!!
-  mutate(growth = growth/10)
+  mutate(growth = growth/10,
+             sp = as.factor(sp))
 
@@
 # Study region boundary polygon
-scbi_study_region <- tibble(x = c(0,400,400,0,0), y = c(0,0,640,640,0)) %>% 
+scbi_study_region <- 
+ #tibble(x = c(0,400,400,0,0), y = c(0,0,640,640,0)) %>% 
+  tibble(x = c(0,400,400,0,0), y = c(0,0,400,400,0)) %>% 
   sf_polygon()
@@
 scbi_cv_grid <- spatialBlock(
-  speciesData = scbi_growth_df, theRange = 100, k = 28, yOffset = 0.9999, verbose = FALSE
+  speciesData = scbi_growth_df, theRange = 100, k = 20, yOffset = 0.9999, verbose = FALSE
 )
@@
 posterior_param_scbi <- focal_vs_comp_scbi %>% 
-  fit_bayesian_model(prior_param = NULL, run_shuffle = TRUE)
+  fit_bayesian_model(prior_param = NULL, run_shuffle = FALSE)
 toc()
@@
 
-```{r, eval = run_scbi}
+```{r, eval = FALSE}
 # SCBI
@@
 ```
+
+** SCBI
+
+```{r, eval = run_bw}
+posterior_plots_scbi <- plot_posterior_parameters(
+  posterior_param = posterior_param_scbi,
+  sp_to_plot = c("libe", "caca", "astr", "litu", "havi", "cagl", "cato")
+)
+```
+
+```{r, eval = run_bw}
+posterior_plots_scbi[["lambda"]]
+```

### 2020-10-02 README.Rmd
Opps little edit

@@
 
-```{r, eval = run_bw}
+```{r, eval = run_scbi}
 posterior_plots_scbi <- plot_posterior_parameters(
@@
 
-```{r, eval = run_bw}
+```{r, eval = run_scbi}
 posterior_plots_scbi[["lambda"]]

### 2020-10-06 README.Rmd
Readme knit with SCBI
Small edits for clarity

@@
 
-* Plot posterior distributions of all parameters
+**Big Woods**
 
@@
 This is often not of interest.
-So to display these parameters as we care about them we have to sample from the baseline distrubiton and from the comparison one to get the posterior distribution of interest.
+So to display these parameters as we care about them we have to sample from the baseline distribution and from the comparison one to get the posterior distribution of interest.
 
@@
 
-** SCBI
+**SCBI**
 
@@
   posterior_param = posterior_param_scbi,
-  sp_to_plot = c("libe", "caca", "astr", "litu", "havi", "cagl", "cato")
+  sp_to_plot = c("caca", "litu", "cagl", "cato")
 )

### 2020-10-06 README.Rmd
Change species for SCBI example

@@
   posterior_param = posterior_param_scbi,
-  sp_to_plot = c("caca", "litu", "cagl", "cato")
+  sp_to_plot = c("quru", "litu", "cagl", "cato")
 )
@@
 posterior_plots_scbi[["lambda"]]
+#comment to change
 ```

### 2020-10-10 README.Rmd
everything runs!

@@
 
-For SCBI we will just look at a 16 ha plot to speed calculations.
+For SCBI we will just look at a 9 ha plot to speed calculations.
 
@@
   mutate(date = mdy(date)) %>%
-  filter(gy <= 400)
+  filter(gx < 300, gy > 300, gy < 600)
 
@@
          date = mdy(date))  %>%
-  filter(gy <= 400)
+  filter(gx < 300, gy > 300, gy < 600)
 ```
@@
  #tibble(x = c(0,400,400,0,0), y = c(0,0,640,640,0)) %>% 
-  tibble(x = c(0,400,400,0,0), y = c(0,0,400,400,0)) %>% 
+  tibble(x = c(0,300,300,0,0), y = c(300,300,600,600,3)) %>% 
   sf_polygon()
@@
 scbi_cv_grid <- spatialBlock(
-  speciesData = scbi_growth_df, theRange = 100, k = 20, yOffset = 0.9999, verbose = FALSE
+  speciesData = scbi_growth_df, theRange = 100, yOffset = 0.9999, k = 9, verbose = FALSE
 )
@@
 # SCBI
-scbi_growth_df <- focal_vs_comp_scbi %>% 
+scbi_growth_df_noCV <- focal_vs_comp_scbi %>% 
   predict_bayesian_model(posterior_param = posterior_param_scbi) %>% 
@@
 # Observed vs predicted growth  
-ggplot(scbi_growth_df, aes(x = growth, y = growth_hat)) +
+ggplot(scbi_growth_df_noCV, aes(x = growth, y = growth_hat)) +
   geom_point(size = 0.5, color = rgb(0, 0, 0, 0.25)) +
@@
 reslab <- expression(paste('Residual (cm ',y^{-1},')'))
-scbi_growth_df %>% 
+scbi_growth_df_noCV %>% 
   st_as_sf() %>%
@@
   ggplot() + 
-  geom_sf(aes(col = error_compress), size = 0.4) + 
+  geom_sf(aes(col = error_compress), size = 1) + 
   theme_bw() + 
@@
 
-```{r, eval = FALSE}
+```{r, eval = run_scbi}
 # SCBI
@@
 posterior_plots_scbi[["lambda"]]
-#comment to change
 ```

### 2020-10-16 README.Rmd
RUns great on SCBI!
I feel good about the readme now. Did some editing of the text. Looks great to me!

@@
 
-We present an example analysis using:
+We present an example analysis using data from two [ForestGEO](https://forestgeo.si.edu/) plots:
 
-* [Michigan Big Woods](https://doi.org/10.7302/wx55-kt18) research plot data
-* [Smithsonian Conservation Biology Institute (SCBI) ForestGEO](https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/tree/master/tree_main_census) research plot data
+* [Michigan Big Woods](https://doi.org/10.7302/wx55-kt18) research plot data. For more information see [here](http://hdl.handle.net/2027.42/156251).
+* [Smithsonian Conservation Biology Institute (SCBI) ForestGEO](https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/tree/master/tree_main_census) research plot data.
+For more information see [here](https://doi.org/10.1890/13-0010.1).
 
@@
 
-The data for the Big Woods plor are included in the package.
+The data for the Big Woods plot are included in the package.
 For SCBI, we load and preprocess the data corresponding to the two census reflecting the time period over which we consider growth in dbh.
@@
 
-For SCBI we will just look at a 9 ha plot to speed calculations.
+For SCBI we will look at just a 9 ha section of the plot to speed calculations.
 
@@
 
-We then combine the two census data frames into a single data frame  that now includes a numerical variable `growth` reflecting the average annual growth of dbh in cm.
+For each plot we combine the two census data frames into a single data frame  that now includes a numerical variable `growth` reflecting the average annual growth of dbh in cm.
 Furthermore, variables that (in theory) remain unchanged between censuses appear only once, such as `gx`, `gy`, species-related variables.
@@
 
-# we need to filter out the resprouts
+# we need to filter out the re-sprouts
 bw_census_2014 <- bw_census_2014 %>% 
@@
   compute_growth(census_df1, census_df2, id) %>%
-  # they are mesuaring in mm while we are measuring in cm!!!
+  # Convert growth from cm to mm to make result comperable
   mutate(growth = growth/10,
@@
 In order to control for study region "edge effects" (cite Waller?), we need a function that adds "buffers" of trees.
-In our case, since our model of interspecies competition relies on a spatial definition of who is a "neighboring competitor" and certain explanatory varibles such as biomass are cummulative, we need to ensure that all trees being modeled are not biased to have different neighbor structure, in particular the trees at the boundary of study regions.
+In our case, since our model of interspecific competition relies on a spatial definition of who is a "neighboring competitor" and certain explanatory variables such as biomass are cumulative, we need to ensure that all trees being modeled are not biased to have different neighbor structure, in particular the trees at the boundary of study regions.
 1.
@@
 
-Thus, before we can incorporate the above information to `growth_df`, we need to define two constants:
+Thus, before we can incorporate the above information to `growth_df`, we need to define two constants.
 
@@
 
+The first, `cv_fold_size` is the length of the folds, each fold is a square.
+The next, `max_dist` is the maximum distance for a tree's competitive neighborhood.
+Trees within this distance of each other are assumed to compete while those farther than this distance apart do not.
+Both these values are in the same units as `gx` and `gz`, most often this will be meters.
+
+<!--
+DA comment: I am not sure about the below so I took it out.
+We don't assume that competitive interactions are inversely related to distance.
+We assume that competitive interactions are equivalent between distance 0 and 7.5, but than zero after that.
+Canham has the inversely related structure but we don't.
+
+Following Tobler's first law of geography that "everything is related to everything else, but near things are more related than distant things." @tobler1970firstlawofgeo, we assume that the degree of spatial autocorrelation is inversely-related to distance.
+However, we further assume that once trees are more than `r max_dist` meters apart, this autocorrelation is negligeable.
+-->
+
+Other studies have estimated the value of `max_dist`; we use an average of estimated values [@canham2004, @canham2006, @uriarte2004, @tatsumi2013].
+
 #### Defining buffers
 
-For a focal tree of interest, the `max_dist` of `r max_dist` meters acts as a radius defining a neighborhood within which all trees are considered competitors.
-In other words, all trees within `r max_dist` meters of the focal tree are considered competitor trees.
-Other studies have estimated this distance; we used `r max_dist` meters as an average of estimated values [@canham2004, @canham2006, @uriarte2004, @tatsumi2013].
-Following Tobler's first law of geography that "everything is related to everything else, but near things are more related than distant things." @tobler1970firstlawofgeo, we assume that the degree of spatial autocorrelation is inversely-related to distance.
-However, we further assume that once trees are more than `r max_dist` meters apart, this autocorrelation is negligeable.
-
-We use the function `add_buffer_variable` to identifiy which trees are inside of a buffer region of size `max_dist`. The user will need to specify a study region and convert to a `sf_polygon`.
+We use the function `add_buffer_variable` to identify which trees are inside of a buffer region of size `max_dist`. The user will need to specify a study region and convert to a `sf_polygon`.
 
@@
 
-We use the [`blockCV`](https://github.com/rvalavi/blockCV) package to define the spatial grid, whose elements will act as the folds in our leave-one-out (by "one" we meen "one grid block") cross-validation scheme.
+We use the [`blockCV`](https://github.com/rvalavi/blockCV) package to define the spatial grid, whose elements will act as the folds in our leave-one-out (by "one" we mean "one grid block") cross-validation scheme.
+The upshot here is we add `foldID` to `growth_df` which identifies which fold each individual is in, and the creation of a `cv_grid_sf` object which gives the geometry of the cross validation grid.
 
@@
 
-This requires the `growth_df` data frame; `max_dist`, the scalar defining competitive range; `cv_fold_size`, defining the size of the spatial cross-validation blocks; and the `id` variable as inputs.
+This requires the `growth_df` data frame; `max_dist`, the scalar defining competitive range; `cv_grid_sf`, giving the cross validation grid; and the `id` variable as inputs.
 
@@
 This function needs only the `focal_vs_comp` as an input.
-Other options allow the user to specify prior parameters and run a species identity shuffle (see below).
+Other options allow the user to specify prior parameters and run a competitor-species identity shuffle (see below).
 
@@
 This `posterior_param` output can be used to get predicted growths for each individual (with `predict_bayesain_model`) to test how well the model performs.
-Or this `posterior_param` output can be plots (either the betas or lambdas) to understand what controls individual growth.
+Or this `posterior_param` output can be plotted (either the betas or lambdas with `plot_posterior_params`) to understand what controls growth and the nature of the competitive ineractions.
 
@@
 We do this by shuffling the identity of competitors (but not of focal trees or spatial locations or sizes) and fitting the model again.
-We then compare RMSEs to see whether competitor identity matters to competitive effects
+We then compare RMSEs to see whether competitor identity matters to competitive interactions.
 
@@
 
-The RMSE is lower for the non-shuffled version. This gives support for the idea that competitor identiy does matter for competitive interactions.
+The RMSE is lower for the non-shuffled version. This gives support for the idea that competitor identity does matter for competitive interactions.
 
@@
 
+Here we provide examples of seeing how well the model fits and then a spatial plot of the residuals to see whether there are any consistent spatial patterns to them.
+
 ```{r, eval = run_scbi}
@@
   right_join(scbi_growth_df, by = c("focal_ID" = "stemID"))
+
+scbi_growth_df_noCV %>%
+  rmse(truth = growth, estimate = growth_hat) %>%
+  pull(.estimate)
 
@@
 We can use the spatial block structure we defined above to deal with with.
-The function `run_cv` goes through each fold in the `cv_grid` and fits the model on all the other folds.
+The function `run_cv` goes through each fold in the `cv_grid_sf` and fits the model on all the other folds.
 Then applies that fit to the focal fold.
@@
 
+Here also this RMSE is much higher than that for the above SCBI model fit without cross validation.
+
 ### Visualize posterior distributions
@@
 The betas tell us about how fast each species grows and how this depends on DBH.
-The lambdas, which are often of more interest, are the species-specific competition coefficents The full lambda matrix gives competition strength between species.
-There is a rich literature how this matrix (cite).
+The lambdas, often of more interest, are the species-specific competition coefficients.
+The full lambda matrix gives competition strength between species.
+There is a rich literature on this matrix (cite).
 
 Because of the structure of the `bw_fit_model` object we cannot simply draw these curves based on the posterior distribution.
-`bw_fit_model` gives the parameteres *compared* to a baseline.
-This is often not of interest.
-So to display these parameters as we care about them we have to sample from the baseline distribution and from the comparison one to get the posterior distribution of interest.
+`bw_fit_model` gives the parameters *compared* to a baseline.
+This is not of direct interest.
+So to display these parameters, as we care about them, we have to sample from the baseline distribution and from the comparison one to get the posterior distribution of interest.
 

### 2020-10-23 README.Rmd
little example
starting to add small example

@@
 
-run_scbi <- TRUE
-run_bw <- TRUE
+run_scbi <- FALSE
+run_bw <- FALSE
+small_example <- TRUE
 ```
@@
 bw_cv_grid <- spatialBlock(
-  speciesData = bw_growth_df, theRange = 100, verbose = FALSE,
+  speciesData = bw_growth_df, theRange = cv_fold_size, verbose = FALSE,
   # Some guess work in figuring this out:
@@
 scbi_cv_grid <- spatialBlock(
-  speciesData = scbi_growth_df, theRange = 100, yOffset = 0.9999, k = 9, verbose = FALSE
+  speciesData = scbi_growth_df, theRange = cv_fold_size, yOffset = 0.9999, k = 9, verbose = FALSE
 )
@@
 ```
+
+### Using small example
+
+```{r, eval = small_example}
+# Read in census files
+data(census_df1_ex, census_df2_ex)
+
+# Filter out resprouts
+census_df2_ex_no_r <- census_df2_ex %>%
+  filter(!str_detect(codes, 'R'))
+
+id <- 'ID'
+
+ex_growth_df <-
+  # Merge both censuses and compute growth:
+  compute_growth(census_df1_ex, census_df2_ex_no_r, id) %>%
+  mutate(
+    sp = to_any_case(sp),
+    sp = as.factor(sp)) %>%
+  # drop stemID
+  select(-ID)
+
+data("ex_study_region")
+
+# set max dist
+max_dist <- 1
+
+# add buffer
+ex_growth_df <- ex_growth_df %>%
+  add_buffer_variable(direction = "in", size = max_dist, region = ex_study_region)
+
+ggplot() +
+  geom_sf(data = ex_growth_df, aes(col = buffer), size = 2)
+
+set.seed(76)
+ex_cv_grid <- spatialBlock(
+  speciesData = ex_growth_df, theRange = 5, verbose = FALSE, k = 2
+)
+
+```
+
+bw_cv_grid <- spatialBlock(
+  speciesData = bw_growth_df, theRange = cv_fold_size, verbose = FALSE,
+  # Some guess work in figuring this out:
+  k = 28, xOffset = 0.5, yOffset = 0
+)
+
+# Add foldID to data
+bw_growth_df <- bw_growth_df %>% 
+  mutate(
+    foldID = bw_cv_grid$foldID
+  )
+
+# Visualize grid. Why does fold 19 repeat?
+bw_cv_grid$plots +
+  geom_sf(data = bw_growth_df %>% sample_frac(0.2), aes(col=factor(foldID)), size = 0.1)
+
+bw_growth_df <- bw_growth_df %>%
+  filter(!foldID %in% c(19, 23, 21, 17, 8, 19)) %>%
+  mutate(foldID = as.character(foldID))
+
+# Deliverable
+ggplot() +
+  geom_sf(data = bw_growth_df, aes(col = foldID, alpha = buffer))
+
+bw_cv_grid_sf <- bw_cv_grid$blocks %>%
+  st_as_sf()

### 2020-10-23 data-raw/process_data_raw.R

@@ -73,7 +73,39 @@ bw_census_2014 <-
   )
 use_data(bw_census_2014, overwrite = TRUE)
 
+# Small example for whole workflow
+census_df1_ex <- tibble(
+  ID = 1:10,
+  sp = rep(c('sugar maple', 'American beech'),5),
+  gx = c(0.75, 1.5, 1.75, 3, 3.25, 5.5, 8, 8.5, 8.75, 8.75),
+  gy = c(2.5, 2.5, 2.25, 1.5, 1.75, 4.5, 1.5, 0.75, 1.5, 1.75),
+  date = ymd('20150601'),
+  codes = 'M',
+  dbh = c(5, 20, 15, 12, 35, 6, 22, 14, 42, 4)
+)
+use_data(census_df1_ex, overwrite = TRUE)
 
+
+census_df2_ex <- tibble(
+  ID = c(1:9,11,12),
+  sp = c(rep(c('sugar maple', 'American beech'),4),'sugar maple','sugar maple','sugar maple'),
+  gx = c(0.75, 1.5, 1.75, 3, 3.25, 5.5, 8, 8.5, 8.75, 6.5, 2.5),
+  gy = c(2.5, 2.5, 2.25, 1.5, 1.75, 4.5, 1.5, 0.75, 1.5, 3, 4.5),
+  date = ymd('20200601'),
+  codes = c(rep('M',5),'R',rep('M',5)),
+  dbh = c(6, 24, 20, 14, 42, 2, 25, 19, 49, 2, 2)
+)
+use_data(census_df2_ex, overwrite = TRUE)
+
+ex_study_region <-
+  tibble(
+    # Study region boundary
+    x = c(0,10,10,0,0),
+    y = c(0,0,5,5,0)
+  ) %>%
+  # Convert to sf object
+  sf_polygon()
+use_data(ex_study_region, overwrite = TRUE)
 
 # Example growth_df data frame used to illustrate focal_vs_comp()
 growth_df_ex <- tibble(

### 2020-10-27 README.Rmd
working on small example

@@
 set.seed(76)
-ex_cv_grid <- spatialBlock(
-  speciesData = ex_growth_df, theRange = 5, verbose = FALSE, k = 2
-)
-
-```
-
-bw_cv_grid <- spatialBlock(
-  speciesData = bw_growth_df, theRange = cv_fold_size, verbose = FALSE,
-  # Some guess work in figuring this out:
-  k = 28, xOffset = 0.5, yOffset = 0
-)
+#ex_cv_grid <- spatialBlock(
+#  speciesData = ex_growth_df, theRange = 5, verbose = FALSE, k = 2
+#)
 
 # Add foldID to data
-bw_growth_df <- bw_growth_df %>% 
+ex_growth_df <- ex_growth_df %>% 
   mutate(
-    foldID = bw_cv_grid$foldID
+    foldID = ex_cv_grid$foldID
   )
@@
 # Visualize grid. Why does fold 19 repeat?
-bw_cv_grid$plots +
-  geom_sf(data = bw_growth_df %>% sample_frac(0.2), aes(col=factor(foldID)), size = 0.1)
+ex_cv_grid$plots +
+  geom_sf(data = ex_growth_df, aes(col=factor(foldID)), size = 1)
 
-bw_growth_df <- bw_growth_df %>%
-  filter(!foldID %in% c(19, 23, 21, 17, 8, 19)) %>%
-  mutate(foldID = as.character(foldID))
+ex_cv_grid <- ex_cv_grid$blocks %>%
+  st_as_sf()
 
-# Deliverable
-ggplot() +
-  geom_sf(data = bw_growth_df, aes(col = foldID, alpha = buffer))
-
-bw_cv_grid_sf <- bw_cv_grid$blocks %>%
-  st_as_sf()
+```

### 2020-10-28 README.Rmd
Working on small example and documentation

@@
 
+Need to write this up for the readme.
+
 ```{r, eval = small_example}
@@
     sp = to_any_case(sp),
-    sp = as.factor(sp)) %>%
-  # drop stemID
-  select(-ID)
+    sp = as.factor(sp))
 
@@
 
-set.seed(76)
-#ex_cv_grid <- spatialBlock(
-#  speciesData = ex_growth_df, theRange = 5, verbose = FALSE, k = 2
-#)
+# Bert's solution: Manually create a blocks sf object
+fold1 <- rbind(c(0, 0), c(5, 0), c(5, 5), c(0, 5), c(0, 0))
+fold2 <- rbind(c(5, 0), c(10, 0), c(10, 5), c(5, 5), c(5, 0))
+blocks <- bind_rows(
+  sf_polygon(fold1),
+  sf_polygon(fold2) ) %>%
+  mutate(foldID = c(1, 2))
+
+# Plot
+ggplot() +
+  geom_sf(data = ex_growth_df, aes(col = buffer), size = 2) +
+  geom_sf(data = blocks, fill = "transparent")
+
+# fit spatialBlock()
+ex_cv_grid <- spatialBlock(
+  speciesData = ex_growth_df,
+  verbose = FALSE,
+  k = 2,
+  # Note new arguments
+  selection = "systematic",
+  blocks = blocks#,
+ # foldsCol = "foldID"
+)
 
@@
 ex_growth_df <- ex_growth_df %>%
-  mutate(
-    foldID = ex_cv_grid$foldID
-  )
+  mutate(foldID = ex_cv_grid$foldID %>% as.factor())
 
-# Visualize grid. Why does fold 19 repeat?
+# Visualize grid
 ex_cv_grid$plots +
-  geom_sf(data = ex_growth_df, aes(col=factor(foldID)), size = 1)
+  geom_sf(data = ex_growth_df, aes(col = foldID), size = 2)
 
-ex_cv_grid <- ex_cv_grid$blocks %>%
+# Deliverable
+ggplot() +
+  geom_sf(data = ex_growth_df, aes(col = foldID, shape = buffer))
+
+ex_cv_grid_sf <- ex_cv_grid$blocks %>%
   st_as_sf()
 
+# focal v comp
+focal_vs_comp_ex <- ex_growth_df %>% 
+  create_focal_vs_comp(max_dist, cv_grid_sf = ex_cv_grid_sf, id = "ID")
+
+# fit the model
+posterior_param_ex <- focal_vs_comp_ex %>% 
+  fit_bayesian_model(prior_param = NULL, run_shuffle = FALSE)
+
+# get predictions
+predictions <- focal_vs_comp_ex %>%
+  predict_bayesian_model(posterior_param = posterior_param_ex) %>%
+  right_join(ex_growth_df, by = c("focal_ID" = "ID"))
+predictions %>%
+  rmse(truth = growth, estimate = growth_hat) %>%
+  pull(.estimate)
+
+# plot posteriors
+posterior_param_ex %>%
+  plot_posterior_parameters()
+
+# run model with CV
+ex_bw <- focal_vs_comp_ex %>% 
+  run_cv(max_dist = max_dist, cv_grid = ex_cv_grid) %>%
+  right_join(ex_growth_df, by = c("focal_ID" = "ID"))
+
+ex_bw %>%
+  rmse(truth = growth, estimate = growth_hat) %>%
+  pull(.estimate)
+
 ```

### 2020-10-28 R/datasets.R

@@ -168,6 +168,94 @@
 "bw_study_region"
 
 
+#' Example input census data for package use
+#'
+#' This is example forest census data to be analyzed with this package.
+#'
+#' @format A \code{tibble}
+#'\describe{
+#'   \item{ID}{Tree identification number. This identifies an individual tree and
+#'   can be used to connect trees between the two censuses.}
+#'   \item{sp}{Speices of the individual}
+#'   \item{gx}{x-coordinate meters from reference point}
+#'   \item{gy}{y-coordinate meters from reference point}
+#'   \item{date}{Date the stem was measured}
+#'   \item{codes}{Code for additional information on the stem: M means the main stem
+#'   of the individual tree and R means the stem was lost, but the tag was moved to another stem greater than DBH cutoff,
+#'   this stands for resprout.}
+#'   \item{dbh}{Diameter at breast hieght of the tree in cm}
+#' }
+#' @seealso \code{\link{census_df2_ex}}
+#' @examples
+#' data(census_df1_ex,census_df2_ex)
+#' require(dplyr)
+#' # Filter out resprouts
+#' census_df2_ex_no_r <- census_df2_ex %>%
+#'  filter(!str_detect(codes, 'R'))
+#' id <- 'ID'
+#' ex_growth_df <-
+#'  # Merge both censuses and compute growth:
+#'  compute_growth(census_df1_ex, census_df2_ex_no_r, id) %>%
+#'  mutate(
+#'    sp = to_any_case(sp),
+#'    sp = as.factor(sp))
+"census_df1_ex"
+
+
+#' Example input census data for package use
+#'
+#' This is an example second census to be analyzed with the package.
+#'
+#' @format A \code{tibble}
+#'\describe{
+#'   \item{ID}{Tree identification number. This identifies an individual tree and
+#'   can be used to connect trees between the two censuses.}
+#'   \item{sp}{Speices of the individual}
+#'   \item{gx}{x-coordinate meters from reference point}
+#'   \item{gy}{y-coordinate meters from reference point}
+#'   \item{date}{Date the stem was measured}
+#'   \item{codes}{Code for additional information on the stem: M means the main stem
+#'   of the individual tree and R means the stem was lost, but the tag was moved to another stem greater than DBH cutoff,
[…]

### 2020-10-28 spatial_block_issue.R

@@ -36,20 +36,19 @@ ex_growth_df <- ex_growth_df %>%
 ggplot() +
   geom_sf(data = ex_growth_df, aes(col = buffer), size = 2)
 
-ex_cv_grid <- spatialBlock(
-  speciesData = ex_growth_df, theRange = 5, verbose = FALSE, k = 2
-)
-
-
+#ex_cv_grid <- spatialBlock(
+#  speciesData = ex_growth_df, theRange = 5, verbose = FALSE, k = 2
+#)
 
 
 # Bert's solution: Manually create a blocks sf object
 fold1 <- rbind(c(0, 0), c(5, 0), c(5, 5), c(0, 5), c(0, 0))
 fold2 <- rbind(c(5, 0), c(10, 0), c(10, 5), c(5, 5), c(5, 0))
 blocks <- bind_rows(
-  sf_polygon(fold1), sf_polygon(fold2)
-) %>%
+  sf_polygon(fold1),
+  sf_polygon(fold2) ) %>%
   mutate(foldID = c(1, 2))
+
 
 # Plot
 ggplot() +
@@ -62,9 +61,9 @@ ex_cv_grid <- spatialBlock(
   verbose = FALSE,
   k = 2,
   # Note new arguments
-  selection = "predefined",
-  blocks = blocks,
-  foldsCol = "foldID"
+  selection = "systematic",
+  blocks = blocks#,
+ # foldsCol = "foldID"
 )
 
 # Add foldID to data
@@ -79,8 +78,3 @@ ex_cv_grid$plots +
 ggplot() +
   geom_sf(data = ex_growth_df, aes(col = foldID, shape = buffer))
 
-# No need for this since we already defined sf object blocks
-# ex_cv_grid_sf <- ex_cv_grid$blocks %>%
-#   st_as_sf()
-#
-# ex_cv_grid_sf

### 2020-11-13 README.Rmd
rebuilt package to update datasets and documentation. Also re-did readme to include only small example

@@
 
+This package is designed to work for spaitally mapped, repeat censused forests plots.
+The package has commands to fit models of tree growth based on neighborhood competition.
+This can be used to estimate species-specific competition coefficients.
+The models can be fit with a spatial cross validation scheme to avoid overfitting.
+Additionally these models can test whether the idenity of competitors matters, by shuffling competitor identity and testing whether model preformance changes.
+See Allen and Kim (2020) for a full descripition.
+
 ## Example analysis
-
-We present an example analysis using data from two [ForestGEO](https://forestgeo.si.edu/) plots:
-
-* [Michigan Big Woods](https://doi.org/10.7302/wx55-kt18) research plot data. For more information see [here](http://hdl.handle.net/2027.42/156251).
-* [Smithsonian Conservation Biology Institute (SCBI) ForestGEO](https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/tree/master/tree_main_census) research plot data.
-For more information see [here](https://doi.org/10.1890/13-0010.1).
 
@@
 library(forestecology)
-library(snakecase)
-library(skimr)
 library(sf)
 library(sfheaders)
-# devtools::install_github("rvalavi/blockCV")
 library(blockCV)
-library(tictoc)
 library(yardstick)
-library(viridis)
-library(lubridate)
-
-run_scbi <- FALSE
-run_bw <- FALSE
-small_example <- TRUE
 ```
 
-### Load & preprocess data
-
-The data for the Big Woods plot are included in the package.
-For SCBI, we load and preprocess the data corresponding to the two census reflecting the time period over which we consider growth in dbh.
-We load them into R as "tibble" data frames thereby ensuring a standardized input/output format that can be used across all `tidyverse` packages @tidyverse.
-Furthermore, we ensure that the different variables have the correct names, types (`dbl`, `data`, `factor`).
-
-**Big Woods**:
-
-```{r, eval = run_bw}
-# Big Woods
-# Read in census data from 2008 & 2014
-data(bw_census_2008, bw_census_2014, bw_species)
-
-# Append additional species data
-bw_census_2008 <- bw_census_2008 %>%
-  left_join(bw_species, by = "sp") %>%
-  select(-c(genus, species, latin))
-
-```
-
-**SCBI**:
-
-For SCBI we will look at just a 9 ha section of the plot to speed calculations.
-
-```{r, eval = run_scbi}
-# SCBI
-scbi_2013 <- 
-  "https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/raw/master/tree_main_census/data/census-csv-files/scbi.stem2.csv" %>% 
-  read_csv() %>% 
-  select(
-    treeID, stemID, sp, quadrat, gx, gy, dbh, 
-    date = ExactDate, codes, status
-  ) %>%
-  mutate(date = mdy(date)) %>%
-  filter(gx < 300, gy > 300, gy < 600)
-
-scbi_2018 <-
-  "https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/raw/master/tree_main_census/data/census-csv-files/scbi.stem3.csv" %>% 
-  read_csv() %>% 
-  select(
-    treeID, stemID, sp, quadrat, gx, gy, dbh, 
-    date = ExactDate, codes, status
-  ) %>% 
-  mutate(dbh = as.numeric(dbh),
-         date = mdy(date))  %>%
-  filter(gx < 300, gy > 300, gy < 600)
-```
-
-### Compute growth
-
-For each plot we combine the two census data frames into a single data frame  that now includes a numerical variable `growth` reflecting the average annual growth of dbh in cm.
-Furthermore, variables that (in theory) remain unchanged between censuses appear only once, such as `gx`, `gy`, species-related variables.
-Variables that should change between censuses are tagged with `1/2` indicating earlier/later, such as `dbh1/dbh2`, `codes1/codes2`.
-
-The resulting data frames are named with some variation of `growth_df`.
-
-**Big Woods**:
-
-```{r, eval = run_bw}
-# Big Woods
-id <- "treeID"
-
-# we need to filter out the re-sprouts
-bw_census_2014 <- bw_census_2014 %>% 
-  filter(!str_detect(codes, 'R'))
-
-bw_growth_df <-
-  # Merge both censuses and compute growth:
-  compute_growth(bw_census_2008, bw_census_2014, id) %>%
-  mutate(
-    sp = to_any_case(sp),
-    sp = as.factor(sp),
-    species = sp,
-    family = as.factor(family),
-    trait_group = as.factor(trait_group)) %>%
-  # drop stemID
-  select(-stemID)
-
-```
-
-**SCBI**:
-
-```{r, eval = run_scbi}
-# SCBI
-census_df1 <- scbi_2013 
-census_df2 <- scbi_2018
-id <- "stemID"
-
-scbi_growth_df <- 
-  # Merge both censuses and compute growth:
-  compute_growth(census_df1, census_df2, id) %>%
-  # Convert growth from cm to mm to make result comperable
-  mutate(growth = growth/10,
-             sp = as.factor(sp))
-
-```
-
-**Comparison**:
-
-```{r, eval = run_scbi & run_bw}
-# Both Big Woods & SCBI
-growth_df <- bind_rows(
-  bw_growth_df %>% st_drop_geometry %>% select(growth) %>% mutate(site = "bw"),
-  scbi_growth_df %>% st_drop_geometry %>% select(growth) %>% mutate(site = "scbi")
-)
-ggplot(growth_df, aes(x = growth, y = ..density.., fill = site)) +
-  geom_histogram(alpha = 0.5, position = "identity", binwidth = 0.05) +
-  labs(x = "Average annual growth in dbh (cm per yr)") +
-  coord_cartesian(xlim = c(-0.5, 1))
-```
-
-### Add spatial information
-
-We now add spatial information to the `growth_df` data frame. Specifically:
-
-1.
-In order to control for study region "edge effects" (cite Waller?), we need a function that adds "buffers" of trees.
-In our case, since our model of interspecific competition relies on a spatial definition of who is a "neighboring competitor" and certain explanatory variables such as biomass are cumulative, we need to ensure that all trees being modeled are not biased to have different neighbor structure, in particular the trees at the boundary of study regions.
-1.
-Assign each tree to a "fold" for cross-validation purposes.
-Conventional cross-validation schemes assign observations to folds by resampling individual observations at random.
-However, underlying this scheme is an assumption that the observations are independent.
-In the case of forest census data, observations exhibit spatial autocorrelation, and thus this dependence must be incorporated in our resampling scheme @roberts2017 @pohjankukka2017.
-Packages that have implemented spatial cross-validation include @valavi2019
-
-Thus, before we can incorporate the above information to `growth_df`, we need to define two constants.
+Here we provide an example with a small dateset. First we combine the two census data files into a single `tibble` with the growth of each surviving individual.
 
 ```{r}
-cv_fold_size <- 100
-max_dist <- 7.5
-```
-
-The first, `cv_fold_size` is the length of the folds, each fold is a square.
-The next, `max_dist` is the maximum distance for a tree's competitive neighborhood.
-Trees within this distance of each other are assumed to compete while those farther than this distance apart do not.
-Both these values are in the same units as `gx` and `gz`, most often this will be meters.
-
-<!--
-DA comment: I am not sure about the below so I took it out.
-We don't assume that competitive interactions are inversely related to distance.
-We assume that competitive interactions are equivalent between distance 0 and 7.5, but than zero after that.
-Canham has the inversely related structure but we don't.
-
-Following Tobler's first law of geography that "everything is related to everything else, but near things are more related than distant things." @tobler1970firstlawofgeo, we assume that the degree of spatial autocorrelation is inversely-related to distance.
-However, we further assume that once trees are more than `r max_dist` meters apart, this autocorrelation is negligeable.
--->
-
-Other studies have estimated the value of `max_dist`; we use an average of estimated values [@canham2004, @canham2006, @uriarte2004, @tatsumi2013].
-
-#### Defining buffers
-
-We use the function `add_buffer_variable` to identify which trees are inside of a buffer region of size `max_dist`. The user will need to specify a study region and convert to a `sf_polygon`.
-
-**Example of buffer at work**:
-
-```{r}
-# Boundary polygon
-square_boundary <- tibble(
-  x = c(0,0,1,1),
-  y = c(0,1,1,0)
-) %>% 
-  sf_polygon()
-
-# "Trees" in polygon
-trees_df <- tibble(
-    x = runif(100),
-    y = runif(100)
-  ) %>%
-  sf_point()
-
-# Buffer polygon
-trees_df <- trees_df %>%
-  add_buffer_variable(direction = "in", size = 0.1, region = square_boundary)
-
-ggplot() +
-  geom_sf(data = trees_df, aes(col = buffer))
-
-```
-
-**Big Woods**:
-
-```{r, eval = run_bw}
-# Bigwoods
-data(bw_study_region)
-
-# Add buffer variable to data frame
-bw_growth_df <- bw_growth_df %>%
-  add_buffer_variable(direction = "in", size = max_dist, region = bw_study_region)
-
-ggplot() +
-  geom_sf(data = bw_growth_df, aes(col = buffer))
-
-```
-
-**SCBI**:
-
-```{r, eval = run_scbi}
-# SCBI
-# Study region boundary polygon
-scbi_study_region <- 
- #tibble(x = c(0,400,400,0,0), y = c(0,0,640,640,0)) %>% 
-  tibble(x = c(0,300,300,0,0), y = c(300,300,600,600,3)) %>% 
-  sf_polygon()
-
-# Add buffer variable to data frame
-scbi_growth_df <- scbi_growth_df %>%
-  add_buffer_variable(direction = "in", size = max_dist, region = scbi_study_region)
-
-ggplot() +
-  geom_sf(data = scbi_growth_df, aes(col = buffer))
-```
-
-#### Defining spatial cross-validation folds
-
-We use the [`blockCV`](https://github.com/rvalavi/blockCV) package to define the spatial grid, whose elements will act as the folds in our leave-one-out (by "one" we mean "one grid block") cross-validation scheme.
-The upshot here is we add `foldID` to `growth_df` which identifies which fold each individual is in, and the creation of a `cv_grid_sf` object which gives the geometry of the cross validation grid.
-
-**Big Woods**:
-
-```{r, eval = run_bw}
-# Big Woods
-set.seed(76)
-bw_cv_grid <- spatialBlock(
-  speciesData = bw_growth_df, theRange = cv_fold_size, verbose = FALSE,
-  # Some guess work in figuring this out:
-  k = 28, xOffset = 0.5, yOffset = 0
-)
-
-# Add foldID to data
-bw_growth_df <- bw_growth_df %>% 
-  mutate(
-    foldID = bw_cv_grid$foldID
-  )
-
-# Visualize grid. Why does fold 19 repeat?
-bw_cv_grid$plots +
-  geom_sf(data = bw_growth_df %>% sample_frac(0.2), aes(col=factor(foldID)), size = 0.1)
-
-bw_growth_df <- bw_growth_df %>%
-  filter(!foldID %in% c(19, 23, 21, 17, 8, 19)) %>%
-  mutate(foldID = as.character(foldID))
-
-# Deliverable
-ggplot() +
-  geom_sf(data = bw_growth_df, aes(col = foldID, alpha = buffer))
-
-bw_cv_grid_sf <- bw_cv_grid$blocks %>%
-  st_as_sf()
-```
-
-**SCBI**:
-
-```{r, eval = run_scbi}
-# SCBI
-scbi_cv_grid <- spatialBlock(
-  speciesData = scbi_growth_df, theRange = cv_fold_size, yOffset = 0.9999, k = 9, verbose = FALSE
-)
-
-# Add foldID to data
-scbi_growth_df <- scbi_growth_df %>% 
-  mutate(
-    foldID = scbi_cv_grid$foldID
-  )
-
-# Visualize grid
-scbi_cv_grid$plots +
-  geom_sf(data = scbi_growth_df, aes(col=factor(foldID)), size = 0.1)
-
-scbi_cv_grid_sf <- scbi_cv_grid$blocks %>%
-  st_as_sf()
-```
-
-### Focal versus competitor trees
-
-Next we create the `focal_vs_comp` data frame which connects each tree in `growth_df` to the trees in its competitive neighborhood range (as defined by `max_dist`).
-So for example, if `growth_df` consisted of two focal trees with two and three neighbors respectively, `focal_vs_comp` would consist of 5 rows.
-
-This requires the `growth_df` data frame; `max_dist`, the scalar defining competitive range; `cv_grid_sf`, giving the cross validation grid; and the `id` variable as inputs.
-
-```{r, eval = run_bw}
-# Big Woods
-if (!file.exists("focal_vs_comp_bw.Rdata")) {
-  tic()
-  focal_vs_comp_bw <- bw_growth_df %>% 
-    create_focal_vs_comp(max_dist, cv_grid_sf = bw_cv_grid_sf, id = "treeID")
-  toc()
-  save(focal_vs_comp_bw, file = "focal_vs_comp_bw.Rdata")
-} else {
-  load("focal_vs_comp_bw.Rdata")
-}
-```
-
-```{r, eval = run_bw}
-# Big Woods
-glimpse(focal_vs_comp_bw)
-```
-
-**SCBI**:
-
-```{r, eval = run_scbi}
-# SCBI
-tic()
-focal_vs_comp_scbi <- scbi_growth_df %>% 
-  create_focal_vs_comp(max_dist, cv_grid_sf = scbi_cv_grid_sf, id = "stemID")
-toc()
-```
-
-```{r, eval = run_scbi}
-# SCBI
-glimpse(focal_vs_comp_scbi)
-```
-
-### Model fit and prediction
-
-**Big Woods**:
-
-Now we are ready to fit the competition model with `fit_bayesian_model`.
-This function needs only the `focal_vs_comp` as an input.
-Other options allow the user to specify prior parameters and run a competitor-species identity shuffle (see below).
-
-```{r, eval = run_bw}
-# Big Woods
-tic()
-posterior_param_bw <- focal_vs_comp_bw %>% 
-  fit_bayesian_model(prior_param = NULL, run_shuffle = FALSE)
-toc()
-```
-
-This output has the posterior parameters for the specified competition model.
-This `posterior_param` output can be used to get predicted growths for each individual (with `predict_bayesain_model`) to test how well the model performs.
-Or this `posterior_param` output can be plotted (either the betas or lambdas with `plot_posterior_params`) to understand what controls growth and the nature of the competitive ineractions.
-
-Here we calculate the RMSE
-
-```{r, eval = run_bw}
-# Big Woods
-predictions <- focal_vs_comp_bw %>%
-  predict_bayesian_model(posterior_param = posterior_param_bw) %>%
-  right_join(bw_growth_df, by = c("focal_ID" = "treeID"))
-predictions %>%
-  rmse(truth = growth, estimate = growth_hat) %>%
-  pull(.estimate)
-```
-
-Now we test whether the identity of the competitor matters.
-We do this by shuffling the identity of competitors (but not of focal trees or spatial locations or sizes) and fitting the model again.
-We then compare RMSEs to see whether competitor identity matters to competitive interactions.
-
-```{r, eval = run_bw}
-posterior_param_bw_shuffle <- focal_vs_comp_bw %>%
-  fit_bayesian_model(prior_param = NULL, run_shuffle = TRUE)
-
-# b) Make predictions and compute RMSE
-predictions_shuffle <- focal_vs_comp_bw %>%
-  predict_bayesian_model(posterior_param = posterior_param_bw_shuffle) %>%
-  right_join(bw_growth_df, by = c("focal_ID" = "treeID"))
-predictions_shuffle %>%
-  rmse(truth = growth, estimate = growth_hat) %>%
-  pull(.estimate)
-```
-
-The RMSE is lower for the non-shuffled version. This gives support for the idea that competitor identity does matter for competitive interactions.
-
-**SCBI**:
-
-```{r, eval = run_scbi}
-# SCBI
-tic()
-posterior_param_scbi <- focal_vs_comp_scbi %>% 
-  fit_bayesian_model(prior_param = NULL, run_shuffle = FALSE)
-toc()
-```
-
-Here we provide examples of seeing how well the model fits and then a spatial plot of the residuals to see whether there are any consistent spatial patterns to them.
-
-```{r, eval = run_scbi}
-# SCBI
-scbi_growth_df_noCV <- focal_vs_comp_scbi %>% 
-  predict_bayesian_model(posterior_param = posterior_param_scbi) %>% 
-  right_join(scbi_growth_df, by = c("focal_ID" = "stemID"))
-
-scbi_growth_df_noCV %>%
-  rmse(truth = growth, estimate = growth_hat) %>%
-  pull(.estimate)
-
-# Observed vs predicted growth  
-ggplot(scbi_growth_df_noCV, aes(x = growth, y = growth_hat)) +
-  geom_point(size = 0.5, color = rgb(0, 0, 0, 0.25)) +
-  stat_smooth(method = 'lm') +
-  geom_abline(slope = 1, intercept = 0) +
-  coord_fixed() + 
-  labs(
-    x = "Observed growth in dbh", y = "Predicted growth in dbh", 
-    title = "Predicted vs Observed Growth"
-  )
-
-reslab <- expression(paste('Residual (cm ',y^{-1},')'))
-scbi_growth_df_noCV %>% 
-  st_as_sf() %>%
-  # Need to investigate missingness
-  filter(!is.na(growth_hat)) %>% 
-  mutate(
-    error = growth - growth_hat,
-    error_bin = cut_number(error, n = 5), 
-    error_compress = ifelse(error < -0.75, -0.75, ifelse(error > 0.75, 0.75, error))
-  ) %>% 
-  ggplot() + 
-  geom_sf(aes(col = error_compress), size = 1) + 
-  theme_bw() + 
-  scale_color_gradient2(
-    low = "#ef8a62", mid = "#f7f7f7", high = "#67a9cf", 
-    name = reslab,
-    breaks = seq(from = -0.75, to = 0.75, by = 0.25),
-    labels = c('< -0.75', '-0.5', '0.25', '0', '0.25', '0.5', '> 0.75')) +
-  labs(x = "Meter", y = "Meter")
-```
-
-### Run spatial cross-validation
-
-For the above results we fit the model to the entire data set, and then make predictions across the entire data set from that fit.
-This could lead to overfitting because we are using the training data to also test the model.
-If model error is spatially correlated this could be a large issue (cite important sources here!).
-We can use the spatial block structure we defined above to deal with with.
-The function `run_cv` goes through each fold in the `cv_grid_sf` and fits the model on all the other folds.
-Then applies that fit to the focal fold.
-It is a wrapper for `fit_bayesain_model` and `predict_bayesain_model` but fits a separate model for each fold.
-
-```{r, eval = run_bw}
-# big woods
-tic()
-cv_bw <- focal_vs_comp_bw %>%
-  run_cv(max_dist = max_dist, cv_grid = bw_cv_grid) %>%
-  right_join(bw_growth_df, by = c("focal_ID" = "treeID"))
-toc()
-
-cv_bw %>%
-  rmse(truth = growth, estimate = growth_hat) %>%
-  pull(.estimate)
-```
-
-As expected this RMSE is higher than that when the model is fit without cross validation. See Allen and Kim (2020) for more discussion of this.
-
-```{r, eval = run_scbi}
-# SCBI
-tic()
-cv_scbi <- focal_vs_comp_scbi %>%
-  run_cv(max_dist = max_dist, cv_grid = scbi_cv_grid) %>%
-  right_join(scbi_growth_df, by = c("focal_ID" = "treeID"))
-toc()
-
-cv_scbi %>%
-  rmse(truth = growth, estimate = growth_hat) %>%
-  pull(.estimate)
-```
-
-Here also this RMSE is much higher than that for the above SCBI model fit without cross validation.
-
-### Visualize posterior distributions
-
-**Big Woods**
-
-We might be interested in the posterior distributions of parameters.
-The betas tell us about how fast each species grows and how this depends on DBH.
-The lambdas, often of more interest, are the species-specific competition coefficients.
-The full lambda matrix gives competition strength between species.
-There is a rich literature on this matrix (cite).
-
-Because of the structure of the `bw_fit_model` object we cannot simply draw these curves based on the posterior distribution.
-`bw_fit_model` gives the parameters *compared* to a baseline.
-This is not of direct interest.
-So to display these parameters, as we care about them, we have to sample from the baseline distribution and from the comparison one to get the posterior distribution of interest.
-
-Here we re-run the Big Woods model but using the family as the group for comparison.
-This makes the posterior distributions easier to follow.
-Also, surprisingly, grouping by family performed just as well as grouping by species (see Allen and Kim 2020).
-First we re-run `create_focal_vs_comp` and `fit_bayesian_model` with the grouping variable as family.
-
-```{r, eval = run_bw}
-
-focal_vs_comp_bw <- bw_growth_df %>%
-  # mutate(sp = trait_group) %>%
-  mutate(sp = family) %>%
-  create_focal_vs_comp(max_dist = max_dist, cv_grid_sf = bw_cv_grid_sf, id = "treeID")
-
-# a) Fit model (compute posterior parameters) with no permutation shuffling
-posterior_param_bw <- focal_vs_comp_bw %>%
-  fit_bayesian_model(prior_param = NULL, run_shuffle = FALSE)
-
-```
-
-Now the output of `fit_bayesian_model` is passed to `plot_posterior_parameters`.
-
-```{r, eval = run_bw}
-# b) Recreate Fig5 from Allen (2020): Posterior distributions of selected lambdas
-posterior_plots <- plot_posterior_parameters(
-  posterior_param = posterior_param_bw,
-  sp_to_plot = c("cornaceae", "fagaceae", "hamamelidaceae", "juglandaceae", "lauraceae", "rosaceae", "sapindaceae", "ulmaceae")
-)
-```
-
-The output is a list with three plots stored. The element `beta_0` gives the growth intercept, i.e., how fast an individual of each group grows independent of DBH).
-
-```{r, eval = run_bw}
-posterior_plots[["beta_0"]]
-```
-
-Next `beta_dbh` gives the DBH-growth slope for each group.
-
-```{r, eval = run_bw}
-posterior_plots[["beta_dbh"]]
-```
-
-Finally `lambda` gives the competition coeffiencts.
-
-```{r, eval = run_bw}
-posterior_plots[["lambda"]]
-```
-
-**SCBI**
-
-```{r, eval = run_scbi}
-posterior_plots_scbi <- plot_posterior_parameters(
-  posterior_param = posterior_param_scbi,
-  sp_to_plot = c("quru", "litu", "cagl", "cato")
-)
-```
-
-```{r, eval = run_scbi}
-posterior_plots_scbi[["lambda"]]
-```
-
-### Using small example
-
-Need to write this up for the readme.
-
-```{r, eval = small_example}
 # Read in census files
@@
     sp = as.factor(sp)) 
+```
 
+All growth models in the package assume that two individuals compete if they are less than a specified distance away.
+Here we set this distance and put all individuals within this distance of a plot boundray in a buffer (since we do not have full information about their competitors).
+
+```{r}
 data("ex_study_region")
@@
   geom_sf(data = ex_growth_df, aes(col = buffer), size = 2)
+```
 
+We then set blocks for the cross validation scheme.
+
+```{r}
 # Bert's solution: Manually create a blocks sf object
@@
   st_as_sf()
+```
+
+We then create a `focal_v_comp` tibble which has a row for each competing pair of individuals (each pair of individuals within `max_dist` of one another).
+
+```{r}
 
@@
   create_focal_vs_comp(max_dist, cv_grid_sf = ex_cv_grid_sf, id = "ID")
+```
 
+With `focal_v_comp` we can then run the growth model and get predictions from it.
+Here we run the model without cross validation.
+Posterior estimates can be plotted to give, for example, the lambda matrix of competition coefficents.
+
+```{r}
 # fit the model
@@
 # plot posteriors
-posterior_param_ex %>%
+plot_ex <- posterior_param_ex %>%
   plot_posterior_parameters()
 
+plot_ex[["lambda"]]
+
+```
+
+Here we repeat the model fitting but with cross-validation. Note the increase in RMSE.
+
+```{r}
 # run model with CV
@@
 ```
+
+For fuller examples on actual datasets see ___.

### 2020-11-20 README.Rmd

@@
   out.width = "100%",
-  fig.width = 16/2,
-  fig.height = 9/2,
+  fig.width = 16 / 2,
+  fig.height = 9 / 2,
   message = FALSE,
   warning = FALSE,
-  cache = TRUE
+  cache = FALSE
 )
@@
 
-This package is designed to work for spaitally mapped, repeat censused forests plots.
-The package has commands to fit models of tree growth based on neighborhood competition.
-This can be used to estimate species-specific competition coefficients.
-The models can be fit with a spatial cross validation scheme to avoid overfitting.
-Additionally these models can test whether the idenity of competitors matters, by shuffling competitor identity and testing whether model preformance changes.
-See Allen and Kim (2020) for a full descripition.
+This package is designed to work for spatially mapped, repeat censused forests plots.
+The package has commands to fit models of tree growth based on neighborhood competition which can be used to estimate species-specific competition coefficients.
+The models fits can then be evaluated using a spatial cross-validation scheme to detect possible overfitting.
+Additionally these models can test whether the species identity of competitors matters using a permutation test-style shuffling of competitor identity (under the null hypothesis) and subsequently evaluating if model performance changes.
+See Allen and Kim (2020) [A permutation test and spatial cross-validation approach to assess models of interspecific competition between trees](https://doi.org/10.1371/journal.pone.0229930) for a full description.
 
 ## Example analysis
+
+Here we provide an example with a small dataset.
 
@@
 library(yardstick)
+library(snakecase)
 ```
 
-Here we provide an example with a small dateset. First we combine the two census data files into a single `tibble` with the growth of each surviving individual.
+First we combine two example census data files into a single `tibble` and compute the growth of all surviving individuals.
 
@@
 census_df2_ex_no_r <- census_df2_ex %>%
-  filter(!str_detect(codes, 'R'))
+  filter(!str_detect(codes, "R"))
 
-id <- 'ID'
+# Name of variable in tibble that uniquely identifies each stem:
+id <- "ID"
 
+# Merge both censuses and compute growth:
 ex_growth_df <-
-  # Merge both censuses and compute growth:
   compute_growth(census_df1_ex, census_df2_ex_no_r, id) %>%
@@
     sp = to_any_case(sp),
-    sp = as.factor(sp)) 
+    # Must be a factor
+    sp = as.factor(sp)
+  )
 ```
@@
 All growth models in the package assume that two individuals compete if they are less than a specified distance away.
-Here we set this distance and put all individuals within this distance of a plot boundray in a buffer (since we do not have full information about their competitors).
+Here we set this distance in `max_dist` and put all individuals within this distance of a plot boundary in a buffer (since we do not have full information about their competitors).
 
 ```{r}
+# Load study region:
 data("ex_study_region")
 
-# set max dist
+# Set max dist
 max_dist <- 1
 
-# add buffer
+# Add buffer
 ex_growth_df <- ex_growth_df %>%
@@
 
-We then set blocks for the cross validation scheme.
+We then set blocks/folds for the spatial cross-validation scheme.
+We this both manually and using the [`blockCV`](https://github.com/rvalavi/blockCV) package.
+Note that here we define the geometric structure of the folds manually using `sfheaders::sf_polygon()`:
 
 ```{r}
-# Bert's solution: Manually create a blocks sf object
 fold1 <- rbind(c(0, 0), c(5, 0), c(5, 5), c(0, 5), c(0, 0))
@@
   sf_polygon(fold1),
-  sf_polygon(fold2) ) %>%
+  sf_polygon(fold2)
+) %>%
   mutate(foldID = c(1, 2))
+```
 
-# Plot
-ggplot() +
-  geom_sf(data = ex_growth_df, aes(col = buffer), size = 2) +
-  geom_sf(data = blocks, fill = "transparent")
+Next we assign each point to folds using `blockCV::spatialBlock()`
 
-# fit spatialBlock()
+```{r}
 ex_cv_grid <- spatialBlock(
@@
   k = 2,
-  # Note new arguments
   selection = "systematic",
-  blocks = blocks#,
- # foldsCol = "foldID"
+  blocks = blocks,
+  showBlocks = FALSE
 )
@@
 
-# Visualize grid
-ex_cv_grid$plots +
-  geom_sf(data = ex_growth_df, aes(col = foldID), size = 2)
-
 # Deliverable
 ggplot() +
-  geom_sf(data = ex_growth_df, aes(col = foldID, shape = buffer))
+  geom_sf(data = ex_growth_df, aes(col = buffer, shape = foldID), size = 2) +
+  geom_sf(data = blocks, fill = "transparent")
 
+# Convert to sf object
 ex_cv_grid_sf <- ex_cv_grid$blocks %>%
@@
 
-We then create a `focal_v_comp` tibble which has a row for each competing pair of individuals (each pair of individuals within `max_dist` of one another).
+We then create a `focal_v_comp` tibble which has a row for each competing pair of individuals (each pair of individuals within `max_dist` of one another):
 
 ```{r}
-
-# focal v comp
 focal_vs_comp_ex <- ex_growth_df %>%
@@
 
-With `focal_v_comp` we can then run the growth model and get predictions from it.
+With `focal_v_comp`, we can then run the growth model and get predictions from it.
 Here we run the model without cross validation.
-Posterior estimates can be plotted to give, for example, the lambda matrix of competition coefficents.
+Posterior estimates can be plotted to give, for example, the lambda matrix of competition coefficients.
 
 ```{r}
-# fit the model
+# Fit model
 posterior_param_ex <- focal_vs_comp_ex %>%
@@
 
-# get predictions
+# Get predicted dbh values values
 predictions <- focal_vs_comp_ex %>%
@@
   right_join(ex_growth_df, by = c("focal_ID" = "ID"))
+
+# Compute RMSE of true vs predicted dbh values
 predictions %>%
@@
 
-# plot posteriors
+# Plot posteriors
 plot_ex <- posterior_param_ex %>%
   plot_posterior_parameters()
-
 plot_ex[["lambda"]]
-
 ```
 
-Here we repeat the model fitting but with cross-validation. Note the increase in RMSE.
+Here we repeat the process but with cross-validation. Note the increase in RMSE, reflecting the fact that our original estimate of model error was overly optimistic.
 
 ```{r}
-# run model with CV
+# Fit model with cross-validation
 ex_bw <- focal_vs_comp_ex %>%
@@
 
+# Compute RMSE of true vs predicted dbh values
 ex_bw %>%
@@
   pull(.estimate)
-
 ```

### 2021-01-08 etc/paper/forestecology.Rmd
moved paper to non-vignettes folder

---
title: "`forestecology` package for modeling interspecies competition between trees"

# to produce blinded version set to 1 and change blinded <- TRUE on line 54 below
blinded: 0

authors: 
- name: Albert Y. Kim 
  thanks: "Albert Y. Kim is Assistant Professor, Statistical & Data Sciences, Smith College, Northampton, MA 01063 (e-mail: akim04@smith.edu). This work was not supported by any grant. The authors thank numerous colleagues and students for their support."
  affiliation: Program in Statistical & Data Sciences, Smith College

- name: David Allen
  affiliation: Biology Department, Middlebury College

keywords:
- forestecology, trees

abstract: 
- "We provide a computational exercise suitable for early introduction in an undergraduate statistics or data science course that allows students to 'play the whole game' of data science: performing both data collection and data analysis. While many teaching resources exist for data analysis, such resources are not as abundant for data collection given the inherent difficulty of the task. Our proposed exercise centers around student use of Google Calendar to collect data with the goal of answering the question 'How do I spend my time?' On the one hand, the exercise involves answering a question with near universal appeal, but on the other hand, the data collection mechanism is not beyond the reach of a typical undergraduate student. A further benefit of the exercise is that it provides an opportunity for discussions on ethical questions and considerations that data providers and data analysts face in today's age of large-scale internet-based data collection."

bibliography: paper.bib
output: rticles::asa_article
link-citations: yes
header-includes:
- \usepackage{xcolor, soul, xspace, float, subfig}
---

```{r setup, include = FALSE, cache = FALSE}
# knitr settings
knitr::opts_chunk$set(
  # Code output:
  warning = FALSE,
  message = FALSE,
  echo = TRUE,
  cache = TRUE,
  # Figure:
  out.width = "100%",
  fig.path = "Figures/",
  fig.width = 16 / 2.5,
  fig.height = 9 / 2.5,
  fig.align = "center",
  fig.show = "hold",
  # Etc:
  collapse = TRUE,
  comment = "##"
  # tidy = FALSE
)

# Internally used packages
library(knitr)
library(ggplot2)
library(viridis)

# Random number generator seed value
set.seed(76)

# Set ggplot defaults output:
if (!knitr::is_html_output()) {
  # Grey theme:
  theme_set(theme_light())
  # Color scheme
  # TODO: REVERT THIS
  # scale_colour_discrete <- scale_colour_viridis_d
}

# Set output width
options(width = 70)
```

# Introduction

Repeat-censused forest plots offer excellent data to test neighborhood models of tree competition @allen_permutation_2020 @canham_neighborhood_2006 @uriarte_spatially_2004. Here we describe an R package, `forestecology`, to do that. This package implements the methods in @allen_permutation_2020. It provides: a convenient way specify and fit models of tree growth based on neighborhood competition; a spatial cross validation method to test and compare model fits @roberts_cross-validation_2017; and an ANOVA-like method to assess whether the competitor identity matters in these models. The model is written to work with ForestGEO plot data @andersonteixeira_ctfs-forestgeo_2015, but we envision that it could easily be modified to work with data from other forest plots, e.g. the US Forest Service Forest Inventory and Analysis plots @smith_forest_2002.

# Example

We demonstrate the `forestecology` package's features on two data sets, both based on inventory censuses of two sites from the Smithsonian Institution’s ForestGEO international network of 72 long‐term forest dynamics research sites @andersonteixeira_ctfs-forestgeo_2015. First, the Michigan Big Woods Forest Dynamics Plot located at the Edwin S. George Reserve in Pinckney, MI, USA. The 23 ha plot is situated in mature oak-hickory forest. The canopy is dominated by white oak (Quercus alba), northern red oak (Quercus rubra), black oak (Quercus velutina), shagbark hickory (Carya ovata) and pignut hickory (Carya glabra). The most common understory tree is witch-hazel (Hamamelis virginiana) @allen_michigan_2020. In the example below, we will preface any data frames from this plot in with `bw_`.

Second, the Smithsonian Conservation Biology Institute (SCBI) large forest dynamics plot, located at the Smithsonian's National Zoo and Conservation Biology Institute in Front Royal, VA, USA. The 25.6 ha (640 x 400 m) plot is located at the intersection of three of the major physiographic provinces of the eastern US: the Blue Ridge, Ridge and Valley, and Piedmont provinces and is adjacent to the northern end of Shenandoah National Park. The forest type is typical mature secondary eastern mixed deciduous forest, with a canopy dominated by tulip poplar (Liriodendron tulipifera), oaks (Quercus spp.), and hickories (Carya spp.), and an understory composed mainly of spicebush (Lindera benzoin), paw-paw (Asimina triloba), American hornbeam (Carpinus caroliniana), and witch hazel (Hamamelis virginiana) @bourg_initial_2013. In the example below, we will preface any data frames from this plot in with `scbi_`.

We load all the necessary packages.

```{r load-packages}
library(forestecology)

# Load tidyverse packages:
library(tidyverse)
library(lubridate)

# Load spatial packages:
# devtools::install_github("rvalavi/blockCV")
library(blockCV)
library(sf)
library(sfheaders)

# Load other packages:
library(snakecase)
library(yardstick)
```

## Preprocess census data

We start by preprocessing the census data for both sites. While ForestGEO data protocols ensure a high degree of standardization between site, minor variations still exist @andersonteixeira_ctfs-forestgeo_2015. While the Big Woods data comes pre-loaded in the `forestecology` package, we load the SCBI data as they are saved in .csv files in the SCBI-ForestGEO-Data repository on GitHub @gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020.  In both cases, we load the census data as R as "tibble" data frames thereby ensuring a standardized input/output format that can be used across all `tidyverse` packages @wickham_welcome_2019.

Furthermore, we ensure that the different variables have the correct names, types (`dbl`, `data`, `factor`).

### Big Woods

We load census data from 2008 and 2014 saved in the package, then merge species data (genus, species, linnean classification, family, etc).

```{r bw-load-data}
data(bw_census_2008, bw_census_2014, bw_species)

# Append additional species data
bw_census_2008 <- bw_census_2008 %>%
  left_join(bw_species, by = "sp") %>%
  select(-c(genus, species, latin))
```

### SCBI {#scbi-data}

We load census data from 2008 and 2014 from `.csv` files saved from GitHub on November 20, 2020. Furthermore, we perform two additional pre-processing steps. First, in order to speed up computation for purposes of this example, we only consider a 9 ha subsection of the 25.6 ha of the SCBI site: `gx` from 0--300 instead of 0--400 and `gy` from  300--600 instead of 0--640. Second, in order to standardize comparisons between Big Woods and SCBI, we convert the units of dbh from mm to cm. ^[A rule of thumb to ascertain if dbh is in mm or cm is to verify if the smallest non-zero and non-missing measurement is 1 or 10. If the former, then cm. If the later, then mm. This is because ForestGEO protocols state that only trees with dbh greater or equal to 1cm should be included in censuses. ]

```{r scbi-load-data}
scbi_2013 <- read_csv("scbi.stem2.csv") %>%
  select(treeID, stemID, sp, ExactDate, gx, gy, dbh, codes, status) %>%
  mutate(
    date = ExactDate,
    dbh = as.numeric(dbh),
    date = mdy(date)
  ) %>%
  filter(gx < 300, between(gy, 300, 600)) %>% 
  mutate(dbh = dbh / 10)

scbi_2018 <- read_csv("scbi.stem3.csv") %>%
  select(treeID, stemID, sp, ExactDate, gx, gy, dbh, codes, status) %>%
  mutate(
    date = ExactDate,
    dbh = as.numeric(dbh),
    date = mdy(date)
  ) %>%
  filter(gx < 300, between(gy, 300, 600)) %>% 
  mutate(dbh = dbh / 10)
```

## Compute annual growth

For each plot we then compute average annual growth between the two censuses using the `compute_growth()` function. This function takes the two census data frames as well as a character indicating which variable in both data frames uniquely identifies each stem. This function returns a single data frame that includes a numerical variable `growth` reflecting the average annual dbh growth (in cm) of all trees alive at both time points. Furthermore, variables that (in theory) remain unchanged between censuses appear only once, such as location variables `gx` and `gy`; as well as species-related variables. Variables that should change between censuses are suffixed with `1` and `2` indicating the earlier and later censuses, such as `dbh1/dbh2` and `codes1/codes2`. Here the resulting data frames are named with some variation of `growth_df`.

After computing the average annual growth for each tree, we ensure to convert all variables denote species from type character to factors; this is to ensure that issues of rare species being accounted for in both training and test sets in our upcoming cross-validation step (see Section REF)

### Big Woods

In the case of Big Woods data, we first remove all trees that were re-sprouts in the later (2014) census. Additionally, we have included three classification of tree species: `species`, `family`, and `trait_group`. DESCRIBE THESE

```{r bw-compute-growth}
bw_census_2014 <- bw_census_2014 %>%
  filter(!str_detect(codes, "R"))

bw_growth_df <-
  compute_growth(bw_census_2008, bw_census_2014, id = "treeID") %>%
  # Convert all variables denoting species to factors
  mutate(
    sp = sp %>% to_any_case() %>% as.factor(),
    species = sp,
    family = as.factor(family),
    trait_group = as.factor(trait_group)
  ) %>%
  # Drop unnecessary variables
  select(-stemID)
```

### SCBI

```{r scbi-compute-growth}
scbi_growth_df <-
  compute_growth(scbi_2013, scbi_2018, "stemID") %>%
  # Convert all variables denoting species to factors
  mutate(sp = as.factor(sp))
```

### Comparison

Figure \ref{fig:growth-histogram} displays histograms comparing the distribution of average annual growth at both sites. Observe that average annual growth appears higher at the Big Woods site.

```{r growth-histogram, out.width="100%", fig.cap="Distribution of average annual growth in DBH for both sites."}
# Both Big Woods & SCBI
bind_rows(
  bw_growth_df %>% st_drop_geometry() %>% select(growth) %>% mutate(Site = "Big Woods"),
  scbi_growth_df %>% st_drop_geometry() %>% select(growth) %>% mutate(Site = "SCBI")
) %>% 
  ggplot(aes(x = growth, y = ..density.., fill = Site)) +
  geom_histogram(alpha = 0.5, position = "identity", binwidth = 0.05) +
  labs(x = "Average annual growth in dbh (cm per yr)") +
  coord_cartesian(xlim = c(-0.5, 1))
```

## Add spatial information {#spatial-information}

We now encode spatial information to the `growth_df` data frames. First, in order to control for study region edge effects, we add "buffers" to the periphery of the study region (cite Waller?). Our model of interspecific competition relies on a spatial definition of who the competitor trees are for focal trees of interest. Since certain explanatory variables such as biomass are cumulative, we must ensure that all trees being modeled are not biased to have different neighbor structures. This is a particular concern for trees at the boundary of study regions, which will not have the same number of neighbors as trees in the internal part of the study region.

Second, our ultimate method for model assessment will rely on estimates of model error as generated by cross-validation. Conventional cross-validation schemes assign observations to folds by resampling individual observations at random. However, underlying this scheme is an assumption that the observations are independent. In the case of forest census data, observations exhibit spatial autocorrelation, and thus this dependence must be incorporated in our resampling scheme in spatial cross-validation @roberts2017 @pohjankukka2017. We will therefore associate portions of the study region to spatial folds.

To these two ends, we define two constants, both of which are in the same units as the `gx` and `gy` variables (most often meters).

```{r}
max_dist <- 7.5
cv_fold_size <- 100
```

The first constant is `max_dist` which defines the maximum distance for a tree's competitive neighborhood. Trees within this distance of each other are assumed to compete while those farther than this distance apart do not. Put differently, all trees within `max_dist` of a focal tree will be considered its competitors (see below). Other studies have estimated the value of `max_dist`; we use an average of estimated values @canham_neighborhood_2004, @uriarte_spatially_2004, @tatsumi2013, @canham_neighborhood_2006.

Furthermore, `max_dist` will define the size of all buffers considered, which will be encoded as a binary variable `buffer` as computed by the `add_buffer_variable()` function. This function takes as input the main `growth_df` data frame, the `size` of the buffer which we set as `max_dist`, and the boundary of the study region encoded as a simple features polygon @pebesma_simple_2018. DESCRIBE SF PACKAGE. In the Big Woods example below we will use a pre-loaded simple features polygon while for the SCBI example we present example code on how to manually construct one.

The second constant is `cv_fold_size` which defines the length and width of the spatial folds (note that for now the spatial folds are restricted be squares). We will then use this constant to associate each observed tree to one of $k$ folds in the respective study region. In the Big Woods example below we will use the `blockCV` R package that has implemented spatial cross-validation while for the SCBI we will do this manually @valavi2019.

### Big Woods

First, we indicate which trees are part of the buffer. This necessitates information about the study region boundary. In this case, we use a `sf_polygon` object `bw_study_region` which comes pre-loaded in the `forestecology` packages. After loading `bw_study_region`, we illustrate the results of the `add_buffer_variable()` function in Figure \ref{fig:bw-define-buffer}. Trees on the periphery denote with lighter colors are part of the buffer and will not be considered as "focal" trees of interest going forward; they will only be considered as competitor trees.

```{r bw-define-buffer, out.width="100%", fig.cap="Buffer region for Big Woods study region."}
data(bw_study_region)

bw_growth_df <- bw_growth_df %>%
  add_buffer_variable(direction = "in", size = max_dist, region = bw_study_region)

ggplot() +
  geom_sf(data = bw_growth_df %>% sample_frac(0.2), aes(col = buffer), size = 0.5)
```

Second, we associate each tree to spatial cross validation folds. In this case, we use the `spatialBlock()` function from the `blockCV` package to define the spatial grid which

THIS IS A MESS. We use the  @valavi_blockcv_2019, whose elements will act as the folds in our leave-one-out (by "one" we mean "one grid block") cross-validation scheme. The upshot here is we add `foldID` to `growth_df` which identifies which fold each individual is in, and the creation of a `cv_grid_sf` object which gives the geometry of the cross validation grid.

```{r bw-define-cv-folds}
set.seed(76)
bw_spatialBlock <- spatialBlock(
  speciesData = bw_growth_df, theRange = cv_fold_size,
  k = 28, xOffset = 0.5, yOffset = 0,
  verbose = FALSE
)

# Add foldID to each tree
bw_growth_df <- bw_growth_df %>%
  mutate(foldID = bw_spatialBlock$foldID)

# Visualize grid. Why does fold 19 repeat?
bw_spatialBlock$plots +
  geom_sf(data = bw_growth_df %>% sample_frac(0.2), aes(col = factor(foldID)), size = 0.1)

# Remove empty folds
bw_growth_df <- bw_growth_df %>%
  filter(!foldID %in% c(19, 23, 21, 17, 8, 19)) %>%
  mutate(foldID = as.character(foldID))
```

Separately, we save the spatial cross-validation grid as an `sf_polygon` object `bw_cv_grid`

```{r}
bw_cv_grid <- bw_spatialBlock$blocks %>%
  st_as_sf()
```

### SCBI

First, we indicate which trees are part of the buffer. In this case however we manually define the study region boundary based on the subregion we defined in Section \ref{scbi-data} and create an `sf_polygon` object using the `sf_polygon()` function from the `sfheaders` package. Figure \ref{fig:scbi-define-buffer} displays the resulting buffer trees.

```{r scbi-define-buffer, out.width="50%", fig.cap="Buffer region for SCBI study region."}
scbi_study_region <- tibble(
  x = c(0, 300, 300, 0, 0),
  y = c(300, 300, 600, 600, 300)
) %>%
  sf_polygon()

scbi_growth_df <- scbi_growth_df %>%
  add_buffer_variable(direction = "in", size = max_dist, region = scbi_study_region)

ggplot() +
  geom_sf(data = scbi_growth_df, aes(col = buffer), size = 0.5)
```

Second, we associate each tree to spatial cross validation folds. In this case we manually define a spatial crossvaliation grid. Figure \ref{fig:scbi-define-cv-folds} displays the resulting cross-validation folds along with the buffer from Figure \ref{fig:scbi-define-buffer}.

Here we manually define the spatial cross-validation grid as an `sf_polygon` object `scbi_cv_grid`

```{r}
fold1 <- rbind(c(0, 300), c(150, 300), c(150, 600), c(0, 600), c(0, 300)) %>%
  sf_polygon() %>% 
  mutate(folds = 1)
fold2 <- rbind(c(150, 300), c(300, 300), c(300, 600), c(150, 600), c(150, 300)) %>%
  sf_polygon() %>% 
  mutate(folds = 2)
scbi_cv_grid <- bind_rows(fold1, fold2)
```

```{r scbi-define-cv-folds, out.width="50%", fig.cap="Buffer region for SCBI study region."}
scbi_spatialBlock <- spatialBlock(
  speciesData = scbi_growth_df,
  k = 2,
  verbose = FALSE,
  showBlocks = FALSE,
  # Note new arguments:
  selection = "systematic",
  blocks = scbi_cv_grid
)

# Add foldID to each tree
scbi_growth_df <- scbi_growth_df %>%
  mutate(foldID = scbi_spatialBlock$foldID)

ggplot() +
  geom_sf_text(data = scbi_growth_df, aes(label = foldID, col = buffer)) +
  geom_sf(data = scbi_cv_grid, fill = "transparent")
```

## Define focal versus competitor trees

Next we define `focal_vs_comp` data frames which connects each focal tree in the `growth_df` data frames to the trees in its competitive neighborhood range as defined by the `max_dist` constant. So for example, if `growth_df` consisted of two focal trees with two and three neighbors with `max_dist` respectively, `focal_vs_comp` would be a data frame of 5 rows connecting each focal tree to it's competitors. The `create_focal_vs_comp()` function makes this connection taking as inputs the `growth_df` data frame; the `max_dist` constant defining competitive range; `cv_grid_sf`, giving the cross validation grid; and the `id` variable.

### Big Woods

```{r bw-focal-vs-comp}
focal_vs_comp_bw <- bw_growth_df %>%
  create_focal_vs_comp(max_dist, cv_grid_sf = bw_cv_grid, id = "treeID")
```

TODO: Figure out how to show this data frame's contents.

```{r}
head(focal_vs_comp_bw)
```

### SCBI

```{r scbi-focal-vs-comp}
focal_vs_comp_scbi <- scbi_growth_df %>%
  create_focal_vs_comp(max_dist, cv_grid_sf = scbi_cv_grid, id = "stemID")
```

TODO: Figure out how to show this data frame's contents.

```{r}
head(focal_vs_comp_scbi)
```

## Fit model and make predictions {#model-fit-predict}

Next we fit the following linear model to the dbh of each focal tree. Let $i = 1, \ldots, n_j$ index all $n_j$ trees of ``focal'' species group $j$; let $j = 1, \ldots, J$ index all $J$ focal species groups; and let $k = 1, \ldots, K$ index all $K$ ``competitor'' species groups. We modeled the growth in diameter per year $y_{ij}$ (in centimeters per year) of the $i^{th}$ tree of focal species group $j$ as a linear model $f$ of the following covariates $\vec{x}_{ij}$

$$
\newcommand{\dbh}{\text{DBH}}
\newcommand{\biomass}{\text{biomass}}
\newcommand{\BA}{\text{BA}}
y_{ij} = f(\vec{x}_{ij}) + \epsilon_{ij} = \beta_{0,j} + \beta_{\dbh,j} \cdot \dbh_{ij} + \sum_{k=1}^{K} \lambda_{jk} \cdot \BA_{ijk} + \epsilon_{ij}
$$

<!--where $\beta_{0,j}$ is the diameter-independent growth rate for group $j$; $\dbh_{ij}$ is the diameter at breast height (in centimeters) of the focal tree at the earlier census; $\beta_{\dbh,j}$ is the amount of the growth rate changed depending on diameter for group $j$;  $\BA_{ijk}$ is the sum of the basal area of all trees of competitor species group $k$; $\lambda_{jk}$ is the change in growth for individuals of group $j$ from nearby competitors of group $k$; and $\epsilon_{ij}$ is a random error term distributed $\text{Normal}(0, \sigma^2)$.-->

We estimate the model's parameters using Bayesian linear regression implemented in the `fit_bayesian_model()` function. TODO: define all parameters

For this linear model's case, there exists a closed form solution as described here. As such, the `fit_bayesian_model()` function using matrix algebra to obtain all parameter estimates, rather than computationally expensive Monte Carlo approximations. The inputs to this function are a `focal_vs_comp` data frame, `prior_param` a list of priors, and a boolean flag `run_shuffle` on whether or not to run competitor-species identity permutations which we will demonstrate below on the Michigan Big Woods data. This function returns the posterior means of all parameters.

Using these posterior means, we then use the posterior predictive distribution to obtain fitted/predicted values $\widehat{y}$ of the dbh for each focal tree using the `predict_bayesian_model()`. These $\widehat{y}$ can then be compared to the observed $y$ dbh's to compute the root mean-square error, a measure of a model's predictive error which has the same units as the observed data $y$.

### Big Woods

For the Michigan Big Woods data we present two use cases of the model fitting and prediction scheme. The first use case is the simplest where we assess the fit of the model using root mean squared error. The second use case then answers the question of whether species competitor identity matters using permutation test.

For the first use case, we fit the linear model specified in Equation XXX to our data frame of type `focal_vs_comp`. This input/outputs of the `fit_bayesian_model()` function are lists of the prior/posterior means of parameters of the linear regression specified in XXX. Generally speaking, there are two classes of regression parameters: $\beta$ main effects and $\lambda$ competitive effects. In the upcoming Section \ref{viz-posterior-distributions}, we will present code visualizing this posterior distributions.

```{r bw-model-fit}
posterior_param_bw <- focal_vs_comp_bw %>%
  fit_bayesian_model(prior_param = NULL)
```

This output of posterior parameters for the specified competition model are then used along with the posterior predictive distribution encoded in `predict_bayesian_model()` to return predicted growths for each individual tree. We join these predicted growths to the original growth data frame.

```{r bw-model-predict}
predictions <- focal_vs_comp_bw %>%
  predict_bayesian_model(posterior_param = posterior_param_bw) %>%
  right_join(bw_growth_df, by = c("focal_ID" = "treeID"))
```

We then use the `rmse()` function from the `yardstick` package to obtain the root mean squared error of the observed versus fitted values of growth.

```{r bw-model-rmse}
predictions %>%
  yardstick::rmse(truth = growth, estimate = growth_hat) %>%
  pull(.estimate)
```

The second use case is near identical to the first, but with a small change in the code to test whether the identity of the competitor matters. By adding a `run_shuffle = TRUE` argument to `fit_bayesian_model()`, for each focal tree its competitor trees' species identity will be "shuffled" randomly much like in a permutation test. By shuffling these species labels we are effectively fitting the model under a null model that competitor species identity does not matter. If the "shuffled" RMSE's are consistently lower than the unshuffled RMSE corresponding to the observed data, then we have evidence to suggest that competitor identity matters to competitive interactions.

```{r bw-model-shuffle-fit}
posterior_param_bw_shuffle <- focal_vs_comp_bw %>%
  fit_bayesian_model(prior_param = NULL, run_shuffle = TRUE)
```

```{r bw-model-shuffle-predict}
predictions_shuffle <- focal_vs_comp_bw %>%
  predict_bayesian_model(posterior_param = posterior_param_bw_shuffle) %>%
  right_join(bw_growth_df, by = c("focal_ID" = "treeID"))
```

```{r bw-model-shuffle-rmse}
predictions_shuffle %>%
  rmse(truth = growth, estimate = growth_hat) %>%
  pull(.estimate)
```

The RMSE is fact lower for the non-shuffled version, indicative of a better model fit. This gives support for the idea that competitor identity does matter for competitive interactions. In @allen_permutation_2020 we run this shuffle a large number of times to construct a full permutation distribution to show that this difference is robust to resampling variation.

### SCBI

In the case of the SCBI data, we once again perform the same model fitting and computing of fitted growths as with the Big Woods data, but this time we map the residuals of the observed minus fitted values to look for spatial patterns.

```{r scbi-model-fit}
posterior_param_scbi <- focal_vs_comp_scbi %>%
  fit_bayesian_model(prior_param = NULL, run_shuffle = FALSE)
```

```{r scbi-model-predict}
scbi_growth_df_noCV <- focal_vs_comp_scbi %>%
  predict_bayesian_model(posterior_param = posterior_param_scbi) %>%
  right_join(scbi_growth_df, by = c("focal_ID" = "stemID"))
```

```{r scbi-model-rmse}
scbi_growth_df_noCV %>%
  rmse(truth = growth, estimate = growth_hat) %>%
  pull(.estimate)
```

In Figures \ref{fig:scbi-model-residuals} and \ref{fig:scbi-model-residuals-2} we present the residuals.

```{r scbi-model-residuals, out.width="100%", fig.cap="Spatial distribution of residuals for model applied to SCBI data."}
ggplot(scbi_growth_df_noCV, aes(x = growth, y = growth_hat)) +
  geom_point(size = 0.5, color = rgb(0, 0, 0, 0.25)) +
  stat_smooth(method = "lm") +
  geom_abline(slope = 1, intercept = 0) +
  coord_fixed() +
  labs(
    x = "Observed growth in dbh", y = "Predicted growth in dbh",
    title = "Predicted vs Observed Growth"
  )
```

```{r scbi-model-residuals-2, out.width="100%", fig.cap="Spatial distribution of residuals for model applied to SCBI data part 2."}
scbi_growth_df_noCV %>%
  st_as_sf() %>%
  # TODO: Need to investigate missingness
  filter(!is.na(growth_hat)) %>%
  mutate(
    error = growth - growth_hat,
    error_bin = cut_number(error, n = 5),
    error_compress = ifelse(error < -0.75, -0.75, ifelse(error > 0.75, 0.75, error))
  ) %>%
  ggplot() +
  geom_sf(aes(col = error_compress), size = 1) +
  theme_bw() +
  scale_color_gradient2(
    low = "#ef8a62", mid = "#f7f7f7", high = "#67a9cf",
    name = expression(paste("Residual (cm ", y^{-1}, ")")),
    breaks = seq(from = -0.75, to = 0.75, by = 0.25),
    labels = c("< -0.75", "-0.5", "0.25", "0", "0.25", "0.5", "> 0.75")
  ) +
  labs(x = "Meter", y = "Meter")
```

## Run spatial cross-validation

The model fits and predictions in Section \ref{model-fit-predict} all suffer from a common failing: they use the same data to both fit the model and to assess the model's performance using the RMSE. As argued by @roberts_cross-validation_2017, this can lead to overly optimistic assessments of model quality as the models can be overfit, in particular in situations where spatial-autocorrelation is present. To mitigate the effects of such overfitting, we use a spatially block cross-validation algorithm implemented in the `run_cv()`. This function at its core uses the same model fitting implemented in the `fit_bayesian_model()` function, however trains the model on $k-1$ spatial folds of the train and returns fitted values for the test data. Recall that the spatial blocking scheme wass encoded in Section \ref{spatial-information}.

### Big Woods

Applying this spatially cross-validated model fit yields an RMSE is higher than that when the model is fit without cross validation. In other words, our model fits in \ref{model-fit-predict} were overly optimistic in the model's fitting power, whereas a cross-validated results yield an estimate that is closer to the truth. See @allen_permutation_2020 for more discussion of this.

```{r bw-spatial-cv}
cv_bw <- focal_vs_comp_bw %>%
  run_cv(max_dist = max_dist, cv_grid = bw_cv_grid) %>%
  right_join(bw_growth_df, by = c("focal_ID" = "treeID"))

cv_bw %>%
  rmse(truth = growth, estimate = growth_hat) %>%
  pull(.estimate)
```

### SCBI

Observe once again that this RMSE is much higher than that for the above SCBI model fit without cross-validation.

```{r scbi-spatial-cv}
cv_scbi <- focal_vs_comp_scbi %>%
  run_cv(max_dist = max_dist, cv_grid = scbi_cv_grid) %>%
  right_join(scbi_growth_df, by = c("focal_ID" = "treeID"))

cv_scbi %>%
  rmse(truth = growth, estimate = growth_hat) %>%
  pull(.estimate)
```

## Visualize posterior distributions {#viz-posterior-distributions}

Lastly, we return to the model fits from Section \ref{model-fit-predict} and present tools to visually explore the posterior distributions of all parameters in our model. There are two main groups of parameters to consider. The $\beta$ coefficients tell us about how fast each species grows and how this depends on DBH while the full matrix of $\lambda$ values describe the competitive effects between pairs of species. There is a rich literature on this matrix (cite).

DO WE NEED TO DESCRIBE MECHANICS? Because of the structure of the `bw_fit_model` object we cannot simply draw these curves based on the posterior distribution. `bw_fit_model()` gives the parameters *compared* to a baseline. This is not of direct interest. So to display these parameters, as we care about them, we have to sample from the baseline distribution and from the comparison one to get the posterior distribution of interest.

### Big Woods

Here we re-run the model fit to the Big Woods data from Section \ref{model-fit-predict}, but this time use "family" as the group for comparison which has. This makes the posterior distributions easier to follow. Also, surprisingly, grouping by family performed just as well as grouping by species @allen_permutation_2020. First we re-run `create_focal_vs_comp()` and `fit_bayesian_model()` with no permutation shuffling with the grouping variable as family.

```{r bw-posterior-viz-0}
focal_vs_comp_bw <- bw_growth_df %>%
  mutate(sp = family) %>%
  create_focal_vs_comp(max_dist = max_dist, cv_grid_sf = bw_cv_grid, id = "treeID")

posterior_param_bw <- focal_vs_comp_bw %>%
  fit_bayesian_model(prior_param = NULL, run_shuffle = FALSE)
```

Now the posterior parameter outputs of `fit_bayesian_model()` are passed to `plot_posterior_parameters()` to generate visualizations of the posterior parameters. These visualizations are displayed in Figure 5 of @allen_permutation_2020. For simplicity we only plot a subset of the species families.

```{r bw-posterior-viz-1}
posterior_plots <- plot_posterior_parameters(
  posterior_param = posterior_param_bw,
  sp_to_plot = c("cornaceae", "fagaceae", "hamamelidaceae", "juglandaceae", 
                 "lauraceae", "rosaceae", "sapindaceae", "ulmaceae")
)
```

The output is a list with three plots stored. Figure \ref{fig:bw-posterior-viz-beta0} The element `beta_0` gives the baseline growth intercept $\beta_0$, i.e., how fast an individual of each group grows independent of DBH).

```{r bw-posterior-viz-beta0, out.width="50%", fig.cap="Posterior distribution of beta0."}
posterior_plots[["beta_0"]]
```

Figure \ref{fig:bw-posterior-viz-beta-dbh} Next `beta_dbh` gives the slope for DBH slope $\beta_{dbh,i}$ for each group.

```{r bw-posterior-viz-beta-dbh, out.width="100%", fig.cap="Posterior distribution of betadbh."}
posterior_plots[["beta_dbh"]]
```

Finally Figure \ref{fig:bw-posterior-viz-lambda} `lambda` gives the competition coefficients $\lambda$.

```{r bw-posterior-viz-lambda, out.width="100%", fig.cap="Posterior distribution of lambda."}
posterior_plots[["lambda"]]
```

### SCBI

We revisit the posterior parameters for the SCBI from Section {model-fit-predict}, but this time only focus on the $\lambda$ competition coefficients.

```{r scbi-posterior-viz-1}
posterior_plots_scbi <- plot_posterior_parameters(
  posterior_param = posterior_param_scbi,
  sp_to_plot = c("quru", "litu", "cagl", "cato")
)
```

```{r scbi-posterior-viz-lambda, out.width="100%", fig.cap="Posterior distribution of lambda."}
posterior_plots_scbi[["lambda"]]
```

Add explanation here.

HEY BERT PICK IT UP HERE

# Conclusions

# Acknowledgments

# References

### 2021-01-08 etc/paper/forestecology.tex
output pdf and tex files

% !TeX program = pdfLaTeX
\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx,psfrag,epsf}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{textcomp}
\usepackage[hyphens]{url} % not crucial - just used below for the URL
\usepackage{hyperref}
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

%% load any required packages here

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}

% Pandoc citation processing

\usepackage{xcolor, soul, xspace, float, subfig}

\begin{document}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\if0\blind
{
  \title{\bf \texttt{forestecology} package for modeling interspecies competition
between trees}

  \author{
        Albert Y. Kim \thanks{Albert Y. Kim is Assistant Professor, Statistical \& Data Sciences,
Smith College, Northampton, MA 01063 (e-mail:
\href{mailto:akim04@smith.edu}{\nolinkurl{akim04@smith.edu}}). This work
was not supported by any grant. The authors thank numerous colleagues
and students for their support.} \\
    Program in Statistical \& Data Sciences, Smith College\\
     and \\     David Allen \\
    Biology Department, Middlebury College\\
      }
  \maketitle
} \fi

\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf \texttt{forestecology} package for modeling interspecies competition
between trees}
  \end{center}
  \medskip
} \fi

\bigskip
\begin{abstract}
We provide a computational exercise suitable for early introduction in
an undergraduate statistics or data science course that allows students
to `play the whole game' of data science: performing both data
collection and data analysis. While many teaching resources exist for
data analysis, such resources are not as abundant for data collection
given the inherent difficulty of the task. Our proposed exercise centers
around student use of Google Calendar to collect data with the goal of
answering the question `How do I spend my time?' On the one hand, the
exercise involves answering a question with near universal appeal, but
on the other hand, the data collection mechanism is not beyond the reach
of a typical undergraduate student. A further benefit of the exercise is
that it provides an opportunity for discussions on ethical questions and
considerations that data providers and data analysts face in today's age
of large-scale internet-based data collection.
\end{abstract}

\noindent%
{\it Keywords:} forestecology, trees
\vfill

\newpage
\spacingset{1.45} % DON'T change the spacing!

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Repeat-censused forest plots offer excellent data to test neighborhood
models of tree competition \citet{allen_permutation_2020}
\citet{canham_neighborhood_2006} \citet{uriarte_spatially_2004}. Here we
describe an R package, \texttt{forestecology}, to do that. This package
implements the methods in \citet{allen_permutation_2020}. It provides: a
convenient way specify and fit models of tree growth based on
neighborhood competition; a spatial cross validation method to test and
compare model fits \citet{roberts_cross-validation_2017}; and an
ANOVA-like method to assess whether the competitor identity matters in
these models. The model is written to work with ForestGEO plot data
\citet{andersonteixeira_ctfs-forestgeo_2015}, but we envision that it
could easily be modified to work with data from other forest plots,
e.g.~the US Forest Service Forest Inventory and Analysis plots
\citet{smith_forest_2002}.

\hypertarget{example}{%
\section{Example}\label{example}}

We demonstrate the \texttt{forestecology} package's features on two data
sets, both based on inventory censuses of two sites from the Smithsonian
Institution's ForestGEO international network of 72 long‐term forest
dynamics research sites \citet{andersonteixeira_ctfs-forestgeo_2015}.
First, the Michigan Big Woods Forest Dynamics Plot located at the Edwin
S. George Reserve in Pinckney, MI, USA. The 23 ha plot is situated in
mature oak-hickory forest. The canopy is dominated by white oak (Quercus
alba), northern red oak (Quercus rubra), black oak (Quercus velutina),
shagbark hickory (Carya ovata) and pignut hickory (Carya glabra). The
most common understory tree is witch-hazel (Hamamelis virginiana)
\citet{allen_michigan_2020}. In the example below, we will preface any
data frames from this plot in with \texttt{bw\_}.

Second, the Smithsonian Conservation Biology Institute (SCBI) large
forest dynamics plot, located at the Smithsonian's National Zoo and
Conservation Biology Institute in Front Royal, VA, USA. The 25.6 ha (640
x 400 m) plot is located at the intersection of three of the major
physiographic provinces of the eastern US: the Blue Ridge, Ridge and
Valley, and Piedmont provinces and is adjacent to the northern end of
Shenandoah National Park. The forest type is typical mature secondary
eastern mixed deciduous forest, with a canopy dominated by tulip poplar
(Liriodendron tulipifera), oaks (Quercus spp.), and hickories (Carya
spp.), and an understory composed mainly of spicebush (Lindera benzoin),
paw-paw (Asimina triloba), American hornbeam (Carpinus caroliniana), and
witch hazel (Hamamelis virginiana) \citet{bourg_initial_2013}. In the
example below, we will preface any data frames from this plot in with
\texttt{scbi\_}.

We load all the necessary packages.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(forestecology)}

\CommentTok{# Load tidyverse packages:}
\KeywordTok{library}\NormalTok{(tidyverse)}
\KeywordTok{library}\NormalTok{(lubridate)}

\CommentTok{# Load spatial packages:}
\CommentTok{# devtools::install_github("rvalavi/blockCV")}
\KeywordTok{library}\NormalTok{(blockCV)}
\KeywordTok{library}\NormalTok{(sf)}
\KeywordTok{library}\NormalTok{(sfheaders)}

\CommentTok{# Load other packages:}
\KeywordTok{library}\NormalTok{(snakecase)}
\KeywordTok{library}\NormalTok{(yardstick)}
\end{Highlighting}
\end{Shaded}

\hypertarget{preprocess-census-data}{%
\subsection{Preprocess census data}\label{preprocess-census-data}}

We start by preprocessing the census data for both sites. While
ForestGEO data protocols ensure a high degree of standardization between
site, minor variations still exist
\citet{andersonteixeira_ctfs-forestgeo_2015}. While the Big Woods data
comes pre-loaded in the \texttt{forestecology} package, we load the SCBI
data as they are saved in .csv files in the SCBI-ForestGEO-Data
repository on GitHub
\citet{gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020}. In both
cases, we load the census data as R as ``tibble'' data frames thereby
ensuring a standardized input/output format that can be used across all
\texttt{tidyverse} packages \citet{wickham_welcome_2019}.

Furthermore, we ensure that the different variables have the correct
names, types (\texttt{dbl}, \texttt{data}, \texttt{factor}).

\hypertarget{big-woods}{%
\subsubsection{Big Woods}\label{big-woods}}

We load census data from 2008 and 2014 saved in the package, then merge
species data (genus, species, linnean classification, family, etc).

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{data}\NormalTok{(bw_census_}\DecValTok{2008}\NormalTok{, bw_census_}\DecValTok{2014}\NormalTok{, bw_species)}

\CommentTok{# Append additional species data}
\NormalTok{bw_census_}\DecValTok{2008}\NormalTok{ <-}\StringTok{ }\NormalTok{bw_census_}\DecValTok{2008} \OperatorTok{%>%}
\StringTok{  }\KeywordTok{left_join}\NormalTok{(bw_species, }\DataTypeTok{by =} \StringTok{"sp"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\KeywordTok{c}\NormalTok{(genus, species, latin))}
\end{Highlighting}
\end{Shaded}

\hypertarget{scbi-data}{%
\subsubsection{SCBI}\label{scbi-data}}

We load census data from 2008 and 2014 from \texttt{.csv} files saved
from GitHub on November 20, 2020. Furthermore, we perform two additional
pre-processing steps. First, in order to speed up computation for
purposes of this example, we only consider a 9 ha subsection of the 25.6
ha of the SCBI site: \texttt{gx} from 0--300 instead of 0--400 and
\texttt{gy} from 300--600 instead of 0--640. Second, in order to
standardize comparisons between Big Woods and SCBI, we convert the units
of dbh from mm to cm. \footnote{A rule of thumb to ascertain if dbh is
  in mm or cm is to verify if the smallest non-zero and non-missing
  measurement is 1 or 10. If the former, then cm. If the later, then mm.
  This is because ForestGEO protocols state that only trees with dbh
  greater or equal to 1cm should be included in censuses. }

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{scbi_}\DecValTok{2013}\NormalTok{ <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem2.csv"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(treeID, stemID, sp, ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
    \DataTypeTok{date =}\NormalTok{ ExactDate,}
    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh),}
    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date)}
\NormalTok{  ) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{dbh =}\NormalTok{ dbh }\OperatorTok{/}\StringTok{ }\DecValTok{10}\NormalTok{)}

\NormalTok{scbi_}\DecValTok{2018}\NormalTok{ <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem3.csv"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(treeID, stemID, sp, ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
    \DataTypeTok{date =}\NormalTok{ ExactDate,}
    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh),}
    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date)}
\NormalTok{  ) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{dbh =}\NormalTok{ dbh }\OperatorTok{/}\StringTok{ }\DecValTok{10}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{compute-annual-growth}{%
\subsection{Compute annual growth}\label{compute-annual-growth}}

For each plot we then compute average annual growth between the two
censuses using the \texttt{compute\_growth()} function. This function
takes the two census data frames as well as a character indicating which
variable in both data frames uniquely identifies each stem. This
function returns a single data frame that includes a numerical variable
\texttt{growth} reflecting the average annual dbh growth (in cm) of all
trees alive at both time points. Furthermore, variables that (in theory)
remain unchanged between censuses appear only once, such as location
variables \texttt{gx} and \texttt{gy}; as well as species-related
variables. Variables that should change between censuses are suffixed
with \texttt{1} and \texttt{2} indicating the earlier and later
censuses, such as \texttt{dbh1/dbh2} and \texttt{codes1/codes2}. Here
the resulting data frames are named with some variation of
\texttt{growth\_df}.

After computing the average annual growth for each tree, we ensure to
convert all variables denote species from type character to factors;
this is to ensure that issues of rare species being accounted for in
both training and test sets in our upcoming cross-validation step (see
Section REF)

\hypertarget{big-woods-1}{%
\subsubsection{Big Woods}\label{big-woods-1}}

In the case of Big Woods data, we first remove all trees that were
re-sprouts in the later (2014) census. Additionally, we have included
three classification of tree species: \texttt{species}, \texttt{family},
and \texttt{trait\_group}. DESCRIBE THESE

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{bw_census_}\DecValTok{2014}\NormalTok{ <-}\StringTok{ }\NormalTok{bw_census_}\DecValTok{2014} \OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{))}

\NormalTok{bw_growth_df <-}
\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(bw_census_}\DecValTok{2008}\NormalTok{, bw_census_}\DecValTok{2014}\NormalTok{, }\DataTypeTok{id =} \StringTok{"treeID"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\CommentTok{# Convert all variables denoting species to factors}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
    \DataTypeTok{sp =}\NormalTok{ sp }\OperatorTok{%>%}\StringTok{ }\KeywordTok{to_any_case}\NormalTok{() }\OperatorTok{%>%}\StringTok{ }\KeywordTok{as.factor}\NormalTok{(),}
    \DataTypeTok{species =}\NormalTok{ sp,}
    \DataTypeTok{family =} \KeywordTok{as.factor}\NormalTok{(family),}
    \DataTypeTok{trait_group =} \KeywordTok{as.factor}\NormalTok{(trait_group)}
\NormalTok{  ) }\OperatorTok{%>%}
\StringTok{  }\CommentTok{# Drop unnecessary variables}
\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\NormalTok{stemID)}
\end{Highlighting}
\end{Shaded}

\hypertarget{scbi}{%
\subsubsection{SCBI}\label{scbi}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{scbi_growth_df <-}
\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(scbi_}\DecValTok{2013}\NormalTok{, scbi_}\DecValTok{2018}\NormalTok{, }\StringTok{"stemID"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\CommentTok{# Convert all variables denoting species to factors}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{as.factor}\NormalTok{(sp))}
\end{Highlighting}
\end{Shaded}

\hypertarget{comparison}{%
\subsubsection{Comparison}\label{comparison}}

Figure \ref{fig:growth-histogram} displays histograms comparing the
distribution of average annual growth at both sites. Observe that
average annual growth appears higher at the Big Woods site.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Both Big Woods & SCBI}
\KeywordTok{bind_rows}\NormalTok{(}
\NormalTok{  bw_growth_df }\OperatorTok{%>%}\StringTok{ }\KeywordTok{st_drop_geometry}\NormalTok{() }\OperatorTok{%>%}\StringTok{ }\KeywordTok{select}\NormalTok{(growth) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{Site =} \StringTok{"Big Woods"}\NormalTok{),}
\NormalTok{  scbi_growth_df }\OperatorTok{%>%}\StringTok{ }\KeywordTok{st_drop_geometry}\NormalTok{() }\OperatorTok{%>%}\StringTok{ }\KeywordTok{select}\NormalTok{(growth) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{Site =} \StringTok{"SCBI"}\NormalTok{)}
\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ growth, }\DataTypeTok{y =}\NormalTok{ ..density.., }\DataTypeTok{fill =}\NormalTok{ Site)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_histogram}\NormalTok{(}\DataTypeTok{alpha =} \FloatTok{0.5}\NormalTok{, }\DataTypeTok{position =} \StringTok{"identity"}\NormalTok{, }\DataTypeTok{binwidth =} \FloatTok{0.05}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{x =} \StringTok{"Average annual growth in dbh (cm per yr)"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{coord_cartesian}\NormalTok{(}\DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(}\OperatorTok{-}\FloatTok{0.5}\NormalTok{, }\DecValTok{1}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{Figures/growth-histogram-1}

}

\caption{Distribution of average annual growth in DBH for both sites.}\label{fig:growth-histogram}
\end{figure}

\hypertarget{spatial-information}{%
\subsection{Add spatial information}\label{spatial-information}}

We now encode spatial information to the \texttt{growth\_df} data
frames. First, in order to control for study region edge effects, we add
``buffers'' to the periphery of the study region (cite Waller?). Our
model of interspecific competition relies on a spatial definition of who
the competitor trees are for focal trees of interest. Since certain
explanatory variables such as biomass are cumulative, we must ensure
that all trees being modeled are not biased to have different neighbor
structures. This is a particular concern for trees at the boundary of
study regions, which will not have the same number of neighbors as trees
in the internal part of the study region.

Second, our ultimate method for model assessment will rely on estimates
of model error as generated by cross-validation. Conventional
cross-validation schemes assign observations to folds by resampling
individual observations at random. However, underlying this scheme is an
assumption that the observations are independent. In the case of forest
census data, observations exhibit spatial autocorrelation, and thus this
dependence must be incorporated in our resampling scheme in spatial
cross-validation \citet{roberts2017} \citet{pohjankukka2017}. We will
therefore associate portions of the study region to spatial folds.

To these two ends, we define two constants, both of which are in the
same units as the \texttt{gx} and \texttt{gy} variables (most often
meters).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{max_dist <-}\StringTok{ }\FloatTok{7.5}
\NormalTok{cv_fold_size <-}\StringTok{ }\DecValTok{100}
\end{Highlighting}
\end{Shaded}

The first constant is \texttt{max\_dist} which defines the maximum
distance for a tree's competitive neighborhood. Trees within this
distance of each other are assumed to compete while those farther than
this distance apart do not. Put differently, all trees within
\texttt{max\_dist} of a focal tree will be considered its competitors
(see below). Other studies have estimated the value of
\texttt{max\_dist}; we use an average of estimated values
\citet{canham_neighborhood_2004}, \citet{uriarte_spatially_2004},
\citet{tatsumi2013}, \citet{canham_neighborhood_2006}.

Furthermore, \texttt{max\_dist} will define the size of all buffers
considered, which will be encoded as a binary variable \texttt{buffer}
as computed by the \texttt{add\_buffer\_variable()} function. This
function takes as input the main \texttt{growth\_df} data frame, the
\texttt{size} of the buffer which we set as \texttt{max\_dist}, and the
boundary of the study region encoded as a simple features polygon
\citet{pebesma_simple_2018}. DESCRIBE SF PACKAGE. In the Big Woods
example below we will use a pre-loaded simple features polygon while for
the SCBI example we present example code on how to manually construct
one.

The second constant is \texttt{cv\_fold\_size} which defines the length
and width of the spatial folds (note that for now the spatial folds are
restricted be squares). We will then use this constant to associate each
observed tree to one of \(k\) folds in the respective study region. In
the Big Woods example below we will use the \texttt{blockCV} R package
that has implemented spatial cross-validation while for the SCBI we will
do this manually \citet{valavi2019}.

\hypertarget{big-woods-2}{%
\subsubsection{Big Woods}\label{big-woods-2}}

First, we indicate which trees are part of the buffer. This necessitates
information about the study region boundary. In this case, we use a
\texttt{sf\_polygon} object \texttt{bw\_study\_region} which comes
pre-loaded in the \texttt{forestecology} packages. After loading
\texttt{bw\_study\_region}, we illustrate the results of the
\texttt{add\_buffer\_variable()} function in Figure
\ref{fig:bw-define-buffer}. Trees on the periphery denote with lighter
colors are part of the buffer and will not be considered as ``focal''
trees of interest going forward; they will only be considered as
competitor trees.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{data}\NormalTok{(bw_study_region)}

\NormalTok{bw_growth_df <-}\StringTok{ }\NormalTok{bw_growth_df }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{direction =} \StringTok{"in"}\NormalTok{, }\DataTypeTok{size =}\NormalTok{ max_dist, }\DataTypeTok{region =}\NormalTok{ bw_study_region)}

\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ bw_growth_df }\OperatorTok{%>%}\StringTok{ }\KeywordTok{sample_frac}\NormalTok{(}\FloatTok{0.2}\NormalTok{), }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{col =}\NormalTok{ buffer), }\DataTypeTok{size =} \FloatTok{0.5}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{Figures/bw-define-buffer-1}

}

\caption{Buffer region for Big Woods study region.}\label{fig:bw-define-buffer}
\end{figure}

Second, we associate each tree to spatial cross validation folds. In
this case, we use the \texttt{spatialBlock()} function from the
\texttt{blockCV} package to define the spatial grid which

THIS IS A MESS. We use the \citet{valavi_blockcv_2019}, whose elements
will act as the folds in our leave-one-out (by ``one'' we mean ``one
grid block'') cross-validation scheme. The upshot here is we add
\texttt{foldID} to \texttt{growth\_df} which identifies which fold each
individual is in, and the creation of a \texttt{cv\_grid\_sf} object
which gives the geometry of the cross validation grid.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{set.seed}\NormalTok{(}\DecValTok{76}\NormalTok{)}
\NormalTok{bw_spatialBlock <-}\StringTok{ }\KeywordTok{spatialBlock}\NormalTok{(}
  \DataTypeTok{speciesData =}\NormalTok{ bw_growth_df, }\DataTypeTok{theRange =}\NormalTok{ cv_fold_size,}
  \DataTypeTok{k =} \DecValTok{28}\NormalTok{, }\DataTypeTok{xOffset =} \FloatTok{0.5}\NormalTok{, }\DataTypeTok{yOffset =} \DecValTok{0}\NormalTok{,}
  \DataTypeTok{verbose =} \OtherTok{FALSE}
\NormalTok{)}

\CommentTok{# Add foldID to each tree}
\NormalTok{bw_growth_df <-}\StringTok{ }\NormalTok{bw_growth_df }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ bw_spatialBlock}\OperatorTok{$}\NormalTok{foldID)}

\CommentTok{# Visualize grid. Why does fold 19 repeat?}
\NormalTok{bw_spatialBlock}\OperatorTok{$}\NormalTok{plots }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ bw_growth_df }\OperatorTok{%>%}\StringTok{ }\KeywordTok{sample_frac}\NormalTok{(}\FloatTok{0.2}\NormalTok{), }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{col =} \KeywordTok{factor}\NormalTok{(foldID)), }\DataTypeTok{size =} \FloatTok{0.1}\NormalTok{)}

\CommentTok{# Remove empty folds}
\NormalTok{bw_growth_df <-}\StringTok{ }\NormalTok{bw_growth_df }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\NormalTok{foldID }\OperatorTok{%in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{19}\NormalTok{, }\DecValTok{23}\NormalTok{, }\DecValTok{21}\NormalTok{, }\DecValTok{17}\NormalTok{, }\DecValTok{8}\NormalTok{, }\DecValTok{19}\NormalTok{)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =} \KeywordTok{as.character}\NormalTok{(foldID))}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics[width=1\linewidth]{Figures/bw-define-cv-folds-1} \includegraphics[width=1\linewidth]{Figures/bw-define-cv-folds-2} \end{center}

Separately, we save the spatial cross-validation grid as an
\texttt{sf\_polygon} object \texttt{bw\_cv\_grid}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{bw_cv_grid <-}\StringTok{ }\NormalTok{bw_spatialBlock}\OperatorTok{$}\NormalTok{blocks }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{st_as_sf}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\hypertarget{scbi-1}{%
\subsubsection{SCBI}\label{scbi-1}}

First, we indicate which trees are part of the buffer. In this case
however we manually define the study region boundary based on the
subregion we defined in Section \ref{scbi-data} and create an
\texttt{sf\_polygon} object using the \texttt{sf\_polygon()} function
from the \texttt{sfheaders} package. Figure \ref{fig:scbi-define-buffer}
displays the resulting buffer trees.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{scbi_study_region <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}
  \DataTypeTok{x =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{),}
  \DataTypeTok{y =} \KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{300}\NormalTok{)}
\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{sf_polygon}\NormalTok{()}

\NormalTok{scbi_growth_df <-}\StringTok{ }\NormalTok{scbi_growth_df }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{direction =} \StringTok{"in"}\NormalTok{, }\DataTypeTok{size =}\NormalTok{ max_dist, }\DataTypeTok{region =}\NormalTok{ scbi_study_region)}

\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ scbi_growth_df, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{col =}\NormalTok{ buffer), }\DataTypeTok{size =} \FloatTok{0.5}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{figure}

{\centering \includegraphics[width=0.5\linewidth]{Figures/scbi-define-buffer-1}

}

\caption{Buffer region for SCBI study region.}\label{fig:scbi-define-buffer}
\end{figure}

Second, we associate each tree to spatial cross validation folds. In
this case we manually define a spatial crossvaliation grid. Figure
\ref{fig:scbi-define-cv-folds} displays the resulting cross-validation
folds along with the buffer from Figure \ref{fig:scbi-define-buffer}.

Here we manually define the spatial cross-validation grid as an
\texttt{sf\_polygon} object \texttt{scbi\_cv\_grid}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{fold1 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{600}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{600}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{sf_polygon}\NormalTok{() }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \DecValTok{1}\NormalTok{)}
\NormalTok{fold2 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{600}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{sf_polygon}\NormalTok{() }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \DecValTok{2}\NormalTok{)}
\NormalTok{scbi_cv_grid <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(fold1, fold2)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{scbi_spatialBlock <-}\StringTok{ }\KeywordTok{spatialBlock}\NormalTok{(}
  \DataTypeTok{speciesData =}\NormalTok{ scbi_growth_df,}
  \DataTypeTok{k =} \DecValTok{2}\NormalTok{,}
  \DataTypeTok{verbose =} \OtherTok{FALSE}\NormalTok{,}
  \DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{,}
  \CommentTok{# Note new arguments:}
  \DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{,}
  \DataTypeTok{blocks =}\NormalTok{ scbi_cv_grid}
\NormalTok{)}

\CommentTok{# Add foldID to each tree}
\NormalTok{scbi_growth_df <-}\StringTok{ }\NormalTok{scbi_growth_df }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ scbi_spatialBlock}\OperatorTok{$}\NormalTok{foldID)}

\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_sf_text}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ scbi_growth_df, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{label =}\NormalTok{ foldID, }\DataTypeTok{col =}\NormalTok{ buffer)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ scbi_cv_grid, }\DataTypeTok{fill =} \StringTok{"transparent"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{figure}

{\centering \includegraphics[width=0.5\linewidth]{Figures/scbi-define-cv-folds-1}

}

\caption{Buffer region for SCBI study region.}\label{fig:scbi-define-cv-folds}
\end{figure}

\hypertarget{define-focal-versus-competitor-trees}{%
\subsection{Define focal versus competitor
trees}\label{define-focal-versus-competitor-trees}}

Next we define \texttt{focal\_vs\_comp} data frames which connects each
focal tree in the \texttt{growth\_df} data frames to the trees in its
competitive neighborhood range as defined by the \texttt{max\_dist}
constant. So for example, if \texttt{growth\_df} consisted of two focal
trees with two and three neighbors with \texttt{max\_dist} respectively,
\texttt{focal\_vs\_comp} would be a data frame of 5 rows connecting each
focal tree to it's competitors. The \texttt{create\_focal\_vs\_comp()}
function makes this connection taking as inputs the \texttt{growth\_df}
data frame; the \texttt{max\_dist} constant defining competitive range;
\texttt{cv\_grid\_sf}, giving the cross validation grid; and the
\texttt{id} variable.

\hypertarget{big-woods-3}{%
\subsubsection{Big Woods}\label{big-woods-3}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_bw <-}\StringTok{ }\NormalTok{bw_growth_df }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(max_dist, }\DataTypeTok{cv_grid_sf =}\NormalTok{ bw_cv_grid, }\DataTypeTok{id =} \StringTok{"treeID"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

TODO: Figure out how to show this data frame's contents.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{head}\NormalTok{(focal_vs_comp_bw)}
\CommentTok{## # A tibble: 6 x 10}
\CommentTok{##   focal_ID focal_sp   dbh foldID                  geometry growth}
\CommentTok{##      <dbl> <fct>    <dbl> <chr>                    <POINT>  <dbl>}
\CommentTok{## 1        1 white_o~  41.2 12                   (8.7 107.5)  0.404}
\CommentTok{## 2        1 white_o~  41.2 12                   (8.7 107.5)  0.404}
\CommentTok{## 3        1 white_o~  41.2 12                   (8.7 107.5)  0.404}
\CommentTok{## 4        1 white_o~  41.2 12                   (8.7 107.5)  0.404}
\CommentTok{## 5        1 white_o~  41.2 12                   (8.7 107.5)  0.404}
\CommentTok{## 6        1 white_o~  41.2 12                   (8.7 107.5)  0.404}
\CommentTok{## # ... with 4 more variables: comp_ID <dbl>, dist <dbl>,}
\CommentTok{## #   comp_sp <fct>, comp_basal_area <dbl>}
\end{Highlighting}
\end{Shaded}

\hypertarget{scbi-2}{%
\subsubsection{SCBI}\label{scbi-2}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{scbi_growth_df }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(max_dist, }\DataTypeTok{cv_grid_sf =}\NormalTok{ scbi_cv_grid, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

TODO: Figure out how to show this data frame's contents.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{head}\NormalTok{(focal_vs_comp_scbi)}
\CommentTok{## # A tibble: 6 x 10}
\CommentTok{##   focal_ID focal_sp   dbh foldID                  geometry growth}
\CommentTok{##      <dbl> <fct>    <dbl>  <int>                   <POINT>  <dbl>}
\CommentTok{## 1        4 nysy      13.6      1              (14.2 428.5)  0.103}
\CommentTok{## 2        4 nysy      13.6      1              (14.2 428.5)  0.103}
\CommentTok{## 3        4 nysy      13.6      1              (14.2 428.5)  0.103}
\CommentTok{## 4        4 nysy      13.6      1              (14.2 428.5)  0.103}
\CommentTok{## 5        4 nysy      13.6      1              (14.2 428.5)  0.103}
\CommentTok{## 6        4 nysy      13.6      1              (14.2 428.5)  0.103}
\CommentTok{## # ... with 4 more variables: comp_ID <dbl>, dist <dbl>,}
\CommentTok{## #   comp_sp <fct>, comp_basal_area <dbl>}
\end{Highlighting}
\end{Shaded}

\hypertarget{model-fit-predict}{%
\subsection{Fit model and make predictions}\label{model-fit-predict}}

Next we fit the following linear model to the dbh of each focal tree.
Let \(i = 1, \ldots, n_j\) index all \(n_j\) trees of
\texttt{focal\textquotesingle{}\textquotesingle{}\ species\ group\ \$j\$;\ let\ \$j\ =\ 1,\ \textbackslash{}ldots,\ J\$\ index\ all\ \$J\$\ focal\ species\ groups;\ and\ let\ \$k\ =\ 1,\ \textbackslash{}ldots,\ K\$\ index\ all\ \$K\$}competitor''
species groups. We modeled the growth in diameter per year \(y_{ij}\)
(in centimeters per year) of the \(i^{th}\) tree of focal species group
\(j\) as a linear model \(f\) of the following covariates
\(\vec{x}_{ij}\)

\[
\newcommand{\dbh}{\text{DBH}}
\newcommand{\biomass}{\text{biomass}}
\newcommand{\BA}{\text{BA}}
y_{ij} = f(\vec{x}_{ij}) + \epsilon_{ij} = \beta_{0,j} + \beta_{\dbh,j} \cdot \dbh_{ij} + \sum_{k=1}^{K} \lambda_{jk} \cdot \BA_{ijk} + \epsilon_{ij}
\]

We estimate the model's parameters using Bayesian linear regression
implemented in the \texttt{fit\_bayesian\_model()} function. TODO:
define all parameters

For this linear model's case, there exists a closed form solution as
described here. As such, the \texttt{fit\_bayesian\_model()} function
using matrix algebra to obtain all parameter estimates, rather than
computationally expensive Monte Carlo approximations. The inputs to this
function are a \texttt{focal\_vs\_comp} data frame,
\texttt{prior\_param} a list of priors, and a boolean flag
\texttt{run\_shuffle} on whether or not to run competitor-species
identity permutations which we will demonstrate below on the Michigan
Big Woods data. This function returns the posterior means of all
parameters.

Using these posterior means, we then use the posterior predictive
distribution to obtain fitted/predicted values \(\widehat{y}\) of the
dbh for each focal tree using the \texttt{predict\_bayesian\_model()}.
These \(\widehat{y}\) can then be compared to the observed \(y\) dbh's
to compute the root mean-square error, a measure of a model's predictive
error which has the same units as the observed data \(y\).

\hypertarget{big-woods-4}{%
\subsubsection{Big Woods}\label{big-woods-4}}

For the Michigan Big Woods data we present two use cases of the model
fitting and prediction scheme. The first use case is the simplest where
we assess the fit of the model using root mean squared error. The second
use case then answers the question of whether species competitor
identity matters using permutation test.

For the first use case, we fit the linear model specified in Equation
XXX to our data frame of type \texttt{focal\_vs\_comp}. This
input/outputs of the \texttt{fit\_bayesian\_model()} function are lists
of the prior/posterior means of parameters of the linear regression
specified in XXX. Generally speaking, there are two classes of
regression parameters: \(\beta\) main effects and \(\lambda\)
competitive effects. In the upcoming Section
\ref{viz-posterior-distributions}, we will present code visualizing this
posterior distributions.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{posterior_param_bw <-}\StringTok{ }\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{fit_bayesian_model}\NormalTok{(}\DataTypeTok{prior_param =} \OtherTok{NULL}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

This output of posterior parameters for the specified competition model
are then used along with the posterior predictive distribution encoded
in \texttt{predict\_bayesian\_model()} to return predicted growths for
each individual tree. We join these predicted growths to the original
growth data frame.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{predictions <-}\StringTok{ }\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{predict_bayesian_model}\NormalTok{(}\DataTypeTok{posterior_param =}\NormalTok{ posterior_param_bw) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{right_join}\NormalTok{(bw_growth_df, }\DataTypeTok{by =} \KeywordTok{c}\NormalTok{(}\StringTok{"focal_ID"}\NormalTok{ =}\StringTok{ "treeID"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

We then use the \texttt{rmse()} function from the \texttt{yardstick}
package to obtain the root mean squared error of the observed versus
fitted values of growth.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{predictions }\OperatorTok{%>%}
\StringTok{  }\NormalTok{yardstick}\OperatorTok{::}\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
\CommentTok{## [1] 0.148145}
\end{Highlighting}
\end{Shaded}

The second use case is near identical to the first, but with a small
change in the code to test whether the identity of the competitor
matters. By adding a \texttt{run\_shuffle\ =\ TRUE} argument to
\texttt{fit\_bayesian\_model()}, for each focal tree its competitor
trees' species identity will be ``shuffled'' randomly much like in a
permutation test. By shuffling these species labels we are effectively
fitting the model under a null model that competitor species identity
does not matter. If the ``shuffled'' RMSE's are consistently lower than
the unshuffled RMSE corresponding to the observed data, then we have
evidence to suggest that competitor identity matters to competitive
interactions.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{posterior_param_bw_shuffle <-}\StringTok{ }\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{fit_bayesian_model}\NormalTok{(}\DataTypeTok{prior_param =} \OtherTok{NULL}\NormalTok{, }\DataTypeTok{run_shuffle =} \OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{predictions_shuffle <-}\StringTok{ }\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{predict_bayesian_model}\NormalTok{(}\DataTypeTok{posterior_param =}\NormalTok{ posterior_param_bw_shuffle) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{right_join}\NormalTok{(bw_growth_df, }\DataTypeTok{by =} \KeywordTok{c}\NormalTok{(}\StringTok{"focal_ID"}\NormalTok{ =}\StringTok{ "treeID"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{predictions_shuffle }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
\CommentTok{## [1] 0.1505383}
\end{Highlighting}
\end{Shaded}

The RMSE is fact lower for the non-shuffled version, indicative of a
better model fit. This gives support for the idea that competitor
identity does matter for competitive interactions. In
\citet{allen_permutation_2020} we run this shuffle a large number of
times to construct a full permutation distribution to show that this
difference is robust to resampling variation.

\hypertarget{scbi-3}{%
\subsubsection{SCBI}\label{scbi-3}}

In the case of the SCBI data, we once again perform the same model
fitting and computing of fitted growths as with the Big Woods data, but
this time we map the residuals of the observed minus fitted values to
look for spatial patterns.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{posterior_param_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{fit_bayesian_model}\NormalTok{(}\DataTypeTok{prior_param =} \OtherTok{NULL}\NormalTok{, }\DataTypeTok{run_shuffle =} \OtherTok{FALSE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{scbi_growth_df_noCV <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{predict_bayesian_model}\NormalTok{(}\DataTypeTok{posterior_param =}\NormalTok{ posterior_param_scbi) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{right_join}\NormalTok{(scbi_growth_df, }\DataTypeTok{by =} \KeywordTok{c}\NormalTok{(}\StringTok{"focal_ID"}\NormalTok{ =}\StringTok{ "stemID"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{scbi_growth_df_noCV }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
\CommentTok{## [1] 0.1280644}
\end{Highlighting}
\end{Shaded}

In Figures \ref{fig:scbi-model-residuals} and
\ref{fig:scbi-model-residuals-2} we present the residuals.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{ggplot}\NormalTok{(scbi_growth_df_noCV, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ growth, }\DataTypeTok{y =}\NormalTok{ growth_hat)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_point}\NormalTok{(}\DataTypeTok{size =} \FloatTok{0.5}\NormalTok{, }\DataTypeTok{color =} \KeywordTok{rgb}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\FloatTok{0.25}\NormalTok{)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{stat_smooth}\NormalTok{(}\DataTypeTok{method =} \StringTok{"lm"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_abline}\NormalTok{(}\DataTypeTok{slope =} \DecValTok{1}\NormalTok{, }\DataTypeTok{intercept =} \DecValTok{0}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{coord_fixed}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{labs}\NormalTok{(}
    \DataTypeTok{x =} \StringTok{"Observed growth in dbh"}\NormalTok{, }\DataTypeTok{y =} \StringTok{"Predicted growth in dbh"}\NormalTok{,}
    \DataTypeTok{title =} \StringTok{"Predicted vs Observed Growth"}
\NormalTok{  )}
\end{Highlighting}
\end{Shaded}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{Figures/scbi-model-residuals-1}

}

\caption{Spatial distribution of residuals for model applied to SCBI data.}\label{fig:scbi-model-residuals}
\end{figure}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{scbi_growth_df_noCV }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{st_as_sf}\NormalTok{() }\OperatorTok{%>%}
\StringTok{  }\CommentTok{# }\AlertTok{TODO}\CommentTok{: Need to investigate missingness}
\StringTok{  }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{is.na}\NormalTok{(growth_hat)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
    \DataTypeTok{error =}\NormalTok{ growth }\OperatorTok{-}\StringTok{ }\NormalTok{growth_hat,}
    \DataTypeTok{error_bin =} \KeywordTok{cut_number}\NormalTok{(error, }\DataTypeTok{n =} \DecValTok{5}\NormalTok{),}
    \DataTypeTok{error_compress =} \KeywordTok{ifelse}\NormalTok{(error }\OperatorTok{<}\StringTok{ }\FloatTok{-0.75}\NormalTok{, }\FloatTok{-0.75}\NormalTok{, }\KeywordTok{ifelse}\NormalTok{(error }\OperatorTok{>}\StringTok{ }\FloatTok{0.75}\NormalTok{, }\FloatTok{0.75}\NormalTok{, error))}
\NormalTok{  ) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{col =}\NormalTok{ error_compress), }\DataTypeTok{size =} \DecValTok{1}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme_bw}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_color_gradient2}\NormalTok{(}
    \DataTypeTok{low =} \StringTok{"#ef8a62"}\NormalTok{, }\DataTypeTok{mid =} \StringTok{"#f7f7f7"}\NormalTok{, }\DataTypeTok{high =} \StringTok{"#67a9cf"}\NormalTok{,}
    \DataTypeTok{name =} \KeywordTok{expression}\NormalTok{(}\KeywordTok{paste}\NormalTok{(}\StringTok{"Residual (cm "}\NormalTok{, y}\OperatorTok{^}\NormalTok{\{}\OperatorTok{-}\DecValTok{1}\NormalTok{\}, }\StringTok{")"}\NormalTok{)),}
    \DataTypeTok{breaks =} \KeywordTok{seq}\NormalTok{(}\DataTypeTok{from =} \FloatTok{-0.75}\NormalTok{, }\DataTypeTok{to =} \FloatTok{0.75}\NormalTok{, }\DataTypeTok{by =} \FloatTok{0.25}\NormalTok{),}
    \DataTypeTok{labels =} \KeywordTok{c}\NormalTok{(}\StringTok{"< -0.75"}\NormalTok{, }\StringTok{"-0.5"}\NormalTok{, }\StringTok{"0.25"}\NormalTok{, }\StringTok{"0"}\NormalTok{, }\StringTok{"0.25"}\NormalTok{, }\StringTok{"0.5"}\NormalTok{, }\StringTok{"> 0.75"}\NormalTok{)}
\NormalTok{  ) }\OperatorTok{+}
\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{x =} \StringTok{"Meter"}\NormalTok{, }\DataTypeTok{y =} \StringTok{"Meter"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{Figures/scbi-model-residuals-2-1}

}

\caption{Spatial distribution of residuals for model applied to SCBI data part 2.}\label{fig:scbi-model-residuals-2}
\end{figure}

\hypertarget{run-spatial-cross-validation}{%
\subsection{Run spatial
cross-validation}\label{run-spatial-cross-validation}}

The model fits and predictions in Section \ref{model-fit-predict} all
suffer from a common failing: they use the same data to both fit the
model and to assess the model's performance using the RMSE. As argued by
\citet{roberts_cross-validation_2017}, this can lead to overly
optimistic assessments of model quality as the models can be overfit, in
particular in situations where spatial-autocorrelation is present. To
mitigate the effects of such overfitting, we use a spatially block
cross-validation algorithm implemented in the \texttt{run\_cv()}. This
function at its core uses the same model fitting implemented in the
\texttt{fit\_bayesian\_model()} function, however trains the model on
\(k-1\) spatial folds of the train and returns fitted values for the
test data. Recall that the spatial blocking scheme wass encoded in
Section \ref{spatial-information}.

\hypertarget{big-woods-5}{%
\subsubsection{Big Woods}\label{big-woods-5}}

Applying this spatially cross-validated model fit yields an RMSE is
higher than that when the model is fit without cross validation. In
other words, our model fits in \ref{model-fit-predict} were overly
optimistic in the model's fitting power, whereas a cross-validated
results yield an estimate that is closer to the truth. See
\citet{allen_permutation_2020} for more discussion of this.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{cv_bw <-}\StringTok{ }\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{max_dist =}\NormalTok{ max_dist, }\DataTypeTok{cv_grid =}\NormalTok{ bw_cv_grid) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{right_join}\NormalTok{(bw_growth_df, }\DataTypeTok{by =} \KeywordTok{c}\NormalTok{(}\StringTok{"focal_ID"}\NormalTok{ =}\StringTok{ "treeID"}\NormalTok{))}

\NormalTok{cv_bw }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
\CommentTok{## [1] 0.1533511}
\end{Highlighting}
\end{Shaded}

\hypertarget{scbi-4}{%
\subsubsection{SCBI}\label{scbi-4}}

Observe once again that this RMSE is much higher than that for the above
SCBI model fit without cross-validation.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{cv_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{max_dist =}\NormalTok{ max_dist, }\DataTypeTok{cv_grid =}\NormalTok{ scbi_cv_grid) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{right_join}\NormalTok{(scbi_growth_df, }\DataTypeTok{by =} \KeywordTok{c}\NormalTok{(}\StringTok{"focal_ID"}\NormalTok{ =}\StringTok{ "treeID"}\NormalTok{))}

\NormalTok{cv_scbi }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
\CommentTok{## [1] 0.1494775}
\end{Highlighting}
\end{Shaded}

\hypertarget{viz-posterior-distributions}{%
\subsection{Visualize posterior
distributions}\label{viz-posterior-distributions}}

Lastly, we return to the model fits from Section \ref{model-fit-predict}
and present tools to visually explore the posterior distributions of all
parameters in our model. There are two main groups of parameters to
consider. The \(\beta\) coefficients tell us about how fast each species
grows and how this depends on DBH while the full matrix of \(\lambda\)
values describe the competitive effects between pairs of species. There
is a rich literature on this matrix (cite).

DO WE NEED TO DESCRIBE MECHANICS? Because of the structure of the
\texttt{bw\_fit\_model} object we cannot simply draw these curves based
on the posterior distribution. \texttt{bw\_fit\_model()} gives the
parameters \emph{compared} to a baseline. This is not of direct
interest. So to display these parameters, as we care about them, we have
to sample from the baseline distribution and from the comparison one to
get the posterior distribution of interest.

\hypertarget{big-woods-6}{%
\subsubsection{Big Woods}\label{big-woods-6}}

Here we re-run the model fit to the Big Woods data from Section
\ref{model-fit-predict}, but this time use ``family'' as the group for
comparison which has. This makes the posterior distributions easier to
follow. Also, surprisingly, grouping by family performed just as well as
grouping by species \citet{allen_permutation_2020}. First we re-run
\texttt{create\_focal\_vs\_comp()} and \texttt{fit\_bayesian\_model()}
with no permutation shuffling with the grouping variable as family.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_bw <-}\StringTok{ }\NormalTok{bw_growth_df }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =}\NormalTok{ family) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{max_dist =}\NormalTok{ max_dist, }\DataTypeTok{cv_grid_sf =}\NormalTok{ bw_cv_grid, }\DataTypeTok{id =} \StringTok{"treeID"}\NormalTok{)}

\NormalTok{posterior_param_bw <-}\StringTok{ }\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{fit_bayesian_model}\NormalTok{(}\DataTypeTok{prior_param =} \OtherTok{NULL}\NormalTok{, }\DataTypeTok{run_shuffle =} \OtherTok{FALSE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Now the posterior parameter outputs of \texttt{fit\_bayesian\_model()}
are passed to \texttt{plot\_posterior\_parameters()} to generate
visualizations of the posterior parameters. These visualizations are
displayed in Figure 5 of \citet{allen_permutation_2020}. For simplicity
we only plot a subset of the species families.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{posterior_plots <-}\StringTok{ }\KeywordTok{plot_posterior_parameters}\NormalTok{(}
  \DataTypeTok{posterior_param =}\NormalTok{ posterior_param_bw,}
  \DataTypeTok{sp_to_plot =} \KeywordTok{c}\NormalTok{(}\StringTok{"cornaceae"}\NormalTok{, }\StringTok{"fagaceae"}\NormalTok{, }\StringTok{"hamamelidaceae"}\NormalTok{, }\StringTok{"juglandaceae"}\NormalTok{, }
                 \StringTok{"lauraceae"}\NormalTok{, }\StringTok{"rosaceae"}\NormalTok{, }\StringTok{"sapindaceae"}\NormalTok{, }\StringTok{"ulmaceae"}\NormalTok{)}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

The output is a list with three plots stored. Figure
\ref{fig:bw-posterior-viz-beta0} The element \texttt{beta\_0} gives the
baseline growth intercept \(\beta_0\), i.e., how fast an individual of
each group grows independent of DBH).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{posterior_plots[[}\StringTok{"beta_0"}\NormalTok{]]}
\end{Highlighting}
\end{Shaded}

\begin{figure}

{\centering \includegraphics[width=0.5\linewidth]{Figures/bw-posterior-viz-beta0-1}

}

\caption{Posterior distribution of beta0.}\label{fig:bw-posterior-viz-beta0}
\end{figure}

Figure \ref{fig:bw-posterior-viz-beta-dbh} Next \texttt{beta\_dbh} gives
the slope for DBH slope \(\beta_{dbh,i}\) for each group.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{posterior_plots[[}\StringTok{"beta_dbh"}\NormalTok{]]}
\end{Highlighting}
\end{Shaded}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{Figures/bw-posterior-viz-beta-dbh-1}

}

\caption{Posterior distribution of betadbh.}\label{fig:bw-posterior-viz-beta-dbh}
\end{figure}

Finally Figure \ref{fig:bw-posterior-viz-lambda} \texttt{lambda} gives
the competition coefficients \(\lambda\).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{posterior_plots[[}\StringTok{"lambda"}\NormalTok{]]}
\end{Highlighting}
\end{Shaded}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{Figures/bw-posterior-viz-lambda-1}

}

\caption{Posterior distribution of lambda.}\label{fig:bw-posterior-viz-lambda}
\end{figure}

\hypertarget{scbi-5}{%
\subsubsection{SCBI}\label{scbi-5}}

We revisit the posterior parameters for the SCBI from Section
\{model-fit-predict\}, but this time only focus on the \(\lambda\)
competition coefficients.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{posterior_plots_scbi <-}\StringTok{ }\KeywordTok{plot_posterior_parameters}\NormalTok{(}
  \DataTypeTok{posterior_param =}\NormalTok{ posterior_param_scbi,}
  \DataTypeTok{sp_to_plot =} \KeywordTok{c}\NormalTok{(}\StringTok{"quru"}\NormalTok{, }\StringTok{"litu"}\NormalTok{, }\StringTok{"cagl"}\NormalTok{, }\StringTok{"cato"}\NormalTok{)}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{posterior_plots_scbi[[}\StringTok{"lambda"}\NormalTok{]]}
\end{Highlighting}
\end{Shaded}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{Figures/scbi-posterior-viz-lambda-1}

}

\caption{Posterior distribution of lambda.}\label{fig:scbi-posterior-viz-lambda}
\end{figure}

Add explanation here.

HEY BERT PICK IT UP HERE

\hypertarget{conclusions}{%
\section{Conclusions}\label{conclusions}}

\hypertarget{acknowledgments}{%
\section{Acknowledgments}\label{acknowledgments}}

\bibliographystyle{agsm}
\bibliography{paper.bib}

\end{document}

### 2021-01-08 etc/paper/forestecology.Rmd
Setup MEE formatting

@@
 - name: Albert Y. Kim 
-  thanks: "Albert Y.
-Kim is Assistant Professor, Statistical & Data Sciences, Smith College, Northampton, MA 01063 (e-mail: akim04@smith.edu).
-This work was not supported by any grant.
-The authors thank numerous colleagues and students for their support."
+  thanks: "Albert Y. Kim is Assistant Professor, Statistical & Data Sciences, Smith College, Northampton, MA 01063 (e-mail: akim04@smith.edu)."
   affiliation: Program in Statistical & Data Sciences, Smith College
@@
 
+abstract: 
+- "Move abstract below here after completed."
+
 keywords:
-- forestecology, trees
-
-abstract: 
-- "We provide a computational exercise suitable for early introduction in an undergraduate statistics or data science course that allows students to 'play the whole game' of data science: performing both data collection and data analysis.
-While many teaching resources exist for data analysis, such resources are not as abundant for data collection given the inherent difficulty of the task.
-Our proposed exercise centers around student use of Google Calendar to collect data with the goal of answering the question 'How do I spend my time?' On the one hand, the exercise involves answering a question with near universal appeal, but on the other hand, the data collection mechanism is not beyond the reach of a typical undergraduate student.
-A further benefit of the exercise is that it provides an opportunity for discussions on ethical questions and considerations that data providers and data analysts face in today's age of large-scale internet-based data collection."
+- forest ecology, competition, R, Rstats, tidyverse, sf, cross-validation,
 
@@
 header-includes:
-- \usepackage{xcolor, soul, xspace, float, subfig}
+- \usepackage{xcolor, soul, xspace, float, subfig, lineno}
 ---
+
+\linenumbers
 
@@
 ```
+
+# Abstract (350 words)
+
+1. set the context for and purpose of the work 
+1. indicate the approach and the methods 
+1. outline the main results 
+1. identify the conclusions and wider implications
 
@@
 
-# Conclusions
+# Discussion
 

### 2021-01-08 etc/paper/forestecology.tex
Setup MEE formatting

@@
 
-\usepackage{xcolor, soul, xspace, float, subfig}
+\usepackage{xcolor, soul, xspace, float, subfig, lineno}
 
@@
 Smith College, Northampton, MA 01063 (e-mail:
-\href{mailto:akim04@smith.edu}{\nolinkurl{akim04@smith.edu}}). This work
-was not supported by any grant. The authors thank numerous colleagues
-and students for their support.} \\
+\href{mailto:akim04@smith.edu}{\nolinkurl{akim04@smith.edu}}).} \\
     Program in Statistical \& Data Sciences, Smith College\\
@@
 \begin{abstract}
-We provide a computational exercise suitable for early introduction in
-an undergraduate statistics or data science course that allows students
-to `play the whole game' of data science: performing both data
-collection and data analysis. While many teaching resources exist for
-data analysis, such resources are not as abundant for data collection
-given the inherent difficulty of the task. Our proposed exercise centers
-around student use of Google Calendar to collect data with the goal of
-answering the question `How do I spend my time?' On the one hand, the
-exercise involves answering a question with near universal appeal, but
-on the other hand, the data collection mechanism is not beyond the reach
-of a typical undergraduate student. A further benefit of the exercise is
-that it provides an opportunity for discussions on ethical questions and
-considerations that data providers and data analysts face in today's age
-of large-scale internet-based data collection.
+Move abstract below here after completed.
 \end{abstract}
@@
 \noindent%
-{\it Keywords:} forestecology, trees
+{\it Keywords:} forest ecology, competition, R, Rstats, tidyverse, sf, cross-validation,
 \vfill
@@
 \spacingset{1.45} % DON'T change the spacing!
+
+\linenumbers
+
+\hypertarget{abstract-350-words}{%
+\section{Abstract (350 words)}\label{abstract-350-words}}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  set the context for and purpose of the work
+\item
+  indicate the approach and the methods
+\item
+  outline the main results
+\item
+  identify the conclusions and wider implications
+\end{enumerate}
 
@@
 
-\hypertarget{conclusions}{%
-\section{Conclusions}\label{conclusions}}
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
 

### 2021-01-09 README.Rmd
renamed plot_posterior_parameters() and changed its sp behavior

@@
 plot_ex <- posterior_param_ex %>%
-  plot_posterior_parameters()
+  plot_bayesian_model_parameters()
 plot_ex[["lambda"]]
@@
 ex_bw <- focal_vs_comp_ex %>%
-  run_cv(max_dist = max_dist, cv_grid = ex_cv_grid) %>%
+  run_cv(max_dist = max_dist, cv_grid = ex_cv_grid_sf) %>%
   right_join(ex_growth_df, by = c("focal_ID" = "ID"))

### 2021-01-09 R/example_datasets.R

@@ -300,7 +300,7 @@
 #'
 #' # Plot posterior parameters
 #' posterior_param_ex %>%
-#'   plot_posterior_parameters()
+#'   plot_bayesian_model_parameters()
 "posterior_param_ex"
 
 

### 2021-01-09 R/model_functions.R

@@ -322,7 +322,7 @@ create_bayesian_model_data <- function(focal_vs_comp, run_shuffle = FALSE) {
 #' library(ggridges)
 #' # Load in posterior parameter example
 #' data(posterior_param_ex)
-#' plots <- plot_posterior_parameters(posterior_param_ex)
+#' plots <- plot_bayesian_model_parameters(posterior_param_ex)
 #'
 #' # Plot beta_0, growth intercepts
 #' plots[[1]]
@@ -332,10 +332,7 @@ create_bayesian_model_data <- function(focal_vs_comp, run_shuffle = FALSE) {
 #'
 #' # Plot lambdas, competition coefficents
 #' plots[[3]]
-plot_posterior_parameters <- function(posterior_param, sp_to_plot = NULL) {
-  if (FALSE) {
-    posterior_param <- posterior_param_bw
-  }
+plot_bayesian_model_parameters <- function(posterior_param, sp_to_plot = NULL) {
 
   # Identify all species and baseline category of species used for regression
   sp_list <- posterior_param$sp_list
@@ -407,6 +404,15 @@ plot_posterior_parameters <- function(posterior_param, sp_to_plot = NULL) {
     # mutate(type = str_to_title(type)) %>%
     select(-offset)
 
+  # When we only want to plot a subset of species:
+  if (!is.null(sp_to_plot)) {
+    sp_to_plot <- sort(sp_to_plot)
+
+    posterior_beta_0 <- posterior_beta_0 %>%
+      filter(type %in% sp_to_plot)
+  }
+
+
   plot_list[["beta_0"]] <-
     ggplot(posterior_beta_0, aes(x = value, y = fct_rev(type))) +
     geom_density_ridges() +
@@ -437,6 +443,14 @@ plot_posterior_parameters <- function(posterior_param, sp_to_plot = NULL) {
     ) %>%
     # mutate(type = str_to_title(type)) %>%
     select(-offset)
+
+  # When we only want to plot a subset of species:
+  if (!is.null(sp_to_plot)) {
+    sp_to_plot <- sort(sp_to_plot)
+
+    posterior_beta_dbh <- posterior_beta_dbh %>%
+      filter(type %in% sp_to_plot)
+  }
 
   plot_list[["beta_dbh"]] <-
     ggplot(posterior_beta_dbh, aes(x = value, y = fct_rev(type))) +

### 2021-01-09 etc/paper/forestecology.Rmd
cleaned manuscript

@@
 
+The code that generates Figures are included in the supplementary materials.
+
 We load all the necessary packages.
@@
 # Load spatial packages:
-# devtools::install_github("rvalavi/blockCV")
 library(blockCV)
@@
 
-```{r growth-histogram, out.width="100%", fig.cap="Distribution of average annual growth in DBH for both sites."}
+```{r growth-histogram, out.width="100%", fig.cap="Distribution of average annual growth in DBH for both sites.", echo = FALSE}
 # Both Big Woods & SCBI
@@
 bw_growth_df <- bw_growth_df %>%
-  add_buffer_variable(direction = "in", size = max_dist, region = bw_study_region)
+  add_buffer_variable(direction = "in", size = max_dist, 
+                      region = bw_study_region)
 
 ggplot() +
-  geom_sf(data = bw_growth_df %>% sample_frac(0.2), aes(col = buffer), size = 0.5)
+  geom_sf(data = bw_growth_df %>% sample_frac(0.2), 
+          aes(col = buffer), size = 0.5)
 ```
@@
 
-```{r bw-define-cv-folds}
+```{r bw-define-cv-folds, out.width="100%", fig.cap="blockCV function output."}
 set.seed(76)
 bw_spatialBlock <- spatialBlock(
-  speciesData = bw_growth_df, theRange = cv_fold_size,
-  k = 28, xOffset = 0.5, yOffset = 0,
-  verbose = FALSE
+  speciesData = bw_growth_df, theRange = cv_fold_size, k = 28, xOffset = 0.5, 
+  yOffset = 0, verbose = FALSE, showBlocks = FALSE
 )
+```
 
 # Add foldID to each tree
+```{r}
 bw_growth_df <- bw_growth_df %>%
   mutate(foldID = bw_spatialBlock$foldID)
+```
 
+```{r bw-define-cv-folds-with-trees, out.width="100%", fig.cap="Inspect blocks closely."}
 # Visualize grid. Why does fold 19 repeat?
-bw_spatialBlock$plots +
-  geom_sf(data = bw_growth_df %>% sample_frac(0.2), aes(col = factor(foldID)), size = 0.1)
+ggplot() +
+  geom_sf(data = bw_spatialBlock$blocks %>% st_as_sf()) + 
+  geom_sf(data = bw_growth_df %>% sample_frac(0.2), 
+          aes(col = factor(foldID)), size = 0.1) + 
+  theme(legend.position="bottom")
+```
 
 # Remove empty folds
+
+```{r}
 bw_growth_df <- bw_growth_df %>%
@@
 scbi_growth_df <- scbi_growth_df %>%
-  add_buffer_variable(direction = "in", size = max_dist, region = scbi_study_region)
+  add_buffer_variable(direction = "in", size = max_dist, 
+                      region = scbi_study_region)
 
@@
 ```{r}
-fold1 <- rbind(c(0, 300), c(150, 300), c(150, 600), c(0, 600), c(0, 300)) %>%
+fold1 <- rbind(c(0, 300), c(150, 300), c(150, 600), c(0, 600)) %>%
   sf_polygon() %>% 
   mutate(folds = 1)
-fold2 <- rbind(c(150, 300), c(300, 300), c(300, 600), c(150, 600), c(150, 300)) %>%
+fold2 <- rbind(c(150, 300), c(300, 300), c(300, 600), c(150, 600)) %>%
   sf_polygon() %>% 
@@
 scbi_spatialBlock <- spatialBlock(
-  speciesData = scbi_growth_df,
-  k = 2,
-  verbose = FALSE,
-  showBlocks = FALSE,
+  speciesData = scbi_growth_df, k = 2, verbose = FALSE, showBlocks = FALSE,
   # Note new arguments:
-  selection = "systematic",
-  blocks = scbi_cv_grid
+  selection = "systematic", blocks = scbi_cv_grid
 )
@@
 ggplot() +
-  geom_sf_text(data = scbi_growth_df, aes(label = foldID, col = buffer)) +
+  geom_sf_text(data = scbi_growth_df, 
+               aes(label = foldID, col = buffer)) +
   geom_sf(data = scbi_cv_grid, fill = "transparent")
@@
 
-```{r}
-head(focal_vs_comp_bw)
-```
-
 ### SCBI
@@
 
-```{r}
-head(focal_vs_comp_scbi)
-```
-
 ## Fit model and make predictions {#model-fit-predict}
@@
 Next we fit the following linear model to the dbh of each focal tree.
-Let $i = 1, \ldots, n_j$ index all $n_j$ trees of ``focal'' species group $j$; let $j = 1, \ldots, J$ index all $J$ focal species groups; and let $k = 1, \ldots, K$ index all $K$ ``competitor'' species groups.
+Let $i = 1, \ldots, n_j$ index all $n_j$ trees of "focal" species group $j$; let $j = 1, \ldots, J$ index all $J$ focal species groups; and let $k = 1, \ldots, K$ index all $K$ "competitor" species groups.
 We modeled the growth in diameter per year $y_{ij}$ (in centimeters per year) of the $i^{th}$ tree of focal species group $j$ as a linear model $f$ of the following covariates $\vec{x}_{ij}$
@@
 
-```{r scbi-model-residuals, out.width="100%", fig.cap="Spatial distribution of residuals for model applied to SCBI data."}
+```{r scbi-model-residuals, out.width="100%", fig.cap="Predicted versus observed growth.", echo = FALSE}
 ggplot(scbi_growth_df_noCV, aes(x = growth, y = growth_hat)) +
@@
 
-```{r scbi-model-residuals-2, out.width="100%", fig.cap="Spatial distribution of residuals for model applied to SCBI data part 2."}
+```{r scbi-model-residuals-2, out.width="80%", fig.cap="Spatial distribution of residuals for model applied to SCBI data.", echo = FALSE}
 scbi_growth_df_noCV %>%
@@
   mutate(sp = family) %>%
-  create_focal_vs_comp(max_dist = max_dist, cv_grid_sf = bw_cv_grid, id = "treeID")
+  create_focal_vs_comp(max_dist = max_dist, cv_grid_sf = bw_cv_grid, 
+                       id = "treeID")
 
@@
 
-Now the posterior parameter outputs of `fit_bayesian_model()` are passed to `plot_posterior_parameters()` to generate visualizations of the posterior parameters.
+Now the posterior parameter outputs of `fit_bayesian_model()` are passed to `plot_bayesian_model_parameters()` to generate visualizations of the posterior parameters.
 These visualizations are displayed in Figure 5 of @allen_permutation_2020.
@@
 ```{r bw-posterior-viz-1}
-posterior_plots <- plot_posterior_parameters(
+posterior_plots_bw <- plot_bayesian_model_parameters(
   posterior_param = posterior_param_bw,
@@
 
-```{r bw-posterior-viz-beta0, out.width="50%", fig.cap="Posterior distribution of beta0."}
-posterior_plots[["beta_0"]]
+```{r bw-posterior-viz-beta0, out.width="100%", fig.cap="Posterior distribution of beta0."}
+posterior_plots_bw[["beta_0"]]
 ```
@@
 ```{r bw-posterior-viz-beta-dbh, out.width="100%", fig.cap="Posterior distribution of betadbh."}
-posterior_plots[["beta_dbh"]]
+posterior_plots_bw[["beta_dbh"]]
 ```
@@
 
-```{r bw-posterior-viz-lambda, out.width="100%", fig.cap="Posterior distribution of lambda."}
-posterior_plots[["lambda"]]
+```{r bw-posterior-viz-lambda, out.width="100%", fig.cap="Posterior distribution of lambda's for Big Woods."}
+posterior_plots_bw[["lambda"]]
 ```
@@
 ```{r scbi-posterior-viz-1}
-posterior_plots_scbi <- plot_posterior_parameters(
+posterior_plots_scbi <- plot_bayesian_model_parameters(
   posterior_param = posterior_param_scbi,
@@
 
-```{r scbi-posterior-viz-lambda, out.width="100%", fig.cap="Posterior distribution of lambda."}
+```{r scbi-posterior-viz-lambda, out.width="100%", fig.cap="Posterior distribution of lambda's for SCBI."}
 posterior_plots_scbi[["lambda"]]

### 2021-01-09 etc/paper/forestecology.Rmd
first pass of cleaning and abstract

@@
 
-1. set the context for and purpose of the work 
-1. indicate the approach and the methods 
-1. outline the main results 
-1. identify the conclusions and wider implications
+1. set the context for and purpose of the work: models for interspecific competition. Robert and spatial crossvalidation 
+1. indicate the approach and the methods: use tidyverse, sf, and tidymodels packages. 
+1.
+outline the main results: Recreate the figure in PLOSOne paper.
+Code would've been way more complicated in base.
+We don't have to worry about how the component functions work, only that sequence/converyor belt is correct and output is correct.
+scientist is abstracted away from ugly programming details.
+1. identify the conclusions and wider implications: the fact truths hold was already covered in PLOSOne paper, this package implements this and can serve as a blueprint for others who model
 
@@
   geom_sf(data = bw_growth_df %>% sample_frac(0.2), 
-          aes(col = factor(foldID)), size = 0.1) + 
-  theme(legend.position="bottom")
+          aes(col = factor(foldID)), size = 0.1, show.legend = FALSE) + 
+  geom_sf_text(data = bw_spatialBlock$blocks %>% st_as_sf(), 
+               aes(label = folds))
 ```

### 2021-01-09 etc/paper/forestecology.tex

@@
 \item
-  set the context for and purpose of the work
+  set the context for and purpose of the work: models for interspecific
+  competition. Robert and spatial crossvalidation
 \item
-  indicate the approach and the methods
+  indicate the approach and the methods: use tidyverse, sf, and
+  tidymodels packages.
 \item
-  outline the main results
+  outline the main results: Recreate the figure in PLOSOne paper. Code
+  would've been way more complicated in base. We don't have to worry
+  about how the component functions work, only that sequence/converyor
+  belt is correct and output is correct. scientist is abstracted away
+  from ugly programming details.
 \item
-  identify the conclusions and wider implications
+  identify the conclusions and wider implications: the fact truths hold
+  was already covered in PLOSOne paper, this package implements this and
+  can serve as a blueprint for others who model
 \end{enumerate}
@@
 
+The code that generates Figures are included in the supplementary
+materials.
+
 We load all the necessary packages.
@@
 \CommentTok{# Load spatial packages:}
-\CommentTok{# devtools::install_github("rvalavi/blockCV")}
 \KeywordTok{library}\NormalTok{(blockCV)}
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{# Both Big Woods & SCBI}
-\KeywordTok{bind_rows}\NormalTok{(}
-\NormalTok{  bw_growth_df }\OperatorTok{%>%}\StringTok{ }\KeywordTok{st_drop_geometry}\NormalTok{() }\OperatorTok{%>%}\StringTok{ }\KeywordTok{select}\NormalTok{(growth) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{Site =} \StringTok{"Big Woods"}\NormalTok{),}
-\NormalTok{  scbi_growth_df }\OperatorTok{%>%}\StringTok{ }\KeywordTok{st_drop_geometry}\NormalTok{() }\OperatorTok{%>%}\StringTok{ }\KeywordTok{select}\NormalTok{(growth) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{Site =} \StringTok{"SCBI"}\NormalTok{)}
-\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ growth, }\DataTypeTok{y =}\NormalTok{ ..density.., }\DataTypeTok{fill =}\NormalTok{ Site)) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_histogram}\NormalTok{(}\DataTypeTok{alpha =} \FloatTok{0.5}\NormalTok{, }\DataTypeTok{position =} \StringTok{"identity"}\NormalTok{, }\DataTypeTok{binwidth =} \FloatTok{0.05}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{x =} \StringTok{"Average annual growth in dbh (cm per yr)"}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{coord_cartesian}\NormalTok{(}\DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(}\OperatorTok{-}\FloatTok{0.5}\NormalTok{, }\DecValTok{1}\NormalTok{))}
-\end{Highlighting}
-\end{Shaded}
-
 \begin{figure}
@@
 \NormalTok{bw_growth_df <-}\StringTok{ }\NormalTok{bw_growth_df }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{direction =} \StringTok{"in"}\NormalTok{, }\DataTypeTok{size =}\NormalTok{ max_dist, }\DataTypeTok{region =}\NormalTok{ bw_study_region)}
+\StringTok{  }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{direction =} \StringTok{"in"}\NormalTok{, }\DataTypeTok{size =}\NormalTok{ max_dist, }
+                      \DataTypeTok{region =}\NormalTok{ bw_study_region)}
 
 \KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ bw_growth_df }\OperatorTok{%>%}\StringTok{ }\KeywordTok{sample_frac}\NormalTok{(}\FloatTok{0.2}\NormalTok{), }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{col =}\NormalTok{ buffer), }\DataTypeTok{size =} \FloatTok{0.5}\NormalTok{)}
+\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ bw_growth_df }\OperatorTok{%>%}\StringTok{ }\KeywordTok{sample_frac}\NormalTok{(}\FloatTok{0.2}\NormalTok{), }
+          \KeywordTok{aes}\NormalTok{(}\DataTypeTok{col =}\NormalTok{ buffer), }\DataTypeTok{size =} \FloatTok{0.5}\NormalTok{)}
 \end{Highlighting}
@@
 \NormalTok{bw_spatialBlock <-}\StringTok{ }\KeywordTok{spatialBlock}\NormalTok{(}
-  \DataTypeTok{speciesData =}\NormalTok{ bw_growth_df, }\DataTypeTok{theRange =}\NormalTok{ cv_fold_size,}
-  \DataTypeTok{k =} \DecValTok{28}\NormalTok{, }\DataTypeTok{xOffset =} \FloatTok{0.5}\NormalTok{, }\DataTypeTok{yOffset =} \DecValTok{0}\NormalTok{,}
-  \DataTypeTok{verbose =} \OtherTok{FALSE}
+  \DataTypeTok{speciesData =}\NormalTok{ bw_growth_df, }\DataTypeTok{theRange =}\NormalTok{ cv_fold_size, }\DataTypeTok{k =} \DecValTok{28}\NormalTok{, }\DataTypeTok{xOffset =} \FloatTok{0.5}\NormalTok{, }
+  \DataTypeTok{yOffset =} \DecValTok{0}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{showBlocks =} \OtherTok{FALSE}
 \NormalTok{)}
-
-\CommentTok{# Add foldID to each tree}
+\end{Highlighting}
+\end{Shaded}
+
+\hypertarget{add-foldid-to-each-tree}{%
+\section{Add foldID to each tree}\label{add-foldid-to-each-tree}}
+
+\begin{Shaded}
+\begin{Highlighting}[]
 \NormalTok{bw_growth_df <-}\StringTok{ }\NormalTok{bw_growth_df }\OperatorTok{%>%}
 \StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ bw_spatialBlock}\OperatorTok{$}\NormalTok{foldID)}
-
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
 \CommentTok{# Visualize grid. Why does fold 19 repeat?}
-\NormalTok{bw_spatialBlock}\OperatorTok{$}\NormalTok{plots }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ bw_growth_df }\OperatorTok{%>%}\StringTok{ }\KeywordTok{sample_frac}\NormalTok{(}\FloatTok{0.2}\NormalTok{), }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{col =} \KeywordTok{factor}\NormalTok{(foldID)), }\DataTypeTok{size =} \FloatTok{0.1}\NormalTok{)}
-
-\CommentTok{# Remove empty folds}
+\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
+\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ bw_spatialBlock}\OperatorTok{$}\NormalTok{blocks }\OperatorTok{%>%}\StringTok{ }\KeywordTok{st_as_sf}\NormalTok{()) }\OperatorTok{+}\StringTok{ }
+\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ bw_growth_df }\OperatorTok{%>%}\StringTok{ }\KeywordTok{sample_frac}\NormalTok{(}\FloatTok{0.2}\NormalTok{), }
+          \KeywordTok{aes}\NormalTok{(}\DataTypeTok{col =} \KeywordTok{factor}\NormalTok{(foldID)), }\DataTypeTok{size =} \FloatTok{0.1}\NormalTok{, }\DataTypeTok{show.legend =} \OtherTok{FALSE}\NormalTok{) }\OperatorTok{+}\StringTok{ }
+\StringTok{  }\KeywordTok{geom_sf_text}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ bw_spatialBlock}\OperatorTok{$}\NormalTok{blocks }\OperatorTok{%>%}\StringTok{ }\KeywordTok{st_as_sf}\NormalTok{(), }
+               \KeywordTok{aes}\NormalTok{(}\DataTypeTok{label =}\NormalTok{ folds))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{Figures/bw-define-cv-folds-with-trees-1}
+
+}
+
+\caption{Inspect blocks closely.}\label{fig:bw-define-cv-folds-with-trees}
+\end{figure}
+
+\hypertarget{remove-empty-folds}{%
+\section{Remove empty folds}\label{remove-empty-folds}}
+
+\begin{Shaded}
+\begin{Highlighting}[]
 \NormalTok{bw_growth_df <-}\StringTok{ }\NormalTok{bw_growth_df }\OperatorTok{%>%}
@@
 \end{Shaded}
-
-\begin{center}\includegraphics[width=1\linewidth]{Figures/bw-define-cv-folds-1} \includegraphics[width=1\linewidth]{Figures/bw-define-cv-folds-2} \end{center}
 
@@
 \NormalTok{scbi_growth_df <-}\StringTok{ }\NormalTok{scbi_growth_df }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{direction =} \StringTok{"in"}\NormalTok{, }\DataTypeTok{size =}\NormalTok{ max_dist, }\DataTypeTok{region =}\NormalTok{ scbi_study_region)}
+\StringTok{  }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{direction =} \StringTok{"in"}\NormalTok{, }\DataTypeTok{size =}\NormalTok{ max_dist, }
+                      \DataTypeTok{region =}\NormalTok{ scbi_study_region)}
 
@@
 \begin{Highlighting}[]
-\NormalTok{fold1 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{600}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{600}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{)) }\OperatorTok{%>%}
+\NormalTok{fold1 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{600}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{600}\NormalTok{)) }\OperatorTok{%>%}
 \StringTok{  }\KeywordTok{sf_polygon}\NormalTok{() }\OperatorTok{%>%}\StringTok{ }
 \StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \DecValTok{1}\NormalTok{)}
-\NormalTok{fold2 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{600}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{)) }\OperatorTok{%>%}
+\NormalTok{fold2 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{600}\NormalTok{)) }\OperatorTok{%>%}
 \StringTok{  }\KeywordTok{sf_polygon}\NormalTok{() }\OperatorTok{%>%}\StringTok{ }
@@
 \NormalTok{scbi_spatialBlock <-}\StringTok{ }\KeywordTok{spatialBlock}\NormalTok{(}
-  \DataTypeTok{speciesData =}\NormalTok{ scbi_growth_df,}
-  \DataTypeTok{k =} \DecValTok{2}\NormalTok{,}
-  \DataTypeTok{verbose =} \OtherTok{FALSE}\NormalTok{,}
-  \DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{,}
+  \DataTypeTok{speciesData =}\NormalTok{ scbi_growth_df, }\DataTypeTok{k =} \DecValTok{2}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{,}
   \CommentTok{# Note new arguments:}
-  \DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{,}
-  \DataTypeTok{blocks =}\NormalTok{ scbi_cv_grid}
+  \DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{, }\DataTypeTok{blocks =}\NormalTok{ scbi_cv_grid}
 \NormalTok{)}
@@
 \KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_sf_text}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ scbi_growth_df, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{label =}\NormalTok{ foldID, }\DataTypeTok{col =}\NormalTok{ buffer)) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{geom_sf_text}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ scbi_growth_df, }
+               \KeywordTok{aes}\NormalTok{(}\DataTypeTok{label =}\NormalTok{ foldID, }\DataTypeTok{col =}\NormalTok{ buffer)) }\OperatorTok{+}
 \StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ scbi_cv_grid, }\DataTypeTok{fill =} \StringTok{"transparent"}\NormalTok{)}
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{head}\NormalTok{(focal_vs_comp_bw)}
-\CommentTok{## # A tibble: 6 x 10}
-\CommentTok{##   focal_ID focal_sp   dbh foldID                  geometry growth}
-\CommentTok{##      <dbl> <fct>    <dbl> <chr>                    <POINT>  <dbl>}
-\CommentTok{## 1        1 white_o~  41.2 12                   (8.7 107.5)  0.404}
-\CommentTok{## 2        1 white_o~  41.2 12                   (8.7 107.5)  0.404}
-\CommentTok{## 3        1 white_o~  41.2 12                   (8.7 107.5)  0.404}
-\CommentTok{## 4        1 white_o~  41.2 12                   (8.7 107.5)  0.404}
-\CommentTok{## 5        1 white_o~  41.2 12                   (8.7 107.5)  0.404}
-\CommentTok{## 6        1 white_o~  41.2 12                   (8.7 107.5)  0.404}
-\CommentTok{## # ... with 4 more variables: comp_ID <dbl>, dist <dbl>,}
-\CommentTok{## #   comp_sp <fct>, comp_basal_area <dbl>}
-\end{Highlighting}
-\end{Shaded}
-
 \hypertarget{scbi-2}{%
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{head}\NormalTok{(focal_vs_comp_scbi)}
-\CommentTok{## # A tibble: 6 x 10}
-\CommentTok{##   focal_ID focal_sp   dbh foldID                  geometry growth}
-\CommentTok{##      <dbl> <fct>    <dbl>  <int>                   <POINT>  <dbl>}
-\CommentTok{## 1        4 nysy      13.6      1              (14.2 428.5)  0.103}
-\CommentTok{## 2        4 nysy      13.6      1              (14.2 428.5)  0.103}
-\CommentTok{## 3        4 nysy      13.6      1              (14.2 428.5)  0.103}
-\CommentTok{## 4        4 nysy      13.6      1              (14.2 428.5)  0.103}
-\CommentTok{## 5        4 nysy      13.6      1              (14.2 428.5)  0.103}
-\CommentTok{## 6        4 nysy      13.6      1              (14.2 428.5)  0.103}
-\CommentTok{## # ... with 4 more variables: comp_ID <dbl>, dist <dbl>,}
-\CommentTok{## #   comp_sp <fct>, comp_basal_area <dbl>}
-\end{Highlighting}
-\end{Shaded}
-
 \hypertarget{model-fit-predict}{%
@@
 Next we fit the following linear model to the dbh of each focal tree.
-Let \(i = 1, \ldots, n_j\) index all \(n_j\) trees of
-\texttt{focal\textquotesingle{}\textquotesingle{}\ species\ group\ \$j\$;\ let\ \$j\ =\ 1,\ \textbackslash{}ldots,\ J\$\ index\ all\ \$J\$\ focal\ species\ groups;\ and\ let\ \$k\ =\ 1,\ \textbackslash{}ldots,\ K\$\ index\ all\ \$K\$}competitor''
+Let \(i = 1, \ldots, n_j\) index all \(n_j\) trees of ``focal'' species
+group \(j\); let \(j = 1, \ldots, J\) index all \(J\) focal species
+groups; and let \(k = 1, \ldots, K\) index all \(K\) ``competitor''
 species groups. We modeled the growth in diameter per year \(y_{ij}\)
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{ggplot}\NormalTok{(scbi_growth_df_noCV, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ growth, }\DataTypeTok{y =}\NormalTok{ growth_hat)) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_point}\NormalTok{(}\DataTypeTok{size =} \FloatTok{0.5}\NormalTok{, }\DataTypeTok{color =} \KeywordTok{rgb}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\FloatTok{0.25}\NormalTok{)) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{stat_smooth}\NormalTok{(}\DataTypeTok{method =} \StringTok{"lm"}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_abline}\NormalTok{(}\DataTypeTok{slope =} \DecValTok{1}\NormalTok{, }\DataTypeTok{intercept =} \DecValTok{0}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{coord_fixed}\NormalTok{() }\OperatorTok{+}
-\StringTok{  }\KeywordTok{labs}\NormalTok{(}
-    \DataTypeTok{x =} \StringTok{"Observed growth in dbh"}\NormalTok{, }\DataTypeTok{y =} \StringTok{"Predicted growth in dbh"}\NormalTok{,}
-    \DataTypeTok{title =} \StringTok{"Predicted vs Observed Growth"}
-\NormalTok{  )}
-\end{Highlighting}
-\end{Shaded}
-
 \begin{figure}
@@
 
-\caption{Spatial distribution of residuals for model applied to SCBI data.}\label{fig:scbi-model-residuals}
+\caption{Predicted versus observed growth.}\label{fig:scbi-model-residuals}
 \end{figure}
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{scbi_growth_df_noCV }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{st_as_sf}\NormalTok{() }\OperatorTok{%>%}
-\StringTok{  }\CommentTok{# }\AlertTok{TODO}\CommentTok{: Need to investigate missingness}
-\StringTok{  }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{is.na}\NormalTok{(growth_hat)) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
-    \DataTypeTok{error =}\NormalTok{ growth }\OperatorTok{-}\StringTok{ }\NormalTok{growth_hat,}
-    \DataTypeTok{error_bin =} \KeywordTok{cut_number}\NormalTok{(error, }\DataTypeTok{n =} \DecValTok{5}\NormalTok{),}
-    \DataTypeTok{error_compress =} \KeywordTok{ifelse}\NormalTok{(error }\OperatorTok{<}\StringTok{ }\FloatTok{-0.75}\NormalTok{, }\FloatTok{-0.75}\NormalTok{, }\KeywordTok{ifelse}\NormalTok{(error }\OperatorTok{>}\StringTok{ }\FloatTok{0.75}\NormalTok{, }\FloatTok{0.75}\NormalTok{, error))}
-\NormalTok{  ) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{col =}\NormalTok{ error_compress), }\DataTypeTok{size =} \DecValTok{1}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{theme_bw}\NormalTok{() }\OperatorTok{+}
-\StringTok{  }\KeywordTok{scale_color_gradient2}\NormalTok{(}
-    \DataTypeTok{low =} \StringTok{"#ef8a62"}\NormalTok{, }\DataTypeTok{mid =} \StringTok{"#f7f7f7"}\NormalTok{, }\DataTypeTok{high =} \StringTok{"#67a9cf"}\NormalTok{,}
-    \DataTypeTok{name =} \KeywordTok{expression}\NormalTok{(}\KeywordTok{paste}\NormalTok{(}\StringTok{"Residual (cm "}\NormalTok{, y}\OperatorTok{^}\NormalTok{\{}\OperatorTok{-}\DecValTok{1}\NormalTok{\}, }\StringTok{")"}\NormalTok{)),}
-    \DataTypeTok{breaks =} \KeywordTok{seq}\NormalTok{(}\DataTypeTok{from =} \FloatTok{-0.75}\NormalTok{, }\DataTypeTok{to =} \FloatTok{0.75}\NormalTok{, }\DataTypeTok{by =} \FloatTok{0.25}\NormalTok{),}
-    \DataTypeTok{labels =} \KeywordTok{c}\NormalTok{(}\StringTok{"< -0.75"}\NormalTok{, }\StringTok{"-0.5"}\NormalTok{, }\StringTok{"0.25"}\NormalTok{, }\StringTok{"0"}\NormalTok{, }\StringTok{"0.25"}\NormalTok{, }\StringTok{"0.5"}\NormalTok{, }\StringTok{"> 0.75"}\NormalTok{)}
-\NormalTok{  ) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{x =} \StringTok{"Meter"}\NormalTok{, }\DataTypeTok{y =} \StringTok{"Meter"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
 \begin{figure}
 
-{\centering \includegraphics[width=1\linewidth]{Figures/scbi-model-residuals-2-1}
-
-}
-
-\caption{Spatial distribution of residuals for model applied to SCBI data part 2.}\label{fig:scbi-model-residuals-2}
+{\centering \includegraphics[width=0.8\linewidth]{Figures/scbi-model-residuals-2-1}
+
+}
+
+\caption{Spatial distribution of residuals for model applied to SCBI data.}\label{fig:scbi-model-residuals-2}
 \end{figure}
@@
 \StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =}\NormalTok{ family) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{max_dist =}\NormalTok{ max_dist, }\DataTypeTok{cv_grid_sf =}\NormalTok{ bw_cv_grid, }\DataTypeTok{id =} \StringTok{"treeID"}\NormalTok{)}
+\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{max_dist =}\NormalTok{ max_dist, }\DataTypeTok{cv_grid_sf =}\NormalTok{ bw_cv_grid, }
+                       \DataTypeTok{id =} \StringTok{"treeID"}\NormalTok{)}
 
@@
 Now the posterior parameter outputs of \texttt{fit\_bayesian\_model()}
-are passed to \texttt{plot\_posterior\_parameters()} to generate
+are passed to \texttt{plot\_bayesian\_model\_parameters()} to generate
 visualizations of the posterior parameters. These visualizations are
@@
 \begin{Highlighting}[]
-\NormalTok{posterior_plots <-}\StringTok{ }\KeywordTok{plot_posterior_parameters}\NormalTok{(}
+\NormalTok{posterior_plots_bw <-}\StringTok{ }\KeywordTok{plot_bayesian_model_parameters}\NormalTok{(}
   \DataTypeTok{posterior_param =}\NormalTok{ posterior_param_bw,}
@@
 \begin{Highlighting}[]
-\NormalTok{posterior_plots[[}\StringTok{"beta_0"}\NormalTok{]]}
+\NormalTok{posterior_plots_bw[[}\StringTok{"beta_0"}\NormalTok{]]}
 \end{Highlighting}
@@
 
-{\centering \includegraphics[width=0.5\linewidth]{Figures/bw-posterior-viz-beta0-1}
+{\centering \includegraphics[width=1\linewidth]{Figures/bw-posterior-viz-beta0-1}
 
@@
 \begin{Highlighting}[]
-\NormalTok{posterior_plots[[}\StringTok{"beta_dbh"}\NormalTok{]]}
+\NormalTok{posterior_plots_bw[[}\StringTok{"beta_dbh"}\NormalTok{]]}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{posterior_plots[[}\StringTok{"lambda"}\NormalTok{]]}
+\NormalTok{posterior_plots_bw[[}\StringTok{"lambda"}\NormalTok{]]}
 \end{Highlighting}
@@
 
-\caption{Posterior distribution of lambda.}\label{fig:bw-posterior-viz-lambda}
+\caption{Posterior distribution of lambda's for Big Woods.}\label{fig:bw-posterior-viz-lambda}
 \end{figure}
@@
 \begin{Highlighting}[]
-\NormalTok{posterior_plots_scbi <-}\StringTok{ }\KeywordTok{plot_posterior_parameters}\NormalTok{(}
+\NormalTok{posterior_plots_scbi <-}\StringTok{ }\KeywordTok{plot_bayesian_model_parameters}\NormalTok{(}
   \DataTypeTok{posterior_param =}\NormalTok{ posterior_param_scbi,}
@@
 
-\caption{Posterior distribution of lambda.}\label{fig:scbi-posterior-viz-lambda}
+\caption{Posterior distribution of lambda's for SCBI.}\label{fig:scbi-posterior-viz-lambda}
 \end{figure}

### 2021-01-09 etc/paper/forestecology.Rmd
Amplified ideas in abstract

--- 2021-01-09 etc/paper/forestecology.Rmd
@@
 
-1. set the context for and purpose of the work 
-1. indicate the approach and the methods 
-1. outline the main results 
-1. identify the conclusions and wider implications
+1. Set the context for and purpose of the work: The scientific question/problem and the desiderata
+    + fitting models for interspecific competition and assessing them using spatial crossvalidation
+    + Leverage ForestGEO protocols providing standardization
+1. Indicate the approach and the methods: Use tidyverse, simple features, and tidymodels packages. 
+    + tidyverse: as stated in tidy tools manifesto: standadized data structures, functional programming with pipe, designed for humans
+    + sf: tidyverse-friendly package that makes wrangling and visualizing spatial data much easier
+    + tidymodels: given our spatial-crossvalidation, use tidymodels framework is a collection of packages for modeling and machine learning using tidyverse principles
+1. Outline the main results: We replicate from scratch the figure in PLOSOne paper using Big Woods data, conduct similar analysis for SCBI data.
+    + Code would've been way more complicated in base. 
+    + We don't have to worry about how the component functions work, only that sequence/converyor belt is correct and output is correct.
+    + Scientist is abstracted away from ugly programming details.
+1. Identify the conclusions and wider implications:
+    + New scientific conclusions from SCBI data
+    + this can serve as blue print for other modeling situations.
+    + Modularly switch out our bayesian lm() functions to anything you want.
 

### 2021-01-09 etc/paper/forestecology.tex
Amplified ideas in abstract

@@
 \item
-  set the context for and purpose of the work: models for interspecific
-  competition. Robert and spatial crossvalidation
-\item
-  indicate the approach and the methods: use tidyverse, sf, and
-  tidymodels packages.
-\item
-  outline the main results: Recreate the figure in PLOSOne paper. Code
-  would've been way more complicated in base. We don't have to worry
-  about how the component functions work, only that sequence/converyor
-  belt is correct and output is correct. scientist is abstracted away
-  from ugly programming details.
-\item
-  identify the conclusions and wider implications: the fact truths hold
-  was already covered in PLOSOne paper, this package implements this and
-  can serve as a blueprint for others who model
+  Set the context for and purpose of the work: The scientific
+  question/problem and the desiderata
+
+  \begin{itemize}
+  \tightlist
+  \item
+    fitting models for interspecific competition and assessing them
+    using spatial crossvalidation
+  \item
+    Leverage ForestGEO protocols providing standardization
+  \end{itemize}
+\item
+  Indicate the approach and the methods: Use tidyverse, simple features,
+  and tidymodels packages.
+
+  \begin{itemize}
+  \tightlist
+  \item
+    tidyverse: as stated in tidy tools manifesto: standadized data
+    structures, functional programming with pipe, designed for humans
+  \item
+    sf: tidyverse-friendly package that makes wrangling and visualizing
+    spatial data much easier
+  \item
+    tidymodels: given our spatial-crossvalidation, use tidymodels
+    framework is a collection of packages for modeling and machine
+    learning using tidyverse principles
+  \end{itemize}
+\item
+  Outline the main results: We replicate from scratch the figure in
+  PLOSOne paper using Big Woods data, conduct similar analysis for SCBI
+  data.
+
+  \begin{itemize}
+  \tightlist
+  \item
+    Code would've been way more complicated in base.
+  \item
+    We don't have to worry about how the component functions work, only
+    that sequence/converyor belt is correct and output is correct.
+  \item
+    Scientist is abstracted away from ugly programming details.
+  \end{itemize}
+\item
+  Identify the conclusions and wider implications:
+
+  \begin{itemize}
+  \tightlist
+  \item
+    New scientific conclusions from SCBI data
+  \item
+    this can serve as blue print for other modeling situations.
+  \item
+    Modularly switch out our bayesian lm() functions to anything you
+    want.
+  \end{itemize}
 \end{enumerate}

### 2021-01-12 etc/paper/forestecology.Rmd
small edits

@@
     + tidymodels: given our spatial-crossvalidation, use tidymodels framework is a collection of packages for modeling and machine learning using tidyverse principles
+    + Dave: hmm, I think here we might need to mention the methods we implement.
+Something like, the package provides functions to specific a linear, bayesian neighborhood competition model of growth, then fit the model, and compare competing models with spatiall cross validation.
+Or somethign like that.
 1. Outline the main results: We replicate from scratch the figure in PLOSOne paper using Big Woods data, conduct similar analysis for SCBI data.
@@
 This package implements the methods in @allen_permutation_2020.
-It provides: a convenient way specify and fit models of tree growth based on neighborhood competition; a spatial cross validation method to test and compare model fits @roberts_cross-validation_2017; and an ANOVA-like method to assess whether the competitor identity matters in these models.
+It provides: a convenient way to specify and fit models of tree growth based on neighborhood competition; a spatial cross validation method to test and compare model fits @roberts_cross-validation_2017; and an ANOVA-like method to assess whether the competitor identity matters in these models.
 The model is written to work with ForestGEO plot data @andersonteixeira_ctfs-forestgeo_2015, but we envision that it could easily be modified to work with data from other forest plots, e.g.
@@
 The 23 ha plot is situated in mature oak-hickory forest.
-The canopy is dominated by white oak (Quercus alba), northern red oak (Quercus rubra), black oak (Quercus velutina), shagbark hickory (Carya ovata) and pignut hickory (Carya glabra).
-The most common understory tree is witch-hazel (Hamamelis virginiana) @allen_michigan_2020.
+The canopy is dominated by white oak (*Quercus alba*), northern red oak (*Q.
+rubra*), black oak (*Q.
+velutina*), shagbark hickory (*Carya ovata*) and pignut hickory (*C.
+glabra*) @allen_michigan_2020.
 In the example below, we will preface any data frames from this plot in with `bw_`.
@@
 The 25.6 ha (640 x 400 m) plot is located at the intersection of three of the major physiographic provinces of the eastern US: the Blue Ridge, Ridge and Valley, and Piedmont provinces and is adjacent to the northern end of Shenandoah National Park.
-The forest type is typical mature secondary eastern mixed deciduous forest, with a canopy dominated by tulip poplar (Liriodendron tulipifera), oaks (Quercus spp.), and hickories (Carya spp.), and an understory composed mainly of spicebush (Lindera benzoin), paw-paw (Asimina triloba), American hornbeam (Carpinus caroliniana), and witch hazel (Hamamelis virginiana) @bourg_initial_2013.
+The forest type is typical mature secondary eastern mixed deciduous forest, with a canopy dominated by tulip poplar (*Liriodendron tulipifera*), oaks (*Quercus* spp.), and hickories (*Carya* spp.), and an understory composed mainly of spicebush (*Lindera benzoin*), paw-paw (*Asimina triloba*), American hornbeam (*Carpinus caroliniana*), and witch hazel (*Hamamelis virginiana*) @bourg_initial_2013.
 In the example below, we will preface any data frames from this plot in with `scbi_`.
@@
 In the case of Big Woods data, we first remove all trees that were re-sprouts in the later (2014) census.
-Additionally, we have included three classification of tree species: `species`, `family`, and `trait_group`.
-DESCRIBE THESE
+Additionally, we have included two classification of tree species: `species` and `family`.
+To illustrate model comparison, we test competition models in which individuals are grouped by species and by family.
 
@@
 
-Figure \ref{fig:growth-histogram} displays histograms comparing the distribution of average annual growth at both sites. Observe that average annual growth appears higher at the Big Woods site.
+DAVE: I THINK THIS FIGURE IS UNNEEDED.
+COULD BE REMOVED FOR SPACE REASONS.
+Figure \ref{fig:growth-histogram} displays histograms comparing the distribution of average annual growth at both sites.
+Observe that average annual growth appears higher at the Big Woods site.
 
@@
 Our model of interspecific competition relies on a spatial definition of who the competitor trees are for focal trees of interest.
-Since certain explanatory variables such as biomass are cumulative, we must ensure that all trees being modeled are not biased to have different neighbor structures.
+Since certain explanatory variables such as basal area are cumulative, we must ensure that all trees being modeled are not biased to have different neighbor structures.
 This is a particular concern for trees at the boundary of study regions, which will not have the same number of neighbors as trees in the internal part of the study region.
@@
 However, underlying this scheme is an assumption that the observations are independent.
-In the case of forest census data, observations exhibit spatial autocorrelation, and thus this dependence must be incorporated in our resampling scheme in spatial cross-validation @roberts2017 @pohjankukka2017.
-We will therefore associate portions of the study region to spatial folds.
+In the case of forest census data, observations exhibit spatial autocorrelation, and thus this dependence must be incorporated in our resampling scheme in spatial cross-validation @roberts_cross-validation_2017 @pohjankukka_estimating_2017 We will therefore associate portions of the study region to spatial folds.
 
@@
 We will then use this constant to associate each observed tree to one of $k$ folds in the respective study region.
-In the Big Woods example below we will use the `blockCV` R package that has implemented spatial cross-validation while for the SCBI we will do this manually @valavi2019.
+In the Big Woods example below we will use the `blockCV` R package that has implemented spatial cross-validation while for the SCBI we will do this manually @valavi_blockcv_2019
 

### 2021-01-13 etc/paper/forestecology.Rmd
tweaks to abstract

--- 2021-01-09 etc/paper/forestecology.Rmd
@@
 1. Set the context for and purpose of the work: The scientific question/problem and the desiderata
-    + fitting models for interspecific competition and assessing them using spatial crossvalidation
+    + (Eventually) modularly fitting models for interspecific competition and assessing them using spatial crossvalidation
     + Leverage ForestGEO protocols providing standardization
@@
     + New scientific conclusions from SCBI data
+    + Modularly switch out our bayesian lm() functions to anything you want.
     + this can serve as blue print for other modeling situations.
-    + Modularly switch out our bayesian lm() functions to anything you want.
 

### 2021-01-13 etc/paper/forestecology.tex
Reknit manuscript

@@
   \item
-    fitting models for interspecific competition and assessing them
-    using spatial crossvalidation
+    (Eventually) modularly fitting models for interspecific competition
+    and assessing them using spatial crossvalidation
   \item
@@
     learning using tidyverse principles
+  \item
+    Dave: hmm, I think here we might need to mention the methods we
+    implement. Something like, the package provides functions to
+    specific a linear, bayesian neighborhood competition model of
+    growth, then fit the model, and compare competing models with
+    spatiall cross validation. Or somethign like that.
   \end{itemize}
@@
   \item
-    this can serve as blue print for other modeling situations.
-  \item
     Modularly switch out our bayesian lm() functions to anything you
     want.
+  \item
+    this can serve as blue print for other modeling situations.
   \end{itemize}
@@
 implements the methods in \citet{allen_permutation_2020}. It provides: a
-convenient way specify and fit models of tree growth based on
+convenient way to specify and fit models of tree growth based on
 neighborhood competition; a spatial cross validation method to test and
@@
 S. George Reserve in Pinckney, MI, USA. The 23 ha plot is situated in
-mature oak-hickory forest. The canopy is dominated by white oak (Quercus
-alba), northern red oak (Quercus rubra), black oak (Quercus velutina),
-shagbark hickory (Carya ovata) and pignut hickory (Carya glabra). The
-most common understory tree is witch-hazel (Hamamelis virginiana)
-\citet{allen_michigan_2020}. In the example below, we will preface any
-data frames from this plot in with \texttt{bw\_}.
+mature oak-hickory forest. The canopy is dominated by white oak
+(\emph{Quercus alba}), northern red oak (\emph{Q. rubra}), black oak
+(\emph{Q. velutina}), shagbark hickory (\emph{Carya ovata}) and pignut
+hickory (\emph{C. glabra}) \citet{allen_michigan_2020}. In the example
+below, we will preface any data frames from this plot in with
+\texttt{bw\_}.
 
@@
 eastern mixed deciduous forest, with a canopy dominated by tulip poplar
-(Liriodendron tulipifera), oaks (Quercus spp.), and hickories (Carya
-spp.), and an understory composed mainly of spicebush (Lindera benzoin),
-paw-paw (Asimina triloba), American hornbeam (Carpinus caroliniana), and
-witch hazel (Hamamelis virginiana) \citet{bourg_initial_2013}. In the
-example below, we will preface any data frames from this plot in with
+(\emph{Liriodendron tulipifera}), oaks (\emph{Quercus} spp.), and
+hickories (\emph{Carya} spp.), and an understory composed mainly of
+spicebush (\emph{Lindera benzoin}), paw-paw (\emph{Asimina triloba}),
+American hornbeam (\emph{Carpinus caroliniana}), and witch hazel
+(\emph{Hamamelis virginiana}) \citet{bourg_initial_2013}. In the example
+below, we will preface any data frames from this plot in with
 \texttt{scbi\_}.
@@
 re-sprouts in the later (2014) census. Additionally, we have included
-three classification of tree species: \texttt{species}, \texttt{family},
-and \texttt{trait\_group}. DESCRIBE THESE
+two classification of tree species: \texttt{species} and
+\texttt{family}. To illustrate model comparison, we test competition
+models in which individuals are grouped by species and by family.
 
@@
 
-Figure \ref{fig:growth-histogram} displays histograms comparing the
-distribution of average annual growth at both sites. Observe that
+DAVE: I THINK THIS FIGURE IS UNNEEDED. COULD BE REMOVED FOR SPACE
+REASONS. Figure \ref{fig:growth-histogram} displays histograms comparing
+the distribution of average annual growth at both sites. Observe that
 average annual growth appears higher at the Big Woods site.
@@
 the competitor trees are for focal trees of interest. Since certain
-explanatory variables such as biomass are cumulative, we must ensure
+explanatory variables such as basal area are cumulative, we must ensure
 that all trees being modeled are not biased to have different neighbor
@@
 dependence must be incorporated in our resampling scheme in spatial
-cross-validation \citet{roberts2017} \citet{pohjankukka2017}. We will
-therefore associate portions of the study region to spatial folds.
+cross-validation \citet{roberts_cross-validation_2017}
+\citet{pohjankukka_estimating_2017} We will therefore associate portions
+of the study region to spatial folds.
 
@@
 that has implemented spatial cross-validation while for the SCBI we will
-do this manually \citet{valavi2019}.
+do this manually \citet{valavi_blockcv_2019}
 

### 2021-01-15 README.Rmd
adjust output of predict method

@@
 posterior_param_ex <- focal_vs_comp_ex %>%
-  fit_bayesian_model(prior_param = NULL, run_shuffle = FALSE)
+  fit(prior_param = NULL, run_shuffle = FALSE)
 
@@
 predictions <- focal_vs_comp_ex %>%
-  predict_bayesian_model(posterior_param = posterior_param_ex) %>%
+  predict(posterior_param = posterior_param_ex) %>%
   right_join(ex_growth_df, by = c("focal_ID" = "ID"))
@@
 predictions %>%
-  rmse(truth = growth, estimate = growth_hat) %>%
+  rmse(truth = growth, estimate = .pred) %>%
   pull(.estimate)
@@
 ex_bw %>%
-  rmse(truth = growth, estimate = growth_hat) %>%
+  rmse(truth = growth, estimate = .pred) %>%
   pull(.estimate)

### 2021-01-15 etc/paper/forestecology.Rmd
fixed header error
adjust output of predict method

@@
 
-# Add foldID to each tree
+Then add foldID to each tree
+
 ```{r}
@@
 
-# Remove empty folds
+Then remove empty folds
 
@@
 predictions %>%
-  yardstick::rmse(truth = growth, estimate = growth_hat) %>%
+  yardstick::rmse(truth = growth, estimate = .pred) %>%
   pull(.estimate)
@@
 predictions_shuffle %>%
-  rmse(truth = growth, estimate = growth_hat) %>%
+  rmse(truth = growth, estimate = .pred) %>%
   pull(.estimate)
@@
 scbi_growth_df_noCV %>%
-  rmse(truth = growth, estimate = growth_hat) %>%
+  rmse(truth = growth, estimate = .pred) %>%
   pull(.estimate)
@@
 ```{r scbi-model-residuals, out.width="100%", fig.cap="Predicted versus observed growth.", echo = FALSE}
-ggplot(scbi_growth_df_noCV, aes(x = growth, y = growth_hat)) +
+ggplot(scbi_growth_df_noCV, aes(x = growth, y = .pred)) +
   geom_point(size = 0.5, color = rgb(0, 0, 0, 0.25)) +
@@
   # TODO: Need to investigate missingness
-  filter(!is.na(growth_hat)) %>%
+  filter(!is.na(.pred)) %>%
   mutate(
-    error = growth - growth_hat,
+    error = growth - .pred,
     error_bin = cut_number(error, n = 5),
@@
 cv_bw %>%
-  rmse(truth = growth, estimate = growth_hat) %>%
+  rmse(truth = growth, estimate = .pred) %>%
   pull(.estimate)
@@
 cv_scbi %>%
-  rmse(truth = growth, estimate = growth_hat) %>%
+  rmse(truth = growth, estimate = .pred) %>%
   pull(.estimate)

### 2021-01-15 R/data_processing_functions.R

@@ -219,7 +219,7 @@ create_focal_vs_comp <- function(growth_df, max_dist, cv_grid_sf, id) {
     # Convert list to tibble:
     bind_rows() %>%
     arrange(focal_ID, comp_ID) %>%
-    mutate(growth_hat = NA) %>%
+    mutate(.pred = NA) %>%
     select(
       # Relating to focal tree:
       focal_ID, focal_sp, dbh, foldID, geometry, growth,

### 2021-01-15 R/forestecology.R

@@ -20,7 +20,7 @@ if (getRversion() >= "2.15.1") {
       "dbh08", "dbh14", "family_phylo", "trait_group", "buffer", "code14",
       "species", "focal_ID", "ID", "spCode", "fold", "basal_area", "dbh",
       "focal_ID_neighbor", "focal_bw_ID", "comp_species", "comp_biomass",
-      "inside", "growth", "biomass_total", "model_formula", "growth_hat",
+      "inside", "growth", "biomass_total", "model_formula", ".pred",
       "comp_ID", "x", "y", "x_fold", "y_fold", "neighbor",
       "notion_of_competitor_species",
       "ExactDate", "ExactDate2", "code", "code_2018", "dbh1", "dbh2", "family",

### 2021-01-15 README.Rmd
debug examples, update readme

@@
 # Get predicted dbh values values
-predictions <- focal_vs_comp_ex %>%
-  predict(posterior_param = posterior_param_ex) %>%
-  right_join(ex_growth_df, by = c("focal_ID" = "ID"))
+predictions <- 
+  posterior_param_ex$fe_data %>%
+  select(focal_ID, focal_sp, dbh, growth) %>%
+  distinct() %>%
+  bind_cols(predict(posterior_param_ex))
 
@@
 # Fit model with cross-validation
-ex_bw <- focal_vs_comp_ex %>%
+ex_bw <- posterior_param_ex %>%
   run_cv(max_dist = max_dist, cv_grid = ex_cv_grid_sf) %>%
-  right_join(ex_growth_df, by = c("focal_ID" = "ID"))
+  bind_cols(
+    posterior_param_ex$fe_data %>%
+    select(focal_ID, focal_sp, dbh, growth) %>%
+    distinct()
+  )
 

### 2021-01-15 R/example_datasets.R

@@ -291,9 +291,13 @@
 #' library(yardstick)
 #'
 #' # Compare model predictions to observation
-#' predictions <- focal_vs_comp_ex %>%
-#'   predict(posterior_param = posterior_param_ex) %>%
-#'   right_join(ex_growth_df, by = c("focal_ID" = "ID"))
+#' predictions <-
+#'   predict(posterior_param_ex, newdata = focal_vs_comp_ex) %>%
+#'     bind_cols(
+#'       focal_vs_comp_ex %>%
+#'         select(focal_ID, focal_sp, dbh, growth) %>%
+#'         distinct()
+#'     )
 #' predictions %>%
 #'   rmse(truth = growth, estimate = .pred) %>%
 #'   pull(.estimate)

### 2021-01-15 R/model_functions.R

@@ -105,25 +105,13 @@ fit.fe_data <- function(object, prior_param = NULL, run_shuffle = FALSE, ...) {
 
 
 
-# For now, we haven't solidified what `fe_data` objects should look like.
-# Since they subclass `tbl_df`, we'll define the fit method for them using
-# that class.
-#' @rdname fit.fe_data
-#' @method fit tbl_df
-#' @export fit.tbl_df
-#' @export
-fit.tbl_df <- fit.fe_data
-
-
-
-
-
 #' Make predictions based on fitted Bayesian model
 #'
 #' @description Applies fitted model from [fit.fe_data] and
 #'   returns posterior predicted values.
 #'
 #' @inheritParams fit.fe_data
+#' @param newdata Data to predict on. If NULL, defaults to the training data.
 #' @inheritParams create_focal_vs_comp
 #'
 #' @import dplyr
@@ -146,18 +134,25 @@ fit.tbl_df <- fit.fe_data
 #' # and growth data to compare to
 #' data(posterior_param_ex, ex_growth_df)
 #'
-#' predictions <- focal_vs_comp_ex %>%
-#'   predict(posterior_param = posterior_param_ex) %>%
-#'   right_join(ex_growth_df, by = c("focal_ID" = "ID"))
+#' predictions <-
+#'   predict(posterior_param_ex, newdata = focal_vs_comp_ex) %>%
+#'     bind_cols(
+#'       focal_vs_comp_ex %>%
+#'         select(focal_ID, focal_sp, dbh, growth) %>%
+#'         distinct()
+#'     )
+#'
 #' predictions %>%
 #'   ggplot(aes(growth, .pred)) +
 #'   geom_point() +
 #'   geom_abline(slope = 1, intercept = 0)
-predict.fe_bayes_lr <- function(object, ...) {
+predict.fe_bayes_lr <- function(object, newdata = NULL, ...) {
   constr_fe_bayes_lr(object)
 
+  newdata <- if (is.null(newdata)) {object$fe_data} else {newdata}
+
   # Create linear regression model formula object
-  sp_list <- focal_vs_comp$focal_sp %>%
+  sp_list <- newdata$focal_sp %>%
     levels() %>%
     sort()
   model_formula <- sp_list %>%
@@ -166,7 +161,7 @@ predict.fe_bayes_lr <- function(object, ...) {
     as.formula()
 
   # Create matrices & vectors for Bayesian regression
-  focal_trees <- focal_vs_comp %>%
+  focal_trees <- newdata %>%
[…]

### 2021-01-18 README.Rmd
migrate to github actions for CI

--- 2021-01-09 README.Rmd
@@
 
-[![Travis Build Status](https://travis-ci.org/rudeboybert/forestecology.svg?branch=master)](https://travis-ci.org/rudeboybert/forestecology)
+[![Build Status](https://github.com/rudeboybert/forestecology/workflows/R-CMD-check/badge.svg)](https://github.com/rudeboybert/forestecology/actions)
 [![Lifecycle: stable](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)

### 2021-01-19 README.Rmd
class output of fit function, convert predict to S3 method

@@
 posterior_param_ex <- focal_vs_comp_ex %>%
-  fit_bayesian_model(prior_param = NULL, run_shuffle = FALSE)
+  fe_bayes_lm(prior_param = NULL, run_shuffle = FALSE)
 

### 2021-01-19 R/example_datasets.R

@@ -268,7 +268,7 @@
 #' library(dplyr)
 #'
 #' posterior_param_ex <- focal_vs_comp_ex %>%
-#'   fit_bayesian_model(prior_param = NULL, run_shuffle = FALSE)
+#'   fe_bayes_lm(prior_param = NULL, run_shuffle = FALSE)
 "focal_vs_comp_ex"
 
 
@@ -292,7 +292,7 @@
 #'
 #' # Compare model predictions to observation
 #' predictions <- focal_vs_comp_ex %>%
-#'   predict_bayesian_model(posterior_param = posterior_param_ex) %>%
+#'   predict(posterior_param = posterior_param_ex) %>%
 #'   right_join(ex_growth_df, by = c("focal_ID" = "ID"))
 #' predictions %>%
 #'   rmse(truth = growth, estimate = growth_hat) %>%

### 2021-01-19 R/model_functions.R

@@ -18,7 +18,7 @@
 #' @importFrom tidyr spread
 #' @source Closed-form solutions of Bayesian linear regression \url{https://doi.org/10.1371/journal.pone.0229930.s004}
 #' @return A list of `{a_star, b_star, mu_star, V_star}` posterior hyperparameters
-#' @seealso \code{\link{predict_bayesian_model}}
+#' @seealso \code{\link{predict.fe_bayes_lm}}
 #' @export
 #'
 #' @examples
@@ -28,8 +28,8 @@
 #' data(focal_vs_comp_ex)
 #'
 #' posterior_param_ex <- focal_vs_comp_ex %>%
-#'   fit_bayesian_model(prior_param = NULL, run_shuffle = FALSE)
-fit_bayesian_model <- function(focal_vs_comp, prior_param = NULL, run_shuffle = FALSE) {
+#'   fe_bayes_lm(prior_param = NULL, run_shuffle = FALSE)
+fe_bayes_lm <- function(focal_vs_comp, prior_param = NULL, run_shuffle = FALSE) {
   if (FALSE) {
     focal_vs_comp <- focal_vs_comp_bw
     run_shuffle <- FALSE
@@ -93,7 +93,11 @@ fit_bayesian_model <- function(focal_vs_comp, prior_param = NULL, run_shuffle = 
     V_star = V_star,
     sp_list = sp_list
   )
-  return(posterior_hyperparameters)
+
+  structure(
+    posterior_hyperparameters,
+    class = c("fe_bayes_lm", class(posterior_hyperparameters))
+  )
 }
 
 
@@ -102,17 +106,17 @@ fit_bayesian_model <- function(focal_vs_comp, prior_param = NULL, run_shuffle = 
 
 #' Make predictions based on fitted Bayesian model
 #'
-#' @description Applies fitted model from \code{\link{fit_bayesian_model}} and
+#' @description Applies fitted model from \code{\link{fe_bayes_lm}} and
 #'   returns posterior predicted values.
-#' @inheritParams fit_bayesian_model
-#' @param posterior_param Output of \code{\link{fit_bayesian_model}}: A list of
+#' @inheritParams fe_bayes_lm
+#' @param posterior_param Output of \code{\link{fe_bayes_lm}}: A list of
 #'   `{a_star, b_star, mu_star, V_star}` posterior hyperparameters
 #' @inheritParams create_focal_vs_comp
 #' @import dplyr
 #' @importFrom stats model.matrix
 #' @importFrom tidyr nest
[…]

### 2021-01-19 data-raw/process_data_raw.R

@@ -154,7 +154,7 @@ use_data(focal_vs_comp_ex, overwrite = TRUE)
 
 # fit the model
 posterior_param_ex <- focal_vs_comp_ex %>%
-  fit_bayesian_model()
+  fe_bayes_lm()
 use_data(posterior_param_ex, overwrite = TRUE)
 
 

### 2021-01-19 R/example_datasets.R
bug fixes + R CMD check

@@ -291,16 +291,17 @@
 #' library(yardstick)
 #'
 #' # Compare model predictions to observation
-#' predictions <- focal_vs_comp_ex %>%
-#'   predict(posterior_param = posterior_param_ex) %>%
+#' predictions <- posterior_param_ex %>%
+#'   predict(focal_vs_comp = focal_vs_comp_ex) %>%
 #'   right_join(ex_growth_df, by = c("focal_ID" = "ID"))
+#'
 #' predictions %>%
 #'   rmse(truth = growth, estimate = growth_hat) %>%
 #'   pull(.estimate)
 #'
 #' # Plot posterior parameters
 #' posterior_param_ex %>%
-#'   plot_bayesian_model_parameters()
+#'   autoplot()
 "posterior_param_ex"
 
 

### 2021-01-19 R/model_functions.R

@@ -108,13 +108,17 @@ fe_bayes_lm <- function(focal_vs_comp, prior_param = NULL, run_shuffle = FALSE) 
 #'
 #' @description Applies fitted model from \code{\link{fe_bayes_lm}} and
 #'   returns posterior predicted values.
-#' @inheritParams fe_bayes_lm
-#' @param posterior_param Output of \code{\link{fe_bayes_lm}}: A list of
+#'
+#' @param object Output of \code{\link{fe_bayes_lm}}: A list of
 #'   `{a_star, b_star, mu_star, V_star}` posterior hyperparameters
+#' @inheritParams fe_bayes_lm
 #' @inheritParams create_focal_vs_comp
+#' @param ... Currently igniored—only included for consistency with generic.
+#'
 #' @import dplyr
 #' @importFrom stats model.matrix
 #' @importFrom tidyr nest
+#'
 #' @return \code{focal_vs_comp} with new column of predicted \code{growth_hat}
 #' @seealso \code{\link{fe_bayes_lm}}
 #' @source Closed-form solutions of Bayesian linear regression \url{https://doi.org/10.1371/journal.pone.0229930.s004}
@@ -129,17 +133,18 @@ fe_bayes_lm <- function(focal_vs_comp, prior_param = NULL, run_shuffle = FALSE) 
 #' # and growth data to compare to
 #' data(posterior_param_ex, ex_growth_df)
 #'
-#' predictions <- focal_vs_comp_ex %>%
-#'   predict(posterior_param = posterior_param_ex) %>%
+#' predictions <- posterior_param_ex %>%
+#'   predict(focal_vs_comp = focal_vs_comp_ex) %>%
 #'   right_join(ex_growth_df, by = c("focal_ID" = "ID"))
+#'
 #' predictions %>%
 #'   ggplot(aes(growth, growth_hat)) +
 #'   geom_point() +
 #'   geom_abline(slope = 1, intercept = 0)
-predict.fe_bayes_lm <- function(focal_vs_comp, posterior_param) {
+predict.fe_bayes_lm <- function(object, focal_vs_comp, ...) {
   if (FALSE) {
     focal_vs_comp <- focal_vs_comp_bw
-    posterior_param <- bw_fit_model
+    object <- bw_fit_model
   }
 
   # Create linear regression model formula object
@@ -161,7 +166,7 @@ predict.fe_bayes_lm <- function(focal_vs_comp, posterior_param) {
[…]

### 2021-01-19 R/plot.R

@@ -4,10 +4,11 @@ ggplot2::autoplot
 
 #' Plot Bayesian model parameters
 #'
-#' @param posterior_param Output of \code{\link{fe_bayes_lm}}
+#' @param object Output of \code{\link{fe_bayes_lm}}
 #' @param type A single character string for plot type with possible values
 #' "intercepts", "slopes", or "competition".
 #' @param sp_to_plot Vector of subset of species to plot
+#' @inheritParams predict.fe_bayes_lm
 #'
 #' @import ggridges
 #' @importFrom mvnfast rmvt
@@ -34,19 +35,20 @@ ggplot2::autoplot
 #' # Plot lambdas, competition coefficients
 #' autoplot(posterior_param_ex, type = "competition")
 #' @export
-autoplot.fe_bayes_lm <- function(posterior_param,
+autoplot.fe_bayes_lm <- function(object,
                                  type = "intercepts",
-                                 sp_to_plot = NULL) {
+                                 sp_to_plot = NULL,
+                                 ...) {
 
   # Identify all species and baseline category of species used for regression
-  sp_list <- posterior_param$sp_list
+  sp_list <- object$sp_list
   baseline_species <- sp_list %>%
     sort() %>%
     .[1]
 
   # Simulate observations from posterior
   beta_lambda_posterior_df <- simulate_beta_lambda_posterior(
-    posterior_param,
+    object,
     sp_list,
     baseline_species
   )

### 2021-01-20 README.Rmd
bug fixes + R CMD check
Fix and rebuild README

@@
 # Get predicted dbh values values
-predictions <- focal_vs_comp_ex %>%
-  predict(posterior_param = posterior_param_ex) %>%
+predictions <- posterior_param_ex %>%
+  predict(focal_vs_comp = focal_vs_comp_ex) %>%
   right_join(ex_growth_df, by = c("focal_ID" = "ID"))
@@
 # Plot posteriors
-plot_ex <- posterior_param_ex %>%
-  plot_bayesian_model_parameters()
-plot_ex[["lambda"]]
+posterior_param_ex %>%
+  autoplot(type = "intercepts")
+posterior_param_ex %>%
+  autoplot(type = "dbh_slopes")
+posterior_param_ex %>%
+  autoplot(type = "competition")
 ```

### 2021-01-20 R/example_datasets.R
Renamed object class from fe_bayes_lm to comp_bayes_lm

@@ -268,7 +268,7 @@
 #' library(dplyr)
 #'
 #' posterior_param_ex <- focal_vs_comp_ex %>%
-#'   fe_bayes_lm(prior_param = NULL, run_shuffle = FALSE)
+#'   comp_bayes_lm(prior_param = NULL, run_shuffle = FALSE)
 "focal_vs_comp_ex"
 
 

### 2021-01-20 R/model_functions.R

@@ -18,7 +18,7 @@
 #' @importFrom tidyr spread
 #' @source Closed-form solutions of Bayesian linear regression \url{https://doi.org/10.1371/journal.pone.0229930.s004}
 #' @return A list of `{a_star, b_star, mu_star, V_star}` posterior hyperparameters
-#' @seealso \code{\link{predict.fe_bayes_lm}}
+#' @seealso \code{\link{predict.comp_bayes_lm}}
 #' @export
 #'
 #' @examples
@@ -28,8 +28,8 @@
 #' data(focal_vs_comp_ex)
 #'
 #' posterior_param_ex <- focal_vs_comp_ex %>%
-#'   fe_bayes_lm(prior_param = NULL, run_shuffle = FALSE)
-fe_bayes_lm <- function(focal_vs_comp, prior_param = NULL, run_shuffle = FALSE) {
+#'   comp_bayes_lm(prior_param = NULL, run_shuffle = FALSE)
+comp_bayes_lm <- function(focal_vs_comp, prior_param = NULL, run_shuffle = FALSE) {
   if (FALSE) {
     focal_vs_comp <- focal_vs_comp_bw
     run_shuffle <- FALSE
@@ -96,7 +96,7 @@ fe_bayes_lm <- function(focal_vs_comp, prior_param = NULL, run_shuffle = FALSE) 
 
   structure(
     posterior_hyperparameters,
-    class = c("fe_bayes_lm", class(posterior_hyperparameters))
+    class = c("comp_bayes_lm", class(posterior_hyperparameters))
   )
 }
 
@@ -106,12 +106,12 @@ fe_bayes_lm <- function(focal_vs_comp, prior_param = NULL, run_shuffle = FALSE) 
 
 #' Make predictions based on fitted Bayesian model
 #'
-#' @description Applies fitted model from \code{\link{fe_bayes_lm}} and
+#' @description Applies fitted model from \code{\link{comp_bayes_lm}} and
 #'   returns posterior predicted values.
 #'
-#' @param object Output of \code{\link{fe_bayes_lm}}: A list of
+#' @param object Output of \code{\link{comp_bayes_lm}}: A list of
 #'   `{a_star, b_star, mu_star, V_star}` posterior hyperparameters
-#' @inheritParams fe_bayes_lm
+#' @inheritParams comp_bayes_lm
 #' @inheritParams create_focal_vs_comp
 #' @param ... Currently igniored—only included for consistency with generic.
 #'
[…]

### 2021-01-20 R/plot.R

@@ -4,11 +4,11 @@ ggplot2::autoplot
 
 #' Plot Bayesian model parameters
 #'
-#' @param object Output of \code{\link{fe_bayes_lm}}
+#' @param object Output of \code{\link{comp_bayes_lm}}
 #' @param type A single character string for plot type with possible values
 #' "intercepts", "dbh_slopes", or "competition".
 #' @param sp_to_plot Vector of subset of species to plot
-#' @inheritParams predict.fe_bayes_lm
+#' @inheritParams predict.comp_bayes_lm
 #'
 #' @import ggridges
 #' @importFrom mvnfast rmvt
@@ -35,7 +35,7 @@ ggplot2::autoplot
 #' # Plot lambdas, competition coefficients
 #' autoplot(posterior_param_ex, type = "competition")
 #' @export
-autoplot.fe_bayes_lm <- function(object,
+autoplot.comp_bayes_lm <- function(object,
                                  type = "intercepts",
                                  sp_to_plot = NULL,
                                  ...) {

### 2021-01-20 data-raw/process_data_raw.R

@@ -154,7 +154,7 @@ use_data(focal_vs_comp_ex, overwrite = TRUE)
 
 # fit the model
 posterior_param_ex <- focal_vs_comp_ex %>%
-  fe_bayes_lm()
+  comp_bayes_lm()
 use_data(posterior_param_ex, overwrite = TRUE)
 
 

### 2021-01-20 etc/paper/forestecology.tex
fixed header error
Re-knit manuscript
Added Simon as author

@@
     Biology Department, Middlebury College\\
+     and \\     Simon P. Couch \\
+    Mathematics Department, Reed College\\
       }
@@
 
-\hypertarget{add-foldid-to-each-tree}{%
-\section{Add foldID to each tree}\label{add-foldid-to-each-tree}}
+Then add foldID to each tree
 
@@
 
-\hypertarget{remove-empty-folds}{%
-\section{Remove empty folds}\label{remove-empty-folds}}
+Then remove empty folds
 
@@
 \StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
-\CommentTok{## [1] 0.1533511}
+\CommentTok{## [1] 0.1532316}
 \end{Highlighting}

### 2021-01-24 README.Rmd
Renamed object class from fe_bayes_lm to comp_bayes_lm
update structure of fitted model
now contains the prior parameters, posterior parameters, and the formula object.

renames posterior_param_ex -> comp_bayes_lm_ex

--- 2021-01-19 README.Rmd
@@
 # Fit model
-posterior_param_ex <- focal_vs_comp_ex %>%
-  fe_bayes_lm(prior_param = NULL, run_shuffle = FALSE)
+comp_bayes_lm_ex <- focal_vs_comp_ex %>%
+  comp_bayes_lm(prior_param = NULL, run_shuffle = FALSE)
 
 # Get predicted dbh values values
-predictions <- focal_vs_comp_ex %>%
-  predict(posterior_param = posterior_param_ex) %>%
+predictions <- comp_bayes_lm_ex %>%
+  predict(focal_vs_comp = focal_vs_comp_ex) %>%
   right_join(ex_growth_df, by = c("focal_ID" = "ID"))
@@
 # Plot posteriors
-plot_ex <- posterior_param_ex %>%
-  plot_bayesian_model_parameters()
-plot_ex[["lambda"]]
+comp_bayes_lm_ex %>%
+  autoplot(type = "intercepts")
+comp_bayes_lm_ex %>%
+  autoplot(type = "dbh_slopes")
+comp_bayes_lm_ex %>%
+  autoplot(type = "competition")
 ```

### 2021-01-24 R/example_datasets.R

@@ -275,7 +275,7 @@
 #' @examples
 #' library(dplyr)
 #'
-#' posterior_param_ex <- focal_vs_comp_ex %>%
+#' comp_bayes_lm_ex <- focal_vs_comp_ex %>%
 #'   comp_bayes_lm(prior_param = NULL, run_shuffle = FALSE)
 "focal_vs_comp_ex"
 
@@ -299,7 +299,7 @@
 #' library(yardstick)
 #'
 #' # Compare model predictions to observation
-#' predictions <- posterior_param_ex %>%
+#' predictions <- comp_bayes_lm_ex %>%
 #'   predict(focal_vs_comp = focal_vs_comp_ex) %>%
 #'   right_join(ex_growth_df, by = c("focal_ID" = "ID"))
 #'
@@ -308,9 +308,9 @@
 #'   pull(.estimate)
 #'
 #' # Plot posterior parameters
-#' posterior_param_ex %>%
+#' comp_bayes_lm_ex %>%
 #'   autoplot()
-"posterior_param_ex"
+"comp_bayes_lm_ex"
 
 
 

### 2021-01-24 R/model_functions.R

@@ -27,7 +27,7 @@
 #' # Load in focal versus comp
 #' data(focal_vs_comp_ex)
 #'
-#' posterior_param_ex <- focal_vs_comp_ex %>%
+#' comp_bayes_lm_ex <- focal_vs_comp_ex %>%
 #'   comp_bayes_lm(prior_param = NULL, run_shuffle = FALSE)
 comp_bayes_lm <- function(focal_vs_comp, prior_param = NULL, run_shuffle = FALSE) {
   if (FALSE) {
@@ -57,19 +57,13 @@ comp_bayes_lm <- function(focal_vs_comp, prior_param = NULL, run_shuffle = FALSE
 
   # Set priors. If no prior_param specified:
   if (is.null(prior_param)) {
-    # Prior parameters for sigma2:
-    a_0 <- 250
-    b_0 <- 25
-    # Prior parameters for betas and lambdas:
-    mu_0 <- rep(0, ncol(X)) %>%
-      matrix(ncol = 1)
-    V_0 <- ncol(X) %>% diag()
-  } else {
-    a_0 <- prior_param$a_0
-    b_0 <- prior_param$b_0
-    mu_0 <- prior_param$mu_0
-    V_0 <- prior_param$V_0
+    prior_param <- default_prior_params(X)
   }
+
+  a_0 <- prior_param$a_0
+  b_0 <- prior_param$b_0
+  mu_0 <- prior_param$mu_0
+  V_0 <- prior_param$V_0
 
   # Compute posteriors
   # Posterior parameters for betas and lambdas:
@@ -86,7 +80,7 @@ comp_bayes_lm <- function(focal_vs_comp, prior_param = NULL, run_shuffle = FALSE
     as.vector()
 
   # Return posterior parameters
-  posterior_hyperparameters <- list(
+  post_params <- list(
     a_star = a_star,
     b_star = b_star,
     mu_star = mu_star,
@@ -94,10 +88,60 @@ comp_bayes_lm <- function(focal_vs_comp, prior_param = NULL, run_shuffle = FALSE
     sp_list = sp_list
   )
 
-  structure(
-    posterior_hyperparameters,
-    class = c("comp_bayes_lm", class(posterior_hyperparameters))
+  out <- list(
+    prior_params = prior_param,
+    post_params = post_params,
+    terms = model_formula
   )
+
+  structure(
+    out,
+    class = c("comp_bayes_lm", class(out))
+  )
+}
+
+
+
+
+
+default_prior_params <- function(X) {
[…]

### 2021-01-24 R/plot.R

@@ -24,16 +24,16 @@ ggplot2::autoplot
 #' library(ggridges)
 #'
 #' # Load in posterior parameter example
-#' data(posterior_param_ex)
+#' data(comp_bayes_lm_ex)
 #'
 #' # Plot beta_0, growth intercepts
-#' autoplot(posterior_param_ex, type = "intercepts")
+#' autoplot(comp_bayes_lm_ex, type = "intercepts")
 #'
 #' # Plot beta_dbh, growth-dbh slope
-#' autoplot(posterior_param_ex, type = "dbh_slopes")
+#' autoplot(comp_bayes_lm_ex, type = "dbh_slopes")
 #'
 #' # Plot lambdas, competition coefficients
-#' autoplot(posterior_param_ex, type = "competition")
+#' autoplot(comp_bayes_lm_ex, type = "competition")
 #' @export
 autoplot.comp_bayes_lm <- function(object,
                                  type = "intercepts",
@@ -41,14 +41,14 @@ autoplot.comp_bayes_lm <- function(object,
                                  ...) {
 
   # Identify all species and baseline category of species used for regression
-  sp_list <- object$sp_list
+  sp_list <- object$post_params$sp_list
   baseline_species <- sp_list %>%
     sort() %>%
     .[1]
 
   # Simulate observations from posterior
   beta_lambda_posterior_df <- simulate_beta_lambda_posterior(
-    object,
+    object$post_params,
     sp_list,
     baseline_species
   )

### 2021-01-25 README.Rmd
transition predict to output vector

@@
 # Get predicted dbh values values
-predictions <- comp_bayes_lm_ex %>%
-  predict(focal_vs_comp = focal_vs_comp_ex) %>%
-  right_join(ex_growth_df, by = c("focal_ID" = "ID"))
+predictions <- focal_vs_comp_ex %>%
+  mutate(growth_hat = predict(comp_bayes_lm_ex, focal_vs_comp_ex))
 

### 2021-01-25 R/example_datasets.R

@@ -303,9 +303,8 @@
 #' library(yardstick)
 #'
 #' # Compare model predictions to observation
-#' predictions <- comp_bayes_lm_ex %>%
-#'   predict(focal_vs_comp = focal_vs_comp_ex) %>%
-#'   right_join(ex_growth_df, by = c("focal_ID" = "ID"))
+#' predictions <- focal_vs_comp_ex %>%
+#'   mutate(growth_hat = predict(comp_bayes_lm_ex, focal_vs_comp_ex))
 #'
 #' predictions %>%
 #'   rmse(truth = growth, estimate = growth_hat) %>%

### 2021-01-25 R/model_functions.R

@@ -161,7 +161,7 @@ print.comp_bayes_lm <- function(x, ...) {
 #' @importFrom stats model.matrix
 #' @importFrom tidyr nest
 #'
-#' @return \code{focal_vs_comp} with new column of predicted \code{growth_hat}
+#' @return A vector of predictions with length equal to the input data.
 #' @seealso \code{\link{comp_bayes_lm}}
 #' @source Closed-form solutions of Bayesian linear regression \url{https://doi.org/10.1371/journal.pone.0229930.s004}
 #' @export
@@ -175,9 +175,8 @@ print.comp_bayes_lm <- function(x, ...) {
 #' # and growth data to compare to
 #' data(comp_bayes_lm_ex, ex_growth_df)
 #'
-#' predictions <- comp_bayes_lm_ex %>%
-#'   predict(focal_vs_comp = focal_vs_comp_ex) %>%
-#'   right_join(ex_growth_df, by = c("focal_ID" = "ID"))
+#' predictions <- focal_vs_comp_ex %>%
+#'   mutate(growth_hat = predict(comp_bayes_lm_ex, focal_vs_comp_ex))
 #'
 #' predictions %>%
 #'   ggplot(aes(growth, growth_hat)) +
@@ -201,11 +200,7 @@ predict.comp_bayes_lm <- function(object, focal_vs_comp, ...) {
 
   # Make posterior predictions
   mu_star <- object$post_params$mu_star
-  focal_trees <- focal_trees %>%
-    mutate(growth_hat = as.vector(X %*% mu_star)) %>%
-    select(focal_ID, growth_hat)
-
-  focal_trees
+  as.vector(X %*% mu_star)
 }
 
 
@@ -222,6 +217,7 @@ predict.comp_bayes_lm <- function(object, focal_vs_comp, ...) {
 #' @import dplyr
 #' @import sf
 #' @import sfheaders
+#' @importFrom tibble enframe
 #' @return \code{focal_vs_comp} with new column of predicted \code{growth_hat}
 #' @export
 #'
@@ -282,7 +278,9 @@ fit_one_fold <- function(fold, focal_vs_comp, max_dist,
     comp_bayes_lm(prior_param = prior_param, run_shuffle = run_shuffle)
 
   comp_bayes_lm_fold %>%
-    predict(focal_vs_comp = test)
+    predict(focal_vs_comp = test) %>%
+    enframe(name = NULL, value = "growth_hat") %>%
+    mutate(focal_ID = test$focal_ID)
 }
 
 

### 2021-01-25 tests/testthat/test-readme.R

@@ -69,9 +69,8 @@ test_that("readme code works", {
 
   expect_true(check_comp_bayes_lm(comp_bayes_lm_ex))
 
-  predictions <- comp_bayes_lm_ex %>%
-    predict(focal_vs_comp = focal_vs_comp_ex) %>%
-    right_join(ex_growth_df, by = c("focal_ID" = "ID"))
+  predictions <- focal_vs_comp_ex %>%
+    mutate(growth_hat = predict(comp_bayes_lm_ex, focal_vs_comp_ex))
 
   expect_true(
     check_inherits(

### 2021-01-25 README.Rmd
First pass at refactoring README

@@
 # install.packages("remotes")
-# remotes::install_github("rudeboybert/forestecology")
+remotes::install_github("rudeboybert/forestecology")
 ```
@@
 
-Here we provide an example with a small dataset.
+We present an example analysis using toy data pre-loaded into the package.
 
@@
 
-First we combine two example census data files into a single `tibble` and compute the growth of all surviving individuals.
+### Preprocess census data
+
+We first combine data from two repeat censuses into a single `tibble` and compute the growth of all surviving individuals.
+Note that per Allen and Kim (2020) we first remove resprouts from the second census.
 
 ```{r}
-# Read in census files
 data(census_df1_ex, census_df2_ex)
@@
 
-# Name of variable in tibble that uniquely identifies each stem:
-id <- "ID"
-
-# Merge both censuses and compute growth:
-ex_growth_df <-
-  compute_growth(census_df1_ex, census_df2_ex_no_r, id) %>%
-  mutate(
-    sp = to_any_case(sp),
-    # Must be a factor
-    sp = as.factor(sp)
-  )
+# Merge both censuses and compute growth. Note we require any species variable
+# to be coded as a factor
+census_combined_ex <-
+  compute_growth(census_df1_ex, census_df2_ex_no_r, id = "ID") %>%
+  mutate(sp = to_any_case(sp) %>% as.factor())
 ```
 
-All growth models in the package assume that two individuals compete if they are less than a specified distance away.
-Here we set this distance in `max_dist` and put all individuals within this distance of a plot boundary in a buffer (since we do not have full information about their competitors).
+### Add spatial information
+
+Our growth model assumes that two individual trees compete if they are less than a pre-specified distance `max_dist` apart.
+Furthermore, we define a buffer region of size `max_dist` from the boundary of the study region.
+All trees outside this buffer region will be our "focal" trees of interest since we have complete competitor information on them.
+All trees inside this buffer region will only be considered as "competitor" trees to "focal" trees.
 
 ```{r}
-# Load study region:
 data("ex_study_region")
 
-# Set max dist
+# Set competitor distance
 max_dist <- 1
 
-# Add buffer
-ex_growth_df <- ex_growth_df %>%
+# Create sf representation of buffer region
+buffer_region <- ex_study_region %>% 
+  compute_buffer_region(direction = "in", size = max_dist)
+
+# Add buffer variable
+census_combined_ex <- census_combined_ex %>%
   add_buffer_variable(direction = "in", size = max_dist, region = ex_study_region)
 
-ggplot() +
-  geom_sf(data = ex_growth_df, aes(col = buffer), size = 2)
+base_plot <- ggplot() +
+  geom_sf(data = ex_study_region, fill = "transparent") +
+  geom_sf(data = buffer_region, fill = "transparent", linetype = "dashed")
+
+base_plot + 
+  geom_sf(data = census_combined_ex, aes(col = buffer), size = 2)
 ```
 
-We then set blocks/folds for the spatial cross-validation scheme.
-We this both manually and using the [`blockCV`](https://github.com/rvalavi/blockCV) package.
-Note that here we define the geometric structure of the folds manually using `sfheaders::sf_polygon()`:
+We then manually define the blocks/folds for our spatial cross-validation scheme and convert them to an `sf` object using `sfheaders::sf_polygon()`:
 
@@
 fold2 <- rbind(c(5, 0), c(10, 0), c(10, 5), c(5, 5), c(5, 0))
+
 blocks <- bind_rows(
@@
 
-Next we assign each point to folds using `blockCV::spatialBlock()`
+Next we assign each tree to blocks/folds using the `spatialBlock()` function from the [`blockCV`](https://github.com/rvalavi/blockCV) package.
 
@@
 ex_cv_grid <- spatialBlock(
-  speciesData = ex_growth_df,
-  verbose = FALSE,
-  k = 2,
-  selection = "systematic",
-  blocks = blocks,
-  showBlocks = FALSE
+  speciesData = census_combined_ex, k = 2, selection = "systematic", blocks = blocks,
+  showBlocks = FALSE, verbose = FALSE
 )
 
-# Add foldID to data
-ex_growth_df <- ex_growth_df %>%
-  mutate(foldID = ex_cv_grid$foldID %>% as.factor())
+# Add fold information to data
+census_combined_ex <- census_combined_ex %>%
+  mutate(foldID = ex_cv_grid$foldID %>% factor())
 
-# Deliverable
-ggplot() +
-  geom_sf(data = ex_growth_df, aes(col = buffer, shape = foldID), size = 2) +
-  geom_sf(data = blocks, fill = "transparent")
+base_plot + 
+  geom_sf(data = census_combined_ex, aes(col = buffer, shape = foldID), size = 2) +
+  geom_sf(data = blocks, fill = "transparent", col = "orange")
+```
 
-# Convert to sf object
+We also extract the spatial cross-validation grid itself as an `sf` object for later use.
+
+```{r}
 ex_cv_grid_sf <- ex_cv_grid$blocks %>%
   st_as_sf()
+
+ggplot() +
+  geom_sf(data = ex_cv_grid_sf)
 ```
 
-We then create a `focal_v_comp` tibble which has a row for each competing pair of individuals (each pair of individuals within `max_dist` of one another):
+### Define focal versus competitor trees
+
+We then create a list-column data frame (via `tidyr::nest()`) where each row represents a focal tree of interest and the `comp` variable contains nested information on each of it's competitors
+
+which has a row for each competing pair of individuals (each pair of individuals within `max_dist` of one another):
 
 ```{r}
-focal_vs_comp_ex <- ex_growth_df %>%
+focal_vs_comp_ex <- census_combined_ex %>%
   create_focal_vs_comp(max_dist, cv_grid_sf = ex_cv_grid_sf, id = "ID")
+
+# Nested view
+focal_vs_comp_ex
+
+# Unnested view:
+focal_vs_comp_ex %>% 
+  unnest(cols = "comp")
 ```
 
-With `focal_v_comp`, we can then run the growth model and get predictions from it.
-Here we run the model without cross validation.
-Posterior estimates can be plotted to give, for example, the lambda matrix of competition coefficients.
+### Fit model and make predictions
+
+We then fit our competitor growth model as specified in Allen and Kim (2020)
 
 ```{r}
-# Fit model
 comp_bayes_lm_ex <- focal_vs_comp_ex %>%
   comp_bayes_lm(prior_param = NULL, run_shuffle = FALSE)
+```
 
-# Get predicted dbh values values
-predictions <- focal_vs_comp_ex %>%
+Since the resulting output is an S3 object of class `comp_bayes_lm`, we can both print and visualize its output: the posterior distribution of all the linear regression parameters.
+
+```{r}
+# Object summary
+comp_bayes_lm_ex
+
+# Visualizations
+autoplot(comp_bayes_lm_ex, type = "intercepts")
+autoplot(comp_bayes_lm_ex, type = "dbh_slopes")
+autoplot(comp_bayes_lm_ex, type = "competition")
+```
+
+Furthermore, we append fitted/predicted growth values for each tree using `predict.comp_bayes_lm()` to our data frame and then compute the root mean squared error (RMSE).
+
+```{r}
+focal_vs_comp_ex <- focal_vs_comp_ex %>%
   mutate(growth_hat = predict(comp_bayes_lm_ex, focal_vs_comp_ex))
 
-# Compute RMSE of true vs predicted dbh values
-predictions %>%
-  rmse(truth = growth, estimate = growth_hat) %>%
+# Compute RMSE
+focal_vs_comp_ex %>%
+  yardstick::rmse(truth = growth, estimate = growth_hat) %>%
   pull(.estimate)
-
-# Plot posteriors
-comp_bayes_lm_ex %>%
-  autoplot(type = "intercepts")
-comp_bayes_lm_ex %>%
-  autoplot(type = "dbh_slopes")
-comp_bayes_lm_ex %>%
-  autoplot(type = "competition")
 ```
 
-Here we repeat the process but with cross-validation. Note the increase in RMSE, reflecting the fact that our original estimate of model error was overly optimistic.
+### Run spatial cross-validation
+
+Here we repeat the process but with running spatial cross-validation: we fit the data to one fold and apply it to the other.
+Note the increase in RMSE, reflecting the fact that our original estimate of model error was overly optimistic as it did not account for spatial autocorrelation.
 
 ```{r}
-# Fit model with cross-validation
 ex_bw <- focal_vs_comp_ex %>%
   run_cv(max_dist = max_dist, cv_grid = ex_cv_grid_sf) %>%
-  right_join(ex_growth_df, by = c("focal_ID" = "ID"))
+  right_join(census_combined_ex, by = c("focal_ID" = "ID"))
 
-# Compute RMSE of true vs predicted dbh values
+# Compute RMSE of spatially cross-validated model
 ex_bw %>%
@@
 
-For fuller examples on actual datasets see ___.
+<!--For fuller examples on actual datasets see ___.-->

### 2021-01-25 tests/testthat/test-readme.R

@@ -57,6 +57,7 @@ test_that("readme code works", {
   focal_vs_comp_ex <- ex_growth_df %>%
     create_focal_vs_comp(max_dist, cv_grid_sf = ex_cv_grid_sf, id = "ID")
 
+  # Checks each column in focal_vs_comp is of appropriate type
   expect_true(check_inherits(focal_vs_comp_ex, "data.frame"))
   expect_true(
     check_focal_vs_comp(focal_vs_comp_ex) %>%
@@ -66,6 +67,7 @@ test_that("readme code works", {
 
   comp_bayes_lm_ex <- focal_vs_comp_ex %>%
     comp_bayes_lm(prior_param = NULL, run_shuffle = FALSE)
+
 
   expect_true(check_comp_bayes_lm(comp_bayes_lm_ex))
 

### 2021-01-25 README.Rmd
rebuilt README

@@
 library(snakecase)
+library(patchwork)
 ```
@@
 
-# Visualizations
-autoplot(comp_bayes_lm_ex, type = "intercepts")
-autoplot(comp_bayes_lm_ex, type = "dbh_slopes")
-autoplot(comp_bayes_lm_ex, type = "competition")
+# Visualizations (combined with patchwork pkg)
+p1 <- autoplot(comp_bayes_lm_ex, type = "intercepts")
+p2 <- autoplot(comp_bayes_lm_ex, type = "dbh_slopes")
+p3 <- autoplot(comp_bayes_lm_ex, type = "competition")
+(p1 | p2) / p3
 ```

### 2021-01-25 etc/paper/forestecology.Rmd
Added Simon as author
First pass at refactoring README
Renamed all objects to follow type_instance convention. Ex: growth_bw and growth_ex

--- 2021-01-13 etc/paper/forestecology.Rmd
@@
   affiliation: Biology Department, Middlebury College
+
+- name: Simon P. Couch
+  affiliation: Mathematics Department, Reed College
 
@@
 ```{r bw-load-data}
-data(bw_census_2008, bw_census_2014, bw_species)
+data(census_2008_bw, census_2014_bw, species_bw)
 
 # Append additional species data
-bw_census_2008 <- bw_census_2008 %>%
-  left_join(bw_species, by = "sp") %>%
+census_2008_bw <- census_2008_bw %>%
+  left_join(species_bw, by = "sp") %>%
   select(-c(genus, species, latin))
@@
 
+MERGE THIS SUBSECTION with previous?
+
 For each plot we then compute average annual growth between the two censuses using the `compute_growth()` function.
@@
 ```{r bw-compute-growth}
-bw_census_2014 <- bw_census_2014 %>%
+census_2014_bw <- census_2014_bw %>%
   filter(!str_detect(codes, "R"))
@@
 bw_growth_df <-
-  compute_growth(bw_census_2008, bw_census_2014, id = "treeID") %>%
+  compute_growth(census_2008_bw, census_2014_bw, id = "treeID") %>%
   # Convert all variables denoting species to factors
@@
 This necessitates information about the study region boundary.
-In this case, we use a `sf_polygon` object `bw_study_region` which comes pre-loaded in the `forestecology` packages.
-After loading `bw_study_region`, we illustrate the results of the `add_buffer_variable()` function in Figure \ref{fig:bw-define-buffer}.
+In this case, we use a `sf_polygon` object `study_region_bw` which comes pre-loaded in the `forestecology` packages.
+After loading `study_region_bw`, we illustrate the results of the `add_buffer_variable()` function in Figure \ref{fig:bw-define-buffer}.
 Trees on the periphery denote with lighter colors are part of the buffer and will not be considered as "focal" trees of interest going forward; they will only be considered as competitor trees.
@@
 ```{r bw-define-buffer, out.width="100%", fig.cap="Buffer region for Big Woods study region."}
-data(bw_study_region)
+data(study_region_bw)
 
@@
   add_buffer_variable(direction = "in", size = max_dist, 
-                      region = bw_study_region)
+                      region = study_region_bw)
 

### 2021-01-25 R/data_processing_functions.R

@@ -3,9 +3,9 @@
 #' Based on two tree censuses compute the average annual growth in dbh for all
 #' trees that were alive at (TODO: fill this in).
 #'
-#' @param census_df1 A data frame of the first census.
-#' @param census_df2 A data frame of the second (later) census
-#' @param id Name of variable common to \code{census_df1} and \code{census_df2}
+#' @param census_1 A data frame of the first census.
+#' @param census_2 A data frame of the second (later) census
+#' @param id Name of variable common to \code{census_1} and \code{census_2}
 #' allowing you to join/merge both data frames.
 #'
 #' @return growth_df An sf data frame with \code{growth}: average annual growth in dbh.
@@ -16,13 +16,13 @@
 #' library(dplyr)
 #' library(stringr)
 #' # Load in data from two forst censuses
-#' data(census_df1_ex, census_df2_ex)
+#' data(census_1_ex, census_2_ex)
 #' # Filter out resprouts in second census
-#' census_df2_ex_no_r <- census_df2_ex %>%
+#' census_2_ex_no_r <- census_2_ex %>%
 #'   filter(!str_detect(codes, "R"))
 #' id <- "ID"
-#' ex_growth_df <- compute_growth(census_df1_ex, census_df2_ex_no_r, id)
-compute_growth <- function(census_df1, census_df2, id) {
+#' growth_ex <- compute_growth(census_1_ex, census_2_ex_no_r, id)
+compute_growth <- function(census_1, census_2, id) {
 
   # TODO: Write following checks
   # - Both census data frames have variables: id, dbh, date, and codes.
@@ -31,15 +31,15 @@ compute_growth <- function(census_df1, census_df2, id) {
   # - Prompt use with message: "Assuming dbh are in cm"
 
   # Limit second census data to only those variables that can change
-  census_df2 <- census_df2 %>%
+  census_2 <- census_2 %>%
     select(all_of(id), dbh, date, codes)
 
-  growth_df <- census_df1 %>%
+  growth_df <- census_1 %>%
     filter(dbh > 0) %>%
     # TODO: Hey Dave, don't we want inner_join here then?
     # left_join because we want all trees from census 1 (competitors) but
     # only want trees from census 2 that were alive in 1 (to see how much they grew)
[…]

### 2021-01-25 R/example_datasets.R

@@ -15,24 +15,24 @@
 #'   this stands for resprout.}
 #'   \item{dbh}{Diameter at breast height of the tree in cm}
 #' }
-#' @seealso \code{\link{census_df2_ex}}
+#' @seealso \code{\link{census_2_ex}}
 #' @examples
-#' data(census_df1_ex, census_df2_ex)
+#' data(census_1_ex, census_2_ex)
 #' library(dplyr)
 #' library(stringr)
 #' library(snakecase)
 #' # Filter out resprouts
-#' census_df2_ex_no_r <- census_df2_ex %>%
+#' census_2_ex_no_r <- census_2_ex %>%
 #'   filter(!str_detect(codes, "R"))
 #' id <- "ID"
-#' ex_growth_df <-
+#' growth_ex <-
 #'   # Merge both censuses and compute growth:
-#'   compute_growth(census_df1_ex, census_df2_ex_no_r, id) %>%
+#'   compute_growth(census_1_ex, census_2_ex_no_r, id) %>%
 #'   mutate(
 #'     sp = to_any_case(sp),
 #'     sp = as.factor(sp)
 #'   )
-"census_df1_ex"
+"census_1_ex"
 
 
 
@@ -55,24 +55,24 @@
 #'   this stands for resprout.}
 #'   \item{dbh}{Diameter at breast height of the tree in cm}
 #' }
-#' @seealso \code{\link{census_df1_ex}}
+#' @seealso \code{\link{census_1_ex}}
 #' @examples
-#' data(census_df1_ex, census_df2_ex)
+#' data(census_1_ex, census_2_ex)
 #' library(dplyr)
 #' library(stringr)
 #' library(snakecase)
 #' # Filter out resprouts
-#' census_df2_ex_no_r <- census_df2_ex %>%
+#' census_2_ex_no_r <- census_2_ex %>%
 #'   filter(!str_detect(codes, "R"))
 #' id <- "ID"
-#' ex_growth_df <-
+#' growth_ex <-
 #'   # Merge both censuses and compute growth:
-#'   compute_growth(census_df1_ex, census_df2_ex_no_r, id) %>%
+#'   compute_growth(census_1_ex, census_2_ex_no_r, id) %>%
 #'   mutate(
 #'     sp = to_any_case(sp),
 #'     sp = as.factor(sp)
 #'   )
-"census_df2_ex"
+"census_2_ex"
 
 
 
@@ -83,20 +83,20 @@
[…]

### 2021-01-25 R/model_functions.R

@@ -173,7 +173,7 @@ print.comp_bayes_lm <- function(x, ...) {
 #'
 #' # Load in posterior parameter example
 #' # and growth data to compare to
-#' data(comp_bayes_lm_ex, ex_growth_df)
+#' data(comp_bayes_lm_ex, growth_ex)
 #'
 #' predictions <- focal_vs_comp_ex %>%
 #'   mutate(growth_hat = predict(comp_bayes_lm_ex, focal_vs_comp_ex))

### 2021-01-25 data-raw/process_data_raw.R

@@ -9,8 +9,9 @@ library(blockCV)
 
 
 
-# Boundary polygon of bigwoods region ------------------------------------------
-bw_study_region <-
+# Big Woods data ----
+## Boundary of study region ----
+study_region_bw <-
   tibble(
     # Study region boundary
     x = c(-100, -100, -300, -300, -200, -200, 300, 300, 400, 400, 500, 500, -100),
@@ -18,12 +19,10 @@ bw_study_region <-
   ) %>%
   # Convert to sf object
   sf_polygon()
-use_data(bw_study_region, overwrite = TRUE)
+use_data(study_region_bw, overwrite = TRUE)
 
 
-
-# Species info -----------------------------------------------------------------
-# Get trait_cluster variable for all bigwood species
+## Species, family, and trait_cluster classification of all species
 families <- "data-raw/species_list.csv" %>%
   read_csv() %>%
   mutate(
@@ -34,13 +33,12 @@ families <- "data-raw/species_list.csv" %>%
       spcode == "Highbush Blueberry" ~ "highbush blueberry",
       TRUE ~ spcode
     ),
-    # Rename:
     trait_group = traitclust2
   ) %>%
   mutate_at(c("spcode", "family", "trait_group"), to_any_case) %>%
   select(spcode, family, trait_group)
 
-bw_species <-
+species_bw <-
   "https://deepblue.lib.umich.edu/data/downloads/000000086" %>%
   read_delim(delim = "\t") %>%
   # convert all to snake case:
@@ -51,12 +49,11 @@ bw_species <-
   ) %>%
   select(sp = spcode, genus, species, latin, family) %>%
   left_join(families, by = c("sp" = "spcode", "family"))
-use_data(bw_species, overwrite = TRUE)
+use_data(species_bw, overwrite = TRUE)
 
 
-
-# Import bigwoods data
-bw_census_2008 <-
+# Import census data
+census_2008_bw <-
   "https://deepblue.lib.umich.edu/data/downloads/z603qx485" %>%
   read_delim(delim = "\t") %>%
   mutate(spcode = to_any_case(spcode)) %>%
[…]

### 2021-01-25 etc/old_code/pkg_todo.R

@@ -19,8 +19,8 @@ library(ggridges)
 
 # Load & preprocess data -------------------------------------------------------
 # Append additional species data
-bw_census_2008 <- bw_census_2008 %>%
-  left_join(bw_species, by = "sp") %>%
+census_2008_bw <- census_2008_bw %>%
+  left_join(species_bw, by = "sp") %>%
   select(-c(genus, species, latin))
 
 # Maybe?
@@ -30,9 +30,9 @@ bw_census_2008 <- bw_census_2008 %>%
 
 
 # Compute growth ---------------------------------------------------------------
-census_2008 <- bw_census_2008
+census_2008 <- census_2008_bw
 # we need to filter out the resprouts
-census_2014 <- bw_census_2014 %>%
+census_2014 <- census_2014_bw %>%
   filter(!str_detect(codes, "R"))
 
 # How to designate unique identifier?
@@ -66,7 +66,7 @@ max_dist <- 7.5
 
 # Add buffer variable to data frame
 bw_growth_df <- bw_growth_df %>%
-  add_buffer_variable(direction = "in", size = max_dist, region = bw_study_region)
+  add_buffer_variable(direction = "in", size = max_dist, region = study_region_bw)
 
 ggplot() +
   geom_sf(data = bw_growth_df, aes(col = buffer))

### 2021-01-25 etc/old_code/spatial_block_issue.R

@@ -7,37 +7,37 @@ library(sf)
 library(sfheaders)
 
 # Read in census files
-data(census_df1_ex, census_df2_ex)
+data(census_1_ex, census_2_ex)
 
 # Filter out resprouts
-census_df2_ex_no_r <- census_df2_ex %>%
+census_2_ex_no_r <- census_2_ex %>%
   filter(!str_detect(codes, 'R'))
 
 id <- 'ID'
 
-ex_growth_df <-
+growth_ex <-
   # Merge both censuses and compute growth:
-  compute_growth(census_df1_ex, census_df2_ex_no_r, id) %>%
+  compute_growth(census_1_ex, census_2_ex_no_r, id) %>%
   mutate(
     sp = to_any_case(sp),
     sp = as.factor(sp)) %>%
   # drop stemID
   select(-ID)
 
-data(ex_study_region)
+data(study_region_ex)
 
 # set max dist
 max_dist <- 1
 
 # add buffer
-ex_growth_df <- ex_growth_df %>%
-  add_buffer_variable(direction = "in", size = max_dist, region = ex_study_region)
+growth_ex <- growth_ex %>%
+  add_buffer_variable(direction = "in", size = max_dist, region = study_region_ex)
 
 ggplot() +
-  geom_sf(data = ex_growth_df, aes(col = buffer), size = 2)
+  geom_sf(data = growth_ex, aes(col = buffer), size = 2)
 
-#ex_cv_grid <- spatialBlock(
-#  speciesData = ex_growth_df, theRange = 5, verbose = FALSE, k = 2
+#cv_grid_ex <- spatialBlock(
+#  speciesData = growth_ex, theRange = 5, verbose = FALSE, k = 2
 #)
 
 
@@ -52,12 +52,12 @@ blocks <- bind_rows(
 
 # Plot
 ggplot() +
-  geom_sf(data = ex_growth_df, aes(col = buffer), size = 2) +
+  geom_sf(data = growth_ex, aes(col = buffer), size = 2) +
   geom_sf(data = blocks, fill = "transparent")
 
 # fit spatialBlock()
-ex_cv_grid <- spatialBlock(
-  speciesData = ex_growth_df,
+cv_grid_ex <- spatialBlock(
+  speciesData = growth_ex,
   verbose = FALSE,
   k = 2,
   # Note new arguments
@@ -67,14 +67,14 @@ ex_cv_grid <- spatialBlock(
 )
 
[…]

### 2021-01-25 tests/testthat/test-readme.R

@@ -12,13 +12,13 @@ test_that("readme code works", {
   library(yardstick)
   library(snakecase)
 
-  census_df2_ex_no_r <- census_df2_ex %>%
+  census_2_ex_no_r <- census_2_ex %>%
     filter(!str_detect(codes, "R"))
 
   id <- "ID"
 
-  ex_growth_df <-
-    compute_growth(census_df1_ex, census_df2_ex_no_r, id) %>%
+  growth_ex <-
+    compute_growth(census_1_ex, census_2_ex_no_r, id) %>%
     mutate(
       sp = to_any_case(sp),
       sp = as.factor(sp)
@@ -26,10 +26,10 @@ test_that("readme code works", {
 
   max_dist <- 1
 
-  ex_growth_df <- ex_growth_df %>%
-    add_buffer_variable(direction = "in", size = max_dist, region = ex_study_region)
+  growth_ex <- growth_ex %>%
+    add_buffer_variable(direction = "in", size = max_dist, region = study_region_ex)
 
-  expect_true(check_inherits(ex_growth_df, "data.frame"))
+  expect_true(check_inherits(growth_ex, "data.frame"))
 
   fold1 <- rbind(c(0, 0), c(5, 0), c(5, 5), c(0, 5), c(0, 0))
   fold2 <- rbind(c(5, 0), c(10, 0), c(10, 5), c(5, 5), c(5, 0))
@@ -39,8 +39,8 @@ test_that("readme code works", {
   ) %>%
     mutate(foldID = c(1, 2))
 
-  ex_cv_grid <- spatialBlock(
-    speciesData = ex_growth_df,
+  cv_grid_ex <- spatialBlock(
+    speciesData = growth_ex,
     verbose = FALSE,
     k = 2,
     selection = "systematic",
@@ -48,14 +48,14 @@ test_that("readme code works", {
     showBlocks = FALSE
   )
 
-  ex_growth_df <- ex_growth_df %>%
-    mutate(foldID = ex_cv_grid$foldID %>% as.factor())
+  growth_ex <- growth_ex %>%
+    mutate(foldID = cv_grid_ex$foldID %>% as.factor())
 
-  ex_cv_grid_sf <- ex_cv_grid$blocks %>%
+  cv_grid_sf_ex <- cv_grid_ex$blocks %>%
     st_as_sf()
 
-  focal_vs_comp_ex <- ex_growth_df %>%
[…]

### 2021-01-25 README.Rmd
Refactored README upto run_cv()

@@
 
-### Preprocess census data
+### Compute growth of trees based on census data
 
-We first combine data from two repeat censuses into a single `tibble` and compute the growth of all surviving individuals.
-Note that per Allen and Kim (2020) we first remove resprouts from the second census.
+We first combine data from two repeat censuses into a single `growth` data frame that has the average annual growth of all trees alive at both censuses that aren't resprouts at the second census per Allen and Kim (2020).
 
 ```{r}
-data(census_df1_ex, census_df2_ex)
-
-# Filter out resprouts
-census_df2_ex_no_r <- census_df2_ex %>%
-  filter(!str_detect(codes, "R"))
-
-# Merge both censuses and compute growth. Note we require any species variable
-# to be coded as a factor
-census_combined_ex <-
-  compute_growth(census_df1_ex, census_df2_ex_no_r, id = "ID") %>%
-  mutate(sp = to_any_case(sp) %>% as.factor())
+growth_ex <-
+  compute_growth(
+    census_1 = census_1_ex,
+    census_2 = census_2_ex %>% filter(!str_detect(codes, "R")),
+    id = "ID"
+    ) %>%
+  mutate(sp = to_any_case(sp) %>% factor())
 ```
@@
 
-Our growth model assumes that two individual trees compete if they are less than a pre-specified distance `max_dist` apart.
-Furthermore, we define a buffer region of size `max_dist` from the boundary of the study region.
-All trees outside this buffer region will be our "focal" trees of interest since we have complete competitor information on them.
-All trees inside this buffer region will only be considered as "competitor" trees to "focal" trees.
+Our growth model assumes that two individual trees compete if they are less than a pre-specified distance `comp_dist` apart.
+Furthermore, we define a buffer region of size `comp_dist` from the boundary of the study region.
 
 ```{r}
-data("ex_study_region")
+# Set competitor distance
+comp_dist <- 1
 
-# Set competitor distance
-max_dist <- 1
+# Add buffer variable to growth data frame
+growth_ex <- growth_ex %>%
+  add_buffer_variable(direction = "in", size = comp_dist, region = study_region_ex)
 
-# Create sf representation of buffer region
-buffer_region <- ex_study_region %>% 
-  compute_buffer_region(direction = "in", size = max_dist)
+# Optional: Create sf representation of buffer region
+buffer_region <- study_region_ex %>% 
+  compute_buffer_region(direction = "in", size = comp_dist)
+```
 
-# Add buffer variable
-census_combined_ex <- census_combined_ex %>%
-  add_buffer_variable(direction = "in", size = max_dist, region = ex_study_region)
+In the visualization below, the solid line represents the boundary of the study region while the dashed line delimits the buffer region within.
+All trees outside this buffer region (in red) will be our "focal" trees of interest in our model since we have complete competitor information on all of them.
+All trees inside this buffer region (in blue) will only be considered as "competitor" trees to "focal" trees.
 
+```{r}
 base_plot <- ggplot() +
-  geom_sf(data = ex_study_region, fill = "transparent") +
+  geom_sf(data = study_region_ex, fill = "transparent") +
   geom_sf(data = buffer_region, fill = "transparent", linetype = "dashed")
@@
 base_plot + 
-  geom_sf(data = census_combined_ex, aes(col = buffer), size = 2)
+  geom_sf(data = growth_ex, aes(col = buffer), size = 2)
 ```
 
-We then manually define the blocks/folds for our spatial cross-validation scheme and convert them to an `sf` object using `sfheaders::sf_polygon()`:
+Next we add information pertaining to our spatial cross-validation scheme.
+We first manually define the spatial blocks that will act as our cross-validation folds and convert them to an `sf` object using `sf_polygon()` function from the `sfheaders` package.
 
@@
 
-Next we assign each tree to blocks/folds using the `spatialBlock()` function from the [`blockCV`](https://github.com/rvalavi/blockCV) package.
+Next we assign each tree to the correct folds using the `foldID` variable of the output returned by the `spatialBlock()` function from the [`blockCV`](https://github.com/rvalavi/blockCV) package.
 
 ```{r}
-ex_cv_grid <- spatialBlock(
-  speciesData = census_combined_ex, k = 2, selection = "systematic", blocks = blocks,
+SpatialBlock_ex <- spatialBlock(
+  speciesData = growth_ex, k = 2, selection = "systematic", blocks = blocks,
   showBlocks = FALSE, verbose = FALSE
@@
 
-# Add fold information to data
-census_combined_ex <- census_combined_ex %>%
-  mutate(foldID = ex_cv_grid$foldID %>% factor())
-
-base_plot + 
-  geom_sf(data = census_combined_ex, aes(col = buffer, shape = foldID), size = 2) +
-  geom_sf(data = blocks, fill = "transparent", col = "orange")
+growth_ex <- growth_ex %>%
+  mutate(foldID = SpatialBlock_ex$foldID %>% factor())
 ```
@@
 ```{r}
-ex_cv_grid_sf <- ex_cv_grid$blocks %>%
+cv_grid_sf_ex <- SpatialBlock_ex$blocks %>%
   st_as_sf()
-
 ggplot() +
-  geom_sf(data = ex_cv_grid_sf)
+  geom_sf(data = cv_grid_sf_ex)
 ```
 
-### Define focal versus competitor trees
-
-We then create a list-column data frame (via `tidyr::nest()`) where each row represents a focal tree of interest and the `comp` variable contains nested information on each of it's competitors
-
-which has a row for each competing pair of individuals (each pair of individuals within `max_dist` of one another):
+In the visualization below, the spatial blocks that act as our cross-validation folds are delineated in orange.
+Correspondingly the shape of each point indicates which fold each tree has been assigned to.
 
 ```{r}
-focal_vs_comp_ex <- census_combined_ex %>%
-  create_focal_vs_comp(max_dist, cv_grid_sf = ex_cv_grid_sf, id = "ID")
+base_plot + 
+  geom_sf(data = growth_ex, aes(col = buffer, shape = foldID), size = 2) +
+  geom_sf(data = blocks, fill = "transparent", col = "orange")
+```
 
-# Nested view
+### Compute focal versus competitor tree information
+
+Based on our `growth` data frame, we now explicitly define all "focal" trees and their respective "competitor" trees in a `focal_vs_comp` data frame.
+This data frame has rows corresponding to each focal tree and all information about its competitors are saved in list-column variable `comp`.
+We implemented this nested format using `tidyr::nest()` in order to minimize redundancy, given that the same tree can act as a competitor multiple times.
+
+```{r}
+focal_vs_comp_ex <- growth_ex %>%
+  create_focal_vs_comp(comp_dist, cv_grid_sf = cv_grid_sf_ex, id = "ID")
 focal_vs_comp_ex
+```
 
-# Unnested view:
+Using `unnest()` we can fully expand the competitor information saved in the `focal_vs_comp` data frame.
+For example the tree with `focal_ID` equal to 2 located at (1.5, 2.5) has two competitors within `comp_dist` distance from it.
+
+```{r}
 focal_vs_comp_ex %>% 
@@
 
-We then fit our competitor growth model as specified in Allen and Kim (2020)
+We then fit our competitor growth model as specified in Allen and Kim (2020).
 
@@
 comp_bayes_lm_ex <- focal_vs_comp_ex %>%
-  comp_bayes_lm(prior_param = NULL, run_shuffle = FALSE)
+  comp_bayes_lm(prior_param = NULL)
 ```
 
-Since the resulting output is an S3 object of class `comp_bayes_lm`, we can both print and visualize its output: the posterior distribution of all the linear regression parameters.
+Since the resulting output is an S3 object of class `comp_bayes_lm` we defined, we can both print and visualize its output: the posterior distribution of all linear regression parameters: the intercept, the slope for dbh for each species, and a matrix of all species pairs competitive effects on growth.
 
 ```{r}
-# Object summary
+# Print
 comp_bayes_lm_ex
@@
 
-Furthermore, we append fitted/predicted growth values for each tree using `predict.comp_bayes_lm()` to our data frame and then compute the root mean squared error (RMSE).
+Furthermore, we can apply a generic `predict()` function to the resulting `comp_bayes_lm` object to obtain fitted/predicted values of this model.
+We append these `growth_hat` values to our `focal_vs_comp` data frame.
 
@@
   mutate(growth_hat = predict(comp_bayes_lm_ex, focal_vs_comp_ex))
+focal_vs_comp_ex
+```
+We then compute the root mean squared error (RMSE) of the observed versus fitted growths as a measure of our model's fit.
 
-# Compute RMSE
+```{r}
 focal_vs_comp_ex %>%
@@
 
-Here we repeat the process but with running spatial cross-validation: we fit the data to one fold and apply it to the other.
+Whereas in our example above we fit our model to the entirety of the data and then generated fitted/predicted growths on this same data, we now apply the same model with spatial cross-validation.
+All the trees in a given fold will be given a turn as the "test" data while the trees is all remaining folds will be the "training" data.
+We then fit the model to the training data, but compute fitted/predicted growths for the separate and independent data.
+
+```{r}
+ex_bw <- focal_vs_comp_ex %>%
+  run_cv(comp_dist = comp_dist, cv_grid = cv_grid_sf_ex) %>%
+  right_join(growth_ex, by = c("focal_ID" = "ID"))
+```
+
 Note the increase in RMSE, reflecting the fact that our original estimate of model error was overly optimistic as it did not account for spatial autocorrelation.
@@
 ```{r}
-ex_bw <- focal_vs_comp_ex %>%
-  run_cv(max_dist = max_dist, cv_grid = ex_cv_grid_sf) %>%
-  right_join(census_combined_ex, by = c("focal_ID" = "ID"))
-
-# Compute RMSE of spatially cross-validated model
 ex_bw %>%

### 2021-01-25 etc/paper/forestecology.Rmd
Refactored README upto run_cv()

@@
 ```{r}
-max_dist <- 7.5
+comp_dist <- 7.5
 cv_fold_size <- 100
@@
 
-The first constant is `max_dist` which defines the maximum distance for a tree's competitive neighborhood.
+The first constant is `comp_dist` which defines the maximum distance for a tree's competitive neighborhood.
 Trees within this distance of each other are assumed to compete while those farther than this distance apart do not.
-Put differently, all trees within `max_dist` of a focal tree will be considered its competitors (see below).
-Other studies have estimated the value of `max_dist`; we use an average of estimated values @canham_neighborhood_2004, @uriarte_spatially_2004, @tatsumi2013, @canham_neighborhood_2006.
+Put differently, all trees within `comp_dist` of a focal tree will be considered its competitors (see below).
+Other studies have estimated the value of `comp_dist`; we use an average of estimated values @canham_neighborhood_2004, @uriarte_spatially_2004, @tatsumi2013, @canham_neighborhood_2006.
 
-Furthermore, `max_dist` will define the size of all buffers considered, which will be encoded as a binary variable `buffer` as computed by the `add_buffer_variable()` function.
-This function takes as input the main `growth_df` data frame, the `size` of the buffer which we set as `max_dist`, and the boundary of the study region encoded as a simple features polygon @pebesma_simple_2018.
+Furthermore, `comp_dist` will define the size of all buffers considered, which will be encoded as a binary variable `buffer` as computed by the `add_buffer_variable()` function.
+This function takes as input the main `growth_df` data frame, the `size` of the buffer which we set as `comp_dist`, and the boundary of the study region encoded as a simple features polygon @pebesma_simple_2018.
 DESCRIBE SF PACKAGE.
@@
 bw_growth_df <- bw_growth_df %>%
-  add_buffer_variable(direction = "in", size = max_dist, 
+  add_buffer_variable(direction = "in", size = comp_dist, 
                       region = study_region_bw)
@@
 scbi_growth_df <- scbi_growth_df %>%
-  add_buffer_variable(direction = "in", size = max_dist, 
+  add_buffer_variable(direction = "in", size = comp_dist, 
                       region = scbi_study_region)
@@
 
-Next we define `focal_vs_comp` data frames which connects each focal tree in the `growth_df` data frames to the trees in its competitive neighborhood range as defined by the `max_dist` constant.
-So for example, if `growth_df` consisted of two focal trees with two and three neighbors with `max_dist` respectively, `focal_vs_comp` would be a data frame of 5 rows connecting each focal tree to it's competitors.
-The `create_focal_vs_comp()` function makes this connection taking as inputs the `growth_df` data frame; the `max_dist` constant defining competitive range; `cv_grid_sf`, giving the cross validation grid; and the `id` variable.
+Next we define `focal_vs_comp` data frames which connects each focal tree in the `growth_df` data frames to the trees in its competitive neighborhood range as defined by the `comp_dist` constant.
+So for example, if `growth_df` consisted of two focal trees with two and three neighbors with `comp_dist` respectively, `focal_vs_comp` would be a data frame of 5 rows connecting each focal tree to it's competitors.
+The `create_focal_vs_comp()` function makes this connection taking as inputs the `growth_df` data frame; the `comp_dist` constant defining competitive range; `cv_grid_sf`, giving the cross validation grid; and the `id` variable.
 
@@
 focal_vs_comp_bw <- bw_growth_df %>%
-  create_focal_vs_comp(max_dist, cv_grid_sf = bw_cv_grid, id = "treeID")
+  create_focal_vs_comp(comp_dist, cv_grid_sf = bw_cv_grid, id = "treeID")
 ```
@@
 focal_vs_comp_scbi <- scbi_growth_df %>%
-  create_focal_vs_comp(max_dist, cv_grid_sf = scbi_cv_grid, id = "stemID")
+  create_focal_vs_comp(comp_dist, cv_grid_sf = scbi_cv_grid, id = "stemID")
 ```
@@
 cv_bw <- focal_vs_comp_bw %>%
-  run_cv(max_dist = max_dist, cv_grid = bw_cv_grid) %>%
+  run_cv(comp_dist = comp_dist, cv_grid = bw_cv_grid) %>%
   right_join(bw_growth_df, by = c("focal_ID" = "treeID"))
@@
 cv_scbi <- focal_vs_comp_scbi %>%
-  run_cv(max_dist = max_dist, cv_grid = scbi_cv_grid) %>%
+  run_cv(comp_dist = comp_dist, cv_grid = scbi_cv_grid) %>%
   right_join(scbi_growth_df, by = c("focal_ID" = "treeID"))
@@
   mutate(sp = family) %>%
-  create_focal_vs_comp(max_dist = max_dist, cv_grid_sf = bw_cv_grid, 
+  create_focal_vs_comp(comp_dist = comp_dist, cv_grid_sf = bw_cv_grid, 
                        id = "treeID")

### 2021-01-25 R/data_processing_functions.R

@@ -5,7 +5,7 @@
 #'
 #' @param census_1 A data frame of the first census.
 #' @param census_2 A data frame of the second (later) census
-#' @param id Name of variable common to \code{census_1} and \code{census_2}
+#' @param id Name of variable that uniquely identifies each tree common to \code{census_1} and \code{census_2}
 #' allowing you to join/merge both data frames.
 #'
 #' @return growth_df An sf data frame with \code{growth}: average annual growth in dbh.
@@ -61,7 +61,7 @@ compute_growth <- function(census_1, census_2, id) {
 #'
 #' @param growth_df A \code{\link{compute_growth}} output converted to \code{sf} object
 #' @param cv_grid_sf An sf object of a \code{blockCV} block output
-#' @param max_dist Distance to determine which neighboring trees to a focal tree are competitors.
+#' @param comp_dist Distance to determine which neighboring trees to a focal tree are competitors.
 #' @param id A character string of the variable name in \code{growth_df} uniquely identifying each tree
 #' @return \code{focal_vs_comp} data frame of all focal trees and for each focal
 #'   tree all possible competitor trees. In particular, for each competitor tree
@@ -82,7 +82,7 @@ compute_growth <- function(census_1, census_2, id) {
 #'   `"R"` in the `codes2` variable (OK if a resprout at first census)
 #' }
 #' For each focal tree, "competitor trees" are all trees that (1) were alive at
-#' the first census and (2) within \code{max_dist} distance of the focal tree.
+#' the first census and (2) within \code{comp_dist} distance of the focal tree.
 #' @note In order to speed computation, in particular of distances between all
 #'   focal/competitor tree pairs, we use the cross-validation \code{blockCV}
 #'   object to divide the study region into smaller subsets.
@@ -96,7 +96,7 @@ compute_growth <- function(census_1, census_2, id) {
 #' library(tibble)
 #'
 #' # Create fold information sf object. TODO: clean this
-#' cv_grid_ex <-
+#' SpatialBlock_ex <-
 #'   tibble(
 #'     # Study region boundary
 #'     x = c(0, 0, 5, 5),
@@ -106,7 +106,7 @@ compute_growth <- function(census_1, census_2, id) {
[…]

### 2021-01-25 R/example_datasets.R

@@ -151,7 +151,7 @@
 #' ) %>%
 #'   mutate(foldID = c(1, 2))
 #'
-#' cv_grid_ex <- spatialBlock(
+#' SpatialBlock_ex <- spatialBlock(
 #'   speciesData = growth_ex,
 #'   verbose = FALSE,
 #'   k = 2,
@@ -161,7 +161,7 @@
 #'
 #' # Add foldID to data
 #' growth_spatial_ex <- growth_spatial_ex %>%
-#'   mutate(foldID = cv_grid_ex$foldID %>% as.factor())
+#'   mutate(foldID = SpatialBlock_ex$foldID %>% as.factor())
 "growth_ex"
 
 
@@ -195,7 +195,7 @@
 #' library(dplyr)
 #' library(sf)
 #'
-#' max_dist <- 1
+#' comp_dist <- 1
 #'
 #' ggplot() +
 #'   geom_sf(data = growth_spatial_ex, aes(col = buffer), size = 2)
@@ -205,7 +205,7 @@
 #'
 #' # Create the focal versus comp data frame
 #' focal_vs_comp_ex <- growth_spatial_ex %>%
-#'   create_focal_vs_comp(max_dist, cv_grid_sf = cv_grid_sf_ex, id = "ID")
+#'   create_focal_vs_comp(comp_dist, cv_grid_sf = cv_grid_sf_ex, id = "ID")
 "growth_spatial_ex"
 
 
@@ -228,14 +228,14 @@
 #' library(ggplot2)
 #' library(sf)
 #'
-#' max_dist <- 1
+#' comp_dist <- 1
 #'
 #' ggplot(cv_grid_sf_ex) +
 #'   geom_sf() +
 #'   geom_sf(data = growth_spatial_ex)
 #'
 #' focal_vs_comp_ex <- growth_spatial_ex %>%
-#'   create_focal_vs_comp(max_dist, cv_grid_sf = cv_grid_sf_ex, id = "ID")
+#'   create_focal_vs_comp(comp_dist, cv_grid_sf = cv_grid_sf_ex, id = "ID")
 "cv_grid_sf_ex"
 
 

### 2021-01-25 R/model_functions.R

@@ -223,9 +223,9 @@ predict.comp_bayes_lm <- function(object, focal_vs_comp, ...) {
 #'
 #' @examples
 #' 1 + 1
-run_cv <- function(focal_vs_comp, max_dist, cv_grid, prior_param = NULL, run_shuffle = FALSE) {
+run_cv <- function(focal_vs_comp, comp_dist, cv_grid, prior_param = NULL, run_shuffle = FALSE) {
   check_focal_vs_comp(focal_vs_comp)
-  check_inherits(max_dist, "numeric")
+  check_inherits(comp_dist, "numeric")
   check_inherits(cv_grid, "sf")
   if (!is.null(prior_param)) {check_prior_params(prior_param)}
   check_inherits(run_shuffle, "logical")
@@ -240,14 +240,14 @@ run_cv <- function(focal_vs_comp, max_dist, cv_grid, prior_param = NULL, run_shu
     folds,
     fit_one_fold,
     focal_vs_comp,
-    max_dist,
+    comp_dist,
     cv_grid,
     prior_param,
     run_shuffle
   )
 }
 
-fit_one_fold <- function(fold, focal_vs_comp, max_dist,
+fit_one_fold <- function(fold, focal_vs_comp, comp_dist,
                          cv_grid, prior_param, run_shuffle) {
   # Define test set and "full" training set (we will remove buffer region below)
   test <- focal_vs_comp %>%
@@ -269,7 +269,7 @@ fit_one_fold <- function(fold, focal_vs_comp, max_dist,
   # Remove trees in training set that are part of test set and buffer region to test set
   train <- train_full %>%
     st_as_sf() %>%
-    add_buffer_variable(direction = "out", size = max_dist, region = test_fold_boundary) %>%
+    add_buffer_variable(direction = "out", size = comp_dist, region = test_fold_boundary) %>%
     filter(buffer) %>%
     as_tibble()
 

### 2021-01-25 data-raw/process_data_raw.R

@@ -136,7 +136,7 @@ blocks <- bind_rows(
 ) %>%
   mutate(foldID = c(1, 2))
 
-cv_grid_ex <- spatialBlock(
+SpatialBlock_ex <- spatialBlock(
   speciesData = growth_ex,
   k = 2,
   selection = "systematic",
@@ -146,19 +146,19 @@ cv_grid_ex <- spatialBlock(
 
 # Add foldID to data
 growth_spatial_ex <- growth_spatial_ex %>%
-  mutate(foldID = cv_grid_ex$foldID %>% as.factor())
+  mutate(foldID = SpatialBlock_ex$foldID %>% as.factor())
 use_data(growth_spatial_ex, overwrite = TRUE)
 
 
 ## Create spatial objects ----
-cv_grid_sf_ex <- cv_grid_ex$blocks %>%
+cv_grid_sf_ex <- SpatialBlock_ex$blocks %>%
   st_as_sf()
 use_data(cv_grid_sf_ex, overwrite = TRUE)
 
 
 ## Create focal_vs_comp data frame  ----
 focal_vs_comp_ex <- growth_spatial_ex %>%
-  create_focal_vs_comp(max_dist = 1, cv_grid_sf = cv_grid_sf_ex, id = "ID")
+  create_focal_vs_comp(comp_dist = 1, cv_grid_sf = cv_grid_sf_ex, id = "ID")
 use_data(focal_vs_comp_ex, overwrite = TRUE)
 
 

### 2021-01-25 etc/old_code/pkg_todo.R

@@ -62,11 +62,11 @@ bw_growth_df <-
 
 # Add buffers ---------------------------------------------------------------
 # This number acts as buffer size, but also determining neighbors
-max_dist <- 7.5
+comp_dist <- 7.5
 
 # Add buffer variable to data frame
 bw_growth_df <- bw_growth_df %>%
-  add_buffer_variable(direction = "in", size = max_dist, region = study_region_bw)
+  add_buffer_variable(direction = "in", size = comp_dist, region = study_region_bw)
 
 ggplot() +
   geom_sf(data = bw_growth_df, aes(col = buffer))
@@ -122,7 +122,7 @@ bw_cv_grid_sf <- bw_cv_grid$blocks %>%
 
 # Bayesian modeling ------------------------------------------------------------
 focal_vs_comp_bw <- bw_growth_df %>%
-  create_focal_vs_comp(max_dist = max_dist, cv_grid_sf = bw_cv_grid_sf, id = "treeID")
+  create_focal_vs_comp(comp_dist = comp_dist, cv_grid_sf = bw_cv_grid_sf, id = "treeID")
 
 # a) Fit model (compute posterior parameters) with no permutation shuffling
 posterior_param_bw <- focal_vs_comp_bw %>%
@@ -159,7 +159,7 @@ if(FALSE){
 focal_vs_comp_bw <- bw_growth_df %>%
   # mutate(sp = trait_group) %>%
   mutate(sp = family) %>%
-  create_focal_vs_comp(max_dist = max_dist, cv_grid_sf = bw_cv_grid_sf, id = "treeID")
+  create_focal_vs_comp(comp_dist = comp_dist, cv_grid_sf = bw_cv_grid_sf, id = "treeID")
 
 # a) Fit model (compute posterior parameters) with no permutation shuffling
 posterior_param_bw <- focal_vs_comp_bw %>%
@@ -183,7 +183,7 @@ str_c("results/", format(Sys.time(), "%Y-%m-%d"), "_posterior_lambda_family.png"
 # Cross-validation -------------------------------------------------------------
 tic()
 cv_bw <- focal_vs_comp_bw %>%
-  run_cv(max_dist = max_dist, cv_grid = bw_cv_grid) %>%
+  run_cv(comp_dist = comp_dist, cv_grid = bw_cv_grid) %>%
   right_join(bw_growth_df, by = c("focal_ID" = "treeID"))
[…]

### 2021-01-25 etc/old_code/spatial_block_issue.R

@@ -27,16 +27,16 @@ growth_ex <-
 data(study_region_ex)
 
 # set max dist
-max_dist <- 1
+comp_dist <- 1
 
 # add buffer
 growth_ex <- growth_ex %>%
-  add_buffer_variable(direction = "in", size = max_dist, region = study_region_ex)
+  add_buffer_variable(direction = "in", size = comp_dist, region = study_region_ex)
 
 ggplot() +
   geom_sf(data = growth_ex, aes(col = buffer), size = 2)
 
-#cv_grid_ex <- spatialBlock(
+#SpatialBlock_ex <- spatialBlock(
 #  speciesData = growth_ex, theRange = 5, verbose = FALSE, k = 2
 #)
 
@@ -56,7 +56,7 @@ ggplot() +
   geom_sf(data = blocks, fill = "transparent")
 
 # fit spatialBlock()
-cv_grid_ex <- spatialBlock(
+SpatialBlock_ex <- spatialBlock(
   speciesData = growth_ex,
   verbose = FALSE,
   k = 2,
@@ -68,10 +68,10 @@ cv_grid_ex <- spatialBlock(
 
 # Add foldID to data
 growth_ex <- growth_ex %>%
-  mutate(foldID = cv_grid_ex$foldID %>% as.factor())
+  mutate(foldID = SpatialBlock_ex$foldID %>% as.factor())
 
 # Visualize grid
-cv_grid_ex$plots +
+SpatialBlock_ex$plots +
   geom_sf(data = growth_ex, aes(col = foldID), size = 2)
 
 # Deliverable

### 2021-01-25 tests/testthat/test-readme.R

@@ -24,10 +24,10 @@ test_that("readme code works", {
       sp = as.factor(sp)
     )
 
-  max_dist <- 1
+  comp_dist <- 1
 
   growth_ex <- growth_ex %>%
-    add_buffer_variable(direction = "in", size = max_dist, region = study_region_ex)
+    add_buffer_variable(direction = "in", size = comp_dist, region = study_region_ex)
 
   expect_true(check_inherits(growth_ex, "data.frame"))
 
@@ -39,7 +39,7 @@ test_that("readme code works", {
   ) %>%
     mutate(foldID = c(1, 2))
 
-  cv_grid_ex <- spatialBlock(
+  SpatialBlock_ex <- spatialBlock(
     speciesData = growth_ex,
     verbose = FALSE,
     k = 2,
@@ -49,13 +49,13 @@ test_that("readme code works", {
   )
 
   growth_ex <- growth_ex %>%
-    mutate(foldID = cv_grid_ex$foldID %>% as.factor())
+    mutate(foldID = SpatialBlock_ex$foldID %>% as.factor())
 
-  cv_grid_sf_ex <- cv_grid_ex$blocks %>%
+  cv_grid_sf_ex <- SpatialBlock_ex$blocks %>%
     st_as_sf()
 
   focal_vs_comp_ex <- growth_ex %>%
-    create_focal_vs_comp(max_dist, cv_grid_sf = cv_grid_sf_ex, id = "ID")
+    create_focal_vs_comp(comp_dist, cv_grid_sf = cv_grid_sf_ex, id = "ID")
 
   # Checks each column in focal_vs_comp is of appropriate type
   expect_true(check_inherits(focal_vs_comp_ex, "data.frame"))
@@ -91,7 +91,7 @@ test_that("readme code works", {
   expect_true(
     check_inherits(
       focal_vs_comp_ex %>%
-        run_cv(max_dist = max_dist, cv_grid = cv_grid_sf_ex) %>%
+        run_cv(comp_dist = comp_dist, cv_grid = cv_grid_sf_ex) %>%
         right_join(growth_ex, by = c("focal_ID" = "ID")),
     "data.frame"
     )

### 2021-01-25 README.Rmd
edited fit_one_fold() output to be complete focal_vs_comp data frame itself, not just y_hat

@@
 ```{r}
-ex_bw <- focal_vs_comp_ex %>%
-  run_cv(comp_dist = comp_dist, cv_grid = cv_grid_sf_ex) %>%
-  right_join(growth_ex, by = c("focal_ID" = "ID"))
+focal_vs_comp_ex <- focal_vs_comp_ex %>%
+  run_cv(comp_dist = comp_dist, cv_grid = cv_grid_sf_ex)
 ```
@@
 ```{r}
-ex_bw %>%
-  rmse(truth = growth, estimate = growth_hat) %>%
+focal_vs_comp_ex %>%
+  yardstick::rmse(truth = growth, estimate = growth_hat) %>%
   pull(.estimate)

### 2021-01-25 R/model_functions.R

@@ -273,14 +273,13 @@ fit_one_fold <- function(fold, focal_vs_comp, comp_dist,
     filter(buffer) %>%
     as_tibble()
 
-  # Fit model on training data and predict on test
+  # Fit model on training data
   comp_bayes_lm_fold <- train %>%
     comp_bayes_lm(prior_param = prior_param, run_shuffle = run_shuffle)
 
-  comp_bayes_lm_fold %>%
-    predict(focal_vs_comp = test) %>%
-    enframe(name = NULL, value = "growth_hat") %>%
-    mutate(focal_ID = test$focal_ID)
+  # Compute predicted values and append to test
+  test %>%
+    mutate(growth_hat = predict(comp_bayes_lm_fold, test))
 }
 
 

### 2021-01-25 README.Rmd
Rebuilt README

@@
     ) %>%
-  mutate(sp = to_any_case(sp) %>% factor())
+  mutate(sp = snakecase::to_any_case(sp) %>% factor())
 ```
@@
 
-Next we add information pertaining to our spatial cross-validation scheme.
-We first manually define the spatial blocks that will act as our cross-validation folds and convert them to an `sf` object using `sf_polygon()` function from the `sfheaders` package.
+Next we add information pertaining to our spatial cross-validation scheme. We first manually define the spatial blocks that will act as our cross-validation folds and convert them to an `sf` object.
 
@@
 
-blocks <- bind_rows(
-  sf_polygon(fold1),
-  sf_polygon(fold2)
+blocks_ex <- bind_rows(
+  sfheaders::sf_polygon(fold1),
+  sfheaders::sf_polygon(fold2)
 ) %>%
-  mutate(foldID = c(1, 2))
+  mutate(folds = c(1, 2) %>% factor())
 ```
@@
 ```{r}
-SpatialBlock_ex <- spatialBlock(
-  speciesData = growth_ex, k = 2, selection = "systematic", blocks = blocks,
+SpatialBlock_ex <- blockCV::spatialBlock(
+  speciesData = growth_ex, k = 2, selection = "systematic", blocks = blocks_ex,
   showBlocks = FALSE, verbose = FALSE
@@
   mutate(foldID = SpatialBlock_ex$foldID %>% factor())
-```
-
-We also extract the spatial cross-validation grid itself as an `sf` object for later use.
-
-```{r}
-cv_grid_sf_ex <- SpatialBlock_ex$blocks %>%
-  st_as_sf()
-ggplot() +
-  geom_sf(data = cv_grid_sf_ex)
 ```
@@
   geom_sf(data = growth_ex, aes(col = buffer, shape = foldID), size = 2) +
-  geom_sf(data = blocks, fill = "transparent", col = "orange")
+  geom_sf(data = blocks_ex, fill = "transparent", col = "orange")
 ```
@@
 focal_vs_comp_ex <- growth_ex %>%
-  create_focal_vs_comp(comp_dist, cv_grid_sf = cv_grid_sf_ex, id = "ID")
+  create_focal_vs_comp(comp_dist, cv_grid_sf = blocks_ex, id = "ID")
 focal_vs_comp_ex
@@
 focal_vs_comp_ex %>% 
-  unnest(cols = "comp")
+  tidyr::unnest(cols = "comp")
 ```

### 2021-01-25 README.Rmd
Added opening example

@@
 ### Compute growth of trees based on census data
+
+The starting point of our analysis are data from two repeat censuses `census_1_ex` and `census_2_ex`. For example consider the forest census data in `census_1_ex`
+
+```{r}
+census_1_ex
+```
+
+We convert the `census_1_ex` data frame to an object of type `sf` and then plot using `geom_sf()`.
+
+```{r}
+ggplot() +
+  geom_sf(
+    data = census_1_ex %>% sf::st_as_sf(coords = c("gx", "gy")),
+    aes(col = sp, size = dbh)
+  )
+```
 
@@
 
-# Visualizations (combined with patchwork pkg)
+# Posterior distributions (plots combined with patchwork pkg)
 p1 <- autoplot(comp_bayes_lm_ex, type = "intercepts")

### 2021-01-26 README.Rmd
small readme edits
mostly spellcheck, add library(tidyr)

@@
 The package has commands to fit models of tree growth based on neighborhood competition which can be used to estimate species-specific competition coefficients.
-The models fits can then be evaluated using a spatial cross-validation scheme to detect possible overfitting.
-Additionally these models can test whether the species identity of competitors matters using a permutation test-style shuffling of competitor identity (under the null hypothesis) and subsequently evaluating if model performance changes.
+The model fits can then be evaluated using a spatial cross-validation scheme to detect possible overfitting.
+Additionally, these models can test whether the species identity of competitors matters using a permutation test-style shuffling of competitor identity (under the null hypothesis) and subsequently evaluating if model performance changes.
 See Allen and Kim (2020) [A permutation test and spatial cross-validation approach to assess models of interspecific competition between trees](https://doi.org/10.1371/journal.pone.0229930) for a full description.
@@
 library(patchwork)
+library(tidyr)
 ```
@@
 
-The starting point of our analysis are data from two repeat censuses `census_1_ex` and `census_2_ex`. For example consider the forest census data in `census_1_ex`
+The starting point of our analysis are data from two repeat censuses `census_1_ex` and `census_2_ex`. For example, consider the forest census data in `census_1_ex`.
 
@@
 
-In the visualization below, the spatial blocks that act as our cross-validation folds are delineated in orange.
-Correspondingly the shape of each point indicates which fold each tree has been assigned to.
+In the visualization below, the spatial blocks that act as our cross-validation folds are delineated in orange. The shape of each point indicates which fold each tree has been assigned to.
 
@@
 Based on our `growth` data frame, we now explicitly define all "focal" trees and their respective "competitor" trees in a `focal_vs_comp` data frame.
-This data frame has rows corresponding to each focal tree and all information about its competitors are saved in list-column variable `comp`.
-We implemented this nested format using `tidyr::nest()` in order to minimize redundancy, given that the same tree can act as a competitor multiple times.
+This data frame has rows corresponding to each focal tree, and all information about its competitors are saved in the list-column variable `comp`.
+We implemented this nested format using `nest()` in order to minimize redundancy, given that the same tree can act as a competitor multiple times.
 
@@
 Using `unnest()` we can fully expand the competitor information saved in the `focal_vs_comp` data frame.
-For example the tree with `focal_ID` equal to 2 located at (1.5, 2.5) has two competitors within `comp_dist` distance from it.
+For example, the tree with `focal_ID` equal to 2 located at (1.5, 2.5) has two competitors within `comp_dist` distance from it.
 
@@
 focal_vs_comp_ex %>% 
-  tidyr::unnest(cols = "comp")
+  unnest(cols = "comp")
 ```
@@
 
-Since the resulting output is an S3 object of class `comp_bayes_lm` we defined, we can both print and visualize its output: the posterior distribution of all linear regression parameters: the intercept, the slope for dbh for each species, and a matrix of all species pairs competitive effects on growth.
+The resulting output is an `comp_bayes_lm` object containing the posterior distribution of all linear regression parameters, the intercept, the slope for dbh for each species, and a matrix of all species pairs competitive effects on growth.
+The S3 object class is associated with several methods.
 
@@
 
-Furthermore, we can apply a generic `predict()` function to the resulting `comp_bayes_lm` object to obtain fitted/predicted values of this model.
+Furthermore, we can apply a `predict()` method to the resulting `comp_bayes_lm` object to obtain fitted/predicted values of this model.
 We append these `growth_hat` values to our `focal_vs_comp` data frame.
@@
 ```
+
 We then compute the root mean squared error (RMSE) of the observed versus fitted growths as a measure of our model's fit.
@@
 
-Whereas in our example above we fit our model to the entirety of the data and then generated fitted/predicted growths on this same data, we now apply the same model with spatial cross-validation.
-All the trees in a given fold will be given a turn as the "test" data while the trees is all remaining folds will be the "training" data.
+Whereas in our example above we fit our model to the entirety of the data and then generate fitted/predicted growths on this same data, we now apply the same model with spatial cross-validation.
+All the trees in a given fold will be given a turn as the "test" data while the trees in all remaining folds will be the "training" data.
 We then fit the model to the training data, but compute fitted/predicted growths for the separate and independent data.

### 2021-01-29 README.Rmd
incorporated simon's suggestions

@@
 
-We present an example analysis using toy data pre-loaded into the package.
+We present an example analysis using toy data pre-loaded into the package where we will
+
+1. Compute growth of trees based on census data
+1. Add spatial information
+1. Compute focal versus competitor tree information
+1. Fit model and make predictions
+1. Run spatial cross-validation
 
@@
 library(forestecology)
-library(sf)
-library(sfheaders)
 library(blockCV)
-library(yardstick)
-library(snakecase)
 library(patchwork)
-library(tidyr)
 ```
@@
     ) %>%
-  mutate(sp = snakecase::to_any_case(sp) %>% factor())
+  mutate(sp = to_any_case(sp) %>% factor())
 ```
@@
 
-Next we add information pertaining to our spatial cross-validation scheme. We first manually define the spatial blocks that will act as our cross-validation folds and convert them to an `sf` object.
+Next we add information pertaining to our spatial cross-validation scheme.
+We first manually define the spatial blocks that will act as our cross-validation folds and convert them to an `sf` object using the `sf_polygon()` function from the `sfheaders` package.
 
@@
 blocks_ex <- bind_rows(
-  sfheaders::sf_polygon(fold1),
-  sfheaders::sf_polygon(fold2)
+  sf_polygon(fold1),
+  sf_polygon(fold2)
 ) %>%
@@
 focal_vs_comp_ex %>%
-  yardstick::rmse(truth = growth, estimate = growth_hat) %>%
+  rmse(truth = growth, estimate = growth_hat) %>%
   pull(.estimate)

### 2021-01-29 R/model_functions.R

@@ -16,7 +16,7 @@
 #' @importFrom stats model.matrix
 #' @importFrom tidyr unnest
 #' @importFrom tidyr spread
-#' @source Closed-form solutions of Bayesian linear regression \url{https://doi.org/10.1371/journal.pone.0229930.s004}
+#' @source Closed-form solutions of Bayesian linear regression \url{https://doi.org/10.1371/journal.pone.0229930.s004)} (Source: \url{https://doi.org/10.1371/journal.pone.0229930.s004})
 #' @return A list of `{a_star, b_star, mu_star, V_star}` posterior hyperparameters
 #' @seealso \code{\link{predict.comp_bayes_lm}}
 #' @export
@@ -123,8 +123,7 @@ default_prior_params <- function(X) {
 print.comp_bayes_lm <- function(x, ...) {
   cat(
     paste0(
-      "Parameters of a Bayesian linear regression model with a multivariate Normal likelihood\n",
-      "(for details see https://doi.org/10.1371/journal.pone.0229930.s004):\n\n"
+      "Bayesian linear regression model parameters with a multivariate Normal likelihood. See ?comp_bayes_lm for details:\n\n"
     )
   )
 
@@ -144,9 +143,7 @@ print.comp_bayes_lm <- function(x, ...) {
     paste0(
       "\n",
       "Model formula:\n",
-      x[[3]] %>% as.character() %>% .[2],
-      " ~ ",
-      x[[3]] %>% as.character() %>% .[3],
+      x[[3]] %>% as.character(),
       "\n"
     )
   )

### 2021-01-29 R/re_exported_functions.R

@@ -1,0 +1,11 @@
+##' @importFrom snakecase to_any_case
+##' @export
+snakecase::to_any_case
+
+##' @importFrom yardstick rmse
+##' @export
+yardstick::rmse
+
+##' @importFrom sfheaders sf_polygon
+##' @export
+sfheaders::sf_polygon

### 2021-02-01 README.Rmd

@@
 )
+set.seed(76)
 ```
@@
 
-We present an example analysis using toy data pre-loaded into the package where we will
+We present an example analysis using toy data pre-loaded into the package where we will:
 

### 2021-02-01 README.Rmd
Fixed all testthat errors related to factors

@@
   compute_growth(
-    census_1 = census_1_ex,
-    census_2 = census_2_ex %>% filter(!str_detect(codes, "R")),
+    census_1 = census_1_ex %>% 
+      mutate(sp = to_any_case(sp) %>% factor()),
+    census_2 = census_2_ex %>% 
+      filter(!str_detect(codes, "R")) %>% 
+      mutate(sp = to_any_case(sp) %>% factor()),
     id = "ID"
-    ) %>%
-  mutate(sp = to_any_case(sp) %>% factor())
+  )
 ```

### 2021-02-01 R/data_processing_functions.R

@@ -23,14 +23,15 @@
 #' library(dplyr)
 #' library(stringr)
 #'
-#' # Load in data from two forest censuses
-#' data(census_1_ex, census_2_ex)
-#'
-#' # Filter out resprouts in second census
-#' census_2_ex_no_r <- census_2_ex %>%
-#'   filter(!str_detect(codes, "R"))
-#'
-#' growth_ex <- compute_growth(census_1_ex, census_2_ex_no_r, id = "ID")
+#' growth_ex <-
+#'   compute_growth(
+#'     census_1 = census_1_ex %>%
+#'       mutate(sp = to_any_case(sp) %>% factor()),
+#'     census_2 = census_2_ex %>%
+#'       filter(!str_detect(codes, "R")) %>%
+#'       mutate(sp = to_any_case(sp) %>% factor()),
+#'     id = "ID"
+#'   )
 compute_growth <- function(census_1, census_2, id) {
   # 0. Check inputs
   check_inherits(census_1, "data.frame")
@@ -44,7 +45,7 @@ compute_growth <- function(census_1, census_2, id) {
     census_1
   )
   purrr::map2(
-    c("ID", "sp", "gx", "gy", "date", "codes", "dbh"),
+    c(id, "sp", "gx", "gy", "date", "codes", "dbh"),
     c("numeric", "factor", "numeric", "numeric", "Date", "character", "numeric"),
     check_column,
     census_2

### 2021-02-01 R/example_datasets.R

@@ -20,20 +20,17 @@
 #' @family example data objects
 #'
 #' @examples
-#' data(census_1_ex, census_2_ex)
 #' library(dplyr)
 #' library(stringr)
-#' library(snakecase)
-#' # Filter out resprouts
-#' census_2_ex_no_r <- census_2_ex %>%
-#'   filter(!str_detect(codes, "R"))
-#' id <- "ID"
+#'
 #' growth_ex <-
-#'   # Merge both censuses and compute growth:
-#'   compute_growth(census_1_ex, census_2_ex_no_r, id) %>%
-#'   mutate(
-#'     sp = to_any_case(sp),
-#'     sp = as.factor(sp)
+#'   compute_growth(
+#'     census_1 = census_1_ex %>%
+#'       mutate(sp = to_any_case(sp) %>% factor()),
+#'     census_2 = census_2_ex %>%
+#'       filter(!str_detect(codes, "R")) %>%
+#'       mutate(sp = to_any_case(sp) %>% factor()),
+#'     id = "ID"
 #'   )
 "census_1_ex"
 
@@ -62,20 +59,17 @@
 #' @family example data objects
 #'
 #' @examples
-#' data(census_1_ex, census_2_ex)
 #' library(dplyr)
 #' library(stringr)
-#' library(snakecase)
-#' # Filter out resprouts
-#' census_2_ex_no_r <- census_2_ex %>%
-#'   filter(!str_detect(codes, "R"))
-#' id <- "ID"
+#'
 #' growth_ex <-
-#'   # Merge both censuses and compute growth:
-#'   compute_growth(census_1_ex, census_2_ex_no_r, id) %>%
-#'   mutate(
-#'     sp = to_any_case(sp),
-#'     sp = as.factor(sp)
+#'   compute_growth(
+#'     census_1 = census_1_ex %>%
+#'       mutate(sp = to_any_case(sp) %>% factor()),
+#'     census_2 = census_2_ex %>%
+#'       filter(!str_detect(codes, "R")) %>%
+#'       mutate(sp = to_any_case(sp) %>% factor()),
+#'     id = "ID"
 #'   )
 "census_2_ex"
 

### 2021-02-01 R/model_functions.R

@@ -359,6 +359,7 @@ fit_one_fold <- function(fold, focal_vs_comp, comp_dist,
 #' @inheritParams comp_bayes_lm
 #'
 #' @importFrom tidyr unnest
+#' @export
 #'
 #' @return Data frame for internal package use.
 #'

### 2021-02-01 data-raw/process_data_raw.R

@@ -179,12 +179,13 @@ growth_toy <- tibble(
   gy = c(4, 3, 2, 1, 2)
 ) %>%
   mutate(
+    sp = factor(sp),
     dbh1 = c(40, 25, 30, 35, 20),
     codes1 = rep("M", n()),
     codes2 = rep("M", n()),
     growth = c(1, 2, 1, 3, 2),
     dbh2 = dbh1 + growth,
-    foldID = rep(1, n()) %>% as.character(),
+    foldID = rep(1, n()) %>% as.factor(),
     buffer = rep(FALSE, n())
   ) %>%
   # Convert data frame to sf object

### 2021-02-01 tests/testthat/test-readme.R

@@ -12,11 +12,13 @@ test_that("readme code works", {
 
   growth_ex <-
     compute_growth(
-      census_1 = census_1_ex,
-      census_2 = census_2_ex %>% filter(!str_detect(codes, "R")),
+      census_1 = census_1_ex %>%
+        mutate(sp = to_any_case(sp) %>% factor()),
+      census_2 = census_2_ex %>%
+        filter(!str_detect(codes, "R")) %>%
+        mutate(sp = to_any_case(sp) %>% factor()),
       id = "ID"
     ) %>%
-    mutate(sp = to_any_case(sp) %>% factor()) %>%
     add_buffer_variable(direction = "in", size = comp_dist, region = study_region_ex)
 
   expect_true(check_inherits(growth_ex, "data.frame"))

### 2021-02-02 etc/paper/forestecology.Rmd
incorporated simon's suggestions

@@
 ```{r load-packages}
-library(forestecology)
-
-# Load tidyverse packages:
 library(tidyverse)
 library(lubridate)
-
-# Load spatial packages:
+library(sf)
+library(forestecology)
 library(blockCV)
-library(sf)
-library(sfheaders)
-
-# Load other packages:
-library(snakecase)
-library(yardstick)
 ```
@@
 ```{r bw-load-data}
-data(census_2008_bw, census_2014_bw, species_bw)
-
-# Append additional species data
 census_2008_bw <- census_2008_bw %>%
@@
 ```{r scbi-load-data}
-scbi_2013 <- read_csv("scbi.stem2.csv") %>%
+census_2013_scbi <- read_csv("scbi.stem2.csv") %>%
   select(treeID, stemID, sp, ExactDate, gx, gy, dbh, codes, status) %>%
   mutate(
-    date = ExactDate,
-    dbh = as.numeric(dbh),
-    date = mdy(date)
+    date = ExactDate %>% mdy(),
+    dbh = as.numeric(dbh)
   ) %>%
@@
 
-scbi_2018 <- read_csv("scbi.stem3.csv") %>%
+census_2018_scbi <- read_csv("scbi.stem3.csv") %>%
   select(treeID, stemID, sp, ExactDate, gx, gy, dbh, codes, status) %>%
   mutate(
-    date = ExactDate,
-    dbh = as.numeric(dbh),
-    date = mdy(date)
+    date = ExactDate %>% mdy(),
+    dbh = as.numeric(dbh)
   ) %>%
@@
 ```{r bw-compute-growth}
-census_2014_bw <- census_2014_bw %>%
-  filter(!str_detect(codes, "R"))
-
-bw_growth_df <-
-  compute_growth(census_2008_bw, census_2014_bw, id = "treeID") %>%
-  # Convert all variables denoting species to factors
-  mutate(
-    sp = sp %>% to_any_case() %>% as.factor(),
-    species = sp,
-    family = as.factor(family),
-    trait_group = as.factor(trait_group)
-  ) %>%
-  # Drop unnecessary variables
-  select(-stemID)
+growth_bw <-
+  compute_growth(
+    census_1 = census_2008_bw %>% 
+      mutate(sp = to_any_case(sp) %>% factor()),
+    census_2 = census_2014_bw %>% 
+      filter(!str_detect(codes, "R")) %>% 
+      mutate(sp = to_any_case(sp) %>% factor()),
+    id = "treeID"
+  )
 ```
@@
 ```{r scbi-compute-growth}
-scbi_growth_df <-
-  compute_growth(scbi_2013, scbi_2018, "stemID") %>%
-  # Convert all variables denoting species to factors
-  mutate(sp = as.factor(sp))
+growth_scbi <-
+  compute_growth(
+    census_1 = census_2013_scbi %>% 
+      mutate(sp = to_any_case(sp) %>% factor()),
+    census_2 = census_2018_scbi %>% 
+      filter(!str_detect(codes, "R")) %>% 
+      mutate(sp = to_any_case(sp) %>% factor()),
+    id = "stemID"
+  )
 ```
@@
 bind_rows(
-  bw_growth_df %>% st_drop_geometry() %>% select(growth) %>% mutate(Site = "Big Woods"),
-  scbi_growth_df %>% st_drop_geometry() %>% select(growth) %>% mutate(Site = "SCBI")
+  growth_bw %>% st_drop_geometry() %>% select(growth) %>% mutate(Site = "Big Woods"),
+  growth_scbi %>% st_drop_geometry() %>% select(growth) %>% mutate(Site = "SCBI")
 ) %>% 
@@
 
-bw_growth_df <- bw_growth_df %>%
-  add_buffer_variable(direction = "in", size = comp_dist, 
-                      region = study_region_bw)
+growth_bw <- growth_bw %>%
+  add_buffer_variable(direction = "in", size = comp_dist, region = study_region_bw)
 
 ggplot() +
-  geom_sf(data = bw_growth_df %>% sample_frac(0.2), 
-          aes(col = buffer), size = 0.5)
+  geom_sf(data = growth_bw %>% sample_frac(0.2), aes(col = buffer), size = 0.5)
 ```
@@
 bw_spatialBlock <- spatialBlock(
-  speciesData = bw_growth_df, theRange = cv_fold_size, k = 28, xOffset = 0.5, 
+  speciesData = growth_bw, theRange = cv_fold_size, k = 28, xOffset = 0.5, 
   yOffset = 0, verbose = FALSE, showBlocks = FALSE
@@
 ```{r}
-bw_growth_df <- bw_growth_df %>%
+growth_bw <- growth_bw %>%
   mutate(foldID = bw_spatialBlock$foldID)
@@
   geom_sf(data = bw_spatialBlock$blocks %>% st_as_sf()) + 
-  geom_sf(data = bw_growth_df %>% sample_frac(0.2), 
+  geom_sf(data = growth_bw %>% sample_frac(0.2), 
           aes(col = factor(foldID)), size = 0.1, show.legend = FALSE) + 
@@
 ```{r}
-bw_growth_df <- bw_growth_df %>%
+growth_bw <- growth_bw %>%
   filter(!foldID %in% c(19, 23, 21, 17, 8, 19)) %>%
-  mutate(foldID = as.character(foldID))
+  mutate(foldID = factor(foldID))
 ```
 
-Separately, we save the spatial cross-validation grid as an `sf_polygon` object `bw_cv_grid`
+Separately, we save the spatial cross-validation grid as an `sf_polygon` object `blocks_bw`
 
 ```{r}
-bw_cv_grid <- bw_spatialBlock$blocks %>%
+blocks_bw <- bw_spatialBlock$blocks %>%
   st_as_sf()
@@
 ```{r scbi-define-buffer, out.width="50%", fig.cap="Buffer region for SCBI study region."}
-scbi_study_region <- tibble(
+study_region_scbi <- tibble(
   x = c(0, 300, 300, 0, 0),
@@
 
-scbi_growth_df <- scbi_growth_df %>%
-  add_buffer_variable(direction = "in", size = comp_dist, 
-                      region = scbi_study_region)
+growth_scbi <- growth_scbi %>%
+  add_buffer_variable(direction = "in", size = comp_dist, region = study_region_scbi)
 
 ggplot() +
-  geom_sf(data = scbi_growth_df, aes(col = buffer), size = 0.5)
+  geom_sf(data = growth_scbi, aes(col = buffer), size = 0.5)
 ```
@@
 ```{r}
-fold1 <- rbind(c(0, 300), c(150, 300), c(150, 600), c(0, 600)) %>%
-  sf_polygon() %>% 
-  mutate(folds = 1)
-fold2 <- rbind(c(150, 300), c(300, 300), c(300, 600), c(150, 600)) %>%
-  sf_polygon() %>% 
-  mutate(folds = 2)
-scbi_cv_grid <- bind_rows(fold1, fold2)
+fold1 <- rbind(c(0, 300), c(150, 300), c(150, 600), c(0, 600))
+fold2 <- rbind(c(150, 300), c(300, 300), c(300, 600), c(150, 600))
+
+blocks_scbi <- bind_rows(
+  sf_polygon(fold1),
+  sf_polygon(fold2)
+) %>%
+  mutate(folds = c(1, 2) %>% factor())
 ```
@@
 ```{r scbi-define-cv-folds, out.width="50%", fig.cap="Buffer region for SCBI study region."}
-scbi_spatialBlock <- spatialBlock(
-  speciesData = scbi_growth_df, k = 2, verbose = FALSE, showBlocks = FALSE,
-  # Note new arguments:
-  selection = "systematic", blocks = scbi_cv_grid
+SpatialBlock_scbi <- spatialBlock(
+  speciesData = growth_scbi, k = 2, selection = "systematic", blocks = blocks_scbi,
+  showBlocks = FALSE, verbose = FALSE
 )
@@
 # Add foldID to each tree
-scbi_growth_df <- scbi_growth_df %>%
-  mutate(foldID = scbi_spatialBlock$foldID)
+growth_scbi <- growth_scbi %>%
+  mutate(foldID = SpatialBlock_scbi$foldID %>% factor())
 
 ggplot() +
-  geom_sf_text(data = scbi_growth_df, 
-               aes(label = foldID, col = buffer)) +
-  geom_sf(data = scbi_cv_grid, fill = "transparent")
+  geom_sf_text(data = growth_scbi, aes(label = foldID, col = buffer)) +
+  geom_sf(data = blocks_scbi, fill = "transparent")
 ```
@@
 ```{r bw-focal-vs-comp}
-focal_vs_comp_bw <- bw_growth_df %>%
-  create_focal_vs_comp(comp_dist, cv_grid_sf = bw_cv_grid, id = "treeID")
+focal_vs_comp_bw <- growth_bw %>%
+  create_focal_vs_comp(comp_dist, cv_grid_sf = blocks_bw, id = "treeID")
 ```
@@
 ```{r scbi-focal-vs-comp}
-focal_vs_comp_scbi <- scbi_growth_df %>%
-  create_focal_vs_comp(comp_dist, cv_grid_sf = scbi_cv_grid, id = "stemID")
+focal_vs_comp_scbi <- growth_scbi %>%
+  create_focal_vs_comp(comp_dist, cv_grid_sf = blocks_scbi, id = "stemID")
 ```
@@
 ```{r bw-model-fit}
-posterior_param_bw <- focal_vs_comp_bw %>%
-  fit_bayesian_model(prior_param = NULL)
+comp_bayes_lm_bw <- focal_vs_comp_bw %>%
+  comp_bayes_lm(prior_param = NULL)
 ```
@@
 ```{r bw-model-predict}
-predictions <- focal_vs_comp_bw %>%
-  predict_bayesian_model(posterior_param = posterior_param_bw) %>%
-  right_join(bw_growth_df, by = c("focal_ID" = "treeID"))
+focal_vs_comp_bw <- focal_vs_comp_bw %>%
+  mutate(growth_hat = predict(comp_bayes_lm_bw, focal_vs_comp_bw))
 ```
@@
 ```{r bw-model-rmse}
-predictions %>%
-  yardstick::rmse(truth = growth, estimate = growth_hat) %>%
+focal_vs_comp_bw %>%
+  rmse(truth = growth, estimate = growth_hat) %>%
   pull(.estimate)
@@
 ```{r bw-model-shuffle-fit}
-posterior_param_bw_shuffle <- focal_vs_comp_bw %>%
-  fit_bayesian_model(prior_param = NULL, run_shuffle = TRUE)
+comp_bayes_lm_bw_shuffle <- focal_vs_comp_bw %>%
+  comp_bayes_lm(prior_param = NULL, run_shuffle = TRUE)
 ```
@@
 ```{r bw-model-shuffle-predict}
-predictions_shuffle <- focal_vs_comp_bw %>%
-  predict_bayesian_model(posterior_param = posterior_param_bw_shuffle) %>%
-  right_join(bw_growth_df, by = c("focal_ID" = "treeID"))
+focal_vs_comp_bw <- focal_vs_comp_bw %>%
+  mutate(growth_hat_shuffle = predict(comp_bayes_lm_bw_shuffle, focal_vs_comp_bw))
 ```
@@
 ```{r bw-model-shuffle-rmse}
-predictions_shuffle %>%
-  rmse(truth = growth, estimate = growth_hat) %>%
+focal_vs_comp_bw %>%
+  rmse(truth = growth, estimate = growth_hat_shuffle) %>%
   pull(.estimate)
@@
 ```{r scbi-model-fit}
-posterior_param_scbi <- focal_vs_comp_scbi %>%
-  fit_bayesian_model(prior_param = NULL, run_shuffle = FALSE)
+comp_bayes_lm_scbi <- focal_vs_comp_scbi %>%
+  comp_bayes_lm(prior_param = NULL)
 ```
@@
 ```{r scbi-model-predict}
-scbi_growth_df_noCV <- focal_vs_comp_scbi %>%
-  predict_bayesian_model(posterior_param = posterior_param_scbi) %>%
-  right_join(scbi_growth_df, by = c("focal_ID" = "stemID"))
+focal_vs_comp_scbi <- focal_vs_comp_scbi %>%
+  mutate(growth_hat = predict(comp_bayes_lm_scbi, focal_vs_comp_scbi))
 ```
@@
 ```{r scbi-model-rmse}
-scbi_growth_df_noCV %>%
+focal_vs_comp_scbi %>%
   rmse(truth = growth, estimate = growth_hat) %>%
@@
 ```{r scbi-model-residuals, out.width="100%", fig.cap="Predicted versus observed growth.", echo = FALSE}
-ggplot(scbi_growth_df_noCV, aes(x = growth, y = growth_hat)) +
+ggplot(focal_vs_comp_scbi, aes(x = growth, y = growth_hat)) +
   geom_point(size = 0.5, color = rgb(0, 0, 0, 0.25)) +
@@
 ```{r scbi-model-residuals-2, out.width="80%", fig.cap="Spatial distribution of residuals for model applied to SCBI data.", echo = FALSE}
-scbi_growth_df_noCV %>%
+focal_vs_comp_scbi %>%
   st_as_sf() %>%
@@
 ```{r bw-spatial-cv}
-cv_bw <- focal_vs_comp_bw %>%
-  run_cv(comp_dist = comp_dist, cv_grid = bw_cv_grid) %>%
-  right_join(bw_growth_df, by = c("focal_ID" = "treeID"))
+focal_vs_comp_bw <- focal_vs_comp_bw %>%
+  run_cv(comp_dist = comp_dist, cv_grid = blocks_bw)
 
-cv_bw %>%
+focal_vs_comp_bw %>%
   rmse(truth = growth, estimate = growth_hat) %>%
@@
 ```{r scbi-spatial-cv}
-cv_scbi <- focal_vs_comp_scbi %>%
-  run_cv(comp_dist = comp_dist, cv_grid = scbi_cv_grid) %>%
-  right_join(scbi_growth_df, by = c("focal_ID" = "treeID"))
+focal_vs_comp_scbi <- focal_vs_comp_scbi %>%
+  run_cv(comp_dist = comp_dist, cv_grid = blocks_scbi)
 
-cv_scbi %>%
+focal_vs_comp_scbi %>%
   rmse(truth = growth, estimate = growth_hat) %>%
@@
 ```{r bw-posterior-viz-0}
-focal_vs_comp_bw <- bw_growth_df %>%
-  mutate(sp = family) %>%
-  create_focal_vs_comp(comp_dist = comp_dist, cv_grid_sf = bw_cv_grid, 
-                       id = "treeID")
+focal_vs_comp_bw <- growth_bw %>%
+  mutate(sp = family %>% factor()) %>%
+  create_focal_vs_comp(comp_dist = comp_dist, cv_grid_sf = blocks_bw, id = "treeID")
 
-posterior_param_bw <- focal_vs_comp_bw %>%
-  fit_bayesian_model(prior_param = NULL, run_shuffle = FALSE)
+comp_bayes_lm_bw <- focal_vs_comp_bw %>%
+  comp_bayes_lm(prior_param = NULL)
 ```
@@
 ```{r bw-posterior-viz-1}
-posterior_plots_bw <- plot_bayesian_model_parameters(
-  posterior_param = posterior_param_bw,
-  sp_to_plot = c("cornaceae", "fagaceae", "hamamelidaceae", "juglandaceae", 
+sp_to_plot <- c("cornaceae", "fagaceae", "hamamelidaceae", "juglandaceae", 
                  "lauraceae", "rosaceae", "sapindaceae", "ulmaceae")
-)
 ```
@@
 ```{r bw-posterior-viz-beta0, out.width="100%", fig.cap="Posterior distribution of beta0."}
-posterior_plots_bw[["beta_0"]]
+plot1 <- autoplot(comp_bayes_lm_bw, type = "intercepts")
+plot1
 ```
@@
 ```{r bw-posterior-viz-beta-dbh, out.width="100%", fig.cap="Posterior distribution of betadbh."}
-posterior_plots_bw[["beta_dbh"]]
+plot2 <- autoplot(comp_bayes_lm_bw, type = "dbh_slopes")
+plot2
 ```
@@
 ```{r bw-posterior-viz-lambda, out.width="100%", fig.cap="Posterior distribution of lambda's for Big Woods."}
-posterior_plots_bw[["lambda"]]
+plot3 <- autoplot(comp_bayes_lm_bw, type = "competition")
+plot3
 ```
@@
 ```{r scbi-posterior-viz-1}
-posterior_plots_scbi <- plot_bayesian_model_parameters(
-  posterior_param = posterior_param_scbi,
-  sp_to_plot = c("quru", "litu", "cagl", "cato")
-)
+sp_to_plot <- c("quru", "litu", "cagl", "cato")
 ```
@@
 ```{r scbi-posterior-viz-lambda, out.width="100%", fig.cap="Posterior distribution of lambda's for SCBI."}
-posterior_plots_scbi[["lambda"]]
+plot3 <- autoplot(comp_bayes_lm_bw, type = "competition")
+plot3
 ```

### 2021-02-02 etc/paper/forestecology.Rmd
Changed title

--- 2021-01-25 etc/paper/forestecology.Rmd
@@
 ---
-title: "`forestecology` package for modeling interspecies competition between trees"
+title: "The `forestecology` R package for modeling interspecies competition between trees"
 
@@
 predictions %>%
-  yardstick::rmse(truth = growth, estimate = growth_hat) %>%
+  rmse(truth = growth, estimate = growth_hat) %>%
   pull(.estimate)

### 2021-02-02 etc/paper/forestecology.Rmd
Ugly first stab at new abstract

@@
 # Abstract (350 words)
+
+When modeling growth of trees, the effect of interspecies competition is often included.
+
+We present the `forestecology` package providing methods .
+
+Using R's S3 object-oriented system, we implement a class and methods for a specific linear, Bayesian neighborhood competition model of tree growth.
+Furthermore, this model can be cast in a permutation test framework to evaluate a null hypothesis of no impact of competitor species' identity
+
+The package is setup such that it can be easily extended to other models.
+
+The package follows a tidyverse-like structure whereby verb-named functions can be "piped" to form the sequence of steps of analysis from start to finish.
+Furthermore, most inputs/outputs are of `sf` class from the simple features package, thereby facilitating all wrangling and visualization of geospatial data.
+
+We leverage the ForestGEO network's data collection protocols and standardized data formatting to ensure cross-site compatibility of analysis.
+We apply our model to the Smithsonian Conservation Biology Institute's large forest dynamics plot.
 

### 2021-02-12 etc/paper/forestecology.Rmd
MVP of new abstract in MEE format

--- 2021-02-02 etc/paper/forestecology.Rmd
@@
 
-1. Set the context for and purpose of the work: The scientific question/problem and the desiderata
-    + (Eventually) modularly fitting models for interspecific competition and assessing them using spatial crossvalidation
-    + Leverage ForestGEO protocols providing standardization
-1. Indicate the approach and the methods: Use tidyverse, simple features, and tidymodels packages. 
-    + tidyverse: as stated in tidy tools manifesto: standadized data structures, functional programming with pipe, designed for humans
-    + sf: tidyverse-friendly package that makes wrangling and visualizing spatial data much easier
-    + tidymodels: given our spatial-crossvalidation, use tidymodels framework is a collection of packages for modeling and machine learning using tidyverse principles
-    + Dave: hmm, I think here we might need to mention the methods we implement.
-Something like, the package provides functions to specific a linear, bayesian neighborhood competition model of growth, then fit the model, and compare competing models with spatiall cross validation.
-Or somethign like that.
-1. Outline the main results: We replicate from scratch the figure in PLOSOne paper using Big Woods data, conduct similar analysis for SCBI data.
-    + Code would've been way more complicated in base. 
-    + We don't have to worry about how the component functions work, only that sequence/converyor belt is correct and output is correct.
-    + Scientist is abstracted away from ugly programming details.
-1. Identify the conclusions and wider implications:
-    + New scientific conclusions from SCBI data
-    + Modularly switch out our bayesian lm() functions to anything you want.
-    + this can serve as blue print for other modeling situations.
+1. Set the context for and purpose of the work: 
+1. Indicate the approach and the methods: 
+1. Outline the main results: 
+1. Identify the conclusions and wider implications
+
+1.
+When modeling growth of trees forest ecologists often incorporate the effect of interspecies competition.
+Many such models are based on a neighborhood effect assumption whereby all trees within a fixed distance of all focal trees are considered competitors.
+Methods are needed to evaluate the effect of interspecies competition and to assess their quality.
+1.
+We present the `forestecology` package providing methods for both 1) evaluating the out-of-sample performance of our model using spatial-crossvaliation and 2) testing a null hypothesis of of no impact of competitor species' identity on the growth of trees using a permutation test.
+We implement a class and methods using R's S3 object-oriented system, for a specific linear, Bayesian neighborhood competition model of tree growth.
+1.
+We demonstrate the package's functions using data from the Smithsonian Conservation Biology Institute's large forest dynamics plot, part of the ForestGEO network of reseach sites.
+Given ForestGEO's data collection protocols and data formatting standards, the package cross-compatibility of code.
+We show both that 1) competitor species identity matters and 2) that not spatially cross-validating leads to error estimates that are overly optimistic.
+1.
+The package follows `tidyverse`-like structure whereby verb-named functions can be modularly "piped" in sequence to intuitively display the sequence of steps of analysis from start to finish.
+Additionally, most inputs/outputs of functions assume an are of `sf` class from the simple features package, thereby facilitating all wrangling and visualization of geospatial data.
+Lastly, even though our package is currently limited to one specific model, the package is setup such that it can be easily extended to other models.
 

### 2021-02-12 etc/paper/forestecology.Rmd

@@
 ```{r bw-model-rmse}
-predictions %>%
+focal_vs_comp_bw %>%
   rmse(truth = growth, estimate = growth_hat) %>%

### 2021-02-12 etc/paper/forestecology.tex

@@
 \begin{Highlighting}[]
-\KeywordTok{library}\NormalTok{(forestecology)}
-
-\CommentTok{# Load tidyverse packages:}
 \KeywordTok{library}\NormalTok{(tidyverse)}
 \KeywordTok{library}\NormalTok{(lubridate)}
-
-\CommentTok{# Load spatial packages:}
+\KeywordTok{library}\NormalTok{(sf)}
+\KeywordTok{library}\NormalTok{(forestecology)}
 \KeywordTok{library}\NormalTok{(blockCV)}
-\KeywordTok{library}\NormalTok{(sf)}
-\KeywordTok{library}\NormalTok{(sfheaders)}
-
-\CommentTok{# Load other packages:}
-\KeywordTok{library}\NormalTok{(snakecase)}
-\KeywordTok{library}\NormalTok{(yardstick)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\KeywordTok{data}\NormalTok{(bw_census_}\DecValTok{2008}\NormalTok{, bw_census_}\DecValTok{2014}\NormalTok{, bw_species)}
-
-\CommentTok{# Append additional species data}
-\NormalTok{bw_census_}\DecValTok{2008}\NormalTok{ <-}\StringTok{ }\NormalTok{bw_census_}\DecValTok{2008} \OperatorTok{%>%}
-\StringTok{  }\KeywordTok{left_join}\NormalTok{(bw_species, }\DataTypeTok{by =} \StringTok{"sp"}\NormalTok{) }\OperatorTok{%>%}
+\NormalTok{census_}\DecValTok{2008}\NormalTok{_bw <-}\StringTok{ }\NormalTok{census_}\DecValTok{2008}\NormalTok{_bw }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{left_join}\NormalTok{(species_bw, }\DataTypeTok{by =} \StringTok{"sp"}\NormalTok{) }\OperatorTok{%>%}
 \StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\KeywordTok{c}\NormalTok{(genus, species, latin))}
@@
 \begin{Highlighting}[]
-\NormalTok{scbi_}\DecValTok{2013}\NormalTok{ <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem2.csv"}\NormalTok{) }\OperatorTok{%>%}
+\NormalTok{census_}\DecValTok{2013}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem2.csv"}\NormalTok{) }\OperatorTok{%>%}
 \StringTok{  }\KeywordTok{select}\NormalTok{(treeID, stemID, sp, ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
 \StringTok{  }\KeywordTok{mutate}\NormalTok{(}
-    \DataTypeTok{date =}\NormalTok{ ExactDate,}
-    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh),}
-    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date)}
+    \DataTypeTok{date =}\NormalTok{ ExactDate }\OperatorTok{%>%}\StringTok{ }\KeywordTok{mdy}\NormalTok{(),}
+    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}
 \NormalTok{  ) }\OperatorTok{%>%}
@@
 
-\NormalTok{scbi_}\DecValTok{2018}\NormalTok{ <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem3.csv"}\NormalTok{) }\OperatorTok{%>%}
+\NormalTok{census_}\DecValTok{2018}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem3.csv"}\NormalTok{) }\OperatorTok{%>%}
 \StringTok{  }\KeywordTok{select}\NormalTok{(treeID, stemID, sp, ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
 \StringTok{  }\KeywordTok{mutate}\NormalTok{(}
-    \DataTypeTok{date =}\NormalTok{ ExactDate,}
-    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh),}
-    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date)}
+    \DataTypeTok{date =}\NormalTok{ ExactDate }\OperatorTok{%>%}\StringTok{ }\KeywordTok{mdy}\NormalTok{(),}
+    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}
 \NormalTok{  ) }\OperatorTok{%>%}
@@
 \subsection{Compute annual growth}\label{compute-annual-growth}}
+
+MERGE THIS SUBSECTION with previous?
 
@@
 \begin{Highlighting}[]
-\NormalTok{bw_census_}\DecValTok{2014}\NormalTok{ <-}\StringTok{ }\NormalTok{bw_census_}\DecValTok{2014} \OperatorTok{%>%}
-\StringTok{  }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{))}
-
-\NormalTok{bw_growth_df <-}
-\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(bw_census_}\DecValTok{2008}\NormalTok{, bw_census_}\DecValTok{2014}\NormalTok{, }\DataTypeTok{id =} \StringTok{"treeID"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\CommentTok{# Convert all variables denoting species to factors}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
-    \DataTypeTok{sp =}\NormalTok{ sp }\OperatorTok{%>%}\StringTok{ }\KeywordTok{to_any_case}\NormalTok{() }\OperatorTok{%>%}\StringTok{ }\KeywordTok{as.factor}\NormalTok{(),}
-    \DataTypeTok{species =}\NormalTok{ sp,}
-    \DataTypeTok{family =} \KeywordTok{as.factor}\NormalTok{(family),}
-    \DataTypeTok{trait_group =} \KeywordTok{as.factor}\NormalTok{(trait_group)}
-\NormalTok{  ) }\OperatorTok{%>%}
-\StringTok{  }\CommentTok{# Drop unnecessary variables}
-\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\NormalTok{stemID)}
+\NormalTok{growth_bw <-}
+\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(}
+    \DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2008}\NormalTok{_bw }\OperatorTok{%>%}\StringTok{ }
+\StringTok{      }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{to_any_case}\NormalTok{(sp) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{()),}
+    \DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2014}\NormalTok{_bw }\OperatorTok{%>%}\StringTok{ }
+\StringTok{      }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{      }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{to_any_case}\NormalTok{(sp) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{()),}
+    \DataTypeTok{id =} \StringTok{"treeID"}
+\NormalTok{  )}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{scbi_growth_df <-}
-\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(scbi_}\DecValTok{2013}\NormalTok{, scbi_}\DecValTok{2018}\NormalTok{, }\StringTok{"stemID"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\CommentTok{# Convert all variables denoting species to factors}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{as.factor}\NormalTok{(sp))}
+\NormalTok{growth_scbi <-}
+\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(}
+    \DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi }\OperatorTok{%>%}\StringTok{ }
+\StringTok{      }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{to_any_case}\NormalTok{(sp) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{()),}
+    \DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}\StringTok{ }
+\StringTok{      }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{      }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{to_any_case}\NormalTok{(sp) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{()),}
+    \DataTypeTok{id =} \StringTok{"stemID"}
+\NormalTok{  )}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{max_dist <-}\StringTok{ }\FloatTok{7.5}
+\NormalTok{comp_dist <-}\StringTok{ }\FloatTok{7.5}
 \NormalTok{cv_fold_size <-}\StringTok{ }\DecValTok{100}
@@
 
-The first constant is \texttt{max\_dist} which defines the maximum
+The first constant is \texttt{comp\_dist} which defines the maximum
 distance for a tree's competitive neighborhood. Trees within this
@@
 this distance apart do not. Put differently, all trees within
-\texttt{max\_dist} of a focal tree will be considered its competitors
+\texttt{comp\_dist} of a focal tree will be considered its competitors
 (see below). Other studies have estimated the value of
-\texttt{max\_dist}; we use an average of estimated values
+\texttt{comp\_dist}; we use an average of estimated values
 \citet{canham_neighborhood_2004}, \citet{uriarte_spatially_2004},
@@
 
-Furthermore, \texttt{max\_dist} will define the size of all buffers
+Furthermore, \texttt{comp\_dist} will define the size of all buffers
 considered, which will be encoded as a binary variable \texttt{buffer}
@@
 function takes as input the main \texttt{growth\_df} data frame, the
-\texttt{size} of the buffer which we set as \texttt{max\_dist}, and the
+\texttt{size} of the buffer which we set as \texttt{comp\_dist}, and the
 boundary of the study region encoded as a simple features polygon
@@
 information about the study region boundary. In this case, we use a
-\texttt{sf\_polygon} object \texttt{bw\_study\_region} which comes
+\texttt{sf\_polygon} object \texttt{study\_region\_bw} which comes
 pre-loaded in the \texttt{forestecology} packages. After loading
-\texttt{bw\_study\_region}, we illustrate the results of the
+\texttt{study\_region\_bw}, we illustrate the results of the
 \texttt{add\_buffer\_variable()} function in Figure
@@
 \begin{Highlighting}[]
-\KeywordTok{data}\NormalTok{(bw_study_region)}
-
-\NormalTok{bw_growth_df <-}\StringTok{ }\NormalTok{bw_growth_df }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{direction =} \StringTok{"in"}\NormalTok{, }\DataTypeTok{size =}\NormalTok{ max_dist, }
-                      \DataTypeTok{region =}\NormalTok{ bw_study_region)}
+\KeywordTok{data}\NormalTok{(study_region_bw)}
+
+\NormalTok{growth_bw <-}\StringTok{ }\NormalTok{growth_bw }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{direction =} \StringTok{"in"}\NormalTok{, }\DataTypeTok{size =}\NormalTok{ comp_dist, }\DataTypeTok{region =}\NormalTok{ study_region_bw)}
 
 \KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ bw_growth_df }\OperatorTok{%>%}\StringTok{ }\KeywordTok{sample_frac}\NormalTok{(}\FloatTok{0.2}\NormalTok{), }
-          \KeywordTok{aes}\NormalTok{(}\DataTypeTok{col =}\NormalTok{ buffer), }\DataTypeTok{size =} \FloatTok{0.5}\NormalTok{)}
+\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_bw }\OperatorTok{%>%}\StringTok{ }\KeywordTok{sample_frac}\NormalTok{(}\FloatTok{0.2}\NormalTok{), }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{col =}\NormalTok{ buffer), }\DataTypeTok{size =} \FloatTok{0.5}\NormalTok{)}
 \end{Highlighting}
@@
 \NormalTok{bw_spatialBlock <-}\StringTok{ }\KeywordTok{spatialBlock}\NormalTok{(}
-  \DataTypeTok{speciesData =}\NormalTok{ bw_growth_df, }\DataTypeTok{theRange =}\NormalTok{ cv_fold_size, }\DataTypeTok{k =} \DecValTok{28}\NormalTok{, }\DataTypeTok{xOffset =} \FloatTok{0.5}\NormalTok{, }
+  \DataTypeTok{speciesData =}\NormalTok{ growth_bw, }\DataTypeTok{theRange =}\NormalTok{ cv_fold_size, }\DataTypeTok{k =} \DecValTok{28}\NormalTok{, }\DataTypeTok{xOffset =} \FloatTok{0.5}\NormalTok{, }
   \DataTypeTok{yOffset =} \DecValTok{0}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{showBlocks =} \OtherTok{FALSE}
@@
 \begin{Highlighting}[]
-\NormalTok{bw_growth_df <-}\StringTok{ }\NormalTok{bw_growth_df }\OperatorTok{%>%}
+\NormalTok{growth_bw <-}\StringTok{ }\NormalTok{growth_bw }\OperatorTok{%>%}
 \StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ bw_spatialBlock}\OperatorTok{$}\NormalTok{foldID)}
@@
 \StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ bw_spatialBlock}\OperatorTok{$}\NormalTok{blocks }\OperatorTok{%>%}\StringTok{ }\KeywordTok{st_as_sf}\NormalTok{()) }\OperatorTok{+}\StringTok{ }
-\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ bw_growth_df }\OperatorTok{%>%}\StringTok{ }\KeywordTok{sample_frac}\NormalTok{(}\FloatTok{0.2}\NormalTok{), }
+\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_bw }\OperatorTok{%>%}\StringTok{ }\KeywordTok{sample_frac}\NormalTok{(}\FloatTok{0.2}\NormalTok{), }
           \KeywordTok{aes}\NormalTok{(}\DataTypeTok{col =} \KeywordTok{factor}\NormalTok{(foldID)), }\DataTypeTok{size =} \FloatTok{0.1}\NormalTok{, }\DataTypeTok{show.legend =} \OtherTok{FALSE}\NormalTok{) }\OperatorTok{+}\StringTok{ }
@@
 \begin{Highlighting}[]
-\NormalTok{bw_growth_df <-}\StringTok{ }\NormalTok{bw_growth_df }\OperatorTok{%>%}
+\NormalTok{growth_bw <-}\StringTok{ }\NormalTok{growth_bw }\OperatorTok{%>%}
 \StringTok{  }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\NormalTok{foldID }\OperatorTok{%in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{19}\NormalTok{, }\DecValTok{23}\NormalTok{, }\DecValTok{21}\NormalTok{, }\DecValTok{17}\NormalTok{, }\DecValTok{8}\NormalTok{, }\DecValTok{19}\NormalTok{)) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =} \KeywordTok{as.character}\NormalTok{(foldID))}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =} \KeywordTok{factor}\NormalTok{(foldID))}
 \end{Highlighting}
@@
 Separately, we save the spatial cross-validation grid as an
-\texttt{sf\_polygon} object \texttt{bw\_cv\_grid}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{bw_cv_grid <-}\StringTok{ }\NormalTok{bw_spatialBlock}\OperatorTok{$}\NormalTok{blocks }\OperatorTok{%>%}
+\texttt{sf\_polygon} object \texttt{blocks\_bw}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{blocks_bw <-}\StringTok{ }\NormalTok{bw_spatialBlock}\OperatorTok{$}\NormalTok{blocks }\OperatorTok{%>%}
 \StringTok{  }\KeywordTok{st_as_sf}\NormalTok{()}
@@
 \begin{Highlighting}[]
-\NormalTok{scbi_study_region <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}
+\NormalTok{study_region_scbi <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}
   \DataTypeTok{x =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{),}
@@
 
-\NormalTok{scbi_growth_df <-}\StringTok{ }\NormalTok{scbi_growth_df }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{direction =} \StringTok{"in"}\NormalTok{, }\DataTypeTok{size =}\NormalTok{ max_dist, }
-                      \DataTypeTok{region =}\NormalTok{ scbi_study_region)}
+\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{direction =} \StringTok{"in"}\NormalTok{, }\DataTypeTok{size =}\NormalTok{ comp_dist, }\DataTypeTok{region =}\NormalTok{ study_region_scbi)}
 
 \KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ scbi_growth_df, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{col =}\NormalTok{ buffer), }\DataTypeTok{size =} \FloatTok{0.5}\NormalTok{)}
+\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_scbi, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{col =}\NormalTok{ buffer), }\DataTypeTok{size =} \FloatTok{0.5}\NormalTok{)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{fold1 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{600}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{600}\NormalTok{)) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{sf_polygon}\NormalTok{() }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \DecValTok{1}\NormalTok{)}
-\NormalTok{fold2 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{600}\NormalTok{)) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{sf_polygon}\NormalTok{() }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \DecValTok{2}\NormalTok{)}
-\NormalTok{scbi_cv_grid <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(fold1, fold2)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{scbi_spatialBlock <-}\StringTok{ }\KeywordTok{spatialBlock}\NormalTok{(}
-  \DataTypeTok{speciesData =}\NormalTok{ scbi_growth_df, }\DataTypeTok{k =} \DecValTok{2}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{,}
-  \CommentTok{# Note new arguments:}
-  \DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{, }\DataTypeTok{blocks =}\NormalTok{ scbi_cv_grid}
+\NormalTok{fold1 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{600}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+\NormalTok{fold2 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+
+\NormalTok{blocks_scbi <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}
+  \KeywordTok{sf_polygon}\NormalTok{(fold1),}
+  \KeywordTok{sf_polygon}\NormalTok{(fold2)}
+\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{SpatialBlock_scbi <-}\StringTok{ }\KeywordTok{spatialBlock}\NormalTok{(}
+  \DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =} \DecValTok{2}\NormalTok{, }\DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi,}
+  \DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}
 \NormalTok{)}
@@
 \CommentTok{# Add foldID to each tree}
-\NormalTok{scbi_growth_df <-}\StringTok{ }\NormalTok{scbi_growth_df }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ scbi_spatialBlock}\OperatorTok{$}\NormalTok{foldID)}
+\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ SpatialBlock_scbi}\OperatorTok{$}\NormalTok{foldID }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
 
 \KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_sf_text}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ scbi_growth_df, }
-               \KeywordTok{aes}\NormalTok{(}\DataTypeTok{label =}\NormalTok{ foldID, }\DataTypeTok{col =}\NormalTok{ buffer)) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ scbi_cv_grid, }\DataTypeTok{fill =} \StringTok{"transparent"}\NormalTok{)}
+\StringTok{  }\KeywordTok{geom_sf_text}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_scbi, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{label =}\NormalTok{ foldID, }\DataTypeTok{col =}\NormalTok{ buffer)) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ blocks_scbi, }\DataTypeTok{fill =} \StringTok{"transparent"}\NormalTok{)}
 \end{Highlighting}
@@
 focal tree in the \texttt{growth\_df} data frames to the trees in its
-competitive neighborhood range as defined by the \texttt{max\_dist}
+competitive neighborhood range as defined by the \texttt{comp\_dist}
 constant. So for example, if \texttt{growth\_df} consisted of two focal
-trees with two and three neighbors with \texttt{max\_dist} respectively,
-\texttt{focal\_vs\_comp} would be a data frame of 5 rows connecting each
-focal tree to it's competitors. The \texttt{create\_focal\_vs\_comp()}
-function makes this connection taking as inputs the \texttt{growth\_df}
-data frame; the \texttt{max\_dist} constant defining competitive range;
-\texttt{cv\_grid\_sf}, giving the cross validation grid; and the
-\texttt{id} variable.
+trees with two and three neighbors with \texttt{comp\_dist}
+respectively, \texttt{focal\_vs\_comp} would be a data frame of 5 rows
+connecting each focal tree to it's competitors. The
+\texttt{create\_focal\_vs\_comp()} function makes this connection taking
+as inputs the \texttt{growth\_df} data frame; the \texttt{comp\_dist}
+constant defining competitive range; \texttt{cv\_grid\_sf}, giving the
+cross validation grid; and the \texttt{id} variable.
 
@@
 \begin{Highlighting}[]
-\NormalTok{focal_vs_comp_bw <-}\StringTok{ }\NormalTok{bw_growth_df }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(max_dist, }\DataTypeTok{cv_grid_sf =}\NormalTok{ bw_cv_grid, }\DataTypeTok{id =} \StringTok{"treeID"}\NormalTok{)}
+\NormalTok{focal_vs_comp_bw <-}\StringTok{ }\NormalTok{growth_bw }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(comp_dist, }\DataTypeTok{cv_grid_sf =}\NormalTok{ blocks_bw, }\DataTypeTok{id =} \StringTok{"treeID"}\NormalTok{)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{scbi_growth_df }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(max_dist, }\DataTypeTok{cv_grid_sf =}\NormalTok{ scbi_cv_grid, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{)}
+\NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(comp_dist, }\DataTypeTok{cv_grid_sf =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{posterior_param_bw <-}\StringTok{ }\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{fit_bayesian_model}\NormalTok{(}\DataTypeTok{prior_param =} \OtherTok{NULL}\NormalTok{)}
+\NormalTok{comp_bayes_lm_bw <-}\StringTok{ }\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{comp_bayes_lm}\NormalTok{(}\DataTypeTok{prior_param =} \OtherTok{NULL}\NormalTok{)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{predictions <-}\StringTok{ }\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{predict_bayesian_model}\NormalTok{(}\DataTypeTok{posterior_param =}\NormalTok{ posterior_param_bw) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{right_join}\NormalTok{(bw_growth_df, }\DataTypeTok{by =} \KeywordTok{c}\NormalTok{(}\StringTok{"focal_ID"}\NormalTok{ =}\StringTok{ "treeID"}\NormalTok{))}
+\NormalTok{focal_vs_comp_bw <-}\StringTok{ }\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{growth_hat =} \KeywordTok{predict}\NormalTok{(comp_bayes_lm_bw, focal_vs_comp_bw))}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{predictions }\OperatorTok{%>%}
-\StringTok{  }\NormalTok{yardstick}\OperatorTok{::}\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
+\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
 \StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
@@
 \begin{Highlighting}[]
-\NormalTok{posterior_param_bw_shuffle <-}\StringTok{ }\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{fit_bayesian_model}\NormalTok{(}\DataTypeTok{prior_param =} \OtherTok{NULL}\NormalTok{, }\DataTypeTok{run_shuffle =} \OtherTok{TRUE}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{predictions_shuffle <-}\StringTok{ }\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{predict_bayesian_model}\NormalTok{(}\DataTypeTok{posterior_param =}\NormalTok{ posterior_param_bw_shuffle) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{right_join}\NormalTok{(bw_growth_df, }\DataTypeTok{by =} \KeywordTok{c}\NormalTok{(}\StringTok{"focal_ID"}\NormalTok{ =}\StringTok{ "treeID"}\NormalTok{))}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{predictions_shuffle }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
+\NormalTok{comp_bayes_lm_bw_shuffle <-}\StringTok{ }\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{comp_bayes_lm}\NormalTok{(}\DataTypeTok{prior_param =} \OtherTok{NULL}\NormalTok{, }\DataTypeTok{run_shuffle =} \OtherTok{TRUE}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{focal_vs_comp_bw <-}\StringTok{ }\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{growth_hat_shuffle =} \KeywordTok{predict}\NormalTok{(comp_bayes_lm_bw_shuffle, focal_vs_comp_bw))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat_shuffle) }\OperatorTok{%>%}
 \StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
@@
 \begin{Highlighting}[]
-\NormalTok{posterior_param_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{fit_bayesian_model}\NormalTok{(}\DataTypeTok{prior_param =} \OtherTok{NULL}\NormalTok{, }\DataTypeTok{run_shuffle =} \OtherTok{FALSE}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{scbi_growth_df_noCV <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{predict_bayesian_model}\NormalTok{(}\DataTypeTok{posterior_param =}\NormalTok{ posterior_param_scbi) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{right_join}\NormalTok{(scbi_growth_df, }\DataTypeTok{by =} \KeywordTok{c}\NormalTok{(}\StringTok{"focal_ID"}\NormalTok{ =}\StringTok{ "stemID"}\NormalTok{))}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{scbi_growth_df_noCV }\OperatorTok{%>%}
+\NormalTok{comp_bayes_lm_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{comp_bayes_lm}\NormalTok{(}\DataTypeTok{prior_param =} \OtherTok{NULL}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{growth_hat =} \KeywordTok{predict}\NormalTok{(comp_bayes_lm_scbi, focal_vs_comp_scbi))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
 \StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
 \StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
-\CommentTok{## [1] 0.1280644}
+\CommentTok{## [1] 0.1281398}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{cv_bw <-}\StringTok{ }\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{max_dist =}\NormalTok{ max_dist, }\DataTypeTok{cv_grid =}\NormalTok{ bw_cv_grid) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{right_join}\NormalTok{(bw_growth_df, }\DataTypeTok{by =} \KeywordTok{c}\NormalTok{(}\StringTok{"focal_ID"}\NormalTok{ =}\StringTok{ "treeID"}\NormalTok{))}
-
-\NormalTok{cv_bw }\OperatorTok{%>%}
+\NormalTok{focal_vs_comp_bw <-}\StringTok{ }\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{cv_grid =}\NormalTok{ blocks_bw)}
+
+\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
 \StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
@@
 \begin{Highlighting}[]
-\NormalTok{cv_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{max_dist =}\NormalTok{ max_dist, }\DataTypeTok{cv_grid =}\NormalTok{ scbi_cv_grid) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{right_join}\NormalTok{(scbi_growth_df, }\DataTypeTok{by =} \KeywordTok{c}\NormalTok{(}\StringTok{"focal_ID"}\NormalTok{ =}\StringTok{ "treeID"}\NormalTok{))}
-
-\NormalTok{cv_scbi }\OperatorTok{%>%}
+\NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{cv_grid =}\NormalTok{ blocks_scbi)}
+
+\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
 \StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
 \StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
-\CommentTok{## [1] 0.1494775}
+\CommentTok{## [1] 0.144608}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{focal_vs_comp_bw <-}\StringTok{ }\NormalTok{bw_growth_df }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =}\NormalTok{ family) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{max_dist =}\NormalTok{ max_dist, }\DataTypeTok{cv_grid_sf =}\NormalTok{ bw_cv_grid, }
-                       \DataTypeTok{id =} \StringTok{"treeID"}\NormalTok{)}
-
-\NormalTok{posterior_param_bw <-}\StringTok{ }\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{fit_bayesian_model}\NormalTok{(}\DataTypeTok{prior_param =} \OtherTok{NULL}\NormalTok{, }\DataTypeTok{run_shuffle =} \OtherTok{FALSE}\NormalTok{)}
+\NormalTok{focal_vs_comp_bw <-}\StringTok{ }\NormalTok{growth_bw }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =}\NormalTok{ family }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{()) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{cv_grid_sf =}\NormalTok{ blocks_bw, }\DataTypeTok{id =} \StringTok{"treeID"}\NormalTok{)}
+
+\NormalTok{comp_bayes_lm_bw <-}\StringTok{ }\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{comp_bayes_lm}\NormalTok{(}\DataTypeTok{prior_param =} \OtherTok{NULL}\NormalTok{)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{posterior_plots_bw <-}\StringTok{ }\KeywordTok{plot_bayesian_model_parameters}\NormalTok{(}
-  \DataTypeTok{posterior_param =}\NormalTok{ posterior_param_bw,}
-  \DataTypeTok{sp_to_plot =} \KeywordTok{c}\NormalTok{(}\StringTok{"cornaceae"}\NormalTok{, }\StringTok{"fagaceae"}\NormalTok{, }\StringTok{"hamamelidaceae"}\NormalTok{, }\StringTok{"juglandaceae"}\NormalTok{, }
+\NormalTok{sp_to_plot <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"cornaceae"}\NormalTok{, }\StringTok{"fagaceae"}\NormalTok{, }\StringTok{"hamamelidaceae"}\NormalTok{, }\StringTok{"juglandaceae"}\NormalTok{, }
                  \StringTok{"lauraceae"}\NormalTok{, }\StringTok{"rosaceae"}\NormalTok{, }\StringTok{"sapindaceae"}\NormalTok{, }\StringTok{"ulmaceae"}\NormalTok{)}
-\NormalTok{)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{posterior_plots_bw[[}\StringTok{"beta_0"}\NormalTok{]]}
+\NormalTok{plot1 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_bw, }\DataTypeTok{type =} \StringTok{"intercepts"}\NormalTok{)}
+\NormalTok{plot1}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{posterior_plots_bw[[}\StringTok{"beta_dbh"}\NormalTok{]]}
+\NormalTok{plot2 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_bw, }\DataTypeTok{type =} \StringTok{"dbh_slopes"}\NormalTok{)}
+\NormalTok{plot2}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{posterior_plots_bw[[}\StringTok{"lambda"}\NormalTok{]]}
+\NormalTok{plot3 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_bw, }\DataTypeTok{type =} \StringTok{"competition"}\NormalTok{)}
+\NormalTok{plot3}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{posterior_plots_scbi <-}\StringTok{ }\KeywordTok{plot_bayesian_model_parameters}\NormalTok{(}
-  \DataTypeTok{posterior_param =}\NormalTok{ posterior_param_scbi,}
-  \DataTypeTok{sp_to_plot =} \KeywordTok{c}\NormalTok{(}\StringTok{"quru"}\NormalTok{, }\StringTok{"litu"}\NormalTok{, }\StringTok{"cagl"}\NormalTok{, }\StringTok{"cato"}\NormalTok{)}
-\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{posterior_plots_scbi[[}\StringTok{"lambda"}\NormalTok{]]}
+\NormalTok{sp_to_plot <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"quru"}\NormalTok{, }\StringTok{"litu"}\NormalTok{, }\StringTok{"cagl"}\NormalTok{, }\StringTok{"cato"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{plot3 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_bw, }\DataTypeTok{type =} \StringTok{"competition"}\NormalTok{)}
+\NormalTok{plot3}
 \end{Highlighting}

### 2021-02-12 etc/paper/forestecology.Rmd
Done "compute the growth of trees based on census data" section

@@
 ```{r setup, include = FALSE, cache = FALSE}
+# Internally used packages
+library(viridis)
+library(knitr)
+library(ggplot2)
+library(here)
+
 # knitr settings
-knitr::opts_chunk$set(
+opts_chunk$set(
   # Code output:
@@
 
-# Internally used packages
-library(knitr)
-library(ggplot2)
-library(viridis)
-
 # Random number generator seed value
@@
 # Set ggplot defaults output:
-if (!knitr::is_html_output()) {
+if (!is_html_output()) {
   # Grey theme:
-  theme_set(theme_light())
+  ggplot2::theme_set(theme_light())
   # Color scheme
-  # TODO: REVERT THIS
-  # scale_colour_discrete <- scale_colour_viridis_d
+  scale_colour_discrete <- scale_colour_viridis_d
 }
@@
 
-# Abstract (350 words)
-
-1. Set the context for and purpose of the work: 
-1. Indicate the approach and the methods: 
-1. Outline the main results: 
-1. Identify the conclusions and wider implications
+# Abstract (350 words) {-}
 
@@
 
+The `forestecology` is designed with "tidy" data principles in mind as @wickham_welcome_2019.
+
+Given that our data is of geo-spatial nature, we represent our data using the "simple features" `sf` package class of objects @pebesma_simple_2018 whereby.
+While previously the `sp` package serves such purposes @pebesma_sp_2005, the `sf` package is designed to interface with the `tidyverse` suite of packages.
+
 # Example
 
-We demonstrate the `forestecology` package's features on two data sets, both based on inventory censuses of two sites from the Smithsonian Institution’s ForestGEO international network of 72 long‐term forest dynamics research sites @andersonteixeira_ctfs-forestgeo_2015.
-First, the Michigan Big Woods Forest Dynamics Plot located at the Edwin S.
-George Reserve in Pinckney, MI, USA.
-The 23 ha plot is situated in mature oak-hickory forest.
-The canopy is dominated by white oak (*Quercus alba*), northern red oak (*Q.
-rubra*), black oak (*Q.
-velutina*), shagbark hickory (*Carya ovata*) and pignut hickory (*C.
-glabra*) @allen_michigan_2020.
-In the example below, we will preface any data frames from this plot in with `bw_`.
-
-Second, the Smithsonian Conservation Biology Institute (SCBI) large forest dynamics plot, located at the Smithsonian's National Zoo and Conservation Biology Institute in Front Royal, VA, USA.
+We demonstrate the `forestecology` package's features on the Smithsonian Conservation Biology Institute (SCBI) large forest dynamics plot, located at the Smithsonian's National Zoo and Conservation Biology Institute in Front Royal, VA, USA.
 The 25.6 ha (640 x 400 m) plot is located at the intersection of three of the major physiographic provinces of the eastern US: the Blue Ridge, Ridge and Valley, and Piedmont provinces and is adjacent to the northern end of Shenandoah National Park.
 The forest type is typical mature secondary eastern mixed deciduous forest, with a canopy dominated by tulip poplar (*Liriodendron tulipifera*), oaks (*Quercus* spp.), and hickories (*Carya* spp.), and an understory composed mainly of spicebush (*Lindera benzoin*), paw-paw (*Asimina triloba*), American hornbeam (*Carpinus caroliniana*), and witch hazel (*Hamamelis virginiana*) @bourg_initial_2013.
-In the example below, we will preface any data frames from this plot in with `scbi_`.
 
-The code that generates Figures are included in the supplementary materials.
+A high-level overview of the steps of our analysis pipeline is as follows:
 
-We load all the necessary packages.
+1. Compute the growth of trees based on census data
+1. Add spatial information:
+    1. Define buffer region trees.
+    1. Add spatial cross-validation block information.
+1. Compute focal versus competitor tree information.
+1. Fit model and make predictions.
+1. Additionally: Evaluate model performance using spatial cross-validation.
+1. Additionally: Evaluate the effect of competitor species identity using permutation tests.
+
+We load all necessary packages.
 
@@
 library(forestecology)
-library(blockCV)
 ```
 
-## Preprocess census data
+## Compute the growth of trees based on census data
 
-We start by preprocessing the census data for both sites.
-While ForestGEO data protocols ensure a high degree of standardization between site, minor variations still exist @andersonteixeira_ctfs-forestgeo_2015.
-While the Big Woods data comes pre-loaded in the `forestecology` package, we load the SCBI data as they are saved in .csv files in the SCBI-ForestGEO-Data repository on GitHub @gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020.
-In both cases, we load the census data as R as "tibble" data frames thereby ensuring a standardized input/output format that can be used across all `tidyverse` packages @wickham_welcome_2019.
+The first step in the our analysis sequence is to compute the growth of trees using data from two censuses.
+The `compute_growth()` function computes growth assuming census data that follows ForestGEO standards.
+Despite such standards, minor variations will still exist between sites thereby necessitating some data wrangling and checking.
+For example, the SCBI site records all dbh's in millimeters, whereas the Michigan Big Woods site records them in centimeters @andersonteixeira_ctfs-forestgeo_2015 @allen_michigan_2020.
+The data format of other sites may be such that our `compute_growth()` function doesn't work at all.
+However, in the end all that matters is that the growth of all trees is saved in a data frame of class `sf` whereby the geolocation of each tree is presented in a `geometry` variable of type `<POINT>` and at a minimum the data contains the following variables: a variable uniquely identifying each tree-stem, `sp` of type `fct` factor identifying species, `dbh1` and `dbh2` of type `dbl` quantifying the dbh at earlier and later census, and `growth` of type `dbl` double quantifying the average annual growth in centimeters.
 
-Furthermore, we ensure that the different variables have the correct names, types (`dbl`, `data`, `factor`).
-
-### Big Woods
-
-We load census data from 2008 and 2014 saved in the package, then merge species data (genus, species, linnean classification, family, etc).
-
-```{r bw-load-data}
-census_2008_bw <- census_2008_bw %>%
-  left_join(species_bw, by = "sp") %>%
-  select(-c(genus, species, latin))
-```
-
-### SCBI {#scbi-data}
-
-We load census data from 2008 and 2014 from `.csv` files saved from GitHub on November 20, 2020.
-Furthermore, we perform two additional pre-processing steps.
-First, in order to speed up computation for purposes of this example, we only consider a 9 ha subsection of the 25.6 ha of the SCBI site: `gx` from 0--300 instead of 0--400 and `gy` from  300--600 instead of 0--640.
-Second, in order to standardize comparisons between Big Woods and SCBI, we convert the units of dbh from mm to cm.
-^[A rule of thumb to ascertain if dbh is in mm or cm is to verify if the smallest non-zero and non-missing measurement is 1 or 10.
-If the former, then cm.
-If the later, then mm.
-This is because ForestGEO protocols state that only trees with dbh greater or equal to 1cm should be included in censuses.
-]
+We load both 2008 and 2014 SCBI census data `.csv` files as they existed on GitHub on November 20, 2020.
+After selecting only relevant variables, we perform a few additional data wrangling steps: convert the variable with the date of measurement to be of type `date`, convert dbh to be in centimeters^[A rule of thumb to determine the units of dbh is check if the smallest non-zero and non-missing measurement is 1 or 10.
+If the former, then centimeters.
+If the later, then millimeters.
+This is because ForestGEO protocols state that only trees with dbh greater or equal to 1cm should be included in censuses.], convert the `sp` variable containing species information from type `chr` character to `fct` factor (we will discuss the need for this in Section \ref{spatial-cross-validation}).
+Furthermore, in order to speed up computation for purposes of this examplt, we only consider a 9 ha subsection of the 25.6 ha of the SCBI site: `gx` from 0--300 instead of 0--400 and `gy` from 300--600 instead of 0--640.
 
@@
 census_2013_scbi <- read_csv("scbi.stem2.csv") %>%
-  select(treeID, stemID, sp, ExactDate, gx, gy, dbh, codes, status) %>%
+  select(stemID, sp, date = ExactDate, gx, gy, dbh, codes, status) %>%
   mutate(
-    date = ExactDate %>% mdy(),
-    dbh = as.numeric(dbh)
+    date = mdy(date),
+    dbh = as.numeric(dbh)/10,
+    sp = factor(sp)
   ) %>%
-  filter(gx < 300, between(gy, 300, 600)) %>% 
-  mutate(dbh = dbh / 10)
+  filter(gx < 300, between(gy, 300, 600))
 
 census_2018_scbi <- read_csv("scbi.stem3.csv") %>%
-  select(treeID, stemID, sp, ExactDate, gx, gy, dbh, codes, status) %>%
+  select(stemID, sp, date = ExactDate, gx, gy, dbh, codes, status) %>%
   mutate(
-    date = ExactDate %>% mdy(),
-    dbh = as.numeric(dbh)
+    date = mdy(date),
+    dbh = as.numeric(dbh)/10,
+    sp = factor(sp)
   ) %>%
-  filter(gx < 300, between(gy, 300, 600)) %>% 
-  mutate(dbh = dbh / 10)
+  filter(gx < 300, between(gy, 300, 600))
 ```
 
-## Compute annual growth
-
-MERGE THIS SUBSECTION with previous?
-
-For each plot we then compute average annual growth between the two censuses using the `compute_growth()` function.
-This function takes the two census data frames as well as a character indicating which variable in both data frames uniquely identifies each stem.
-This function returns a single data frame that includes a numerical variable `growth` reflecting the average annual dbh growth (in cm) of all trees alive at both time points.
-Furthermore, variables that (in theory) remain unchanged between censuses appear only once, such as location variables `gx` and `gy`; as well as species-related variables.
-Variables that should change between censuses are suffixed with `1` and `2` indicating the earlier and later censuses, such as `dbh1/dbh2` and `codes1/codes2`.
-Here the resulting data frames are named with some variation of `growth_df`.
-
-After computing the average annual growth for each tree, we ensure to convert all variables denote species from type character to factors; this is to ensure that issues of rare species being accounted for in both training and test sets in our upcoming cross-validation step (see Section REF)
-
-### Big Woods
-
-In the case of Big Woods data, we first remove all trees that were re-sprouts in the later (2014) census.
-Additionally, we have included two classification of tree species: `species` and `family`.
-To illustrate model comparison, we test competition models in which individuals are grouped by species and by family.
-
-```{r bw-compute-growth}
-growth_bw <-
-  compute_growth(
-    census_1 = census_2008_bw %>% 
-      mutate(sp = to_any_case(sp) %>% factor()),
-    census_2 = census_2014_bw %>% 
-      filter(!str_detect(codes, "R")) %>% 
-      mutate(sp = to_any_case(sp) %>% factor()),
-    id = "treeID"
-  )
-```
-
-### SCBI
+These two data frames are then used as the two primary arguments to the `compute_growth()` function, along with the `id` argument whereby the user specifies the name of the variable that uniquely identifies each tree-stem under consideration (note this does not include resprouts in the later census):
 
@@
   compute_growth(
-    census_1 = census_2013_scbi %>% 
-      mutate(sp = to_any_case(sp) %>% factor()),
-    census_2 = census_2018_scbi %>% 
-      filter(!str_detect(codes, "R")) %>% 
-      mutate(sp = to_any_case(sp) %>% factor()),
+    census_1 = census_2013_scbi,
+    census_2 = census_2018_scbi %>% filter(!str_detect(codes, "R")),
     id = "stemID"
   )
+growth_scbi
 ```
 
-### Comparison
+The output `growth_scbi` is a single data frame of class `sf` that includes a numerical `growth` reflecting the average annual growth in dbh (in cm) for all trees that were alive at both time points as well a `geometry` variable encoding each tree's geolocation.
+Furthermore, variables that (in theory) remain unchanged between censuses appear only once, such as location variables `gx` and `gy`; as well as species-related variables.
+Variables that should change between censuses are suffixed with `1` and `2` indicating the earlier and later censuses, such as `dbh1/dbh2` and `codes1/codes2`.
 
-DAVE: I THINK THIS FIGURE IS UNNEEDED.
-COULD BE REMOVED FOR SPACE REASONS.
-Figure \ref{fig:growth-histogram} displays histograms comparing the distribution of average annual growth at both sites.
-Observe that average annual growth appears higher at the Big Woods site.
+Given that `growth_scbi` is of class `sf`, it can be easily plotted in `ggplot2` using the `geom_sf()` geometry as seen in Figure \ref{fig:scbi-trees}.
 
-```{r growth-histogram, out.width="100%", fig.cap="Distribution of average annual growth in DBH for both sites.", echo = FALSE}
-# Both Big Woods & SCBI
-bind_rows(
-  growth_bw %>% st_drop_geometry() %>% select(growth) %>% mutate(Site = "Big Woods"),
-  growth_scbi %>% st_drop_geometry() %>% select(growth) %>% mutate(Site = "SCBI")
-) %>% 
-  ggplot(aes(x = growth, y = ..density.., fill = Site)) +
-  geom_histogram(alpha = 0.5, position = "identity", binwidth = 0.05) +
-  labs(x = "Average annual growth in dbh (cm per yr)") +
-  coord_cartesian(xlim = c(-0.5, 1))
+TODO: Rescale points in this plot:
+
+```{r scbi-trees, out.width="66%", fig.cap="Growth of trees at SCBI."}
+ggplot() +
+  geom_sf(data = growth_scbi, aes(size = growth)) + 
+  scale_size(breaks = c(0.01, 0.1, 1), range = c(0.1, 1))
 ```
@@
 
-## Run spatial cross-validation
+## Run spatial cross-validation {#spatial-cross-validation}
 

### 2021-02-12 etc/paper/forestecology.tex
Done "compute the growth of trees based on census data" section

@@
 {
-  \title{\bf \texttt{forestecology} package for modeling interspecies competition
-between trees}
+  \title{\bf The \texttt{forestecology} R package for modeling interspecies
+competition between trees}
 
@@
   \begin{center}
-    {\LARGE\bf \texttt{forestecology} package for modeling interspecies competition
-between trees}
+    {\LARGE\bf The \texttt{forestecology} R package for modeling interspecies
+competition between trees}
   \end{center}
@@
 \hypertarget{abstract-350-words}{%
-\section{Abstract (350 words)}\label{abstract-350-words}}
+\section*{Abstract (350 words)}\label{abstract-350-words}}
+\addcontentsline{toc}{section}{Abstract (350 words)}
 
@@
 \item
-  Set the context for and purpose of the work: The scientific
-  question/problem and the desiderata
-
-  \begin{itemize}
-  \tightlist
-  \item
-    (Eventually) modularly fitting models for interspecific competition
-    and assessing them using spatial crossvalidation
-  \item
-    Leverage ForestGEO protocols providing standardization
-  \end{itemize}
-\item
-  Indicate the approach and the methods: Use tidyverse, simple features,
-  and tidymodels packages.
-
-  \begin{itemize}
-  \tightlist
-  \item
-    tidyverse: as stated in tidy tools manifesto: standadized data
-    structures, functional programming with pipe, designed for humans
-  \item
-    sf: tidyverse-friendly package that makes wrangling and visualizing
-    spatial data much easier
-  \item
-    tidymodels: given our spatial-crossvalidation, use tidymodels
-    framework is a collection of packages for modeling and machine
-    learning using tidyverse principles
-  \item
-    Dave: hmm, I think here we might need to mention the methods we
-    implement. Something like, the package provides functions to
-    specific a linear, bayesian neighborhood competition model of
-    growth, then fit the model, and compare competing models with
-    spatiall cross validation. Or somethign like that.
-  \end{itemize}
-\item
-  Outline the main results: We replicate from scratch the figure in
-  PLOSOne paper using Big Woods data, conduct similar analysis for SCBI
-  data.
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Code would've been way more complicated in base.
-  \item
-    We don't have to worry about how the component functions work, only
-    that sequence/converyor belt is correct and output is correct.
-  \item
-    Scientist is abstracted away from ugly programming details.
-  \end{itemize}
-\item
-  Identify the conclusions and wider implications:
-
-  \begin{itemize}
-  \tightlist
-  \item
-    New scientific conclusions from SCBI data
-  \item
-    Modularly switch out our bayesian lm() functions to anything you
-    want.
-  \item
-    this can serve as blue print for other modeling situations.
-  \end{itemize}
+  When modeling growth of trees forest ecologists often incorporate the
+  effect of interspecies competition. Many such models are based on a
+  neighborhood effect assumption whereby all trees within a fixed
+  distance of all focal trees are considered competitors. Methods are
+  needed to evaluate the effect of interspecies competition and to
+  assess their quality.
+\item
+  We present the \texttt{forestecology} package providing methods for
+  both 1) evaluating the out-of-sample performance of our model using
+  spatial-crossvaliation and 2) testing a null hypothesis of of no
+  impact of competitor species' identity on the growth of trees using a
+  permutation test. We implement a class and methods using R's S3
+  object-oriented system, for a specific linear, Bayesian neighborhood
+  competition model of tree growth.
+\item
+  We demonstrate the package's functions using data from the Smithsonian
+  Conservation Biology Institute's large forest dynamics plot, part of
+  the ForestGEO network of reseach sites. Given ForestGEO's data
+  collection protocols and data formatting standards, the package
+  cross-compatibility of code. We show both that 1) competitor species
+  identity matters and 2) that not spatially cross-validating leads to
+  error estimates that are overly optimistic.
+\item
+  The package follows \texttt{tidyverse}-like structure whereby
+  verb-named functions can be modularly ``piped'' in sequence to
+  intuitively display the sequence of steps of analysis from start to
+  finish. Additionally, most inputs/outputs of functions assume an are
+  of \texttt{sf} class from the simple features package, thereby
+  facilitating all wrangling and visualization of geospatial data.
+  Lastly, even though our package is currently limited to one specific
+  model, the package is setup such that it can be easily extended to
+  other models.
 \end{enumerate}
@@
 
+The \texttt{forestecology} is designed with ``tidy'' data principles in
+mind as \citet{wickham_welcome_2019}.
+
+Given that our data is of geo-spatial nature, we represent our data
+using the ``simple features'' \texttt{sf} package class of objects
+\citet{pebesma_simple_2018} whereby. While previously the \texttt{sp}
+package serves such purposes \citet{pebesma_sp_2005}, the \texttt{sf}
+package is designed to interface with the \texttt{tidyverse} suite of
+packages.
+
 \hypertarget{example}{%
@@
 
-We demonstrate the \texttt{forestecology} package's features on two data
-sets, both based on inventory censuses of two sites from the Smithsonian
-Institution's ForestGEO international network of 72 long‐term forest
-dynamics research sites \citet{andersonteixeira_ctfs-forestgeo_2015}.
-First, the Michigan Big Woods Forest Dynamics Plot located at the Edwin
-S. George Reserve in Pinckney, MI, USA. The 23 ha plot is situated in
-mature oak-hickory forest. The canopy is dominated by white oak
-(\emph{Quercus alba}), northern red oak (\emph{Q. rubra}), black oak
-(\emph{Q. velutina}), shagbark hickory (\emph{Carya ovata}) and pignut
-hickory (\emph{C. glabra}) \citet{allen_michigan_2020}. In the example
-below, we will preface any data frames from this plot in with
-\texttt{bw\_}.
-
-Second, the Smithsonian Conservation Biology Institute (SCBI) large
-forest dynamics plot, located at the Smithsonian's National Zoo and
-Conservation Biology Institute in Front Royal, VA, USA. The 25.6 ha (640
-x 400 m) plot is located at the intersection of three of the major
-physiographic provinces of the eastern US: the Blue Ridge, Ridge and
-Valley, and Piedmont provinces and is adjacent to the northern end of
-Shenandoah National Park. The forest type is typical mature secondary
-eastern mixed deciduous forest, with a canopy dominated by tulip poplar
+We demonstrate the \texttt{forestecology} package's features on the
+Smithsonian Conservation Biology Institute (SCBI) large forest dynamics
+plot, located at the Smithsonian's National Zoo and Conservation Biology
+Institute in Front Royal, VA, USA. The 25.6 ha (640 x 400 m) plot is
+located at the intersection of three of the major physiographic
+provinces of the eastern US: the Blue Ridge, Ridge and Valley, and
+Piedmont provinces and is adjacent to the northern end of Shenandoah
+National Park. The forest type is typical mature secondary eastern mixed
+deciduous forest, with a canopy dominated by tulip poplar
 (\emph{Liriodendron tulipifera}), oaks (\emph{Quercus} spp.), and
@@
 American hornbeam (\emph{Carpinus caroliniana}), and witch hazel
-(\emph{Hamamelis virginiana}) \citet{bourg_initial_2013}. In the example
-below, we will preface any data frames from this plot in with
-\texttt{scbi\_}.
-
-The code that generates Figures are included in the supplementary
-materials.
-
-We load all the necessary packages.
+(\emph{Hamamelis virginiana}) \citet{bourg_initial_2013}.
+
+A high-level overview of the steps of our analysis pipeline is as
+follows:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Compute the growth of trees based on census data
+\item
+  Add spatial information:
+
+  \begin{enumerate}
+  \def\labelenumii{\arabic{enumii}.}
+  \tightlist
+  \item
+    Define buffer region trees.
+  \item
+    Add spatial cross-validation block information.
+  \end{enumerate}
+\item
+  Compute focal versus competitor tree information.
+\item
+  Fit model and make predictions.
+\item
+  Additionally: Evaluate model performance using spatial
+  cross-validation.
+\item
+  Additionally: Evaluate the effect of competitor species identity using
+  permutation tests.
+\end{enumerate}
+
+We load all necessary packages.
 
@@
 \KeywordTok{library}\NormalTok{(forestecology)}
-\KeywordTok{library}\NormalTok{(blockCV)}
-\end{Highlighting}
-\end{Shaded}
-
-\hypertarget{preprocess-census-data}{%
-\subsection{Preprocess census data}\label{preprocess-census-data}}
-
-We start by preprocessing the census data for both sites. While
-ForestGEO data protocols ensure a high degree of standardization between
-site, minor variations still exist
-\citet{andersonteixeira_ctfs-forestgeo_2015}. While the Big Woods data
-comes pre-loaded in the \texttt{forestecology} package, we load the SCBI
-data as they are saved in .csv files in the SCBI-ForestGEO-Data
-repository on GitHub
-\citet{gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020}. In both
-cases, we load the census data as R as ``tibble'' data frames thereby
-ensuring a standardized input/output format that can be used across all
-\texttt{tidyverse} packages \citet{wickham_welcome_2019}.
-
-Furthermore, we ensure that the different variables have the correct
-names, types (\texttt{dbl}, \texttt{data}, \texttt{factor}).
-
-\hypertarget{big-woods}{%
-\subsubsection{Big Woods}\label{big-woods}}
-
-We load census data from 2008 and 2014 saved in the package, then merge
-species data (genus, species, linnean classification, family, etc).
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{census_}\DecValTok{2008}\NormalTok{_bw <-}\StringTok{ }\NormalTok{census_}\DecValTok{2008}\NormalTok{_bw }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{left_join}\NormalTok{(species_bw, }\DataTypeTok{by =} \StringTok{"sp"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\KeywordTok{c}\NormalTok{(genus, species, latin))}
-\end{Highlighting}
-\end{Shaded}
-
-\hypertarget{scbi-data}{%
-\subsubsection{SCBI}\label{scbi-data}}
-
-We load census data from 2008 and 2014 from \texttt{.csv} files saved
-from GitHub on November 20, 2020. Furthermore, we perform two additional
-pre-processing steps. First, in order to speed up computation for
-purposes of this example, we only consider a 9 ha subsection of the 25.6
-ha of the SCBI site: \texttt{gx} from 0--300 instead of 0--400 and
-\texttt{gy} from 300--600 instead of 0--640. Second, in order to
-standardize comparisons between Big Woods and SCBI, we convert the units
-of dbh from mm to cm. \footnote{A rule of thumb to ascertain if dbh is
-  in mm or cm is to verify if the smallest non-zero and non-missing
-  measurement is 1 or 10. If the former, then cm. If the later, then mm.
-  This is because ForestGEO protocols state that only trees with dbh
-  greater or equal to 1cm should be included in censuses. }
+\end{Highlighting}
+\end{Shaded}
+
+\hypertarget{compute-the-growth-of-trees-based-on-census-data}{%
+\subsection{Compute the growth of trees based on census
+data}\label{compute-the-growth-of-trees-based-on-census-data}}
+
+The first step in the our analysis sequence is to compute the growth of
+trees using data from two censuses. The \texttt{compute\_growth()}
+function computes growth assuming census data that follows ForestGEO
+standards. Despite such standards, minor variations will still exist
+between sites thereby necessitating some data wrangling and checking.
+For example, the SCBI site records all dbh's in millimeters, whereas the
+Michigan Big Woods site records them in centimeters
+\citet{andersonteixeira_ctfs-forestgeo_2015}
+\citet{allen_michigan_2020}. The data format of other sites may be such
+that our \texttt{compute\_growth()} function doesn't work at all.
+However, in the end all that matters is that the growth of all trees is
+saved in a data frame of class \texttt{sf} whereby the geolocation of
+each tree is presented in a \texttt{geometry} variable of type
+\texttt{\textless{}POINT\textgreater{}} and at a minimum the data
+contains the following variables: a variable uniquely identifying each
+tree-stem, \texttt{sp} of type \texttt{fct} factor identifying species,
+\texttt{dbh1} and \texttt{dbh2} of type \texttt{dbl} quantifying the dbh
+at earlier and later census, and \texttt{growth} of type \texttt{dbl}
+double quantifying the average annual growth in centimeters.
+
+We load both 2008 and 2014 SCBI census data \texttt{.csv} files as they
+existed on GitHub on November 20, 2020. After selecting only relevant
+variables, we perform a few additional data wrangling steps: convert the
+variable with the date of measurement to be of type \texttt{date},
+convert dbh to be in centimeters\footnote{A rule of thumb to determine
+  the units of dbh is check if the smallest non-zero and non-missing
+  measurement is 1 or 10. If the former, then centimeters. If the later,
+  then millimeters. This is because ForestGEO protocols state that only
+  trees with dbh greater or equal to 1cm should be included in censuses.},
+convert the \texttt{sp} variable containing species information from
+type \texttt{chr} character to \texttt{fct} factor (we will discuss the
+need for this in Section \ref{spatial-cross-validation}). Furthermore,
+in order to speed up computation for purposes of this examplt, we only
+consider a 9 ha subsection of the 25.6 ha of the SCBI site: \texttt{gx}
+from 0--300 instead of 0--400 and \texttt{gy} from 300--600 instead of
+0--640.
 
@@
 \NormalTok{census_}\DecValTok{2013}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem2.csv"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{select}\NormalTok{(treeID, stemID, sp, ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
 \StringTok{  }\KeywordTok{mutate}\NormalTok{(}
-    \DataTypeTok{date =}\NormalTok{ ExactDate }\OperatorTok{%>%}\StringTok{ }\KeywordTok{mdy}\NormalTok{(),}
-    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}
+    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date),}
+    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}\NormalTok{,}
+    \DataTypeTok{sp =} \KeywordTok{factor}\NormalTok{(sp)}
 \NormalTok{  ) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{dbh =}\NormalTok{ dbh }\OperatorTok{/}\StringTok{ }\DecValTok{10}\NormalTok{)}
+\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
 
 \NormalTok{census_}\DecValTok{2018}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem3.csv"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{select}\NormalTok{(treeID, stemID, sp, ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
 \StringTok{  }\KeywordTok{mutate}\NormalTok{(}
-    \DataTypeTok{date =}\NormalTok{ ExactDate }\OperatorTok{%>%}\StringTok{ }\KeywordTok{mdy}\NormalTok{(),}
-    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}
+    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date),}
+    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}\NormalTok{,}
+    \DataTypeTok{sp =} \KeywordTok{factor}\NormalTok{(sp)}
 \NormalTok{  ) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{dbh =}\NormalTok{ dbh }\OperatorTok{/}\StringTok{ }\DecValTok{10}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\hypertarget{compute-annual-growth}{%
-\subsection{Compute annual growth}\label{compute-annual-growth}}
-
-MERGE THIS SUBSECTION with previous?
-
-For each plot we then compute average annual growth between the two
-censuses using the \texttt{compute\_growth()} function. This function
-takes the two census data frames as well as a character indicating which
-variable in both data frames uniquely identifies each stem. This
-function returns a single data frame that includes a numerical variable
-\texttt{growth} reflecting the average annual dbh growth (in cm) of all
-trees alive at both time points. Furthermore, variables that (in theory)
-remain unchanged between censuses appear only once, such as location
-variables \texttt{gx} and \texttt{gy}; as well as species-related
-variables. Variables that should change between censuses are suffixed
-with \texttt{1} and \texttt{2} indicating the earlier and later
-censuses, such as \texttt{dbh1/dbh2} and \texttt{codes1/codes2}. Here
-the resulting data frames are named with some variation of
-\texttt{growth\_df}.
-
-After computing the average annual growth for each tree, we ensure to
-convert all variables denote species from type character to factors;
-this is to ensure that issues of rare species being accounted for in
-both training and test sets in our upcoming cross-validation step (see
-Section REF)
-
-\hypertarget{big-woods-1}{%
-\subsubsection{Big Woods}\label{big-woods-1}}
-
-In the case of Big Woods data, we first remove all trees that were
-re-sprouts in the later (2014) census. Additionally, we have included
-two classification of tree species: \texttt{species} and
-\texttt{family}. To illustrate model comparison, we test competition
-models in which individuals are grouped by species and by family.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{growth_bw <-}
-\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(}
-    \DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2008}\NormalTok{_bw }\OperatorTok{%>%}\StringTok{ }
-\StringTok{      }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{to_any_case}\NormalTok{(sp) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{()),}
-    \DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2014}\NormalTok{_bw }\OperatorTok{%>%}\StringTok{ }
-\StringTok{      }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
-\StringTok{      }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{to_any_case}\NormalTok{(sp) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{()),}
-    \DataTypeTok{id =} \StringTok{"treeID"}
-\NormalTok{  )}
-\end{Highlighting}
-\end{Shaded}
-
-\hypertarget{scbi}{%
-\subsubsection{SCBI}\label{scbi}}
+\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+These two data frames are then used as the two primary arguments to the
+\texttt{compute\_growth()} function, along with the \texttt{id} argument
+whereby the user specifies the name of the variable that uniquely
+identifies each tree-stem under consideration (note this does not
+include resprouts in the later census):
 
@@
 \StringTok{  }\KeywordTok{compute_growth}\NormalTok{(}
-    \DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi }\OperatorTok{%>%}\StringTok{ }
-\StringTok{      }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{to_any_case}\NormalTok{(sp) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{()),}
-    \DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}\StringTok{ }
-\StringTok{      }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
-\StringTok{      }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{to_any_case}\NormalTok{(sp) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{()),}
+    \DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi,}
+    \DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)),}
     \DataTypeTok{id =} \StringTok{"stemID"}
 \NormalTok{  )}
-\end{Highlighting}
-\end{Shaded}
-
-\hypertarget{comparison}{%
-\subsubsection{Comparison}\label{comparison}}
-
-DAVE: I THINK THIS FIGURE IS UNNEEDED. COULD BE REMOVED FOR SPACE
-REASONS. Figure \ref{fig:growth-histogram} displays histograms comparing
-the distribution of average annual growth at both sites. Observe that
-average annual growth appears higher at the Big Woods site.
+\NormalTok{growth_scbi}
+\CommentTok{## Simple feature collection with 7954 features and 8 fields}
+\CommentTok{## geometry type:  POINT}
+\CommentTok{## dimension:      XY}
+\CommentTok{## bbox:           xmin: 0.2 ymin: 300 xmax: 299.9 ymax: 600}
+\CommentTok{## CRS:            NA}
+\CommentTok{## # A tibble: 7,954 x 9}
+\CommentTok{##    stemID sp     dbh1 codes1 status  dbh2 codes2  growth     geometry}
+\CommentTok{##     <dbl> <fct> <dbl> <chr>  <chr>  <dbl> <chr>    <dbl>      <POINT>}
+\CommentTok{##  1      4 nysy  13.6  M      A       14.2 M       0.103  (14.2 428.5)}
+\CommentTok{##  2      5 havi   8.8  M      A        9.6 M;P     0.150   (9.4 436.4)}
+\CommentTok{##  3      6 havi   3.25 NULL   A        4   M       0.140     (1.3 434)}
+\CommentTok{##  4     77 qual  65.2  M      A       66   M       0.141  (34.7 307.2)}
+\CommentTok{##  5     79 tiam  47.7  M      A       46.8 M      -0.161    (40 381.1)}
+\CommentTok{##  6     80 caca   5.15 M      A        6.5 M       0.253  (38.7 421.7)}
+\CommentTok{##  7     96 libe   2.3  J;M    A        3.7 M       0.262      (60 310)}
+\CommentTok{##  8    100 caca   5.09 NULL   A       NA   DN     NA      (52.5 476.3)}
+\CommentTok{##  9    101 litu  65.4  M      A       68.4 M       0.552  (47.1 567.3)}
+\CommentTok{## 10    102 astr   1.99 NULL   A        2.5 M       0.0954 (40.8 575.5)}
+\CommentTok{## # ... with 7,944 more rows}
+\end{Highlighting}
+\end{Shaded}
+
+The output \texttt{growth\_scbi} is a single data frame of class
+\texttt{sf} that includes a numerical \texttt{growth} reflecting the
+average annual growth in dbh (in cm) for all trees that were alive at
+both time points as well a \texttt{geometry} variable encoding each
+tree's geolocation. Furthermore, variables that (in theory) remain
+unchanged between censuses appear only once, such as location variables
+\texttt{gx} and \texttt{gy}; as well as species-related variables.
+Variables that should change between censuses are suffixed with
+\texttt{1} and \texttt{2} indicating the earlier and later censuses,
+such as \texttt{dbh1/dbh2} and \texttt{codes1/codes2}.
+
+Given that \texttt{growth\_scbi} is of class \texttt{sf}, it can be
+easily plotted in \texttt{ggplot2} using the \texttt{geom\_sf()}
+geometry as seen in Figure \ref{fig:scbi-trees}.
+
+TODO: Rescale points in this plot:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
+\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_scbi, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{size =}\NormalTok{ growth)) }\OperatorTok{+}\StringTok{ }
+\StringTok{  }\KeywordTok{scale_size}\NormalTok{(}\DataTypeTok{breaks =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.01}\NormalTok{, }\FloatTok{0.1}\NormalTok{, }\DecValTok{1}\NormalTok{), }\DataTypeTok{range =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\DecValTok{1}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{Figures/growth-histogram-1}
-
-}
-
-\caption{Distribution of average annual growth in DBH for both sites.}\label{fig:growth-histogram}
+{\centering \includegraphics[width=0.66\linewidth]{Figures/scbi-trees-1}
+
+}
+
+\caption{Growth of trees at SCBI.}\label{fig:scbi-trees}
 \end{figure}
@@
 
-\hypertarget{big-woods-2}{%
-\subsubsection{Big Woods}\label{big-woods-2}}
+\hypertarget{big-woods}{%
+\subsubsection{Big Woods}\label{big-woods}}
 
@@
 
-\hypertarget{scbi-1}{%
-\subsubsection{SCBI}\label{scbi-1}}
+\hypertarget{scbi}{%
+\subsubsection{SCBI}\label{scbi}}
 
@@
 
-\hypertarget{big-woods-3}{%
-\subsubsection{Big Woods}\label{big-woods-3}}
+\hypertarget{big-woods-1}{%
+\subsubsection{Big Woods}\label{big-woods-1}}
 
@@
 
-\hypertarget{scbi-2}{%
-\subsubsection{SCBI}\label{scbi-2}}
+\hypertarget{scbi-1}{%
+\subsubsection{SCBI}\label{scbi-1}}
 
@@
 
-\hypertarget{big-woods-4}{%
-\subsubsection{Big Woods}\label{big-woods-4}}
+\hypertarget{big-woods-2}{%
+\subsubsection{Big Woods}\label{big-woods-2}}
 
@@
 
-\hypertarget{scbi-3}{%
-\subsubsection{SCBI}\label{scbi-3}}
+\hypertarget{scbi-2}{%
+\subsubsection{SCBI}\label{scbi-2}}
 
@@
 
-\hypertarget{run-spatial-cross-validation}{%
+\hypertarget{spatial-cross-validation}{%
 \subsection{Run spatial
-cross-validation}\label{run-spatial-cross-validation}}
+cross-validation}\label{spatial-cross-validation}}
 
@@
 
-\hypertarget{big-woods-5}{%
-\subsubsection{Big Woods}\label{big-woods-5}}
+\hypertarget{big-woods-3}{%
+\subsubsection{Big Woods}\label{big-woods-3}}
 
@@
 
-\hypertarget{scbi-4}{%
-\subsubsection{SCBI}\label{scbi-4}}
+\hypertarget{scbi-3}{%
+\subsubsection{SCBI}\label{scbi-3}}
 
@@
 
-\hypertarget{big-woods-6}{%
-\subsubsection{Big Woods}\label{big-woods-6}}
+\hypertarget{big-woods-4}{%
+\subsubsection{Big Woods}\label{big-woods-4}}
 
@@
 
-\hypertarget{scbi-5}{%
-\subsubsection{SCBI}\label{scbi-5}}
+\hypertarget{scbi-4}{%
+\subsubsection{SCBI}\label{scbi-4}}
 

### 2021-02-15 etc/paper/forestecology.Rmd
Replicated PLOS One simulations

@@
 1. Fit model and make predictions.
+1. Visualize model results.
 1. Additionally: Evaluate model performance using spatial cross-validation.
@@
 ## Add spatial information {#spatial-information}
+
+The next step in the our analysis sequence is to add additional spatial information to our main data frame: defining a buffer region of trees as well as adding spatial cross-validation block information.
+
+HEY BERT PICK IT UP HERE
 
@@
 
-## Run spatial cross-validation {#spatial-cross-validation}
-
-The model fits and predictions in Section \ref{model-fit-predict} all suffer from a common failing: they use the same data to both fit the model and to assess the model's performance using the RMSE.
-As argued by @roberts_cross-validation_2017, this can lead to overly optimistic assessments of model quality as the models can be overfit, in particular in situations where spatial-autocorrelation is present.
-To mitigate the effects of such overfitting, we use a spatially block cross-validation algorithm implemented in the `run_cv()`.
-This function at its core uses the same model fitting implemented in the `fit_bayesian_model()` function, however trains the model on $k-1$ spatial folds of the train and returns fitted values for the test data.
-Recall that the spatial blocking scheme wass encoded in Section \ref{spatial-information}.
-
-### Big Woods
-
-Applying this spatially cross-validated model fit yields an RMSE is higher than that when the model is fit without cross validation.
-In other words, our model fits in \ref{model-fit-predict} were overly optimistic in the model's fitting power, whereas a cross-validated results yield an estimate that is closer to the truth.
-See @allen_permutation_2020 for more discussion of this.
-
-```{r bw-spatial-cv}
-focal_vs_comp_bw <- focal_vs_comp_bw %>%
-  run_cv(comp_dist = comp_dist, cv_grid = blocks_bw)
-
-focal_vs_comp_bw %>%
-  rmse(truth = growth, estimate = growth_hat) %>%
-  pull(.estimate)
-```
-
-### SCBI
-
-Observe once again that this RMSE is much higher than that for the above SCBI model fit without cross-validation.
-
-```{r scbi-spatial-cv}
-focal_vs_comp_scbi <- focal_vs_comp_scbi %>%
-  run_cv(comp_dist = comp_dist, cv_grid = blocks_scbi)
-
-focal_vs_comp_scbi %>%
-  rmse(truth = growth, estimate = growth_hat) %>%
-  pull(.estimate)
-```
-
-## Visualize posterior distributions {#viz-posterior-distributions}
+## Visualize model results {#viz-posterior-distributions}
 
@@
 
+## Run spatial cross-validation {#spatial-cross-validation}
+
+The model fits and predictions in Section \ref{model-fit-predict} all suffer from a common failing: they use the same data to both fit the model and to assess the model's performance using the RMSE.
+As argued by @roberts_cross-validation_2017, this can lead to overly optimistic assessments of model quality as the models can be overfit, in particular in situations where spatial-autocorrelation is present.
+To mitigate the effects of such overfitting, we use a spatially block cross-validation algorithm implemented in the `run_cv()`.
+This function at its core uses the same model fitting implemented in the `fit_bayesian_model()` function, however trains the model on $k-1$ spatial folds of the train and returns fitted values for the test data.
+Recall that the spatial blocking scheme wass encoded in Section \ref{spatial-information}.
+
+### Big Woods
+
+Applying this spatially cross-validated model fit yields an RMSE is higher than that when the model is fit without cross validation.
+In other words, our model fits in \ref{model-fit-predict} were overly optimistic in the model's fitting power, whereas a cross-validated results yield an estimate that is closer to the truth.
+See @allen_permutation_2020 for more discussion of this.
+
+```{r bw-spatial-cv}
+focal_vs_comp_bw <- focal_vs_comp_bw %>%
+  run_cv(comp_dist = comp_dist, cv_grid = blocks_bw)
+
+focal_vs_comp_bw %>%
+  rmse(truth = growth, estimate = growth_hat) %>%
+  pull(.estimate)
+```
+
+### SCBI
+
+Observe once again that this RMSE is much higher than that for the above SCBI model fit without cross-validation.
+
+```{r scbi-spatial-cv}
+focal_vs_comp_scbi <- focal_vs_comp_scbi %>%
+  run_cv(comp_dist = comp_dist, cv_grid = blocks_scbi)
+
+focal_vs_comp_scbi %>%
+  rmse(truth = growth, estimate = growth_hat) %>%
+  pull(.estimate)
+```
+
 # Discussion

### 2021-02-15 etc/PLOSOne/replicate_plosone.R

@@ -72,7 +72,7 @@ focal_vs_comp_bw <- growth_bw %>%
 
 # Fit model and make predictions -----------------------------------------------
 # Number of permutation shuffles:
-num_shuffle <- 9
+num_shuffle <- 49
 
 # For all three possible notions of species
 species_notion_vector <- c("trait_group", "family", "sp")

### 2021-02-15 README.Rmd
tweaked run_cv input
Set default direction = "in" for add_buffer_variable

@@
 growth_ex <- growth_ex %>%
-  add_buffer_variable(direction = "in", size = comp_dist, region = study_region_ex)
+  add_buffer_variable(size = comp_dist, region = study_region_ex)
 
@@
 buffer_region <- study_region_ex %>% 
-  compute_buffer_region(direction = "in", size = comp_dist)
+  compute_buffer_region(size = comp_dist)
 ```
@@
 focal_vs_comp_ex <- focal_vs_comp_ex %>%
-  run_cv(comp_dist = comp_dist, cv_grid = cv_grid_sf_ex)
+  run_cv(comp_dist = comp_dist, cv_grid = blocks_ex)
 ```

### 2021-02-15 R/spatial_functions.R

@@ -2,7 +2,7 @@
 #'
 #' @inheritParams add_buffer_variable
 #' @param region An \code{sf} polygon object of region to be buffered
-#' @param direction "in" for buffers that are contained within \code{region}, "out" for buffers that contain \code{region}.
+#' @param direction "in" for buffers that are contained within \code{region} (default), "out" for buffers that contain \code{region}.
 #'
 #' @return An \code{sf} polygon object of buffer
 #'
@@ -118,7 +118,7 @@ compute_buffer_region <- function(region, direction = "in", size) {
 #'   compute_buffer_region(direction = "in", size = size)
 #' p +
 #'   geom_sf(data = buffer_boundary, col = "red", fill = "transparent")
-add_buffer_variable <- function(growth_df, direction, size, region) {
+add_buffer_variable <- function(growth_df, direction = "in", size, region) {
   buffer_boundary <- region %>%
     compute_buffer_region(direction, size = size)
 

### 2021-02-15 etc/paper/forestecology.Rmd
Done "add spatial information" section

--- 2021-02-12 etc/paper/forestecology.Rmd
@@
   # Color scheme
-  scale_colour_discrete <- scale_colour_viridis_d
+  # scale_colour_discrete <- scale_colour_viridis_d
 }
@@
 library(forestecology)
+library(blockCV)
 ```
@@
 
-We now encode spatial information to the `growth_df` data frames.
-First, in order to control for study region edge effects, we add "buffers" to the periphery of the study region (cite Waller?).
-Our model of interspecific competition relies on a spatial definition of who the competitor trees are for focal trees of interest.
-Since certain explanatory variables such as basal area are cumulative, we must ensure that all trees being modeled are not biased to have different neighbor structures.
-This is a particular concern for trees at the boundary of study regions, which will not have the same number of neighbors as trees in the internal part of the study region.
+The next step in our analysis sequence is to add spatial information to our main `growth_scbi` data frame.
+The first element of spatial information we add is a "buffer region" to the periphery of the study region.
+Since some of our model's explanatory variables such as competitor basal area are cumulative, we must ensure that all trees being modeled are not biased to have different neighbor structures.
+This is of particular concern for trees at the boundary of study regions, which will not have the same number of neighbors that act as competitors as trees in the internal part of the study region.
+In order to account for such edge effects only trees who are not part of this buffer region, i.e.
+are part of the interior of the study region, will have their growths modeled @waller_applied_2004.
 
-Second, our ultimate method for model assessment will rely on estimates of model error as generated by cross-validation.
-Conventional cross-validation schemes assign observations to folds by resampling individual observations at random.
-However, underlying this scheme is an assumption that the observations are independent.
-In the case of forest census data, observations exhibit spatial autocorrelation, and thus this dependence must be incorporated in our resampling scheme in spatial cross-validation @roberts_cross-validation_2017 @pohjankukka_estimating_2017 We will therefore associate portions of the study region to spatial folds.
-
-To these two ends, we define two constants, both of which are in the same units as the `gx` and `gy` variables (most often meters).
+Our model of interspecific competition relies on a spatial definition of who the competitor trees are for focal trees of interest: all trees within a distance `comp_dist` of a focal tree are considered its competitors (assuming the same units as the `gx` and `gy` location variables).
+In our case we set this value below at 7.5m, a value informed by @canham_neighborhood_2004 @uriarte_spatially_2004 @canham_neighborhood_2006.
+Using this value along with a manually constructed `sf` object representation of the study region's boundary, we apply the `add_buffer_variable()` to our `growth_scbi` data frame to add a `buffer` boolean variable: all trees who have `buffer` set to `FALSE` will be our focal trees whose growths are modeled, whereas those with `buffer` set to `TRUE` will only be considered as competitor trees whose growth will not be modeled.
 
 ```{r}
+# Define buffer region using competitive distance range
 comp_dist <- 7.5
-cv_fold_size <- 100
-```
 
-The first constant is `comp_dist` which defines the maximum distance for a tree's competitive neighborhood.
-Trees within this distance of each other are assumed to compete while those farther than this distance apart do not.
-Put differently, all trees within `comp_dist` of a focal tree will be considered its competitors (see below).
-Other studies have estimated the value of `comp_dist`; we use an average of estimated values @canham_neighborhood_2004, @uriarte_spatially_2004, @tatsumi2013, @canham_neighborhood_2006.
-
-Furthermore, `comp_dist` will define the size of all buffers considered, which will be encoded as a binary variable `buffer` as computed by the `add_buffer_variable()` function.
-This function takes as input the main `growth_df` data frame, the `size` of the buffer which we set as `comp_dist`, and the boundary of the study region encoded as a simple features polygon @pebesma_simple_2018.
-DESCRIBE SF PACKAGE.
-In the Big Woods example below we will use a pre-loaded simple features polygon while for the SCBI example we present example code on how to manually construct one.
-
-The second constant is `cv_fold_size` which defines the length and width of the spatial folds (note that for now the spatial folds are restricted be squares).
-We will then use this constant to associate each observed tree to one of $k$ folds in the respective study region.
-In the Big Woods example below we will use the `blockCV` R package that has implemented spatial cross-validation while for the SCBI we will do this manually @valavi_blockcv_2019
-
-### Big Woods
-
-First, we indicate which trees are part of the buffer.
-This necessitates information about the study region boundary.
-In this case, we use a `sf_polygon` object `study_region_bw` which comes pre-loaded in the `forestecology` packages.
-After loading `study_region_bw`, we illustrate the results of the `add_buffer_variable()` function in Figure \ref{fig:bw-define-buffer}.
-Trees on the periphery denote with lighter colors are part of the buffer and will not be considered as "focal" trees of interest going forward; they will only be considered as competitor trees.
-
-```{r bw-define-buffer, out.width="100%", fig.cap="Buffer region for Big Woods study region."}
-data(study_region_bw)
-
-growth_bw <- growth_bw %>%
-  add_buffer_variable(direction = "in", size = comp_dist, region = study_region_bw)
-
-ggplot() +
-  geom_sf(data = growth_bw %>% sample_frac(0.2), aes(col = buffer), size = 0.5)
-```
-
-Second, we associate each tree to spatial cross validation folds. In this case, we use the `spatialBlock()` function from the `blockCV` package to define the spatial grid which
-
-THIS IS A MESS.
-We use the  @valavi_blockcv_2019, whose elements will act as the folds in our leave-one-out (by "one" we mean "one grid block") cross-validation scheme.
-The upshot here is we add `foldID` to `growth_df` which identifies which fold each individual is in, and the creation of a `cv_grid_sf` object which gives the geometry of the cross validation grid.
-
-```{r bw-define-cv-folds, out.width="100%", fig.cap="blockCV function output."}
-set.seed(76)
-bw_spatialBlock <- spatialBlock(
-  speciesData = growth_bw, theRange = cv_fold_size, k = 28, xOffset = 0.5, 
-  yOffset = 0, verbose = FALSE, showBlocks = FALSE
-)
-```
-
-Then add foldID to each tree
-
-```{r}
-growth_bw <- growth_bw %>%
-  mutate(foldID = bw_spatialBlock$foldID)
-```
-
-```{r bw-define-cv-folds-with-trees, out.width="100%", fig.cap="Inspect blocks closely."}
-# Visualize grid. Why does fold 19 repeat?
-ggplot() +
-  geom_sf(data = bw_spatialBlock$blocks %>% st_as_sf()) + 
-  geom_sf(data = growth_bw %>% sample_frac(0.2), 
-          aes(col = factor(foldID)), size = 0.1, show.legend = FALSE) + 
-  geom_sf_text(data = bw_spatialBlock$blocks %>% st_as_sf(), 
-               aes(label = folds))
-```
-
-Then remove empty folds
-
-```{r}
-growth_bw <- growth_bw %>%
-  filter(!foldID %in% c(19, 23, 21, 17, 8, 19)) %>%
-  mutate(foldID = factor(foldID))
-```
-
-Separately, we save the spatial cross-validation grid as an `sf_polygon` object `blocks_bw`
-
-```{r}
-blocks_bw <- bw_spatialBlock$blocks %>%
-  st_as_sf()
-```
-
-### SCBI
-
-First, we indicate which trees are part of the buffer.
-In this case however we manually define the study region boundary based on the subregion we defined in Section \ref{scbi-data} and create an `sf_polygon` object using the `sf_polygon()` function from the `sfheaders` package.
-Figure \ref{fig:scbi-define-buffer} displays the resulting buffer trees.
-
-```{r scbi-define-buffer, out.width="50%", fig.cap="Buffer region for SCBI study region."}
 study_region_scbi <- tibble(
@@
 growth_scbi <- growth_scbi %>%
-  add_buffer_variable(direction = "in", size = comp_dist, region = study_region_scbi)
-
-ggplot() +
-  geom_sf(data = growth_scbi, aes(col = buffer), size = 0.5)
+  add_buffer_variable(size = comp_dist, region = study_region_scbi)
 ```
 
-Second, we associate each tree to spatial cross validation folds.
-In this case we manually define a spatial crossvaliation grid.
-Figure \ref{fig:scbi-define-cv-folds} displays the resulting cross-validation folds along with the buffer from Figure \ref{fig:scbi-define-buffer}.
-
-Here we manually define the spatial cross-validation grid as an `sf_polygon` object `scbi_cv_grid`
+The second element of spatial information are blocks corresponding to folds of a spatial cross-validation algorithm used to estimate model error.
+Conventional cross-validation algorithms assign observations to folds by randomly resampling individual observations.
+However, underlying this algorithm is an assumption that the observations are independent.
+In the case of forest census data, observations exhibit spatial autocorrelation.
+This spatial dependence is incorporated into the cross-validation algorithm by randomly resampling spatial blocks of trees @roberts_cross-validation_2017 @pohjankukka_estimating_2017.
+We therefore associate each observed tree to one of $k$ spatial folds.
+In the example below, we first manually define two folds that partition the study region as an `sf` object.
+We then use the output of the `spatialBlock()` function from the `blockCV` package to associate each tree in `growth_scbi` to the correct fold (saved in the `foldID` variable) @valavi_blockcv_2019.
+\footnote{In the Appendix we present an example where the folds themselves are also created using the `spatialBlock()` function given a specified `cv_block_size`.}
 
 ```{r}
+# Manually define spatial blocks to act as folds
 fold1 <- rbind(c(0, 300), c(150, 300), c(150, 600), c(0, 600))
@@
 
-blocks_scbi <- bind_rows(
-  sf_polygon(fold1),
-  sf_polygon(fold2)
-) %>%
+blocks_scbi <- bind_rows(sf_polygon(fold1), sf_polygon(fold2)) %>%
   mutate(folds = c(1, 2) %>% factor())
+
+# Associate each observation to a fold
+SpatialBlock_scbi <- spatialBlock(
+  speciesData = growth_scbi, k = 2, selection = "systematic", 
+  blocks = blocks_scbi, showBlocks = FALSE, verbose = FALSE
+)
+
+growth_scbi <- growth_scbi %>%
+  mutate(foldID = SpatialBlock_scbi$foldID %>% factor())
 ```
 
-```{r scbi-define-cv-folds, out.width="50%", fig.cap="Buffer region for SCBI study region."}
-SpatialBlock_scbi <- spatialBlock(
-  speciesData = growth_scbi, k = 2, selection = "systematic", blocks = blocks_scbi,
-  showBlocks = FALSE, verbose = FALSE
-)
+Figure \ref{fig:scbi-spatial-information} illustrates the net effect of adding these two elements of information to the `growth_scbi` data frame.
+The location of each tree is marked with an integer indicating which fold it belongs to, where the folds are marked with solid lines.
+The color of each digit indicates whether the tree is part of the buffer region (and thus will only be considered as a competitor tree in our model) or is part of the interior of the study region (and thus is a focal tree whose growth is of modeled interest).
 
-# Add foldID to each tree
-growth_scbi <- growth_scbi %>%
-  mutate(foldID = SpatialBlock_scbi$foldID %>% factor())
-
+```{r scbi-spatial-information, out.width="66%", fig.cap="Buffer region and cross-validation block information for SCBI data."}
 ggplot() +
-  geom_sf_text(data = growth_scbi, aes(label = foldID, col = buffer)) +
+  geom_sf_text(data = growth_scbi, aes(label = foldID, col = buffer), alpha = 0.2) +
   geom_sf(data = blocks_scbi, fill = "transparent")
@@
 
+TODO: Do we talk about `add_buffer_variable(direction = "out")` when each fold gets its turn being the training data?
+
 ## Define focal versus competitor trees
+
+HEY BERT PICK IT UP HERE
 

### 2021-02-15 etc/paper/forestecology.tex
Done "add spatial information" section

@@
 \item
+  Visualize model results.
+\item
   Additionally: Evaluate model performance using spatial
@@
 \KeywordTok{library}\NormalTok{(forestecology)}
+\KeywordTok{library}\NormalTok{(blockCV)}
 \end{Highlighting}
@@
 
-We now encode spatial information to the \texttt{growth\_df} data
-frames. First, in order to control for study region edge effects, we add
-``buffers'' to the periphery of the study region (cite Waller?). Our
-model of interspecific competition relies on a spatial definition of who
-the competitor trees are for focal trees of interest. Since certain
-explanatory variables such as basal area are cumulative, we must ensure
-that all trees being modeled are not biased to have different neighbor
-structures. This is a particular concern for trees at the boundary of
-study regions, which will not have the same number of neighbors as trees
-in the internal part of the study region.
-
-Second, our ultimate method for model assessment will rely on estimates
-of model error as generated by cross-validation. Conventional
-cross-validation schemes assign observations to folds by resampling
-individual observations at random. However, underlying this scheme is an
-assumption that the observations are independent. In the case of forest
-census data, observations exhibit spatial autocorrelation, and thus this
-dependence must be incorporated in our resampling scheme in spatial
-cross-validation \citet{roberts_cross-validation_2017}
-\citet{pohjankukka_estimating_2017} We will therefore associate portions
-of the study region to spatial folds.
-
-To these two ends, we define two constants, both of which are in the
-same units as the \texttt{gx} and \texttt{gy} variables (most often
-meters).
-
-\begin{Shaded}
-\begin{Highlighting}[]
+The next step in our analysis sequence is to add spatial information to
+our main \texttt{growth\_scbi} data frame. The first element of spatial
+information we add is a ``buffer region'' to the periphery of the study
+region. Since some of our model's explanatory variables such as
+competitor basal area are cumulative, we must ensure that all trees
+being modeled are not biased to have different neighbor structures. This
+is of particular concern for trees at the boundary of study regions,
+which will not have the same number of neighbors that act as competitors
+as trees in the internal part of the study region. In order to account
+for such edge effects only trees who are not part of this buffer region,
+i.e.~are part of the interior of the study region, will have their
+growths modeled \citet{waller_applied_2004}.
+
+Our model of interspecific competition relies on a spatial definition of
+who the competitor trees are for focal trees of interest: all trees
+within a distance \texttt{comp\_dist} of a focal tree are considered its
+competitors (assuming the same units as the \texttt{gx} and \texttt{gy}
+location variables). In our case we set this value below at 7.5m, a
+value informed by \citet{canham_neighborhood_2004}
+\citet{uriarte_spatially_2004} \citet{canham_neighborhood_2006}. Using
+this value along with a manually constructed \texttt{sf} object
+representation of the study region's boundary, we apply the
+\texttt{add\_buffer\_variable()} to our \texttt{growth\_scbi} data frame
+to add a \texttt{buffer} boolean variable: all trees who have
+\texttt{buffer} set to \texttt{FALSE} will be our focal trees whose
+growths are modeled, whereas those with \texttt{buffer} set to
+\texttt{TRUE} will only be considered as competitor trees whose growth
+will not be modeled.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{# Define buffer region using competitive distance range}
 \NormalTok{comp_dist <-}\StringTok{ }\FloatTok{7.5}
-\NormalTok{cv_fold_size <-}\StringTok{ }\DecValTok{100}
-\end{Highlighting}
-\end{Shaded}
-
-The first constant is \texttt{comp\_dist} which defines the maximum
-distance for a tree's competitive neighborhood. Trees within this
-distance of each other are assumed to compete while those farther than
-this distance apart do not. Put differently, all trees within
-\texttt{comp\_dist} of a focal tree will be considered its competitors
-(see below). Other studies have estimated the value of
-\texttt{comp\_dist}; we use an average of estimated values
-\citet{canham_neighborhood_2004}, \citet{uriarte_spatially_2004},
-\citet{tatsumi2013}, \citet{canham_neighborhood_2006}.
-
-Furthermore, \texttt{comp\_dist} will define the size of all buffers
-considered, which will be encoded as a binary variable \texttt{buffer}
-as computed by the \texttt{add\_buffer\_variable()} function. This
-function takes as input the main \texttt{growth\_df} data frame, the
-\texttt{size} of the buffer which we set as \texttt{comp\_dist}, and the
-boundary of the study region encoded as a simple features polygon
-\citet{pebesma_simple_2018}. DESCRIBE SF PACKAGE. In the Big Woods
-example below we will use a pre-loaded simple features polygon while for
-the SCBI example we present example code on how to manually construct
-one.
-
-The second constant is \texttt{cv\_fold\_size} which defines the length
-and width of the spatial folds (note that for now the spatial folds are
-restricted be squares). We will then use this constant to associate each
-observed tree to one of \(k\) folds in the respective study region. In
-the Big Woods example below we will use the \texttt{blockCV} R package
-that has implemented spatial cross-validation while for the SCBI we will
-do this manually \citet{valavi_blockcv_2019}
-
-\hypertarget{big-woods}{%
-\subsubsection{Big Woods}\label{big-woods}}
-
-First, we indicate which trees are part of the buffer. This necessitates
-information about the study region boundary. In this case, we use a
-\texttt{sf\_polygon} object \texttt{study\_region\_bw} which comes
-pre-loaded in the \texttt{forestecology} packages. After loading
-\texttt{study\_region\_bw}, we illustrate the results of the
-\texttt{add\_buffer\_variable()} function in Figure
-\ref{fig:bw-define-buffer}. Trees on the periphery denote with lighter
-colors are part of the buffer and will not be considered as ``focal''
-trees of interest going forward; they will only be considered as
-competitor trees.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{data}\NormalTok{(study_region_bw)}
-
-\NormalTok{growth_bw <-}\StringTok{ }\NormalTok{growth_bw }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{direction =} \StringTok{"in"}\NormalTok{, }\DataTypeTok{size =}\NormalTok{ comp_dist, }\DataTypeTok{region =}\NormalTok{ study_region_bw)}
-
-\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_bw }\OperatorTok{%>%}\StringTok{ }\KeywordTok{sample_frac}\NormalTok{(}\FloatTok{0.2}\NormalTok{), }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{col =}\NormalTok{ buffer), }\DataTypeTok{size =} \FloatTok{0.5}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{Figures/bw-define-buffer-1}
-
-}
-
-\caption{Buffer region for Big Woods study region.}\label{fig:bw-define-buffer}
-\end{figure}
-
-Second, we associate each tree to spatial cross validation folds. In
-this case, we use the \texttt{spatialBlock()} function from the
-\texttt{blockCV} package to define the spatial grid which
-
-THIS IS A MESS. We use the \citet{valavi_blockcv_2019}, whose elements
-will act as the folds in our leave-one-out (by ``one'' we mean ``one
-grid block'') cross-validation scheme. The upshot here is we add
-\texttt{foldID} to \texttt{growth\_df} which identifies which fold each
-individual is in, and the creation of a \texttt{cv\_grid\_sf} object
-which gives the geometry of the cross validation grid.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{set.seed}\NormalTok{(}\DecValTok{76}\NormalTok{)}
-\NormalTok{bw_spatialBlock <-}\StringTok{ }\KeywordTok{spatialBlock}\NormalTok{(}
-  \DataTypeTok{speciesData =}\NormalTok{ growth_bw, }\DataTypeTok{theRange =}\NormalTok{ cv_fold_size, }\DataTypeTok{k =} \DecValTok{28}\NormalTok{, }\DataTypeTok{xOffset =} \FloatTok{0.5}\NormalTok{, }
-  \DataTypeTok{yOffset =} \DecValTok{0}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{showBlocks =} \OtherTok{FALSE}
-\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-Then add foldID to each tree
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{growth_bw <-}\StringTok{ }\NormalTok{growth_bw }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ bw_spatialBlock}\OperatorTok{$}\NormalTok{foldID)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{# Visualize grid. Why does fold 19 repeat?}
-\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ bw_spatialBlock}\OperatorTok{$}\NormalTok{blocks }\OperatorTok{%>%}\StringTok{ }\KeywordTok{st_as_sf}\NormalTok{()) }\OperatorTok{+}\StringTok{ }
-\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_bw }\OperatorTok{%>%}\StringTok{ }\KeywordTok{sample_frac}\NormalTok{(}\FloatTok{0.2}\NormalTok{), }
-          \KeywordTok{aes}\NormalTok{(}\DataTypeTok{col =} \KeywordTok{factor}\NormalTok{(foldID)), }\DataTypeTok{size =} \FloatTok{0.1}\NormalTok{, }\DataTypeTok{show.legend =} \OtherTok{FALSE}\NormalTok{) }\OperatorTok{+}\StringTok{ }
-\StringTok{  }\KeywordTok{geom_sf_text}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ bw_spatialBlock}\OperatorTok{$}\NormalTok{blocks }\OperatorTok{%>%}\StringTok{ }\KeywordTok{st_as_sf}\NormalTok{(), }
-               \KeywordTok{aes}\NormalTok{(}\DataTypeTok{label =}\NormalTok{ folds))}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{Figures/bw-define-cv-folds-with-trees-1}
-
-}
-
-\caption{Inspect blocks closely.}\label{fig:bw-define-cv-folds-with-trees}
-\end{figure}
-
-Then remove empty folds
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{growth_bw <-}\StringTok{ }\NormalTok{growth_bw }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\NormalTok{foldID }\OperatorTok{%in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{19}\NormalTok{, }\DecValTok{23}\NormalTok{, }\DecValTok{21}\NormalTok{, }\DecValTok{17}\NormalTok{, }\DecValTok{8}\NormalTok{, }\DecValTok{19}\NormalTok{)) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =} \KeywordTok{factor}\NormalTok{(foldID))}
-\end{Highlighting}
-\end{Shaded}
-
-Separately, we save the spatial cross-validation grid as an
-\texttt{sf\_polygon} object \texttt{blocks\_bw}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{blocks_bw <-}\StringTok{ }\NormalTok{bw_spatialBlock}\OperatorTok{$}\NormalTok{blocks }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{st_as_sf}\NormalTok{()}
-\end{Highlighting}
-\end{Shaded}
-
-\hypertarget{scbi}{%
-\subsubsection{SCBI}\label{scbi}}
-
-First, we indicate which trees are part of the buffer. In this case
-however we manually define the study region boundary based on the
-subregion we defined in Section \ref{scbi-data} and create an
-\texttt{sf\_polygon} object using the \texttt{sf\_polygon()} function
-from the \texttt{sfheaders} package. Figure \ref{fig:scbi-define-buffer}
-displays the resulting buffer trees.
-
-\begin{Shaded}
-\begin{Highlighting}[]
+
 \NormalTok{study_region_scbi <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}
@@
 \NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{direction =} \StringTok{"in"}\NormalTok{, }\DataTypeTok{size =}\NormalTok{ comp_dist, }\DataTypeTok{region =}\NormalTok{ study_region_scbi)}
-
-\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_scbi, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{col =}\NormalTok{ buffer), }\DataTypeTok{size =} \FloatTok{0.5}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.5\linewidth]{Figures/scbi-define-buffer-1}
-
-}
-
-\caption{Buffer region for SCBI study region.}\label{fig:scbi-define-buffer}
-\end{figure}
-
-Second, we associate each tree to spatial cross validation folds. In
-this case we manually define a spatial crossvaliation grid. Figure
-\ref{fig:scbi-define-cv-folds} displays the resulting cross-validation
-folds along with the buffer from Figure \ref{fig:scbi-define-buffer}.
-
-Here we manually define the spatial cross-validation grid as an
-\texttt{sf\_polygon} object \texttt{scbi\_cv\_grid}
-
-\begin{Shaded}
-\begin{Highlighting}[]
+\StringTok{  }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{size =}\NormalTok{ comp_dist, }\DataTypeTok{region =}\NormalTok{ study_region_scbi)}
+\end{Highlighting}
+\end{Shaded}
+
+The second element of spatial information are blocks corresponding to
+folds of a spatial cross-validation algorithm used to estimate model
+error. Conventional cross-validation algorithms assign observations to
+folds by randomly resampling individual observations. However,
+underlying this algorithm is an assumption that the observations are
+independent. In the case of forest census data, observations exhibit
+spatial autocorrelation. This spatial dependence is incorporated into
+the cross-validation algorithm by randomly resampling spatial blocks of
+trees \citet{roberts_cross-validation_2017}
+\citet{pohjankukka_estimating_2017}. We therefore associate each
+observed tree to one of \(k\) spatial folds. In the example below, we
+first manually define two folds that partition the study region as an
+\texttt{sf} object. We then use the output of the
+\texttt{spatialBlock()} function from the \texttt{blockCV} package to
+associate each tree in \texttt{growth\_scbi} to the correct fold (saved
+in the \texttt{foldID} variable) \citet{valavi_blockcv_2019}.
+\textbackslash footnote\{In the Appendix we present an example where the
+folds themselves are also created using the \texttt{spatialBlock()}
+function given a specified \texttt{cv\_block\_size}.\}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{# Manually define spatial blocks to act as folds}
 \NormalTok{fold1 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{600}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{600}\NormalTok{))}
@@
 \StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
+
+\CommentTok{# Associate each observation to a fold}
 \NormalTok{SpatialBlock_scbi <-}\StringTok{ }\KeywordTok{spatialBlock}\NormalTok{(}
-  \DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =} \DecValTok{2}\NormalTok{, }\DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi,}
-  \DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}
+  \DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =} \DecValTok{2}\NormalTok{, }\DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{, }
+  \DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}
 \NormalTok{)}
 
-\CommentTok{# Add foldID to each tree}
 \NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
 \StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ SpatialBlock_scbi}\OperatorTok{$}\NormalTok{foldID }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
-
+\end{Highlighting}
+\end{Shaded}
+
+Figure \ref{fig:scbi-spatial-information} illustrates the net effect of
+adding these two elements of information to the \texttt{growth\_scbi}
+data frame. The location of each tree is marked with an integer
+indicating which fold it belongs to, where the folds are marked with
+solid lines. The color of each digit indicates whether the tree is part
+of the buffer region (and thus will only be considered as a competitor
+tree in our model) or is part of the interior of the study region (and
+thus is a focal tree whose growth is of modeled interest).
+
+\begin{Shaded}
+\begin{Highlighting}[]
 \KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_sf_text}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_scbi, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{label =}\NormalTok{ foldID, }\DataTypeTok{col =}\NormalTok{ buffer)) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{geom_sf_text}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_scbi, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{label =}\NormalTok{ foldID, }\DataTypeTok{col =}\NormalTok{ buffer), }\DataTypeTok{alpha =} \FloatTok{0.2}\NormalTok{) }\OperatorTok{+}
 \StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ blocks_scbi, }\DataTypeTok{fill =} \StringTok{"transparent"}\NormalTok{)}
@@
 
-{\centering \includegraphics[width=0.5\linewidth]{Figures/scbi-define-cv-folds-1}
+{\centering \includegraphics[width=0.66\linewidth]{Figures/scbi-spatial-information-1}
 
@@
 
-\caption{Buffer region for SCBI study region.}\label{fig:scbi-define-cv-folds}
+\caption{Buffer region and cross-validation block information for SCBI data.}\label{fig:scbi-spatial-information}
 \end{figure}
+
+TODO: Do we talk about
+\texttt{add\_buffer\_variable(direction\ =\ "out")} when each fold gets
+its turn being the training data?
 
@@
 trees}\label{define-focal-versus-competitor-trees}}
+
+HEY BERT PICK IT UP HERE
 
@@
 
-\hypertarget{big-woods-1}{%
-\subsubsection{Big Woods}\label{big-woods-1}}
+\hypertarget{big-woods}{%
+\subsubsection{Big Woods}\label{big-woods}}
 
@@
 
-\hypertarget{big-woods-2}{%
-\subsubsection{Big Woods}\label{big-woods-2}}
+\hypertarget{big-woods-1}{%
+\subsubsection{Big Woods}\label{big-woods-1}}
 
@@
 \end{figure}
+
+\hypertarget{viz-posterior-distributions}{%
+\subsection{Visualize model results}\label{viz-posterior-distributions}}
+
+Lastly, we return to the model fits from Section \ref{model-fit-predict}
+and present tools to visually explore the posterior distributions of all
+parameters in our model. There are two main groups of parameters to
+consider. The \(\beta\) coefficients tell us about how fast each species
+grows and how this depends on DBH while the full matrix of \(\lambda\)
+values describe the competitive effects between pairs of species. There
+is a rich literature on this matrix (cite).
+
+DO WE NEED TO DESCRIBE MECHANICS? Because of the structure of the
+\texttt{bw\_fit\_model} object we cannot simply draw these curves based
+on the posterior distribution. \texttt{bw\_fit\_model()} gives the
+parameters \emph{compared} to a baseline. This is not of direct
+interest. So to display these parameters, as we care about them, we have
+to sample from the baseline distribution and from the comparison one to
+get the posterior distribution of interest.
+
+\hypertarget{big-woods-2}{%
+\subsubsection{Big Woods}\label{big-woods-2}}
+
+Here we re-run the model fit to the Big Woods data from Section
+\ref{model-fit-predict}, but this time use ``family'' as the group for
+comparison which has. This makes the posterior distributions easier to
+follow. Also, surprisingly, grouping by family performed just as well as
+grouping by species \citet{allen_permutation_2020}. First we re-run
+\texttt{create\_focal\_vs\_comp()} and \texttt{fit\_bayesian\_model()}
+with no permutation shuffling with the grouping variable as family.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{focal_vs_comp_bw <-}\StringTok{ }\NormalTok{growth_bw }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =}\NormalTok{ family }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{()) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{cv_grid_sf =}\NormalTok{ blocks_bw, }\DataTypeTok{id =} \StringTok{"treeID"}\NormalTok{)}
+
+\NormalTok{comp_bayes_lm_bw <-}\StringTok{ }\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{comp_bayes_lm}\NormalTok{(}\DataTypeTok{prior_param =} \OtherTok{NULL}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+Now the posterior parameter outputs of \texttt{fit\_bayesian\_model()}
+are passed to \texttt{plot\_bayesian\_model\_parameters()} to generate
+visualizations of the posterior parameters. These visualizations are
+displayed in Figure 5 of \citet{allen_permutation_2020}. For simplicity
+we only plot a subset of the species families.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{sp_to_plot <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"cornaceae"}\NormalTok{, }\StringTok{"fagaceae"}\NormalTok{, }\StringTok{"hamamelidaceae"}\NormalTok{, }\StringTok{"juglandaceae"}\NormalTok{, }
+                 \StringTok{"lauraceae"}\NormalTok{, }\StringTok{"rosaceae"}\NormalTok{, }\StringTok{"sapindaceae"}\NormalTok{, }\StringTok{"ulmaceae"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+The output is a list with three plots stored. Figure
+\ref{fig:bw-posterior-viz-beta0} The element \texttt{beta\_0} gives the
+baseline growth intercept \(\beta_0\), i.e., how fast an individual of
+each group grows independent of DBH).
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{plot1 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_bw, }\DataTypeTok{type =} \StringTok{"intercepts"}\NormalTok{)}
+\NormalTok{plot1}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{Figures/bw-posterior-viz-beta0-1}
+
+}
+
+\caption{Posterior distribution of beta0.}\label{fig:bw-posterior-viz-beta0}
+\end{figure}
+
+Figure \ref{fig:bw-posterior-viz-beta-dbh} Next \texttt{beta\_dbh} gives
+the slope for DBH slope \(\beta_{dbh,i}\) for each group.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{plot2 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_bw, }\DataTypeTok{type =} \StringTok{"dbh_slopes"}\NormalTok{)}
+\NormalTok{plot2}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{Figures/bw-posterior-viz-beta-dbh-1}
+
+}
+
+\caption{Posterior distribution of betadbh.}\label{fig:bw-posterior-viz-beta-dbh}
+\end{figure}
+
+Finally Figure \ref{fig:bw-posterior-viz-lambda} \texttt{lambda} gives
+the competition coefficients \(\lambda\).
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{plot3 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_bw, }\DataTypeTok{type =} \StringTok{"competition"}\NormalTok{)}
+\NormalTok{plot3}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{Figures/bw-posterior-viz-lambda-1}
+
+}
+
+\caption{Posterior distribution of lambda's for Big Woods.}\label{fig:bw-posterior-viz-lambda}
+\end{figure}
+
+\hypertarget{scbi-2}{%
+\subsubsection{SCBI}\label{scbi-2}}
+
+We revisit the posterior parameters for the SCBI from Section
+\{model-fit-predict\}, but this time only focus on the \(\lambda\)
+competition coefficients.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{sp_to_plot <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"quru"}\NormalTok{, }\StringTok{"litu"}\NormalTok{, }\StringTok{"cagl"}\NormalTok{, }\StringTok{"cato"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{plot3 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_bw, }\DataTypeTok{type =} \StringTok{"competition"}\NormalTok{)}
+\NormalTok{plot3}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{Figures/scbi-posterior-viz-lambda-1}
+
+}
+
+\caption{Posterior distribution of lambda's for SCBI.}\label{fig:scbi-posterior-viz-lambda}
+\end{figure}
+
+Add explanation here.
+
+HEY BERT PICK IT UP HERE
 
@@
 
-\hypertarget{viz-posterior-distributions}{%
-\subsection{Visualize posterior
-distributions}\label{viz-posterior-distributions}}
-
-Lastly, we return to the model fits from Section \ref{model-fit-predict}
-and present tools to visually explore the posterior distributions of all
-parameters in our model. There are two main groups of parameters to
-consider. The \(\beta\) coefficients tell us about how fast each species
-grows and how this depends on DBH while the full matrix of \(\lambda\)
-values describe the competitive effects between pairs of species. There
-is a rich literature on this matrix (cite).
-
-DO WE NEED TO DESCRIBE MECHANICS? Because of the structure of the
-\texttt{bw\_fit\_model} object we cannot simply draw these curves based
-on the posterior distribution. \texttt{bw\_fit\_model()} gives the
-parameters \emph{compared} to a baseline. This is not of direct
-interest. So to display these parameters, as we care about them, we have
-to sample from the baseline distribution and from the comparison one to
-get the posterior distribution of interest.
-
-\hypertarget{big-woods-4}{%
-\subsubsection{Big Woods}\label{big-woods-4}}
-
-Here we re-run the model fit to the Big Woods data from Section
-\ref{model-fit-predict}, but this time use ``family'' as the group for
-comparison which has. This makes the posterior distributions easier to
-follow. Also, surprisingly, grouping by family performed just as well as
-grouping by species \citet{allen_permutation_2020}. First we re-run
-\texttt{create\_focal\_vs\_comp()} and \texttt{fit\_bayesian\_model()}
-with no permutation shuffling with the grouping variable as family.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{focal_vs_comp_bw <-}\StringTok{ }\NormalTok{growth_bw }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =}\NormalTok{ family }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{()) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{cv_grid_sf =}\NormalTok{ blocks_bw, }\DataTypeTok{id =} \StringTok{"treeID"}\NormalTok{)}
-
-\NormalTok{comp_bayes_lm_bw <-}\StringTok{ }\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{comp_bayes_lm}\NormalTok{(}\DataTypeTok{prior_param =} \OtherTok{NULL}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-Now the posterior parameter outputs of \texttt{fit\_bayesian\_model()}
-are passed to \texttt{plot\_bayesian\_model\_parameters()} to generate
-visualizations of the posterior parameters. These visualizations are
-displayed in Figure 5 of \citet{allen_permutation_2020}. For simplicity
-we only plot a subset of the species families.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{sp_to_plot <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"cornaceae"}\NormalTok{, }\StringTok{"fagaceae"}\NormalTok{, }\StringTok{"hamamelidaceae"}\NormalTok{, }\StringTok{"juglandaceae"}\NormalTok{, }
-                 \StringTok{"lauraceae"}\NormalTok{, }\StringTok{"rosaceae"}\NormalTok{, }\StringTok{"sapindaceae"}\NormalTok{, }\StringTok{"ulmaceae"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-The output is a list with three plots stored. Figure
-\ref{fig:bw-posterior-viz-beta0} The element \texttt{beta\_0} gives the
-baseline growth intercept \(\beta_0\), i.e., how fast an individual of
-each group grows independent of DBH).
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{plot1 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_bw, }\DataTypeTok{type =} \StringTok{"intercepts"}\NormalTok{)}
-\NormalTok{plot1}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{Figures/bw-posterior-viz-beta0-1}
-
-}
-
-\caption{Posterior distribution of beta0.}\label{fig:bw-posterior-viz-beta0}
-\end{figure}
-
-Figure \ref{fig:bw-posterior-viz-beta-dbh} Next \texttt{beta\_dbh} gives
-the slope for DBH slope \(\beta_{dbh,i}\) for each group.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{plot2 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_bw, }\DataTypeTok{type =} \StringTok{"dbh_slopes"}\NormalTok{)}
-\NormalTok{plot2}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{Figures/bw-posterior-viz-beta-dbh-1}
-
-}
-
-\caption{Posterior distribution of betadbh.}\label{fig:bw-posterior-viz-beta-dbh}
-\end{figure}
-
-Finally Figure \ref{fig:bw-posterior-viz-lambda} \texttt{lambda} gives
-the competition coefficients \(\lambda\).
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{plot3 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_bw, }\DataTypeTok{type =} \StringTok{"competition"}\NormalTok{)}
-\NormalTok{plot3}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{Figures/bw-posterior-viz-lambda-1}
-
-}
-
-\caption{Posterior distribution of lambda's for Big Woods.}\label{fig:bw-posterior-viz-lambda}
-\end{figure}
-
-\hypertarget{scbi-4}{%
-\subsubsection{SCBI}\label{scbi-4}}
-
-We revisit the posterior parameters for the SCBI from Section
-\{model-fit-predict\}, but this time only focus on the \(\lambda\)
-competition coefficients.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{sp_to_plot <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"quru"}\NormalTok{, }\StringTok{"litu"}\NormalTok{, }\StringTok{"cagl"}\NormalTok{, }\StringTok{"cato"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{plot3 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_bw, }\DataTypeTok{type =} \StringTok{"competition"}\NormalTok{)}
-\NormalTok{plot3}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{Figures/scbi-posterior-viz-lambda-1}
-
-}
-
-\caption{Posterior distribution of lambda's for SCBI.}\label{fig:scbi-posterior-viz-lambda}
-\end{figure}
-
-Add explanation here.
-
-HEY BERT PICK IT UP HERE
-
 \hypertarget{discussion}{%

### 2021-02-15 etc/PLOSOne/replicate_plosone.R

@@ -27,7 +27,7 @@ growth_bw <-
 
 # Add spatial information -----------------------------------------------------
 comp_dist <- 7.5
-cv_fold_size <- 100
+cv_block_size <- 100
 data(study_region_bw)
 
 ## Add buffer ----
@@ -41,7 +41,7 @@ ggplot() +
 ## Add spatial folds ----
 set.seed(76)
 bw_spatialBlock <- spatialBlock(
-  speciesData = growth_bw, theRange = cv_fold_size, k = 28, xOffset = 0.5,
+  speciesData = growth_bw, theRange = cv_block_size, k = 28, xOffset = 0.5,
   yOffset = 0, verbose = FALSE, showBlocks = FALSE
 )
 growth_bw <- growth_bw %>%

### 2021-02-16 etc/paper/forestecology.Rmd
Done "Identify all focal and corresponding competitor trees" section

@@
 library(here)
+library(tibble)
 
@@
   collapse = TRUE,
-  comment = "##"
-  # tidy = FALSE
+  comment = "##",
+  # tidy = FALSE,
+  tibble.print_max = 5,
+  tibble.print_min = 5
 )
@@
 
+## Model
+
+Describe model specifics.
+
 # Example
@@
     1. Add spatial cross-validation block information.
-1. Compute focal versus competitor tree information.
+1. Identify all focal and corresponding competitor trees.
 1. Fit model and make predictions.
-1. Visualize model results.
 1. Additionally: Evaluate model performance using spatial cross-validation.
@@
 Despite such standards, minor variations will still exist between sites thereby necessitating some data wrangling and checking.
-For example, the SCBI site records all dbh's in millimeters, whereas the Michigan Big Woods site records them in centimeters @andersonteixeira_ctfs-forestgeo_2015 @allen_michigan_2020.
+For example, the SCBI site records all DBH's in millimeters, whereas the Michigan Big Woods site records them in centimeters @andersonteixeira_ctfs-forestgeo_2015 @allen_michigan_2020.
 The data format of other sites may be such that our `compute_growth()` function doesn't work at all.
-However, in the end all that matters is that the growth of all trees is saved in a data frame of class `sf` whereby the geolocation of each tree is presented in a `geometry` variable of type `<POINT>` and at a minimum the data contains the following variables: a variable uniquely identifying each tree-stem, `sp` of type `fct` factor identifying species, `dbh1` and `dbh2` of type `dbl` quantifying the dbh at earlier and later census, and `growth` of type `dbl` double quantifying the average annual growth in centimeters.
+However, in the end all that matters is that the growth of all trees is saved in a data frame of class `sf` whereby the geolocation of each tree is presented in a `geometry` variable of type `<POINT>` and at a minimum the data contains the following variables: a variable uniquely identifying each tree-stem, `sp` of type `fct` factor identifying species, `dbh1` and `dbh2` of type `dbl` quantifying the DBH at earlier and later census, and `growth` of type `dbl` double quantifying the average annual growth in centimeters.
 
 We load both 2008 and 2014 SCBI census data `.csv` files as they existed on GitHub on November 20, 2020.
-After selecting only relevant variables, we perform a few additional data wrangling steps: convert the variable with the date of measurement to be of type `date`, convert dbh to be in centimeters^[A rule of thumb to determine the units of dbh is check if the smallest non-zero and non-missing measurement is 1 or 10.
+After selecting only relevant variables, we perform a few additional data wrangling steps: convert the variable with the date of measurement to be of type `date`, convert DBH to be in centimeters^[A rule of thumb to determine the units of DBH is check if the smallest non-zero and non-missing measurement is 1 or 10.
 If the former, then centimeters.
 If the later, then millimeters.
-This is because ForestGEO protocols state that only trees with dbh greater or equal to 1cm should be included in censuses.], convert the `sp` variable containing species information from type `chr` character to `fct` factor (we will discuss the need for this in Section \ref{spatial-cross-validation}).
-Furthermore, in order to speed up computation for purposes of this examplt, we only consider a 9 ha subsection of the 25.6 ha of the SCBI site: `gx` from 0--300 instead of 0--400 and `gy` from 300--600 instead of 0--640.
+This is because ForestGEO protocols state that only trees with DBH greater or equal to 1cm should be included in censuses.], convert the `sp` variable containing species information from type `chr` character to `fct` factor (we will discuss the need for this in Section \ref{spatial-cross-validation}).
+Furthermore, in order to speed up computation for purposes of this example, we only consider a 9 ha subsection of the 25.6 ha of the SCBI site: `gx` from 0--300 instead of 0--400 and `gy` from 300--600 instead of 0--640.
 
@@
 
-The output `growth_scbi` is a single data frame of class `sf` that includes a numerical `growth` reflecting the average annual growth in dbh (in cm) for all trees that were alive at both time points as well a `geometry` variable encoding each tree's geolocation.
+The output `growth_scbi` is a single data frame of class `sf` that includes a numerical `growth` reflecting the average annual growth in DBH (in cm) for all trees that were alive at both time points as well a `geometry` variable encoding each tree's geolocation.
 Furthermore, variables that (in theory) remain unchanged between censuses appear only once, such as location variables `gx` and `gy`; as well as species-related variables.
@@
 ggplot() +
-  geom_sf_text(data = growth_scbi, aes(label = foldID, col = buffer), alpha = 0.2) +
+  geom_sf_text(data = growth_scbi, aes(label = foldID, col = buffer), 
+               alpha = 0.2) +
   geom_sf(data = blocks_scbi, fill = "transparent")
@@
 
-## Define focal versus competitor trees
+## Identify all focal and corresponding competitor trees
 
-HEY BERT PICK IT UP HERE
-
-Next we define `focal_vs_comp` data frames which connects each focal tree in the `growth_df` data frames to the trees in its competitive neighborhood range as defined by the `comp_dist` constant.
-So for example, if `growth_df` consisted of two focal trees with two and three neighbors with `comp_dist` respectively, `focal_vs_comp` would be a data frame of 5 rows connecting each focal tree to it's competitors.
-The `create_focal_vs_comp()` function makes this connection taking as inputs the `growth_df` data frame; the `comp_dist` constant defining competitive range; `cv_grid_sf`, giving the cross validation grid; and the `id` variable.
-
-### Big Woods
-
-```{r bw-focal-vs-comp}
-focal_vs_comp_bw <- growth_bw %>%
-  create_focal_vs_comp(comp_dist, cv_grid_sf = blocks_bw, id = "treeID")
-```
-
-TODO: Figure out how to show this data frame's contents.
-
-### SCBI
+The next step in our analysis sequence is to identify all focal trees and their corresponding competitor trees.
+The `create_focal_vs_comp()` functions performs these tasks and returns a new data frame of type `sf` containing this information.
+On top of the previously discussed arguments `comp_dist` defining the competition neighborhood and `id` indicating which variable in the data frame uniquely identifies each tree-stem, this function also requires an `sf` object representation of the spatial cross-validation blocks/folds; in our case, this was manually encoded in the `blocks_scbi` in Section \ref{spatial-information} while in our Appendix we present an example where this was performed using `spatialBlock()` from the `blockCV` package.
+We present the resulting data frame below with the `foldID` variable omitted for compactness of presentation.
 
@@
   create_focal_vs_comp(comp_dist, cv_grid_sf = blocks_scbi, id = "stemID")
+focal_vs_comp_scbi %>% 
+  select(-foldID)
 ```
 
-TODO: Figure out how to show this data frame's contents.
+TODO: Below reconcile the number of rows as they off by one from `growth_scbi` pipe `filter(!is.na(growth) & !buffer)`. perhaps by removing NA's in the `growth_scbi` stage.
+
+The resulting data frame `focal_vs_comp_scbi` has `r nrow(focal_vs_comp_scbi)` rows, representing the subset of the `r nrow(growth_scbi)` trees in `growth_scbi` that will be considered as focal trees and thus have their growths modeled.
+Recall from Section \ref{spatial-information} this consists all trees that are not part of the buffer region in Figure \ref{fig:scbi-spatial-information}.
+Two new variables `focal_ID` and `focal_sp` related to tree-stem identification and species information.
+Most notably however is a new variable `comp` which contains information on all competitor trees for a given focal tree saved in list-column format, a feature of the `tidyr` package @tidyr_package.
+For example, we drill-down on the tree with `focal_ID` equal to 4.
+It has 20 competitor trees each described by 4 variables as indicated by the fact that `comp` is a `<tibble [20 × 4]>`.
+
+```{r scbi-focal-vs-comp-2}
+focal_vs_comp_scbi %>% 
+  filter(focal_ID == 4) %>% 
+  select(focal_ID, dbh, comp)
+```
+
+Using the `unnest()` function from the `tidyr` package, we can flatten list-column into regular columns.
+We observe that for the same focal tree with DBH equal to 13.65cm, we have information on all 20 competitor trees whose `dist` distance to the focal tree is less than or equal to `r comp_dist`, including their unique tree-stem ID number, their species, and their basal area (in m$^2$) calculated as $\frac{\pi \times (\text{DBH/2})^2}{10000}$ where $DBH$ is the value from the earlier census in cm.
+Saving our focal versus competitor information in list-column minimizes redundancy since we do not repeat information on the focal tree 20 times.
+
+```{r, eval = FALSE}
+focal_vs_comp_scbi %>% 
+  filter(focal_ID == 4) %>% 
+  select(focal_ID, dbh, comp) %>% 
+  unnest(cols = "comp")
+```
+```{r scbi-focal-vs-comp-3, echo = FALSE}
+# Not sure why we need to do things this way:
+focal_vs_comp_scbi %>% 
+  filter(focal_ID == 4) %>% 
+  select(focal_ID, dbh, comp) %>% 
+  unnest(cols = "comp") %>% 
+  print(n = 5)
+```
 
@@
 
-Next we fit the following linear model to the dbh of each focal tree.
+HEY BERT PICK IT UP HERE
+
+Next we fit the following linear model to the DBH of each focal tree.
 Let $i = 1, \ldots, n_j$ index all $n_j$ trees of "focal" species group $j$; let $j = 1, \ldots, J$ index all $J$ focal species groups; and let $k = 1, \ldots, K$ index all $K$ "competitor" species groups.
@@
 
-Using these posterior means, we then use the posterior predictive distribution to obtain fitted/predicted values $\widehat{y}$ of the dbh for each focal tree using the `predict_bayesian_model()`.
-These $\widehat{y}$ can then be compared to the observed $y$ dbh's to compute the root mean-square error, a measure of a model's predictive error which has the same units as the observed data $y$.
+Using these posterior means, we then use the posterior predictive distribution to obtain fitted/predicted values $\widehat{y}$ of the DBH for each focal tree using the `predict_bayesian_model()`.
+These $\widehat{y}$ can then be compared to the observed $y$ DBH's to compute the root mean-square error, a measure of a model's predictive error which has the same units as the observed data $y$.
 
@@
 
-```{r bw-model-fit}
+```{r bw-model-fit, eval=FALSE}
 comp_bayes_lm_bw <- focal_vs_comp_bw %>%
@@
 
-```{r bw-model-predict}
+```{r bw-model-predict, eval=FALSE}
 focal_vs_comp_bw <- focal_vs_comp_bw %>%
@@
 
-```{r bw-model-rmse}
+```{r bw-model-rmse, eval=FALSE}
 focal_vs_comp_bw %>%
@@
 
-```{r bw-model-shuffle-fit}
+```{r bw-model-shuffle-fit, eval=FALSE}
 comp_bayes_lm_bw_shuffle <- focal_vs_comp_bw %>%
@@
 
-```{r bw-model-shuffle-predict}
+```{r bw-model-shuffle-predict, eval=FALSE}
 focal_vs_comp_bw <- focal_vs_comp_bw %>%
@@
 
-```{r bw-model-shuffle-rmse}
+```{r bw-model-shuffle-rmse, eval=FALSE}
 focal_vs_comp_bw %>%
@@
 
-```{r scbi-model-fit}
+```{r scbi-model-fit, eval=FALSE}
 comp_bayes_lm_scbi <- focal_vs_comp_scbi %>%
@@
 
-```{r scbi-model-predict}
+```{r scbi-model-predict, eval=FALSE}
 focal_vs_comp_scbi <- focal_vs_comp_scbi %>%
@@
 
-```{r scbi-model-rmse}
+```{r scbi-model-rmse, eval=FALSE}
 focal_vs_comp_scbi %>%
@@
 
-```{r scbi-model-residuals, out.width="100%", fig.cap="Predicted versus observed growth.", echo = FALSE}
+```{r scbi-model-residuals, out.width="100%", fig.cap="Predicted versus observed growth.", echo = FALSE, eval=FALSE}
 ggplot(focal_vs_comp_scbi, aes(x = growth, y = growth_hat)) +
@@
   labs(
-    x = "Observed growth in dbh", y = "Predicted growth in dbh",
+    x = "Observed growth in DBH", y = "Predicted growth in DBH",
     title = "Predicted vs Observed Growth"
@@
 
-```{r scbi-model-residuals-2, out.width="80%", fig.cap="Spatial distribution of residuals for model applied to SCBI data.", echo = FALSE}
+```{r scbi-model-residuals-2, out.width="80%", fig.cap="Spatial distribution of residuals for model applied to SCBI data.", echo = FALSE, eval=FALSE}
 focal_vs_comp_scbi %>%
@@
 
-```{r bw-posterior-viz-0}
+```{r bw-posterior-viz-0, eval=FALSE}
 focal_vs_comp_bw <- growth_bw %>%
@@
 
-```{r bw-posterior-viz-1}
+```{r bw-posterior-viz-1, eval=FALSE}
 sp_to_plot <- c("cornaceae", "fagaceae", "hamamelidaceae", "juglandaceae", 
@@
 
-```{r bw-posterior-viz-beta0, out.width="100%", fig.cap="Posterior distribution of beta0."}
+```{r bw-posterior-viz-beta0, out.width="100%", fig.cap="Posterior distribution of beta0.", eval=FALSE}
 plot1 <- autoplot(comp_bayes_lm_bw, type = "intercepts")
@@
 
-```{r bw-posterior-viz-beta-dbh, out.width="100%", fig.cap="Posterior distribution of betadbh."}
+```{r bw-posterior-viz-beta-dbh, out.width="100%", fig.cap="Posterior distribution of betadbh.", eval=FALSE}
 plot2 <- autoplot(comp_bayes_lm_bw, type = "dbh_slopes")
@@
 
-```{r bw-posterior-viz-lambda, out.width="100%", fig.cap="Posterior distribution of lambda's for Big Woods."}
+```{r bw-posterior-viz-lambda, out.width="100%", fig.cap="Posterior distribution of lambda's for Big Woods.", eval=FALSE}
 plot3 <- autoplot(comp_bayes_lm_bw, type = "competition")
@@
 
-```{r scbi-posterior-viz-1}
+```{r scbi-posterior-viz-1, eval=FALSE}
 sp_to_plot <- c("quru", "litu", "cagl", "cato")
@@
 
-```{r scbi-posterior-viz-lambda, out.width="100%", fig.cap="Posterior distribution of lambda's for SCBI."}
+```{r scbi-posterior-viz-lambda, out.width="100%", fig.cap="Posterior distribution of lambda's for SCBI.", eval=FALSE}
 plot3 <- autoplot(comp_bayes_lm_bw, type = "competition")
@@
 
-```{r bw-spatial-cv}
+```{r bw-spatial-cv, eval=FALSE}
 focal_vs_comp_bw <- focal_vs_comp_bw %>%
@@
 
-```{r scbi-spatial-cv}
+```{r scbi-spatial-cv, eval=FALSE}
 focal_vs_comp_scbi <- focal_vs_comp_scbi %>%

### 2021-02-16 etc/paper/forestecology.tex
Done "Identify all focal and corresponding competitor trees" section

@@
 
+\hypertarget{model}{%
+\subsection{Model}\label{model}}
+
+Describe model specifics.
+
 \hypertarget{example}{%
@@
 \item
-  Compute focal versus competitor tree information.
+  Identify all focal and corresponding competitor trees.
 \item
   Fit model and make predictions.
-\item
-  Visualize model results.
 \item
@@
 between sites thereby necessitating some data wrangling and checking.
-For example, the SCBI site records all dbh's in millimeters, whereas the
+For example, the SCBI site records all DBH's in millimeters, whereas the
 Michigan Big Woods site records them in centimeters
@@
 tree-stem, \texttt{sp} of type \texttt{fct} factor identifying species,
-\texttt{dbh1} and \texttt{dbh2} of type \texttt{dbl} quantifying the dbh
+\texttt{dbh1} and \texttt{dbh2} of type \texttt{dbl} quantifying the DBH
 at earlier and later census, and \texttt{growth} of type \texttt{dbl}
@@
 variable with the date of measurement to be of type \texttt{date},
-convert dbh to be in centimeters\footnote{A rule of thumb to determine
-  the units of dbh is check if the smallest non-zero and non-missing
+convert DBH to be in centimeters\footnote{A rule of thumb to determine
+  the units of DBH is check if the smallest non-zero and non-missing
   measurement is 1 or 10. If the former, then centimeters. If the later,
   then millimeters. This is because ForestGEO protocols state that only
-  trees with dbh greater or equal to 1cm should be included in censuses.},
+  trees with DBH greater or equal to 1cm should be included in censuses.},
 convert the \texttt{sp} variable containing species information from
@@
 need for this in Section \ref{spatial-cross-validation}). Furthermore,
-in order to speed up computation for purposes of this examplt, we only
+in order to speed up computation for purposes of this example, we only
 consider a 9 ha subsection of the 25.6 ha of the SCBI site: \texttt{gx}
@@
 \CommentTok{## 5     79 tiam  47.7  M      A       46.8 M      -0.161   (40 381.1)}
-\CommentTok{##  6     80 caca   5.15 M      A        6.5 M       0.253  (38.7 421.7)}
-\CommentTok{##  7     96 libe   2.3  J;M    A        3.7 M       0.262      (60 310)}
-\CommentTok{##  8    100 caca   5.09 NULL   A       NA   DN     NA      (52.5 476.3)}
-\CommentTok{##  9    101 litu  65.4  M      A       68.4 M       0.552  (47.1 567.3)}
-\CommentTok{## 10    102 astr   1.99 NULL   A        2.5 M       0.0954 (40.8 575.5)}
-\CommentTok{## # ... with 7,944 more rows}
+\CommentTok{## # ... with 7,949 more rows}
 \end{Highlighting}
@@
 \texttt{sf} that includes a numerical \texttt{growth} reflecting the
-average annual growth in dbh (in cm) for all trees that were alive at
+average annual growth in DBH (in cm) for all trees that were alive at
 both time points as well a \texttt{geometry} variable encoding each
@@
 
-\NormalTok{blocks_scbi <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}
-  \KeywordTok{sf_polygon}\NormalTok{(fold1),}
-  \KeywordTok{sf_polygon}\NormalTok{(fold2)}
-\NormalTok{) }\OperatorTok{%>%}
+\NormalTok{blocks_scbi <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}\KeywordTok{sf_polygon}\NormalTok{(fold1), }\KeywordTok{sf_polygon}\NormalTok{(fold2)) }\OperatorTok{%>%}
 \StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
@@
 \KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_sf_text}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_scbi, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{label =}\NormalTok{ foldID, }\DataTypeTok{col =}\NormalTok{ buffer), }\DataTypeTok{alpha =} \FloatTok{0.2}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{geom_sf_text}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_scbi, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{label =}\NormalTok{ foldID, }\DataTypeTok{col =}\NormalTok{ buffer), }
+               \DataTypeTok{alpha =} \FloatTok{0.2}\NormalTok{) }\OperatorTok{+}
 \StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ blocks_scbi, }\DataTypeTok{fill =} \StringTok{"transparent"}\NormalTok{)}
@@
 
-\hypertarget{define-focal-versus-competitor-trees}{%
-\subsection{Define focal versus competitor
-trees}\label{define-focal-versus-competitor-trees}}
+\hypertarget{identify-all-focal-and-corresponding-competitor-trees}{%
+\subsection{Identify all focal and corresponding competitor
+trees}\label{identify-all-focal-and-corresponding-competitor-trees}}
 
-HEY BERT PICK IT UP HERE
-
-Next we define \texttt{focal\_vs\_comp} data frames which connects each
-focal tree in the \texttt{growth\_df} data frames to the trees in its
-competitive neighborhood range as defined by the \texttt{comp\_dist}
-constant. So for example, if \texttt{growth\_df} consisted of two focal
-trees with two and three neighbors with \texttt{comp\_dist}
-respectively, \texttt{focal\_vs\_comp} would be a data frame of 5 rows
-connecting each focal tree to it's competitors. The
-\texttt{create\_focal\_vs\_comp()} function makes this connection taking
-as inputs the \texttt{growth\_df} data frame; the \texttt{comp\_dist}
-constant defining competitive range; \texttt{cv\_grid\_sf}, giving the
-cross validation grid; and the \texttt{id} variable.
-
-\hypertarget{big-woods}{%
-\subsubsection{Big Woods}\label{big-woods}}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{focal_vs_comp_bw <-}\StringTok{ }\NormalTok{growth_bw }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(comp_dist, }\DataTypeTok{cv_grid_sf =}\NormalTok{ blocks_bw, }\DataTypeTok{id =} \StringTok{"treeID"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-TODO: Figure out how to show this data frame's contents.
-
-\hypertarget{scbi}{%
-\subsubsection{SCBI}\label{scbi}}
+The next step in our analysis sequence is to identify all focal trees
+and their corresponding competitor trees. The
+\texttt{create\_focal\_vs\_comp()} functions performs these tasks and
+returns a new data frame of type \texttt{sf} containing this
+information. On top of the previously discussed arguments
+\texttt{comp\_dist} defining the competition neighborhood and
+\texttt{id} indicating which variable in the data frame uniquely
+identifies each tree-stem, this function also requires an \texttt{sf}
+object representation of the spatial cross-validation blocks/folds; in
+our case, this was manually encoded in the \texttt{blocks\_scbi} in
+Section \ref{spatial-information} while in our Appendix we present an
+example where this was performed using \texttt{spatialBlock()} from the
+\texttt{blockCV} package. We present the resulting data frame below with
+the \texttt{foldID} variable omitted for compactness of presentation.
 
@@
 \StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(comp_dist, }\DataTypeTok{cv_grid_sf =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{)}
+\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\NormalTok{foldID)}
+\CommentTok{## # A tibble: 6,296 x 6}
+\CommentTok{##   focal_ID focal_sp   dbh     geometry growth comp             }
+\CommentTok{##      <dbl> <fct>    <dbl>      <POINT>  <dbl> <list>           }
+\CommentTok{## 1        4 nysy     13.6  (14.2 428.5)  0.103 <tibble [20 x 4]>}
+\CommentTok{## 2        5 havi      8.8   (9.4 436.4)  0.150 <tibble [32 x 4]>}
+\CommentTok{## 3       79 tiam     47.7    (40 381.1) -0.161 <tibble [20 x 4]>}
+\CommentTok{## 4       80 caca      5.15 (38.7 421.7)  0.253 <tibble [12 x 4]>}
+\CommentTok{## 5       96 libe      2.3      (60 310)  0.262 <tibble [14 x 4]>}
+\CommentTok{## # ... with 6,291 more rows}
 \end{Highlighting}
@@
 
-TODO: Figure out how to show this data frame's contents.
+TODO: Below reconcile the number of rows as they off by one from
+\texttt{growth\_scbi} pipe \texttt{filter(!is.na(growth)\ \&\ !buffer)}.
+perhaps by removing NA's in the \texttt{growth\_scbi} stage.
+
+The resulting data frame \texttt{focal\_vs\_comp\_scbi} has 6296 rows,
+representing the subset of the 7954 trees in \texttt{growth\_scbi} that
+will be considered as focal trees and thus have their growths modeled.
+Recall from Section \ref{spatial-information} this consists all trees
+that are not part of the buffer region in Figure
+\ref{fig:scbi-spatial-information}. Two new variables \texttt{focal\_ID}
+and \texttt{focal\_sp} related to tree-stem identification and species
+information. Most notably however is a new variable \texttt{comp} which
+contains information on all competitor trees for a given focal tree
+saved in list-column format, a feature of the \texttt{tidyr} package
+\citet{tidyr_package}. For example, we drill-down on the tree with
+\texttt{focal\_ID} equal to 4. It has 20 competitor trees each described
+by 4 variables as indicated by the fact that \texttt{comp} is a
+\texttt{\textless{}tibble\ {[}20\ ×\ 4{]}\textgreater{}}.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{filter}\NormalTok{(focal_ID }\OperatorTok{==}\StringTok{ }\DecValTok{4}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{select}\NormalTok{(focal_ID, dbh, comp)}
+\CommentTok{## # A tibble: 1 x 3}
+\CommentTok{##   focal_ID   dbh comp             }
+\CommentTok{##      <dbl> <dbl> <list>           }
+\CommentTok{## 1        4  13.6 <tibble [20 x 4]>}
+\end{Highlighting}
+\end{Shaded}
+
+Using the \texttt{unnest()} function from the \texttt{tidyr} package, we
+can flatten list-column into regular columns. We observe that for the
+same focal tree with DBH equal to 13.65cm, we have information on all 20
+competitor trees whose \texttt{dist} distance to the focal tree is less
+than or equal to 7.5, including their unique tree-stem ID number, their
+species, and their basal area (in m\(^2\)) calculated as
+\(\frac{\pi \times (\text{DBH/2})^2}{10000}\) where \(DBH\) is the value
+from the earlier census in cm. Saving our focal versus competitor
+information in list-column minimizes redundancy since we do not repeat
+information on the focal tree 20 times.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{filter}\NormalTok{(focal_ID }\OperatorTok{==}\StringTok{ }\DecValTok{4}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{select}\NormalTok{(focal_ID, dbh, comp) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{unnest}\NormalTok{(}\DataTypeTok{cols =} \StringTok{"comp"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # A tibble: 20 x 6
+##   focal_ID   dbh comp_ID  dist comp_sp comp_basal_area
+##      <dbl> <dbl>   <dbl> <dbl> <fct>             <dbl>
+## 1        4  13.6    1836  7.48 tiam            0.0176 
+## 2        4  13.6    1847  2.81 nysy            0.00332
+## 3        4  13.6    1848  1.62 nysy            0.00396
+## 4        4  13.6    1849  2.62 nysy            0.00535
+## 5        4  13.6    1850  2.98 havi            0.00472
+## # ... with 15 more rows
+\end{verbatim}
 
@@
 
-Next we fit the following linear model to the dbh of each focal tree.
+HEY BERT PICK IT UP HERE
+
+Next we fit the following linear model to the DBH of each focal tree.
 Let \(i = 1, \ldots, n_j\) index all \(n_j\) trees of ``focal'' species
@@
 distribution to obtain fitted/predicted values \(\widehat{y}\) of the
-dbh for each focal tree using the \texttt{predict\_bayesian\_model()}.
-These \(\widehat{y}\) can then be compared to the observed \(y\) dbh's
+DBH for each focal tree using the \texttt{predict\_bayesian\_model()}.
+These \(\widehat{y}\) can then be compared to the observed \(y\) DBH's
 to compute the root mean-square error, a measure of a model's predictive
@@
 \StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
-\CommentTok{## [1] 0.148145}
 \end{Highlighting}
@@
 \StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
-\CommentTok{## [1] 0.1505383}
 \end{Highlighting}
@@
 \StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
-\CommentTok{## [1] 0.1281398}
 \end{Highlighting}
@@
 \ref{fig:scbi-model-residuals-2} we present the residuals.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{Figures/scbi-model-residuals-1}
-
-}
-
-\caption{Predicted versus observed growth.}\label{fig:scbi-model-residuals}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.8\linewidth]{Figures/scbi-model-residuals-2-1}
-
-}
-
-\caption{Spatial distribution of residuals for model applied to SCBI data.}\label{fig:scbi-model-residuals-2}
-\end{figure}
 
@@
 
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{Figures/bw-posterior-viz-beta0-1}
-
-}
-
-\caption{Posterior distribution of beta0.}\label{fig:bw-posterior-viz-beta0}
-\end{figure}
-
 Figure \ref{fig:bw-posterior-viz-beta-dbh} Next \texttt{beta\_dbh} gives
@@
 \end{Shaded}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{Figures/bw-posterior-viz-beta-dbh-1}
-
-}
-
-\caption{Posterior distribution of betadbh.}\label{fig:bw-posterior-viz-beta-dbh}
-\end{figure}
 
@@
 
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{Figures/bw-posterior-viz-lambda-1}
-
-}
-
-\caption{Posterior distribution of lambda's for Big Woods.}\label{fig:bw-posterior-viz-lambda}
-\end{figure}
-
-\hypertarget{scbi-2}{%
-\subsubsection{SCBI}\label{scbi-2}}
+\hypertarget{scbi-1}{%
+\subsubsection{SCBI}\label{scbi-1}}
 
@@
 \end{Shaded}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{Figures/scbi-posterior-viz-lambda-1}
-
-}
-
-\caption{Posterior distribution of lambda's for SCBI.}\label{fig:scbi-posterior-viz-lambda}
-\end{figure}
 
@@
 
-\hypertarget{big-woods-3}{%
-\subsubsection{Big Woods}\label{big-woods-3}}
+\hypertarget{big-woods-2}{%
+\subsubsection{Big Woods}\label{big-woods-2}}
 
@@
 \StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
-\CommentTok{## [1] 0.1532316}
 \end{Highlighting}
@@
 \StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
-\CommentTok{## [1] 0.144608}
 \end{Highlighting}

### 2021-02-18 README.Rmd
Done "Identify all focal and corresponding competitor trees" section
update README

@@
 1. Add spatial information
-1. Compute focal versus competitor tree information
+1. Identify all focal and corresponding competitor trees
 1. Fit model and make predictions
@@
 focal_vs_comp_ex <- focal_vs_comp_ex %>%
-  mutate(growth_hat = predict(comp_bayes_lm_ex, focal_vs_comp_ex))
+  mutate(growth_hat = predict(comp_bayes_lm_ex, newdata = focal_vs_comp_ex))
 focal_vs_comp_ex

### 2021-02-18 etc/paper/forestecology.Rmd
Done "Fit model & make predictions" & "Visualize model results" section

@@
 
-## Model
+## Model for growth of tree {#competition-model}
 
-Describe model specifics.
+While there are a littany of models one can consider, in PLOSOne we considered a simple one.
+
+Describe model specifics. Bayesian linear regression model.
+
+Next we fit the following linear model to the DBH of each focal tree.
+Let $i = 1, \ldots, n_j$ index all $n_j$ trees of "focal" species group $j$; let $j = 1, \ldots, J$ index all $J$ focal species groups; and let $k = 1, \ldots, K$ index all $K$ "competitor" species groups.
+We modeled the growth in diameter per year $y_{ij}$ (in centimeters per year) of the $i^{th}$ tree of focal species group $j$ as a linear model $f$ of the following covariates $\vec{x}_{ij}$
+
+$$
+\newcommand{\dbh}{\text{DBH}}
+\newcommand{\biomass}{\text{biomass}}
+\newcommand{\BA}{\text{BA}}
+y_{ij} = f(\vec{x}_{ij}) + \epsilon_{ij} = \beta_{0,j} + \beta_{\dbh,j} \cdot \dbh_{ij} + \sum_{k=1}^{K} \lambda_{jk} \cdot \BA_{ijk} + \epsilon_{ij}
+$$
+
+<!--where $\beta_{0,j}$ is the diameter-independent growth rate for group $j$; $\dbh_{ij}$ is the diameter at breast height (in centimeters) of the focal tree at the earlier census; $\beta_{\dbh,j}$ is the amount of the growth rate changed depending on diameter for group $j$;  $\BA_{ijk}$ is the sum of the basal area of all trees of competitor species group $k$; $\lambda_{jk}$ is the change in growth for individuals of group $j$ from nearby competitors of group $k$; and $\epsilon_{ij}$ is a random error term distributed $\text{Normal}(0, \sigma^2)$.-->
+
+Link to https://doi.org/10.1371/journal.pone.0229930.s004
+
+For this linear model's case, there exists a closed form solution as described here.
+As such, the `fit_bayesian_model()` function using matrix algebra to obtain all parameter estimates, rather than computationally expensive Monte Carlo approximations.
+The inputs to this function are a `focal_vs_comp` data frame, `prior_param` a list of priors, and a boolean flag `run_shuffle` on whether or not to run competitor-species identity permutations which we will demonstrate below on the Michigan Big Woods data.
+This function returns the posterior means of all parameters.
 
@@
 1. Identify all focal and corresponding competitor trees.
-1. Fit model and make predictions.
-1. Additionally: Evaluate model performance using spatial cross-validation.
-1. Additionally: Evaluate the effect of competitor species identity using permutation tests.
+1. Fit model, make predictions, and visualize model results
+
+Additionally,
+
+1. Evaluate the effect of competitor species identity using permutation tests.
+1. Evaluate model performance using spatial cross-validation.
 
@@
 library(blockCV)
+library(patchwork)
 ```
@@
 ggplot() +
-  geom_sf_text(data = growth_scbi, aes(label = foldID, col = buffer), 
-               alpha = 0.2) +
+  geom_sf_text(data = growth_scbi, aes(label = foldID, col = buffer), alpha = 0.2) +
   geom_sf(data = blocks_scbi, fill = "transparent")
@@
 
-## Identify all focal and corresponding competitor trees
+## Identify all focal and corresponding competitor trees {#focal-vs-comp}
 
@@
 
-HEY BERT PICK IT UP HERE
+Now that we've identified all focal and corresponding competitor trees and saved this information in a data frame of type `focal_vs_comp`, the next step in our analysis sequence is to fit a model for the growth of all focal trees.
+Currently the `forestecology` package can only fit the competition Bayesian linear regression model outlined in Section \ref{competition-model} using the `comp_bayes_lm()` function.
+However, any model implemented in a function that similarly takes an input data frame of type `focal_vs_comp` as an argument can be easily swapped in.
+In the case of our specific competition Bayesian linear regression model, we can also specify prior distributions on all parameters of interest (here chosen to be the defaults).
 
-Next we fit the following linear model to the DBH of each focal tree.
-Let $i = 1, \ldots, n_j$ index all $n_j$ trees of "focal" species group $j$; let $j = 1, \ldots, J$ index all $J$ focal species groups; and let $k = 1, \ldots, K$ index all $K$ "competitor" species groups.
-We modeled the growth in diameter per year $y_{ij}$ (in centimeters per year) of the $i^{th}$ tree of focal species group $j$ as a linear model $f$ of the following covariates $\vec{x}_{ij}$
-
-$$
-\newcommand{\dbh}{\text{DBH}}
-\newcommand{\biomass}{\text{biomass}}
-\newcommand{\BA}{\text{BA}}
-y_{ij} = f(\vec{x}_{ij}) + \epsilon_{ij} = \beta_{0,j} + \beta_{\dbh,j} \cdot \dbh_{ij} + \sum_{k=1}^{K} \lambda_{jk} \cdot \BA_{ijk} + \epsilon_{ij}
-$$
-
-<!--where $\beta_{0,j}$ is the diameter-independent growth rate for group $j$; $\dbh_{ij}$ is the diameter at breast height (in centimeters) of the focal tree at the earlier census; $\beta_{\dbh,j}$ is the amount of the growth rate changed depending on diameter for group $j$;  $\BA_{ijk}$ is the sum of the basal area of all trees of competitor species group $k$; $\lambda_{jk}$ is the change in growth for individuals of group $j$ from nearby competitors of group $k$; and $\epsilon_{ij}$ is a random error term distributed $\text{Normal}(0, \sigma^2)$.-->
-
-We estimate the model's parameters using Bayesian linear regression implemented in the `fit_bayesian_model()` function. TODO: define all parameters
-
-For this linear model's case, there exists a closed form solution as described here.
-As such, the `fit_bayesian_model()` function using matrix algebra to obtain all parameter estimates, rather than computationally expensive Monte Carlo approximations.
-The inputs to this function are a `focal_vs_comp` data frame, `prior_param` a list of priors, and a boolean flag `run_shuffle` on whether or not to run competitor-species identity permutations which we will demonstrate below on the Michigan Big Woods data.
-This function returns the posterior means of all parameters.
-
-Using these posterior means, we then use the posterior predictive distribution to obtain fitted/predicted values $\widehat{y}$ of the DBH for each focal tree using the `predict_bayesian_model()`.
-These $\widehat{y}$ can then be compared to the observed $y$ DBH's to compute the root mean-square error, a measure of a model's predictive error which has the same units as the observed data $y$.
-
-### Big Woods
-
-For the Michigan Big Woods data we present two use cases of the model fitting and prediction scheme.
-The first use case is the simplest where we assess the fit of the model using root mean squared error.
-The second use case then answers the question of whether species competitor identity matters using permutation test.
-
-For the first use case, we fit the linear model specified in Equation XXX to our data frame of type `focal_vs_comp`.
-This input/outputs of the `fit_bayesian_model()` function are lists of the prior/posterior means of parameters of the linear regression specified in XXX.
-Generally speaking, there are two classes of regression parameters: $\beta$ main effects and $\lambda$ competitive effects.
-In the upcoming Section \ref{viz-posterior-distributions}, we will present code visualizing this posterior distributions.
-
-```{r bw-model-fit, eval=FALSE}
-comp_bayes_lm_bw <- focal_vs_comp_bw %>%
-  comp_bayes_lm(prior_param = NULL)
-```
-
-This output of posterior parameters for the specified competition model are then used along with the posterior predictive distribution encoded in `predict_bayesian_model()` to return predicted growths for each individual tree.
-We join these predicted growths to the original growth data frame.
-
-```{r bw-model-predict, eval=FALSE}
-focal_vs_comp_bw <- focal_vs_comp_bw %>%
-  mutate(growth_hat = predict(comp_bayes_lm_bw, focal_vs_comp_bw))
-```
-
-We then use the `rmse()` function from the `yardstick` package to obtain the root mean squared error of the observed versus fitted values of growth.
-
-```{r bw-model-rmse, eval=FALSE}
-focal_vs_comp_bw %>%
-  rmse(truth = growth, estimate = growth_hat) %>%
-  pull(.estimate)
-```
-
-The second use case is near identical to the first, but with a small change in the code to test whether the identity of the competitor matters.
-By adding a `run_shuffle = TRUE` argument to `fit_bayesian_model()`, for each focal tree its competitor trees' species identity will be "shuffled" randomly much like in a permutation test.
-By shuffling these species labels we are effectively fitting the model under a null model that competitor species identity does not matter.
-If the "shuffled" RMSE's are consistently lower than the unshuffled RMSE corresponding to the observed data, then we have evidence to suggest that competitor identity matters to competitive interactions.
-
-```{r bw-model-shuffle-fit, eval=FALSE}
-comp_bayes_lm_bw_shuffle <- focal_vs_comp_bw %>%
-  comp_bayes_lm(prior_param = NULL, run_shuffle = TRUE)
-```
-
-```{r bw-model-shuffle-predict, eval=FALSE}
-focal_vs_comp_bw <- focal_vs_comp_bw %>%
-  mutate(growth_hat_shuffle = predict(comp_bayes_lm_bw_shuffle, focal_vs_comp_bw))
-```
-
-```{r bw-model-shuffle-rmse, eval=FALSE}
-focal_vs_comp_bw %>%
-  rmse(truth = growth, estimate = growth_hat_shuffle) %>%
-  pull(.estimate)
-```
-
-The RMSE is fact lower for the non-shuffled version, indicative of a better model fit.
-This gives support for the idea that competitor identity does matter for competitive interactions.
-In @allen_permutation_2020 we run this shuffle a large number of times to construct a full permutation distribution to show that this difference is robust to resampling variation.
-
-### SCBI
-
-In the case of the SCBI data, we once again perform the same model fitting and computing of fitted growths as with the Big Woods data, but this time we map the residuals of the observed minus fitted values to look for spatial patterns.
-
-```{r scbi-model-fit, eval=FALSE}
+```{r scbi-model-fit}
 comp_bayes_lm_scbi <- focal_vs_comp_scbi %>%
@@
 
-```{r scbi-model-predict, eval=FALSE}
-focal_vs_comp_scbi <- focal_vs_comp_scbi %>%
-  mutate(growth_hat = predict(comp_bayes_lm_scbi, focal_vs_comp_scbi))
+The returned `comp_bayes_lm_scbi` output is an object of S3 class type `comp_bayes_lm` which contains the posterior values of all parameters in our competition Bayesian linear regression.
+This class of object includes generic methods implemented for `print()`, `predict()`, and `ggplot2::autoplot()`.
+First the generic for `print()` displays the names of all prior & posterior parameters (accessible via `comp_bayes_lm_scbi$prior_params` and `comp_bayes_lm_scbi$post_params`) along with the model formula:
+
+```{r}
+comp_bayes_lm_scbi
 ```
 
-```{r scbi-model-rmse, eval=FALSE}
-focal_vs_comp_scbi %>%
+Next, the generic for `predict()` takes as inputs the posterior parameter values in `comp_bayes_lm_scbi` and the predictor variables in `newdata` and outputs a vector of fitted/predicted values $\widehat{y}$ of the DBH for each focal tree (as returned by the posterior predictive distribution).
+
+```{r scbi-model-predict}
+focal_vs_comp_scbi <- focal_vs_comp_scbi %>%
+  mutate(growth_hat = predict(comp_bayes_lm_scbi, newdata = focal_vs_comp_scbi))
+```
+```{r}
+focal_vs_comp_scbi
+```
+
+We can then compare the observed and fitted/predicted growths to compute the root mean squared error of our model fit.
+
+```{r scbi-model-rmse}
+model_rmse <- focal_vs_comp_scbi %>%
   rmse(truth = growth, estimate = growth_hat) %>%
   pull(.estimate)
+model_rmse
 ```
 
-In Figures \ref{fig:scbi-model-residuals} and \ref{fig:scbi-model-residuals-2} we present the residuals.
+Lastly, the generic for `ggplot2::autoplot()` allows us to
 
-```{r scbi-model-residuals, out.width="100%", fig.cap="Predicted versus observed growth.", echo = FALSE, eval=FALSE}
-ggplot(focal_vs_comp_scbi, aes(x = growth, y = growth_hat)) +
-  geom_point(size = 0.5, color = rgb(0, 0, 0, 0.25)) +
-  stat_smooth(method = "lm") +
-  geom_abline(slope = 1, intercept = 0) +
-  coord_fixed() +
-  labs(
-    x = "Observed growth in DBH", y = "Predicted growth in DBH",
-    title = "Predicted vs Observed Growth"
-  )
+```{r scbi-posterior-viz, out.width="100%", fig.cap="Posterior distributions of all parameters for a subset of species."}
+# Plot posteriors for only a subset of species
+sp_to_plot <- c("quru", "litu", "cagl", "cato")
+
+plot1 <- autoplot(comp_bayes_lm_scbi, type = "intercepts", sp_to_plot = sp_to_plot)
+plot2 <- autoplot(comp_bayes_lm_scbi, type = "dbh_slopes", sp_to_plot = sp_to_plot)
+plot3 <- autoplot(comp_bayes_lm_scbi, type = "competition", sp_to_plot = sp_to_plot)
+
+# Combine plots using patchwork
+(plot1 | plot2) / plot3
 ```
 
-```{r scbi-model-residuals-2, out.width="80%", fig.cap="Spatial distribution of residuals for model applied to SCBI data.", echo = FALSE, eval=FALSE}
-focal_vs_comp_scbi %>%
-  st_as_sf() %>%
-  # TODO: Need to investigate missingness
-  filter(!is.na(growth_hat)) %>%
-  mutate(
-    error = growth - growth_hat,
-    error_bin = cut_number(error, n = 5),
-    error_compress = ifelse(error < -0.75, -0.75, ifelse(error > 0.75, 0.75, error))
-  ) %>%
-  ggplot() +
-  geom_sf(aes(col = error_compress), size = 1) +
-  theme_bw() +
-  scale_color_gradient2(
-    low = "#ef8a62", mid = "#f7f7f7", high = "#67a9cf",
-    name = expression(paste("Residual (cm ", y^{-1}, ")")),
-    breaks = seq(from = -0.75, to = 0.75, by = 0.25),
-    labels = c("< -0.75", "-0.5", "0.25", "0", "0.25", "0.5", "> 0.75")
-  ) +
-  labs(x = "Meter", y = "Meter")
-```
-
-## Visualize model results {#viz-posterior-distributions}
-
-Lastly, we return to the model fits from Section \ref{model-fit-predict} and present tools to visually explore the posterior distributions of all parameters in our model.
-There are two main groups of parameters to consider.
-The $\beta$ coefficients tell us about how fast each species grows and how this depends on DBH while the full matrix of $\lambda$ values describe the competitive effects between pairs of species.
-There is a rich literature on this matrix (cite).
-
-DO WE NEED TO DESCRIBE MECHANICS?
-Because of the structure of the `bw_fit_model` object we cannot simply draw these curves based on the posterior distribution.
-`bw_fit_model()` gives the parameters *compared* to a baseline.
-This is not of direct interest.
-So to display these parameters, as we care about them, we have to sample from the baseline distribution and from the comparison one to get the posterior distribution of interest.
-
-### Big Woods
-
-Here we re-run the model fit to the Big Woods data from Section \ref{model-fit-predict}, but this time use "family" as the group for comparison which has.
-This makes the posterior distributions easier to follow.
-Also, surprisingly, grouping by family performed just as well as grouping by species @allen_permutation_2020.
-First we re-run `create_focal_vs_comp()` and `fit_bayesian_model()` with no permutation shuffling with the grouping variable as family.
-
-```{r bw-posterior-viz-0, eval=FALSE}
-focal_vs_comp_bw <- growth_bw %>%
-  mutate(sp = family %>% factor()) %>%
-  create_focal_vs_comp(comp_dist = comp_dist, cv_grid_sf = blocks_bw, id = "treeID")
-
-comp_bayes_lm_bw <- focal_vs_comp_bw %>%
-  comp_bayes_lm(prior_param = NULL)
-```
-
-Now the posterior parameter outputs of `fit_bayesian_model()` are passed to `plot_bayesian_model_parameters()` to generate visualizations of the posterior parameters.
-These visualizations are displayed in Figure 5 of @allen_permutation_2020.
-For simplicity we only plot a subset of the species families.
-
-```{r bw-posterior-viz-1, eval=FALSE}
-sp_to_plot <- c("cornaceae", "fagaceae", "hamamelidaceae", "juglandaceae", 
-                 "lauraceae", "rosaceae", "sapindaceae", "ulmaceae")
-```
-
-The output is a list with three plots stored.
-Figure \ref{fig:bw-posterior-viz-beta0} The element `beta_0` gives the baseline growth intercept $\beta_0$, i.e., how fast an individual of each group grows independent of DBH).
-
-```{r bw-posterior-viz-beta0, out.width="100%", fig.cap="Posterior distribution of beta0.", eval=FALSE}
-plot1 <- autoplot(comp_bayes_lm_bw, type = "intercepts")
-plot1
-```
-
-Figure \ref{fig:bw-posterior-viz-beta-dbh} Next `beta_dbh` gives the slope for DBH slope $\beta_{dbh,i}$ for each group.
-
-```{r bw-posterior-viz-beta-dbh, out.width="100%", fig.cap="Posterior distribution of betadbh.", eval=FALSE}
-plot2 <- autoplot(comp_bayes_lm_bw, type = "dbh_slopes")
-plot2
-```
-
-Finally Figure \ref{fig:bw-posterior-viz-lambda} `lambda` gives the competition coefficients $\lambda$.
-
-```{r bw-posterior-viz-lambda, out.width="100%", fig.cap="Posterior distribution of lambda's for Big Woods.", eval=FALSE}
-plot3 <- autoplot(comp_bayes_lm_bw, type = "competition")
-plot3
-```
-
-### SCBI
-
-We revisit the posterior parameters for the SCBI from Section {model-fit-predict}, but this time only focus on the $\lambda$ competition coefficients.
-
-```{r scbi-posterior-viz-1, eval=FALSE}
-sp_to_plot <- c("quru", "litu", "cagl", "cato")
-```
-
-```{r scbi-posterior-viz-lambda, out.width="100%", fig.cap="Posterior distribution of lambda's for SCBI.", eval=FALSE}
-plot3 <- autoplot(comp_bayes_lm_bw, type = "competition")
-plot3
-```
-
-Add explanation here.
+## Evaluate the effect of competitor species identity using permutation tests
 
@@
 
-## Run spatial cross-validation {#spatial-cross-validation}
+```{r scbi-permutation-model-fit, eval = FALSE}
+comp_bayes_lm_scbi <- focal_vs_comp_scbi %>%
+  comp_bayes_lm(prior_param = NULL, run_shuffle = TRUE)
+
+focal_vs_comp_scbi <- focal_vs_comp_scbi %>%
+  mutate(
+    growth_hat = predict(comp_bayes_lm_scbi, newdata = focal_vs_comp_scbi)
+  )
+
+model_rmse <- focal_vs_comp_scbi %>%
+  rmse(truth = growth, estimate = growth_hat) %>%
+  pull(.estimate)
+model_rmse
+```
+
+## Evaluate model performance using spatial cross-validation {#spatial-cross-validation}
 
@@
 
-### Big Woods
-
-Applying this spatially cross-validated model fit yields an RMSE is higher than that when the model is fit without cross validation.
-In other words, our model fits in \ref{model-fit-predict} were overly optimistic in the model's fitting power, whereas a cross-validated results yield an estimate that is closer to the truth.
-See @allen_permutation_2020 for more discussion of this.
-
-```{r bw-spatial-cv, eval=FALSE}
-focal_vs_comp_bw <- focal_vs_comp_bw %>%
-  run_cv(comp_dist = comp_dist, cv_grid = blocks_bw)
-
-focal_vs_comp_bw %>%
-  rmse(truth = growth, estimate = growth_hat) %>%
-  pull(.estimate)
-```
-
-### SCBI
-
 Observe once again that this RMSE is much higher than that for the above SCBI model fit without cross-validation.
 
-```{r scbi-spatial-cv, eval=FALSE}
+```{r scbi-spatial-cv, eval=TRUE}
 focal_vs_comp_scbi <- focal_vs_comp_scbi %>%

### 2021-02-18 etc/paper/forestecology.tex
Done "Fit model & make predictions" & "Visualize model results" section

@@
 
-\hypertarget{model}{%
-\subsection{Model}\label{model}}
+\hypertarget{competition-model}{%
+\subsection{Model for growth of tree}\label{competition-model}}
 
-Describe model specifics.
+While there are a littany of models one can consider, in PLOSOne we
+considered a simple one.
+
+Describe model specifics. Bayesian linear regression model.
+
+Next we fit the following linear model to the DBH of each focal tree.
+Let \(i = 1, \ldots, n_j\) index all \(n_j\) trees of ``focal'' species
+group \(j\); let \(j = 1, \ldots, J\) index all \(J\) focal species
+groups; and let \(k = 1, \ldots, K\) index all \(K\) ``competitor''
+species groups. We modeled the growth in diameter per year \(y_{ij}\)
+(in centimeters per year) of the \(i^{th}\) tree of focal species group
+\(j\) as a linear model \(f\) of the following covariates
+\(\vec{x}_{ij}\)
+
+\[
+\newcommand{\dbh}{\text{DBH}}
+\newcommand{\biomass}{\text{biomass}}
+\newcommand{\BA}{\text{BA}}
+y_{ij} = f(\vec{x}_{ij}) + \epsilon_{ij} = \beta_{0,j} + \beta_{\dbh,j} \cdot \dbh_{ij} + \sum_{k=1}^{K} \lambda_{jk} \cdot \BA_{ijk} + \epsilon_{ij}
+\]
+
+Link to \url{https://doi.org/10.1371/journal.pone.0229930.s004}
+
+For this linear model's case, there exists a closed form solution as
+described here. As such, the \texttt{fit\_bayesian\_model()} function
+using matrix algebra to obtain all parameter estimates, rather than
+computationally expensive Monte Carlo approximations. The inputs to this
+function are a \texttt{focal\_vs\_comp} data frame,
+\texttt{prior\_param} a list of priors, and a boolean flag
+\texttt{run\_shuffle} on whether or not to run competitor-species
+identity permutations which we will demonstrate below on the Michigan
+Big Woods data. This function returns the posterior means of all
+parameters.
 
@@
 \item
-  Fit model and make predictions.
+  Fit model, make predictions, and visualize model results
+\end{enumerate}
+
+Additionally,
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
 \item
-  Additionally: Evaluate model performance using spatial
-  cross-validation.
+  Evaluate the effect of competitor species identity using permutation
+  tests.
 \item
-  Additionally: Evaluate the effect of competitor species identity using
-  permutation tests.
+  Evaluate model performance using spatial cross-validation.
 \end{enumerate}
@@
 \KeywordTok{library}\NormalTok{(blockCV)}
+\KeywordTok{library}\NormalTok{(patchwork)}
 \end{Highlighting}
@@
 \KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_sf_text}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_scbi, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{label =}\NormalTok{ foldID, }\DataTypeTok{col =}\NormalTok{ buffer), }
-               \DataTypeTok{alpha =} \FloatTok{0.2}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{geom_sf_text}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_scbi, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{label =}\NormalTok{ foldID, }\DataTypeTok{col =}\NormalTok{ buffer), }\DataTypeTok{alpha =} \FloatTok{0.2}\NormalTok{) }\OperatorTok{+}
 \StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ blocks_scbi, }\DataTypeTok{fill =} \StringTok{"transparent"}\NormalTok{)}
@@
 
-\hypertarget{identify-all-focal-and-corresponding-competitor-trees}{%
+\hypertarget{focal-vs-comp}{%
 \subsection{Identify all focal and corresponding competitor
-trees}\label{identify-all-focal-and-corresponding-competitor-trees}}
+trees}\label{focal-vs-comp}}
 
@@
 
-HEY BERT PICK IT UP HERE
-
-Next we fit the following linear model to the DBH of each focal tree.
-Let \(i = 1, \ldots, n_j\) index all \(n_j\) trees of ``focal'' species
-group \(j\); let \(j = 1, \ldots, J\) index all \(J\) focal species
-groups; and let \(k = 1, \ldots, K\) index all \(K\) ``competitor''
-species groups. We modeled the growth in diameter per year \(y_{ij}\)
-(in centimeters per year) of the \(i^{th}\) tree of focal species group
-\(j\) as a linear model \(f\) of the following covariates
-\(\vec{x}_{ij}\)
-
-\[
-\newcommand{\dbh}{\text{DBH}}
-\newcommand{\biomass}{\text{biomass}}
-\newcommand{\BA}{\text{BA}}
-y_{ij} = f(\vec{x}_{ij}) + \epsilon_{ij} = \beta_{0,j} + \beta_{\dbh,j} \cdot \dbh_{ij} + \sum_{k=1}^{K} \lambda_{jk} \cdot \BA_{ijk} + \epsilon_{ij}
-\]
-
-We estimate the model's parameters using Bayesian linear regression
-implemented in the \texttt{fit\_bayesian\_model()} function. TODO:
-define all parameters
-
-For this linear model's case, there exists a closed form solution as
-described here. As such, the \texttt{fit\_bayesian\_model()} function
-using matrix algebra to obtain all parameter estimates, rather than
-computationally expensive Monte Carlo approximations. The inputs to this
-function are a \texttt{focal\_vs\_comp} data frame,
-\texttt{prior\_param} a list of priors, and a boolean flag
-\texttt{run\_shuffle} on whether or not to run competitor-species
-identity permutations which we will demonstrate below on the Michigan
-Big Woods data. This function returns the posterior means of all
-parameters.
-
-Using these posterior means, we then use the posterior predictive
-distribution to obtain fitted/predicted values \(\widehat{y}\) of the
-DBH for each focal tree using the \texttt{predict\_bayesian\_model()}.
-These \(\widehat{y}\) can then be compared to the observed \(y\) DBH's
-to compute the root mean-square error, a measure of a model's predictive
-error which has the same units as the observed data \(y\).
-
-\hypertarget{big-woods}{%
-\subsubsection{Big Woods}\label{big-woods}}
-
-For the Michigan Big Woods data we present two use cases of the model
-fitting and prediction scheme. The first use case is the simplest where
-we assess the fit of the model using root mean squared error. The second
-use case then answers the question of whether species competitor
-identity matters using permutation test.
-
-For the first use case, we fit the linear model specified in Equation
-XXX to our data frame of type \texttt{focal\_vs\_comp}. This
-input/outputs of the \texttt{fit\_bayesian\_model()} function are lists
-of the prior/posterior means of parameters of the linear regression
-specified in XXX. Generally speaking, there are two classes of
-regression parameters: \(\beta\) main effects and \(\lambda\)
-competitive effects. In the upcoming Section
-\ref{viz-posterior-distributions}, we will present code visualizing this
-posterior distributions.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{comp_bayes_lm_bw <-}\StringTok{ }\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{comp_bayes_lm}\NormalTok{(}\DataTypeTok{prior_param =} \OtherTok{NULL}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-This output of posterior parameters for the specified competition model
-are then used along with the posterior predictive distribution encoded
-in \texttt{predict\_bayesian\_model()} to return predicted growths for
-each individual tree. We join these predicted growths to the original
-growth data frame.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{focal_vs_comp_bw <-}\StringTok{ }\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{growth_hat =} \KeywordTok{predict}\NormalTok{(comp_bayes_lm_bw, focal_vs_comp_bw))}
-\end{Highlighting}
-\end{Shaded}
-
-We then use the \texttt{rmse()} function from the \texttt{yardstick}
-package to obtain the root mean squared error of the observed versus
-fitted values of growth.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
-\end{Highlighting}
-\end{Shaded}
-
-The second use case is near identical to the first, but with a small
-change in the code to test whether the identity of the competitor
-matters. By adding a \texttt{run\_shuffle\ =\ TRUE} argument to
-\texttt{fit\_bayesian\_model()}, for each focal tree its competitor
-trees' species identity will be ``shuffled'' randomly much like in a
-permutation test. By shuffling these species labels we are effectively
-fitting the model under a null model that competitor species identity
-does not matter. If the ``shuffled'' RMSE's are consistently lower than
-the unshuffled RMSE corresponding to the observed data, then we have
-evidence to suggest that competitor identity matters to competitive
-interactions.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{comp_bayes_lm_bw_shuffle <-}\StringTok{ }\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{comp_bayes_lm}\NormalTok{(}\DataTypeTok{prior_param =} \OtherTok{NULL}\NormalTok{, }\DataTypeTok{run_shuffle =} \OtherTok{TRUE}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{focal_vs_comp_bw <-}\StringTok{ }\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{growth_hat_shuffle =} \KeywordTok{predict}\NormalTok{(comp_bayes_lm_bw_shuffle, focal_vs_comp_bw))}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat_shuffle) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
-\end{Highlighting}
-\end{Shaded}
-
-The RMSE is fact lower for the non-shuffled version, indicative of a
-better model fit. This gives support for the idea that competitor
-identity does matter for competitive interactions. In
-\citet{allen_permutation_2020} we run this shuffle a large number of
-times to construct a full permutation distribution to show that this
-difference is robust to resampling variation.
-
-\hypertarget{scbi}{%
-\subsubsection{SCBI}\label{scbi}}
-
-In the case of the SCBI data, we once again perform the same model
-fitting and computing of fitted growths as with the Big Woods data, but
-this time we map the residuals of the observed minus fitted values to
-look for spatial patterns.
+Now that we've identified all focal and corresponding competitor trees
+and saved this information in a data frame of type
+\texttt{focal\_vs\_comp}, the next step in our analysis sequence is to
+fit a model for the growth of all focal trees. Currently the
+\texttt{forestecology} package can only fit the competition Bayesian
+linear regression model outlined in Section \ref{competition-model}
+using the \texttt{comp\_bayes\_lm()} function. However, any model
+implemented in a function that similarly takes an input data frame of
+type \texttt{focal\_vs\_comp} as an argument can be easily swapped in.
+In the case of our specific competition Bayesian linear regression
+model, we can also specify prior distributions on all parameters of
+interest (here chosen to be the defaults).
 
@@
 
+The returned \texttt{comp\_bayes\_lm\_scbi} output is an object of S3
+class type \texttt{comp\_bayes\_lm} which contains the posterior values
+of all parameters in our competition Bayesian linear regression. This
+class of object includes generic methods implemented for
+\texttt{print()}, \texttt{predict()}, and \texttt{ggplot2::autoplot()}.
+First the generic for \texttt{print()} displays the names of all prior
+\& posterior parameters (accessible via
+\texttt{comp\_bayes\_lm\_scbi\$prior\_params} and
+\texttt{comp\_bayes\_lm\_scbi\$post\_params}) along with the model
+formula:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{comp_bayes_lm_scbi}
+\CommentTok{## Bayesian linear regression model parameters with a multivariate Normal likelihood. See ?comp_bayes_lm for details:}
+\CommentTok{## }
+\CommentTok{##   parameter_type           prior posterior}
+\CommentTok{## 1 Inverse-Gamma on sigma^2 a_0   a_star   }
+\CommentTok{## 2 Inverse-Gamma on sigma^2 b_0   b_star   }
+\CommentTok{## 3 Multivariate t on beta   mu_0  mu_star  }
+\CommentTok{## 4 Multivariate t on beta   V_0   V_star   }
+\CommentTok{## }
+\CommentTok{## Model formula:}
+\CommentTok{## growth ~ sp + dbh + dbh * sp + acne * sp + acpl * sp + acru * sp + acsp * sp + aial * sp + amar * sp + astr * sp + beth * sp + caca * sp + caco * sp + cade * sp + cagl * sp + caovl * sp + casp * sp + cato * sp + ceca * sp + ceoc * sp + chvi * sp + coal * sp + coam * sp + cofl * sp + crpr * sp + crsp * sp + divi * sp + elum * sp + eual * sp + fagr * sp + fram * sp + frni * sp + frpe * sp + frsp * sp + havi * sp + ilve * sp + juci * sp + juni * sp + juvi * sp + libe * sp + litu * sp + loma * sp + nysy * sp + pato * sp + pipu * sp + pist * sp + pivi * sp + ploc * sp + prav * sp + prpe * sp + prse * sp + prsp * sp + qual * sp + quco * sp + qufa * sp + qumi * sp + qupr * sp + quru * sp + qusp * sp + quve * sp + rhpe * sp + romu * sp + rops * sp + rual * sp + rupe * sp + ruph * sp + saal * sp + saca * sp + tiam * sp + ulam * sp + ulru * sp + ulsp * sp + unk * sp + viac * sp + vipr * sp + vire * sp}
+\end{Highlighting}
+\end{Shaded}
+
+Next, the generic for \texttt{predict()} takes as inputs the posterior
+parameter values in \texttt{comp\_bayes\_lm\_scbi} and the predictor
+variables in \texttt{newdata} and outputs a vector of fitted/predicted
+values \(\widehat{y}\) of the DBH for each focal tree (as returned by
+the posterior predictive distribution).
+
 \begin{Shaded}
@@
 \NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{growth_hat =} \KeywordTok{predict}\NormalTok{(comp_bayes_lm_scbi, focal_vs_comp_scbi))}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{growth_hat =} \KeywordTok{predict}\NormalTok{(comp_bayes_lm_scbi, }\DataTypeTok{newdata =}\NormalTok{ focal_vs_comp_scbi))}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
+\NormalTok{focal_vs_comp_scbi}
+\CommentTok{## # A tibble: 6,296 x 8}
+\CommentTok{##    focal_ID focal_sp   dbh foldID                  geometry  growth}
+\CommentTok{##       <dbl> <fct>    <dbl> <fct>                    <POINT>   <dbl>}
+\CommentTok{##  1        4 nysy     13.6  1                   (14.2 428.5)  0.103 }
+\CommentTok{##  2        5 havi      8.8  1                    (9.4 436.4)  0.150 }
+\CommentTok{##  3       79 tiam     47.7  1                     (40 381.1) -0.161 }
+\CommentTok{##  4       80 caca      5.15 1                   (38.7 421.7)  0.253 }
+\CommentTok{##  5       96 libe      2.3  1                       (60 310)  0.262 }
+\CommentTok{##  6      101 litu     65.4  1                   (47.1 567.3)  0.552 }
+\CommentTok{##  7      102 astr      1.99 1                   (40.8 575.5)  0.0954}
+\CommentTok{##  8      126 cato     37.4  1                   (60.6 400.2)  0.165 }
+\CommentTok{##  9      127 caca      8.72 1                   (72.7 514.1)  0.0370}
+\CommentTok{## 10      139 astr      1.71 1                   (96.7 315.1)  0.0549}
+\CommentTok{## # ... with 6,286 more rows, and 2 more variables: comp <list>,}
+\CommentTok{## #   growth_hat <dbl>}
 \end{Highlighting}
@@
 
-In Figures \ref{fig:scbi-model-residuals} and
-\ref{fig:scbi-model-residuals-2} we present the residuals.
-
-\hypertarget{viz-posterior-distributions}{%
-\subsection{Visualize model results}\label{viz-posterior-distributions}}
-
-Lastly, we return to the model fits from Section \ref{model-fit-predict}
-and present tools to visually explore the posterior distributions of all
-parameters in our model. There are two main groups of parameters to
-consider. The \(\beta\) coefficients tell us about how fast each species
-grows and how this depends on DBH while the full matrix of \(\lambda\)
-values describe the competitive effects between pairs of species. There
-is a rich literature on this matrix (cite).
-
-DO WE NEED TO DESCRIBE MECHANICS? Because of the structure of the
-\texttt{bw\_fit\_model} object we cannot simply draw these curves based
-on the posterior distribution. \texttt{bw\_fit\_model()} gives the
-parameters \emph{compared} to a baseline. This is not of direct
-interest. So to display these parameters, as we care about them, we have
-to sample from the baseline distribution and from the comparison one to
-get the posterior distribution of interest.
-
-\hypertarget{big-woods-1}{%
-\subsubsection{Big Woods}\label{big-woods-1}}
-
-Here we re-run the model fit to the Big Woods data from Section
-\ref{model-fit-predict}, but this time use ``family'' as the group for
-comparison which has. This makes the posterior distributions easier to
-follow. Also, surprisingly, grouping by family performed just as well as
-grouping by species \citet{allen_permutation_2020}. First we re-run
-\texttt{create\_focal\_vs\_comp()} and \texttt{fit\_bayesian\_model()}
-with no permutation shuffling with the grouping variable as family.
+We can then compare the observed and fitted/predicted growths to compute
+the root mean squared error of our model fit.
 
@@
 \begin{Highlighting}[]
-\NormalTok{focal_vs_comp_bw <-}\StringTok{ }\NormalTok{growth_bw }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =}\NormalTok{ family }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{()) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{cv_grid_sf =}\NormalTok{ blocks_bw, }\DataTypeTok{id =} \StringTok{"treeID"}\NormalTok{)}
-
-\NormalTok{comp_bayes_lm_bw <-}\StringTok{ }\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{comp_bayes_lm}\NormalTok{(}\DataTypeTok{prior_param =} \OtherTok{NULL}\NormalTok{)}
+\NormalTok{model_rmse <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
+\NormalTok{model_rmse}
+\CommentTok{## [1] 0.1281398}
 \end{Highlighting}
@@
 
-Now the posterior parameter outputs of \texttt{fit\_bayesian\_model()}
-are passed to \texttt{plot\_bayesian\_model\_parameters()} to generate
-visualizations of the posterior parameters. These visualizations are
-displayed in Figure 5 of \citet{allen_permutation_2020}. For simplicity
-we only plot a subset of the species families.
+Lastly, the generic for \texttt{ggplot2::autoplot()} allows us to
 
@@
 \begin{Highlighting}[]
-\NormalTok{sp_to_plot <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"cornaceae"}\NormalTok{, }\StringTok{"fagaceae"}\NormalTok{, }\StringTok{"hamamelidaceae"}\NormalTok{, }\StringTok{"juglandaceae"}\NormalTok{, }
-                 \StringTok{"lauraceae"}\NormalTok{, }\StringTok{"rosaceae"}\NormalTok{, }\StringTok{"sapindaceae"}\NormalTok{, }\StringTok{"ulmaceae"}\NormalTok{)}
+\CommentTok{# Plot posteriors for only a subset of species}
+\NormalTok{sp_to_plot <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"quru"}\NormalTok{, }\StringTok{"litu"}\NormalTok{, }\StringTok{"cagl"}\NormalTok{, }\StringTok{"cato"}\NormalTok{)}
+
+\NormalTok{plot1 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_scbi, }\DataTypeTok{type =} \StringTok{"intercepts"}\NormalTok{, }\DataTypeTok{sp_to_plot =}\NormalTok{ sp_to_plot)}
+\NormalTok{plot2 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_scbi, }\DataTypeTok{type =} \StringTok{"dbh_slopes"}\NormalTok{, }\DataTypeTok{sp_to_plot =}\NormalTok{ sp_to_plot)}
+\NormalTok{plot3 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_scbi, }\DataTypeTok{type =} \StringTok{"competition"}\NormalTok{, }\DataTypeTok{sp_to_plot =}\NormalTok{ sp_to_plot)}
+
+\CommentTok{# Combine plots using patchwork}
+\NormalTok{(plot1 }\OperatorTok{|}\StringTok{ }\NormalTok{plot2) }\OperatorTok{/}\StringTok{ }\NormalTok{plot3}
 \end{Highlighting}
@@
 
-The output is a list with three plots stored. Figure
-\ref{fig:bw-posterior-viz-beta0} The element \texttt{beta\_0} gives the
-baseline growth intercept \(\beta_0\), i.e., how fast an individual of
-each group grows independent of DBH).
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{Figures/scbi-posterior-viz-1}
+
+}
+
+\caption{Posterior distributions of all parameters for a subset of species.}\label{fig:scbi-posterior-viz}
+\end{figure}
+
+\hypertarget{evaluate-the-effect-of-competitor-species-identity-using-permutation-tests}{%
+\subsection{Evaluate the effect of competitor species identity using
+permutation
+tests}\label{evaluate-the-effect-of-competitor-species-identity-using-permutation-tests}}
+
+HEY BERT PICK IT UP HERE
 
@@
 \begin{Highlighting}[]
-\NormalTok{plot1 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_bw, }\DataTypeTok{type =} \StringTok{"intercepts"}\NormalTok{)}
-\NormalTok{plot1}
+\NormalTok{comp_bayes_lm_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{comp_bayes_lm}\NormalTok{(}\DataTypeTok{prior_param =} \OtherTok{NULL}\NormalTok{, }\DataTypeTok{run_shuffle =} \OtherTok{TRUE}\NormalTok{)}
+
+\NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
+    \DataTypeTok{growth_hat =} \KeywordTok{predict}\NormalTok{(comp_bayes_lm_scbi, }\DataTypeTok{newdata =}\NormalTok{ focal_vs_comp_scbi)}
+\NormalTok{  )}
+
+\NormalTok{model_rmse <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
+\NormalTok{model_rmse}
 \end{Highlighting}
@@
 
-Figure \ref{fig:bw-posterior-viz-beta-dbh} Next \texttt{beta\_dbh} gives
-the slope for DBH slope \(\beta_{dbh,i}\) for each group.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{plot2 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_bw, }\DataTypeTok{type =} \StringTok{"dbh_slopes"}\NormalTok{)}
-\NormalTok{plot2}
-\end{Highlighting}
-\end{Shaded}
-
-Finally Figure \ref{fig:bw-posterior-viz-lambda} \texttt{lambda} gives
-the competition coefficients \(\lambda\).
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{plot3 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_bw, }\DataTypeTok{type =} \StringTok{"competition"}\NormalTok{)}
-\NormalTok{plot3}
-\end{Highlighting}
-\end{Shaded}
-
-\hypertarget{scbi-1}{%
-\subsubsection{SCBI}\label{scbi-1}}
-
-We revisit the posterior parameters for the SCBI from Section
-\{model-fit-predict\}, but this time only focus on the \(\lambda\)
-competition coefficients.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{sp_to_plot <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"quru"}\NormalTok{, }\StringTok{"litu"}\NormalTok{, }\StringTok{"cagl"}\NormalTok{, }\StringTok{"cato"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{plot3 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_bw, }\DataTypeTok{type =} \StringTok{"competition"}\NormalTok{)}
-\NormalTok{plot3}
-\end{Highlighting}
-\end{Shaded}
-
-Add explanation here.
-
-HEY BERT PICK IT UP HERE
-
 \hypertarget{spatial-cross-validation}{%
-\subsection{Run spatial
+\subsection{Evaluate model performance using spatial
 cross-validation}\label{spatial-cross-validation}}
@@
 
-\hypertarget{big-woods-2}{%
-\subsubsection{Big Woods}\label{big-woods-2}}
-
-Applying this spatially cross-validated model fit yields an RMSE is
-higher than that when the model is fit without cross validation. In
-other words, our model fits in \ref{model-fit-predict} were overly
-optimistic in the model's fitting power, whereas a cross-validated
-results yield an estimate that is closer to the truth. See
-\citet{allen_permutation_2020} for more discussion of this.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{focal_vs_comp_bw <-}\StringTok{ }\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{cv_grid =}\NormalTok{ blocks_bw)}
-
-\NormalTok{focal_vs_comp_bw }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
-\end{Highlighting}
-\end{Shaded}
-
-\hypertarget{scbi-2}{%
-\subsubsection{SCBI}\label{scbi-2}}
-
 Observe once again that this RMSE is much higher than that for the above
@@
 \StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
+\CommentTok{## [1] 0.1446232}
 \end{Highlighting}

### 2021-02-19 etc/paper/forestecology.Rmd
impletemented ideas from Mensh 2017

@@
 
-A high-level overview of the steps of our analysis pipeline is as follows:
+There are four steps in our modeling pipeline:
 
-1. Compute the growth of trees based on census data
+1. Compute the growth of trees based on census data.
 1. Add spatial information:
@@
 1. Identify all focal and corresponding competitor trees.
-1. Fit model, make predictions, and visualize model results
+1. Fit model and make predictions.
 
-Additionally,
+Using these four steps, we can address the two following questions of ecological interest:
 
@@
 
-```{r scbi-trees, out.width="66%", fig.cap="Growth of trees at SCBI."}
+```{r scbi-trees, out.width="66%", fig.cap="Compute growth of trees based on census data from the Smithsonian Conservation Biology Institute (SCBI)."}
 ggplot() +
@@
 # Manually define spatial blocks to act as folds
-fold1 <- rbind(c(0, 300), c(150, 300), c(150, 600), c(0, 600))
-fold2 <- rbind(c(150, 300), c(300, 300), c(300, 600), c(150, 600))
+fold1 <- rbind(c(0, 300), c(150, 300), c(150, 450), c(0, 450))
+fold2 <- rbind(c(150, 300), c(300, 300), c(300, 450), c(150, 450))
+fold3 <- rbind(c(0, 450), c(150, 450), c(150, 600), c(0, 600))
+fold4 <- rbind(c(150, 450), c(300, 450), c(300, 600), c(150, 600))
+n_fold <- 4
 
-blocks_scbi <- bind_rows(sf_polygon(fold1), sf_polygon(fold2)) %>%
-  mutate(folds = c(1, 2) %>% factor())
+blocks_scbi <- bind_rows(
+  sf_polygon(fold1), sf_polygon(fold2), sf_polygon(fold3), sf_polygon(fold4)
+) %>%
+  mutate(folds = c(1:n_fold) %>% factor())
 
@@
 SpatialBlock_scbi <- spatialBlock(
-  speciesData = growth_scbi, k = 2, selection = "systematic", 
+  speciesData = growth_scbi, k = n_fold, selection = "systematic", 
   blocks = blocks_scbi, showBlocks = FALSE, verbose = FALSE
@@
 
-```{r scbi-spatial-information, out.width="66%", fig.cap="Buffer region and cross-validation block information for SCBI data."}
+```{r scbi-spatial-information, out.width="66%", fig.cap="Add spatial information: Buffer region and spatial cross-validation blocks (1 through 4).
+All trees in the interior of the study region will be the focal trees whose growth will modeled."}
 ggplot() +
-  geom_sf_text(data = growth_scbi, aes(label = foldID, col = buffer), alpha = 0.2) +
+  geom_sf_text(data = growth_scbi, aes(label = foldID, col = buffer)) +
   geom_sf(data = blocks_scbi, fill = "transparent")
 ```
-
-TODO: Do we talk about `add_buffer_variable(direction = "out")` when each fold gets its turn being the training data?
 
@@
 ```
+```{r scbi-focal-vs-comp-map, out.width="66%", echo = FALSE, fig.cap="Identify all focal and corresponding competitor trees: All 20 competitor trees of focal tree 4."}
+comp_radius <- growth_scbi %>% 
+  filter(stemID == 4) %>% 
+  st_buffer(dist = comp_dist)
+
+comp_radius_box <- growth_scbi %>% 
+  filter(stemID == 4) %>% 
+  st_buffer(dist = comp_dist + 10) %>% 
+  st_bbox() %>% 
+  st_as_sfc() %>% 
+  st_intersects(growth_scbi, ., sparse = FALSE)
+
+growth_scbi_subset <- growth_scbi %>% 
+  mutate(inside = comp_radius_box) %>% 
+  filter(inside)
+
+ggplot() +
+  geom_sf(data = growth_scbi_subset, aes(col = sp, size = dbh1)) +
+  geom_sf(data = comp_radius, linetype = "dashed", alpha = 0, inherit.aes = FALSE) +
+  coord_sf(xlim = st_bbox(comp_radius)[c("xmin", "xmax")], ylim = st_bbox(comp_radius)[c("ymin", "ymax")]) +
+  geom_sf_text(data = growth_scbi %>% filter(stemID == 4), aes(label = stemID), size = 12) +
+  labs(size = "DBH at first census", col = "species")
+```
 
@@
 
-```{r scbi-posterior-viz, out.width="100%", fig.cap="Posterior distributions of all parameters for a subset of species."}
+```{r scbi-posterior-viz, out.width="100%", fig.cap="Fit model: Posterior distributions of all parameters for three species."}
 # Plot posteriors for only a subset of species
-sp_to_plot <- c("quru", "litu", "cagl", "cato")
+sp_to_plot <- c("litu", "quru", "cagl")
 
@@
 ```
+
+TODO: Discuss meaning of results.
 
@@
 
+TODO: Do we talk about `add_buffer_variable(direction = "out")` when each fold gets its turn being the training data?
+
 The model fits and predictions in Section \ref{model-fit-predict} all suffer from a common failing: they use the same data to both fit the model and to assess the model's performance using the RMSE.
@@
 
-```{r scbi-spatial-cv, eval=TRUE}
+```{r scbi-spatial-cv}
 focal_vs_comp_scbi <- focal_vs_comp_scbi %>%
   run_cv(comp_dist = comp_dist, cv_grid = blocks_scbi)
+```
 
+```{r}
 focal_vs_comp_scbi %>%

### 2021-02-19 etc/paper/forestecology.tex
impletemented ideas from Mensh 2017

@@
 
-A high-level overview of the steps of our analysis pipeline is as
-follows:
+There are four steps in our modeling pipeline:
 
@@
 \item
-  Compute the growth of trees based on census data
+  Compute the growth of trees based on census data.
 \item
@@
 \item
-  Fit model, make predictions, and visualize model results
+  Fit model and make predictions.
 \end{enumerate}
 
-Additionally,
+Using these four steps, we can address the two following questions of
+ecological interest:
 
@@
 \CommentTok{##  5     79 tiam  47.7  M      A       46.8 M      -0.161    (40 381.1)}
-\CommentTok{## # ... with 7,949 more rows}
+\CommentTok{##  6     80 caca   5.15 M      A        6.5 M       0.253  (38.7 421.7)}
+\CommentTok{##  7     96 libe   2.3  J;M    A        3.7 M       0.262      (60 310)}
+\CommentTok{##  8    100 caca   5.09 NULL   A       NA   DN     NA      (52.5 476.3)}
+\CommentTok{##  9    101 litu  65.4  M      A       68.4 M       0.552  (47.1 567.3)}
+\CommentTok{## 10    102 astr   1.99 NULL   A        2.5 M       0.0954 (40.8 575.5)}
+\CommentTok{## # ... with 7,944 more rows}
 \end{Highlighting}
@@
 
-\caption{Growth of trees at SCBI.}\label{fig:scbi-trees}
+\caption{Compute growth of trees based on census data from the Smithsonian Conservation Biology Institute (SCBI).}\label{fig:scbi-trees}
 \end{figure}
@@
 \CommentTok{# Manually define spatial blocks to act as folds}
-\NormalTok{fold1 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{600}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{600}\NormalTok{))}
-\NormalTok{fold2 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+\NormalTok{fold1 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{450}\NormalTok{))}
+\NormalTok{fold2 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{450}\NormalTok{))}
+\NormalTok{fold3 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{600}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+\NormalTok{fold4 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+\NormalTok{n_fold <-}\StringTok{ }\DecValTok{4}
 
-\NormalTok{blocks_scbi <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}\KeywordTok{sf_polygon}\NormalTok{(fold1), }\KeywordTok{sf_polygon}\NormalTok{(fold2)) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
+\NormalTok{blocks_scbi <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}
+  \KeywordTok{sf_polygon}\NormalTok{(fold1), }\KeywordTok{sf_polygon}\NormalTok{(fold2), }\KeywordTok{sf_polygon}\NormalTok{(fold3), }\KeywordTok{sf_polygon}\NormalTok{(fold4)}
+\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\NormalTok{n_fold) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
 
@@
 \NormalTok{SpatialBlock_scbi <-}\StringTok{ }\KeywordTok{spatialBlock}\NormalTok{(}
-  \DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =} \DecValTok{2}\NormalTok{, }\DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{, }
+  \DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =}\NormalTok{ n_fold, }\DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{, }
   \DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}
@@
 \KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_sf_text}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_scbi, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{label =}\NormalTok{ foldID, }\DataTypeTok{col =}\NormalTok{ buffer), }\DataTypeTok{alpha =} \FloatTok{0.2}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{geom_sf_text}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_scbi, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{label =}\NormalTok{ foldID, }\DataTypeTok{col =}\NormalTok{ buffer)) }\OperatorTok{+}
 \StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ blocks_scbi, }\DataTypeTok{fill =} \StringTok{"transparent"}\NormalTok{)}
@@
 
-\caption{Buffer region and cross-validation block information for SCBI data.}\label{fig:scbi-spatial-information}
+\caption{Add spatial information: Buffer region and spatial cross-validation blocks (1 through 4).
+All trees in the interior of the study region will be the focal trees whose growth will modeled.}\label{fig:scbi-spatial-information}
 \end{figure}
-
-TODO: Do we talk about
-\texttt{add\_buffer\_variable(direction\ =\ "out")} when each fold gets
-its turn being the training data?
 
@@
 \CommentTok{##  5       96 libe      2.3      (60 310)  0.262  <tibble [14 x 4]>}
-\CommentTok{## # ... with 6,291 more rows}
+\CommentTok{##  6      101 litu     65.4  (47.1 567.3)  0.552  <tibble [19 x 4]>}
+\CommentTok{##  7      102 astr      1.99 (40.8 575.5)  0.0954 <tibble [44 x 4]>}
+\CommentTok{##  8      126 cato     37.4  (60.6 400.2)  0.165  <tibble [16 x 4]>}
+\CommentTok{##  9      127 caca      8.72 (72.7 514.1)  0.0370 <tibble [14 x 4]>}
+\CommentTok{## 10      139 astr      1.71 (96.7 315.1)  0.0549 <tibble [48 x 4]>}
+\CommentTok{## # ... with 6,286 more rows}
 \end{Highlighting}
@@
 \end{verbatim}
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.66\linewidth]{Figures/scbi-focal-vs-comp-map-1}
+
+}
+
+\caption{Identify all focal and corresponding competitor trees: All 20 competitor trees of focal tree 4.}\label{fig:scbi-focal-vs-comp-map}
+\end{figure}
 
@@
 \CommentTok{##  5       96 libe      2.3  1                       (60 310)  0.262 }
-\CommentTok{##  6      101 litu     65.4  1                   (47.1 567.3)  0.552 }
-\CommentTok{##  7      102 astr      1.99 1                   (40.8 575.5)  0.0954}
+\CommentTok{##  6      101 litu     65.4  3                   (47.1 567.3)  0.552 }
+\CommentTok{##  7      102 astr      1.99 3                   (40.8 575.5)  0.0954}
 \CommentTok{##  8      126 cato     37.4  1                   (60.6 400.2)  0.165 }
-\CommentTok{##  9      127 caca      8.72 1                   (72.7 514.1)  0.0370}
+\CommentTok{##  9      127 caca      8.72 3                   (72.7 514.1)  0.0370}
 \CommentTok{## 10      139 astr      1.71 1                   (96.7 315.1)  0.0549}
@@
 \CommentTok{# Plot posteriors for only a subset of species}
-\NormalTok{sp_to_plot <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"quru"}\NormalTok{, }\StringTok{"litu"}\NormalTok{, }\StringTok{"cagl"}\NormalTok{, }\StringTok{"cato"}\NormalTok{)}
+\NormalTok{sp_to_plot <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"litu"}\NormalTok{, }\StringTok{"quru"}\NormalTok{, }\StringTok{"cagl"}\NormalTok{)}
 
@@
 
-\caption{Posterior distributions of all parameters for a subset of species.}\label{fig:scbi-posterior-viz}
+\caption{Fit model: Posterior distributions of all parameters for three species.}\label{fig:scbi-posterior-viz}
 \end{figure}
+
+TODO: Discuss meaning of results.
 
@@
 
+TODO: Do we talk about
+\texttt{add\_buffer\_variable(direction\ =\ "out")} when each fold gets
+its turn being the training data?
+
 The model fits and predictions in Section \ref{model-fit-predict} all
@@
 \StringTok{  }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{cv_grid =}\NormalTok{ blocks_scbi)}
+\end{Highlighting}
+\end{Shaded}
 
+\begin{Shaded}
+\begin{Highlighting}[]
 \NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
@@
 \StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
-\CommentTok{## [1] 0.1446232}
+\CommentTok{## [1] 0.1402209}
 \end{Highlighting}

### 2021-02-22 etc/paper/forestecology.Rmd
Rewrite "permutation test" and "spatial cross-validation" sections

@@
 ---
-title: "The `forestecology` R package for modeling interspecies competition between trees"
+title: "The forestecology R package for fitting and assessing models of interspecies competition on the growth of trees"
 
@@
 - name: Albert Y. Kim 
-  thanks: "Albert Y. Kim is Assistant Professor, Statistical & Data Sciences, Smith College, Northampton, MA 01063 (e-mail: akim04@smith.edu)."
   affiliation: Program in Statistical & Data Sciences, Smith College
@@
 abstract: 
-- "Move abstract below here after completed."
+- |
+  | 1.
+When modeling growth of trees forest ecologists often incorporate the effect of interspecies competition.
+Many such models are based on a neighborhood effect assumption whereby all trees within a fixed distance of all focal trees are considered competitors.
+Methods are needed to evaluate the effect of interspecies competition and to assess their quality.
+  | 2.
+We present the `forestecology` package providing methods for both 1) evaluating the out-of-sample performance of our model using spatial-crossvaliation and 2) testing a null hypothesis of of no impact of competitor species' identity on the growth of trees using a permutation test.
+We implement a class and methods using R's S3 object-oriented system, for a specific linear, Bayesian neighborhood competition model of tree growth.
+  | 3.
+We demonstrate the package's functions using data from the Smithsonian Conservation Biology Institute's large forest dynamics plot, part of the ForestGEO network of reseach sites.
+Given ForestGEO's data collection protocols and data formatting standards, the package cross-compatibility of code.
+We show both that 1) competitor species identity matters and 2) that not spatially cross-validating leads to error estimates that are overly optimistic.
+  | 4.
+The package follows `tidyverse`-like structure whereby verb-named functions can be modularly "piped" in sequence to intuitively display the sequence of steps of analysis from start to finish.
+Additionally, most inputs/outputs of functions assume an are of `sf` class from the simple features package, thereby facilitating all wrangling and visualization of geospatial data.
+Lastly, even though our package is currently limited to one specific model, the package is setup such that it can be easily extended to other models.
 
 keywords:
-- forest ecology, competition, R, Rstats, tidyverse, sf, cross-validation,
+- forest ecology, competition, R, Rstats, tidyverse, sf, cross-validation, spatial statistics
 
@@
 
-# Abstract (350 words) {-}
-
-1.
-When modeling growth of trees forest ecologists often incorporate the effect of interspecies competition.
-Many such models are based on a neighborhood effect assumption whereby all trees within a fixed distance of all focal trees are considered competitors.
-Methods are needed to evaluate the effect of interspecies competition and to assess their quality.
-1.
-We present the `forestecology` package providing methods for both 1) evaluating the out-of-sample performance of our model using spatial-crossvaliation and 2) testing a null hypothesis of of no impact of competitor species' identity on the growth of trees using a permutation test.
-We implement a class and methods using R's S3 object-oriented system, for a specific linear, Bayesian neighborhood competition model of tree growth.
-1.
-We demonstrate the package's functions using data from the Smithsonian Conservation Biology Institute's large forest dynamics plot, part of the ForestGEO network of reseach sites.
-Given ForestGEO's data collection protocols and data formatting standards, the package cross-compatibility of code.
-We show both that 1) competitor species identity matters and 2) that not spatially cross-validating leads to error estimates that are overly optimistic.
-1.
-The package follows `tidyverse`-like structure whereby verb-named functions can be modularly "piped" in sequence to intuitively display the sequence of steps of analysis from start to finish.
-Additionally, most inputs/outputs of functions assume an are of `sf` class from the simple features package, thereby facilitating all wrangling and visualization of geospatial data.
-Lastly, even though our package is currently limited to one specific model, the package is setup such that it can be easily extended to other models.
-
 # Introduction
@@
 
-1. Compute the growth of trees based on census data.
-1. Add spatial information:
+- Step 1: Compute the growth of trees based on census data.
+- Step 2: Add spatial information:
     1. Define buffer region trees.
     1. Add spatial cross-validation block information.
-1. Identify all focal and corresponding competitor trees.
-1. Fit model and make predictions.
+- Step 3: Identify all focal and corresponding competitor trees.
+- Step 4: Fit model, as well as:
+    1. Compute fitted/predicted values.
+    1. Visualize posterior distributions.
 
@@
 
-## Compute the growth of trees based on census data
+## Step 1: Compute the growth of trees based on census data
 
@@
 
-```{r scbi-trees, out.width="66%", fig.cap="Compute growth of trees based on census data from the Smithsonian Conservation Biology Institute (SCBI)."}
+```{r scbi-trees, out.width="66%", fig.cap="Step 1 - Compute growth of trees based on census data: All trees from a 9 ha subsection of the Smithsonian Conservation Biology Institute (SCBI) forest plot."}
 ggplot() +
@@
 
-## Add spatial information {#spatial-information}
+## Step 2: Add spatial information {#spatial-information}
 
@@
 
-```{r scbi-spatial-information, out.width="66%", fig.cap="Add spatial information: Buffer region and spatial cross-validation blocks (1 through 4).
+```{r scbi-spatial-information, out.width="66%", fig.cap="Step 2 - Add spatial information: Buffer region and spatial cross-validation blocks (1 through 4).
 All trees in the interior of the study region will be the focal trees whose growth will modeled."}
 ggplot() +
-  geom_sf_text(data = growth_scbi, aes(label = foldID, col = buffer)) +
-  geom_sf(data = blocks_scbi, fill = "transparent")
+  geom_sf(data = blocks_scbi, fill = "transparent", linetype = "dashed") +
+  geom_sf_text(data = growth_scbi %>% sample_n(1000), aes(label = foldID, col = buffer))
 ```
 
-## Identify all focal and corresponding competitor trees {#focal-vs-comp}
+## Step 3: Identify all focal and corresponding competitor trees {#focal-vs-comp}
 
@@
 Saving our focal versus competitor information in list-column minimizes redundancy since we do not repeat information on the focal tree 20 times.
+The spatial distribution of these trees is visualized in Figure \ref{fig:scbi-focal-vs-comp-map}: the dashed circle extends `r comp_dist` m away from the focal tree while all 20 competitor trees are within this circle.
 
@@
 ```
-```{r scbi-focal-vs-comp-map, out.width="66%", echo = FALSE, fig.cap="Identify all focal and corresponding competitor trees: All 20 competitor trees of focal tree 4."}
+```{r scbi-focal-vs-comp-map, out.width="66%", echo = FALSE, fig.cap="Step 3 - Identify all focal and corresponding competitor trees: All 20 competitor trees of focal tree 4."}
 comp_radius <- growth_scbi %>% 
@@
 
-## Fit model and make predictions {#model-fit-predict}
+## Step 4: Fit model {#model-fit-predict}
 
@@
 
-Lastly, the generic for `ggplot2::autoplot()` allows us to
+Lastly, the generic for `ggplot2::autoplot()` allows us to plot the posterior distribution of all parameters in Figure \ref{fig:scbi-posterior-viz}.
 
-```{r scbi-posterior-viz, out.width="100%", fig.cap="Fit model: Posterior distributions of all parameters for three species."}
+```{r scbi-posterior-viz, out.width="100%", fig.cap= "Step 4 - Fit model: Posterior distributions of all parameters for three species."}
 # Plot posteriors for only a subset of species
@@
 
-HEY BERT PICK IT UP HERE
+We use the four steps of our analysis sequence we answer the first of our two questions: does competitor species identity matter when modeling the growth of focal trees?
+We answer this via a permutation test: Under a null hypothesis where competitor species identity does not matter we can permute this variable within each focal tree, compute the RMSE (the test statistic of interest), repeat this process several times to construct a null distribution of RMSE, and compare the observed RMSE to assess significance.
+Going back to our example in Section \ref{focal-vs-comp} inspecting all 20 competitor trees to focal tree with `focal_ID` 4, the permutation test involves randomly resampling the `comp_sp` variable with replacement, leaving all other competitor variables intact.
+The resampling with replacement is nested within each focal tree in order to preserve the neighborhood structure of our competitor model.
+To run the permutation test, we use the same code as in Section \ref{model-fit-predict}, but adding a `run_shuffle = TRUE` argument to `comp_bayes_lm()`.
 
-```{r scbi-permutation-model-fit, eval = FALSE}
-comp_bayes_lm_scbi <- focal_vs_comp_scbi %>%
+```{r scbi-permutation-model-fit}
+comp_bayes_lm_scbi_shuffle <- focal_vs_comp_scbi %>%
   comp_bayes_lm(prior_param = NULL, run_shuffle = TRUE)
@@
 focal_vs_comp_scbi <- focal_vs_comp_scbi %>%
-  mutate(
-    growth_hat = predict(comp_bayes_lm_scbi, newdata = focal_vs_comp_scbi)
-  )
+  mutate(growth_hat_shuffle = predict(comp_bayes_lm_scbi_shuffle, newdata = focal_vs_comp_scbi))
 
-model_rmse <- focal_vs_comp_scbi %>%
-  rmse(truth = growth, estimate = growth_hat) %>%
+model_rmse_shuffle <- focal_vs_comp_scbi %>%
+  rmse(truth = growth, estimate = growth_hat_shuffle) %>%
   pull(.estimate)
-model_rmse
+model_rmse_shuffle
 ```
+
+The resulting RMSE of `r model_rmse_shuffle` based on the permutation test is `r ifelse(model_rmse_shuffle < model_rmse, "smaller", "larger")` than the earlier RMSE of `r model_rmse`, suggesting that models that do incorporate competitor species identity better fit the data.
+We conduct a fuller simulation in Section below.
 
@@
 
-TODO: Do we talk about `add_buffer_variable(direction = "out")` when each fold gets its turn being the training data?
-
+Using the four steps of our analysis sequence again, we answer the second of our two questions: how can we obtain an accurate estimate of model performance/error?
 The model fits and predictions in Section \ref{model-fit-predict} all suffer from a common failing: they use the same data to both fit the model and to assess the model's performance using the RMSE.
 As argued by @roberts_cross-validation_2017, this can lead to overly optimistic assessments of model quality as the models can be overfit, in particular in situations where spatial-autocorrelation is present.
-To mitigate the effects of such overfitting, we use a spatially block cross-validation algorithm implemented in the `run_cv()`.
-This function at its core uses the same model fitting implemented in the `fit_bayesian_model()` function, however trains the model on $k-1$ spatial folds of the train and returns fitted values for the test data.
-Recall that the spatial blocking scheme wass encoded in Section \ref{spatial-information}.
+To mitigate the effects of such overfitting, we use a spatially block cross-validation algorithm.
 
-Observe once again that this RMSE is much higher than that for the above SCBI model fit without cross-validation.
+To this end, we use the `foldID` variable defined in Section \ref{spatial-information} whereby all focal trees are assigned to one of `r n_fold` spatially contiguous blocks that act as folds in our cross-validation routine.
+Figure \ref{fig:scbi-spatial-cross-validation-schematic} presents a schematic illustrating this scheme for fold 1 as the test set and folds 2, 3, and 4 as the training sets.
+We fit the model to all focal trees in the training set, apply the model to all focal trees in the test set to compute fitted/predicted values, and compute the RMSE of the observed versus predicted growths.
+We repeat this procedure 3 more times with each of the three remaining folds acting as the test set.
+Furthermore, in order to maintain spatial independence between the test and training set, a buffer that extend outwards from the boundary of the test set is computed; all trees falling within this buffer will not be part of the training set.
+
+```{r scbi-spatial-cross-validation-schematic, out.width="66%", fig.cap= "Schematic of spatial cross-validation: Using the k = 1 fold as the test set, assigning each focal tree to training set, test set, and buffer.", echo = FALSE}
+fold_number <- 1
+
+test_fold_buffer <-  blocks_scbi %>% 
+  filter(folds == fold_number)
+
+test_vs_train <- focal_vs_comp_scbi %>% 
+  st_as_sf() %>% 
+  add_buffer_variable(
+    direction = "out", size = comp_dist, region = test_fold_buffer
+  ) %>% 
+  mutate(
+    fold = case_when(
+      foldID == fold_number ~ "test",
+      foldID != fold_number & buffer ~ "training",
+      foldID != fold_number & !buffer ~ "buffer"
+    ),
+    fold = factor(fold, levels = c("test", "buffer", "training"))
+  )
+
+ggplot() +
+  geom_sf(data = test_fold_buffer, col = "red", fill = "transparent") +
+  geom_sf(data = blocks_scbi, fill = "transparent", linetype = "dashed") +
+  geom_sf_text(data = test_vs_train %>% st_as_sf() %>% sample_n(1000), aes(label = foldID, col = fold))
+```
+
+This algorithm is implemented in the `run_cv()` function, which acts as a wrapper function to the `comp_bayes_lm()` function that fits the model and the `predict()` generic that returns fitted values.
 
@@
 
+Observe once again that this RMSE is much higher than that for the above SCBI model fit without cross-validation.
+
 ```{r}
-focal_vs_comp_scbi %>%
+model_rmse_cv <- focal_vs_comp_scbi %>%
   rmse(truth = growth, estimate = growth_hat) %>%
   pull(.estimate)
+model_rmse_cv
 ```
 
+The resulting RMSE of `r model_rmse_cv` computed using cross-validation is `r ifelse(model_rmse_cv > model_rmse, "larger", "smaller")` than the earlier RMSE of `r model_rmse`, suggesting that models that do not take the inherent spatial autocorrelation of the data into account generate error estimates that are overly optimistic; in our case RMSE's that are too low.
+
 # Discussion
+
+- show image of PLOSOne
+- run time considerations
 

### 2021-02-22 etc/paper/forestecology.tex
Rewrite "permutation test" and "spatial cross-validation" sections

@@
 {
-  \title{\bf The \texttt{forestecology} R package for modeling interspecies
-competition between trees}
+  \title{\bf The forestecology R package for fitting and assessing models of
+interspecies competition on the growth of trees}
 
   \author{
-        Albert Y. Kim \thanks{Albert Y. Kim is Assistant Professor, Statistical \& Data Sciences,
-Smith College, Northampton, MA 01063 (e-mail:
-\href{mailto:akim04@smith.edu}{\nolinkurl{akim04@smith.edu}}).} \\
+        Albert Y. Kim \\
     Program in Statistical \& Data Sciences, Smith College\\
@@
   \begin{center}
-    {\LARGE\bf The \texttt{forestecology} R package for modeling interspecies
-competition between trees}
+    {\LARGE\bf The forestecology R package for fitting and assessing models of
+interspecies competition on the growth of trees}
   \end{center}
@@
 \begin{abstract}
-Move abstract below here after completed.
+1. When modeling growth of trees forest ecologists often incorporate the
+effect of interspecies competition. Many such models are based on a
+neighborhood effect assumption whereby all trees within a fixed distance
+of all focal trees are considered competitors. Methods are needed to
+evaluate the effect of interspecies competition and to assess their
+quality.\\
+2. We present the \texttt{forestecology} package providing methods for
+both 1) evaluating the out-of-sample performance of our model using
+spatial-crossvaliation and 2) testing a null hypothesis of of no impact
+of competitor species' identity on the growth of trees using a
+permutation test. We implement a class and methods using R's S3
+object-oriented system, for a specific linear, Bayesian neighborhood
+competition model of tree growth.\\
+3. We demonstrate the package's functions using data from the
+Smithsonian Conservation Biology Institute's large forest dynamics plot,
+part of the ForestGEO network of reseach sites. Given ForestGEO's data
+collection protocols and data formatting standards, the package
+cross-compatibility of code. We show both that 1) competitor species
+identity matters and 2) that not spatially cross-validating leads to
+error estimates that are overly optimistic.\\
+4. The package follows \texttt{tidyverse}-like structure whereby
+verb-named functions can be modularly ``piped'' in sequence to
+intuitively display the sequence of steps of analysis from start to
+finish. Additionally, most inputs/outputs of functions assume an are of
+\texttt{sf} class from the simple features package, thereby facilitating
+all wrangling and visualization of geospatial data. Lastly, even though
+our package is currently limited to one specific model, the package is
+setup such that it can be easily extended to other models.
 \end{abstract}
@@
 {\it Keywords:} forest ecology, competition, R, Rstats, tidyverse, sf, cross-validation,
+spatial statistics
 \vfill
@@
 \linenumbers
-
-\hypertarget{abstract-350-words}{%
-\section*{Abstract (350 words)}\label{abstract-350-words}}
-\addcontentsline{toc}{section}{Abstract (350 words)}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  When modeling growth of trees forest ecologists often incorporate the
-  effect of interspecies competition. Many such models are based on a
-  neighborhood effect assumption whereby all trees within a fixed
-  distance of all focal trees are considered competitors. Methods are
-  needed to evaluate the effect of interspecies competition and to
-  assess their quality.
-\item
-  We present the \texttt{forestecology} package providing methods for
-  both 1) evaluating the out-of-sample performance of our model using
-  spatial-crossvaliation and 2) testing a null hypothesis of of no
-  impact of competitor species' identity on the growth of trees using a
-  permutation test. We implement a class and methods using R's S3
-  object-oriented system, for a specific linear, Bayesian neighborhood
-  competition model of tree growth.
-\item
-  We demonstrate the package's functions using data from the Smithsonian
-  Conservation Biology Institute's large forest dynamics plot, part of
-  the ForestGEO network of reseach sites. Given ForestGEO's data
-  collection protocols and data formatting standards, the package
-  cross-compatibility of code. We show both that 1) competitor species
-  identity matters and 2) that not spatially cross-validating leads to
-  error estimates that are overly optimistic.
-\item
-  The package follows \texttt{tidyverse}-like structure whereby
-  verb-named functions can be modularly ``piped'' in sequence to
-  intuitively display the sequence of steps of analysis from start to
-  finish. Additionally, most inputs/outputs of functions assume an are
-  of \texttt{sf} class from the simple features package, thereby
-  facilitating all wrangling and visualization of geospatial data.
-  Lastly, even though our package is currently limited to one specific
-  model, the package is setup such that it can be easily extended to
-  other models.
-\end{enumerate}
 
@@
 
+\begin{itemize}
+\tightlist
+\item
+  Step 1: Compute the growth of trees based on census data.
+\item
+  Step 2: Add spatial information:
+
   \begin{enumerate}
   \def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Compute the growth of trees based on census data.
-\item
-  Add spatial information:
-
-  \begin{enumerate}
-  \def\labelenumii{\arabic{enumii}.}
   \tightlist
@@
 \item
-  Identify all focal and corresponding competitor trees.
+  Step 3: Identify all focal and corresponding competitor trees.
 \item
-  Fit model and make predictions.
+  Step 4: Fit model, as well as:
+
+  \begin{enumerate}
+  \def\labelenumi{\arabic{enumi}.}
+  \tightlist
+  \item
+    Compute fitted/predicted values.
+  \item
+    Visualize posterior distributions.
   \end{enumerate}
+\end{itemize}
 
@@
 
-\hypertarget{compute-the-growth-of-trees-based-on-census-data}{%
-\subsection{Compute the growth of trees based on census
-data}\label{compute-the-growth-of-trees-based-on-census-data}}
+\hypertarget{step-1-compute-the-growth-of-trees-based-on-census-data}{%
+\subsection{Step 1: Compute the growth of trees based on census
+data}\label{step-1-compute-the-growth-of-trees-based-on-census-data}}
 
@@
 
-\caption{Compute growth of trees based on census data from the Smithsonian Conservation Biology Institute (SCBI).}\label{fig:scbi-trees}
+\caption{Step 1 - Compute growth of trees based on census data: All trees from a 9 ha subsection of the Smithsonian Conservation Biology Institute (SCBI) forest plot.}\label{fig:scbi-trees}
 \end{figure}
@@
 \hypertarget{spatial-information}{%
-\subsection{Add spatial information}\label{spatial-information}}
+\subsection{Step 2: Add spatial information}\label{spatial-information}}
 
@@
 \KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_sf_text}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_scbi, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{label =}\NormalTok{ foldID, }\DataTypeTok{col =}\NormalTok{ buffer)) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ blocks_scbi, }\DataTypeTok{fill =} \StringTok{"transparent"}\NormalTok{)}
+\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ blocks_scbi, }\DataTypeTok{fill =} \StringTok{"transparent"}\NormalTok{, }\DataTypeTok{linetype =} \StringTok{"dashed"}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{geom_sf_text}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{sample_n}\NormalTok{(}\DecValTok{1000}\NormalTok{), }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{label =}\NormalTok{ foldID, }\DataTypeTok{col =}\NormalTok{ buffer))}
 \end{Highlighting}
@@
 
-\caption{Add spatial information: Buffer region and spatial cross-validation blocks (1 through 4).
+\caption{Step 2 - Add spatial information: Buffer region and spatial cross-validation blocks (1 through 4).
 All trees in the interior of the study region will be the focal trees whose growth will modeled.}\label{fig:scbi-spatial-information}
@@
 \hypertarget{focal-vs-comp}{%
-\subsection{Identify all focal and corresponding competitor
+\subsection{Step 3: Identify all focal and corresponding competitor
 trees}\label{focal-vs-comp}}
@@
 information in list-column minimizes redundancy since we do not repeat
-information on the focal tree 20 times.
+information on the focal tree 20 times. The spatial distribution of
+these trees is visualized in Figure \ref{fig:scbi-focal-vs-comp-map}:
+the dashed circle extends 7.5 m away from the focal tree while all 20
+competitor trees are within this circle.
 
@@
 
-\caption{Identify all focal and corresponding competitor trees: All 20 competitor trees of focal tree 4.}\label{fig:scbi-focal-vs-comp-map}
+\caption{Step 3 - Identify all focal and corresponding competitor trees: All 20 competitor trees of focal tree 4.}\label{fig:scbi-focal-vs-comp-map}
 \end{figure}
@@
 \hypertarget{model-fit-predict}{%
-\subsection{Fit model and make predictions}\label{model-fit-predict}}
+\subsection{Step 4: Fit model}\label{model-fit-predict}}
 
@@
 
-Lastly, the generic for \texttt{ggplot2::autoplot()} allows us to
+Lastly, the generic for \texttt{ggplot2::autoplot()} allows us to plot
+the posterior distribution of all parameters in Figure
+\ref{fig:scbi-posterior-viz}.
 
@@
 
-\caption{Fit model: Posterior distributions of all parameters for three species.}\label{fig:scbi-posterior-viz}
+\caption{Step 4 - Fit model: Posterior distributions of all parameters for three species.}\label{fig:scbi-posterior-viz}
 \end{figure}
@@
 
-HEY BERT PICK IT UP HERE
+We use the four steps of our analysis sequence we answer the first of
+our two questions: does competitor species identity matter when modeling
+the growth of focal trees? We answer this via a permutation test: Under
+a null hypothesis where competitor species identity does not matter we
+can permute this variable within each focal tree, compute the RMSE (the
+test statistic of interest), repeat this process several times to
+construct a null distribution of RMSE, and compare the observed RMSE to
+assess significance. Going back to our example in Section
+\ref{focal-vs-comp} inspecting all 20 competitor trees to focal tree
+with \texttt{focal\_ID} 4, the permutation test involves randomly
+resampling the \texttt{comp\_sp} variable with replacement, leaving all
+other competitor variables intact. The resampling with replacement is
+nested within each focal tree in order to preserve the neighborhood
+structure of our competitor model. To run the permutation test, we use
+the same code as in Section \ref{model-fit-predict}, but adding a
+\texttt{run\_shuffle\ =\ TRUE} argument to \texttt{comp\_bayes\_lm()}.
 
@@
 \begin{Highlighting}[]
-\NormalTok{comp_bayes_lm_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
+\NormalTok{comp_bayes_lm_scbi_shuffle <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
 \StringTok{  }\KeywordTok{comp_bayes_lm}\NormalTok{(}\DataTypeTok{prior_param =} \OtherTok{NULL}\NormalTok{, }\DataTypeTok{run_shuffle =} \OtherTok{TRUE}\NormalTok{)}
@@
 \NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
-    \DataTypeTok{growth_hat =} \KeywordTok{predict}\NormalTok{(comp_bayes_lm_scbi, }\DataTypeTok{newdata =}\NormalTok{ focal_vs_comp_scbi)}
-\NormalTok{  )}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{growth_hat_shuffle =} \KeywordTok{predict}\NormalTok{(comp_bayes_lm_scbi_shuffle, }\DataTypeTok{newdata =}\NormalTok{ focal_vs_comp_scbi))}
 
-\NormalTok{model_rmse <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
+\NormalTok{model_rmse_shuffle <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat_shuffle) }\OperatorTok{%>%}
 \StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
-\NormalTok{model_rmse}
+\NormalTok{model_rmse_shuffle}
+\CommentTok{## [1] 0.131693}
 \end{Highlighting}
 \end{Shaded}
+
+The resulting RMSE of 0.131693 based on the permutation test is larger
+than the earlier RMSE of 0.1281398, suggesting that models that do
+incorporate competitor species identity better fit the data. We conduct
+a fuller simulation in Section below.
 
@@
 
-TODO: Do we talk about
-\texttt{add\_buffer\_variable(direction\ =\ "out")} when each fold gets
-its turn being the training data?
+Using the four steps of our analysis sequence again, we answer the
+second of our two questions: how can we obtain an accurate estimate of
+model performance/error? The model fits and predictions in Section
+\ref{model-fit-predict} all suffer from a common failing: they use the
+same data to both fit the model and to assess the model's performance
+using the RMSE. As argued by \citet{roberts_cross-validation_2017}, this
+can lead to overly optimistic assessments of model quality as the models
+can be overfit, in particular in situations where
+spatial-autocorrelation is present. To mitigate the effects of such
+overfitting, we use a spatially block cross-validation algorithm.
 
-The model fits and predictions in Section \ref{model-fit-predict} all
-suffer from a common failing: they use the same data to both fit the
-model and to assess the model's performance using the RMSE. As argued by
-\citet{roberts_cross-validation_2017}, this can lead to overly
-optimistic assessments of model quality as the models can be overfit, in
-particular in situations where spatial-autocorrelation is present. To
-mitigate the effects of such overfitting, we use a spatially block
-cross-validation algorithm implemented in the \texttt{run\_cv()}. This
-function at its core uses the same model fitting implemented in the
-\texttt{fit\_bayesian\_model()} function, however trains the model on
-\(k-1\) spatial folds of the train and returns fitted values for the
-test data. Recall that the spatial blocking scheme wass encoded in
-Section \ref{spatial-information}.
+To this end, we use the \texttt{foldID} variable defined in Section
+\ref{spatial-information} whereby all focal trees are assigned to one of
+4 spatially contiguous blocks that act as folds in our cross-validation
+routine. Figure \ref{fig:scbi-spatial-cross-validation-schematic}
+presents a schematic illustrating this scheme for fold 1 as the test set
+and folds 2, 3, and 4 as the training sets. We fit the model to all
+focal trees in the training set, apply the model to all focal trees in
+the test set to compute fitted/predicted values, and compute the RMSE of
+the observed versus predicted growths. We repeat this procedure 3 more
+times with each of the three remaining folds acting as the test set.
+Furthermore, in order to maintain spatial independence between the test
+and training set, a buffer that extend outwards from the boundary of the
+test set is computed; all trees falling within this buffer will not be
+part of the training set.
 
-Observe once again that this RMSE is much higher than that for the above
-SCBI model fit without cross-validation.
+\begin{figure}
+
+{\centering \includegraphics[width=0.66\linewidth]{Figures/scbi-spatial-cross-validation-schematic-1}
+
+}
+
+\caption{Schematic of spatial cross-validation: Using the k = 1 fold as the test set, assigning each focal tree to training set, test set, and buffer.}\label{fig:scbi-spatial-cross-validation-schematic}
+\end{figure}
+
+This algorithm is implemented in the \texttt{run\_cv()} function, which
+acts as a wrapper function to the \texttt{comp\_bayes\_lm()} function
+that fits the model and the \texttt{predict()} generic that returns
+fitted values.
 
@@
 
+Observe once again that this RMSE is much higher than that for the above
+SCBI model fit without cross-validation.
+
 \begin{Shaded}
 \begin{Highlighting}[]
-\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
+\NormalTok{model_rmse_cv <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
 \StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
 \StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
+\NormalTok{model_rmse_cv}
 \CommentTok{## [1] 0.1402209}
@@
 
+The resulting RMSE of 0.1402209 computed using cross-validation is
+larger than the earlier RMSE of 0.1281398, suggesting that models that
+do not take the inherent spatial autocorrelation of the data into
+account generate error estimates that are overly optimistic; in our case
+RMSE's that are too low.
+
 \hypertarget{discussion}{%
 \section{Discussion}\label{discussion}}
+
+\begin{itemize}
+\tightlist
+\item
+  show image of PLOSOne
+\item
+  run time considerations
+\end{itemize}
 

### 2021-02-22 paper/forestecology.Rmd
Restructured paper folder
Done second pass of Section 2: Example

--- 2021-02-22 etc/paper/forestecology.Rmd
@@
 ---
-title: "The forestecology R package for fitting and assessing models of interspecies competition on the growth of trees"
+title: "The forestecology R package for fitting and assessing models of interspecies competitive effects on the growth of trees"
 
@@
   | 1.
-When modeling growth of trees forest ecologists often incorporate the effect of interspecies competition.
-Many such models are based on a neighborhood effect assumption whereby all trees within a fixed distance of all focal trees are considered competitors.
-Methods are needed to evaluate the effect of interspecies competition and to assess their quality.
+Many models for the growth of trees that incorporate the effect of interspecies competition are based on a neighborhood effect assumption whereby all trees within a fixed distance of all focal trees are considered competitors.
+Methods and tools are needed to quantify this competitive effect and assess the quality of all resulting models
   | 2.
-We present the `forestecology` package providing methods for both 1) evaluating the out-of-sample performance of our model using spatial-crossvaliation and 2) testing a null hypothesis of of no impact of competitor species' identity on the growth of trees using a permutation test.
-We implement a class and methods using R's S3 object-oriented system, for a specific linear, Bayesian neighborhood competition model of tree growth.
+We present the `forestecology` package providing methods for both 1) evaluating the effect of competitor species identity using permutation tests and 2) evaluating model performance using spatial cross-validation.
+Following @allen_permutation_2020, we implement a Bayesian linear regression competition model.
   | 3.
-We demonstrate the package's functions using data from the Smithsonian Conservation Biology Institute's large forest dynamics plot, part of the ForestGEO network of reseach sites.
-Given ForestGEO's data collection protocols and data formatting standards, the package cross-compatibility of code.
-We show both that 1) competitor species identity matters and 2) that not spatially cross-validating leads to error estimates that are overly optimistic.
+We demonstrate the package's functionality using data from the Smithsonian Conservation Biology Institute's large forest dynamics plot, part of the ForestGEO global network of reseach sites.
+Given ForestGEO's data collection protocols and data formatting standards, the package was designed with cross-site compatibility in mind.
+We demonstrate that both 1) competitor species identity matters and 2) that not spatially cross-validating leads to error estimates that are overly optimistic.
   | 4.
-The package follows `tidyverse`-like structure whereby verb-named functions can be modularly "piped" in sequence to intuitively display the sequence of steps of analysis from start to finish.
-Additionally, most inputs/outputs of functions assume an are of `sf` class from the simple features package, thereby facilitating all wrangling and visualization of geospatial data.
-Lastly, even though our package is currently limited to one specific model, the package is setup such that it can be easily extended to other models.
+The package features 1) `tidyverse`-like structure whereby verb-named functions can be modularly "piped" in sequence, 2) functions with standardized inputs/outputs of simple features `sf` package class, and 3) R S3 object-oriented implementation of the Bayesian linear regression model.
+These three facts allow for clear articulation of all the steps in the sequence of analysis and easy wrangling and visualization of the geospatial forestry data.
+Furthermore, while the package only has Bayesian linear regression implemented, the package was designed with extensibility to other methods in mind.
 
@@
 
-We demonstrate the `forestecology` package's features on the Smithsonian Conservation Biology Institute (SCBI) large forest dynamics plot, located at the Smithsonian's National Zoo and Conservation Biology Institute in Front Royal, VA, USA.
+We demonstrate the `forestecology` package's functionality on data from the Smithsonian Conservation Biology Institute (SCBI) large forest dynamics plot, located at the Smithsonian's National Zoo and Conservation Biology Institute in Front Royal, VA, USA.
 The 25.6 ha (640 x 400 m) plot is located at the intersection of three of the major physiographic provinces of the eastern US: the Blue Ridge, Ridge and Valley, and Piedmont provinces and is adjacent to the northern end of Shenandoah National Park.
@@
 
-There are four steps in our modeling pipeline:
+The `forestecology` package attempts has the following two ecological goals: 1) to evaluate the effect of competitor species identity using permutation tests and 2) to evaluate model performance using spatial cross-validation.
+To achieve these two goals, we outline a basic analysis sequence comprising of these four main steps:
 
-- Step 1: Compute the growth of trees based on census data.
-- Step 2: Add spatial information:
-    1. Define buffer region trees.
+1. Compute the growth of trees based on census data.
+1. Add spatial information:
+    1. Define a buffer region of trees.
     1. Add spatial cross-validation block information.
-- Step 3: Identify all focal and corresponding competitor trees.
-- Step 4: Fit model, as well as:
+1. Identify all focal trees and their competitors.
+1. Apply model, which includes:
+    1. Fit model.
     1. Compute fitted/predicted values.
@@
 
-Using these four steps, we can address the two following questions of ecological interest:
-
-1. Evaluate the effect of competitor species identity using permutation tests.
-1. Evaluate model performance using spatial cross-validation.
-
-We load all necessary packages.
+We start by loading all necessary packages.
 
@@
 library(sf)
+library(patchwork)
 library(forestecology)
 library(blockCV)
-library(patchwork)
 ```
@@
 The first step in the our analysis sequence is to compute the growth of trees using data from two censuses.
-The `compute_growth()` function computes growth assuming census data that follows ForestGEO standards.
-Despite such standards, minor variations will still exist between sites thereby necessitating some data wrangling and checking.
+The `compute_growth()` function computes growth assuming census data that roughly follows ForestGEO standards.
+Despite such standards, minor variations will still exist between sites, thereby necessitating some data wrangling and checking.
 For example, the SCBI site records all DBH's in millimeters, whereas the Michigan Big Woods site records them in centimeters @andersonteixeira_ctfs-forestgeo_2015 @allen_michigan_2020.
-The data format of other sites may be such that our `compute_growth()` function doesn't work at all.
-However, in the end all that matters is that the growth of all trees is saved in a data frame of class `sf` whereby the geolocation of each tree is presented in a `geometry` variable of type `<POINT>` and at a minimum the data contains the following variables: a variable uniquely identifying each tree-stem, `sp` of type `fct` factor identifying species, `dbh1` and `dbh2` of type `dbl` quantifying the DBH at earlier and later census, and `growth` of type `dbl` double quantifying the average annual growth in centimeters.
 
 We load both 2008 and 2014 SCBI census data `.csv` files as they existed on GitHub on November 20, 2020.
-After selecting only relevant variables, we perform a few additional data wrangling steps: convert the variable with the date of measurement to be of type `date`, convert DBH to be in centimeters^[A rule of thumb to determine the units of DBH is check if the smallest non-zero and non-missing measurement is 1 or 10.
+After selecting only the relevant variables, we perform a few additional data wrangling steps: convert the character variable with the date of measurement to be of explicit type `date`, convert DBH to be in centimeters^[A rule of thumb to determine the units of DBH is to check if the smallest non-zero and non-missing measurement is 1 or 10.
 If the former, then centimeters.
 If the later, then millimeters.
-This is because ForestGEO protocols state that only trees with DBH greater or equal to 1cm should be included in censuses.], convert the `sp` variable containing species information from type `chr` character to `fct` factor (we will discuss the need for this in Section \ref{spatial-cross-validation}).
-Furthermore, in order to speed up computation for purposes of this example, we only consider a 9 ha subsection of the 25.6 ha of the SCBI site: `gx` from 0--300 instead of 0--400 and `gy` from 300--600 instead of 0--640.
+This is because ForestGEO protocols state that only trees with DBH greater or equal to 1cm should be included in censuses.], convert the `sp` variable containing species information from type `chr` character to `fct` factor.^[In our spatial cross-validation algorithm in Section \ref{spatial-cross-validation} issues can occur when rare species do not occur in the training set, but then are encountered in the test set.
+This risk is mitigated by representing `sp` as a factor variable, which has a complete list of all levels of the categorical variable.]Furthermore, in order to speed up computation for purposes of this example, we only consider a 9 ha subsection of the 25.6 ha of the SCBI site: `gx` from 0--300 instead of 0--400 and `gy` from 300--600 instead of 0--640.
 
@@
 
-These two data frames are then used as the two primary arguments to the `compute_growth()` function, along with the `id` argument whereby the user specifies the name of the variable that uniquely identifies each tree-stem under consideration (note this does not include resprouts in the later census):
+These two data frames are then used as arguments to the `compute_growth()` function, along with the `id` argument that specifies the variable that uniquely identifies each tree-stem.
+Note furthermore that we discard all resprouts in the later census (those with `code == R`).
 
@@
 
-The output `growth_scbi` is a single data frame of class `sf` that includes a numerical `growth` reflecting the average annual growth in DBH (in cm) for all trees that were alive at both time points as well a `geometry` variable encoding each tree's geolocation.
-Furthermore, variables that (in theory) remain unchanged between censuses appear only once, such as location variables `gx` and `gy`; as well as species-related variables.
+The output `growth_scbi` is a single data frame of class `sf` that includes `growth`, the average annual growth in DBH (in cm) for all trees that were alive at both time points, as well a `geometry` variable encoding each tree's geolocation.
+Furthermore, the variables that should remain unchanged between censuses appear only once, such as location variables `gx` and `gy`; as well as species-related variables.
 Variables that should change between censuses are suffixed with `1` and `2` indicating the earlier and later censuses, such as `dbh1/dbh2` and `codes1/codes2`.
+
+The data format of other sites may be such that our `compute_growth()` function doesn't work at all.
+However, in the end all that matters is that the growth of all trees is saved in a data frame of class `sf` whereby the geolocation of each tree is presented in a `geometry` variable of type `<POINT>` and at a minimum the data contains the variables above.
 
@@
 
-TODO: Rescale points in this plot:
-
-```{r scbi-trees, out.width="66%", fig.cap="Step 1 - Compute growth of trees based on census data: All trees from a 9 ha subsection of the Smithsonian Conservation Biology Institute (SCBI) forest plot."}
+```{r scbi-trees, out.width="66%", fig.cap="Compute growth of trees based on census data: Map with growth of all trees from a 9 ha subsection of the Smithsonian Conservation Biology Institute (SCBI) forest plot."}
+# TODO: Rescale points in this plot
 ggplot() +
@@
 The first element of spatial information we add is a "buffer region" to the periphery of the study region.
-Since some of our model's explanatory variables such as competitor basal area are cumulative, we must ensure that all trees being modeled are not biased to have different neighbor structures.
-This is of particular concern for trees at the boundary of study regions, which will not have the same number of neighbors that act as competitors as trees in the internal part of the study region.
-In order to account for such edge effects only trees who are not part of this buffer region, i.e.
+Since some of our model's explanatory variables are cumulative (such as competitor basal area), we must ensure that all trees being modeled are not biased to have different neighbor structures.
+This is of concern for trees at the boundary of study regions, which will not have the same number of neighbors as trees in the internal part of the study region.
+In order to account for such edge effects, only trees that are not part of this buffer region, i.e.
 are part of the interior of the study region, will have their growths modeled @waller_applied_2004.
@@
 Our model of interspecific competition relies on a spatial definition of who the competitor trees are for focal trees of interest: all trees within a distance `comp_dist` of a focal tree are considered its competitors (assuming the same units as the `gx` and `gy` location variables).
-In our case we set this value below at 7.5m, a value informed by @canham_neighborhood_2004 @uriarte_spatially_2004 @canham_neighborhood_2006.
-Using this value along with a manually constructed `sf` object representation of the study region's boundary, we apply the `add_buffer_variable()` to our `growth_scbi` data frame to add a `buffer` boolean variable: all trees who have `buffer` set to `FALSE` will be our focal trees whose growths are modeled, whereas those with `buffer` set to `TRUE` will only be considered as competitor trees whose growth will not be modeled.
+In our case we set this value at 7.5m, a value informed by @canham_neighborhood_2004 @uriarte_spatially_2004 @canham_neighborhood_2006.
+Using this value along with a manually constructed `sf` object representation of the study region's boundary, we apply the `add_buffer_variable()` to our `growth_scbi` data frame to add a `buffer` boolean variable: all trees who have `buffer` set to `FALSE` will be our focal trees whose growths are modeled, whereas those with `TRUE` will only be considered as competitor trees whose growth will not be modeled.
 
@@
 
-The second element of spatial information are blocks corresponding to folds of a spatial cross-validation algorithm used to estimate model error.
+The second element of spatial information are blocks corresponding to folds of a spatial cross-validation algorithm used to estimate out-of-sample model error.
 Conventional cross-validation algorithms assign observations to folds by randomly resampling individual observations.
-However, underlying this algorithm is an assumption that the observations are independent.
+However, an assumption that the observations are independent underlies any cross-validation algorithm.
 In the case of forest census data, observations exhibit spatial autocorrelation.
-This spatial dependence is incorporated into the cross-validation algorithm by randomly resampling spatial blocks of trees @roberts_cross-validation_2017 @pohjankukka_estimating_2017.
-We therefore associate each observed tree to one of $k$ spatial folds.
-In the example below, we first manually define two folds that partition the study region as an `sf` object.
+We therefore incorporate this spatial dependence into the cross-validation algorithm by randomly resampling spatial blocks of trees @roberts_cross-validation_2017 @pohjankukka_estimating_2017.
+
+In the example below, we first manually define four folds that partition the study region as an `sf` object.
 We then use the output of the `spatialBlock()` function from the `blockCV` package to associate each tree in `growth_scbi` to the correct fold (saved in the `foldID` variable) @valavi_blockcv_2019.
-\footnote{In the Appendix we present an example where the folds themselves are also created using the `spatialBlock()` function given a specified `cv_block_size`.}
+^[In the Appendix we present an example where the folds themselves are also created using the `spatialBlock()` function given a specified `cv_block_size`.]
 
@@
 # Manually define spatial blocks to act as folds
+n_fold <- 4
 fold1 <- rbind(c(0, 300), c(150, 300), c(150, 450), c(0, 450))
@@
 fold4 <- rbind(c(150, 450), c(300, 450), c(300, 600), c(150, 600))
-n_fold <- 4
 
@@
 
-```{r scbi-spatial-information, out.width="66%", fig.cap="Step 2 - Add spatial information: Buffer region and spatial cross-validation blocks (1 through 4).
-All trees in the interior of the study region will be the focal trees whose growth will modeled."}
+```{r scbi-spatial-information, out.width="66%", fig.cap="Add spatial information: Buffer region and spatial cross-validation blocks (1 through 4).
+All trees in the interior of the study region (i.e.
+not part of buffer) will be the focal trees whose growth will be modeled."}
 ggplot() +
@@
 The next step in our analysis sequence is to identify all focal trees and their corresponding competitor trees.
-The `create_focal_vs_comp()` functions performs these tasks and returns a new data frame of type `sf` containing this information.
-On top of the previously discussed arguments `comp_dist` defining the competition neighborhood and `id` indicating which variable in the data frame uniquely identifies each tree-stem, this function also requires an `sf` object representation of the spatial cross-validation blocks/folds; in our case, this was manually encoded in the `blocks_scbi` in Section \ref{spatial-information} while in our Appendix we present an example where this was performed using `spatialBlock()` from the `blockCV` package.
+The `create_focal_vs_comp()` functions performs these tasks and returns a new data frame of type `sf`.
+On top of the previous arguments `comp_dist` defining the competition neighborhood and `id` indicating which variable uniquely identifies each tree-stem, this function also requires an `sf` object representation of the spatial cross-validation blocks/folds; in our case, this was manually encoded in the `blocks_scbi` in Section \ref{spatial-information} while in our Appendix we present an example where this was performed using `spatialBlock()` from the `blockCV` package.
 We present the resulting data frame below with the `foldID` variable omitted for compactness of presentation.
@@
 ```{r scbi-focal-vs-comp}
+# TODO: Below reconcile the number of rows as they off by one from growth_scbi %>% 
+# filter(!is.na(growth) & !buffer). Perhaps by removing NA's in the growth_scbi stage?
 focal_vs_comp_scbi <- growth_scbi %>%
@@
 
-TODO: Below reconcile the number of rows as they off by one from `growth_scbi` pipe `filter(!is.na(growth) & !buffer)`. perhaps by removing NA's in the `growth_scbi` stage.
-
-The resulting data frame `focal_vs_comp_scbi` has `r nrow(focal_vs_comp_scbi)` rows, representing the subset of the `r nrow(growth_scbi)` trees in `growth_scbi` that will be considered as focal trees and thus have their growths modeled.
-Recall from Section \ref{spatial-information} this consists all trees that are not part of the buffer region in Figure \ref{fig:scbi-spatial-information}.
-Two new variables `focal_ID` and `focal_sp` related to tree-stem identification and species information.
-Most notably however is a new variable `comp` which contains information on all competitor trees for a given focal tree saved in list-column format, a feature of the `tidyr` package @tidyr_package.
-For example, we drill-down on the tree with `focal_ID` equal to 4.
-It has 20 competitor trees each described by 4 variables as indicated by the fact that `comp` is a `<tibble [20 × 4]>`.
+The resulting data frame `focal_vs_comp_scbi` has `r nrow(focal_vs_comp_scbi)` rows, representing the subset of the `r nrow(growth_scbi)` trees in `growth_scbi` that will be considered as focal trees.
+Two new variables `focal_ID` and `focal_sp` relate to tree-stem identification and species information.
+Most notably however is a new variable `comp` which contains information on all competitor trees for a given focal tree, saved in `tidyr` package list-column format @tidyr_package.
+For example, we drill-down on the tree with `focal_ID` 4, which has 20 competitor trees each described by 4 variables as indicated by the fact that `comp` is a `<tibble [20 × 4]>`.
 
@@
 
-Using the `unnest()` function from the `tidyr` package, we can flatten list-column into regular columns.
-We observe that for the same focal tree with DBH equal to 13.65cm, we have information on all 20 competitor trees whose `dist` distance to the focal tree is less than or equal to `r comp_dist`, including their unique tree-stem ID number, their species, and their basal area (in m$^2$) calculated as $\frac{\pi \times (\text{DBH/2})^2}{10000}$ where $DBH$ is the value from the earlier census in cm.
-Saving our focal versus competitor information in list-column minimizes redundancy since we do not repeat information on the focal tree 20 times.
 The spatial distribution of these trees is visualized in Figure \ref{fig:scbi-focal-vs-comp-map}: the dashed circle extends `r comp_dist` m away from the focal tree while all 20 competitor trees are within this circle.
 
-```{r, eval = FALSE}
-focal_vs_comp_scbi %>% 
-  filter(focal_ID == 4) %>% 
-  select(focal_ID, dbh, comp) %>% 
-  unnest(cols = "comp")
-```
-```{r scbi-focal-vs-comp-3, echo = FALSE}
-# Not sure why we need to do things this way:
-focal_vs_comp_scbi %>% 
-  filter(focal_ID == 4) %>% 
-  select(focal_ID, dbh, comp) %>% 
-  unnest(cols = "comp") %>% 
-  print(n = 5)
-```
-```{r scbi-focal-vs-comp-map, out.width="66%", echo = FALSE, fig.cap="Step 3 - Identify all focal and corresponding competitor trees: All 20 competitor trees of focal tree 4."}
+```{r scbi-focal-vs-comp-map, out.width="66%", echo = FALSE, fig.cap="Identify all focal and corresponding competitor trees: All 20 competitor trees of focal tree 4."}
 comp_radius <- growth_scbi %>% 
@@
 
+Using the `unnest()` function from the `tidyr` package, we can flatten list-column into regular columns.
+We observe that for the same focal tree, we have information on all 20 competitor trees whose `dist` distance to the focal tree is $\leq$ `r comp_dist`: their unique tree-stem ID number, their species, and their basal area (in m$^2$) calculated as $\frac{\pi \times (\text{DBH/2})^2}{10000}$ where $DBH$ is the value from the earlier of the two censuses in cm.
+Saving our focal versus competitor information in list-column minimizes redundancy since we do not repeat information on the focal tree 20 times.
+
+```{r, eval = FALSE}
+focal_vs_comp_scbi %>% 
+  filter(focal_ID == 4) %>% 
+  select(focal_ID, dbh, comp) %>% 
+  unnest(cols = "comp")
+```
+```{r scbi-focal-vs-comp-3, echo = FALSE}
+# Not sure why we need to do things this way:
+focal_vs_comp_scbi %>% 
+  filter(focal_ID == 4) %>% 
+  select(focal_ID, dbh, comp) %>% 
+  unnest(cols = "comp") %>% 
+  print(n = 5)
+```
+
 ## Step 4: Fit model {#model-fit-predict}
 
-Now that we've identified all focal and corresponding competitor trees and saved this information in a data frame of type `focal_vs_comp`, the next step in our analysis sequence is to fit a model for the growth of all focal trees.
+Now that we've identified all focal and corresponding competitor trees and saved this information in a data frame of type `focal_vs_comp`, the final step in our analysis sequence is to fit a model for the growth of all focal trees.
 Currently the `forestecology` package can only fit the competition Bayesian linear regression model outlined in Section \ref{competition-model} using the `comp_bayes_lm()` function.
 However, any model implemented in a function that similarly takes an input data frame of type `focal_vs_comp` as an argument can be easily swapped in.
-In the case of our specific competition Bayesian linear regression model, we can also specify prior distributions on all parameters of interest (here chosen to be the defaults).
+For our specific competition Bayesian linear regression model, we also specify prior distributions on all parameters of interest (here chosen to be the defaults as specified in `?comp_bayes_lm`).
 
 ```{r scbi-model-fit}
+# TODO: Add information about default priors in ?comp_bayes_lm()
 comp_bayes_lm_scbi <- focal_vs_comp_scbi %>%
@@
 This class of object includes generic methods implemented for `print()`, `predict()`, and `ggplot2::autoplot()`.
-First the generic for `print()` displays the names of all prior & posterior parameters (accessible via `comp_bayes_lm_scbi$prior_params` and `comp_bayes_lm_scbi$post_params`) along with the model formula:
+First the generic for `print()` displays the names of all prior & posterior parameters along with the model formula:
 
@@
 
-Next, the generic for `predict()` takes as inputs the posterior parameter values in `comp_bayes_lm_scbi` and the predictor variables in `newdata` and outputs a vector of fitted/predicted values $\widehat{y}$ of the DBH for each focal tree (as returned by the posterior predictive distribution).
+Next, the generic for `predict()` takes as inputs the posterior parameter values in `comp_bayes_lm_scbi` and the predictor variables in `newdata` and outputs a vector of fitted/predicted values $\widehat{y}$ of the DBH for each focal tree computed from the posterior predictive distribution.
 
@@
 
-We can then compare the observed and fitted/predicted growths to compute the root mean squared error of our model fit.
+We then compare the observed and fitted/predicted growths to compute the root mean squared error (RMSE) of our model fit.
 
@@
 
-Lastly, the generic for `ggplot2::autoplot()` allows us to plot the posterior distribution of all parameters in Figure \ref{fig:scbi-posterior-viz}.
+Lastly, the generic for `ggplot2::autoplot()` allows us to plot the posterior distribution of all parameters in Figure \ref{fig:scbi-posterior-viz} (for compactness we only show posteriors for 3 species).
 
-```{r scbi-posterior-viz, out.width="100%", fig.cap= "Step 4 - Fit model: Posterior distributions of all parameters for three species."}
+```{r scbi-posterior-viz, out.width="100%", fig.cap= "Fit model: Posterior distributions of all parameters for three species."}
 # Plot posteriors for only a subset of species
@@
 
-We use the four steps of our analysis sequence we answer the first of our two questions: does competitor species identity matter when modeling the growth of focal trees?
-We answer this via a permutation test: Under a null hypothesis where competitor species identity does not matter we can permute this variable within each focal tree, compute the RMSE (the test statistic of interest), repeat this process several times to construct a null distribution of RMSE, and compare the observed RMSE to assess significance.
-Going back to our example in Section \ref{focal-vs-comp} inspecting all 20 competitor trees to focal tree with `focal_ID` 4, the permutation test involves randomly resampling the `comp_sp` variable with replacement, leaving all other competitor variables intact.
-The resampling with replacement is nested within each focal tree in order to preserve the neighborhood structure of our competitor model.
-To run the permutation test, we use the same code as in Section \ref{model-fit-predict}, but adding a `run_shuffle = TRUE` argument to `comp_bayes_lm()`.
+In order to evaluate the effect of competitor species identity, we use the four steps of our analysis sequence answer along with a permutation test: Under a null hypothesis where competitor species identity does not matter, we can permute/shuffle this variable within each focal tree, compute the RMSE (the test statistic of interest), repeat this process several times to construct a null distribution of the RMSE, and compare it to the observed RMSE to assess significance.
+Going back to our example in Section \ref{focal-vs-comp} of focal tree with `focal_ID` 4 and its 20 competitors, the permutation test randomly resamples the `comp_sp` variable with replacement, leaving all other variables intact.
+The resampling with replacement is nested within each focal tree in order to preserve the neighborhood structure of our competition model.
+To run the permutation test, we use the same`comp_bayes_lm()` function as in Section \ref{model-fit-predict}, but with a `run_shuffle = TRUE` argument.
 
@@
 
-Using the four steps of our analysis sequence again, we answer the second of our two questions: how can we obtain an accurate estimate of model performance/error?
+We answer the second of our two questions: how can we obtain an accurate estimate of model performance/error?
 The model fits and predictions in Section \ref{model-fit-predict} all suffer from a common failing: they use the same data to both fit the model and to assess the model's performance using the RMSE.
@@
 To this end, we use the `foldID` variable defined in Section \ref{spatial-information} whereby all focal trees are assigned to one of `r n_fold` spatially contiguous blocks that act as folds in our cross-validation routine.
-Figure \ref{fig:scbi-spatial-cross-validation-schematic} presents a schematic illustrating this scheme for fold 1 as the test set and folds 2, 3, and 4 as the training sets.
+Figure \ref{fig:scbi-spatial-cross-validation-schematic} presents a schematic illustrating this scheme for fold 1 (bottom-left) as the test set and folds 2, 3, and 4 as the training sets.
 We fit the model to all focal trees in the training set, apply the model to all focal trees in the test set to compute fitted/predicted values, and compute the RMSE of the observed versus predicted growths.
-We repeat this procedure 3 more times with each of the three remaining folds acting as the test set.
-Furthermore, in order to maintain spatial independence between the test and training set, a buffer that extend outwards from the boundary of the test set is computed; all trees falling within this buffer will not be part of the training set.
+We repeat this procedure 3 more times with each of the three remaining folds acting as the test set and then average all four resulting RMSE's.
+Furthermore, in order to maintain spatial independence between the test and training set, a buffer that extend outwards from the boundary of the test set is computed; all trees falling within this buffer are excluded from the training set.
 
@@
 ggplot() +
-  geom_sf(data = test_fold_buffer, col = "red", fill = "transparent") +
+  geom_sf(data = compute_buffer_region(test_fold_buffer, direction = "out", size = comp_dist), col = "red", fill = "transparent") +
   geom_sf(data = blocks_scbi, fill = "transparent", linetype = "dashed") +
-  geom_sf_text(data = test_vs_train %>% st_as_sf() %>% sample_n(1000), aes(label = foldID, col = fold))
+  geom_sf(data = test_vs_train %>% st_as_sf() %>% sample_n(1000), aes(col = fold, shape = fold))
 ```
 
-This algorithm is implemented in the `run_cv()` function, which acts as a wrapper function to the `comp_bayes_lm()` function that fits the model and the `predict()` generic that returns fitted values.
+This algorithm is implemented in the `run_cv()` function, which is a wrapper function to the `comp_bayes_lm()` function that fits the model and the `predict()` generic that returns fitted/predicted values.
+We compare these values to the observed growth values to again compute our RMSE.
 
@@
 ```
-
-Observe once again that this RMSE is much higher than that for the above SCBI model fit without cross-validation.
-
 ```{r}
@@
 
-- show image of PLOSOne
+- Run full simulation on SCBI data
 - run time considerations
@@
 
+Sophie Li for her feedback on package interface.
+
 # References

### 2021-02-22 paper/forestecology.tex
Done second pass of Section 2: Example

--- 2021-02-22 etc/paper/forestecology.tex
@@
   \title{\bf The forestecology R package for fitting and assessing models of
-interspecies competition on the growth of trees}
+interspecies competitive effects on the growth of trees}
 
@@
     {\LARGE\bf The forestecology R package for fitting and assessing models of
-interspecies competition on the growth of trees}
+interspecies competitive effects on the growth of trees}
   \end{center}
@@
 \begin{abstract}
-1. When modeling growth of trees forest ecologists often incorporate the
-effect of interspecies competition. Many such models are based on a
-neighborhood effect assumption whereby all trees within a fixed distance
-of all focal trees are considered competitors. Methods are needed to
-evaluate the effect of interspecies competition and to assess their
-quality.\\
+1. Many models for the growth of trees that incorporate the effect of
+interspecies competition are based on a neighborhood effect assumption
+whereby all trees within a fixed distance of all focal trees are
+considered competitors. Methods and tools are needed to quantify this
+competitive effect and assess the quality of all resulting models\\
 2. We present the \texttt{forestecology} package providing methods for
-both 1) evaluating the out-of-sample performance of our model using
-spatial-crossvaliation and 2) testing a null hypothesis of of no impact
-of competitor species' identity on the growth of trees using a
-permutation test. We implement a class and methods using R's S3
-object-oriented system, for a specific linear, Bayesian neighborhood
-competition model of tree growth.\\
-3. We demonstrate the package's functions using data from the
+both 1) evaluating the effect of competitor species identity using
+permutation tests and 2) evaluating model performance using spatial
+cross-validation. Following \citet{allen_permutation_2020}, we implement
+a Bayesian linear regression competition model.\\
+3. We demonstrate the package's functionality using data from the
 Smithsonian Conservation Biology Institute's large forest dynamics plot,
-part of the ForestGEO network of reseach sites. Given ForestGEO's data
-collection protocols and data formatting standards, the package
-cross-compatibility of code. We show both that 1) competitor species
-identity matters and 2) that not spatially cross-validating leads to
-error estimates that are overly optimistic.\\
-4. The package follows \texttt{tidyverse}-like structure whereby
-verb-named functions can be modularly ``piped'' in sequence to
-intuitively display the sequence of steps of analysis from start to
-finish. Additionally, most inputs/outputs of functions assume an are of
-\texttt{sf} class from the simple features package, thereby facilitating
-all wrangling and visualization of geospatial data. Lastly, even though
-our package is currently limited to one specific model, the package is
-setup such that it can be easily extended to other models.
+part of the ForestGEO global network of reseach sites. Given ForestGEO's
+data collection protocols and data formatting standards, the package was
+designed with cross-site compatibility in mind. We demonstrate that both
+1) competitor species identity matters and 2) that not spatially
+cross-validating leads to error estimates that are overly optimistic.\\
+4. The package features 1) \texttt{tidyverse}-like structure whereby
+verb-named functions can be modularly ``piped'' in sequence, 2)
+functions with standardized inputs/outputs of simple features
+\texttt{sf} package class, and 3) R S3 object-oriented implementation of
+the Bayesian linear regression model. These three facts allow for clear
+articulation of all the steps in the sequence of analysis and easy
+wrangling and visualization of the geospatial forestry data.
+Furthermore, while the package only has Bayesian linear regression
+implemented, the package was designed with extensibility to other
+methods in mind.
 \end{abstract}
@@
 
-We demonstrate the \texttt{forestecology} package's features on the
-Smithsonian Conservation Biology Institute (SCBI) large forest dynamics
-plot, located at the Smithsonian's National Zoo and Conservation Biology
-Institute in Front Royal, VA, USA. The 25.6 ha (640 x 400 m) plot is
-located at the intersection of three of the major physiographic
-provinces of the eastern US: the Blue Ridge, Ridge and Valley, and
-Piedmont provinces and is adjacent to the northern end of Shenandoah
-National Park. The forest type is typical mature secondary eastern mixed
-deciduous forest, with a canopy dominated by tulip poplar
+We demonstrate the \texttt{forestecology} package's functionality on
+data from the Smithsonian Conservation Biology Institute (SCBI) large
+forest dynamics plot, located at the Smithsonian's National Zoo and
+Conservation Biology Institute in Front Royal, VA, USA. The 25.6 ha (640
+x 400 m) plot is located at the intersection of three of the major
+physiographic provinces of the eastern US: the Blue Ridge, Ridge and
+Valley, and Piedmont provinces and is adjacent to the northern end of
+Shenandoah National Park. The forest type is typical mature secondary
+eastern mixed deciduous forest, with a canopy dominated by tulip poplar
 (\emph{Liriodendron tulipifera}), oaks (\emph{Quercus} spp.), and
@@
 
-There are four steps in our modeling pipeline:
-
-\begin{itemize}
-\tightlist
-\item
-  Step 1: Compute the growth of trees based on census data.
-\item
-  Step 2: Add spatial information:
+The \texttt{forestecology} package attempts has the following two
+ecological goals: 1) to evaluate the effect of competitor species
+identity using permutation tests and 2) to evaluate model performance
+using spatial cross-validation. To achieve these two goals, we outline a
+basic analysis sequence comprising of these four main steps:
 
@@
 \item
-    Define buffer region trees.
+  Compute the growth of trees based on census data.
+\item
+  Add spatial information:
+
+  \begin{enumerate}
+  \def\labelenumii{\arabic{enumii}.}
+  \tightlist
+  \item
+    Define a buffer region of trees.
   \item
@@
 \item
-  Step 3: Identify all focal and corresponding competitor trees.
+  Identify all focal trees and their competitors.
 \item
-  Step 4: Fit model, as well as:
+  Apply model, which includes:
 
   \begin{enumerate}
-  \def\labelenumi{\arabic{enumi}.}
+  \def\labelenumii{\arabic{enumii}.}
   \tightlist
+  \item
+    Fit model.
   \item
@@
   \end{enumerate}
-\end{itemize}
-
-Using these four steps, we can address the two following questions of
-ecological interest:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Evaluate the effect of competitor species identity using permutation
-  tests.
-\item
-  Evaluate model performance using spatial cross-validation.
 \end{enumerate}
 
-We load all necessary packages.
+We start by loading all necessary packages.
 
@@
 \KeywordTok{library}\NormalTok{(sf)}
+\KeywordTok{library}\NormalTok{(patchwork)}
 \KeywordTok{library}\NormalTok{(forestecology)}
 \KeywordTok{library}\NormalTok{(blockCV)}
-\KeywordTok{library}\NormalTok{(patchwork)}
 \end{Highlighting}
@@
 trees using data from two censuses. The \texttt{compute\_growth()}
-function computes growth assuming census data that follows ForestGEO
-standards. Despite such standards, minor variations will still exist
-between sites thereby necessitating some data wrangling and checking.
-For example, the SCBI site records all DBH's in millimeters, whereas the
-Michigan Big Woods site records them in centimeters
+function computes growth assuming census data that roughly follows
+ForestGEO standards. Despite such standards, minor variations will still
+exist between sites, thereby necessitating some data wrangling and
+checking. For example, the SCBI site records all DBH's in millimeters,
+whereas the Michigan Big Woods site records them in centimeters
 \citet{andersonteixeira_ctfs-forestgeo_2015}
-\citet{allen_michigan_2020}. The data format of other sites may be such
-that our \texttt{compute\_growth()} function doesn't work at all.
-However, in the end all that matters is that the growth of all trees is
-saved in a data frame of class \texttt{sf} whereby the geolocation of
-each tree is presented in a \texttt{geometry} variable of type
-\texttt{\textless{}POINT\textgreater{}} and at a minimum the data
-contains the following variables: a variable uniquely identifying each
-tree-stem, \texttt{sp} of type \texttt{fct} factor identifying species,
-\texttt{dbh1} and \texttt{dbh2} of type \texttt{dbl} quantifying the DBH
-at earlier and later census, and \texttt{growth} of type \texttt{dbl}
-double quantifying the average annual growth in centimeters.
+\citet{allen_michigan_2020}.
 
 We load both 2008 and 2014 SCBI census data \texttt{.csv} files as they
-existed on GitHub on November 20, 2020. After selecting only relevant
-variables, we perform a few additional data wrangling steps: convert the
-variable with the date of measurement to be of type \texttt{date},
-convert DBH to be in centimeters\footnote{A rule of thumb to determine
-  the units of DBH is check if the smallest non-zero and non-missing
-  measurement is 1 or 10. If the former, then centimeters. If the later,
-  then millimeters. This is because ForestGEO protocols state that only
-  trees with DBH greater or equal to 1cm should be included in censuses.},
-convert the \texttt{sp} variable containing species information from
-type \texttt{chr} character to \texttt{fct} factor (we will discuss the
-need for this in Section \ref{spatial-cross-validation}). Furthermore,
-in order to speed up computation for purposes of this example, we only
-consider a 9 ha subsection of the 25.6 ha of the SCBI site: \texttt{gx}
-from 0--300 instead of 0--400 and \texttt{gy} from 300--600 instead of
-0--640.
+existed on GitHub on November 20, 2020. After selecting only the
+relevant variables, we perform a few additional data wrangling steps:
+convert the character variable with the date of measurement to be of
+explicit type \texttt{date}, convert DBH to be in centimeters\footnote{A
+  rule of thumb to determine the units of DBH is to check if the
+  smallest non-zero and non-missing measurement is 1 or 10. If the
+  former, then centimeters. If the later, then millimeters. This is
+  because ForestGEO protocols state that only trees with DBH greater or
+  equal to 1cm should be included in censuses.}, convert the \texttt{sp}
+variable containing species information from type \texttt{chr} character
+to \texttt{fct} factor.\footnote{In our spatial cross-validation
+  algorithm in Section \ref{spatial-cross-validation} issues can occur
+  when rare species do not occur in the training set, but then are
+  encountered in the test set. This risk is mitigated by representing
+  \texttt{sp} as a factor variable, which has a complete list of all
+  levels of the categorical variable.}Furthermore, in order to speed up
+computation for purposes of this example, we only consider a 9 ha
+subsection of the 25.6 ha of the SCBI site: \texttt{gx} from 0--300
+instead of 0--400 and \texttt{gy} from 300--600 instead of 0--640.
 
@@
 
-These two data frames are then used as the two primary arguments to the
+These two data frames are then used as arguments to the
 \texttt{compute\_growth()} function, along with the \texttt{id} argument
-whereby the user specifies the name of the variable that uniquely
-identifies each tree-stem under consideration (note this does not
-include resprouts in the later census):
+that specifies the variable that uniquely identifies each tree-stem.
+Note furthermore that we discard all resprouts in the later census
+(those with \texttt{code\ ==\ R}).
 
@@
 The output \texttt{growth\_scbi} is a single data frame of class
-\texttt{sf} that includes a numerical \texttt{growth} reflecting the
-average annual growth in DBH (in cm) for all trees that were alive at
-both time points as well a \texttt{geometry} variable encoding each
-tree's geolocation. Furthermore, variables that (in theory) remain
-unchanged between censuses appear only once, such as location variables
-\texttt{gx} and \texttt{gy}; as well as species-related variables.
-Variables that should change between censuses are suffixed with
-\texttt{1} and \texttt{2} indicating the earlier and later censuses,
-such as \texttt{dbh1/dbh2} and \texttt{codes1/codes2}.
+\texttt{sf} that includes \texttt{growth}, the average annual growth in
+DBH (in cm) for all trees that were alive at both time points, as well a
+\texttt{geometry} variable encoding each tree's geolocation.
+Furthermore, the variables that should remain unchanged between censuses
+appear only once, such as location variables \texttt{gx} and
+\texttt{gy}; as well as species-related variables. Variables that should
+change between censuses are suffixed with \texttt{1} and \texttt{2}
+indicating the earlier and later censuses, such as \texttt{dbh1/dbh2}
+and \texttt{codes1/codes2}.
+
+The data format of other sites may be such that our
+\texttt{compute\_growth()} function doesn't work at all. However, in the
+end all that matters is that the growth of all trees is saved in a data
+frame of class \texttt{sf} whereby the geolocation of each tree is
+presented in a \texttt{geometry} variable of type
+\texttt{\textless{}POINT\textgreater{}} and at a minimum the data
+contains the variables above.
 
@@
 
-TODO: Rescale points in this plot:
-
 \begin{Shaded}
 \begin{Highlighting}[]
+\CommentTok{# }\AlertTok{TODO}\CommentTok{: Rescale points in this plot}
 \KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
@@
 
-\caption{Step 1 - Compute growth of trees based on census data: All trees from a 9 ha subsection of the Smithsonian Conservation Biology Institute (SCBI) forest plot.}\label{fig:scbi-trees}
+\caption{Compute growth of trees based on census data: Map with growth of all trees from a 9 ha subsection of the Smithsonian Conservation Biology Institute (SCBI) forest plot.}\label{fig:scbi-trees}
 \end{figure}
@@
 information we add is a ``buffer region'' to the periphery of the study
-region. Since some of our model's explanatory variables such as
-competitor basal area are cumulative, we must ensure that all trees
-being modeled are not biased to have different neighbor structures. This
-is of particular concern for trees at the boundary of study regions,
-which will not have the same number of neighbors that act as competitors
-as trees in the internal part of the study region. In order to account
-for such edge effects only trees who are not part of this buffer region,
-i.e.~are part of the interior of the study region, will have their
-growths modeled \citet{waller_applied_2004}.
+region. Since some of our model's explanatory variables are cumulative
+(such as competitor basal area), we must ensure that all trees being
+modeled are not biased to have different neighbor structures. This is of
+concern for trees at the boundary of study regions, which will not have
+the same number of neighbors as trees in the internal part of the study
+region. In order to account for such edge effects, only trees that are
+not part of this buffer region, i.e.~are part of the interior of the
+study region, will have their growths modeled
+\citet{waller_applied_2004}.
 
@@
 competitors (assuming the same units as the \texttt{gx} and \texttt{gy}
-location variables). In our case we set this value below at 7.5m, a
-value informed by \citet{canham_neighborhood_2004}
+location variables). In our case we set this value at 7.5m, a value
+informed by \citet{canham_neighborhood_2004}
 \citet{uriarte_spatially_2004} \citet{canham_neighborhood_2006}. Using
@@
 \texttt{buffer} set to \texttt{FALSE} will be our focal trees whose
-growths are modeled, whereas those with \texttt{buffer} set to
-\texttt{TRUE} will only be considered as competitor trees whose growth
-will not be modeled.
+growths are modeled, whereas those with \texttt{TRUE} will only be
+considered as competitor trees whose growth will not be modeled.
 
@@
 The second element of spatial information are blocks corresponding to
-folds of a spatial cross-validation algorithm used to estimate model
-error. Conventional cross-validation algorithms assign observations to
-folds by randomly resampling individual observations. However,
-underlying this algorithm is an assumption that the observations are
-independent. In the case of forest census data, observations exhibit
-spatial autocorrelation. This spatial dependence is incorporated into
-the cross-validation algorithm by randomly resampling spatial blocks of
-trees \citet{roberts_cross-validation_2017}
-\citet{pohjankukka_estimating_2017}. We therefore associate each
-observed tree to one of \(k\) spatial folds. In the example below, we
-first manually define two folds that partition the study region as an
-\texttt{sf} object. We then use the output of the
+folds of a spatial cross-validation algorithm used to estimate
+out-of-sample model error. Conventional cross-validation algorithms
+assign observations to folds by randomly resampling individual
+observations. However, an assumption that the observations are
+independent underlies any cross-validation algorithm. In the case of
+forest census data, observations exhibit spatial autocorrelation. We
+therefore incorporate this spatial dependence into the cross-validation
+algorithm by randomly resampling spatial blocks of trees
+\citet{roberts_cross-validation_2017}
+\citet{pohjankukka_estimating_2017}.
+
+In the example below, we first manually define four folds that partition
+the study region as an \texttt{sf} object. We then use the output of the
 \texttt{spatialBlock()} function from the \texttt{blockCV} package to
 associate each tree in \texttt{growth\_scbi} to the correct fold (saved
-in the \texttt{foldID} variable) \citet{valavi_blockcv_2019}.
-\textbackslash footnote\{In the Appendix we present an example where the
-folds themselves are also created using the \texttt{spatialBlock()}
-function given a specified \texttt{cv\_block\_size}.\}
+in the \texttt{foldID} variable) \citet{valavi_blockcv_2019}. \footnote{In
+  the Appendix we present an example where the folds themselves are also
+  created using the \texttt{spatialBlock()} function given a specified
+  \texttt{cv\_block\_size}.}
 
@@
 \CommentTok{# Manually define spatial blocks to act as folds}
+\NormalTok{n_fold <-}\StringTok{ }\DecValTok{4}
 \NormalTok{fold1 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{450}\NormalTok{))}
@@
 \NormalTok{fold4 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{600}\NormalTok{))}
-\NormalTok{n_fold <-}\StringTok{ }\DecValTok{4}
 
@@
 
-\caption{Step 2 - Add spatial information: Buffer region and spatial cross-validation blocks (1 through 4).
-All trees in the interior of the study region will be the focal trees whose growth will modeled.}\label{fig:scbi-spatial-information}
+\caption{Add spatial information: Buffer region and spatial cross-validation blocks (1 through 4).
+All trees in the interior of the study region (i.e.
+not part of buffer) will be the focal trees whose growth will be modeled.}\label{fig:scbi-spatial-information}
 \end{figure}
@@
 \texttt{create\_focal\_vs\_comp()} functions performs these tasks and
-returns a new data frame of type \texttt{sf} containing this
-information. On top of the previously discussed arguments
-\texttt{comp\_dist} defining the competition neighborhood and
-\texttt{id} indicating which variable in the data frame uniquely
-identifies each tree-stem, this function also requires an \texttt{sf}
-object representation of the spatial cross-validation blocks/folds; in
-our case, this was manually encoded in the \texttt{blocks\_scbi} in
-Section \ref{spatial-information} while in our Appendix we present an
-example where this was performed using \texttt{spatialBlock()} from the
+returns a new data frame of type \texttt{sf}. On top of the previous
+arguments \texttt{comp\_dist} defining the competition neighborhood and
+\texttt{id} indicating which variable uniquely identifies each
+tree-stem, this function also requires an \texttt{sf} object
+representation of the spatial cross-validation blocks/folds; in our
+case, this was manually encoded in the \texttt{blocks\_scbi} in Section
+\ref{spatial-information} while in our Appendix we present an example
+where this was performed using \texttt{spatialBlock()} from the
 \texttt{blockCV} package. We present the resulting data frame below with
@@
 \begin{Highlighting}[]
+\CommentTok{# }\AlertTok{TODO}\CommentTok{: Below reconcile the number of rows as they off by one from growth_scbi %>% }
+\CommentTok{# filter(!is.na(growth) & !buffer). Perhaps by removing NA's in the growth_scbi stage?}
 \NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
@@
 
-TODO: Below reconcile the number of rows as they off by one from
-\texttt{growth\_scbi} pipe \texttt{filter(!is.na(growth)\ \&\ !buffer)}.
-perhaps by removing NA's in the \texttt{growth\_scbi} stage.
-
 The resulting data frame \texttt{focal\_vs\_comp\_scbi} has 6296 rows,
 representing the subset of the 7954 trees in \texttt{growth\_scbi} that
-will be considered as focal trees and thus have their growths modeled.
-Recall from Section \ref{spatial-information} this consists all trees
-that are not part of the buffer region in Figure
-\ref{fig:scbi-spatial-information}. Two new variables \texttt{focal\_ID}
-and \texttt{focal\_sp} related to tree-stem identification and species
+will be considered as focal trees. Two new variables \texttt{focal\_ID}
+and \texttt{focal\_sp} relate to tree-stem identification and species
 information. Most notably however is a new variable \texttt{comp} which
-contains information on all competitor trees for a given focal tree
-saved in list-column format, a feature of the \texttt{tidyr} package
+contains information on all competitor trees for a given focal tree,
+saved in \texttt{tidyr} package list-column format
 \citet{tidyr_package}. For example, we drill-down on the tree with
-\texttt{focal\_ID} equal to 4. It has 20 competitor trees each described
-by 4 variables as indicated by the fact that \texttt{comp} is a
+\texttt{focal\_ID} 4, which has 20 competitor trees each described by 4
+variables as indicated by the fact that \texttt{comp} is a
 \texttt{\textless{}tibble\ {[}20\ ×\ 4{]}\textgreater{}}.
@@
 
+The spatial distribution of these trees is visualized in Figure
+\ref{fig:scbi-focal-vs-comp-map}: the dashed circle extends 7.5 m away
+from the focal tree while all 20 competitor trees are within this
+circle.
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.66\linewidth]{Figures/scbi-focal-vs-comp-map-1}
+
+}
+
+\caption{Identify all focal and corresponding competitor trees: All 20 competitor trees of focal tree 4.}\label{fig:scbi-focal-vs-comp-map}
+\end{figure}
+
 Using the \texttt{unnest()} function from the \texttt{tidyr} package, we
 can flatten list-column into regular columns. We observe that for the
-same focal tree with DBH equal to 13.65cm, we have information on all 20
-competitor trees whose \texttt{dist} distance to the focal tree is less
-than or equal to 7.5, including their unique tree-stem ID number, their
-species, and their basal area (in m\(^2\)) calculated as
-\(\frac{\pi \times (\text{DBH/2})^2}{10000}\) where \(DBH\) is the value
-from the earlier census in cm. Saving our focal versus competitor
-information in list-column minimizes redundancy since we do not repeat
-information on the focal tree 20 times. The spatial distribution of
-these trees is visualized in Figure \ref{fig:scbi-focal-vs-comp-map}:
-the dashed circle extends 7.5 m away from the focal tree while all 20
-competitor trees are within this circle.
+same focal tree, we have information on all 20 competitor trees whose
+\texttt{dist} distance to the focal tree is \(\leq\) 7.5: their unique
+tree-stem ID number, their species, and their basal area (in m\(^2\))
+calculated as \(\frac{\pi \times (\text{DBH/2})^2}{10000}\) where
+\(DBH\) is the value from the earlier of the two censuses in cm. Saving
+our focal versus competitor information in list-column minimizes
+redundancy since we do not repeat information on the focal tree 20
+times.
 
@@
 
-\begin{figure}
-
-{\centering \includegraphics[width=0.66\linewidth]{Figures/scbi-focal-vs-comp-map-1}
-
-}
-
-\caption{Step 3 - Identify all focal and corresponding competitor trees: All 20 competitor trees of focal tree 4.}\label{fig:scbi-focal-vs-comp-map}
-\end{figure}
-
 \hypertarget{model-fit-predict}{%
@@
 and saved this information in a data frame of type
-\texttt{focal\_vs\_comp}, the next step in our analysis sequence is to
+\texttt{focal\_vs\_comp}, the final step in our analysis sequence is to
 fit a model for the growth of all focal trees. Currently the
@@
 type \texttt{focal\_vs\_comp} as an argument can be easily swapped in.
-In the case of our specific competition Bayesian linear regression
-model, we can also specify prior distributions on all parameters of
-interest (here chosen to be the defaults).
+For our specific competition Bayesian linear regression model, we also
+specify prior distributions on all parameters of interest (here chosen
+to be the defaults as specified in \texttt{?comp\_bayes\_lm}).
 
@@
 \begin{Highlighting}[]
+\CommentTok{# }\AlertTok{TODO}\CommentTok{: Add information about default priors in ?comp_bayes_lm()}
 \NormalTok{comp_bayes_lm_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
@@
 First the generic for \texttt{print()} displays the names of all prior
-\& posterior parameters (accessible via
-\texttt{comp\_bayes\_lm\_scbi\$prior\_params} and
-\texttt{comp\_bayes\_lm\_scbi\$post\_params}) along with the model
-formula:
+\& posterior parameters along with the model formula:
 
@@
 variables in \texttt{newdata} and outputs a vector of fitted/predicted
-values \(\widehat{y}\) of the DBH for each focal tree (as returned by
-the posterior predictive distribution).
+values \(\widehat{y}\) of the DBH for each focal tree computed from the
+posterior predictive distribution.
 
@@
 
-We can then compare the observed and fitted/predicted growths to compute
-the root mean squared error of our model fit.
+We then compare the observed and fitted/predicted growths to compute the
+root mean squared error (RMSE) of our model fit.
 
@@
 the posterior distribution of all parameters in Figure
-\ref{fig:scbi-posterior-viz}.
+\ref{fig:scbi-posterior-viz} (for compactness we only show posteriors
+for 3 species).
 
@@
 
-\caption{Step 4 - Fit model: Posterior distributions of all parameters for three species.}\label{fig:scbi-posterior-viz}
+\caption{Fit model: Posterior distributions of all parameters for three species.}\label{fig:scbi-posterior-viz}
 \end{figure}
@@
 
-We use the four steps of our analysis sequence we answer the first of
-our two questions: does competitor species identity matter when modeling
-the growth of focal trees? We answer this via a permutation test: Under
-a null hypothesis where competitor species identity does not matter we
-can permute this variable within each focal tree, compute the RMSE (the
-test statistic of interest), repeat this process several times to
-construct a null distribution of RMSE, and compare the observed RMSE to
-assess significance. Going back to our example in Section
-\ref{focal-vs-comp} inspecting all 20 competitor trees to focal tree
-with \texttt{focal\_ID} 4, the permutation test involves randomly
-resampling the \texttt{comp\_sp} variable with replacement, leaving all
-other competitor variables intact. The resampling with replacement is
+In order to evaluate the effect of competitor species identity, we use
+the four steps of our analysis sequence answer along with a permutation
+test: Under a null hypothesis where competitor species identity does not
+matter, we can permute/shuffle this variable within each focal tree,
+compute the RMSE (the test statistic of interest), repeat this process
+several times to construct a null distribution of the RMSE, and compare
+it to the observed RMSE to assess significance. Going back to our
+example in Section \ref{focal-vs-comp} of focal tree with
+\texttt{focal\_ID} 4 and its 20 competitors, the permutation test
+randomly resamples the \texttt{comp\_sp} variable with replacement,
+leaving all other variables intact. The resampling with replacement is
 nested within each focal tree in order to preserve the neighborhood
-structure of our competitor model. To run the permutation test, we use
-the same code as in Section \ref{model-fit-predict}, but adding a
-\texttt{run\_shuffle\ =\ TRUE} argument to \texttt{comp\_bayes\_lm()}.
+structure of our competition model. To run the permutation test, we use
+the same\texttt{comp\_bayes\_lm()} function as in Section
+\ref{model-fit-predict}, but with a \texttt{run\_shuffle\ =\ TRUE}
+argument.
 
@@
 \NormalTok{model_rmse_shuffle}
-\CommentTok{## [1] 0.131693}
+\CommentTok{## [1] 0.131083}
 \end{Highlighting}
@@
 
-The resulting RMSE of 0.131693 based on the permutation test is larger
+The resulting RMSE of 0.131083 based on the permutation test is larger
 than the earlier RMSE of 0.1281398, suggesting that models that do
@@
 
-Using the four steps of our analysis sequence again, we answer the
-second of our two questions: how can we obtain an accurate estimate of
-model performance/error? The model fits and predictions in Section
-\ref{model-fit-predict} all suffer from a common failing: they use the
-same data to both fit the model and to assess the model's performance
-using the RMSE. As argued by \citet{roberts_cross-validation_2017}, this
-can lead to overly optimistic assessments of model quality as the models
-can be overfit, in particular in situations where
-spatial-autocorrelation is present. To mitigate the effects of such
-overfitting, we use a spatially block cross-validation algorithm.
+We answer the second of our two questions: how can we obtain an accurate
+estimate of model performance/error? The model fits and predictions in
+Section \ref{model-fit-predict} all suffer from a common failing: they
+use the same data to both fit the model and to assess the model's
+performance using the RMSE. As argued by
+\citet{roberts_cross-validation_2017}, this can lead to overly
+optimistic assessments of model quality as the models can be overfit, in
+particular in situations where spatial-autocorrelation is present. To
+mitigate the effects of such overfitting, we use a spatially block
+cross-validation algorithm.
 
@@
 routine. Figure \ref{fig:scbi-spatial-cross-validation-schematic}
-presents a schematic illustrating this scheme for fold 1 as the test set
-and folds 2, 3, and 4 as the training sets. We fit the model to all
-focal trees in the training set, apply the model to all focal trees in
-the test set to compute fitted/predicted values, and compute the RMSE of
-the observed versus predicted growths. We repeat this procedure 3 more
-times with each of the three remaining folds acting as the test set.
+presents a schematic illustrating this scheme for fold 1 (bottom-left)
+as the test set and folds 2, 3, and 4 as the training sets. We fit the
+model to all focal trees in the training set, apply the model to all
+focal trees in the test set to compute fitted/predicted values, and
+compute the RMSE of the observed versus predicted growths. We repeat
+this procedure 3 more times with each of the three remaining folds
+acting as the test set and then average all four resulting RMSE's.
 Furthermore, in order to maintain spatial independence between the test
 and training set, a buffer that extend outwards from the boundary of the
-test set is computed; all trees falling within this buffer will not be
-part of the training set.
+test set is computed; all trees falling within this buffer are excluded
+from the training set.
 
@@
 This algorithm is implemented in the \texttt{run\_cv()} function, which
-acts as a wrapper function to the \texttt{comp\_bayes\_lm()} function
-that fits the model and the \texttt{predict()} generic that returns
-fitted values.
+is a wrapper function to the \texttt{comp\_bayes\_lm()} function that
+fits the model and the \texttt{predict()} generic that returns
+fitted/predicted values. We compare these values to the observed growth
+values to again compute our RMSE.
 
@@
 \end{Shaded}
-
-Observe once again that this RMSE is much higher than that for the above
-SCBI model fit without cross-validation.
 
@@
 \item
-  show image of PLOSOne
+  Run full simulation on SCBI data
 \item
@@
 
+Sophie Li for her feedback on package interface.
+
 \bibliographystyle{agsm}

### 2021-02-22 paper/plosone_results/replicate_plosone_scbi.R

@@ -1,0 +1,257 @@
+library(tidyverse)
+library(lubridate)
+library(here)
+library(sf)
+library(viridis)
+library(forestecology)
+library(blockCV)
+library(tictoc)
+
+
+# Compute growth of trees based on census data ---------------------------------
+census_2013_scbi <- here("paper/scbi.stem2.csv") %>%
+  read_csv() %>%
+  select(stemID, sp, date = ExactDate, gx, gy, dbh, codes, status) %>%
+  mutate(
+    date = mdy(date),
+    dbh = as.numeric(dbh)/10,
+    sp = factor(sp)
+  ) %>%
+  filter(gx < 300, between(gy, 300, 600))
+
+census_2018_scbi <- here("paper/scbi.stem3.csv") %>%
+  read_csv() %>%
+  select(stemID, sp, date = ExactDate, gx, gy, dbh, codes, status) %>%
+  mutate(
+    date = mdy(date),
+    dbh = as.numeric(dbh)/10,
+    sp = factor(sp)
+  ) %>%
+  filter(gx < 300, between(gy, 300, 600))
+
+growth_scbi <-
+  compute_growth(
+    census_1 = census_2013_scbi,
+    census_2 = census_2018_scbi %>% filter(!str_detect(codes, "R")),
+    id = "stemID"
+  )
+
+
+
+# Add spatial information -----------------------------------------------------
+# Define buffer region using competitive distance range
+comp_dist <- 7.5
+
+study_region_scbi <- tibble(
+  x = c(0, 300, 300, 0, 0),
+  y = c(300, 300, 600, 600, 300)
+) %>%
+  sf_polygon()
+
+growth_scbi <- growth_scbi %>%
+  add_buffer_variable(size = comp_dist, region = study_region_scbi)
+
+# Manually define spatial blocks to act as folds
+fold1 <- rbind(c(0, 300), c(150, 300), c(150, 450), c(0, 450))
+fold2 <- rbind(c(150, 300), c(300, 300), c(300, 450), c(150, 450))
+fold3 <- rbind(c(0, 450), c(150, 450), c(150, 600), c(0, 600))
+fold4 <- rbind(c(150, 450), c(300, 450), c(300, 600), c(150, 600))
+n_fold <- 4
+
+blocks_scbi <- bind_rows(
+  sf_polygon(fold1), sf_polygon(fold2), sf_polygon(fold3), sf_polygon(fold4)
+) %>%
[…]

### 2021-02-23 paper/forestecology.Rmd
first push of small edits

@@
 
-While there are a littany of models one can consider, in PLOSOne we considered a simple one.
+While there are a litany of models one can consider, in @allen_permutation_2020 we considered a simple one.
 
-Describe model specifics. Bayesian linear regression model.
-
-Next we fit the following linear model to the DBH of each focal tree.
+We fit the following linear model to the DBH of each focal tree.
 Let $i = 1, \ldots, n_j$ index all $n_j$ trees of "focal" species group $j$; let $j = 1, \ldots, J$ index all $J$ focal species groups; and let $k = 1, \ldots, K$ index all $K$ "competitor" species groups.

### 2021-02-23 paper/forestecology.Rmd
second small push

@@
 
-The `forestecology` package attempts has the following two ecological goals: 1) to evaluate the effect of competitor species identity using permutation tests and 2) to evaluate model performance using spatial cross-validation.
-To achieve these two goals, we outline a basic analysis sequence comprising of these four main steps:
+The `forestecology` package has the following ecological goals: 1) to evaluate the effect of competitor species identity using permutation tests and 2) to evaluate model performance using spatial cross-validation.
+To achieve these goals, we outline a basic analysis sequence comprising of these four main steps:
 
@@
 Despite such standards, minor variations will still exist between sites, thereby necessitating some data wrangling and checking.
-For example, the SCBI site records all DBH's in millimeters, whereas the Michigan Big Woods site records them in centimeters @andersonteixeira_ctfs-forestgeo_2015 @allen_michigan_2020.
+For example, the SCBI site records all DBH's in millimeters @bourg_initial_2013, whereas the Michigan Big Woods site records them in centimeters @allen_michigan_2020.
 

### 2021-02-23 paper/forestecology.tex
second small push

@@
 
-While there are a littany of models one can consider, in PLOSOne we
-considered a simple one.
+While there are a litany of models one can consider, in
+\citet{allen_permutation_2020} we considered a simple one.
 
-Describe model specifics. Bayesian linear regression model.
-
-Next we fit the following linear model to the DBH of each focal tree.
-Let \(i = 1, \ldots, n_j\) index all \(n_j\) trees of ``focal'' species
+We fit the following linear model to the DBH of each focal tree. Let
+\(i = 1, \ldots, n_j\) index all \(n_j\) trees of ``focal'' species
 group \(j\); let \(j = 1, \ldots, J\) index all \(J\) focal species
@@
 
-The \texttt{forestecology} package attempts has the following two
-ecological goals: 1) to evaluate the effect of competitor species
-identity using permutation tests and 2) to evaluate model performance
-using spatial cross-validation. To achieve these two goals, we outline a
-basic analysis sequence comprising of these four main steps:
+The \texttt{forestecology} package has the following ecological goals:
+1) to evaluate the effect of competitor species identity using
+permutation tests and 2) to evaluate model performance using spatial
+cross-validation. To achieve these goals, we outline a basic analysis
+sequence comprising of these four main steps:
 
@@
 exist between sites, thereby necessitating some data wrangling and
-checking. For example, the SCBI site records all DBH's in millimeters,
-whereas the Michigan Big Woods site records them in centimeters
-\citet{andersonteixeira_ctfs-forestgeo_2015}
-\citet{allen_michigan_2020}.
+checking. For example, the SCBI site records all DBH's in millimeters
+\citet{bourg_initial_2013}, whereas the Michigan Big Woods site records
+them in centimeters \citet{allen_michigan_2020}.
 

### 2021-02-23 paper/forestecology.Rmd
Here are my edits. A couple of things - This is great! Very nice clear explanation of each step to do this analysis. - I am not 100% sure, but MEE might want a little bit more on why the underlying method is important. I could be wrong. Maybe the PLOS One paper serves that purpose and all we are doing here is presenting the package and explaning how to use it. - With the above said, I think what you have here is great and no reason to do a major restructure. But just make a clear push in the intro for the underlying method. - I have a couple of places hwere I hvae left commetns/notes. YOu can

@@
 the US Forest Service Forest Inventory and Analysis plots @smith_forest_2002.
+Or more generally to model interactions of any community of mapped sessile organisms.
 
@@
 
-We demonstrate the `forestecology` package's functionality on data from the Smithsonian Conservation Biology Institute (SCBI) large forest dynamics plot, located at the Smithsonian's National Zoo and Conservation Biology Institute in Front Royal, VA, USA.
+We demonstrate the `forestecology` package's functionality on data from the Smithsonian Conservation Biology Institute (SCBI) large forest dynamics plot, located at the Smithsonian's National Zoo and Conservation Biology Institute in Front Royal, VA, USA @bourg_initial_2013.
 The 25.6 ha (640 x 400 m) plot is located at the intersection of three of the major physiographic provinces of the eastern US: the Blue Ridge, Ridge and Valley, and Piedmont provinces and is adjacent to the northern end of Shenandoah National Park.
@@
 
-The `forestecology` package has the following ecological goals: 1) to evaluate the effect of competitor species identity using permutation tests and 2) to evaluate model performance using spatial cross-validation.
+The `forestecology` package has the following ecological goals: 1) to evaluate the effect of competitor species identity on using permutation tests and 2) to evaluate model performance using spatial cross-validation.
 To achieve these goals, we outline a basic analysis sequence comprising of these four main steps:
 
-1. Compute the growth of trees based on census data.
+1. Compute the growth of stems based on two censuses.
 1. Add spatial information:
@@
 The first step in the our analysis sequence is to compute the growth of trees using data from two censuses.
-The `compute_growth()` function computes growth assuming census data that roughly follows ForestGEO standards.
+The `compute_growth()` function computes average annual growth assuming census data that roughly follows ForestGEO standards.
 Despite such standards, minor variations will still exist between sites, thereby necessitating some data wrangling and checking.
-For example, the SCBI site records all DBH's in millimeters @bourg_initial_2013, whereas the Michigan Big Woods site records them in centimeters @allen_michigan_2020.
+For example, the SCBI site records all diameters at breast height (DBH) in millimeters @bourg_initial_2013, whereas the Michigan Big Woods site records them in centimeters @allen_michigan_2020.
 
@@
 This is because ForestGEO protocols state that only trees with DBH greater or equal to 1cm should be included in censuses.], convert the `sp` variable containing species information from type `chr` character to `fct` factor.^[In our spatial cross-validation algorithm in Section \ref{spatial-cross-validation} issues can occur when rare species do not occur in the training set, but then are encountered in the test set.
-This risk is mitigated by representing `sp` as a factor variable, which has a complete list of all levels of the categorical variable.]Furthermore, in order to speed up computation for purposes of this example, we only consider a 9 ha subsection of the 25.6 ha of the SCBI site: `gx` from 0--300 instead of 0--400 and `gy` from 300--600 instead of 0--640.
+This risk is mitigated by representing `sp` as a factor variable, which has a complete list of all levels of the categorical variable.] Furthermore, in order to speed up computation for purposes of this example, we only consider a 9 ha subsection of the 25.6 ha of the SCBI site: `gx` from 0--300 instead of 0--400 and `gy` from 300--600 instead of 0--640.
 
@@
 These two data frames are then used as arguments to the `compute_growth()` function, along with the `id` argument that specifies the variable that uniquely identifies each tree-stem.
-Note furthermore that we discard all resprouts in the later census (those with `code == R`).
+Note furthermore that we discard all resprouts in the later census (those with `code == R`), since we are only interested in the diameter growth of surviving, and not resprouted, stems.
 
@@
 
-The output `growth_scbi` is a single data frame of class `sf` that includes `growth`, the average annual growth in DBH (in cm) for all trees that were alive at both time points, as well a `geometry` variable encoding each tree's geolocation.
+The output `growth_scbi` is a single data frame of class `sf` that includes `growth`, the average annual growth in DBH (cm y^-1^) for all stems that were alive at both time points, as well a `geometry` variable encoding each tree's geolocation.
 Furthermore, the variables that should remain unchanged between censuses appear only once, such as location variables `gx` and `gy`; as well as species-related variables.
@@
 The data format of other sites may be such that our `compute_growth()` function doesn't work at all.
-However, in the end all that matters is that the growth of all trees is saved in a data frame of class `sf` whereby the geolocation of each tree is presented in a `geometry` variable of type `<POINT>` and at a minimum the data contains the variables above.
+However, in the end all that matters is that the growth of all stems is saved in a data frame of class `sf` whereby the geolocation of each tree is presented in a `geometry` variable of type `<POINT>` and at a minimum the data contains the variables above (DAVE NOTE: THis isn't exaclty clear to me what variables above means.
+Maybe list all the variables needed.
+Do they have to have exact names?)
 
@@
 
-The next step in our analysis sequence is to add spatial information to our main `growth_scbi` data frame.
+The next step in our analysis sequence is to add additional spatial information to our main `growth_scbi` data frame.
 The first element of spatial information we add is a "buffer region" to the periphery of the study region.
 Since some of our model's explanatory variables are cumulative (such as competitor basal area), we must ensure that all trees being modeled are not biased to have different neighbor structures.
-This is of concern for trees at the boundary of study regions, which will not have the same number of neighbors as trees in the internal part of the study region.
+This is of concern for trees at the boundary of study regions, for which all neighbors will not be included in the censused stems.
 In order to account for such edge effects, only trees that are not part of this buffer region, i.e.
-are part of the interior of the study region, will have their growths modeled @waller_applied_2004.
+are part of the interior of the study region, will have their growth modeled @waller_applied_2004.
 
@@
 In our case we set this value at 7.5m, a value informed by @canham_neighborhood_2004 @uriarte_spatially_2004 @canham_neighborhood_2006.
-Using this value along with a manually constructed `sf` object representation of the study region's boundary, we apply the `add_buffer_variable()` to our `growth_scbi` data frame to add a `buffer` boolean variable: all trees who have `buffer` set to `FALSE` will be our focal trees whose growths are modeled, whereas those with `TRUE` will only be considered as competitor trees whose growth will not be modeled.
+Using this value along with a manually constructed `sf` object representation of the study region's boundary, we apply the `add_buffer_variable()` to our `growth_scbi` data frame to add a `buffer` boolean variable: all trees who have `buffer` set to `FALSE` will be our focal trees whose growth will be modeled, whereas those with `TRUE` will only be considered as competitor trees whose growth will not.
 
@@
 We therefore incorporate this spatial dependence into the cross-validation algorithm by randomly resampling spatial blocks of trees @roberts_cross-validation_2017 @pohjankukka_estimating_2017.
+(DAVE NOTE: I don't want to change somethign I don't understand, but "randomly resampling" seems wrong to me.
+I don't see anything random about our block structure.
+Seems to me like "incorpoate this spatial dependence into the cross-validation algorithm with our spatial blocks of trees.")
 
@@
 ^[In the Appendix we present an example where the folds themselves are also created using the `spatialBlock()` function given a specified `cv_block_size`.]
+
+DAVE NOTE::there are a couple of places where you convert to factor.
+Earlier with `sp` and then twice in the code block below.
+ANy reason we can't just do this "under the hood" in the package code?
+Also any reason with swtiched from undercase with underscores to `SpatialBlock_scbi` in the naming?
 
@@
 We then compare the observed and fitted/predicted growths to compute the root mean squared error (RMSE) of our model fit.
+DAVE NOTE: I rounded all the RMSEs, in ecology convention to not show WAY more digits than resolution of the measurement.
 
@@
   rmse(truth = growth, estimate = growth_hat) %>%
-  pull(.estimate)
+  pull(.estimate) %>%
+  round(3)
 model_rmse
@@
 
-TODO: Discuss meaning of results.
+These plots give the posterior distributions of parameters from equation 1 (DAVE NOTE:SOrry couldn't figure out how to make equation references work).
+For many pacakge users they will be of interest because they give insight into the species-specific competitive interactions.
+Setting `type = "intercepts"` gives posterior distributions for $\beta_{0,j}$ and `type = "dbh_slopes"` for $\beta_{dbh,j}$.
+These give species specific growth independent of competition.
+The values of more interest are plotted with `type = 'competition'` which gives the posterior distribution for $\lambda_{j,k}$ species-specific competition coefficients (i.e., the $\lambda$-matrix).
+Negative values indicate a competitor species which slows the growth of a focal species.
+Here, for example, we see that tulip trees (`litu`) have a strong negative effect on the growth of conspecifics but relatively little effect on neighbors of the other two species.
 
@@
   rmse(truth = growth, estimate = growth_hat_shuffle) %>%
-  pull(.estimate)
+  pull(.estimate) %>%
+  round(3)
 model_rmse_shuffle
@@
 The resulting RMSE of `r model_rmse_shuffle` based on the permutation test is `r ifelse(model_rmse_shuffle < model_rmse, "smaller", "larger")` than the earlier RMSE of `r model_rmse`, suggesting that models that do incorporate competitor species identity better fit the data.
-We conduct a fuller simulation in Section below.
+We conduct an analysis of the full SCBI plot in Section below.
 
@@
 We repeat this procedure 3 more times with each of the three remaining folds acting as the test set and then average all four resulting RMSE's.
-Furthermore, in order to maintain spatial independence between the test and training set, a buffer that extend outwards from the boundary of the test set is computed; all trees falling within this buffer are excluded from the training set.
+Furthermore, in order to maintain spatial independence between the test and training set, a fold buffer that extend outwards from the boundary of the test set is computed; all trees falling within this fold buffer are excluded from the training set.
+DAVE NOTE: I repalced "buffer" with "fold buffer" just to keep clear that this is different than our global buffer.
+But if you think that distinction is not necessary feel free to switch back!
 
-```{r scbi-spatial-cross-validation-schematic, out.width="66%", fig.cap= "Schematic of spatial cross-validation: Using the k = 1 fold as the test set, assigning each focal tree to training set, test set, and buffer.", echo = FALSE}
+```{r scbi-spatial-cross-validation-schematic, out.width="66%", fig.cap= "Schematic of spatial cross-validation: Using the k = 1 fold as the test set, assigning each focal tree to training set, test set, and fold buffer.", echo = FALSE}
 fold_number <- 1
@@
       foldID != fold_number & buffer ~ "training",
-      foldID != fold_number & !buffer ~ "buffer"
+      foldID != fold_number & !buffer ~ "fold buffer"
     ),
-    fold = factor(fold, levels = c("test", "buffer", "training"))
+    fold = factor(fold, levels = c("test", "fold buffer", "training"))
   )
@@
   rmse(truth = growth, estimate = growth_hat) %>%
-  pull(.estimate)
+  pull(.estimate) %>%
+  round(3)
 model_rmse_cv

### 2021-02-23 paper/forestecology.tex
Here are my edits. A couple of things - This is great! Very nice clear explanation of each step to do this analysis. - I am not 100% sure, but MEE might want a little bit more on why the underlying method is important. I could be wrong. Maybe the PLOS One paper serves that purpose and all we are doing here is presenting the package and explaning how to use it. - With the above said, I think what you have here is great and no reason to do a major restructure. But just make a clear push in the intro for the underlying method. - I have a couple of places hwere I hvae left commetns/notes. YOu can

@@
 e.g.~the US Forest Service Forest Inventory and Analysis plots
-\citet{smith_forest_2002}.
+\citet{smith_forest_2002}. Or more generally to model interactions of
+any community of mapped sessile organisms.
 
@@
 forest dynamics plot, located at the Smithsonian's National Zoo and
-Conservation Biology Institute in Front Royal, VA, USA. The 25.6 ha (640
-x 400 m) plot is located at the intersection of three of the major
-physiographic provinces of the eastern US: the Blue Ridge, Ridge and
-Valley, and Piedmont provinces and is adjacent to the northern end of
-Shenandoah National Park. The forest type is typical mature secondary
-eastern mixed deciduous forest, with a canopy dominated by tulip poplar
-(\emph{Liriodendron tulipifera}), oaks (\emph{Quercus} spp.), and
-hickories (\emph{Carya} spp.), and an understory composed mainly of
-spicebush (\emph{Lindera benzoin}), paw-paw (\emph{Asimina triloba}),
-American hornbeam (\emph{Carpinus caroliniana}), and witch hazel
-(\emph{Hamamelis virginiana}) \citet{bourg_initial_2013}.
+Conservation Biology Institute in Front Royal, VA, USA
+\citet{bourg_initial_2013}. The 25.6 ha (640 x 400 m) plot is located at
+the intersection of three of the major physiographic provinces of the
+eastern US: the Blue Ridge, Ridge and Valley, and Piedmont provinces and
+is adjacent to the northern end of Shenandoah National Park. The forest
+type is typical mature secondary eastern mixed deciduous forest, with a
+canopy dominated by tulip poplar (\emph{Liriodendron tulipifera}), oaks
+(\emph{Quercus} spp.), and hickories (\emph{Carya} spp.), and an
+understory composed mainly of spicebush (\emph{Lindera benzoin}),
+paw-paw (\emph{Asimina triloba}), American hornbeam (\emph{Carpinus
+caroliniana}), and witch hazel (\emph{Hamamelis virginiana})
+\citet{bourg_initial_2013}.
 
 The \texttt{forestecology} package has the following ecological goals:
-1) to evaluate the effect of competitor species identity using
+1) to evaluate the effect of competitor species identity on using
 permutation tests and 2) to evaluate model performance using spatial
@@
 \item
-  Compute the growth of trees based on census data.
+  Compute the growth of stems based on two censuses.
 \item
@@
 trees using data from two censuses. The \texttt{compute\_growth()}
-function computes growth assuming census data that roughly follows
-ForestGEO standards. Despite such standards, minor variations will still
-exist between sites, thereby necessitating some data wrangling and
-checking. For example, the SCBI site records all DBH's in millimeters
+function computes average annual growth assuming census data that
+roughly follows ForestGEO standards. Despite such standards, minor
+variations will still exist between sites, thereby necessitating some
+data wrangling and checking. For example, the SCBI site records all
+diameters at breast height (DBH) in millimeters
 \citet{bourg_initial_2013}, whereas the Michigan Big Woods site records
@@
   \texttt{sp} as a factor variable, which has a complete list of all
-  levels of the categorical variable.}Furthermore, in order to speed up
+  levels of the categorical variable.} Furthermore, in order to speed up
 computation for purposes of this example, we only consider a 9 ha
@@
 Note furthermore that we discard all resprouts in the later census
-(those with \texttt{code\ ==\ R}).
+(those with \texttt{code\ ==\ R}), since we are only interested in the
+diameter growth of surviving, and not resprouted, stems.
 
@@
 \texttt{sf} that includes \texttt{growth}, the average annual growth in
-DBH (in cm) for all trees that were alive at both time points, as well a
-\texttt{geometry} variable encoding each tree's geolocation.
-Furthermore, the variables that should remain unchanged between censuses
-appear only once, such as location variables \texttt{gx} and
-\texttt{gy}; as well as species-related variables. Variables that should
-change between censuses are suffixed with \texttt{1} and \texttt{2}
-indicating the earlier and later censuses, such as \texttt{dbh1/dbh2}
-and \texttt{codes1/codes2}.
+DBH (cm y\textsuperscript{-1}) for all stems that were alive at both
+time points, as well a \texttt{geometry} variable encoding each tree's
+geolocation. Furthermore, the variables that should remain unchanged
+between censuses appear only once, such as location variables
+\texttt{gx} and \texttt{gy}; as well as species-related variables.
+Variables that should change between censuses are suffixed with
+\texttt{1} and \texttt{2} indicating the earlier and later censuses,
+such as \texttt{dbh1/dbh2} and \texttt{codes1/codes2}.
 
@@
 \texttt{compute\_growth()} function doesn't work at all. However, in the
-end all that matters is that the growth of all trees is saved in a data
+end all that matters is that the growth of all stems is saved in a data
 frame of class \texttt{sf} whereby the geolocation of each tree is
@@
 \texttt{\textless{}POINT\textgreater{}} and at a minimum the data
-contains the variables above.
+contains the variables above (DAVE NOTE: THis isn't exaclty clear to me
+what variables above means. Maybe list all the variables needed. Do they
+have to have exact names?)
 
@@
 
-The next step in our analysis sequence is to add spatial information to
-our main \texttt{growth\_scbi} data frame. The first element of spatial
-information we add is a ``buffer region'' to the periphery of the study
-region. Since some of our model's explanatory variables are cumulative
-(such as competitor basal area), we must ensure that all trees being
-modeled are not biased to have different neighbor structures. This is of
-concern for trees at the boundary of study regions, which will not have
-the same number of neighbors as trees in the internal part of the study
-region. In order to account for such edge effects, only trees that are
+The next step in our analysis sequence is to add additional spatial
+information to our main \texttt{growth\_scbi} data frame. The first
+element of spatial information we add is a ``buffer region'' to the
+periphery of the study region. Since some of our model's explanatory
+variables are cumulative (such as competitor basal area), we must ensure
+that all trees being modeled are not biased to have different neighbor
+structures. This is of concern for trees at the boundary of study
+regions, for which all neighbors will not be included in the censused
+stems. In order to account for such edge effects, only trees that are
 not part of this buffer region, i.e.~are part of the interior of the
-study region, will have their growths modeled
+study region, will have their growth modeled
 \citet{waller_applied_2004}.
@@
 \texttt{buffer} set to \texttt{FALSE} will be our focal trees whose
-growths are modeled, whereas those with \texttt{TRUE} will only be
-considered as competitor trees whose growth will not be modeled.
+growth will be modeled, whereas those with \texttt{TRUE} will only be
+considered as competitor trees whose growth will not.
 
@@
 \citet{roberts_cross-validation_2017}
-\citet{pohjankukka_estimating_2017}.
+\citet{pohjankukka_estimating_2017}. (DAVE NOTE: I don't want to change
+somethign I don't understand, but ``randomly resampling'' seems wrong to
+me. I don't see anything random about our block structure. Seems to me
+like ``incorpoate this spatial dependence into the cross-validation
+algorithm with our spatial blocks of trees.'')
 
@@
   \texttt{cv\_block\_size}.}
+
+DAVE NOTE::there are a couple of places where you convert to factor.
+Earlier with \texttt{sp} and then twice in the code block below. ANy
+reason we can't just do this ``under the hood'' in the package code?
+Also any reason with swtiched from undercase with underscores to
+\texttt{SpatialBlock\_scbi} in the naming?
 
@@
 We then compare the observed and fitted/predicted growths to compute the
-root mean squared error (RMSE) of our model fit.
+root mean squared error (RMSE) of our model fit. DAVE NOTE: I rounded
+all the RMSEs, in ecology convention to not show WAY more digits than
+resolution of hte measurement.
 
@@
 \StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
+\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{round}\NormalTok{(}\DecValTok{3}\NormalTok{)}
 \NormalTok{model_rmse}
-\CommentTok{## [1] 0.1281398}
+\CommentTok{## [1] 0.128}
 \end{Highlighting}
@@
 
-TODO: Discuss meaning of results.
+These plots give the posterior distributions of parameters from equation
+1 (DAVE NOTE:SOrry couldn't figure out how to make equation references
+work). For many pacakge users they will be of interest because they give
+insight into the species-specific competitive interactions. Setting
+\texttt{type\ =\ "intercepts"} gives posterior distributions for
+\(\beta_{0,j}\) and \texttt{type\ =\ "dbh\_slopes"} for
+\(\beta_{dbh,j}\). These give species specific growth independent of
+competition. The values of more interest are plotted with
+\texttt{type\ =\ \textquotesingle{}competition\textquotesingle{}} which
+gives the posterior distribution for \(\lambda_{j,k}\) species-specific
+competition coefficients (i.e., the \(\lambda\)-matrix). Negative values
+indicate a competitor species which slows the growth of a focal species.
+Here, for example, we see that tulip trees (\texttt{litu}) have a strong
+negative effect on the growth of conspecifics but relatively little
+effect on neighbors of the other two species.
 
@@
 \StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat_shuffle) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
+\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{round}\NormalTok{(}\DecValTok{3}\NormalTok{)}
 \NormalTok{model_rmse_shuffle}
-\CommentTok{## [1] 0.131083}
+\CommentTok{## [1] 0.131}
 \end{Highlighting}
@@
 
-The resulting RMSE of 0.131083 based on the permutation test is larger
-than the earlier RMSE of 0.1281398, suggesting that models that do
-incorporate competitor species identity better fit the data. We conduct
-a fuller simulation in Section below.
+The resulting RMSE of 0.131 based on the permutation test is larger than
+the earlier RMSE of 0.128, suggesting that models that do incorporate
+competitor species identity better fit the data. We conduct an analysis
+of the full SCBI plot in Section below.
 
@@
 Furthermore, in order to maintain spatial independence between the test
-and training set, a buffer that extend outwards from the boundary of the
-test set is computed; all trees falling within this buffer are excluded
-from the training set.
+and training set, a fold buffer that extend outwards from the boundary
+of the test set is computed; all trees falling within this fold buffer
+are excluded from the training set. DAVE NOTE: I repalced ``buffer''
+with ``fold buffer'' just to keep clear that this is different than our
+global buffer. But if you think that distinction is not necessary feel
+free to switch back!
 
@@
 
-\caption{Schematic of spatial cross-validation: Using the k = 1 fold as the test set, assigning each focal tree to training set, test set, and buffer.}\label{fig:scbi-spatial-cross-validation-schematic}
+\caption{Schematic of spatial cross-validation: Using the k = 1 fold as the test set, assigning each focal tree to training set, test set, and fold buffer.}\label{fig:scbi-spatial-cross-validation-schematic}
 \end{figure}
@@
 \StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
+\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{round}\NormalTok{(}\DecValTok{3}\NormalTok{)}
 \NormalTok{model_rmse_cv}
-\CommentTok{## [1] 0.1402209}
+\CommentTok{## [1] 0.14}
 \end{Highlighting}
@@
 
-The resulting RMSE of 0.1402209 computed using cross-validation is
-larger than the earlier RMSE of 0.1281398, suggesting that models that
-do not take the inherent spatial autocorrelation of the data into
-account generate error estimates that are overly optimistic; in our case
-RMSE's that are too low.
+The resulting RMSE of 0.14 computed using cross-validation is larger
+than the earlier RMSE of 0.128, suggesting that models that do not take
+the inherent spatial autocorrelation of the data into account generate
+error estimates that are overly optimistic; in our case RMSE's that are
+too low.
 

### 2021-02-23 paper/forestecology.Rmd
missed on thing

@@
 
-The `forestecology` package has the following ecological goals: 1) to evaluate the effect of competitor species identity on using permutation tests and 2) to evaluate model performance using spatial cross-validation.
+The `forestecology` package has the following ecological goals: 1) to evaluate the effect of competitor species identity using permutation tests and 2) to evaluate model performance using spatial cross-validation.
 To achieve these goals, we outline a basic analysis sequence comprising of these four main steps:

### 2021-02-23 paper/forestecology.tex
missed on thing

@@
 The \texttt{forestecology} package has the following ecological goals:
-1) to evaluate the effect of competitor species identity on using
+1) to evaluate the effect of competitor species identity using
 permutation tests and 2) to evaluate model performance using spatial
@@
 all the RMSEs, in ecology convention to not show WAY more digits than
-resolution of hte measurement.
+resolution of the measurement.
 

### 2021-02-25 paper/forestecology.Rmd

@@
   | 1.
-Many models for the growth of trees that incorporate the effect of interspecies competition are based on a neighborhood effect assumption whereby all trees within a fixed distance of all focal trees are considered competitors.
-Methods and tools are needed to quantify this competitive effect and assess the quality of all resulting models
+Many models for the growth of trees that incorporate interspecies competition are based on a neighborhood effect assumption whereby all trees within a fixed distance of a focal tree are considered competitors.
+Methods and tools are needed to quantify this competitive effect and assess the quality of all resulting models.
   | 2.
@@
   | 4.
-The package features 1) `tidyverse`-like structure whereby verb-named functions can be modularly "piped" in sequence, 2) functions with standardized inputs/outputs of simple features `sf` package class, and 3) R S3 object-oriented implementation of the Bayesian linear regression model.
+The package features 1) `tidyverse`-like structure whereby verb-named functions can be modularly "piped" in sequence, 2) functions with standardized inputs/outputs of simple features `sf` package class, and 3) an S3 object-oriented implementation of the Bayesian linear regression model.
 These three facts allow for clear articulation of all the steps in the sequence of analysis and easy wrangling and visualization of the geospatial forestry data.
@@
   collapse = TRUE,
-  comment = "##",
-  # tidy = FALSE,
-  tibble.print_max = 5,
-  tibble.print_min = 5
+  comment = "##"
 )
@@
 # Set output width
-options(width = 70)
+options(
+  width = 70, 
+  # Decimal precision
+  digits = 3, 
+  # Number of rows for tibble printing
+  tibble.print_max = 5, tibble.print_min = 5
+)
 ```
@@
 
-$$
-\newcommand{\dbh}{\text{DBH}}
-\newcommand{\biomass}{\text{biomass}}
-\newcommand{\BA}{\text{BA}}
-y_{ij} = f(\vec{x}_{ij}) + \epsilon_{ij} = \beta_{0,j} + \beta_{\dbh,j} \cdot \dbh_{ij} + \sum_{k=1}^{K} \lambda_{jk} \cdot \BA_{ijk} + \epsilon_{ij}
-$$
+\begin{equation}
+\label{eq:model}
+y_{ij} = f(\vec{x}_{ij}) + \epsilon_{ij} = \beta_{0,j} + \beta_{\text{dbh},j} \cdot \text{dbh}_{ij} + \sum_{k=1}^{K} \lambda_{jk} \cdot \text{BA}_{ijk} + \epsilon_{ij}
+\end{equation}
 
@@
 
-The output `growth_scbi` is a single data frame of class `sf` that includes `growth`, the average annual growth in DBH (cm y^-1^) for all stems that were alive at both time points, as well a `geometry` variable encoding each tree's geolocation.
+The output `growth_scbi` is a single data frame of class `sf` that includes variables `growth`, the average annual growth in DBH (cm y^-1^) for all stems that were alive at both time points, and `geometry`, the `sf` package's encoding of geolocations of type `<POINT>`.
 Furthermore, the variables that should remain unchanged between censuses appear only once, such as location variables `gx` and `gy`; as well as species-related variables.
 Variables that should change between censuses are suffixed with `1` and `2` indicating the earlier and later censuses, such as `dbh1/dbh2` and `codes1/codes2`.
-
 The data format of other sites may be such that our `compute_growth()` function doesn't work at all.
-However, in the end all that matters is that the growth of all stems is saved in a data frame of class `sf` whereby the geolocation of each tree is presented in a `geometry` variable of type `<POINT>` and at a minimum the data contains the variables above (DAVE NOTE: THis isn't exaclty clear to me what variables above means.
-Maybe list all the variables needed.
-Do they have to have exact names?)
+However, in the end all that matters is that the growth of all stems is saved in a data frame of class `sf` and at a minimum the data contains the variables above.
+TODO: enumerate what these variables are and say if they need exact names
 
@@
 In the case of forest census data, observations exhibit spatial autocorrelation.
-We therefore incorporate this spatial dependence into the cross-validation algorithm by randomly resampling spatial blocks of trees @roberts_cross-validation_2017 @pohjankukka_estimating_2017.
-(DAVE NOTE: I don't want to change somethign I don't understand, but "randomly resampling" seems wrong to me.
-I don't see anything random about our block structure.
-Seems to me like "incorpoate this spatial dependence into the cross-validation algorithm with our spatial blocks of trees.")
+We therefore incorporate this spatial dependence into the cross-validation algorithm with our spatial blocks of trees.
+@roberts_cross-validation_2017 @pohjankukka_estimating_2017.
 
@@
 ^[In the Appendix we present an example where the folds themselves are also created using the `spatialBlock()` function given a specified `cv_block_size`.]
-
-DAVE NOTE::there are a couple of places where you convert to factor.
-Earlier with `sp` and then twice in the code block below.
-ANy reason we can't just do this "under the hood" in the package code?
-Also any reason with swtiched from undercase with underscores to `SpatialBlock_scbi` in the naming?
 
@@
 # Associate each observation to a fold
-SpatialBlock_scbi <- spatialBlock(
+spatial_block_scbi <- spatialBlock(
   speciesData = growth_scbi, k = n_fold, selection = "systematic", 
@@
 growth_scbi <- growth_scbi %>%
-  mutate(foldID = SpatialBlock_scbi$foldID %>% factor())
+  mutate(foldID = spatial_block_scbi$foldID %>% factor())
 ```
@@
 We then compare the observed and fitted/predicted growths to compute the root mean squared error (RMSE) of our model fit.
-DAVE NOTE: I rounded all the RMSEs, in ecology convention to not show WAY more digits than resolution of the measurement.
 
@@
   rmse(truth = growth, estimate = growth_hat) %>%
-  pull(.estimate) %>%
-  round(3)
+  pull(.estimate)
 model_rmse
@@
 
-These plots give the posterior distributions of parameters from equation 1 (DAVE NOTE:SOrry couldn't figure out how to make equation references work).
+These plots give the posterior distributions of parameters from Equation \ref{eq:model}.
 For many pacakge users they will be of interest because they give insight into the species-specific competitive interactions.
@@
   mutate(growth_hat_shuffle = predict(comp_bayes_lm_scbi_shuffle, newdata = focal_vs_comp_scbi))
-
+```
+```{r}
 model_rmse_shuffle <- focal_vs_comp_scbi %>%
   rmse(truth = growth, estimate = growth_hat_shuffle) %>%
-  pull(.estimate) %>%
-  round(3)
+  pull(.estimate)
 model_rmse_shuffle
@@
 Furthermore, in order to maintain spatial independence between the test and training set, a fold buffer that extend outwards from the boundary of the test set is computed; all trees falling within this fold buffer are excluded from the training set.
-DAVE NOTE: I repalced "buffer" with "fold buffer" just to keep clear that this is different than our global buffer.
-But if you think that distinction is not necessary feel free to switch back!
 
@@
   rmse(truth = growth, estimate = growth_hat) %>%
-  pull(.estimate) %>%
-  round(3)
+  pull(.estimate)
 model_rmse_cv

### 2021-02-25 paper/forestecology.tex

@@
 \begin{abstract}
-1. Many models for the growth of trees that incorporate the effect of
-interspecies competition are based on a neighborhood effect assumption
-whereby all trees within a fixed distance of all focal trees are
-considered competitors. Methods and tools are needed to quantify this
-competitive effect and assess the quality of all resulting models\\
+1. Many models for the growth of trees that incorporate interspecies
+competition are based on a neighborhood effect assumption whereby all
+trees within a fixed distance of a focal tree are considered
+competitors. Methods and tools are needed to quantify this competitive
+effect and assess the quality of all resulting models.\\
 2. We present the \texttt{forestecology} package providing methods for
@@
 functions with standardized inputs/outputs of simple features
-\texttt{sf} package class, and 3) R S3 object-oriented implementation of
-the Bayesian linear regression model. These three facts allow for clear
-articulation of all the steps in the sequence of analysis and easy
+\texttt{sf} package class, and 3) an S3 object-oriented implementation
+of the Bayesian linear regression model. These three facts allow for
+clear articulation of all the steps in the sequence of analysis and easy
 wrangling and visualization of the geospatial forestry data.
@@
 
-\[
-\newcommand{\dbh}{\text{DBH}}
-\newcommand{\biomass}{\text{biomass}}
-\newcommand{\BA}{\text{BA}}
-y_{ij} = f(\vec{x}_{ij}) + \epsilon_{ij} = \beta_{0,j} + \beta_{\dbh,j} \cdot \dbh_{ij} + \sum_{k=1}^{K} \lambda_{jk} \cdot \BA_{ijk} + \epsilon_{ij}
-\]
+\begin{equation}
+\label{eq:model}
+y_{ij} = f(\vec{x}_{ij}) + \epsilon_{ij} = \beta_{0,j} + \beta_{\text{dbh},j} \cdot \text{dbh}_{ij} + \sum_{k=1}^{K} \lambda_{jk} \cdot \text{BA}_{ijk} + \epsilon_{ij}
+\end{equation}
 
@@
 \CommentTok{## dimension:      XY}
-\CommentTok{## bbox:           xmin: 0.2 ymin: 300 xmax: 299.9 ymax: 600}
+\CommentTok{## bbox:           xmin: 0.2 ymin: 300 xmax: 300 ymax: 600}
 \CommentTok{## CRS:            NA}
@@
 \CommentTok{##    <dbl> <fct> <dbl> <chr>  <chr>  <dbl> <chr>   <dbl>    <POINT>}
-\CommentTok{##  1      4 nysy  13.6  M      A       14.2 M       0.103  (14.2 428.5)}
-\CommentTok{##  2      5 havi   8.8  M      A        9.6 M;P     0.150   (9.4 436.4)}
+\CommentTok{## 1      4 nysy  13.6  M      A       14.2 M       0.103 (14.2 428)}
+\CommentTok{## 2      5 havi   8.8  M      A        9.6 M;P     0.150  (9.4 436)}
 \CommentTok{## 3      6 havi   3.25 NULL   A        4   M       0.140  (1.3 434)}
-\CommentTok{##  4     77 qual  65.2  M      A       66   M       0.141  (34.7 307.2)}
-\CommentTok{##  5     79 tiam  47.7  M      A       46.8 M      -0.161    (40 381.1)}
-\CommentTok{##  6     80 caca   5.15 M      A        6.5 M       0.253  (38.7 421.7)}
-\CommentTok{##  7     96 libe   2.3  J;M    A        3.7 M       0.262      (60 310)}
-\CommentTok{##  8    100 caca   5.09 NULL   A       NA   DN     NA      (52.5 476.3)}
-\CommentTok{##  9    101 litu  65.4  M      A       68.4 M       0.552  (47.1 567.3)}
-\CommentTok{## 10    102 astr   1.99 NULL   A        2.5 M       0.0954 (40.8 575.5)}
-\CommentTok{## # ... with 7,944 more rows}
+\CommentTok{## 4     77 qual  65.2  M      A       66   M       0.141 (34.7 307)}
+\CommentTok{## 5     79 tiam  47.7  M      A       46.8 M      -0.161   (40 381)}
+\CommentTok{## # ... with 7,949 more rows}
 \end{Highlighting}
@@
 The output \texttt{growth\_scbi} is a single data frame of class
-\texttt{sf} that includes \texttt{growth}, the average annual growth in
-DBH (cm y\textsuperscript{-1}) for all stems that were alive at both
-time points, as well a \texttt{geometry} variable encoding each tree's
-geolocation. Furthermore, the variables that should remain unchanged
-between censuses appear only once, such as location variables
-\texttt{gx} and \texttt{gy}; as well as species-related variables.
-Variables that should change between censuses are suffixed with
-\texttt{1} and \texttt{2} indicating the earlier and later censuses,
-such as \texttt{dbh1/dbh2} and \texttt{codes1/codes2}.
-
+\texttt{sf} that includes variables \texttt{growth}, the average annual
+growth in DBH (cm y\textsuperscript{-1}) for all stems that were alive
+at both time points, and \texttt{geometry}, the \texttt{sf} package's
+encoding of geolocations of type
+\texttt{\textless{}POINT\textgreater{}}. Furthermore, the variables that
+should remain unchanged between censuses appear only once, such as
+location variables \texttt{gx} and \texttt{gy}; as well as
+species-related variables. Variables that should change between censuses
+are suffixed with \texttt{1} and \texttt{2} indicating the earlier and
+later censuses, such as \texttt{dbh1/dbh2} and \texttt{codes1/codes2}.
 The data format of other sites may be such that our
@@
 end all that matters is that the growth of all stems is saved in a data
-frame of class \texttt{sf} whereby the geolocation of each tree is
-presented in a \texttt{geometry} variable of type
-\texttt{\textless{}POINT\textgreater{}} and at a minimum the data
-contains the variables above (DAVE NOTE: THis isn't exaclty clear to me
-what variables above means. Maybe list all the variables needed. Do they
-have to have exact names?)
+frame of class \texttt{sf} and at a minimum the data contains the
+variables above. TODO: enumerate what these variables are and say if
+they need exact names
 
@@
 therefore incorporate this spatial dependence into the cross-validation
-algorithm by randomly resampling spatial blocks of trees
+algorithm with our spatial blocks of trees.
 \citet{roberts_cross-validation_2017}
-\citet{pohjankukka_estimating_2017}. (DAVE NOTE: I don't want to change
-somethign I don't understand, but ``randomly resampling'' seems wrong to
-me. I don't see anything random about our block structure. Seems to me
-like ``incorpoate this spatial dependence into the cross-validation
-algorithm with our spatial blocks of trees.'')
+\citet{pohjankukka_estimating_2017}.
 
@@
   \texttt{cv\_block\_size}.}
-
-DAVE NOTE::there are a couple of places where you convert to factor.
-Earlier with \texttt{sp} and then twice in the code block below. ANy
-reason we can't just do this ``under the hood'' in the package code?
-Also any reason with swtiched from undercase with underscores to
-\texttt{SpatialBlock\_scbi} in the naming?
 
@@
 \CommentTok{# Associate each observation to a fold}
-\NormalTok{SpatialBlock_scbi <-}\StringTok{ }\KeywordTok{spatialBlock}\NormalTok{(}
+\NormalTok{spatial_block_scbi <-}\StringTok{ }\KeywordTok{spatialBlock}\NormalTok{(}
   \DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =}\NormalTok{ n_fold, }\DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{, }
@@
 \NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ SpatialBlock_scbi}\OperatorTok{$}\NormalTok{foldID }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ spatial_block_scbi}\OperatorTok{$}\NormalTok{foldID }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
 \end{Highlighting}
@@
 \CommentTok{##      <dbl> <fct>    <dbl>    <POINT>  <dbl> <list>           }
-\CommentTok{##  1        4 nysy     13.6  (14.2 428.5)  0.103  <tibble [20 x 4]>}
-\CommentTok{##  2        5 havi      8.8   (9.4 436.4)  0.150  <tibble [32 x 4]>}
-\CommentTok{##  3       79 tiam     47.7    (40 381.1) -0.161  <tibble [20 x 4]>}
-\CommentTok{##  4       80 caca      5.15 (38.7 421.7)  0.253  <tibble [12 x 4]>}
+\CommentTok{## 1        4 nysy     13.6  (14.2 428)  0.103 <tibble [20 x 4]>}
+\CommentTok{## 2        5 havi      8.8   (9.4 436)  0.150 <tibble [32 x 4]>}
+\CommentTok{## 3       79 tiam     47.7    (40 381) -0.161 <tibble [20 x 4]>}
+\CommentTok{## 4       80 caca      5.15 (38.7 422)  0.253 <tibble [12 x 4]>}
 \CommentTok{## 5       96 libe      2.3    (60 310)  0.262 <tibble [14 x 4]>}
-\CommentTok{##  6      101 litu     65.4  (47.1 567.3)  0.552  <tibble [19 x 4]>}
-\CommentTok{##  7      102 astr      1.99 (40.8 575.5)  0.0954 <tibble [44 x 4]>}
-\CommentTok{##  8      126 cato     37.4  (60.6 400.2)  0.165  <tibble [16 x 4]>}
-\CommentTok{##  9      127 caca      8.72 (72.7 514.1)  0.0370 <tibble [14 x 4]>}
-\CommentTok{## 10      139 astr      1.71 (96.7 315.1)  0.0549 <tibble [48 x 4]>}
-\CommentTok{## # ... with 6,286 more rows}
+\CommentTok{## # ... with 6,291 more rows}
 \end{Highlighting}
@@
 \CommentTok{##      <dbl> <fct>    <dbl> <fct>                    <POINT>  <dbl>}
-\CommentTok{##  1        4 nysy     13.6  1                   (14.2 428.5)  0.103 }
-\CommentTok{##  2        5 havi      8.8  1                    (9.4 436.4)  0.150 }
-\CommentTok{##  3       79 tiam     47.7  1                     (40 381.1) -0.161 }
-\CommentTok{##  4       80 caca      5.15 1                   (38.7 421.7)  0.253 }
-\CommentTok{##  5       96 libe      2.3  1                       (60 310)  0.262 }
-\CommentTok{##  6      101 litu     65.4  3                   (47.1 567.3)  0.552 }
-\CommentTok{##  7      102 astr      1.99 3                   (40.8 575.5)  0.0954}
-\CommentTok{##  8      126 cato     37.4  1                   (60.6 400.2)  0.165 }
-\CommentTok{##  9      127 caca      8.72 3                   (72.7 514.1)  0.0370}
-\CommentTok{## 10      139 astr      1.71 1                   (96.7 315.1)  0.0549}
-\CommentTok{## # ... with 6,286 more rows, and 2 more variables: comp <list>,}
+\CommentTok{## 1        4 nysy     13.6  1                     (14.2 428)  0.103}
+\CommentTok{## 2        5 havi      8.8  1                      (9.4 436)  0.150}
+\CommentTok{## 3       79 tiam     47.7  1                       (40 381) -0.161}
+\CommentTok{## 4       80 caca      5.15 1                     (38.7 422)  0.253}
+\CommentTok{## 5       96 libe      2.3  1                       (60 310)  0.262}
+\CommentTok{## # ... with 6,291 more rows, and 2 more variables: comp <list>,}
 \CommentTok{## #   growth_hat <dbl>}
@@
 We then compare the observed and fitted/predicted growths to compute the
-root mean squared error (RMSE) of our model fit. DAVE NOTE: I rounded
-all the RMSEs, in ecology convention to not show WAY more digits than
-resolution of the measurement.
+root mean squared error (RMSE) of our model fit.
 
@@
 \StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{round}\NormalTok{(}\DecValTok{3}\NormalTok{)}
+\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
 \NormalTok{model_rmse}
@@
 
-These plots give the posterior distributions of parameters from equation
-1 (DAVE NOTE:SOrry couldn't figure out how to make equation references
-work). For many pacakge users they will be of interest because they give
-insight into the species-specific competitive interactions. Setting
-\texttt{type\ =\ "intercepts"} gives posterior distributions for
+These plots give the posterior distributions of parameters from Equation
+\ref{eq:model}. For many pacakge users they will be of interest because
+they give insight into the species-specific competitive interactions.
+Setting \texttt{type\ =\ "intercepts"} gives posterior distributions for
 \(\beta_{0,j}\) and \texttt{type\ =\ "dbh\_slopes"} for
@@
 \StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{growth_hat_shuffle =} \KeywordTok{predict}\NormalTok{(comp_bayes_lm_scbi_shuffle, }\DataTypeTok{newdata =}\NormalTok{ focal_vs_comp_scbi))}
+\end{Highlighting}
+\end{Shaded}
 
+\begin{Shaded}
+\begin{Highlighting}[]
 \NormalTok{model_rmse_shuffle <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
 \StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat_shuffle) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{round}\NormalTok{(}\DecValTok{3}\NormalTok{)}
+\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
 \NormalTok{model_rmse_shuffle}
@@
 of the test set is computed; all trees falling within this fold buffer
-are excluded from the training set. DAVE NOTE: I repalced ``buffer''
-with ``fold buffer'' just to keep clear that this is different than our
-global buffer. But if you think that distinction is not necessary feel
-free to switch back!
+are excluded from the training set.
 
@@
 \StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{round}\NormalTok{(}\DecValTok{3}\NormalTok{)}
+\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
 \NormalTok{model_rmse_cv}

### 2021-02-28 paper/forestecology.Rmd
simon's edits to the Example section!

@@
 We demonstrate the `forestecology` package's functionality on data from the Smithsonian Conservation Biology Institute (SCBI) large forest dynamics plot, located at the Smithsonian's National Zoo and Conservation Biology Institute in Front Royal, VA, USA @bourg_initial_2013.
-The 25.6 ha (640 x 400 m) plot is located at the intersection of three of the major physiographic provinces of the eastern US: the Blue Ridge, Ridge and Valley, and Piedmont provinces and is adjacent to the northern end of Shenandoah National Park.
+The 25.6 ha (640 x 400 m) plot is located at the intersection of three of the major physiographic provinces of the eastern US---the Blue Ridge, Ridge and Valley, and Piedmont provinces---and is adjacent to the northern end of Shenandoah National Park.
 The forest type is typical mature secondary eastern mixed deciduous forest, with a canopy dominated by tulip poplar (*Liriodendron tulipifera*), oaks (*Quercus* spp.), and hickories (*Carya* spp.), and an understory composed mainly of spicebush (*Lindera benzoin*), paw-paw (*Asimina triloba*), American hornbeam (*Carpinus caroliniana*), and witch hazel (*Hamamelis virginiana*) @bourg_initial_2013.
@@
 
-These two data frames are then used as arguments to the `compute_growth()` function, along with the `id` argument that specifies the variable that uniquely identifies each tree-stem.
+These two data frames are then supplied as arguments to the `compute_growth()` function, along with the `id` argument that specifies the variable that uniquely identifies each tree-stem.
 Note furthermore that we discard all resprouts in the later census (those with `code == R`), since we are only interested in the diameter growth of surviving, and not resprouted, stems.
@@
 The data format of other sites may be such that our `compute_growth()` function doesn't work at all.
-However, in the end all that matters is that the growth of all stems is saved in a data frame of class `sf` and at a minimum the data contains the variables above.
+However, in the end, all that matters is that the growth of all stems is saved in a data frame of class `sf` and, at a minimum, the data contains the variables above.
 TODO: enumerate what these variables are and say if they need exact names
@@
 Conventional cross-validation algorithms assign observations to folds by randomly resampling individual observations.
-However, an assumption that the observations are independent underlies any cross-validation algorithm.
+However, many of these algorithms assume that the observations are independent of each other.
 In the case of forest census data, observations exhibit spatial autocorrelation.
@@
 These plots give the posterior distributions of parameters from Equation \ref{eq:model}.
-For many pacakge users they will be of interest because they give insight into the species-specific competitive interactions.
+For many package users they will be of interest because they give insight into the species-specific competitive interactions.
 Setting `type = "intercepts"` gives posterior distributions for $\beta_{0,j}$ and `type = "dbh_slopes"` for $\beta_{dbh,j}$.
@@
 The resampling with replacement is nested within each focal tree in order to preserve the neighborhood structure of our competition model.
-To run the permutation test, we use the same`comp_bayes_lm()` function as in Section \ref{model-fit-predict}, but with a `run_shuffle = TRUE` argument.
+To run the permutation test, we use the same `comp_bayes_lm()` function as in Section \ref{model-fit-predict}, but with a `run_shuffle = TRUE` argument.
 

### 2021-03-01 paper/forestecology.Rmd
Incorporated @simonpcouch comments

--- 2021-02-25 paper/forestecology.Rmd
@@
 Variables that should change between censuses are suffixed with `1` and `2` indicating the earlier and later censuses, such as `dbh1/dbh2` and `codes1/codes2`.
-The data format of other sites may be such that our `compute_growth()` function doesn't work at all.
+Site data that does not align with this convention will need to be transformed for use with the `compute_growth()` function.
 However, in the end all that matters is that the growth of all stems is saved in a data frame of class `sf` and at a minimum the data contains the variables above.
@@
 In our case we set this value at 7.5m, a value informed by @canham_neighborhood_2004 @uriarte_spatially_2004 @canham_neighborhood_2006.
-Using this value along with a manually constructed `sf` object representation of the study region's boundary, we apply the `add_buffer_variable()` to our `growth_scbi` data frame to add a `buffer` boolean variable: all trees who have `buffer` set to `FALSE` will be our focal trees whose growth will be modeled, whereas those with `TRUE` will only be considered as competitor trees whose growth will not.
+Using this value along with a manually constructed `sf` object representation of the study region's boundary via its vertices, we apply the `add_buffer_variable()` to our `growth_scbi` data frame to add a `buffer` boolean variable: all trees who have `buffer` set to `FALSE` will be our focal trees whose growth will be modeled, whereas those with `TRUE` will only be considered as competitor trees whose growth will not.
 
@@
 The `create_focal_vs_comp()` functions performs these tasks and returns a new data frame of type `sf`.
-On top of the previous arguments `comp_dist` defining the competition neighborhood and `id` indicating which variable uniquely identifies each tree-stem, this function also requires an `sf` object representation of the spatial cross-validation blocks/folds; in our case, this was manually encoded in the `blocks_scbi` in Section \ref{spatial-information} while in our Appendix we present an example where this was performed using `spatialBlock()` from the `blockCV` package.
+On top of the previous arguments `comp_dist` defining the competition neighborhood and `id` indicating which variable uniquely identifies each tree-stem, this function also requires an `sf` object representation of the spatial cross-validation blocks/folds.
+In this example, the blocks were manually encoded in `blocks_scbi` by specifying it's vertices in Section \ref{spatial-information}^[We present an alternative method for defining spatial cross-validation blocks is using the `spatialBlock()` function from the `blockCV` package in the Appendix].
 We present the resulting data frame below with the `foldID` variable omitted for compactness of presentation.
@@
 Most notably however is a new variable `comp` which contains information on all competitor trees for a given focal tree, saved in `tidyr` package list-column format @tidyr_package.
-For example, we drill-down on the tree with `focal_ID` 4, which has 20 competitor trees each described by 4 variables as indicated by the fact that `comp` is a `<tibble [20 × 4]>`.
+For example, we drill-down on the tree with `focal_ID` 4, which has 20 competitor trees each described by 4 variables as indicated by the fact that `comp` is a `tibble [20 × 4]`.
 
@@
 Currently the `forestecology` package can only fit the competition Bayesian linear regression model outlined in Section \ref{competition-model} using the `comp_bayes_lm()` function.
-However, any model implemented in a function that similarly takes an input data frame of type `focal_vs_comp` as an argument can be easily swapped in.
+However, any model implemented in a function that similarly takes an input data frame of type `focal_vs_comp` as an argument can also be used.
 For our specific competition Bayesian linear regression model, we also specify prior distributions on all parameters of interest (here chosen to be the defaults as specified in `?comp_bayes_lm`).

### 2021-03-01 paper/forestecology.Rmd

@@
 Site data that does not align with this convention will need to be transformed for use with the `compute_growth()` function.
-However, in the end all that matters is that the growth of all stems is saved in a data frame of class `sf` and at a minimum the data contains the variables above.
+However, in the end, all that matters is that the growth of all stems is saved in a data frame of class `sf` and, at a minimum, the data contains the variables above.
 TODO: enumerate what these variables are and say if they need exact names

### 2021-03-01 paper/forestecology.tex

@@
 {
-  \title{\bf The forestecology R package for fitting and assessing models of
-interspecies competitive effects on the growth of trees}
+  \title{\bf The forestecology R package for fitting and assessing
+models of interspecies competitive effects on the growth of trees}
 
@@
   \begin{center}
-    {\LARGE\bf The forestecology R package for fitting and assessing models of
-interspecies competitive effects on the growth of trees}
+    {\LARGE\bf The forestecology R package for fitting and assessing
+models of interspecies competitive effects on the growth of trees}
   \end{center}
@@
 \noindent%
-{\it Keywords:} forest ecology, competition, R, Rstats, tidyverse, sf, cross-validation,
-spatial statistics
+{\it Keywords:} forest ecology, competition, R, Rstats, tidyverse, sf,
+cross-validation, spatial statistics
 \vfill
@@
 the intersection of three of the major physiographic provinces of the
-eastern US: the Blue Ridge, Ridge and Valley, and Piedmont provinces and
-is adjacent to the northern end of Shenandoah National Park. The forest
-type is typical mature secondary eastern mixed deciduous forest, with a
-canopy dominated by tulip poplar (\emph{Liriodendron tulipifera}), oaks
-(\emph{Quercus} spp.), and hickories (\emph{Carya} spp.), and an
-understory composed mainly of spicebush (\emph{Lindera benzoin}),
-paw-paw (\emph{Asimina triloba}), American hornbeam (\emph{Carpinus
-caroliniana}), and witch hazel (\emph{Hamamelis virginiana})
-\citet{bourg_initial_2013}.
+eastern US---the Blue Ridge, Ridge and Valley, and Piedmont
+provinces---and is adjacent to the northern end of Shenandoah National
+Park. The forest type is typical mature secondary eastern mixed
+deciduous forest, with a canopy dominated by tulip poplar
+(\emph{Liriodendron tulipifera}), oaks (\emph{Quercus} spp.), and
+hickories (\emph{Carya} spp.), and an understory composed mainly of
+spicebush (\emph{Lindera benzoin}), paw-paw (\emph{Asimina triloba}),
+American hornbeam (\emph{Carpinus caroliniana}), and witch hazel
+(\emph{Hamamelis virginiana}) \citet{bourg_initial_2013}.
 
@@
 \begin{Highlighting}[]
-\KeywordTok{library}\NormalTok{(tidyverse)}
-\KeywordTok{library}\NormalTok{(lubridate)}
-\KeywordTok{library}\NormalTok{(sf)}
-\KeywordTok{library}\NormalTok{(patchwork)}
-\KeywordTok{library}\NormalTok{(forestecology)}
-\KeywordTok{library}\NormalTok{(blockCV)}
+\FunctionTok{library}\NormalTok{(tidyverse)}
+\FunctionTok{library}\NormalTok{(lubridate)}
+\FunctionTok{library}\NormalTok{(sf)}
+\FunctionTok{library}\NormalTok{(patchwork)}
+\FunctionTok{library}\NormalTok{(forestecology)}
+\FunctionTok{library}\NormalTok{(blockCV)}
 \end{Highlighting}
@@
 convert the character variable with the date of measurement to be of
-explicit type \texttt{date}, convert DBH to be in centimeters\footnote{A
-  rule of thumb to determine the units of DBH is to check if the
-  smallest non-zero and non-missing measurement is 1 or 10. If the
-  former, then centimeters. If the later, then millimeters. This is
-  because ForestGEO protocols state that only trees with DBH greater or
-  equal to 1cm should be included in censuses.}, convert the \texttt{sp}
-variable containing species information from type \texttt{chr} character
-to \texttt{fct} factor.\footnote{In our spatial cross-validation
-  algorithm in Section \ref{spatial-cross-validation} issues can occur
-  when rare species do not occur in the training set, but then are
-  encountered in the test set. This risk is mitigated by representing
-  \texttt{sp} as a factor variable, which has a complete list of all
-  levels of the categorical variable.} Furthermore, in order to speed up
-computation for purposes of this example, we only consider a 9 ha
-subsection of the 25.6 ha of the SCBI site: \texttt{gx} from 0--300
+explicit type \texttt{date}, convert DBH to be in
+centimeters\footnote{A rule of thumb to determine the units of DBH is to
+  check if the smallest non-zero and non-missing measurement is 1 or 10.
+  If the former, then centimeters. If the later, then millimeters. This
+  is because ForestGEO protocols state that only trees with DBH greater
+  or equal to 1cm should be included in censuses.}, convert the
+\texttt{sp} variable containing species information from type
+\texttt{chr} character to \texttt{fct} factor.\footnote{In our spatial
+  cross-validation algorithm in Section \ref{spatial-cross-validation}
+  issues can occur when rare species do not occur in the training set,
+  but then are encountered in the test set. This risk is mitigated by
+  representing \texttt{sp} as a factor variable, which has a complete
+  list of all levels of the categorical variable.} Furthermore, in order
+to speed up computation for purposes of this example, we only consider a
+9 ha subsection of the 25.6 ha of the SCBI site: \texttt{gx} from 0--300
 instead of 0--400 and \texttt{gy} from 300--600 instead of 0--640.
@@
 \begin{Highlighting}[]
-\NormalTok{census_}\DecValTok{2013}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem2.csv"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
-    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date),}
-    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}\NormalTok{,}
-    \DataTypeTok{sp =} \KeywordTok{factor}\NormalTok{(sp)}
-\NormalTok{  ) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+\NormalTok{census\_2013\_scbi }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"scbi.stem2.csv"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{select}\NormalTok{(stemID, sp, }\AttributeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}
+    \AttributeTok{date =} \FunctionTok{mdy}\NormalTok{(date),}
+    \AttributeTok{dbh =} \FunctionTok{as.numeric}\NormalTok{(dbh)}\SpecialCharTok{/}\DecValTok{10}\NormalTok{,}
+    \AttributeTok{sp =} \FunctionTok{factor}\NormalTok{(sp)}
+\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{filter}\NormalTok{(gx }\SpecialCharTok{\textless{}} \DecValTok{300}\NormalTok{, }\FunctionTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
 
-\NormalTok{census_}\DecValTok{2018}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem3.csv"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
-    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date),}
-    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}\NormalTok{,}
-    \DataTypeTok{sp =} \KeywordTok{factor}\NormalTok{(sp)}
-\NormalTok{  ) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+\NormalTok{census\_2018\_scbi }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"scbi.stem3.csv"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{select}\NormalTok{(stemID, sp, }\AttributeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}
+    \AttributeTok{date =} \FunctionTok{mdy}\NormalTok{(date),}
+    \AttributeTok{dbh =} \FunctionTok{as.numeric}\NormalTok{(dbh)}\SpecialCharTok{/}\DecValTok{10}\NormalTok{,}
+    \AttributeTok{sp =} \FunctionTok{factor}\NormalTok{(sp)}
+\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{filter}\NormalTok{(gx }\SpecialCharTok{\textless{}} \DecValTok{300}\NormalTok{, }\FunctionTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
 \end{Highlighting}
@@
 
-These two data frames are then used as arguments to the
+These two data frames are then supplied as arguments to the
 \texttt{compute\_growth()} function, along with the \texttt{id} argument
@@
 \begin{Highlighting}[]
-\NormalTok{growth_scbi <-}
-\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(}
-    \DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi,}
-    \DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)),}
-    \DataTypeTok{id =} \StringTok{"stemID"}
+\NormalTok{growth\_scbi }\OtherTok{\textless{}{-}}
+  \FunctionTok{compute\_growth}\NormalTok{(}
+    \AttributeTok{census\_1 =}\NormalTok{ census\_2013\_scbi,}
+    \AttributeTok{census\_2 =}\NormalTok{ census\_2018\_scbi }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{str\_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)),}
+    \AttributeTok{id =} \StringTok{"stemID"}
 \NormalTok{  )}
-\NormalTok{growth_scbi}
-\CommentTok{## Simple feature collection with 7954 features and 8 fields}
-\CommentTok{## geometry type:  POINT}
-\CommentTok{## dimension:      XY}
-\CommentTok{## bbox:           xmin: 0.2 ymin: 300 xmax: 300 ymax: 600}
-\CommentTok{## CRS:            NA}
-\CommentTok{## # A tibble: 7,954 x 9}
-\CommentTok{##   stemID sp     dbh1 codes1 status  dbh2 codes2 growth   geometry}
-\CommentTok{##    <dbl> <fct> <dbl> <chr>  <chr>  <dbl> <chr>   <dbl>    <POINT>}
-\CommentTok{## 1      4 nysy  13.6  M      A       14.2 M       0.103 (14.2 428)}
-\CommentTok{## 2      5 havi   8.8  M      A        9.6 M;P     0.150  (9.4 436)}
-\CommentTok{## 3      6 havi   3.25 NULL   A        4   M       0.140  (1.3 434)}
-\CommentTok{## 4     77 qual  65.2  M      A       66   M       0.141 (34.7 307)}
-\CommentTok{## 5     79 tiam  47.7  M      A       46.8 M      -0.161   (40 381)}
-\CommentTok{## # ... with 7,949 more rows}
+\NormalTok{growth\_scbi}
+\DocumentationTok{\#\# Simple feature collection with 7954 features and 8 fields}
+\DocumentationTok{\#\# geometry type:  POINT}
+\DocumentationTok{\#\# dimension:      XY}
+\DocumentationTok{\#\# bbox:           xmin: 0.2 ymin: 300 xmax: 300 ymax: 600}
+\DocumentationTok{\#\# CRS:            NA}
+\DocumentationTok{\#\# \# A tibble: 7,954 x 9}
+\DocumentationTok{\#\#   stemID sp     dbh1 codes1 status  dbh2 codes2 growth}
+\DocumentationTok{\#\#    \textless{}dbl\textgreater{} \textless{}fct\textgreater{} \textless{}dbl\textgreater{} \textless{}chr\textgreater{}  \textless{}chr\textgreater{}  \textless{}dbl\textgreater{} \textless{}chr\textgreater{}   \textless{}dbl\textgreater{}}
+\DocumentationTok{\#\# 1      4 nysy  13.6  M      A       14.2 M       0.103}
+\DocumentationTok{\#\# 2      5 havi   8.8  M      A        9.6 M;P     0.150}
+\DocumentationTok{\#\# 3      6 havi   3.25 NULL   A        4   M       0.140}
+\DocumentationTok{\#\# 4     77 qual  65.2  M      A       66   M       0.141}
+\DocumentationTok{\#\# 5     79 tiam  47.7  M      A       46.8 M      {-}0.161}
+\DocumentationTok{\#\# \# ... with 7,949 more rows, and 1 more variable: geometry \textless{}POINT\textgreater{}}
 \end{Highlighting}
@@
 \texttt{compute\_growth()} function doesn't work at all. However, in the
-end all that matters is that the growth of all stems is saved in a data
-frame of class \texttt{sf} and at a minimum the data contains the
+end, all that matters is that the growth of all stems is saved in a data
+frame of class \texttt{sf} and, at a minimum, the data contains the
 variables above. TODO: enumerate what these variables are and say if
@@
 \begin{Highlighting}[]
-\CommentTok{# }\AlertTok{TODO}\CommentTok{: Rescale points in this plot}
-\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_scbi, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{size =}\NormalTok{ growth)) }\OperatorTok{+}\StringTok{ }
-\StringTok{  }\KeywordTok{scale_size}\NormalTok{(}\DataTypeTok{breaks =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.01}\NormalTok{, }\FloatTok{0.1}\NormalTok{, }\DecValTok{1}\NormalTok{), }\DataTypeTok{range =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\DecValTok{1}\NormalTok{))}
+\CommentTok{\# }\AlertTok{TODO}\CommentTok{: Rescale points in this plot}
+\FunctionTok{ggplot}\NormalTok{() }\SpecialCharTok{+}
+  \FunctionTok{geom\_sf}\NormalTok{(}\AttributeTok{data =}\NormalTok{ growth\_scbi, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{size =}\NormalTok{ growth)) }\SpecialCharTok{+} 
+  \FunctionTok{scale\_size}\NormalTok{(}\AttributeTok{breaks =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.01}\NormalTok{, }\FloatTok{0.1}\NormalTok{, }\DecValTok{1}\NormalTok{), }\AttributeTok{range =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\DecValTok{1}\NormalTok{))}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\CommentTok{# Define buffer region using competitive distance range}
-\NormalTok{comp_dist <-}\StringTok{ }\FloatTok{7.5}
+\CommentTok{\# Define buffer region using competitive distance range}
+\NormalTok{comp\_dist }\OtherTok{\textless{}{-}} \FloatTok{7.5}
 
-\NormalTok{study_region_scbi <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}
-  \DataTypeTok{x =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{),}
-  \DataTypeTok{y =} \KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{300}\NormalTok{)}
-\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{sf_polygon}\NormalTok{()}
+\NormalTok{study\_region\_scbi }\OtherTok{\textless{}{-}} \FunctionTok{tibble}\NormalTok{(}
+  \AttributeTok{x =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{),}
+  \AttributeTok{y =} \FunctionTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{300}\NormalTok{)}
+\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{sf\_polygon}\NormalTok{()}
 
-\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{size =}\NormalTok{ comp_dist, }\DataTypeTok{region =}\NormalTok{ study_region_scbi)}
+\NormalTok{growth\_scbi }\OtherTok{\textless{}{-}}\NormalTok{ growth\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{add\_buffer\_variable}\NormalTok{(}\AttributeTok{size =}\NormalTok{ comp\_dist, }\AttributeTok{region =}\NormalTok{ study\_region\_scbi)}
 \end{Highlighting}
@@
 assign observations to folds by randomly resampling individual
-observations. However, an assumption that the observations are
-independent underlies any cross-validation algorithm. In the case of
-forest census data, observations exhibit spatial autocorrelation. We
-therefore incorporate this spatial dependence into the cross-validation
-algorithm with our spatial blocks of trees.
-\citet{roberts_cross-validation_2017}
+observations. However, many of these algorithms assume that the
+observations are independent of each other. In the case of forest census
+data, observations exhibit spatial autocorrelation. We therefore
+incorporate this spatial dependence into the cross-validation algorithm
+with our spatial blocks of trees. \citet{roberts_cross-validation_2017}
 \citet{pohjankukka_estimating_2017}.
@@
 associate each tree in \texttt{growth\_scbi} to the correct fold (saved
-in the \texttt{foldID} variable) \citet{valavi_blockcv_2019}. \footnote{In
-  the Appendix we present an example where the folds themselves are also
-  created using the \texttt{spatialBlock()} function given a specified
-  \texttt{cv\_block\_size}.}
+in the \texttt{foldID} variable) \citet{valavi_blockcv_2019}.
+\footnote{In the Appendix we present an example where the folds
+  themselves are also created using the \texttt{spatialBlock()} function
+  given a specified \texttt{cv\_block\_size}.}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# Manually define spatial blocks to act as folds}
-\NormalTok{n_fold <-}\StringTok{ }\DecValTok{4}
-\NormalTok{fold1 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{450}\NormalTok{))}
-\NormalTok{fold2 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{450}\NormalTok{))}
-\NormalTok{fold3 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{600}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{600}\NormalTok{))}
-\NormalTok{fold4 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+\CommentTok{\# Manually define spatial blocks to act as folds}
+\NormalTok{n\_fold }\OtherTok{\textless{}{-}} \DecValTok{4}
+\NormalTok{fold1 }\OtherTok{\textless{}{-}} \FunctionTok{rbind}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{450}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{450}\NormalTok{))}
+\NormalTok{fold2 }\OtherTok{\textless{}{-}} \FunctionTok{rbind}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{450}\NormalTok{))}
+\NormalTok{fold3 }\OtherTok{\textless{}{-}} \FunctionTok{rbind}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{450}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{450}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{600}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+\NormalTok{fold4 }\OtherTok{\textless{}{-}} \FunctionTok{rbind}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{450}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{600}\NormalTok{))}
 
-\NormalTok{blocks_scbi <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}
-  \KeywordTok{sf_polygon}\NormalTok{(fold1), }\KeywordTok{sf_polygon}\NormalTok{(fold2), }\KeywordTok{sf_polygon}\NormalTok{(fold3), }\KeywordTok{sf_polygon}\NormalTok{(fold4)}
-\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\NormalTok{n_fold) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
+\NormalTok{blocks\_scbi }\OtherTok{\textless{}{-}} \FunctionTok{bind\_rows}\NormalTok{(}
+  \FunctionTok{sf\_polygon}\NormalTok{(fold1), }\FunctionTok{sf\_polygon}\NormalTok{(fold2), }\FunctionTok{sf\_polygon}\NormalTok{(fold3), }\FunctionTok{sf\_polygon}\NormalTok{(fold4)}
+\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{folds =} \FunctionTok{c}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\NormalTok{n\_fold) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{factor}\NormalTok{())}
 
-\CommentTok{# Associate each observation to a fold}
-\NormalTok{spatial_block_scbi <-}\StringTok{ }\KeywordTok{spatialBlock}\NormalTok{(}
-  \DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =}\NormalTok{ n_fold, }\DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{, }
-  \DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}
+\CommentTok{\# Associate each observation to a fold}
+\NormalTok{spatial\_block\_scbi }\OtherTok{\textless{}{-}} \FunctionTok{spatialBlock}\NormalTok{(}
+  \AttributeTok{speciesData =}\NormalTok{ growth\_scbi, }\AttributeTok{k =}\NormalTok{ n\_fold, }\AttributeTok{selection =} \StringTok{"systematic"}\NormalTok{, }
+  \AttributeTok{blocks =}\NormalTok{ blocks\_scbi, }\AttributeTok{showBlocks =} \ConstantTok{FALSE}\NormalTok{, }\AttributeTok{verbose =} \ConstantTok{FALSE}
 \NormalTok{)}
 
-\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ spatial_block_scbi}\OperatorTok{$}\NormalTok{foldID }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
+\NormalTok{growth\_scbi }\OtherTok{\textless{}{-}}\NormalTok{ growth\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{foldID =}\NormalTok{ spatial\_block\_scbi}\SpecialCharTok{$}\NormalTok{foldID }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{factor}\NormalTok{())}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ blocks_scbi, }\DataTypeTok{fill =} \StringTok{"transparent"}\NormalTok{, }\DataTypeTok{linetype =} \StringTok{"dashed"}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_sf_text}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{sample_n}\NormalTok{(}\DecValTok{1000}\NormalTok{), }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{label =}\NormalTok{ foldID, }\DataTypeTok{col =}\NormalTok{ buffer))}
+\FunctionTok{ggplot}\NormalTok{() }\SpecialCharTok{+}
+  \FunctionTok{geom\_sf}\NormalTok{(}\AttributeTok{data =}\NormalTok{ blocks\_scbi, }\AttributeTok{fill =} \StringTok{"transparent"}\NormalTok{, }\AttributeTok{linetype =} \StringTok{"dashed"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{geom\_sf\_text}\NormalTok{(}\AttributeTok{data =}\NormalTok{ growth\_scbi }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{sample\_n}\NormalTok{(}\DecValTok{1000}\NormalTok{), }\FunctionTok{aes}\NormalTok{(}\AttributeTok{label =}\NormalTok{ foldID, }\AttributeTok{col =}\NormalTok{ buffer))}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\CommentTok{# }\AlertTok{TODO}\CommentTok{: Below reconcile the number of rows as they off by one from growth_scbi %>% }
-\CommentTok{# filter(!is.na(growth) & !buffer). Perhaps by removing NA's in the growth_scbi stage?}
-\NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(comp_dist, }\DataTypeTok{cv_grid_sf =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{)}
-\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\NormalTok{foldID)}
-\CommentTok{## # A tibble: 6,296 x 6}
-\CommentTok{##   focal_ID focal_sp   dbh   geometry growth comp             }
-\CommentTok{##      <dbl> <fct>    <dbl>    <POINT>  <dbl> <list>           }
-\CommentTok{## 1        4 nysy     13.6  (14.2 428)  0.103 <tibble [20 x 4]>}
-\CommentTok{## 2        5 havi      8.8   (9.4 436)  0.150 <tibble [32 x 4]>}
-\CommentTok{## 3       79 tiam     47.7    (40 381) -0.161 <tibble [20 x 4]>}
-\CommentTok{## 4       80 caca      5.15 (38.7 422)  0.253 <tibble [12 x 4]>}
-\CommentTok{## 5       96 libe      2.3    (60 310)  0.262 <tibble [14 x 4]>}
-\CommentTok{## # ... with 6,291 more rows}
+\CommentTok{\# }\AlertTok{TODO}\CommentTok{: Below reconcile the number of rows as they off by one from growth\_scbi \%\textgreater{}\% }
+\CommentTok{\# filter(!is.na(growth) \& !buffer). Perhaps by removing NA\textquotesingle{}s in the growth\_scbi stage?}
+\NormalTok{focal\_vs\_comp\_scbi }\OtherTok{\textless{}{-}}\NormalTok{ growth\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{create\_focal\_vs\_comp}\NormalTok{(comp\_dist, }\AttributeTok{cv\_grid\_sf =}\NormalTok{ blocks\_scbi, }\AttributeTok{id =} \StringTok{"stemID"}\NormalTok{)}
+\NormalTok{focal\_vs\_comp\_scbi }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{select}\NormalTok{(}\SpecialCharTok{{-}}\NormalTok{foldID)}
+\DocumentationTok{\#\# \# A tibble: 6,296 x 6}
+\DocumentationTok{\#\#   focal\_ID focal\_sp   dbh   geometry growth comp             }
+\DocumentationTok{\#\#      \textless{}dbl\textgreater{} \textless{}fct\textgreater{}    \textless{}dbl\textgreater{}    \textless{}POINT\textgreater{}  \textless{}dbl\textgreater{} \textless{}list\textgreater{}           }
+\DocumentationTok{\#\# 1        4 nysy     13.6  (14.2 428)  0.103 \textless{}tibble [20 x 4]\textgreater{}}
+\DocumentationTok{\#\# 2        5 havi      8.8   (9.4 436)  0.150 \textless{}tibble [32 x 4]\textgreater{}}
+\DocumentationTok{\#\# 3       79 tiam     47.7    (40 381) {-}0.161 \textless{}tibble [20 x 4]\textgreater{}}
+\DocumentationTok{\#\# 4       80 caca      5.15 (38.7 422)  0.253 \textless{}tibble [12 x 4]\textgreater{}}
+\DocumentationTok{\#\# 5       96 libe      2.3    (60 310)  0.262 \textless{}tibble [14 x 4]\textgreater{}}
+\DocumentationTok{\#\# \# ... with 6,291 more rows}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{filter}\NormalTok{(focal_ID }\OperatorTok{==}\StringTok{ }\DecValTok{4}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{select}\NormalTok{(focal_ID, dbh, comp)}
-\CommentTok{## # A tibble: 1 x 3}
-\CommentTok{##   focal_ID   dbh comp             }
-\CommentTok{##      <dbl> <dbl> <list>           }
-\CommentTok{## 1        4  13.6 <tibble [20 x 4]>}
+\NormalTok{focal\_vs\_comp\_scbi }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{filter}\NormalTok{(focal\_ID }\SpecialCharTok{==} \DecValTok{4}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{select}\NormalTok{(focal\_ID, dbh, comp)}
+\DocumentationTok{\#\# \# A tibble: 1 x 3}
+\DocumentationTok{\#\#   focal\_ID   dbh comp             }
+\DocumentationTok{\#\#      \textless{}dbl\textgreater{} \textless{}dbl\textgreater{} \textless{}list\textgreater{}           }
+\DocumentationTok{\#\# 1        4  13.6 \textless{}tibble [20 x 4]\textgreater{}}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{filter}\NormalTok{(focal_ID }\OperatorTok{==}\StringTok{ }\DecValTok{4}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{select}\NormalTok{(focal_ID, dbh, comp) }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{unnest}\NormalTok{(}\DataTypeTok{cols =} \StringTok{"comp"}\NormalTok{)}
+\NormalTok{focal\_vs\_comp\_scbi }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{filter}\NormalTok{(focal\_ID }\SpecialCharTok{==} \DecValTok{4}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{select}\NormalTok{(focal\_ID, dbh, comp) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{unnest}\NormalTok{(}\AttributeTok{cols =} \StringTok{"comp"}\NormalTok{)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\CommentTok{# }\AlertTok{TODO}\CommentTok{: Add information about default priors in ?comp_bayes_lm()}
-\NormalTok{comp_bayes_lm_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{comp_bayes_lm}\NormalTok{(}\DataTypeTok{prior_param =} \OtherTok{NULL}\NormalTok{)}
+\CommentTok{\# }\AlertTok{TODO}\CommentTok{: Add information about default priors in ?comp\_bayes\_lm()}
+\NormalTok{comp\_bayes\_lm\_scbi }\OtherTok{\textless{}{-}}\NormalTok{ focal\_vs\_comp\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{comp\_bayes\_lm}\NormalTok{(}\AttributeTok{prior\_param =} \ConstantTok{NULL}\NormalTok{)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{comp_bayes_lm_scbi}
-\CommentTok{## Bayesian linear regression model parameters with a multivariate Normal likelihood. See ?comp_bayes_lm for details:}
-\CommentTok{## }
-\CommentTok{##   parameter_type           prior posterior}
-\CommentTok{## 1 Inverse-Gamma on sigma^2 a_0   a_star   }
-\CommentTok{## 2 Inverse-Gamma on sigma^2 b_0   b_star   }
-\CommentTok{## 3 Multivariate t on beta   mu_0  mu_star  }
-\CommentTok{## 4 Multivariate t on beta   V_0   V_star   }
-\CommentTok{## }
-\CommentTok{## Model formula:}
-\CommentTok{## growth ~ sp + dbh + dbh * sp + acne * sp + acpl * sp + acru * sp + acsp * sp + aial * sp + amar * sp + astr * sp + beth * sp + caca * sp + caco * sp + cade * sp + cagl * sp + caovl * sp + casp * sp + cato * sp + ceca * sp + ceoc * sp + chvi * sp + coal * sp + coam * sp + cofl * sp + crpr * sp + crsp * sp + divi * sp + elum * sp + eual * sp + fagr * sp + fram * sp + frni * sp + frpe * sp + frsp * sp + havi * sp + ilve * sp + juci * sp + juni * sp + juvi * sp + libe * sp + litu * sp + loma * sp + nysy * sp + pato * sp + pipu * sp + pist * sp + pivi * sp + ploc * sp + prav * sp + prpe * sp + prse * sp + prsp * sp + qual * sp + quco * sp + qufa * sp + qumi * sp + qupr * sp + quru * sp + qusp * sp + quve * sp + rhpe * sp + romu * sp + rops * sp + rual * sp + rupe * sp + ruph * sp + saal * sp + saca * sp + tiam * sp + ulam * sp + ulru * sp + ulsp * sp + unk * sp + viac * sp + vipr * sp + vire * sp}
+\NormalTok{comp\_bayes\_lm\_scbi}
+\DocumentationTok{\#\# Bayesian linear regression model parameters with a multivariate Normal likelihood. See ?comp\_bayes\_lm for details:}
+\DocumentationTok{\#\# }
+\DocumentationTok{\#\#   parameter\_type           prior posterior}
+\DocumentationTok{\#\# 1 Inverse{-}Gamma on sigma\^{}2 a\_0   a\_star   }
+\DocumentationTok{\#\# 2 Inverse{-}Gamma on sigma\^{}2 b\_0   b\_star   }
+\DocumentationTok{\#\# 3 Multivariate t on beta   mu\_0  mu\_star  }
+\DocumentationTok{\#\# 4 Multivariate t on beta   V\_0   V\_star   }
+\DocumentationTok{\#\# }
+\DocumentationTok{\#\# Model formula:}
+\DocumentationTok{\#\# growth \textasciitilde{} sp + dbh + dbh * sp + acne * sp + acpl * sp + acru * sp + acsp * sp + aial * sp + amar * sp + astr * sp + beth * sp + caca * sp + caco * sp + cade * sp + cagl * sp + caovl * sp + casp * sp + cato * sp + ceca * sp + ceoc * sp + chvi * sp + coal * sp + coam * sp + cofl * sp + crpr * sp + crsp * sp + divi * sp + elum * sp + eual * sp + fagr * sp + fram * sp + frni * sp + frpe * sp + frsp * sp + havi * sp + ilve * sp + juci * sp + juni * sp + juvi * sp + libe * sp + litu * sp + loma * sp + nysy * sp + pato * sp + pipu * sp + pist * sp + pivi * sp + ploc * sp + prav * sp + prpe * sp + prse * sp + prsp * sp + qual * sp + quco * sp + qufa * sp + qumi * sp + qupr * sp + quru * sp + qusp * sp + quve * sp + rhpe * sp + romu * sp + rops * sp + rual * sp + rupe * sp + ruph * sp + saal * sp + saca * sp + tiam * sp + ulam * sp + ulru * sp + ulsp * sp + unk * sp + viac * sp + vipr * sp + vire * sp}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{growth_hat =} \KeywordTok{predict}\NormalTok{(comp_bayes_lm_scbi, }\DataTypeTok{newdata =}\NormalTok{ focal_vs_comp_scbi))}
+\NormalTok{focal\_vs\_comp\_scbi }\OtherTok{\textless{}{-}}\NormalTok{ focal\_vs\_comp\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{growth\_hat =} \FunctionTok{predict}\NormalTok{(comp\_bayes\_lm\_scbi, }\AttributeTok{newdata =}\NormalTok{ focal\_vs\_comp\_scbi))}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{focal_vs_comp_scbi}
-\CommentTok{## # A tibble: 6,296 x 8}
-\CommentTok{##   focal_ID focal_sp   dbh foldID                  geometry growth}
-\CommentTok{##      <dbl> <fct>    <dbl> <fct>                    <POINT>  <dbl>}
-\CommentTok{## 1        4 nysy     13.6  1                     (14.2 428)  0.103}
-\CommentTok{## 2        5 havi      8.8  1                      (9.4 436)  0.150}
-\CommentTok{## 3       79 tiam     47.7  1                       (40 381) -0.161}
-\CommentTok{## 4       80 caca      5.15 1                     (38.7 422)  0.253}
-\CommentTok{## 5       96 libe      2.3  1                       (60 310)  0.262}
-\CommentTok{## # ... with 6,291 more rows, and 2 more variables: comp <list>,}
-\CommentTok{## #   growth_hat <dbl>}
+\NormalTok{focal\_vs\_comp\_scbi}
+\DocumentationTok{\#\# \# A tibble: 6,296 x 8}
+\DocumentationTok{\#\#   focal\_ID focal\_sp   dbh foldID   geometry growth}
+\DocumentationTok{\#\#      \textless{}dbl\textgreater{} \textless{}fct\textgreater{}    \textless{}dbl\textgreater{} \textless{}fct\textgreater{}     \textless{}POINT\textgreater{}  \textless{}dbl\textgreater{}}
+\DocumentationTok{\#\# 1        4 nysy     13.6  1      (14.2 428)  0.103}
+\DocumentationTok{\#\# 2        5 havi      8.8  1       (9.4 436)  0.150}
+\DocumentationTok{\#\# 3       79 tiam     47.7  1        (40 381) {-}0.161}
+\DocumentationTok{\#\# 4       80 caca      5.15 1      (38.7 422)  0.253}
+\DocumentationTok{\#\# 5       96 libe      2.3  1        (60 310)  0.262}
+\DocumentationTok{\#\# \# ... with 6,291 more rows, and 2 more variables: comp \textless{}list\textgreater{},}
+\DocumentationTok{\#\# \#   growth\_hat \textless{}dbl\textgreater{}}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{model_rmse <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
-\NormalTok{model_rmse}
-\CommentTok{## [1] 0.128}
+\NormalTok{model\_rmse }\OtherTok{\textless{}{-}}\NormalTok{ focal\_vs\_comp\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{rmse}\NormalTok{(}\AttributeTok{truth =}\NormalTok{ growth, }\AttributeTok{estimate =}\NormalTok{ growth\_hat) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{pull}\NormalTok{(.estimate)}
+\NormalTok{model\_rmse}
+\DocumentationTok{\#\# [1] 0.128}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\CommentTok{# Plot posteriors for only a subset of species}
-\NormalTok{sp_to_plot <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"litu"}\NormalTok{, }\StringTok{"quru"}\NormalTok{, }\StringTok{"cagl"}\NormalTok{)}
+\CommentTok{\# Plot posteriors for only a subset of species}
+\NormalTok{sp\_to\_plot }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\StringTok{"litu"}\NormalTok{, }\StringTok{"quru"}\NormalTok{, }\StringTok{"cagl"}\NormalTok{)}
 
-\NormalTok{plot1 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_scbi, }\DataTypeTok{type =} \StringTok{"intercepts"}\NormalTok{, }\DataTypeTok{sp_to_plot =}\NormalTok{ sp_to_plot)}
-\NormalTok{plot2 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_scbi, }\DataTypeTok{type =} \StringTok{"dbh_slopes"}\NormalTok{, }\DataTypeTok{sp_to_plot =}\NormalTok{ sp_to_plot)}
-\NormalTok{plot3 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_scbi, }\DataTypeTok{type =} \StringTok{"competition"}\NormalTok{, }\DataTypeTok{sp_to_plot =}\NormalTok{ sp_to_plot)}
+\NormalTok{plot1 }\OtherTok{\textless{}{-}} \FunctionTok{autoplot}\NormalTok{(comp\_bayes\_lm\_scbi, }\AttributeTok{type =} \StringTok{"intercepts"}\NormalTok{, }\AttributeTok{sp\_to\_plot =}\NormalTok{ sp\_to\_plot)}
+\NormalTok{plot2 }\OtherTok{\textless{}{-}} \FunctionTok{autoplot}\NormalTok{(comp\_bayes\_lm\_scbi, }\AttributeTok{type =} \StringTok{"dbh\_slopes"}\NormalTok{, }\AttributeTok{sp\_to\_plot =}\NormalTok{ sp\_to\_plot)}
+\NormalTok{plot3 }\OtherTok{\textless{}{-}} \FunctionTok{autoplot}\NormalTok{(comp\_bayes\_lm\_scbi, }\AttributeTok{type =} \StringTok{"competition"}\NormalTok{, }\AttributeTok{sp\_to\_plot =}\NormalTok{ sp\_to\_plot)}
 
-\CommentTok{# Combine plots using patchwork}
-\NormalTok{(plot1 }\OperatorTok{|}\StringTok{ }\NormalTok{plot2) }\OperatorTok{/}\StringTok{ }\NormalTok{plot3}
+\CommentTok{\# Combine plots using patchwork}
+\NormalTok{(plot1 }\SpecialCharTok{|}\NormalTok{ plot2) }\SpecialCharTok{/}\NormalTok{ plot3}
 \end{Highlighting}
@@
 These plots give the posterior distributions of parameters from Equation
-\ref{eq:model}. For many pacakge users they will be of interest because
+\ref{eq:model}. For many package users they will be of interest because
 they give insight into the species-specific competitive interactions.
@@
 structure of our competition model. To run the permutation test, we use
-the same\texttt{comp\_bayes\_lm()} function as in Section
+the same \texttt{comp\_bayes\_lm()} function as in Section
 \ref{model-fit-predict}, but with a \texttt{run\_shuffle\ =\ TRUE}
@@
 \begin{Highlighting}[]
-\NormalTok{comp_bayes_lm_scbi_shuffle <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{comp_bayes_lm}\NormalTok{(}\DataTypeTok{prior_param =} \OtherTok{NULL}\NormalTok{, }\DataTypeTok{run_shuffle =} \OtherTok{TRUE}\NormalTok{)}
+\NormalTok{comp\_bayes\_lm\_scbi\_shuffle }\OtherTok{\textless{}{-}}\NormalTok{ focal\_vs\_comp\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{comp\_bayes\_lm}\NormalTok{(}\AttributeTok{prior\_param =} \ConstantTok{NULL}\NormalTok{, }\AttributeTok{run\_shuffle =} \ConstantTok{TRUE}\NormalTok{)}
 
-\NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{growth_hat_shuffle =} \KeywordTok{predict}\NormalTok{(comp_bayes_lm_scbi_shuffle, }\DataTypeTok{newdata =}\NormalTok{ focal_vs_comp_scbi))}
+\NormalTok{focal\_vs\_comp\_scbi }\OtherTok{\textless{}{-}}\NormalTok{ focal\_vs\_comp\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{growth\_hat\_shuffle =} \FunctionTok{predict}\NormalTok{(comp\_bayes\_lm\_scbi\_shuffle, }\AttributeTok{newdata =}\NormalTok{ focal\_vs\_comp\_scbi))}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{model_rmse_shuffle <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat_shuffle) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
-\NormalTok{model_rmse_shuffle}
-\CommentTok{## [1] 0.131}
+\NormalTok{model\_rmse\_shuffle }\OtherTok{\textless{}{-}}\NormalTok{ focal\_vs\_comp\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{rmse}\NormalTok{(}\AttributeTok{truth =}\NormalTok{ growth, }\AttributeTok{estimate =}\NormalTok{ growth\_hat\_shuffle) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{pull}\NormalTok{(.estimate)}
+\NormalTok{model\_rmse\_shuffle}
+\DocumentationTok{\#\# [1] 0.131}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{cv_grid =}\NormalTok{ blocks_scbi)}
+\NormalTok{focal\_vs\_comp\_scbi }\OtherTok{\textless{}{-}}\NormalTok{ focal\_vs\_comp\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{run\_cv}\NormalTok{(}\AttributeTok{comp\_dist =}\NormalTok{ comp\_dist, }\AttributeTok{cv\_grid =}\NormalTok{ blocks\_scbi)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{model_rmse_cv <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
-\NormalTok{model_rmse_cv}
-\CommentTok{## [1] 0.14}
+\NormalTok{model\_rmse\_cv }\OtherTok{\textless{}{-}}\NormalTok{ focal\_vs\_comp\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{rmse}\NormalTok{(}\AttributeTok{truth =}\NormalTok{ growth, }\AttributeTok{estimate =}\NormalTok{ growth\_hat) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{pull}\NormalTok{(.estimate)}
+\NormalTok{model\_rmse\_cv}
+\DocumentationTok{\#\# [1] 0.14}
 \end{Highlighting}

### 2021-03-03 paper/forestecology.Rmd
Done most of introduction

@@
 
-Repeat-censused forest plots offer excellent data to test neighborhood models of tree competition @allen_permutation_2020 @canham_neighborhood_2006 @uriarte_spatially_2004.
-Here we describe an R package, `forestecology`, to do that.
-This package implements the methods in @allen_permutation_2020.
-It provides: a convenient way to specify and fit models of tree growth based on neighborhood competition; a spatial cross validation method to test and compare model fits @roberts_cross-validation_2017; and an ANOVA-like method to assess whether the competitor identity matters in these models.
-The model is written to work with ForestGEO plot data @andersonteixeira_ctfs-forestgeo_2015, but we envision that it could easily be modified to work with data from other forest plots, e.g.
-the US Forest Service Forest Inventory and Analysis plots @smith_forest_2002.
-Or more generally to model interactions of any community of mapped sessile organisms.
+BERT: Dave could you write a few sentences (and add references where appropriate) starting with the big picture of modeling tree growth, zooming in to neighhorhood models, and then connecting to the "@allen_permutation_2020 considers the following model" I write below?
 
-The `forestecology` is designed with "tidy" data principles in mind as @wickham_welcome_2019.
+Repeat-censused forest plots offer excellent data to test neighborhood models of the effect of competition on the growth of trees @canham_neighborhood_2006 @uriarte_spatially_2004.
 
-Given that our data is of geo-spatial nature, we represent our data using the "simple features" `sf` package class of objects @pebesma_simple_2018 whereby.
-While previously the `sp` package serves such purposes @pebesma_sp_2005, the `sf` package is designed to interface with the `tidyverse` suite of packages.
-
-## Model for growth of tree {#competition-model}
-
-While there are a litany of models one can consider, in @allen_permutation_2020 we considered a simple one.
-
-We fit the following linear model to the DBH of each focal tree.
-Let $i = 1, \ldots, n_j$ index all $n_j$ trees of "focal" species group $j$; let $j = 1, \ldots, J$ index all $J$ focal species groups; and let $k = 1, \ldots, K$ index all $K$ "competitor" species groups.
-We modeled the growth in diameter per year $y_{ij}$ (in centimeters per year) of the $i^{th}$ tree of focal species group $j$ as a linear model $f$ of the following covariates $\vec{x}_{ij}$
+@allen_permutation_2020 considers the following model: Let $i = 1, \ldots, n_j$ index all $n_j$ trees of "focal" species group $j$; let $j = 1, \ldots, J$ index all $J$ focal species groups; and let $k = 1, \ldots, K$ index all $K$ "competitor" species groups.
+We model the average annual growth in diameter $y_{ij}$ (in centimeters per year) of the $i^{th}$ tree of focal species group $j$ as a linear model $f$ of the covariates $\vec{x}_{ij}$
 
@@
 
-<!--where $\beta_{0,j}$ is the diameter-independent growth rate for group $j$; $\dbh_{ij}$ is the diameter at breast height (in centimeters) of the focal tree at the earlier census; $\beta_{\dbh,j}$ is the amount of the growth rate changed depending on diameter for group $j$;  $\BA_{ijk}$ is the sum of the basal area of all trees of competitor species group $k$; $\lambda_{jk}$ is the change in growth for individuals of group $j$ from nearby competitors of group $k$; and $\epsilon_{ij}$ is a random error term distributed $\text{Normal}(0, \sigma^2)$.-->
+where $\beta_{0,j}$ is the diameter-independent growth rate for group $j$; $\text{dbh}_{ij}$ is the diameter at breast height (in centimeters) of the focal tree at the earlier census; $\beta_{\text{dbh},j}$ is the amount of the growth rate changed depending on diameter for group $j$;  $\text{BA}_{ijk}$ is the sum of the basal area of all trees of competitor species group $k$; $\lambda_{jk}$ is the change in growth for individuals of group $j$ from nearby competitors of group $k$; and $\epsilon_{ij}$ is a random error term distributed $\text{Normal}(0, \sigma^2)$.
+They estimate all parameters via Bayesian linear regression while exploiting Normal/Inverse Gamma conjugacy to derive closed-form solutions to all posterior distributions via linear algebra^[See S1 Appendix of @allen_permutation_2020, available at https://doi.org/10.1371/journal.pone.0229930.s004].
+These closed-form solutions for the posterior distributions are in contrast to approximations of all posteriors via computationally expensive Markov Chain Monte Carlo algorithms.
 
-Link to https://doi.org/10.1371/journal.pone.0229930.s004
+In order to evaluate whether competitor species identity matters, @allen_permutation_2020 run a permutation test where under the null hypothesis the species identity of all competitors of a focal tree can be permuted/shuffled:
 
-For this linear model's case, there exists a closed form solution as described here.
-As such, the `fit_bayesian_model()` function using matrix algebra to obtain all parameter estimates, rather than computationally expensive Monte Carlo approximations.
-The inputs to this function are a `focal_vs_comp` data frame, `prior_param` a list of priors, and a boolean flag `run_shuffle` on whether or not to run competitor-species identity permutations which we will demonstrate below on the Michigan Big Woods data.
-This function returns the posterior means of all parameters.
+\begin{eqnarray}
+\label{eq:permutation-hypothesis-test}
+&&H_0: \lambda_{jk} = \lambda_{j} \mbox{ for all } k = 1, \ldots, K\\
+\text{vs.}&&H_A: \text{at least one } \lambda_{jk} \mbox{ is different}
+\end{eqnarray}
 
-# Example
+where the null hypothesis $H_0$ reflects a hypothesis of no species grouping-specific effects of competition while the alternative hypothesis $H_A$ reflects a hypothesis of species grouping-specific effects of competition.
+Furthermore, in order to account for the spatial autocorrelation inherent to forest data in their estimates of out-of-sample model error, @allen_permutation_2020 use spatial cross-validation.
+Estimates of model error that do not account for this spatial dependency tend to underestiamte the true model error [@roberts_cross-validation_2017].
 
-We demonstrate the `forestecology` package's functionality on data from the Smithsonian Conservation Biology Institute (SCBI) large forest dynamics plot, located at the Smithsonian's National Zoo and Conservation Biology Institute in Front Royal, VA, USA @bourg_initial_2013.
+We introduce the `forestecology` R package providing methods and data for forest ecology model fitting and assessment, available on CRAN (https://cran.r-project.org/web/packages/forestecology/index.html) with the corresponding source code available on GitHub (https://github.com/rudeboybert/forestecology).
+The package implements all aspects of the model in Equation \ref{eq:model}: model fitting and generating fitted/predicted values, evaluating the effect of competitor species identity using permutation tests, and evaluating model performance using spatial cross-validation.
+
+The package designed with "tidy" design principles in mind [@wickham_welcome_2019].
+Much like many of the `tidyverse` component packages, `forestecology` is designed with verb-named functions that can be modularly composed in sequence using the pipe `%>%` operator [@bache_pipe_2020].
+As we articulate in Section \ref{casestudy}, these functions delineate the key steps in our analysis sequence.
+Furthermore, the inputs and outputs of nearly all of our functions use the same "simple features for R" data structures as implemented in the `sf` package for standardized support for spatial vector data [@pebesma_simple_2018].
+The `sf` package is a `tidyverse`-friendly evolution of the `sp` package of classes and methods for spatial data in R [@pebesma_sp_2005].
+As such, wrangling and visualization spatial data such as ours becomes much easier.
+
+Currently the package only implements the Bayesian linear regression model of tree growth based on neighborhood competition detailed in Equation \ref{eq:model}.
+As we demonstrate in Section \ref{model-fit-predict} however, the fitting of this model is self-contained in a single function `comp_bayes_lm()`.
+This function returns an object of S3 class type `comp_bayes_lm` with generic methods implemented for `print()` to inspect the output, `predict()` to generate fitted/predicted values, and `ggplot2::autoplot()` to visualize all results.
+Therefore the package can be modularly extended to fit other models as long as they are coded into a function similar type as `comp_bayes_lm()` as has equivalent generic methods implemented.
+
+We present a case-study of the `forestecology` package's use on data from the Smithsonian Conservation Biology Institute's (SCBI) large forest dynamics plot in Front Royal, Virginia, USA in Section \ref{casestudy}, which is part of the ForestGEO global network of research sites [@bourg_initial_2013, @andersonteixeira_ctfs-forestgeo_2015].
+The package is designed with ForestGEO plot data in mind, but we envision that it could easily be modified to work with data from other forest plots, e.g.
+the US Forest Service Forest Inventory and Analysis plots or more generally to model interactions of any community of mapped sessile organisms [@smith_forest_2002].
+
+# forestecology workflow: a case study {#casestudy}
+
+We demonstrate the `forestecology` package's functionality on data from the Smithsonian Conservation Biology Institute (SCBI) large forest dynamics plot, located at the Smithsonian's National Zoo and Conservation Biology Institute in Front Royal, VA, USA [@bourg_initial_2013].
 The 25.6 ha (640 x 400 m) plot is located at the intersection of three of the major physiographic provinces of the eastern US---the Blue Ridge, Ridge and Valley, and Piedmont provinces---and is adjacent to the northern end of Shenandoah National Park.
-The forest type is typical mature secondary eastern mixed deciduous forest, with a canopy dominated by tulip poplar (*Liriodendron tulipifera*), oaks (*Quercus* spp.), and hickories (*Carya* spp.), and an understory composed mainly of spicebush (*Lindera benzoin*), paw-paw (*Asimina triloba*), American hornbeam (*Carpinus caroliniana*), and witch hazel (*Hamamelis virginiana*) @bourg_initial_2013.
+The forest type is typical mature secondary eastern mixed deciduous forest, with a canopy dominated by tulip poplar (*Liriodendron tulipifera*), oaks (*Quercus* spp.), and hickories (*Carya* spp.), and an understory composed mainly of spicebush (*Lindera benzoin*), paw-paw (*Asimina triloba*), American hornbeam (*Carpinus caroliniana*), and witch hazel (*Hamamelis virginiana*) [@bourg_initial_2013].
 
@@
 Despite such standards, minor variations will still exist between sites, thereby necessitating some data wrangling and checking.
-For example, the SCBI site records all diameters at breast height (DBH) in millimeters @bourg_initial_2013, whereas the Michigan Big Woods site records them in centimeters @allen_michigan_2020.
+For example, the SCBI site records all diameters at breast height (DBH) in millimeters [@bourg_initial_2013], whereas the Michigan Big Woods site records them in centimeters [@allen_michigan_2020].
 
@@
 In order to account for such edge effects, only trees that are not part of this buffer region, i.e.
-are part of the interior of the study region, will have their growth modeled @waller_applied_2004.
+are part of the interior of the study region, will have their growth modeled [@waller_applied_2004].
 
 Our model of interspecific competition relies on a spatial definition of who the competitor trees are for focal trees of interest: all trees within a distance `comp_dist` of a focal tree are considered its competitors (assuming the same units as the `gx` and `gy` location variables).
-In our case we set this value at 7.5m, a value informed by @canham_neighborhood_2004 @uriarte_spatially_2004 @canham_neighborhood_2006.
+In our case we set this value at 7.5m, a value informed by other studies [@canham_neighborhood_2004, @uriarte_spatially_2004, @canham_neighborhood_2006].
 Using this value along with a manually constructed `sf` object representation of the study region's boundary via its vertices, we apply the `add_buffer_variable()` to our `growth_scbi` data frame to add a `buffer` boolean variable: all trees who have `buffer` set to `FALSE` will be our focal trees whose growth will be modeled, whereas those with `TRUE` will only be considered as competitor trees whose growth will not.
@@
 In the case of forest census data, observations exhibit spatial autocorrelation.
-We therefore incorporate this spatial dependence into the cross-validation algorithm with our spatial blocks of trees.
-@roberts_cross-validation_2017 @pohjankukka_estimating_2017.
-
+We therefore incorporate this spatial dependence into the cross-validation algorithm with our spatial blocks of trees [@roberts_cross-validation_2017 @pohjankukka_estimating_2017]
 In the example below, we first manually define four folds that partition the study region as an `sf` object.
-We then use the output of the `spatialBlock()` function from the `blockCV` package to associate each tree in `growth_scbi` to the correct fold (saved in the `foldID` variable) @valavi_blockcv_2019.
+We then use the output of the `spatialBlock()` function from the `blockCV` package to associate each tree in `growth_scbi` to the correct fold (saved in the `foldID` variable) [@valavi_blockcv_2019].
 ^[In the Appendix we present an example where the folds themselves are also created using the `spatialBlock()` function given a specified `cv_block_size`.]
@@
 Two new variables `focal_ID` and `focal_sp` relate to tree-stem identification and species information.
-Most notably however is a new variable `comp` which contains information on all competitor trees for a given focal tree, saved in `tidyr` package list-column format @tidyr_package.
+Most notably however is a new variable `comp` which contains information on all competitor trees for a given focal tree, saved in `tidyr` package list-column format [@tidyr_package].
 For example, we drill-down on the tree with `focal_ID` 4, which has 20 competitor trees each described by 4 variables as indicated by the fact that `comp` is a `tibble [20 × 4]`.
@@
 
-Sophie Li for her feedback on package interface.
+The authors thank Ryan Giordano and Jonathan Che for their help with the statistical methodology and Sophie Li for their feedback on package interface.
 

### 2021-03-03 paper/forestecology.tex
Done most of introduction

--- 2021-02-25 paper/forestecology.tex
@@
 
+BERT: Dave could you write a few sentences (and add references where
+appropriate) starting with the big picture of modeling tree growth,
+zooming in to neighhorhood models, and then connecting to the
+``\citet{allen_permutation_2020} considers the following model'' I write
+below?
+
 Repeat-censused forest plots offer excellent data to test neighborhood
-models of tree competition \citet{allen_permutation_2020}
-\citet{canham_neighborhood_2006} \citet{uriarte_spatially_2004}. Here we
-describe an R package, \texttt{forestecology}, to do that. This package
-implements the methods in \citet{allen_permutation_2020}. It provides: a
-convenient way to specify and fit models of tree growth based on
-neighborhood competition; a spatial cross validation method to test and
-compare model fits \citet{roberts_cross-validation_2017}; and an
-ANOVA-like method to assess whether the competitor identity matters in
-these models. The model is written to work with ForestGEO plot data
-\citet{andersonteixeira_ctfs-forestgeo_2015}, but we envision that it
-could easily be modified to work with data from other forest plots,
-e.g.~the US Forest Service Forest Inventory and Analysis plots
-\citet{smith_forest_2002}. Or more generally to model interactions of
-any community of mapped sessile organisms.
+models of the effect of competition on the growth of trees
+\citet{canham_neighborhood_2006} \citet{uriarte_spatially_2004}.
 
-The \texttt{forestecology} is designed with ``tidy'' data principles in
-mind as \citet{wickham_welcome_2019}.
-
-Given that our data is of geo-spatial nature, we represent our data
-using the ``simple features'' \texttt{sf} package class of objects
-\citet{pebesma_simple_2018} whereby. While previously the \texttt{sp}
-package serves such purposes \citet{pebesma_sp_2005}, the \texttt{sf}
-package is designed to interface with the \texttt{tidyverse} suite of
-packages.
-
-\hypertarget{competition-model}{%
-\subsection{Model for growth of tree}\label{competition-model}}
-
-While there are a litany of models one can consider, in
-\citet{allen_permutation_2020} we considered a simple one.
-
-We fit the following linear model to the DBH of each focal tree. Let
+\citet{allen_permutation_2020} considers the following model: Let
 \(i = 1, \ldots, n_j\) index all \(n_j\) trees of ``focal'' species
@@
 groups; and let \(k = 1, \ldots, K\) index all \(K\) ``competitor''
-species groups. We modeled the growth in diameter per year \(y_{ij}\)
-(in centimeters per year) of the \(i^{th}\) tree of focal species group
-\(j\) as a linear model \(f\) of the following covariates
+species groups. We model the average annual growth in diameter
+\(y_{ij}\) (in centimeters per year) of the \(i^{th}\) tree of focal
+species group \(j\) as a linear model \(f\) of the covariates
 \(\vec{x}_{ij}\)
@@
 
-Link to \url{https://doi.org/10.1371/journal.pone.0229930.s004}
+where \(\beta_{0,j}\) is the diameter-independent growth rate for group
+\(j\); \(\text{dbh}_{ij}\) is the diameter at breast height (in
+centimeters) of the focal tree at the earlier census;
+\(\beta_{\text{dbh},j}\) is the amount of the growth rate changed
+depending on diameter for group \(j\); \(\text{BA}_{ijk}\) is the sum of
+the basal area of all trees of competitor species group \(k\);
+\(\lambda_{jk}\) is the change in growth for individuals of group \(j\)
+from nearby competitors of group \(k\); and \(\epsilon_{ij}\) is a
+random error term distributed \(\text{Normal}(0, \sigma^2)\). They
+estimate all parameters via Bayesian linear regression while exploiting
+Normal/Inverse Gamma conjugacy to derive closed-form solutions to all
+posterior distributions via linear algebra\footnote{See S1 Appendix of
+  \citet{allen_permutation_2020}, available at
+  \url{https://doi.org/10.1371/journal.pone.0229930.s004}}. These
+closed-form solutions for the posterior distributions are in contrast to
+approximations of all posteriors via computationally expensive Markov
+Chain Monte Carlo algorithms.
 
-For this linear model's case, there exists a closed form solution as
-described here. As such, the \texttt{fit\_bayesian\_model()} function
-using matrix algebra to obtain all parameter estimates, rather than
-computationally expensive Monte Carlo approximations. The inputs to this
-function are a \texttt{focal\_vs\_comp} data frame,
-\texttt{prior\_param} a list of priors, and a boolean flag
-\texttt{run\_shuffle} on whether or not to run competitor-species
-identity permutations which we will demonstrate below on the Michigan
-Big Woods data. This function returns the posterior means of all
-parameters.
+In order to evaluate whether competitor species identity matters,
+\citet{allen_permutation_2020} run a permutation test where under the
+null hypothesis the species identity of all competitors of a focal tree
+can be permuted/shuffled:
 
-\hypertarget{example}{%
-\section{Example}\label{example}}
+\begin{eqnarray}
+\label{eq:permutation-hypothesis-test}
+&&H_0: \lambda_{jk} = \lambda_{j} \mbox{ for all } k = 1, \ldots, K\\
+\text{vs.}&&H_A: \text{at least one } \lambda_{jk} \mbox{ is different}
+\end{eqnarray}
+
+where the null hypothesis \(H_0\) reflects a hypothesis of no species
+grouping-specific effects of competition while the alternative
+hypothesis \(H_A\) reflects a hypothesis of species grouping-specific
+effects of competition. Furthermore, in order to account for the spatial
+autocorrelation inherent to forest data in their estimates of
+out-of-sample model error, \citet{allen_permutation_2020} use spatial
+cross-validation. Estimates of model error that do not account for this
+spatial dependency tend to underestiamte the true model error
+\citep{roberts_cross-validation_2017}.
+
+We introduce the \texttt{forestecology} R package providing methods and
+data for forest ecology model fitting and assessment, available on CRAN
+(\url{https://cran.r-project.org/web/packages/forestecology/index.html})
+with the corresponding source code available on GitHub
+(\url{https://github.com/rudeboybert/forestecology}). The package
+implements all aspects of the model in Equation \ref{eq:model}: model
+fitting and generating fitted/predicted values, evaluating the effect of
+competitor species identity using permutation tests, and evaluating
+model performance using spatial cross-validation.
+
+The package designed with ``tidy'' design principles in mind
+\citep{wickham_welcome_2019}. Much like many of the \texttt{tidyverse}
+component packages, \texttt{forestecology} is designed with verb-named
+functions that can be modularly composed in sequence using the pipe
+\texttt{\%\textgreater{}\%} operator \citep{bache_pipe_2020}. As we
+articulate in Section \ref{casestudy}, these functions delineate the key
+steps in our analysis sequence. Furthermore, the inputs and outputs of
+nearly all of our functions use the same ``simple features for R'' data
+structures as implemented in the \texttt{sf} package for standardized
+support for spatial vector data \citep{pebesma_simple_2018}. The
+\texttt{sf} package is a \texttt{tidyverse}-friendly evolution of the
+\texttt{sp} package of classes and methods for spatial data in R
+\citep{pebesma_sp_2005}. As such, wrangling and visualization spatial
+data such as ours becomes much easier.
+
+Currently the package only implements the Bayesian linear regression
+model of tree growth based on neighborhood competition detailed in
+Equation \ref{eq:model}. As we demonstrate in Section
+\ref{model-fit-predict} however, the fitting of this model is
+self-contained in a single function \texttt{comp\_bayes\_lm()}. This
+function returns an object of S3 class type \texttt{comp\_bayes\_lm}
+with generic methods implemented for \texttt{print()} to inspect the
+output, \texttt{predict()} to generate fitted/predicted values, and
+\texttt{ggplot2::autoplot()} to visualize all results. Therefore the
+package can be modularly extended to fit other models as long as they
+are coded into a function similar type as \texttt{comp\_bayes\_lm()} as
+has equivalent generic methods implemented.
+
+We present a case-study of the \texttt{forestecology} package's use on
+data from the Smithsonian Conservation Biology Institute's (SCBI) large
+forest dynamics plot in Front Royal, Virginia, USA in Section
+\ref{casestudy}, which is part of the ForestGEO global network of
+research sites
+\citep[\citet{andersonteixeira_ctfs-forestgeo_2015}]{bourg_initial_2013}.
+The package is designed with ForestGEO plot data in mind, but we
+envision that it could easily be modified to work with data from other
+forest plots, e.g.~the US Forest Service Forest Inventory and Analysis
+plots or more generally to model interactions of any community of mapped
+sessile organisms \citep{smith_forest_2002}.
+
+\hypertarget{casestudy}{%
+\section{forestecology workflow: a case study}\label{casestudy}}
 
@@
 Conservation Biology Institute in Front Royal, VA, USA
-\citet{bourg_initial_2013}. The 25.6 ha (640 x 400 m) plot is located at
+\citep{bourg_initial_2013}. The 25.6 ha (640 x 400 m) plot is located at
 the intersection of three of the major physiographic provinces of the
-eastern US: the Blue Ridge, Ridge and Valley, and Piedmont provinces and
-is adjacent to the northern end of Shenandoah National Park. The forest
-type is typical mature secondary eastern mixed deciduous forest, with a
-canopy dominated by tulip poplar (\emph{Liriodendron tulipifera}), oaks
-(\emph{Quercus} spp.), and hickories (\emph{Carya} spp.), and an
-understory composed mainly of spicebush (\emph{Lindera benzoin}),
-paw-paw (\emph{Asimina triloba}), American hornbeam (\emph{Carpinus
-caroliniana}), and witch hazel (\emph{Hamamelis virginiana})
-\citet{bourg_initial_2013}.
+eastern US---the Blue Ridge, Ridge and Valley, and Piedmont
+provinces---and is adjacent to the northern end of Shenandoah National
+Park. The forest type is typical mature secondary eastern mixed
+deciduous forest, with a canopy dominated by tulip poplar
+(\emph{Liriodendron tulipifera}), oaks (\emph{Quercus} spp.), and
+hickories (\emph{Carya} spp.), and an understory composed mainly of
+spicebush (\emph{Lindera benzoin}), paw-paw (\emph{Asimina triloba}),
+American hornbeam (\emph{Carpinus caroliniana}), and witch hazel
+(\emph{Hamamelis virginiana}) \citep{bourg_initial_2013}.
 
@@
 diameters at breast height (DBH) in millimeters
-\citet{bourg_initial_2013}, whereas the Michigan Big Woods site records
-them in centimeters \citet{allen_michigan_2020}.
+\citep{bourg_initial_2013}, whereas the Michigan Big Woods site records
+them in centimeters \citep{allen_michigan_2020}.
 
@@
 later censuses, such as \texttt{dbh1/dbh2} and \texttt{codes1/codes2}.
-The data format of other sites may be such that our
-\texttt{compute\_growth()} function doesn't work at all. However, in the
-end all that matters is that the growth of all stems is saved in a data
-frame of class \texttt{sf} and at a minimum the data contains the
-variables above. TODO: enumerate what these variables are and say if
-they need exact names
+Site data that does not align with this convention will need to be
+transformed for use with the \texttt{compute\_growth()} function.
+However, in the end, all that matters is that the growth of all stems is
+saved in a data frame of class \texttt{sf} and, at a minimum, the data
+contains the variables above. TODO: enumerate what these variables are
+and say if they need exact names
 
@@
 study region, will have their growth modeled
-\citet{waller_applied_2004}.
+\citep{waller_applied_2004}.
 
@@
 location variables). In our case we set this value at 7.5m, a value
-informed by \citet{canham_neighborhood_2004}
-\citet{uriarte_spatially_2004} \citet{canham_neighborhood_2006}. Using
-this value along with a manually constructed \texttt{sf} object
-representation of the study region's boundary, we apply the
-\texttt{add\_buffer\_variable()} to our \texttt{growth\_scbi} data frame
-to add a \texttt{buffer} boolean variable: all trees who have
+informed by other studies \citep[\citet{uriarte_spatially_2004},
+\citet{canham_neighborhood_2006}]{canham_neighborhood_2004}. Using this
+value along with a manually constructed \texttt{sf} object
+representation of the study region's boundary via its vertices, we apply
+the \texttt{add\_buffer\_variable()} to our \texttt{growth\_scbi} data
+frame to add a \texttt{buffer} boolean variable: all trees who have
 \texttt{buffer} set to \texttt{FALSE} will be our focal trees whose
@@
 assign observations to folds by randomly resampling individual
-observations. However, an assumption that the observations are
-independent underlies any cross-validation algorithm. In the case of
-forest census data, observations exhibit spatial autocorrelation. We
-therefore incorporate this spatial dependence into the cross-validation
-algorithm with our spatial blocks of trees.
-\citet{roberts_cross-validation_2017}
-\citet{pohjankukka_estimating_2017}.
-
-In the example below, we first manually define four folds that partition
+observations. However, many of these algorithms assume that the
+observations are independent of each other. In the case of forest census
+data, observations exhibit spatial autocorrelation. We therefore
+incorporate this spatial dependence into the cross-validation algorithm
+with our spatial blocks of trees \citep[
+\citet{pohjankukka_estimating_2017}]{roberts_cross-validation_2017} In
+the example below, we first manually define four folds that partition
 the study region as an \texttt{sf} object. We then use the output of the
@@
 associate each tree in \texttt{growth\_scbi} to the correct fold (saved
-in the \texttt{foldID} variable) \citet{valavi_blockcv_2019}. \footnote{In
+in the \texttt{foldID} variable) \citep{valavi_blockcv_2019}. \footnote{In
   the Appendix we present an example where the folds themselves are also
@@
 tree-stem, this function also requires an \texttt{sf} object
-representation of the spatial cross-validation blocks/folds; in our
-case, this was manually encoded in the \texttt{blocks\_scbi} in Section
-\ref{spatial-information} while in our Appendix we present an example
-where this was performed using \texttt{spatialBlock()} from the
-\texttt{blockCV} package. We present the resulting data frame below with
-the \texttt{foldID} variable omitted for compactness of presentation.
+representation of the spatial cross-validation blocks/folds. In this
+example, the blocks were manually encoded in \texttt{blocks\_scbi} by
+specifying it's vertices in Section \ref{spatial-information}\footnote{We
+  present an alternative method for defining spatial cross-validation
+  blocks is using the \texttt{spatialBlock()} function from the
+  \texttt{blockCV} package in the Appendix}. We present the resulting
+data frame below with the \texttt{foldID} variable omitted for
+compactness of presentation.
 
@@
 saved in \texttt{tidyr} package list-column format
-\citet{tidyr_package}. For example, we drill-down on the tree with
+\citep{tidyr_package}. For example, we drill-down on the tree with
 \texttt{focal\_ID} 4, which has 20 competitor trees each described by 4
 variables as indicated by the fact that \texttt{comp} is a
-\texttt{\textless{}tibble\ {[}20\ ×\ 4{]}\textgreater{}}.
+\texttt{tibble\ {[}20\ ×\ 4{]}}.
 
@@
 implemented in a function that similarly takes an input data frame of
-type \texttt{focal\_vs\_comp} as an argument can be easily swapped in.
-For our specific competition Bayesian linear regression model, we also
-specify prior distributions on all parameters of interest (here chosen
-to be the defaults as specified in \texttt{?comp\_bayes\_lm}).
+type \texttt{focal\_vs\_comp} as an argument can also be used. For our
+specific competition Bayesian linear regression model, we also specify
+prior distributions on all parameters of interest (here chosen to be the
+defaults as specified in \texttt{?comp\_bayes\_lm}).
 
@@
 
-Sophie Li for her feedback on package interface.
+The authors thank Ryan Giordano and Jonathan Che for their help with the
+statistical methodology and Sophie Li for their feedback on package
+interface.
 

### 2021-03-03 paper/forestecology.Rmd
renamed all cv_grid_sf object names to blocks

@@
 focal_vs_comp_scbi <- growth_scbi %>%
-  create_focal_vs_comp(comp_dist, cv_grid_sf = blocks_scbi, id = "stemID")
+  create_focal_vs_comp(comp_dist, blocks = blocks_scbi, id = "stemID")
 focal_vs_comp_scbi %>% 

### 2021-03-03 R/data_processing_functions.R

@@ -89,7 +89,7 @@ compute_growth <- function(census_1, census_2, id) {
 #' Create focal versus competitor trees data frame
 #'
 #' @param growth_df A [compute_growth()] output converted to \code{sf} object
-#' @param cv_grid_sf An sf object of a \code{blockCV} block output
+#' @param blocks An sf object of a \code{blockCV} block output
 #' @param comp_dist Distance to determine which neighboring trees to a focal tree are competitors.
 #' @param id A character string of the variable name in \code{growth_df} uniquely identifying each tree
 #'
@@ -155,21 +155,21 @@ compute_growth <- function(census_1, census_2, id) {
 #'
 #' # Return corresponding data frame
 #' growth_toy %>%
-#'   create_focal_vs_comp(comp_dist = 1.5, cv_grid_sf = SpatialBlock_ex, id = "ID")
+#'   create_focal_vs_comp(comp_dist = 1.5, blocks = SpatialBlock_ex, id = "ID")
 #'
 #' # Load in growth_df with spatial data
 #' # See ?growth_ex for attaching spatial data to growth_df
 #' data(growth_spatial_ex)
 #' # Load in cv_grid
-#' data(cv_grid_sf_ex)
+#' data(blocks_ex)
 #'
 #' focal_vs_comp_ex <- growth_spatial_ex %>%
-#'   create_focal_vs_comp(comp_dist = 1, cv_grid_sf = cv_grid_sf_ex, id = "ID")
-create_focal_vs_comp <- function(growth_df, comp_dist, cv_grid_sf, id) {
+#'   create_focal_vs_comp(comp_dist = 1, blocks = blocks_ex, id = "ID")
+create_focal_vs_comp <- function(growth_df, comp_dist, blocks, id) {
   # 0. Check inputs
   check_inherits(growth_df, "data.frame")
   check_inherits(comp_dist, "numeric")
-  check_inherits(cv_grid_sf, "sf")
+  check_inherits(blocks, "sf")
   check_inherits(id, "character")
 
   if (!id %in% colnames(growth_df)) {
@@ -216,7 +216,7 @@ create_focal_vs_comp <- function(growth_df, comp_dist, cv_grid_sf, id) {
 
   for (i in 1:length(all_folds)) {
     # Identify this fold's boundary
-    fold_boundary <- cv_grid_sf %>%
+    fold_boundary <- blocks %>%
       filter(folds == all_folds[i])
 
     # Identify focal trees in this fold

### 2021-03-03 R/example_datasets.R

@@ -214,7 +214,7 @@
 #'
 #' # Create the focal versus comp data frame
 #' focal_vs_comp_ex <- growth_spatial_ex %>%
-#'   create_focal_vs_comp(comp_dist, cv_grid_sf = cv_grid_sf_ex, id = "ID")
+#'   create_focal_vs_comp(comp_dist, blocks = blocks_ex, id = "ID")
 "growth_spatial_ex"
 
 
@@ -242,13 +242,13 @@
 #'
 #' comp_dist <- 1
 #'
-#' ggplot(cv_grid_sf_ex) +
+#' ggplot(blocks_ex) +
 #'   geom_sf() +
 #'   geom_sf(data = growth_spatial_ex)
 #'
 #' focal_vs_comp_ex <- growth_spatial_ex %>%
-#'   create_focal_vs_comp(comp_dist, cv_grid_sf = cv_grid_sf_ex, id = "ID")
-"cv_grid_sf_ex"
+#'   create_focal_vs_comp(comp_dist, blocks = blocks_ex, id = "ID")
+"blocks_ex"
 
 
 

### 2021-03-03 R/model_functions.R

@@ -280,7 +280,7 @@ predict.comp_bayes_lm <- function(object, newdata, ...) {
 #' run_cv(
 #'   focal_vs_comp_ex,
 #'   comp_dist = 1,
-#'   cv_grid = cv_grid_sf_ex
+#'   cv_grid = blocks_ex
 #' )
 run_cv <- function(focal_vs_comp, comp_dist, cv_grid, prior_param = NULL, run_shuffle = FALSE) {
   check_focal_vs_comp(focal_vs_comp)

### 2021-03-03 data-raw/process_data_raw.R

@@ -151,14 +151,14 @@ use_data(growth_spatial_ex, overwrite = TRUE)
 
 
 ## Create spatial objects ----
-cv_grid_sf_ex <- SpatialBlock_ex$blocks %>%
+blocks_ex <- SpatialBlock_ex$blocks %>%
   st_as_sf()
-use_data(cv_grid_sf_ex, overwrite = TRUE)
+use_data(blocks_ex, overwrite = TRUE)
 
 
 ## Create focal_vs_comp data frame  ----
 focal_vs_comp_ex <- growth_spatial_ex %>%
-  create_focal_vs_comp(comp_dist = 1, cv_grid_sf = cv_grid_sf_ex, id = "ID")
+  create_focal_vs_comp(comp_dist = 1, blocks = blocks_ex, id = "ID")
 use_data(focal_vs_comp_ex, overwrite = TRUE)
 
 

### 2021-03-03 paper/simulation_results/replicate_plosone_fig1_bw.R

@@ -66,7 +66,7 @@ blocks_bw <- bw_spatialBlock$blocks %>%
 
 # Compute focal versus competitor tree information -----------------------------
 focal_vs_comp_bw <- growth_bw %>%
-  create_focal_vs_comp(comp_dist, cv_grid_sf = blocks_bw, id = "treeID")
+  create_focal_vs_comp(comp_dist, blocks = blocks_bw, id = "treeID")
 
 
 
@@ -95,7 +95,7 @@ for(i in 1:length(species_notion_vector)){
   # Focal vs comp main dataframe for analysis
   focal_vs_comp_bw <- growth_bw %>%
     mutate(sp = .data[[species_notion_vector[i]]] %>% factor()) %>%
-    create_focal_vs_comp(comp_dist, cv_grid_sf = blocks_bw, id = "treeID")
+    create_focal_vs_comp(comp_dist, blocks = blocks_bw, id = "treeID")
 
 
   # 1. Compute observed test statistic: RMSE with no cross-validation ----

### 2021-03-03 paper/simulation_results/replicate_plosone_fig1_scbi.R

@@ -76,7 +76,7 @@ growth_scbi <- growth_scbi %>%
 
 # Compute focal versus competitor tree information -----------------------------
 focal_vs_comp_scbi <- growth_scbi %>%
-  create_focal_vs_comp(comp_dist, cv_grid_sf = blocks_scbi, id = "stemID")
+  create_focal_vs_comp(comp_dist, blocks = blocks_scbi, id = "stemID")
 
 
 

### 2021-03-03 tests/testthat/test-readme.R

@@ -46,7 +46,7 @@ test_that("readme code works", {
     mutate(foldID = SpatialBlock_ex$foldID %>% factor())
 
   focal_vs_comp_ex <- growth_ex %>%
-    create_focal_vs_comp(comp_dist, cv_grid_sf = blocks_ex, id = "ID")
+    create_focal_vs_comp(comp_dist, blocks = blocks_ex, id = "ID")
 
   # Checks each column in focal_vs_comp is of appropriate type
   expect_true(check_inherits(focal_vs_comp_ex, "data.frame"))
@@ -73,7 +73,7 @@ test_that("readme code works", {
   expect_true(
     check_inherits(
       focal_vs_comp_ex %>%
-        run_cv(comp_dist = comp_dist, cv_grid = cv_grid_sf_ex) %>%
+        run_cv(comp_dist = comp_dist, cv_grid = blocks_ex) %>%
         right_join(growth_ex, by = c("focal_ID" = "ID")),
       "data.frame"
     )
@@ -89,7 +89,7 @@ test_that("readme code works", {
   )
 
   focal_vs_comp_ex <- focal_vs_comp_ex %>%
-    run_cv(comp_dist = comp_dist, cv_grid = cv_grid_sf_ex)
+    run_cv(comp_dist = comp_dist, cv_grid = blocks_ex)
 
   expect_equal(
     focal_vs_comp_ex %>%

### 2021-03-03 paper/forestecology.Rmd
convert all sp to factors internally

@@
 If the later, then millimeters.
-This is because ForestGEO protocols state that only trees with DBH greater or equal to 1cm should be included in censuses.], convert the `sp` variable containing species information from type `chr` character to `fct` factor.^[In our spatial cross-validation algorithm in Section \ref{spatial-cross-validation} issues can occur when rare species do not occur in the training set, but then are encountered in the test set.
-This risk is mitigated by representing `sp` as a factor variable, which has a complete list of all levels of the categorical variable.] Furthermore, in order to speed up computation for purposes of this example, we only consider a 9 ha subsection of the 25.6 ha of the SCBI site: `gx` from 0--300 instead of 0--400 and `gy` from 300--600 instead of 0--640.
+This is because ForestGEO protocols state that only trees with DBH greater or equal to 1cm should be included in censuses.] Furthermore, in order to speed up computation for purposes of this example, we only consider a 9 ha subsection of the 25.6 ha of the SCBI site: `gx` from 0--300 instead of 0--400 and `gy` from 300--600 instead of 0--640.
 
@@
     date = mdy(date),
-    dbh = as.numeric(dbh)/10,
-    sp = factor(sp)
+    dbh = as.numeric(dbh)/10
   ) %>%
@@
 However, in the end, all that matters is that the growth of all stems is saved in a data frame of class `sf` and, at a minimum, the data contains the variables above.
+Lastly, the species variable `sp` is automatically converted to a factor^[In our spatial cross-validation algorithm in Section \ref{spatial-cross-validation} issues can occur when rare species do not occur in the training set, but then are encountered in the test set.
+This risk is mitigated by representing `sp` as a factor variable, which has a complete list of all levels of the categorical variable.].
+
 TODO: enumerate what these variables are and say if they need exact names

### 2021-03-03 paper/forestecology.tex
convert all sp to factors internally

@@
   because ForestGEO protocols state that only trees with DBH greater or
-  equal to 1cm should be included in censuses.}, convert the \texttt{sp}
-variable containing species information from type \texttt{chr} character
-to \texttt{fct} factor.\footnote{In our spatial cross-validation
-  algorithm in Section \ref{spatial-cross-validation} issues can occur
-  when rare species do not occur in the training set, but then are
-  encountered in the test set. This risk is mitigated by representing
-  \texttt{sp} as a factor variable, which has a complete list of all
-  levels of the categorical variable.} Furthermore, in order to speed up
-computation for purposes of this example, we only consider a 9 ha
-subsection of the 25.6 ha of the SCBI site: \texttt{gx} from 0--300
+  equal to 1cm should be included in censuses.} Furthermore, in order to
+speed up computation for purposes of this example, we only consider a 9
+ha subsection of the 25.6 ha of the SCBI site: \texttt{gx} from 0--300
 instead of 0--400 and \texttt{gy} from 300--600 instead of 0--640.
@@
     \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date),}
-    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}\NormalTok{,}
-    \DataTypeTok{sp =} \KeywordTok{factor}\NormalTok{(sp)}
+    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}
 \NormalTok{  ) }\OperatorTok{%>%}
@@
 saved in a data frame of class \texttt{sf} and, at a minimum, the data
-contains the variables above. TODO: enumerate what these variables are
-and say if they need exact names
+contains the variables above. Lastly, the species variable \texttt{sp}
+is automatically converted to a factor\footnote{In our spatial
+  cross-validation algorithm in Section \ref{spatial-cross-validation}
+  issues can occur when rare species do not occur in the training set,
+  but then are encountered in the test set. This risk is mitigated by
+  representing \texttt{sp} as a factor variable, which has a complete
+  list of all levels of the categorical variable.}.
+
+TODO: enumerate what these variables are and say if they need exact
+names
 
@@
 \NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(comp_dist, }\DataTypeTok{cv_grid_sf =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{)}
+\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{)}
 \NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}\StringTok{ }

### 2021-03-03 R/data_processing_functions.R

@@ -39,14 +39,14 @@ compute_growth <- function(census_1, census_2, id) {
   check_inherits(id, "character")
 
   purrr::map2(
-    c(id, "sp", "gx", "gy", "date", "codes", "dbh"),
-    c("numeric", "factor", "numeric", "numeric", "Date", "character", "numeric"),
+    c(id, "gx", "gy", "date", "codes", "dbh"),
+    c("numeric", "numeric", "numeric", "Date", "character", "numeric"),
     check_column,
     census_1
   )
   purrr::map2(
-    c(id, "sp", "gx", "gy", "date", "codes", "dbh"),
-    c("numeric", "factor", "numeric", "numeric", "Date", "character", "numeric"),
+    c(id, "gx", "gy", "date", "codes", "dbh"),
+    c("numeric", "numeric", "numeric", "Date", "character", "numeric"),
     check_column,
     census_2
   )
@@ -74,7 +74,9 @@ compute_growth <- function(census_1, census_2, id) {
       n_days = difftime(date2, date1),
       n_days = as.numeric(n_days),
       n_years = n_days / 365.25,
-      growth = (dbh2 - dbh1) / n_years
+      growth = (dbh2 - dbh1) / n_years,
+      # Convert species to factor:
+      sp = factor(sp)
     ) %>%
     select(-c(n_days, n_years, date1, date2)) %>%
     st_as_sf(coords = c("gx", "gy"))

### 2021-03-03 paper/simulation_results/replicate_plosone_fig1_bw.R

@@ -16,10 +16,10 @@ census_2008_bw <- census_2008_bw %>%
 growth_bw <-
   compute_growth(
     census_1 = census_2008_bw %>%
-      mutate(sp = to_any_case(sp) %>% factor()),
+      mutate(sp = to_any_case(sp)),
     census_2 = census_2014_bw %>%
       filter(!str_detect(codes, "R")) %>%
-      mutate(sp = to_any_case(sp) %>% factor()),
+      mutate(sp = to_any_case(sp)),
     id = "treeID"
   )
 
@@ -94,7 +94,7 @@ for(i in 1:length(species_notion_vector)){
 
   # Focal vs comp main dataframe for analysis
   focal_vs_comp_bw <- growth_bw %>%
-    mutate(sp = .data[[species_notion_vector[i]]] %>% factor()) %>%
+    mutate(sp = .data[[species_notion_vector[i]]]) %>%
     create_focal_vs_comp(comp_dist, blocks = blocks_bw, id = "treeID")
 
 

### 2021-03-03 paper/simulation_results/replicate_plosone_fig1_scbi.R

@@ -14,8 +14,7 @@ census_2013_scbi <- here("paper/scbi.stem2.csv") %>%
   select(stemID, sp, date = ExactDate, gx, gy, dbh, codes, status) %>%
   mutate(
     date = mdy(date),
-    dbh = as.numeric(dbh)/10,
-    sp = factor(sp)
+    dbh = as.numeric(dbh)/10
   ) %>%
   filter(gx < 300, between(gy, 300, 600))
 
@@ -24,8 +23,7 @@ census_2018_scbi <- here("paper/scbi.stem3.csv") %>%
   select(stemID, sp, date = ExactDate, gx, gy, dbh, codes, status) %>%
   mutate(
     date = mdy(date),
-    dbh = as.numeric(dbh)/10,
-    sp = factor(sp)
+    dbh = as.numeric(dbh)/10
   ) %>%
   filter(gx < 300, between(gy, 300, 600))
 

### 2021-03-03 README.Rmd
renamed all cv_grid_sf object names to blocks
replace all cv_grid arguments with blocks

@@
 focal_vs_comp_ex <- growth_ex %>%
-  create_focal_vs_comp(comp_dist, cv_grid_sf = blocks_ex, id = "ID")
+  create_focal_vs_comp(comp_dist, blocks = blocks_ex, id = "ID")
 focal_vs_comp_ex
@@
 focal_vs_comp_ex <- focal_vs_comp_ex %>%
-  run_cv(comp_dist = comp_dist, cv_grid = blocks_ex)
+  run_cv(comp_dist = comp_dist, blocks = blocks_ex)
 ```

### 2021-03-03 paper/forestecology.Rmd
replace all cv_grid arguments with blocks

@@
 focal_vs_comp_scbi <- focal_vs_comp_scbi %>%
-  run_cv(comp_dist = comp_dist, cv_grid = blocks_scbi)
+  run_cv(comp_dist = comp_dist, blocks = blocks_scbi)
 ```

### 2021-03-03 R/data_processing_functions.R

@@ -162,7 +162,7 @@ compute_growth <- function(census_1, census_2, id) {
 #' # Load in growth_df with spatial data
 #' # See ?growth_ex for attaching spatial data to growth_df
 #' data(growth_spatial_ex)
-#' # Load in cv_grid
+#' # Load in blocks
 #' data(blocks_ex)
 #'
 #' focal_vs_comp_ex <- growth_spatial_ex %>%

### 2021-03-03 R/model_functions.R

@@ -262,7 +262,6 @@ predict.comp_bayes_lm <- function(object, newdata, ...) {
 #'
 #' @inheritParams comp_bayes_lm
 #' @inheritParams create_focal_vs_comp
-#' @param cv_grid \code{sf} polygon output from \code{\link[blockCV]{spatialBlock}}
 #'
 #' @import dplyr
 #' @import sf
@@ -280,12 +279,12 @@ predict.comp_bayes_lm <- function(object, newdata, ...) {
 #' run_cv(
 #'   focal_vs_comp_ex,
 #'   comp_dist = 1,
-#'   cv_grid = blocks_ex
+#'   blocks = blocks_ex
 #' )
-run_cv <- function(focal_vs_comp, comp_dist, cv_grid, prior_param = NULL, run_shuffle = FALSE) {
+run_cv <- function(focal_vs_comp, comp_dist, blocks, prior_param = NULL, run_shuffle = FALSE) {
   check_focal_vs_comp(focal_vs_comp)
   check_inherits(comp_dist, "numeric")
-  check_inherits(cv_grid, "sf")
+  check_inherits(blocks, "sf")
   if (!is.null(prior_param)) {
     check_prior_params(prior_param)
   }
@@ -302,14 +301,14 @@ run_cv <- function(focal_vs_comp, comp_dist, cv_grid, prior_param = NULL, run_sh
     fit_one_fold,
     focal_vs_comp,
     comp_dist,
-    cv_grid,
+    blocks,
     prior_param,
     run_shuffle
   )
 }
 
 fit_one_fold <- function(fold, focal_vs_comp, comp_dist,
-                         cv_grid, prior_param, run_shuffle) {
+                         blocks, prior_param, run_shuffle) {
   # Define test set and "full" training set (we will remove buffer region below)
   test <- focal_vs_comp %>%
     filter(foldID == fold)
@@ -322,7 +321,7 @@ fit_one_fold <- function(fold, focal_vs_comp, comp_dist,
   }
 
   # Define sf object of boundary of test fold
-  test_fold_boundary <- cv_grid %>%
+  test_fold_boundary <- blocks %>%
     subset(folds == fold) %>%
     st_bbox() %>%
     st_as_sfc()

### 2021-03-03 paper/simulation_results/replicate_plosone_fig1_bw.R

@@ -112,7 +112,7 @@ for(i in 1:length(species_notion_vector)){
 
   # 2. Compute observed test statistic: RMSE with cross-validation ----
   observed_RMSE_CV[i] <- focal_vs_comp_bw %>%
-    run_cv(comp_dist = comp_dist, cv_grid = blocks_bw) %>%
+    run_cv(comp_dist = comp_dist, blocks = blocks_bw) %>%
     rmse(truth = growth, estimate = growth_hat) %>%
     pull(.estimate)
 
@@ -141,7 +141,7 @@ for(i in 1:length(species_notion_vector)){
   for(j in 1:num_shuffle){
     # Compute and save RMSE
     shuffle_RMSE_CV[[i]][j] <- focal_vs_comp_bw %>%
-      run_cv(comp_dist = comp_dist, cv_grid = blocks_bw, run_shuffle = TRUE) %>%
+      run_cv(comp_dist = comp_dist, blocks = blocks_bw, run_shuffle = TRUE) %>%
       rmse(truth = growth, estimate = growth_hat) %>%
       pull(.estimate)
 

### 2021-03-03 paper/simulation_results/replicate_plosone_fig1_scbi.R

@@ -113,7 +113,7 @@ observed_RMSE <- focal_vs_comp_scbi %>%
 
 # 2. Compute observed test statistic: RMSE with cross-validation ----
 observed_RMSE_CV <- focal_vs_comp_scbi %>%
-  run_cv(comp_dist = comp_dist, cv_grid = blocks_scbi) %>%
+  run_cv(comp_dist = comp_dist, blocks = blocks_scbi) %>%
   rmse(truth = growth, estimate = growth_hat) %>%
   pull(.estimate)
 
@@ -142,7 +142,7 @@ shuffle_RMSE_CV <- numeric(length = num_shuffle)
 for(j in 1:num_shuffle){
   # Compute and save RMSE
   shuffle_RMSE_CV[j] <- focal_vs_comp_scbi %>%
-    run_cv(comp_dist = comp_dist, cv_grid = blocks_scbi, run_shuffle = TRUE) %>%
+    run_cv(comp_dist = comp_dist, blocks = blocks_scbi, run_shuffle = TRUE) %>%
     rmse(truth = growth, estimate = growth_hat) %>%
     pull(.estimate)
 

### 2021-03-03 tests/testthat/test-readme.R

@@ -73,7 +73,7 @@ test_that("readme code works", {
   expect_true(
     check_inherits(
       focal_vs_comp_ex %>%
-        run_cv(comp_dist = comp_dist, cv_grid = blocks_ex) %>%
+        run_cv(comp_dist = comp_dist, blocks = blocks_ex) %>%
         right_join(growth_ex, by = c("focal_ID" = "ID")),
       "data.frame"
     )
@@ -89,7 +89,7 @@ test_that("readme code works", {
   )
 
   focal_vs_comp_ex <- focal_vs_comp_ex %>%
-    run_cv(comp_dist = comp_dist, cv_grid = blocks_ex)
+    run_cv(comp_dist = comp_dist, blocks = blocks_ex)
 
   expect_equal(
     focal_vs_comp_ex %>%

### 2021-03-04 paper/forestecology.Rmd
resolved all TODO's and started following submission guidelines

@@
 
-- name: David Allen
+- name: David N. Allen
   affiliation: Biology Department, Middlebury College
@@
 
+spacing: double
+
 bibliography: paper.bib
@@
 header-includes:
-- \usepackage{xcolor, soul, xspace, float, subfig, lineno}
+- \usepackage{xcolor, soul, xspace, float, subfig, lineno, setspace}
+- \linenumbers
 ---
-
-\linenumbers
 
@@
 ```
+
+\doublespacing
 
@@
 The output `growth_scbi` is a single data frame of class `sf` that includes variables `growth`, the average annual growth in DBH (cm y^-1^) for all stems that were alive at both time points, and `geometry`, the `sf` package's encoding of geolocations of type `<POINT>`.
-Furthermore, the variables that should remain unchanged between censuses appear only once, such as location variables `gx` and `gy`; as well as species-related variables.
+In addition the species variable `sp` is converted to a factor if it wasn't already by `compute_growth()`.^[In our spatial cross-validation algorithm in Section \ref{spatial-cross-validation} issues can occur when rare species do not occur in the training set, but then are encountered in the test set.
+This risk is mitigated by representing `sp` as a factor variable, which has a complete list of all levels of the categorical variable.] Furthermore, the variables that should remain unchanged between censuses appear only once, such as location variables `gx` and `gy`; as well as species-related variables.
 Variables that should change between censuses are suffixed with `1` and `2` indicating the earlier and later censuses, such as `dbh1/dbh2` and `codes1/codes2`.
+
 Site data that does not align with this convention will need to be transformed for use with the `compute_growth()` function.
-However, in the end, all that matters is that the growth of all stems is saved in a data frame of class `sf` and, at a minimum, the data contains the variables above.
-Lastly, the species variable `sp` is automatically converted to a factor^[In our spatial cross-validation algorithm in Section \ref{spatial-cross-validation} issues can occur when rare species do not occur in the training set, but then are encountered in the test set.
-This risk is mitigated by representing `sp` as a factor variable, which has a complete list of all levels of the categorical variable.].
+However, in the end, all that matters is that the growth of all stems is saved in a data frame of class `sf` and, at a minimum, contains the variable uniquely identifying each stem, `sp`, `dbh1`, `growth`, `geometry`.
 
-TODO: enumerate what these variables are and say if they need exact names
+Given that `growth_scbi` is of class `sf`, it can be easily plotted in `ggplot2` using the `geom_sf()` geometry as seen in Figure \ref{fig:scbi-trees} where we plot a random sample of 500 out of the `r nrow(growth_scbi)` trees.
 
-Given that `growth_scbi` is of class `sf`, it can be easily plotted in `ggplot2` using the `geom_sf()` geometry as seen in Figure \ref{fig:scbi-trees}.
-
-```{r scbi-trees, out.width="66%", fig.cap="Compute growth of trees based on census data: Map with growth of all trees from a 9 ha subsection of the Smithsonian Conservation Biology Institute (SCBI) forest plot."}
-# TODO: Rescale points in this plot
+```{r scbi-trees, out.width="66%", fig.cap="Compute growth of trees based on census data: Map with growth of a random sample of 500 trees from a 9 ha subsection of the Smithsonian Conservation Biology Institute (SCBI) forest plot."}
 ggplot() +
-  geom_sf(data = growth_scbi, aes(size = growth)) + 
-  scale_size(breaks = c(0.01, 0.1, 1), range = c(0.1, 1))
+  geom_sf(data = growth_scbi %>% sample_n(500), aes(size = growth)) + 
+  scale_size_binned(limits = c(0.1, 1))
 ```
@@
 The next step in our analysis sequence is to identify all focal trees and their corresponding competitor trees.
+More specifically, it identifies all trees that are not part of the buffer region, have a valid `growth` measurement, and have at least one neighbor within `r comp_dist` of it.
 The `create_focal_vs_comp()` functions performs these tasks and returns a new data frame of type `sf`.
@@
 ```{r scbi-focal-vs-comp}
-# TODO: Below reconcile the number of rows as they off by one from growth_scbi %>% 
-# filter(!is.na(growth) & !buffer). Perhaps by removing NA's in the growth_scbi stage?
 focal_vs_comp_scbi <- growth_scbi %>%
@@
   select(-foldID)
+```
+
+```{r, eval = FALSE, echo = FALSE}
+# Internal: after removing trees in buffer and have NA for growth, there is one
+# tree that is not included in focal_vs_comp_scbi b/c it has no neighbors within
+# comp_dist = 7.5m of it
+growth_scbi %>% 
+  filter(!buffer & !is.na(growth)) %>% 
+  filter(!stemID %in% focal_vs_comp_scbi$focal_ID)
+
+growth_scbi %>% 
+  mutate(flag = stemID == 18823) %>% 
+  ggplot() +
+  geom_sf(aes(col = flag))
 ```
@@
 
-```{r, eval = FALSE}
+```{r, scbi-focal-vs-comp-3}
 focal_vs_comp_scbi %>% 
@@
   unnest(cols = "comp")
-```
-```{r scbi-focal-vs-comp-3, echo = FALSE}
-# Not sure why we need to do things this way:
-focal_vs_comp_scbi %>% 
-  filter(focal_ID == 4) %>% 
-  select(focal_ID, dbh, comp) %>% 
-  unnest(cols = "comp") %>% 
-  print(n = 5)
 ```
@@
 ```{r scbi-model-fit}
-# TODO: Add information about default priors in ?comp_bayes_lm()
 comp_bayes_lm_scbi <- focal_vs_comp_scbi %>%
@@
 
+The `forestecology` package provides
+
 - Run full simulation on SCBI data
@@
 
-The authors thank Ryan Giordano and Jonathan Che for their help with the statistical methodology and Sophie Li for their feedback on package interface.
+The authors thank Ryan Giordano and Jonathan Che for their help with the statistical methodology and Sophie Li for their feedback on package interface. The authors declare no conflicts of interest.
+
+# Author's contributions
+
+A.Y.K.
+and D.N.A.
+conceived the ideas and coded most of the package in the early design stages.
+A.Y.K.
+led the writing of the manuscript.
+S.P.C.
+refactored much of the package's code in the later design stages to align with R and "tidy" programming best practices [@wickham_welcome_2019].
+All authors contributed critically to the drafts and gave final approval for publication.
+
+# Data accessibility
+
+We intend to archive all data and code on GitHub at the following repository https://github.com/rudeboybert/forestecology. This repository will be versioned and archived on Zenodo upon acceptance.
 

### 2021-03-04 paper/forestecology.tex
resolved all TODO's and started following submission guidelines

@@
 
-\usepackage{xcolor, soul, xspace, float, subfig, lineno}
+\usepackage{xcolor, soul, xspace, float, subfig, lineno, setspace}
+\linenumbers
 
@@
     Program in Statistical \& Data Sciences, Smith College\\
-     and \\     David Allen \\
+     and \\     David N. Allen \\
     Biology Department, Middlebury College\\
@@
 
-\linenumbers
+\doublespacing
 
@@
 encoding of geolocations of type
-\texttt{\textless{}POINT\textgreater{}}. Furthermore, the variables that
+\texttt{\textless{}POINT\textgreater{}}. In addition the species
+variable \texttt{sp} is converted to a factor if it wasn't already by
+\texttt{compute\_growth()}.\footnote{In our spatial cross-validation
+  algorithm in Section \ref{spatial-cross-validation} issues can occur
+  when rare species do not occur in the training set, but then are
+  encountered in the test set. This risk is mitigated by representing
+  \texttt{sp} as a factor variable, which has a complete list of all
+  levels of the categorical variable.} Furthermore, the variables that
 should remain unchanged between censuses appear only once, such as
@@
 later censuses, such as \texttt{dbh1/dbh2} and \texttt{codes1/codes2}.
+
 Site data that does not align with this convention will need to be
@@
 However, in the end, all that matters is that the growth of all stems is
-saved in a data frame of class \texttt{sf} and, at a minimum, the data
-contains the variables above. Lastly, the species variable \texttt{sp}
-is automatically converted to a factor\footnote{In our spatial
-  cross-validation algorithm in Section \ref{spatial-cross-validation}
-  issues can occur when rare species do not occur in the training set,
-  but then are encountered in the test set. This risk is mitigated by
-  representing \texttt{sp} as a factor variable, which has a complete
-  list of all levels of the categorical variable.}.
-
-TODO: enumerate what these variables are and say if they need exact
-names
+saved in a data frame of class \texttt{sf} and, at a minimum, contains
+the variable uniquely identifying each stem, \texttt{sp}, \texttt{dbh1},
+\texttt{growth}, \texttt{geometry}.
 
@@
 easily plotted in \texttt{ggplot2} using the \texttt{geom\_sf()}
-geometry as seen in Figure \ref{fig:scbi-trees}.
+geometry as seen in Figure \ref{fig:scbi-trees} where we plot a random
+sample of 500 out of the 7954 trees.
 
@@
 \begin{Highlighting}[]
-\CommentTok{# }\AlertTok{TODO}\CommentTok{: Rescale points in this plot}
 \KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_scbi, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{size =}\NormalTok{ growth)) }\OperatorTok{+}\StringTok{ }
-\StringTok{  }\KeywordTok{scale_size}\NormalTok{(}\DataTypeTok{breaks =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.01}\NormalTok{, }\FloatTok{0.1}\NormalTok{, }\DecValTok{1}\NormalTok{), }\DataTypeTok{range =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\DecValTok{1}\NormalTok{))}
+\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{sample_n}\NormalTok{(}\DecValTok{500}\NormalTok{), }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{size =}\NormalTok{ growth)) }\OperatorTok{+}\StringTok{ }
+\StringTok{  }\KeywordTok{scale_size_binned}\NormalTok{(}\DataTypeTok{limits =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\DecValTok{1}\NormalTok{))}
 \end{Highlighting}
@@
 
-\caption{Compute growth of trees based on census data: Map with growth of all trees from a 9 ha subsection of the Smithsonian Conservation Biology Institute (SCBI) forest plot.}\label{fig:scbi-trees}
+\caption{Compute growth of trees based on census data: Map with growth of a random sample of 500 trees from a 9 ha subsection of the Smithsonian Conservation Biology Institute (SCBI) forest plot.}\label{fig:scbi-trees}
 \end{figure}
@@
 The next step in our analysis sequence is to identify all focal trees
-and their corresponding competitor trees. The
-\texttt{create\_focal\_vs\_comp()} functions performs these tasks and
-returns a new data frame of type \texttt{sf}. On top of the previous
-arguments \texttt{comp\_dist} defining the competition neighborhood and
-\texttt{id} indicating which variable uniquely identifies each
-tree-stem, this function also requires an \texttt{sf} object
-representation of the spatial cross-validation blocks/folds. In this
-example, the blocks were manually encoded in \texttt{blocks\_scbi} by
-specifying it's vertices in Section \ref{spatial-information}\footnote{We
-  present an alternative method for defining spatial cross-validation
-  blocks is using the \texttt{spatialBlock()} function from the
-  \texttt{blockCV} package in the Appendix}. We present the resulting
-data frame below with the \texttt{foldID} variable omitted for
-compactness of presentation.
+and their corresponding competitor trees. More specifically, it
+identifies all trees that are not part of the buffer region, have a
+valid \texttt{growth} measurement, and have at least one neighbor within
+7.5 of it. The \texttt{create\_focal\_vs\_comp()} functions performs
+these tasks and returns a new data frame of type \texttt{sf}. On top of
+the previous arguments \texttt{comp\_dist} defining the competition
+neighborhood and \texttt{id} indicating which variable uniquely
+identifies each tree-stem, this function also requires an \texttt{sf}
+object representation of the spatial cross-validation blocks/folds. In
+this example, the blocks were manually encoded in \texttt{blocks\_scbi}
+by specifying it's vertices in Section
+\ref{spatial-information}\footnote{We present an alternative method for
+  defining spatial cross-validation blocks is using the
+  \texttt{spatialBlock()} function from the \texttt{blockCV} package in
+  the Appendix}. We present the resulting data frame below with the
+\texttt{foldID} variable omitted for compactness of presentation.
 
@@
 \begin{Highlighting}[]
-\CommentTok{# }\AlertTok{TODO}\CommentTok{: Below reconcile the number of rows as they off by one from growth_scbi %>% }
-\CommentTok{# filter(!is.na(growth) & !buffer). Perhaps by removing NA's in the growth_scbi stage?}
 \NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
@@
 \StringTok{  }\KeywordTok{unnest}\NormalTok{(}\DataTypeTok{cols =} \StringTok{"comp"}\NormalTok{)}
+\CommentTok{## # A tibble: 20 x 6}
+\CommentTok{##   focal_ID   dbh comp_ID  dist comp_sp comp_basal_area}
+\CommentTok{##      <dbl> <dbl>   <dbl> <dbl> <fct>             <dbl>}
+\CommentTok{## 1        4  13.6    1836  7.48 tiam            0.0176 }
+\CommentTok{## 2        4  13.6    1847  2.81 nysy            0.00332}
+\CommentTok{## 3        4  13.6    1848  1.62 nysy            0.00396}
+\CommentTok{## 4        4  13.6    1849  2.62 nysy            0.00535}
+\CommentTok{## 5        4  13.6    1850  2.98 havi            0.00472}
+\CommentTok{## # ... with 15 more rows}
 \end{Highlighting}
 \end{Shaded}
-
-\begin{verbatim}
-## # A tibble: 20 x 6
-##   focal_ID   dbh comp_ID  dist comp_sp comp_basal_area
-##      <dbl> <dbl>   <dbl> <dbl> <fct>             <dbl>
-## 1        4  13.6    1836  7.48 tiam            0.0176 
-## 2        4  13.6    1847  2.81 nysy            0.00332
-## 3        4  13.6    1848  1.62 nysy            0.00396
-## 4        4  13.6    1849  2.62 nysy            0.00535
-## 5        4  13.6    1850  2.98 havi            0.00472
-## # ... with 15 more rows
-\end{verbatim}
 
@@
 \begin{Highlighting}[]
-\CommentTok{# }\AlertTok{TODO}\CommentTok{: Add information about default priors in ?comp_bayes_lm()}
 \NormalTok{comp_bayes_lm_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
@@
 \NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{cv_grid =}\NormalTok{ blocks_scbi)}
+\StringTok{  }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi)}
 \end{Highlighting}
@@
 
+The \texttt{forestecology} package provides
+
 \begin{itemize}
@@
 statistical methodology and Sophie Li for their feedback on package
-interface.
+interface. The authors declare no conflicts of interest.
+
+\hypertarget{authors-contributions}{%
+\section{Author's contributions}\label{authors-contributions}}
+
+A.Y.K. and D.N.A. conceived the ideas and coded most of the package in
+the early design stages. A.Y.K. led the writing of the manuscript.
+S.P.C. refactored much of the package's code in the later design stages
+to align with R and ``tidy'' programming best practices
+\citep{wickham_welcome_2019}. All authors contributed critically to the
+drafts and gave final approval for publication.
+
+\hypertarget{data-accessibility}{%
+\section{Data accessibility}\label{data-accessibility}}
+
+We intend to archive all data and code on GitHub at the following
+repository \url{https://github.com/rudeboybert/forestecology}. This
+repository will be versioned and archived on Zenodo upon acceptance.
 

### 2021-03-04 paper/forestecology.Rmd
more formatting

@@
   affiliation: Program in Statistical & Data Sciences, Smith College
+  thanks: "Assistant Professor, Statistical & Data Sciences, Smith College, Northampton, MA 01063 (e-mail: akim04@smith.edu)."
 
@@
 keywords:
-- forest ecology, competition, R, Rstats, tidyverse, sf, cross-validation, spatial statistics
-
-spacing: double
+- forest ecology, inter-species competition, tree growth, R, Rstats, tidyverse, sf, cross-validation, spatial statistics
 
@@
   collapse = TRUE,
-  comment = "##"
+  comment = "##",
+  tidy.opts = list(width.cutoff=72),
+  tidy = TRUE
 )
@@
 options(
-  width = 70, 
+  width = 72, 
   # Decimal precision
@@
   # Number of rows for tibble printing
-  tibble.print_max = 5, tibble.print_min = 5
+  tibble.print_max = 5, 
+  tibble.print_min = 5
 )
@@
 
-## Step 1: Compute the growth of trees based on census data
+## Step 1: Compute the growth of trees based on census data {compute-growth}
 
@@
 
-We load both 2008 and 2014 SCBI census data `.csv` files as they existed on GitHub on November 20, 2020.
+We load both 2008 and 2014 SCBI census data `.csv` files as they existed on GitHub on November 20, 2020 [@gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020].
 After selecting only the relevant variables, we perform a few additional data wrangling steps: convert the character variable with the date of measurement to be of explicit type `date`, convert DBH to be in centimeters^[A rule of thumb to determine the units of DBH is to check if the smallest non-zero and non-missing measurement is 1 or 10.
@@
 We then use the output of the `spatialBlock()` function from the `blockCV` package to associate each tree in `growth_scbi` to the correct fold (saved in the `foldID` variable) [@valavi_blockcv_2019].
-^[In the Appendix we present an example where the folds themselves are also created using the `spatialBlock()` function given a specified `cv_block_size`.]
+^[In the Supporting Information we present an example where the folds themselves are also created using the `spatialBlock()` function given a specified `cv_block_size`.]
 
@@
 On top of the previous arguments `comp_dist` defining the competition neighborhood and `id` indicating which variable uniquely identifies each tree-stem, this function also requires an `sf` object representation of the spatial cross-validation blocks/folds.
-In this example, the blocks were manually encoded in `blocks_scbi` by specifying it's vertices in Section \ref{spatial-information}^[We present an alternative method for defining spatial cross-validation blocks is using the `spatialBlock()` function from the `blockCV` package in the Appendix].
+In this example, the blocks were manually encoded in `blocks_scbi` by specifying it's vertices in Section \ref{spatial-information}^[We present an alternative method for defining spatial cross-validation blocks is using the `spatialBlock()` function from the `blockCV` package in the Supporting Information.].
 We present the resulting data frame below with the `foldID` variable omitted for compactness of presentation.
@@
 The resulting RMSE of `r model_rmse_shuffle` based on the permutation test is `r ifelse(model_rmse_shuffle < model_rmse, "smaller", "larger")` than the earlier RMSE of `r model_rmse`, suggesting that models that do incorporate competitor species identity better fit the data.
-We conduct an analysis of the full SCBI plot in Section below.
+We conduct an analysis of the full SCBI plot in Section \ref{results} below.
 
@@
 The resulting RMSE of `r model_rmse_cv` computed using cross-validation is `r ifelse(model_rmse_cv > model_rmse, "larger", "smaller")` than the earlier RMSE of `r model_rmse`, suggesting that models that do not take the inherent spatial autocorrelation of the data into account generate error estimates that are overly optimistic; in our case RMSE's that are too low.
+We conduct an analysis of the full SCBI plot in Section \ref{results} below.
+
+# Results {#results}
+
+Figure \ref{fig:scbi-simulation}
+
+```{r scbi-simulation, out.width="100%", fig.cap="Root mean squared error of models for standard, permuted, and spatial cross-validated error estimates.", echo = FALSE}
+include_graphics("simulation_results/2021-03-03_scbi_49_shuffles.png")
+```
 
@@
 
-We intend to archive all data and code on GitHub at the following repository https://github.com/rudeboybert/forestecology. This repository will be versioned and archived on Zenodo upon acceptance.
+We intend to archive all data and source code for the `forestecology` package as well as this manuscript on GitHub at https://github.com/rudeboybert/forestecology.
+This repository will be versioned and archived on Zenodo upon acceptance.
+The 2008 and 2014 Smithsonian Conservation Biology Institute census data loaded in Section \ref{compute-growth} and saved in `scbi.stem2.csv` and `scbi.stem3.csv` are available on GitHub at https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/tree/master/tree_main_census/data/census-csv-files and have been versioned and archived on Zenodo at https://doi.org/10.5281/zenodo.2649301 [@gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020].
 

### 2021-03-04 paper/forestecology.tex
more formatting

@@
   \author{
-        Albert Y. Kim \\
+        Albert Y. Kim \thanks{Assistant Professor, Statistical \& Data Sciences, Smith College,
+Northampton, MA 01063 (e-mail:
+\href{mailto:akim04@smith.edu}{\nolinkurl{akim04@smith.edu}}).} \\
     Program in Statistical \& Data Sciences, Smith College\\
@@
 \noindent%
-{\it Keywords:} forest ecology, competition, R, Rstats, tidyverse, sf, cross-validation,
-spatial statistics
+{\it Keywords:} forest ecology, inter-species competition, tree growth, R, Rstats,
+tidyverse, sf, cross-validation, spatial statistics
 \vfill
@@
 
-\hypertarget{step-1-compute-the-growth-of-trees-based-on-census-data}{%
-\subsection{Step 1: Compute the growth of trees based on census
-data}\label{step-1-compute-the-growth-of-trees-based-on-census-data}}
+\hypertarget{step-1-compute-the-growth-of-trees-based-on-census-data-compute-growth}{%
+\subsection{Step 1: Compute the growth of trees based on census data
+\{compute-growth\}}\label{step-1-compute-the-growth-of-trees-based-on-census-data-compute-growth}}
 
@@
 We load both 2008 and 2014 SCBI census data \texttt{.csv} files as they
-existed on GitHub on November 20, 2020. After selecting only the
-relevant variables, we perform a few additional data wrangling steps:
-convert the character variable with the date of measurement to be of
-explicit type \texttt{date}, convert DBH to be in centimeters\footnote{A
-  rule of thumb to determine the units of DBH is to check if the
-  smallest non-zero and non-missing measurement is 1 or 10. If the
-  former, then centimeters. If the later, then millimeters. This is
-  because ForestGEO protocols state that only trees with DBH greater or
-  equal to 1cm should be included in censuses.} Furthermore, in order to
-speed up computation for purposes of this example, we only consider a 9
-ha subsection of the 25.6 ha of the SCBI site: \texttt{gx} from 0--300
+existed on GitHub on November 20, 2020
+\citep{gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020}. After
+selecting only the relevant variables, we perform a few additional data
+wrangling steps: convert the character variable with the date of
+measurement to be of explicit type \texttt{date}, convert DBH to be in
+centimeters\footnote{A rule of thumb to determine the units of DBH is to
+  check if the smallest non-zero and non-missing measurement is 1 or 10.
+  If the former, then centimeters. If the later, then millimeters. This
+  is because ForestGEO protocols state that only trees with DBH greater
+  or equal to 1cm should be included in censuses.} Furthermore, in order
+to speed up computation for purposes of this example, we only consider a
+9 ha subsection of the 25.6 ha of the SCBI site: \texttt{gx} from 0--300
 instead of 0--400 and \texttt{gy} from 300--600 instead of 0--640.
@@
 \CommentTok{## # A tibble: 7,954 x 9}
-\CommentTok{##   stemID sp     dbh1 codes1 status  dbh2 codes2 growth   geometry}
-\CommentTok{##    <dbl> <fct> <dbl> <chr>  <chr>  <dbl> <chr>   <dbl>    <POINT>}
-\CommentTok{## 1      4 nysy  13.6  M      A       14.2 M       0.103 (14.2 428)}
-\CommentTok{## 2      5 havi   8.8  M      A        9.6 M;P     0.150  (9.4 436)}
-\CommentTok{## 3      6 havi   3.25 NULL   A        4   M       0.140  (1.3 434)}
-\CommentTok{## 4     77 qual  65.2  M      A       66   M       0.141 (34.7 307)}
-\CommentTok{## 5     79 tiam  47.7  M      A       46.8 M      -0.161   (40 381)}
-\CommentTok{## # ... with 7,949 more rows}
+\CommentTok{##   stemID sp     dbh1 codes1 status  dbh2 codes2 growth}
+\CommentTok{##    <dbl> <fct> <dbl> <chr>  <chr>  <dbl> <chr>   <dbl>}
+\CommentTok{## 1      4 nysy  13.6  M      A       14.2 M       0.103}
+\CommentTok{## 2      5 havi   8.8  M      A        9.6 M;P     0.150}
+\CommentTok{## 3      6 havi   3.25 NULL   A        4   M       0.140}
+\CommentTok{## 4     77 qual  65.2  M      A       66   M       0.141}
+\CommentTok{## 5     79 tiam  47.7  M      A       46.8 M      -0.161}
+\CommentTok{## # ... with 7,949 more rows, and 1 more variable: geometry <POINT>}
 \end{Highlighting}
@@
 in the \texttt{foldID} variable) \citep{valavi_blockcv_2019}. \footnote{In
-  the Appendix we present an example where the folds themselves are also
-  created using the \texttt{spatialBlock()} function given a specified
-  \texttt{cv\_block\_size}.}
+  the Supporting Information we present an example where the folds
+  themselves are also created using the \texttt{spatialBlock()} function
+  given a specified \texttt{cv\_block\_size}.}
 
@@
   \texttt{spatialBlock()} function from the \texttt{blockCV} package in
-  the Appendix}. We present the resulting data frame below with the
-\texttt{foldID} variable omitted for compactness of presentation.
+  the Supporting Information.}. We present the resulting data frame
+below with the \texttt{foldID} variable omitted for compactness of
+presentation.
 
@@
 \CommentTok{## Model formula:}
-\CommentTok{## growth ~ sp + dbh + dbh * sp + acne * sp + acpl * sp + acru * sp + acsp * sp + aial * sp + amar * sp + astr * sp + beth * sp + caca * sp + caco * sp + cade * sp + cagl * sp + caovl * sp + casp * sp + cato * sp + ceca * sp + ceoc * sp + chvi * sp + coal * sp + coam * sp + cofl * sp + crpr * sp + crsp * sp + divi * sp + elum * sp + eual * sp + fagr * sp + fram * sp + frni * sp + frpe * sp + frsp * sp + havi * sp + ilve * sp + juci * sp + juni * sp + juvi * sp + libe * sp + litu * sp + loma * sp + nysy * sp + pato * sp + pipu * sp + pist * sp + pivi * sp + ploc * sp + prav * sp + prpe * sp + prse * sp + prsp * sp + qual * sp + quco * sp + qufa * sp + qumi * sp + qupr * sp + quru * sp + qusp * sp + quve * sp + rhpe * sp + romu * sp + rops * sp + rual * sp + rupe * sp + ruph * sp + saal * sp + saca * sp + tiam * sp + ulam * sp + ulru * sp + ulsp * sp + unk * sp + viac * sp + vipr * sp + vire * sp}
+\CommentTok{## growth ~ sp + dbh + dbh * sp + acne * sp + acru * sp + amar * sp + astr * sp + caca * sp + caco * sp + cade * sp + cagl * sp + caovl * sp + cato * sp + ceca * sp + ceoc * sp + chvi * sp + cofl * sp + crpr * sp + crsp * sp + divi * sp + elum * sp + fagr * sp + fram * sp + frni * sp + frpe * sp + havi * sp + ilve * sp + juci * sp + juni * sp + libe * sp + litu * sp + nysy * sp + pist * sp + pivi * sp + ploc * sp + prav * sp + prse * sp + qual * sp + quco * sp + qufa * sp + qumi * sp + qupr * sp + quru * sp + quve * sp + rops * sp + saal * sp + saca * sp + tiam * sp + ulam * sp + ulru * sp + unk * sp + vipr * sp}
 \end{Highlighting}
@@
 \CommentTok{## # A tibble: 6,296 x 8}
-\CommentTok{##   focal_ID focal_sp   dbh foldID   geometry growth}
-\CommentTok{##      <dbl> <fct>    <dbl> <fct>     <POINT>  <dbl>}
-\CommentTok{## 1        4 nysy     13.6  1      (14.2 428)  0.103}
-\CommentTok{## 2        5 havi      8.8  1       (9.4 436)  0.150}
-\CommentTok{## 3       79 tiam     47.7  1        (40 381) -0.161}
-\CommentTok{## 4       80 caca      5.15 1      (38.7 422)  0.253}
-\CommentTok{## 5       96 libe      2.3  1        (60 310)  0.262}
-\CommentTok{## # ... with 6,291 more rows, and 2 more variables: comp <list>,}
-\CommentTok{## #   growth_hat <dbl>}
+\CommentTok{##   focal_ID focal_sp   dbh foldID   geometry growth comp             }
+\CommentTok{##      <dbl> <fct>    <dbl> <fct>     <POINT>  <dbl> <list>           }
+\CommentTok{## 1        4 nysy     13.6  1      (14.2 428)  0.103 <tibble [20 x 4]>}
+\CommentTok{## 2        5 havi      8.8  1       (9.4 436)  0.150 <tibble [32 x 4]>}
+\CommentTok{## 3       79 tiam     47.7  1        (40 381) -0.161 <tibble [20 x 4]>}
+\CommentTok{## 4       80 caca      5.15 1      (38.7 422)  0.253 <tibble [12 x 4]>}
+\CommentTok{## 5       96 libe      2.3  1        (60 310)  0.262 <tibble [14 x 4]>}
+\CommentTok{## # ... with 6,291 more rows, and 1 more variable: growth_hat <dbl>}
 \end{Highlighting}
@@
 competitor species identity better fit the data. We conduct an analysis
-of the full SCBI plot in Section below.
+of the full SCBI plot in Section \ref{results} below.
 
@@
 error estimates that are overly optimistic; in our case RMSE's that are
-too low.
+too low. We conduct an analysis of the full SCBI plot in Section
+\ref{results} below.
+
+\hypertarget{results}{%
+\section{Results}\label{results}}
+
+Figure \ref{fig:scbi-simulation}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{simulation_results/2021-03-03_scbi_49_shuffles}
+
+}
+
+\caption{Root mean squared error of models for standard, permuted, and spatial cross-validated error estimates.}\label{fig:scbi-simulation}
+\end{figure}
 
@@
 
-We intend to archive all data and code on GitHub at the following
-repository \url{https://github.com/rudeboybert/forestecology}. This
-repository will be versioned and archived on Zenodo upon acceptance.
+We intend to archive all data and source code for the
+\texttt{forestecology} package as well as this manuscript on GitHub at
+\url{https://github.com/rudeboybert/forestecology}. This repository will
+be versioned and archived on Zenodo upon acceptance. The 2008 and 2014
+Smithsonian Conservation Biology Institute census data loaded in Section
+\ref{compute-growth} and saved in \texttt{scbi.stem2.csv} and
+\texttt{scbi.stem3.csv} are available on GitHub at
+\url{https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/tree/master/tree_main_census/data/census-csv-files}
+and have been versioned and archived on Zenodo at
+\url{https://doi.org/10.5281/zenodo.2649301}
+\citep{gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020}.
 

### 2021-03-04 paper/forestecology.Rmd
more formatting

@@
 ---
-title: "The forestecology R package for fitting and assessing models of interspecies competitive effects on the growth of trees"
+title: The forestecology R package for fitting and assessing models of interspecies competitive effects on the growth of trees
 
-# to produce blinded version set to 1 and change blinded <- TRUE on line 54 below
 blinded: 0
@@
   thanks: "Assistant Professor, Statistical & Data Sciences, Smith College, Northampton, MA 01063 (e-mail: akim04@smith.edu)."
-
 - name: David N. Allen
   affiliation: Biology Department, Middlebury College
-
 - name: Simon P. Couch
@@
 keywords:
-- forest ecology, inter-species competition, tree growth, R, Rstats, tidyverse, sf, cross-validation, spatial statistics
+- forest ecology, competition, tree growth, R, tidyverse, sf, cross-validation, spatial statistics
 
@@
 header-includes:
-- \usepackage{xcolor, soul, xspace, float, subfig, lineno, setspace}
+- \usepackage{xcolor, soul, xspace, float, subfig, lineno, setspace, fancyhdr}
+# Add line numbers:
 - \linenumbers
+# Add running header:
+- \pagestyle{fancy}
+- \fancyhead[LO,LE]{forestecology R package}
+- \renewcommand{\sectionmark}[1]{\markright{#1}{}}
 ---
@@
 
-- Run full simulation on SCBI data
-- run time considerations
-
 # Acknowledgments
@@
 
-A.Y.K.
-and D.N.A.
-conceived the ideas and coded most of the package in the early design stages.
-A.Y.K.
-led the writing of the manuscript.
-S.P.C.
-refactored much of the package's code in the later design stages to align with R and "tidy" programming best practices [@wickham_welcome_2019].
+AYK and DNA conceived the ideas and coded most of the package in the early design stages.
+AYK led the writing of the manuscript.
+SPC refactored much of the package's code in the later design stages to align with R and "tidy" programming best practices [@wickham_welcome_2019].
 All authors contributed critically to the drafts and gave final approval for publication.

### 2021-03-04 paper/forestecology.tex
more formatting

@@
 
-\usepackage{xcolor, soul, xspace, float, subfig, lineno, setspace}
+\usepackage{xcolor, soul, xspace, float, subfig, lineno, setspace, fancyhdr}
 \linenumbers
+\pagestyle{fancy}
+\fancyhead[LO,LE]{forestecology R package}
+\renewcommand{\sectionmark}[1]{\markright{#1}{}}
 
@@
 \noindent%
-{\it Keywords:} forest ecology, inter-species competition, tree growth, R, Rstats,
-tidyverse, sf, cross-validation, spatial statistics
+{\it Keywords:} forest ecology, competition, tree growth, R, tidyverse, sf,
+cross-validation, spatial statistics
 \vfill
@@
 
-\begin{itemize}
-\tightlist
-\item
-  Run full simulation on SCBI data
-\item
-  run time considerations
-\end{itemize}
-
 \hypertarget{acknowledgments}{%
@@
 
-A.Y.K. and D.N.A. conceived the ideas and coded most of the package in
-the early design stages. A.Y.K. led the writing of the manuscript.
-S.P.C. refactored much of the package's code in the later design stages
-to align with R and ``tidy'' programming best practices
+AYK and DNA conceived the ideas and coded most of the package in the
+early design stages. AYK led the writing of the manuscript. SPC
+refactored much of the package's code in the later design stages to
+align with R and ``tidy'' programming best practices
 \citep{wickham_welcome_2019}. All authors contributed critically to the

### 2021-03-04 paper/forestecology.Rmd
Updated simulation plot

@@
   ggplot2::theme_set(theme_light())
-  # Color scheme
+  # Color scheme:
   # scale_colour_discrete <- scale_colour_viridis_d
@@
 blocks_scbi <- bind_rows(
-  sf_polygon(fold1), sf_polygon(fold2), sf_polygon(fold3), sf_polygon(fold4)
+  sf_polygon(fold1), sf_polygon(fold2), sf_polygon(fold3), 
+  sf_polygon(fold4)
 ) %>%
@@
   geom_sf(data = blocks_scbi, fill = "transparent", linetype = "dashed") +
-  geom_sf_text(data = growth_scbi %>% sample_n(1000), aes(label = foldID, col = buffer))
+  geom_sf_text(data = growth_scbi %>% sample_n(1000), 
+               aes(label = foldID, col = buffer))
 ```
@@
 
-plot1 <- autoplot(comp_bayes_lm_scbi, type = "intercepts", sp_to_plot = sp_to_plot)
-plot2 <- autoplot(comp_bayes_lm_scbi, type = "dbh_slopes", sp_to_plot = sp_to_plot)
-plot3 <- autoplot(comp_bayes_lm_scbi, type = "competition", sp_to_plot = sp_to_plot)
+plot1 <- autoplot(comp_bayes_lm_scbi, type = "intercepts", 
+                  sp_to_plot = sp_to_plot)
+plot2 <- autoplot(comp_bayes_lm_scbi, type = "dbh_slopes", 
+                  sp_to_plot = sp_to_plot)
+plot3 <- autoplot(comp_bayes_lm_scbi, type = "competition", 
+                  sp_to_plot = sp_to_plot)
 
@@
 focal_vs_comp_scbi <- focal_vs_comp_scbi %>%
-  mutate(growth_hat_shuffle = predict(comp_bayes_lm_scbi_shuffle, newdata = focal_vs_comp_scbi))
+  mutate(
+    growth_hat_shuffle = predict(comp_bayes_lm_scbi_shuffle, 
+                                 newdata = focal_vs_comp_scbi)
+  )
 ```

### 2021-03-04 paper/forestecology.tex
Updated simulation plot

@@
 \NormalTok{blocks_scbi <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}
-  \KeywordTok{sf_polygon}\NormalTok{(fold1), }\KeywordTok{sf_polygon}\NormalTok{(fold2), }\KeywordTok{sf_polygon}\NormalTok{(fold3), }\KeywordTok{sf_polygon}\NormalTok{(fold4)}
+  \KeywordTok{sf_polygon}\NormalTok{(fold1), }\KeywordTok{sf_polygon}\NormalTok{(fold2), }\KeywordTok{sf_polygon}\NormalTok{(fold3), }
+  \KeywordTok{sf_polygon}\NormalTok{(fold4)}
 \NormalTok{) }\OperatorTok{%>%}
@@
 \StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ blocks_scbi, }\DataTypeTok{fill =} \StringTok{"transparent"}\NormalTok{, }\DataTypeTok{linetype =} \StringTok{"dashed"}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_sf_text}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{sample_n}\NormalTok{(}\DecValTok{1000}\NormalTok{), }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{label =}\NormalTok{ foldID, }\DataTypeTok{col =}\NormalTok{ buffer))}
+\StringTok{  }\KeywordTok{geom_sf_text}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{sample_n}\NormalTok{(}\DecValTok{1000}\NormalTok{), }
+               \KeywordTok{aes}\NormalTok{(}\DataTypeTok{label =}\NormalTok{ foldID, }\DataTypeTok{col =}\NormalTok{ buffer))}
 \end{Highlighting}
@@
 
-\NormalTok{plot1 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_scbi, }\DataTypeTok{type =} \StringTok{"intercepts"}\NormalTok{, }\DataTypeTok{sp_to_plot =}\NormalTok{ sp_to_plot)}
-\NormalTok{plot2 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_scbi, }\DataTypeTok{type =} \StringTok{"dbh_slopes"}\NormalTok{, }\DataTypeTok{sp_to_plot =}\NormalTok{ sp_to_plot)}
-\NormalTok{plot3 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_scbi, }\DataTypeTok{type =} \StringTok{"competition"}\NormalTok{, }\DataTypeTok{sp_to_plot =}\NormalTok{ sp_to_plot)}
+\NormalTok{plot1 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_scbi, }\DataTypeTok{type =} \StringTok{"intercepts"}\NormalTok{, }
+                  \DataTypeTok{sp_to_plot =}\NormalTok{ sp_to_plot)}
+\NormalTok{plot2 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_scbi, }\DataTypeTok{type =} \StringTok{"dbh_slopes"}\NormalTok{, }
+                  \DataTypeTok{sp_to_plot =}\NormalTok{ sp_to_plot)}
+\NormalTok{plot3 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_scbi, }\DataTypeTok{type =} \StringTok{"competition"}\NormalTok{, }
+                  \DataTypeTok{sp_to_plot =}\NormalTok{ sp_to_plot)}
 
@@
 \NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{growth_hat_shuffle =} \KeywordTok{predict}\NormalTok{(comp_bayes_lm_scbi_shuffle, }\DataTypeTok{newdata =}\NormalTok{ focal_vs_comp_scbi))}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
+    \DataTypeTok{growth_hat_shuffle =} \KeywordTok{predict}\NormalTok{(comp_bayes_lm_scbi_shuffle, }
+                                 \DataTypeTok{newdata =}\NormalTok{ focal_vs_comp_scbi)}
+\NormalTok{  )}
 \end{Highlighting}

### 2021-03-04 paper/simulation_results/replicate_plosone_fig1_scbi.R

@@ -186,7 +186,7 @@ model_comp_shuffle <- model_comp %>%
 
 cv_plot <- ggplot() +
   geom_vline(data = model_comp_observed, aes(xintercept = RMSE, col = CV), linetype = "dashed", show.legend = F) +
-  geom_histogram(data = model_comp_shuffle, aes(x = RMSE, fill = CV), bins = 200) +
+  geom_histogram(data = model_comp_shuffle, aes(x = RMSE, fill = CV), bins = 50) +
   labs(
     fill = "Cross-validated?",
     x = expression(paste("RMSE (cm ", y^{-1}, ")"))
@@ -198,5 +198,5 @@ cv_plot
 
 filename %>%
   str_c(".png") %>%
-  ggsave(plot = cv_plot)
+  ggsave(plot = cv_plot, width = 16/2, height = 9/2)
 

### 2021-03-06 paper/forestecology.Rmd
first batch of edits

@@
 - |
-  | 1.
-Many models for the growth of trees that incorporate interspecies competition are based on a neighborhood effect assumption whereby all trees within a fixed distance of a focal tree are considered competitors.
-Methods and tools are needed to quantify this competitive effect and assess the quality of all resulting models.
+  | 1. Neighborhood competition models are powerful tools to measure the effect of interspecific competition. Statistical methods to ease the application of these models are currently lacking.
   | 2.
-We present the `forestecology` package providing methods for both 1) evaluating the effect of competitor species identity using permutation tests and 2) evaluating model performance using spatial cross-validation.
-Following @allen_permutation_2020, we implement a Bayesian linear regression competition model.
+We present the `forestecology` package providing methods to i) specify neighborhood competition models, ii) evalulate the effect of competitor species identity using permutation tests, and iii) measure model performance using spatial cross-validation.
+Following @allen_permutation_2020, we implement a Bayesian linear regression neighborhood competition model.
   | 3.
@@
 Given ForestGEO's data collection protocols and data formatting standards, the package was designed with cross-site compatibility in mind.
-We demonstrate that both 1) competitor species identity matters and 2) that not spatially cross-validating leads to error estimates that are overly optimistic.
+We demonstrate that both i) competitor species identity matters and ii) that not spatially cross-validating leads to error estimates that are overly optimistic.
   | 4.
-The package features 1) `tidyverse`-like structure whereby verb-named functions can be modularly "piped" in sequence, 2) functions with standardized inputs/outputs of simple features `sf` package class, and 3) an S3 object-oriented implementation of the Bayesian linear regression model.
+The package features i) `tidyverse`-like structure whereby verb-named functions can be modularly "piped" in sequence, ii) functions with standardized inputs/outputs of simple features `sf` package class, and iii) an S3 object-oriented implementation of the Bayesian linear regression model.
 These three facts allow for clear articulation of all the steps in the sequence of analysis and easy wrangling and visualization of the geospatial forestry data.
@@
 keywords:
-- forest ecology, competition, tree growth, R, tidyverse, sf, cross-validation, spatial statistics
+- forest ecology, interspecific competition, neigbhorhood competition, tree growth, R, ForestGEO, cross-validation, spatial statistics,
 
@@
 
-BERT: Dave could you write a few sentences (and add references where appropriate) starting with the big picture of modeling tree growth, zooming in to neighhorhood models, and then connecting to the "@allen_permutation_2020 considers the following model" I write below?
-
-Repeat-censused forest plots offer excellent data to test neighborhood models of the effect of competition on the growth of trees @canham_neighborhood_2006 @uriarte_spatially_2004.
+Repeat-censused forest plots offer excellent data to test neighborhood models of the effect of competition on the growth of trees (@canham_neighborhood_2004).
+Neighborhood models of competition have been used to: test whether the species identity of a competitor matters (@uriarte_spatially_2004); measure species-specific competition coefficients (@das_effect_2012 @tatsumi_estimating_2016); test competing models to see what structures competitive interactions, e.g.
+traits or phylogeny (@allen_permutation_2020 @uriarte_trait_2010); and inform selective logging practices (@canham_neighborhood_2006).
+Although these are well-described methods, no methods are currently available for their easy application.
+Here we address this in an R package.
+We largely follow the methods presented in @allen_permutation_2020.
+The package is written to model stem radial growth between two censuses based on neighborhood competition.
+It assumes a single large plot using the ForestGEO plot data structure as a guide (@andersonteixeira_ctfs-forestgeo_2015).
 
@@
 
-Figure \ref{fig:scbi-simulation}
+Figure \ref{fig:scbi-simulation} highlights the importance of spatial cross validating when interpreting results from neighborhood compeition models.
+Without cross-validation the competitor idenity did matter, the non-permuted competitor species had a much lower RMSE than the permuted one.
+But once we include the spatial cross validation this improvement disaprees.
+So in this 9 ha subplot of the SCBI plot competitive interactions do not depend on the identity of the competitor, which is the opposite of what has been observed in other places (@allen_permutation_2020 @uriarte_spatially_2004).
 
@@
 
-# Discussion
+# Conclusion
 
-The `forestecology` package provides
+The `forestecology` package provides an accessible way to fit and test models of neighborhood competition.
+Currently it is written to work with data from ForestGEO plots.
+But it could easily be modified to work on any single large, mapped forest plot in which at least two measurements of each individual have been taken.
+With a little bit of work the package could also be applied to forest inventory data in which several small plots are spatially separated, e.g., USFS Forest Inventory.
+In a future version of the package we also hope to make it possible to model plant mortality in addition to plant growth.
+The package follows the guidelines for `tidy` data, `sf` spatial structure., and S3 open-oriented model structure.
+These make for ease of use.
+We hope that the package will further the use of neighborhood competition models to better understand what structures plant competition.
 
@@
 
-AYK and DNA conceived the ideas and coded most of the package in the early design stages.
-AYK led the writing of the manuscript.
-SPC refactored much of the package's code in the later design stages to align with R and "tidy" programming best practices [@wickham_welcome_2019].
-All authors contributed critically to the drafts and gave final approval for publication.
+AYK and DNA conceived the ideas and coded a draft of the package.
+AYK wrote an initial manuscript draft.
+SPC rewrote much of the package's code to align with R and "tidy" best practices [@wickham_welcome_2019].
+All authors contributed to subsequent drafts and gave final approval for manuscript.
 

### 2021-03-06 paper/forestecology.tex
first batch of edits

@@
 \begin{abstract}
-1. Many models for the growth of trees that incorporate interspecies
-competition are based on a neighborhood effect assumption whereby all
-trees within a fixed distance of a focal tree are considered
-competitors. Methods and tools are needed to quantify this competitive
-effect and assess the quality of all resulting models.\\
-2. We present the \texttt{forestecology} package providing methods for
-both 1) evaluating the effect of competitor species identity using
-permutation tests and 2) evaluating model performance using spatial
-cross-validation. Following \citet{allen_permutation_2020}, we implement
-a Bayesian linear regression competition model.\\
+1. Neighborhood competition models are powerful tools to measure the
+effect of interspecific competition. Statistical methods to ease the
+application of these models are currently lacking.\\
+2. We present the \texttt{forestecology} package providing methods to i)
+specify neighborhood competition models, ii) evalulate the effect of
+competitor species identity using permutation tests, and iii) measure
+model performance using spatial cross-validation. Following
+\citet{allen_permutation_2020}, we implement a Bayesian linear
+regression neighborhood competition model.\\
 3. We demonstrate the package's functionality using data from the
@@
 designed with cross-site compatibility in mind. We demonstrate that both
-1) competitor species identity matters and 2) that not spatially
+i) competitor species identity matters and ii) that not spatially
 cross-validating leads to error estimates that are overly optimistic.\\
-4. The package features 1) \texttt{tidyverse}-like structure whereby
-verb-named functions can be modularly ``piped'' in sequence, 2)
+4. The package features i) \texttt{tidyverse}-like structure whereby
+verb-named functions can be modularly ``piped'' in sequence, ii)
 functions with standardized inputs/outputs of simple features
-\texttt{sf} package class, and 3) an S3 object-oriented implementation
+\texttt{sf} package class, and iii) an S3 object-oriented implementation
 of the Bayesian linear regression model. These three facts allow for
@@
 \noindent%
-{\it Keywords:} forest ecology, competition, tree growth, R, tidyverse, sf,
-cross-validation, spatial statistics
+{\it Keywords:} forest ecology, interspecific competition, neigbhorhood competition,
+tree growth, R, ForestGEO, cross-validation, spatial statistics,
 \vfill
@@
 
-BERT: Dave could you write a few sentences (and add references where
-appropriate) starting with the big picture of modeling tree growth,
-zooming in to neighhorhood models, and then connecting to the
-``\citet{allen_permutation_2020} considers the following model'' I write
-below?
-
 Repeat-censused forest plots offer excellent data to test neighborhood
 models of the effect of competition on the growth of trees
-\citet{canham_neighborhood_2006} \citet{uriarte_spatially_2004}.
+(\citet{canham_neighborhood_2004}). Neighborhood models of competition
+have been used to: test whether the species identity of a competitor
+matters (\citet{uriarte_spatially_2004}); measure species-specific
+competition coefficients (\citet{das_effect_2012}
+\citet{tatsumi_estimating_2016}); test competing models to see what
+structures competitive interactions, e.g.~traits or phylogeny
+(\citet{allen_permutation_2020} \citet{uriarte_trait_2010}); and inform
+selective logging practices (\citet{canham_neighborhood_2006}). Although
+these are well-described methods, no methods are currently available for
+their easy application. Here we address this in an R package. We largely
+follow the methods presented in \citet{allen_permutation_2020}. The
+package is written to model stem radial growth between two censuses
+based on neighborhood competition. It assumes a single large plot using
+the ForestGEO plot data structure as a guide
+(\citet{andersonteixeira_ctfs-forestgeo_2015}).
 
@@
 \CommentTok{## # A tibble: 7,954 x 9}
-\CommentTok{##   stemID sp     dbh1 codes1 status  dbh2 codes2 growth}
-\CommentTok{##    <dbl> <fct> <dbl> <chr>  <chr>  <dbl> <chr>   <dbl>}
-\CommentTok{## 1      4 nysy  13.6  M      A       14.2 M       0.103}
-\CommentTok{## 2      5 havi   8.8  M      A        9.6 M;P     0.150}
-\CommentTok{## 3      6 havi   3.25 NULL   A        4   M       0.140}
-\CommentTok{## 4     77 qual  65.2  M      A       66   M       0.141}
-\CommentTok{## 5     79 tiam  47.7  M      A       46.8 M      -0.161}
-\CommentTok{## # ... with 7,949 more rows, and 1 more variable: geometry <POINT>}
+\CommentTok{##   stemID sp     dbh1 codes1 status  dbh2 codes2 growth   geometry}
+\CommentTok{##    <dbl> <fct> <dbl> <chr>  <chr>  <dbl> <chr>   <dbl>    <POINT>}
+\CommentTok{## 1      4 nysy  13.6  M      A       14.2 M       0.103 (14.2 428)}
+\CommentTok{## 2      5 havi   8.8  M      A        9.6 M;P     0.150  (9.4 436)}
+\CommentTok{## 3      6 havi   3.25 NULL   A        4   M       0.140  (1.3 434)}
+\CommentTok{## 4     77 qual  65.2  M      A       66   M       0.141 (34.7 307)}
+\CommentTok{## 5     79 tiam  47.7  M      A       46.8 M      -0.161   (40 381)}
+\CommentTok{## # ... with 7,949 more rows}
 \end{Highlighting}
@@
 \CommentTok{##   focal_ID focal_sp   dbh foldID                  geometry growth comp }
-\CommentTok{##      <dbl> <fct>    <dbl> <fct>     <POINT>  <dbl> <list>           }
-\CommentTok{## 1        4 nysy     13.6  1      (14.2 428)  0.103 <tibble [20 x 4]>}
-\CommentTok{## 2        5 havi      8.8  1       (9.4 436)  0.150 <tibble [32 x 4]>}
-\CommentTok{## 3       79 tiam     47.7  1        (40 381) -0.161 <tibble [20 x 4]>}
-\CommentTok{## 4       80 caca      5.15 1      (38.7 422)  0.253 <tibble [12 x 4]>}
-\CommentTok{## 5       96 libe      2.3  1        (60 310)  0.262 <tibble [14 x 4]>}
+\CommentTok{##      <dbl> <fct>    <dbl> <fct>                    <POINT>  <dbl> <lis>}
+\CommentTok{## 1        4 nysy     13.6  1                     (14.2 428)  0.103 <tib~}
+\CommentTok{## 2        5 havi      8.8  1                      (9.4 436)  0.150 <tib~}
+\CommentTok{## 3       79 tiam     47.7  1                       (40 381) -0.161 <tib~}
+\CommentTok{## 4       80 caca      5.15 1                     (38.7 422)  0.253 <tib~}
+\CommentTok{## 5       96 libe      2.3  1                       (60 310)  0.262 <tib~}
 \CommentTok{## # ... with 6,291 more rows, and 1 more variable: growth_hat <dbl>}
@@
 
-Figure \ref{fig:scbi-simulation}
+Figure \ref{fig:scbi-simulation} highlights the importance of spatial
+cross validating when interpreting results from neighborhood compeition
+models. Without cross-validation the competitor idenity did matter, the
+non-permuted competitor species had a much lower RMSE than the permuted
+one. But once we include the spatial cross validation this improvement
+disaprees. So in this 9 ha subplot of the SCBI plot competitive
+interactions do not depend on the identity of the competitor, which is
+the opposite of what has been observed in other places
+(\citet{allen_permutation_2020} \citet{uriarte_spatially_2004}).
 
@@
 
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
-The \texttt{forestecology} package provides
+\hypertarget{conclusion}{%
+\section{Conclusion}\label{conclusion}}
+
+The \texttt{forestecology} package provides an accessible way to fit and
+test models of neighborhood competition. Currently it is written to work
+with data from ForestGEO plots. But it could easily be modified to work
+on any single large, mapped forest plot in which at least two
+measurements of each individual have been taken. With a little bit of
+work the package could also be applied to forest inventory data in which
+several small plots are spatially separated, e.g., USFS Forest
+Inventory. In a future version of the package we also hope to make it
+possible to model plant mortality in addition to plant growth. The
+package follows the guidelines for \texttt{tidy} data, \texttt{sf}
+spatial structure., and S3 open-oriented model structure. These make for
+ease of use. We hope that the package will further the use of
+neighborhood competition models to better understand what structures
+plant competition.
 
@@
 
-AYK and DNA conceived the ideas and coded most of the package in the
-early design stages. AYK led the writing of the manuscript. SPC
-refactored much of the package's code in the later design stages to
-align with R and ``tidy'' programming best practices
-\citep{wickham_welcome_2019}. All authors contributed critically to the
-drafts and gave final approval for publication.
+AYK and DNA conceived the ideas and coded a draft of the package. AYK
+wrote an initial manuscript draft. SPC rewrote much of the package's
+code to align with R and ``tidy'' best practices
+\citep{wickham_welcome_2019}. All authors contributed to subsequent
+drafts and gave final approval for manuscript.
 

### 2021-03-08 paper/forestecology.Rmd
My last edits

@@
 ---
-title: The forestecology R package for fitting and assessing models of interspecies competitive effects on the growth of trees
+title: The forestecology R package for fitting and assessing nieghborhood models of the effect of interspecific competition on the growth of trees
 
@@
 Given ForestGEO's data collection protocols and data formatting standards, the package was designed with cross-site compatibility in mind.
-We demonstrate that both i) competitor species identity matters and ii) that not spatially cross-validating leads to error estimates that are overly optimistic.
+We highlight the importance of spatial cross validation when interpreting model results.
   | 4.
@@
 keywords:
-- forest ecology, interspecific competition, neigbhorhood competition, tree growth, R, ForestGEO, cross-validation, spatial statistics,
+- forest ecology, interspecific competition, neigbhorhood competition, tree growth, R, ForestGEO, spatial cross validation
 
@@
 
-Repeat-censused forest plots offer excellent data to test neighborhood models of the effect of competition on the growth of trees (@canham_neighborhood_2004).
+Repeat-censused forest plots offer excellent opportunities to test neighborhood models of the effect of competition on the growth of trees (@canham_neighborhood_2004).
 Neighborhood models of competition have been used to: test whether the species identity of a competitor matters (@uriarte_spatially_2004); measure species-specific competition coefficients (@das_effect_2012 @tatsumi_estimating_2016); test competing models to see what structures competitive interactions, e.g.
-traits or phylogeny (@allen_permutation_2020 @uriarte_trait_2010); and inform selective logging practices (@canham_neighborhood_2006).
+traits or phylogeny (@allen_permutation_2020; @uriarte_trait_2010); and inform selective logging practices (@canham_neighborhood_2006).
 Although these are well-described methods, no methods are currently available for their easy application.
@@
 The package is written to model stem radial growth between two censuses based on neighborhood competition.
-It assumes a single large plot using the ForestGEO plot data structure as a guide (@andersonteixeira_ctfs-forestgeo_2015).
 
@@
 The resulting RMSE of `r model_rmse_shuffle` based on the permutation test is `r ifelse(model_rmse_shuffle < model_rmse, "smaller", "larger")` than the earlier RMSE of `r model_rmse`, suggesting that models that do incorporate competitor species identity better fit the data.
-We conduct an analysis of the full SCBI plot in Section \ref{results} below.
 
@@
 The resulting RMSE of `r model_rmse_cv` computed using cross-validation is `r ifelse(model_rmse_cv > model_rmse, "larger", "smaller")` than the earlier RMSE of `r model_rmse`, suggesting that models that do not take the inherent spatial autocorrelation of the data into account generate error estimates that are overly optimistic; in our case RMSE's that are too low.
-We conduct an analysis of the full SCBI plot in Section \ref{results} below.
 
-# Results {#results}
+# Importance of spatial cross validation
 
-Figure \ref{fig:scbi-simulation} highlights the importance of spatial cross validating when interpreting results from neighborhood compeition models.
-Without cross-validation the competitor idenity did matter, the non-permuted competitor species had a much lower RMSE than the permuted one.
-But once we include the spatial cross validation this improvement disaprees.
+The `run_cv` function also accepts the `run_shuffle` argument.
+This permutes the competitor species, as described above, but does so when calculating predicted growth with the cross validated scheme.
+Figure \ref{fig:scbi-simulation} compares model performance when permuting competitor species and calculating RMSE with and without cross validation.
+Without cross-validation the competitor identity did matter, the non-permuted competitor species had a much lower RMSE than the permuted one.
+But once we include the spatial cross validation this improvement disappears.
 So in this 9 ha subplot of the SCBI plot competitive interactions do not depend on the identity of the competitor, which is the opposite of what has been observed in other places (@allen_permutation_2020 @uriarte_spatially_2004).
+This highlights the importance of cross validation, without it the model was overfit.
 
-```{r scbi-simulation, out.width="100%", fig.cap="Root mean squared error of models for standard, permuted, and spatial cross-validated error estimates.", echo = FALSE}
+```{r scbi-simulation, out.width="100%", fig.cap="Root mean squared error of models for standard, permuted, and spatial cross-validated error estimates.
+The dotted lines show non-permuted competitor identity, while the histogras so the RMSE for 49 permutations.
+The colors indicate whether cross validaton was used.", echo = FALSE}
 include_graphics("simulation_results/2021-03-03_scbi_49_shuffles.png")
@@
 In a future version of the package we also hope to make it possible to model plant mortality in addition to plant growth.
-The package follows the guidelines for `tidy` data, `sf` spatial structure., and S3 open-oriented model structure.
-These make for ease of use.
-We hope that the package will further the use of neighborhood competition models to better understand what structures plant competition.
+The package follows the guidelines for `tidy` data, `sf` spatial structure, and S3 open-oriented model structure.
+We hope that the package will increase the use of neighborhood competition models to better understand what structures plant competition.
 

### 2021-03-08 paper/forestecology.tex
My last edits

@@
 {
-  \title{\bf The forestecology R package for fitting and assessing models of
-interspecies competitive effects on the growth of trees}
+  \title{\bf The forestecology R package for fitting and assessing nieghborhood
+models of the effect of interspecific competition on the growth of trees}
 
@@
   \begin{center}
-    {\LARGE\bf The forestecology R package for fitting and assessing models of
-interspecies competitive effects on the growth of trees}
+    {\LARGE\bf The forestecology R package for fitting and assessing nieghborhood
+models of the effect of interspecific competition on the growth of trees}
   \end{center}
@@
 data collection protocols and data formatting standards, the package was
-designed with cross-site compatibility in mind. We demonstrate that both
-i) competitor species identity matters and ii) that not spatially
-cross-validating leads to error estimates that are overly optimistic.\\
+designed with cross-site compatibility in mind. We highlight the
+importance of spatial cross validation when interpreting model
+results.\\
 4. The package features i) \texttt{tidyverse}-like structure whereby
@@
 {\it Keywords:} forest ecology, interspecific competition, neigbhorhood competition,
-tree growth, R, ForestGEO, cross-validation, spatial statistics,
+tree growth, R, ForestGEO, spatial cross validation
 \vfill
@@
 
-Repeat-censused forest plots offer excellent data to test neighborhood
-models of the effect of competition on the growth of trees
+Repeat-censused forest plots offer excellent opportunities to test
+neighborhood models of the effect of competition on the growth of trees
 (\citet{canham_neighborhood_2004}). Neighborhood models of competition
@@
 structures competitive interactions, e.g.~traits or phylogeny
-(\citet{allen_permutation_2020} \citet{uriarte_trait_2010}); and inform
+(\citet{allen_permutation_2020}; \citet{uriarte_trait_2010}); and inform
 selective logging practices (\citet{canham_neighborhood_2006}). Although
@@
 package is written to model stem radial growth between two censuses
-based on neighborhood competition. It assumes a single large plot using
-the ForestGEO plot data structure as a guide
-(\citet{andersonteixeira_ctfs-forestgeo_2015}).
+based on neighborhood competition.
 
@@
 the earlier RMSE of 0.128, suggesting that models that do incorporate
-competitor species identity better fit the data. We conduct an analysis
-of the full SCBI plot in Section \ref{results} below.
+competitor species identity better fit the data.
 
@@
 error estimates that are overly optimistic; in our case RMSE's that are
-too low. We conduct an analysis of the full SCBI plot in Section
-\ref{results} below.
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-Figure \ref{fig:scbi-simulation} highlights the importance of spatial
-cross validating when interpreting results from neighborhood compeition
-models. Without cross-validation the competitor idenity did matter, the
-non-permuted competitor species had a much lower RMSE than the permuted
-one. But once we include the spatial cross validation this improvement
-disaprees. So in this 9 ha subplot of the SCBI plot competitive
-interactions do not depend on the identity of the competitor, which is
-the opposite of what has been observed in other places
-(\citet{allen_permutation_2020} \citet{uriarte_spatially_2004}).
+too low.
+
+\hypertarget{importance-of-spatial-cross-validation}{%
+\section{Importance of spatial cross
+validation}\label{importance-of-spatial-cross-validation}}
+
+The \texttt{run\_cv} function also accepts the \texttt{run\_shuffle}
+argument. This permutes the competitor species, as described above, but
+does so when calculating predicted growth with the cross validated
+scheme. Figure \ref{fig:scbi-simulation} compares model performance when
+permuting competitor species and calculating RMSE with and without cross
+validation. Without cross-validation the competitor identity did matter,
+the non-permuted competitor species had a much lower RMSE than the
+permuted one. But once we include the spatial cross validation this
+improvement disappears. So in this 9 ha subplot of the SCBI plot
+competitive interactions do not depend on the identity of the
+competitor, which is the opposite of what has been observed in other
+places (\citet{allen_permutation_2020} \citet{uriarte_spatially_2004}).
+This highlights the importance of cross validation, without it the model
+was overfit.
 
@@
 
-\caption{Root mean squared error of models for standard, permuted, and spatial cross-validated error estimates.}\label{fig:scbi-simulation}
+\caption{Root mean squared error of models for standard, permuted, and spatial cross-validated error estimates.
+The dotted lines show non-permuted competitor identity, while the histogras so the RMSE for 49 permutations.
+The colors indicate whether cross validaton was used.}\label{fig:scbi-simulation}
 \end{figure}
@@
 package follows the guidelines for \texttt{tidy} data, \texttt{sf}
-spatial structure., and S3 open-oriented model structure. These make for
-ease of use. We hope that the package will further the use of
-neighborhood competition models to better understand what structures
-plant competition.
+spatial structure, and S3 open-oriented model structure. We hope that
+the package will increase the use of neighborhood competition models to
+better understand what structures plant competition.
 

### 2021-03-08 paper/forestecology.Rmd
typos

@@
 ---
-title: The forestecology R package for fitting and assessing nieghborhood models of the effect of interspecific competition on the growth of trees
+title: The forestecology R package for fitting and assessing neighborhood models of the effect of interspecific competition on the growth of trees
 
@@
 Furthermore, in order to account for the spatial autocorrelation inherent to forest data in their estimates of out-of-sample model error, @allen_permutation_2020 use spatial cross-validation.
-Estimates of model error that do not account for this spatial dependency tend to underestiamte the true model error [@roberts_cross-validation_2017].
+Estimates of model error that do not account for this spatial dependency tend to underestimate the true model error [@roberts_cross-validation_2017].
 

### 2021-03-08 paper/forestecology.tex
typos

@@
 {
-  \title{\bf The forestecology R package for fitting and assessing nieghborhood
+  \title{\bf The forestecology R package for fitting and assessing neighborhood
 models of the effect of interspecific competition on the growth of trees}
@@
   \begin{center}
-    {\LARGE\bf The forestecology R package for fitting and assessing nieghborhood
+    {\LARGE\bf The forestecology R package for fitting and assessing neighborhood
 models of the effect of interspecific competition on the growth of trees}
@@
 cross-validation. Estimates of model error that do not account for this
-spatial dependency tend to underestiamte the true model error
+spatial dependency tend to underestimate the true model error
 \citep{roberts_cross-validation_2017}.
@@
 \CommentTok{##   focal_ID focal_sp   dbh foldID   geometry growth comp             }
-\CommentTok{##      <dbl> <fct>    <dbl> <fct>                    <POINT>  <dbl> <lis>}
-\CommentTok{## 1        4 nysy     13.6  1                     (14.2 428)  0.103 <tib~}
-\CommentTok{## 2        5 havi      8.8  1                      (9.4 436)  0.150 <tib~}
-\CommentTok{## 3       79 tiam     47.7  1                       (40 381) -0.161 <tib~}
-\CommentTok{## 4       80 caca      5.15 1                     (38.7 422)  0.253 <tib~}
-\CommentTok{## 5       96 libe      2.3  1                       (60 310)  0.262 <tib~}
+\CommentTok{##      <dbl> <fct>    <dbl> <fct>     <POINT>  <dbl> <list>           }
+\CommentTok{## 1        4 nysy     13.6  1      (14.2 428)  0.103 <tibble [20 x 4]>}
+\CommentTok{## 2        5 havi      8.8  1       (9.4 436)  0.150 <tibble [32 x 4]>}
+\CommentTok{## 3       79 tiam     47.7  1        (40 381) -0.161 <tibble [20 x 4]>}
+\CommentTok{## 4       80 caca      5.15 1      (38.7 422)  0.253 <tibble [12 x 4]>}
+\CommentTok{## 5       96 libe      2.3  1        (60 310)  0.262 <tibble [14 x 4]>}
 \CommentTok{## # ... with 6,291 more rows, and 1 more variable: growth_hat <dbl>}

### 2021-03-09 paper/forestecology.Rmd
4210 -> 4069

@@
 
-## Step 1: Compute the growth of trees based on census data {compute-growth}
+## Step 1: Compute the growth of trees based on census data {#compute-growth}
 
 The first step in the our analysis sequence is to compute the growth of trees using data from two censuses.
-The `compute_growth()` function computes average annual growth assuming census data that roughly follows ForestGEO standards.
+`compute_growth()` computes average annual growth assuming census data that roughly follows ForestGEO standards.
 Despite such standards, minor variations will still exist between sites, thereby necessitating some data wrangling and checking.
@@
 We load both 2008 and 2014 SCBI census data `.csv` files as they existed on GitHub on November 20, 2020 [@gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020].
-After selecting only the relevant variables, we perform a few additional data wrangling steps: convert the character variable with the date of measurement to be of explicit type `date`, convert DBH to be in centimeters^[A rule of thumb to determine the units of DBH is to check if the smallest non-zero and non-missing measurement is 1 or 10.
-If the former, then centimeters.
-If the later, then millimeters.
-This is because ForestGEO protocols state that only trees with DBH greater or equal to 1cm should be included in censuses.] Furthermore, in order to speed up computation for purposes of this example, we only consider a 9 ha subsection of the 25.6 ha of the SCBI site: `gx` from 0--300 instead of 0--400 and `gy` from 300--600 instead of 0--640.
+After selecting only the relevant variables, we perform a few additional data wrangling steps: convert the character variable with the date of measurement to be of explicit type `date`, convert DBH to be in centimeters.
+Furthermore, in order to speed up computation for purposes of this example, we only consider a 9 ha subsection of the 25.6 ha of the SCBI site: `gx` from 0--300 instead of 0--400 and `gy` from 300--600 instead of 0--640.
 
@@
 
-These two data frames are then supplied as arguments to the `compute_growth()` function, along with the `id` argument that specifies the variable that uniquely identifies each tree-stem.
+These two data frames are then supplied as arguments to `compute_growth()`, along with the `id` argument that specifies the variable that uniquely identifies each tree-stem.
 Note furthermore that we discard all resprouts in the later census (those with `code == R`), since we are only interested in the diameter growth of surviving, and not resprouted, stems.
@@
 
-Site data that does not align with this convention will need to be transformed for use with the `compute_growth()` function.
+Site data that does not align with this convention will need to be transformed for use with `compute_growth()`.
 However, in the end, all that matters is that the growth of all stems is saved in a data frame of class `sf` and, at a minimum, contains the variable uniquely identifying each stem, `sp`, `dbh1`, `growth`, `geometry`.
@@
 
-The next step in our analysis sequence is to add additional spatial information to our main `growth_scbi` data frame.
+The next step in our analysis sequence is to add additional spatial information to `growth_scbi`.
 The first element of spatial information we add is a "buffer region" to the periphery of the study region.
@@
 In our case we set this value at 7.5m, a value informed by other studies [@canham_neighborhood_2004, @uriarte_spatially_2004, @canham_neighborhood_2006].
-Using this value along with a manually constructed `sf` object representation of the study region's boundary via its vertices, we apply the `add_buffer_variable()` to our `growth_scbi` data frame to add a `buffer` boolean variable: all trees who have `buffer` set to `FALSE` will be our focal trees whose growth will be modeled, whereas those with `TRUE` will only be considered as competitor trees whose growth will not.
+Using this value along with a manually constructed `sf` object representation of the study region's boundary via its vertices, we apply the `add_buffer_variable()` to `growth_scbi` to add a `buffer` boolean variable: all trees who have `buffer` set to `FALSE` will be our focal trees whose growth will be modeled, whereas those with `TRUE` will only be considered as competitor trees whose growth will not.
 
@@
 We then use the output of the `spatialBlock()` function from the `blockCV` package to associate each tree in `growth_scbi` to the correct fold (saved in the `foldID` variable) [@valavi_blockcv_2019].
-^[In the Supporting Information we present an example where the folds themselves are also created using the `spatialBlock()` function given a specified `cv_block_size`.]
+^[In the Supporting Information we present an example where the folds themselves are also created using `spatialBlock()` given a specified `cv_block_size`.]
 
@@
 
-Figure \ref{fig:scbi-spatial-information} illustrates the net effect of adding these two elements of information to the `growth_scbi` data frame.
+Figure \ref{fig:scbi-spatial-information} illustrates the net effect of adding these two elements of information to `growth_scbi`.
 The location of each tree is marked with an integer indicating which fold it belongs to, where the folds are marked with solid lines.
@@
 On top of the previous arguments `comp_dist` defining the competition neighborhood and `id` indicating which variable uniquely identifies each tree-stem, this function also requires an `sf` object representation of the spatial cross-validation blocks/folds.
-In this example, the blocks were manually encoded in `blocks_scbi` by specifying it's vertices in Section \ref{spatial-information}^[We present an alternative method for defining spatial cross-validation blocks is using the `spatialBlock()` function from the `blockCV` package in the Supporting Information.].
+In this example, the blocks were manually encoded in `blocks_scbi` by specifying it's vertices in Section \ref{spatial-information}.
 We present the resulting data frame below with the `foldID` variable omitted for compactness of presentation.
@@
 
-The resulting data frame `focal_vs_comp_scbi` has `r nrow(focal_vs_comp_scbi)` rows, representing the subset of the `r nrow(growth_scbi)` trees in `growth_scbi` that will be considered as focal trees.
+The resulting `focal_vs_comp_scbi` has `r nrow(focal_vs_comp_scbi)` rows, representing the subset of the `r nrow(growth_scbi)` trees in `growth_scbi` that will be considered as focal trees.
 Two new variables `focal_ID` and `focal_sp` relate to tree-stem identification and species information.
@@
 
-Using the `unnest()` function from the `tidyr` package, we can flatten list-column into regular columns.
+Using `unnest()` from the `tidyr` package, we can flatten list-column into regular columns.
 We observe that for the same focal tree, we have information on all 20 competitor trees whose `dist` distance to the focal tree is $\leq$ `r comp_dist`: their unique tree-stem ID number, their species, and their basal area (in m$^2$) calculated as $\frac{\pi \times (\text{DBH/2})^2}{10000}$ where $DBH$ is the value from the earlier of the two censuses in cm.
@@
 
-Now that we've identified all focal and corresponding competitor trees and saved this information in a data frame of type `focal_vs_comp`, the final step in our analysis sequence is to fit a model for the growth of all focal trees.
-Currently the `forestecology` package can only fit the competition Bayesian linear regression model outlined in Section \ref{competition-model} using the `comp_bayes_lm()` function.
-However, any model implemented in a function that similarly takes an input data frame of type `focal_vs_comp` as an argument can also be used.
+The final step is to fit a model for the growth of all focal trees.
+Currently the `forestecology` package can only fit the competition Bayesian linear regression model outlined in Section \ref{competition-model} using `comp_bayes_lm()`.
+However, any model implemented in a function that similarly takes an input data frame similar to `focal_vs_comp` as an argument can also be used.
 For our specific competition Bayesian linear regression model, we also specify prior distributions on all parameters of interest (here chosen to be the defaults as specified in `?comp_bayes_lm`).
@@
 The resampling with replacement is nested within each focal tree in order to preserve the neighborhood structure of our competition model.
-To run the permutation test, we use the same `comp_bayes_lm()` function as in Section \ref{model-fit-predict}, but with a `run_shuffle = TRUE` argument.
+To run the permutation test, we use `comp_bayes_lm()` as in Section \ref{model-fit-predict}, but with a `run_shuffle = TRUE` argument.
 
@@
 
-This algorithm is implemented in the `run_cv()` function, which is a wrapper function to the `comp_bayes_lm()` function that fits the model and the `predict()` generic that returns fitted/predicted values.
+This algorithm is implemented in `run_cv()`, which is a wrapper function to `comp_bayes_lm()` that fits the model and the `predict()` generic that returns fitted/predicted values.
 We compare these values to the observed growth values to again compute our RMSE.
@@
 
-The `run_cv` function also accepts the `run_shuffle` argument.
+`run_cv()` also accepts the `run_shuffle` argument.
 This permutes the competitor species, as described above, but does so when calculating predicted growth with the cross validated scheme.

### 2021-03-09 paper/forestecology.Rmd
4069 -> 3712

@@
 Given ForestGEO's data collection protocols and data formatting standards, the package was designed with cross-site compatibility in mind.
-We highlight the importance of spatial cross validation when interpreting model results.
+We highlight the importance of spatial cross-validation when interpreting model results.
   | 4.
@@
 keywords:
-- forest ecology, interspecific competition, neigbhorhood competition, tree growth, R, ForestGEO, spatial cross validation
+- forest ecology, interspecific competition, neigbhorhood competition, tree growth, R, ForestGEO, spatial cross-validation
 
@@
 # Internally used packages
+library(tidyverse)
+library(conflicted)
 library(viridis)
 library(knitr)
-library(ggplot2)
 library(here)
-library(tibble)
 
@@
 )
+
+# Resolve conflicts in function names
+filter <- dplyr::filter
+select <- dplyr::select
 ```
@@
 traits or phylogeny (@allen_permutation_2020; @uriarte_trait_2010); and inform selective logging practices (@canham_neighborhood_2006).
-Although these are well-described methods, no methods are currently available for their easy application.
+Although these are well-described methods, few methods are currently available for easy application.
 Here we address this in an R package.
@@
 @allen_permutation_2020 considers the following model: Let $i = 1, \ldots, n_j$ index all $n_j$ trees of "focal" species group $j$; let $j = 1, \ldots, J$ index all $J$ focal species groups; and let $k = 1, \ldots, K$ index all $K$ "competitor" species groups.
-We model the average annual growth in diameter $y_{ij}$ (in centimeters per year) of the $i^{th}$ tree of focal species group $j$ as a linear model $f$ of the covariates $\vec{x}_{ij}$
+We model the average annual growth in diameter at breast height (DBH) $y_{ij}$ (in centimeters per year) of the $i^{th}$ tree of focal species group $j$ as a linear model $f$ of the covariates $\vec{x}_{ij}$
 
@@
 
-where $\beta_{0,j}$ is the diameter-independent growth rate for group $j$; $\text{dbh}_{ij}$ is the diameter at breast height (in centimeters) of the focal tree at the earlier census; $\beta_{\text{dbh},j}$ is the amount of the growth rate changed depending on diameter for group $j$;  $\text{BA}_{ijk}$ is the sum of the basal area of all trees of competitor species group $k$; $\lambda_{jk}$ is the change in growth for individuals of group $j$ from nearby competitors of group $k$; and $\epsilon_{ij}$ is a random error term distributed $\text{Normal}(0, \sigma^2)$.
-They estimate all parameters via Bayesian linear regression while exploiting Normal/Inverse Gamma conjugacy to derive closed-form solutions to all posterior distributions via linear algebra^[See S1 Appendix of @allen_permutation_2020, available at https://doi.org/10.1371/journal.pone.0229930.s004].
+where $\beta_{0,j}$ is the diameter-independent growth rate for group $j$; $\text{dbh}_{ij}$ is the DBH the focal tree at the earlier census; $\beta_{\text{dbh},j}$ is the amount of the growth rate changed depending on diameter for group $j$;  $\text{BA}_{ijk}$ is the sum of the basal area of all trees of competitor species group $k$; $\lambda_{jk}$ is the change in growth for individuals of group $j$ from nearby competitors of group $k$; and $\epsilon_{ij}$ is a random error term distributed $\text{Normal}(0, \sigma^2)$.
+They estimate all parameters via Bayesian linear regression while exploiting Normal/Inverse Gamma conjugacy to derive closed-form solutions to all posterior distributions^[See S1 Appendix of @allen_permutation_2020, available at https://doi.org/10.1371/journal.pone.0229930.s004].
 These closed-form solutions for the posterior distributions are in contrast to approximations of all posteriors via computationally expensive Markov Chain Monte Carlo algorithms.
@@
 
-We introduce the `forestecology` R package providing methods and data for forest ecology model fitting and assessment, available on CRAN (https://cran.r-project.org/web/packages/forestecology/index.html) with the corresponding source code available on GitHub (https://github.com/rudeboybert/forestecology).
+We introduce the `forestecology` R package providing methods and data for forest ecology model fitting and assessment, available on CRAN (https://cran.r-project.org/web/packages/forestecology/index.html) and on GitHub (https://github.com/rudeboybert/forestecology).
 The package implements all aspects of the model in Equation \ref{eq:model}: model fitting and generating fitted/predicted values, evaluating the effect of competitor species identity using permutation tests, and evaluating model performance using spatial cross-validation.
@@
 As we demonstrate in Section \ref{model-fit-predict} however, the fitting of this model is self-contained in a single function `comp_bayes_lm()`.
-This function returns an object of S3 class type `comp_bayes_lm` with generic methods implemented for `print()` to inspect the output, `predict()` to generate fitted/predicted values, and `ggplot2::autoplot()` to visualize all results.
-Therefore the package can be modularly extended to fit other models as long as they are coded into a function similar type as `comp_bayes_lm()` as has equivalent generic methods implemented.
+This function returns an object of S3 class type `comp_bayes_lm` with generic methods implemented to print, make predictions using, and plot all results.
+Therefore the package can be modularly extended to fit other models as long as they are coded into a function similar to `comp_bayes_lm()` and has equivalent generic methods implemented.
 
@@
 The 25.6 ha (640 x 400 m) plot is located at the intersection of three of the major physiographic provinces of the eastern US---the Blue Ridge, Ridge and Valley, and Piedmont provinces---and is adjacent to the northern end of Shenandoah National Park.
-The forest type is typical mature secondary eastern mixed deciduous forest, with a canopy dominated by tulip poplar (*Liriodendron tulipifera*), oaks (*Quercus* spp.), and hickories (*Carya* spp.), and an understory composed mainly of spicebush (*Lindera benzoin*), paw-paw (*Asimina triloba*), American hornbeam (*Carpinus caroliniana*), and witch hazel (*Hamamelis virginiana*) [@bourg_initial_2013].
 
@@
 
-The first step in the our analysis sequence is to compute the growth of trees using data from two censuses.
+The first step is to compute the growth of trees using data from two censuses.
 `compute_growth()` computes average annual growth assuming census data that roughly follows ForestGEO standards.
@@
 
-We load both 2008 and 2014 SCBI census data `.csv` files as they existed on GitHub on November 20, 2020 [@gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020].
-After selecting only the relevant variables, we perform a few additional data wrangling steps: convert the character variable with the date of measurement to be of explicit type `date`, convert DBH to be in centimeters.
-Furthermore, in order to speed up computation for purposes of this example, we only consider a 9 ha subsection of the 25.6 ha of the SCBI site: `gx` from 0--300 instead of 0--400 and `gy` from 300--600 instead of 0--640.
+We first load both 2008 and 2014 SCBI census data `.csv` files as they existed on GitHub on November 20, 2020 [@gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020] and perform some data wrangling to both data sets.
+We then only consider a 9 ha subsection of the 25.6 ha of the SCBI site, `gx` from 0--300 instead of 0--400 and `gy` from 300--600 instead of 0--640, in order to speed up computation for purposes of this example.
 
@@
   mutate(
+    # Convert date from character to date
     date = mdy(date),
+    # Convert dbh to be in cm
     dbh = as.numeric(dbh)/10
@@
 
-The output `growth_scbi` is a single data frame of class `sf` that includes variables `growth`, the average annual growth in DBH (cm y^-1^) for all stems that were alive at both time points, and `geometry`, the `sf` package's encoding of geolocations of type `<POINT>`.
-In addition the species variable `sp` is converted to a factor if it wasn't already by `compute_growth()`.^[In our spatial cross-validation algorithm in Section \ref{spatial-cross-validation} issues can occur when rare species do not occur in the training set, but then are encountered in the test set.
-This risk is mitigated by representing `sp` as a factor variable, which has a complete list of all levels of the categorical variable.] Furthermore, the variables that should remain unchanged between censuses appear only once, such as location variables `gx` and `gy`; as well as species-related variables.
-Variables that should change between censuses are suffixed with `1` and `2` indicating the earlier and later censuses, such as `dbh1/dbh2` and `codes1/codes2`.
-
-Site data that does not align with this convention will need to be transformed for use with `compute_growth()`.
-However, in the end, all that matters is that the growth of all stems is saved in a data frame of class `sf` and, at a minimum, contains the variable uniquely identifying each stem, `sp`, `dbh1`, `growth`, `geometry`.
+The output `growth_scbi` is a single data frame of class `sf` that includes variables `growth`, the average annual growth in DBH (cm $\cdot$ y^-1^) for all stems that were alive at both time points, and `geometry`, the `sf` package's encoding of geolocations of type `<POINT>`.
+In addition the species variable `sp` is returned as a factor.^[In our spatial cross-validation algorithm in Section \ref{spatial-cross-validation} issues can occur when rare species do not occur in the training set, but then are encountered in the test set.
+This risk is mitigated by representing `sp` as a factor variable, which has a complete list of all levels of the categorical variable.]
 
@@
 
-The next step in our analysis sequence is to add additional spatial information to `growth_scbi`.
-The first element of spatial information we add is a "buffer region" to the periphery of the study region.
-Since some of our model's explanatory variables are cumulative (such as competitor basal area), we must ensure that all trees being modeled are not biased to have different neighbor structures.
+The next step is to add additional spatial information to `growth_scbi`.
+The first element we add is a "buffer region" to the periphery of the study region.
+Since some of our model's explanatory variables are cumulative, we must ensure that all trees being modeled are not biased to have different neighbor structures.
 This is of concern for trees at the boundary of study regions, for which all neighbors will not be included in the censused stems.
@@
 
-Our model of interspecific competition relies on a spatial definition of who the competitor trees are for focal trees of interest: all trees within a distance `comp_dist` of a focal tree are considered its competitors (assuming the same units as the `gx` and `gy` location variables).
+Our model of interspecific competition relies on a spatial definition of who the competitor trees are for focal trees of interest: all trees within a distance `comp_dist` of a focal tree are considered its competitors.
 In our case we set this value at 7.5m, a value informed by other studies [@canham_neighborhood_2004, @uriarte_spatially_2004, @canham_neighborhood_2006].
-Using this value along with a manually constructed `sf` object representation of the study region's boundary via its vertices, we apply the `add_buffer_variable()` to `growth_scbi` to add a `buffer` boolean variable: all trees who have `buffer` set to `FALSE` will be our focal trees whose growth will be modeled, whereas those with `TRUE` will only be considered as competitor trees whose growth will not.
+Using this value along with a manually constructed `sf` object representation of the study region's boundary, we apply the `add_buffer_variable()` to `growth_scbi` to add a `buffer` boolean variable.
+All trees with `buffer` as `FALSE` will be our focal trees whose growth will be modeled, whereas those with `TRUE` will only be considered as competitor trees.
 
 ```{r}
-# Define buffer region using competitive distance range
+# Define competitive distance range
 comp_dist <- 7.5
 
+# Manually construct study region boundary
 study_region_scbi <- tibble(
@@
 Conventional cross-validation algorithms assign observations to folds by randomly resampling individual observations.
-However, many of these algorithms assume that the observations are independent of each other.
+However, many of these algorithms assume that the observations are independent.
 In the case of forest census data, observations exhibit spatial autocorrelation.
-We therefore incorporate this spatial dependence into the cross-validation algorithm with our spatial blocks of trees [@roberts_cross-validation_2017 @pohjankukka_estimating_2017]
+We therefore incorporate this spatial dependence into the cross-validation algorithm with our spatial blocks of trees [@roberts_cross-validation_2017 @pohjankukka_estimating_2017].
+
 In the example below, we first manually define four folds that partition the study region as an `sf` object.
-We then use the output of the `spatialBlock()` function from the `blockCV` package to associate each tree in `growth_scbi` to the correct fold (saved in the `foldID` variable) [@valavi_blockcv_2019].
-^[In the Supporting Information we present an example where the folds themselves are also created using `spatialBlock()` given a specified `cv_block_size`.]
+We then use the output of the `spatialBlock()` function from the `blockCV` package to associate each tree in `growth_scbi` to the correct fold `foldID` [@valavi_blockcv_2019].
+^[In the Supporting Information we present an example where the folds themselves are also created automatically using `spatialBlock()` given a specified `cv_block_size`, as opposed to manually as in the example.]
 
@@
 
-Figure \ref{fig:scbi-spatial-information} illustrates the net effect of adding these two elements of information to `growth_scbi`.
-The location of each tree is marked with an integer indicating which fold it belongs to, where the folds are marked with solid lines.
+Figure \ref{fig:scbi-spatial-information} illustrates the net effect of adding these two elements of spatial information to `growth_scbi`.
+The location of each tree is marked with an integer indicating its fold, where the folds are marked with solid lines.
 The color of each digit indicates whether the tree is part of the buffer region (and thus will only be considered as a competitor tree in our model) or is part of the interior of the study region (and thus is a focal tree whose growth is of modeled interest).
@@
 
-The next step in our analysis sequence is to identify all focal trees and their corresponding competitor trees.
-More specifically, it identifies all trees that are not part of the buffer region, have a valid `growth` measurement, and have at least one neighbor within `r comp_dist` of it.
-The `create_focal_vs_comp()` functions performs these tasks and returns a new data frame of type `sf`.
-On top of the previous arguments `comp_dist` defining the competition neighborhood and `id` indicating which variable uniquely identifies each tree-stem, this function also requires an `sf` object representation of the spatial cross-validation blocks/folds.
-In this example, the blocks were manually encoded in `blocks_scbi` by specifying it's vertices in Section \ref{spatial-information}.
-We present the resulting data frame below with the `foldID` variable omitted for compactness of presentation.
+The next step is to identify all focal trees and their corresponding competitor trees.
+More specifically, identify all trees that are not part of the buffer region, have a valid `growth` measurement, and have at least one neighbor within `r comp_dist`m.
+`create_focal_vs_comp()` returns a new data frame of type `sf`.
+On top of previously detailed arguments `comp_dist` and `id`, `create_focal_vs_comp()` also requires an `sf` object representation of the spatial cross-validation blocks/folds as seen in Section \ref{spatial-information}.
+We present the resulting data frame below with the `foldID` variable omitted for compactness.
 
@@
 Two new variables `focal_ID` and `focal_sp` relate to tree-stem identification and species information.
-Most notably however is a new variable `comp` which contains information on all competitor trees for a given focal tree, saved in `tidyr` package list-column format [@tidyr_package].
-For example, we drill-down on the tree with `focal_ID` 4, which has 20 competitor trees each described by 4 variables as indicated by the fact that `comp` is a `tibble [20 × 4]`.
+Most notably however is a new variable `comp` which contains information on all competitor trees saved in `tidyr` package list-column format [@tidyr_package].
+We flatten the `comp` list-column for the tree with `focal_ID` 4 in the first row, here a `tibble [20 × 4]`, into regular columns using `unnest()` from the `tidyr` package.
 
-```{r scbi-focal-vs-comp-2}
+```{r, scbi-focal-vs-comp-3}
 focal_vs_comp_scbi %>% 
   filter(focal_ID == 4) %>% 
-  select(focal_ID, dbh, comp)
+  select(focal_ID, dbh, comp) %>% 
+  unnest(cols = "comp")
 ```
 
+We observe that for this focal tree, we have 4 variables of information on its 20 competitor trees: their unique tree-stem ID number, their distance to the focal tree (all $\leq$ `r comp_dist`), their species, and their basal area (in m$^2$) calculated as $\frac{\pi \times (\text{DBH/2})^2}{10000}$ where $DBH$ is the value from the earlier of the two censuses in cm.
+Saving our focal versus competitor information in list-column minimizes redundancy since we do not repeat information on the focal tree 20 times.
 The spatial distribution of these trees is visualized in Figure \ref{fig:scbi-focal-vs-comp-map}: the dashed circle extends `r comp_dist` m away from the focal tree while all 20 competitor trees are within this circle.
@@
 
-Using `unnest()` from the `tidyr` package, we can flatten list-column into regular columns.
-We observe that for the same focal tree, we have information on all 20 competitor trees whose `dist` distance to the focal tree is $\leq$ `r comp_dist`: their unique tree-stem ID number, their species, and their basal area (in m$^2$) calculated as $\frac{\pi \times (\text{DBH/2})^2}{10000}$ where $DBH$ is the value from the earlier of the two censuses in cm.
-Saving our focal versus competitor information in list-column minimizes redundancy since we do not repeat information on the focal tree 20 times.
-
-```{r, scbi-focal-vs-comp-3}
-focal_vs_comp_scbi %>% 
-  filter(focal_ID == 4) %>% 
-  select(focal_ID, dbh, comp) %>% 
-  unnest(cols = "comp")
-```
-
 ## Step 4: Fit model {#model-fit-predict}
@@
 The final step is to fit a model for the growth of all focal trees.
-Currently the `forestecology` package can only fit the competition Bayesian linear regression model outlined in Section \ref{competition-model} using `comp_bayes_lm()`.
-However, any model implemented in a function that similarly takes an input data frame similar to `focal_vs_comp` as an argument can also be used.
-For our specific competition Bayesian linear regression model, we also specify prior distributions on all parameters of interest (here chosen to be the defaults as specified in `?comp_bayes_lm`).
+We fit the competition Bayesian linear regression model outlined in Equation \ref{eq:model} using `comp_bayes_lm()`, which has an option to specify prior distributions on all parameters of interest (chosen to be the defaults specified in `?comp_bayes_lm`).
 
@@
 
-The returned `comp_bayes_lm_scbi` output is an object of S3 class type `comp_bayes_lm` which contains the posterior values of all parameters in our competition Bayesian linear regression.
-This class of object includes generic methods implemented for `print()`, `predict()`, and `ggplot2::autoplot()`.
-First the generic for `print()` displays the names of all prior & posterior parameters along with the model formula:
+The returned `comp_bayes_lm_scbi` output is an object of S3 class type `comp_bayes_lm` containing the posterior values of all parameters in our competition Bayesian linear regression.
+This class of object includes three generic methods.
+First, the generic for `print()` displays the names of all prior & posterior parameters along with the model formula:
 
+<!--Hack alert. I can't figure out how to ensure output of print stay within 72 character line limit-->
 ```{r}
@@
 
-Next, the generic for `predict()` takes as inputs the posterior parameter values in `comp_bayes_lm_scbi` and the predictor variables in `newdata` and outputs a vector of fitted/predicted values $\widehat{y}$ of the DBH for each focal tree computed from the posterior predictive distribution.
+Next, the generic for `predict()` takes the posterior parameter values in `comp_bayes_lm_scbi` and the predictor variables in `newdata` and outputs a vector `growth_hat` of fitted/predicted values $\widehat{y}$ of the DBH for each focal tree computed from the posterior predictive distribution.
 
@@
 ```{r}
-focal_vs_comp_scbi
+focal_vs_comp_scbi %>% 
+  select(focal_ID, focal_sp, dbh, growth, growth_hat)
 ```
@@
 
-Lastly, the generic for `ggplot2::autoplot()` allows us to plot the posterior distribution of all parameters in Figure \ref{fig:scbi-posterior-viz} (for compactness we only show posteriors for 3 species).
+Lastly, the generic for `ggplot2::autoplot()` allows us to plot the posterior distribution of all parameters in Figure \ref{fig:scbi-posterior-viz} (for compactness we include only 3 species).
 
@@
 
-These plots give the posterior distributions of parameters from Equation \ref{eq:model}.
+These plots visualize the posterior distributions of parameters from Equation \ref{eq:model}.
 For many package users they will be of interest because they give insight into the species-specific competitive interactions.
-Setting `type = "intercepts"` gives posterior distributions for $\beta_{0,j}$ and `type = "dbh_slopes"` for $\beta_{dbh,j}$.
-These give species specific growth independent of competition.
-The values of more interest are plotted with `type = 'competition'` which gives the posterior distribution for $\lambda_{j,k}$ species-specific competition coefficients (i.e., the $\lambda$-matrix).
-Negative values indicate a competitor species which slows the growth of a focal species.
-Here, for example, we see that tulip trees (`litu`) have a strong negative effect on the growth of conspecifics but relatively little effect on neighbors of the other two species.
+Setting `type = "intercepts"` returns species-specific posterior distributions for $\beta_{0,j}$ and `type = "dbh_slopes"` for $\beta_{dbh,j}$.
+Setting `type = "competition"` returns competition coefficients $\lambda_{j,k}$ where negative values indicate a competitor species which slows the growth of a focal species.
+Here, for example, we see that `litu` tulip poplars have a strong negative effect on the growth of conspecifics but relatively lesser effect on neighbors of the other two species.
+
+Currently the `forestecology` package can only fit the competition Bayesian linear regression model outlined in Equation \ref{eq:model}.
+However, it can be extended to any model implemented in a function similar to `comp_bayes_lm()` that uses data frames of similar format to `focal_vs_comp` as input.
 
@@
 
-In order to evaluate the effect of competitor species identity, we use the four steps of our analysis sequence answer along with a permutation test: Under a null hypothesis where competitor species identity does not matter, we can permute/shuffle this variable within each focal tree, compute the RMSE (the test statistic of interest), repeat this process several times to construct a null distribution of the RMSE, and compare it to the observed RMSE to assess significance.
-Going back to our example in Section \ref{focal-vs-comp} of focal tree with `focal_ID` 4 and its 20 competitors, the permutation test randomly resamples the `comp_sp` variable with replacement, leaving all other variables intact.
-The resampling with replacement is nested within each focal tree in order to preserve the neighborhood structure of our competition model.
-To run the permutation test, we use `comp_bayes_lm()` as in Section \ref{model-fit-predict}, but with a `run_shuffle = TRUE` argument.
+To evaluate the effect of competitor species identity, we use the four steps of our analysis sequence answer along with a permutation test: Under a null hypothesis where competitor species identity does not matter, we permute/shuffle this variable within each focal tree, compute the RMSE (the test statistic of interest), repeat this process several times to construct a null distribution of the RMSE, and compare it to the observed RMSE to assess significance.
+Going back to our example in Section \ref{focal-vs-comp} of focal tree with `focal_ID` 4 and its 20 competitors, the permutation test randomly resamples only the `comp_sp` variable with replacement, leaving all other variables intact.
+The resampling with replacement is nested within each focal tree in order to preserve neighborhood structure.
+We once again use `comp_bayes_lm()` as in Section \ref{model-fit-predict}, but with `run_shuffle = TRUE`.
 
@@
 To this end, we use the `foldID` variable defined in Section \ref{spatial-information} whereby all focal trees are assigned to one of `r n_fold` spatially contiguous blocks that act as folds in our cross-validation routine.
-Figure \ref{fig:scbi-spatial-cross-validation-schematic} presents a schematic illustrating this scheme for fold 1 (bottom-left) as the test set and folds 2, 3, and 4 as the training sets.
+Figure \ref{fig:scbi-spatial-cross-validation-schematic} presents a schematic illustrating this scheme for fold 1 (bottom-left) as the test set and folds 2--4 as the training set.
 We fit the model to all focal trees in the training set, apply the model to all focal trees in the test set to compute fitted/predicted values, and compute the RMSE of the observed versus predicted growths.
 We repeat this procedure 3 more times with each of the three remaining folds acting as the test set and then average all four resulting RMSE's.
-Furthermore, in order to maintain spatial independence between the test and training set, a fold buffer that extend outwards from the boundary of the test set is computed; all trees falling within this fold buffer are excluded from the training set.
+Furthermore, in order to maintain spatial independence between the test and training set, a "fold buffer" that extend outwards from the boundary of the test set is computed; all trees falling within this fold buffer are excluded from the training set.
 
@@
 
-This algorithm is implemented in `run_cv()`, which is a wrapper function to `comp_bayes_lm()` that fits the model and the `predict()` generic that returns fitted/predicted values.
-We compare these values to the observed growth values to again compute our RMSE.
+This algorithm is implemented in `run_cv()`, which is a wrapper function to both `comp_bayes_lm()` that fits the model and `predict()` that returns fitted/predicted values.
+We once again compute the RMSE.
 
@@
 
-# Importance of spatial cross validation
+# Importance of spatial cross-validation
 
@@
 This permutes the competitor species, as described above, but does so when calculating predicted growth with the cross validated scheme.
-Figure \ref{fig:scbi-simulation} compares model performance when permuting competitor species and calculating RMSE with and without cross validation.
+Figure \ref{fig:scbi-simulation} compares model performance when permuting competitor species and calculating RMSE with and without cross-validation.
 Without cross-validation the competitor identity did matter, the non-permuted competitor species had a much lower RMSE than the permuted one.
-But once we include the spatial cross validation this improvement disappears.
-So in this 9 ha subplot of the SCBI plot competitive interactions do not depend on the identity of the competitor, which is the opposite of what has been observed in other places (@allen_permutation_2020 @uriarte_spatially_2004).
-This highlights the importance of cross validation, without it the model was overfit.
+But once we include the spatial cross-validation, this improvement disappears.
+These results suggest that in this 9 ha subplot of the SCBI plot competitive interactions do not depend on the identity of the competitor, which is the opposite of what has been observed in other locations (@allen_permutation_2020 @uriarte_spatially_2004).
+This highlights the importance of cross-validation, without it the model was overfit.
 
@@
 The `forestecology` package provides an accessible way to fit and test models of neighborhood competition.
-Currently it is written to work with data from ForestGEO plots.
-But it could easily be modified to work on any single large, mapped forest plot in which at least two measurements of each individual have been taken.
-With a little bit of work the package could also be applied to forest inventory data in which several small plots are spatially separated, e.g., USFS Forest Inventory.
-In a future version of the package we also hope to make it possible to model plant mortality in addition to plant growth.
-The package follows the guidelines for `tidy` data, `sf` spatial structure, and S3 open-oriented model structure.
+Currently it is written to work with data from ForestGEO plots, but it could easily be modified to work on any single large, mapped forest plot in which at least two measurements of each individual have been taken.
+With some data wrangling the package could also be applied to forest inventory data in which several small plots are spatially separated, e.g., USFS Forest Inventory.
+In future versions of `forestecology` we also hope to make it possible to model plant mortality in addition to plant growth.
+The package follows the guidelines for `tidy` data, leverages the `sf` package for spatial data, and S3 open-oriented model structure.
 We hope that the package will increase the use of neighborhood competition models to better understand what structures plant competition.
@@
 This repository will be versioned and archived on Zenodo upon acceptance.
-The 2008 and 2014 Smithsonian Conservation Biology Institute census data loaded in Section \ref{compute-growth} and saved in `scbi.stem2.csv` and `scbi.stem3.csv` are available on GitHub at https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/tree/master/tree_main_census/data/census-csv-files and have been versioned and archived on Zenodo at https://doi.org/10.5281/zenodo.2649301 [@gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020].
+The 2008 and 2014 Smithsonian Conservation Biology Institute census data loaded in Section \ref{compute-growth}are available on GitHub at https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/tree/master/tree_main_census/data/census-csv-files and are versioned and archived on Zenodo at https://doi.org/10.5281/zenodo.2649301 [@gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020].
 

### 2021-03-09 paper/forestecology.Rmd
About to start slash & burn 712 words

@@
 )
-
-# Resolve conflicts in function names
-filter <- dplyr::filter
-select <- dplyr::select
 ```
@@
 library(blockCV)
+
+# Resolve conflicting functions
+filter <- dplyr::filter
+select <- dplyr::select
 ```
@@
 
-<!--Hack alert. I can't figure out how to ensure output of print stay within 72 character line limit-->
 ```{r}

### 2021-03-09 paper/forestecology.tex
About to start slash & burn 712 words

@@
 designed with cross-site compatibility in mind. We highlight the
-importance of spatial cross validation when interpreting model
+importance of spatial cross-validation when interpreting model
 results.\\
@@
 {\it Keywords:} forest ecology, interspecific competition, neigbhorhood competition,
-tree growth, R, ForestGEO, spatial cross validation
+tree growth, R, ForestGEO, spatial cross-validation
 \vfill
@@
 selective logging practices (\citet{canham_neighborhood_2006}). Although
-these are well-described methods, no methods are currently available for
-their easy application. Here we address this in an R package. We largely
+these are well-described methods, few methods are currently available
+for easy application. Here we address this in an R package. We largely
 follow the methods presented in \citet{allen_permutation_2020}. The
@@
 groups; and let \(k = 1, \ldots, K\) index all \(K\) ``competitor''
-species groups. We model the average annual growth in diameter
-\(y_{ij}\) (in centimeters per year) of the \(i^{th}\) tree of focal
-species group \(j\) as a linear model \(f\) of the covariates
+species groups. We model the average annual growth in diameter at breast
+height (DBH) \(y_{ij}\) (in centimeters per year) of the \(i^{th}\) tree
+of focal species group \(j\) as a linear model \(f\) of the covariates
 \(\vec{x}_{ij}\)
@@
 where \(\beta_{0,j}\) is the diameter-independent growth rate for group
-\(j\); \(\text{dbh}_{ij}\) is the diameter at breast height (in
-centimeters) of the focal tree at the earlier census;
-\(\beta_{\text{dbh},j}\) is the amount of the growth rate changed
-depending on diameter for group \(j\); \(\text{BA}_{ijk}\) is the sum of
-the basal area of all trees of competitor species group \(k\);
-\(\lambda_{jk}\) is the change in growth for individuals of group \(j\)
-from nearby competitors of group \(k\); and \(\epsilon_{ij}\) is a
+\(j\); \(\text{dbh}_{ij}\) is the DBH the focal tree at the earlier
+census; \(\beta_{\text{dbh},j}\) is the amount of the growth rate
+changed depending on diameter for group \(j\); \(\text{BA}_{ijk}\) is
+the sum of the basal area of all trees of competitor species group
+\(k\); \(\lambda_{jk}\) is the change in growth for individuals of group
+\(j\) from nearby competitors of group \(k\); and \(\epsilon_{ij}\) is a
 random error term distributed \(\text{Normal}(0, \sigma^2)\). They
@@
 Normal/Inverse Gamma conjugacy to derive closed-form solutions to all
-posterior distributions via linear algebra\footnote{See S1 Appendix of
+posterior distributions\footnote{See S1 Appendix of
   \citet{allen_permutation_2020}, available at
@@
 (\url{https://cran.r-project.org/web/packages/forestecology/index.html})
-with the corresponding source code available on GitHub
-(\url{https://github.com/rudeboybert/forestecology}). The package
-implements all aspects of the model in Equation \ref{eq:model}: model
-fitting and generating fitted/predicted values, evaluating the effect of
-competitor species identity using permutation tests, and evaluating
-model performance using spatial cross-validation.
+and on GitHub (\url{https://github.com/rudeboybert/forestecology}). The
+package implements all aspects of the model in Equation \ref{eq:model}:
+model fitting and generating fitted/predicted values, evaluating the
+effect of competitor species identity using permutation tests, and
+evaluating model performance using spatial cross-validation.
 
@@
 function returns an object of S3 class type \texttt{comp\_bayes\_lm}
-with generic methods implemented for \texttt{print()} to inspect the
-output, \texttt{predict()} to generate fitted/predicted values, and
-\texttt{ggplot2::autoplot()} to visualize all results. Therefore the
-package can be modularly extended to fit other models as long as they
-are coded into a function similar type as \texttt{comp\_bayes\_lm()} as
-has equivalent generic methods implemented.
+with generic methods implemented to print, make predictions using, and
+plot all results. Therefore the package can be modularly extended to fit
+other models as long as they are coded into a function similar to
+\texttt{comp\_bayes\_lm()} and has equivalent generic methods
+implemented.
 
@@
 provinces---and is adjacent to the northern end of Shenandoah National
-Park. The forest type is typical mature secondary eastern mixed
-deciduous forest, with a canopy dominated by tulip poplar
-(\emph{Liriodendron tulipifera}), oaks (\emph{Quercus} spp.), and
-hickories (\emph{Carya} spp.), and an understory composed mainly of
-spicebush (\emph{Lindera benzoin}), paw-paw (\emph{Asimina triloba}),
-American hornbeam (\emph{Carpinus caroliniana}), and witch hazel
-(\emph{Hamamelis virginiana}) \citep{bourg_initial_2013}.
+Park.
 
@@
 \KeywordTok{library}\NormalTok{(blockCV)}
-\end{Highlighting}
-\end{Shaded}
-
-\hypertarget{step-1-compute-the-growth-of-trees-based-on-census-data-compute-growth}{%
-\subsection{Step 1: Compute the growth of trees based on census data
-\{compute-growth\}}\label{step-1-compute-the-growth-of-trees-based-on-census-data-compute-growth}}
-
-The first step in the our analysis sequence is to compute the growth of
-trees using data from two censuses. The \texttt{compute\_growth()}
-function computes average annual growth assuming census data that
-roughly follows ForestGEO standards. Despite such standards, minor
-variations will still exist between sites, thereby necessitating some
-data wrangling and checking. For example, the SCBI site records all
-diameters at breast height (DBH) in millimeters
+
+\CommentTok{# Resolve conflicting functions}
+\NormalTok{filter <-}\StringTok{ }\NormalTok{dplyr}\OperatorTok{::}\NormalTok{filter}
+\NormalTok{select <-}\StringTok{ }\NormalTok{dplyr}\OperatorTok{::}\NormalTok{select}
+\end{Highlighting}
+\end{Shaded}
+
+\hypertarget{compute-growth}{%
+\subsection{Step 1: Compute the growth of trees based on census
+data}\label{compute-growth}}
+
+The first step is to compute the growth of trees using data from two
+censuses. \texttt{compute\_growth()} computes average annual growth
+assuming census data that roughly follows ForestGEO standards. Despite
+such standards, minor variations will still exist between sites, thereby
+necessitating some data wrangling and checking. For example, the SCBI
+site records all diameters at breast height (DBH) in millimeters
 \citep{bourg_initial_2013}, whereas the Michigan Big Woods site records
@@
 
-We load both 2008 and 2014 SCBI census data \texttt{.csv} files as they
-existed on GitHub on November 20, 2020
-\citep{gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020}. After
-selecting only the relevant variables, we perform a few additional data
-wrangling steps: convert the character variable with the date of
-measurement to be of explicit type \texttt{date}, convert DBH to be in
-centimeters\footnote{A rule of thumb to determine the units of DBH is to
-  check if the smallest non-zero and non-missing measurement is 1 or 10.
-  If the former, then centimeters. If the later, then millimeters. This
-  is because ForestGEO protocols state that only trees with DBH greater
-  or equal to 1cm should be included in censuses.} Furthermore, in order
-to speed up computation for purposes of this example, we only consider a
-9 ha subsection of the 25.6 ha of the SCBI site: \texttt{gx} from 0--300
-instead of 0--400 and \texttt{gy} from 300--600 instead of 0--640.
+We first load both 2008 and 2014 SCBI census data \texttt{.csv} files as
+they existed on GitHub on November 20, 2020
+\citep{gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020} and perform
+some data wrangling to both data sets. We then only consider a 9 ha
+subsection of the 25.6 ha of the SCBI site, \texttt{gx} from 0--300
+instead of 0--400 and \texttt{gy} from 300--600 instead of 0--640, in
+order to speed up computation for purposes of this example.
 
@@
 \StringTok{  }\KeywordTok{mutate}\NormalTok{(}
+    \CommentTok{# Convert date from character to date}
     \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date),}
+    \CommentTok{# Convert dbh to be in cm}
     \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}
@@
 
-These two data frames are then supplied as arguments to the
-\texttt{compute\_growth()} function, along with the \texttt{id} argument
-that specifies the variable that uniquely identifies each tree-stem.
-Note furthermore that we discard all resprouts in the later census
-(those with \texttt{code\ ==\ R}), since we are only interested in the
-diameter growth of surviving, and not resprouted, stems.
+These two data frames are then supplied as arguments to
+\texttt{compute\_growth()}, along with the \texttt{id} argument that
+specifies the variable that uniquely identifies each tree-stem. Note
+furthermore that we discard all resprouts in the later census (those
+with \texttt{code\ ==\ R}), since we are only interested in the diameter
+growth of surviving, and not resprouted, stems.
 
@@
 \texttt{sf} that includes variables \texttt{growth}, the average annual
-growth in DBH (cm y\textsuperscript{-1}) for all stems that were alive
-at both time points, and \texttt{geometry}, the \texttt{sf} package's
-encoding of geolocations of type
+growth in DBH (cm \(\cdot\) y\textsuperscript{-1}) for all stems that
+were alive at both time points, and \texttt{geometry}, the \texttt{sf}
+package's encoding of geolocations of type
 \texttt{\textless{}POINT\textgreater{}}. In addition the species
-variable \texttt{sp} is converted to a factor if it wasn't already by
-\texttt{compute\_growth()}.\footnote{In our spatial cross-validation
-  algorithm in Section \ref{spatial-cross-validation} issues can occur
-  when rare species do not occur in the training set, but then are
-  encountered in the test set. This risk is mitigated by representing
-  \texttt{sp} as a factor variable, which has a complete list of all
-  levels of the categorical variable.} Furthermore, the variables that
-should remain unchanged between censuses appear only once, such as
-location variables \texttt{gx} and \texttt{gy}; as well as
-species-related variables. Variables that should change between censuses
-are suffixed with \texttt{1} and \texttt{2} indicating the earlier and
-later censuses, such as \texttt{dbh1/dbh2} and \texttt{codes1/codes2}.
-
-Site data that does not align with this convention will need to be
-transformed for use with the \texttt{compute\_growth()} function.
-However, in the end, all that matters is that the growth of all stems is
-saved in a data frame of class \texttt{sf} and, at a minimum, contains
-the variable uniquely identifying each stem, \texttt{sp}, \texttt{dbh1},
-\texttt{growth}, \texttt{geometry}.
+variable \texttt{sp} is returned as a factor.\footnote{In our spatial
+  cross-validation algorithm in Section \ref{spatial-cross-validation}
+  issues can occur when rare species do not occur in the training set,
+  but then are encountered in the test set. This risk is mitigated by
+  representing \texttt{sp} as a factor variable, which has a complete
+  list of all levels of the categorical variable.}
 
@@
 
-The next step in our analysis sequence is to add additional spatial
-information to our main \texttt{growth\_scbi} data frame. The first
-element of spatial information we add is a ``buffer region'' to the
-periphery of the study region. Since some of our model's explanatory
-variables are cumulative (such as competitor basal area), we must ensure
-that all trees being modeled are not biased to have different neighbor
-structures. This is of concern for trees at the boundary of study
-regions, for which all neighbors will not be included in the censused
-stems. In order to account for such edge effects, only trees that are
-not part of this buffer region, i.e.~are part of the interior of the
-study region, will have their growth modeled
-\citep{waller_applied_2004}.
+The next step is to add additional spatial information to
+\texttt{growth\_scbi}. The first element we add is a ``buffer region''
+to the periphery of the study region. Since some of our model's
+explanatory variables are cumulative, we must ensure that all trees
+being modeled are not biased to have different neighbor structures. This
+is of concern for trees at the boundary of study regions, for which all
+neighbors will not be included in the censused stems. In order to
+account for such edge effects, only trees that are not part of this
+buffer region, i.e.~are part of the interior of the study region, will
+have their growth modeled \citep{waller_applied_2004}.
 
@@
 within a distance \texttt{comp\_dist} of a focal tree are considered its
-competitors (assuming the same units as the \texttt{gx} and \texttt{gy}
-location variables). In our case we set this value at 7.5m, a value
-informed by other studies \citep[\citet{uriarte_spatially_2004},
+competitors. In our case we set this value at 7.5m, a value informed by
+other studies \citep[\citet{uriarte_spatially_2004},
 \citet{canham_neighborhood_2006}]{canham_neighborhood_2004}. Using this
 value along with a manually constructed \texttt{sf} object
-representation of the study region's boundary via its vertices, we apply
-the \texttt{add\_buffer\_variable()} to our \texttt{growth\_scbi} data
-frame to add a \texttt{buffer} boolean variable: all trees who have
-\texttt{buffer} set to \texttt{FALSE} will be our focal trees whose
-growth will be modeled, whereas those with \texttt{TRUE} will only be
-considered as competitor trees whose growth will not.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{# Define buffer region using competitive distance range}
+representation of the study region's boundary, we apply the
+\texttt{add\_buffer\_variable()} to \texttt{growth\_scbi} to add a
+\texttt{buffer} boolean variable. All trees with \texttt{buffer} as
+\texttt{FALSE} will be our focal trees whose growth will be modeled,
+whereas those with \texttt{TRUE} will only be considered as competitor
+trees.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{# Define competitive distance range}
 \NormalTok{comp_dist <-}\StringTok{ }\FloatTok{7.5}
 
+\CommentTok{# Manually construct study region boundary}
 \NormalTok{study_region_scbi <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}
@@
 observations. However, many of these algorithms assume that the
-observations are independent of each other. In the case of forest census
-data, observations exhibit spatial autocorrelation. We therefore
-incorporate this spatial dependence into the cross-validation algorithm
-with our spatial blocks of trees \citep[
-\citet{pohjankukka_estimating_2017}]{roberts_cross-validation_2017} In
-the example below, we first manually define four folds that partition
+observations are independent. In the case of forest census data,
+observations exhibit spatial autocorrelation. We therefore incorporate
+this spatial dependence into the cross-validation algorithm with our
+spatial blocks of trees \citep[
+\citet{pohjankukka_estimating_2017}]{roberts_cross-validation_2017}.
+
+In the example below, we first manually define four folds that partition
 the study region as an \texttt{sf} object. We then use the output of the
 \texttt{spatialBlock()} function from the \texttt{blockCV} package to
-associate each tree in \texttt{growth\_scbi} to the correct fold (saved
-in the \texttt{foldID} variable) \citep{valavi_blockcv_2019}. \footnote{In
-  the Supporting Information we present an example where the folds
-  themselves are also created using the \texttt{spatialBlock()} function
-  given a specified \texttt{cv\_block\_size}.}
+associate each tree in \texttt{growth\_scbi} to the correct fold
+\texttt{foldID} \citep{valavi_blockcv_2019}. \footnote{In the Supporting
+  Information we present an example where the folds themselves are also
+  created automatically using \texttt{spatialBlock()} given a specified
+  \texttt{cv\_block\_size}, as opposed to manually as in the example.}
 
@@
 Figure \ref{fig:scbi-spatial-information} illustrates the net effect of
-adding these two elements of information to the \texttt{growth\_scbi}
-data frame. The location of each tree is marked with an integer
-indicating which fold it belongs to, where the folds are marked with
-solid lines. The color of each digit indicates whether the tree is part
-of the buffer region (and thus will only be considered as a competitor
-tree in our model) or is part of the interior of the study region (and
-thus is a focal tree whose growth is of modeled interest).
+adding these two elements of spatial information to
+\texttt{growth\_scbi}. The location of each tree is marked with an
+integer indicating its fold, where the folds are marked with solid
+lines. The color of each digit indicates whether the tree is part of the
+buffer region (and thus will only be considered as a competitor tree in
+our model) or is part of the interior of the study region (and thus is a
+focal tree whose growth is of modeled interest).
 
@@
 
-The next step in our analysis sequence is to identify all focal trees
-and their corresponding competitor trees. More specifically, it
-identifies all trees that are not part of the buffer region, have a
-valid \texttt{growth} measurement, and have at least one neighbor within
-7.5 of it. The \texttt{create\_focal\_vs\_comp()} functions performs
-these tasks and returns a new data frame of type \texttt{sf}. On top of
-the previous arguments \texttt{comp\_dist} defining the competition
-neighborhood and \texttt{id} indicating which variable uniquely
-identifies each tree-stem, this function also requires an \texttt{sf}
-object representation of the spatial cross-validation blocks/folds. In
-this example, the blocks were manually encoded in \texttt{blocks\_scbi}
-by specifying it's vertices in Section
-\ref{spatial-information}\footnote{We present an alternative method for
-  defining spatial cross-validation blocks is using the
-  \texttt{spatialBlock()} function from the \texttt{blockCV} package in
-  the Supporting Information.}. We present the resulting data frame
-below with the \texttt{foldID} variable omitted for compactness of
-presentation.
+The next step is to identify all focal trees and their corresponding
+competitor trees. More specifically, identify all trees that are not
+part of the buffer region, have a valid \texttt{growth} measurement, and
+have at least one neighbor within 7.5m.
+\texttt{create\_focal\_vs\_comp()} returns a new data frame of type
+\texttt{sf}. On top of previously detailed arguments \texttt{comp\_dist}
+and \texttt{id}, \texttt{create\_focal\_vs\_comp()} also requires an
+\texttt{sf} object representation of the spatial cross-validation
+blocks/folds as seen in Section \ref{spatial-information}. We present
+the resulting data frame below with the \texttt{foldID} variable omitted
+for compactness.
 
@@
 
-The resulting data frame \texttt{focal\_vs\_comp\_scbi} has 6296 rows,
-representing the subset of the 7954 trees in \texttt{growth\_scbi} that
-will be considered as focal trees. Two new variables \texttt{focal\_ID}
-and \texttt{focal\_sp} relate to tree-stem identification and species
+The resulting \texttt{focal\_vs\_comp\_scbi} has 6296 rows, representing
+the subset of the 7954 trees in \texttt{growth\_scbi} that will be
+considered as focal trees. Two new variables \texttt{focal\_ID} and
+\texttt{focal\_sp} relate to tree-stem identification and species
 information. Most notably however is a new variable \texttt{comp} which
-contains information on all competitor trees for a given focal tree,
-saved in \texttt{tidyr} package list-column format
-\citep{tidyr_package}. For example, we drill-down on the tree with
-\texttt{focal\_ID} 4, which has 20 competitor trees each described by 4
-variables as indicated by the fact that \texttt{comp} is a
-\texttt{tibble\ {[}20\ ×\ 4{]}}.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{filter}\NormalTok{(focal_ID }\OperatorTok{==}\StringTok{ }\DecValTok{4}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{select}\NormalTok{(focal_ID, dbh, comp)}
-\CommentTok{## # A tibble: 1 x 3}
-\CommentTok{##   focal_ID   dbh comp             }
-\CommentTok{##      <dbl> <dbl> <list>           }
-\CommentTok{## 1        4  13.6 <tibble [20 x 4]>}
-\end{Highlighting}
-\end{Shaded}
-
-The spatial distribution of these trees is visualized in Figure
-\ref{fig:scbi-focal-vs-comp-map}: the dashed circle extends 7.5 m away
-from the focal tree while all 20 competitor trees are within this
-circle.
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.66\linewidth]{Figures/scbi-focal-vs-comp-map-1}
-
-}
-
-\caption{Identify all focal and corresponding competitor trees: All 20 competitor trees of focal tree 4.}\label{fig:scbi-focal-vs-comp-map}
-\end{figure}
-
-Using the \texttt{unnest()} function from the \texttt{tidyr} package, we
-can flatten list-column into regular columns. We observe that for the
-same focal tree, we have information on all 20 competitor trees whose
-\texttt{dist} distance to the focal tree is \(\leq\) 7.5: their unique
-tree-stem ID number, their species, and their basal area (in m\(^2\))
-calculated as \(\frac{\pi \times (\text{DBH/2})^2}{10000}\) where
-\(DBH\) is the value from the earlier of the two censuses in cm. Saving
-our focal versus competitor information in list-column minimizes
-redundancy since we do not repeat information on the focal tree 20
-times.
+contains information on all competitor trees saved in \texttt{tidyr}
+package list-column format \citep{tidyr_package}. We flatten the
+\texttt{comp} list-column for the tree with \texttt{focal\_ID} 4 in the
+first row, here a \texttt{tibble\ {[}20\ ×\ 4{]}}, into regular columns
+using \texttt{unnest()} from the \texttt{tidyr} package.
 
@@
 
+We observe that for this focal tree, we have 4 variables of information
+on its 20 competitor trees: their unique tree-stem ID number, their
+distance to the focal tree (all \(\leq\) 7.5), their species, and their
+basal area (in m\(^2\)) calculated as
+\(\frac{\pi \times (\text{DBH/2})^2}{10000}\) where \(DBH\) is the value
+from the earlier of the two censuses in cm. Saving our focal versus
+competitor information in list-column minimizes redundancy since we do
+not repeat information on the focal tree 20 times. The spatial
+distribution of these trees is visualized in Figure
+\ref{fig:scbi-focal-vs-comp-map}: the dashed circle extends 7.5 m away
+from the focal tree while all 20 competitor trees are within this
+circle.
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.66\linewidth]{Figures/scbi-focal-vs-comp-map-1}
+
+}
+
+\caption{Identify all focal and corresponding competitor trees: All 20 competitor trees of focal tree 4.}\label{fig:scbi-focal-vs-comp-map}
+\end{figure}
+
 \hypertarget{model-fit-predict}{%
@@
 
-Now that we've identified all focal and corresponding competitor trees
-and saved this information in a data frame of type
-\texttt{focal\_vs\_comp}, the final step in our analysis sequence is to
-fit a model for the growth of all focal trees. Currently the
-\texttt{forestecology} package can only fit the competition Bayesian
-linear regression model outlined in Section \ref{competition-model}
-using the \texttt{comp\_bayes\_lm()} function. However, any model
-implemented in a function that similarly takes an input data frame of
-type \texttt{focal\_vs\_comp} as an argument can also be used. For our
-specific competition Bayesian linear regression model, we also specify
-prior distributions on all parameters of interest (here chosen to be the
-defaults as specified in \texttt{?comp\_bayes\_lm}).
+The final step is to fit a model for the growth of all focal trees. We
+fit the competition Bayesian linear regression model outlined in
+Equation \ref{eq:model} using \texttt{comp\_bayes\_lm()}, which has an
+option to specify prior distributions on all parameters of interest
+(chosen to be the defaults specified in \texttt{?comp\_bayes\_lm}).
 
@@
 The returned \texttt{comp\_bayes\_lm\_scbi} output is an object of S3
-class type \texttt{comp\_bayes\_lm} which contains the posterior values
-of all parameters in our competition Bayesian linear regression. This
-class of object includes generic methods implemented for
-\texttt{print()}, \texttt{predict()}, and \texttt{ggplot2::autoplot()}.
-First the generic for \texttt{print()} displays the names of all prior
-\& posterior parameters along with the model formula:
+class type \texttt{comp\_bayes\_lm} containing the posterior values of
+all parameters in our competition Bayesian linear regression. This class
+of object includes three generic methods. First, the generic for
+\texttt{print()} displays the names of all prior \& posterior parameters
+along with the model formula:
 
@@
 \NormalTok{comp_bayes_lm_scbi}
-\CommentTok{## Bayesian linear regression model parameters with a multivariate Normal likelihood. See ?comp_bayes_lm for details:}
+\CommentTok{## Bayesian linear regression model parameters with a multivariate Normal}
+\CommentTok{## likelihood. See ?comp_bayes_lm for details:}
 \CommentTok{## }
@@
 \CommentTok{## Model formula:}
-\CommentTok{## growth ~ sp + dbh + dbh * sp + acne * sp + acru * sp + amar * sp + astr * sp + caca * sp + caco * sp + cade * sp + cagl * sp + caovl * sp + cato * sp + ceca * sp + ceoc * sp + chvi * sp + cofl * sp + crpr * sp + crsp * sp + divi * sp + elum * sp + fagr * sp + fram * sp + frni * sp + frpe * sp + havi * sp + ilve * sp + juci * sp + juni * sp + libe * sp + litu * sp + nysy * sp + pist * sp + pivi * sp + ploc * sp + prav * sp + prse * sp + qual * sp + quco * sp + qufa * sp + qumi * sp + qupr * sp + quru * sp + quve * sp + rops * sp + saal * sp + saca * sp + tiam * sp + ulam * sp + ulru * sp + unk * sp + vipr * sp}
-\end{Highlighting}
-\end{Shaded}
-
-Next, the generic for \texttt{predict()} takes as inputs the posterior
-parameter values in \texttt{comp\_bayes\_lm\_scbi} and the predictor
-variables in \texttt{newdata} and outputs a vector of fitted/predicted
-values \(\widehat{y}\) of the DBH for each focal tree computed from the
-posterior predictive distribution.
+\CommentTok{## growth ~ sp + dbh + dbh * sp + acne * sp + acru * sp + amar * sp + astr}
+\CommentTok{## * sp + caca * sp + caco * sp + cade * sp + cagl * sp + caovl * sp + cato}
+\CommentTok{## * sp + ceca * sp + ceoc * sp + chvi * sp + cofl * sp + crpr * sp + crsp}
+\CommentTok{## * sp + divi * sp + elum * sp + fagr * sp + fram * sp + frni * sp + frpe}
+\CommentTok{## * sp + havi * sp + ilve * sp + juci * sp + juni * sp + libe * sp + litu}
+\CommentTok{## * sp + nysy * sp + pist * sp + pivi * sp + ploc * sp + prav * sp + prse}
+\CommentTok{## * sp + qual * sp + quco * sp + qufa * sp + qumi * sp + qupr * sp + quru}
+\CommentTok{## * sp + quve * sp + rops * sp + saal * sp + saca * sp + tiam * sp + ulam}
+\CommentTok{## * sp + ulru * sp + unk * sp + vipr * sp}
+\end{Highlighting}
+\end{Shaded}
+
+Next, the generic for \texttt{predict()} takes the posterior parameter
+values in \texttt{comp\_bayes\_lm\_scbi} and the predictor variables in
+\texttt{newdata} and outputs a vector \texttt{growth\_hat} of
+fitted/predicted values \(\widehat{y}\) of the DBH for each focal tree
+computed from the posterior predictive distribution.
 
@@
 \begin{Highlighting}[]
-\NormalTok{focal_vs_comp_scbi}
-\CommentTok{## # A tibble: 6,296 x 8}
-\CommentTok{##   focal_ID focal_sp   dbh foldID   geometry growth comp             }
-\CommentTok{##      <dbl> <fct>    <dbl> <fct>     <POINT>  <dbl> <list>           }
-\CommentTok{## 1        4 nysy     13.6  1      (14.2 428)  0.103 <tibble [20 x 4]>}
-\CommentTok{## 2        5 havi      8.8  1       (9.4 436)  0.150 <tibble [32 x 4]>}
-\CommentTok{## 3       79 tiam     47.7  1        (40 381) -0.161 <tibble [20 x 4]>}
-\CommentTok{## 4       80 caca      5.15 1      (38.7 422)  0.253 <tibble [12 x 4]>}
-\CommentTok{## 5       96 libe      2.3  1        (60 310)  0.262 <tibble [14 x 4]>}
-\CommentTok{## # ... with 6,291 more rows, and 1 more variable: growth_hat <dbl>}
+\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{select}\NormalTok{(focal_ID, focal_sp, dbh, growth, growth_hat)}
+\CommentTok{## # A tibble: 6,296 x 5}
+\CommentTok{##   focal_ID focal_sp   dbh growth growth_hat}
+\CommentTok{##      <dbl> <fct>    <dbl>  <dbl>      <dbl>}
+\CommentTok{## 1        4 nysy     13.6   0.103     0.0809}
+\CommentTok{## 2        5 havi      8.8   0.150     0.112 }
+\CommentTok{## 3       79 tiam     47.7  -0.161     0.229 }
+\CommentTok{## 4       80 caca      5.15  0.253     0.121 }
+\CommentTok{## 5       96 libe      2.3   0.262     0.142 }
+\CommentTok{## # ... with 6,291 more rows}
 \end{Highlighting}
@@
 the posterior distribution of all parameters in Figure
-\ref{fig:scbi-posterior-viz} (for compactness we only show posteriors
-for 3 species).
+\ref{fig:scbi-posterior-viz} (for compactness we include only 3
+species).
 
@@
 
-These plots give the posterior distributions of parameters from Equation
-\ref{eq:model}. For many package users they will be of interest because
-they give insight into the species-specific competitive interactions.
-Setting \texttt{type\ =\ "intercepts"} gives posterior distributions for
-\(\beta_{0,j}\) and \texttt{type\ =\ "dbh\_slopes"} for
-\(\beta_{dbh,j}\). These give species specific growth independent of
-competition. The values of more interest are plotted with
-\texttt{type\ =\ \textquotesingle{}competition\textquotesingle{}} which
-gives the posterior distribution for \(\lambda_{j,k}\) species-specific
-competition coefficients (i.e., the \(\lambda\)-matrix). Negative values
-indicate a competitor species which slows the growth of a focal species.
-Here, for example, we see that tulip trees (\texttt{litu}) have a strong
-negative effect on the growth of conspecifics but relatively little
-effect on neighbors of the other two species.
+These plots visualize the posterior distributions of parameters from
+Equation \ref{eq:model}. For many package users they will be of interest
+because they give insight into the species-specific competitive
+interactions. Setting \texttt{type\ =\ "intercepts"} returns
+species-specific posterior distributions for \(\beta_{0,j}\) and
+\texttt{type\ =\ "dbh\_slopes"} for \(\beta_{dbh,j}\). Setting
+\texttt{type\ =\ "competition"} returns competition coefficients
+\(\lambda_{j,k}\) where negative values indicate a competitor species
+which slows the growth of a focal species. Here, for example, we see
+that \texttt{litu} tulip poplars have a strong negative effect on the
+growth of conspecifics but relatively lesser effect on neighbors of the
+other two species.
+
+Currently the \texttt{forestecology} package can only fit the
+competition Bayesian linear regression model outlined in Equation
+\ref{eq:model}. However, it can be extended to any model implemented in
+a function similar to \texttt{comp\_bayes\_lm()} that uses data frames
+of similar format to \texttt{focal\_vs\_comp} as input.
 
@@
 
-In order to evaluate the effect of competitor species identity, we use
-the four steps of our analysis sequence answer along with a permutation
-test: Under a null hypothesis where competitor species identity does not
-matter, we can permute/shuffle this variable within each focal tree,
-compute the RMSE (the test statistic of interest), repeat this process
-several times to construct a null distribution of the RMSE, and compare
-it to the observed RMSE to assess significance. Going back to our
-example in Section \ref{focal-vs-comp} of focal tree with
-\texttt{focal\_ID} 4 and its 20 competitors, the permutation test
-randomly resamples the \texttt{comp\_sp} variable with replacement,
-leaving all other variables intact. The resampling with replacement is
-nested within each focal tree in order to preserve the neighborhood
-structure of our competition model. To run the permutation test, we use
-the same \texttt{comp\_bayes\_lm()} function as in Section
-\ref{model-fit-predict}, but with a \texttt{run\_shuffle\ =\ TRUE}
-argument.
+To evaluate the effect of competitor species identity, we use the four
+steps of our analysis sequence answer along with a permutation test:
+Under a null hypothesis where competitor species identity does not
+matter, we permute/shuffle this variable within each focal tree, compute
+the RMSE (the test statistic of interest), repeat this process several
+times to construct a null distribution of the RMSE, and compare it to
+the observed RMSE to assess significance. Going back to our example in
+Section \ref{focal-vs-comp} of focal tree with \texttt{focal\_ID} 4 and
+its 20 competitors, the permutation test randomly resamples only the
+\texttt{comp\_sp} variable with replacement, leaving all other variables
+intact. The resampling with replacement is nested within each focal tree
+in order to preserve neighborhood structure. We once again use
+\texttt{comp\_bayes\_lm()} as in Section \ref{model-fit-predict}, but
+with \texttt{run\_shuffle\ =\ TRUE}.
 
@@
 presents a schematic illustrating this scheme for fold 1 (bottom-left)
-as the test set and folds 2, 3, and 4 as the training sets. We fit the
-model to all focal trees in the training set, apply the model to all
-focal trees in the test set to compute fitted/predicted values, and
-compute the RMSE of the observed versus predicted growths. We repeat
-this procedure 3 more times with each of the three remaining folds
-acting as the test set and then average all four resulting RMSE's.
-Furthermore, in order to maintain spatial independence between the test
-and training set, a fold buffer that extend outwards from the boundary
-of the test set is computed; all trees falling within this fold buffer
-are excluded from the training set.
+as the test set and folds 2--4 as the training set. We fit the model to
+all focal trees in the training set, apply the model to all focal trees
+in the test set to compute fitted/predicted values, and compute the RMSE
+of the observed versus predicted growths. We repeat this procedure 3
+more times with each of the three remaining folds acting as the test set
+and then average all four resulting RMSE's. Furthermore, in order to
+maintain spatial independence between the test and training set, a
+``fold buffer'' that extend outwards from the boundary of the test set
+is computed; all trees falling within this fold buffer are excluded from
+the training set.
 
@@
 
-This algorithm is implemented in the \texttt{run\_cv()} function, which
-is a wrapper function to the \texttt{comp\_bayes\_lm()} function that
-fits the model and the \texttt{predict()} generic that returns
-fitted/predicted values. We compare these values to the observed growth
-values to again compute our RMSE.
+This algorithm is implemented in \texttt{run\_cv()}, which is a wrapper
+function to both \texttt{comp\_bayes\_lm()} that fits the model and
+\texttt{predict()} that returns fitted/predicted values. We once again
+compute the RMSE.
 
@@
 \hypertarget{importance-of-spatial-cross-validation}{%
-\section{Importance of spatial cross
-validation}\label{importance-of-spatial-cross-validation}}
-
-The \texttt{run\_cv} function also accepts the \texttt{run\_shuffle}
-argument. This permutes the competitor species, as described above, but
-does so when calculating predicted growth with the cross validated
-scheme. Figure \ref{fig:scbi-simulation} compares model performance when
-permuting competitor species and calculating RMSE with and without cross
-validation. Without cross-validation the competitor identity did matter,
-the non-permuted competitor species had a much lower RMSE than the
-permuted one. But once we include the spatial cross validation this
-improvement disappears. So in this 9 ha subplot of the SCBI plot
-competitive interactions do not depend on the identity of the
-competitor, which is the opposite of what has been observed in other
-places (\citet{allen_permutation_2020} \citet{uriarte_spatially_2004}).
-This highlights the importance of cross validation, without it the model
-was overfit.
+\section{Importance of spatial
+cross-validation}\label{importance-of-spatial-cross-validation}}
+
+\texttt{run\_cv()} also accepts the \texttt{run\_shuffle} argument. This
+permutes the competitor species, as described above, but does so when
+calculating predicted growth with the cross validated scheme. Figure
+\ref{fig:scbi-simulation} compares model performance when permuting
+competitor species and calculating RMSE with and without
+cross-validation. Without cross-validation the competitor identity did
+matter, the non-permuted competitor species had a much lower RMSE than
+the permuted one. But once we include the spatial cross-validation, this
+improvement disappears. These results suggest that in this 9 ha subplot
+of the SCBI plot competitive interactions do not depend on the identity
+of the competitor, which is the opposite of what has been observed in
+other locations (\citet{allen_permutation_2020}
+\citet{uriarte_spatially_2004}). This highlights the importance of
+cross-validation, without it the model was overfit.
 
@@
 test models of neighborhood competition. Currently it is written to work
-with data from ForestGEO plots. But it could easily be modified to work
+with data from ForestGEO plots, but it could easily be modified to work
 on any single large, mapped forest plot in which at least two
-measurements of each individual have been taken. With a little bit of
-work the package could also be applied to forest inventory data in which
-several small plots are spatially separated, e.g., USFS Forest
-Inventory. In a future version of the package we also hope to make it
-possible to model plant mortality in addition to plant growth. The
-package follows the guidelines for \texttt{tidy} data, \texttt{sf}
-spatial structure, and S3 open-oriented model structure. We hope that
-the package will increase the use of neighborhood competition models to
-better understand what structures plant competition.
+measurements of each individual have been taken. With some data
+wrangling the package could also be applied to forest inventory data in
+which several small plots are spatially separated, e.g., USFS Forest
+Inventory. In future versions of \texttt{forestecology} we also hope to
+make it possible to model plant mortality in addition to plant growth.
+The package follows the guidelines for \texttt{tidy} data, leverages the
+\texttt{sf} package for spatial data, and S3 open-oriented model
+structure. We hope that the package will increase the use of
+neighborhood competition models to better understand what structures
+plant competition.
 
@@
 Smithsonian Conservation Biology Institute census data loaded in Section
-\ref{compute-growth} and saved in \texttt{scbi.stem2.csv} and
-\texttt{scbi.stem3.csv} are available on GitHub at
+\ref{compute-growth}are available on GitHub at
 \url{https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/tree/master/tree_main_census/data/census-csv-files}
-and have been versioned and archived on Zenodo at
+and are versioned and archived on Zenodo at
 \url{https://doi.org/10.5281/zenodo.2649301}

### 2021-03-09 paper/forestecology.Rmd
3712 -> 3502

@@
 As we articulate in Section \ref{casestudy}, these functions delineate the key steps in our analysis sequence.
-Furthermore, the inputs and outputs of nearly all of our functions use the same "simple features for R" data structures as implemented in the `sf` package for standardized support for spatial vector data [@pebesma_simple_2018].
-The `sf` package is a `tidyverse`-friendly evolution of the `sp` package of classes and methods for spatial data in R [@pebesma_sp_2005].
-As such, wrangling and visualization spatial data such as ours becomes much easier.
+Furthermore, the inputs and outputs of nearly all of our functions use the same "simple features for R" data structures as implemented in the `sf` package for standardized and `tidyverse`-friendly support for spatial vector data [@pebesma_simple_2018]
 
@@
 
-We present a case-study of the `forestecology` package's use on data from the Smithsonian Conservation Biology Institute's (SCBI) large forest dynamics plot in Front Royal, Virginia, USA in Section \ref{casestudy}, which is part of the ForestGEO global network of research sites [@bourg_initial_2013, @andersonteixeira_ctfs-forestgeo_2015].
-The package is designed with ForestGEO plot data in mind, but we envision that it could easily be modified to work with data from other forest plots, e.g.
-the US Forest Service Forest Inventory and Analysis plots or more generally to model interactions of any community of mapped sessile organisms [@smith_forest_2002].
-
 # forestecology workflow: a case study {#casestudy}
 
-We demonstrate the `forestecology` package's functionality on data from the Smithsonian Conservation Biology Institute (SCBI) large forest dynamics plot, located at the Smithsonian's National Zoo and Conservation Biology Institute in Front Royal, VA, USA [@bourg_initial_2013].
+We present a case-study of the `forestecology` package's functionality on data from the Smithsonian Conservation Biology Institute (SCBI) large forest dynamics plot in Front Royal, VA, USA, which is part of the ForestGEO global network of research sites [@bourg_initial_2013, @andersonteixeira_ctfs-forestgeo_2015] [@bourg_initial_2013].
 The 25.6 ha (640 x 400 m) plot is located at the intersection of three of the major physiographic provinces of the eastern US---the Blue Ridge, Ridge and Valley, and Piedmont provinces---and is adjacent to the northern end of Shenandoah National Park.
@@
 We then use the output of the `spatialBlock()` function from the `blockCV` package to associate each tree in `growth_scbi` to the correct fold `foldID` [@valavi_blockcv_2019].
-^[In the Supporting Information we present an example where the folds themselves are also created automatically using `spatialBlock()` given a specified `cv_block_size`, as opposed to manually as in the example.]
+^[In the Supporting Information we present an example where the folds themselves are also created automatically using `spatialBlock()` given a specified `cv_block_size`, as opposed to manually as in the example.] Figure \ref{fig:scbi-spatial-information} illustrates the net effect of adding these two elements of spatial information to `growth_scbi`.
 
@@
 
-Figure \ref{fig:scbi-spatial-information} illustrates the net effect of adding these two elements of spatial information to `growth_scbi`.
-The location of each tree is marked with an integer indicating its fold, where the folds are marked with solid lines.
-The color of each digit indicates whether the tree is part of the buffer region (and thus will only be considered as a competitor tree in our model) or is part of the interior of the study region (and thus is a focal tree whose growth is of modeled interest).
-
 ```{r scbi-spatial-information, out.width="66%", fig.cap="Add spatial information: Buffer region and spatial cross-validation blocks (1 through 4).
-All trees in the interior of the study region (i.e.
-not part of buffer) will be the focal trees whose growth will be modeled."}
+The location of each tree is marked with an integer indicating its fold, with folds delineated with solid lines.
+The color of each digit indicates whether the tree is part of the buffer region (and thus will only be considered as a competitor tree in our model) or is part of the interior of the study region (and thus is a focal tree whose growth is of modeled interest)."}
 ggplot() +
@@
 On top of previously detailed arguments `comp_dist` and `id`, `create_focal_vs_comp()` also requires an `sf` object representation of the spatial cross-validation blocks/folds as seen in Section \ref{spatial-information}.
-We present the resulting data frame below with the `foldID` variable omitted for compactness.
 
@@
 focal_vs_comp_scbi %>% 
-  select(-foldID)
+  select(focal_ID, focal_sp, geometry, growth, comp)
 ```
@@
 
-```{r scbi-posterior-viz, out.width="100%", fig.cap= "Fit model: Posterior distributions of all parameters for three species."}
+```{r scbi-posterior-viz, out.width="100%", fig.cap= "Fit model: Posterior distributions of all parameters for three species.", fig.width = 7.5}
 # Plot posteriors for only a subset of species
@@
 The `forestecology` package provides an accessible way to fit and test models of neighborhood competition.
-Currently it is written to work with data from ForestGEO plots, but it could easily be modified to work on any single large, mapped forest plot in which at least two measurements of each individual have been taken.
-With some data wrangling the package could also be applied to forest inventory data in which several small plots are spatially separated, e.g., USFS Forest Inventory.
+While the package is designed with ForestGEO plot data in mind, we envision that it can be modified to work on i) any single large, mapped forest plot in which at least two measurements of each individual have been taken, e.g.
+the US Forest Service Forest Inventory and Analysis plots or ii) more generally to model interactions of any community of mapped sessile organisms [@smith_forest_2002].
 In future versions of `forestecology` we also hope to make it possible to model plant mortality in addition to plant growth.

### 2021-03-09 paper/forestecology.Rmd
3712 -> 3421 by moving text to figure captions

@@
 In addition the species variable `sp` is returned as a factor.^[In our spatial cross-validation algorithm in Section \ref{spatial-cross-validation} issues can occur when rare species do not occur in the training set, but then are encountered in the test set.
-This risk is mitigated by representing `sp` as a factor variable, which has a complete list of all levels of the categorical variable.]
-
-Given that `growth_scbi` is of class `sf`, it can be easily plotted in `ggplot2` using the `geom_sf()` geometry as seen in Figure \ref{fig:scbi-trees} where we plot a random sample of 500 out of the `r nrow(growth_scbi)` trees.
+This risk is mitigated by representing `sp` as a factor variable, which has a complete list of all levels of the categorical variable.] Given that `growth_scbi` is of class `sf`, it can be easily plotted in `ggplot2` using the `geom_sf()` geometry as seen in Figure \ref{fig:scbi-trees}.
 
@@
 Saving our focal versus competitor information in list-column minimizes redundancy since we do not repeat information on the focal tree 20 times.
-The spatial distribution of these trees is visualized in Figure \ref{fig:scbi-focal-vs-comp-map}: the dashed circle extends `r comp_dist` m away from the focal tree while all 20 competitor trees are within this circle.
+The spatial distribution of these trees is visualized in Figure \ref{fig:scbi-focal-vs-comp-map}.
 
-```{r scbi-focal-vs-comp-map, out.width="66%", echo = FALSE, fig.cap="Identify all focal and corresponding competitor trees: All 20 competitor trees of focal tree 4."}
+```{r scbi-focal-vs-comp-map, out.width="66%", echo = FALSE, fig.cap="Identify all focal and corresponding competitor trees: The dashed circle extends 7.5m away from the focal tree 4 while all 20 competitor trees are within this circle."}
 comp_radius <- growth_scbi %>% 
@@
 
-Lastly, the generic for `ggplot2::autoplot()` allows us to plot the posterior distribution of all parameters in Figure \ref{fig:scbi-posterior-viz} (for compactness we include only 3 species).
+Lastly, the generic for `ggplot2::autoplot()` allows us to plot the posterior distribution of all parameters in Figure \ref{fig:scbi-posterior-viz}.
 
-```{r scbi-posterior-viz, out.width="100%", fig.cap= "Fit model: Posterior distributions of all parameters for three species.", fig.width = 7.5}
+```{r scbi-posterior-viz, out.width="100%", fig.cap= "Fit model: Posterior distributions of all parameters. For compactness we include only three species.", fig.width = 8, fig.height = 6}
 # Plot posteriors for only a subset of species
@@
 To this end, we use the `foldID` variable defined in Section \ref{spatial-information} whereby all focal trees are assigned to one of `r n_fold` spatially contiguous blocks that act as folds in our cross-validation routine.
-Figure \ref{fig:scbi-spatial-cross-validation-schematic} presents a schematic illustrating this scheme for fold 1 (bottom-left) as the test set and folds 2--4 as the training set.
+Figure \ref{fig:scbi-spatial-cross-validation-schematic} presents a schematic illustrating this scheme.
 We fit the model to all focal trees in the training set, apply the model to all focal trees in the test set to compute fitted/predicted values, and compute the RMSE of the observed versus predicted growths.
@@
 
-```{r scbi-spatial-cross-validation-schematic, out.width="66%", fig.cap= "Schematic of spatial cross-validation: Using the k = 1 fold as the test set, assigning each focal tree to training set, test set, and fold buffer.", echo = FALSE}
+```{r scbi-spatial-cross-validation-schematic, out.width="66%", fig.cap= "Schematic of spatial cross-validation: Using the k = 1 fold (bottom-left) as the test set, k = 2 through 4 as the training set, along with a \"fold buffer.\"", echo = FALSE}
 fold_number <- 1
@@
 
-The authors thank Ryan Giordano and Jonathan Che for their help with the statistical methodology and Sophie Li for their feedback on package interface. The authors declare no conflicts of interest.
+The authors thank Sophie Li for their feedback on the package interface. The authors declare no conflicts of interest.
 
@@
 
-We intend to archive all data and source code for the `forestecology` package as well as this manuscript on GitHub at https://github.com/rudeboybert/forestecology.
-This repository will be versioned and archived on Zenodo upon acceptance.
-The 2008 and 2014 Smithsonian Conservation Biology Institute census data loaded in Section \ref{compute-growth}are available on GitHub at https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/tree/master/tree_main_census/data/census-csv-files and are versioned and archived on Zenodo at https://doi.org/10.5281/zenodo.2649301 [@gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020].
+We intend to archive all data and source code for this manuscript on GitHub at https://github.com/rudeboybert/forestecology.
+This repository will be archived on Zenodo upon acceptance.
+The example Smithsonian Conservation Biology Institute census data used are available on GitHub at https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/tree/master/tree_main_census/data/census-csv-files and are archived on Zenodo at https://doi.org/10.5281/zenodo.2649301 [@gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020].
 

### 2021-03-09 paper/forestecology.tex
3712 -> 3421 by moving text to figure captions

@@
 structures as implemented in the \texttt{sf} package for standardized
-support for spatial vector data \citep{pebesma_simple_2018}. The
-\texttt{sf} package is a \texttt{tidyverse}-friendly evolution of the
-\texttt{sp} package of classes and methods for spatial data in R
-\citep{pebesma_sp_2005}. As such, wrangling and visualization spatial
-data such as ours becomes much easier.
+and \texttt{tidyverse}-friendly support for spatial vector data
+\citep{pebesma_simple_2018}
 
@@
 
-We present a case-study of the \texttt{forestecology} package's use on
-data from the Smithsonian Conservation Biology Institute's (SCBI) large
-forest dynamics plot in Front Royal, Virginia, USA in Section
-\ref{casestudy}, which is part of the ForestGEO global network of
-research sites
-\citep[\citet{andersonteixeira_ctfs-forestgeo_2015}]{bourg_initial_2013}.
-The package is designed with ForestGEO plot data in mind, but we
-envision that it could easily be modified to work with data from other
-forest plots, e.g.~the US Forest Service Forest Inventory and Analysis
-plots or more generally to model interactions of any community of mapped
-sessile organisms \citep{smith_forest_2002}.
-
 \hypertarget{casestudy}{%
@@
 
-We demonstrate the \texttt{forestecology} package's functionality on
-data from the Smithsonian Conservation Biology Institute (SCBI) large
-forest dynamics plot, located at the Smithsonian's National Zoo and
-Conservation Biology Institute in Front Royal, VA, USA
+We present a case-study of the \texttt{forestecology} package's
+functionality on data from the Smithsonian Conservation Biology
+Institute (SCBI) large forest dynamics plot in Front Royal, VA, USA,
+which is part of the ForestGEO global network of research sites
+\citep[\citet{andersonteixeira_ctfs-forestgeo_2015}]{bourg_initial_2013}
 \citep{bourg_initial_2013}. The 25.6 ha (640 x 400 m) plot is located at
@@
   representing \texttt{sp} as a factor variable, which has a complete
-  list of all levels of the categorical variable.}
-
-Given that \texttt{growth\_scbi} is of class \texttt{sf}, it can be
-easily plotted in \texttt{ggplot2} using the \texttt{geom\_sf()}
-geometry as seen in Figure \ref{fig:scbi-trees} where we plot a random
-sample of 500 out of the 7954 trees.
+  list of all levels of the categorical variable.} Given that
+\texttt{growth\_scbi} is of class \texttt{sf}, it can be easily plotted
+in \texttt{ggplot2} using the \texttt{geom\_sf()} geometry as seen in
+Figure \ref{fig:scbi-trees}.
 
@@
   \texttt{cv\_block\_size}, as opposed to manually as in the example.}
+Figure \ref{fig:scbi-spatial-information} illustrates the net effect of
+adding these two elements of spatial information to
+\texttt{growth\_scbi}.
 
@@
 
-Figure \ref{fig:scbi-spatial-information} illustrates the net effect of
-adding these two elements of spatial information to
-\texttt{growth\_scbi}. The location of each tree is marked with an
-integer indicating its fold, where the folds are marked with solid
-lines. The color of each digit indicates whether the tree is part of the
-buffer region (and thus will only be considered as a competitor tree in
-our model) or is part of the interior of the study region (and thus is a
-focal tree whose growth is of modeled interest).
-
 \begin{Shaded}
@@
 \caption{Add spatial information: Buffer region and spatial cross-validation blocks (1 through 4).
-All trees in the interior of the study region (i.e.
-not part of buffer) will be the focal trees whose growth will be modeled.}\label{fig:scbi-spatial-information}
+The location of each tree is marked with an integer indicating its fold, with folds delineated with solid lines.
+The color of each digit indicates whether the tree is part of the buffer region (and thus will only be considered as a competitor tree in our model) or is part of the interior of the study region (and thus is a focal tree whose growth is of modeled interest).}\label{fig:scbi-spatial-information}
 \end{figure}
@@
 \texttt{sf} object representation of the spatial cross-validation
-blocks/folds as seen in Section \ref{spatial-information}. We present
-the resulting data frame below with the \texttt{foldID} variable omitted
-for compactness.
+blocks/folds as seen in Section \ref{spatial-information}.
 
@@
 \NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\NormalTok{foldID)}
-\CommentTok{## # A tibble: 6,296 x 6}
-\CommentTok{##   focal_ID focal_sp   dbh   geometry growth comp             }
-\CommentTok{##      <dbl> <fct>    <dbl>    <POINT>  <dbl> <list>           }
-\CommentTok{## 1        4 nysy     13.6  (14.2 428)  0.103 <tibble [20 x 4]>}
-\CommentTok{## 2        5 havi      8.8   (9.4 436)  0.150 <tibble [32 x 4]>}
-\CommentTok{## 3       79 tiam     47.7    (40 381) -0.161 <tibble [20 x 4]>}
-\CommentTok{## 4       80 caca      5.15 (38.7 422)  0.253 <tibble [12 x 4]>}
-\CommentTok{## 5       96 libe      2.3    (60 310)  0.262 <tibble [14 x 4]>}
+\StringTok{  }\KeywordTok{select}\NormalTok{(focal_ID, focal_sp, geometry, growth, comp)}
+\CommentTok{## # A tibble: 6,296 x 5}
+\CommentTok{##   focal_ID focal_sp   geometry growth comp             }
+\CommentTok{##      <dbl> <fct>       <POINT>  <dbl> <list>           }
+\CommentTok{## 1        4 nysy     (14.2 428)  0.103 <tibble [20 x 4]>}
+\CommentTok{## 2        5 havi      (9.4 436)  0.150 <tibble [32 x 4]>}
+\CommentTok{## 3       79 tiam       (40 381) -0.161 <tibble [20 x 4]>}
+\CommentTok{## 4       80 caca     (38.7 422)  0.253 <tibble [12 x 4]>}
+\CommentTok{## 5       96 libe       (60 310)  0.262 <tibble [14 x 4]>}
 \CommentTok{## # ... with 6,291 more rows}
@@
 distribution of these trees is visualized in Figure
-\ref{fig:scbi-focal-vs-comp-map}: the dashed circle extends 7.5 m away
-from the focal tree while all 20 competitor trees are within this
-circle.
+\ref{fig:scbi-focal-vs-comp-map}.
 
@@
 
-\caption{Identify all focal and corresponding competitor trees: All 20 competitor trees of focal tree 4.}\label{fig:scbi-focal-vs-comp-map}
+\caption{Identify all focal and corresponding competitor trees: The dashed circle extends 7.5m away from the focal tree 4 while all 20 competitor trees are within this circle.}\label{fig:scbi-focal-vs-comp-map}
 \end{figure}
@@
 the posterior distribution of all parameters in Figure
-\ref{fig:scbi-posterior-viz} (for compactness we include only 3
-species).
+\ref{fig:scbi-posterior-viz}.
 
@@
 
-\caption{Fit model: Posterior distributions of all parameters for three species.}\label{fig:scbi-posterior-viz}
+\caption{Fit model: Posterior distributions of all parameters. For compactness we include only three species.}\label{fig:scbi-posterior-viz}
 \end{figure}
@@
 routine. Figure \ref{fig:scbi-spatial-cross-validation-schematic}
-presents a schematic illustrating this scheme for fold 1 (bottom-left)
-as the test set and folds 2--4 as the training set. We fit the model to
-all focal trees in the training set, apply the model to all focal trees
-in the test set to compute fitted/predicted values, and compute the RMSE
-of the observed versus predicted growths. We repeat this procedure 3
-more times with each of the three remaining folds acting as the test set
-and then average all four resulting RMSE's. Furthermore, in order to
+presents a schematic illustrating this scheme. We fit the model to all
+focal trees in the training set, apply the model to all focal trees in
+the test set to compute fitted/predicted values, and compute the RMSE of
+the observed versus predicted growths. We repeat this procedure 3 more
+times with each of the three remaining folds acting as the test set and
+then average all four resulting RMSE's. Furthermore, in order to
 maintain spatial independence between the test and training set, a
@@
 
-\caption{Schematic of spatial cross-validation: Using the k = 1 fold as the test set, assigning each focal tree to training set, test set, and fold buffer.}\label{fig:scbi-spatial-cross-validation-schematic}
+\caption{Schematic of spatial cross-validation: Using the k = 1 fold (bottom-left) as the test set, k = 2 through 4 as the training set, along with a "fold buffer."}\label{fig:scbi-spatial-cross-validation-schematic}
 \end{figure}
@@
 The \texttt{forestecology} package provides an accessible way to fit and
-test models of neighborhood competition. Currently it is written to work
-with data from ForestGEO plots, but it could easily be modified to work
-on any single large, mapped forest plot in which at least two
-measurements of each individual have been taken. With some data
-wrangling the package could also be applied to forest inventory data in
-which several small plots are spatially separated, e.g., USFS Forest
-Inventory. In future versions of \texttt{forestecology} we also hope to
-make it possible to model plant mortality in addition to plant growth.
-The package follows the guidelines for \texttt{tidy} data, leverages the
-\texttt{sf} package for spatial data, and S3 open-oriented model
-structure. We hope that the package will increase the use of
+test models of neighborhood competition. While the package is designed
+with ForestGEO plot data in mind, we envision that it can be modified to
+work on i) any single large, mapped forest plot in which at least two
+measurements of each individual have been taken, e.g.~the US Forest
+Service Forest Inventory and Analysis plots or ii) more generally to
+model interactions of any community of mapped sessile organisms
+\citep{smith_forest_2002}. In future versions of \texttt{forestecology}
+we also hope to make it possible to model plant mortality in addition to
+plant growth. The package follows the guidelines for \texttt{tidy} data,
+leverages the \texttt{sf} package for spatial data, and S3 open-oriented
+model structure. We hope that the package will increase the use of
 neighborhood competition models to better understand what structures
@@
 
-The authors thank Ryan Giordano and Jonathan Che for their help with the
-statistical methodology and Sophie Li for their feedback on package
-interface. The authors declare no conflicts of interest.
+The authors thank Sophie Li for their feedback on the package interface.
+The authors declare no conflicts of interest.
 
@@
 
-We intend to archive all data and source code for the
-\texttt{forestecology} package as well as this manuscript on GitHub at
-\url{https://github.com/rudeboybert/forestecology}. This repository will
-be versioned and archived on Zenodo upon acceptance. The 2008 and 2014
-Smithsonian Conservation Biology Institute census data loaded in Section
-\ref{compute-growth}are available on GitHub at
+We intend to archive all data and source code for this manuscript on
+GitHub at \url{https://github.com/rudeboybert/forestecology}. This
+repository will be archived on Zenodo upon acceptance. The example
+Smithsonian Conservation Biology Institute census data used are
+available on GitHub at
 \url{https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/tree/master/tree_main_census/data/census-csv-files}
-and are versioned and archived on Zenodo at
+and are archived on Zenodo at
 \url{https://doi.org/10.5281/zenodo.2649301}

### 2021-03-09 paper/forestecology.Rmd
3421 -> 3139 via only quick pass of text

@@
 @allen_permutation_2020 considers the following model: Let $i = 1, \ldots, n_j$ index all $n_j$ trees of "focal" species group $j$; let $j = 1, \ldots, J$ index all $J$ focal species groups; and let $k = 1, \ldots, K$ index all $K$ "competitor" species groups.
-We model the average annual growth in diameter at breast height (DBH) $y_{ij}$ (in centimeters per year) of the $i^{th}$ tree of focal species group $j$ as a linear model $f$ of the covariates $\vec{x}_{ij}$
+They use the following linear model the average annual growth in diameter at breast height (DBH) $y_{ij}$ (in centimeters/year) of the $i^{th}$ tree of focal species group $j$:
 
@@
 
-In order to evaluate whether competitor species identity matters, @allen_permutation_2020 run a permutation test where under the null hypothesis the species identity of all competitors of a focal tree can be permuted/shuffled:
+In order to evaluate whether competitor species identity matters, @allen_permutation_2020 run a permutation test where a null hypothesis of no species grouping-specific effects of competition is assumed, and thus the species identity of all competitors can be permuted/shuffled:
 
@@
 
-where the null hypothesis $H_0$ reflects a hypothesis of no species grouping-specific effects of competition while the alternative hypothesis $H_A$ reflects a hypothesis of species grouping-specific effects of competition.
 Furthermore, in order to account for the spatial autocorrelation inherent to forest data in their estimates of out-of-sample model error, @allen_permutation_2020 use spatial cross-validation.
@@
 We introduce the `forestecology` R package providing methods and data for forest ecology model fitting and assessment, available on CRAN (https://cran.r-project.org/web/packages/forestecology/index.html) and on GitHub (https://github.com/rudeboybert/forestecology).
-The package implements all aspects of the model in Equation \ref{eq:model}: model fitting and generating fitted/predicted values, evaluating the effect of competitor species identity using permutation tests, and evaluating model performance using spatial cross-validation.
+The package implements all aspects of the model in Equation \ref{eq:model}: model fitting and generating predicted values, evaluating the effect of competitor species identity using permutation tests, and evaluating model performance using spatial cross-validation.
 
 The package designed with "tidy" design principles in mind [@wickham_welcome_2019].
-Much like many of the `tidyverse` component packages, `forestecology` is designed with verb-named functions that can be modularly composed in sequence using the pipe `%>%` operator [@bache_pipe_2020].
-As we articulate in Section \ref{casestudy}, these functions delineate the key steps in our analysis sequence.
-Furthermore, the inputs and outputs of nearly all of our functions use the same "simple features for R" data structures as implemented in the `sf` package for standardized and `tidyverse`-friendly support for spatial vector data [@pebesma_simple_2018]
+Much like all `tidyverse` packages, `forestecology` is has verb-named functions that can be modularly composed using the pipe `%>%` operator to complete all the necessary steps in the analysis sequence [@bache_pipe_2020].
+Furthermore, the inputs and outputs of most of our functions use the same "simple features for R" data structures from the `sf` package for standardized and `tidyverse`-friendly wrangling and visualizing of spatial data [@pebesma_simple_2018]
 
-Currently the package only implements the Bayesian linear regression model of tree growth based on neighborhood competition detailed in Equation \ref{eq:model}.
-As we demonstrate in Section \ref{model-fit-predict} however, the fitting of this model is self-contained in a single function `comp_bayes_lm()`.
-This function returns an object of S3 class type `comp_bayes_lm` with generic methods implemented to print, make predictions using, and plot all results.
-Therefore the package can be modularly extended to fit other models as long as they are coded into a function similar to `comp_bayes_lm()` and has equivalent generic methods implemented.
+Currently the package only implements the Bayesian linear regression model detailed in Equation \ref{eq:model}.
+As we demonstrate in Section \ref{model-fit-predict} however, the fitting of this model is self-contained in a single function `comp_bayes_lm()` which returns an object of S3 class type `comp_bayes_lm`.
+This class has generic methods implemented to print, make predictions using, and plot all results.
+Therefore the package can be modularly extended to fit other models as long as they are coded similarly as `comp_bayes_lm()` and have equivalent generic methods implemented.
 
@@
 
-We present a case-study of the `forestecology` package's functionality on data from the Smithsonian Conservation Biology Institute (SCBI) large forest dynamics plot in Front Royal, VA, USA, which is part of the ForestGEO global network of research sites [@bourg_initial_2013, @andersonteixeira_ctfs-forestgeo_2015] [@bourg_initial_2013].
+We present a case-study of the `forestecology` package's functionality on data from the Smithsonian Conservation Biology Institute (SCBI) large forest dynamics plot in Front Royal, VA, USA, part of the ForestGEO global network of research sites [@bourg_initial_2013, @andersonteixeira_ctfs-forestgeo_2015] [@bourg_initial_2013].
 The 25.6 ha (640 x 400 m) plot is located at the intersection of three of the major physiographic provinces of the eastern US---the Blue Ridge, Ridge and Valley, and Piedmont provinces---and is adjacent to the northern end of Shenandoah National Park.
 
-The `forestecology` package has the following ecological goals: 1) to evaluate the effect of competitor species identity using permutation tests and 2) to evaluate model performance using spatial cross-validation.
-To achieve these goals, we outline a basic analysis sequence comprising of these four main steps:
+The package has the following ecological goals: to evaluate i) the effect of competitor species identity using permutation tests and ii) model performance using spatial cross-validation.
+We outline a basic analysis sequence comprising of the following four main steps:
 
@@
     1. Fit model.
-    1. Compute fitted/predicted values.
+    1. Compute predicted values.
     1. Visualize posterior distributions.
@@
 The first step is to compute the growth of trees using data from two censuses.
-`compute_growth()` computes average annual growth assuming census data that roughly follows ForestGEO standards.
-Despite such standards, minor variations will still exist between sites, thereby necessitating some data wrangling and checking.
-For example, the SCBI site records all diameters at breast height (DBH) in millimeters [@bourg_initial_2013], whereas the Michigan Big Woods site records them in centimeters [@allen_michigan_2020].
+`compute_growth()` computes the average annual growth based on census data that roughly follows ForestGEO standards.
+Despite such standards, minor variations will still exist between sites, thereby necessitating some data wrangling.
+For example, the SCBI site records all DBH values in millimeters [@bourg_initial_2013], whereas the Michigan Big Woods site records them in centimeters [@allen_michigan_2020].
 
-We first load both 2008 and 2014 SCBI census data `.csv` files as they existed on GitHub on November 20, 2020 [@gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020] and perform some data wrangling to both data sets.
-We then only consider a 9 ha subsection of the 25.6 ha of the SCBI site, `gx` from 0--300 instead of 0--400 and `gy` from 300--600 instead of 0--640, in order to speed up computation for purposes of this example.
+We first load both 2008 and 2014 SCBI census `.csv` files as they existed on GitHub on 2020/11/20 and perform some data wrangling to both data sets [@gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020].
+We then only consider a 9 ha subsection of the 25.6 ha of the SCBI site in order to speed up computation for this example: `gx` from 0--300 instead of 0--400 and `gy` from 300--600 instead of 0--640.
 
@@
 
-These two data frames are then supplied as arguments to `compute_growth()`, along with the `id` argument that specifies the variable that uniquely identifies each tree-stem.
-Note furthermore that we discard all resprouts in the later census (those with `code == R`), since we are only interested in the diameter growth of surviving, and not resprouted, stems.
+These two data frames are then used as inputs to `compute_growth()`, along with the `id` argument specifying the variable that uniquely identifies each tree-stem.
+Note that we also discard all resprouts in the later census with `code == R`, since we are only interested in the growth of surviving, and not resprouted, stems.
 
@@
 
-The output `growth_scbi` is a single data frame of class `sf` that includes variables `growth`, the average annual growth in DBH (cm $\cdot$ y^-1^) for all stems that were alive at both time points, and `geometry`, the `sf` package's encoding of geolocations of type `<POINT>`.
-In addition the species variable `sp` is returned as a factor.^[In our spatial cross-validation algorithm in Section \ref{spatial-cross-validation} issues can occur when rare species do not occur in the training set, but then are encountered in the test set.
-This risk is mitigated by representing `sp` as a factor variable, which has a complete list of all levels of the categorical variable.] Given that `growth_scbi` is of class `sf`, it can be easily plotted in `ggplot2` using the `geom_sf()` geometry as seen in Figure \ref{fig:scbi-trees}.
+The output `growth_scbi` is a data frame of class `sf` that includes variables the average annual `growth` in DBH (cm $\cdot$ y^-1^) for all stems that were alive at both time points, the `sf` package's encoding of geolocations of `geometry` type `<POINT>`, and the species variable `sp` converted to a factor.
+Given that `growth_scbi` is of class `sf`, it can be easily plotted in `ggplot2` using `geom_sf()` as seen in Figure \ref{fig:scbi-trees}.
 
@@
 
-The next step is to add additional spatial information to `growth_scbi`.
+The next step is to add spatial information to `growth_scbi`.
 The first element we add is a "buffer region" to the periphery of the study region.
 Since some of our model's explanatory variables are cumulative, we must ensure that all trees being modeled are not biased to have different neighbor structures.
-This is of concern for trees at the boundary of study regions, for which all neighbors will not be included in the censused stems.
-In order to account for such edge effects, only trees that are not part of this buffer region, i.e.
+This is of concern for trees at the boundary of the study region whose neighbors will not all be included in the censused stems.
+To account for such edge effects only trees that are not part of this buffer region, i.e.
 are part of the interior of the study region, will have their growth modeled [@waller_applied_2004].
 
-Our model of interspecific competition relies on a spatial definition of who the competitor trees are for focal trees of interest: all trees within a distance `comp_dist` of a focal tree are considered its competitors.
-In our case we set this value at 7.5m, a value informed by other studies [@canham_neighborhood_2004, @uriarte_spatially_2004, @canham_neighborhood_2006].
-Using this value along with a manually constructed `sf` object representation of the study region's boundary, we apply the `add_buffer_variable()` to `growth_scbi` to add a `buffer` boolean variable.
-All trees with `buffer` as `FALSE` will be our focal trees whose growth will be modeled, whereas those with `TRUE` will only be considered as competitor trees.
+Our model of interspecific competition relies on a spatial definition of who competitor trees are: all trees within a distance `comp_dist` of a focal tree.
+In our case we set `comp_dist` to 7.5m, a value informed by other studies [@canham_neighborhood_2004, @uriarte_spatially_2004, @canham_neighborhood_2006].
+We use `comp_dist` and a manually constructed `sf` representation of the study region's boundary as inputs to `add_buffer_variable()` to add a `buffer` boolean variable to `growth_scbi`.
+All trees with `buffer` equal to `FALSE` will be our focal trees whose growth will be modeled, whereas those with `TRUE` will only act as competitor trees.
 
@@
 
-The second element of spatial information are blocks corresponding to folds of a spatial cross-validation algorithm used to estimate out-of-sample model error.
-Conventional cross-validation algorithms assign observations to folds by randomly resampling individual observations.
-However, many of these algorithms assume that the observations are independent.
-In the case of forest census data, observations exhibit spatial autocorrelation.
-We therefore incorporate this spatial dependence into the cross-validation algorithm with our spatial blocks of trees [@roberts_cross-validation_2017 @pohjankukka_estimating_2017].
+The second element of spatial information we add are blocks corresponding to folds of a spatial cross-validation algorithm.
+Conventional cross-validation algorithms assign observations to folds by randomly resampling individual observations that are assumed independent.
+In the case of forest census data however, observations exhibit spatial autocorrelation.
+We therefore incorporate this dependence into the cross-validation algorithm by resampling spatial blocks of trees [@roberts_cross-validation_2017 @pohjankukka_estimating_2017].
 
-In the example below, we first manually define four folds that partition the study region as an `sf` object.
-We then use the output of the `spatialBlock()` function from the `blockCV` package to associate each tree in `growth_scbi` to the correct fold `foldID` [@valavi_blockcv_2019].
-^[In the Supporting Information we present an example where the folds themselves are also created automatically using `spatialBlock()` given a specified `cv_block_size`, as opposed to manually as in the example.] Figure \ref{fig:scbi-spatial-information} illustrates the net effect of adding these two elements of spatial information to `growth_scbi`.
+We first manually an `sf` object defining four folds that partition the study region.
+We then use the output of the `spatialBlock()` function from the `blockCV` package to associate each tree in `growth_scbi` to the correct `foldID` [@valavi_blockcv_2019].^[In the Supporting Information we present an example where the folds themselves are created automatically, as opposed to manually as in the example.] This `foldID` variable will be used in Section \ref{spatial-cross-validation}.
+
+Figure \ref{fig:scbi-spatial-information} illustrates the net effect of adding these two elements of spatial information to `growth_scbi`.
 
@@
 More specifically, identify all trees that are not part of the buffer region, have a valid `growth` measurement, and have at least one neighbor within `r comp_dist`m.
-`create_focal_vs_comp()` returns a new data frame of type `sf`.
-On top of previously detailed arguments `comp_dist` and `id`, `create_focal_vs_comp()` also requires an `sf` object representation of the spatial cross-validation blocks/folds as seen in Section \ref{spatial-information}.
+We do this using `create_focal_vs_comp()`, which takes the previously detailed arguments `comp_dist` and `id` as well as the `sf` representation of the spatial cross-validation blocks and returns a new data frame `focal_vs_comp_scbi`.
 
@@
 The resulting `focal_vs_comp_scbi` has `r nrow(focal_vs_comp_scbi)` rows, representing the subset of the `r nrow(growth_scbi)` trees in `growth_scbi` that will be considered as focal trees.
-Two new variables `focal_ID` and `focal_sp` relate to tree-stem identification and species information.
-Most notably however is a new variable `comp` which contains information on all competitor trees saved in `tidyr` package list-column format [@tidyr_package].
-We flatten the `comp` list-column for the tree with `focal_ID` 4 in the first row, here a `tibble [20 × 4]`, into regular columns using `unnest()` from the `tidyr` package.
+The variables `focal_ID` and `focal_sp` relate to tree-stem identification and species information.
+Most notably however is the variable `comp`, which contains information on all competitor trees saved in `tidyr` package list-column format [@tidyr_package].
+To inspect this information, we flatten the `comp` list-column for the tree with `focal_ID` 4 in the first row, here a `tibble [20 × 4]`, into regular columns using `unnest()` from the `tidyr` package.
 
@@
 
-We observe that for this focal tree, we have 4 variables of information on its 20 competitor trees: their unique tree-stem ID number, their distance to the focal tree (all $\leq$ `r comp_dist`), their species, and their basal area (in m$^2$) calculated as $\frac{\pi \times (\text{DBH/2})^2}{10000}$ where $DBH$ is the value from the earlier of the two censuses in cm.
-Saving our focal versus competitor information in list-column minimizes redundancy since we do not repeat information on the focal tree 20 times.
-The spatial distribution of these trees is visualized in Figure \ref{fig:scbi-focal-vs-comp-map}.
+We observe 4 variables describing 20 competitor trees: their unique tree-stem ID, their distance to the focal tree (all $\leq$ `r comp_dist`), their species, and their basal area (in m$^2$) calculated as $\frac{\pi \times (\text{DBH/2})^2}{10000}$ for $DBH$ in cm from the earlier census.
+Saving competitor information in list-column format minimizes redundancy since we do not need to repeat information on the focal tree 20 times.
+We visualize the spatial distribution of these trees in Figure \ref{fig:scbi-focal-vs-comp-map}.
 
@@
 
-The final step is to fit a model for the growth of all focal trees.
-We fit the competition Bayesian linear regression model outlined in Equation \ref{eq:model} using `comp_bayes_lm()`, which has an option to specify prior distributions on all parameters of interest (chosen to be the defaults specified in `?comp_bayes_lm`).
+The final step is to fit the competition Bayesian linear regression model for tree growth outlined in Equation \ref{eq:model} using `comp_bayes_lm()`.
+This function has an option to specify prior distributions on all parameters of interest, chosen here to be the defaults detailed in `?comp_bayes_lm`.
 
@@
 
-The returned `comp_bayes_lm_scbi` output is an object of S3 class type `comp_bayes_lm` containing the posterior values of all parameters in our competition Bayesian linear regression.
-This class of object includes three generic methods.
-First, the generic for `print()` displays the names of all prior & posterior parameters along with the model formula:
+The resutling `comp_bayes_lm_scbi` is an object of S3 class type `comp_bayes_lm` containing the posterior values of all parameters of our model.
+Furthermore, this class of object includes generics for three methods.
+First, the generic for `print()` displays the names of all prior and posterior parameters along with the model formula:
 
@@
 
-Next, the generic for `predict()` takes the posterior parameter values in `comp_bayes_lm_scbi` and the predictor variables in `newdata` and outputs a vector `growth_hat` of fitted/predicted values $\widehat{y}$ of the DBH for each focal tree computed from the posterior predictive distribution.
+Next, the generic for `predict()` takes the posterior parameter values in `comp_bayes_lm_scbi` and a `newdata` data frame and outputs a vector `growth_hat` of predicted values $\widehat{y}$ of the DBH for each focal tree computed from the posterior predictive distribution.
 
@@
 
-We then compare the observed and fitted/predicted growths to compute the root mean squared error (RMSE) of our model fit.
+We then compare the observed and predicted growths to compute the root mean squared error (RMSE) of our model fit.
 
@@
 
-Lastly, the generic for `ggplot2::autoplot()` allows us to plot the posterior distribution of all parameters in Figure \ref{fig:scbi-posterior-viz}.
+Lastly, the generic for `ggplot2::autoplot()` allows us to visualize the posterior distribution of all parameters, as seen in Figure \ref{fig:scbi-posterior-viz}.
+Setting `type` to  `"intercepts"` and `"dbh_slopes"` returns species-specific posterior distributions for $\beta_{0,j}$ and $\beta_{dbh,j}$ respectively, while setting `type = "competition"` returns competition coefficients $\lambda_{j,k}$.
 
@@
 
-# Combine plots using patchwork
+# Combine plots using the patchwork package
 (plot1 | plot2) / plot3
@@
 
-These plots visualize the posterior distributions of parameters from Equation \ref{eq:model}.
-For many package users they will be of interest because they give insight into the species-specific competitive interactions.
-Setting `type = "intercepts"` returns species-specific posterior distributions for $\beta_{0,j}$ and `type = "dbh_slopes"` for $\beta_{dbh,j}$.
-Setting `type = "competition"` returns competition coefficients $\lambda_{j,k}$ where negative values indicate a competitor species which slows the growth of a focal species.
-Here, for example, we see that `litu` tulip poplars have a strong negative effect on the growth of conspecifics but relatively lesser effect on neighbors of the other two species.
+For many users the visualizations relating to $\lambda_{j,k}$ will be of particular interest as they provide insight into species competitive interactions where negative values indicate a competitor species which slows the growth of a focal species.
+Here, for example, we see that tulip poplars (`litu`) have a strong negative effect on the growth of conspecifics but relatively lesser effect on neighbors of the other two species.
 
 Currently the `forestecology` package can only fit the competition Bayesian linear regression model outlined in Equation \ref{eq:model}.
-However, it can be extended to any model implemented in a function similar to `comp_bayes_lm()` that uses data frames of similar format to `focal_vs_comp` as input.
+However, it can be extended to any model as long as it is implemented in a function similar to `comp_bayes_lm()`.
 
-## Evaluate the effect of competitor species identity using permutation tests
+## Evaluate the effect of competitor species identity using permutation tests {#permutation-test}
 
-To evaluate the effect of competitor species identity, we use the four steps of our analysis sequence answer along with a permutation test: Under a null hypothesis where competitor species identity does not matter, we permute/shuffle this variable within each focal tree, compute the RMSE (the test statistic of interest), repeat this process several times to construct a null distribution of the RMSE, and compare it to the observed RMSE to assess significance.
-Going back to our example in Section \ref{focal-vs-comp} of focal tree with `focal_ID` 4 and its 20 competitors, the permutation test randomly resamples only the `comp_sp` variable with replacement, leaving all other variables intact.
-The resampling with replacement is nested within each focal tree in order to preserve neighborhood structure.
-We once again use `comp_bayes_lm()` as in Section \ref{model-fit-predict}, but with `run_shuffle = TRUE`.
+To evaluate the effect of competitor species identity, we use the above four steps along with a permutation test.
+Under a null hypothesis where competitor species identity does not matter, we permute the competitor species identity within each focal tree, compute the RMSE test statistic, repeat this process several times to construct a null distribution, and finally compare it to the observed RMSE to assess significance.
+Going back to our example in Section \ref{focal-vs-comp} of focal tree with `focal_ID` 4 and its 20 competitors, the permutation test randomly resamples only the `comp_sp` variable without replacement, leaving all other variables intact.
+The resampling without replacement is nested within each focal tree in order to preserve neighborhood structure.
+We once again use `comp_bayes_lm()`, but this time setting `run_shuffle = TRUE`.
 
@@
 
-The resulting RMSE of `r model_rmse_shuffle` based on the permutation test is `r ifelse(model_rmse_shuffle < model_rmse, "smaller", "larger")` than the earlier RMSE of `r model_rmse`, suggesting that models that do incorporate competitor species identity better fit the data.
+The resulting permutation test RMSE of `r model_rmse_shuffle` is `r ifelse(model_rmse_shuffle < model_rmse, "smaller", "larger")` than the earlier RMSE of `r model_rmse`, suggesting that models that do incorporate competitor species identity better fit the data.
 
@@
 
-We answer the second of our two questions: how can we obtain an accurate estimate of model performance/error?
-The model fits and predictions in Section \ref{model-fit-predict} all suffer from a common failing: they use the same data to both fit the model and to assess the model's performance using the RMSE.
-As argued by @roberts_cross-validation_2017, this can lead to overly optimistic assessments of model quality as the models can be overfit, in particular in situations where spatial-autocorrelation is present.
-To mitigate the effects of such overfitting, we use a spatially block cross-validation algorithm.
+To evaluate model performance, we use spatial cross-validation.
+The model fits and predictions in Section \ref{model-fit-predict} use the same data to both fit the model and to assess the model's performance.
+Given the inherent spatial-autocorrelation of our data, this can potentially lead to potentially overfit models [@roberts_cross-validation_2017].
+To mitigate this potential overfitting, we use the spatial cross-validation blocking scheme encoded in the `foldID` variable from Section \ref{spatial-information} and visualized in Figure \ref{fig:scbi-spatial-information}.
 
-To this end, we use the `foldID` variable defined in Section \ref{spatial-information} whereby all focal trees are assigned to one of `r n_fold` spatially contiguous blocks that act as folds in our cross-validation routine.
-Figure \ref{fig:scbi-spatial-cross-validation-schematic} presents a schematic illustrating this scheme.
-We fit the model to all focal trees in the training set, apply the model to all focal trees in the test set to compute fitted/predicted values, and compute the RMSE of the observed versus predicted growths.
-We repeat this procedure 3 more times with each of the three remaining folds acting as the test set and then average all four resulting RMSE's.
-Furthermore, in order to maintain spatial independence between the test and training set, a "fold buffer" that extend outwards from the boundary of the test set is computed; all trees falling within this fold buffer are excluded from the training set.
+At each iteration of our cross-validation, one fold will act as the test set with the remaining three acting as the training set.
+We fit the model to all focal trees in the training set, apply the model to all focal trees in the test set, compute predicted values, and compute the RMSE.
+Furthermore, to maintain spatial independence between the test and training set, a "fold buffer" that extend outwards from the boundary of the test set is computed; all trees falling within this fold buffer are excluded from the training set (see Figure \ref{fig:scbi-spatial-cross-validation-schematic}).
 
-```{r scbi-spatial-cross-validation-schematic, out.width="66%", fig.cap= "Schematic of spatial cross-validation: Using the k = 1 fold (bottom-left) as the test set, k = 2 through 4 as the training set, along with a \"fold buffer.\"", echo = FALSE}
+```{r scbi-spatial-cross-validation-schematic, out.width="66%", fig.cap= "Schematic of spatial cross-validation: Using the k = 1 fold (bottom-left) as the test set, k = 2 through 4 as the training set, along with a fold buffer.", echo = FALSE}
 fold_number <- 1
@@
 
-This algorithm is implemented in `run_cv()`, which is a wrapper function to both `comp_bayes_lm()` that fits the model and `predict()` that returns fitted/predicted values.
-We once again compute the RMSE.
+This is repeated for each of the four folds acting as the test set and then the four resulting RMSE's are averaged to provide a single estimate of model error.
+This algorithm is implemented in `run_cv()`, which acts as a wrapper function to both `comp_bayes_lm()` that fits the model and `predict()` that returns predicted values.
 
@@
 
-The resulting RMSE of `r model_rmse_cv` computed using cross-validation is `r ifelse(model_rmse_cv > model_rmse, "larger", "smaller")` than the earlier RMSE of `r model_rmse`, suggesting that models that do not take the inherent spatial autocorrelation of the data into account generate error estimates that are overly optimistic; in our case RMSE's that are too low.
+The resulting RMSE of `r model_rmse_cv` computed using cross-validation is `r ifelse(model_rmse_cv > model_rmse, "larger", "smaller")` than the earlier RMSE of `r model_rmse`, suggesting that models that do not take spatial autocorrelation account generate moel error estimates that are overly optimistic.
+In our case, RMSE values that are too low.
 
@@
 
-`run_cv()` also accepts the `run_shuffle` argument.
-This permutes the competitor species, as described above, but does so when calculating predicted growth with the cross validated scheme.
-Figure \ref{fig:scbi-simulation} compares model performance when permuting competitor species and calculating RMSE with and without cross-validation.
-Without cross-validation the competitor identity did matter, the non-permuted competitor species had a much lower RMSE than the permuted one.
-But once we include the spatial cross-validation, this improvement disappears.
-These results suggest that in this 9 ha subplot of the SCBI plot competitive interactions do not depend on the identity of the competitor, which is the opposite of what has been observed in other locations (@allen_permutation_2020 @uriarte_spatially_2004).
-This highlights the importance of cross-validation, without it the model was overfit.
+`run_cv()` also accepts the `run_shuffle` argument in order to permute competitor species identity as described in Section \ref{permutation-test}.
+Figure \ref{fig:scbi-simulation} compares model performance for 49 permutations of competitor species and calculating RMSE, both with and without cross-validation.
+Without cross-validation, competitor species identity did matter as the observed RMSE was lower than the permutation null distribution of RMSE.
+However, once we incorporate spatial cross-validation, this improvement disappears.
+These results suggest that in this 9 ha subplot of the SCBI plot competitive interactions do not depend on the identity of the competitor, which is the opposite of what has been observed in other locations [@allen_permutation_2020 @uriarte_spatially_2004].
+Furthermore, the larger cross-validated RMSE's are indicative of the importance of being vigilant against model overfitting.
 
@@
 While the package is designed with ForestGEO plot data in mind, we envision that it can be modified to work on i) any single large, mapped forest plot in which at least two measurements of each individual have been taken, e.g.
-the US Forest Service Forest Inventory and Analysis plots or ii) more generally to model interactions of any community of mapped sessile organisms [@smith_forest_2002].
-In future versions of `forestecology` we also hope to make it possible to model plant mortality in addition to plant growth.
+the US Forest Service Forest Inventory and Analysis plots, or ii) more generally to model interactions of any community of mapped sessile organisms [@smith_forest_2002].
+In future versions of `forestecology` we also hope to include models that account for tree mortality in addition to tree growth.
 The package follows the guidelines for `tidy` data, leverages the `sf` package for spatial data, and S3 open-oriented model structure.

### 2021-03-09 paper/forestecology.tex
3421 -> 3139 via only quick pass of text

@@
 groups; and let \(k = 1, \ldots, K\) index all \(K\) ``competitor''
-species groups. We model the average annual growth in diameter at breast
-height (DBH) \(y_{ij}\) (in centimeters per year) of the \(i^{th}\) tree
-of focal species group \(j\) as a linear model \(f\) of the covariates
-\(\vec{x}_{ij}\)
+species groups. They use the following linear model the average annual
+growth in diameter at breast height (DBH) \(y_{ij}\) (in
+centimeters/year) of the \(i^{th}\) tree of focal species group \(j\):
 
@@
 In order to evaluate whether competitor species identity matters,
-\citet{allen_permutation_2020} run a permutation test where under the
-null hypothesis the species identity of all competitors of a focal tree
-can be permuted/shuffled:
+\citet{allen_permutation_2020} run a permutation test where a null
+hypothesis of no species grouping-specific effects of competition is
+assumed, and thus the species identity of all competitors can be
+permuted/shuffled:
 
@@
 
-where the null hypothesis \(H_0\) reflects a hypothesis of no species
-grouping-specific effects of competition while the alternative
-hypothesis \(H_A\) reflects a hypothesis of species grouping-specific
-effects of competition. Furthermore, in order to account for the spatial
-autocorrelation inherent to forest data in their estimates of
-out-of-sample model error, \citet{allen_permutation_2020} use spatial
-cross-validation. Estimates of model error that do not account for this
-spatial dependency tend to underestimate the true model error
+Furthermore, in order to account for the spatial autocorrelation
+inherent to forest data in their estimates of out-of-sample model error,
+\citet{allen_permutation_2020} use spatial cross-validation. Estimates
+of model error that do not account for this spatial dependency tend to
+underestimate the true model error
 \citep{roberts_cross-validation_2017}.
@@
 package implements all aspects of the model in Equation \ref{eq:model}:
-model fitting and generating fitted/predicted values, evaluating the
-effect of competitor species identity using permutation tests, and
-evaluating model performance using spatial cross-validation.
+model fitting and generating predicted values, evaluating the effect of
+competitor species identity using permutation tests, and evaluating
+model performance using spatial cross-validation.
 
 The package designed with ``tidy'' design principles in mind
-\citep{wickham_welcome_2019}. Much like many of the \texttt{tidyverse}
-component packages, \texttt{forestecology} is designed with verb-named
-functions that can be modularly composed in sequence using the pipe
-\texttt{\%\textgreater{}\%} operator \citep{bache_pipe_2020}. As we
-articulate in Section \ref{casestudy}, these functions delineate the key
-steps in our analysis sequence. Furthermore, the inputs and outputs of
-nearly all of our functions use the same ``simple features for R'' data
-structures as implemented in the \texttt{sf} package for standardized
-and \texttt{tidyverse}-friendly support for spatial vector data
+\citep{wickham_welcome_2019}. Much like all \texttt{tidyverse} packages,
+\texttt{forestecology} is has verb-named functions that can be modularly
+composed using the pipe \texttt{\%\textgreater{}\%} operator to complete
+all the necessary steps in the analysis sequence
+\citep{bache_pipe_2020}. Furthermore, the inputs and outputs of most of
+our functions use the same ``simple features for R'' data structures
+from the \texttt{sf} package for standardized and
+\texttt{tidyverse}-friendly wrangling and visualizing of spatial data
 \citep{pebesma_simple_2018}
@@
 Currently the package only implements the Bayesian linear regression
-model of tree growth based on neighborhood competition detailed in
-Equation \ref{eq:model}. As we demonstrate in Section
+model detailed in Equation \ref{eq:model}. As we demonstrate in Section
 \ref{model-fit-predict} however, the fitting of this model is
-self-contained in a single function \texttt{comp\_bayes\_lm()}. This
-function returns an object of S3 class type \texttt{comp\_bayes\_lm}
-with generic methods implemented to print, make predictions using, and
+self-contained in a single function \texttt{comp\_bayes\_lm()} which
+returns an object of S3 class type \texttt{comp\_bayes\_lm}. This class
+has generic methods implemented to print, make predictions using, and
 plot all results. Therefore the package can be modularly extended to fit
-other models as long as they are coded into a function similar to
-\texttt{comp\_bayes\_lm()} and has equivalent generic methods
+other models as long as they are coded similarly as
+\texttt{comp\_bayes\_lm()} and have equivalent generic methods
 implemented.
@@
 Institute (SCBI) large forest dynamics plot in Front Royal, VA, USA,
-which is part of the ForestGEO global network of research sites
+part of the ForestGEO global network of research sites
 \citep[\citet{andersonteixeira_ctfs-forestgeo_2015}]{bourg_initial_2013}
@@
 
-The \texttt{forestecology} package has the following ecological goals:
-1) to evaluate the effect of competitor species identity using
-permutation tests and 2) to evaluate model performance using spatial
-cross-validation. To achieve these goals, we outline a basic analysis
-sequence comprising of these four main steps:
+The package has the following ecological goals: to evaluate i) the
+effect of competitor species identity using permutation tests and ii)
+model performance using spatial cross-validation. We outline a basic
+analysis sequence comprising of the following four main steps:
 
@@
   \item
-    Compute fitted/predicted values.
+    Compute predicted values.
   \item
@@
 The first step is to compute the growth of trees using data from two
-censuses. \texttt{compute\_growth()} computes average annual growth
-assuming census data that roughly follows ForestGEO standards. Despite
+censuses. \texttt{compute\_growth()} computes the average annual growth
+based on census data that roughly follows ForestGEO standards. Despite
 such standards, minor variations will still exist between sites, thereby
-necessitating some data wrangling and checking. For example, the SCBI
-site records all diameters at breast height (DBH) in millimeters
-\citep{bourg_initial_2013}, whereas the Michigan Big Woods site records
-them in centimeters \citep{allen_michigan_2020}.
+necessitating some data wrangling. For example, the SCBI site records
+all DBH values in millimeters \citep{bourg_initial_2013}, whereas the
+Michigan Big Woods site records them in centimeters
+\citep{allen_michigan_2020}.
 
-We first load both 2008 and 2014 SCBI census data \texttt{.csv} files as
-they existed on GitHub on November 20, 2020
-\citep{gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020} and perform
-some data wrangling to both data sets. We then only consider a 9 ha
-subsection of the 25.6 ha of the SCBI site, \texttt{gx} from 0--300
-instead of 0--400 and \texttt{gy} from 300--600 instead of 0--640, in
-order to speed up computation for purposes of this example.
+We first load both 2008 and 2014 SCBI census \texttt{.csv} files as they
+existed on GitHub on 2020/11/20 and perform some data wrangling to both
+data sets \citep{gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020}.
+We then only consider a 9 ha subsection of the 25.6 ha of the SCBI site
+in order to speed up computation for this example: \texttt{gx} from
+0--300 instead of 0--400 and \texttt{gy} from 300--600 instead of
+0--640.
 
@@
 
-These two data frames are then supplied as arguments to
-\texttt{compute\_growth()}, along with the \texttt{id} argument that
-specifies the variable that uniquely identifies each tree-stem. Note
-furthermore that we discard all resprouts in the later census (those
-with \texttt{code\ ==\ R}), since we are only interested in the diameter
-growth of surviving, and not resprouted, stems.
+These two data frames are then used as inputs to
+\texttt{compute\_growth()}, along with the \texttt{id} argument
+specifying the variable that uniquely identifies each tree-stem. Note
+that we also discard all resprouts in the later census with
+\texttt{code\ ==\ R}, since we are only interested in the growth of
+surviving, and not resprouted, stems.
 
@@
 
-The output \texttt{growth\_scbi} is a single data frame of class
-\texttt{sf} that includes variables \texttt{growth}, the average annual
-growth in DBH (cm \(\cdot\) y\textsuperscript{-1}) for all stems that
-were alive at both time points, and \texttt{geometry}, the \texttt{sf}
-package's encoding of geolocations of type
-\texttt{\textless{}POINT\textgreater{}}. In addition the species
-variable \texttt{sp} is returned as a factor.\footnote{In our spatial
-  cross-validation algorithm in Section \ref{spatial-cross-validation}
-  issues can occur when rare species do not occur in the training set,
-  but then are encountered in the test set. This risk is mitigated by
-  representing \texttt{sp} as a factor variable, which has a complete
-  list of all levels of the categorical variable.} Given that
+The output \texttt{growth\_scbi} is a data frame of class \texttt{sf}
+that includes variables the average annual \texttt{growth} in DBH (cm
+\(\cdot\) y\textsuperscript{-1}) for all stems that were alive at both
+time points, the \texttt{sf} package's encoding of geolocations of
+\texttt{geometry} type \texttt{\textless{}POINT\textgreater{}}, and the
+species variable \texttt{sp} converted to a factor. Given that
 \texttt{growth\_scbi} is of class \texttt{sf}, it can be easily plotted
-in \texttt{ggplot2} using the \texttt{geom\_sf()} geometry as seen in
-Figure \ref{fig:scbi-trees}.
+in \texttt{ggplot2} using \texttt{geom\_sf()} as seen in Figure
+\ref{fig:scbi-trees}.
 
@@
 
-The next step is to add additional spatial information to
-\texttt{growth\_scbi}. The first element we add is a ``buffer region''
-to the periphery of the study region. Since some of our model's
-explanatory variables are cumulative, we must ensure that all trees
-being modeled are not biased to have different neighbor structures. This
-is of concern for trees at the boundary of study regions, for which all
-neighbors will not be included in the censused stems. In order to
-account for such edge effects, only trees that are not part of this
-buffer region, i.e.~are part of the interior of the study region, will
-have their growth modeled \citep{waller_applied_2004}.
+The next step is to add spatial information to \texttt{growth\_scbi}.
+The first element we add is a ``buffer region'' to the periphery of the
+study region. Since some of our model's explanatory variables are
+cumulative, we must ensure that all trees being modeled are not biased
+to have different neighbor structures. This is of concern for trees at
+the boundary of the study region whose neighbors will not all be
+included in the censused stems. To account for such edge effects only
+trees that are not part of this buffer region, i.e.~are part of the
+interior of the study region, will have their growth modeled
+\citep{waller_applied_2004}.
 
 Our model of interspecific competition relies on a spatial definition of
-who the competitor trees are for focal trees of interest: all trees
-within a distance \texttt{comp\_dist} of a focal tree are considered its
-competitors. In our case we set this value at 7.5m, a value informed by
-other studies \citep[\citet{uriarte_spatially_2004},
-\citet{canham_neighborhood_2006}]{canham_neighborhood_2004}. Using this
-value along with a manually constructed \texttt{sf} object
-representation of the study region's boundary, we apply the
-\texttt{add\_buffer\_variable()} to \texttt{growth\_scbi} to add a
-\texttt{buffer} boolean variable. All trees with \texttt{buffer} as
-\texttt{FALSE} will be our focal trees whose growth will be modeled,
-whereas those with \texttt{TRUE} will only be considered as competitor
-trees.
+who competitor trees are: all trees within a distance
+\texttt{comp\_dist} of a focal tree. In our case we set
+\texttt{comp\_dist} to 7.5m, a value informed by other studies
+\citep[\citet{uriarte_spatially_2004},
+\citet{canham_neighborhood_2006}]{canham_neighborhood_2004}. We use
+\texttt{comp\_dist} and a manually constructed \texttt{sf}
+representation of the study region's boundary as inputs to
+\texttt{add\_buffer\_variable()} to add a \texttt{buffer} boolean
+variable to \texttt{growth\_scbi}. All trees with \texttt{buffer} equal
+to \texttt{FALSE} will be our focal trees whose growth will be modeled,
+whereas those with \texttt{TRUE} will only act as competitor trees.
 
@@
 
-The second element of spatial information are blocks corresponding to
-folds of a spatial cross-validation algorithm used to estimate
-out-of-sample model error. Conventional cross-validation algorithms
-assign observations to folds by randomly resampling individual
-observations. However, many of these algorithms assume that the
-observations are independent. In the case of forest census data,
-observations exhibit spatial autocorrelation. We therefore incorporate
-this spatial dependence into the cross-validation algorithm with our
-spatial blocks of trees \citep[
+The second element of spatial information we add are blocks
+corresponding to folds of a spatial cross-validation algorithm.
+Conventional cross-validation algorithms assign observations to folds by
+randomly resampling individual observations that are assumed
+independent. In the case of forest census data however, observations
+exhibit spatial autocorrelation. We therefore incorporate this
+dependence into the cross-validation algorithm by resampling spatial
+blocks of trees \citep[
 \citet{pohjankukka_estimating_2017}]{roberts_cross-validation_2017}.
 
-In the example below, we first manually define four folds that partition
-the study region as an \texttt{sf} object. We then use the output of the
+We first manually an \texttt{sf} object defining four folds that
+partition the study region. We then use the output of the
 \texttt{spatialBlock()} function from the \texttt{blockCV} package to
-associate each tree in \texttt{growth\_scbi} to the correct fold
-\texttt{foldID} \citep{valavi_blockcv_2019}. \footnote{In the Supporting
-  Information we present an example where the folds themselves are also
-  created automatically using \texttt{spatialBlock()} given a specified
-  \texttt{cv\_block\_size}, as opposed to manually as in the example.}
+associate each tree in \texttt{growth\_scbi} to the correct
+\texttt{foldID} \citep{valavi_blockcv_2019}.\footnote{In the Supporting
+  Information we present an example where the folds themselves are
+  created automatically, as opposed to manually as in the example.} This
+\texttt{foldID} variable will be used in Section
+\ref{spatial-cross-validation}.
+
 Figure \ref{fig:scbi-spatial-information} illustrates the net effect of
@@
 part of the buffer region, have a valid \texttt{growth} measurement, and
-have at least one neighbor within 7.5m.
-\texttt{create\_focal\_vs\_comp()} returns a new data frame of type
-\texttt{sf}. On top of previously detailed arguments \texttt{comp\_dist}
-and \texttt{id}, \texttt{create\_focal\_vs\_comp()} also requires an
-\texttt{sf} object representation of the spatial cross-validation
-blocks/folds as seen in Section \ref{spatial-information}.
+have at least one neighbor within 7.5m. We do this using
+\texttt{create\_focal\_vs\_comp()}, which takes the previously detailed
+arguments \texttt{comp\_dist} and \texttt{id} as well as the \texttt{sf}
+representation of the spatial cross-validation blocks and returns a new
+data frame \texttt{focal\_vs\_comp\_scbi}.
 
@@
 the subset of the 7954 trees in \texttt{growth\_scbi} that will be
-considered as focal trees. Two new variables \texttt{focal\_ID} and
+considered as focal trees. The variables \texttt{focal\_ID} and
 \texttt{focal\_sp} relate to tree-stem identification and species
-information. Most notably however is a new variable \texttt{comp} which
+information. Most notably however is the variable \texttt{comp}, which
 contains information on all competitor trees saved in \texttt{tidyr}
-package list-column format \citep{tidyr_package}. We flatten the
-\texttt{comp} list-column for the tree with \texttt{focal\_ID} 4 in the
-first row, here a \texttt{tibble\ {[}20\ ×\ 4{]}}, into regular columns
-using \texttt{unnest()} from the \texttt{tidyr} package.
+package list-column format \citep{tidyr_package}. To inspect this
+information, we flatten the \texttt{comp} list-column for the tree with
+\texttt{focal\_ID} 4 in the first row, here a
+\texttt{tibble\ {[}20\ ×\ 4{]}}, into regular columns using
+\texttt{unnest()} from the \texttt{tidyr} package.
 
@@
 
-We observe that for this focal tree, we have 4 variables of information
-on its 20 competitor trees: their unique tree-stem ID number, their
-distance to the focal tree (all \(\leq\) 7.5), their species, and their
-basal area (in m\(^2\)) calculated as
-\(\frac{\pi \times (\text{DBH/2})^2}{10000}\) where \(DBH\) is the value
-from the earlier of the two censuses in cm. Saving our focal versus
-competitor information in list-column minimizes redundancy since we do
-not repeat information on the focal tree 20 times. The spatial
-distribution of these trees is visualized in Figure
-\ref{fig:scbi-focal-vs-comp-map}.
+We observe 4 variables describing 20 competitor trees: their unique
+tree-stem ID, their distance to the focal tree (all \(\leq\) 7.5), their
+species, and their basal area (in m\(^2\)) calculated as
+\(\frac{\pi \times (\text{DBH/2})^2}{10000}\) for \(DBH\) in cm from the
+earlier census. Saving competitor information in list-column format
+minimizes redundancy since we do not need to repeat information on the
+focal tree 20 times. We visualize the spatial distribution of these
+trees in Figure \ref{fig:scbi-focal-vs-comp-map}.
 
@@
 
-The final step is to fit a model for the growth of all focal trees. We
-fit the competition Bayesian linear regression model outlined in
-Equation \ref{eq:model} using \texttt{comp\_bayes\_lm()}, which has an
-option to specify prior distributions on all parameters of interest
-(chosen to be the defaults specified in \texttt{?comp\_bayes\_lm}).
+The final step is to fit the competition Bayesian linear regression
+model for tree growth outlined in Equation \ref{eq:model} using
+\texttt{comp\_bayes\_lm()}. This function has an option to specify prior
+distributions on all parameters of interest, chosen here to be the
+defaults detailed in \texttt{?comp\_bayes\_lm}.
 
@@
 
-The returned \texttt{comp\_bayes\_lm\_scbi} output is an object of S3
-class type \texttt{comp\_bayes\_lm} containing the posterior values of
-all parameters in our competition Bayesian linear regression. This class
-of object includes three generic methods. First, the generic for
-\texttt{print()} displays the names of all prior \& posterior parameters
-along with the model formula:
+The resutling \texttt{comp\_bayes\_lm\_scbi} is an object of S3 class
+type \texttt{comp\_bayes\_lm} containing the posterior values of all
+parameters of our model. Furthermore, this class of object includes
+generics for three methods. First, the generic for \texttt{print()}
+displays the names of all prior and posterior parameters along with the
+model formula:
 
@@
 Next, the generic for \texttt{predict()} takes the posterior parameter
-values in \texttt{comp\_bayes\_lm\_scbi} and the predictor variables in
-\texttt{newdata} and outputs a vector \texttt{growth\_hat} of
-fitted/predicted values \(\widehat{y}\) of the DBH for each focal tree
-computed from the posterior predictive distribution.
+values in \texttt{comp\_bayes\_lm\_scbi} and a \texttt{newdata} data
+frame and outputs a vector \texttt{growth\_hat} of predicted values
+\(\widehat{y}\) of the DBH for each focal tree computed from the
+posterior predictive distribution.
 
@@
 
-We then compare the observed and fitted/predicted growths to compute the
-root mean squared error (RMSE) of our model fit.
+We then compare the observed and predicted growths to compute the root
+mean squared error (RMSE) of our model fit.
 
@@
 
-Lastly, the generic for \texttt{ggplot2::autoplot()} allows us to plot
-the posterior distribution of all parameters in Figure
-\ref{fig:scbi-posterior-viz}.
+Lastly, the generic for \texttt{ggplot2::autoplot()} allows us to
+visualize the posterior distribution of all parameters, as seen in
+Figure \ref{fig:scbi-posterior-viz}. Setting \texttt{type} to
+\texttt{"intercepts"} and \texttt{"dbh\_slopes"} returns
+species-specific posterior distributions for \(\beta_{0,j}\) and
+\(\beta_{dbh,j}\) respectively, while setting
+\texttt{type\ =\ "competition"} returns competition coefficients
+\(\lambda_{j,k}\).
 
@@
 
-\CommentTok{# Combine plots using patchwork}
+\CommentTok{# Combine plots using the patchwork package}
 \NormalTok{(plot1 }\OperatorTok{|}\StringTok{ }\NormalTok{plot2) }\OperatorTok{/}\StringTok{ }\NormalTok{plot3}
@@
 
-These plots visualize the posterior distributions of parameters from
-Equation \ref{eq:model}. For many package users they will be of interest
-because they give insight into the species-specific competitive
-interactions. Setting \texttt{type\ =\ "intercepts"} returns
-species-specific posterior distributions for \(\beta_{0,j}\) and
-\texttt{type\ =\ "dbh\_slopes"} for \(\beta_{dbh,j}\). Setting
-\texttt{type\ =\ "competition"} returns competition coefficients
-\(\lambda_{j,k}\) where negative values indicate a competitor species
-which slows the growth of a focal species. Here, for example, we see
-that \texttt{litu} tulip poplars have a strong negative effect on the
+For many users the visualizations relating to \(\lambda_{j,k}\) will be
+of particular interest as they provide insight into species competitive
+interactions where negative values indicate a competitor species which
+slows the growth of a focal species. Here, for example, we see that
+tulip poplars (\texttt{litu}) have a strong negative effect on the
 growth of conspecifics but relatively lesser effect on neighbors of the
@@
 competition Bayesian linear regression model outlined in Equation
-\ref{eq:model}. However, it can be extended to any model implemented in
-a function similar to \texttt{comp\_bayes\_lm()} that uses data frames
-of similar format to \texttt{focal\_vs\_comp} as input.
+\ref{eq:model}. However, it can be extended to any model as long as it
+is implemented in a function similar to \texttt{comp\_bayes\_lm()}.
 
-\hypertarget{evaluate-the-effect-of-competitor-species-identity-using-permutation-tests}{%
+\hypertarget{permutation-test}{%
 \subsection{Evaluate the effect of competitor species identity using
-permutation
-tests}\label{evaluate-the-effect-of-competitor-species-identity-using-permutation-tests}}
+permutation tests}\label{permutation-test}}
 
-To evaluate the effect of competitor species identity, we use the four
-steps of our analysis sequence answer along with a permutation test:
-Under a null hypothesis where competitor species identity does not
-matter, we permute/shuffle this variable within each focal tree, compute
-the RMSE (the test statistic of interest), repeat this process several
-times to construct a null distribution of the RMSE, and compare it to
-the observed RMSE to assess significance. Going back to our example in
-Section \ref{focal-vs-comp} of focal tree with \texttt{focal\_ID} 4 and
-its 20 competitors, the permutation test randomly resamples only the
-\texttt{comp\_sp} variable with replacement, leaving all other variables
-intact. The resampling with replacement is nested within each focal tree
-in order to preserve neighborhood structure. We once again use
-\texttt{comp\_bayes\_lm()} as in Section \ref{model-fit-predict}, but
-with \texttt{run\_shuffle\ =\ TRUE}.
+To evaluate the effect of competitor species identity, we use the above
+four steps along with a permutation test. Under a null hypothesis where
+competitor species identity does not matter, we permute the competitor
+species identity within each focal tree, compute the RMSE test
+statistic, repeat this process several times to construct a null
+distribution, and finally compare it to the observed RMSE to assess
+significance. Going back to our example in Section \ref{focal-vs-comp}
+of focal tree with \texttt{focal\_ID} 4 and its 20 competitors, the
+permutation test randomly resamples only the \texttt{comp\_sp} variable
+without replacement, leaving all other variables intact. The resampling
+without replacement is nested within each focal tree in order to
+preserve neighborhood structure. We once again use
+\texttt{comp\_bayes\_lm()}, but this time setting
+\texttt{run\_shuffle\ =\ TRUE}.
 
@@
 
-The resulting RMSE of 0.131 based on the permutation test is larger than
-the earlier RMSE of 0.128, suggesting that models that do incorporate
-competitor species identity better fit the data.
+The resulting permutation test RMSE of 0.131 is larger than the earlier
+RMSE of 0.128, suggesting that models that do incorporate competitor
+species identity better fit the data.
 
@@
 
-We answer the second of our two questions: how can we obtain an accurate
-estimate of model performance/error? The model fits and predictions in
-Section \ref{model-fit-predict} all suffer from a common failing: they
-use the same data to both fit the model and to assess the model's
-performance using the RMSE. As argued by
-\citet{roberts_cross-validation_2017}, this can lead to overly
-optimistic assessments of model quality as the models can be overfit, in
-particular in situations where spatial-autocorrelation is present. To
-mitigate the effects of such overfitting, we use a spatially block
-cross-validation algorithm.
+To evaluate model performance, we use spatial cross-validation. The
+model fits and predictions in Section \ref{model-fit-predict} use the
+same data to both fit the model and to assess the model's performance.
+Given the inherent spatial-autocorrelation of our data, this can
+potentially lead to potentially overfit models
+\citep{roberts_cross-validation_2017}. To mitigate this potential
+overfitting, we use the spatial cross-validation blocking scheme encoded
+in the \texttt{foldID} variable from Section \ref{spatial-information}
+and visualized in Figure \ref{fig:scbi-spatial-information}.
 
-To this end, we use the \texttt{foldID} variable defined in Section
-\ref{spatial-information} whereby all focal trees are assigned to one of
-4 spatially contiguous blocks that act as folds in our cross-validation
-routine. Figure \ref{fig:scbi-spatial-cross-validation-schematic}
-presents a schematic illustrating this scheme. We fit the model to all
-focal trees in the training set, apply the model to all focal trees in
-the test set to compute fitted/predicted values, and compute the RMSE of
-the observed versus predicted growths. We repeat this procedure 3 more
-times with each of the three remaining folds acting as the test set and
-then average all four resulting RMSE's. Furthermore, in order to
-maintain spatial independence between the test and training set, a
-``fold buffer'' that extend outwards from the boundary of the test set
-is computed; all trees falling within this fold buffer are excluded from
-the training set.
+At each iteration of our cross-validation, one fold will act as the test
+set with the remaining three acting as the training set. We fit the
+model to all focal trees in the training set, apply the model to all
+focal trees in the test set, compute predicted values, and compute the
+RMSE. Furthermore, to maintain spatial independence between the test and
+training set, a ``fold buffer'' that extend outwards from the boundary
+of the test set is computed; all trees falling within this fold buffer
+are excluded from the training set (see Figure
+\ref{fig:scbi-spatial-cross-validation-schematic}).
 
@@
 
-\caption{Schematic of spatial cross-validation: Using the k = 1 fold (bottom-left) as the test set, k = 2 through 4 as the training set, along with a "fold buffer."}\label{fig:scbi-spatial-cross-validation-schematic}
+\caption{Schematic of spatial cross-validation: Using the k = 1 fold (bottom-left) as the test set, k = 2 through 4 as the training set, along with a fold buffer.}\label{fig:scbi-spatial-cross-validation-schematic}
 \end{figure}
 
-This algorithm is implemented in \texttt{run\_cv()}, which is a wrapper
-function to both \texttt{comp\_bayes\_lm()} that fits the model and
-\texttt{predict()} that returns fitted/predicted values. We once again
-compute the RMSE.
+This is repeated for each of the four folds acting as the test set and
+then the four resulting RMSE's are averaged to provide a single estimate
+of model error. This algorithm is implemented in \texttt{run\_cv()},
+which acts as a wrapper function to both \texttt{comp\_bayes\_lm()} that
+fits the model and \texttt{predict()} that returns predicted values.
 
@@
 than the earlier RMSE of 0.128, suggesting that models that do not take
-the inherent spatial autocorrelation of the data into account generate
-error estimates that are overly optimistic; in our case RMSE's that are
-too low.
+spatial autocorrelation account generate moel error estimates that are
+overly optimistic. In our case, RMSE values that are too low.
 
@@
 
-\texttt{run\_cv()} also accepts the \texttt{run\_shuffle} argument. This
-permutes the competitor species, as described above, but does so when
-calculating predicted growth with the cross validated scheme. Figure
-\ref{fig:scbi-simulation} compares model performance when permuting
-competitor species and calculating RMSE with and without
-cross-validation. Without cross-validation the competitor identity did
-matter, the non-permuted competitor species had a much lower RMSE than
-the permuted one. But once we include the spatial cross-validation, this
-improvement disappears. These results suggest that in this 9 ha subplot
-of the SCBI plot competitive interactions do not depend on the identity
-of the competitor, which is the opposite of what has been observed in
-other locations (\citet{allen_permutation_2020}
-\citet{uriarte_spatially_2004}). This highlights the importance of
-cross-validation, without it the model was overfit.
+\texttt{run\_cv()} also accepts the \texttt{run\_shuffle} argument in
+order to permute competitor species identity as described in Section
+\ref{permutation-test}. Figure \ref{fig:scbi-simulation} compares model
+performance for 49 permutations of competitor species and calculating
+RMSE, both with and without cross-validation. Without cross-validation,
+competitor species identity did matter as the observed RMSE was lower
+than the permutation null distribution of RMSE. However, once we
+incorporate spatial cross-validation, this improvement disappears. These
+results suggest that in this 9 ha subplot of the SCBI plot competitive
+interactions do not depend on the identity of the competitor, which is
+the opposite of what has been observed in other locations \citep[
+\citet{uriarte_spatially_2004}]{allen_permutation_2020}. Furthermore,
+the larger cross-validated RMSE's are indicative of the importance of
+being vigilant against model overfitting.
 
@@
 measurements of each individual have been taken, e.g.~the US Forest
-Service Forest Inventory and Analysis plots or ii) more generally to
+Service Forest Inventory and Analysis plots, or ii) more generally to
 model interactions of any community of mapped sessile organisms
 \citep{smith_forest_2002}. In future versions of \texttt{forestecology}
-we also hope to make it possible to model plant mortality in addition to
-plant growth. The package follows the guidelines for \texttt{tidy} data,
-leverages the \texttt{sf} package for spatial data, and S3 open-oriented
-model structure. We hope that the package will increase the use of
-neighborhood competition models to better understand what structures
-plant competition.
+we also hope to include models that account for tree mortality in
+addition to tree growth. The package follows the guidelines for
+\texttt{tidy} data, leverages the \texttt{sf} package for spatial data,
+and S3 open-oriented model structure. We hope that the package will
+increase the use of neighborhood competition models to better understand
+what structures plant competition.
 

### 2021-03-10 paper/forestecology.Rmd
3139 -> 3060. Gah!

@@
 keywords:
-- forest ecology, interspecific competition, neigbhorhood competition, tree growth, R, ForestGEO, spatial cross-validation
+- forest ecology, interspecific competition, neighborhood competition, tree growth, R, ForestGEO, spatial cross-validation
 
@@
 Although these are well-described methods, few methods are currently available for easy application.
-Here we address this in an R package.
-We largely follow the methods presented in @allen_permutation_2020.
-The package is written to model stem radial growth between two censuses based on neighborhood competition.
 
-@allen_permutation_2020 considers the following model: Let $i = 1, \ldots, n_j$ index all $n_j$ trees of "focal" species group $j$; let $j = 1, \ldots, J$ index all $J$ focal species groups; and let $k = 1, \ldots, K$ index all $K$ "competitor" species groups.
-They use the following linear model the average annual growth in diameter at breast height (DBH) $y_{ij}$ (in centimeters/year) of the $i^{th}$ tree of focal species group $j$:
+We address this shortcoming with the `forestecology` R package providing methods and data for forest ecology model fitting and assessment available on CRAN (https://cran.r-project.org/web/packages/forestecology/index.html) and on GitHub (https://github.com/rudeboybert/forestecology).
+The package is written to model stem radial growth between two censuses based on neighborhood competition largely following the methods in @allen_permutation_2020.
+
+Let $i = 1, \ldots, n_j$ index all $n_j$ trees of "focal" species group $j$; let $j = 1, \ldots, J$ index all $J$ focal species groups; and let $k = 1, \ldots, K$ index all $K$ "competitor" species groups.
+The average annual growth in diameter at breast height (DBH) $y_{ij}$ (in centimeters/year) of the $i^{th}$ tree of focal species group $j$ is modeled as
 
@@
 \label{eq:model}
-y_{ij} = f(\vec{x}_{ij}) + \epsilon_{ij} = \beta_{0,j} + \beta_{\text{dbh},j} \cdot \text{dbh}_{ij} + \sum_{k=1}^{K} \lambda_{jk} \cdot \text{BA}_{ijk} + \epsilon_{ij}
+y_{ij} = \beta_{0,j} + \beta_{\text{dbh},j} \cdot \text{dbh}_{ij} + \sum_{k=1}^{K} \lambda_{jk} \cdot \text{BA}_{ijk} + \epsilon_{ij}
 \end{equation}
 
-where $\beta_{0,j}$ is the diameter-independent growth rate for group $j$; $\text{dbh}_{ij}$ is the DBH the focal tree at the earlier census; $\beta_{\text{dbh},j}$ is the amount of the growth rate changed depending on diameter for group $j$;  $\text{BA}_{ijk}$ is the sum of the basal area of all trees of competitor species group $k$; $\lambda_{jk}$ is the change in growth for individuals of group $j$ from nearby competitors of group $k$; and $\epsilon_{ij}$ is a random error term distributed $\text{Normal}(0, \sigma^2)$.
+where $\beta_{0,j}$ is the diameter-independent growth rate for group $j$; $\text{dbh}_{ij}$ is the DBH of the focal tree at the earlier census for group $j$ and $\beta_{\text{dbh},j}$ quantifies its corresponding growth rate change;  $\text{BA}_{ijk}$ is the sum of the basal area of all trees of competitor species group $k$ and $\lambda_{jk}$ quantifies the corresponding change in growth for individuals of group $j$ from these competitors; and $\epsilon_{ij}$ is a random error term distributed $\text{Normal}(0, \sigma^2)$.
 They estimate all parameters via Bayesian linear regression while exploiting Normal/Inverse Gamma conjugacy to derive closed-form solutions to all posterior distributions^[See S1 Appendix of @allen_permutation_2020, available at https://doi.org/10.1371/journal.pone.0229930.s004].
-These closed-form solutions for the posterior distributions are in contrast to approximations of all posteriors via computationally expensive Markov Chain Monte Carlo algorithms.
+These closed-form solutions are in contrast to approximations of all posteriors via computationally expensive Markov Chain Monte Carlo algorithms.
 
-In order to evaluate whether competitor species identity matters, @allen_permutation_2020 run a permutation test where a null hypothesis of no species grouping-specific effects of competition is assumed, and thus the species identity of all competitors can be permuted/shuffled:
+To evaluate whether competitor species identity matters, @allen_permutation_2020 run a permutation test where a null hypothesis of no species grouping-specific effects of competition is assumed, and thus the species identity of all competitors can be permuted:
 
@@
 &&H_0: \lambda_{jk} = \lambda_{j} \mbox{ for all } k = 1, \ldots, K\\
-\text{vs.}&&H_A: \text{at least one } \lambda_{jk} \mbox{ is different}
+\text{vs.}&&H_A: \text{at least one } \lambda_{jk} \mbox{ is different} \nonumber
 \end{eqnarray}
 
-Furthermore, in order to account for the spatial autocorrelation inherent to forest data in their estimates of out-of-sample model error, @allen_permutation_2020 use spatial cross-validation.
+Furthermore, to account for the spatial autocorrelation of forestry data in their estimates of out-of-sample model error, @allen_permutation_2020 use spatial cross-validation.
 Estimates of model error that do not account for this spatial dependency tend to underestimate the true model error [@roberts_cross-validation_2017].
 
-We introduce the `forestecology` R package providing methods and data for forest ecology model fitting and assessment, available on CRAN (https://cran.r-project.org/web/packages/forestecology/index.html) and on GitHub (https://github.com/rudeboybert/forestecology).
-The package implements all aspects of the model in Equation \ref{eq:model}: model fitting and generating predicted values, evaluating the effect of competitor species identity using permutation tests, and evaluating model performance using spatial cross-validation.
-
-The package designed with "tidy" design principles in mind [@wickham_welcome_2019].
-Much like all `tidyverse` packages, `forestecology` is has verb-named functions that can be modularly composed using the pipe `%>%` operator to complete all the necessary steps in the analysis sequence [@bache_pipe_2020].
-Furthermore, the inputs and outputs of most of our functions use the same "simple features for R" data structures from the `sf` package for standardized and `tidyverse`-friendly wrangling and visualizing of spatial data [@pebesma_simple_2018]
+The package is designed with "tidy" design principles in mind [@wickham_welcome_2019].
+Much like all `tidyverse` packages, `forestecology` has verb-named functions that can be modularly composed using the pipe `%>%` operator to sequentially complete all the necessary analysis steps [@bache_pipe_2020].
+Furthermore, the inputs and outputs of most of our functions use the same "simple features for R" data structures from the `sf` package, a package for standardized and `tidyverse`-friendly wrangling and visualizing of spatial data [@pebesma_simple_2018]
 
@@
 As we demonstrate in Section \ref{model-fit-predict} however, the fitting of this model is self-contained in a single function `comp_bayes_lm()` which returns an object of S3 class type `comp_bayes_lm`.
-This class has generic methods implemented to print, make predictions using, and plot all results.
-Therefore the package can be modularly extended to fit other models as long as they are coded similarly as `comp_bayes_lm()` and have equivalent generic methods implemented.
+This class has generic methods implemented to print, make predictions, and plot all results.
+Therefore the package can be modularly extended to fit other models as long as they are coded similarly to `comp_bayes_lm()` and have equivalent generic methods implemented.
 
@@
 
-We present a case-study of the `forestecology` package's functionality on data from the Smithsonian Conservation Biology Institute (SCBI) large forest dynamics plot in Front Royal, VA, USA, part of the ForestGEO global network of research sites [@bourg_initial_2013, @andersonteixeira_ctfs-forestgeo_2015] [@bourg_initial_2013].
+We present a case-study of `forestecology`'s functionality on data from the Smithsonian Conservation Biology Institute (SCBI) large forest dynamics plot in Front Royal, VA, USA, part of the ForestGEO global network of research sites [@bourg_initial_2013, @andersonteixeira_ctfs-forestgeo_2015] [@bourg_initial_2013].
 The 25.6 ha (640 x 400 m) plot is located at the intersection of three of the major physiographic provinces of the eastern US---the Blue Ridge, Ridge and Valley, and Piedmont provinces---and is adjacent to the northern end of Shenandoah National Park.
@@
 The package has the following ecological goals: to evaluate i) the effect of competitor species identity using permutation tests and ii) model performance using spatial cross-validation.
-We outline a basic analysis sequence comprising of the following four main steps:
+We outline the four step basic analysis sequence:
 
@@
 Despite such standards, minor variations will still exist between sites, thereby necessitating some data wrangling.
-For example, the SCBI site records all DBH values in millimeters [@bourg_initial_2013], whereas the Michigan Big Woods site records them in centimeters [@allen_michigan_2020].
+For example, the SCBI site records all DBH values in millimeters [@bourg_initial_2013], whereas the Michigan Big Woods site used in @allen_permutation_2020 records them in centimeters [@allen_michigan_2020].
 
-We first load both 2008 and 2014 SCBI census `.csv` files as they existed on GitHub on 2020/11/20 and perform some data wrangling to both data sets [@gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020].
+We first load both 2008 and 2014 SCBI census `.csv` files as they existed on GitHub on 2020/11/20 and perform minor data wrangling to both data sets [@gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020].
 We then only consider a 9 ha subsection of the 25.6 ha of the SCBI site in order to speed up computation for this example: `gx` from 0--300 instead of 0--400 and `gy` from 300--600 instead of 0--640.
@@
 These two data frames are then used as inputs to `compute_growth()`, along with the `id` argument specifying the variable that uniquely identifies each tree-stem.
-Note that we also discard all resprouts in the later census with `code == R`, since we are only interested in the growth of surviving, and not resprouted, stems.
+Note that we also discard all resprouts with `code == R` in the later census, since we are only interested in the growth of surviving, and not resprouted, stems.
 
@@
   )
-growth_scbi
+growth_scbi %>% 
+  select(stemID, sp, dbh1, dbh2, growth, geometry)
 ```
 
-The output `growth_scbi` is a data frame of class `sf` that includes variables the average annual `growth` in DBH (cm $\cdot$ y^-1^) for all stems that were alive at both time points, the `sf` package's encoding of geolocations of `geometry` type `<POINT>`, and the species variable `sp` converted to a factor.
+The output `growth_scbi` is a data frame of class `sf` that includes among other variables the species variable `sp` converted to a factor, the average annual `growth` in DBH (cm $\cdot$ y^-1^) for all stems that were alive at both time points, and the `sf` package's encoding of geolocations of `geometry` type `<POINT>`.
 Given that `growth_scbi` is of class `sf`, it can be easily plotted in `ggplot2` using `geom_sf()` as seen in Figure \ref{fig:scbi-trees}.
 
-```{r scbi-trees, out.width="66%", fig.cap="Compute growth of trees based on census data: Map with growth of a random sample of 500 trees from a 9 ha subsection of the Smithsonian Conservation Biology Institute (SCBI) forest plot."}
+```{r scbi-trees, out.width="66%", fig.cap="Compute growth of trees based on census data.
+A map of the growth of a random sample of 500 trees from a 9 ha subsection of the Smithsonian Conservation Biology Institute (SCBI) forest plot."}
 ggplot() +
@@
 
-The next step is to add spatial information to `growth_scbi`.
-The first element we add is a "buffer region" to the periphery of the study region.
+We then add spatial information to `growth_scbi`.
+We first add a "buffer region" to the periphery of the study region.
 Since some of our model's explanatory variables are cumulative, we must ensure that all trees being modeled are not biased to have different neighbor structures.
-This is of concern for trees at the boundary of the study region whose neighbors will not all be included in the censused stems.
-To account for such edge effects only trees that are not part of this buffer region, i.e.
+This is of concern for trees at the boundary of the study region who will not have all their neighbors included in the census stems.
+To account for such edge effects, only trees that are not part of this buffer region, i.e.
 are part of the interior of the study region, will have their growth modeled [@waller_applied_2004].
@@
 Our model of interspecific competition relies on a spatial definition of who competitor trees are: all trees within a distance `comp_dist` of a focal tree.
-In our case we set `comp_dist` to 7.5m, a value informed by other studies [@canham_neighborhood_2004, @uriarte_spatially_2004, @canham_neighborhood_2006].
+We set `comp_dist` to 7.5m, a value informed by other studies [@canham_neighborhood_2004, @uriarte_spatially_2004, @canham_neighborhood_2006].
 We use `comp_dist` and a manually constructed `sf` representation of the study region's boundary as inputs to `add_buffer_variable()` to add a `buffer` boolean variable to `growth_scbi`.
@@
 The second element of spatial information we add are blocks corresponding to folds of a spatial cross-validation algorithm.
-Conventional cross-validation algorithms assign observations to folds by randomly resampling individual observations that are assumed independent.
+Conventional cross-validation algorithms assign individual observations to folds by randomly resampling them all while assuming they are statistically independent.
 In the case of forest census data however, observations exhibit spatial autocorrelation.
@@
 
-We first manually an `sf` object defining four folds that partition the study region.
+We first manually define an `sf` object defining four folds that partition the study region.
 We then use the output of the `spatialBlock()` function from the `blockCV` package to associate each tree in `growth_scbi` to the correct `foldID` [@valavi_blockcv_2019].^[In the Supporting Information we present an example where the folds themselves are created automatically, as opposed to manually as in the example.] This `foldID` variable will be used in Section \ref{spatial-cross-validation}.
@@
 # Associate each observation to a fold
-spatial_block_scbi <- spatialBlock(
-  speciesData = growth_scbi, k = n_fold, selection = "systematic", 
-  blocks = blocks_scbi, showBlocks = FALSE, verbose = FALSE
-)
+spatial_block_scbi <- 
+  spatialBlock(speciesData = growth_scbi, k = n_fold, 
+               selection = "systematic", blocks = blocks_scbi, 
+               showBlocks = FALSE, verbose = FALSE)
 
@@
 
-```{r scbi-spatial-information, out.width="66%", fig.cap="Add spatial information: Buffer region and spatial cross-validation blocks (1 through 4).
-The location of each tree is marked with an integer indicating its fold, with folds delineated with solid lines.
-The color of each digit indicates whether the tree is part of the buffer region (and thus will only be considered as a competitor tree in our model) or is part of the interior of the study region (and thus is a focal tree whose growth is of modeled interest)."}
+```{r scbi-spatial-information, out.width="66%", fig.cap="Add spatial information.
+A buffer region and spatial cross-validation blocks 1 througth 4.
+The location of each tree is marked with its fold number where the folds are delineated with solid lines.
+The color of each digit indicates whether the tree is part of the buffer region (thus will only be considered as a competitor tree) or is part of the interior of the study region (thus is a focal tree whose growth is of modeled interest)."}
 ggplot() +
@@
 
-The next step is to identify all focal trees and their corresponding competitor trees.
+We then identify all focal trees and their corresponding competitor trees.
 More specifically, identify all trees that are not part of the buffer region, have a valid `growth` measurement, and have at least one neighbor within `r comp_dist`m.
-We do this using `create_focal_vs_comp()`, which takes the previously detailed arguments `comp_dist` and `id` as well as the `sf` representation of the spatial cross-validation blocks and returns a new data frame `focal_vs_comp_scbi`.
+We do this using `create_focal_vs_comp()`, which takes the previously detailed `comp_dist` and `id` arguments as well as the `sf` representation of the spatial cross-validation blocks and returns a new data frame `focal_vs_comp_scbi`.
 
@@
 
-We observe 4 variables describing 20 competitor trees: their unique tree-stem ID, their distance to the focal tree (all $\leq$ `r comp_dist`), their species, and their basal area (in m$^2$) calculated as $\frac{\pi \times (\text{DBH/2})^2}{10000}$ for $DBH$ in cm from the earlier census.
+We observe 4 variables describing 20 competitor trees: the unique tree-stem ID, the distance to the focal tree (all $\leq$ `r comp_dist`), the species, and the basal area (in m$^2$) calculated as $\frac{\pi \times (\text{DBH/2})^2}{10000}$ for the DBH in cm from the earlier census.
 Saving competitor information in list-column format minimizes redundancy since we do not need to repeat information on the focal tree 20 times.
@@
 
-```{r scbi-focal-vs-comp-map, out.width="66%", echo = FALSE, fig.cap="Identify all focal and corresponding competitor trees: The dashed circle extends 7.5m away from the focal tree 4 while all 20 competitor trees are within this circle."}
+```{r scbi-focal-vs-comp-map, out.width="66%", echo = FALSE, fig.cap="Identify all focal and corresponding competitor trees.
+The dashed circle extends 7.5m away from the focal tree 4 while all 20 competitor trees are within this circle."}
 comp_radius <- growth_scbi %>% 
@@
 
-The final step is to fit the competition Bayesian linear regression model for tree growth outlined in Equation \ref{eq:model} using `comp_bayes_lm()`.
-This function has an option to specify prior distributions on all parameters of interest, chosen here to be the defaults detailed in `?comp_bayes_lm`.
+Lastly, we then fit the competition Bayesian linear regression model for tree growth outlined in Equation \ref{eq:model} using `comp_bayes_lm()`.
+This function has an option to specify prior distributions of all parameters of interest, chosen here to be the defaults detailed in `?comp_bayes_lm`.
 
@@
 
-The resutling `comp_bayes_lm_scbi` is an object of S3 class type `comp_bayes_lm` containing the posterior values of all parameters of our model.
+The resulting`comp_bayes_lm_scbi` is an object of S3 class type `comp_bayes_lm` containing the posterior values of all parameters.
 Furthermore, this class of object includes generics for three methods.
-First, the generic for `print()` displays the names of all prior and posterior parameters along with the model formula:
+First, the generic for `print()` displays the names of all prior and posterior parameters and the model formula:
 
@@
 
-Next, the generic for `predict()` takes the posterior parameter values in `comp_bayes_lm_scbi` and a `newdata` data frame and outputs a vector `growth_hat` of predicted values $\widehat{y}$ of the DBH for each focal tree computed from the posterior predictive distribution.
+Next, the generic for `predict()` takes the posterior parameter values in `comp_bayes_lm_scbi` and a `newdata` data frame, and outputs a vector `growth_hat` of predicted values $\widehat{y}$ of the DBH computed from the posterior predictive distribution.
 
@@
 
-We then compare the observed and predicted growths to compute the root mean squared error (RMSE) of our model fit.
+We can now compare the observed and predicted growths to compute the root mean squared error (RMSE) of our model fit.
 
@@
 
-```{r scbi-posterior-viz, out.width="100%", fig.cap= "Fit model: Posterior distributions of all parameters. For compactness we include only three species.", fig.width = 8, fig.height = 6}
+```{r scbi-posterior-viz, out.width="100%", fig.cap= "Fit model. Posterior distributions of all parameters. For compactness we include only three species.", fig.width = 8, fig.height = 6}
 # Plot posteriors for only a subset of species
@@
 
-For many users the visualizations relating to $\lambda_{j,k}$ will be of particular interest as they provide insight into species competitive interactions where negative values indicate a competitor species which slows the growth of a focal species.
-Here, for example, we see that tulip poplars (`litu`) have a strong negative effect on the growth of conspecifics but relatively lesser effect on neighbors of the other two species.
+For many users the visualizations relating to $\lambda_{j,k}$ will be of particular interest as they provide insight into species competitive interactions, where negative values indicate a competitor species which slows the growth of a focal species.
+Here, for example, we see that tulip poplars (litu) have a strong negative effect on the growth of conspecifics but relatively lesser effect on pignut hickory (cagl) and red oak (quru) neighbors.
 
@@
 
-To evaluate the effect of competitor species identity, we use the above four steps along with a permutation test.
-Under a null hypothesis where competitor species identity does not matter, we permute the competitor species identity within each focal tree, compute the RMSE test statistic, repeat this process several times to construct a null distribution, and finally compare it to the observed RMSE to assess significance.
-Going back to our example in Section \ref{focal-vs-comp} of focal tree with `focal_ID` 4 and its 20 competitors, the permutation test randomly resamples only the `comp_sp` variable without replacement, leaving all other variables intact.
-The resampling without replacement is nested within each focal tree in order to preserve neighborhood structure.
-We once again use `comp_bayes_lm()`, but this time setting `run_shuffle = TRUE`.
+To evaluate the effect of competitor species identity, we use the above four steps along with the permutation test in Equation \ref{eq:permutation-hypothesis-test}.
+Under a null hypothesis where competitor species identity does not matter, we can permute the competitor species identities within each focal tree, compute the RMSE test statistic, repeat this process several times to construct a null distribution, and finally compare it to the observed RMSE to assess significance.
+Going back to our example in Section \ref{focal-vs-comp} of focal tree with `focal_ID` 4 and its 20 competitors, the permutation test only randomly resamples the `comp_sp` variable without replacement, leaving all other variables intact.
+This resampling is nested within each focal tree in order to preserve neighborhood structure.
+We perform this permutation test, we once again use `comp_bayes_lm()` but by setting `run_shuffle = TRUE`.
 
@@
 focal_vs_comp_scbi <- focal_vs_comp_scbi %>%
-  mutate(
-    growth_hat_shuffle = predict(comp_bayes_lm_scbi_shuffle, 
-                                 newdata = focal_vs_comp_scbi)
-  )
+  mutate(growth_hat_shuffle = predict(comp_bayes_lm_scbi_shuffle, 
+                                 newdata = focal_vs_comp_scbi))
 ```
@@
 To evaluate model performance, we use spatial cross-validation.
-The model fits and predictions in Section \ref{model-fit-predict} use the same data to both fit the model and to assess the model's performance.
-Given the inherent spatial-autocorrelation of our data, this can potentially lead to potentially overfit models [@roberts_cross-validation_2017].
-To mitigate this potential overfitting, we use the spatial cross-validation blocking scheme encoded in the `foldID` variable from Section \ref{spatial-information} and visualized in Figure \ref{fig:scbi-spatial-information}.
+The example model fit in Section \ref{model-fit-predict} uses the same data to both fit the model and assess its performance.
+Given the inherent spatial-autocorrelation of our data, this can potentially lead to overfit models [@roberts_cross-validation_2017].
+To mitigate the risk of overfitting, we make use of the spatial cross-validation blocking scheme encoded in the `foldID` variable from Section \ref{spatial-information} and visualized in Figure \ref{fig:scbi-spatial-information}.
 
-At each iteration of our cross-validation, one fold will act as the test set with the remaining three acting as the training set.
+At each iteration of our cross-validation, one fold will act as the test set and the remaining three acting as the training set.
 We fit the model to all focal trees in the training set, apply the model to all focal trees in the test set, compute predicted values, and compute the RMSE.
-Furthermore, to maintain spatial independence between the test and training set, a "fold buffer" that extend outwards from the boundary of the test set is computed; all trees falling within this fold buffer are excluded from the training set (see Figure \ref{fig:scbi-spatial-cross-validation-schematic}).
+Furthermore, to maintain spatial independence between the test and training set, a "fold buffer" that extend `r comp_dist`m outwards from the boundary of the test set is computed; all trees falling within this fold buffer are excluded from the training set (see Figure \ref{fig:scbi-spatial-cross-validation-schematic}).
 
-```{r scbi-spatial-cross-validation-schematic, out.width="66%", fig.cap= "Schematic of spatial cross-validation: Using the k = 1 fold (bottom-left) as the test set, k = 2 through 4 as the training set, along with a fold buffer.", echo = FALSE}
+```{r scbi-spatial-cross-validation-schematic, out.width="66%", fig.cap= "Schematic of spatial cross-validation.
+Using the k = 1 fold (bottom-left) as the test set, k = 2 through 4 as the training set, along with a \"fold buffer\" extending outwards from the test set to maintain spatial independence between it and the training set.", echo = FALSE}
 fold_number <- 1
@@
 
-This is repeated for each of the four folds acting as the test set and then the four resulting RMSE's are averaged to provide a single estimate of model error.
+This is repeated for each of the four folds acting as the test set, then the four resulting RMSE's are averaged to provide a single estimate of model error.
 This algorithm is implemented in `run_cv()`, which acts as a wrapper function to both `comp_bayes_lm()` that fits the model and `predict()` that returns predicted values.
@@
 
-The resulting RMSE of `r model_rmse_cv` computed using cross-validation is `r ifelse(model_rmse_cv > model_rmse, "larger", "smaller")` than the earlier RMSE of `r model_rmse`, suggesting that models that do not take spatial autocorrelation account generate moel error estimates that are overly optimistic.
+The resulting RMSE of `r model_rmse_cv` computed using cross-validation is `r ifelse(model_rmse_cv > model_rmse, "larger", "smaller")` than the earlier RMSE of `r model_rmse`, suggesting that models that do not take spatial autocorrelation account generate model error estimates that are overly optimistic.
 In our case, RMSE values that are too low.
@@
 `run_cv()` also accepts the `run_shuffle` argument in order to permute competitor species identity as described in Section \ref{permutation-test}.
-Figure \ref{fig:scbi-simulation} compares model performance for 49 permutations of competitor species and calculating RMSE, both with and without cross-validation.
-Without cross-validation, competitor species identity did matter as the observed RMSE was lower than the permutation null distribution of RMSE.
+Figure \ref{fig:scbi-simulation} compares model performance for 49 permutations of competitor species and RMSE calculations, both with and without cross-validation.
+Without cross-validation, competitor species identity did matter as the observed RMSE was significantly lower than the permutation null distribution of RMSE.
 However, once we incorporate spatial cross-validation, this improvement disappears.
 These results suggest that in this 9 ha subplot of the SCBI plot competitive interactions do not depend on the identity of the competitor, which is the opposite of what has been observed in other locations [@allen_permutation_2020 @uriarte_spatially_2004].
-Furthermore, the larger cross-validated RMSE's are indicative of the importance of being vigilant against model overfitting.
+Furthermore, the larger cross-validated RMSE's are indicative of the importance of being vigilant against potential model overfitting.
 
@@
 The `forestecology` package provides an accessible way to fit and test models of neighborhood competition.
-While the package is designed with ForestGEO plot data in mind, we envision that it can be modified to work on i) any single large, mapped forest plot in which at least two measurements of each individual have been taken, e.g.
-the US Forest Service Forest Inventory and Analysis plots, or ii) more generally to model interactions of any community of mapped sessile organisms [@smith_forest_2002].
+While the package is designed with ForestGEO plot data in mind, we envision that it can be modified to work on any single large, mapped forest plot in which at least two measurements of each individual have been taken, e.g.
+the US Forest Service Forest Inventory and Analysis plots.
+Furthemore, we also envision the package being used more generally to model interactions of any community of mapped sessile organisms [@smith_forest_2002].
 In future versions of `forestecology` we also hope to include models that account for tree mortality in addition to tree growth.
-The package follows the guidelines for `tidy` data, leverages the `sf` package for spatial data, and S3 open-oriented model structure.
+The package follows the `tidy` data design principles, leverages the `sf` package for spatial data, and S3 open-oriented model implementation structure.
 We hope that the package will increase the use of neighborhood competition models to better understand what structures plant competition.
@@
 This repository will be archived on Zenodo upon acceptance.
-The example Smithsonian Conservation Biology Institute census data used are available on GitHub at https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/tree/master/tree_main_census/data/census-csv-files and are archived on Zenodo at https://doi.org/10.5281/zenodo.2649301 [@gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020].
+The example Smithsonian Conservation Biology Institute census data are available on GitHub at https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/tree/master/tree_main_census/data/census-csv-files and are archived on Zenodo at https://doi.org/10.5281/zenodo.2649301 [@gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020].
 

### 2021-03-10 paper/forestecology.Rmd
3060 -> 3000 🏁🏁🏁

@@
 
-We address this shortcoming with the `forestecology` R package providing methods and data for forest ecology model fitting and assessment available on CRAN (https://cran.r-project.org/web/packages/forestecology/index.html) and on GitHub (https://github.com/rudeboybert/forestecology).
-The package is written to model stem radial growth between two censuses based on neighborhood competition largely following the methods in @allen_permutation_2020.
+We address this shortcoming with the `forestecology` R package providing methods and data for forest ecology model fitting and assessment, available on CRAN (https://cran.r-project.org/web/packages/forestecology/index.html) and on GitHub (https://github.com/rudeboybert/forestecology).
+The package is written to model stem radial growth between two censuses based on neighborhood competition, largely following the methods in @allen_permutation_2020.
 
@@
 where $\beta_{0,j}$ is the diameter-independent growth rate for group $j$; $\text{dbh}_{ij}$ is the DBH of the focal tree at the earlier census for group $j$ and $\beta_{\text{dbh},j}$ quantifies its corresponding growth rate change;  $\text{BA}_{ijk}$ is the sum of the basal area of all trees of competitor species group $k$ and $\lambda_{jk}$ quantifies the corresponding change in growth for individuals of group $j$ from these competitors; and $\epsilon_{ij}$ is a random error term distributed $\text{Normal}(0, \sigma^2)$.
-They estimate all parameters via Bayesian linear regression while exploiting Normal/Inverse Gamma conjugacy to derive closed-form solutions to all posterior distributions^[See S1 Appendix of @allen_permutation_2020, available at https://doi.org/10.1371/journal.pone.0229930.s004].
-These closed-form solutions are in contrast to approximations of all posteriors via computationally expensive Markov Chain Monte Carlo algorithms.
+They estimate all parameters via Bayesian linear regression, while exploiting Normal/Inverse Gamma conjugacy to derive closed-form solutions to all posterior distributions^[See S1 Appendix of @allen_permutation_2020, available at https://doi.org/10.1371/journal.pone.0229930.s004].
+These closed-form solutions are not as computationally expensive as approximations from Markov Chain Monte Carlo algorithms.
 
-To evaluate whether competitor species identity matters, @allen_permutation_2020 run a permutation test where a null hypothesis of no species grouping-specific effects of competition is assumed, and thus the species identity of all competitors can be permuted:
+To evaluate whether competitor species identity matters, @allen_permutation_2020 run a permutation test where a null hypothesis of no species grouping-specific effects of competition is assumed, thus the species identity of all competitors can be permuted:
 
@@
 
-Furthermore, to account for the spatial autocorrelation of forestry data in their estimates of out-of-sample model error, @allen_permutation_2020 use spatial cross-validation.
-Estimates of model error that do not account for this spatial dependency tend to underestimate the true model error [@roberts_cross-validation_2017].
+Furthermore, to account for the spatial autocorrelation in their estimates of out-of-sample model error, @allen_permutation_2020 use spatial cross-validation.
+Estimates of model error that do not account for this dependence tend to underestimate the true model error [@roberts_cross-validation_2017].
 
 The package is designed with "tidy" design principles in mind [@wickham_welcome_2019].
-Much like all `tidyverse` packages, `forestecology` has verb-named functions that can be modularly composed using the pipe `%>%` operator to sequentially complete all the necessary analysis steps [@bache_pipe_2020].
-Furthermore, the inputs and outputs of most of our functions use the same "simple features for R" data structures from the `sf` package, a package for standardized and `tidyverse`-friendly wrangling and visualizing of spatial data [@pebesma_simple_2018]
+Much like all `tidyverse` packages, `forestecology` has verb-named functions that can be modularly composed using the pipe `%>%` operator to sequentially complete all necessary analysis steps [@bache_pipe_2020].
+Furthermore, the inputs and outputs of most functions use the same "simple features for R" data structures from the `sf` package, a package for standardized and `tidyverse`-friendly wrangling and visualizing of spatial data [@pebesma_simple_2018].
 
@@
 
-The package has the following ecological goals: to evaluate i) the effect of competitor species identity using permutation tests and ii) model performance using spatial cross-validation.
-We outline the four step basic analysis sequence:
+The package has the following goals: to evaluate i) the effect of competitor species identity using permutation tests and ii) model performance using spatial cross-validation.
+We outline the four-step basic analysis sequence:
 
@@
 
-We start by loading all necessary packages.
+We start by loading all packages.
 
@@
 
-The first step is to compute the growth of trees using data from two censuses.
+We first compute the growth of trees using data from two censuses.
 `compute_growth()` computes the average annual growth based on census data that roughly follows ForestGEO standards.
@@
 
-We first load both 2008 and 2014 SCBI census `.csv` files as they existed on GitHub on 2020/11/20 and perform minor data wrangling to both data sets [@gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020].
-We then only consider a 9 ha subsection of the 25.6 ha of the SCBI site in order to speed up computation for this example: `gx` from 0--300 instead of 0--400 and `gy` from 300--600 instead of 0--640.
+We load both 2008 and 2014 SCBI census `.csv` files as they existed on GitHub on 2020/11/20 and perform minor data wrangling [@gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020].
+We then only consider a 9 ha subsection of the 25.6 ha of the site to speed up computation for this example: `gx` from 0--300 instead of 0--400 and `gy` from 300--600 instead of 0--640.
 
@@
 
-These two data frames are then used as inputs to `compute_growth()`, along with the `id` argument specifying the variable that uniquely identifies each tree-stem.
-Note that we also discard all resprouts with `code == R` in the later census, since we are only interested in the growth of surviving, and not resprouted, stems.
+These two data frames are then used as inputs to `compute_growth()`, along with `id` specifying the variable that uniquely identifies each tree-stem.
+We also discard all resprouts with `code == R` in the later census, since we are only interested in the growth of surviving, and not resprouted, stems.
 
@@
 n_fold <- 4
-fold1 <- rbind(c(0, 300), c(150, 300), c(150, 450), c(0, 450))
-fold2 <- rbind(c(150, 300), c(300, 300), c(300, 450), c(150, 450))
-fold3 <- rbind(c(0, 450), c(150, 450), c(150, 600), c(0, 600))
-fold4 <- rbind(c(150, 450), c(300, 450), c(300, 600), c(150, 600))
+fold1 <- cbind(c(0, 150, 150, 0), c(300, 300, 450, 450))
+fold2 <- cbind(c(150, 300, 300, 150), c(300, 300, 450, 450))
+fold3 <- cbind(c(0, 150, 150, 0), c(450, 450, 600, 600))
+fold4 <- cbind(c(150, 300, 300, 150), c(450, 450, 600, 600))
 
@@
 
-Lastly, we then fit the competition Bayesian linear regression model for tree growth outlined in Equation \ref{eq:model} using `comp_bayes_lm()`.
-This function has an option to specify prior distributions of all parameters of interest, chosen here to be the defaults detailed in `?comp_bayes_lm`.
+Lastly, we fit the competition Bayesian linear regression model for tree growth outlined in Equation \ref{eq:model} using `comp_bayes_lm()`.
+This function has an option to specify prior distributions of all parameters, chosen here to be the defaults detailed in `?comp_bayes_lm`.
 
@@
 The resulting`comp_bayes_lm_scbi` is an object of S3 class type `comp_bayes_lm` containing the posterior values of all parameters.
-Furthermore, this class of object includes generics for three methods.
+Furthermore, this class includes generics for three methods.
 First, the generic for `print()` displays the names of all prior and posterior parameters and the model formula:
@@
 
-Next, the generic for `predict()` takes the posterior parameter values in `comp_bayes_lm_scbi` and a `newdata` data frame, and outputs a vector `growth_hat` of predicted values $\widehat{y}$ of the DBH computed from the posterior predictive distribution.
+Next, the generic for `predict()` takes the posterior parameter values in `comp_bayes_lm_scbi` and a `newdata` data frame, and outputs a vector `growth_hat` of predicted DBH values $\widehat{y_{ij}}$ computed from the posterior predictive distribution.
 
@@
 
-We can now compare the observed and predicted growths to compute the root mean squared error (RMSE) of our model fit.
+We can now compare the observed and predicted growths to compute the root mean squared error (RMSE) of our model:
 
@@
 
-Lastly, the generic for `ggplot2::autoplot()` allows us to visualize the posterior distribution of all parameters, as seen in Figure \ref{fig:scbi-posterior-viz}.
+Lastly, the generic for `ggplot2::autoplot()` allows us to visualize all posterior distributions, as seen in Figure \ref{fig:scbi-posterior-viz}.
 Setting `type` to `"intercepts"` and `"dbh_slopes"` returns species-specific posterior distributions for $\beta_{0,j}$ and $\beta_{dbh,j}$ respectively, while setting `type = "competition"` returns competition coefficients $\lambda_{j,k}$.
@@
 
-For many users the visualizations relating to $\lambda_{j,k}$ will be of particular interest as they provide insight into species competitive interactions, where negative values indicate a competitor species which slows the growth of a focal species.
+For many users the visualizations of $\lambda_{j,k}$ will be of particular interest as they provide insight into species competitive interactions, where negative values indicate a competitor species which slows the growth of a focal species.
 Here, for example, we see that tulip poplars (litu) have a strong negative effect on the growth of conspecifics but relatively lesser effect on pignut hickory (cagl) and red oak (quru) neighbors.
 
-Currently the `forestecology` package can only fit the competition Bayesian linear regression model outlined in Equation \ref{eq:model}.
+Currently the `forestecology` package can only fit the competition Bayesian linear regression model in Equation \ref{eq:model}.
 However, it can be extended to any model as long as it is implemented in a function similar to `comp_bayes_lm()`.
@@
 To evaluate the effect of competitor species identity, we use the above four steps along with the permutation test in Equation \ref{eq:permutation-hypothesis-test}.
-Under a null hypothesis where competitor species identity does not matter, we can permute the competitor species identities within each focal tree, compute the RMSE test statistic, repeat this process several times to construct a null distribution, and finally compare it to the observed RMSE to assess significance.
+Under a null hypothesis where competitor species identity does not matter, we can permute the competitor species identities within each focal tree, compute the RMSE test statistic, repeat this process several times to construct a null distribution, and compare it to the observed RMSE to assess significance.
 Going back to our example in Section \ref{focal-vs-comp} of focal tree with `focal_ID` 4 and its 20 competitors, the permutation test only randomly resamples the `comp_sp` variable without replacement, leaving all other variables intact.
 This resampling is nested within each focal tree in order to preserve neighborhood structure.
-We perform this permutation test, we once again use `comp_bayes_lm()` but by setting `run_shuffle = TRUE`.
+We perform this permutation test once again using `comp_bayes_lm()` but by setting `run_shuffle = TRUE`.
 
@@
 To evaluate model performance, we use spatial cross-validation.
-The example model fit in Section \ref{model-fit-predict} uses the same data to both fit the model and assess its performance.
-Given the inherent spatial-autocorrelation of our data, this can potentially lead to overfit models [@roberts_cross-validation_2017].
-To mitigate the risk of overfitting, we make use of the spatial cross-validation blocking scheme encoded in the `foldID` variable from Section \ref{spatial-information} and visualized in Figure \ref{fig:scbi-spatial-information}.
+The model fit in Section \ref{model-fit-predict} uses the same data to both fit and assess model performance.
+Given the spatial-autocorrelation of our data, this can potentially lead to overfit models [@roberts_cross-validation_2017].
+To mitigate this risk, we use the spatial cross-validation blocking scheme encoded in the `foldID` variable from Section \ref{spatial-information} and visualized in Figure \ref{fig:scbi-spatial-information}.
 
-At each iteration of our cross-validation, one fold will act as the test set and the remaining three acting as the training set.
+At each iteration of the cross-validation, one fold acts as the test set and the remaining three act as the training set.
 We fit the model to all focal trees in the training set, apply the model to all focal trees in the test set, compute predicted values, and compute the RMSE.
-Furthermore, to maintain spatial independence between the test and training set, a "fold buffer" that extend `r comp_dist`m outwards from the boundary of the test set is computed; all trees falling within this fold buffer are excluded from the training set (see Figure \ref{fig:scbi-spatial-cross-validation-schematic}).
+Furthermore, to maintain spatial independence between the test and training sets, a "fold buffer" that extends `r comp_dist`m outwards from the boundary of the test set is considered; all trees within this "fold buffer" are excluded from the training set (see Figure \ref{fig:scbi-spatial-cross-validation-schematic}).
 
@@
 
-This is repeated for each of the four folds acting as the test set, then the four resulting RMSE's are averaged to provide a single estimate of model error.
+This process is repeated for each of the four folds acting as the test set, then the four RMSE's are averaged to provide a single estimate of model error.
 This algorithm is implemented in `run_cv()`, which acts as a wrapper function to both `comp_bayes_lm()` that fits the model and `predict()` that returns predicted values.
@@
 
-The resulting RMSE of `r model_rmse_cv` computed using cross-validation is `r ifelse(model_rmse_cv > model_rmse, "larger", "smaller")` than the earlier RMSE of `r model_rmse`, suggesting that models that do not take spatial autocorrelation account generate model error estimates that are overly optimistic.
-In our case, RMSE values that are too low.
+The resulting RMSE of `r model_rmse_cv` computed using cross-validation is `r ifelse(model_rmse_cv > model_rmse, "larger", "smaller")` than the earlier RMSE of `r model_rmse`, suggesting that models that do not account for spatial autocorrelation generate model error estimates that are overly optimistic, i.e.
+RMSE values that are too low.
 
@@
 Figure \ref{fig:scbi-simulation} compares model performance for 49 permutations of competitor species and RMSE calculations, both with and without cross-validation.
-Without cross-validation, competitor species identity did matter as the observed RMSE was significantly lower than the permutation null distribution of RMSE.
+Without cross-validation, competitor species identity does matter as the observed RMSE was significantly lower than the permutation null distribution of RMSE.
 However, once we incorporate spatial cross-validation, this improvement disappears.
-These results suggest that in this 9 ha subplot of the SCBI plot competitive interactions do not depend on the identity of the competitor, which is the opposite of what has been observed in other locations [@allen_permutation_2020 @uriarte_spatially_2004].
+These results suggest that in this 9 ha subplot of the SCBI plot, competitive interactions do not depend on the identity of the competitor, which is the opposite of what has been observed in other locations [@allen_permutation_2020 @uriarte_spatially_2004].
 Furthermore, the larger cross-validated RMSE's are indicative of the importance of being vigilant against potential model overfitting.
@@
 
-The authors thank Sophie Li for their feedback on the package interface. The authors declare no conflicts of interest.
+We thank Sophie Li for their feedback on the package interface. The authors declare no conflicts of interest.
 
@@
 
-We intend to archive all data and source code for this manuscript on GitHub at https://github.com/rudeboybert/forestecology.
-This repository will be archived on Zenodo upon acceptance.
+We intend to archive all data and source code for this manuscript on GitHub at https://github.com/rudeboybert/forestecology and on Zenodo upon acceptance.
 The example Smithsonian Conservation Biology Institute census data are available on GitHub at https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/tree/master/tree_main_census/data/census-csv-files and are archived on Zenodo at https://doi.org/10.5281/zenodo.2649301 [@gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020].

### 2021-03-10 paper/forestecology.tex
3060 -> 3000 🏁🏁🏁

@@
 \noindent%
-{\it Keywords:} forest ecology, interspecific competition, neigbhorhood competition,
+{\it Keywords:} forest ecology, interspecific competition, neighborhood competition,
 tree growth, R, ForestGEO, spatial cross-validation
@@
 these are well-described methods, few methods are currently available
-for easy application. Here we address this in an R package. We largely
-follow the methods presented in \citet{allen_permutation_2020}. The
+for easy application.
+
+We address this shortcoming with the \texttt{forestecology} R package
+providing methods and data for forest ecology model fitting and
+assessment, available on CRAN
+(\url{https://cran.r-project.org/web/packages/forestecology/index.html})
+and on GitHub (\url{https://github.com/rudeboybert/forestecology}). The
 package is written to model stem radial growth between two censuses
-based on neighborhood competition.
+based on neighborhood competition, largely following the methods in
+\citet{allen_permutation_2020}.
 
-\citet{allen_permutation_2020} considers the following model: Let
-\(i = 1, \ldots, n_j\) index all \(n_j\) trees of ``focal'' species
+Let \(i = 1, \ldots, n_j\) index all \(n_j\) trees of ``focal'' species
 group \(j\); let \(j = 1, \ldots, J\) index all \(J\) focal species
 groups; and let \(k = 1, \ldots, K\) index all \(K\) ``competitor''
-species groups. They use the following linear model the average annual
-growth in diameter at breast height (DBH) \(y_{ij}\) (in
-centimeters/year) of the \(i^{th}\) tree of focal species group \(j\):
+species groups. The average annual growth in diameter at breast height
+(DBH) \(y_{ij}\) (in centimeters/year) of the \(i^{th}\) tree of focal
+species group \(j\) is modeled as
 
@@
 \label{eq:model}
-y_{ij} = f(\vec{x}_{ij}) + \epsilon_{ij} = \beta_{0,j} + \beta_{\text{dbh},j} \cdot \text{dbh}_{ij} + \sum_{k=1}^{K} \lambda_{jk} \cdot \text{BA}_{ijk} + \epsilon_{ij}
+y_{ij} = \beta_{0,j} + \beta_{\text{dbh},j} \cdot \text{dbh}_{ij} + \sum_{k=1}^{K} \lambda_{jk} \cdot \text{BA}_{ijk} + \epsilon_{ij}
 \end{equation}
@@
 where \(\beta_{0,j}\) is the diameter-independent growth rate for group
-\(j\); \(\text{dbh}_{ij}\) is the DBH the focal tree at the earlier
-census; \(\beta_{\text{dbh},j}\) is the amount of the growth rate
-changed depending on diameter for group \(j\); \(\text{BA}_{ijk}\) is
-the sum of the basal area of all trees of competitor species group
-\(k\); \(\lambda_{jk}\) is the change in growth for individuals of group
-\(j\) from nearby competitors of group \(k\); and \(\epsilon_{ij}\) is a
-random error term distributed \(\text{Normal}(0, \sigma^2)\). They
-estimate all parameters via Bayesian linear regression while exploiting
+\(j\); \(\text{dbh}_{ij}\) is the DBH of the focal tree at the earlier
+census for group \(j\) and \(\beta_{\text{dbh},j}\) quantifies its
+corresponding growth rate change; \(\text{BA}_{ijk}\) is the sum of the
+basal area of all trees of competitor species group \(k\) and
+\(\lambda_{jk}\) quantifies the corresponding change in growth for
+individuals of group \(j\) from these competitors; and \(\epsilon_{ij}\)
+is a random error term distributed \(\text{Normal}(0, \sigma^2)\). They
+estimate all parameters via Bayesian linear regression, while exploiting
 Normal/Inverse Gamma conjugacy to derive closed-form solutions to all
@@
   \url{https://doi.org/10.1371/journal.pone.0229930.s004}}. These
-closed-form solutions for the posterior distributions are in contrast to
-approximations of all posteriors via computationally expensive Markov
-Chain Monte Carlo algorithms.
+closed-form solutions are not as computationally expensive as
+approximations from Markov Chain Monte Carlo algorithms.
 
-In order to evaluate whether competitor species identity matters,
+To evaluate whether competitor species identity matters,
 \citet{allen_permutation_2020} run a permutation test where a null
 hypothesis of no species grouping-specific effects of competition is
-assumed, and thus the species identity of all competitors can be
-permuted/shuffled:
+assumed, thus the species identity of all competitors can be permuted:
 
@@
 &&H_0: \lambda_{jk} = \lambda_{j} \mbox{ for all } k = 1, \ldots, K\\
-\text{vs.}&&H_A: \text{at least one } \lambda_{jk} \mbox{ is different}
+\text{vs.}&&H_A: \text{at least one } \lambda_{jk} \mbox{ is different} \nonumber
 \end{eqnarray}
 
-Furthermore, in order to account for the spatial autocorrelation
-inherent to forest data in their estimates of out-of-sample model error,
-\citet{allen_permutation_2020} use spatial cross-validation. Estimates
-of model error that do not account for this spatial dependency tend to
-underestimate the true model error
+Furthermore, to account for the spatial autocorrelation in their
+estimates of out-of-sample model error, \citet{allen_permutation_2020}
+use spatial cross-validation. Estimates of model error that do not
+account for this dependence tend to underestimate the true model error
 \citep{roberts_cross-validation_2017}.
 
-We introduce the \texttt{forestecology} R package providing methods and
-data for forest ecology model fitting and assessment, available on CRAN
-(\url{https://cran.r-project.org/web/packages/forestecology/index.html})
-and on GitHub (\url{https://github.com/rudeboybert/forestecology}). The
-package implements all aspects of the model in Equation \ref{eq:model}:
-model fitting and generating predicted values, evaluating the effect of
-competitor species identity using permutation tests, and evaluating
-model performance using spatial cross-validation.
-
-The package designed with ``tidy'' design principles in mind
+The package is designed with ``tidy'' design principles in mind
 \citep{wickham_welcome_2019}. Much like all \texttt{tidyverse} packages,
-\texttt{forestecology} is has verb-named functions that can be modularly
-composed using the pipe \texttt{\%\textgreater{}\%} operator to complete
-all the necessary steps in the analysis sequence
-\citep{bache_pipe_2020}. Furthermore, the inputs and outputs of most of
-our functions use the same ``simple features for R'' data structures
-from the \texttt{sf} package for standardized and
+\texttt{forestecology} has verb-named functions that can be modularly
+composed using the pipe \texttt{\%\textgreater{}\%} operator to
+sequentially complete all necessary analysis steps
+\citep{bache_pipe_2020}. Furthermore, the inputs and outputs of most
+functions use the same ``simple features for R'' data structures from
+the \texttt{sf} package, a package for standardized and
 \texttt{tidyverse}-friendly wrangling and visualizing of spatial data
-\citep{pebesma_simple_2018}
+\citep{pebesma_simple_2018}.
 
@@
 returns an object of S3 class type \texttt{comp\_bayes\_lm}. This class
-has generic methods implemented to print, make predictions using, and
-plot all results. Therefore the package can be modularly extended to fit
-other models as long as they are coded similarly as
-\texttt{comp\_bayes\_lm()} and have equivalent generic methods
-implemented.
+has generic methods implemented to print, make predictions, and plot all
+results. Therefore the package can be modularly extended to fit other
+models as long as they are coded similarly to \texttt{comp\_bayes\_lm()}
+and have equivalent generic methods implemented.
 
@@
 
-We present a case-study of the \texttt{forestecology} package's
-functionality on data from the Smithsonian Conservation Biology
-Institute (SCBI) large forest dynamics plot in Front Royal, VA, USA,
-part of the ForestGEO global network of research sites
+We present a case-study of \texttt{forestecology}'s functionality on
+data from the Smithsonian Conservation Biology Institute (SCBI) large
+forest dynamics plot in Front Royal, VA, USA, part of the ForestGEO
+global network of research sites
 \citep[\citet{andersonteixeira_ctfs-forestgeo_2015}]{bourg_initial_2013}
@@
 
-The package has the following ecological goals: to evaluate i) the
-effect of competitor species identity using permutation tests and ii)
-model performance using spatial cross-validation. We outline a basic
-analysis sequence comprising of the following four main steps:
+The package has the following goals: to evaluate i) the effect of
+competitor species identity using permutation tests and ii) model
+performance using spatial cross-validation. We outline the four-step
+basic analysis sequence:
 
@@
 
-The first step is to compute the growth of trees using data from two
-censuses. \texttt{compute\_growth()} computes the average annual growth
-based on census data that roughly follows ForestGEO standards. Despite
-such standards, minor variations will still exist between sites, thereby
+We first compute the growth of trees using data from two censuses.
+\texttt{compute\_growth()} computes the average annual growth based on
+census data that roughly follows ForestGEO standards. Despite such
+standards, minor variations will still exist between sites, thereby
 necessitating some data wrangling. For example, the SCBI site records
 all DBH values in millimeters \citep{bourg_initial_2013}, whereas the
-Michigan Big Woods site records them in centimeters
-\citep{allen_michigan_2020}.
+Michigan Big Woods site used in \citet{allen_permutation_2020} records
+them in centimeters \citep{allen_michigan_2020}.
 
-We first load both 2008 and 2014 SCBI census \texttt{.csv} files as they
-existed on GitHub on 2020/11/20 and perform some data wrangling to both
-data sets \citep{gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020}.
-We then only consider a 9 ha subsection of the 25.6 ha of the SCBI site
-in order to speed up computation for this example: \texttt{gx} from
-0--300 instead of 0--400 and \texttt{gy} from 300--600 instead of
-0--640.
+We load both 2008 and 2014 SCBI census \texttt{.csv} files as they
+existed on GitHub on 2020/11/20 and perform minor data wrangling
+\citep{gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020}. We then
+only consider a 9 ha subsection of the 25.6 ha of the site to speed up
+computation for this example: \texttt{gx} from 0--300 instead of 0--400
+and \texttt{gy} from 300--600 instead of 0--640.
 
@@
 These two data frames are then used as inputs to
-\texttt{compute\_growth()}, along with the \texttt{id} argument
-specifying the variable that uniquely identifies each tree-stem. Note
-that we also discard all resprouts in the later census with
-\texttt{code\ ==\ R}, since we are only interested in the growth of
-surviving, and not resprouted, stems.
+\texttt{compute\_growth()}, along with \texttt{id} specifying the
+variable that uniquely identifies each tree-stem. We also discard all
+resprouts with \texttt{code\ ==\ R} in the later census, since we are
+only interested in the growth of surviving, and not resprouted, stems.
 
@@
 \NormalTok{  )}
-\NormalTok{growth_scbi}
-\CommentTok{## Simple feature collection with 7954 features and 8 fields}
+\NormalTok{growth_scbi }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{select}\NormalTok{(stemID, sp, dbh1, dbh2, growth, geometry)}
+\CommentTok{## Simple feature collection with 7954 features and 5 fields}
 \CommentTok{## geometry type:  POINT}
@@
 \CommentTok{## CRS:            NA}
-\CommentTok{## # A tibble: 7,954 x 9}
-\CommentTok{##   stemID sp     dbh1 codes1 status  dbh2 codes2 growth}
-\CommentTok{##    <dbl> <fct> <dbl> <chr>  <chr>  <dbl> <chr>   <dbl>}
-\CommentTok{## 1      4 nysy  13.6  M      A       14.2 M       0.103}
-\CommentTok{## 2      5 havi   8.8  M      A        9.6 M;P     0.150}
-\CommentTok{## 3      6 havi   3.25 NULL   A        4   M       0.140}
-\CommentTok{## 4     77 qual  65.2  M      A       66   M       0.141}
-\CommentTok{## 5     79 tiam  47.7  M      A       46.8 M      -0.161}
-\CommentTok{## # ... with 7,949 more rows, and 1 more variable: geometry <POINT>}
+\CommentTok{## # A tibble: 7,954 x 6}
+\CommentTok{##   stemID sp     dbh1  dbh2 growth   geometry}
+\CommentTok{##    <dbl> <fct> <dbl> <dbl>  <dbl>    <POINT>}
+\CommentTok{## 1      4 nysy  13.6   14.2  0.103 (14.2 428)}
+\CommentTok{## 2      5 havi   8.8    9.6  0.150  (9.4 436)}
+\CommentTok{## 3      6 havi   3.25   4    0.140  (1.3 434)}
+\CommentTok{## 4     77 qual  65.2   66    0.141 (34.7 307)}
+\CommentTok{## 5     79 tiam  47.7   46.8 -0.161   (40 381)}
+\CommentTok{## # ... with 7,949 more rows}
 \end{Highlighting}
@@
 The output \texttt{growth\_scbi} is a data frame of class \texttt{sf}
-that includes variables the average annual \texttt{growth} in DBH (cm
+that includes among other variables the species variable \texttt{sp}
+converted to a factor, the average annual \texttt{growth} in DBH (cm
 \(\cdot\) y\textsuperscript{-1}) for all stems that were alive at both
-time points, the \texttt{sf} package's encoding of geolocations of
-\texttt{geometry} type \texttt{\textless{}POINT\textgreater{}}, and the
-species variable \texttt{sp} converted to a factor. Given that
-\texttt{growth\_scbi} is of class \texttt{sf}, it can be easily plotted
-in \texttt{ggplot2} using \texttt{geom\_sf()} as seen in Figure
+time points, and the \texttt{sf} package's encoding of geolocations of
+\texttt{geometry} type \texttt{\textless{}POINT\textgreater{}}. Given
+that \texttt{growth\_scbi} is of class \texttt{sf}, it can be easily
+plotted in \texttt{ggplot2} using \texttt{geom\_sf()} as seen in Figure
 \ref{fig:scbi-trees}.
@@
 
-\caption{Compute growth of trees based on census data: Map with growth of a random sample of 500 trees from a 9 ha subsection of the Smithsonian Conservation Biology Institute (SCBI) forest plot.}\label{fig:scbi-trees}
+\caption{Compute growth of trees based on census data.
+A map of the growth of a random sample of 500 trees from a 9 ha subsection of the Smithsonian Conservation Biology Institute (SCBI) forest plot.}\label{fig:scbi-trees}
 \end{figure}
@@
 
-The next step is to add spatial information to \texttt{growth\_scbi}.
-The first element we add is a ``buffer region'' to the periphery of the
-study region. Since some of our model's explanatory variables are
-cumulative, we must ensure that all trees being modeled are not biased
-to have different neighbor structures. This is of concern for trees at
-the boundary of the study region whose neighbors will not all be
-included in the censused stems. To account for such edge effects only
-trees that are not part of this buffer region, i.e.~are part of the
-interior of the study region, will have their growth modeled
-\citep{waller_applied_2004}.
+We then add spatial information to \texttt{growth\_scbi}. We first add a
+``buffer region'' to the periphery of the study region. Since some of
+our model's explanatory variables are cumulative, we must ensure that
+all trees being modeled are not biased to have different neighbor
+structures. This is of concern for trees at the boundary of the study
+region who will not have all their neighbors included in the census
+stems. To account for such edge effects, only trees that are not part of
+this buffer region, i.e.~are part of the interior of the study region,
+will have their growth modeled \citep{waller_applied_2004}.
 
@@
 who competitor trees are: all trees within a distance
-\texttt{comp\_dist} of a focal tree. In our case we set
-\texttt{comp\_dist} to 7.5m, a value informed by other studies
-\citep[\citet{uriarte_spatially_2004},
+\texttt{comp\_dist} of a focal tree. We set \texttt{comp\_dist} to 7.5m,
+a value informed by other studies \citep[\citet{uriarte_spatially_2004},
 \citet{canham_neighborhood_2006}]{canham_neighborhood_2004}. We use
@@
 corresponding to folds of a spatial cross-validation algorithm.
-Conventional cross-validation algorithms assign observations to folds by
-randomly resampling individual observations that are assumed
-independent. In the case of forest census data however, observations
-exhibit spatial autocorrelation. We therefore incorporate this
-dependence into the cross-validation algorithm by resampling spatial
-blocks of trees \citep[
+Conventional cross-validation algorithms assign individual observations
+to folds by randomly resampling them all while assuming they are
+statistically independent. In the case of forest census data however,
+observations exhibit spatial autocorrelation. We therefore incorporate
+this dependence into the cross-validation algorithm by resampling
+spatial blocks of trees \citep[
 \citet{pohjankukka_estimating_2017}]{roberts_cross-validation_2017}.
 
-We first manually an \texttt{sf} object defining four folds that
+We first manually define an \texttt{sf} object defining four folds that
 partition the study region. We then use the output of the
@@
 \NormalTok{n_fold <-}\StringTok{ }\DecValTok{4}
-\NormalTok{fold1 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{450}\NormalTok{))}
-\NormalTok{fold2 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{450}\NormalTok{))}
-\NormalTok{fold3 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{600}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{600}\NormalTok{))}
-\NormalTok{fold4 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+\NormalTok{fold1 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{0}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{))}
+\NormalTok{fold2 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{150}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{))}
+\NormalTok{fold3 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{0}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+\NormalTok{fold4 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{150}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{))}
 
@@
 \CommentTok{# Associate each observation to a fold}
-\NormalTok{spatial_block_scbi <-}\StringTok{ }\KeywordTok{spatialBlock}\NormalTok{(}
-  \DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =}\NormalTok{ n_fold, }\DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{, }
-  \DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}
-\NormalTok{)}
+\NormalTok{spatial_block_scbi <-}\StringTok{ }
+\StringTok{  }\KeywordTok{spatialBlock}\NormalTok{(}\DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =}\NormalTok{ n_fold, }
+               \DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }
+               \DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}\NormalTok{)}
 
@@
 
-\caption{Add spatial information: Buffer region and spatial cross-validation blocks (1 through 4).
-The location of each tree is marked with an integer indicating its fold, with folds delineated with solid lines.
-The color of each digit indicates whether the tree is part of the buffer region (and thus will only be considered as a competitor tree in our model) or is part of the interior of the study region (and thus is a focal tree whose growth is of modeled interest).}\label{fig:scbi-spatial-information}
+\caption{Add spatial information.
+A buffer region and spatial cross-validation blocks 1 througth 4.
+The location of each tree is marked with its fold number where the folds are delineated with solid lines.
+The color of each digit indicates whether the tree is part of the buffer region (thus will only be considered as a competitor tree) or is part of the interior of the study region (thus is a focal tree whose growth is of modeled interest).}\label{fig:scbi-spatial-information}
 \end{figure}
@@
 
-The next step is to identify all focal trees and their corresponding
-competitor trees. More specifically, identify all trees that are not
-part of the buffer region, have a valid \texttt{growth} measurement, and
-have at least one neighbor within 7.5m. We do this using
+We then identify all focal trees and their corresponding competitor
+trees. More specifically, identify all trees that are not part of the
+buffer region, have a valid \texttt{growth} measurement, and have at
+least one neighbor within 7.5m. We do this using
 \texttt{create\_focal\_vs\_comp()}, which takes the previously detailed
-arguments \texttt{comp\_dist} and \texttt{id} as well as the \texttt{sf}
+\texttt{comp\_dist} and \texttt{id} arguments as well as the \texttt{sf}
 representation of the spatial cross-validation blocks and returns a new
@@
 
-We observe 4 variables describing 20 competitor trees: their unique
-tree-stem ID, their distance to the focal tree (all \(\leq\) 7.5), their
-species, and their basal area (in m\(^2\)) calculated as
-\(\frac{\pi \times (\text{DBH/2})^2}{10000}\) for \(DBH\) in cm from the
+We observe 4 variables describing 20 competitor trees: the unique
+tree-stem ID, the distance to the focal tree (all \(\leq\) 7.5), the
+species, and the basal area (in m\(^2\)) calculated as
+\(\frac{\pi \times (\text{DBH/2})^2}{10000}\) for the DBH in cm from the
 earlier census. Saving competitor information in list-column format
@@
 
-\caption{Identify all focal and corresponding competitor trees: The dashed circle extends 7.5m away from the focal tree 4 while all 20 competitor trees are within this circle.}\label{fig:scbi-focal-vs-comp-map}
+\caption{Identify all focal and corresponding competitor trees.
+The dashed circle extends 7.5m away from the focal tree 4 while all 20 competitor trees are within this circle.}\label{fig:scbi-focal-vs-comp-map}
 \end{figure}
@@
 
-The final step is to fit the competition Bayesian linear regression
-model for tree growth outlined in Equation \ref{eq:model} using
+Lastly, we fit the competition Bayesian linear regression model for tree
+growth outlined in Equation \ref{eq:model} using
 \texttt{comp\_bayes\_lm()}. This function has an option to specify prior
-distributions on all parameters of interest, chosen here to be the
-defaults detailed in \texttt{?comp\_bayes\_lm}.
+distributions of all parameters, chosen here to be the defaults detailed
+in \texttt{?comp\_bayes\_lm}.
 
@@
 
-The resutling \texttt{comp\_bayes\_lm\_scbi} is an object of S3 class
+The resulting\texttt{comp\_bayes\_lm\_scbi} is an object of S3 class
 type \texttt{comp\_bayes\_lm} containing the posterior values of all
-parameters of our model. Furthermore, this class of object includes
-generics for three methods. First, the generic for \texttt{print()}
-displays the names of all prior and posterior parameters along with the
-model formula:
+parameters. Furthermore, this class includes generics for three methods.
+First, the generic for \texttt{print()} displays the names of all prior
+and posterior parameters and the model formula:
 
@@
 values in \texttt{comp\_bayes\_lm\_scbi} and a \texttt{newdata} data
-frame and outputs a vector \texttt{growth\_hat} of predicted values
-\(\widehat{y}\) of the DBH for each focal tree computed from the
-posterior predictive distribution.
+frame, and outputs a vector \texttt{growth\_hat} of predicted DBH values
+\(\widehat{y_{ij}}\) computed from the posterior predictive
+distribution.
 
@@
 
-We then compare the observed and predicted growths to compute the root
-mean squared error (RMSE) of our model fit.
+We can now compare the observed and predicted growths to compute the
+root mean squared error (RMSE) of our model:
 
@@
 Lastly, the generic for \texttt{ggplot2::autoplot()} allows us to
-visualize the posterior distribution of all parameters, as seen in
-Figure \ref{fig:scbi-posterior-viz}. Setting \texttt{type} to
+visualize all posterior distributions, as seen in Figure
+\ref{fig:scbi-posterior-viz}. Setting \texttt{type} to
 \texttt{"intercepts"} and \texttt{"dbh\_slopes"} returns
@@
 
-\caption{Fit model: Posterior distributions of all parameters. For compactness we include only three species.}\label{fig:scbi-posterior-viz}
+\caption{Fit model. Posterior distributions of all parameters. For compactness we include only three species.}\label{fig:scbi-posterior-viz}
 \end{figure}
 
-For many users the visualizations relating to \(\lambda_{j,k}\) will be
-of particular interest as they provide insight into species competitive
-interactions where negative values indicate a competitor species which
+For many users the visualizations of \(\lambda_{j,k}\) will be of
+particular interest as they provide insight into species competitive
+interactions, where negative values indicate a competitor species which
 slows the growth of a focal species. Here, for example, we see that
-tulip poplars (\texttt{litu}) have a strong negative effect on the
-growth of conspecifics but relatively lesser effect on neighbors of the
-other two species.
+tulip poplars (litu) have a strong negative effect on the growth of
+conspecifics but relatively lesser effect on pignut hickory (cagl) and
+red oak (quru) neighbors.
 
 Currently the \texttt{forestecology} package can only fit the
-competition Bayesian linear regression model outlined in Equation
-\ref{eq:model}. However, it can be extended to any model as long as it
-is implemented in a function similar to \texttt{comp\_bayes\_lm()}.
+competition Bayesian linear regression model in Equation \ref{eq:model}.
+However, it can be extended to any model as long as it is implemented in
+a function similar to \texttt{comp\_bayes\_lm()}.
 
@@
 To evaluate the effect of competitor species identity, we use the above
-four steps along with a permutation test. Under a null hypothesis where
-competitor species identity does not matter, we permute the competitor
-species identity within each focal tree, compute the RMSE test
-statistic, repeat this process several times to construct a null
-distribution, and finally compare it to the observed RMSE to assess
+four steps along with the permutation test in Equation
+\ref{eq:permutation-hypothesis-test}. Under a null hypothesis where
+competitor species identity does not matter, we can permute the
+competitor species identities within each focal tree, compute the RMSE
+test statistic, repeat this process several times to construct a null
+distribution, and compare it to the observed RMSE to assess
 significance. Going back to our example in Section \ref{focal-vs-comp}
 of focal tree with \texttt{focal\_ID} 4 and its 20 competitors, the
-permutation test randomly resamples only the \texttt{comp\_sp} variable
-without replacement, leaving all other variables intact. The resampling
-without replacement is nested within each focal tree in order to
-preserve neighborhood structure. We once again use
-\texttt{comp\_bayes\_lm()}, but this time setting
+permutation test only randomly resamples the \texttt{comp\_sp} variable
+without replacement, leaving all other variables intact. This resampling
+is nested within each focal tree in order to preserve neighborhood
+structure. We perform this permutation test once again using
+\texttt{comp\_bayes\_lm()} but by setting
 \texttt{run\_shuffle\ =\ TRUE}.
@@
 \NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
-    \DataTypeTok{growth_hat_shuffle =} \KeywordTok{predict}\NormalTok{(comp_bayes_lm_scbi_shuffle, }
-                                 \DataTypeTok{newdata =}\NormalTok{ focal_vs_comp_scbi)}
-\NormalTok{  )}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{growth_hat_shuffle =} \KeywordTok{predict}\NormalTok{(comp_bayes_lm_scbi_shuffle, }
+                                 \DataTypeTok{newdata =}\NormalTok{ focal_vs_comp_scbi))}
 \end{Highlighting}
@@
 To evaluate model performance, we use spatial cross-validation. The
-model fits and predictions in Section \ref{model-fit-predict} use the
-same data to both fit the model and to assess the model's performance.
-Given the inherent spatial-autocorrelation of our data, this can
-potentially lead to potentially overfit models
-\citep{roberts_cross-validation_2017}. To mitigate this potential
-overfitting, we use the spatial cross-validation blocking scheme encoded
-in the \texttt{foldID} variable from Section \ref{spatial-information}
-and visualized in Figure \ref{fig:scbi-spatial-information}.
+model fit in Section \ref{model-fit-predict} uses the same data to both
+fit and assess model performance. Given the spatial-autocorrelation of
+our data, this can potentially lead to overfit models
+\citep{roberts_cross-validation_2017}. To mitigate this risk, we use the
+spatial cross-validation blocking scheme encoded in the \texttt{foldID}
+variable from Section \ref{spatial-information} and visualized in Figure
+\ref{fig:scbi-spatial-information}.
 
-At each iteration of our cross-validation, one fold will act as the test
-set with the remaining three acting as the training set. We fit the
-model to all focal trees in the training set, apply the model to all
-focal trees in the test set, compute predicted values, and compute the
-RMSE. Furthermore, to maintain spatial independence between the test and
-training set, a ``fold buffer'' that extend outwards from the boundary
-of the test set is computed; all trees falling within this fold buffer
-are excluded from the training set (see Figure
+At each iteration of the cross-validation, one fold acts as the test set
+and the remaining three act as the training set. We fit the model to all
+focal trees in the training set, apply the model to all focal trees in
+the test set, compute predicted values, and compute the RMSE.
+Furthermore, to maintain spatial independence between the test and
+training sets, a ``fold buffer'' that extends 7.5m outwards from the
+boundary of the test set is considered; all trees within this ``fold
+buffer'' are excluded from the training set (see Figure
 \ref{fig:scbi-spatial-cross-validation-schematic}).
@@
 
-\caption{Schematic of spatial cross-validation: Using the k = 1 fold (bottom-left) as the test set, k = 2 through 4 as the training set, along with a fold buffer.}\label{fig:scbi-spatial-cross-validation-schematic}
+\caption{Schematic of spatial cross-validation.
+Using the k = 1 fold (bottom-left) as the test set, k = 2 through 4 as the training set, along with a "fold buffer" extending outwards from the test set to maintain spatial independence between it and the training set.}\label{fig:scbi-spatial-cross-validation-schematic}
 \end{figure}
 
-This is repeated for each of the four folds acting as the test set and
-then the four resulting RMSE's are averaged to provide a single estimate
-of model error. This algorithm is implemented in \texttt{run\_cv()},
-which acts as a wrapper function to both \texttt{comp\_bayes\_lm()} that
-fits the model and \texttt{predict()} that returns predicted values.
+This process is repeated for each of the four folds acting as the test
+set, then the four RMSE's are averaged to provide a single estimate of
+model error. This algorithm is implemented in \texttt{run\_cv()}, which
+acts as a wrapper function to both \texttt{comp\_bayes\_lm()} that fits
+the model and \texttt{predict()} that returns predicted values.
 
@@
 The resulting RMSE of 0.14 computed using cross-validation is larger
-than the earlier RMSE of 0.128, suggesting that models that do not take
-spatial autocorrelation account generate moel error estimates that are
-overly optimistic. In our case, RMSE values that are too low.
+than the earlier RMSE of 0.128, suggesting that models that do not
+account for spatial autocorrelation generate model error estimates that
+are overly optimistic, i.e.~RMSE values that are too low.
 
@@
 \ref{permutation-test}. Figure \ref{fig:scbi-simulation} compares model
-performance for 49 permutations of competitor species and calculating
-RMSE, both with and without cross-validation. Without cross-validation,
-competitor species identity did matter as the observed RMSE was lower
-than the permutation null distribution of RMSE. However, once we
-incorporate spatial cross-validation, this improvement disappears. These
-results suggest that in this 9 ha subplot of the SCBI plot competitive
-interactions do not depend on the identity of the competitor, which is
-the opposite of what has been observed in other locations \citep[
+performance for 49 permutations of competitor species and RMSE
+calculations, both with and without cross-validation. Without
+cross-validation, competitor species identity does matter as the
+observed RMSE was significantly lower than the permutation null
+distribution of RMSE. However, once we incorporate spatial
+cross-validation, this improvement disappears. These results suggest
+that in this 9 ha subplot of the SCBI plot, competitive interactions do
+not depend on the identity of the competitor, which is the opposite of
+what has been observed in other locations \citep[
 \citet{uriarte_spatially_2004}]{allen_permutation_2020}. Furthermore,
 the larger cross-validated RMSE's are indicative of the importance of
-being vigilant against model overfitting.
+being vigilant against potential model overfitting.
 
@@
 with ForestGEO plot data in mind, we envision that it can be modified to
-work on i) any single large, mapped forest plot in which at least two
+work on any single large, mapped forest plot in which at least two
 measurements of each individual have been taken, e.g.~the US Forest
-Service Forest Inventory and Analysis plots, or ii) more generally to
-model interactions of any community of mapped sessile organisms
-\citep{smith_forest_2002}. In future versions of \texttt{forestecology}
-we also hope to include models that account for tree mortality in
-addition to tree growth. The package follows the guidelines for
-\texttt{tidy} data, leverages the \texttt{sf} package for spatial data,
-and S3 open-oriented model structure. We hope that the package will
-increase the use of neighborhood competition models to better understand
-what structures plant competition.
+Service Forest Inventory and Analysis plots. Furthemore, we also
+envision the package being used more generally to model interactions of
+any community of mapped sessile organisms \citep{smith_forest_2002}. In
+future versions of \texttt{forestecology} we also hope to include models
+that account for tree mortality in addition to tree growth. The package
+follows the \texttt{tidy} data design principles, leverages the
+\texttt{sf} package for spatial data, and S3 open-oriented model
+implementation structure. We hope that the package will increase the use
+of neighborhood competition models to better understand what structures
+plant competition.
 
@@
 
-The authors thank Sophie Li for their feedback on the package interface.
-The authors declare no conflicts of interest.
+We thank Sophie Li for their feedback on the package interface. The
+authors declare no conflicts of interest.
 
@@
 We intend to archive all data and source code for this manuscript on
-GitHub at \url{https://github.com/rudeboybert/forestecology}. This
-repository will be archived on Zenodo upon acceptance. The example
-Smithsonian Conservation Biology Institute census data used are
-available on GitHub at
+GitHub at \url{https://github.com/rudeboybert/forestecology} and on
+Zenodo upon acceptance. The example Smithsonian Conservation Biology
+Institute census data are available on GitHub at
 \url{https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/tree/master/tree_main_census/data/census-csv-files}

### 2021-03-10 paper/forestecology.Rmd
typos

@@
   | 2.
-We present the `forestecology` package providing methods to i) specify neighborhood competition models, ii) evalulate the effect of competitor species identity using permutation tests, and iii) measure model performance using spatial cross-validation.
+We present the `forestecology` package providing methods to i) specify neighborhood competition models, ii) evaluate the effect of competitor species identity using permutation tests, and iii) measure model performance using spatial cross-validation.
 Following @allen_permutation_2020, we implement a Bayesian linear regression neighborhood competition model.
   | 3.
-We demonstrate the package's functionality using data from the Smithsonian Conservation Biology Institute's large forest dynamics plot, part of the ForestGEO global network of reseach sites.
+We demonstrate the package's functionality using data from the Smithsonian Conservation Biology Institute's large forest dynamics plot, part of the ForestGEO global network of research sites.
 Given ForestGEO's data collection protocols and data formatting standards, the package was designed with cross-site compatibility in mind.
@@
 the US Forest Service Forest Inventory and Analysis plots.
-Furthemore, we also envision the package being used more generally to model interactions of any community of mapped sessile organisms [@smith_forest_2002].
+Furthermore, we also envision the package being used more generally to model interactions of any community of mapped sessile organisms [@smith_forest_2002].
 In future versions of `forestecology` we also hope to include models that account for tree mortality in addition to tree growth.

### 2021-03-10 paper/forestecology.Rmd
Updated CRAN link

@@
 
-We address this shortcoming with the `forestecology` R package providing methods and data for forest ecology model fitting and assessment, available on CRAN (https://cran.r-project.org/web/packages/forestecology/index.html) and on GitHub (https://github.com/rudeboybert/forestecology).
+We address this shortcoming with the `forestecology` R package providing methods and data for forest ecology model fitting and assessment, available on CRAN (https://cran.r-project.org/package=forestecology) and on GitHub (https://github.com/rudeboybert/forestecology).
 The package is written to model stem radial growth between two censuses based on neighborhood competition, largely following the methods in @allen_permutation_2020.

### 2021-03-10 paper/forestecology.Rmd
According to the koRpus method this is at 2994! Feel free to reject these changes if you don't like them. Or happy to chat about them tomorrow.

@@
 The package features i) `tidyverse`-like structure whereby verb-named functions can be modularly "piped" in sequence, ii) functions with standardized inputs/outputs of simple features `sf` package class, and iii) an S3 object-oriented implementation of the Bayesian linear regression model.
-These three facts allow for clear articulation of all the steps in the sequence of analysis and easy wrangling and visualization of the geospatial forestry data.
+These three facts allow for clear articulation of all the steps in the sequence of analysis and easy wrangling and visualization of the geospatial data.
 Furthermore, while the package only has Bayesian linear regression implemented, the package was designed with extensibility to other methods in mind.
@@
 
-Repeat-censused forest plots offer excellent opportunities to test neighborhood models of the effect of competition on the growth of trees (@canham_neighborhood_2004).
-Neighborhood models of competition have been used to: test whether the species identity of a competitor matters (@uriarte_spatially_2004); measure species-specific competition coefficients (@das_effect_2012 @tatsumi_estimating_2016); test competing models to see what structures competitive interactions, e.g.
-traits or phylogeny (@allen_permutation_2020; @uriarte_trait_2010); and inform selective logging practices (@canham_neighborhood_2006).
+Repeat-censused forest plots offer excellent opportunities to test neighborhood models of the effect of competition on the growth of trees [@canham_neighborhood_2004].
+Neighborhood models of competition have been used to: test whether the species identity of a competitor matters [@uriarte_spatially_2004; measure species-specific competition coefficients [@das_effect_2012 @tatsumi_estimating_2016]; test competing models to see what structures competitive interactions, e.g.
+traits or phylogeny [@allen_permutation_2020; @uriarte_trait_2010]; and inform selective logging practices [@canham_neighborhood_2006].
 Although these are well-described methods, few methods are currently available for easy application.
@@
 We address this shortcoming with the `forestecology` R package providing methods and data for forest ecology model fitting and assessment, available on CRAN (https://cran.r-project.org/package=forestecology) and on GitHub (https://github.com/rudeboybert/forestecology).
-The package is written to model stem radial growth between two censuses based on neighborhood competition, largely following the methods in @allen_permutation_2020.
+The package is written to model stem diameter growth between two censuses based on neighborhood competition, largely following the methods in @allen_permutation_2020.
 
-Let $i = 1, \ldots, n_j$ index all $n_j$ trees of "focal" species group $j$; let $j = 1, \ldots, J$ index all $J$ focal species groups; and let $k = 1, \ldots, K$ index all $K$ "competitor" species groups.
-The average annual growth in diameter at breast height (DBH) $y_{ij}$ (in centimeters/year) of the $i^{th}$ tree of focal species group $j$ is modeled as
+Let $i = 1, \ldots, n_j$ index all $n_j$ trees of "focal" species $j$; let $j = 1, \ldots, J$ index all $J$ focal species; and let $k = 1, \ldots, K$ index all $K$ "competitor" species.
+The average annual growth in diameter at breast height (DBH) $y_{ij}$ (in centimeters/year) of the $i^{th}$ tree of focal species $j$ is modeled as
 
@@
 
-where $\beta_{0,j}$ is the diameter-independent growth rate for group $j$; $\text{dbh}_{ij}$ is the DBH of the focal tree at the earlier census for group $j$ and $\beta_{\text{dbh},j}$ quantifies its corresponding growth rate change;  $\text{BA}_{ijk}$ is the sum of the basal area of all trees of competitor species group $k$ and $\lambda_{jk}$ quantifies the corresponding change in growth for individuals of group $j$ from these competitors; and $\epsilon_{ij}$ is a random error term distributed $\text{Normal}(0, \sigma^2)$.
-They estimate all parameters via Bayesian linear regression, while exploiting Normal/Inverse Gamma conjugacy to derive closed-form solutions to all posterior distributions^[See S1 Appendix of @allen_permutation_2020, available at https://doi.org/10.1371/journal.pone.0229930.s004].
+where $\beta_{0,j}$ is the diameter-independent growth rate of species $j$; $\text{dbh}_{ij}$ is the DBH of the focal tree at the earlier census and $\beta_{\text{dbh},j}$ the slope of that species's diameter-growth relationship;  $\text{BA}_{ijk}$ is the sum of the basal area of all trees of competitor species $k$ and $\lambda_{jk}$ quantifies the corresponding change in growth for individuals of group $j$ from these competitors; and $\epsilon_{ij}$ is a random error term distributed $\text{Normal}(0, \sigma^2)$.
+@allen_permutation_2020 estimate all parameters via Bayesian linear regression, while exploiting Normal/Inverse Gamma conjugacy to derive closed-form solutions to all posterior distributions^[See S1 Appendix of @allen_permutation_2020, available at https://doi.org/10.1371/journal.pone.0229930.s004].
 These closed-form solutions are not as computationally expensive as approximations from Markov Chain Monte Carlo algorithms.
@@
 
-We present a case-study of `forestecology`'s functionality on data from the Smithsonian Conservation Biology Institute (SCBI) large forest dynamics plot in Front Royal, VA, USA, part of the ForestGEO global network of research sites [@bourg_initial_2013, @andersonteixeira_ctfs-forestgeo_2015] [@bourg_initial_2013].
+We present a case-study of `forestecology`'s functionality on data from the Smithsonian Conservation Biology Institute (SCBI) large forest dynamics plot in Front Royal, VA, USA, part of the ForestGEO global network of research sites [@bourg_initial_2013, @andersonteixeira_ctfs-forestgeo_2015].
 The 25.6 ha (640 x 400 m) plot is located at the intersection of three of the major physiographic provinces of the eastern US---the Blue Ridge, Ridge and Valley, and Piedmont provinces---and is adjacent to the northern end of Shenandoah National Park.
@@
   geom_sf(data = growth_scbi %>% sample_n(500), aes(size = growth)) + 
-  scale_size_binned(limits = c(0.1, 1))
+  scale_size_binned(limits = c(0.1, 1)) +
+  labs(size = expression(paste(Growth, " (cm ",y^{-1},')')) )
 ```
@@
 
-We observe 4 variables describing 20 competitor trees: the unique tree-stem ID, the distance to the focal tree (all $\leq$ `r comp_dist`), the species, and the basal area (in m$^2$) calculated as $\frac{\pi \times (\text{DBH/2})^2}{10000}$ for the DBH in cm from the earlier census.
+We observe 4 variables describing 20 competitor trees: the unique tree-stem ID, the distance to the focal tree (all $\leq$ `r comp_dist` m), the species, and the basal area (in m$^2$) calculated as $\frac{\pi \times (\text{DBH/2})^2}{10000}$ for the DBH in cm from the earlier census.
 Saving competitor information in list-column format minimizes redundancy since we do not need to repeat information on the focal tree 20 times.
@@
 
-For many users the visualizations of $\lambda_{j,k}$ will be of particular interest as they provide insight into species competitive interactions, where negative values indicate a competitor species which slows the growth of a focal species.
+For many users the visualizations of $\lambda_{j,k}$ will be of particular interest as they provide insight into species-specific competitive interactions, where negative values indicate a competitor species which slows the growth of a focal species.
 Here, for example, we see that tulip poplars (litu) have a strong negative effect on the growth of conspecifics but relatively lesser effect on pignut hickory (cagl) and red oak (quru) neighbors.
@@
 These results suggest that in this 9 ha subplot of the SCBI plot, competitive interactions do not depend on the identity of the competitor, which is the opposite of what has been observed in other locations [@allen_permutation_2020 @uriarte_spatially_2004].
-Furthermore, the larger cross-validated RMSE's are indicative of the importance of being vigilant against potential model overfitting.
+This provides a striking example of the importance of cross-validation, as without it the over-fit model gives rise to an incorrect conclusion.
 
@@
 The `forestecology` package provides an accessible way to fit and test models of neighborhood competition.
-While the package is designed with ForestGEO plot data in mind, we envision that it can be modified to work on any single large, mapped forest plot in which at least two measurements of each individual have been taken, e.g.
-the US Forest Service Forest Inventory and Analysis plots.
-Furthermore, we also envision the package being used more generally to model interactions of any community of mapped sessile organisms [@smith_forest_2002].
+While the package is designed with ForestGEO plot data in mind, we envision that it can be modified to work on any single large, mapped forest plot in which at least two measurements of each individual have been taken.
+Furthermore, we also envision the package being used more generally for inventory plot-structure with many spatially separated plots, e.g.
+the US Forest Service Forest Inventory and Analysis plots [@smith_forest_2002].
 In future versions of `forestecology` we also hope to include models that account for tree mortality in addition to tree growth.

### 2021-03-10 paper/forestecology.tex

@@
 2. We present the \texttt{forestecology} package providing methods to i)
-specify neighborhood competition models, ii) evalulate the effect of
+specify neighborhood competition models, ii) evaluate the effect of
 competitor species identity using permutation tests, and iii) measure
@@
 Smithsonian Conservation Biology Institute's large forest dynamics plot,
-part of the ForestGEO global network of reseach sites. Given ForestGEO's
-data collection protocols and data formatting standards, the package was
-designed with cross-site compatibility in mind. We highlight the
-importance of spatial cross-validation when interpreting model
+part of the ForestGEO global network of research sites. Given
+ForestGEO's data collection protocols and data formatting standards, the
+package was designed with cross-site compatibility in mind. We highlight
+the importance of spatial cross-validation when interpreting model
 results.\\
@@
 clear articulation of all the steps in the sequence of analysis and easy
-wrangling and visualization of the geospatial forestry data.
-Furthermore, while the package only has Bayesian linear regression
-implemented, the package was designed with extensibility to other
-methods in mind.
+wrangling and visualization of the geospatial data. Furthermore, while
+the package only has Bayesian linear regression implemented, the package
+was designed with extensibility to other methods in mind.
 \end{abstract}
@@
 neighborhood models of the effect of competition on the growth of trees
-(\citet{canham_neighborhood_2004}). Neighborhood models of competition
+\citep{canham_neighborhood_2004}. Neighborhood models of competition
 have been used to: test whether the species identity of a competitor
-matters (\citet{uriarte_spatially_2004}); measure species-specific
-competition coefficients (\citet{das_effect_2012}
-\citet{tatsumi_estimating_2016}); test competing models to see what
-structures competitive interactions, e.g.~traits or phylogeny
-(\citet{allen_permutation_2020}; \citet{uriarte_trait_2010}); and inform
-selective logging practices (\citet{canham_neighborhood_2006}). Although
+matters {[}\citet{uriarte_spatially_2004}; measure species-specific
+competition coefficients \citep[
+\citet{tatsumi_estimating_2016}]{das_effect_2012}; test competing models
+to see what structures competitive interactions, e.g.~traits or
+phylogeny \citep{allen_permutation_2020, uriarte_trait_2010}; and inform
+selective logging practices \citep{canham_neighborhood_2006}. Although
 these are well-described methods, few methods are currently available
@@
 assessment, available on CRAN
-(\url{https://cran.r-project.org/web/packages/forestecology/index.html})
-and on GitHub (\url{https://github.com/rudeboybert/forestecology}). The
-package is written to model stem radial growth between two censuses
-based on neighborhood competition, largely following the methods in
+(\url{https://cran.r-project.org/package=forestecology}) and on GitHub
+(\url{https://github.com/rudeboybert/forestecology}). The package is
+written to model stem diameter growth between two censuses based on
+neighborhood competition, largely following the methods in
 \citet{allen_permutation_2020}.
@@
 Let \(i = 1, \ldots, n_j\) index all \(n_j\) trees of ``focal'' species
-group \(j\); let \(j = 1, \ldots, J\) index all \(J\) focal species
-groups; and let \(k = 1, \ldots, K\) index all \(K\) ``competitor''
-species groups. The average annual growth in diameter at breast height
-(DBH) \(y_{ij}\) (in centimeters/year) of the \(i^{th}\) tree of focal
-species group \(j\) is modeled as
+\(j\); let \(j = 1, \ldots, J\) index all \(J\) focal species; and let
+\(k = 1, \ldots, K\) index all \(K\) ``competitor'' species. The average
+annual growth in diameter at breast height (DBH) \(y_{ij}\) (in
+centimeters/year) of the \(i^{th}\) tree of focal species \(j\) is
+modeled as
 
@@
 
-where \(\beta_{0,j}\) is the diameter-independent growth rate for group
+where \(\beta_{0,j}\) is the diameter-independent growth rate of species
 \(j\); \(\text{dbh}_{ij}\) is the DBH of the focal tree at the earlier
-census for group \(j\) and \(\beta_{\text{dbh},j}\) quantifies its
-corresponding growth rate change; \(\text{BA}_{ijk}\) is the sum of the
-basal area of all trees of competitor species group \(k\) and
-\(\lambda_{jk}\) quantifies the corresponding change in growth for
-individuals of group \(j\) from these competitors; and \(\epsilon_{ij}\)
-is a random error term distributed \(\text{Normal}(0, \sigma^2)\). They
-estimate all parameters via Bayesian linear regression, while exploiting
-Normal/Inverse Gamma conjugacy to derive closed-form solutions to all
-posterior distributions\footnote{See S1 Appendix of
-  \citet{allen_permutation_2020}, available at
+census and \(\beta_{\text{dbh},j}\) the slope of that species's
+diameter-growth relationship; \(\text{BA}_{ijk}\) is the sum of the
+basal area of all trees of competitor species \(k\) and \(\lambda_{jk}\)
+quantifies the corresponding change in growth for individuals of group
+\(j\) from these competitors; and \(\epsilon_{ij}\) is a random error
+term distributed \(\text{Normal}(0, \sigma^2)\).
+\citet{allen_permutation_2020} estimate all parameters via Bayesian
+linear regression, while exploiting Normal/Inverse Gamma conjugacy to
+derive closed-form solutions to all posterior distributions\footnote{See
+  S1 Appendix of \citet{allen_permutation_2020}, available at
   \url{https://doi.org/10.1371/journal.pone.0229930.s004}}. These
@@
 global network of research sites
-\citep[\citet{andersonteixeira_ctfs-forestgeo_2015}]{bourg_initial_2013}
-\citep{bourg_initial_2013}. The 25.6 ha (640 x 400 m) plot is located at
-the intersection of three of the major physiographic provinces of the
-eastern US---the Blue Ridge, Ridge and Valley, and Piedmont
-provinces---and is adjacent to the northern end of Shenandoah National
-Park.
+\citep[\citet{andersonteixeira_ctfs-forestgeo_2015}]{bourg_initial_2013}.
+The 25.6 ha (640 x 400 m) plot is located at the intersection of three
+of the major physiographic provinces of the eastern US---the Blue Ridge,
+Ridge and Valley, and Piedmont provinces---and is adjacent to the
+northern end of Shenandoah National Park.
 
@@
 \StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{sample_n}\NormalTok{(}\DecValTok{500}\NormalTok{), }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{size =}\NormalTok{ growth)) }\OperatorTok{+}\StringTok{ }
-\StringTok{  }\KeywordTok{scale_size_binned}\NormalTok{(}\DataTypeTok{limits =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\DecValTok{1}\NormalTok{))}
+\StringTok{  }\KeywordTok{scale_size_binned}\NormalTok{(}\DataTypeTok{limits =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\DecValTok{1}\NormalTok{)) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{size =} \KeywordTok{expression}\NormalTok{(}\KeywordTok{paste}\NormalTok{(Growth, }\StringTok{" (cm "}\NormalTok{,y}\OperatorTok{^}\NormalTok{\{}\OperatorTok{-}\DecValTok{1}\NormalTok{\},}\StringTok{')'}\NormalTok{)) )}
 \end{Highlighting}
@@
 We observe 4 variables describing 20 competitor trees: the unique
-tree-stem ID, the distance to the focal tree (all \(\leq\) 7.5), the
+tree-stem ID, the distance to the focal tree (all \(\leq\) 7.5 m), the
 species, and the basal area (in m\(^2\)) calculated as
@@
 \NormalTok{comp_bayes_lm_scbi}
-\CommentTok{## Bayesian linear regression model parameters with a multivariate Normal}
-\CommentTok{## likelihood. See ?comp_bayes_lm for details:}
+\CommentTok{## Bayesian linear regression model parameters with a multivariate Normal likelihood. See ?comp_bayes_lm for details:}
 \CommentTok{## }
@@
 \CommentTok{## Model formula:}
-\CommentTok{## growth ~ sp + dbh + dbh * sp + acne * sp + acru * sp + amar * sp + astr}
-\CommentTok{## * sp + caca * sp + caco * sp + cade * sp + cagl * sp + caovl * sp + cato}
-\CommentTok{## * sp + ceca * sp + ceoc * sp + chvi * sp + cofl * sp + crpr * sp + crsp}
-\CommentTok{## * sp + divi * sp + elum * sp + fagr * sp + fram * sp + frni * sp + frpe}
-\CommentTok{## * sp + havi * sp + ilve * sp + juci * sp + juni * sp + libe * sp + litu}
-\CommentTok{## * sp + nysy * sp + pist * sp + pivi * sp + ploc * sp + prav * sp + prse}
-\CommentTok{## * sp + qual * sp + quco * sp + qufa * sp + qumi * sp + qupr * sp + quru}
-\CommentTok{## * sp + quve * sp + rops * sp + saal * sp + saca * sp + tiam * sp + ulam}
-\CommentTok{## * sp + ulru * sp + unk * sp + vipr * sp}
+\CommentTok{## growth ~ sp + dbh + dbh * sp + acne * sp + acru * sp + amar * sp + astr * sp + caca * sp + caco * sp + cade * sp + cagl * sp + caovl * sp + cato * sp + ceca * sp + ceoc * sp + chvi * sp + cofl * sp + crpr * sp + crsp * sp + divi * sp + elum * sp + fagr * sp + fram * sp + frni * sp + frpe * sp + havi * sp + ilve * sp + juci * sp + juni * sp + libe * sp + litu * sp + nysy * sp + pist * sp + pivi * sp + ploc * sp + prav * sp + prse * sp + qual * sp + quco * sp + qufa * sp + qumi * sp + qupr * sp + quru * sp + quve * sp + rops * sp + saal * sp + saca * sp + tiam * sp + ulam * sp + ulru * sp + unk * sp + vipr * sp}
 \end{Highlighting}
@@
 For many users the visualizations of \(\lambda_{j,k}\) will be of
-particular interest as they provide insight into species competitive
-interactions, where negative values indicate a competitor species which
-slows the growth of a focal species. Here, for example, we see that
-tulip poplars (litu) have a strong negative effect on the growth of
-conspecifics but relatively lesser effect on pignut hickory (cagl) and
-red oak (quru) neighbors.
+particular interest as they provide insight into species-specific
+competitive interactions, where negative values indicate a competitor
+species which slows the growth of a focal species. Here, for example, we
+see that tulip poplars (litu) have a strong negative effect on the
+growth of conspecifics but relatively lesser effect on pignut hickory
+(cagl) and red oak (quru) neighbors.
 
@@
 what has been observed in other locations \citep[
-\citet{uriarte_spatially_2004}]{allen_permutation_2020}. Furthermore,
-the larger cross-validated RMSE's are indicative of the importance of
-being vigilant against potential model overfitting.
+\citet{uriarte_spatially_2004}]{allen_permutation_2020}. This provides a
+striking example of the importance of cross-validation, as without it
+the over-fit model gives rise to an incorrect conclusion.
 
@@
 work on any single large, mapped forest plot in which at least two
-measurements of each individual have been taken, e.g.~the US Forest
-Service Forest Inventory and Analysis plots. Furthemore, we also
-envision the package being used more generally to model interactions of
-any community of mapped sessile organisms \citep{smith_forest_2002}. In
-future versions of \texttt{forestecology} we also hope to include models
-that account for tree mortality in addition to tree growth. The package
-follows the \texttt{tidy} data design principles, leverages the
+measurements of each individual have been taken. Furthermore, we also
+envision the package being used more generally for inventory
+plot-structure with many spatially separated plots, e.g.~the US Forest
+Service Forest Inventory and Analysis plots \citep{smith_forest_2002}.
+In future versions of \texttt{forestecology} we also hope to include
+models that account for tree mortality in addition to tree growth. The
+package follows the \texttt{tidy} data design principles, leverages the
 \texttt{sf} package for spatial data, and S3 open-oriented model

### 2021-03-10 paper/forestecology.tex
Re-knit manuscript just in case

@@
 \CommentTok{## Model formula:}
-\CommentTok{## growth ~ sp + dbh + dbh * sp + acne * sp + acru * sp + amar * sp + astr * sp + caca * sp + caco * sp + cade * sp + cagl * sp + caovl * sp + cato * sp + ceca * sp + ceoc * sp + chvi * sp + cofl * sp + crpr * sp + crsp * sp + divi * sp + elum * sp + fagr * sp + fram * sp + frni * sp + frpe * sp + havi * sp + ilve * sp + juci * sp + juni * sp + libe * sp + litu * sp + nysy * sp + pist * sp + pivi * sp + ploc * sp + prav * sp + prse * sp + qual * sp + quco * sp + qufa * sp + qumi * sp + qupr * sp + quru * sp + quve * sp + rops * sp + saal * sp + saca * sp + tiam * sp + ulam * sp + ulru * sp + unk * sp + vipr * sp}
+\CommentTok{## growth ~ sp + dbh + dbh * sp + acne * sp + acru * sp + amar * sp + astr}
+\CommentTok{## * sp + caca * sp + caco * sp + cade * sp + cagl * sp + caovl * sp + cato}
+\CommentTok{## * sp + ceca * sp + ceoc * sp + chvi * sp + cofl * sp + crpr * sp + crsp}
+\CommentTok{## * sp + divi * sp + elum * sp + fagr * sp + fram * sp + frni * sp + frpe}
+\CommentTok{## * sp + havi * sp + ilve * sp + juci * sp + juni * sp + libe * sp + litu}
+\CommentTok{## * sp + nysy * sp + pist * sp + pivi * sp + ploc * sp + prav * sp + prse}
+\CommentTok{## * sp + qual * sp + quco * sp + qufa * sp + qumi * sp + qupr * sp + quru}
+\CommentTok{## * sp + quve * sp + rops * sp + saal * sp + saca * sp + tiam * sp + ulam}
+\CommentTok{## * sp + ulru * sp + unk * sp + vipr * sp}
 \end{Highlighting}

### 2021-03-11 paper/paper.Rmd
Switched image png -> pdf

--- 2021-03-10 paper/forestecology.Rmd
@@
 
-```{r scbi-simulation, out.width="100%", fig.cap="Root mean squared error of models for standard, permuted, and spatial cross-validated error estimates.
+```{r scbi-simulation, out.width="100%", fig.cap="Comparison of root mean squared error of models for standard, permuted, and spatial cross-validated error estimates.
 The dotted lines show non-permuted competitor identity, while the histogras so the RMSE for 49 permutations.
 The colors indicate whether cross validaton was used.", echo = FALSE}
-include_graphics("simulation_results/2021-03-03_scbi_49_shuffles.png")
+include_graphics("simulation_results/2021-03-03_scbi_49_shuffles.pdf")
 ```

### 2021-03-12 README.Rmd
README update
Adding code coverage

@@
 [![Build Status](https://github.com/rudeboybert/forestecology/workflows/R-CMD-check/badge.svg)](https://github.com/rudeboybert/forestecology/actions)
-[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
+[![Codecov test coverage](https://codecov.io/gh/rudeboybert/forestecology/branch/master/graph/badge.svg)](https://codecov.io/gh/rudeboybert/forestecology?branch=master)
 [![CRAN status](https://www.r-pkg.org/badges/version/forestecology)](https://cran.r-project.org/package=forestecology)
@@
 library(forestecology)
+library(patchwork)
 library(blockCV)
-library(patchwork)
+
+# Resolve conflicting functions
+filter <- dplyr::filter
 ```
@@
 ```
-
-<!--For fuller examples on actual datasets see ___.-->

### 2021-03-14 README.Rmd
- Added a little info to the citation info. - In the README removed the step where we make sp a factor, because now that happens under-the-hood.

@@
     census_1 = census_1_ex %>% 
-      mutate(sp = to_any_case(sp) %>% factor()),
+      mutate(sp = to_any_case(sp)),
     census_2 = census_2_ex %>% 
       filter(!str_detect(codes, "R")) %>% 
-      mutate(sp = to_any_case(sp) %>% factor()),
+      mutate(sp = to_any_case(sp)),
     id = "ID"

### 2021-05-10 paper/paper.Rmd
Adding future work rubric

--- 2021-03-10 paper/forestecology.Rmd
@@
 
-```{r scbi-trees, out.width="66%", fig.cap="Compute growth of trees based on census data.
+```{r scbi-trees, out.width="66%", fig.cap="Step 1 - Compute growth of trees based on census data.
 A map of the growth of a random sample of 500 trees from a 9 ha subsection of the Smithsonian Conservation Biology Institute (SCBI) forest plot."}
@@
 
-```{r scbi-spatial-information, out.width="66%", fig.cap="Add spatial information.
-A buffer region and spatial cross-validation blocks 1 througth 4.
+```{r scbi-spatial-information, out.width="66%", fig.cap="Step 2 - Add spatial information.
+A buffer region and spatial cross-validation blocks 1 through 4.
 The location of each tree is marked with its fold number where the folds are delineated with solid lines.
@@
 
-```{r scbi-focal-vs-comp-map, out.width="66%", echo = FALSE, fig.cap="Identify all focal and corresponding competitor trees.
-The dashed circle extends 7.5m away from the focal tree 4 while all 20 competitor trees are within this circle."}
+```{r scbi-focal-vs-comp-map, out.width="66%", echo = FALSE, fig.cap="Step 3 - Identify all focal and corresponding competitor trees.
+The dashed circle extends 7.5m away from the focal tree 4 while all 20 competitor trees are within this circle.", fig.height=4}
 comp_radius <- growth_scbi %>% 
@@
 
-```{r scbi-posterior-viz, out.width="100%", fig.cap= "Fit model. Posterior distributions of all parameters. For compactness we include only three species.", fig.width = 8, fig.height = 6}
+```{r scbi-posterior-viz, out.width="100%", fig.cap= "Step 4 - Fit model. Posterior distributions of all parameters. For compactness we include only three species.", fig.width = 8, fig.height = 6}
 # Plot posteriors for only a subset of species
@@
 This algorithm is implemented in `run_cv()`, which acts as a wrapper function to both `comp_bayes_lm()` that fits the model and `predict()` that returns predicted values.
+Furthermore, `comp_dist` can now be specified as a vector of competitive distances to consider.
 
@@
 
-```{r scbi-simulation, out.width="100%", fig.cap="Root mean squared error of models for standard, permuted, and spatial cross-validated error estimates.
-The dotted lines show non-permuted competitor identity, while the histogras so the RMSE for 49 permutations.
-The colors indicate whether cross validaton was used.", echo = FALSE}
-include_graphics("simulation_results/2021-03-03_scbi_49_shuffles.png")
+```{r scbi-simulation, out.width="100%", fig.cap="Comparison of root mean squared error of models for standard, permuted, and spatially cross-validated error estimates.
+The dotted lines show observed RMSE while the histograms show the null distribution of RMSE for 49 permutations under the null hypothesis of no competitor species identity effects.
+The colors indicate whether spatial cross-validation was used or not.", echo = FALSE}
+include_graphics("simulation_results/2021-03-03_scbi_49_shuffles.pdf")
 ```
 
-# Conclusion
+# Conclusion and future work
 

### 2021-05-10 paper/paper.Rmd
First pass at Conclusion -> Future Work

@@
 The `forestecology` package provides an accessible way to fit and test models of neighborhood competition.
+The package follows the `tidy` data design principles, leverages the `sf` package for spatial data, and S3 open-oriented model implementation structure.
+We hope that the package will increase the use of neighborhood competition models to better understand what structures plant competition.
+
 While the package is designed with ForestGEO plot data in mind, we envision that it can be modified to work on any single large, mapped forest plot in which at least two measurements of each individual have been taken.
@@
 the US Forest Service Forest Inventory and Analysis plots [@smith_forest_2002].
-In future versions of `forestecology` we also hope to include models that account for tree mortality in addition to tree growth.
-The package follows the `tidy` data design principles, leverages the `sf` package for spatial data, and S3 open-oriented model implementation structure.
-We hope that the package will increase the use of neighborhood competition models to better understand what structures plant competition.
+
+We hope to extend the `forestecology` package's functionality to account for a larger variety of models for tree growth and corresponding tests. Some immediate improvements that can made include
+
+> To address the reviewer's critique that the model is not flexible enough and does not include applications of existing competition theory, we could add another option for run_shuffle.
+In this case the shuffle would not shuffle identity of competitors which share with the focal species.
+The idea being there is a special kind of competition with conspecifics (because they share your exact resource needs).
+The result is two lambda values, one for competition between conspecifics and one for competition between heterospecifics.
+
+A recently emerging body of literature suggests that traits of species rather than species identity themselves better inform competition models [@kunstler_competitive_2012, @lasky_trait-mediated_2014].
+To account for such clustering, instead of using Equation \ref{eq:model}, we present a model of which is a modification of @kunstler_plant_2016 in Equation \ref{eq:model2}.
+Let $Y_{ik}$ be the growth of the $k^{\text{th}}$ individual of species $i$ and let $j$ be the species of competitors in the neighborhood.
+How much growth changes per unit of $\text{BA}$ of all neighbors is quantified by two terms: an overall competition coefficient $C$ and an $i$-$j$ species pair specific coefficient $c_{ij}$.
+which depends on the traits of the $i$ and $j$ species.
+Say we have $l$ traits and $t_{li}$ is the value of trait $l$ for species $i$.
+$a_{\text{tol},l}$ quantifies how trait $l$ affects the way a species tolerates (or is sensitive to) competition.
+$a_{\text{eff},l}$ quantifies how trait $l$ is how much trait $l$ affects how strong a competitor is, how much it effects the reduction in growth of its neighbors.
+Finally $a_{\text{sim},l}$ is about whether competition is stronger (or weaker) if the two species are similar.
+This is based on an idea that two species that are similar use the same resources and should compete more.
+
+\label{eq:model2}
+\begin{eqnarray}
+y_{ik} &=& \beta_{0,j} + \beta_{\text{dbh},j} \cdot \text{dbh}_{ij} + \sum_{j} (C + c_{ij}) \cdot \text{BA}_{j} + \epsilon_{ij}\\
+\text{where } c_{ij} &=& \sum_{l} a_{\text{tol},l}\cdot t_{li} + a_{\text{eff},l}\cdot t_{lj} + a_{\text{sim},l}|t_{li} - t_{lj}|\nonumber
+\end{eqnarray}
+
+Lastly, the `forestecology` current uses the `blockCV` package behind the scenes to create the spatial blocks acting as folds for our spatial cross-validation algorithm detailed in Sections \ref{spatial-information} and \ref{spatial-cross-validation} [@valavi_blockcv_2019].
+This back-end functionality could be substituted with the `spatialsample` package for spatial resampling infrastructure; a `tidymodels` package under active development as of 2021 [@spatialsample_package, @tidymodels_package].
 

### 2021-05-10 paper/paper.tex
Switched image png -> pdf
First pass at Conclusion -> Future Work

--- 2021-03-10 paper/forestecology.tex
@@
 
-\caption{Compute growth of trees based on census data.
+\caption{Step 1 - Compute growth of trees based on census data.
 A map of the growth of a random sample of 500 trees from a 9 ha subsection of the Smithsonian Conservation Biology Institute (SCBI) forest plot.}\label{fig:scbi-trees}
@@
 
-\caption{Add spatial information.
-A buffer region and spatial cross-validation blocks 1 througth 4.
+\caption{Step 2 - Add spatial information.
+A buffer region and spatial cross-validation blocks 1 through 4.
 The location of each tree is marked with its fold number where the folds are delineated with solid lines.
@@
 
-\caption{Identify all focal and corresponding competitor trees.
+\caption{Step 3 - Identify all focal and corresponding competitor trees.
 The dashed circle extends 7.5m away from the focal tree 4 while all 20 competitor trees are within this circle.}\label{fig:scbi-focal-vs-comp-map}
@@
 
-\caption{Fit model. Posterior distributions of all parameters. For compactness we include only three species.}\label{fig:scbi-posterior-viz}
+\caption{Step 4 - Fit model. Posterior distributions of all parameters. For compactness we include only three species.}\label{fig:scbi-posterior-viz}
 \end{figure}
@@
 the model and \texttt{predict()} that returns predicted values.
+Furthermore, \texttt{comp\_dist} can now be specified as a vector of
+competitive distances to consider.
 
@@
 
-\caption{Root mean squared error of models for standard, permuted, and spatial cross-validated error estimates.
-The dotted lines show non-permuted competitor identity, while the histogras so the RMSE for 49 permutations.
-The colors indicate whether cross validaton was used.}\label{fig:scbi-simulation}
+\caption{Comparison of root mean squared error of models for standard, permuted, and spatially cross-validated error estimates.
+The dotted lines show observed RMSE while the histograms show the null distribution of RMSE for 49 permutations under the null hypothesis of no competitor species identity effects.
+The colors indicate whether spatial cross-validation was used or not.}\label{fig:scbi-simulation}
 \end{figure}
 
-\hypertarget{conclusion}{%
-\section{Conclusion}\label{conclusion}}
+\hypertarget{conclusion-and-future-work}{%
+\section{Conclusion and future work}\label{conclusion-and-future-work}}
 
 The \texttt{forestecology} package provides an accessible way to fit and
-test models of neighborhood competition. While the package is designed
-with ForestGEO plot data in mind, we envision that it can be modified to
-work on any single large, mapped forest plot in which at least two
-measurements of each individual have been taken. Furthermore, we also
-envision the package being used more generally for inventory
-plot-structure with many spatially separated plots, e.g.~the US Forest
-Service Forest Inventory and Analysis plots \citep{smith_forest_2002}.
-In future versions of \texttt{forestecology} we also hope to include
-models that account for tree mortality in addition to tree growth. The
-package follows the \texttt{tidy} data design principles, leverages the
-\texttt{sf} package for spatial data, and S3 open-oriented model
-implementation structure. We hope that the package will increase the use
-of neighborhood competition models to better understand what structures
-plant competition.
+test models of neighborhood competition. The package follows the
+\texttt{tidy} data design principles, leverages the \texttt{sf} package
+for spatial data, and S3 open-oriented model implementation structure.
+We hope that the package will increase the use of neighborhood
+competition models to better understand what structures plant
+competition.
+
+While the package is designed with ForestGEO plot data in mind, we
+envision that it can be modified to work on any single large, mapped
+forest plot in which at least two measurements of each individual have
+been taken. Furthermore, we also envision the package being used more
+generally for inventory plot-structure with many spatially separated
+plots, e.g.~the US Forest Service Forest Inventory and Analysis plots
+\citep{smith_forest_2002}.
+
+We hope to extend the \texttt{forestecology} package's functionality to
+account for a larger variety of models for tree growth and corresponding
+tests. Some immediate
+
+\begin{quote}
+To address the reviewer's critique that the model is not flexible enough
+and does not include applications of existing competition theory, we
+could add another option for run\_shuffle. In this case the shuffle
+would not shuffle identity of competitors which share with the focal
+species. The idea being there is a special kind of competition with
+conspecifics (because they share your exact resource needs). The result
+is two lambda values, one for competition between conspecifics and one
+for competition between heterospecifics.
+\end{quote}
+
+A recently emerging body of literature suggests that traits of species
+rather than species identity themselves better inform competition models
+\citep[\citet{lasky_trait-mediated_2014}]{kunstler_competitive_2012}. To
+account for such clustering, instead of using Equation \ref{eq:model},
+we could use a model of which is a modification of
+\citet{kunstler_plant_2016}. Let \(Y_{ik}\) be the growth of the
+\(k^{\text{th}}\) individual of species \(i\) and let \(j\) be the
+species of competitors in the neighborhood. How much growth changes per
+unit of \(\text{BA}\) of all neighbors is quantified by two terms: an
+overall competition coefficient \(C\) and an \(i\)-\(j\) species pair
+specific coefficient \(c_{ij}\). which depends on the traits of the
+\(i\) and \(j\) species. Say we have \(l\) traits and \(t_{li}\) is the
+value of trait \(l\) for species \(i\). \(a_{\text{tol},l}\) quantifies
+how trait \(l\) affects the way a species tolerates (or is sensitive to)
+competition. \(a_{\text{eff},l}\) quantifies how trait \(l\) is how much
+trait \(l\) affects how strong a competitor is, how much it effects the
+reduction in growth of its neighbors. Finally \(a_{\text{sim},l}\) is
+about whether competition is stronger (or weaker) if the two species are
+similar. This is based on an idea that two species that are similar use
+the same resources and should compete more.
+
+\label{eq:model2} \begin{eqnarray}
+y_{ik} &=& \beta_{0,j} + \beta_{\text{dbh},j} \cdot \text{dbh}_{ij} + \sum_{j} (C + c_{ij}) \cdot \text{BA}_{j} + \epsilon_{ij}\\
+\text{where } c_{ij} &=& \sum_{l} a_{\text{tol},l}\cdot t_{li} + a_{\text{eff},l}\cdot t_{lj} + a_{\text{sim},l}|t_{li} - t_{lj}|\nonumber
+\end{eqnarray}
+
+Lastly, the \texttt{forestecology} current uses the \texttt{blockCV}
+package behind the scenes to create the spatial blocks acting as folds
+for our spatial cross-validation algorithm detailed in Sections
+\ref{spatial-information} and \ref{spatial-cross-validation}
+\citep{valavi_blockcv_2019}. This back-end functionality could be
+substituted with the \texttt{spatialsample} package for spatial
+resampling infrastructure; a \texttt{tidymodels} package under active
+development as of 2021
+\citep[\citet{tidymodels_package}]{spatialsample_package}.
 

### 2021-05-12 paper/paper.Rmd
Moved code to Appendix

--- 2021-05-10 paper/paper.Rmd
@@
 We first manually define an `sf` object defining four folds that partition the study region.
-We then use the output of the `spatialBlock()` function from the `blockCV` package to associate each tree in `growth_scbi` to the correct `foldID` [@valavi_blockcv_2019].^[In the Supporting Information we present an example where the folds themselves are created automatically, as opposed to manually as in the example.] This `foldID` variable will be used in Section \ref{spatial-cross-validation}.
+We then use the output of the `spatialBlock()` function from the `blockCV` package to associate each tree in `growth_scbi` to the correct `foldID` [@valavi_blockcv_2019].
+This `foldID` variable will be used in Section \ref{spatial-cross-validation}.
 
@@
 This algorithm is implemented in `run_cv()`, which acts as a wrapper function to both `comp_bayes_lm()` that fits the model and `predict()` that returns predicted values.
-Furthermore, `comp_dist` can now be specified as a vector of competitive distances to consider.
 
@@
 
-# Conclusion and future work
+# Conclusion
 
@@
 
+# Appendix
+
+This code replicates Figure \ref{fig:scbi-simulation}: A comparison of root mean squared error of
+models for standard, permuted, and spatial cross- validated error estimates.
+
+```{r, eval = FALSE}
+library(tidyverse)
+library(lubridate)
+library(here)
+library(sf)
+library(viridis)
+library(forestecology)
+library(blockCV)
+library(tictoc)
+
+# Compute growth of trees based on census data -------------------------
+census_2013_scbi <- here("paper/scbi.stem2.csv") %>%
+  read_csv() %>%
+  select(stemID, sp, date = ExactDate, gx, gy, dbh, codes, status) %>%
+  mutate(
+    date = mdy(date),
+    dbh = as.numeric(dbh)/10
+  ) %>%
+  filter(gx < 300, between(gy, 300, 600))
+
+census_2018_scbi <- here("paper/scbi.stem3.csv") %>%
+  read_csv() %>%
+  select(stemID, sp, date = ExactDate, gx, gy, dbh, codes, status) %>%
+  mutate(
+    date = mdy(date),
+    dbh = as.numeric(dbh)/10
+  ) %>%
+  filter(gx < 300, between(gy, 300, 600))
+
+growth_scbi <-
+  compute_growth(
+    census_1 = census_2013_scbi,
+    census_2 = census_2018_scbi %>% filter(!str_detect(codes, "R")),
+    id = "stemID"
+  )
+
+# Add spatial information ----------------------------------------------
+# Define buffer region using competitive distance range
+comp_dist <- 7.5
+
+study_region_scbi <- tibble(
+  x = c(0, 300, 300, 0, 0),
+  y = c(300, 300, 600, 600, 300)
+) %>%
+  sf_polygon()
+
+growth_scbi <- growth_scbi %>%
+  add_buffer_variable(size = comp_dist, region = study_region_scbi)
+
+# Manually define spatial blocks to act as folds
+fold1 <- rbind(c(0, 300), c(150, 300), c(150, 450), c(0, 450))
+fold2 <- rbind(c(150, 300), c(300, 300), c(300, 450), c(150, 450))
+fold3 <- rbind(c(0, 450), c(150, 450), c(150, 600), c(0, 600))
+fold4 <- rbind(c(150, 450), c(300, 450), c(300, 600), c(150, 600))
+n_fold <- 4
+
+blocks_scbi <- bind_rows(
+  sf_polygon(fold1), sf_polygon(fold2), 
+  sf_polygon(fold3), sf_polygon(fold4)
+) %>%
+  mutate(folds = c(1:n_fold) %>% factor())
+
+# Associate each observation to a fold
+SpatialBlock_scbi <- spatialBlock(
+  speciesData = growth_scbi, k = n_fold, selection = "systematic",
+  blocks = blocks_scbi, showBlocks = FALSE, verbose = FALSE
+)
+
+growth_scbi <- growth_scbi %>%
+  mutate(foldID = SpatialBlock_scbi$foldID %>% factor())
+
+# Compute focal versus competitor tree information ---------------------
+focal_vs_comp_scbi <- growth_scbi %>%
+  create_focal_vs_comp(comp_dist, blocks = blocks_scbi, id = "stemID")
+
+# Fit model and make predictions ---------------------------------------
+# Number of permutation shuffles:
+num_shuffle <- 49
+
+# Save results here
+run_time <- 0
+observed_RMSE <- 0
+observed_RMSE_CV <- 0
+shuffle_RMSE <- vector("list", 1)
+shuffle_RMSE_CV <- vector("list", 1)
+filename <- here("paper/simulation_results/") %>%
+  str_c("2021-03-03_scbi_", num_shuffle, "_shuffles")
+
+# Run all simulations
+# 0. Setup simulation for this species type ----
+# Start clock
+tic()
+
+# 1. Compute observed test statistic: RMSE with no cross-validation ----
+# Fit model (compute posterior parameters)
+comp_bayes_lm_scbi <- focal_vs_comp_scbi %>%
+  comp_bayes_lm(prior_param = NULL, run_shuffle = FALSE)
+
+# Make predictions and compute RMSE
+observed_RMSE <- focal_vs_comp_scbi %>%
+  mutate(growth_hat = 
+           predict(comp_bayes_lm_scbi, focal_vs_comp_scbi)) %>%
+  rmse(truth = growth, estimate = growth_hat) %>%
+  pull(.estimate)
+
+# 2. Compute observed test statistic: RMSE with cross-validation -------
+observed_RMSE_CV <- focal_vs_comp_scbi %>%
+  run_cv(comp_dist = comp_dist, blocks = blocks_scbi) %>%
+  rmse(truth = growth, estimate = growth_hat) %>%
+  pull(.estimate)
+
+# 3. Permutation distribution: RMSE with no cross-validation -----------
+# Compute num_shuffle permutation test statistics
+shuffle_RMSE <- numeric(length = num_shuffle)
+for(j in 1:num_shuffle){
+  # Fit model (compute posterior parameters) with shuffling
+  comp_bayes_lm_scbi <- focal_vs_comp_scbi %>%
+    comp_bayes_lm(prior_param = NULL, run_shuffle = TRUE)
+
+  # Make predictions and compute RMSE
+  shuffle_RMSE[j] <-  focal_vs_comp_scbi %>%
+    mutate(growth_hat = 
+             predict(comp_bayes_lm_scbi, focal_vs_comp_scbi)) %>%
+    rmse(truth = growth, estimate = growth_hat) %>%
+    pull(.estimate)
+}
+
+# 4. Permutation distribution: RMSE with cross-validation --------------
+# Compute num_shuffle permutation test statistics
+shuffle_RMSE_CV <- numeric(length = num_shuffle)
+
+# Compute num_shuffle permutation test statistics
+for(j in 1:num_shuffle){
+  # Compute and save RMSE
+  shuffle_RMSE_CV[j] <- focal_vs_comp_scbi %>%
+    run_cv(comp_dist = comp_dist, blocks = blocks_scbi, 
+           run_shuffle = TRUE) %>%
+    rmse(truth = growth, estimate = growth_hat) %>%
+    pull(.estimate)
+
+  # Status update
+  str_c("Shuffle with permutation ", j, " at ", Sys.time()) %>% print()
+}
+
+# 5. Save results ----
+clock <- toc(quiet = TRUE)
+run_time <- clock$toc - clock$tic
+
+model_comp_tbl <- tibble(
+  run_time = run_time,
+  observed_RMSE = observed_RMSE,
+  observed_RMSE_CV = observed_RMSE_CV,
+  shuffle_RMSE = shuffle_RMSE,
+  shuffle_RMSE_CV = shuffle_RMSE_CV,
+)
+save(model_comp_tbl, file = filename %>% str_c(".RData"))
+
+# Visualize results ----------------------------------------------------
+model_comp <- bind_rows(
+  model_comp_tbl %>% 
+    select(run_time, observed = observed_RMSE, 
+           shuffle = shuffle_RMSE) %>% 
+    mutate(CV = FALSE),
+  model_comp_tbl %>% 
+    select(run_time, observed = observed_RMSE_CV, 
+           shuffle = shuffle_RMSE_CV) %>% 
+    mutate(CV = TRUE)
+) %>%
+  gather(type, RMSE, -c(run_time, CV))
+
+model_comp_observed <- model_comp %>%
+  filter(type == "observed") %>%
+  unnest(cols = c(RMSE))
+model_comp_shuffle <- model_comp %>%
+  filter(type == "shuffle") %>%
+  unnest(cols = c(RMSE))
+
+cv_plot <- ggplot() +
+  geom_vline(data = model_comp_observed, 
+             aes(xintercept = RMSE, col = CV), 
+             linetype = "dashed", show.legend = F) +
+  geom_histogram(data = model_comp_shuffle, 
+                 aes(x = RMSE, fill = CV), bins = 50) +
+  labs(
+    fill = "Cross-validated?",
+    x = expression(paste("RMSE (cm ", y^{-1}, ")"))
+  ) +
+  scale_color_viridis(discrete = TRUE, option = "D") +
+  scale_fill_viridis(discrete = TRUE) +
+  theme_light()
+cv_plot
+
+filename %>%
+  str_c(".pdf") %>%
+  ggsave(plot = cv_plot, width = 16/2, height = 9/2)
+```
+
 # References

### 2021-05-13 paper/paper.Rmd
Dave future work edit

--- 2021-05-10 paper/paper.Rmd
@@
 The `forestecology` package provides an accessible way to fit and test models of neighborhood competition.
-While the package is designed with ForestGEO plot data in mind, we envision that it can be modified to work on any single large, mapped forest plot in which at least two measurements of each individual have been taken.
-Furthermore, we also envision the package being used more generally for inventory plot-structure with many spatially separated plots, e.g.
-the US Forest Service Forest Inventory and Analysis plots [@smith_forest_2002].
-In future versions of `forestecology` we also hope to include models that account for tree mortality in addition to tree growth.
 The package follows the `tidy` data design principles, leverages the `sf` package for spatial data, and S3 open-oriented model implementation structure.
 We hope that the package will increase the use of neighborhood competition models to better understand what structures plant competition.
+
+While the package is designed with ForestGEO plot data in mind, we envision that it can be modified to work on any single large, mapped forest plot in which at least two measurements of each individual have been taken.
+Furthermore, we hope that future versions of the package will be flexible to other plot layouts, for example inventory plot-structure with many spatially separated plots like the US Forest Service Forest Inventory and Analysis plots [@smith_forest_2002].
+
+We also hope to extend the `forestecology` package's functionality to account for a larger variety of models for tree growth.
+One clear future direction would be to allow competition based on species trait values rather than species identity.
+There is evidence that traits predict competitive outcomes [@kunstler_competitive_2012, @lasky_trait-mediated_2014, @uriarte_trait_2010].
+Thus an extenstion of the model would allow $lambda$ values to be a function of the traits of competing species.
+
+Lastly, the `forestecology` current uses the `blockCV` package behind the scenes to create the spatial blocks acting as folds for our spatial cross-validation algorithm detailed in Sections \ref{spatial-information} and \ref{spatial-cross-validation} [@valavi_blockcv_2019].
+This back-end functionality could be substituted with the `spatialsample` package for spatial resampling infrastructure; a `tidymodels` package under active development as of 2021 [@spatialsample_package, @tidymodels_package].
 

### 2021-05-13 paper/paper.tex
Dave future work edit

--- 2021-03-10 paper/forestecology.tex
@@
 
-\caption{Root mean squared error of models for standard, permuted, and spatial cross-validated error estimates.
-The dotted lines show non-permuted competitor identity, while the histogras so the RMSE for 49 permutations.
-The colors indicate whether cross validaton was used.}\label{fig:scbi-simulation}
+\caption{Comparison of root mean squared error of models for standard, permuted, and spatially cross-validated error estimates.
+The dotted lines show observed RMSE while the histograms show the null distribution of RMSE for 49 permutations under the null hypothesis of no competitor species identity effects.
+The colors indicate whether spatial cross-validation was used or not.}\label{fig:scbi-simulation}
 \end{figure}
 
-\hypertarget{conclusion}{%
-\section{Conclusion}\label{conclusion}}
+\hypertarget{conclusion-and-future-work}{%
+\section{Conclusion and future work}\label{conclusion-and-future-work}}
 
 The \texttt{forestecology} package provides an accessible way to fit and
-test models of neighborhood competition. While the package is designed
-with ForestGEO plot data in mind, we envision that it can be modified to
-work on any single large, mapped forest plot in which at least two
-measurements of each individual have been taken. Furthermore, we also
-envision the package being used more generally for inventory
-plot-structure with many spatially separated plots, e.g.~the US Forest
+test models of neighborhood competition. The package follows the
+\texttt{tidy} data design principles, leverages the \texttt{sf} package
+for spatial data, and S3 open-oriented model implementation structure.
+We hope that the package will increase the use of neighborhood
+competition models to better understand what structures plant
+competition.
+
+While the package is designed with ForestGEO plot data in mind, we
+envision that it can be modified to work on any single large, mapped
+forest plot in which at least two measurements of each individual have
+been taken. Furthermore, we hope that future versions of the package
+will be flexible to other plot layouts, for example inventory
+plot-structure with many spatially separated plots like the US Forest
 Service Forest Inventory and Analysis plots \citep{smith_forest_2002}.
-In future versions of \texttt{forestecology} we also hope to include
-models that account for tree mortality in addition to tree growth. The
-package follows the \texttt{tidy} data design principles, leverages the
-\texttt{sf} package for spatial data, and S3 open-oriented model
-implementation structure. We hope that the package will increase the use
-of neighborhood competition models to better understand what structures
-plant competition.
+
+We also hope to extend the \texttt{forestecology} package's
+functionality to account for a larger variety of models for tree growth.
+One clear future direction would be to allow competition based on
+species trait values rather than species identity. There is evidence
+that traits predict competitive outcomes
+\citep[\citet{lasky_trait-mediated_2014},
+\citet{uriarte_trait_2010}]{kunstler_competitive_2012}. Thus an
+extenstion of the model would allow \(lambda\) values to be a function
+of the traits of competing species.
+
+Lastly, the \texttt{forestecology} current uses the \texttt{blockCV}
+package behind the scenes to create the spatial blocks acting as folds
+for our spatial cross-validation algorithm detailed in Sections
+\ref{spatial-information} and \ref{spatial-cross-validation}
+\citep{valavi_blockcv_2019}. This back-end functionality could be
+substituted with the \texttt{spatialsample} package for spatial
+resampling infrastructure; a \texttt{tidymodels} package under active
+development as of 2021
+\citep[\citet{tidymodels_package}]{spatialsample_package}.
 

### 2021-05-13 paper/paper.Rmd

--- 2021-05-12 paper/paper.Rmd
@@
 
-# Conclusion
+# Conclusion and future work
 
 The `forestecology` package provides an accessible way to fit and test models of neighborhood competition.
-While the package is designed with ForestGEO plot data in mind, we envision that it can be modified to work on any single large, mapped forest plot in which at least two measurements of each individual have been taken.
-Furthermore, we also envision the package being used more generally for inventory plot-structure with many spatially separated plots, e.g.
-the US Forest Service Forest Inventory and Analysis plots [@smith_forest_2002].
-In future versions of `forestecology` we also hope to include models that account for tree mortality in addition to tree growth.
 The package follows the `tidy` data design principles, leverages the `sf` package for spatial data, and S3 open-oriented model implementation structure.
 We hope that the package will increase the use of neighborhood competition models to better understand what structures plant competition.
+
+While the package is designed with ForestGEO plot data in mind, we envision that it can be modified to work on any single large, mapped forest plot in which at least two measurements of each individual have been taken.
+Furthermore, we hope that future versions of the package will be flexible to other plot layouts, for example inventory plot-structure with many spatially separated plots like the US Forest Service Forest Inventory and Analysis plots [@smith_forest_2002].
+
+We also hope to extend the `forestecology` package's functionality to account for a larger variety of models for tree growth.
+One clear future direction would be to allow competition based on species trait values rather than species identity.
+There is evidence that traits predict competitive outcomes [@kunstler_competitive_2012, @lasky_trait-mediated_2014, @uriarte_trait_2010].
+Thus an extenstion of the model would allow $lambda$ values to be a function of the traits of competing species.
+
+Lastly, the `forestecology` current uses the `blockCV` package behind the scenes to create the spatial blocks acting as folds for our spatial cross-validation algorithm detailed in Sections \ref{spatial-information} and \ref{spatial-cross-validation} [@valavi_blockcv_2019].
+This back-end functionality could be substituted with the `spatialsample` package for spatial resampling infrastructure; a `tidymodels` package under active development as of 2021 [@spatialsample_package, @tidymodels_package].
 

### 2021-05-13 paper/paper.Rmd
named all chunks

@@
 
-```{r}
+```{r add-buffer}
 # Define competitive distance range
@@
 
-```{r}
+```{r manually-define-folds}
 # Manually define spatial blocks to act as folds
@@
 
-```{r, eval = FALSE, echo = FALSE}
+```{r internal, eval = FALSE, echo = FALSE}
 # Internal: after removing trees in buffer and have NA for growth, there is one
@@
 
-```{r, scbi-focal-vs-comp-3}
+```{r demonstrate-unnest, scbi-focal-vs-comp-3}
 focal_vs_comp_scbi %>% 
@@
 
-```{r}
+```{r show-comp-bayes}
 comp_bayes_lm_scbi
@@
 ```
-```{r}
+```{r show-focal-vs-comp}
 focal_vs_comp_scbi %>% 
@@
 ```
-```{r}
+```{r show-rmse-shuffle}
 model_rmse_shuffle <- focal_vs_comp_scbi %>%
@@
 ```
-```{r}
+```{r show-rmse-cv}
 model_rmse_cv <- focal_vs_comp_scbi %>%
@@
 
-```{r, eval = FALSE}
+```{r appendix-code, eval = FALSE, cache = FALSE}
 library(tidyverse)

### 2021-05-17 paper/paper.tex
tweak

@@
 \CommentTok{# Manually construct study region boundary}
-\NormalTok{study_region_scbi <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}
-  \DataTypeTok{x =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{),}
-  \DataTypeTok{y =} \KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{300}\NormalTok{)}
-\NormalTok{) }\OperatorTok{%>%}
+\NormalTok{study_region_scbi <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}\DataTypeTok{x =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{), }\DataTypeTok{y =} \KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{, }
+    \DecValTok{600}\NormalTok{, }\DecValTok{300}\NormalTok{)) }\OperatorTok{%>%}
 \StringTok{    }\KeywordTok{sf_polygon}\NormalTok{()}
@@
 associate each tree in \texttt{growth\_scbi} to the correct
-\texttt{foldID} \citep{valavi_blockcv_2019}.\footnote{In the Supporting
-  Information we present an example where the folds themselves are
-  created automatically, as opposed to manually as in the example.} This
-\texttt{foldID} variable will be used in Section
-\ref{spatial-cross-validation}.
+\texttt{foldID} \citep{valavi_blockcv_2019}. This \texttt{foldID}
+variable will be used in Section \ref{spatial-cross-validation}.
 
@@
 
-\NormalTok{blocks_scbi <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}
-  \KeywordTok{sf_polygon}\NormalTok{(fold1), }\KeywordTok{sf_polygon}\NormalTok{(fold2), }\KeywordTok{sf_polygon}\NormalTok{(fold3), }
-  \KeywordTok{sf_polygon}\NormalTok{(fold4)}
-\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\NormalTok{n_fold) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
+\NormalTok{blocks_scbi <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}\KeywordTok{sf_polygon}\NormalTok{(fold1), }\KeywordTok{sf_polygon}\NormalTok{(fold2), }\KeywordTok{sf_polygon}\NormalTok{(fold3), }
+    \KeywordTok{sf_polygon}\NormalTok{(fold4)) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\NormalTok{n_fold) }\OperatorTok{%>%}
+\StringTok{        }\KeywordTok{factor}\NormalTok{())}
 
 \CommentTok{# Associate each observation to a fold}
-\NormalTok{spatial_block_scbi <-}\StringTok{ }
-\StringTok{  }\KeywordTok{spatialBlock}\NormalTok{(}\DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =}\NormalTok{ n_fold, }
-               \DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }
-               \DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}\NormalTok{)}
+\NormalTok{spatial_block_scbi <-}\StringTok{ }\KeywordTok{spatialBlock}\NormalTok{(}\DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =}\NormalTok{ n_fold, }
+    \DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}\NormalTok{)}
 
 \NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ spatial_block_scbi}\OperatorTok{$}\NormalTok{foldID }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ spatial_block_scbi}\OperatorTok{$}\NormalTok{foldID }\OperatorTok{%>%}
+\StringTok{        }\KeywordTok{factor}\NormalTok{())}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{filter}\NormalTok{(focal_ID }\OperatorTok{==}\StringTok{ }\DecValTok{4}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{select}\NormalTok{(focal_ID, dbh, comp) }\OperatorTok{%>%}\StringTok{ }
+\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{filter}\NormalTok{(focal_ID }\OperatorTok{==}\StringTok{ }\DecValTok{4}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{select}\NormalTok{(focal_ID, dbh, comp) }\OperatorTok{%>%}
 \StringTok{    }\KeywordTok{unnest}\NormalTok{(}\DataTypeTok{cols =} \StringTok{"comp"}\NormalTok{)}
@@
 \begin{Highlighting}[]
-\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}\StringTok{ }
+\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
 \StringTok{    }\KeywordTok{select}\NormalTok{(focal_ID, focal_sp, dbh, growth, growth_hat)}
@@
 \citet{uriarte_trait_2010}]{kunstler_competitive_2012}. Thus an
-extenstion of the model would allow \(lambda\) values to be a function
-of the traits of competing species.
+extension of the model would allow \(\lambda\) values from Equation
+\ref{eq:model} to be a function of the traits of competing species.
 
@@
 
+\hypertarget{appendix}{%
+\section{Appendix}\label{appendix}}
+
+This code replicates Figure \ref{fig:scbi-simulation}: A comparison of
+root mean squared error of models for standard, permuted, and spatial
+cross- validated error estimates.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{library}\NormalTok{(tidyverse)}
+\KeywordTok{library}\NormalTok{(lubridate)}
+\KeywordTok{library}\NormalTok{(here)}
+\KeywordTok{library}\NormalTok{(sf)}
+\KeywordTok{library}\NormalTok{(viridis)}
+\KeywordTok{library}\NormalTok{(forestecology)}
+\KeywordTok{library}\NormalTok{(blockCV)}
+\KeywordTok{library}\NormalTok{(tictoc)}
+
+\CommentTok{# Compute growth of trees based on census data -------------------------}
+\NormalTok{census_}\DecValTok{2013}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{here}\NormalTok{(}\StringTok{"paper/scbi.stem2.csv"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{read_csv}\NormalTok{() }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date), }\DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+
+\NormalTok{census_}\DecValTok{2018}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{here}\NormalTok{(}\StringTok{"paper/scbi.stem3.csv"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{read_csv}\NormalTok{() }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date), }\DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+
+\NormalTok{growth_scbi <-}\StringTok{ }\KeywordTok{compute_growth}\NormalTok{(}\DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi, }\DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)), }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{)}
+
+\CommentTok{# Add spatial information ----------------------------------------------}
+\CommentTok{# Define buffer region using competitive distance range}
+\NormalTok{comp_dist <-}\StringTok{ }\FloatTok{7.5}
+
+\NormalTok{study_region_scbi <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}\DataTypeTok{x =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{), }\DataTypeTok{y =} \KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{, }
+    \DecValTok{600}\NormalTok{, }\DecValTok{300}\NormalTok{)) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{sf_polygon}\NormalTok{()}
+
+\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{size =}\NormalTok{ comp_dist, }\DataTypeTok{region =}\NormalTok{ study_region_scbi)}
+
+\CommentTok{# Manually define spatial blocks to act as folds}
+\NormalTok{fold1 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{450}\NormalTok{))}
+\NormalTok{fold2 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{450}\NormalTok{))}
+\NormalTok{fold3 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{600}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+\NormalTok{fold4 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+\NormalTok{n_fold <-}\StringTok{ }\DecValTok{4}
+
+\NormalTok{blocks_scbi <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}\KeywordTok{sf_polygon}\NormalTok{(fold1), }\KeywordTok{sf_polygon}\NormalTok{(fold2), }\KeywordTok{sf_polygon}\NormalTok{(fold3), }
+    \KeywordTok{sf_polygon}\NormalTok{(fold4)) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\NormalTok{n_fold) }\OperatorTok{%>%}
+\StringTok{        }\KeywordTok{factor}\NormalTok{())}
+
+\CommentTok{# Associate each observation to a fold}
+\NormalTok{SpatialBlock_scbi <-}\StringTok{ }\KeywordTok{spatialBlock}\NormalTok{(}\DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =}\NormalTok{ n_fold, }
+    \DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}\NormalTok{)}
+
+\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ SpatialBlock_scbi}\OperatorTok{$}\NormalTok{foldID }\OperatorTok{%>%}
+\StringTok{        }\KeywordTok{factor}\NormalTok{())}
+
+\CommentTok{# Compute focal versus competitor tree information ---------------------}
+\NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{)}
+
+\CommentTok{# Fit model and make predictions ---------------------------------------}
+\CommentTok{# Number of permutation shuffles:}
+\NormalTok{num_shuffle <-}\StringTok{ }\DecValTok{49}
+
+\CommentTok{# Save results here}
+\NormalTok{run_time <-}\StringTok{ }\DecValTok{0}
+\NormalTok{observed_RMSE <-}\StringTok{ }\DecValTok{0}
+\NormalTok{observed_RMSE_CV <-}\StringTok{ }\DecValTok{0}
+\NormalTok{shuffle_RMSE <-}\StringTok{ }\KeywordTok{vector}\NormalTok{(}\StringTok{"list"}\NormalTok{, }\DecValTok{1}\NormalTok{)}
+\NormalTok{shuffle_RMSE_CV <-}\StringTok{ }\KeywordTok{vector}\NormalTok{(}\StringTok{"list"}\NormalTok{, }\DecValTok{1}\NormalTok{)}
+\NormalTok{filename <-}\StringTok{ }\KeywordTok{here}\NormalTok{(}\StringTok{"paper/simulation_results/"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{str_c}\NormalTok{(}\StringTok{"2021-03-03_scbi_"}\NormalTok{, num_shuffle, }\StringTok{"_shuffles"}\NormalTok{)}
+
+\CommentTok{# Run all simulations 0. Setup simulation for this species type ----}
+\CommentTok{# Start clock}
+\KeywordTok{tic}\NormalTok{()}
+
+\CommentTok{# 1. Compute observed test statistic: RMSE with no cross-validation ----}
+\CommentTok{# Fit model (compute posterior parameters)}
+\NormalTok{comp_bayes_lm_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{comp_bayes_lm}\NormalTok{(}\DataTypeTok{prior_param =} \OtherTok{NULL}\NormalTok{, }\DataTypeTok{run_shuffle =} \OtherTok{FALSE}\NormalTok{)}
+
+\CommentTok{# Make predictions and compute RMSE}
+\NormalTok{observed_RMSE <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{growth_hat =} \KeywordTok{predict}\NormalTok{(comp_bayes_lm_scbi, focal_vs_comp_scbi)) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{pull}\NormalTok{(.estimate)}
+
+\CommentTok{# 2. Compute observed test statistic: RMSE with cross-validation -------}
+\NormalTok{observed_RMSE_CV <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{pull}\NormalTok{(.estimate)}
+
+\CommentTok{# 3. Permutation distribution: RMSE with no cross-validation -----------}
+\CommentTok{# Compute num_shuffle permutation test statistics}
+\NormalTok{shuffle_RMSE <-}\StringTok{ }\KeywordTok{numeric}\NormalTok{(}\DataTypeTok{length =}\NormalTok{ num_shuffle)}
+\ControlFlowTok{for}\NormalTok{ (j }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\NormalTok{num_shuffle) \{}
+    \CommentTok{# Fit model (compute posterior parameters) with shuffling}
+\NormalTok{    comp_bayes_lm_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
+\StringTok{        }\KeywordTok{comp_bayes_lm}\NormalTok{(}\DataTypeTok{prior_param =} \OtherTok{NULL}\NormalTok{, }\DataTypeTok{run_shuffle =} \OtherTok{TRUE}\NormalTok{)}
+
+    \CommentTok{# Make predictions and compute RMSE}
+\NormalTok{    shuffle_RMSE[j] <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
+\StringTok{        }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{growth_hat =} \KeywordTok{predict}\NormalTok{(comp_bayes_lm_scbi, focal_vs_comp_scbi)) }\OperatorTok{%>%}
+\StringTok{        }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
+\StringTok{        }\KeywordTok{pull}\NormalTok{(.estimate)}
+\NormalTok{\}}
+
+\CommentTok{# 4. Permutation distribution: RMSE with cross-validation --------------}
+\CommentTok{# Compute num_shuffle permutation test statistics}
+\NormalTok{shuffle_RMSE_CV <-}\StringTok{ }\KeywordTok{numeric}\NormalTok{(}\DataTypeTok{length =}\NormalTok{ num_shuffle)}
+
+\CommentTok{# Compute num_shuffle permutation test statistics}
+\ControlFlowTok{for}\NormalTok{ (j }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\NormalTok{num_shuffle) \{}
+    \CommentTok{# Compute and save RMSE}
+\NormalTok{    shuffle_RMSE_CV[j] <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
+\StringTok{        }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{run_shuffle =} \OtherTok{TRUE}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{        }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
+\StringTok{        }\KeywordTok{pull}\NormalTok{(.estimate)}
+
+    \CommentTok{# Status update}
+    \KeywordTok{str_c}\NormalTok{(}\StringTok{"Shuffle with permutation "}\NormalTok{, j, }\StringTok{" at "}\NormalTok{, }\KeywordTok{Sys.time}\NormalTok{()) }\OperatorTok{%>%}
+\StringTok{        }\KeywordTok{print}\NormalTok{()}
+\NormalTok{\}}
+
+\CommentTok{# 5. Save results ----}
+\NormalTok{clock <-}\StringTok{ }\KeywordTok{toc}\NormalTok{(}\DataTypeTok{quiet =} \OtherTok{TRUE}\NormalTok{)}
+\NormalTok{run_time <-}\StringTok{ }\NormalTok{clock}\OperatorTok{$}\NormalTok{toc }\OperatorTok{-}\StringTok{ }\NormalTok{clock}\OperatorTok{$}\NormalTok{tic}
+
+\NormalTok{model_comp_tbl <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}\DataTypeTok{run_time =}\NormalTok{ run_time, }\DataTypeTok{observed_RMSE =}\NormalTok{ observed_RMSE, }
+    \DataTypeTok{observed_RMSE_CV =}\NormalTok{ observed_RMSE_CV, }\DataTypeTok{shuffle_RMSE =}\NormalTok{ shuffle_RMSE, }\DataTypeTok{shuffle_RMSE_CV =}\NormalTok{ shuffle_RMSE_CV, }
+\NormalTok{    )}
+\KeywordTok{save}\NormalTok{(model_comp_tbl, }\DataTypeTok{file =}\NormalTok{ filename }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{str_c}\NormalTok{(}\StringTok{".RData"}\NormalTok{))}
+
+\CommentTok{# Visualize results ----------------------------------------------------}
+\NormalTok{model_comp <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(model_comp_tbl }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{select}\NormalTok{(run_time, }\DataTypeTok{observed =}\NormalTok{ observed_RMSE, }\DataTypeTok{shuffle =}\NormalTok{ shuffle_RMSE) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{CV =} \OtherTok{FALSE}\NormalTok{), model_comp_tbl }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{select}\NormalTok{(run_time, }\DataTypeTok{observed =}\NormalTok{ observed_RMSE_CV, }\DataTypeTok{shuffle =}\NormalTok{ shuffle_RMSE_CV) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{CV =} \OtherTok{TRUE}\NormalTok{)) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{gather}\NormalTok{(type, RMSE, }\OperatorTok{-}\KeywordTok{c}\NormalTok{(run_time, CV))}
+
+\NormalTok{model_comp_observed <-}\StringTok{ }\NormalTok{model_comp }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{filter}\NormalTok{(type }\OperatorTok{==}\StringTok{ "observed"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{unnest}\NormalTok{(}\DataTypeTok{cols =} \KeywordTok{c}\NormalTok{(RMSE))}
+\NormalTok{model_comp_shuffle <-}\StringTok{ }\NormalTok{model_comp }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{filter}\NormalTok{(type }\OperatorTok{==}\StringTok{ "shuffle"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{unnest}\NormalTok{(}\DataTypeTok{cols =} \KeywordTok{c}\NormalTok{(RMSE))}
+
+\NormalTok{cv_plot <-}\StringTok{ }\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}\StringTok{ }\KeywordTok{geom_vline}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ model_comp_observed, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{xintercept =}\NormalTok{ RMSE, }
+    \DataTypeTok{col =}\NormalTok{ CV), }\DataTypeTok{linetype =} \StringTok{"dashed"}\NormalTok{, }\DataTypeTok{show.legend =}\NormalTok{ F) }\OperatorTok{+}\StringTok{ }\KeywordTok{geom_histogram}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ model_comp_shuffle, }
+    \KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ RMSE, }\DataTypeTok{fill =}\NormalTok{ CV), }\DataTypeTok{bins =} \DecValTok{50}\NormalTok{) }\OperatorTok{+}\StringTok{ }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{fill =} \StringTok{"Cross-validated?"}\NormalTok{, }
+    \DataTypeTok{x =} \KeywordTok{expression}\NormalTok{(}\KeywordTok{paste}\NormalTok{(}\StringTok{"RMSE (cm "}\NormalTok{, y}\OperatorTok{^}\NormalTok{\{}
+        \DecValTok{-1}
+\NormalTok{    \}, }\StringTok{")"}\NormalTok{))) }\OperatorTok{+}\StringTok{ }\KeywordTok{scale_color_viridis}\NormalTok{(}\DataTypeTok{discrete =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{option =} \StringTok{"D"}\NormalTok{) }\OperatorTok{+}\StringTok{ }\KeywordTok{scale_fill_viridis}\NormalTok{(}\DataTypeTok{discrete =} \OtherTok{TRUE}\NormalTok{) }\OperatorTok{+}\StringTok{ }
+\StringTok{    }\KeywordTok{theme_light}\NormalTok{()}
+\NormalTok{cv_plot}
+
+\NormalTok{filename }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{str_c}\NormalTok{(}\StringTok{".pdf"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{ggsave}\NormalTok{(}\DataTypeTok{plot =}\NormalTok{ cv_plot, }\DataTypeTok{width =} \DecValTok{16}\OperatorTok{/}\DecValTok{2}\NormalTok{, }\DataTypeTok{height =} \DecValTok{9}\OperatorTok{/}\DecValTok{2}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
 \bibliographystyle{agsm}

### 2021-05-17 paper/paper_previous.tex

--- 2021-03-10 paper/forestecology.tex
@@
 
-\caption{Root mean squared error of models for standard, permuted, and spatial cross-validated error estimates.
-The dotted lines show non-permuted competitor identity, while the histogras so the RMSE for 49 permutations.
-The colors indicate whether cross validaton was used.}\label{fig:scbi-simulation}
+\caption{Comparison of root mean squared error of models for standard, permuted, and spatially cross-validated error estimates.
+The dotted lines show observed RMSE while the histograms show the null distribution of RMSE for 49 permutations under the null hypothesis of no competitor species identity effects.
+The colors indicate whether spatial cross-validation was used or not.}\label{fig:scbi-simulation}
 \end{figure}

### 2021-07-19 paper/paper_previous.Rmd

--- 2021-03-10 paper/forestecology.Rmd
@@
 
-```{r scbi-simulation, out.width="100%", fig.cap="Root mean squared error of models for standard, permuted, and spatial cross-validated error estimates.
-The dotted lines show non-permuted competitor identity, while the histogras so the RMSE for 49 permutations.
-The colors indicate whether cross validaton was used.", echo = FALSE}
-include_graphics("simulation_results/2021-03-03_scbi_49_shuffles.png")
+```{r scbi-simulation, out.width="100%", fig.cap="Comparison of root mean squared error of models for standard, permuted, and spatially cross-validated error estimates.
+The dotted lines show observed RMSE while the histograms show the null distribution of RMSE for 49 permutations under the null hypothesis of no competitor species identity effects.
+The colors indicate whether spatial cross-validation was used or not.", echo = FALSE}
+include_graphics("simulation_results/2021-03-03_scbi_49_shuffles.pdf")
 ```

### 2021-07-21 paper/paper.Rmd
tweak
Trying out multiple competitive distances. Wrote code but have not yet run it

@@
 Our model of interspecific competition relies on a spatial definition of who competitor trees are: all trees within a distance `comp_dist` of a focal tree.
-We set `comp_dist` to 7.5m, a value informed by other studies [@canham_neighborhood_2004, @uriarte_spatially_2004, @canham_neighborhood_2006].
+Here we set `comp_dist` to 7.5m, a value informed by other studies [@canham_neighborhood_2004, @uriarte_spatially_2004, @canham_neighborhood_2006], but the package could also be used to compare multiple distances and see which is best supported (see Appendix 2 \ref{mult-dist}).
 We use `comp_dist` and a manually constructed `sf` representation of the study region's boundary as inputs to `add_buffer_variable()` to add a `buffer` boolean variable to `growth_scbi`.
@@
 There is evidence that traits predict competitive outcomes [@kunstler_competitive_2012, @lasky_trait-mediated_2014, @uriarte_trait_2010].
-Thus an extenstion of the model would allow $lambda$ values to be a function of the traits of competing species.
+Thus an extension of the model would allow $\lambda$ values from Equation \ref{eq:model} to be a function of the traits of competing species.
 
@@
 
-# Appendix
+# Appendix 1: Replicate RMSE comparison
 
@@
 
-```{r appendix-code, eval = FALSE, cache = FALSE}
+```{r appendix1-code, eval = FALSE, cache = FALSE}
 library(tidyverse)
@@
 
+# Appendix 2: Compare different competitive distances {#mult-dist}
+
+Above we set the cut off distance for two stems to compete at 7.5 m. Our package can be used to find which distance is best supported by the data. Here we give an example.
+
+```{r appendix2-code}
+
+mult_dist_comp <- tibble(
+  dist = c(2.5, 5, 7.5, 10, 12.5 ),
+  rmse = 0
+)
+
+for (i in 1:length(mult_dist_comp$dist))
+{
+  comp_dist <- mult_dist_comp$dist[i]
+  growth_scbi <- growth_scbi %>%
+    add_buffer_variable(size = comp_dist, region = study_region_scbi)
+
+  focal_vs_comp_scbi <- growth_scbi %>%
+    create_focal_vs_comp(comp_dist, blocks = blocks_scbi, id = "stemID")
+  focal_vs_comp_scbi %>% 
+    select(focal_ID, focal_sp, geometry, growth, comp)
+
+  focal_vs_comp_scbi <- focal_vs_comp_scbi %>%
+    run_cv(comp_dist = mult_dists[i], blocks = blocks_scbi)
+
+  mult_dist_comp$rmse[i] <- focal_vs_comp_scbi %>%
+    rmse(truth = growth, estimate = growth_hat) %>%
+    pull(.estimate)
+}
+
+mult_dist_comp %>%
+  ggplot(aes(dist,rmse)) +
+  geom_point() +
+  xlab('Distance (m)') +
+  ylab('RMSE (cm)')
+
+```
+
 # References

### 2021-07-21 paper/paper.tex
Trying out multiple competitive distances. Wrote code but have not yet run it

--- 2021-05-13 paper/paper.tex
@@
 who competitor trees are: all trees within a distance
-\texttt{comp\_dist} of a focal tree. We set \texttt{comp\_dist} to 7.5m,
-a value informed by other studies \citep[\citet{uriarte_spatially_2004},
-\citet{canham_neighborhood_2006}]{canham_neighborhood_2004}. We use
+\texttt{comp\_dist} of a focal tree. Here we set \texttt{comp\_dist} to
+7.5m, a value informed by other studies
+\citep[\citet{uriarte_spatially_2004},
+\citet{canham_neighborhood_2006}]{canham_neighborhood_2004}, but the
+package could also be used to compare multiple distances and see which
+is best supported (see Appendix 2 \ref{mult-dist}). We use
 \texttt{comp\_dist} and a manually constructed \texttt{sf}
@@
 
+\hypertarget{appendix-1-replicate-rmse-comparison}{%
+\section{Appendix 1: Replicate RMSE
+comparison}\label{appendix-1-replicate-rmse-comparison}}
+
+This code replicates Figure \ref{fig:scbi-simulation}: A comparison of
+root mean squared error of models for standard, permuted, and spatial
+cross- validated error estimates.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{library}\NormalTok{(tidyverse)}
+\KeywordTok{library}\NormalTok{(lubridate)}
+\KeywordTok{library}\NormalTok{(here)}
+\KeywordTok{library}\NormalTok{(sf)}
+\KeywordTok{library}\NormalTok{(viridis)}
+\KeywordTok{library}\NormalTok{(forestecology)}
+\KeywordTok{library}\NormalTok{(blockCV)}
+\KeywordTok{library}\NormalTok{(tictoc)}
+
+\CommentTok{# Compute growth of trees based on census data -------------------------}
+\NormalTok{census_}\DecValTok{2013}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{here}\NormalTok{(}\StringTok{"paper/scbi.stem2.csv"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{read_csv}\NormalTok{() }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
+    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date),}
+    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}
+\NormalTok{  ) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+
+\NormalTok{census_}\DecValTok{2018}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{here}\NormalTok{(}\StringTok{"paper/scbi.stem3.csv"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{read_csv}\NormalTok{() }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
+    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date),}
+    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}
+\NormalTok{  ) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+
+\NormalTok{growth_scbi <-}
+\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(}
+    \DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi,}
+    \DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)),}
+    \DataTypeTok{id =} \StringTok{"stemID"}
+\NormalTok{  )}
+
+\CommentTok{# Add spatial information ----------------------------------------------}
+\CommentTok{# Define buffer region using competitive distance range}
+\NormalTok{comp_dist <-}\StringTok{ }\FloatTok{7.5}
+
+\NormalTok{study_region_scbi <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}
+  \DataTypeTok{x =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{),}
+  \DataTypeTok{y =} \KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{300}\NormalTok{)}
+\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{sf_polygon}\NormalTok{()}
+
+\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{size =}\NormalTok{ comp_dist, }\DataTypeTok{region =}\NormalTok{ study_region_scbi)}
+
+\CommentTok{# Manually define spatial blocks to act as folds}
+\NormalTok{fold1 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{450}\NormalTok{))}
+\NormalTok{fold2 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{450}\NormalTok{))}
+\NormalTok{fold3 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{600}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+\NormalTok{fold4 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+\NormalTok{n_fold <-}\StringTok{ }\DecValTok{4}
+
+\NormalTok{blocks_scbi <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}
+  \KeywordTok{sf_polygon}\NormalTok{(fold1), }\KeywordTok{sf_polygon}\NormalTok{(fold2), }
+  \KeywordTok{sf_polygon}\NormalTok{(fold3), }\KeywordTok{sf_polygon}\NormalTok{(fold4)}
+\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\NormalTok{n_fold) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
+
+\CommentTok{# Associate each observation to a fold}
+\NormalTok{SpatialBlock_scbi <-}\StringTok{ }\KeywordTok{spatialBlock}\NormalTok{(}
+  \DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =}\NormalTok{ n_fold, }\DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{,}
+  \DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}
+\NormalTok{)}
+
+\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ SpatialBlock_scbi}\OperatorTok{$}\NormalTok{foldID }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
+
+\CommentTok{# Compute focal versus competitor tree information ---------------------}
+\NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{)}
+
+\CommentTok{# Fit model and make predictions ---------------------------------------}
+\CommentTok{# Number of permutation shuffles:}
+\NormalTok{num_shuffle <-}\StringTok{ }\DecValTok{49}
+
+\CommentTok{# Save results here}
+\NormalTok{run_time <-}\StringTok{ }\DecValTok{0}
+\NormalTok{observed_RMSE <-}\StringTok{ }\DecValTok{0}
+\NormalTok{observed_RMSE_CV <-}\StringTok{ }\DecValTok{0}
+\NormalTok{shuffle_RMSE <-}\StringTok{ }\KeywordTok{vector}\NormalTok{(}\StringTok{"list"}\NormalTok{, }\DecValTok{1}\NormalTok{)}
+\NormalTok{shuffle_RMSE_CV <-}\StringTok{ }\KeywordTok{vector}\NormalTok{(}\StringTok{"list"}\NormalTok{, }\DecValTok{1}\NormalTok{)}
+\NormalTok{filename <-}\StringTok{ }\KeywordTok{here}\NormalTok{(}\StringTok{"paper/simulation_results/"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{str_c}\NormalTok{(}\StringTok{"2021-03-03_scbi_"}\NormalTok{, num_shuffle, }\StringTok{"_shuffles"}\NormalTok{)}
+
+\CommentTok{# Run all simulations}
+\CommentTok{# 0. Setup simulation for this species type ----}
+\CommentTok{# Start clock}
+\KeywordTok{tic}\NormalTok{()}
+
+\CommentTok{# 1. Compute observed test statistic: RMSE with no cross-validation ----}
+\CommentTok{# Fit model (compute posterior parameters)}
+\NormalTok{comp_bayes_lm_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{comp_bayes_lm}\NormalTok{(}\DataTypeTok{prior_param =} \OtherTok{NULL}\NormalTok{, }\DataTypeTok{run_shuffle =} \OtherTok{FALSE}\NormalTok{)}
+
+\CommentTok{# Make predictions and compute RMSE}
+\NormalTok{observed_RMSE <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{growth_hat =} 
+           \KeywordTok{predict}\NormalTok{(comp_bayes_lm_scbi, focal_vs_comp_scbi)) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
+
+\CommentTok{# 2. Compute observed test statistic: RMSE with cross-validation -------}
+\NormalTok{observed_RMSE_CV <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
+
+\CommentTok{# 3. Permutation distribution: RMSE with no cross-validation -----------}
+\CommentTok{# Compute num_shuffle permutation test statistics}
+\NormalTok{shuffle_RMSE <-}\StringTok{ }\KeywordTok{numeric}\NormalTok{(}\DataTypeTok{length =}\NormalTok{ num_shuffle)}
+\ControlFlowTok{for}\NormalTok{(j }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\NormalTok{num_shuffle)\{}
+  \CommentTok{# Fit model (compute posterior parameters) with shuffling}
+\NormalTok{  comp_bayes_lm_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{comp_bayes_lm}\NormalTok{(}\DataTypeTok{prior_param =} \OtherTok{NULL}\NormalTok{, }\DataTypeTok{run_shuffle =} \OtherTok{TRUE}\NormalTok{)}
+
+  \CommentTok{# Make predictions and compute RMSE}
+\NormalTok{  shuffle_RMSE[j] <-}\StringTok{  }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{growth_hat =} 
+             \KeywordTok{predict}\NormalTok{(comp_bayes_lm_scbi, focal_vs_comp_scbi)) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{pull}\NormalTok{(.estimate)}
+\NormalTok{\}}
+
+\CommentTok{# 4. Permutation distribution: RMSE with cross-validation --------------}
+\CommentTok{# Compute num_shuffle permutation test statistics}
+\NormalTok{shuffle_RMSE_CV <-}\StringTok{ }\KeywordTok{numeric}\NormalTok{(}\DataTypeTok{length =}\NormalTok{ num_shuffle)}
+
+\CommentTok{# Compute num_shuffle permutation test statistics}
+\ControlFlowTok{for}\NormalTok{(j }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\NormalTok{num_shuffle)\{}
+  \CommentTok{# Compute and save RMSE}
+\NormalTok{  shuffle_RMSE_CV[j] <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }
+           \DataTypeTok{run_shuffle =} \OtherTok{TRUE}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{pull}\NormalTok{(.estimate)}
+
+  \CommentTok{# Status update}
+  \KeywordTok{str_c}\NormalTok{(}\StringTok{"Shuffle with permutation "}\NormalTok{, j, }\StringTok{" at "}\NormalTok{, }\KeywordTok{Sys.time}\NormalTok{()) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{print}\NormalTok{()}
+\NormalTok{\}}
+
+\CommentTok{# 5. Save results ----}
+\NormalTok{clock <-}\StringTok{ }\KeywordTok{toc}\NormalTok{(}\DataTypeTok{quiet =} \OtherTok{TRUE}\NormalTok{)}
+\NormalTok{run_time <-}\StringTok{ }\NormalTok{clock}\OperatorTok{$}\NormalTok{toc }\OperatorTok{-}\StringTok{ }\NormalTok{clock}\OperatorTok{$}\NormalTok{tic}
+
+\NormalTok{model_comp_tbl <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}
+  \DataTypeTok{run_time =}\NormalTok{ run_time,}
+  \DataTypeTok{observed_RMSE =}\NormalTok{ observed_RMSE,}
+  \DataTypeTok{observed_RMSE_CV =}\NormalTok{ observed_RMSE_CV,}
+  \DataTypeTok{shuffle_RMSE =}\NormalTok{ shuffle_RMSE,}
+  \DataTypeTok{shuffle_RMSE_CV =}\NormalTok{ shuffle_RMSE_CV,}
+\NormalTok{)}
+\KeywordTok{save}\NormalTok{(model_comp_tbl, }\DataTypeTok{file =}\NormalTok{ filename }\OperatorTok{%>%}\StringTok{ }\KeywordTok{str_c}\NormalTok{(}\StringTok{".RData"}\NormalTok{))}
+
+\CommentTok{# Visualize results ----------------------------------------------------}
+\NormalTok{model_comp <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}
+\NormalTok{  model_comp_tbl }\OperatorTok{%>%}\StringTok{ }
+\StringTok{    }\KeywordTok{select}\NormalTok{(run_time, }\DataTypeTok{observed =}\NormalTok{ observed_RMSE, }
+           \DataTypeTok{shuffle =}\NormalTok{ shuffle_RMSE) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{CV =} \OtherTok{FALSE}\NormalTok{),}
+\NormalTok{  model_comp_tbl }\OperatorTok{%>%}\StringTok{ }
+\StringTok{    }\KeywordTok{select}\NormalTok{(run_time, }\DataTypeTok{observed =}\NormalTok{ observed_RMSE_CV, }
+           \DataTypeTok{shuffle =}\NormalTok{ shuffle_RMSE_CV) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{CV =} \OtherTok{TRUE}\NormalTok{)}
+\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{gather}\NormalTok{(type, RMSE, }\OperatorTok{-}\KeywordTok{c}\NormalTok{(run_time, CV))}
+
+\NormalTok{model_comp_observed <-}\StringTok{ }\NormalTok{model_comp }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{filter}\NormalTok{(type }\OperatorTok{==}\StringTok{ "observed"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{unnest}\NormalTok{(}\DataTypeTok{cols =} \KeywordTok{c}\NormalTok{(RMSE))}
+\NormalTok{model_comp_shuffle <-}\StringTok{ }\NormalTok{model_comp }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{filter}\NormalTok{(type }\OperatorTok{==}\StringTok{ "shuffle"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{unnest}\NormalTok{(}\DataTypeTok{cols =} \KeywordTok{c}\NormalTok{(RMSE))}
+
+\NormalTok{cv_plot <-}\StringTok{ }\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
+\StringTok{  }\KeywordTok{geom_vline}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ model_comp_observed, }
+             \KeywordTok{aes}\NormalTok{(}\DataTypeTok{xintercept =}\NormalTok{ RMSE, }\DataTypeTok{col =}\NormalTok{ CV), }
+             \DataTypeTok{linetype =} \StringTok{"dashed"}\NormalTok{, }\DataTypeTok{show.legend =}\NormalTok{ F) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{geom_histogram}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ model_comp_shuffle, }
+                 \KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ RMSE, }\DataTypeTok{fill =}\NormalTok{ CV), }\DataTypeTok{bins =} \DecValTok{50}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{labs}\NormalTok{(}
+    \DataTypeTok{fill =} \StringTok{"Cross-validated?"}\NormalTok{,}
+    \DataTypeTok{x =} \KeywordTok{expression}\NormalTok{(}\KeywordTok{paste}\NormalTok{(}\StringTok{"RMSE (cm "}\NormalTok{, y}\OperatorTok{^}\NormalTok{\{}\OperatorTok{-}\DecValTok{1}\NormalTok{\}, }\StringTok{")"}\NormalTok{))}
+\NormalTok{  ) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{scale_color_viridis}\NormalTok{(}\DataTypeTok{discrete =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{option =} \StringTok{"D"}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{scale_fill_viridis}\NormalTok{(}\DataTypeTok{discrete =} \OtherTok{TRUE}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{theme_light}\NormalTok{()}
+\NormalTok{cv_plot}
+
+\NormalTok{filename }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{str_c}\NormalTok{(}\StringTok{".pdf"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{ggsave}\NormalTok{(}\DataTypeTok{plot =}\NormalTok{ cv_plot, }\DataTypeTok{width =} \DecValTok{16}\OperatorTok{/}\DecValTok{2}\NormalTok{, }\DataTypeTok{height =} \DecValTok{9}\OperatorTok{/}\DecValTok{2}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\hypertarget{mult-dist}{%
+\section{Appendix 2: Compare different competitive
+distances}\label{mult-dist}}
+
 \bibliographystyle{agsm}

### 2021-07-21 paper/paper.Rmd
runs and gives result, but looks funky!

@@
 ```{r appendix2-code}
-
 mult_dist_comp <- tibble(
-  dist = c(2.5, 5, 7.5, 10, 12.5 ),
+  dist = c(2.5, 5, 7.5, 10, 12.5),
   rmse = 0
@@
   focal_vs_comp_scbi <- growth_scbi %>%
-    create_focal_vs_comp(comp_dist, blocks = blocks_scbi, id = "stemID")
+    create_focal_vs_comp(comp_dist = comp_dist, blocks = blocks_scbi, id = "stemID")
   focal_vs_comp_scbi %>% 
@@
   focal_vs_comp_scbi <- focal_vs_comp_scbi %>%
-    run_cv(comp_dist = mult_dists[i], blocks = blocks_scbi)
+    run_cv(comp_dist = comp_dist, blocks = blocks_scbi)
 

### 2021-07-21 paper/paper.tex
runs and gives result, but looks funky!

@@
 
+Above we set the cut off distance for two stems to compete at 7.5 m. Our
+package can be used to find which distance is best supported by the
+data. Here we give an example.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{mult_dist_comp <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}
+  \DataTypeTok{dist =} \KeywordTok{c}\NormalTok{(}\FloatTok{2.5}\NormalTok{, }\DecValTok{5}\NormalTok{, }\FloatTok{7.5}\NormalTok{, }\DecValTok{10}\NormalTok{, }\FloatTok{12.5}\NormalTok{),}
+  \DataTypeTok{rmse =} \DecValTok{0}
+\NormalTok{)}
+
+\ControlFlowTok{for}\NormalTok{ (i }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(mult_dist_comp}\OperatorTok{$}\NormalTok{dist))}
+\NormalTok{\{}
+\NormalTok{  comp_dist <-}\StringTok{ }\NormalTok{mult_dist_comp}\OperatorTok{$}\NormalTok{dist[i]}
+\NormalTok{  growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{size =}\NormalTok{ comp_dist, }\DataTypeTok{region =}\NormalTok{ study_region_scbi)}
+
+\NormalTok{  focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{)}
+\NormalTok{  focal_vs_comp_scbi }\OperatorTok{%>%}\StringTok{ }
+\StringTok{    }\KeywordTok{select}\NormalTok{(focal_ID, focal_sp, geometry, growth, comp)}
+
+\NormalTok{  focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi)}
+
+\NormalTok{  mult_dist_comp}\OperatorTok{$}\NormalTok{rmse[i] <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{pull}\NormalTok{(.estimate)}
+\NormalTok{\}}
+
+\NormalTok{mult_dist_comp }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(dist,rmse)) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{geom_point}\NormalTok{() }\OperatorTok{+}
+\StringTok{  }\KeywordTok{xlab}\NormalTok{(}\StringTok{'Distance (m)'}\NormalTok{) }\OperatorTok{+}
+\StringTok{  }\KeywordTok{ylab}\NormalTok{(}\StringTok{'RMSE (cm)'}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{center}\includegraphics[width=1\linewidth]{Figures/appendix2-code-1} \end{center}
+
 \bibliographystyle{agsm}

### 2021-07-26 paper/paper.tex
compare multiple competitive distances

@@
 
-Above we set the cut off distance for two stems to compete at 7.5 m. Our
-package can be used to find which distance is best supported by the
-data. Here we give an example.
-
-\begin{Shaded}
-\begin{Highlighting}[]
+For all the above analyses we took the cut off distance
+(\texttt{comp\_dist}) for two stems to compete as 7.5 m. This distance
+has been estimated in previous neighborhood competition studies in
+forests \citep[\citet{uriarte_spatially_2004},
+\citet{canham_neighborhood_2006}]{canham_neighborhood_2004}. We used 7.5
+m in \citet{allen_permutation_2020} as an average of the values
+estimated in other studies. But our package can be used to find which
+distance is best supported by the data. Here we provide an example using
+another section of the SCBI plot to provide an additional example of the
+CV block layout. To speed computation we do not consider species
+differences in competitive effects and treat all species as the same.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+
+\NormalTok{census_}\DecValTok{2013}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem2.csv"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
+    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date),}
+    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}
+\NormalTok{  ) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{400}\NormalTok{, gy }\OperatorTok{<}\StringTok{ }\DecValTok{400}\NormalTok{)}
+
+\NormalTok{census_}\DecValTok{2018}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem3.csv"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
+    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date),}
+    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}
+\NormalTok{  ) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{400}\NormalTok{, gy }\OperatorTok{<}\StringTok{ }\DecValTok{400}\NormalTok{)}
+
+\NormalTok{growth_scbi <-}
+\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(}
+    \DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi,}
+    \DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)),}
+    \DataTypeTok{id =} \StringTok{"stemID"}
+\NormalTok{  ) }\OperatorTok{%>%}
+\StringTok{  }\CommentTok{# make all species the same, needs to be factor with at least two levels}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{factor}\NormalTok{(}\StringTok{'A'}\NormalTok{, }\DataTypeTok{levels =} \KeywordTok{c}\NormalTok{(}\StringTok{'A'}\NormalTok{, }\StringTok{'B'}\NormalTok{)) )}
+
+\NormalTok{study_region_scbi <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}
+  \DataTypeTok{x =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{),}
+  \DataTypeTok{y =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{0}\NormalTok{)}
+\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{sf_polygon}\NormalTok{()}
+
+\NormalTok{n_fold <-}\StringTok{ }\DecValTok{4}
+\NormalTok{fold1 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{0}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{200}\NormalTok{))}
+\NormalTok{fold2 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{200}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{200}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{200}\NormalTok{))}
+\NormalTok{fold3 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{0}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{200}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{400}\NormalTok{))}
+\NormalTok{fold4 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{200}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{200}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{200}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{400}\NormalTok{))}
+
+\NormalTok{blocks_scbi <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}
+  \KeywordTok{sf_polygon}\NormalTok{(fold1), }\KeywordTok{sf_polygon}\NormalTok{(fold2), }\KeywordTok{sf_polygon}\NormalTok{(fold3), }
+  \KeywordTok{sf_polygon}\NormalTok{(fold4)}
+\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\NormalTok{n_fold) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
+
+\CommentTok{# Associate each observation to a fold}
+\NormalTok{spatial_block_scbi <-}\StringTok{ }
+\StringTok{  }\KeywordTok{spatialBlock}\NormalTok{(}\DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =}\NormalTok{ n_fold, }
+               \DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }
+               \DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}\NormalTok{)}
+
+\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ spatial_block_scbi}\OperatorTok{$}\NormalTok{foldID }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
+
 \NormalTok{mult_dist_comp <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}
-  \DataTypeTok{dist =} \KeywordTok{c}\NormalTok{(}\FloatTok{2.5}\NormalTok{, }\DecValTok{5}\NormalTok{, }\FloatTok{7.5}\NormalTok{, }\DecValTok{10}\NormalTok{, }\FloatTok{12.5}\NormalTok{),}
+  \DataTypeTok{dist =} \KeywordTok{c}\NormalTok{(}\DecValTok{5}\NormalTok{ ,}\FloatTok{6.25}\NormalTok{,  }\FloatTok{7.5}\NormalTok{, }\FloatTok{8.75}\NormalTok{,  }\DecValTok{10}\NormalTok{),}
   \DataTypeTok{rmse =} \DecValTok{0}
@@
 \NormalTok{  comp_dist <-}\StringTok{ }\NormalTok{mult_dist_comp}\OperatorTok{$}\NormalTok{dist[i]}
+
 \NormalTok{  growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
@@
 \NormalTok{  focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{)}
-\NormalTok{  focal_vs_comp_scbi }\OperatorTok{%>%}\StringTok{ }
-\StringTok{    }\KeywordTok{select}\NormalTok{(focal_ID, focal_sp, geometry, growth, comp)}
-
-\NormalTok{  focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{) }\OperatorTok{%>%}
 \StringTok{    }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi)}
@@
 
+Here we see that a cut off distance of approximately 6 m minimizes RMSE.
+
 \bibliographystyle{agsm}

### 2021-08-02 paper/paper.Rmd

--- 2021-07-21 paper/paper.Rmd
@@
 
-Above we set the cut off distance for two stems to compete at 7.5 m. Our package can be used to find which distance is best supported by the data. Here we give an example.
+For all the above analyses we took the cut off distance (`comp_dist`) for two stems to compete as 7.5 m.
+This distance has been estimated in previous neighborhood competition studies in forests [@canham_neighborhood_2004, @uriarte_spatially_2004, @canham_neighborhood_2006].
+We used 7.5 m in @allen_permutation_2020 as an average of the values estimated in other studies.
+But our package can be used to find which distance is best supported by the data.
+Here we provide an example using another section of the SCBI plot to provide an additional example of the CV block layout.
+To speed computation we do not consider species differences in competitive effects and treat all species as the same.
 
@@
 
+census_2013_scbi <- read_csv("scbi.stem2.csv") %>%
+  select(stemID, sp, date = ExactDate, gx, gy, dbh, codes, status) %>%
+  mutate(
+    date = mdy(date),
+    dbh = as.numeric(dbh)/10
+  ) %>%
+  filter(gx < 400, gy < 400)
+
+census_2018_scbi <- read_csv("scbi.stem3.csv") %>%
+  select(stemID, sp, date = ExactDate, gx, gy, dbh, codes, status) %>%
+  mutate(
+    date = mdy(date),
+    dbh = as.numeric(dbh)/10
+  ) %>%
+  filter(gx < 400, gy < 400)
+
+growth_scbi <-
+  compute_growth(
+    census_1 = census_2013_scbi,
+    census_2 = census_2018_scbi %>% filter(!str_detect(codes, "R")),
+    id = "stemID"
+  ) %>%
+  # make all species the same, needs to be factor with at least two levels
+  mutate(sp = factor('A', levels = c('A', 'B')) )
+
+study_region_scbi <- tibble(
+  x = c(0, 400, 400, 0, 0),
+  y = c(0, 0, 400, 400, 0)
+) %>%
+  sf_polygon()
+
+n_fold <- 4
+fold1 <- cbind(c(0, 200, 200, 0), c(0, 0, 200, 200))
+fold2 <- cbind(c(200, 400, 400, 200), c(0, 0, 200, 200))
+fold3 <- cbind(c(0, 200, 200, 0), c(200, 200, 400, 400))
+fold4 <- cbind(c(200, 400, 400, 200), c(200, 200, 400, 400))
+
+blocks_scbi <- bind_rows(
+  sf_polygon(fold1), sf_polygon(fold2), sf_polygon(fold3), 
+  sf_polygon(fold4)
+) %>%
+  mutate(folds = c(1:n_fold) %>% factor())
+
+# Associate each observation to a fold
+spatial_block_scbi <- 
+  spatialBlock(speciesData = growth_scbi, k = n_fold, 
+               selection = "systematic", blocks = blocks_scbi, 
+               showBlocks = FALSE, verbose = FALSE)
+
+growth_scbi <- growth_scbi %>%
+  mutate(foldID = spatial_block_scbi$foldID %>% factor())
+
 mult_dist_comp <- tibble(
-  dist = c(2.5, 5, 7.5, 10, 12.5 ),
+  dist = c(5 ,6.25,  7.5, 8.75,  10),
   rmse = 0
@@
   comp_dist <- mult_dist_comp$dist[i]
+
   growth_scbi <- growth_scbi %>%
@@
   focal_vs_comp_scbi <- growth_scbi %>%
-    create_focal_vs_comp(comp_dist, blocks = blocks_scbi, id = "stemID")
-  focal_vs_comp_scbi %>% 
-    select(focal_ID, focal_sp, geometry, growth, comp)
-
-  focal_vs_comp_scbi <- focal_vs_comp_scbi %>%
-    run_cv(comp_dist = mult_dists[i], blocks = blocks_scbi)
+    create_focal_vs_comp(comp_dist = comp_dist, blocks = blocks_scbi, id = "stemID") %>%
+    run_cv(comp_dist = comp_dist, blocks = blocks_scbi)
 
@@
 
+Here we see that a cut off distance of approximately 6 m minimizes RMSE.
+
 # References

### 2021-08-02 paper/paper.Rmd
Started edits
Added text about sf package

@@
 Much like all `tidyverse` packages, `forestecology` has verb-named functions that can be modularly composed using the pipe `%>%` operator to sequentially complete all necessary analysis steps [@bache_pipe_2020].
-Furthermore, the inputs and outputs of most functions use the same "simple features for R" data structures from the `sf` package, a package for standardized and `tidyverse`-friendly wrangling and visualizing of spatial data [@pebesma_simple_2018].
 
-Currently the package only implements the Bayesian linear regression model detailed in Equation \ref{eq:model}.
+Furthermore, the inputs and outputs of most functions use the same "simple features for R" data structures for spatial data from the `sf` package [@pebesma_simple_2018].
+Previously `sp` package classes were commonly used for storing spatial data and interfacing with geospatial libraries [@bivand_sp_2013]; the `sf` package aims to improve on the `sp` package by:
+
+1. Using simple feature access as the base standard for representing and encoding spatial data, rather than shapefiles [@herring_2011]. 
+1. Leveraging improvements in external libraries for reading and writing spatial data (GDAL) and for geometrical operations (GEOS) [@warmerdam_geospatial_2008, @geos_2017].
+1. Integrating closely with the popular `tidyverse` suite of packages for data science [@wickham_welcome_2019].
+
+By using the `sf` package classes to represent spatial data rather than the `sp` package, the implementation and use of the `forestecology` package's spatial algorithms was greatly simplified.
+
+Lastly, currently the package only implements the Bayesian linear regression model detailed in Equation \ref{eq:model}.
 As we demonstrate in Section \ref{model-fit-predict} however, the fitting of this model is self-contained in a single function `comp_bayes_lm()` which returns an object of S3 class type `comp_bayes_lm`.
@@
 The `forestecology` package provides an accessible way to fit and test models of neighborhood competition.
-The package follows the `tidy` data design principles, leverages the `sf` package for spatial data, and S3 open-oriented model implementation structure.
+The package follows the `tidy` data design principles, leverages the `sf` package for spatial data, and S3 open-oriented model implementation structure [@pebesma_simple_2018].
 We hope that the package will increase the use of neighborhood competition models to better understand what structures plant competition.

### 2021-08-02 paper/paper.Rmd
code line breaks

--- 2021-08-02 paper/paper.Rmd
@@
 blocks_scbi <- bind_rows(
-  sf_polygon(fold1), sf_polygon(fold2), sf_polygon(fold3), 
-  sf_polygon(fold4)
+  sf_polygon(fold1), sf_polygon(fold2), 
+  sf_polygon(fold3), sf_polygon(fold4)
 ) %>%
@@
 spatial_block_scbi <- 
-  spatialBlock(speciesData = growth_scbi, k = n_fold, 
+  spatialBlock(
+    speciesData = growth_scbi, k = n_fold, 
     selection = "systematic", blocks = blocks_scbi, 
-               showBlocks = FALSE, verbose = FALSE)
+    showBlocks = FALSE, verbose = FALSE
+  )
 
@@
 focal_vs_comp_scbi <- focal_vs_comp_scbi %>%
-  mutate(growth_hat = predict(comp_bayes_lm_scbi, newdata = focal_vs_comp_scbi))
+  mutate(growth_hat = predict(comp_bayes_lm_scbi, 
+                              newdata = focal_vs_comp_scbi))
 ```
@@
   shuffle_RMSE[j] <-  focal_vs_comp_scbi %>%
-    mutate(growth_hat = 
-             predict(comp_bayes_lm_scbi, focal_vs_comp_scbi)) %>%
+    mutate(
+      growth_hat = predict(comp_bayes_lm_scbi, focal_vs_comp_scbi)
+    ) %>%
     rmse(truth = growth, estimate = growth_hat) %>%

### 2021-08-02 paper/paper.Rmd

@@
 \label{eq:model}
-y_{ij} = \beta_{0,j} + \beta_{\text{dbh},j} \cdot \text{dbh}_{ij} + \sum_{k=1}^{K} \lambda_{jk} \cdot \text{BA}_{ijk} + \epsilon_{ij}
+y_{ij} = \beta_{0,j} + \beta_{\text{dbh},j} \cdot \text{dbh}_{ij} + \sum_{k=1}^{K} \lambda_{jk} \cdot x^{\text{comp}}_{ijk} + \epsilon_{ij}
 \end{equation}
 
-where $\beta_{0,j}$ is the diameter-independent growth rate of species $j$; $\text{dbh}_{ij}$ is the DBH of the focal tree at the earlier census and $\beta_{\text{dbh},j}$ the slope of that species's diameter-growth relationship;  $\text{BA}_{ijk}$ is the sum of the basal area of all trees of competitor species $k$ and $\lambda_{jk}$ quantifies the corresponding change in growth for individuals of group $j$ from these competitors; and $\epsilon_{ij}$ is a random error term distributed $\text{Normal}(0, \sigma^2)$.
-@allen_permutation_2020 estimate all parameters via Bayesian linear regression, while exploiting Normal/Inverse Gamma conjugacy to derive closed-form solutions to all posterior distributions^[See S1 Appendix of @allen_permutation_2020, available at https://doi.org/10.1371/journal.pone.0229930.s004].
+where $\beta_{0,j}$ is the diameter-independent growth rate of species $j$; $\text{dbh}_{ij}$ is the DBH of the focal tree at the earlier census and $\beta_{\text{dbh},j}$ the slope of that species's diameter-growth relationship;  $x^{\text{comp}}_{ijk}$ is the sum of some numerical explanatory variable of all trees of competitor species $k$ and $\lambda_{jk}$ quantifies the corresponding change in growth for individuals of group $j$ from these competitors; and $\epsilon_{ij}$ is a random error term distributed $\text{Normal}(0, \sigma^2)$.
+
+@allen_permutation_2020 use the sum of the basal area of all trees of competitor species $k$ as $x^{\text{comp}}_{ijk}$.
+Furthermore, they estimate all parameters via Bayesian linear regression, while exploiting Normal/Inverse Gamma conjugacy to derive closed-form solutions to all posterior distributions^[See S1 Appendix of @allen_permutation_2020, available at https://doi.org/10.1371/journal.pone.0229930.s004].
 These closed-form solutions are not as computationally expensive as approximations from Markov Chain Monte Carlo algorithms.
@@
 
-We load both 2008 and 2014 SCBI census `.csv` files as they existed on GitHub on 2020/11/20 and perform minor data wrangling [@gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020].
+We load both 2008 and 2014 SCBI census `.csv` files as they existed on GitHub on 2021/08/02 and perform minor data wrangling [@gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020].
 We then only consider a 9 ha subsection of the 25.6 ha of the site to speed up computation for this example: `gx` from 0--300 instead of 0--400 and `gy` from 300--600 instead of 0--640.
@@
   labs(size = expression(paste(Growth, " (cm ",y^{-1},')')) )
+```
+
+We also load species information as it existed on GitHub on 2021/08/02, which includes family, genus, and species information; as well as classifications of the canopy position (canopy, canopy emergent, understory, shrub layer) and drought tolerance (intolerant, resistant) of the species.
+
+```{r scbi-load-sp-data}
+sp_info <- read_csv("SCBI_ForestGEO_sp_ecology.csv") %>%
+  select(
+    sp = spcode, family, genus, species, canopy_position, 
+    drought_tolerance
+  ) %>% 
+  mutate()
+sp_info
+```
+
+We join this species information to our `growth_scbi` data frame.
+
+```{r}
+growth_scbi <- growth_scbi %>% 
+  left_join(sp_info, by = "sp")
+```
+
+Furthermore, we compute two potential competitor explanatory variables $x^{\text{comp}}_{ijk}$ from Equation \ref{eq:model}.
+First, the basal area of each tree as a function of its DBH in the earlier census.
+Second, the above ground biomass as estimated by allometric equations encoded in the `get_biomass()` function from the `allodb` package [@allodb_package]; this function has DBH, species, and geographic coordinates as arguments.
+
+```{r scbi-compute-comp-explanatory-variables}
+# Install development version of allodb using:
+# remotes::install_github("forestgeo/allodb")
+library(allodb)
+growth_scbi <- growth_scbi %>% 
+  mutate(
+    # Compute basal area:
+    basal_area = 0.0001 * pi * (dbh1 / 2)^2,
+    # Compute above ground biomass:
+    agb = get_biomass(
+      dbh = dbh1,
+      genus = genus,
+      species = species,
+      coords = c(-78.2, 38.9)
+    )
+  )
 ```
@@
 
-```{r appendix2-code}
-
+```{r appendix2-code, eval = FALSE}
 census_2013_scbi <- read_csv("scbi.stem2.csv") %>%
@@
 
-for (i in 1:length(mult_dist_comp$dist))
-{
+for (i in 1:length(mult_dist_comp$dist)){
   comp_dist <- mult_dist_comp$dist[i]
@@
   ylab('RMSE (cm)')
-
 ```

### 2021-08-02 paper/paper.Rmd
Added comp_x_var argument to create_focal_vs_comp()

@@
 focal_vs_comp_scbi <- growth_scbi %>%
-  create_focal_vs_comp(comp_dist, blocks = blocks_scbi, id = "stemID")
+  create_focal_vs_comp(comp_dist, blocks = blocks_scbi, id = "stemID", comp_x_var = "agb")
 focal_vs_comp_scbi %>% 

### 2021-08-03 README.Rmd
CRAN checks pass

@@
     id = "ID"
-  )
+  ) %>% 
+  # Compute basal area:
+  mutate(basal_area = 0.0001 * pi * (dbh1 / 2)^2)
 ```
@@
 focal_vs_comp_ex <- growth_ex %>%
-  create_focal_vs_comp(comp_dist, blocks = blocks_ex, id = "ID")
+  create_focal_vs_comp(comp_dist, blocks = blocks_ex, id = "ID", comp_x_var = "basal_area")
 focal_vs_comp_ex

### 2021-08-03 R/data_processing_functions.R

@@ -159,7 +159,9 @@ compute_growth <- function(census_1, census_2, id) {
 #'
 #' # Return corresponding data frame
 #' growth_toy %>%
-#'   create_focal_vs_comp(comp_dist = 1.5, blocks = SpatialBlock_ex, id = "ID")
+#'   mutate(basal_area = 0.0001 * pi * (dbh1 / 2)^2) %>%
+#'   create_focal_vs_comp(comp_dist = 1.5, blocks = SpatialBlock_ex, id = "ID",
+#'                        comp_x_var = "basal_area")
 #'
 #' # Load in growth_df with spatial data
 #' # See ?growth_ex for attaching spatial data to growth_df
@@ -168,7 +170,9 @@ compute_growth <- function(census_1, census_2, id) {
 #' data(blocks_ex)
 #'
 #' focal_vs_comp_ex <- growth_spatial_ex %>%
-#'   create_focal_vs_comp(comp_dist = 1, blocks = blocks_ex, id = "ID")
+#'   mutate(basal_area = 0.0001 * pi * (dbh1 / 2)^2) %>%
+#'   create_focal_vs_comp(comp_dist = 1, blocks = blocks_ex, id = "ID",
+#'                        comp_x_var = "basal_area")
 create_focal_vs_comp <- function(growth_df, comp_dist, blocks, id, comp_x_var) {
   # 0. Check inputs
   check_inherits(growth_df, "data.frame")

### 2021-08-03 R/example_datasets.R

@@ -214,7 +214,8 @@
 #'
 #' # Create the focal versus comp data frame
 #' focal_vs_comp_ex <- growth_spatial_ex %>%
-#'   create_focal_vs_comp(comp_dist, blocks = blocks_ex, id = "ID")
+#'   mutate(basal_area = 0.0001 * pi * (dbh1 / 2)^2) %>%
+#'   create_focal_vs_comp(comp_dist, blocks = blocks_ex, id = "ID", comp_x_var = "basal_area")
 "growth_spatial_ex"
 
 
@@ -239,6 +240,7 @@
 #' @examples
 #' library(ggplot2)
 #' library(sf)
+#' library(dplyr)
 #'
 #' comp_dist <- 1
 #'
@@ -247,7 +249,8 @@
 #'   geom_sf(data = growth_spatial_ex)
 #'
 #' focal_vs_comp_ex <- growth_spatial_ex %>%
-#'   create_focal_vs_comp(comp_dist, blocks = blocks_ex, id = "ID")
+#'   mutate(basal_area = 0.0001 * pi * (dbh1 / 2)^2) %>%
+#'   create_focal_vs_comp(comp_dist, blocks = blocks_ex, id = "ID", comp_x_var = "basal_area")
 "blocks_ex"
 
 

### 2021-08-03 data-raw/process_data_raw.R

@@ -158,7 +158,8 @@ use_data(blocks_ex, overwrite = TRUE)
 
 ## Create focal_vs_comp data frame  ----
 focal_vs_comp_ex <- growth_spatial_ex %>%
-  create_focal_vs_comp(comp_dist = 1, blocks = blocks_ex, id = "ID")
+  mutate(basal_area = 0.0001 * pi * (dbh1 / 2)^2) %>%
+  create_focal_vs_comp(comp_dist = 1, blocks = blocks_ex, id = "ID", comp_x_var = "basal_area")
 use_data(focal_vs_comp_ex, overwrite = TRUE)
 
 

### 2021-08-04 paper/paper.Rmd
Update paper

--- 2021-08-02 paper/paper.Rmd
@@
 )
+
+# TODO:
+setwd("~/Documents/forestecology/forestecology/paper/")
+
+# TODO: set eval=TRUE for Appendix code blocks
 ```
@@
 growth_scbi <- growth_scbi %>% 
-  left_join(sp_info, by = "sp")
+  left_join(sp_info, by = "sp") %>% 
+  # TODO:
+  mutate(sp = as.factor(sp))
 ```
@@
 focal_vs_comp_scbi <- growth_scbi %>%
-  create_focal_vs_comp(comp_dist, blocks = blocks_scbi, id = "stemID")
+  create_focal_vs_comp(comp_dist, blocks = blocks_scbi, id = "stemID", comp_x_var = "basal_area")
 focal_vs_comp_scbi %>% 
@@
 
-```{r appendix1-code, eval = FALSE, cache = FALSE}
+```{r appendix1-code, eval = FALSE}
 library(tidyverse)
@@
     id = "stemID"
-  )
+  ) %>% 
+  # Compute basal area:
+  mutate(basal_area = 0.0001 * pi * (dbh1 / 2)^2)
 
@@
 focal_vs_comp_scbi <- growth_scbi %>%
-  create_focal_vs_comp(comp_dist, blocks = blocks_scbi, id = "stemID")
+  create_focal_vs_comp(comp_dist, blocks = blocks_scbi, id = "stemID", comp_x_var = "basal_area")
 
@@
   # make all species the same, needs to be factor with at least two levels
-  mutate(sp = factor('A', levels = c('A', 'B')) )
+  mutate(sp = factor('A', levels = c('A', 'B')) ) %>% 
+  # Compute basal area:
+  mutate(basal_area = 0.0001 * pi * (dbh1 / 2)^2)
 
@@
   focal_vs_comp_scbi <- growth_scbi %>%
-    create_focal_vs_comp(comp_dist = comp_dist, blocks = blocks_scbi, id = "stemID") %>%
+    create_focal_vs_comp(comp_dist = comp_dist, blocks = blocks_scbi, id = "stemID", comp_x_var = "basal_area") %>%
     run_cv(comp_dist = comp_dist, blocks = blocks_scbi)

### 2021-08-04 paper/paper.tex
Update paper

@@
 \label{eq:model}
-y_{ij} = \beta_{0,j} + \beta_{\text{dbh},j} \cdot \text{dbh}_{ij} + \sum_{k=1}^{K} \lambda_{jk} \cdot \text{BA}_{ijk} + \epsilon_{ij}
+y_{ij} = \beta_{0,j} + \beta_{\text{dbh},j} \cdot \text{dbh}_{ij} + \sum_{k=1}^{K} \lambda_{jk} \cdot x^{\text{comp}}_{ijk} + \epsilon_{ij}
 \end{equation}
@@
 census and \(\beta_{\text{dbh},j}\) the slope of that species's
-diameter-growth relationship; \(\text{BA}_{ijk}\) is the sum of the
-basal area of all trees of competitor species \(k\) and \(\lambda_{jk}\)
-quantifies the corresponding change in growth for individuals of group
-\(j\) from these competitors; and \(\epsilon_{ij}\) is a random error
-term distributed \(\text{Normal}(0, \sigma^2)\).
-\citet{allen_permutation_2020} estimate all parameters via Bayesian
-linear regression, while exploiting Normal/Inverse Gamma conjugacy to
-derive closed-form solutions to all posterior distributions\footnote{See
-  S1 Appendix of \citet{allen_permutation_2020}, available at
+diameter-growth relationship; \(x^{\text{comp}}_{ijk}\) is the sum of
+some numerical explanatory variable of all trees of competitor species
+\(k\) and \(\lambda_{jk}\) quantifies the corresponding change in growth
+for individuals of group \(j\) from these competitors; and
+\(\epsilon_{ij}\) is a random error term distributed
+\(\text{Normal}(0, \sigma^2)\).
+
+\citet{allen_permutation_2020} use the sum of the basal area of all
+trees of competitor species \(k\) as \(x^{\text{comp}}_{ijk}\).
+Furthermore, they estimate all parameters via Bayesian linear
+regression, while exploiting Normal/Inverse Gamma conjugacy to derive
+closed-form solutions to all posterior distributions\footnote{See S1
+  Appendix of \citet{allen_permutation_2020}, available at
   \url{https://doi.org/10.1371/journal.pone.0229930.s004}}. These
@@
 sequentially complete all necessary analysis steps
-\citep{bache_pipe_2020}. Furthermore, the inputs and outputs of most
-functions use the same ``simple features for R'' data structures from
-the \texttt{sf} package, a package for standardized and
-\texttt{tidyverse}-friendly wrangling and visualizing of spatial data
-\citep{pebesma_simple_2018}.
-
-Currently the package only implements the Bayesian linear regression
-model detailed in Equation \ref{eq:model}. As we demonstrate in Section
-\ref{model-fit-predict} however, the fitting of this model is
+\citep{bache_pipe_2020}.
+
+Furthermore, the inputs and outputs of most functions use the same
+``simple features for R'' data structures for spatial data from the
+\texttt{sf} package \citep{pebesma_simple_2018}. Previously \texttt{sp}
+package classes were commonly used for storing spatial data and
+interfacing with geospatial libraries \citep{bivand_sp_2013}; the
+\texttt{sf} package aims to improve on the \texttt{sp} package by:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Using simple feature access as the base standard for representing and
+  encoding spatial data, rather than shapefiles \citep{herring_2011}.
+\item
+  Leveraging improvements in external libraries for reading and writing
+  spatial data (GDAL) and for geometrical operations (GEOS)
+  \citep[\citet{geos_2017}]{warmerdam_geospatial_2008}.
+\item
+  Integrating closely with the popular \texttt{tidyverse} suite of
+  packages for data science \citep{wickham_welcome_2019}.
+\end{enumerate}
+
+By using the \texttt{sf} package classes to represent spatial data
+rather than the \texttt{sp} package, the implementation and use of the
+\texttt{forestecology} package's spatial algorithms was greatly
+simplified.
+
+Lastly, currently the package only implements the Bayesian linear
+regression model detailed in Equation \ref{eq:model}. As we demonstrate
+in Section \ref{model-fit-predict} however, the fitting of this model is
 self-contained in a single function \texttt{comp\_bayes\_lm()} which
@@
 We load both 2008 and 2014 SCBI census \texttt{.csv} files as they
-existed on GitHub on 2020/11/20 and perform minor data wrangling
+existed on GitHub on 2021/08/02 and perform minor data wrangling
 \citep{gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020}. We then
@@
 \begin{Highlighting}[]
-\NormalTok{growth_scbi <-}
-\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(}
-    \DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi,}
-    \DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)),}
-    \DataTypeTok{id =} \StringTok{"stemID"}
-\NormalTok{  )}
-\NormalTok{growth_scbi }\OperatorTok{%>%}\StringTok{ }
+\NormalTok{growth_scbi <-}\StringTok{ }\KeywordTok{compute_growth}\NormalTok{(}\DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi, }\DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)), }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{)}
+\NormalTok{growth_scbi }\OperatorTok{%>%}
 \StringTok{    }\KeywordTok{select}\NormalTok{(stemID, sp, dbh1, dbh2, growth, geometry)}
 \CommentTok{## Simple feature collection with 7954 features and 5 fields}
-\CommentTok{## geometry type:  POINT}
-\CommentTok{## dimension:      XY}
-\CommentTok{## bbox:           xmin: 0.2 ymin: 300 xmax: 300 ymax: 600}
+\CommentTok{## Geometry type: POINT}
+\CommentTok{## Dimension:     XY}
+\CommentTok{## Bounding box:  xmin: 0.2 ymin: 300 xmax: 300 ymax: 600}
 \CommentTok{## CRS:           NA}
@@
 \begin{Highlighting}[]
-\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{sample_n}\NormalTok{(}\DecValTok{500}\NormalTok{), }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{size =}\NormalTok{ growth)) }\OperatorTok{+}\StringTok{ }
-\StringTok{  }\KeywordTok{scale_size_binned}\NormalTok{(}\DataTypeTok{limits =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\DecValTok{1}\NormalTok{)) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{size =} \KeywordTok{expression}\NormalTok{(}\KeywordTok{paste}\NormalTok{(Growth, }\StringTok{" (cm "}\NormalTok{,y}\OperatorTok{^}\NormalTok{\{}\OperatorTok{-}\DecValTok{1}\NormalTok{\},}\StringTok{')'}\NormalTok{)) )}
+\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}\StringTok{ }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{sample_n}\NormalTok{(}\DecValTok{500}\NormalTok{), }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{size =}\NormalTok{ growth)) }\OperatorTok{+}\StringTok{ }\KeywordTok{scale_size_binned}\NormalTok{(}\DataTypeTok{limits =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{,}
+    \DecValTok{1}\NormalTok{)) }\OperatorTok{+}\StringTok{ }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{size =} \KeywordTok{expression}\NormalTok{(}\KeywordTok{paste}\NormalTok{(Growth, }\StringTok{" (cm "}\NormalTok{, y}\OperatorTok{^}\NormalTok{\{}
+    \DecValTok{-1}
+\NormalTok{\}, }\StringTok{")"}\NormalTok{)))}
 \end{Highlighting}
@@
 \end{figure}
+
+We also load species information as it existed on GitHub on 2021/08/02,
+which includes family, genus, and species information; as well as
+classifications of the canopy position (canopy, canopy emergent,
+understory, shrub layer) and drought tolerance (intolerant, resistant)
+of the species.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{sp_info <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"SCBI_ForestGEO_sp_ecology.csv"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{select}\NormalTok{(}\DataTypeTok{sp =}\NormalTok{ spcode, family, genus, species, canopy_position, drought_tolerance) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{()}
+\NormalTok{sp_info}
+\CommentTok{## # A tibble: 65 x 6}
+\CommentTok{##   sp    family        genus       species     canopy_position drought_toleran~}
+\CommentTok{##   <chr> <chr>         <chr>       <chr>       <chr>           <chr>           }
+\CommentTok{## 1 acne  Sapindaceae   Acer        negundo     understory      drought-resista~}
+\CommentTok{## 2 acpl  Sapindaceae   Acer        platanoides canopy          drought-resista~}
+\CommentTok{## 3 acru  Sapindaceae   Acer        rubrum      canopy          drought-resista~}
+\CommentTok{## 4 aial  Simaroubaceae Ailanthus   altissima   canopy          drought-resista~}
+\CommentTok{## 5 amar  Rosaceae      Amelanchier arborea     understory      drought-resista~}
+\CommentTok{## # ... with 60 more rows}
+\end{Highlighting}
+\end{Shaded}
+
+We join this species information to our \texttt{growth\_scbi} data
+frame.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{left_join}\NormalTok{(sp_info, }\DataTypeTok{by =} \StringTok{"sp"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\CommentTok{# }\AlertTok{TODO}\CommentTok{:}
+\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{as.factor}\NormalTok{(sp))}
+\end{Highlighting}
+\end{Shaded}
+
+Furthermore, we compute two potential competitor explanatory variables
+\(x^{\text{comp}}_{ijk}\) from Equation \ref{eq:model}. First, the basal
+area of each tree as a function of its DBH in the earlier census.
+Second, the above ground biomass as estimated by allometric equations
+encoded in the \texttt{get\_biomass()} function from the \texttt{allodb}
+package \citep{allodb_package}; this function has DBH, species, and
+geographic coordinates as arguments.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{# Install development version of allodb using:}
+\CommentTok{# remotes::install_github("forestgeo/allodb")}
+\KeywordTok{library}\NormalTok{(allodb)}
+\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
+    \CommentTok{# Compute basal area:}
+    \DataTypeTok{basal_area =} \FloatTok{0.0001} \OperatorTok{*}\StringTok{ }\NormalTok{pi }\OperatorTok{*}\StringTok{ }\NormalTok{(dbh1 }\OperatorTok{/}\StringTok{ }\DecValTok{2}\NormalTok{)}\OperatorTok{^}\DecValTok{2}\NormalTok{,}
+    \CommentTok{# Compute above ground biomass:}
+    \DataTypeTok{agb =} \KeywordTok{get_biomass}\NormalTok{(}
+      \DataTypeTok{dbh =}\NormalTok{ dbh1,}
+      \DataTypeTok{genus =}\NormalTok{ genus,}
+      \DataTypeTok{species =}\NormalTok{ species,}
+      \DataTypeTok{coords =} \KeywordTok{c}\NormalTok{(}\OperatorTok{-}\FloatTok{78.2}\NormalTok{, }\FloatTok{38.9}\NormalTok{)}
+\NormalTok{    )}
+\NormalTok{  )}
+\end{Highlighting}
+\end{Shaded}
 
@@
 \CommentTok{# Manually construct study region boundary}
-\NormalTok{study_region_scbi <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}
-  \DataTypeTok{x =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{),}
-  \DataTypeTok{y =} \KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{300}\NormalTok{)}
-\NormalTok{) }\OperatorTok{%>%}
+\NormalTok{study_region_scbi <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}\DataTypeTok{x =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{), }\DataTypeTok{y =} \KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{,}
+    \DecValTok{600}\NormalTok{, }\DecValTok{300}\NormalTok{)) }\OperatorTok{%>%}
 \StringTok{    }\KeywordTok{sf_polygon}\NormalTok{()}
@@
 
-\NormalTok{blocks_scbi <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}
-  \KeywordTok{sf_polygon}\NormalTok{(fold1), }\KeywordTok{sf_polygon}\NormalTok{(fold2), }\KeywordTok{sf_polygon}\NormalTok{(fold3), }
-  \KeywordTok{sf_polygon}\NormalTok{(fold4)}
-\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\NormalTok{n_fold) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
+\NormalTok{blocks_scbi <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}\KeywordTok{sf_polygon}\NormalTok{(fold1), }\KeywordTok{sf_polygon}\NormalTok{(fold2), }\KeywordTok{sf_polygon}\NormalTok{(fold3),}
+    \KeywordTok{sf_polygon}\NormalTok{(fold4)) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\NormalTok{n_fold) }\OperatorTok{%>%}
+\StringTok{        }\KeywordTok{factor}\NormalTok{())}
 
 \CommentTok{# Associate each observation to a fold}
-\NormalTok{spatial_block_scbi <-}\StringTok{ }
-\StringTok{  }\KeywordTok{spatialBlock}\NormalTok{(}\DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =}\NormalTok{ n_fold, }
-               \DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }
-               \DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}\NormalTok{)}
+\NormalTok{spatial_block_scbi <-}\StringTok{ }\KeywordTok{spatialBlock}\NormalTok{(}\DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =}\NormalTok{ n_fold,}
+    \DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}\NormalTok{)}
 
 \NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ spatial_block_scbi}\OperatorTok{$}\NormalTok{foldID }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ blocks_scbi, }\DataTypeTok{fill =} \StringTok{"transparent"}\NormalTok{, }\DataTypeTok{linetype =} \StringTok{"dashed"}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_sf_text}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{sample_n}\NormalTok{(}\DecValTok{1000}\NormalTok{), }
-               \KeywordTok{aes}\NormalTok{(}\DataTypeTok{label =}\NormalTok{ foldID, }\DataTypeTok{col =}\NormalTok{ buffer))}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ spatial_block_scbi}\OperatorTok{$}\NormalTok{foldID }\OperatorTok{%>%}
+\StringTok{        }\KeywordTok{factor}\NormalTok{())}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}\StringTok{ }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ blocks_scbi, }\DataTypeTok{fill =} \StringTok{"transparent"}\NormalTok{, }\DataTypeTok{linetype =} \StringTok{"dashed"}\NormalTok{) }\OperatorTok{+}
+\StringTok{    }\KeywordTok{geom_sf_text}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_scbi }\OperatorTok{%>%}
+\StringTok{        }\KeywordTok{sample_n}\NormalTok{(}\DecValTok{1000}\NormalTok{), }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{label =}\NormalTok{ foldID, }\DataTypeTok{col =}\NormalTok{ buffer))}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
+\CommentTok{# }\AlertTok{TODO}\CommentTok{:}
+\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{comp_x_var =}\NormalTok{ basal_area)}
+
 \NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
 \StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{)}
-\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}\StringTok{ }
+\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
 \StringTok{    }\KeywordTok{select}\NormalTok{(focal_ID, focal_sp, geometry, growth, comp)}
@@
 \begin{Highlighting}[]
-\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{filter}\NormalTok{(focal_ID }\OperatorTok{==}\StringTok{ }\DecValTok{4}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{select}\NormalTok{(focal_ID, dbh, comp) }\OperatorTok{%>%}\StringTok{ }
+\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{filter}\NormalTok{(focal_ID }\OperatorTok{==}\StringTok{ }\DecValTok{4}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{select}\NormalTok{(focal_ID, dbh, comp) }\OperatorTok{%>%}
 \StringTok{    }\KeywordTok{unnest}\NormalTok{(}\DataTypeTok{cols =} \StringTok{"comp"}\NormalTok{)}
 \CommentTok{## # A tibble: 20 x 6}
-\CommentTok{##   focal_ID   dbh comp_ID  dist comp_sp comp_basal_area}
+\CommentTok{##   focal_ID   dbh comp_ID  dist comp_sp comp_x_var}
 \CommentTok{##      <dbl> <dbl>   <dbl> <dbl> <fct>        <dbl>}
@@
 
-\NormalTok{plot1 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_scbi, }\DataTypeTok{type =} \StringTok{"intercepts"}\NormalTok{, }
-                  \DataTypeTok{sp_to_plot =}\NormalTok{ sp_to_plot)}
-\NormalTok{plot2 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_scbi, }\DataTypeTok{type =} \StringTok{"dbh_slopes"}\NormalTok{, }
-                  \DataTypeTok{sp_to_plot =}\NormalTok{ sp_to_plot)}
-\NormalTok{plot3 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_scbi, }\DataTypeTok{type =} \StringTok{"competition"}\NormalTok{, }
-                  \DataTypeTok{sp_to_plot =}\NormalTok{ sp_to_plot)}
+\NormalTok{plot1 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_scbi, }\DataTypeTok{type =} \StringTok{"intercepts"}\NormalTok{, }\DataTypeTok{sp_to_plot =}\NormalTok{ sp_to_plot)}
+\NormalTok{plot2 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_scbi, }\DataTypeTok{type =} \StringTok{"dbh_slopes"}\NormalTok{, }\DataTypeTok{sp_to_plot =}\NormalTok{ sp_to_plot)}
+\NormalTok{plot3 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_scbi, }\DataTypeTok{type =} \StringTok{"competition"}\NormalTok{, }\DataTypeTok{sp_to_plot =}\NormalTok{ sp_to_plot)}
 
 \CommentTok{# Combine plots using the patchwork package}
-\NormalTok{(plot1 }\OperatorTok{|}\StringTok{ }\NormalTok{plot2) }\OperatorTok{/}\StringTok{ }\NormalTok{plot3}
+\NormalTok{(plot1 }\OperatorTok{|}\StringTok{ }\NormalTok{plot2)}\OperatorTok{/}\NormalTok{plot3}
 \end{Highlighting}
@@
 \NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{growth_hat_shuffle =} \KeywordTok{predict}\NormalTok{(comp_bayes_lm_scbi_shuffle, }
-                                 \DataTypeTok{newdata =}\NormalTok{ focal_vs_comp_scbi))}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{growth_hat_shuffle =} \KeywordTok{predict}\NormalTok{(comp_bayes_lm_scbi_shuffle, }\DataTypeTok{newdata =}\NormalTok{ focal_vs_comp_scbi))}
 \end{Highlighting}
@@
 \texttt{tidy} data design principles, leverages the \texttt{sf} package
-for spatial data, and S3 open-oriented model implementation structure.
-We hope that the package will increase the use of neighborhood
-competition models to better understand what structures plant
-competition.
+for spatial data, and S3 open-oriented model implementation structure
+\citep{pebesma_simple_2018}. We hope that the package will increase the
+use of neighborhood competition models to better understand what
+structures plant competition.
 
@@
 
-\CommentTok{# Compute growth of trees based on census data -------------------------}
+\CommentTok{# Compute growth of trees based on census data}
+\CommentTok{# -------------------------}
 \NormalTok{census_}\DecValTok{2013}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{here}\NormalTok{(}\StringTok{"paper/scbi.stem2.csv"}\NormalTok{) }\OperatorTok{%>%}
@@
 \StringTok{    }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
-    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date),}
-    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}
-\NormalTok{  ) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date), }\DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}\NormalTok{) }\OperatorTok{%>%}
 \StringTok{    }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
@@
 \StringTok{    }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
-    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date),}
-    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}
-\NormalTok{  ) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date), }\DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}\NormalTok{) }\OperatorTok{%>%}
 \StringTok{    }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
 
-\NormalTok{growth_scbi <-}
-\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(}
-    \DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi,}
-    \DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)),}
-    \DataTypeTok{id =} \StringTok{"stemID"}
-\NormalTok{  )}
-
-\CommentTok{# Add spatial information ----------------------------------------------}
-\CommentTok{# Define buffer region using competitive distance range}
+\NormalTok{growth_scbi <-}\StringTok{ }\KeywordTok{compute_growth}\NormalTok{(}\DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi, }\DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)), }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{)}
+
+\CommentTok{# Add spatial information}
+\CommentTok{# ---------------------------------------------- Define buffer region}
+\CommentTok{# using competitive distance range}
 \NormalTok{comp_dist <-}\StringTok{ }\FloatTok{7.5}
 
-\NormalTok{study_region_scbi <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}
-  \DataTypeTok{x =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{),}
-  \DataTypeTok{y =} \KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{300}\NormalTok{)}
-\NormalTok{) }\OperatorTok{%>%}
+\NormalTok{study_region_scbi <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}\DataTypeTok{x =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{), }\DataTypeTok{y =} \KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{,}
+    \DecValTok{600}\NormalTok{, }\DecValTok{300}\NormalTok{)) }\OperatorTok{%>%}
 \StringTok{    }\KeywordTok{sf_polygon}\NormalTok{()}
@@
 
-\NormalTok{blocks_scbi <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}
-  \KeywordTok{sf_polygon}\NormalTok{(fold1), }\KeywordTok{sf_polygon}\NormalTok{(fold2), }
-  \KeywordTok{sf_polygon}\NormalTok{(fold3), }\KeywordTok{sf_polygon}\NormalTok{(fold4)}
-\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\NormalTok{n_fold) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
+\NormalTok{blocks_scbi <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}\KeywordTok{sf_polygon}\NormalTok{(fold1), }\KeywordTok{sf_polygon}\NormalTok{(fold2), }\KeywordTok{sf_polygon}\NormalTok{(fold3),}
+    \KeywordTok{sf_polygon}\NormalTok{(fold4)) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\NormalTok{n_fold) }\OperatorTok{%>%}
+\StringTok{        }\KeywordTok{factor}\NormalTok{())}
 
 \CommentTok{# Associate each observation to a fold}
-\NormalTok{SpatialBlock_scbi <-}\StringTok{ }\KeywordTok{spatialBlock}\NormalTok{(}
-  \DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =}\NormalTok{ n_fold, }\DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{,}
-  \DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}
-\NormalTok{)}
+\NormalTok{SpatialBlock_scbi <-}\StringTok{ }\KeywordTok{spatialBlock}\NormalTok{(}\DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =}\NormalTok{ n_fold,}
+    \DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}\NormalTok{)}
 
 \NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ SpatialBlock_scbi}\OperatorTok{$}\NormalTok{foldID }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
-
-\CommentTok{# Compute focal versus competitor tree information ---------------------}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ SpatialBlock_scbi}\OperatorTok{$}\NormalTok{foldID }\OperatorTok{%>%}
+\StringTok{        }\KeywordTok{factor}\NormalTok{())}
+
+\CommentTok{# Compute focal versus competitor tree information}
+\CommentTok{# ---------------------}
 \NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
@@
 
-\CommentTok{# Fit model and make predictions ---------------------------------------}
-\CommentTok{# Number of permutation shuffles:}
+\CommentTok{# Fit model and make predictions}
+\CommentTok{# --------------------------------------- Number of permutation}
+\CommentTok{# shuffles:}
 \NormalTok{num_shuffle <-}\StringTok{ }\DecValTok{49}
@@
 
-\CommentTok{# Run all simulations}
-\CommentTok{# 0. Setup simulation for this species type ----}
+\CommentTok{# Run all simulations 0. Setup simulation for this species type ----}
 \CommentTok{# Start clock}
@@
 
-\CommentTok{# 1. Compute observed test statistic: RMSE with no cross-validation ----}
-\CommentTok{# Fit model (compute posterior parameters)}
+\CommentTok{# 1. Compute observed test statistic: RMSE with no cross-validation}
+\CommentTok{# ---- Fit model (compute posterior parameters)}
 \NormalTok{comp_bayes_lm_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
@@
 \NormalTok{observed_RMSE <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{growth_hat =} 
-           \KeywordTok{predict}\NormalTok{(comp_bayes_lm_scbi, focal_vs_comp_scbi)) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{growth_hat =} \KeywordTok{predict}\NormalTok{(comp_bayes_lm_scbi, focal_vs_comp_scbi)) }\OperatorTok{%>%}
 \StringTok{    }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
@@
 
-\CommentTok{# 2. Compute observed test statistic: RMSE with cross-validation -------}
+\CommentTok{# 2. Compute observed test statistic: RMSE with cross-validation}
+\CommentTok{# -------}
 \NormalTok{observed_RMSE_CV <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
@@
 
-\CommentTok{# 3. Permutation distribution: RMSE with no cross-validation -----------}
-\CommentTok{# Compute num_shuffle permutation test statistics}
+\CommentTok{# 3. Permutation distribution: RMSE with no cross-validation}
+\CommentTok{# ----------- Compute num_shuffle permutation test statistics}
 \NormalTok{shuffle_RMSE <-}\StringTok{ }\KeywordTok{numeric}\NormalTok{(}\DataTypeTok{length =}\NormalTok{ num_shuffle)}
-\ControlFlowTok{for}\NormalTok{(j }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\NormalTok{num_shuffle)\{}
+\ControlFlowTok{for}\NormalTok{ (j }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\NormalTok{num_shuffle) \{}
     \CommentTok{# Fit model (compute posterior parameters) with shuffling}
@@
 \NormalTok{    shuffle_RMSE[j] <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{growth_hat =} 
-             \KeywordTok{predict}\NormalTok{(comp_bayes_lm_scbi, focal_vs_comp_scbi)) }\OperatorTok{%>%}
+\StringTok{        }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{growth_hat =} \KeywordTok{predict}\NormalTok{(comp_bayes_lm_scbi, focal_vs_comp_scbi)) }\OperatorTok{%>%}
 \StringTok{        }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
@@
 
-\CommentTok{# 4. Permutation distribution: RMSE with cross-validation --------------}
+\CommentTok{# 4. Permutation distribution: RMSE with cross-validation}
+\CommentTok{# -------------- Compute num_shuffle permutation test statistics}
+\NormalTok{shuffle_RMSE_CV <-}\StringTok{ }\KeywordTok{numeric}\NormalTok{(}\DataTypeTok{length =}\NormalTok{ num_shuffle)}
+
 \CommentTok{# Compute num_shuffle permutation test statistics}
-\NormalTok{shuffle_RMSE_CV <-}\StringTok{ }\KeywordTok{numeric}\NormalTok{(}\DataTypeTok{length =}\NormalTok{ num_shuffle)}
-
-\CommentTok{# Compute num_shuffle permutation test statistics}
-\ControlFlowTok{for}\NormalTok{(j }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\NormalTok{num_shuffle)\{}
+\ControlFlowTok{for}\NormalTok{ (j }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\NormalTok{num_shuffle) \{}
     \CommentTok{# Compute and save RMSE}
 \NormalTok{    shuffle_RMSE_CV[j] <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }
-           \DataTypeTok{run_shuffle =} \OtherTok{TRUE}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{        }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{run_shuffle =} \OtherTok{TRUE}\NormalTok{) }\OperatorTok{%>%}
 \StringTok{        }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
@@
     \CommentTok{# Status update}
-  \KeywordTok{str_c}\NormalTok{(}\StringTok{"Shuffle with permutation "}\NormalTok{, j, }\StringTok{" at "}\NormalTok{, }\KeywordTok{Sys.time}\NormalTok{()) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{print}\NormalTok{()}
+    \KeywordTok{str_c}\NormalTok{(}\StringTok{"Shuffle with permutation "}\NormalTok{, j, }\StringTok{" at "}\NormalTok{, }\KeywordTok{Sys.time}\NormalTok{()) }\OperatorTok{%>%}
+\StringTok{        }\KeywordTok{print}\NormalTok{()}
 \NormalTok{\}}
@@
 
-\NormalTok{model_comp_tbl <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}
-  \DataTypeTok{run_time =}\NormalTok{ run_time,}
-  \DataTypeTok{observed_RMSE =}\NormalTok{ observed_RMSE,}
-  \DataTypeTok{observed_RMSE_CV =}\NormalTok{ observed_RMSE_CV,}
-  \DataTypeTok{shuffle_RMSE =}\NormalTok{ shuffle_RMSE,}
-  \DataTypeTok{shuffle_RMSE_CV =}\NormalTok{ shuffle_RMSE_CV,}
-\NormalTok{)}
-\KeywordTok{save}\NormalTok{(model_comp_tbl, }\DataTypeTok{file =}\NormalTok{ filename }\OperatorTok{%>%}\StringTok{ }\KeywordTok{str_c}\NormalTok{(}\StringTok{".RData"}\NormalTok{))}
-
-\CommentTok{# Visualize results ----------------------------------------------------}
-\NormalTok{model_comp <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}
-\NormalTok{  model_comp_tbl }\OperatorTok{%>%}\StringTok{ }
-\StringTok{    }\KeywordTok{select}\NormalTok{(run_time, }\DataTypeTok{observed =}\NormalTok{ observed_RMSE, }
-           \DataTypeTok{shuffle =}\NormalTok{ shuffle_RMSE) }\OperatorTok{%>%}\StringTok{ }
-\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{CV =} \OtherTok{FALSE}\NormalTok{),}
-\NormalTok{  model_comp_tbl }\OperatorTok{%>%}\StringTok{ }
-\StringTok{    }\KeywordTok{select}\NormalTok{(run_time, }\DataTypeTok{observed =}\NormalTok{ observed_RMSE_CV, }
-           \DataTypeTok{shuffle =}\NormalTok{ shuffle_RMSE_CV) }\OperatorTok{%>%}\StringTok{ }
-\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{CV =} \OtherTok{TRUE}\NormalTok{)}
-\NormalTok{) }\OperatorTok{%>%}
+\NormalTok{model_comp_tbl <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}\DataTypeTok{run_time =}\NormalTok{ run_time, }\DataTypeTok{observed_RMSE =}\NormalTok{ observed_RMSE,}
+    \DataTypeTok{observed_RMSE_CV =}\NormalTok{ observed_RMSE_CV, }\DataTypeTok{shuffle_RMSE =}\NormalTok{ shuffle_RMSE, }\DataTypeTok{shuffle_RMSE_CV =}\NormalTok{ shuffle_RMSE_CV,}
+\NormalTok{    )}
+\KeywordTok{save}\NormalTok{(model_comp_tbl, }\DataTypeTok{file =}\NormalTok{ filename }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{str_c}\NormalTok{(}\StringTok{".RData"}\NormalTok{))}
+
+\CommentTok{# Visualize results}
+\CommentTok{# ----------------------------------------------------}
+\NormalTok{model_comp <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(model_comp_tbl }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{select}\NormalTok{(run_time, }\DataTypeTok{observed =}\NormalTok{ observed_RMSE, }\DataTypeTok{shuffle =}\NormalTok{ shuffle_RMSE) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{CV =} \OtherTok{FALSE}\NormalTok{), model_comp_tbl }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{select}\NormalTok{(run_time, }\DataTypeTok{observed =}\NormalTok{ observed_RMSE_CV, }\DataTypeTok{shuffle =}\NormalTok{ shuffle_RMSE_CV) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{CV =} \OtherTok{TRUE}\NormalTok{)) }\OperatorTok{%>%}
 \StringTok{    }\KeywordTok{gather}\NormalTok{(type, RMSE, }\OperatorTok{-}\KeywordTok{c}\NormalTok{(run_time, CV))}
@@
 
-\NormalTok{cv_plot <-}\StringTok{ }\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_vline}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ model_comp_observed, }
-             \KeywordTok{aes}\NormalTok{(}\DataTypeTok{xintercept =}\NormalTok{ RMSE, }\DataTypeTok{col =}\NormalTok{ CV), }
-             \DataTypeTok{linetype =} \StringTok{"dashed"}\NormalTok{, }\DataTypeTok{show.legend =}\NormalTok{ F) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_histogram}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ model_comp_shuffle, }
-                 \KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ RMSE, }\DataTypeTok{fill =}\NormalTok{ CV), }\DataTypeTok{bins =} \DecValTok{50}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{labs}\NormalTok{(}
-    \DataTypeTok{fill =} \StringTok{"Cross-validated?"}\NormalTok{,}
-    \DataTypeTok{x =} \KeywordTok{expression}\NormalTok{(}\KeywordTok{paste}\NormalTok{(}\StringTok{"RMSE (cm "}\NormalTok{, y}\OperatorTok{^}\NormalTok{\{}\OperatorTok{-}\DecValTok{1}\NormalTok{\}, }\StringTok{")"}\NormalTok{))}
-\NormalTok{  ) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{scale_color_viridis}\NormalTok{(}\DataTypeTok{discrete =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{option =} \StringTok{"D"}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{scale_fill_viridis}\NormalTok{(}\DataTypeTok{discrete =} \OtherTok{TRUE}\NormalTok{) }\OperatorTok{+}
+\NormalTok{cv_plot <-}\StringTok{ }\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}\StringTok{ }\KeywordTok{geom_vline}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ model_comp_observed, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{xintercept =}\NormalTok{ RMSE,}
+    \DataTypeTok{col =}\NormalTok{ CV), }\DataTypeTok{linetype =} \StringTok{"dashed"}\NormalTok{, }\DataTypeTok{show.legend =}\NormalTok{ F) }\OperatorTok{+}\StringTok{ }\KeywordTok{geom_histogram}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ model_comp_shuffle,}
+    \KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ RMSE, }\DataTypeTok{fill =}\NormalTok{ CV), }\DataTypeTok{bins =} \DecValTok{50}\NormalTok{) }\OperatorTok{+}\StringTok{ }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{fill =} \StringTok{"Cross-validated?"}\NormalTok{,}
+    \DataTypeTok{x =} \KeywordTok{expression}\NormalTok{(}\KeywordTok{paste}\NormalTok{(}\StringTok{"RMSE (cm "}\NormalTok{, y}\OperatorTok{^}\NormalTok{\{}
+        \DecValTok{-1}
+\NormalTok{    \}, }\StringTok{")"}\NormalTok{))) }\OperatorTok{+}\StringTok{ }\KeywordTok{scale_color_viridis}\NormalTok{(}\DataTypeTok{discrete =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{option =} \StringTok{"D"}\NormalTok{) }\OperatorTok{+}\StringTok{ }\KeywordTok{scale_fill_viridis}\NormalTok{(}\DataTypeTok{discrete =} \OtherTok{TRUE}\NormalTok{) }\OperatorTok{+}
 \StringTok{    }\KeywordTok{theme_light}\NormalTok{()}
@@
 \begin{Highlighting}[]
-
 \NormalTok{census_}\DecValTok{2013}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem2.csv"}\NormalTok{) }\OperatorTok{%>%}
 \StringTok{    }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
-    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date),}
-    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}
-\NormalTok{  ) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date), }\DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}\NormalTok{) }\OperatorTok{%>%}
 \StringTok{    }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{400}\NormalTok{, gy }\OperatorTok{<}\StringTok{ }\DecValTok{400}\NormalTok{)}
@@
 \StringTok{    }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
-    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date),}
-    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}
-\NormalTok{  ) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date), }\DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}\NormalTok{) }\OperatorTok{%>%}
 \StringTok{    }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{400}\NormalTok{, gy }\OperatorTok{<}\StringTok{ }\DecValTok{400}\NormalTok{)}
 
-\NormalTok{growth_scbi <-}
-\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(}
-    \DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi,}
-    \DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)),}
-    \DataTypeTok{id =} \StringTok{"stemID"}
-\NormalTok{  ) }\OperatorTok{%>%}
-\StringTok{  }\CommentTok{# make all species the same, needs to be factor with at least two levels}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{factor}\NormalTok{(}\StringTok{'A'}\NormalTok{, }\DataTypeTok{levels =} \KeywordTok{c}\NormalTok{(}\StringTok{'A'}\NormalTok{, }\StringTok{'B'}\NormalTok{)) )}
-
-\NormalTok{study_region_scbi <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}
-  \DataTypeTok{x =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{),}
-  \DataTypeTok{y =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{0}\NormalTok{)}
-\NormalTok{) }\OperatorTok{%>%}
+\NormalTok{growth_scbi <-}\StringTok{ }\KeywordTok{compute_growth}\NormalTok{(}\DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi, }\DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)), }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\CommentTok{# make all species the same, needs to be factor with at least two}
+\StringTok{    }\CommentTok{# levels}
+\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{factor}\NormalTok{(}\StringTok{"A"}\NormalTok{, }\DataTypeTok{levels =} \KeywordTok{c}\NormalTok{(}\StringTok{"A"}\NormalTok{, }\StringTok{"B"}\NormalTok{)))}
+
+\NormalTok{study_region_scbi <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}\DataTypeTok{x =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{), }\DataTypeTok{y =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{400}\NormalTok{,}
+    \DecValTok{0}\NormalTok{)) }\OperatorTok{%>%}
 \StringTok{    }\KeywordTok{sf_polygon}\NormalTok{()}
@@
 
-\NormalTok{blocks_scbi <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}
-  \KeywordTok{sf_polygon}\NormalTok{(fold1), }\KeywordTok{sf_polygon}\NormalTok{(fold2), }\KeywordTok{sf_polygon}\NormalTok{(fold3), }
-  \KeywordTok{sf_polygon}\NormalTok{(fold4)}
-\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\NormalTok{n_fold) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
+\NormalTok{blocks_scbi <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}\KeywordTok{sf_polygon}\NormalTok{(fold1), }\KeywordTok{sf_polygon}\NormalTok{(fold2), }\KeywordTok{sf_polygon}\NormalTok{(fold3),}
+    \KeywordTok{sf_polygon}\NormalTok{(fold4)) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\NormalTok{n_fold) }\OperatorTok{%>%}
+\StringTok{        }\KeywordTok{factor}\NormalTok{())}
 
 \CommentTok{# Associate each observation to a fold}
-\NormalTok{spatial_block_scbi <-}\StringTok{ }
-\StringTok{  }\KeywordTok{spatialBlock}\NormalTok{(}\DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =}\NormalTok{ n_fold, }
-               \DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }
-               \DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}\NormalTok{)}
+\NormalTok{spatial_block_scbi <-}\StringTok{ }\KeywordTok{spatialBlock}\NormalTok{(}\DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =}\NormalTok{ n_fold,}
+    \DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}\NormalTok{)}
 
 \NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ spatial_block_scbi}\OperatorTok{$}\NormalTok{foldID }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
-
-\NormalTok{mult_dist_comp <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}
-  \DataTypeTok{dist =} \KeywordTok{c}\NormalTok{(}\DecValTok{5}\NormalTok{ ,}\FloatTok{6.25}\NormalTok{,  }\FloatTok{7.5}\NormalTok{, }\FloatTok{8.75}\NormalTok{,  }\DecValTok{10}\NormalTok{),}
-  \DataTypeTok{rmse =} \DecValTok{0}
-\NormalTok{)}
-
-\ControlFlowTok{for}\NormalTok{ (i }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(mult_dist_comp}\OperatorTok{$}\NormalTok{dist))}
-\NormalTok{\{}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ spatial_block_scbi}\OperatorTok{$}\NormalTok{foldID }\OperatorTok{%>%}
+\StringTok{        }\KeywordTok{factor}\NormalTok{())}
+
+\NormalTok{mult_dist_comp <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}\DataTypeTok{dist =} \KeywordTok{c}\NormalTok{(}\DecValTok{5}\NormalTok{, }\FloatTok{6.25}\NormalTok{, }\FloatTok{7.5}\NormalTok{, }\FloatTok{8.75}\NormalTok{, }\DecValTok{10}\NormalTok{), }\DataTypeTok{rmse =} \DecValTok{0}\NormalTok{)}
+
+\ControlFlowTok{for}\NormalTok{ (i }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(mult_dist_comp}\OperatorTok{$}\NormalTok{dist)) \{}
 \NormalTok{    comp_dist <-}\StringTok{ }\NormalTok{mult_dist_comp}\OperatorTok{$}\NormalTok{dist[i]}
@@
 \NormalTok{    focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{        }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi,}
+            \DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{) }\OperatorTok{%>%}
 \StringTok{        }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi)}
@@
 \NormalTok{mult_dist_comp }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(dist,rmse)) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_point}\NormalTok{() }\OperatorTok{+}
-\StringTok{  }\KeywordTok{xlab}\NormalTok{(}\StringTok{'Distance (m)'}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{ylab}\NormalTok{(}\StringTok{'RMSE (cm)'}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{center}\includegraphics[width=1\linewidth]{Figures/appendix2-code-1} \end{center}
+\StringTok{    }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(dist, rmse)) }\OperatorTok{+}\StringTok{ }\KeywordTok{geom_point}\NormalTok{() }\OperatorTok{+}\StringTok{ }\KeywordTok{xlab}\NormalTok{(}\StringTok{"Distance (m)"}\NormalTok{) }\OperatorTok{+}\StringTok{ }\KeywordTok{ylab}\NormalTok{(}\StringTok{"RMSE (cm)"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 

### 2021-08-06 paper/paper.Rmd
results for appedices 3+4 need to edit text

@@
 )
-
-# TODO:
-setwd("~/Documents/forestecology/forestecology/paper/")
-
-# TODO: set eval=TRUE for Appendix code blocks
 ```
@@
 
+# Appendix 3: Compare predictor variables {#compare-pred-var}
+
+```{r compar-pred-var}
+census_2013_scbi <- read_csv("scbi.stem2.csv") %>%
+  select(stemID, sp, date = ExactDate, gx, gy, dbh, codes, status) %>%
+  mutate(
+    date = mdy(date),
+    dbh = as.numeric(dbh)/10
+  ) %>%
+  filter(gx < 300, between(gy, 300, 600))
+
+census_2018_scbi <- read_csv("scbi.stem3.csv") %>%
+  select(stemID, sp, date = ExactDate, gx, gy, dbh, codes, status) %>%
+  mutate(
+    date = mdy(date),
+    dbh = as.numeric(dbh)/10
+  ) %>%
+  filter(gx < 300, between(gy, 300, 600))
+
+growth_scbi <-
+  compute_growth(
+    census_1 = census_2013_scbi,
+    census_2 = census_2018_scbi %>% filter(!str_detect(codes, "R")),
+    id = "stemID"
+  ) %>%
+  left_join(sp_info, by = "sp") %>% 
+  mutate(sp = as.factor(sp),
+        basal_area = 0.0001 * pi * (dbh1 / 2)^2,
+         agb = get_biomass(
+                  dbh = dbh1,
+                  genus = genus,
+                  species = species,
+                  coords = c(-78.2, 38.9)
+    ))
+
+study_region_scbi <- tibble(
+  x = c(0, 300, 300, 0, 0),
+  y = c(300, 300, 600, 600, 300)
+) %>%
+  sf_polygon()
+
+n_fold <- 4
+fold1 <- cbind(c(0, 150, 150, 0), c(300, 300, 450, 450))
+fold2 <- cbind(c(150, 300, 300, 150), c(300, 300, 450, 450))
+fold3 <- cbind(c(0, 150, 150, 0), c(450, 450, 600, 600))
+fold4 <- cbind(c(150, 300, 300, 150), c(450, 450, 600, 600))
+
+blocks_scbi <- bind_rows(
+  sf_polygon(fold1), sf_polygon(fold2), sf_polygon(fold3), 
+  sf_polygon(fold4)
+) %>%
+  mutate(folds = c(1:n_fold) %>% factor())
+
+# Associate each observation to a fold
+spatial_block_scbi <- 
+  spatialBlock(speciesData = growth_scbi, k = n_fold, 
+               selection = "systematic", blocks = blocks_scbi, 
+               showBlocks = FALSE, verbose = FALSE)
+
+growth_scbi <- growth_scbi %>%
+  mutate(foldID = spatial_block_scbi$foldID %>% factor())
+
+comp_dist <- 7.5
+
+growth_scbi <- growth_scbi %>%
+    add_buffer_variable(size = comp_dist, region = study_region_scbi)
+
+focal_vs_comp_ba <- growth_scbi %>%
+    create_focal_vs_comp(comp_dist = comp_dist, 
+                         blocks = blocks_scbi, 
+                         id = "stemID", 
+                         comp_x_var = "basal_area") %>%
+    run_cv(comp_dist = comp_dist, blocks = blocks_scbi)
+
+focal_vs_comp_agb <- growth_scbi %>%
+    create_focal_vs_comp(comp_dist = comp_dist, 
+                         blocks = blocks_scbi, 
+                         id = "stemID", 
+                         comp_x_var = "agb") %>%
+    run_cv(comp_dist = comp_dist, blocks = blocks_scbi)
+
+  focal_vs_comp_ba %>%
+    rmse(truth = growth, estimate = growth_hat) %>%
+    pull(.estimate)
+
+  focal_vs_comp_agb %>%
+    rmse(truth = growth, estimate = growth_hat) %>%
+    pull(.estimate)
+
+```
+
+# Appendix 4: Compare grouping variables {#compare-group-var}
+
+Creative use of the grouping variable allows for tests of what best explains competitive interactions. Here we compare whether species idenity or canopy position of the species best explains.
+
+```{r compare-group-var}
+
+census_2013_scbi <- read_csv("scbi.stem2.csv") %>%
+  select(stemID, sp, date = ExactDate, gx, gy, dbh, codes, status) %>%
+  mutate(
+    date = mdy(date),
+    dbh = as.numeric(dbh)/10
+  ) %>%
+  filter(gx < 300, between(gy, 300, 600))
+
+census_2018_scbi <- read_csv("scbi.stem3.csv") %>%
+  select(stemID, sp, date = ExactDate, gx, gy, dbh, codes, status) %>%
+  mutate(
+    date = mdy(date),
+    dbh = as.numeric(dbh)/10
+  ) %>%
+  filter(gx < 300, between(gy, 300, 600))
+
+growth_scbi_sp <-
+  compute_growth(
+    census_1 = census_2013_scbi,
+    census_2 = census_2018_scbi %>% filter(!str_detect(codes, "R")),
+    id = "stemID"
+  ) %>%
+  mutate(sp = as.factor(sp),
+        basal_area = 0.0001 * pi * (dbh1 / 2)^2 )
+
+growth_scbi_can_pos <-
+  compute_growth(
+    census_1 = census_2013_scbi,
+    census_2 = census_2018_scbi %>% filter(!str_detect(codes, "R")),
+    id = "stemID"
+  ) %>%
+  left_join(sp_info, by = "sp") %>% 
+  mutate(
+    canopy_position = str_replace(canopy_position, ' ', '_'),
+    canopy_position = str_replace(canopy_position, ',', ''),
+    canopy_position = ifelse(is.na(canopy_position), 'shrub_layer', canopy_position),
+    sp = as.factor(canopy_position),
+    basal_area = 0.0001 * pi * (dbh1 / 2)^2 )
+
+study_region_scbi <- tibble(
+  x = c(0, 300, 300, 0, 0),
+  y = c(300, 300, 600, 600, 300)
+) %>%
+  sf_polygon()
+
+n_fold <- 4
+fold1 <- cbind(c(0, 150, 150, 0), c(300, 300, 450, 450))
+fold2 <- cbind(c(150, 300, 300, 150), c(300, 300, 450, 450))
+fold3 <- cbind(c(0, 150, 150, 0), c(450, 450, 600, 600))
+fold4 <- cbind(c(150, 300, 300, 150), c(450, 450, 600, 600))
+
+blocks_scbi <- bind_rows(
+  sf_polygon(fold1), sf_polygon(fold2), sf_polygon(fold3), 
+  sf_polygon(fold4)
+) %>%
+  mutate(folds = c(1:n_fold) %>% factor())
+
+# Associate each observation to a fold
+spatial_block_scbi <- 
+  spatialBlock(speciesData = growth_scbi, k = n_fold, 
+               selection = "systematic", blocks = blocks_scbi, 
+               showBlocks = FALSE, verbose = FALSE)
+
+growth_scbi_sp <- growth_scbi_sp %>%
+  mutate(foldID = spatial_block_scbi$foldID %>% factor())
+growth_scbi_can_pos <- growth_scbi_can_pos %>%
+  mutate(foldID = spatial_block_scbi$foldID %>% factor())
+
+comp_dist <- 7.5
+
+growth_scbi_sp <- growth_scbi_sp %>%
+    add_buffer_variable(size = comp_dist, region = study_region_scbi)
+growth_scbi_can_pos <- growth_scbi_can_pos %>%
+    add_buffer_variable(size = comp_dist, region = study_region_scbi)
+
+focal_vs_comp_sp <- growth_scbi_sp %>%
+    create_focal_vs_comp(comp_dist = comp_dist, 
+                         blocks = blocks_scbi, 
+                         id = "stemID", 
+                         comp_x_var = "basal_area") %>%
+    run_cv(comp_dist = comp_dist, blocks = blocks_scbi)
+
+focal_vs_comp_can_pos <- growth_scbi_can_pos %>%
+    create_focal_vs_comp(comp_dist = comp_dist, 
+                         blocks = blocks_scbi, 
+                         id = "stemID", 
+                         comp_x_var = "basal_area") %>%
+    run_cv(comp_dist = comp_dist, blocks = blocks_scbi)
+
+  focal_vs_comp_sp %>%
+    rmse(truth = growth, estimate = growth_hat) %>%
+    pull(.estimate)
+
+  focal_vs_comp_can_pos %>%
+    rmse(truth = growth, estimate = growth_hat) %>%
+    pull(.estimate)
+```
+
+We find that species identity has a lower RMSE, so does a better job. But we can still plot the competition posteriors for the canopoy position groupings.
+
+```{r can-pos-comp}
+
+  fit_mod_can_pos <- growth_scbi_can_pos %>%
+    create_focal_vs_comp(comp_dist = comp_dist, 
+                         blocks = blocks_scbi, 
+                         id = "stemID", 
+                         comp_x_var = "basal_area")  %>%
+    comp_bayes_lm(prior_param = NULL)
+
+  autoplot(fit_mod_can_pos, type = "competition")
+
+```
+
 # References

### 2021-08-06 paper/paper.tex
results for appedices 3+4 need to edit text

--- 2021-07-26 paper/paper.tex
@@
 
+We also load species information as it existed on GitHub on 2021/08/02,
+which includes family, genus, and species information; as well as
+classifications of the canopy position (canopy, canopy emergent,
+understory, shrub layer) and drought tolerance (intolerant, resistant)
+of the species.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{sp_info <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"SCBI_ForestGEO_sp_ecology.csv"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{select}\NormalTok{(}
+    \DataTypeTok{sp =}\NormalTok{ spcode, family, genus, species, canopy_position, }
+\NormalTok{    drought_tolerance}
+\NormalTok{  ) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{mutate}\NormalTok{()}
+\NormalTok{sp_info}
+\CommentTok{## # A tibble: 65 x 6}
+\CommentTok{##   sp    family        genus       species     canopy_position drought_toleran~}
+\CommentTok{##   <chr> <chr>         <chr>       <chr>       <chr>           <chr>           }
+\CommentTok{## 1 acne  Sapindaceae   Acer        negundo     understory      drought-resista~}
+\CommentTok{## 2 acpl  Sapindaceae   Acer        platanoides canopy          drought-resista~}
+\CommentTok{## 3 acru  Sapindaceae   Acer        rubrum      canopy          drought-resista~}
+\CommentTok{## 4 aial  Simaroubaceae Ailanthus   altissima   canopy          drought-resista~}
+\CommentTok{## 5 amar  Rosaceae      Amelanchier arborea     understory      drought-resista~}
+\CommentTok{## # ... with 60 more rows}
+\end{Highlighting}
+\end{Shaded}
+
+We join this species information to our \texttt{growth\_scbi} data
+frame.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{left_join}\NormalTok{(sp_info, }\DataTypeTok{by =} \StringTok{"sp"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\CommentTok{# }\AlertTok{TODO}\CommentTok{:}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{as.factor}\NormalTok{(sp))}
+\end{Highlighting}
+\end{Shaded}
+
+Furthermore, we compute two potential competitor explanatory variables
+\(x^{\text{comp}}_{ijk}\) from Equation \ref{eq:model}. First, the basal
+area of each tree as a function of its DBH in the earlier census.
+Second, the above ground biomass as estimated by allometric equations
+encoded in the \texttt{get\_biomass()} function from the \texttt{allodb}
+package \citep{allodb_package}; this function has DBH, species, and
+geographic coordinates as arguments.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{# Install development version of allodb using:}
+\CommentTok{# remotes::install_github("forestgeo/allodb")}
+\KeywordTok{library}\NormalTok{(allodb)}
+\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
+    \CommentTok{# Compute basal area:}
+    \DataTypeTok{basal_area =} \FloatTok{0.0001} \OperatorTok{*}\StringTok{ }\NormalTok{pi }\OperatorTok{*}\StringTok{ }\NormalTok{(dbh1 }\OperatorTok{/}\StringTok{ }\DecValTok{2}\NormalTok{)}\OperatorTok{^}\DecValTok{2}\NormalTok{,}
+    \CommentTok{# Compute above ground biomass:}
+    \DataTypeTok{agb =} \KeywordTok{get_biomass}\NormalTok{(}
+      \DataTypeTok{dbh =}\NormalTok{ dbh1,}
+      \DataTypeTok{genus =}\NormalTok{ genus,}
+      \DataTypeTok{species =}\NormalTok{ species,}
+      \DataTypeTok{coords =} \KeywordTok{c}\NormalTok{(}\OperatorTok{-}\FloatTok{78.2}\NormalTok{, }\FloatTok{38.9}\NormalTok{)}
+\NormalTok{    )}
+\NormalTok{  )}
+\end{Highlighting}
+\end{Shaded}
+
 \hypertarget{spatial-information}{%
@@
 \NormalTok{blocks_scbi <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}
-  \KeywordTok{sf_polygon}\NormalTok{(fold1), }\KeywordTok{sf_polygon}\NormalTok{(fold2), }\KeywordTok{sf_polygon}\NormalTok{(fold3), }
-  \KeywordTok{sf_polygon}\NormalTok{(fold4)}
+  \KeywordTok{sf_polygon}\NormalTok{(fold1), }\KeywordTok{sf_polygon}\NormalTok{(fold2), }
+  \KeywordTok{sf_polygon}\NormalTok{(fold3), }\KeywordTok{sf_polygon}\NormalTok{(fold4)}
 \NormalTok{) }\OperatorTok{%>%}
@@
 \NormalTok{spatial_block_scbi <-}\StringTok{ }
-\StringTok{  }\KeywordTok{spatialBlock}\NormalTok{(}\DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =}\NormalTok{ n_fold, }
+\StringTok{  }\KeywordTok{spatialBlock}\NormalTok{(}
+    \DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =}\NormalTok{ n_fold, }
     \DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }
-               \DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}\NormalTok{)}
+    \DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}
+\NormalTok{  )}
 
@@
 \NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{)}
+\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{, }\DataTypeTok{comp_x_var =} \StringTok{"basal_area"}\NormalTok{)}
 \NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}\StringTok{ }
@@
 \CommentTok{## # A tibble: 20 x 6}
-\CommentTok{##   focal_ID   dbh comp_ID  dist comp_sp comp_basal_area}
+\CommentTok{##   focal_ID   dbh comp_ID  dist comp_sp comp_x_var}
 \CommentTok{##      <dbl> <dbl>   <dbl> <dbl> <fct>        <dbl>}
@@
 \NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{growth_hat =} \KeywordTok{predict}\NormalTok{(comp_bayes_lm_scbi, }\DataTypeTok{newdata =}\NormalTok{ focal_vs_comp_scbi))}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{growth_hat =} \KeywordTok{predict}\NormalTok{(comp_bayes_lm_scbi, }
+                              \DataTypeTok{newdata =}\NormalTok{ focal_vs_comp_scbi))}
 \end{Highlighting}
@@
     \DataTypeTok{id =} \StringTok{"stemID"}
-\NormalTok{  )}
+\NormalTok{  ) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\CommentTok{# Compute basal area:}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{basal_area =} \FloatTok{0.0001} \OperatorTok{*}\StringTok{ }\NormalTok{pi }\OperatorTok{*}\StringTok{ }\NormalTok{(dbh1 }\OperatorTok{/}\StringTok{ }\DecValTok{2}\NormalTok{)}\OperatorTok{^}\DecValTok{2}\NormalTok{)}
 
@@
 \NormalTok{blocks_scbi <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}
-  \KeywordTok{sf_polygon}\NormalTok{(fold1), }\KeywordTok{sf_polygon}\NormalTok{(fold2), }
+  \KeywordTok{sf_polygon}\NormalTok{(fold1), }\KeywordTok{sf_polygon}\NormalTok{(fold2),}
   \KeywordTok{sf_polygon}\NormalTok{(fold3), }\KeywordTok{sf_polygon}\NormalTok{(fold4)}
@@
 \NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{)}
+\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{, }\DataTypeTok{comp_x_var =} \StringTok{"basal_area"}\NormalTok{)}
 
@@
 \NormalTok{  shuffle_RMSE[j] <-}\StringTok{  }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{growth_hat =} 
-             \KeywordTok{predict}\NormalTok{(comp_bayes_lm_scbi, focal_vs_comp_scbi)) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}
+      \DataTypeTok{growth_hat =} \KeywordTok{predict}\NormalTok{(comp_bayes_lm_scbi, focal_vs_comp_scbi)}
+\NormalTok{    ) }\OperatorTok{%>%}
 \StringTok{    }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
@@
 \begin{Highlighting}[]
-
 \NormalTok{census_}\DecValTok{2013}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem2.csv"}\NormalTok{) }\OperatorTok{%>%}
@@
 \StringTok{  }\CommentTok{# make all species the same, needs to be factor with at least two levels}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{factor}\NormalTok{(}\StringTok{'A'}\NormalTok{, }\DataTypeTok{levels =} \KeywordTok{c}\NormalTok{(}\StringTok{'A'}\NormalTok{, }\StringTok{'B'}\NormalTok{)) )}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{factor}\NormalTok{(}\StringTok{'A'}\NormalTok{, }\DataTypeTok{levels =} \KeywordTok{c}\NormalTok{(}\StringTok{'A'}\NormalTok{, }\StringTok{'B'}\NormalTok{)) ) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\CommentTok{# Compute basal area:}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{basal_area =} \FloatTok{0.0001} \OperatorTok{*}\StringTok{ }\NormalTok{pi }\OperatorTok{*}\StringTok{ }\NormalTok{(dbh1 }\OperatorTok{/}\StringTok{ }\DecValTok{2}\NormalTok{)}\OperatorTok{^}\DecValTok{2}\NormalTok{)}
 
@@
 
-\ControlFlowTok{for}\NormalTok{ (i }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(mult_dist_comp}\OperatorTok{$}\NormalTok{dist))}
-\NormalTok{\{}
+\ControlFlowTok{for}\NormalTok{ (i }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(mult_dist_comp}\OperatorTok{$}\NormalTok{dist))\{}
 \NormalTok{  comp_dist <-}\StringTok{ }\NormalTok{mult_dist_comp}\OperatorTok{$}\NormalTok{dist[i]}
@@
 \NormalTok{  focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{, }\DataTypeTok{comp_x_var =} \StringTok{"basal_area"}\NormalTok{) }\OperatorTok{%>%}
 \StringTok{    }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi)}
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{Figures/appendix2-code-1} \end{center}
-
 Here we see that a cut off distance of approximately 6 m minimizes RMSE.
+
+\hypertarget{compare-pred-var}{%
+\section{Appendix 3: Compare predictor
+variables}\label{compare-pred-var}}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{census_}\DecValTok{2013}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem2.csv"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
+    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date),}
+    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}
+\NormalTok{  ) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+
+\NormalTok{census_}\DecValTok{2018}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem3.csv"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
+    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date),}
+    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}
+\NormalTok{  ) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+
+\NormalTok{growth_scbi <-}
+\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(}
+    \DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi,}
+    \DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)),}
+    \DataTypeTok{id =} \StringTok{"stemID"}
+\NormalTok{  ) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{left_join}\NormalTok{(sp_info, }\DataTypeTok{by =} \StringTok{"sp"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{as.factor}\NormalTok{(sp),}
+        \DataTypeTok{basal_area =} \FloatTok{0.0001} \OperatorTok{*}\StringTok{ }\NormalTok{pi }\OperatorTok{*}\StringTok{ }\NormalTok{(dbh1 }\OperatorTok{/}\StringTok{ }\DecValTok{2}\NormalTok{)}\OperatorTok{^}\DecValTok{2}\NormalTok{,}
+         \DataTypeTok{agb =} \KeywordTok{get_biomass}\NormalTok{(}
+                  \DataTypeTok{dbh =}\NormalTok{ dbh1,}
+                  \DataTypeTok{genus =}\NormalTok{ genus,}
+                  \DataTypeTok{species =}\NormalTok{ species,}
+                  \DataTypeTok{coords =} \KeywordTok{c}\NormalTok{(}\OperatorTok{-}\FloatTok{78.2}\NormalTok{, }\FloatTok{38.9}\NormalTok{)}
+\NormalTok{    ))}
+
+\NormalTok{study_region_scbi <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}
+  \DataTypeTok{x =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{),}
+  \DataTypeTok{y =} \KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{300}\NormalTok{)}
+\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{sf_polygon}\NormalTok{()}
+
+\NormalTok{n_fold <-}\StringTok{ }\DecValTok{4}
+\NormalTok{fold1 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{0}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{))}
+\NormalTok{fold2 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{150}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{))}
+\NormalTok{fold3 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{0}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+\NormalTok{fold4 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{150}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+
+\NormalTok{blocks_scbi <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}
+  \KeywordTok{sf_polygon}\NormalTok{(fold1), }\KeywordTok{sf_polygon}\NormalTok{(fold2), }\KeywordTok{sf_polygon}\NormalTok{(fold3), }
+  \KeywordTok{sf_polygon}\NormalTok{(fold4)}
+\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\NormalTok{n_fold) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
+
+\CommentTok{# Associate each observation to a fold}
+\NormalTok{spatial_block_scbi <-}\StringTok{ }
+\StringTok{  }\KeywordTok{spatialBlock}\NormalTok{(}\DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =}\NormalTok{ n_fold, }
+               \DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }
+               \DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}\NormalTok{)}
+
+\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ spatial_block_scbi}\OperatorTok{$}\NormalTok{foldID }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
+
+\NormalTok{comp_dist <-}\StringTok{ }\FloatTok{7.5}
+
+\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{size =}\NormalTok{ comp_dist, }\DataTypeTok{region =}\NormalTok{ study_region_scbi)}
+
+\NormalTok{focal_vs_comp_ba <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }
+                         \DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }
+                         \DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{, }
+                         \DataTypeTok{comp_x_var =} \StringTok{"basal_area"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi)}
+
+\NormalTok{focal_vs_comp_agb <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }
+                         \DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }
+                         \DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{, }
+                         \DataTypeTok{comp_x_var =} \StringTok{"agb"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi)}
+
+\NormalTok{  focal_vs_comp_ba }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{pull}\NormalTok{(.estimate)}
+\CommentTok{## [1] 0.14}
+
+\NormalTok{  focal_vs_comp_agb }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{pull}\NormalTok{(.estimate)}
+\CommentTok{## [1] 1.83}
+\end{Highlighting}
+\end{Shaded}
+
+\hypertarget{compare-group-var}{%
+\section{Appendix 4: Compare grouping
+variables}\label{compare-group-var}}
+
+Creative use of the grouping variable allows for tests of what best
+explains competitive interactions. Here we compare whether species
+idenity or canopy position of the species best explains.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+
+\NormalTok{census_}\DecValTok{2013}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem2.csv"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
+    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date),}
+    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}
+\NormalTok{  ) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+
+\NormalTok{census_}\DecValTok{2018}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem3.csv"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
+    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date),}
+    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}
+\NormalTok{  ) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+
+\NormalTok{growth_scbi_sp <-}
+\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(}
+    \DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi,}
+    \DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)),}
+    \DataTypeTok{id =} \StringTok{"stemID"}
+\NormalTok{  ) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{as.factor}\NormalTok{(sp),}
+        \DataTypeTok{basal_area =} \FloatTok{0.0001} \OperatorTok{*}\StringTok{ }\NormalTok{pi }\OperatorTok{*}\StringTok{ }\NormalTok{(dbh1 }\OperatorTok{/}\StringTok{ }\DecValTok{2}\NormalTok{)}\OperatorTok{^}\DecValTok{2}\NormalTok{ )}
+
+\NormalTok{growth_scbi_can_pos <-}
+\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(}
+    \DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi,}
+    \DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)),}
+    \DataTypeTok{id =} \StringTok{"stemID"}
+\NormalTok{  ) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{left_join}\NormalTok{(sp_info, }\DataTypeTok{by =} \StringTok{"sp"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
+    \DataTypeTok{canopy_position =} \KeywordTok{str_replace}\NormalTok{(canopy_position, }\StringTok{' '}\NormalTok{, }\StringTok{'_'}\NormalTok{),}
+    \DataTypeTok{canopy_position =} \KeywordTok{str_replace}\NormalTok{(canopy_position, }\StringTok{','}\NormalTok{, }\StringTok{''}\NormalTok{),}
+    \DataTypeTok{canopy_position =} \KeywordTok{ifelse}\NormalTok{(}\KeywordTok{is.na}\NormalTok{(canopy_position), }\StringTok{'shrub_layer'}\NormalTok{, canopy_position),}
+    \DataTypeTok{sp =} \KeywordTok{as.factor}\NormalTok{(canopy_position),}
+    \DataTypeTok{basal_area =} \FloatTok{0.0001} \OperatorTok{*}\StringTok{ }\NormalTok{pi }\OperatorTok{*}\StringTok{ }\NormalTok{(dbh1 }\OperatorTok{/}\StringTok{ }\DecValTok{2}\NormalTok{)}\OperatorTok{^}\DecValTok{2}\NormalTok{ )}
+
+\NormalTok{study_region_scbi <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}
+  \DataTypeTok{x =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{),}
+  \DataTypeTok{y =} \KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{300}\NormalTok{)}
+\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{sf_polygon}\NormalTok{()}
+
+\NormalTok{n_fold <-}\StringTok{ }\DecValTok{4}
+\NormalTok{fold1 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{0}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{))}
+\NormalTok{fold2 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{150}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{))}
+\NormalTok{fold3 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{0}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+\NormalTok{fold4 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{150}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+
+\NormalTok{blocks_scbi <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}
+  \KeywordTok{sf_polygon}\NormalTok{(fold1), }\KeywordTok{sf_polygon}\NormalTok{(fold2), }\KeywordTok{sf_polygon}\NormalTok{(fold3), }
+  \KeywordTok{sf_polygon}\NormalTok{(fold4)}
+\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\NormalTok{n_fold) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
+
+\CommentTok{# Associate each observation to a fold}
+\NormalTok{spatial_block_scbi <-}\StringTok{ }
+\StringTok{  }\KeywordTok{spatialBlock}\NormalTok{(}\DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =}\NormalTok{ n_fold, }
+               \DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }
+               \DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}\NormalTok{)}
+
+\NormalTok{growth_scbi_sp <-}\StringTok{ }\NormalTok{growth_scbi_sp }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ spatial_block_scbi}\OperatorTok{$}\NormalTok{foldID }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
+\NormalTok{growth_scbi_can_pos <-}\StringTok{ }\NormalTok{growth_scbi_can_pos }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ spatial_block_scbi}\OperatorTok{$}\NormalTok{foldID }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
+
+\NormalTok{comp_dist <-}\StringTok{ }\FloatTok{7.5}
+
+\NormalTok{growth_scbi_sp <-}\StringTok{ }\NormalTok{growth_scbi_sp }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{size =}\NormalTok{ comp_dist, }\DataTypeTok{region =}\NormalTok{ study_region_scbi)}
+\NormalTok{growth_scbi_can_pos <-}\StringTok{ }\NormalTok{growth_scbi_can_pos }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{size =}\NormalTok{ comp_dist, }\DataTypeTok{region =}\NormalTok{ study_region_scbi)}
+
+\NormalTok{focal_vs_comp_sp <-}\StringTok{ }\NormalTok{growth_scbi_sp }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }
+                         \DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }
+                         \DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{, }
+                         \DataTypeTok{comp_x_var =} \StringTok{"basal_area"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi)}
+
+\NormalTok{focal_vs_comp_can_pos <-}\StringTok{ }\NormalTok{growth_scbi_can_pos }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }
+                         \DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }
+                         \DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{, }
+                         \DataTypeTok{comp_x_var =} \StringTok{"basal_area"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi)}
+
+\NormalTok{  focal_vs_comp_sp }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{pull}\NormalTok{(.estimate)}
+\CommentTok{## [1] 0.14}
+
+\NormalTok{  focal_vs_comp_can_pos }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{pull}\NormalTok{(.estimate)}
+\CommentTok{## [1] 0.144}
+\end{Highlighting}
+\end{Shaded}
+
+We find that species identity has a lower RMSE, so does a better job.
+But we can still plot the competition posteriors for the canopoy
+position groupings.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+
+\NormalTok{  fit_mod_can_pos <-}\StringTok{ }\NormalTok{growth_scbi_can_pos }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }
+                         \DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }
+                         \DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{, }
+                         \DataTypeTok{comp_x_var =} \StringTok{"basal_area"}\NormalTok{)  }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{comp_bayes_lm}\NormalTok{(}\DataTypeTok{prior_param =} \OtherTok{NULL}\NormalTok{)}
+
+  \KeywordTok{autoplot}\NormalTok{(fit_mod_can_pos, }\DataTypeTok{type =} \StringTok{"competition"}\NormalTok{)}
+
+\end{Highlighting}
+\end{Shaded}
+
+\begin{center}\includegraphics[width=1\linewidth]{Figures/can-pos-comp-1} \end{center}
 

### 2021-08-07 paper/paper.Rmd
my edits are done!

@@
 
-We present a case-study of `forestecology`'s functionality on data from the Smithsonian Conservation Biology Institute (SCBI) large forest dynamics plot in Front Royal, VA, USA, part of the ForestGEO global network of research sites [@bourg_initial_2013, @andersonteixeira_ctfs-forestgeo_2015].
+We present a case-study of `forestecology`'s functionality on data from the Smithsonian Conservation Biology Institute (SCBI) large forest dynamics plot in Front Royal, VA, USA, part of the ForestGEO global network of research sites [@bourg_initial_2013, @andersonteixeira_ctfs-forestgeo_2015, @davies_forestgeo_2021].
 The 25.6 ha (640 x 400 m) plot is located at the intersection of three of the major physiographic provinces of the eastern US---the Blue Ridge, Ridge and Valley, and Piedmont provinces---and is adjacent to the northern end of Shenandoah National Park.
@@
 
-We also load species information as it existed on GitHub on 2021/08/02, which includes family, genus, and species information; as well as classifications of the canopy position (canopy, canopy emergent, understory, shrub layer) and drought tolerance (intolerant, resistant) of the species.
+We also load species information as it existed on GitHub on 2021/08/02, which includes family, genus, and species information; as well as classifications of the canopy position (canopy, canopy emergent, understory, shrub layer), drought tolerance (intolerant, resistant) and other characteristics of the species.
 
@@
 We visualize the spatial distribution of these trees in Figure \ref{fig:scbi-focal-vs-comp-map}.
+
+Here we use basal area as the continuous predictor variable but the package is flexible to allow the user to specify any continuous predictor variable (basal area, biomass, tree height, a soil nutrient value, \dots).
+The package can also be used to compare predictor variables and see which best explains tree growth, see Appendix 3 \ref{compare-pred-var} for an example comparing basal area and above ground biomass.
+Similarly, the package can use any categorical variable as a predictor and compare between different categorical variables.
+For example in @allen_permutation_2020 we compare grouping individuals based on species, family, and based on trait-based groups.
+In Appendix 4 \ref{compare-group-var} we give another example and compare grouping individuals by species or by potential canopy position (canopy, understory, shrub layer).
 
@@
 
+In the above code we use the basal area of an individual as a continuous predictor variable.
+But the package allows the user to specify any predictor variable in the `comp_x_var` argument of `create_focal_vs_comp` function.
+Here we use the cross-validated model comparison to see which of two possible predictor variables, basal area or above ground biomass, best explains growth.
+
 ```{r compar-pred-var}
@@
 
+Here we see that basal area does a much better job than above ground biomass.
+
 # Appendix 4: Compare grouping variables {#compare-group-var}
 
-Creative use of the grouping variable allows for tests of what best explains competitive interactions. Here we compare whether species idenity or canopy position of the species best explains.
+The package also allows the user to specify the categorical predictor grouping variable.
+Here we compare two different such variables: species and the potential canopy position of that species.
+If we had individual-level crown classes (@smith_practice_1986 dominant, codominant, intermediate and suppressed) that could also be used.
 
@@
 
-We find that species identity has a lower RMSE, so does a better job. But we can still plot the competition posteriors for the canopoy position groupings.
+We find that species identity has a lower RMSE, so does a better job. But we can still plot the competition posteriors for the canopy position groupings.
 
@@
 
+Unsurprisingly we see that canopy and canopy emergent competitors generally have negative effects on their neighbors, while shrubs and understory competitors have neutral or even positive effects.
+
 # References

### 2021-08-07 paper/paper.tex
my edits are done!

@@
 global network of research sites
-\citep[\citet{andersonteixeira_ctfs-forestgeo_2015}]{bourg_initial_2013}.
-The 25.6 ha (640 x 400 m) plot is located at the intersection of three
-of the major physiographic provinces of the eastern US---the Blue Ridge,
-Ridge and Valley, and Piedmont provinces---and is adjacent to the
-northern end of Shenandoah National Park.
+\citep[\citet{andersonteixeira_ctfs-forestgeo_2015},
+\citet{davies_forestgeo_2021}]{bourg_initial_2013}. The 25.6 ha (640 x
+400 m) plot is located at the intersection of three of the major
+physiographic provinces of the eastern US---the Blue Ridge, Ridge and
+Valley, and Piedmont provinces---and is adjacent to the northern end of
+Shenandoah National Park.
 
@@
 classifications of the canopy position (canopy, canopy emergent,
-understory, shrub layer) and drought tolerance (intolerant, resistant)
-of the species.
+understory, shrub layer), drought tolerance (intolerant, resistant) and
+other characteristics of the species.
 
@@
 trees in Figure \ref{fig:scbi-focal-vs-comp-map}.
+
+Here we use basal area as the continuous predictor variable but the
+package is flexible to allow the user to specify any continuous
+predictor variable (basal area, biomass, tree height, a soil nutrient
+value, \dots). The package can also be used to compare predictor
+variables and see which best explains tree growth, see Appendix 3
+\ref{compare-pred-var} for an example comparing basal area and above
+ground biomass. Similarly, the package can use any categorical variable
+as a predictor and compare between different categorical variables. For
+example in \citet{allen_permutation_2020} we compare grouping
+individuals based on species, family, and based on trait-based groups.
+In Appendix 4 \ref{compare-group-var} we give another example and
+compare grouping individuals by species or by potential canopy position
+(canopy, understory, shrub layer).
 
@@
 
+In the above code we use the basal area of an individual as a continuous
+predictor variable. But the package allows the user to specify any
+predictor variable in the \texttt{comp\_x\_var} argument of
+\texttt{create\_focal\_vs\_comp} function. Here we use the
+cross-validated model comparison to see which of two possible predictor
+variables, basal area or above ground biomass, best explains growth.
+
 \begin{Shaded}
@@
 
+Here we see that basal area does a much better job than above ground
+biomass.
+
 \hypertarget{compare-group-var}{%
@@
 
-Creative use of the grouping variable allows for tests of what best
-explains competitive interactions. Here we compare whether species
-idenity or canopy position of the species best explains.
+The package also allows the user to specify the categorical predictor
+grouping variable. Here we compare two different such variables: species
+and the potential canopy position of that species. If we had
+individual-level crown classes (\citet{smith_practice_1986} dominant,
+codominant, intermediate and suppressed) that could also be used.
 
@@
 We find that species identity has a lower RMSE, so does a better job.
-But we can still plot the competition posteriors for the canopoy
-position groupings.
+But we can still plot the competition posteriors for the canopy position
+groupings.
 
@@
 
+Unsurprisingly we see that canopy and canopy emergent competitors
+generally have negative effects on their neighbors, while shrubs and
+understory competitors have neutral or even positive effects.
+
 \bibliographystyle{agsm}

### 2021-08-09 paper/paper.Rmd
set eval = TRUE + ran styler

@@
   scale_size_binned(limits = c(0.1, 1)) +
-  labs(size = expression(paste(Growth, " (cm ",y^{-1},')')) )
+  labs(size = expression(paste(Growth, " (cm ", y^{-1}, ")")))
 ```
@@
 
-```{r internal, eval = FALSE, echo = FALSE}
+```{r internal, echo = FALSE}
 # Internal: after removing trees in buffer and have NA for growth, there is one
@@
 
-```{r appendix1-code, eval = FALSE}
+```{r appendix1-code, eval=FALSE}
 library(tidyverse)
@@
     date = mdy(date),
-    dbh = as.numeric(dbh)/10
+    dbh = as.numeric(dbh) / 10
   ) %>%
@@
 observed_RMSE <- focal_vs_comp_scbi %>%
-  mutate(growth_hat = 
-           predict(comp_bayes_lm_scbi, focal_vs_comp_scbi)) %>%
+  mutate(
+    growth_hat =
+      predict(comp_bayes_lm_scbi, focal_vs_comp_scbi)
+  ) %>%
   rmse(truth = growth, estimate = growth_hat) %>%
@@
 shuffle_RMSE <- numeric(length = num_shuffle)
-for(j in 1:num_shuffle){
+for (j in 1:num_shuffle) {
   # Fit model (compute posterior parameters) with shuffling
@@
 # Compute num_shuffle permutation test statistics
-for(j in 1:num_shuffle){
+for (j in 1:num_shuffle) {
   # Compute and save RMSE
   shuffle_RMSE_CV[j] <- focal_vs_comp_scbi %>%
-    run_cv(comp_dist = comp_dist, blocks = blocks_scbi, 
-           run_shuffle = TRUE) %>%
+    run_cv(
+      comp_dist = comp_dist, blocks = blocks_scbi,
+      run_shuffle = TRUE
+    ) %>%
     rmse(truth = growth, estimate = growth_hat) %>%
@@
   model_comp_tbl %>%
-    select(run_time, observed = observed_RMSE, 
-           shuffle = shuffle_RMSE) %>% 
+    select(run_time,
+      observed = observed_RMSE,
+      shuffle = shuffle_RMSE
+    ) %>%
     mutate(CV = FALSE),
   model_comp_tbl %>%
-    select(run_time, observed = observed_RMSE_CV, 
-           shuffle = shuffle_RMSE_CV) %>% 
+    select(run_time,
+      observed = observed_RMSE_CV,
+      shuffle = shuffle_RMSE_CV
+    ) %>%
     mutate(CV = TRUE)
@@
 cv_plot <- ggplot() +
-  geom_vline(data = model_comp_observed, 
+  geom_vline(
+    data = model_comp_observed,
     aes(xintercept = RMSE, col = CV),
-             linetype = "dashed", show.legend = F) +
-  geom_histogram(data = model_comp_shuffle, 
-                 aes(x = RMSE, fill = CV), bins = 50) +
+    linetype = "dashed", show.legend = F
+  ) +
+  geom_histogram(
+    data = model_comp_shuffle,
+    aes(x = RMSE, fill = CV), bins = 50
+  ) +
   labs(
     fill = "Cross-validated?",
-    x = expression(paste("RMSE (cm ", y^{-1}, ")"))
+    x = expression(paste("RMSE (cm ", y^{
+      -1
+    }, ")"))
   ) +
@@
   str_c(".pdf") %>%
-  ggsave(plot = cv_plot, width = 16/2, height = 9/2)
+  ggsave(plot = cv_plot, width = 16 / 2, height = 9 / 2)
 ```
@@
 
-```{r appendix2-code, eval = FALSE}
+```{r appendix2-code}
 census_2013_scbi <- read_csv("scbi.stem2.csv") %>%
@@
   # make all species the same, needs to be factor with at least two levels
-  mutate(sp = factor('A', levels = c('A', 'B')) ) %>% 
+  mutate(sp = factor("A", levels = c("A", "B"))) %>%
   # Compute basal area:
@@
 spatial_block_scbi <-
-  spatialBlock(speciesData = growth_scbi, k = n_fold, 
+  spatialBlock(
+    speciesData = growth_scbi, k = n_fold,
     selection = "systematic", blocks = blocks_scbi,
-               showBlocks = FALSE, verbose = FALSE)
+    showBlocks = FALSE, verbose = FALSE
+  )
 
@@
 mult_dist_comp <- tibble(
-  dist = c(5 ,6.25,  7.5, 8.75,  10),
+  dist = c(5, 6.25, 7.5, 8.75, 10),
   rmse = 0
@@
 
-for (i in 1:length(mult_dist_comp$dist)){
+for (i in 1:length(mult_dist_comp$dist)) {
   comp_dist <- mult_dist_comp$dist[i]
@@
 mult_dist_comp %>%
-  ggplot(aes(dist,rmse)) +
+  ggplot(aes(dist, rmse)) +
   geom_point() +
-  xlab('Distance (m)') +
-  ylab('RMSE (cm)')
+  xlab("Distance (m)") +
+  ylab("RMSE (cm)")
 ```
@@
   left_join(sp_info, by = "sp") %>%
-  mutate(sp = as.factor(sp),
+  mutate(
+    sp = as.factor(sp),
     basal_area = 0.0001 * pi * (dbh1 / 2)^2,
@@
       coords = c(-78.2, 38.9)
-    ))
+    )
+  )
 
@@
 focal_vs_comp_ba <- growth_scbi %>%
-    create_focal_vs_comp(comp_dist = comp_dist, 
+  create_focal_vs_comp(
+    comp_dist = comp_dist,
     blocks = blocks_scbi,
     id = "stemID",
-                         comp_x_var = "basal_area") %>%
+    comp_x_var = "basal_area"
+  ) %>%
   run_cv(comp_dist = comp_dist, blocks = blocks_scbi)
@@
 focal_vs_comp_agb <- growth_scbi %>%
-    create_focal_vs_comp(comp_dist = comp_dist, 
+  create_focal_vs_comp(
+    comp_dist = comp_dist,
     blocks = blocks_scbi,
     id = "stemID",
-                         comp_x_var = "agb") %>%
+    comp_x_var = "agb"
+  ) %>%
   run_cv(comp_dist = comp_dist, blocks = blocks_scbi)
@@
 ```{r compare-group-var}
-
 census_2013_scbi <- read_csv("scbi.stem2.csv") %>%
@@
   ) %>%
-  mutate(sp = as.factor(sp),
-        basal_area = 0.0001 * pi * (dbh1 / 2)^2 )
+  mutate(
+    sp = as.factor(sp),
+    basal_area = 0.0001 * pi * (dbh1 / 2)^2
+  )
 
@@
   mutate(
-    canopy_position = str_replace(canopy_position, ' ', '_'),
-    canopy_position = str_replace(canopy_position, ',', ''),
-    canopy_position = ifelse(is.na(canopy_position), 'shrub_layer', canopy_position),
+    canopy_position = str_replace(canopy_position, " ", "_"),
+    canopy_position = str_replace(canopy_position, ",", ""),
+    canopy_position = ifelse(is.na(canopy_position), "shrub_layer", canopy_position),
     sp = as.factor(canopy_position),
-    basal_area = 0.0001 * pi * (dbh1 / 2)^2 )
+    basal_area = 0.0001 * pi * (dbh1 / 2)^2
+  )
 
@@
 focal_vs_comp_sp <- growth_scbi_sp %>%
-    create_focal_vs_comp(comp_dist = comp_dist, 
+  create_focal_vs_comp(
+    comp_dist = comp_dist,
     blocks = blocks_scbi,
     id = "stemID",
-                         comp_x_var = "basal_area") %>%
+    comp_x_var = "basal_area"
+  ) %>%
   run_cv(comp_dist = comp_dist, blocks = blocks_scbi)
@@
 focal_vs_comp_can_pos <- growth_scbi_can_pos %>%
-    create_focal_vs_comp(comp_dist = comp_dist, 
+  create_focal_vs_comp(
+    comp_dist = comp_dist,
     blocks = blocks_scbi,
     id = "stemID",
-                         comp_x_var = "basal_area") %>%
+    comp_x_var = "basal_area"
+  ) %>%
   run_cv(comp_dist = comp_dist, blocks = blocks_scbi)
@@
 ```{r can-pos-comp}
-
 fit_mod_can_pos <- growth_scbi_can_pos %>%
-    create_focal_vs_comp(comp_dist = comp_dist, 
+  create_focal_vs_comp(
+    comp_dist = comp_dist,
     blocks = blocks_scbi,
     id = "stemID",
-                         comp_x_var = "basal_area")  %>%
+    comp_x_var = "basal_area"
+  ) %>%
   comp_bayes_lm(prior_param = NULL)
-
 autoplot(fit_mod_can_pos, type = "competition")
-
 ```

### 2021-08-09 paper/paper.Rmd
ran styler

@@
   comment = "##",
-  tidy.opts = list(width.cutoff=72),
+  tidy.opts = list(width.cutoff = 72),
   tidy = TRUE
@@
     # Convert dbh to be in cm
-    dbh = as.numeric(dbh)/10
+    dbh = as.numeric(dbh) / 10
   ) %>%
@@
   geom_sf(data = blocks_scbi, fill = "transparent", linetype = "dashed") +
-  geom_sf_text(data = growth_scbi %>% sample_n(1000), 
-               aes(label = foldID, col = buffer))
+  geom_sf_text(
+    data = growth_scbi %>% sample_n(1000),
+    aes(label = foldID, col = buffer)
+  )
 ```
@@
   mutate(growth_hat = predict(comp_bayes_lm_scbi,
-                              newdata = focal_vs_comp_scbi))
+    newdata = focal_vs_comp_scbi
+  ))
 ```
@@
 
-plot1 <- autoplot(comp_bayes_lm_scbi, type = "intercepts", 
-                  sp_to_plot = sp_to_plot)
-plot2 <- autoplot(comp_bayes_lm_scbi, type = "dbh_slopes", 
-                  sp_to_plot = sp_to_plot)
-plot3 <- autoplot(comp_bayes_lm_scbi, type = "competition", 
-                  sp_to_plot = sp_to_plot)
+plot1 <- autoplot(comp_bayes_lm_scbi,
+  type = "intercepts",
+  sp_to_plot = sp_to_plot
+)
+plot2 <- autoplot(comp_bayes_lm_scbi,
+  type = "dbh_slopes",
+  sp_to_plot = sp_to_plot
+)
+plot3 <- autoplot(comp_bayes_lm_scbi,
+  type = "competition",
+  sp_to_plot = sp_to_plot
+)
 
@@
   mutate(growth_hat_shuffle = predict(comp_bayes_lm_scbi_shuffle,
-                                 newdata = focal_vs_comp_scbi))
+    newdata = focal_vs_comp_scbi
+  ))
 ```

### 2021-08-10 paper/paper.Rmd
Reknit entire paper from scratch

@@
 
-```{r appendix1-code, eval=FALSE}
+```{r appendix1-code}
 library(tidyverse)

### 2021-08-10 paper/paper.tex
Reknit entire paper from scratch

--- 2021-08-04 paper/paper.tex
@@
     \CommentTok{# Convert dbh to be in cm}
-    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}
+    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh) }\OperatorTok{/}\StringTok{ }\DecValTok{10}
 \NormalTok{  ) }\OperatorTok{%>%}
@@
     \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date),}
-    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}
+    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh) }\OperatorTok{/}\StringTok{ }\DecValTok{10}
 \NormalTok{  ) }\OperatorTok{%>%}
@@
 \KeywordTok{library}\NormalTok{(allodb)}
-\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}\StringTok{ }
+\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
 \StringTok{  }\KeywordTok{mutate}\NormalTok{(}
@@
 \begin{Highlighting}[]
-\CommentTok{# }\AlertTok{TODO}\CommentTok{:}
-\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{comp_x_var =}\NormalTok{ basal_area)}
-
 \NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{)}
+\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{,}
+        \DataTypeTok{comp_x_var =} \StringTok{"basal_area"}\NormalTok{)}
 \NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
@@
 \end{Shaded}
+
+\begin{verbatim}
+## Simple feature collection with 1 feature and 17 fields
+## Geometry type: POINT
+## Dimension:     XY
+## Bounding box:  xmin: 250 ymin: 582 xmax: 250 ymax: 582
+## CRS:           NA
+## # A tibble: 1 x 18
+##   stemID sp     dbh1 codes1 status  dbh2 codes2 growth
+## *  <dbl> <fct> <dbl> <chr>  <chr>  <dbl> <chr>   <dbl>
+## 1  18823 quve   68.8 NULL   A       69.9 NULL    0.215
+## # ... with 10 more variables: geometry <POINT>, family <chr>,
+## #   genus <chr>, species <chr>, canopy_position <chr>,
+## #   drought_tolerance <chr>, basal_area <dbl>, agb <dbl>, buffer <lgl>,
+## #   foldID <fct>
+\end{verbatim}
+
+\begin{center}\includegraphics[width=1\linewidth]{Figures/internal-1} \end{center}
 
@@
 \NormalTok{growth_scbi <-}\StringTok{ }\KeywordTok{compute_growth}\NormalTok{(}\DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi, }\DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)), }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{)}
+\StringTok{    }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)), }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\CommentTok{# Compute basal area:}
+\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{basal_area =} \FloatTok{1e-04} \OperatorTok{*}\StringTok{ }\NormalTok{pi }\OperatorTok{*}\StringTok{ }\NormalTok{(dbh1}\OperatorTok{/}\DecValTok{2}\NormalTok{)}\OperatorTok{^}\DecValTok{2}\NormalTok{)}
 
@@
 \NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{)}
+\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{,}
+        \DataTypeTok{comp_x_var =} \StringTok{"basal_area"}\NormalTok{)}
 
@@
 \NormalTok{\}}
+\CommentTok{## [1] "Shuffle with permutation 1 at 2021-08-09 18:21:51"}
+\CommentTok{## [1] "Shuffle with permutation 2 at 2021-08-09 18:26:48"}
+\CommentTok{## [1] "Shuffle with permutation 3 at 2021-08-09 18:31:46"}
+\CommentTok{## [1] "Shuffle with permutation 4 at 2021-08-09 18:36:41"}
+\CommentTok{## [1] "Shuffle with permutation 5 at 2021-08-09 18:41:31"}
+\CommentTok{## [1] "Shuffle with permutation 6 at 2021-08-09 18:46:03"}
+\CommentTok{## [1] "Shuffle with permutation 7 at 2021-08-09 18:50:32"}
+\CommentTok{## [1] "Shuffle with permutation 8 at 2021-08-09 18:55:09"}
+\CommentTok{## [1] "Shuffle with permutation 9 at 2021-08-09 18:59:38"}
+\CommentTok{## [1] "Shuffle with permutation 10 at 2021-08-09 19:04:06"}
+\CommentTok{## [1] "Shuffle with permutation 11 at 2021-08-09 19:08:36"}
+\CommentTok{## [1] "Shuffle with permutation 12 at 2021-08-09 19:13:01"}
+\CommentTok{## [1] "Shuffle with permutation 13 at 2021-08-09 19:17:25"}
+\CommentTok{## [1] "Shuffle with permutation 14 at 2021-08-09 19:21:47"}
+\CommentTok{## [1] "Shuffle with permutation 15 at 2021-08-09 19:26:20"}
+\CommentTok{## [1] "Shuffle with permutation 16 at 2021-08-09 19:30:42"}
+\CommentTok{## [1] "Shuffle with permutation 17 at 2021-08-09 19:35:05"}
+\CommentTok{## [1] "Shuffle with permutation 18 at 2021-08-09 19:39:28"}
+\CommentTok{## [1] "Shuffle with permutation 19 at 2021-08-09 19:43:51"}
+\CommentTok{## [1] "Shuffle with permutation 20 at 2021-08-09 19:48:14"}
+\CommentTok{## [1] "Shuffle with permutation 21 at 2021-08-09 19:52:37"}
+\CommentTok{## [1] "Shuffle with permutation 22 at 2021-08-09 19:57:01"}
+\CommentTok{## [1] "Shuffle with permutation 23 at 2021-08-09 20:01:29"}
+\CommentTok{## [1] "Shuffle with permutation 24 at 2021-08-09 20:06:09"}
+\CommentTok{## [1] "Shuffle with permutation 25 at 2021-08-09 20:10:42"}
+\CommentTok{## [1] "Shuffle with permutation 26 at 2021-08-09 20:15:06"}
+\CommentTok{## [1] "Shuffle with permutation 27 at 2021-08-09 20:19:29"}
+\CommentTok{## [1] "Shuffle with permutation 28 at 2021-08-09 20:23:52"}
+\CommentTok{## [1] "Shuffle with permutation 29 at 2021-08-09 20:28:15"}
+\CommentTok{## [1] "Shuffle with permutation 30 at 2021-08-09 20:32:37"}
+\CommentTok{## [1] "Shuffle with permutation 31 at 2021-08-09 20:36:59"}
+\CommentTok{## [1] "Shuffle with permutation 32 at 2021-08-09 20:41:21"}
+\CommentTok{## [1] "Shuffle with permutation 33 at 2021-08-09 20:45:43"}
+\CommentTok{## [1] "Shuffle with permutation 34 at 2021-08-09 20:50:06"}
+\CommentTok{## [1] "Shuffle with permutation 35 at 2021-08-09 20:54:29"}
+\CommentTok{## [1] "Shuffle with permutation 36 at 2021-08-09 20:58:51"}
+\CommentTok{## [1] "Shuffle with permutation 37 at 2021-08-09 21:03:16"}
+\CommentTok{## [1] "Shuffle with permutation 38 at 2021-08-09 21:07:38"}
+\CommentTok{## [1] "Shuffle with permutation 39 at 2021-08-09 21:12:01"}
+\CommentTok{## [1] "Shuffle with permutation 40 at 2021-08-09 21:16:24"}
+\CommentTok{## [1] "Shuffle with permutation 41 at 2021-08-09 21:20:47"}
+\CommentTok{## [1] "Shuffle with permutation 42 at 2021-08-09 21:25:11"}
+\CommentTok{## [1] "Shuffle with permutation 43 at 2021-08-09 21:29:34"}
+\CommentTok{## [1] "Shuffle with permutation 44 at 2021-08-09 21:33:57"}
+\CommentTok{## [1] "Shuffle with permutation 45 at 2021-08-09 21:38:20"}
+\CommentTok{## [1] "Shuffle with permutation 46 at 2021-08-09 21:42:42"}
+\CommentTok{## [1] "Shuffle with permutation 47 at 2021-08-09 21:47:05"}
+\CommentTok{## [1] "Shuffle with permutation 48 at 2021-08-09 21:51:27"}
+\CommentTok{## [1] "Shuffle with permutation 49 at 2021-08-09 21:56:05"}
 
@@
 \end{Shaded}
+
+\begin{center}\includegraphics[width=1\linewidth]{Figures/appendix1-code-1} \end{center}
 
@@
 \StringTok{    }\CommentTok{# levels}
-\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{factor}\NormalTok{(}\StringTok{"A"}\NormalTok{, }\DataTypeTok{levels =} \KeywordTok{c}\NormalTok{(}\StringTok{"A"}\NormalTok{, }\StringTok{"B"}\NormalTok{)))}
+\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{factor}\NormalTok{(}\StringTok{"A"}\NormalTok{, }\DataTypeTok{levels =} \KeywordTok{c}\NormalTok{(}\StringTok{"A"}\NormalTok{, }\StringTok{"B"}\NormalTok{))) }\OperatorTok{%>%}
+\StringTok{    }\CommentTok{# Compute basal area:}
+\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{basal_area =} \FloatTok{1e-04} \OperatorTok{*}\StringTok{ }\NormalTok{pi }\OperatorTok{*}\StringTok{ }\NormalTok{(dbh1}\OperatorTok{/}\DecValTok{2}\NormalTok{)}\OperatorTok{^}\DecValTok{2}\NormalTok{)}
 
@@
 \StringTok{        }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi,}
-            \DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{) }\OperatorTok{%>%}
+            \DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{, }\DataTypeTok{comp_x_var =} \StringTok{"basal_area"}\NormalTok{) }\OperatorTok{%>%}
 \StringTok{        }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi)}
@@
 
+\begin{center}\includegraphics[width=1\linewidth]{Figures/appendix2-code-1} \end{center}
+
 Here we see that a cut off distance of approximately 6 m minimizes RMSE.
+
+\hypertarget{compare-pred-var}{%
+\section{Appendix 3: Compare predictor
+variables}\label{compare-pred-var}}
+
+In the above code we use the basal area of an individual as a continuous
+predictor variable. But the package allows the user to specify any
+predictor variable in the \texttt{comp\_x\_var} argument of
+\texttt{create\_focal\_vs\_comp} function. Here we use the
+cross-validated model comparison to see which of two possible predictor
+variables, basal area or above ground biomass, best explains growth.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{census_}\DecValTok{2013}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem2.csv"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date), }\DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+
+\NormalTok{census_}\DecValTok{2018}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem3.csv"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date), }\DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+
+\NormalTok{growth_scbi <-}\StringTok{ }\KeywordTok{compute_growth}\NormalTok{(}\DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi, }\DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)), }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{left_join}\NormalTok{(sp_info, }\DataTypeTok{by =} \StringTok{"sp"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{as.factor}\NormalTok{(sp), }\DataTypeTok{basal_area =} \FloatTok{1e-04} \OperatorTok{*}\StringTok{ }\NormalTok{pi }\OperatorTok{*}\StringTok{ }\NormalTok{(dbh1}\OperatorTok{/}\DecValTok{2}\NormalTok{)}\OperatorTok{^}\DecValTok{2}\NormalTok{, }\DataTypeTok{agb =} \KeywordTok{get_biomass}\NormalTok{(}\DataTypeTok{dbh =}\NormalTok{ dbh1,}
+        \DataTypeTok{genus =}\NormalTok{ genus, }\DataTypeTok{species =}\NormalTok{ species, }\DataTypeTok{coords =} \KeywordTok{c}\NormalTok{(}\OperatorTok{-}\FloatTok{78.2}\NormalTok{, }\FloatTok{38.9}\NormalTok{)))}
+
+\NormalTok{study_region_scbi <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}\DataTypeTok{x =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{), }\DataTypeTok{y =} \KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{,}
+    \DecValTok{600}\NormalTok{, }\DecValTok{300}\NormalTok{)) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{sf_polygon}\NormalTok{()}
+
+\NormalTok{n_fold <-}\StringTok{ }\DecValTok{4}
+\NormalTok{fold1 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{0}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{))}
+\NormalTok{fold2 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{150}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{))}
+\NormalTok{fold3 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{0}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+\NormalTok{fold4 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{150}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+
+\NormalTok{blocks_scbi <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}\KeywordTok{sf_polygon}\NormalTok{(fold1), }\KeywordTok{sf_polygon}\NormalTok{(fold2), }\KeywordTok{sf_polygon}\NormalTok{(fold3),}
+    \KeywordTok{sf_polygon}\NormalTok{(fold4)) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\NormalTok{n_fold) }\OperatorTok{%>%}
+\StringTok{        }\KeywordTok{factor}\NormalTok{())}
+
+\CommentTok{# Associate each observation to a fold}
+\NormalTok{spatial_block_scbi <-}\StringTok{ }\KeywordTok{spatialBlock}\NormalTok{(}\DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =}\NormalTok{ n_fold,}
+    \DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}\NormalTok{)}
+
+\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ spatial_block_scbi}\OperatorTok{$}\NormalTok{foldID }\OperatorTok{%>%}
+\StringTok{        }\KeywordTok{factor}\NormalTok{())}
+
+\NormalTok{comp_dist <-}\StringTok{ }\FloatTok{7.5}
+
+\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{size =}\NormalTok{ comp_dist, }\DataTypeTok{region =}\NormalTok{ study_region_scbi)}
+
+\NormalTok{focal_vs_comp_ba <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{,}
+        \DataTypeTok{comp_x_var =} \StringTok{"basal_area"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi)}
+
+\NormalTok{focal_vs_comp_agb <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{,}
+        \DataTypeTok{comp_x_var =} \StringTok{"agb"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi)}
+
+\NormalTok{focal_vs_comp_ba }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{pull}\NormalTok{(.estimate)}
+\CommentTok{## [1] 0.14}
+
+\NormalTok{focal_vs_comp_agb }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{pull}\NormalTok{(.estimate)}
+\CommentTok{## [1] 2.08}
+\end{Highlighting}
+\end{Shaded}
+
+Here we see that basal area does a much better job than above ground
+biomass.
+
+\hypertarget{compare-group-var}{%
+\section{Appendix 4: Compare grouping
+variables}\label{compare-group-var}}
+
+The package also allows the user to specify the categorical predictor
+grouping variable. Here we compare two different such variables: species
+and the potential canopy position of that species. If we had
+individual-level crown classes (\citet{smith_practice_1986} dominant,
+codominant, intermediate and suppressed) that could also be used.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{census_}\DecValTok{2013}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem2.csv"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date), }\DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+
+\NormalTok{census_}\DecValTok{2018}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem3.csv"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date), }\DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+
+\NormalTok{growth_scbi_sp <-}\StringTok{ }\KeywordTok{compute_growth}\NormalTok{(}\DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi, }\DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)), }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{as.factor}\NormalTok{(sp), }\DataTypeTok{basal_area =} \FloatTok{1e-04} \OperatorTok{*}\StringTok{ }\NormalTok{pi }\OperatorTok{*}\StringTok{ }\NormalTok{(dbh1}\OperatorTok{/}\DecValTok{2}\NormalTok{)}\OperatorTok{^}\DecValTok{2}\NormalTok{)}
+
+\NormalTok{growth_scbi_can_pos <-}\StringTok{ }\KeywordTok{compute_growth}\NormalTok{(}\DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi, }\DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)), }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{left_join}\NormalTok{(sp_info, }\DataTypeTok{by =} \StringTok{"sp"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{canopy_position =} \KeywordTok{str_replace}\NormalTok{(canopy_position, }\StringTok{" "}\NormalTok{, }\StringTok{"_"}\NormalTok{), }\DataTypeTok{canopy_position =} \KeywordTok{str_replace}\NormalTok{(canopy_position,}
+        \StringTok{","}\NormalTok{, }\StringTok{""}\NormalTok{), }\DataTypeTok{canopy_position =} \KeywordTok{ifelse}\NormalTok{(}\KeywordTok{is.na}\NormalTok{(canopy_position), }\StringTok{"shrub_layer"}\NormalTok{,}
+\NormalTok{        canopy_position), }\DataTypeTok{sp =} \KeywordTok{as.factor}\NormalTok{(canopy_position), }\DataTypeTok{basal_area =} \FloatTok{1e-04} \OperatorTok{*}
+\StringTok{        }\NormalTok{pi }\OperatorTok{*}\StringTok{ }\NormalTok{(dbh1}\OperatorTok{/}\DecValTok{2}\NormalTok{)}\OperatorTok{^}\DecValTok{2}\NormalTok{)}
+
+\NormalTok{study_region_scbi <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}\DataTypeTok{x =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{), }\DataTypeTok{y =} \KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{,}
+    \DecValTok{600}\NormalTok{, }\DecValTok{300}\NormalTok{)) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{sf_polygon}\NormalTok{()}
+
+\NormalTok{n_fold <-}\StringTok{ }\DecValTok{4}
+\NormalTok{fold1 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{0}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{))}
+\NormalTok{fold2 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{150}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{))}
+\NormalTok{fold3 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{0}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+\NormalTok{fold4 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{150}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+
+\NormalTok{blocks_scbi <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}\KeywordTok{sf_polygon}\NormalTok{(fold1), }\KeywordTok{sf_polygon}\NormalTok{(fold2), }\KeywordTok{sf_polygon}\NormalTok{(fold3),}
+    \KeywordTok{sf_polygon}\NormalTok{(fold4)) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\NormalTok{n_fold) }\OperatorTok{%>%}
+\StringTok{        }\KeywordTok{factor}\NormalTok{())}
+
+\CommentTok{# Associate each observation to a fold}
+\NormalTok{spatial_block_scbi <-}\StringTok{ }\KeywordTok{spatialBlock}\NormalTok{(}\DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =}\NormalTok{ n_fold,}
+    \DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}\NormalTok{)}
+
+\NormalTok{growth_scbi_sp <-}\StringTok{ }\NormalTok{growth_scbi_sp }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ spatial_block_scbi}\OperatorTok{$}\NormalTok{foldID }\OperatorTok{%>%}
+\StringTok{        }\KeywordTok{factor}\NormalTok{())}
+\NormalTok{growth_scbi_can_pos <-}\StringTok{ }\NormalTok{growth_scbi_can_pos }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ spatial_block_scbi}\OperatorTok{$}\NormalTok{foldID }\OperatorTok{%>%}
+\StringTok{        }\KeywordTok{factor}\NormalTok{())}
+
+\NormalTok{comp_dist <-}\StringTok{ }\FloatTok{7.5}
+
+\NormalTok{growth_scbi_sp <-}\StringTok{ }\NormalTok{growth_scbi_sp }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{size =}\NormalTok{ comp_dist, }\DataTypeTok{region =}\NormalTok{ study_region_scbi)}
+\NormalTok{growth_scbi_can_pos <-}\StringTok{ }\NormalTok{growth_scbi_can_pos }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{size =}\NormalTok{ comp_dist, }\DataTypeTok{region =}\NormalTok{ study_region_scbi)}
+
+\NormalTok{focal_vs_comp_sp <-}\StringTok{ }\NormalTok{growth_scbi_sp }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{,}
+        \DataTypeTok{comp_x_var =} \StringTok{"basal_area"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi)}
+
+\NormalTok{focal_vs_comp_can_pos <-}\StringTok{ }\NormalTok{growth_scbi_can_pos }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{,}
+        \DataTypeTok{comp_x_var =} \StringTok{"basal_area"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi)}
+
+\NormalTok{focal_vs_comp_sp }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{pull}\NormalTok{(.estimate)}
+\CommentTok{## [1] 0.14}
+
+\NormalTok{focal_vs_comp_can_pos }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{pull}\NormalTok{(.estimate)}
+\CommentTok{## [1] 0.144}
+\end{Highlighting}
+\end{Shaded}
+
+We find that species identity has a lower RMSE, so does a better job.
+But we can still plot the competition posteriors for the canopy position
+groupings.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{fit_mod_can_pos <-}\StringTok{ }\NormalTok{growth_scbi_can_pos }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{,}
+        \DataTypeTok{comp_x_var =} \StringTok{"basal_area"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{comp_bayes_lm}\NormalTok{(}\DataTypeTok{prior_param =} \OtherTok{NULL}\NormalTok{)}
+\KeywordTok{autoplot}\NormalTok{(fit_mod_can_pos, }\DataTypeTok{type =} \StringTok{"competition"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{center}\includegraphics[width=1\linewidth]{Figures/can-pos-comp-1} \end{center}
+
+Unsurprisingly we see that canopy and canopy emergent competitors
+generally have negative effects on their neighbors, while shrubs and
+understory competitors have neutral or even positive effects.
 

### 2021-08-10 paper/paper.Rmd
eval = FALSE

--- 2021-08-09 paper/paper.Rmd
@@
 
-```{r internal, echo = FALSE}
+```{r internal, echo = FALSE, eval = FALSE}
 # Internal: after removing trees in buffer and have NA for growth, there is one

### 2021-08-10 paper/paper.tex
No longer eval appendix 1
eval = FALSE

@@
 
-\begin{verbatim}
-## Simple feature collection with 1 feature and 17 fields
-## Geometry type: POINT
-## Dimension:     XY
-## Bounding box:  xmin: 250 ymin: 582 xmax: 250 ymax: 582
-## CRS:           NA
-## # A tibble: 1 x 18
-##   stemID sp     dbh1 codes1 status  dbh2 codes2 growth
-## *  <dbl> <fct> <dbl> <chr>  <chr>  <dbl> <chr>   <dbl>
-## 1  18823 quve   68.8 NULL   A       69.9 NULL    0.215
-## # ... with 10 more variables: geometry <POINT>, family <chr>,
-## #   genus <chr>, species <chr>, canopy_position <chr>,
-## #   drought_tolerance <chr>, basal_area <dbl>, agb <dbl>, buffer <lgl>,
-## #   foldID <fct>
-\end{verbatim}
-
-\begin{center}\includegraphics[width=1\linewidth]{Figures/internal-1} \end{center}
-
 The resulting \texttt{focal\_vs\_comp\_scbi} has 6296 rows, representing
@@
 \NormalTok{\}}
-\CommentTok{## [1] "Shuffle with permutation 1 at 2021-08-09 18:21:51"}
-\CommentTok{## [1] "Shuffle with permutation 2 at 2021-08-09 18:26:48"}
-\CommentTok{## [1] "Shuffle with permutation 3 at 2021-08-09 18:31:46"}
-\CommentTok{## [1] "Shuffle with permutation 4 at 2021-08-09 18:36:41"}
-\CommentTok{## [1] "Shuffle with permutation 5 at 2021-08-09 18:41:31"}
-\CommentTok{## [1] "Shuffle with permutation 6 at 2021-08-09 18:46:03"}
-\CommentTok{## [1] "Shuffle with permutation 7 at 2021-08-09 18:50:32"}
-\CommentTok{## [1] "Shuffle with permutation 8 at 2021-08-09 18:55:09"}
-\CommentTok{## [1] "Shuffle with permutation 9 at 2021-08-09 18:59:38"}
-\CommentTok{## [1] "Shuffle with permutation 10 at 2021-08-09 19:04:06"}
-\CommentTok{## [1] "Shuffle with permutation 11 at 2021-08-09 19:08:36"}
-\CommentTok{## [1] "Shuffle with permutation 12 at 2021-08-09 19:13:01"}
-\CommentTok{## [1] "Shuffle with permutation 13 at 2021-08-09 19:17:25"}
-\CommentTok{## [1] "Shuffle with permutation 14 at 2021-08-09 19:21:47"}
-\CommentTok{## [1] "Shuffle with permutation 15 at 2021-08-09 19:26:20"}
-\CommentTok{## [1] "Shuffle with permutation 16 at 2021-08-09 19:30:42"}
-\CommentTok{## [1] "Shuffle with permutation 17 at 2021-08-09 19:35:05"}
-\CommentTok{## [1] "Shuffle with permutation 18 at 2021-08-09 19:39:28"}
-\CommentTok{## [1] "Shuffle with permutation 19 at 2021-08-09 19:43:51"}
-\CommentTok{## [1] "Shuffle with permutation 20 at 2021-08-09 19:48:14"}
-\CommentTok{## [1] "Shuffle with permutation 21 at 2021-08-09 19:52:37"}
-\CommentTok{## [1] "Shuffle with permutation 22 at 2021-08-09 19:57:01"}
-\CommentTok{## [1] "Shuffle with permutation 23 at 2021-08-09 20:01:29"}
-\CommentTok{## [1] "Shuffle with permutation 24 at 2021-08-09 20:06:09"}
-\CommentTok{## [1] "Shuffle with permutation 25 at 2021-08-09 20:10:42"}
-\CommentTok{## [1] "Shuffle with permutation 26 at 2021-08-09 20:15:06"}
-\CommentTok{## [1] "Shuffle with permutation 27 at 2021-08-09 20:19:29"}
-\CommentTok{## [1] "Shuffle with permutation 28 at 2021-08-09 20:23:52"}
-\CommentTok{## [1] "Shuffle with permutation 29 at 2021-08-09 20:28:15"}
-\CommentTok{## [1] "Shuffle with permutation 30 at 2021-08-09 20:32:37"}
-\CommentTok{## [1] "Shuffle with permutation 31 at 2021-08-09 20:36:59"}
-\CommentTok{## [1] "Shuffle with permutation 32 at 2021-08-09 20:41:21"}
-\CommentTok{## [1] "Shuffle with permutation 33 at 2021-08-09 20:45:43"}
-\CommentTok{## [1] "Shuffle with permutation 34 at 2021-08-09 20:50:06"}
-\CommentTok{## [1] "Shuffle with permutation 35 at 2021-08-09 20:54:29"}
-\CommentTok{## [1] "Shuffle with permutation 36 at 2021-08-09 20:58:51"}
-\CommentTok{## [1] "Shuffle with permutation 37 at 2021-08-09 21:03:16"}
-\CommentTok{## [1] "Shuffle with permutation 38 at 2021-08-09 21:07:38"}
-\CommentTok{## [1] "Shuffle with permutation 39 at 2021-08-09 21:12:01"}
-\CommentTok{## [1] "Shuffle with permutation 40 at 2021-08-09 21:16:24"}
-\CommentTok{## [1] "Shuffle with permutation 41 at 2021-08-09 21:20:47"}
-\CommentTok{## [1] "Shuffle with permutation 42 at 2021-08-09 21:25:11"}
-\CommentTok{## [1] "Shuffle with permutation 43 at 2021-08-09 21:29:34"}
-\CommentTok{## [1] "Shuffle with permutation 44 at 2021-08-09 21:33:57"}
-\CommentTok{## [1] "Shuffle with permutation 45 at 2021-08-09 21:38:20"}
-\CommentTok{## [1] "Shuffle with permutation 46 at 2021-08-09 21:42:42"}
-\CommentTok{## [1] "Shuffle with permutation 47 at 2021-08-09 21:47:05"}
-\CommentTok{## [1] "Shuffle with permutation 48 at 2021-08-09 21:51:27"}
-\CommentTok{## [1] "Shuffle with permutation 49 at 2021-08-09 21:56:05"}
 
@@
 \end{Shaded}
-
-\begin{center}\includegraphics[width=1\linewidth]{Figures/appendix1-code-1} \end{center}
 

### 2021-08-10 paper/paper.Rmd
edits

@@
 
-where $\beta_{0,j}$ is the diameter-independent growth rate of species $j$; $\text{dbh}_{ij}$ is the DBH of the focal tree at the earlier census and $\beta_{\text{dbh},j}$ the slope of that species's diameter-growth relationship;  $x^{\text{comp}}_{ijk}$ is the sum of some numerical explanatory variable of all trees of competitor species $k$ and $\lambda_{jk}$ quantifies the corresponding change in growth for individuals of group $j$ from these competitors; and $\epsilon_{ij}$ is a random error term distributed $\text{Normal}(0, \sigma^2)$.
+where $\beta_{0,j}$ is the diameter-independent growth rate of species $j$; $\text{dbh}_{ij}$ is the DBH of the focal tree at the earlier census and $\beta_{\text{dbh},j}$ the slope of that species's diameter-growth relationship;  $x^{\text{comp}}_{ijk}$ is the sum of some numerical explanatory variable of all trees of competitor species $k$, and $\lambda_{jk}$ quantifies the corresponding change in growth for individuals of species $j$ from these competitors; and $\epsilon_{ij}$ is a random error term distributed $\text{Normal}(0, \sigma^2)$.
 
@@
 By using the `sf` package classes to represent spatial data rather than the `sp` package, the implementation and use of the `forestecology` package's spatial algorithms was greatly simplified.
-
-Lastly, currently the package only implements the Bayesian linear regression model detailed in Equation \ref{eq:model}.
-As we demonstrate in Section \ref{model-fit-predict} however, the fitting of this model is self-contained in a single function `comp_bayes_lm()` which returns an object of S3 class type `comp_bayes_lm`.
-This class has generic methods implemented to print, make predictions, and plot all results.
-Therefore the package can be modularly extended to fit other models as long as they are coded similarly to `comp_bayes_lm()` and have equivalent generic methods implemented.
 
@@
 
-We also load species information as it existed on GitHub on 2021/08/02, which includes family, genus, and species information; as well as classifications of the canopy position (canopy, canopy emergent, understory, shrub layer), drought tolerance (intolerant, resistant) and other characteristics of the species.
+We also load species information as it existed on GitHub on 2021/08/02, which includes family, genus, and species information; as well as classifications of the canopy position (canopy, canopy emergent, understory, shrub layer), drought tolerance (intolerant, resistant), and other characteristics of the species.
 
@@
     drought_tolerance
-  ) %>%
-  mutate()
+  )
 sp_info
@@
 
-We join this species information to our `growth_scbi` data frame.
+We join this species information to our `growth_scbi` data frame and convert the species variable to a factor.
 
@@
   left_join(sp_info, by = "sp") %>%
-  # TODO:
   mutate(sp = as.factor(sp))
@@
 Our model of interspecific competition relies on a spatial definition of who competitor trees are: all trees within a distance `comp_dist` of a focal tree.
-Here we set `comp_dist` to 7.5m, a value informed by other studies [@canham_neighborhood_2004, @uriarte_spatially_2004, @canham_neighborhood_2006], but the package could also be used to compare multiple distances and see which is best supported (see Appendix 2 \ref{mult-dist}).
+Here we set `comp_dist` to 7.5m, a value informed by other studies [@canham_neighborhood_2004, @uriarte_spatially_2004, @canham_neighborhood_2006], but the package could also be used to compare multiple distances and see which is best supported (see Appendix \ref{mult-dist}).
 We use `comp_dist` and a manually constructed `sf` representation of the study region's boundary as inputs to `add_buffer_variable()` to add a `buffer` boolean variable to `growth_scbi`.
@@
 More specifically, identify all trees that are not part of the buffer region, have a valid `growth` measurement, and have at least one neighbor within `r comp_dist`m.
-We do this using `create_focal_vs_comp()`, which takes the previously detailed `comp_dist` and `id` arguments as well as the `sf` representation of the spatial cross-validation blocks and returns a new data frame `focal_vs_comp_scbi`.
+We do this using `create_focal_vs_comp()`, which takes the previously detailed `comp_dist` and `id` arguments, the `sf` representation of the spatial cross-validation blocks `blocks_scbi`, and a specification `comp_x_var` of the `basal_area` variable we use as the competitor explanatory variable $x^{\text{comp}}_{ijk}$ from Equation \ref{eq:model}.
+This function returns a new data frame `focal_vs_comp_scbi`.
 
@@
 
-We observe 4 variables describing 20 competitor trees: the unique tree-stem ID, the distance to the focal tree (all $\leq$ `r comp_dist` m), the species, and the basal area (in m$^2$) calculated as $\frac{\pi \times (\text{DBH/2})^2}{10000}$ for the DBH in cm from the earlier census.
+We observe 4 variables describing 20 competitor trees: the unique tree-stem ID, the distance to the focal tree (all $\leq$ `r comp_dist`m), the species, and the basal area (in m$^2$) calculated as $\frac{\pi \times (\text{DBH/2})^2}{10000}$ for the DBH in cm from the earlier census.
 Saving competitor information in list-column format minimizes redundancy since we do not need to repeat information on the focal tree 20 times.
@@
 Here we use basal area as the continuous predictor variable but the package is flexible to allow the user to specify any continuous predictor variable (basal area, biomass, tree height, a soil nutrient value, \dots).
-The package can also be used to compare predictor variables and see which best explains tree growth, see Appendix 3 \ref{compare-pred-var} for an example comparing basal area and above ground biomass.
+The package can also be used to compare predictor variables and see which best explains tree growth, see Appendix \ref{compare-pred-var} for an example comparing basal area and above ground biomass.
 Similarly, the package can use any categorical variable as a predictor and compare between different categorical variables.
 For example in @allen_permutation_2020 we compare grouping individuals based on species, family, and based on trait-based groups.
-In Appendix 4 \ref{compare-group-var} we give another example and compare grouping individuals by species or by potential canopy position (canopy, understory, shrub layer).
+In Appendix \ref{compare-group-var} we give another example and compare grouping individuals by species or by potential canopy position (canopy, understory, shrub layer).
 
@@
 
-The resulting`comp_bayes_lm_scbi` is an object of S3 class type `comp_bayes_lm` containing the posterior values of all parameters.
+The resulting `comp_bayes_lm_scbi` is an object of S3 class type `comp_bayes_lm` containing the posterior values of all parameters.
 Furthermore, this class includes generics for three methods.
@@
 
+Lastly, currently the package only implements the Bayesian linear regression model detailed in Equation \ref{eq:model}.
+As we demonstrate in Section \ref{model-fit-predict} however, the fitting of this model is self-contained in a single function `comp_bayes_lm()` which returns an object of S3 class type `comp_bayes_lm`.
+This class has generic methods implemented to print, make predictions, and plot all results.
+Therefore the package can be modularly extended to fit other models as long as they are coded similarly to `comp_bayes_lm()` and have equivalent generic methods implemented.
+
 # Acknowledgments
@@
 We intend to archive all data and source code for this manuscript on GitHub at https://github.com/rudeboybert/forestecology and on Zenodo upon acceptance.
-The example Smithsonian Conservation Biology Institute census data are available on GitHub at https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/tree/master/tree_main_census/data/census-csv-files and are archived on Zenodo at https://doi.org/10.5281/zenodo.2649301 [@gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020].
-
-# Appendix 1: Replicate RMSE comparison
+The example Smithsonian Conservation Biology Institute census and species information data are available on GitHub at https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/tree/master/ and are archived on Zenodo at https://doi.org/10.5281/zenodo.2649301 [@gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020].
+
+<div id="refs"></div>
+
+\appendix
+
+# Appendix: Replicate RMSE comparison
 
 This code replicates Figure \ref{fig:scbi-simulation}: A comparison of root mean squared error of
-models for standard, permuted, and spatial cross- validated error estimates.
+models for standard, permuted, and spatial cross-validated error estimates.
 
@@
 
-# Appendix 2: Compare different competitive distances {#mult-dist}
-
-For all the above analyses we took the cut off distance (`comp_dist`) for two stems to compete as 7.5 m.
+# Appendix: Compare different competitive distances {#mult-dist}
+
+For all the above analyses we took the cut off distance (`comp_dist`) for two stems to compete as `r comp_dist`m.
 This distance has been estimated in previous neighborhood competition studies in forests [@canham_neighborhood_2004, @uriarte_spatially_2004, @canham_neighborhood_2006].
-We used 7.5 m in @allen_permutation_2020 as an average of the values estimated in other studies.
+We used `r comp_dist`m in @allen_permutation_2020 as an average of the values estimated in other studies.
 But our package can be used to find which distance is best supported by the data.
@@
 
-Here we see that a cut off distance of approximately 6 m minimizes RMSE.
-
-# Appendix 3: Compare predictor variables {#compare-pred-var}
+Here we see that a cut off distance of approximately 6m minimizes the RMSE.
+
+# Appendix: Compare predictor variables {#compare-pred-var}
 
@@
 
-# Appendix 4: Compare grouping variables {#compare-group-var}
+# Appendix: Compare grouping variables {#compare-group-var}
 

### 2021-08-10 paper/paper.Rmd
More tweaks

@@
 
-Here we use basal area as the continuous predictor variable but the package is flexible to allow the user to specify any continuous predictor variable (basal area, biomass, tree height, a soil nutrient value, \dots).
-The package can also be used to compare predictor variables and see which best explains tree growth, see Appendix \ref{compare-pred-var} for an example comparing basal area and above ground biomass.
-Similarly, the package can use any categorical variable as a predictor and compare between different categorical variables.
+Here we use basal area as the continuous competitor explanatory variable but the package is flexible to allow the user to specify any competitor explanatory variable (basal area, biomass, tree height, a soil nutrient value, \dots).
+The package can also be used to compare competitor explanatory variables and see which best explains tree growth, see Appendix \ref{compare-pred-var} for an example comparing basal area and above ground biomass.
+Similarly, the package can use any categorical variable as an explanatory variable and compare between different categorical variables.
 For example in @allen_permutation_2020 we compare grouping individuals based on species, family, and based on trait-based groups.
@@
 \appendix
+
+# Appendix: Compare different competitive distances {#mult-dist}
+
+For all the above analyses we set the cut off distance (`comp_dist`) for two stems to compete as `r comp_dist`m.
+This distance has been estimated in previous neighborhood competition studies in forests [@canham_neighborhood_2004, @uriarte_spatially_2004, @canham_neighborhood_2006].
+We used `r comp_dist`m in @allen_permutation_2020 as an average of the values estimated in other studies.
+But our package can be used to find which distance is best supported by the data.
+Here we provide an example using another section of the SCBI plot to provide an additional example of the cross-validation block layout.
+To speed computation we do not consider species differences in competitive effects and treat all species as the same.
+
+We observe in Figure \ref{fig:appendix2-plot} that a cut-off distance of approximately 6m minimizes the cross-validation estimated RMSE.
+
+```{r appendix2-code}
+census_2013_scbi <- read_csv("scbi.stem2.csv") %>%
+  select(stemID, sp, date = ExactDate, gx, gy, dbh, codes, status) %>%
+  mutate(
+    date = mdy(date),
+    dbh = as.numeric(dbh) / 10
+  ) %>%
+  filter(gx < 400, gy < 400)
+
+census_2018_scbi <- read_csv("scbi.stem3.csv") %>%
+  select(stemID, sp, date = ExactDate, gx, gy, dbh, codes, status) %>%
+  mutate(
+    date = mdy(date),
+    dbh = as.numeric(dbh) / 10
+  ) %>%
+  filter(gx < 400, gy < 400)
+
+growth_scbi <-
+  compute_growth(
+    census_1 = census_2013_scbi,
+    census_2 = census_2018_scbi %>% filter(!str_detect(codes, "R")),
+    id = "stemID"
+  ) %>%
+  # make all species the same, needs to be factor with at least two levels
+  mutate(sp = factor("A", levels = c("A", "B"))) %>%
+  # Compute basal area:
+  mutate(basal_area = 0.0001 * pi * (dbh1 / 2)^2)
+
+study_region_scbi <- tibble(
+  x = c(0, 400, 400, 0, 0),
+  y = c(0, 0, 400, 400, 0)
+) %>%
+  sf_polygon()
+
+n_fold <- 4
+fold1 <- cbind(c(0, 200, 200, 0), c(0, 0, 200, 200))
+fold2 <- cbind(c(200, 400, 400, 200), c(0, 0, 200, 200))
+fold3 <- cbind(c(0, 200, 200, 0), c(200, 200, 400, 400))
+fold4 <- cbind(c(200, 400, 400, 200), c(200, 200, 400, 400))
+
+blocks_scbi <- bind_rows(
+  sf_polygon(fold1), sf_polygon(fold2), sf_polygon(fold3),
+  sf_polygon(fold4)
+) %>%
+  mutate(folds = c(1:n_fold) %>% factor())
+
+# Associate each observation to a fold
+spatial_block_scbi <-
+  spatialBlock(
+    speciesData = growth_scbi, k = n_fold,
+    selection = "systematic", blocks = blocks_scbi,
+    showBlocks = FALSE, verbose = FALSE
+  )
+
+growth_scbi <- growth_scbi %>%
+  mutate(foldID = spatial_block_scbi$foldID %>% factor())
+
+mult_dist_comp <- tibble(
+  dist = c(5, 6.25, 7.5, 8.75, 10),
+  rmse = 0
+)
+
+for (i in 1:length(mult_dist_comp$dist)) {
+  comp_dist <- mult_dist_comp$dist[i]
+
+  growth_scbi <- growth_scbi %>%
+    add_buffer_variable(size = comp_dist, region = study_region_scbi)
+
+  focal_vs_comp_scbi <- growth_scbi %>%
+    create_focal_vs_comp(comp_dist = comp_dist, blocks = blocks_scbi, id = "stemID", comp_x_var = "basal_area") %>%
+    run_cv(comp_dist = comp_dist, blocks = blocks_scbi)
+
+  mult_dist_comp$rmse[i] <- focal_vs_comp_scbi %>%
+    rmse(truth = growth, estimate = growth_hat) %>%
+    pull(.estimate)
+}
+```
+
+```{r appendix2-plot, out.width="100%", fig.cap="Cross-validated RMSE estimates for 5 `comp_dist` competitive distances.", echo = FALSE}
+mult_dist_comp %>%
+  ggplot(aes(dist, rmse)) +
+  geom_point() +
+  xlab("Distance (m)") +
+  ylab("RMSE (cm)")
+```
+
+# Appendix: Compare competitor explanatory variables {#compare-pred-var}
+
+In the above code we use the basal area of an individual as a continuous competitor explanatory variable.
+But the package allows the user to specify any competitor explanatory variable in the `comp_x_var` argument of `create_focal_vs_comp` function.
+Here we use the cross-validated model comparison to see which of two possible competitor explanatory variables computed in Section \ref{compute-growth}, basal area or above ground biomass, best explains growth.
+
+We observe in Figure \ref{fig:appendix2-code} that a cut-off distance of approximately 6m minimizes the cross-validation estimated RMSE.
+
+```{r compar-pred-var}
+census_2013_scbi <- read_csv("scbi.stem2.csv") %>%
+  select(stemID, sp, date = ExactDate, gx, gy, dbh, codes, status) %>%
+  mutate(
+    date = mdy(date),
+    dbh = as.numeric(dbh) / 10
+  ) %>%
+  filter(gx < 300, between(gy, 300, 600))
+
+census_2018_scbi <- read_csv("scbi.stem3.csv") %>%
+  select(stemID, sp, date = ExactDate, gx, gy, dbh, codes, status) %>%
+  mutate(
+    date = mdy(date),
+    dbh = as.numeric(dbh) / 10
+  ) %>%
+  filter(gx < 300, between(gy, 300, 600))
+
+growth_scbi <-
+  compute_growth(
+    census_1 = census_2013_scbi,
+    census_2 = census_2018_scbi %>% filter(!str_detect(codes, "R")),
+    id = "stemID"
+  ) %>%
+  left_join(sp_info, by = "sp") %>%
+  mutate(
+    sp = as.factor(sp),
+    basal_area = 0.0001 * pi * (dbh1 / 2)^2,
+    agb = get_biomass(
+      dbh = dbh1,
+      genus = genus,
+      species = species,
+      coords = c(-78.2, 38.9)
+    )
+  )
+
+study_region_scbi <- tibble(
+  x = c(0, 300, 300, 0, 0),
+  y = c(300, 300, 600, 600, 300)
+) %>%
+  sf_polygon()
+
+n_fold <- 4
+fold1 <- cbind(c(0, 150, 150, 0), c(300, 300, 450, 450))
+fold2 <- cbind(c(150, 300, 300, 150), c(300, 300, 450, 450))
+fold3 <- cbind(c(0, 150, 150, 0), c(450, 450, 600, 600))
+fold4 <- cbind(c(150, 300, 300, 150), c(450, 450, 600, 600))
+
+blocks_scbi <- bind_rows(
+  sf_polygon(fold1), sf_polygon(fold2), sf_polygon(fold3),
+  sf_polygon(fold4)
+) %>%
+  mutate(folds = c(1:n_fold) %>% factor())
+
+# Associate each observation to a fold
+spatial_block_scbi <-
+  spatialBlock(
+    speciesData = growth_scbi, k = n_fold,
+    selection = "systematic", blocks = blocks_scbi,
+    showBlocks = FALSE, verbose = FALSE
+  )
+
+growth_scbi <- growth_scbi %>%
+  mutate(foldID = spatial_block_scbi$foldID %>% factor())
+
+comp_dist <- 7.5
+
+growth_scbi <- growth_scbi %>%
+  add_buffer_variable(size = comp_dist, region = study_region_scbi)
+
+focal_vs_comp_ba <- growth_scbi %>%
+  create_focal_vs_comp(
+    comp_dist = comp_dist,
+    blocks = blocks_scbi,
+    id = "stemID",
+    comp_x_var = "basal_area"
+  ) %>%
+  run_cv(comp_dist = comp_dist, blocks = blocks_scbi)
+
+focal_vs_comp_agb <- growth_scbi %>%
+  create_focal_vs_comp(
+    comp_dist = comp_dist,
+    blocks = blocks_scbi,
+    id = "stemID",
+    comp_x_var = "agb"
+  ) %>%
+  run_cv(comp_dist = comp_dist, blocks = blocks_scbi)
+
+focal_vs_comp_ba %>%
+  rmse(truth = growth, estimate = growth_hat) %>%
+  pull(.estimate)
+
+focal_vs_comp_agb %>%
+  rmse(truth = growth, estimate = growth_hat) %>%
+  pull(.estimate)
+```
+
+Here we observe that basal area is a better competitor explanatory variable competitor explanatory variables $x^{\text{comp}}_{ijk}$ from Equation \ref{eq:model} than above ground biomass as suggested by the lower estimated RMSE.
+
+# Appendix: Compare grouping variables {#compare-group-var}
+
+The package also allows the user to specify the categorical explanatory grouping variable.
+Here we compare two different such variables: species and the potential canopy position of that species.
+If we had individual-level crown classes (@smith_practice_1986 dominant, codominant, intermediate and suppressed) that could also be used.
+
+```{r compare-group-var}
+census_2013_scbi <- read_csv("scbi.stem2.csv") %>%
+  select(stemID, sp, date = ExactDate, gx, gy, dbh, codes, status) %>%
+  mutate(
+    date = mdy(date),
+    dbh = as.numeric(dbh) / 10
+  ) %>%
+  filter(gx < 300, between(gy, 300, 600))
+
+census_2018_scbi <- read_csv("scbi.stem3.csv") %>%
+  select(stemID, sp, date = ExactDate, gx, gy, dbh, codes, status) %>%
+  mutate(
+    date = mdy(date),
+    dbh = as.numeric(dbh) / 10
+  ) %>%
+  filter(gx < 300, between(gy, 300, 600))
+
+growth_scbi_sp <-
+  compute_growth(
+    census_1 = census_2013_scbi,
+    census_2 = census_2018_scbi %>% filter(!str_detect(codes, "R")),
+    id = "stemID"
+  ) %>%
+  mutate(
+    sp = as.factor(sp),
+    basal_area = 0.0001 * pi * (dbh1 / 2)^2
+  )
+
+growth_scbi_can_pos <-
+  compute_growth(
+    census_1 = census_2013_scbi,
+    census_2 = census_2018_scbi %>% filter(!str_detect(codes, "R")),
+    id = "stemID"
+  ) %>%
+  left_join(sp_info, by = "sp") %>%
+  mutate(
+    canopy_position = str_replace(canopy_position, " ", "_"),
+    canopy_position = str_replace(canopy_position, ",", ""),
+    canopy_position = ifelse(is.na(canopy_position), "shrub_layer", canopy_position),
+    sp = as.factor(canopy_position),
+    basal_area = 0.0001 * pi * (dbh1 / 2)^2
+  )
+
+study_region_scbi <- tibble(
+  x = c(0, 300, 300, 0, 0),
+  y = c(300, 300, 600, 600, 300)
+) %>%
+  sf_polygon()
+
+n_fold <- 4
+fold1 <- cbind(c(0, 150, 150, 0), c(300, 300, 450, 450))
+fold2 <- cbind(c(150, 300, 300, 150), c(300, 300, 450, 450))
+fold3 <- cbind(c(0, 150, 150, 0), c(450, 450, 600, 600))
+fold4 <- cbind(c(150, 300, 300, 150), c(450, 450, 600, 600))
+
+blocks_scbi <- bind_rows(
+  sf_polygon(fold1), sf_polygon(fold2), sf_polygon(fold3),
+  sf_polygon(fold4)
+) %>%
+  mutate(folds = c(1:n_fold) %>% factor())
+
+# Associate each observation to a fold
+spatial_block_scbi <-
+  spatialBlock(
+    speciesData = growth_scbi, k = n_fold,
+    selection = "systematic", blocks = blocks_scbi,
+    showBlocks = FALSE, verbose = FALSE
+  )
+
+growth_scbi_sp <- growth_scbi_sp %>%
+  mutate(foldID = spatial_block_scbi$foldID %>% factor())
+growth_scbi_can_pos <- growth_scbi_can_pos %>%
+  mutate(foldID = spatial_block_scbi$foldID %>% factor())
+
+comp_dist <- 7.5
+
+growth_scbi_sp <- growth_scbi_sp %>%
+  add_buffer_variable(size = comp_dist, region = study_region_scbi)
+growth_scbi_can_pos <- growth_scbi_can_pos %>%
+  add_buffer_variable(size = comp_dist, region = study_region_scbi)
+
+focal_vs_comp_sp <- growth_scbi_sp %>%
+  create_focal_vs_comp(
+    comp_dist = comp_dist,
+    blocks = blocks_scbi,
+    id = "stemID",
+    comp_x_var = "basal_area"
+  ) %>%
+  run_cv(comp_dist = comp_dist, blocks = blocks_scbi)
+
+focal_vs_comp_can_pos <- growth_scbi_can_pos %>%
+  create_focal_vs_comp(
+    comp_dist = comp_dist,
+    blocks = blocks_scbi,
+    id = "stemID",
+    comp_x_var = "basal_area"
+  ) %>%
+  run_cv(comp_dist = comp_dist, blocks = blocks_scbi)
+
+focal_vs_comp_sp %>%
+  rmse(truth = growth, estimate = growth_hat) %>%
+  pull(.estimate)
+
+focal_vs_comp_can_pos %>%
+  rmse(truth = growth, estimate = growth_hat) %>%
+  pull(.estimate)
+```
+
+We find that species identity has a lower RMSE, so does a better job.
+We still however plot the competition posteriors for the canopy position groupings in Figure \ref{fig:can-pos-comp-plot}.
+Unsurprisingly we see that canopy and canopy emergent competitors generally have negative effects on their neighbors, while shrubs and understory competitors have neutral or even positive effects.
+
+```{r can-pos-comp}
+fit_mod_can_pos <- growth_scbi_can_pos %>%
+  create_focal_vs_comp(
+    comp_dist = comp_dist,
+    blocks = blocks_scbi,
+    id = "stemID",
+    comp_x_var = "basal_area"
+  ) %>%
+  comp_bayes_lm(prior_param = NULL)
+```
+
+```{r can-pos-comp-plot, out.width="100%", fig.cap="Posterior distributions of all competition parameters.", echo = FALSE}
+autoplot(fit_mod_can_pos, type = "competition")
+```
 
@@
 
-# Appendix: Compare different competitive distances {#mult-dist}
-
-For all the above analyses we took the cut off distance (`comp_dist`) for two stems to compete as `r comp_dist`m.
-This distance has been estimated in previous neighborhood competition studies in forests [@canham_neighborhood_2004, @uriarte_spatially_2004, @canham_neighborhood_2006].
-We used `r comp_dist`m in @allen_permutation_2020 as an average of the values estimated in other studies.
-But our package can be used to find which distance is best supported by the data.
-Here we provide an example using another section of the SCBI plot to provide an additional example of the CV block layout.
-To speed computation we do not consider species differences in competitive effects and treat all species as the same.
-
-```{r appendix2-code}
-census_2013_scbi <- read_csv("scbi.stem2.csv") %>%
-  select(stemID, sp, date = ExactDate, gx, gy, dbh, codes, status) %>%
-  mutate(
-    date = mdy(date),
-    dbh = as.numeric(dbh) / 10
-  ) %>%
-  filter(gx < 400, gy < 400)
-
-census_2018_scbi <- read_csv("scbi.stem3.csv") %>%
-  select(stemID, sp, date = ExactDate, gx, gy, dbh, codes, status) %>%
-  mutate(
-    date = mdy(date),
-    dbh = as.numeric(dbh) / 10
-  ) %>%
-  filter(gx < 400, gy < 400)
-
-growth_scbi <-
-  compute_growth(
-    census_1 = census_2013_scbi,
-    census_2 = census_2018_scbi %>% filter(!str_detect(codes, "R")),
-    id = "stemID"
-  ) %>%
-  # make all species the same, needs to be factor with at least two levels
-  mutate(sp = factor("A", levels = c("A", "B"))) %>%
-  # Compute basal area:
-  mutate(basal_area = 0.0001 * pi * (dbh1 / 2)^2)
-
-study_region_scbi <- tibble(
-  x = c(0, 400, 400, 0, 0),
-  y = c(0, 0, 400, 400, 0)
-) %>%
-  sf_polygon()
-
-n_fold <- 4
-fold1 <- cbind(c(0, 200, 200, 0), c(0, 0, 200, 200))
-fold2 <- cbind(c(200, 400, 400, 200), c(0, 0, 200, 200))
-fold3 <- cbind(c(0, 200, 200, 0), c(200, 200, 400, 400))
-fold4 <- cbind(c(200, 400, 400, 200), c(200, 200, 400, 400))
-
-blocks_scbi <- bind_rows(
-  sf_polygon(fold1), sf_polygon(fold2), sf_polygon(fold3),
-  sf_polygon(fold4)
-) %>%
-  mutate(folds = c(1:n_fold) %>% factor())
-
-# Associate each observation to a fold
-spatial_block_scbi <-
-  spatialBlock(
-    speciesData = growth_scbi, k = n_fold,
-    selection = "systematic", blocks = blocks_scbi,
-    showBlocks = FALSE, verbose = FALSE
-  )
-
-growth_scbi <- growth_scbi %>%
-  mutate(foldID = spatial_block_scbi$foldID %>% factor())
-
-mult_dist_comp <- tibble(
-  dist = c(5, 6.25, 7.5, 8.75, 10),
-  rmse = 0
-)
-
-for (i in 1:length(mult_dist_comp$dist)) {
-  comp_dist <- mult_dist_comp$dist[i]
-
-  growth_scbi <- growth_scbi %>%
-    add_buffer_variable(size = comp_dist, region = study_region_scbi)
-
-  focal_vs_comp_scbi <- growth_scbi %>%
-    create_focal_vs_comp(comp_dist = comp_dist, blocks = blocks_scbi, id = "stemID", comp_x_var = "basal_area") %>%
-    run_cv(comp_dist = comp_dist, blocks = blocks_scbi)
-
-  mult_dist_comp$rmse[i] <- focal_vs_comp_scbi %>%
-    rmse(truth = growth, estimate = growth_hat) %>%
-    pull(.estimate)
-}
-
-mult_dist_comp %>%
-  ggplot(aes(dist, rmse)) +
-  geom_point() +
-  xlab("Distance (m)") +
-  ylab("RMSE (cm)")
-```
-
-Here we see that a cut off distance of approximately 6m minimizes the RMSE.
-
-# Appendix: Compare predictor variables {#compare-pred-var}
-
-In the above code we use the basal area of an individual as a continuous predictor variable.
-But the package allows the user to specify any predictor variable in the `comp_x_var` argument of `create_focal_vs_comp` function.
-Here we use the cross-validated model comparison to see which of two possible predictor variables, basal area or above ground biomass, best explains growth.
-
-```{r compar-pred-var}
-census_2013_scbi <- read_csv("scbi.stem2.csv") %>%
-  select(stemID, sp, date = ExactDate, gx, gy, dbh, codes, status) %>%
-  mutate(
-    date = mdy(date),
-    dbh = as.numeric(dbh) / 10
-  ) %>%
-  filter(gx < 300, between(gy, 300, 600))
-
-census_2018_scbi <- read_csv("scbi.stem3.csv") %>%
-  select(stemID, sp, date = ExactDate, gx, gy, dbh, codes, status) %>%
-  mutate(
-    date = mdy(date),
-    dbh = as.numeric(dbh) / 10
-  ) %>%
-  filter(gx < 300, between(gy, 300, 600))
-
-growth_scbi <-
-  compute_growth(
-    census_1 = census_2013_scbi,
-    census_2 = census_2018_scbi %>% filter(!str_detect(codes, "R")),
-    id = "stemID"
-  ) %>%
-  left_join(sp_info, by = "sp") %>%
-  mutate(
-    sp = as.factor(sp),
-    basal_area = 0.0001 * pi * (dbh1 / 2)^2,
-    agb = get_biomass(
-      dbh = dbh1,
-      genus = genus,
-      species = species,
-      coords = c(-78.2, 38.9)
-    )
-  )
-
-study_region_scbi <- tibble(
-  x = c(0, 300, 300, 0, 0),
-  y = c(300, 300, 600, 600, 300)
-) %>%
-  sf_polygon()
-
-n_fold <- 4
-fold1 <- cbind(c(0, 150, 150, 0), c(300, 300, 450, 450))
-fold2 <- cbind(c(150, 300, 300, 150), c(300, 300, 450, 450))
-fold3 <- cbind(c(0, 150, 150, 0), c(450, 450, 600, 600))
-fold4 <- cbind(c(150, 300, 300, 150), c(450, 450, 600, 600))
-
-blocks_scbi <- bind_rows(
-  sf_polygon(fold1), sf_polygon(fold2), sf_polygon(fold3),
-  sf_polygon(fold4)
-) %>%
-  mutate(folds = c(1:n_fold) %>% factor())
-
-# Associate each observation to a fold
-spatial_block_scbi <-
-  spatialBlock(
-    speciesData = growth_scbi, k = n_fold,
-    selection = "systematic", blocks = blocks_scbi,
-    showBlocks = FALSE, verbose = FALSE
-  )
-
-growth_scbi <- growth_scbi %>%
-  mutate(foldID = spatial_block_scbi$foldID %>% factor())
-
-comp_dist <- 7.5
-
-growth_scbi <- growth_scbi %>%
-  add_buffer_variable(size = comp_dist, region = study_region_scbi)
-
-focal_vs_comp_ba <- growth_scbi %>%
-  create_focal_vs_comp(
-    comp_dist = comp_dist,
-    blocks = blocks_scbi,
-    id = "stemID",
-    comp_x_var = "basal_area"
-  ) %>%
-  run_cv(comp_dist = comp_dist, blocks = blocks_scbi)
-
-focal_vs_comp_agb <- growth_scbi %>%
-  create_focal_vs_comp(
-    comp_dist = comp_dist,
-    blocks = blocks_scbi,
-    id = "stemID",
-    comp_x_var = "agb"
-  ) %>%
-  run_cv(comp_dist = comp_dist, blocks = blocks_scbi)
-
-focal_vs_comp_ba %>%
-  rmse(truth = growth, estimate = growth_hat) %>%
-  pull(.estimate)
-
-focal_vs_comp_agb %>%
-  rmse(truth = growth, estimate = growth_hat) %>%
-  pull(.estimate)
-```
-
-Here we see that basal area does a much better job than above ground biomass.
-
-# Appendix: Compare grouping variables {#compare-group-var}
-
-The package also allows the user to specify the categorical predictor grouping variable.
-Here we compare two different such variables: species and the potential canopy position of that species.
-If we had individual-level crown classes (@smith_practice_1986 dominant, codominant, intermediate and suppressed) that could also be used.
-
-```{r compare-group-var}
-census_2013_scbi <- read_csv("scbi.stem2.csv") %>%
-  select(stemID, sp, date = ExactDate, gx, gy, dbh, codes, status) %>%
-  mutate(
-    date = mdy(date),
-    dbh = as.numeric(dbh) / 10
-  ) %>%
-  filter(gx < 300, between(gy, 300, 600))
-
-census_2018_scbi <- read_csv("scbi.stem3.csv") %>%
-  select(stemID, sp, date = ExactDate, gx, gy, dbh, codes, status) %>%
-  mutate(
-    date = mdy(date),
-    dbh = as.numeric(dbh) / 10
-  ) %>%
-  filter(gx < 300, between(gy, 300, 600))
-
-growth_scbi_sp <-
-  compute_growth(
-    census_1 = census_2013_scbi,
-    census_2 = census_2018_scbi %>% filter(!str_detect(codes, "R")),
-    id = "stemID"
-  ) %>%
-  mutate(
-    sp = as.factor(sp),
-    basal_area = 0.0001 * pi * (dbh1 / 2)^2
-  )
-
-growth_scbi_can_pos <-
-  compute_growth(
-    census_1 = census_2013_scbi,
-    census_2 = census_2018_scbi %>% filter(!str_detect(codes, "R")),
-    id = "stemID"
-  ) %>%
-  left_join(sp_info, by = "sp") %>%
-  mutate(
-    canopy_position = str_replace(canopy_position, " ", "_"),
-    canopy_position = str_replace(canopy_position, ",", ""),
-    canopy_position = ifelse(is.na(canopy_position), "shrub_layer", canopy_position),
-    sp = as.factor(canopy_position),
-    basal_area = 0.0001 * pi * (dbh1 / 2)^2
-  )
-
-study_region_scbi <- tibble(
-  x = c(0, 300, 300, 0, 0),
-  y = c(300, 300, 600, 600, 300)
-) %>%
-  sf_polygon()
-
-n_fold <- 4
-fold1 <- cbind(c(0, 150, 150, 0), c(300, 300, 450, 450))
-fold2 <- cbind(c(150, 300, 300, 150), c(300, 300, 450, 450))
-fold3 <- cbind(c(0, 150, 150, 0), c(450, 450, 600, 600))
-fold4 <- cbind(c(150, 300, 300, 150), c(450, 450, 600, 600))
-
-blocks_scbi <- bind_rows(
-  sf_polygon(fold1), sf_polygon(fold2), sf_polygon(fold3),
-  sf_polygon(fold4)
-) %>%
-  mutate(folds = c(1:n_fold) %>% factor())
-
-# Associate each observation to a fold
-spatial_block_scbi <-
-  spatialBlock(
-    speciesData = growth_scbi, k = n_fold,
-    selection = "systematic", blocks = blocks_scbi,
-    showBlocks = FALSE, verbose = FALSE
-  )
-
-growth_scbi_sp <- growth_scbi_sp %>%
-  mutate(foldID = spatial_block_scbi$foldID %>% factor())
-growth_scbi_can_pos <- growth_scbi_can_pos %>%
-  mutate(foldID = spatial_block_scbi$foldID %>% factor())
-
-comp_dist <- 7.5
-
-growth_scbi_sp <- growth_scbi_sp %>%
-  add_buffer_variable(size = comp_dist, region = study_region_scbi)
-growth_scbi_can_pos <- growth_scbi_can_pos %>%
-  add_buffer_variable(size = comp_dist, region = study_region_scbi)
-
-focal_vs_comp_sp <- growth_scbi_sp %>%
-  create_focal_vs_comp(
-    comp_dist = comp_dist,
-    blocks = blocks_scbi,
-    id = "stemID",
-    comp_x_var = "basal_area"
-  ) %>%
-  run_cv(comp_dist = comp_dist, blocks = blocks_scbi)
-
-focal_vs_comp_can_pos <- growth_scbi_can_pos %>%
-  create_focal_vs_comp(
-    comp_dist = comp_dist,
-    blocks = blocks_scbi,
-    id = "stemID",
-    comp_x_var = "basal_area"
-  ) %>%
-  run_cv(comp_dist = comp_dist, blocks = blocks_scbi)
-
-focal_vs_comp_sp %>%
-  rmse(truth = growth, estimate = growth_hat) %>%
-  pull(.estimate)
-
-focal_vs_comp_can_pos %>%
-  rmse(truth = growth, estimate = growth_hat) %>%
-  pull(.estimate)
-```
-
-We find that species identity has a lower RMSE, so does a better job. But we can still plot the competition posteriors for the canopy position groupings.
-
-```{r can-pos-comp}
-fit_mod_can_pos <- growth_scbi_can_pos %>%
-  create_focal_vs_comp(
-    comp_dist = comp_dist,
-    blocks = blocks_scbi,
-    id = "stemID",
-    comp_x_var = "basal_area"
-  ) %>%
-  comp_bayes_lm(prior_param = NULL)
-autoplot(fit_mod_can_pos, type = "competition")
-```
-
-Unsurprisingly we see that canopy and canopy emergent competitors generally have negative effects on their neighbors, while shrubs and understory competitors have neutral or even positive effects.
-
 # References

### 2021-08-10 paper/paper.Rmd
Update references

@@
 
-```{r appendix2-plot, out.width="100%", fig.cap="Cross-validated RMSE estimates for 5 `comp_dist` competitive distances.", echo = FALSE}
+```{r appendix2-plot, out.width="100%", fig.cap="Cross-validated RMSE estimates for 5 competitive distances2.", echo = FALSE}
 mult_dist_comp %>%
@@
 Here we use the cross-validated model comparison to see which of two possible competitor explanatory variables computed in Section \ref{compute-growth}, basal area or above ground biomass, best explains growth.
-
-We observe in Figure \ref{fig:appendix2-code} that a cut-off distance of approximately 6m minimizes the cross-validation estimated RMSE.
 

### 2021-08-10 paper/paper.tex
Update references

--- 2021-08-07 paper/paper.tex
@@
 % Pandoc citation processing
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+% for Pandoc 2.8 to 2.10.1
+\newenvironment{cslreferences}%
+  {}%
+  {\par}
+% For Pandoc 2.11+
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
+  % set entry spacing
+  \ifnum #2 > 0
+  \setlength{\parskip}{#2\baselineskip}
+  \fi
+ }%
+ {}
+\usepackage{calc} % for calculating minipage widths
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
@@
 {
-  \title{\bf The forestecology R package for fitting and assessing neighborhood
-models of the effect of interspecific competition on the growth of trees}
+  \title{\bf The forestecology R package for fitting and assessing
+neighborhood models of the effect of interspecific competition on the
+growth of trees}
 
   \author{
-        Albert Y. Kim \thanks{Assistant Professor, Statistical \& Data Sciences, Smith College,
-Northampton, MA 01063 (e-mail:
+        Albert Y. Kim \thanks{Assistant Professor, Statistical \& Data
+Sciences, Smith College, Northampton, MA 01063 (e-mail:
 \href{mailto:akim04@smith.edu}{\nolinkurl{akim04@smith.edu}}).} \\
@@
   \begin{center}
-    {\LARGE\bf The forestecology R package for fitting and assessing neighborhood
-models of the effect of interspecific competition on the growth of trees}
+    {\LARGE\bf The forestecology R package for fitting and assessing
+neighborhood models of the effect of interspecific competition on the
+growth of trees}
   \end{center}
@@
 \noindent%
-{\it Keywords:} forest ecology, interspecific competition, neighborhood competition,
-tree growth, R, ForestGEO, spatial cross-validation
+{\it Keywords:} forest ecology, interspecific competition, neighborhood
+competition, tree growth, R, ForestGEO, spatial cross-validation
 \vfill
@@
 some numerical explanatory variable of all trees of competitor species
-\(k\) and \(\lambda_{jk}\) quantifies the corresponding change in growth
-for individuals of group \(j\) from these competitors; and
+\(k\), and \(\lambda_{jk}\) quantifies the corresponding change in
+growth for individuals of species \(j\) from these competitors; and
 \(\epsilon_{ij}\) is a random error term distributed
@@
 
-Lastly, currently the package only implements the Bayesian linear
-regression model detailed in Equation \ref{eq:model}. As we demonstrate
-in Section \ref{model-fit-predict} however, the fitting of this model is
-self-contained in a single function \texttt{comp\_bayes\_lm()} which
-returns an object of S3 class type \texttt{comp\_bayes\_lm}. This class
-has generic methods implemented to print, make predictions, and plot all
-results. Therefore the package can be modularly extended to fit other
-models as long as they are coded similarly to \texttt{comp\_bayes\_lm()}
-and have equivalent generic methods implemented.
-
 \hypertarget{casestudy}{%
@@
 \begin{Highlighting}[]
-\KeywordTok{library}\NormalTok{(tidyverse)}
-\KeywordTok{library}\NormalTok{(lubridate)}
-\KeywordTok{library}\NormalTok{(sf)}
-\KeywordTok{library}\NormalTok{(patchwork)}
-\KeywordTok{library}\NormalTok{(forestecology)}
-\KeywordTok{library}\NormalTok{(blockCV)}
-
-\CommentTok{# Resolve conflicting functions}
-\NormalTok{filter <-}\StringTok{ }\NormalTok{dplyr}\OperatorTok{::}\NormalTok{filter}
-\NormalTok{select <-}\StringTok{ }\NormalTok{dplyr}\OperatorTok{::}\NormalTok{select}
+\FunctionTok{library}\NormalTok{(tidyverse)}
+\FunctionTok{library}\NormalTok{(lubridate)}
+\FunctionTok{library}\NormalTok{(sf)}
+\FunctionTok{library}\NormalTok{(patchwork)}
+\FunctionTok{library}\NormalTok{(forestecology)}
+\FunctionTok{library}\NormalTok{(blockCV)}
+
+\CommentTok{\# Resolve conflicting functions}
+\NormalTok{filter }\OtherTok{\textless{}{-}}\NormalTok{ dplyr}\SpecialCharTok{::}\NormalTok{filter}
+\NormalTok{select }\OtherTok{\textless{}{-}}\NormalTok{ dplyr}\SpecialCharTok{::}\NormalTok{select}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{census_}\DecValTok{2013}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem2.csv"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
-    \CommentTok{# Convert date from character to date}
-    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date),}
-    \CommentTok{# Convert dbh to be in cm}
-    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}
-\NormalTok{  ) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
-
-\NormalTok{census_}\DecValTok{2018}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem3.csv"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
-    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date),}
-    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}
-\NormalTok{  ) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+\NormalTok{census\_2013\_scbi }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"scbi.stem2.csv"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{select}\NormalTok{(stemID, sp, }\AttributeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}
+    \CommentTok{\# Convert date from character to date}
+    \AttributeTok{date =} \FunctionTok{mdy}\NormalTok{(date),}
+    \CommentTok{\# Convert dbh to be in cm}
+    \AttributeTok{dbh =} \FunctionTok{as.numeric}\NormalTok{(dbh) }\SpecialCharTok{/} \DecValTok{10}
+\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{filter}\NormalTok{(gx }\SpecialCharTok{\textless{}} \DecValTok{300}\NormalTok{, }\FunctionTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+
+\NormalTok{census\_2018\_scbi }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"scbi.stem3.csv"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{select}\NormalTok{(stemID, sp, }\AttributeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}
+    \AttributeTok{date =} \FunctionTok{mdy}\NormalTok{(date),}
+    \AttributeTok{dbh =} \FunctionTok{as.numeric}\NormalTok{(dbh) }\SpecialCharTok{/} \DecValTok{10}
+\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{filter}\NormalTok{(gx }\SpecialCharTok{\textless{}} \DecValTok{300}\NormalTok{, }\FunctionTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{growth_scbi <-}
-\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(}
-    \DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi,}
-    \DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)),}
-    \DataTypeTok{id =} \StringTok{"stemID"}
-\NormalTok{  )}
-\NormalTok{growth_scbi }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{select}\NormalTok{(stemID, sp, dbh1, dbh2, growth, geometry)}
-\CommentTok{## Simple feature collection with 7954 features and 5 fields}
-\CommentTok{## geometry type:  POINT}
-\CommentTok{## dimension:      XY}
-\CommentTok{## bbox:           xmin: 0.2 ymin: 300 xmax: 300 ymax: 600}
-\CommentTok{## CRS:            NA}
-\CommentTok{## # A tibble: 7,954 x 6}
-\CommentTok{##   stemID sp     dbh1  dbh2 growth   geometry}
-\CommentTok{##    <dbl> <fct> <dbl> <dbl>  <dbl>    <POINT>}
-\CommentTok{## 1      4 nysy  13.6   14.2  0.103 (14.2 428)}
-\CommentTok{## 2      5 havi   8.8    9.6  0.150  (9.4 436)}
-\CommentTok{## 3      6 havi   3.25   4    0.140  (1.3 434)}
-\CommentTok{## 4     77 qual  65.2   66    0.141 (34.7 307)}
-\CommentTok{## 5     79 tiam  47.7   46.8 -0.161   (40 381)}
-\CommentTok{## # ... with 7,949 more rows}
+\NormalTok{growth\_scbi }\OtherTok{\textless{}{-}} \FunctionTok{compute\_growth}\NormalTok{(}\AttributeTok{census\_1 =}\NormalTok{ census\_2013\_scbi, }\AttributeTok{census\_2 =}\NormalTok{ census\_2018\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{filter}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{str\_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)), }\AttributeTok{id =} \StringTok{"stemID"}\NormalTok{)}
+\NormalTok{growth\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{select}\NormalTok{(stemID, sp, dbh1, dbh2, growth, geometry)}
+\DocumentationTok{\#\# Simple feature collection with 7954 features and 5 fields}
+\DocumentationTok{\#\# Geometry type: POINT}
+\DocumentationTok{\#\# Dimension:     XY}
+\DocumentationTok{\#\# Bounding box:  xmin: 0.2 ymin: 300 xmax: 300 ymax: 600}
+\DocumentationTok{\#\# CRS:           NA}
+\DocumentationTok{\#\# \# A tibble: 7,954 x 6}
+\DocumentationTok{\#\#   stemID sp     dbh1  dbh2 growth   geometry}
+\DocumentationTok{\#\#    \textless{}dbl\textgreater{} \textless{}fct\textgreater{} \textless{}dbl\textgreater{} \textless{}dbl\textgreater{}  \textless{}dbl\textgreater{}    \textless{}POINT\textgreater{}}
+\DocumentationTok{\#\# 1      4 nysy  13.6   14.2  0.103 (14.2 428)}
+\DocumentationTok{\#\# 2      5 havi   8.8    9.6  0.150  (9.4 436)}
+\DocumentationTok{\#\# 3      6 havi   3.25   4    0.140  (1.3 434)}
+\DocumentationTok{\#\# 4     77 qual  65.2   66    0.141 (34.7 307)}
+\DocumentationTok{\#\# 5     79 tiam  47.7   46.8 {-}0.161   (40 381)}
+\DocumentationTok{\#\# \# ... with 7,949 more rows}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{sample_n}\NormalTok{(}\DecValTok{500}\NormalTok{), }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{size =}\NormalTok{ growth)) }\OperatorTok{+}\StringTok{ }
-\StringTok{  }\KeywordTok{scale_size_binned}\NormalTok{(}\DataTypeTok{limits =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\DecValTok{1}\NormalTok{)) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{size =} \KeywordTok{expression}\NormalTok{(}\KeywordTok{paste}\NormalTok{(Growth, }\StringTok{" (cm "}\NormalTok{,y}\OperatorTok{^}\NormalTok{\{}\OperatorTok{-}\DecValTok{1}\NormalTok{\},}\StringTok{')'}\NormalTok{)) )}
+\FunctionTok{ggplot}\NormalTok{() }\SpecialCharTok{+} \FunctionTok{geom\_sf}\NormalTok{(}\AttributeTok{data =}\NormalTok{ growth\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{sample\_n}\NormalTok{(}\DecValTok{500}\NormalTok{), }\FunctionTok{aes}\NormalTok{(}\AttributeTok{size =}\NormalTok{ growth)) }\SpecialCharTok{+} \FunctionTok{scale\_size\_binned}\NormalTok{(}\AttributeTok{limits =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{,}
+    \DecValTok{1}\NormalTok{)) }\SpecialCharTok{+} \FunctionTok{labs}\NormalTok{(}\AttributeTok{size =} \FunctionTok{expression}\NormalTok{(}\FunctionTok{paste}\NormalTok{(Growth, }\StringTok{" (cm "}\NormalTok{, y}\SpecialCharTok{\^{}}\NormalTok{\{}
+    \SpecialCharTok{{-}}\DecValTok{1}
+\NormalTok{\}, }\StringTok{")"}\NormalTok{)))}
 \end{Highlighting}
@@
 classifications of the canopy position (canopy, canopy emergent,
-understory, shrub layer), drought tolerance (intolerant, resistant) and
+understory, shrub layer), drought tolerance (intolerant, resistant), and
 other characteristics of the species.
@@
 \begin{Highlighting}[]
-\NormalTok{sp_info <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"SCBI_ForestGEO_sp_ecology.csv"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{select}\NormalTok{(}
-    \DataTypeTok{sp =}\NormalTok{ spcode, family, genus, species, canopy_position, }
-\NormalTok{    drought_tolerance}
-\NormalTok{  ) }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{mutate}\NormalTok{()}
-\NormalTok{sp_info}
-\CommentTok{## # A tibble: 65 x 6}
-\CommentTok{##   sp    family        genus       species     canopy_position drought_toleran~}
-\CommentTok{##   <chr> <chr>         <chr>       <chr>       <chr>           <chr>           }
-\CommentTok{## 1 acne  Sapindaceae   Acer        negundo     understory      drought-resista~}
-\CommentTok{## 2 acpl  Sapindaceae   Acer        platanoides canopy          drought-resista~}
-\CommentTok{## 3 acru  Sapindaceae   Acer        rubrum      canopy          drought-resista~}
-\CommentTok{## 4 aial  Simaroubaceae Ailanthus   altissima   canopy          drought-resista~}
-\CommentTok{## 5 amar  Rosaceae      Amelanchier arborea     understory      drought-resista~}
-\CommentTok{## # ... with 60 more rows}
-\end{Highlighting}
-\end{Shaded}
-
-We join this species information to our \texttt{growth\_scbi} data
-frame.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{left_join}\NormalTok{(sp_info, }\DataTypeTok{by =} \StringTok{"sp"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\CommentTok{# }\AlertTok{TODO}\CommentTok{:}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{as.factor}\NormalTok{(sp))}
+\NormalTok{sp\_info }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"SCBI\_ForestGEO\_sp\_ecology.csv"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{select}\NormalTok{(}\AttributeTok{sp =}\NormalTok{ spcode, family, genus, species, canopy\_position, drought\_tolerance)}
+\NormalTok{sp\_info}
+\DocumentationTok{\#\# \# A tibble: 65 x 6}
+\DocumentationTok{\#\#   sp    family        genus       species     canopy\_position drought\_toleran\textasciitilde{}}
+\DocumentationTok{\#\#   \textless{}chr\textgreater{} \textless{}chr\textgreater{}         \textless{}chr\textgreater{}       \textless{}chr\textgreater{}       \textless{}chr\textgreater{}           \textless{}chr\textgreater{}           }
+\DocumentationTok{\#\# 1 acne  Sapindaceae   Acer        negundo     understory      drought{-}resista\textasciitilde{}}
+\DocumentationTok{\#\# 2 acpl  Sapindaceae   Acer        platanoides canopy          drought{-}resista\textasciitilde{}}
+\DocumentationTok{\#\# 3 acru  Sapindaceae   Acer        rubrum      canopy          drought{-}resista\textasciitilde{}}
+\DocumentationTok{\#\# 4 aial  Simaroubaceae Ailanthus   altissima   canopy          drought{-}resista\textasciitilde{}}
+\DocumentationTok{\#\# 5 amar  Rosaceae      Amelanchier arborea     understory      drought{-}resista\textasciitilde{}}
+\DocumentationTok{\#\# \# ... with 60 more rows}
+\end{Highlighting}
+\end{Shaded}
+
+We join this species information to our \texttt{growth\_scbi} data frame
+and convert the species variable to a factor.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{growth\_scbi }\OtherTok{\textless{}{-}}\NormalTok{ growth\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{left\_join}\NormalTok{(sp\_info, }\AttributeTok{by =} \StringTok{"sp"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{sp =} \FunctionTok{as.factor}\NormalTok{(sp))}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\CommentTok{# Install development version of allodb using:}
-\CommentTok{# remotes::install_github("forestgeo/allodb")}
-\KeywordTok{library}\NormalTok{(allodb)}
-\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
-    \CommentTok{# Compute basal area:}
-    \DataTypeTok{basal_area =} \FloatTok{0.0001} \OperatorTok{*}\StringTok{ }\NormalTok{pi }\OperatorTok{*}\StringTok{ }\NormalTok{(dbh1 }\OperatorTok{/}\StringTok{ }\DecValTok{2}\NormalTok{)}\OperatorTok{^}\DecValTok{2}\NormalTok{,}
-    \CommentTok{# Compute above ground biomass:}
-    \DataTypeTok{agb =} \KeywordTok{get_biomass}\NormalTok{(}
-      \DataTypeTok{dbh =}\NormalTok{ dbh1,}
-      \DataTypeTok{genus =}\NormalTok{ genus,}
-      \DataTypeTok{species =}\NormalTok{ species,}
-      \DataTypeTok{coords =} \KeywordTok{c}\NormalTok{(}\OperatorTok{-}\FloatTok{78.2}\NormalTok{, }\FloatTok{38.9}\NormalTok{)}
+\CommentTok{\# Install development version of allodb using:}
+\CommentTok{\# remotes::install\_github("forestgeo/allodb")}
+\FunctionTok{library}\NormalTok{(allodb)}
+\NormalTok{growth\_scbi }\OtherTok{\textless{}{-}}\NormalTok{ growth\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}
+    \CommentTok{\# Compute basal area:}
+    \AttributeTok{basal\_area =} \FloatTok{0.0001} \SpecialCharTok{*}\NormalTok{ pi }\SpecialCharTok{*}\NormalTok{ (dbh1 }\SpecialCharTok{/} \DecValTok{2}\NormalTok{)}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{,}
+    \CommentTok{\# Compute above ground biomass:}
+    \AttributeTok{agb =} \FunctionTok{get\_biomass}\NormalTok{(}
+      \AttributeTok{dbh =}\NormalTok{ dbh1,}
+      \AttributeTok{genus =}\NormalTok{ genus,}
+      \AttributeTok{species =}\NormalTok{ species,}
+      \AttributeTok{coords =} \FunctionTok{c}\NormalTok{(}\SpecialCharTok{{-}}\FloatTok{78.2}\NormalTok{, }\FloatTok{38.9}\NormalTok{)}
 \NormalTok{    )}
@@
 package could also be used to compare multiple distances and see which
-is best supported (see Appendix 2 \ref{mult-dist}). We use
+is best supported (see Appendix \ref{mult-dist}). We use
 \texttt{comp\_dist} and a manually constructed \texttt{sf}
@@
 \begin{Highlighting}[]
-\CommentTok{# Define competitive distance range}
-\NormalTok{comp_dist <-}\StringTok{ }\FloatTok{7.5}
-
-\CommentTok{# Manually construct study region boundary}
-\NormalTok{study_region_scbi <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}
-  \DataTypeTok{x =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{),}
-  \DataTypeTok{y =} \KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{300}\NormalTok{)}
-\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{sf_polygon}\NormalTok{()}
-
-\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{size =}\NormalTok{ comp_dist, }\DataTypeTok{region =}\NormalTok{ study_region_scbi)}
+\CommentTok{\# Define competitive distance range}
+\NormalTok{comp\_dist }\OtherTok{\textless{}{-}} \FloatTok{7.5}
+
+\CommentTok{\# Manually construct study region boundary}
+\NormalTok{study\_region\_scbi }\OtherTok{\textless{}{-}} \FunctionTok{tibble}\NormalTok{(}\AttributeTok{x =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{), }\AttributeTok{y =} \FunctionTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{,}
+    \DecValTok{600}\NormalTok{, }\DecValTok{300}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{sf\_polygon}\NormalTok{()}
+
+\NormalTok{growth\_scbi }\OtherTok{\textless{}{-}}\NormalTok{ growth\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{add\_buffer\_variable}\NormalTok{(}\AttributeTok{size =}\NormalTok{ comp\_dist, }\AttributeTok{region =}\NormalTok{ study\_region\_scbi)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\CommentTok{# Manually define spatial blocks to act as folds}
-\NormalTok{n_fold <-}\StringTok{ }\DecValTok{4}
-\NormalTok{fold1 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{0}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{))}
-\NormalTok{fold2 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{150}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{))}
-\NormalTok{fold3 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{0}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{))}
-\NormalTok{fold4 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{150}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{))}
-
-\NormalTok{blocks_scbi <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}
-  \KeywordTok{sf_polygon}\NormalTok{(fold1), }\KeywordTok{sf_polygon}\NormalTok{(fold2), }
-  \KeywordTok{sf_polygon}\NormalTok{(fold3), }\KeywordTok{sf_polygon}\NormalTok{(fold4)}
-\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\NormalTok{n_fold) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
-
-\CommentTok{# Associate each observation to a fold}
-\NormalTok{spatial_block_scbi <-}\StringTok{ }
-\StringTok{  }\KeywordTok{spatialBlock}\NormalTok{(}
-    \DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =}\NormalTok{ n_fold, }
-    \DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }
-    \DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}
-\NormalTok{  )}
-
-\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ spatial_block_scbi}\OperatorTok{$}\NormalTok{foldID }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ blocks_scbi, }\DataTypeTok{fill =} \StringTok{"transparent"}\NormalTok{, }\DataTypeTok{linetype =} \StringTok{"dashed"}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_sf_text}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{sample_n}\NormalTok{(}\DecValTok{1000}\NormalTok{), }
-               \KeywordTok{aes}\NormalTok{(}\DataTypeTok{label =}\NormalTok{ foldID, }\DataTypeTok{col =}\NormalTok{ buffer))}
+\CommentTok{\# Manually define spatial blocks to act as folds}
+\NormalTok{n\_fold }\OtherTok{\textless{}{-}} \DecValTok{4}
+\NormalTok{fold1 }\OtherTok{\textless{}{-}} \FunctionTok{cbind}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{0}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{))}
+\NormalTok{fold2 }\OtherTok{\textless{}{-}} \FunctionTok{cbind}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{150}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{))}
+\NormalTok{fold3 }\OtherTok{\textless{}{-}} \FunctionTok{cbind}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{0}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+\NormalTok{fold4 }\OtherTok{\textless{}{-}} \FunctionTok{cbind}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{150}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+
+\NormalTok{blocks\_scbi }\OtherTok{\textless{}{-}} \FunctionTok{bind\_rows}\NormalTok{(}\FunctionTok{sf\_polygon}\NormalTok{(fold1), }\FunctionTok{sf\_polygon}\NormalTok{(fold2), }\FunctionTok{sf\_polygon}\NormalTok{(fold3),}
+    \FunctionTok{sf\_polygon}\NormalTok{(fold4)) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{folds =} \FunctionTok{c}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\NormalTok{n\_fold) }\SpecialCharTok{\%\textgreater{}\%}
+        \FunctionTok{factor}\NormalTok{())}
+
+\CommentTok{\# Associate each observation to a fold}
+\NormalTok{spatial\_block\_scbi }\OtherTok{\textless{}{-}} \FunctionTok{spatialBlock}\NormalTok{(}\AttributeTok{speciesData =}\NormalTok{ growth\_scbi, }\AttributeTok{k =}\NormalTok{ n\_fold,}
+    \AttributeTok{selection =} \StringTok{"systematic"}\NormalTok{, }\AttributeTok{blocks =}\NormalTok{ blocks\_scbi, }\AttributeTok{showBlocks =} \ConstantTok{FALSE}\NormalTok{, }\AttributeTok{verbose =} \ConstantTok{FALSE}\NormalTok{)}
+
+\NormalTok{growth\_scbi }\OtherTok{\textless{}{-}}\NormalTok{ growth\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{foldID =}\NormalTok{ spatial\_block\_scbi}\SpecialCharTok{$}\NormalTok{foldID }\SpecialCharTok{\%\textgreater{}\%}
+        \FunctionTok{factor}\NormalTok{())}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{ggplot}\NormalTok{() }\SpecialCharTok{+} \FunctionTok{geom\_sf}\NormalTok{(}\AttributeTok{data =}\NormalTok{ blocks\_scbi, }\AttributeTok{fill =} \StringTok{"transparent"}\NormalTok{, }\AttributeTok{linetype =} \StringTok{"dashed"}\NormalTok{) }\SpecialCharTok{+}
+    \FunctionTok{geom\_sf\_text}\NormalTok{(}\AttributeTok{data =}\NormalTok{ growth\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+        \FunctionTok{sample\_n}\NormalTok{(}\DecValTok{1000}\NormalTok{), }\FunctionTok{aes}\NormalTok{(}\AttributeTok{label =}\NormalTok{ foldID, }\AttributeTok{col =}\NormalTok{ buffer))}
 \end{Highlighting}
@@
 \texttt{create\_focal\_vs\_comp()}, which takes the previously detailed
-\texttt{comp\_dist} and \texttt{id} arguments as well as the \texttt{sf}
-representation of the spatial cross-validation blocks and returns a new
-data frame \texttt{focal\_vs\_comp\_scbi}.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{, }\DataTypeTok{comp_x_var =} \StringTok{"basal_area"}\NormalTok{)}
-\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{select}\NormalTok{(focal_ID, focal_sp, geometry, growth, comp)}
-\CommentTok{## # A tibble: 6,296 x 5}
-\CommentTok{##   focal_ID focal_sp   geometry growth comp             }
-\CommentTok{##      <dbl> <fct>       <POINT>  <dbl> <list>           }
-\CommentTok{## 1        4 nysy     (14.2 428)  0.103 <tibble [20 x 4]>}
-\CommentTok{## 2        5 havi      (9.4 436)  0.150 <tibble [32 x 4]>}
-\CommentTok{## 3       79 tiam       (40 381) -0.161 <tibble [20 x 4]>}
-\CommentTok{## 4       80 caca     (38.7 422)  0.253 <tibble [12 x 4]>}
-\CommentTok{## 5       96 libe       (60 310)  0.262 <tibble [14 x 4]>}
-\CommentTok{## # ... with 6,291 more rows}
+\texttt{comp\_dist} and \texttt{id} arguments, the \texttt{sf}
+representation of the spatial cross-validation blocks
+\texttt{blocks\_scbi}, and a specification \texttt{comp\_x\_var} of the
+\texttt{basal\_area} variable we use as the competitor explanatory
+variable \(x^{\text{comp}}_{ijk}\) from Equation \ref{eq:model}. This
+function returns a new data frame \texttt{focal\_vs\_comp\_scbi}.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{focal\_vs\_comp\_scbi }\OtherTok{\textless{}{-}}\NormalTok{ growth\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{create\_focal\_vs\_comp}\NormalTok{(comp\_dist, }\AttributeTok{blocks =}\NormalTok{ blocks\_scbi, }\AttributeTok{id =} \StringTok{"stemID"}\NormalTok{,}
+        \AttributeTok{comp\_x\_var =} \StringTok{"basal\_area"}\NormalTok{)}
+\NormalTok{focal\_vs\_comp\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{select}\NormalTok{(focal\_ID, focal\_sp, geometry, growth, comp)}
+\DocumentationTok{\#\# \# A tibble: 6,296 x 5}
+\DocumentationTok{\#\#   focal\_ID focal\_sp   geometry growth comp             }
+\DocumentationTok{\#\#      \textless{}dbl\textgreater{} \textless{}fct\textgreater{}       \textless{}POINT\textgreater{}  \textless{}dbl\textgreater{} \textless{}list\textgreater{}           }
+\DocumentationTok{\#\# 1        4 nysy     (14.2 428)  0.103 \textless{}tibble [20 x 4]\textgreater{}}
+\DocumentationTok{\#\# 2        5 havi      (9.4 436)  0.150 \textless{}tibble [32 x 4]\textgreater{}}
+\DocumentationTok{\#\# 3       79 tiam       (40 381) {-}0.161 \textless{}tibble [20 x 4]\textgreater{}}
+\DocumentationTok{\#\# 4       80 caca     (38.7 422)  0.253 \textless{}tibble [12 x 4]\textgreater{}}
+\DocumentationTok{\#\# 5       96 libe       (60 310)  0.262 \textless{}tibble [14 x 4]\textgreater{}}
+\DocumentationTok{\#\# \# ... with 6,291 more rows}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{filter}\NormalTok{(focal_ID }\OperatorTok{==}\StringTok{ }\DecValTok{4}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{select}\NormalTok{(focal_ID, dbh, comp) }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{unnest}\NormalTok{(}\DataTypeTok{cols =} \StringTok{"comp"}\NormalTok{)}
-\CommentTok{## # A tibble: 20 x 6}
-\CommentTok{##   focal_ID   dbh comp_ID  dist comp_sp comp_x_var}
-\CommentTok{##      <dbl> <dbl>   <dbl> <dbl> <fct>        <dbl>}
-\CommentTok{## 1        4  13.6    1836  7.48 tiam       0.0176 }
-\CommentTok{## 2        4  13.6    1847  2.81 nysy       0.00332}
-\CommentTok{## 3        4  13.6    1848  1.62 nysy       0.00396}
-\CommentTok{## 4        4  13.6    1849  2.62 nysy       0.00535}
-\CommentTok{## 5        4  13.6    1850  2.98 havi       0.00472}
-\CommentTok{## # ... with 15 more rows}
+\NormalTok{focal\_vs\_comp\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{filter}\NormalTok{(focal\_ID }\SpecialCharTok{==} \DecValTok{4}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{select}\NormalTok{(focal\_ID, dbh, comp) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{unnest}\NormalTok{(}\AttributeTok{cols =} \StringTok{"comp"}\NormalTok{)}
+\DocumentationTok{\#\# \# A tibble: 20 x 6}
+\DocumentationTok{\#\#   focal\_ID   dbh comp\_ID  dist comp\_sp comp\_x\_var}
+\DocumentationTok{\#\#      \textless{}dbl\textgreater{} \textless{}dbl\textgreater{}   \textless{}dbl\textgreater{} \textless{}dbl\textgreater{} \textless{}fct\textgreater{}        \textless{}dbl\textgreater{}}
+\DocumentationTok{\#\# 1        4  13.6    1836  7.48 tiam       0.0176 }
+\DocumentationTok{\#\# 2        4  13.6    1847  2.81 nysy       0.00332}
+\DocumentationTok{\#\# 3        4  13.6    1848  1.62 nysy       0.00396}
+\DocumentationTok{\#\# 4        4  13.6    1849  2.62 nysy       0.00535}
+\DocumentationTok{\#\# 5        4  13.6    1850  2.98 havi       0.00472}
+\DocumentationTok{\#\# \# ... with 15 more rows}
 \end{Highlighting}
@@
 We observe 4 variables describing 20 competitor trees: the unique
-tree-stem ID, the distance to the focal tree (all \(\leq\) 7.5 m), the
+tree-stem ID, the distance to the focal tree (all \(\leq\) 7.5m), the
 species, and the basal area (in m\(^2\)) calculated as
@@
 
-Here we use basal area as the continuous predictor variable but the
-package is flexible to allow the user to specify any continuous
-predictor variable (basal area, biomass, tree height, a soil nutrient
-value, \dots). The package can also be used to compare predictor
-variables and see which best explains tree growth, see Appendix 3
-\ref{compare-pred-var} for an example comparing basal area and above
-ground biomass. Similarly, the package can use any categorical variable
-as a predictor and compare between different categorical variables. For
-example in \citet{allen_permutation_2020} we compare grouping
-individuals based on species, family, and based on trait-based groups.
-In Appendix 4 \ref{compare-group-var} we give another example and
-compare grouping individuals by species or by potential canopy position
-(canopy, understory, shrub layer).
+Here we use basal area as the continuous competitor explanatory variable
+but the package is flexible to allow the user to specify any competitor
+explanatory variable (basal area, biomass, tree height, a soil nutrient
+value, \dots). The package can also be used to compare competitor
+explanatory variables and see which best explains tree growth, see
+Appendix \ref{compare-pred-var} for an example comparing basal area and
+above ground biomass. Similarly, the package can use any categorical
+variable as an explanatory variable and compare between different
+categorical variables. For example in \citet{allen_permutation_2020} we
+compare grouping individuals based on species, family, and based on
+trait-based groups. In Appendix \ref{compare-group-var} we give another
+example and compare grouping individuals by species or by potential
+canopy position (canopy, understory, shrub layer).
 
@@
 \begin{Highlighting}[]
-\NormalTok{comp_bayes_lm_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{comp_bayes_lm}\NormalTok{(}\DataTypeTok{prior_param =} \OtherTok{NULL}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-The resulting\texttt{comp\_bayes\_lm\_scbi} is an object of S3 class
+\NormalTok{comp\_bayes\_lm\_scbi }\OtherTok{\textless{}{-}}\NormalTok{ focal\_vs\_comp\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{comp\_bayes\_lm}\NormalTok{(}\AttributeTok{prior\_param =} \ConstantTok{NULL}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+The resulting \texttt{comp\_bayes\_lm\_scbi} is an object of S3 class
 type \texttt{comp\_bayes\_lm} containing the posterior values of all
@@
 \begin{Highlighting}[]
-\NormalTok{comp_bayes_lm_scbi}
-\CommentTok{## Bayesian linear regression model parameters with a multivariate Normal}
-\CommentTok{## likelihood. See ?comp_bayes_lm for details:}
-\CommentTok{## }
-\CommentTok{##   parameter_type           prior posterior}
-\CommentTok{## 1 Inverse-Gamma on sigma^2 a_0   a_star   }
-\CommentTok{## 2 Inverse-Gamma on sigma^2 b_0   b_star   }
-\CommentTok{## 3 Multivariate t on beta   mu_0  mu_star  }
-\CommentTok{## 4 Multivariate t on beta   V_0   V_star   }
-\CommentTok{## }
-\CommentTok{## Model formula:}
-\CommentTok{## growth ~ sp + dbh + dbh * sp + acne * sp + acru * sp + amar * sp + astr}
-\CommentTok{## * sp + caca * sp + caco * sp + cade * sp + cagl * sp + caovl * sp + cato}
-\CommentTok{## * sp + ceca * sp + ceoc * sp + chvi * sp + cofl * sp + crpr * sp + crsp}
-\CommentTok{## * sp + divi * sp + elum * sp + fagr * sp + fram * sp + frni * sp + frpe}
-\CommentTok{## * sp + havi * sp + ilve * sp + juci * sp + juni * sp + libe * sp + litu}
-\CommentTok{## * sp + nysy * sp + pist * sp + pivi * sp + ploc * sp + prav * sp + prse}
-\CommentTok{## * sp + qual * sp + quco * sp + qufa * sp + qumi * sp + qupr * sp + quru}
-\CommentTok{## * sp + quve * sp + rops * sp + saal * sp + saca * sp + tiam * sp + ulam}
-\CommentTok{## * sp + ulru * sp + unk * sp + vipr * sp}
+\NormalTok{comp\_bayes\_lm\_scbi}
+\DocumentationTok{\#\# Bayesian linear regression model parameters with a multivariate Normal}
+\DocumentationTok{\#\# likelihood. See ?comp\_bayes\_lm for details:}
+\DocumentationTok{\#\# }
+\DocumentationTok{\#\#   parameter\_type           prior posterior}
+\DocumentationTok{\#\# 1 Inverse{-}Gamma on sigma\^{}2 a\_0   a\_star   }
+\DocumentationTok{\#\# 2 Inverse{-}Gamma on sigma\^{}2 b\_0   b\_star   }
+\DocumentationTok{\#\# 3 Multivariate t on beta   mu\_0  mu\_star  }
+\DocumentationTok{\#\# 4 Multivariate t on beta   V\_0   V\_star   }
+\DocumentationTok{\#\# }
+\DocumentationTok{\#\# Model formula:}
+\DocumentationTok{\#\# growth \textasciitilde{} sp + dbh + dbh * sp + acne * sp + acru * sp + amar * sp + astr}
+\DocumentationTok{\#\# * sp + caca * sp + caco * sp + cade * sp + cagl * sp + caovl * sp + cato}
+\DocumentationTok{\#\# * sp + ceca * sp + ceoc * sp + chvi * sp + cofl * sp + crpr * sp + crsp}
+\DocumentationTok{\#\# * sp + divi * sp + elum * sp + fagr * sp + fram * sp + frni * sp + frpe}
+\DocumentationTok{\#\# * sp + havi * sp + ilve * sp + juci * sp + juni * sp + libe * sp + litu}
+\DocumentationTok{\#\# * sp + nysy * sp + pist * sp + pivi * sp + ploc * sp + prav * sp + prse}
+\DocumentationTok{\#\# * sp + qual * sp + quco * sp + qufa * sp + qumi * sp + qupr * sp + quru}
+\DocumentationTok{\#\# * sp + quve * sp + rops * sp + saal * sp + saca * sp + tiam * sp + ulam}
+\DocumentationTok{\#\# * sp + ulru * sp + unk * sp + vipr * sp}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{growth_hat =} \KeywordTok{predict}\NormalTok{(comp_bayes_lm_scbi, }
-                              \DataTypeTok{newdata =}\NormalTok{ focal_vs_comp_scbi))}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{select}\NormalTok{(focal_ID, focal_sp, dbh, growth, growth_hat)}
-\CommentTok{## # A tibble: 6,296 x 5}
-\CommentTok{##   focal_ID focal_sp   dbh growth growth_hat}
-\CommentTok{##      <dbl> <fct>    <dbl>  <dbl>      <dbl>}
-\CommentTok{## 1        4 nysy     13.6   0.103     0.0809}
-\CommentTok{## 2        5 havi      8.8   0.150     0.112 }
-\CommentTok{## 3       79 tiam     47.7  -0.161     0.229 }
-\CommentTok{## 4       80 caca      5.15  0.253     0.121 }
-\CommentTok{## 5       96 libe      2.3   0.262     0.142 }
-\CommentTok{## # ... with 6,291 more rows}
+\NormalTok{focal\_vs\_comp\_scbi }\OtherTok{\textless{}{-}}\NormalTok{ focal\_vs\_comp\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{growth\_hat =} \FunctionTok{predict}\NormalTok{(comp\_bayes\_lm\_scbi, }\AttributeTok{newdata =}\NormalTok{ focal\_vs\_comp\_scbi))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{focal\_vs\_comp\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{select}\NormalTok{(focal\_ID, focal\_sp, dbh, growth, growth\_hat)}
+\DocumentationTok{\#\# \# A tibble: 6,296 x 5}
+\DocumentationTok{\#\#   focal\_ID focal\_sp   dbh growth growth\_hat}
+\DocumentationTok{\#\#      \textless{}dbl\textgreater{} \textless{}fct\textgreater{}    \textless{}dbl\textgreater{}  \textless{}dbl\textgreater{}      \textless{}dbl\textgreater{}}
+\DocumentationTok{\#\# 1        4 nysy     13.6   0.103     0.0809}
+\DocumentationTok{\#\# 2        5 havi      8.8   0.150     0.112 }
+\DocumentationTok{\#\# 3       79 tiam     47.7  {-}0.161     0.229 }
+\DocumentationTok{\#\# 4       80 caca      5.15  0.253     0.121 }
+\DocumentationTok{\#\# 5       96 libe      2.3   0.262     0.142 }
+\DocumentationTok{\#\# \# ... with 6,291 more rows}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{model_rmse <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
-\NormalTok{model_rmse}
-\CommentTok{## [1] 0.128}
+\NormalTok{model\_rmse }\OtherTok{\textless{}{-}}\NormalTok{ focal\_vs\_comp\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{rmse}\NormalTok{(}\AttributeTok{truth =}\NormalTok{ growth, }\AttributeTok{estimate =}\NormalTok{ growth\_hat) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{pull}\NormalTok{(.estimate)}
+\NormalTok{model\_rmse}
+\DocumentationTok{\#\# [1] 0.128}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\CommentTok{# Plot posteriors for only a subset of species}
-\NormalTok{sp_to_plot <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"litu"}\NormalTok{, }\StringTok{"quru"}\NormalTok{, }\StringTok{"cagl"}\NormalTok{)}
-
-\NormalTok{plot1 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_scbi, }\DataTypeTok{type =} \StringTok{"intercepts"}\NormalTok{, }
-                  \DataTypeTok{sp_to_plot =}\NormalTok{ sp_to_plot)}
-\NormalTok{plot2 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_scbi, }\DataTypeTok{type =} \StringTok{"dbh_slopes"}\NormalTok{, }
-                  \DataTypeTok{sp_to_plot =}\NormalTok{ sp_to_plot)}
-\NormalTok{plot3 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_scbi, }\DataTypeTok{type =} \StringTok{"competition"}\NormalTok{, }
-                  \DataTypeTok{sp_to_plot =}\NormalTok{ sp_to_plot)}
-
-\CommentTok{# Combine plots using the patchwork package}
-\NormalTok{(plot1 }\OperatorTok{|}\StringTok{ }\NormalTok{plot2) }\OperatorTok{/}\StringTok{ }\NormalTok{plot3}
+\CommentTok{\# Plot posteriors for only a subset of species}
+\NormalTok{sp\_to\_plot }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\StringTok{"litu"}\NormalTok{, }\StringTok{"quru"}\NormalTok{, }\StringTok{"cagl"}\NormalTok{)}
+
+\NormalTok{plot1 }\OtherTok{\textless{}{-}} \FunctionTok{autoplot}\NormalTok{(comp\_bayes\_lm\_scbi, }\AttributeTok{type =} \StringTok{"intercepts"}\NormalTok{, }\AttributeTok{sp\_to\_plot =}\NormalTok{ sp\_to\_plot)}
+\NormalTok{plot2 }\OtherTok{\textless{}{-}} \FunctionTok{autoplot}\NormalTok{(comp\_bayes\_lm\_scbi, }\AttributeTok{type =} \StringTok{"dbh\_slopes"}\NormalTok{, }\AttributeTok{sp\_to\_plot =}\NormalTok{ sp\_to\_plot)}
+\NormalTok{plot3 }\OtherTok{\textless{}{-}} \FunctionTok{autoplot}\NormalTok{(comp\_bayes\_lm\_scbi, }\AttributeTok{type =} \StringTok{"competition"}\NormalTok{, }\AttributeTok{sp\_to\_plot =}\NormalTok{ sp\_to\_plot)}
+
+\CommentTok{\# Combine plots using the patchwork package}
+\NormalTok{(plot1 }\SpecialCharTok{|}\NormalTok{ plot2)}\SpecialCharTok{/}\NormalTok{plot3}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{comp_bayes_lm_scbi_shuffle <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{comp_bayes_lm}\NormalTok{(}\DataTypeTok{prior_param =} \OtherTok{NULL}\NormalTok{, }\DataTypeTok{run_shuffle =} \OtherTok{TRUE}\NormalTok{)}
-
-\NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{growth_hat_shuffle =} \KeywordTok{predict}\NormalTok{(comp_bayes_lm_scbi_shuffle, }
-                                 \DataTypeTok{newdata =}\NormalTok{ focal_vs_comp_scbi))}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{model_rmse_shuffle <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat_shuffle) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
-\NormalTok{model_rmse_shuffle}
-\CommentTok{## [1] 0.131}
+\NormalTok{comp\_bayes\_lm\_scbi\_shuffle }\OtherTok{\textless{}{-}}\NormalTok{ focal\_vs\_comp\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{comp\_bayes\_lm}\NormalTok{(}\AttributeTok{prior\_param =} \ConstantTok{NULL}\NormalTok{, }\AttributeTok{run\_shuffle =} \ConstantTok{TRUE}\NormalTok{)}
+
+\NormalTok{focal\_vs\_comp\_scbi }\OtherTok{\textless{}{-}}\NormalTok{ focal\_vs\_comp\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{growth\_hat\_shuffle =} \FunctionTok{predict}\NormalTok{(comp\_bayes\_lm\_scbi\_shuffle, }\AttributeTok{newdata =}\NormalTok{ focal\_vs\_comp\_scbi))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{model\_rmse\_shuffle }\OtherTok{\textless{}{-}}\NormalTok{ focal\_vs\_comp\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{rmse}\NormalTok{(}\AttributeTok{truth =}\NormalTok{ growth, }\AttributeTok{estimate =}\NormalTok{ growth\_hat\_shuffle) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{pull}\NormalTok{(.estimate)}
+\NormalTok{model\_rmse\_shuffle}
+\DocumentationTok{\#\# [1] 0.131}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{model_rmse_cv <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
-\NormalTok{model_rmse_cv}
-\CommentTok{## [1] 0.14}
+\NormalTok{focal\_vs\_comp\_scbi }\OtherTok{\textless{}{-}}\NormalTok{ focal\_vs\_comp\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{run\_cv}\NormalTok{(}\AttributeTok{comp\_dist =}\NormalTok{ comp\_dist, }\AttributeTok{blocks =}\NormalTok{ blocks\_scbi)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{model\_rmse\_cv }\OtherTok{\textless{}{-}}\NormalTok{ focal\_vs\_comp\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{rmse}\NormalTok{(}\AttributeTok{truth =}\NormalTok{ growth, }\AttributeTok{estimate =}\NormalTok{ growth\_hat) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{pull}\NormalTok{(.estimate)}
+\NormalTok{model\_rmse\_cv}
+\DocumentationTok{\#\# [1] 0.14}
 \end{Highlighting}
@@
 
+Lastly, currently the package only implements the Bayesian linear
+regression model detailed in Equation \ref{eq:model}. As we demonstrate
+in Section \ref{model-fit-predict} however, the fitting of this model is
+self-contained in a single function \texttt{comp\_bayes\_lm()} which
+returns an object of S3 class type \texttt{comp\_bayes\_lm}. This class
+has generic methods implemented to print, make predictions, and plot all
+results. Therefore the package can be modularly extended to fit other
+models as long as they are coded similarly to \texttt{comp\_bayes\_lm()}
+and have equivalent generic methods implemented.
+
 \hypertarget{acknowledgments}{%
@@
 Zenodo upon acceptance. The example Smithsonian Conservation Biology
-Institute census data are available on GitHub at
-\url{https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/tree/master/tree_main_census/data/census-csv-files}
+Institute census and species information data are available on GitHub at
+\url{https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/tree/master/}
 and are archived on Zenodo at
@@
 
-\hypertarget{appendix-1-replicate-rmse-comparison}{%
-\section{Appendix 1: Replicate RMSE
-comparison}\label{appendix-1-replicate-rmse-comparison}}
-
-This code replicates Figure \ref{fig:scbi-simulation}: A comparison of
-root mean squared error of models for standard, permuted, and spatial
-cross- validated error estimates.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{library}\NormalTok{(tidyverse)}
-\KeywordTok{library}\NormalTok{(lubridate)}
-\KeywordTok{library}\NormalTok{(here)}
-\KeywordTok{library}\NormalTok{(sf)}
-\KeywordTok{library}\NormalTok{(viridis)}
-\KeywordTok{library}\NormalTok{(forestecology)}
-\KeywordTok{library}\NormalTok{(blockCV)}
-\KeywordTok{library}\NormalTok{(tictoc)}
-
-\CommentTok{# Compute growth of trees based on census data -------------------------}
-\NormalTok{census_}\DecValTok{2013}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{here}\NormalTok{(}\StringTok{"paper/scbi.stem2.csv"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{read_csv}\NormalTok{() }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
-    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date),}
-    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}
-\NormalTok{  ) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
-
-\NormalTok{census_}\DecValTok{2018}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{here}\NormalTok{(}\StringTok{"paper/scbi.stem3.csv"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{read_csv}\NormalTok{() }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
-    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date),}
-    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}
-\NormalTok{  ) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
-
-\NormalTok{growth_scbi <-}
-\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(}
-    \DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi,}
-    \DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)),}
-    \DataTypeTok{id =} \StringTok{"stemID"}
-\NormalTok{  ) }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\CommentTok{# Compute basal area:}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{basal_area =} \FloatTok{0.0001} \OperatorTok{*}\StringTok{ }\NormalTok{pi }\OperatorTok{*}\StringTok{ }\NormalTok{(dbh1 }\OperatorTok{/}\StringTok{ }\DecValTok{2}\NormalTok{)}\OperatorTok{^}\DecValTok{2}\NormalTok{)}
-
-\CommentTok{# Add spatial information ----------------------------------------------}
-\CommentTok{# Define buffer region using competitive distance range}
-\NormalTok{comp_dist <-}\StringTok{ }\FloatTok{7.5}
-
-\NormalTok{study_region_scbi <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}
-  \DataTypeTok{x =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{),}
-  \DataTypeTok{y =} \KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{300}\NormalTok{)}
-\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{sf_polygon}\NormalTok{()}
-
-\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{size =}\NormalTok{ comp_dist, }\DataTypeTok{region =}\NormalTok{ study_region_scbi)}
-
-\CommentTok{# Manually define spatial blocks to act as folds}
-\NormalTok{fold1 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{450}\NormalTok{))}
-\NormalTok{fold2 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{450}\NormalTok{))}
-\NormalTok{fold3 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{600}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{600}\NormalTok{))}
-\NormalTok{fold4 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{600}\NormalTok{))}
-\NormalTok{n_fold <-}\StringTok{ }\DecValTok{4}
-
-\NormalTok{blocks_scbi <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}
-  \KeywordTok{sf_polygon}\NormalTok{(fold1), }\KeywordTok{sf_polygon}\NormalTok{(fold2),}
-  \KeywordTok{sf_polygon}\NormalTok{(fold3), }\KeywordTok{sf_polygon}\NormalTok{(fold4)}
-\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\NormalTok{n_fold) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
-
-\CommentTok{# Associate each observation to a fold}
-\NormalTok{SpatialBlock_scbi <-}\StringTok{ }\KeywordTok{spatialBlock}\NormalTok{(}
-  \DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =}\NormalTok{ n_fold, }\DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{,}
-  \DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}
-\NormalTok{)}
-
-\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ SpatialBlock_scbi}\OperatorTok{$}\NormalTok{foldID }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
-
-\CommentTok{# Compute focal versus competitor tree information ---------------------}
-\NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{, }\DataTypeTok{comp_x_var =} \StringTok{"basal_area"}\NormalTok{)}
-
-\CommentTok{# Fit model and make predictions ---------------------------------------}
-\CommentTok{# Number of permutation shuffles:}
-\NormalTok{num_shuffle <-}\StringTok{ }\DecValTok{49}
-
-\CommentTok{# Save results here}
-\NormalTok{run_time <-}\StringTok{ }\DecValTok{0}
-\NormalTok{observed_RMSE <-}\StringTok{ }\DecValTok{0}
-\NormalTok{observed_RMSE_CV <-}\StringTok{ }\DecValTok{0}
-\NormalTok{shuffle_RMSE <-}\StringTok{ }\KeywordTok{vector}\NormalTok{(}\StringTok{"list"}\NormalTok{, }\DecValTok{1}\NormalTok{)}
-\NormalTok{shuffle_RMSE_CV <-}\StringTok{ }\KeywordTok{vector}\NormalTok{(}\StringTok{"list"}\NormalTok{, }\DecValTok{1}\NormalTok{)}
-\NormalTok{filename <-}\StringTok{ }\KeywordTok{here}\NormalTok{(}\StringTok{"paper/simulation_results/"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{str_c}\NormalTok{(}\StringTok{"2021-03-03_scbi_"}\NormalTok{, num_shuffle, }\StringTok{"_shuffles"}\NormalTok{)}
-
-\CommentTok{# Run all simulations}
-\CommentTok{# 0. Setup simulation for this species type ----}
-\CommentTok{# Start clock}
-\KeywordTok{tic}\NormalTok{()}
-
-\CommentTok{# 1. Compute observed test statistic: RMSE with no cross-validation ----}
-\CommentTok{# Fit model (compute posterior parameters)}
-\NormalTok{comp_bayes_lm_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{comp_bayes_lm}\NormalTok{(}\DataTypeTok{prior_param =} \OtherTok{NULL}\NormalTok{, }\DataTypeTok{run_shuffle =} \OtherTok{FALSE}\NormalTok{)}
-
-\CommentTok{# Make predictions and compute RMSE}
-\NormalTok{observed_RMSE <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{growth_hat =} 
-           \KeywordTok{predict}\NormalTok{(comp_bayes_lm_scbi, focal_vs_comp_scbi)) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
-
-\CommentTok{# 2. Compute observed test statistic: RMSE with cross-validation -------}
-\NormalTok{observed_RMSE_CV <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{pull}\NormalTok{(.estimate)}
-
-\CommentTok{# 3. Permutation distribution: RMSE with no cross-validation -----------}
-\CommentTok{# Compute num_shuffle permutation test statistics}
-\NormalTok{shuffle_RMSE <-}\StringTok{ }\KeywordTok{numeric}\NormalTok{(}\DataTypeTok{length =}\NormalTok{ num_shuffle)}
-\ControlFlowTok{for}\NormalTok{(j }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\NormalTok{num_shuffle)\{}
-  \CommentTok{# Fit model (compute posterior parameters) with shuffling}
-\NormalTok{  comp_bayes_lm_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{comp_bayes_lm}\NormalTok{(}\DataTypeTok{prior_param =} \OtherTok{NULL}\NormalTok{, }\DataTypeTok{run_shuffle =} \OtherTok{TRUE}\NormalTok{)}
-
-  \CommentTok{# Make predictions and compute RMSE}
-\NormalTok{  shuffle_RMSE[j] <-}\StringTok{  }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{mutate}\NormalTok{(}
-      \DataTypeTok{growth_hat =} \KeywordTok{predict}\NormalTok{(comp_bayes_lm_scbi, focal_vs_comp_scbi)}
-\NormalTok{    ) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{pull}\NormalTok{(.estimate)}
-\NormalTok{\}}
-
-\CommentTok{# 4. Permutation distribution: RMSE with cross-validation --------------}
-\CommentTok{# Compute num_shuffle permutation test statistics}
-\NormalTok{shuffle_RMSE_CV <-}\StringTok{ }\KeywordTok{numeric}\NormalTok{(}\DataTypeTok{length =}\NormalTok{ num_shuffle)}
-
-\CommentTok{# Compute num_shuffle permutation test statistics}
-\ControlFlowTok{for}\NormalTok{(j }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\NormalTok{num_shuffle)\{}
-  \CommentTok{# Compute and save RMSE}
-\NormalTok{  shuffle_RMSE_CV[j] <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }
-           \DataTypeTok{run_shuffle =} \OtherTok{TRUE}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{pull}\NormalTok{(.estimate)}
-
-  \CommentTok{# Status update}
-  \KeywordTok{str_c}\NormalTok{(}\StringTok{"Shuffle with permutation "}\NormalTok{, j, }\StringTok{" at "}\NormalTok{, }\KeywordTok{Sys.time}\NormalTok{()) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{print}\NormalTok{()}
-\NormalTok{\}}
-
-\CommentTok{# 5. Save results ----}
-\NormalTok{clock <-}\StringTok{ }\KeywordTok{toc}\NormalTok{(}\DataTypeTok{quiet =} \OtherTok{TRUE}\NormalTok{)}
-\NormalTok{run_time <-}\StringTok{ }\NormalTok{clock}\OperatorTok{$}\NormalTok{toc }\OperatorTok{-}\StringTok{ }\NormalTok{clock}\OperatorTok{$}\NormalTok{tic}
-
-\NormalTok{model_comp_tbl <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}
-  \DataTypeTok{run_time =}\NormalTok{ run_time,}
-  \DataTypeTok{observed_RMSE =}\NormalTok{ observed_RMSE,}
-  \DataTypeTok{observed_RMSE_CV =}\NormalTok{ observed_RMSE_CV,}
-  \DataTypeTok{shuffle_RMSE =}\NormalTok{ shuffle_RMSE,}
-  \DataTypeTok{shuffle_RMSE_CV =}\NormalTok{ shuffle_RMSE_CV,}
-\NormalTok{)}
-\KeywordTok{save}\NormalTok{(model_comp_tbl, }\DataTypeTok{file =}\NormalTok{ filename }\OperatorTok{%>%}\StringTok{ }\KeywordTok{str_c}\NormalTok{(}\StringTok{".RData"}\NormalTok{))}
-
-\CommentTok{# Visualize results ----------------------------------------------------}
-\NormalTok{model_comp <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}
-\NormalTok{  model_comp_tbl }\OperatorTok{%>%}\StringTok{ }
-\StringTok{    }\KeywordTok{select}\NormalTok{(run_time, }\DataTypeTok{observed =}\NormalTok{ observed_RMSE, }
-           \DataTypeTok{shuffle =}\NormalTok{ shuffle_RMSE) }\OperatorTok{%>%}\StringTok{ }
-\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{CV =} \OtherTok{FALSE}\NormalTok{),}
-\NormalTok{  model_comp_tbl }\OperatorTok{%>%}\StringTok{ }
-\StringTok{    }\KeywordTok{select}\NormalTok{(run_time, }\DataTypeTok{observed =}\NormalTok{ observed_RMSE_CV, }
-           \DataTypeTok{shuffle =}\NormalTok{ shuffle_RMSE_CV) }\OperatorTok{%>%}\StringTok{ }
-\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{CV =} \OtherTok{TRUE}\NormalTok{)}
-\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{gather}\NormalTok{(type, RMSE, }\OperatorTok{-}\KeywordTok{c}\NormalTok{(run_time, CV))}
-
-\NormalTok{model_comp_observed <-}\StringTok{ }\NormalTok{model_comp }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{filter}\NormalTok{(type }\OperatorTok{==}\StringTok{ "observed"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{unnest}\NormalTok{(}\DataTypeTok{cols =} \KeywordTok{c}\NormalTok{(RMSE))}
-\NormalTok{model_comp_shuffle <-}\StringTok{ }\NormalTok{model_comp }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{filter}\NormalTok{(type }\OperatorTok{==}\StringTok{ "shuffle"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{unnest}\NormalTok{(}\DataTypeTok{cols =} \KeywordTok{c}\NormalTok{(RMSE))}
-
-\NormalTok{cv_plot <-}\StringTok{ }\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_vline}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ model_comp_observed, }
-             \KeywordTok{aes}\NormalTok{(}\DataTypeTok{xintercept =}\NormalTok{ RMSE, }\DataTypeTok{col =}\NormalTok{ CV), }
-             \DataTypeTok{linetype =} \StringTok{"dashed"}\NormalTok{, }\DataTypeTok{show.legend =}\NormalTok{ F) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_histogram}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ model_comp_shuffle, }
-                 \KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ RMSE, }\DataTypeTok{fill =}\NormalTok{ CV), }\DataTypeTok{bins =} \DecValTok{50}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{labs}\NormalTok{(}
-    \DataTypeTok{fill =} \StringTok{"Cross-validated?"}\NormalTok{,}
-    \DataTypeTok{x =} \KeywordTok{expression}\NormalTok{(}\KeywordTok{paste}\NormalTok{(}\StringTok{"RMSE (cm "}\NormalTok{, y}\OperatorTok{^}\NormalTok{\{}\OperatorTok{-}\DecValTok{1}\NormalTok{\}, }\StringTok{")"}\NormalTok{))}
-\NormalTok{  ) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{scale_color_viridis}\NormalTok{(}\DataTypeTok{discrete =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{option =} \StringTok{"D"}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{scale_fill_viridis}\NormalTok{(}\DataTypeTok{discrete =} \OtherTok{TRUE}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{theme_light}\NormalTok{()}
-\NormalTok{cv_plot}
-
-\NormalTok{filename }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{str_c}\NormalTok{(}\StringTok{".pdf"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{ggsave}\NormalTok{(}\DataTypeTok{plot =}\NormalTok{ cv_plot, }\DataTypeTok{width =} \DecValTok{16}\OperatorTok{/}\DecValTok{2}\NormalTok{, }\DataTypeTok{height =} \DecValTok{9}\OperatorTok{/}\DecValTok{2}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
+\hypertarget{refs}{}
+\begin{CSLReferences}{0}{0}
+\end{CSLReferences}
+
+\appendix
 
 \hypertarget{mult-dist}{%
-\section{Appendix 2: Compare different competitive
+\section{Appendix: Compare different competitive
 distances}\label{mult-dist}}
 
-For all the above analyses we took the cut off distance
-(\texttt{comp\_dist}) for two stems to compete as 7.5 m. This distance
+For all the above analyses we set the cut off distance
+(\texttt{comp\_dist}) for two stems to compete as 7.5m. This distance
 has been estimated in previous neighborhood competition studies in
 forests \citep[\citet{uriarte_spatially_2004},
-\citet{canham_neighborhood_2006}]{canham_neighborhood_2004}. We used 7.5
-m in \citet{allen_permutation_2020} as an average of the values
+\citet{canham_neighborhood_2006}]{canham_neighborhood_2004}. We used
+7.5m in \citet{allen_permutation_2020} as an average of the values
 estimated in other studies. But our package can be used to find which
@@
 another section of the SCBI plot to provide an additional example of the
-CV block layout. To speed computation we do not consider species
-differences in competitive effects and treat all species as the same.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{census_}\DecValTok{2013}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem2.csv"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
-    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date),}
-    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}
-\NormalTok{  ) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{400}\NormalTok{, gy }\OperatorTok{<}\StringTok{ }\DecValTok{400}\NormalTok{)}
-
-\NormalTok{census_}\DecValTok{2018}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem3.csv"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
-    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date),}
-    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}
-\NormalTok{  ) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{400}\NormalTok{, gy }\OperatorTok{<}\StringTok{ }\DecValTok{400}\NormalTok{)}
-
-\NormalTok{growth_scbi <-}
-\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(}
-    \DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi,}
-    \DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)),}
-    \DataTypeTok{id =} \StringTok{"stemID"}
-\NormalTok{  ) }\OperatorTok{%>%}
-\StringTok{  }\CommentTok{# make all species the same, needs to be factor with at least two levels}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{factor}\NormalTok{(}\StringTok{'A'}\NormalTok{, }\DataTypeTok{levels =} \KeywordTok{c}\NormalTok{(}\StringTok{'A'}\NormalTok{, }\StringTok{'B'}\NormalTok{)) ) }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\CommentTok{# Compute basal area:}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{basal_area =} \FloatTok{0.0001} \OperatorTok{*}\StringTok{ }\NormalTok{pi }\OperatorTok{*}\StringTok{ }\NormalTok{(dbh1 }\OperatorTok{/}\StringTok{ }\DecValTok{2}\NormalTok{)}\OperatorTok{^}\DecValTok{2}\NormalTok{)}
-
-\NormalTok{study_region_scbi <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}
-  \DataTypeTok{x =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{),}
-  \DataTypeTok{y =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{0}\NormalTok{)}
-\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{sf_polygon}\NormalTok{()}
-
-\NormalTok{n_fold <-}\StringTok{ }\DecValTok{4}
-\NormalTok{fold1 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{0}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{200}\NormalTok{))}
-\NormalTok{fold2 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{200}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{200}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{200}\NormalTok{))}
-\NormalTok{fold3 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{0}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{200}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{400}\NormalTok{))}
-\NormalTok{fold4 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{200}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{200}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{200}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{400}\NormalTok{))}
-
-\NormalTok{blocks_scbi <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}
-  \KeywordTok{sf_polygon}\NormalTok{(fold1), }\KeywordTok{sf_polygon}\NormalTok{(fold2), }\KeywordTok{sf_polygon}\NormalTok{(fold3), }
-  \KeywordTok{sf_polygon}\NormalTok{(fold4)}
-\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\NormalTok{n_fold) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
-
-\CommentTok{# Associate each observation to a fold}
-\NormalTok{spatial_block_scbi <-}\StringTok{ }
-\StringTok{  }\KeywordTok{spatialBlock}\NormalTok{(}\DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =}\NormalTok{ n_fold, }
-               \DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }
-               \DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}\NormalTok{)}
-
-\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ spatial_block_scbi}\OperatorTok{$}\NormalTok{foldID }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
-
-\NormalTok{mult_dist_comp <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}
-  \DataTypeTok{dist =} \KeywordTok{c}\NormalTok{(}\DecValTok{5}\NormalTok{ ,}\FloatTok{6.25}\NormalTok{,  }\FloatTok{7.5}\NormalTok{, }\FloatTok{8.75}\NormalTok{,  }\DecValTok{10}\NormalTok{),}
-  \DataTypeTok{rmse =} \DecValTok{0}
+cross-validation block layout. To speed computation we do not consider
+species differences in competitive effects and treat all species as the
+same.
+
+We observe in Figure \ref{fig:appendix2-plot} that a cut-off distance of
+approximately 6m minimizes the cross-validation estimated RMSE.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{census\_2013\_scbi }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"scbi.stem2.csv"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{select}\NormalTok{(stemID, sp, }\AttributeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}
+    \AttributeTok{date =} \FunctionTok{mdy}\NormalTok{(date),}
+    \AttributeTok{dbh =} \FunctionTok{as.numeric}\NormalTok{(dbh) }\SpecialCharTok{/} \DecValTok{10}
+\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{filter}\NormalTok{(gx }\SpecialCharTok{\textless{}} \DecValTok{400}\NormalTok{, gy }\SpecialCharTok{\textless{}} \DecValTok{400}\NormalTok{)}
+
+\NormalTok{census\_2018\_scbi }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"scbi.stem3.csv"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{select}\NormalTok{(stemID, sp, }\AttributeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}
+    \AttributeTok{date =} \FunctionTok{mdy}\NormalTok{(date),}
+    \AttributeTok{dbh =} \FunctionTok{as.numeric}\NormalTok{(dbh) }\SpecialCharTok{/} \DecValTok{10}
+\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{filter}\NormalTok{(gx }\SpecialCharTok{\textless{}} \DecValTok{400}\NormalTok{, gy }\SpecialCharTok{\textless{}} \DecValTok{400}\NormalTok{)}
+
+\NormalTok{growth\_scbi }\OtherTok{\textless{}{-}}
+  \FunctionTok{compute\_growth}\NormalTok{(}
+    \AttributeTok{census\_1 =}\NormalTok{ census\_2013\_scbi,}
+    \AttributeTok{census\_2 =}\NormalTok{ census\_2018\_scbi }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{str\_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)),}
+    \AttributeTok{id =} \StringTok{"stemID"}
+\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
+  \CommentTok{\# make all species the same, needs to be factor with at least two levels}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{sp =} \FunctionTok{factor}\NormalTok{(}\StringTok{"A"}\NormalTok{, }\AttributeTok{levels =} \FunctionTok{c}\NormalTok{(}\StringTok{"A"}\NormalTok{, }\StringTok{"B"}\NormalTok{))) }\SpecialCharTok{\%\textgreater{}\%}
+  \CommentTok{\# Compute basal area:}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{basal\_area =} \FloatTok{0.0001} \SpecialCharTok{*}\NormalTok{ pi }\SpecialCharTok{*}\NormalTok{ (dbh1 }\SpecialCharTok{/} \DecValTok{2}\NormalTok{)}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{)}
+
+\NormalTok{study\_region\_scbi }\OtherTok{\textless{}{-}} \FunctionTok{tibble}\NormalTok{(}
+  \AttributeTok{x =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{),}
+  \AttributeTok{y =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{0}\NormalTok{)}
+\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{sf\_polygon}\NormalTok{()}
+
+\NormalTok{n\_fold }\OtherTok{\textless{}{-}} \DecValTok{4}
+\NormalTok{fold1 }\OtherTok{\textless{}{-}} \FunctionTok{cbind}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{0}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{200}\NormalTok{))}
+\NormalTok{fold2 }\OtherTok{\textless{}{-}} \FunctionTok{cbind}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\DecValTok{200}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{200}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{200}\NormalTok{))}
+\NormalTok{fold3 }\OtherTok{\textless{}{-}} \FunctionTok{cbind}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{0}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{200}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{400}\NormalTok{))}
+\NormalTok{fold4 }\OtherTok{\textless{}{-}} \FunctionTok{cbind}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\DecValTok{200}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{200}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{200}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{400}\NormalTok{))}
+
+\NormalTok{blocks\_scbi }\OtherTok{\textless{}{-}} \FunctionTok{bind\_rows}\NormalTok{(}
+  \FunctionTok{sf\_polygon}\NormalTok{(fold1), }\FunctionTok{sf\_polygon}\NormalTok{(fold2), }\FunctionTok{sf\_polygon}\NormalTok{(fold3),}
+  \FunctionTok{sf\_polygon}\NormalTok{(fold4)}
+\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{folds =} \FunctionTok{c}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\NormalTok{n\_fold) }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{factor}\NormalTok{())}
+
+\CommentTok{\# Associate each observation to a fold}
+\NormalTok{spatial\_block\_scbi }\OtherTok{\textless{}{-}}
+  \FunctionTok{spatialBlock}\NormalTok{(}
+    \AttributeTok{speciesData =}\NormalTok{ growth\_scbi, }\AttributeTok{k =}\NormalTok{ n\_fold,}
+    \AttributeTok{selection =} \StringTok{"systematic"}\NormalTok{, }\AttributeTok{blocks =}\NormalTok{ blocks\_scbi,}
+    \AttributeTok{showBlocks =} \ConstantTok{FALSE}\NormalTok{, }\AttributeTok{verbose =} \ConstantTok{FALSE}
+\NormalTok{  )}
+
+\NormalTok{growth\_scbi }\OtherTok{\textless{}{-}}\NormalTok{ growth\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{foldID =}\NormalTok{ spatial\_block\_scbi}\SpecialCharTok{$}\NormalTok{foldID }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{factor}\NormalTok{())}
+
+\NormalTok{mult\_dist\_comp }\OtherTok{\textless{}{-}} \FunctionTok{tibble}\NormalTok{(}
+  \AttributeTok{dist =} \FunctionTok{c}\NormalTok{(}\DecValTok{5}\NormalTok{, }\FloatTok{6.25}\NormalTok{, }\FloatTok{7.5}\NormalTok{, }\FloatTok{8.75}\NormalTok{, }\DecValTok{10}\NormalTok{),}
+  \AttributeTok{rmse =} \DecValTok{0}
 \NormalTok{)}
 
-\ControlFlowTok{for}\NormalTok{ (i }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(mult_dist_comp}\OperatorTok{$}\NormalTok{dist))\{}
-\NormalTok{  comp_dist <-}\StringTok{ }\NormalTok{mult_dist_comp}\OperatorTok{$}\NormalTok{dist[i]}
-
-\NormalTok{  growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{size =}\NormalTok{ comp_dist, }\DataTypeTok{region =}\NormalTok{ study_region_scbi)}
-
-\NormalTok{  focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{, }\DataTypeTok{comp_x_var =} \StringTok{"basal_area"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi)}
-
-\NormalTok{  mult_dist_comp}\OperatorTok{$}\NormalTok{rmse[i] <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{pull}\NormalTok{(.estimate)}
+\ControlFlowTok{for}\NormalTok{ (i }\ControlFlowTok{in} \DecValTok{1}\SpecialCharTok{:}\FunctionTok{length}\NormalTok{(mult\_dist\_comp}\SpecialCharTok{$}\NormalTok{dist)) \{}
+\NormalTok{  comp\_dist }\OtherTok{\textless{}{-}}\NormalTok{ mult\_dist\_comp}\SpecialCharTok{$}\NormalTok{dist[i]}
+
+\NormalTok{  growth\_scbi }\OtherTok{\textless{}{-}}\NormalTok{ growth\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{add\_buffer\_variable}\NormalTok{(}\AttributeTok{size =}\NormalTok{ comp\_dist, }\AttributeTok{region =}\NormalTok{ study\_region\_scbi)}
+
+\NormalTok{  focal\_vs\_comp\_scbi }\OtherTok{\textless{}{-}}\NormalTok{ growth\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{create\_focal\_vs\_comp}\NormalTok{(}\AttributeTok{comp\_dist =}\NormalTok{ comp\_dist, }\AttributeTok{blocks =}\NormalTok{ blocks\_scbi, }\AttributeTok{id =} \StringTok{"stemID"}\NormalTok{, }\AttributeTok{comp\_x\_var =} \StringTok{"basal\_area"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{run\_cv}\NormalTok{(}\AttributeTok{comp\_dist =}\NormalTok{ comp\_dist, }\AttributeTok{blocks =}\NormalTok{ blocks\_scbi)}
+
+\NormalTok{  mult\_dist\_comp}\SpecialCharTok{$}\NormalTok{rmse[i] }\OtherTok{\textless{}{-}}\NormalTok{ focal\_vs\_comp\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{rmse}\NormalTok{(}\AttributeTok{truth =}\NormalTok{ growth, }\AttributeTok{estimate =}\NormalTok{ growth\_hat) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{pull}\NormalTok{(.estimate)}
 \NormalTok{\}}
-
-\NormalTok{mult_dist_comp }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(dist,rmse)) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_point}\NormalTok{() }\OperatorTok{+}
-\StringTok{  }\KeywordTok{xlab}\NormalTok{(}\StringTok{'Distance (m)'}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{ylab}\NormalTok{(}\StringTok{'RMSE (cm)'}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-Here we see that a cut off distance of approximately 6 m minimizes RMSE.
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{Figures/appendix2-plot-1}
+
+}
+
+\caption{Cross-validated RMSE estimates for 5 competitive distances2.}\label{fig:appendix2-plot}
+\end{figure}
 
 \hypertarget{compare-pred-var}{%
-\section{Appendix 3: Compare predictor
+\section{Appendix: Compare competitor explanatory
 variables}\label{compare-pred-var}}
@@
 In the above code we use the basal area of an individual as a continuous
-predictor variable. But the package allows the user to specify any
-predictor variable in the \texttt{comp\_x\_var} argument of
-\texttt{create\_focal\_vs\_comp} function. Here we use the
-cross-validated model comparison to see which of two possible predictor
-variables, basal area or above ground biomass, best explains growth.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{census_}\DecValTok{2013}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem2.csv"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
-    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date),}
-    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}
-\NormalTok{  ) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
-
-\NormalTok{census_}\DecValTok{2018}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem3.csv"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
-    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date),}
-    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}
-\NormalTok{  ) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
-
-\NormalTok{growth_scbi <-}
-\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(}
-    \DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi,}
-    \DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)),}
-    \DataTypeTok{id =} \StringTok{"stemID"}
-\NormalTok{  ) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{left_join}\NormalTok{(sp_info, }\DataTypeTok{by =} \StringTok{"sp"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{as.factor}\NormalTok{(sp),}
-        \DataTypeTok{basal_area =} \FloatTok{0.0001} \OperatorTok{*}\StringTok{ }\NormalTok{pi }\OperatorTok{*}\StringTok{ }\NormalTok{(dbh1 }\OperatorTok{/}\StringTok{ }\DecValTok{2}\NormalTok{)}\OperatorTok{^}\DecValTok{2}\NormalTok{,}
-         \DataTypeTok{agb =} \KeywordTok{get_biomass}\NormalTok{(}
-                  \DataTypeTok{dbh =}\NormalTok{ dbh1,}
-                  \DataTypeTok{genus =}\NormalTok{ genus,}
-                  \DataTypeTok{species =}\NormalTok{ species,}
-                  \DataTypeTok{coords =} \KeywordTok{c}\NormalTok{(}\OperatorTok{-}\FloatTok{78.2}\NormalTok{, }\FloatTok{38.9}\NormalTok{)}
-\NormalTok{    ))}
-
-\NormalTok{study_region_scbi <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}
-  \DataTypeTok{x =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{),}
-  \DataTypeTok{y =} \KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{300}\NormalTok{)}
-\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{sf_polygon}\NormalTok{()}
-
-\NormalTok{n_fold <-}\StringTok{ }\DecValTok{4}
-\NormalTok{fold1 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{0}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{))}
-\NormalTok{fold2 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{150}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{))}
-\NormalTok{fold3 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{0}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{))}
-\NormalTok{fold4 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{150}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{))}
-
-\NormalTok{blocks_scbi <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}
-  \KeywordTok{sf_polygon}\NormalTok{(fold1), }\KeywordTok{sf_polygon}\NormalTok{(fold2), }\KeywordTok{sf_polygon}\NormalTok{(fold3), }
-  \KeywordTok{sf_polygon}\NormalTok{(fold4)}
-\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\NormalTok{n_fold) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
-
-\CommentTok{# Associate each observation to a fold}
-\NormalTok{spatial_block_scbi <-}\StringTok{ }
-\StringTok{  }\KeywordTok{spatialBlock}\NormalTok{(}\DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =}\NormalTok{ n_fold, }
-               \DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }
-               \DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}\NormalTok{)}
-
-\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ spatial_block_scbi}\OperatorTok{$}\NormalTok{foldID }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
-
-\NormalTok{comp_dist <-}\StringTok{ }\FloatTok{7.5}
-
-\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{size =}\NormalTok{ comp_dist, }\DataTypeTok{region =}\NormalTok{ study_region_scbi)}
-
-\NormalTok{focal_vs_comp_ba <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }
-                         \DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }
-                         \DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{, }
-                         \DataTypeTok{comp_x_var =} \StringTok{"basal_area"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi)}
-
-\NormalTok{focal_vs_comp_agb <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }
-                         \DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }
-                         \DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{, }
-                         \DataTypeTok{comp_x_var =} \StringTok{"agb"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi)}
-
-\NormalTok{  focal_vs_comp_ba }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{pull}\NormalTok{(.estimate)}
-\CommentTok{## [1] 0.14}
-
-\NormalTok{  focal_vs_comp_agb }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{pull}\NormalTok{(.estimate)}
-\CommentTok{## [1] 1.83}
-\end{Highlighting}
-\end{Shaded}
-
-Here we see that basal area does a much better job than above ground
-biomass.
+competitor explanatory variable. But the package allows the user to
+specify any competitor explanatory variable in the \texttt{comp\_x\_var}
+argument of \texttt{create\_focal\_vs\_comp} function. Here we use the
+cross-validated model comparison to see which of two possible competitor
+explanatory variables computed in Section \ref{compute-growth}, basal
+area or above ground biomass, best explains growth.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{census\_2013\_scbi }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"scbi.stem2.csv"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{select}\NormalTok{(stemID, sp, }\AttributeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{date =} \FunctionTok{mdy}\NormalTok{(date), }\AttributeTok{dbh =} \FunctionTok{as.numeric}\NormalTok{(dbh)}\SpecialCharTok{/}\DecValTok{10}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{filter}\NormalTok{(gx }\SpecialCharTok{\textless{}} \DecValTok{300}\NormalTok{, }\FunctionTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+
+\NormalTok{census\_2018\_scbi }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"scbi.stem3.csv"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{select}\NormalTok{(stemID, sp, }\AttributeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{date =} \FunctionTok{mdy}\NormalTok{(date), }\AttributeTok{dbh =} \FunctionTok{as.numeric}\NormalTok{(dbh)}\SpecialCharTok{/}\DecValTok{10}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{filter}\NormalTok{(gx }\SpecialCharTok{\textless{}} \DecValTok{300}\NormalTok{, }\FunctionTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+
+\NormalTok{growth\_scbi }\OtherTok{\textless{}{-}} \FunctionTok{compute\_growth}\NormalTok{(}\AttributeTok{census\_1 =}\NormalTok{ census\_2013\_scbi, }\AttributeTok{census\_2 =}\NormalTok{ census\_2018\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{filter}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{str\_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)), }\AttributeTok{id =} \StringTok{"stemID"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{left\_join}\NormalTok{(sp\_info, }\AttributeTok{by =} \StringTok{"sp"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{sp =} \FunctionTok{as.factor}\NormalTok{(sp), }\AttributeTok{basal\_area =} \FloatTok{1e{-}04} \SpecialCharTok{*}\NormalTok{ pi }\SpecialCharTok{*}\NormalTok{ (dbh1}\SpecialCharTok{/}\DecValTok{2}\NormalTok{)}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{, }\AttributeTok{agb =} \FunctionTok{get\_biomass}\NormalTok{(}\AttributeTok{dbh =}\NormalTok{ dbh1,}
+        \AttributeTok{genus =}\NormalTok{ genus, }\AttributeTok{species =}\NormalTok{ species, }\AttributeTok{coords =} \FunctionTok{c}\NormalTok{(}\SpecialCharTok{{-}}\FloatTok{78.2}\NormalTok{, }\FloatTok{38.9}\NormalTok{)))}
+
+\NormalTok{study\_region\_scbi }\OtherTok{\textless{}{-}} \FunctionTok{tibble}\NormalTok{(}\AttributeTok{x =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{), }\AttributeTok{y =} \FunctionTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{,}
+    \DecValTok{600}\NormalTok{, }\DecValTok{300}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{sf\_polygon}\NormalTok{()}
+
+\NormalTok{n\_fold }\OtherTok{\textless{}{-}} \DecValTok{4}
+\NormalTok{fold1 }\OtherTok{\textless{}{-}} \FunctionTok{cbind}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{0}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{))}
+\NormalTok{fold2 }\OtherTok{\textless{}{-}} \FunctionTok{cbind}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{150}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{))}
+\NormalTok{fold3 }\OtherTok{\textless{}{-}} \FunctionTok{cbind}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{0}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+\NormalTok{fold4 }\OtherTok{\textless{}{-}} \FunctionTok{cbind}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{150}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+
+\NormalTok{blocks\_scbi }\OtherTok{\textless{}{-}} \FunctionTok{bind\_rows}\NormalTok{(}\FunctionTok{sf\_polygon}\NormalTok{(fold1), }\FunctionTok{sf\_polygon}\NormalTok{(fold2), }\FunctionTok{sf\_polygon}\NormalTok{(fold3),}
+    \FunctionTok{sf\_polygon}\NormalTok{(fold4)) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{folds =} \FunctionTok{c}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\NormalTok{n\_fold) }\SpecialCharTok{\%\textgreater{}\%}
+        \FunctionTok{factor}\NormalTok{())}
+
+\CommentTok{\# Associate each observation to a fold}
+\NormalTok{spatial\_block\_scbi }\OtherTok{\textless{}{-}} \FunctionTok{spatialBlock}\NormalTok{(}\AttributeTok{speciesData =}\NormalTok{ growth\_scbi, }\AttributeTok{k =}\NormalTok{ n\_fold,}
+    \AttributeTok{selection =} \StringTok{"systematic"}\NormalTok{, }\AttributeTok{blocks =}\NormalTok{ blocks\_scbi, }\AttributeTok{showBlocks =} \ConstantTok{FALSE}\NormalTok{, }\AttributeTok{verbose =} \ConstantTok{FALSE}\NormalTok{)}
+
+\NormalTok{growth\_scbi }\OtherTok{\textless{}{-}}\NormalTok{ growth\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{foldID =}\NormalTok{ spatial\_block\_scbi}\SpecialCharTok{$}\NormalTok{foldID }\SpecialCharTok{\%\textgreater{}\%}
+        \FunctionTok{factor}\NormalTok{())}
+
+\NormalTok{comp\_dist }\OtherTok{\textless{}{-}} \FloatTok{7.5}
+
+\NormalTok{growth\_scbi }\OtherTok{\textless{}{-}}\NormalTok{ growth\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{add\_buffer\_variable}\NormalTok{(}\AttributeTok{size =}\NormalTok{ comp\_dist, }\AttributeTok{region =}\NormalTok{ study\_region\_scbi)}
+
+\NormalTok{focal\_vs\_comp\_ba }\OtherTok{\textless{}{-}}\NormalTok{ growth\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{create\_focal\_vs\_comp}\NormalTok{(}\AttributeTok{comp\_dist =}\NormalTok{ comp\_dist, }\AttributeTok{blocks =}\NormalTok{ blocks\_scbi, }\AttributeTok{id =} \StringTok{"stemID"}\NormalTok{,}
+        \AttributeTok{comp\_x\_var =} \StringTok{"basal\_area"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{run\_cv}\NormalTok{(}\AttributeTok{comp\_dist =}\NormalTok{ comp\_dist, }\AttributeTok{blocks =}\NormalTok{ blocks\_scbi)}
+
+\NormalTok{focal\_vs\_comp\_agb }\OtherTok{\textless{}{-}}\NormalTok{ growth\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{create\_focal\_vs\_comp}\NormalTok{(}\AttributeTok{comp\_dist =}\NormalTok{ comp\_dist, }\AttributeTok{blocks =}\NormalTok{ blocks\_scbi, }\AttributeTok{id =} \StringTok{"stemID"}\NormalTok{,}
+        \AttributeTok{comp\_x\_var =} \StringTok{"agb"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{run\_cv}\NormalTok{(}\AttributeTok{comp\_dist =}\NormalTok{ comp\_dist, }\AttributeTok{blocks =}\NormalTok{ blocks\_scbi)}
+
+\NormalTok{focal\_vs\_comp\_ba }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{rmse}\NormalTok{(}\AttributeTok{truth =}\NormalTok{ growth, }\AttributeTok{estimate =}\NormalTok{ growth\_hat) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{pull}\NormalTok{(.estimate)}
+\DocumentationTok{\#\# [1] 0.14}
+
+\NormalTok{focal\_vs\_comp\_agb }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{rmse}\NormalTok{(}\AttributeTok{truth =}\NormalTok{ growth, }\AttributeTok{estimate =}\NormalTok{ growth\_hat) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{pull}\NormalTok{(.estimate)}
+\DocumentationTok{\#\# [1] 2.08}
+\end{Highlighting}
+\end{Shaded}
+
+Here we observe that basal area is a better competitor explanatory
+variable competitor explanatory variables \(x^{\text{comp}}_{ijk}\) from
+Equation \ref{eq:model} than above ground biomass as suggested by the
+lower estimated RMSE.
 
 \hypertarget{compare-group-var}{%
-\section{Appendix 4: Compare grouping
-variables}\label{compare-group-var}}
-
-The package also allows the user to specify the categorical predictor
+\section{Appendix: Compare grouping variables}\label{compare-group-var}}
+
+The package also allows the user to specify the categorical explanatory
 grouping variable. Here we compare two different such variables: species
@@
 \begin{Highlighting}[]
-
-\NormalTok{census_}\DecValTok{2013}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem2.csv"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
-    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date),}
-    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}
-\NormalTok{  ) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
-
-\NormalTok{census_}\DecValTok{2018}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem3.csv"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
-    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date),}
-    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}
-\NormalTok{  ) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
-
-\NormalTok{growth_scbi_sp <-}
-\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(}
-    \DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi,}
-    \DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)),}
-    \DataTypeTok{id =} \StringTok{"stemID"}
-\NormalTok{  ) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{as.factor}\NormalTok{(sp),}
-        \DataTypeTok{basal_area =} \FloatTok{0.0001} \OperatorTok{*}\StringTok{ }\NormalTok{pi }\OperatorTok{*}\StringTok{ }\NormalTok{(dbh1 }\OperatorTok{/}\StringTok{ }\DecValTok{2}\NormalTok{)}\OperatorTok{^}\DecValTok{2}\NormalTok{ )}
-
-\NormalTok{growth_scbi_can_pos <-}
-\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(}
-    \DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi,}
-    \DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)),}
-    \DataTypeTok{id =} \StringTok{"stemID"}
-\NormalTok{  ) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{left_join}\NormalTok{(sp_info, }\DataTypeTok{by =} \StringTok{"sp"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
-    \DataTypeTok{canopy_position =} \KeywordTok{str_replace}\NormalTok{(canopy_position, }\StringTok{' '}\NormalTok{, }\StringTok{'_'}\NormalTok{),}
-    \DataTypeTok{canopy_position =} \KeywordTok{str_replace}\NormalTok{(canopy_position, }\StringTok{','}\NormalTok{, }\StringTok{''}\NormalTok{),}
-    \DataTypeTok{canopy_position =} \KeywordTok{ifelse}\NormalTok{(}\KeywordTok{is.na}\NormalTok{(canopy_position), }\StringTok{'shrub_layer'}\NormalTok{, canopy_position),}
-    \DataTypeTok{sp =} \KeywordTok{as.factor}\NormalTok{(canopy_position),}
-    \DataTypeTok{basal_area =} \FloatTok{0.0001} \OperatorTok{*}\StringTok{ }\NormalTok{pi }\OperatorTok{*}\StringTok{ }\NormalTok{(dbh1 }\OperatorTok{/}\StringTok{ }\DecValTok{2}\NormalTok{)}\OperatorTok{^}\DecValTok{2}\NormalTok{ )}
-
-\NormalTok{study_region_scbi <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}
-  \DataTypeTok{x =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{),}
-  \DataTypeTok{y =} \KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{300}\NormalTok{)}
-\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{sf_polygon}\NormalTok{()}
-
-\NormalTok{n_fold <-}\StringTok{ }\DecValTok{4}
-\NormalTok{fold1 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{0}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{))}
-\NormalTok{fold2 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{150}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{))}
-\NormalTok{fold3 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{0}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{))}
-\NormalTok{fold4 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{150}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{))}
-
-\NormalTok{blocks_scbi <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}
-  \KeywordTok{sf_polygon}\NormalTok{(fold1), }\KeywordTok{sf_polygon}\NormalTok{(fold2), }\KeywordTok{sf_polygon}\NormalTok{(fold3), }
-  \KeywordTok{sf_polygon}\NormalTok{(fold4)}
-\NormalTok{) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\NormalTok{n_fold) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
-
-\CommentTok{# Associate each observation to a fold}
-\NormalTok{spatial_block_scbi <-}\StringTok{ }
-\StringTok{  }\KeywordTok{spatialBlock}\NormalTok{(}\DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =}\NormalTok{ n_fold, }
-               \DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }
-               \DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}\NormalTok{)}
-
-\NormalTok{growth_scbi_sp <-}\StringTok{ }\NormalTok{growth_scbi_sp }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ spatial_block_scbi}\OperatorTok{$}\NormalTok{foldID }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
-\NormalTok{growth_scbi_can_pos <-}\StringTok{ }\NormalTok{growth_scbi_can_pos }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ spatial_block_scbi}\OperatorTok{$}\NormalTok{foldID }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
-
-\NormalTok{comp_dist <-}\StringTok{ }\FloatTok{7.5}
-
-\NormalTok{growth_scbi_sp <-}\StringTok{ }\NormalTok{growth_scbi_sp }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{size =}\NormalTok{ comp_dist, }\DataTypeTok{region =}\NormalTok{ study_region_scbi)}
-\NormalTok{growth_scbi_can_pos <-}\StringTok{ }\NormalTok{growth_scbi_can_pos }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{size =}\NormalTok{ comp_dist, }\DataTypeTok{region =}\NormalTok{ study_region_scbi)}
-
-\NormalTok{focal_vs_comp_sp <-}\StringTok{ }\NormalTok{growth_scbi_sp }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }
-                         \DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }
-                         \DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{, }
-                         \DataTypeTok{comp_x_var =} \StringTok{"basal_area"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi)}
-
-\NormalTok{focal_vs_comp_can_pos <-}\StringTok{ }\NormalTok{growth_scbi_can_pos }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }
-                         \DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }
-                         \DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{, }
-                         \DataTypeTok{comp_x_var =} \StringTok{"basal_area"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi)}
-
-\NormalTok{  focal_vs_comp_sp }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{pull}\NormalTok{(.estimate)}
-\CommentTok{## [1] 0.14}
-
-\NormalTok{  focal_vs_comp_can_pos }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{pull}\NormalTok{(.estimate)}
-\CommentTok{## [1] 0.144}
-\end{Highlighting}
-\end{Shaded}
-
-We find that species identity has a lower RMSE, so does a better job.
-But we can still plot the competition posteriors for the canopy position
-groupings.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-
-\NormalTok{  fit_mod_can_pos <-}\StringTok{ }\NormalTok{growth_scbi_can_pos }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }
-                         \DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }
-                         \DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{, }
-                         \DataTypeTok{comp_x_var =} \StringTok{"basal_area"}\NormalTok{)  }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{comp_bayes_lm}\NormalTok{(}\DataTypeTok{prior_param =} \OtherTok{NULL}\NormalTok{)}
-
-  \KeywordTok{autoplot}\NormalTok{(fit_mod_can_pos, }\DataTypeTok{type =} \StringTok{"competition"}\NormalTok{)}
-
-\end{Highlighting}
-\end{Shaded}
-
-\begin{center}\includegraphics[width=1\linewidth]{Figures/can-pos-comp-1} \end{center}
-
-Unsurprisingly we see that canopy and canopy emergent competitors
-generally have negative effects on their neighbors, while shrubs and
-understory competitors have neutral or even positive effects.
+\NormalTok{census\_2013\_scbi }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"scbi.stem2.csv"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{select}\NormalTok{(stemID, sp, }\AttributeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{date =} \FunctionTok{mdy}\NormalTok{(date), }\AttributeTok{dbh =} \FunctionTok{as.numeric}\NormalTok{(dbh)}\SpecialCharTok{/}\DecValTok{10}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{filter}\NormalTok{(gx }\SpecialCharTok{\textless{}} \DecValTok{300}\NormalTok{, }\FunctionTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+
+\NormalTok{census\_2018\_scbi }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"scbi.stem3.csv"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{select}\NormalTok{(stemID, sp, }\AttributeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{date =} \FunctionTok{mdy}\NormalTok{(date), }\AttributeTok{dbh =} \FunctionTok{as.numeric}\NormalTok{(dbh)}\SpecialCharTok{/}\DecValTok{10}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{filter}\NormalTok{(gx }\SpecialCharTok{\textless{}} \DecValTok{300}\NormalTok{, }\FunctionTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+
+\NormalTok{growth\_scbi\_sp }\OtherTok{\textless{}{-}} \FunctionTok{compute\_growth}\NormalTok{(}\AttributeTok{census\_1 =}\NormalTok{ census\_2013\_scbi, }\AttributeTok{census\_2 =}\NormalTok{ census\_2018\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{filter}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{str\_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)), }\AttributeTok{id =} \StringTok{"stemID"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{sp =} \FunctionTok{as.factor}\NormalTok{(sp), }\AttributeTok{basal\_area =} \FloatTok{1e{-}04} \SpecialCharTok{*}\NormalTok{ pi }\SpecialCharTok{*}\NormalTok{ (dbh1}\SpecialCharTok{/}\DecValTok{2}\NormalTok{)}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{)}
+
+\NormalTok{growth\_scbi\_can\_pos }\OtherTok{\textless{}{-}} \FunctionTok{compute\_growth}\NormalTok{(}\AttributeTok{census\_1 =}\NormalTok{ census\_2013\_scbi, }\AttributeTok{census\_2 =}\NormalTok{ census\_2018\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{filter}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{str\_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)), }\AttributeTok{id =} \StringTok{"stemID"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{left\_join}\NormalTok{(sp\_info, }\AttributeTok{by =} \StringTok{"sp"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{canopy\_position =} \FunctionTok{str\_replace}\NormalTok{(canopy\_position, }\StringTok{" "}\NormalTok{, }\StringTok{"\_"}\NormalTok{), }\AttributeTok{canopy\_position =} \FunctionTok{str\_replace}\NormalTok{(canopy\_position,}
+        \StringTok{","}\NormalTok{, }\StringTok{""}\NormalTok{), }\AttributeTok{canopy\_position =} \FunctionTok{ifelse}\NormalTok{(}\FunctionTok{is.na}\NormalTok{(canopy\_position), }\StringTok{"shrub\_layer"}\NormalTok{,}
+\NormalTok{        canopy\_position), }\AttributeTok{sp =} \FunctionTok{as.factor}\NormalTok{(canopy\_position), }\AttributeTok{basal\_area =} \FloatTok{1e{-}04} \SpecialCharTok{*}
+\NormalTok{        pi }\SpecialCharTok{*}\NormalTok{ (dbh1}\SpecialCharTok{/}\DecValTok{2}\NormalTok{)}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{)}
+
+\NormalTok{study\_region\_scbi }\OtherTok{\textless{}{-}} \FunctionTok{tibble}\NormalTok{(}\AttributeTok{x =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{), }\AttributeTok{y =} \FunctionTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{,}
+    \DecValTok{600}\NormalTok{, }\DecValTok{300}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{sf\_polygon}\NormalTok{()}
+
+\NormalTok{n\_fold }\OtherTok{\textless{}{-}} \DecValTok{4}
+\NormalTok{fold1 }\OtherTok{\textless{}{-}} \FunctionTok{cbind}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{0}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{))}
+\NormalTok{fold2 }\OtherTok{\textless{}{-}} \FunctionTok{cbind}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{150}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{))}
+\NormalTok{fold3 }\OtherTok{\textless{}{-}} \FunctionTok{cbind}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{0}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+\NormalTok{fold4 }\OtherTok{\textless{}{-}} \FunctionTok{cbind}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{150}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+
+\NormalTok{blocks\_scbi }\OtherTok{\textless{}{-}} \FunctionTok{bind\_rows}\NormalTok{(}\FunctionTok{sf\_polygon}\NormalTok{(fold1), }\FunctionTok{sf\_polygon}\NormalTok{(fold2), }\FunctionTok{sf\_polygon}\NormalTok{(fold3),}
+    \FunctionTok{sf\_polygon}\NormalTok{(fold4)) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{folds =} \FunctionTok{c}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\NormalTok{n\_fold) }\SpecialCharTok{\%\textgreater{}\%}
+        \FunctionTok{factor}\NormalTok{())}
+
+\CommentTok{\# Associate each observation to a fold}
+\NormalTok{spatial\_block\_scbi }\OtherTok{\textless{}{-}} \FunctionTok{spatialBlock}\NormalTok{(}\AttributeTok{speciesData =}\NormalTok{ growth\_scbi, }\AttributeTok{k =}\NormalTok{ n\_fold,}
+    \AttributeTok{selection =} \StringTok{"systematic"}\NormalTok{, }\AttributeTok{blocks =}\NormalTok{ blocks\_scbi, }\AttributeTok{showBlocks =} \ConstantTok{FALSE}\NormalTok{, }\AttributeTok{verbose =} \ConstantTok{FALSE}\NormalTok{)}
+
+\NormalTok{growth\_scbi\_sp }\OtherTok{\textless{}{-}}\NormalTok{ growth\_scbi\_sp }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{foldID =}\NormalTok{ spatial\_block\_scbi}\SpecialCharTok{$}\NormalTok{foldID }\SpecialCharTok{\%\textgreater{}\%}
+        \FunctionTok{factor}\NormalTok{())}
+\NormalTok{growth\_scbi\_can\_pos }\OtherTok{\textless{}{-}}\NormalTok{ growth\_scbi\_can\_pos }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{foldID =}\NormalTok{ spatial\_block\_scbi}\SpecialCharTok{$}\NormalTok{foldID }\SpecialCharTok{\%\textgreater{}\%}
+        \FunctionTok{factor}\NormalTok{())}
+
+\NormalTok{comp\_dist }\OtherTok{\textless{}{-}} \FloatTok{7.5}
+
+\NormalTok{growth\_scbi\_sp }\OtherTok{\textless{}{-}}\NormalTok{ growth\_scbi\_sp }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{add\_buffer\_variable}\NormalTok{(}\AttributeTok{size =}\NormalTok{ comp\_dist, }\AttributeTok{region =}\NormalTok{ study\_region\_scbi)}
+\NormalTok{growth\_scbi\_can\_pos }\OtherTok{\textless{}{-}}\NormalTok{ growth\_scbi\_can\_pos }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{add\_buffer\_variable}\NormalTok{(}\AttributeTok{size =}\NormalTok{ comp\_dist, }\AttributeTok{region =}\NormalTok{ study\_region\_scbi)}
+
+\NormalTok{focal\_vs\_comp\_sp }\OtherTok{\textless{}{-}}\NormalTok{ growth\_scbi\_sp }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{create\_focal\_vs\_comp}\NormalTok{(}\AttributeTok{comp\_dist =}\NormalTok{ comp\_dist, }\AttributeTok{blocks =}\NormalTok{ blocks\_scbi, }\AttributeTok{id =} \StringTok{"stemID"}\NormalTok{,}
+        \AttributeTok{comp\_x\_var =} \StringTok{"basal\_area"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{run\_cv}\NormalTok{(}\AttributeTok{comp\_dist =}\NormalTok{ comp\_dist, }\AttributeTok{blocks =}\NormalTok{ blocks\_scbi)}
+
+\NormalTok{focal\_vs\_comp\_can\_pos }\OtherTok{\textless{}{-}}\NormalTok{ growth\_scbi\_can\_pos }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{create\_focal\_vs\_comp}\NormalTok{(}\AttributeTok{comp\_dist =}\NormalTok{ comp\_dist, }\AttributeTok{blocks =}\NormalTok{ blocks\_scbi, }\AttributeTok{id =} \StringTok{"stemID"}\NormalTok{,}
+        \AttributeTok{comp\_x\_var =} \StringTok{"basal\_area"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{run\_cv}\NormalTok{(}\AttributeTok{comp\_dist =}\NormalTok{ comp\_dist, }\AttributeTok{blocks =}\NormalTok{ blocks\_scbi)}
+
+\NormalTok{focal\_vs\_comp\_sp }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{rmse}\NormalTok{(}\AttributeTok{truth =}\NormalTok{ growth, }\AttributeTok{estimate =}\NormalTok{ growth\_hat) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{pull}\NormalTok{(.estimate)}
+\DocumentationTok{\#\# [1] 0.14}
+
+\NormalTok{focal\_vs\_comp\_can\_pos }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{rmse}\NormalTok{(}\AttributeTok{truth =}\NormalTok{ growth, }\AttributeTok{estimate =}\NormalTok{ growth\_hat) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{pull}\NormalTok{(.estimate)}
+\DocumentationTok{\#\# [1] 0.144}
+\end{Highlighting}
+\end{Shaded}
+
+We find that species identity has a lower RMSE, so does a better job. We
+still however plot the competition posteriors for the canopy position
+groupings in Figure \ref{fig:can-pos-comp-plot}. Unsurprisingly we see
+that canopy and canopy emergent competitors generally have negative
+effects on their neighbors, while shrubs and understory competitors have
+neutral or even positive effects.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{fit\_mod\_can\_pos }\OtherTok{\textless{}{-}}\NormalTok{ growth\_scbi\_can\_pos }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{create\_focal\_vs\_comp}\NormalTok{(}
+    \AttributeTok{comp\_dist =}\NormalTok{ comp\_dist,}
+    \AttributeTok{blocks =}\NormalTok{ blocks\_scbi,}
+    \AttributeTok{id =} \StringTok{"stemID"}\NormalTok{,}
+    \AttributeTok{comp\_x\_var =} \StringTok{"basal\_area"}
+\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{comp\_bayes\_lm}\NormalTok{(}\AttributeTok{prior\_param =} \ConstantTok{NULL}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{Figures/can-pos-comp-plot-1}
+
+}
+
+\caption{Posterior distributions of all competition parameters.}\label{fig:can-pos-comp-plot}
+\end{figure}
+
+\hypertarget{appendix-replicate-rmse-comparison}{%
+\section{Appendix: Replicate RMSE
+comparison}\label{appendix-replicate-rmse-comparison}}
+
+This code replicates Figure \ref{fig:scbi-simulation}: A comparison of
+root mean squared error of models for standard, permuted, and spatial
+cross-validated error estimates.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{library}\NormalTok{(tidyverse)}
+\FunctionTok{library}\NormalTok{(lubridate)}
+\FunctionTok{library}\NormalTok{(here)}
+\FunctionTok{library}\NormalTok{(sf)}
+\FunctionTok{library}\NormalTok{(viridis)}
+\FunctionTok{library}\NormalTok{(forestecology)}
+\FunctionTok{library}\NormalTok{(blockCV)}
+\FunctionTok{library}\NormalTok{(tictoc)}
+
+\CommentTok{\# Compute growth of trees based on census data}
+\CommentTok{\# {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
+\NormalTok{census\_2013\_scbi }\OtherTok{\textless{}{-}} \FunctionTok{here}\NormalTok{(}\StringTok{"paper/scbi.stem2.csv"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{read\_csv}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{select}\NormalTok{(stemID, sp, }\AttributeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{date =} \FunctionTok{mdy}\NormalTok{(date), }\AttributeTok{dbh =} \FunctionTok{as.numeric}\NormalTok{(dbh)}\SpecialCharTok{/}\DecValTok{10}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{filter}\NormalTok{(gx }\SpecialCharTok{\textless{}} \DecValTok{300}\NormalTok{, }\FunctionTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+
+\NormalTok{census\_2018\_scbi }\OtherTok{\textless{}{-}} \FunctionTok{here}\NormalTok{(}\StringTok{"paper/scbi.stem3.csv"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{read\_csv}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{select}\NormalTok{(stemID, sp, }\AttributeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{date =} \FunctionTok{mdy}\NormalTok{(date), }\AttributeTok{dbh =} \FunctionTok{as.numeric}\NormalTok{(dbh)}\SpecialCharTok{/}\DecValTok{10}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{filter}\NormalTok{(gx }\SpecialCharTok{\textless{}} \DecValTok{300}\NormalTok{, }\FunctionTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+
+\NormalTok{growth\_scbi }\OtherTok{\textless{}{-}} \FunctionTok{compute\_growth}\NormalTok{(}\AttributeTok{census\_1 =}\NormalTok{ census\_2013\_scbi, }\AttributeTok{census\_2 =}\NormalTok{ census\_2018\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{filter}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{str\_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)), }\AttributeTok{id =} \StringTok{"stemID"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+    \CommentTok{\# Compute basal area:}
+\FunctionTok{mutate}\NormalTok{(}\AttributeTok{basal\_area =} \FloatTok{1e{-}04} \SpecialCharTok{*}\NormalTok{ pi }\SpecialCharTok{*}\NormalTok{ (dbh1}\SpecialCharTok{/}\DecValTok{2}\NormalTok{)}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{)}
+
+\CommentTok{\# Add spatial information}
+\CommentTok{\# {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-} Define buffer region}
+\CommentTok{\# using competitive distance range}
+\NormalTok{comp\_dist }\OtherTok{\textless{}{-}} \FloatTok{7.5}
+
+\NormalTok{study\_region\_scbi }\OtherTok{\textless{}{-}} \FunctionTok{tibble}\NormalTok{(}\AttributeTok{x =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{), }\AttributeTok{y =} \FunctionTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{,}
+    \DecValTok{600}\NormalTok{, }\DecValTok{300}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{sf\_polygon}\NormalTok{()}
+
+\NormalTok{growth\_scbi }\OtherTok{\textless{}{-}}\NormalTok{ growth\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{add\_buffer\_variable}\NormalTok{(}\AttributeTok{size =}\NormalTok{ comp\_dist, }\AttributeTok{region =}\NormalTok{ study\_region\_scbi)}
+
+\CommentTok{\# Manually define spatial blocks to act as folds}
+\NormalTok{fold1 }\OtherTok{\textless{}{-}} \FunctionTok{rbind}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{450}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{450}\NormalTok{))}
+\NormalTok{fold2 }\OtherTok{\textless{}{-}} \FunctionTok{rbind}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{450}\NormalTok{))}
+\NormalTok{fold3 }\OtherTok{\textless{}{-}} \FunctionTok{rbind}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{450}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{450}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{600}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+\NormalTok{fold4 }\OtherTok{\textless{}{-}} \FunctionTok{rbind}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{450}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+\NormalTok{n\_fold }\OtherTok{\textless{}{-}} \DecValTok{4}
+
+\NormalTok{blocks\_scbi }\OtherTok{\textless{}{-}} \FunctionTok{bind\_rows}\NormalTok{(}\FunctionTok{sf\_polygon}\NormalTok{(fold1), }\FunctionTok{sf\_polygon}\NormalTok{(fold2), }\FunctionTok{sf\_polygon}\NormalTok{(fold3),}
+    \FunctionTok{sf\_polygon}\NormalTok{(fold4)) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{folds =} \FunctionTok{c}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\NormalTok{n\_fold) }\SpecialCharTok{\%\textgreater{}\%}
+        \FunctionTok{factor}\NormalTok{())}
+
+\CommentTok{\# Associate each observation to a fold}
+\NormalTok{SpatialBlock\_scbi }\OtherTok{\textless{}{-}} \FunctionTok{spatialBlock}\NormalTok{(}\AttributeTok{speciesData =}\NormalTok{ growth\_scbi, }\AttributeTok{k =}\NormalTok{ n\_fold,}
+    \AttributeTok{selection =} \StringTok{"systematic"}\NormalTok{, }\AttributeTok{blocks =}\NormalTok{ blocks\_scbi, }\AttributeTok{showBlocks =} \ConstantTok{FALSE}\NormalTok{, }\AttributeTok{verbose =} \ConstantTok{FALSE}\NormalTok{)}
+
+\NormalTok{growth\_scbi }\OtherTok{\textless{}{-}}\NormalTok{ growth\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{foldID =}\NormalTok{ SpatialBlock\_scbi}\SpecialCharTok{$}\NormalTok{foldID }\SpecialCharTok{\%\textgreater{}\%}
+        \FunctionTok{factor}\NormalTok{())}
+
+\CommentTok{\# Compute focal versus competitor tree information}
+\CommentTok{\# {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
+\NormalTok{focal\_vs\_comp\_scbi }\OtherTok{\textless{}{-}}\NormalTok{ growth\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{create\_focal\_vs\_comp}\NormalTok{(comp\_dist, }\AttributeTok{blocks =}\NormalTok{ blocks\_scbi, }\AttributeTok{id =} \StringTok{"stemID"}\NormalTok{,}
+        \AttributeTok{comp\_x\_var =} \StringTok{"basal\_area"}\NormalTok{)}
+
+\CommentTok{\# Fit model and make predictions}
+\CommentTok{\# {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-} Number of permutation}
+\CommentTok{\# shuffles:}
+\NormalTok{num\_shuffle }\OtherTok{\textless{}{-}} \DecValTok{49}
+
+\CommentTok{\# Save results here}
+\NormalTok{run\_time }\OtherTok{\textless{}{-}} \DecValTok{0}
+\NormalTok{observed\_RMSE }\OtherTok{\textless{}{-}} \DecValTok{0}
+\NormalTok{observed\_RMSE\_CV }\OtherTok{\textless{}{-}} \DecValTok{0}
+\NormalTok{shuffle\_RMSE }\OtherTok{\textless{}{-}} \FunctionTok{vector}\NormalTok{(}\StringTok{"list"}\NormalTok{, }\DecValTok{1}\NormalTok{)}
+\NormalTok{shuffle\_RMSE\_CV }\OtherTok{\textless{}{-}} \FunctionTok{vector}\NormalTok{(}\StringTok{"list"}\NormalTok{, }\DecValTok{1}\NormalTok{)}
+\NormalTok{filename }\OtherTok{\textless{}{-}} \FunctionTok{here}\NormalTok{(}\StringTok{"paper/simulation\_results/"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{str\_c}\NormalTok{(}\StringTok{"2021{-}03{-}03\_scbi\_"}\NormalTok{, num\_shuffle, }\StringTok{"\_shuffles"}\NormalTok{)}
+
+\CommentTok{\# Run all simulations 0. Setup simulation for this species type {-}{-}{-}{-}}
+\CommentTok{\# Start clock}
+\FunctionTok{tic}\NormalTok{()}
+
+\CommentTok{\# 1. Compute observed test statistic: RMSE with no cross{-}validation}
+\CommentTok{\# {-}{-}{-}{-} Fit model (compute posterior parameters)}
+\NormalTok{comp\_bayes\_lm\_scbi }\OtherTok{\textless{}{-}}\NormalTok{ focal\_vs\_comp\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{comp\_bayes\_lm}\NormalTok{(}\AttributeTok{prior\_param =} \ConstantTok{NULL}\NormalTok{, }\AttributeTok{run\_shuffle =} \ConstantTok{FALSE}\NormalTok{)}
+
+\CommentTok{\# Make predictions and compute RMSE}
+\NormalTok{observed\_RMSE }\OtherTok{\textless{}{-}}\NormalTok{ focal\_vs\_comp\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{growth\_hat =} \FunctionTok{predict}\NormalTok{(comp\_bayes\_lm\_scbi, focal\_vs\_comp\_scbi)) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{rmse}\NormalTok{(}\AttributeTok{truth =}\NormalTok{ growth, }\AttributeTok{estimate =}\NormalTok{ growth\_hat) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{pull}\NormalTok{(.estimate)}
+
+\CommentTok{\# 2. Compute observed test statistic: RMSE with cross{-}validation}
+\CommentTok{\# {-}{-}{-}{-}{-}{-}{-}}
+\NormalTok{observed\_RMSE\_CV }\OtherTok{\textless{}{-}}\NormalTok{ focal\_vs\_comp\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{run\_cv}\NormalTok{(}\AttributeTok{comp\_dist =}\NormalTok{ comp\_dist, }\AttributeTok{blocks =}\NormalTok{ blocks\_scbi) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{rmse}\NormalTok{(}\AttributeTok{truth =}\NormalTok{ growth, }\AttributeTok{estimate =}\NormalTok{ growth\_hat) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{pull}\NormalTok{(.estimate)}
+
+\CommentTok{\# 3. Permutation distribution: RMSE with no cross{-}validation}
+\CommentTok{\# {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-} Compute num\_shuffle permutation test statistics}
+\NormalTok{shuffle\_RMSE }\OtherTok{\textless{}{-}} \FunctionTok{numeric}\NormalTok{(}\AttributeTok{length =}\NormalTok{ num\_shuffle)}
+\ControlFlowTok{for}\NormalTok{ (j }\ControlFlowTok{in} \DecValTok{1}\SpecialCharTok{:}\NormalTok{num\_shuffle) \{}
+    \CommentTok{\# Fit model (compute posterior parameters) with shuffling}
+\NormalTok{    comp\_bayes\_lm\_scbi }\OtherTok{\textless{}{-}}\NormalTok{ focal\_vs\_comp\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+        \FunctionTok{comp\_bayes\_lm}\NormalTok{(}\AttributeTok{prior\_param =} \ConstantTok{NULL}\NormalTok{, }\AttributeTok{run\_shuffle =} \ConstantTok{TRUE}\NormalTok{)}
+
+    \CommentTok{\# Make predictions and compute RMSE}
+\NormalTok{    shuffle\_RMSE[j] }\OtherTok{\textless{}{-}}\NormalTok{ focal\_vs\_comp\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+        \FunctionTok{mutate}\NormalTok{(}\AttributeTok{growth\_hat =} \FunctionTok{predict}\NormalTok{(comp\_bayes\_lm\_scbi, focal\_vs\_comp\_scbi)) }\SpecialCharTok{\%\textgreater{}\%}
+        \FunctionTok{rmse}\NormalTok{(}\AttributeTok{truth =}\NormalTok{ growth, }\AttributeTok{estimate =}\NormalTok{ growth\_hat) }\SpecialCharTok{\%\textgreater{}\%}
+        \FunctionTok{pull}\NormalTok{(.estimate)}
+\NormalTok{\}}
+
+\CommentTok{\# 4. Permutation distribution: RMSE with cross{-}validation}
+\CommentTok{\# {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-} Compute num\_shuffle permutation test statistics}
+\NormalTok{shuffle\_RMSE\_CV }\OtherTok{\textless{}{-}} \FunctionTok{numeric}\NormalTok{(}\AttributeTok{length =}\NormalTok{ num\_shuffle)}
+
+\CommentTok{\# Compute num\_shuffle permutation test statistics}
+\ControlFlowTok{for}\NormalTok{ (j }\ControlFlowTok{in} \DecValTok{1}\SpecialCharTok{:}\NormalTok{num\_shuffle) \{}
+    \CommentTok{\# Compute and save RMSE}
+\NormalTok{    shuffle\_RMSE\_CV[j] }\OtherTok{\textless{}{-}}\NormalTok{ focal\_vs\_comp\_scbi }\SpecialCharTok{\%\textgreater{}\%}
+        \FunctionTok{run\_cv}\NormalTok{(}\AttributeTok{comp\_dist =}\NormalTok{ comp\_dist, }\AttributeTok{blocks =}\NormalTok{ blocks\_scbi, }\AttributeTok{run\_shuffle =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+        \FunctionTok{rmse}\NormalTok{(}\AttributeTok{truth =}\NormalTok{ growth, }\AttributeTok{estimate =}\NormalTok{ growth\_hat) }\SpecialCharTok{\%\textgreater{}\%}
+        \FunctionTok{pull}\NormalTok{(.estimate)}
+
+    \CommentTok{\# Status update}
+    \FunctionTok{str\_c}\NormalTok{(}\StringTok{"Shuffle with permutation "}\NormalTok{, j, }\StringTok{" at "}\NormalTok{, }\FunctionTok{Sys.time}\NormalTok{()) }\SpecialCharTok{\%\textgreater{}\%}
+        \FunctionTok{print}\NormalTok{()}
+\NormalTok{\}}
+
+\CommentTok{\# 5. Save results {-}{-}{-}{-}}
+\NormalTok{clock }\OtherTok{\textless{}{-}} \FunctionTok{toc}\NormalTok{(}\AttributeTok{quiet =} \ConstantTok{TRUE}\NormalTok{)}
+\NormalTok{run\_time }\OtherTok{\textless{}{-}}\NormalTok{ clock}\SpecialCharTok{$}\NormalTok{toc }\SpecialCharTok{{-}}\NormalTok{ clock}\SpecialCharTok{$}\NormalTok{tic}
+
+\NormalTok{model\_comp\_tbl }\OtherTok{\textless{}{-}} \FunctionTok{tibble}\NormalTok{(}\AttributeTok{run\_time =}\NormalTok{ run\_time, }\AttributeTok{observed\_RMSE =}\NormalTok{ observed\_RMSE,}
+    \AttributeTok{observed\_RMSE\_CV =}\NormalTok{ observed\_RMSE\_CV, }\AttributeTok{shuffle\_RMSE =}\NormalTok{ shuffle\_RMSE, }\AttributeTok{shuffle\_RMSE\_CV =}\NormalTok{ shuffle\_RMSE\_CV,}
+\NormalTok{    )}
+\FunctionTok{save}\NormalTok{(model\_comp\_tbl, }\AttributeTok{file =}\NormalTok{ filename }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{str\_c}\NormalTok{(}\StringTok{".RData"}\NormalTok{))}
+
+\CommentTok{\# Visualize results}
+\CommentTok{\# {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
+\NormalTok{model\_comp }\OtherTok{\textless{}{-}} \FunctionTok{bind\_rows}\NormalTok{(model\_comp\_tbl }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{select}\NormalTok{(run\_time, }\AttributeTok{observed =}\NormalTok{ observed\_RMSE, }\AttributeTok{shuffle =}\NormalTok{ shuffle\_RMSE) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{CV =} \ConstantTok{FALSE}\NormalTok{), model\_comp\_tbl }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{select}\NormalTok{(run\_time, }\AttributeTok{observed =}\NormalTok{ observed\_RMSE\_CV, }\AttributeTok{shuffle =}\NormalTok{ shuffle\_RMSE\_CV) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{CV =} \ConstantTok{TRUE}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{gather}\NormalTok{(type, RMSE, }\SpecialCharTok{{-}}\FunctionTok{c}\NormalTok{(run\_time, CV))}
+
+\NormalTok{model\_comp\_observed }\OtherTok{\textless{}{-}}\NormalTok{ model\_comp }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{filter}\NormalTok{(type }\SpecialCharTok{==} \StringTok{"observed"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{unnest}\NormalTok{(}\AttributeTok{cols =} \FunctionTok{c}\NormalTok{(RMSE))}
+\NormalTok{model\_comp\_shuffle }\OtherTok{\textless{}{-}}\NormalTok{ model\_comp }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{filter}\NormalTok{(type }\SpecialCharTok{==} \StringTok{"shuffle"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{unnest}\NormalTok{(}\AttributeTok{cols =} \FunctionTok{c}\NormalTok{(RMSE))}
+
+\NormalTok{cv\_plot }\OtherTok{\textless{}{-}} \FunctionTok{ggplot}\NormalTok{() }\SpecialCharTok{+} \FunctionTok{geom\_vline}\NormalTok{(}\AttributeTok{data =}\NormalTok{ model\_comp\_observed, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{xintercept =}\NormalTok{ RMSE,}
+    \AttributeTok{col =}\NormalTok{ CV), }\AttributeTok{linetype =} \StringTok{"dashed"}\NormalTok{, }\AttributeTok{show.legend =}\NormalTok{ F) }\SpecialCharTok{+} \FunctionTok{geom\_histogram}\NormalTok{(}\AttributeTok{data =}\NormalTok{ model\_comp\_shuffle,}
+    \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ RMSE, }\AttributeTok{fill =}\NormalTok{ CV), }\AttributeTok{bins =} \DecValTok{50}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{labs}\NormalTok{(}\AttributeTok{fill =} \StringTok{"Cross{-}validated?"}\NormalTok{,}
+    \AttributeTok{x =} \FunctionTok{expression}\NormalTok{(}\FunctionTok{paste}\NormalTok{(}\StringTok{"RMSE (cm "}\NormalTok{, y}\SpecialCharTok{\^{}}\NormalTok{\{}
+        \SpecialCharTok{{-}}\DecValTok{1}
+\NormalTok{    \}, }\StringTok{")"}\NormalTok{))) }\SpecialCharTok{+} \FunctionTok{scale\_color\_viridis}\NormalTok{(}\AttributeTok{discrete =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{option =} \StringTok{"D"}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{scale\_fill\_viridis}\NormalTok{(}\AttributeTok{discrete =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{+}
+    \FunctionTok{theme\_light}\NormalTok{()}
+\NormalTok{cv\_plot}
+
+\NormalTok{filename }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{str\_c}\NormalTok{(}\StringTok{".pdf"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{ggsave}\NormalTok{(}\AttributeTok{plot =}\NormalTok{ cv\_plot, }\AttributeTok{width =} \DecValTok{16}\SpecialCharTok{/}\DecValTok{2}\NormalTok{, }\AttributeTok{height =} \DecValTok{9}\SpecialCharTok{/}\DecValTok{2}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 

### 2021-08-11 paper/paper.Rmd
Update manuscript

@@
 bibliography: paper.bib
-output: rticles::asa_article
+output:
+  rticles::asa_article
 link-citations: yes
@@
 ```{r scbi-spatial-cross-validation-schematic, out.width="66%", fig.cap= "Schematic of spatial cross-validation.
-Using the k = 1 fold (bottom-left) as the test set, k = 2 through 4 as the training set, along with a \"fold buffer\" extending outwards from the test set to maintain spatial independence between it and the training set.", echo = FALSE}
+Using the k = 1 fold (bottom-left) as the test set, k = 2 through 4 as the training set, along with a fold buffer extending outwards from the test set to maintain spatial independence between it and the training set.", echo = FALSE}
 fold_number <- 1
@@
 
-```{r appendix2-plot, out.width="100%", fig.cap="Cross-validated RMSE estimates for 5 competitive distances2.", echo = FALSE}
+```{r appendix2-plot, out.width="100%", fig.cap="Cross-validated RMSE estimates for 5 competitive distances.", echo = FALSE}
 mult_dist_comp %>%

### 2021-08-31 paper/paper_word_output.Rmd
Started word output Rmd

--- 2021-08-11 paper/paper.Rmd
@@
 title: The forestecology R package for fitting and assessing neighborhood models of the effect of interspecific competition on the growth of trees
-
-blinded: 0
-
 authors: 
@@
   affiliation: Mathematics Department, Reed College
-
-abstract: 
-- |
-  | 1. Neighborhood competition models are powerful tools to measure the effect of interspecific competition. Statistical methods to ease the application of these models are currently lacking.
-  | 2.
-We present the `forestecology` package providing methods to i) specify neighborhood competition models, ii) evaluate the effect of competitor species identity using permutation tests, and iii) measure model performance using spatial cross-validation.
-Following @allen_permutation_2020, we implement a Bayesian linear regression neighborhood competition model.
-  | 3.
-We demonstrate the package's functionality using data from the Smithsonian Conservation Biology Institute's large forest dynamics plot, part of the ForestGEO global network of research sites.
-Given ForestGEO's data collection protocols and data formatting standards, the package was designed with cross-site compatibility in mind.
-We highlight the importance of spatial cross-validation when interpreting model results.
-  | 4.
-The package features i) `tidyverse`-like structure whereby verb-named functions can be modularly "piped" in sequence, ii) functions with standardized inputs/outputs of simple features `sf` package class, and iii) an S3 object-oriented implementation of the Bayesian linear regression model.
-These three facts allow for clear articulation of all the steps in the sequence of analysis and easy wrangling and visualization of the geospatial data.
-Furthermore, while the package only has Bayesian linear regression implemented, the package was designed with extensibility to other methods in mind.
-
-keywords:
-- forest ecology, interspecific competition, neighborhood competition, tree growth, R, ForestGEO, spatial cross-validation
-
 bibliography: paper.bib
 output:
-  rticles::asa_article
+  bookdown::word_document2:
+  fig_caption: yes
 link-citations: yes
@@
 # Add line numbers:
-- \linenumbers
+# - \linenumbers
 # Add running header:
@@
 
+# Abstract
+
+1. Neighborhood competition models are powerful tools to measure the effect of interspecific competition. Statistical methods to ease the application of these models are currently lacking.
+2.
+We present the `forestecology` package providing methods to i) specify neighborhood competition models, ii) evaluate the effect of competitor species identity using permutation tests, and iii) measure model performance using spatial cross-validation.
+Following @allen_permutation_2020, we implement a Bayesian linear regression neighborhood competition model.
+3.
+We demonstrate the package's functionality using data from the Smithsonian Conservation Biology Institute's large forest dynamics plot, part of the ForestGEO global network of research sites.
+Given ForestGEO's data collection protocols and data formatting standards, the package was designed with cross-site compatibility in mind.
+We highlight the importance of spatial cross-validation when interpreting model results.
+4.
+The package features i) `tidyverse`-like structure whereby verb-named functions can be modularly "piped" in sequence, ii) functions with standardized inputs/outputs of simple features `sf` package class, and iii) an S3 object-oriented implementation of the Bayesian linear regression model.
+These three facts allow for clear articulation of all the steps in the sequence of analysis and easy wrangling and visualization of the geospatial data.
+Furthermore, while the package only has Bayesian linear regression implemented, the package was designed with extensibility to other methods in mind.
+
+Keywords: forest ecology, interspecific competition, neighborhood competition, tree growth, R, ForestGEO, spatial cross-validation
+
 # Introduction
@@
 
-# Acknowledgments
-
-We thank Sophie Li for their feedback on the package interface. The authors declare no conflicts of interest.
-
-# Author's contributions
-
-AYK and DNA conceived the ideas and coded a draft of the package.
-AYK wrote an initial manuscript draft.
-SPC rewrote much of the package's code to align with R and "tidy" best practices [@wickham_welcome_2019].
-All authors contributed to subsequent drafts and gave final approval for manuscript.
-
-# Data accessibility
-
-We intend to archive all data and source code for this manuscript on GitHub at https://github.com/rudeboybert/forestecology and on Zenodo upon acceptance.
+# Acknowledgements {-}
+
+We thank Sophie Li for their feedback on the package interface.
+
+# Conflict of Interest {-}
+
+We declare no competing interests.
+
+# Author's contributions {-}
+
+**Albert Y.
+Kim**: Conceptualization (equal); Methodology (equal); Software (equal); Writing-original draft (lead); Writing-review & editing (equal).
+**David N.
+Allen**: Conceptualization (equal); Methodology (equal); Software (equal); Writing-original draft (supporting); Writing-review & editing (equal).
+**Simon P.
+Couch**: Software (supporting); Writing-review & editing (supporting).
+
+# Data accessibility {-}
+
+All data and source code for this manuscript on GitHub at https://github.com/rudeboybert/forestecology and are archived on Zenodo at https://doi.org/10.5281/zenodo.5348060.
 The example Smithsonian Conservation Biology Institute census and species information data are available on GitHub at https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/tree/master/ and are archived on Zenodo at https://doi.org/10.5281/zenodo.2649301 [@gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020].

### 2021-08-31 paper/paper_word_output.Rmd
Replaced all , with ; for multiple citation bibtex entries

@@
 Repeat-censused forest plots offer excellent opportunities to test neighborhood models of the effect of competition on the growth of trees [@canham_neighborhood_2004].
-Neighborhood models of competition have been used to: test whether the species identity of a competitor matters [@uriarte_spatially_2004; measure species-specific competition coefficients [@das_effect_2012 @tatsumi_estimating_2016]; test competing models to see what structures competitive interactions, e.g.
+Neighborhood models of competition have been used to: test whether the species identity of a competitor matters [@uriarte_spatially_2004; measure species-specific competition coefficients [@das_effect_2012; @tatsumi_estimating_2016]; test competing models to see what structures competitive interactions, e.g.
 traits or phylogeny [@allen_permutation_2020; @uriarte_trait_2010]; and inform selective logging practices [@canham_neighborhood_2006].
@@
 1. Using simple feature access as the base standard for representing and encoding spatial data, rather than shapefiles [@herring_2011]. 
-1. Leveraging improvements in external libraries for reading and writing spatial data (GDAL) and for geometrical operations (GEOS) [@warmerdam_geospatial_2008, @geos_2017].
+1. Leveraging improvements in external libraries for reading and writing spatial data (GDAL) and for geometrical operations (GEOS) [@warmerdam_geospatial_2008; @geos_2017].
 1. Integrating closely with the popular `tidyverse` suite of packages for data science [@wickham_welcome_2019].
@@
 
-We present a case-study of `forestecology`'s functionality on data from the Smithsonian Conservation Biology Institute (SCBI) large forest dynamics plot in Front Royal, VA, USA, part of the ForestGEO global network of research sites [@bourg_initial_2013, @andersonteixeira_ctfs-forestgeo_2015, @davies_forestgeo_2021].
+We present a case-study of `forestecology`'s functionality on data from the Smithsonian Conservation Biology Institute (SCBI) large forest dynamics plot in Front Royal, VA, USA, part of the ForestGEO global network of research sites [@bourg_initial_2013; @andersonteixeira_ctfs-forestgeo_2015; @davies_forestgeo_2021].
 The 25.6 ha (640 x 400 m) plot is located at the intersection of three of the major physiographic provinces of the eastern US---the Blue Ridge, Ridge and Valley, and Piedmont provinces---and is adjacent to the northern end of Shenandoah National Park.
@@
 Our model of interspecific competition relies on a spatial definition of who competitor trees are: all trees within a distance `comp_dist` of a focal tree.
-Here we set `comp_dist` to 7.5m, a value informed by other studies [@canham_neighborhood_2004, @uriarte_spatially_2004, @canham_neighborhood_2006], but the package could also be used to compare multiple distances and see which is best supported (see Appendix \ref{mult-dist}).
+Here we set `comp_dist` to 7.5m, a value informed by other studies [@canham_neighborhood_2004; @uriarte_spatially_2004; @canham_neighborhood_2006], but the package could also be used to compare multiple distances and see which is best supported (see Appendix \ref{mult-dist}).
 We use `comp_dist` and a manually constructed `sf` representation of the study region's boundary as inputs to `add_buffer_variable()` to add a `buffer` boolean variable to `growth_scbi`.
@@
 In the case of forest census data however, observations exhibit spatial autocorrelation.
-We therefore incorporate this dependence into the cross-validation algorithm by resampling spatial blocks of trees [@roberts_cross-validation_2017 @pohjankukka_estimating_2017].
+We therefore incorporate this dependence into the cross-validation algorithm by resampling spatial blocks of trees [@roberts_cross-validation_2017; @pohjankukka_estimating_2017].
 
@@
 However, once we incorporate spatial cross-validation, this improvement disappears.
-These results suggest that in this 9 ha subplot of the SCBI plot, competitive interactions do not depend on the identity of the competitor, which is the opposite of what has been observed in other locations [@allen_permutation_2020 @uriarte_spatially_2004].
+These results suggest that in this 9 ha subplot of the SCBI plot, competitive interactions do not depend on the identity of the competitor, which is the opposite of what has been observed in other locations [@allen_permutation_2020; @uriarte_spatially_2004].
 This provides a striking example of the importance of cross-validation, as without it the over-fit model gives rise to an incorrect conclusion.
@@
 One clear future direction would be to allow competition based on species trait values rather than species identity.
-There is evidence that traits predict competitive outcomes [@kunstler_competitive_2012, @lasky_trait-mediated_2014, @uriarte_trait_2010].
+There is evidence that traits predict competitive outcomes [@kunstler_competitive_2012; @lasky_trait-mediated_2014; @uriarte_trait_2010].
 Thus an extension of the model would allow $\lambda$ values from Equation \ref{eq:model} to be a function of the traits of competing species.
@@
 Lastly, the `forestecology` current uses the `blockCV` package behind the scenes to create the spatial blocks acting as folds for our spatial cross-validation algorithm detailed in Sections \ref{spatial-information} and \ref{spatial-cross-validation} [@valavi_blockcv_2019].
-This back-end functionality could be substituted with the `spatialsample` package for spatial resampling infrastructure; a `tidymodels` package under active development as of 2021 [@spatialsample_package, @tidymodels_package].
+This back-end functionality could be substituted with the `spatialsample` package for spatial resampling infrastructure; a `tidymodels` package under active development as of 2021 [@spatialsample_package; @tidymodels_package].
 
@@
 For all the above analyses we set the cut off distance (`comp_dist`) for two stems to compete as `r comp_dist`m.
-This distance has been estimated in previous neighborhood competition studies in forests [@canham_neighborhood_2004, @uriarte_spatially_2004, @canham_neighborhood_2006].
+This distance has been estimated in previous neighborhood competition studies in forests [@canham_neighborhood_2004; @uriarte_spatially_2004; @canham_neighborhood_2006].
 We used `r comp_dist`m in @allen_permutation_2020 as an average of the values estimated in other studies.

### 2021-08-31 paper/paper_word_output.Rmd
replace all \ref{} with \@ref()

@@
 The output `growth_scbi` is a data frame of class `sf` that includes among other variables the species variable `sp` converted to a factor, the average annual `growth` in DBH (cm $\cdot$ y^-1^) for all stems that were alive at both time points, and the `sf` package's encoding of geolocations of `geometry` type `<POINT>`.
-Given that `growth_scbi` is of class `sf`, it can be easily plotted in `ggplot2` using `geom_sf()` as seen in Figure \ref{fig:scbi-trees}.
+Given that `growth_scbi` is of class `sf`, it can be easily plotted in `ggplot2` using `geom_sf()` as seen in Figure \@ref(fig:scbi-trees).
 
@@
 
-Furthermore, we compute two potential competitor explanatory variables $x^{\text{comp}}_{ijk}$ from Equation \ref{eq:model}.
+Furthermore, we compute two potential competitor explanatory variables $x^{\text{comp}}_{ijk}$ from Equation \@ref(eq:model).
 First, the basal area of each tree as a function of its DBH in the earlier census.
@@
 Our model of interspecific competition relies on a spatial definition of who competitor trees are: all trees within a distance `comp_dist` of a focal tree.
-Here we set `comp_dist` to 7.5m, a value informed by other studies [@canham_neighborhood_2004; @uriarte_spatially_2004; @canham_neighborhood_2006], but the package could also be used to compare multiple distances and see which is best supported (see Appendix \ref{mult-dist}).
+Here we set `comp_dist` to 7.5m, a value informed by other studies [@canham_neighborhood_2004; @uriarte_spatially_2004; @canham_neighborhood_2006], but the package could also be used to compare multiple distances and see which is best supported (see Appendix \@ref(mult-dist)).
 We use `comp_dist` and a manually constructed `sf` representation of the study region's boundary as inputs to `add_buffer_variable()` to add a `buffer` boolean variable to `growth_scbi`.
@@
 We then use the output of the `spatialBlock()` function from the `blockCV` package to associate each tree in `growth_scbi` to the correct `foldID` [@valavi_blockcv_2019].
-This `foldID` variable will be used in Section \ref{spatial-cross-validation}.
-
-Figure \ref{fig:scbi-spatial-information} illustrates the net effect of adding these two elements of spatial information to `growth_scbi`.
+This `foldID` variable will be used in Section \@ref(spatial-cross-validation).
+
+Figure \@ref(fig:scbi-spatial-information) illustrates the net effect of adding these two elements of spatial information to `growth_scbi`.
 
@@
 More specifically, identify all trees that are not part of the buffer region, have a valid `growth` measurement, and have at least one neighbor within `r comp_dist`m.
-We do this using `create_focal_vs_comp()`, which takes the previously detailed `comp_dist` and `id` arguments, the `sf` representation of the spatial cross-validation blocks `blocks_scbi`, and a specification `comp_x_var` of the `basal_area` variable we use as the competitor explanatory variable $x^{\text{comp}}_{ijk}$ from Equation \ref{eq:model}.
+We do this using `create_focal_vs_comp()`, which takes the previously detailed `comp_dist` and `id` arguments, the `sf` representation of the spatial cross-validation blocks `blocks_scbi`, and a specification `comp_x_var` of the `basal_area` variable we use as the competitor explanatory variable $x^{\text{comp}}_{ijk}$ from Equation \@ref(eq:model).
 This function returns a new data frame `focal_vs_comp_scbi`.
@@
 Saving competitor information in list-column format minimizes redundancy since we do not need to repeat information on the focal tree 20 times.
-We visualize the spatial distribution of these trees in Figure \ref{fig:scbi-focal-vs-comp-map}.
+We visualize the spatial distribution of these trees in Figure \@ref(fig:scbi-focal-vs-comp-map).
 
 Here we use basal area as the continuous competitor explanatory variable but the package is flexible to allow the user to specify any competitor explanatory variable (basal area, biomass, tree height, a soil nutrient value, \dots).
-The package can also be used to compare competitor explanatory variables and see which best explains tree growth, see Appendix \ref{compare-pred-var} for an example comparing basal area and above ground biomass.
+The package can also be used to compare competitor explanatory variables and see which best explains tree growth, see Appendix \@ref(compare-pred-var) for an example comparing basal area and above ground biomass.
 Similarly, the package can use any categorical variable as an explanatory variable and compare between different categorical variables.
 For example in @allen_permutation_2020 we compare grouping individuals based on species, family, and based on trait-based groups.
-In Appendix \ref{compare-group-var} we give another example and compare grouping individuals by species or by potential canopy position (canopy, understory, shrub layer).
+In Appendix \@ref(compare-group-var) we give another example and compare grouping individuals by species or by potential canopy position (canopy, understory, shrub layer).
 
@@
 
-Lastly, we fit the competition Bayesian linear regression model for tree growth outlined in Equation \ref{eq:model} using `comp_bayes_lm()`.
+Lastly, we fit the competition Bayesian linear regression model for tree growth outlined in Equation \@ref(eq:model) using `comp_bayes_lm()`.
 This function has an option to specify prior distributions of all parameters, chosen here to be the defaults detailed in `?comp_bayes_lm`.
@@
 
-Lastly, the generic for `ggplot2::autoplot()` allows us to visualize all posterior distributions, as seen in Figure \ref{fig:scbi-posterior-viz}.
+Lastly, the generic for `ggplot2::autoplot()` allows us to visualize all posterior distributions, as seen in Figure \@ref(fig:scbi-posterior-viz).
 Setting `type` to `"intercepts"` and `"dbh_slopes"` returns species-specific posterior distributions for $\beta_{0,j}$ and $\beta_{dbh,j}$ respectively, while setting `type = "competition"` returns competition coefficients $\lambda_{j,k}$.
@@
 
-Currently the `forestecology` package can only fit the competition Bayesian linear regression model in Equation \ref{eq:model}.
+Currently the `forestecology` package can only fit the competition Bayesian linear regression model in Equation \@ref(eq:model).
 However, it can be extended to any model as long as it is implemented in a function similar to `comp_bayes_lm()`.
@@
 
-To evaluate the effect of competitor species identity, we use the above four steps along with the permutation test in Equation \ref{eq:permutation-hypothesis-test}.
+To evaluate the effect of competitor species identity, we use the above four steps along with the permutation test in Equation \@ref(eq:permutation-hypothesis-test).
 Under a null hypothesis where competitor species identity does not matter, we can permute the competitor species identities within each focal tree, compute the RMSE test statistic, repeat this process several times to construct a null distribution, and compare it to the observed RMSE to assess significance.
-Going back to our example in Section \ref{focal-vs-comp} of focal tree with `focal_ID` 4 and its 20 competitors, the permutation test only randomly resamples the `comp_sp` variable without replacement, leaving all other variables intact.
+Going back to our example in Section \@ref(focal-vs-comp) of focal tree with `focal_ID` 4 and its 20 competitors, the permutation test only randomly resamples the `comp_sp` variable without replacement, leaving all other variables intact.
 This resampling is nested within each focal tree in order to preserve neighborhood structure.
@@
 To evaluate model performance, we use spatial cross-validation.
-The model fit in Section \ref{model-fit-predict} uses the same data to both fit and assess model performance.
+The model fit in Section \@ref(model-fit-predict) uses the same data to both fit and assess model performance.
 Given the spatial-autocorrelation of our data, this can potentially lead to overfit models [@roberts_cross-validation_2017].
-To mitigate this risk, we use the spatial cross-validation blocking scheme encoded in the `foldID` variable from Section \ref{spatial-information} and visualized in Figure \ref{fig:scbi-spatial-information}.
+To mitigate this risk, we use the spatial cross-validation blocking scheme encoded in the `foldID` variable from Section \@ref(spatial-information) and visualized in Figure \@ref(fig:scbi-spatial-information).
 
@@
 We fit the model to all focal trees in the training set, apply the model to all focal trees in the test set, compute predicted values, and compute the RMSE.
-Furthermore, to maintain spatial independence between the test and training sets, a "fold buffer" that extends `r comp_dist`m outwards from the boundary of the test set is considered; all trees within this "fold buffer" are excluded from the training set (see Figure \ref{fig:scbi-spatial-cross-validation-schematic}).
+Furthermore, to maintain spatial independence between the test and training sets, a "fold buffer" that extends `r comp_dist`m outwards from the boundary of the test set is considered; all trees within this "fold buffer" are excluded from the training set (see Figure \@ref(fig:scbi-spatial-cross-validation-schematic).
 
@@
 
-`run_cv()` also accepts the `run_shuffle` argument in order to permute competitor species identity as described in Section \ref{permutation-test}.
-Figure \ref{fig:scbi-simulation} compares model performance for 49 permutations of competitor species and RMSE calculations, both with and without cross-validation.
+`run_cv()` also accepts the `run_shuffle` argument in order to permute competitor species identity as described in Section \@ref(permutation-test).
+Figure \@ref(fig:scbi-simulation) compares model performance for 49 permutations of competitor species and RMSE calculations, both with and without cross-validation.
 Without cross-validation, competitor species identity does matter as the observed RMSE was significantly lower than the permutation null distribution of RMSE.
@@
 There is evidence that traits predict competitive outcomes [@kunstler_competitive_2012; @lasky_trait-mediated_2014; @uriarte_trait_2010].
-Thus an extension of the model would allow $\lambda$ values from Equation \ref{eq:model} to be a function of the traits of competing species.
-
-Lastly, the `forestecology` current uses the `blockCV` package behind the scenes to create the spatial blocks acting as folds for our spatial cross-validation algorithm detailed in Sections \ref{spatial-information} and \ref{spatial-cross-validation} [@valavi_blockcv_2019].
+Thus an extension of the model would allow $\lambda$ values from Equation \@ref(eq:model) to be a function of the traits of competing species.
+
+Lastly, the `forestecology` current uses the `blockCV` package behind the scenes to create the spatial blocks acting as folds for our spatial cross-validation algorithm detailed in Sections \@ref(spatial-information) and \@ref(spatial-cross-validation) [@valavi_blockcv_2019].
 This back-end functionality could be substituted with the `spatialsample` package for spatial resampling infrastructure; a `tidymodels` package under active development as of 2021 [@spatialsample_package; @tidymodels_package].
 
-Lastly, currently the package only implements the Bayesian linear regression model detailed in Equation \ref{eq:model}.
-As we demonstrate in Section \ref{model-fit-predict} however, the fitting of this model is self-contained in a single function `comp_bayes_lm()` which returns an object of S3 class type `comp_bayes_lm`.
+Lastly, currently the package only implements the Bayesian linear regression model detailed in Equation \@ref(eq:model).
+As we demonstrate in Section \@ref(model-fit-predict) however, the fitting of this model is self-contained in a single function `comp_bayes_lm()` which returns an object of S3 class type `comp_bayes_lm`.
 This class has generic methods implemented to print, make predictions, and plot all results.
@@
 
-We observe in Figure \ref{fig:appendix2-plot} that a cut-off distance of approximately 6m minimizes the cross-validation estimated RMSE.
+We observe in Figure \@ref(fig:appendix2-plot) that a cut-off distance of approximately 6m minimizes the cross-validation estimated RMSE.
 
@@
 But the package allows the user to specify any competitor explanatory variable in the `comp_x_var` argument of `create_focal_vs_comp` function.
-Here we use the cross-validated model comparison to see which of two possible competitor explanatory variables computed in Section \ref{compute-growth}, basal area or above ground biomass, best explains growth.
+Here we use the cross-validated model comparison to see which of two possible competitor explanatory variables computed in Section \@ref(compute-growth), basal area or above ground biomass, best explains growth.
 
@@
 
-Here we observe that basal area is a better competitor explanatory variable competitor explanatory variables $x^{\text{comp}}_{ijk}$ from Equation \ref{eq:model} than above ground biomass as suggested by the lower estimated RMSE.
+Here we observe that basal area is a better competitor explanatory variable competitor explanatory variables $x^{\text{comp}}_{ijk}$ from Equation \@ref(eq:model) than above ground biomass as suggested by the lower estimated RMSE.
 
@@
 We find that species identity has a lower RMSE, so does a better job.
-We still however plot the competition posteriors for the canopy position groupings in Figure \ref{fig:can-pos-comp-plot}.
+We still however plot the competition posteriors for the canopy position groupings in Figure \@ref(fig:can-pos-comp-plot).
 Unsurprisingly we see that canopy and canopy emergent competitors generally have negative effects on their neighbors, while shrubs and understory competitors have neutral or even positive effects.
@@
 
-This code replicates Figure \ref{fig:scbi-simulation}: A comparison of root mean squared error of
+This code replicates Figure \@ref(fig:scbi-simulation): A comparison of root mean squared error of
 models for standard, permuted, and spatial cross-validated error estimates.

### 2021-08-31 paper/paper_word_output.Rmd
Fix equation numbering

@@
 
-\begin{equation}
-\label{eq:model}
-y_{ij} = \beta_{0,j} + \beta_{\text{dbh},j} \cdot \text{dbh}_{ij} + \sum_{k=1}^{K} \lambda_{jk} \cdot x^{\text{comp}}_{ijk} + \epsilon_{ij}
-\end{equation}
+\begin{eqnarray}
+y_{ij} = \beta_{0,j} + \beta_{\text{dbh},j} \cdot \text{dbh}_{ij} + \sum_{k=1}^{K} \lambda_{jk} \cdot x^{\text{comp}}_{ijk} + \epsilon_{ij} (\#eq:model)
+\begin{eqnarray}
 
@@
 \begin{eqnarray}
-\label{eq:permutation-hypothesis-test}
-&&H_0: \lambda_{jk} = \lambda_{j} \mbox{ for all } k = 1, \ldots, K\\
-\text{vs.}&&H_A: \text{at least one } \lambda_{jk} \mbox{ is different} \nonumber
+&&H_0: \lambda_{jk} = \lambda_{j} \mbox{ for all } k = 1, \ldots, K (\#eq:permutation-hypothesis-test)\\
+\text{vs.}&&H_A: \text{at least one } \lambda_{jk} \mbox{ is different} \notag
 \end{eqnarray}

### 2021-09-01 paper/paper.Rmd
Cleaned directory

@@
   out.width = "100%",
-  fig.path = "Figures/",
+  fig.path = "figures/",
   fig.width = 16 / 2.5,
@@
   tidy.opts = list(width.cutoff = 72),
-  tidy = TRUE
+  tidy = FALSE
 )
@@
 
+# Abstract {-}
+
+1. Neighborhood competition models are powerful tools to measure the effect of interspecific competition. Statistical methods to ease the application of these models are currently lacking.
+2.
+We present the `forestecology` package providing methods to i) specify neighborhood competition models, ii) evaluate the effect of competitor species identity using permutation tests, and iii) measure model performance using spatial cross-validation.
+Following @allen_permutation_2020, we implement a Bayesian linear regression neighborhood competition model.
+3.
+We demonstrate the package's functionality using data from the Smithsonian Conservation Biology Institute's large forest dynamics plot, part of the ForestGEO global network of research sites.
+Given ForestGEO's data collection protocols and data formatting standards, the package was designed with cross-site compatibility in mind.
+We highlight the importance of spatial cross-validation when interpreting model results.
+4.
+The package features i) `tidyverse`-like structure whereby verb-named functions can be modularly "piped" in sequence, ii) functions with standardized inputs/outputs of simple features `sf` package class, and iii) an S3 object-oriented implementation of the Bayesian linear regression model.
+These three facts allow for clear articulation of all the steps in the sequence of analysis and easy wrangling and visualization of the geospatial data.
+Furthermore, while the package only has Bayesian linear regression implemented, the package was designed with extensibility to other methods in mind.
+
+Keywords: forest ecology, interspecific competition, neighborhood competition, tree growth, R, ForestGEO, spatial cross-validation
+
 # Introduction
@@
 
-\begin{equation}
-\label{eq:model}
-y_{ij} = \beta_{0,j} + \beta_{\text{dbh},j} \cdot \text{dbh}_{ij} + \sum_{k=1}^{K} \lambda_{jk} \cdot x^{\text{comp}}_{ijk} + \epsilon_{ij}
-\end{equation}
+\begin{eqnarray}
+y_{ij} = \beta_{0,j} + \beta_{\text{dbh},j} \cdot \text{dbh}_{ij} + \sum_{k=1}^{K} \lambda_{jk} \cdot x^{\text{comp}}_{ijk} + \epsilon_{ij} (\#eq:model)
+\end{eqnarray}
 
@@
 \begin{eqnarray}
-\label{eq:permutation-hypothesis-test}
-&&H_0: \lambda_{jk} = \lambda_{j} \mbox{ for all } k = 1, \ldots, K\\
+&&H_0: \lambda_{jk} = \lambda_{j} \mbox{ for all } k = 1, \ldots, K (\#eq:permutation-hypothesis-test)\\
 \text{vs.}&&H_A: \text{at least one } \lambda_{jk} \mbox{ is different} \nonumber
@@
 Our model of interspecific competition relies on a spatial definition of who competitor trees are: all trees within a distance `comp_dist` of a focal tree.
-Here we set `comp_dist` to 7.5m, a value informed by other studies [@canham_neighborhood_2004, @uriarte_spatially_2004, @canham_neighborhood_2006], but the package could also be used to compare multiple distances and see which is best supported (see Appendix \ref{mult-dist}).
+Here we set `comp_dist` to 7.5m, a value informed by other studies [@canham_neighborhood_2004; @uriarte_spatially_2004; @canham_neighborhood_2006], but the package could also be used to compare multiple distances and see which is best supported (see Appendix \@ref(mult-dist)).
 We use `comp_dist` and a manually constructed `sf` representation of the study region's boundary as inputs to `add_buffer_variable()` to add a `buffer` boolean variable to `growth_scbi`.
@@
 ggplot() +
-  geom_sf(data = blocks_scbi, fill = "transparent", linetype = "dashed") +
+  geom_sf(
+    data = blocks_scbi, 
+    fill = "transparent", linetype = "dashed"
+  ) +
   geom_sf_text(
@@
 focal_vs_comp_scbi <- growth_scbi %>%
-  create_focal_vs_comp(comp_dist, blocks = blocks_scbi, id = "stemID", comp_x_var = "basal_area")
+  create_focal_vs_comp(
+    comp_dist, blocks = blocks_scbi, id = "stemID", 
+    comp_x_var = "basal_area"
+  )
+
 focal_vs_comp_scbi %>%
@@
 focal_vs_comp_scbi <- focal_vs_comp_scbi %>%
-  mutate(growth_hat = predict(comp_bayes_lm_scbi,
-    newdata = focal_vs_comp_scbi
-  ))
+  mutate(
+    growth_hat = predict(comp_bayes_lm_scbi,
+                         newdata = focal_vs_comp_scbi)
+    )
 ```
@@
 
-plot1 <- autoplot(comp_bayes_lm_scbi,
+plot1 <- autoplot(
+  comp_bayes_lm_scbi,
   type = "intercepts",
@@
 )
-plot2 <- autoplot(comp_bayes_lm_scbi,
+plot2 <- autoplot(
+  comp_bayes_lm_scbi,
   type = "dbh_slopes",
@@
 )
-plot3 <- autoplot(comp_bayes_lm_scbi,
+plot3 <- autoplot(
+  comp_bayes_lm_scbi,
   type = "competition",
@@
 focal_vs_comp_scbi <- focal_vs_comp_scbi %>%
-  mutate(growth_hat_shuffle = predict(comp_bayes_lm_scbi_shuffle,
-    newdata = focal_vs_comp_scbi
-  ))
+  mutate(
+    growth_hat_shuffle = predict(comp_bayes_lm_scbi_shuffle,
+                                 newdata = focal_vs_comp_scbi)
+  )
 ```
@@
 
-`run_cv()` also accepts the `run_shuffle` argument in order to permute competitor species identity as described in Section \ref{permutation-test}.
-Figure \ref{fig:scbi-simulation} compares model performance for 49 permutations of competitor species and RMSE calculations, both with and without cross-validation.
+`run_cv()` also accepts the `run_shuffle` argument in order to permute competitor species identity as described in Section \@ref(permutation-test).
+Figure \@ref(fig:scbi-simulation) compares model performance for 49 permutations of competitor species and RMSE calculations, both with and without cross-validation.
 Without cross-validation, competitor species identity does matter as the observed RMSE was significantly lower than the permutation null distribution of RMSE.
 However, once we incorporate spatial cross-validation, this improvement disappears.
-These results suggest that in this 9 ha subplot of the SCBI plot, competitive interactions do not depend on the identity of the competitor, which is the opposite of what has been observed in other locations [@allen_permutation_2020 @uriarte_spatially_2004].
+These results suggest that in this 9 ha subplot of the SCBI plot, competitive interactions do not depend on the identity of the competitor, which is the opposite of what has been observed in other locations [@allen_permutation_2020; @uriarte_spatially_2004].
 This provides a striking example of the importance of cross-validation, as without it the over-fit model gives rise to an incorrect conclusion.
@@
 One clear future direction would be to allow competition based on species trait values rather than species identity.
-There is evidence that traits predict competitive outcomes [@kunstler_competitive_2012, @lasky_trait-mediated_2014, @uriarte_trait_2010].
-Thus an extension of the model would allow $\lambda$ values from Equation \ref{eq:model} to be a function of the traits of competing species.
-
-Lastly, the `forestecology` current uses the `blockCV` package behind the scenes to create the spatial blocks acting as folds for our spatial cross-validation algorithm detailed in Sections \ref{spatial-information} and \ref{spatial-cross-validation} [@valavi_blockcv_2019].
-This back-end functionality could be substituted with the `spatialsample` package for spatial resampling infrastructure; a `tidymodels` package under active development as of 2021 [@spatialsample_package, @tidymodels_package].
-
-Lastly, currently the package only implements the Bayesian linear regression model detailed in Equation \ref{eq:model}.
-As we demonstrate in Section \ref{model-fit-predict} however, the fitting of this model is self-contained in a single function `comp_bayes_lm()` which returns an object of S3 class type `comp_bayes_lm`.
+There is evidence that traits predict competitive outcomes [@kunstler_competitive_2012; @lasky_trait-mediated_2014; @uriarte_trait_2010].
+Thus an extension of the model would allow $\lambda$ values from Equation \@ref(eq:model) to be a function of the traits of competing species.
+
+Lastly, the `forestecology` current uses the `blockCV` package behind the scenes to create the spatial blocks acting as folds for our spatial cross-validation algorithm detailed in Sections \@ref(spatial-information) and \@ref(spatial-cross-validation) [@valavi_blockcv_2019].
+This back-end functionality could be substituted with the `spatialsample` package for spatial resampling infrastructure; a `tidymodels` package under active development as of 2021 [@spatialsample_package; @tidymodels_package].
+
+Lastly, currently the package only implements the Bayesian linear regression model detailed in Equation \@ref(eq:model).
+As we demonstrate in Section \@ref(model-fit-predict) however, the fitting of this model is self-contained in a single function `comp_bayes_lm()` which returns an object of S3 class type `comp_bayes_lm`.
 This class has generic methods implemented to print, make predictions, and plot all results.
@@
 
-# Acknowledgments
-
-We thank Sophie Li for their feedback on the package interface. The authors declare no conflicts of interest.
-
-# Author's contributions
-
-AYK and DNA conceived the ideas and coded a draft of the package.
-AYK wrote an initial manuscript draft.
-SPC rewrote much of the package's code to align with R and "tidy" best practices [@wickham_welcome_2019].
-All authors contributed to subsequent drafts and gave final approval for manuscript.
-
-# Data accessibility
-
-We intend to archive all data and source code for this manuscript on GitHub at https://github.com/rudeboybert/forestecology and on Zenodo upon acceptance.
+# Acknowledgements {-}
+
+We thank Sophie Li for their feedback on the package interface.
+
+# Conflict of Interest {-}
+
+We declare no competing interests.
+
+# Author's contributions {-}
+
+**Albert Y.
+Kim**: Conceptualization (equal); Methodology (equal); Software (equal); Writing-original draft (lead); Writing-review & editing (equal).
+**David N.
+Allen**: Conceptualization (equal); Methodology (equal); Software (equal); Writing-original draft (supporting); Writing-review & editing (equal).
+**Simon P.
+Couch**: Software (supporting); Writing-review & editing (supporting).
+
+# Data accessibility {-}
+
+All data and source code for this manuscript on GitHub at https://github.com/rudeboybert/forestecology and are archived on Zenodo at https://doi.org/10.5281/zenodo.5348060.
 The example Smithsonian Conservation Biology Institute census and species information data are available on GitHub at https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/tree/master/ and are archived on Zenodo at https://doi.org/10.5281/zenodo.2649301 [@gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020].
 
+# References
+
 <div id="refs"></div>
 
-\appendix
-
 # Appendix: Compare different competitive distances {#mult-dist}
@@
   ) %>%
-  # make all species the same, needs to be factor with at least two levels
+  # make all species the same, needs to be factor with at least 
+  # two levels
   mutate(sp = factor("A", levels = c("A", "B"))) %>%
@@
   focal_vs_comp_scbi <- growth_scbi %>%
-    create_focal_vs_comp(comp_dist = comp_dist, blocks = blocks_scbi, id = "stemID", comp_x_var = "basal_area") %>%
+    create_focal_vs_comp(
+      comp_dist = comp_dist, blocks = blocks_scbi, id = "stemID", 
+      comp_x_var = "basal_area"
+    ) %>%
     run_cv(comp_dist = comp_dist, blocks = blocks_scbi)
@@
     canopy_position = str_replace(canopy_position, ",", ""),
-    canopy_position = ifelse(is.na(canopy_position), "shrub_layer", canopy_position),
+    canopy_position = ifelse(is.na(canopy_position), "shrub_layer", 
+                             canopy_position),
     sp = as.factor(canopy_position),
@@
 
-# Compute growth of trees based on census data -------------------------
+# Compute growth of trees based on census data ------------------------
 census_2013_scbi <- here("paper/scbi.stem2.csv") %>%
@@
 
-# Add spatial information ----------------------------------------------
+# Add spatial information ---------------------------------------------
 # Define buffer region using competitive distance range
@@
 
-# Compute focal versus competitor tree information ---------------------
+# Compute focal versus competitor tree information --------------------
 focal_vs_comp_scbi <- growth_scbi %>%
-  create_focal_vs_comp(comp_dist, blocks = blocks_scbi, id = "stemID", comp_x_var = "basal_area")
-
-# Fit model and make predictions ---------------------------------------
+  create_focal_vs_comp(
+    comp_dist, blocks = blocks_scbi, id = "stemID", 
+    comp_x_var = "basal_area"
+  )
+
+# Fit model and make predictions --------------------------------------
 # Number of permutation shuffles:
@@
 
-# 1. Compute observed test statistic: RMSE with no cross-validation ----
+# 1. Compute observed test statistic: RMSE with no cross-validation ---
 # Fit model (compute posterior parameters)
@@
 
-# 2. Compute observed test statistic: RMSE with cross-validation -------
+# 2. Compute observed test statistic: RMSE with cross-validation ------
 observed_RMSE_CV <- focal_vs_comp_scbi %>%
@@
 
-# 3. Permutation distribution: RMSE with no cross-validation -----------
+# 3. Permutation distribution: RMSE with no cross-validation ----------
 # Compute num_shuffle permutation test statistics
@@
 
-# 4. Permutation distribution: RMSE with cross-validation --------------
+# 4. Permutation distribution: RMSE with cross-validation -------------
 # Compute num_shuffle permutation test statistics
@@
   # Status update
-  str_c("Shuffle with permutation ", j, " at ", Sys.time()) %>% print()
+  str_c("Shuffle with permutation ", j, " at ", Sys.time()) %>% 
+    print()
 }
@@
 
-# Visualize results ----------------------------------------------------
+# Visualize results ---------------------------------------------------
 model_comp <- bind_rows(
@@
     fill = "Cross-validated?",
-    x = expression(paste("RMSE (cm ", y^{
-      -1
-    }, ")"))
+    x = expression(paste("RMSE (cm ", y^{-1}, ")"))
   ) +
@@
 ```
-
-# References

### 2021-09-01 ecology_and_evolution/paper.Rmd
renamed folder

--- 2021-09-01 paper/paper.Rmd
@@
 
+Running head: `forestecology` R package
+
 # Introduction
@@
 We fit the model to all focal trees in the training set, apply the model to all focal trees in the test set, compute predicted values, and compute the RMSE.
-Furthermore, to maintain spatial independence between the test and training sets, a "fold buffer" that extends `r comp_dist`m outwards from the boundary of the test set is considered; all trees within this "fold buffer" are excluded from the training set (see Figure \@ref(fig:scbi-spatial-cross-validation-schematic).
+Furthermore, to maintain spatial independence between the test and training sets, a "fold buffer" that extends `r comp_dist`m outwards from the boundary of the test set is considered; all trees within this "fold buffer" are excluded from the training set (see Figure \@ref(fig:scbi-spatial-cross-validation-schematic)).
 

### 2021-09-01 ecology_and_evolution/paper.tex
renamed folder

--- 2021-08-10 paper/paper.tex
@@
 simplified.
-
-Lastly, currently the package only implements the Bayesian linear
-regression model detailed in Equation \ref{eq:model}. As we demonstrate
-in Section \ref{model-fit-predict} however, the fitting of this model is
-self-contained in a single function \texttt{comp\_bayes\_lm()} which
-returns an object of S3 class type \texttt{comp\_bayes\_lm}. This class
-has generic methods implemented to print, make predictions, and plot all
-results. Therefore the package can be modularly extended to fit other
-models as long as they are coded similarly to \texttt{comp\_bayes\_lm()}
-and have equivalent generic methods implemented.
 
@@
 \NormalTok{sp_info <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"SCBI_ForestGEO_sp_ecology.csv"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{select}\NormalTok{(}\DataTypeTok{sp =}\NormalTok{ spcode, family, genus, species, canopy_position, drought_tolerance) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{mutate}\NormalTok{()}
+\StringTok{    }\KeywordTok{select}\NormalTok{(}\DataTypeTok{sp =}\NormalTok{ spcode, family, genus, species, canopy_position, drought_tolerance)}
 \NormalTok{sp_info}
@@
 
-We join this species information to our \texttt{growth\_scbi} data
-frame.
+We join this species information to our \texttt{growth\_scbi} data frame
+and convert the species variable to a factor.
 
@@
 \StringTok{    }\KeywordTok{left_join}\NormalTok{(sp_info, }\DataTypeTok{by =} \StringTok{"sp"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{    }\CommentTok{# }\AlertTok{TODO}\CommentTok{:}
-\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{as.factor}\NormalTok{(sp))}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{as.factor}\NormalTok{(sp))}
 \end{Highlighting}
@@
 \texttt{create\_focal\_vs\_comp()}, which takes the previously detailed
-\texttt{comp\_dist} and \texttt{id} arguments as well as the \texttt{sf}
-representation of the spatial cross-validation blocks and returns a new
-data frame \texttt{focal\_vs\_comp\_scbi}.
+\texttt{comp\_dist} and \texttt{id} arguments, the \texttt{sf}
+representation of the spatial cross-validation blocks
+\texttt{blocks\_scbi}, and a specification \texttt{comp\_x\_var} of the
+\texttt{basal\_area} variable we use as the competitor explanatory
+variable \(x^{\text{comp}}_{ijk}\) from Equation \ref{eq:model}. This
+function returns a new data frame \texttt{focal\_vs\_comp\_scbi}.
 
@@
 
-The resulting\texttt{comp\_bayes\_lm\_scbi} is an object of S3 class
+The resulting \texttt{comp\_bayes\_lm\_scbi} is an object of S3 class
 type \texttt{comp\_bayes\_lm} containing the posterior values of all
@@
 \caption{Schematic of spatial cross-validation.
-Using the k = 1 fold (bottom-left) as the test set, k = 2 through 4 as the training set, along with a "fold buffer" extending outwards from the test set to maintain spatial independence between it and the training set.}\label{fig:scbi-spatial-cross-validation-schematic}
+Using the k = 1 fold (bottom-left) as the test set, k = 2 through 4 as the training set, along with a fold buffer extending outwards from the test set to maintain spatial independence between it and the training set.}\label{fig:scbi-spatial-cross-validation-schematic}
 \end{figure}
@@
 
-\hypertarget{appendix-1-replicate-rmse-comparison}{%
-\section{Appendix 1: Replicate RMSE
-comparison}\label{appendix-1-replicate-rmse-comparison}}
+\hypertarget{refs}{}
+
+\appendix
+
+\hypertarget{mult-dist}{%
+\section{Appendix: Compare different competitive
+distances}\label{mult-dist}}
+
+For all the above analyses we set the cut off distance
+(\texttt{comp\_dist}) for two stems to compete as 7.5m. This distance
+has been estimated in previous neighborhood competition studies in
+forests \citep[\citet{uriarte_spatially_2004},
+\citet{canham_neighborhood_2006}]{canham_neighborhood_2004}. We used
+7.5m in \citet{allen_permutation_2020} as an average of the values
+estimated in other studies. But our package can be used to find which
+distance is best supported by the data. Here we provide an example using
+another section of the SCBI plot to provide an additional example of the
+cross-validation block layout. To speed computation we do not consider
+species differences in competitive effects and treat all species as the
+same.
+
+We observe in Figure \ref{fig:appendix2-plot} that a cut-off distance of
+approximately 6m minimizes the cross-validation estimated RMSE.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{census_}\DecValTok{2013}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem2.csv"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
+    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date),}
+    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh) }\OperatorTok{/}\StringTok{ }\DecValTok{10}
+\NormalTok{  ) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{400}\NormalTok{, gy }\OperatorTok{<}\StringTok{ }\DecValTok{400}\NormalTok{)}
+
+\NormalTok{census_}\DecValTok{2018}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem3.csv"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
+    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date),}
+    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh) }\OperatorTok{/}\StringTok{ }\DecValTok{10}
+\NormalTok{  ) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{400}\NormalTok{, gy }\OperatorTok{<}\StringTok{ }\DecValTok{400}\NormalTok{)}
+
+\NormalTok{growth_scbi <-}
+\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(}
+    \DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi,}
+    \DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)),}
+    \DataTypeTok{id =} \StringTok{"stemID"}
+\NormalTok{  ) }\OperatorTok{%>%}
+\StringTok{  }\CommentTok{# make all species the same, needs to be factor with at least two levels}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{factor}\NormalTok{(}\StringTok{"A"}\NormalTok{, }\DataTypeTok{levels =} \KeywordTok{c}\NormalTok{(}\StringTok{"A"}\NormalTok{, }\StringTok{"B"}\NormalTok{))) }\OperatorTok{%>%}
+\StringTok{  }\CommentTok{# Compute basal area:}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{basal_area =} \FloatTok{0.0001} \OperatorTok{*}\StringTok{ }\NormalTok{pi }\OperatorTok{*}\StringTok{ }\NormalTok{(dbh1 }\OperatorTok{/}\StringTok{ }\DecValTok{2}\NormalTok{)}\OperatorTok{^}\DecValTok{2}\NormalTok{)}
+
+\NormalTok{study_region_scbi <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}
+  \DataTypeTok{x =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{),}
+  \DataTypeTok{y =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{0}\NormalTok{)}
+\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{sf_polygon}\NormalTok{()}
+
+\NormalTok{n_fold <-}\StringTok{ }\DecValTok{4}
+\NormalTok{fold1 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{0}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{200}\NormalTok{))}
+\NormalTok{fold2 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{200}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{200}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{200}\NormalTok{))}
+\NormalTok{fold3 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{0}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{200}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{400}\NormalTok{))}
+\NormalTok{fold4 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{200}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{200}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{200}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{400}\NormalTok{))}
+
+\NormalTok{blocks_scbi <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}
+  \KeywordTok{sf_polygon}\NormalTok{(fold1), }\KeywordTok{sf_polygon}\NormalTok{(fold2), }\KeywordTok{sf_polygon}\NormalTok{(fold3),}
+  \KeywordTok{sf_polygon}\NormalTok{(fold4)}
+\NormalTok{) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\NormalTok{n_fold) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
+
+\CommentTok{# Associate each observation to a fold}
+\NormalTok{spatial_block_scbi <-}
+\StringTok{  }\KeywordTok{spatialBlock}\NormalTok{(}
+    \DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =}\NormalTok{ n_fold,}
+    \DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi,}
+    \DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}
+\NormalTok{  )}
+
+\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ spatial_block_scbi}\OperatorTok{$}\NormalTok{foldID }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}
+
+\NormalTok{mult_dist_comp <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}
+  \DataTypeTok{dist =} \KeywordTok{c}\NormalTok{(}\DecValTok{5}\NormalTok{, }\FloatTok{6.25}\NormalTok{, }\FloatTok{7.5}\NormalTok{, }\FloatTok{8.75}\NormalTok{, }\DecValTok{10}\NormalTok{),}
+  \DataTypeTok{rmse =} \DecValTok{0}
+\NormalTok{)}
+
+\ControlFlowTok{for}\NormalTok{ (i }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(mult_dist_comp}\OperatorTok{$}\NormalTok{dist)) \{}
+\NormalTok{  comp_dist <-}\StringTok{ }\NormalTok{mult_dist_comp}\OperatorTok{$}\NormalTok{dist[i]}
+
+\NormalTok{  growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{size =}\NormalTok{ comp_dist, }\DataTypeTok{region =}\NormalTok{ study_region_scbi)}
+
+\NormalTok{  focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{, }\DataTypeTok{comp_x_var =} \StringTok{"basal_area"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi)}
+
+\NormalTok{  mult_dist_comp}\OperatorTok{$}\NormalTok{rmse[i] <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{pull}\NormalTok{(.estimate)}
+\NormalTok{\}}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{Figures/appendix2-plot-1}
+
+}
+
+\caption{Cross-validated RMSE estimates for 5 competitive distances.}\label{fig:appendix2-plot}
+\end{figure}
+
+\hypertarget{compare-pred-var}{%
+\section{Appendix: Compare competitor explanatory
+variables}\label{compare-pred-var}}
+
+In the above code we use the basal area of an individual as a continuous
+competitor explanatory variable. But the package allows the user to
+specify any competitor explanatory variable in the \texttt{comp\_x\_var}
+argument of \texttt{create\_focal\_vs\_comp} function. Here we use the
+cross-validated model comparison to see which of two possible competitor
+explanatory variables computed in Section \ref{compute-growth}, basal
+area or above ground biomass, best explains growth.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{census_}\DecValTok{2013}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem2.csv"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date), }\DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+
+\NormalTok{census_}\DecValTok{2018}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem3.csv"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date), }\DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+
+\NormalTok{growth_scbi <-}\StringTok{ }\KeywordTok{compute_growth}\NormalTok{(}\DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi, }\DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)), }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{left_join}\NormalTok{(sp_info, }\DataTypeTok{by =} \StringTok{"sp"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{as.factor}\NormalTok{(sp), }\DataTypeTok{basal_area =} \FloatTok{1e-04} \OperatorTok{*}\StringTok{ }\NormalTok{pi }\OperatorTok{*}\StringTok{ }\NormalTok{(dbh1}\OperatorTok{/}\DecValTok{2}\NormalTok{)}\OperatorTok{^}\DecValTok{2}\NormalTok{, }\DataTypeTok{agb =} \KeywordTok{get_biomass}\NormalTok{(}\DataTypeTok{dbh =}\NormalTok{ dbh1,}
+        \DataTypeTok{genus =}\NormalTok{ genus, }\DataTypeTok{species =}\NormalTok{ species, }\DataTypeTok{coords =} \KeywordTok{c}\NormalTok{(}\OperatorTok{-}\FloatTok{78.2}\NormalTok{, }\FloatTok{38.9}\NormalTok{)))}
+
+\NormalTok{study_region_scbi <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}\DataTypeTok{x =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{), }\DataTypeTok{y =} \KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{,}
+    \DecValTok{600}\NormalTok{, }\DecValTok{300}\NormalTok{)) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{sf_polygon}\NormalTok{()}
+
+\NormalTok{n_fold <-}\StringTok{ }\DecValTok{4}
+\NormalTok{fold1 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{0}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{))}
+\NormalTok{fold2 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{150}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{))}
+\NormalTok{fold3 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{0}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+\NormalTok{fold4 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{150}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+
+\NormalTok{blocks_scbi <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}\KeywordTok{sf_polygon}\NormalTok{(fold1), }\KeywordTok{sf_polygon}\NormalTok{(fold2), }\KeywordTok{sf_polygon}\NormalTok{(fold3),}
+    \KeywordTok{sf_polygon}\NormalTok{(fold4)) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\NormalTok{n_fold) }\OperatorTok{%>%}
+\StringTok{        }\KeywordTok{factor}\NormalTok{())}
+
+\CommentTok{# Associate each observation to a fold}
+\NormalTok{spatial_block_scbi <-}\StringTok{ }\KeywordTok{spatialBlock}\NormalTok{(}\DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =}\NormalTok{ n_fold,}
+    \DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}\NormalTok{)}
+
+\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ spatial_block_scbi}\OperatorTok{$}\NormalTok{foldID }\OperatorTok{%>%}
+\StringTok{        }\KeywordTok{factor}\NormalTok{())}
+
+\NormalTok{comp_dist <-}\StringTok{ }\FloatTok{7.5}
+
+\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{size =}\NormalTok{ comp_dist, }\DataTypeTok{region =}\NormalTok{ study_region_scbi)}
+
+\NormalTok{focal_vs_comp_ba <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{,}
+        \DataTypeTok{comp_x_var =} \StringTok{"basal_area"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi)}
+
+\NormalTok{focal_vs_comp_agb <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{,}
+        \DataTypeTok{comp_x_var =} \StringTok{"agb"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi)}
+
+\NormalTok{focal_vs_comp_ba }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{pull}\NormalTok{(.estimate)}
+\CommentTok{## [1] 0.14}
+
+\NormalTok{focal_vs_comp_agb }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{pull}\NormalTok{(.estimate)}
+\CommentTok{## [1] 2.08}
+\end{Highlighting}
+\end{Shaded}
+
+Here we observe that basal area is a better competitor explanatory
+variable competitor explanatory variables \(x^{\text{comp}}_{ijk}\) from
+Equation \ref{eq:model} than above ground biomass as suggested by the
+lower estimated RMSE.
+
+\hypertarget{compare-group-var}{%
+\section{Appendix: Compare grouping variables}\label{compare-group-var}}
+
+The package also allows the user to specify the categorical explanatory
+grouping variable. Here we compare two different such variables: species
+and the potential canopy position of that species. If we had
+individual-level crown classes (\citet{smith_practice_1986} dominant,
+codominant, intermediate and suppressed) that could also be used.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{census_}\DecValTok{2013}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem2.csv"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date), }\DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+
+\NormalTok{census_}\DecValTok{2018}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem3.csv"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date), }\DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+
+\NormalTok{growth_scbi_sp <-}\StringTok{ }\KeywordTok{compute_growth}\NormalTok{(}\DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi, }\DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)), }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{as.factor}\NormalTok{(sp), }\DataTypeTok{basal_area =} \FloatTok{1e-04} \OperatorTok{*}\StringTok{ }\NormalTok{pi }\OperatorTok{*}\StringTok{ }\NormalTok{(dbh1}\OperatorTok{/}\DecValTok{2}\NormalTok{)}\OperatorTok{^}\DecValTok{2}\NormalTok{)}
+
+\NormalTok{growth_scbi_can_pos <-}\StringTok{ }\KeywordTok{compute_growth}\NormalTok{(}\DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi, }\DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)), }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{left_join}\NormalTok{(sp_info, }\DataTypeTok{by =} \StringTok{"sp"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{canopy_position =} \KeywordTok{str_replace}\NormalTok{(canopy_position, }\StringTok{" "}\NormalTok{, }\StringTok{"_"}\NormalTok{), }\DataTypeTok{canopy_position =} \KeywordTok{str_replace}\NormalTok{(canopy_position,}
+        \StringTok{","}\NormalTok{, }\StringTok{""}\NormalTok{), }\DataTypeTok{canopy_position =} \KeywordTok{ifelse}\NormalTok{(}\KeywordTok{is.na}\NormalTok{(canopy_position), }\StringTok{"shrub_layer"}\NormalTok{,}
+\NormalTok{        canopy_position), }\DataTypeTok{sp =} \KeywordTok{as.factor}\NormalTok{(canopy_position), }\DataTypeTok{basal_area =} \FloatTok{1e-04} \OperatorTok{*}
+\StringTok{        }\NormalTok{pi }\OperatorTok{*}\StringTok{ }\NormalTok{(dbh1}\OperatorTok{/}\DecValTok{2}\NormalTok{)}\OperatorTok{^}\DecValTok{2}\NormalTok{)}
+
+\NormalTok{study_region_scbi <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}\DataTypeTok{x =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{), }\DataTypeTok{y =} \KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{,}
+    \DecValTok{600}\NormalTok{, }\DecValTok{300}\NormalTok{)) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{sf_polygon}\NormalTok{()}
+
+\NormalTok{n_fold <-}\StringTok{ }\DecValTok{4}
+\NormalTok{fold1 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{0}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{))}
+\NormalTok{fold2 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{150}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{))}
+\NormalTok{fold3 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{0}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+\NormalTok{fold4 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{150}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{))}
+
+\NormalTok{blocks_scbi <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}\KeywordTok{sf_polygon}\NormalTok{(fold1), }\KeywordTok{sf_polygon}\NormalTok{(fold2), }\KeywordTok{sf_polygon}\NormalTok{(fold3),}
+    \KeywordTok{sf_polygon}\NormalTok{(fold4)) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\NormalTok{n_fold) }\OperatorTok{%>%}
+\StringTok{        }\KeywordTok{factor}\NormalTok{())}
+
+\CommentTok{# Associate each observation to a fold}
+\NormalTok{spatial_block_scbi <-}\StringTok{ }\KeywordTok{spatialBlock}\NormalTok{(}\DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =}\NormalTok{ n_fold,}
+    \DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}\NormalTok{)}
+
+\NormalTok{growth_scbi_sp <-}\StringTok{ }\NormalTok{growth_scbi_sp }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ spatial_block_scbi}\OperatorTok{$}\NormalTok{foldID }\OperatorTok{%>%}
+\StringTok{        }\KeywordTok{factor}\NormalTok{())}
+\NormalTok{growth_scbi_can_pos <-}\StringTok{ }\NormalTok{growth_scbi_can_pos }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ spatial_block_scbi}\OperatorTok{$}\NormalTok{foldID }\OperatorTok{%>%}
+\StringTok{        }\KeywordTok{factor}\NormalTok{())}
+
+\NormalTok{comp_dist <-}\StringTok{ }\FloatTok{7.5}
+
+\NormalTok{growth_scbi_sp <-}\StringTok{ }\NormalTok{growth_scbi_sp }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{size =}\NormalTok{ comp_dist, }\DataTypeTok{region =}\NormalTok{ study_region_scbi)}
+\NormalTok{growth_scbi_can_pos <-}\StringTok{ }\NormalTok{growth_scbi_can_pos }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{size =}\NormalTok{ comp_dist, }\DataTypeTok{region =}\NormalTok{ study_region_scbi)}
+
+\NormalTok{focal_vs_comp_sp <-}\StringTok{ }\NormalTok{growth_scbi_sp }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{,}
+        \DataTypeTok{comp_x_var =} \StringTok{"basal_area"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi)}
+
+\NormalTok{focal_vs_comp_can_pos <-}\StringTok{ }\NormalTok{growth_scbi_can_pos }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{,}
+        \DataTypeTok{comp_x_var =} \StringTok{"basal_area"}\NormalTok{) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi)}
+
+\NormalTok{focal_vs_comp_sp }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{pull}\NormalTok{(.estimate)}
+\CommentTok{## [1] 0.14}
+
+\NormalTok{focal_vs_comp_can_pos }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
+\StringTok{    }\KeywordTok{pull}\NormalTok{(.estimate)}
+\CommentTok{## [1] 0.144}
+\end{Highlighting}
+\end{Shaded}
+
+We find that species identity has a lower RMSE, so does a better job. We
+still however plot the competition posteriors for the canopy position
+groupings in Figure \ref{fig:can-pos-comp-plot}. Unsurprisingly we see
+that canopy and canopy emergent competitors generally have negative
+effects on their neighbors, while shrubs and understory competitors have
+neutral or even positive effects.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{fit_mod_can_pos <-}\StringTok{ }\NormalTok{growth_scbi_can_pos }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(}
+    \DataTypeTok{comp_dist =}\NormalTok{ comp_dist,}
+    \DataTypeTok{blocks =}\NormalTok{ blocks_scbi,}
+    \DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{,}
+    \DataTypeTok{comp_x_var =} \StringTok{"basal_area"}
+\NormalTok{  ) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{comp_bayes_lm}\NormalTok{(}\DataTypeTok{prior_param =} \OtherTok{NULL}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{Figures/can-pos-comp-plot-1}
+
+}
+
+\caption{Posterior distributions of all competition parameters.}\label{fig:can-pos-comp-plot}
+\end{figure}
+
+\hypertarget{appendix-replicate-rmse-comparison}{%
+\section{Appendix: Replicate RMSE
+comparison}\label{appendix-replicate-rmse-comparison}}
 
@@
 root mean squared error of models for standard, permuted, and spatial
-cross- validated error estimates.
+cross-validated error estimates.
 
@@
 
-\hypertarget{mult-dist}{%
-\section{Appendix 2: Compare different competitive
-distances}\label{mult-dist}}
-
-For all the above analyses we took the cut off distance
-(\texttt{comp\_dist}) for two stems to compete as 7.5 m. This distance
-has been estimated in previous neighborhood competition studies in
-forests \citep[\citet{uriarte_spatially_2004},
-\citet{canham_neighborhood_2006}]{canham_neighborhood_2004}. We used 7.5
-m in \citet{allen_permutation_2020} as an average of the values
-estimated in other studies. But our package can be used to find which
-distance is best supported by the data. Here we provide an example using
-another section of the SCBI plot to provide an additional example of the
-CV block layout. To speed computation we do not consider species
-differences in competitive effects and treat all species as the same.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{census_}\DecValTok{2013}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem2.csv"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date), }\DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{400}\NormalTok{, gy }\OperatorTok{<}\StringTok{ }\DecValTok{400}\NormalTok{)}
-
-\NormalTok{census_}\DecValTok{2018}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem3.csv"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date), }\DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{400}\NormalTok{, gy }\OperatorTok{<}\StringTok{ }\DecValTok{400}\NormalTok{)}
-
-\NormalTok{growth_scbi <-}\StringTok{ }\KeywordTok{compute_growth}\NormalTok{(}\DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi, }\DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)), }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{    }\CommentTok{# make all species the same, needs to be factor with at least two}
-\StringTok{    }\CommentTok{# levels}
-\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{factor}\NormalTok{(}\StringTok{"A"}\NormalTok{, }\DataTypeTok{levels =} \KeywordTok{c}\NormalTok{(}\StringTok{"A"}\NormalTok{, }\StringTok{"B"}\NormalTok{))) }\OperatorTok{%>%}
-\StringTok{    }\CommentTok{# Compute basal area:}
-\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{basal_area =} \FloatTok{1e-04} \OperatorTok{*}\StringTok{ }\NormalTok{pi }\OperatorTok{*}\StringTok{ }\NormalTok{(dbh1}\OperatorTok{/}\DecValTok{2}\NormalTok{)}\OperatorTok{^}\DecValTok{2}\NormalTok{)}
-
-\NormalTok{study_region_scbi <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}\DataTypeTok{x =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{), }\DataTypeTok{y =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{400}\NormalTok{,}
-    \DecValTok{0}\NormalTok{)) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{sf_polygon}\NormalTok{()}
-
-\NormalTok{n_fold <-}\StringTok{ }\DecValTok{4}
-\NormalTok{fold1 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{0}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{200}\NormalTok{))}
-\NormalTok{fold2 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{200}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{200}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{200}\NormalTok{))}
-\NormalTok{fold3 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{0}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{200}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{400}\NormalTok{))}
-\NormalTok{fold4 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{200}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{200}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{200}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{400}\NormalTok{))}
-
-\NormalTok{blocks_scbi <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}\KeywordTok{sf_polygon}\NormalTok{(fold1), }\KeywordTok{sf_polygon}\NormalTok{(fold2), }\KeywordTok{sf_polygon}\NormalTok{(fold3),}
-    \KeywordTok{sf_polygon}\NormalTok{(fold4)) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\NormalTok{n_fold) }\OperatorTok{%>%}
-\StringTok{        }\KeywordTok{factor}\NormalTok{())}
-
-\CommentTok{# Associate each observation to a fold}
-\NormalTok{spatial_block_scbi <-}\StringTok{ }\KeywordTok{spatialBlock}\NormalTok{(}\DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =}\NormalTok{ n_fold,}
-    \DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}\NormalTok{)}
-
-\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ spatial_block_scbi}\OperatorTok{$}\NormalTok{foldID }\OperatorTok{%>%}
-\StringTok{        }\KeywordTok{factor}\NormalTok{())}
-
-\NormalTok{mult_dist_comp <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}\DataTypeTok{dist =} \KeywordTok{c}\NormalTok{(}\DecValTok{5}\NormalTok{, }\FloatTok{6.25}\NormalTok{, }\FloatTok{7.5}\NormalTok{, }\FloatTok{8.75}\NormalTok{, }\DecValTok{10}\NormalTok{), }\DataTypeTok{rmse =} \DecValTok{0}\NormalTok{)}
-
-\ControlFlowTok{for}\NormalTok{ (i }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(mult_dist_comp}\OperatorTok{$}\NormalTok{dist)) \{}
-\NormalTok{    comp_dist <-}\StringTok{ }\NormalTok{mult_dist_comp}\OperatorTok{$}\NormalTok{dist[i]}
-
-\NormalTok{    growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
-\StringTok{        }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{size =}\NormalTok{ comp_dist, }\DataTypeTok{region =}\NormalTok{ study_region_scbi)}
-
-\NormalTok{    focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
-\StringTok{        }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi,}
-            \DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{, }\DataTypeTok{comp_x_var =} \StringTok{"basal_area"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{        }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi)}
-
-\NormalTok{    mult_dist_comp}\OperatorTok{$}\NormalTok{rmse[i] <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
-\StringTok{        }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
-\StringTok{        }\KeywordTok{pull}\NormalTok{(.estimate)}
-\NormalTok{\}}
-
-\NormalTok{mult_dist_comp }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(dist, rmse)) }\OperatorTok{+}\StringTok{ }\KeywordTok{geom_point}\NormalTok{() }\OperatorTok{+}\StringTok{ }\KeywordTok{xlab}\NormalTok{(}\StringTok{"Distance (m)"}\NormalTok{) }\OperatorTok{+}\StringTok{ }\KeywordTok{ylab}\NormalTok{(}\StringTok{"RMSE (cm)"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{center}\includegraphics[width=1\linewidth]{Figures/appendix2-code-1} \end{center}
-
-Here we see that a cut off distance of approximately 6 m minimizes RMSE.
-
-\hypertarget{compare-pred-var}{%
-\section{Appendix 3: Compare predictor
-variables}\label{compare-pred-var}}
-
-In the above code we use the basal area of an individual as a continuous
-predictor variable. But the package allows the user to specify any
-predictor variable in the \texttt{comp\_x\_var} argument of
-\texttt{create\_focal\_vs\_comp} function. Here we use the
-cross-validated model comparison to see which of two possible predictor
-variables, basal area or above ground biomass, best explains growth.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{census_}\DecValTok{2013}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem2.csv"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date), }\DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
-
-\NormalTok{census_}\DecValTok{2018}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem3.csv"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date), }\DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
-
-\NormalTok{growth_scbi <-}\StringTok{ }\KeywordTok{compute_growth}\NormalTok{(}\DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi, }\DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)), }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{left_join}\NormalTok{(sp_info, }\DataTypeTok{by =} \StringTok{"sp"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{as.factor}\NormalTok{(sp), }\DataTypeTok{basal_area =} \FloatTok{1e-04} \OperatorTok{*}\StringTok{ }\NormalTok{pi }\OperatorTok{*}\StringTok{ }\NormalTok{(dbh1}\OperatorTok{/}\DecValTok{2}\NormalTok{)}\OperatorTok{^}\DecValTok{2}\NormalTok{, }\DataTypeTok{agb =} \KeywordTok{get_biomass}\NormalTok{(}\DataTypeTok{dbh =}\NormalTok{ dbh1,}
-        \DataTypeTok{genus =}\NormalTok{ genus, }\DataTypeTok{species =}\NormalTok{ species, }\DataTypeTok{coords =} \KeywordTok{c}\NormalTok{(}\OperatorTok{-}\FloatTok{78.2}\NormalTok{, }\FloatTok{38.9}\NormalTok{)))}
-
-\NormalTok{study_region_scbi <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}\DataTypeTok{x =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{), }\DataTypeTok{y =} \KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{,}
-    \DecValTok{600}\NormalTok{, }\DecValTok{300}\NormalTok{)) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{sf_polygon}\NormalTok{()}
-
-\NormalTok{n_fold <-}\StringTok{ }\DecValTok{4}
-\NormalTok{fold1 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{0}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{))}
-\NormalTok{fold2 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{150}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{))}
-\NormalTok{fold3 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{0}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{))}
-\NormalTok{fold4 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{150}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{))}
-
-\NormalTok{blocks_scbi <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}\KeywordTok{sf_polygon}\NormalTok{(fold1), }\KeywordTok{sf_polygon}\NormalTok{(fold2), }\KeywordTok{sf_polygon}\NormalTok{(fold3),}
-    \KeywordTok{sf_polygon}\NormalTok{(fold4)) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\NormalTok{n_fold) }\OperatorTok{%>%}
-\StringTok{        }\KeywordTok{factor}\NormalTok{())}
-
-\CommentTok{# Associate each observation to a fold}
-\NormalTok{spatial_block_scbi <-}\StringTok{ }\KeywordTok{spatialBlock}\NormalTok{(}\DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =}\NormalTok{ n_fold,}
-    \DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}\NormalTok{)}
-
-\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ spatial_block_scbi}\OperatorTok{$}\NormalTok{foldID }\OperatorTok{%>%}
-\StringTok{        }\KeywordTok{factor}\NormalTok{())}
-
-\NormalTok{comp_dist <-}\StringTok{ }\FloatTok{7.5}
-
-\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{size =}\NormalTok{ comp_dist, }\DataTypeTok{region =}\NormalTok{ study_region_scbi)}
-
-\NormalTok{focal_vs_comp_ba <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{,}
-        \DataTypeTok{comp_x_var =} \StringTok{"basal_area"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi)}
-
-\NormalTok{focal_vs_comp_agb <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{,}
-        \DataTypeTok{comp_x_var =} \StringTok{"agb"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi)}
-
-\NormalTok{focal_vs_comp_ba }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{pull}\NormalTok{(.estimate)}
-\CommentTok{## [1] 0.14}
-
-\NormalTok{focal_vs_comp_agb }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{pull}\NormalTok{(.estimate)}
-\CommentTok{## [1] 2.08}
-\end{Highlighting}
-\end{Shaded}
-
-Here we see that basal area does a much better job than above ground
-biomass.
-
-\hypertarget{compare-group-var}{%
-\section{Appendix 4: Compare grouping
-variables}\label{compare-group-var}}
-
-The package also allows the user to specify the categorical predictor
-grouping variable. Here we compare two different such variables: species
-and the potential canopy position of that species. If we had
-individual-level crown classes (\citet{smith_practice_1986} dominant,
-codominant, intermediate and suppressed) that could also be used.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{census_}\DecValTok{2013}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem2.csv"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date), }\DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
-
-\NormalTok{census_}\DecValTok{2018}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem3.csv"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date), }\DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{))}
-
-\NormalTok{growth_scbi_sp <-}\StringTok{ }\KeywordTok{compute_growth}\NormalTok{(}\DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi, }\DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)), }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{as.factor}\NormalTok{(sp), }\DataTypeTok{basal_area =} \FloatTok{1e-04} \OperatorTok{*}\StringTok{ }\NormalTok{pi }\OperatorTok{*}\StringTok{ }\NormalTok{(dbh1}\OperatorTok{/}\DecValTok{2}\NormalTok{)}\OperatorTok{^}\DecValTok{2}\NormalTok{)}
-
-\NormalTok{growth_scbi_can_pos <-}\StringTok{ }\KeywordTok{compute_growth}\NormalTok{(}\DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi, }\DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)), }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{left_join}\NormalTok{(sp_info, }\DataTypeTok{by =} \StringTok{"sp"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{canopy_position =} \KeywordTok{str_replace}\NormalTok{(canopy_position, }\StringTok{" "}\NormalTok{, }\StringTok{"_"}\NormalTok{), }\DataTypeTok{canopy_position =} \KeywordTok{str_replace}\NormalTok{(canopy_position,}
-        \StringTok{","}\NormalTok{, }\StringTok{""}\NormalTok{), }\DataTypeTok{canopy_position =} \KeywordTok{ifelse}\NormalTok{(}\KeywordTok{is.na}\NormalTok{(canopy_position), }\StringTok{"shrub_layer"}\NormalTok{,}
-\NormalTok{        canopy_position), }\DataTypeTok{sp =} \KeywordTok{as.factor}\NormalTok{(canopy_position), }\DataTypeTok{basal_area =} \FloatTok{1e-04} \OperatorTok{*}
-\StringTok{        }\NormalTok{pi }\OperatorTok{*}\StringTok{ }\NormalTok{(dbh1}\OperatorTok{/}\DecValTok{2}\NormalTok{)}\OperatorTok{^}\DecValTok{2}\NormalTok{)}
-
-\NormalTok{study_region_scbi <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}\DataTypeTok{x =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{), }\DataTypeTok{y =} \KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{,}
-    \DecValTok{600}\NormalTok{, }\DecValTok{300}\NormalTok{)) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{sf_polygon}\NormalTok{()}
-
-\NormalTok{n_fold <-}\StringTok{ }\DecValTok{4}
-\NormalTok{fold1 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{0}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{))}
-\NormalTok{fold2 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{150}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{))}
-\NormalTok{fold3 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{150}\NormalTok{, }\DecValTok{0}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{))}
-\NormalTok{fold4 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{150}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{450}\NormalTok{, }\DecValTok{450}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{))}
-
-\NormalTok{blocks_scbi <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}\KeywordTok{sf_polygon}\NormalTok{(fold1), }\KeywordTok{sf_polygon}\NormalTok{(fold2), }\KeywordTok{sf_polygon}\NormalTok{(fold3),}
-    \KeywordTok{sf_polygon}\NormalTok{(fold4)) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\NormalTok{n_fold) }\OperatorTok{%>%}
-\StringTok{        }\KeywordTok{factor}\NormalTok{())}
-
-\CommentTok{# Associate each observation to a fold}
-\NormalTok{spatial_block_scbi <-}\StringTok{ }\KeywordTok{spatialBlock}\NormalTok{(}\DataTypeTok{speciesData =}\NormalTok{ growth_scbi, }\DataTypeTok{k =}\NormalTok{ n_fold,}
-    \DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}\NormalTok{)}
-
-\NormalTok{growth_scbi_sp <-}\StringTok{ }\NormalTok{growth_scbi_sp }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ spatial_block_scbi}\OperatorTok{$}\NormalTok{foldID }\OperatorTok{%>%}
-\StringTok{        }\KeywordTok{factor}\NormalTok{())}
-\NormalTok{growth_scbi_can_pos <-}\StringTok{ }\NormalTok{growth_scbi_can_pos }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ spatial_block_scbi}\OperatorTok{$}\NormalTok{foldID }\OperatorTok{%>%}
-\StringTok{        }\KeywordTok{factor}\NormalTok{())}
-
-\NormalTok{comp_dist <-}\StringTok{ }\FloatTok{7.5}
-
-\NormalTok{growth_scbi_sp <-}\StringTok{ }\NormalTok{growth_scbi_sp }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{size =}\NormalTok{ comp_dist, }\DataTypeTok{region =}\NormalTok{ study_region_scbi)}
-\NormalTok{growth_scbi_can_pos <-}\StringTok{ }\NormalTok{growth_scbi_can_pos }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{size =}\NormalTok{ comp_dist, }\DataTypeTok{region =}\NormalTok{ study_region_scbi)}
-
-\NormalTok{focal_vs_comp_sp <-}\StringTok{ }\NormalTok{growth_scbi_sp }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{,}
-        \DataTypeTok{comp_x_var =} \StringTok{"basal_area"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi)}
-
-\NormalTok{focal_vs_comp_can_pos <-}\StringTok{ }\NormalTok{growth_scbi_can_pos }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{,}
-        \DataTypeTok{comp_x_var =} \StringTok{"basal_area"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi)}
-
-\NormalTok{focal_vs_comp_sp }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{pull}\NormalTok{(.estimate)}
-\CommentTok{## [1] 0.14}
-
-\NormalTok{focal_vs_comp_can_pos }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{rmse}\NormalTok{(}\DataTypeTok{truth =}\NormalTok{ growth, }\DataTypeTok{estimate =}\NormalTok{ growth_hat) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{pull}\NormalTok{(.estimate)}
-\CommentTok{## [1] 0.144}
-\end{Highlighting}
-\end{Shaded}
-
-We find that species identity has a lower RMSE, so does a better job.
-But we can still plot the competition posteriors for the canopy position
-groupings.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{fit_mod_can_pos <-}\StringTok{ }\NormalTok{growth_scbi_can_pos }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{,}
-        \DataTypeTok{comp_x_var =} \StringTok{"basal_area"}\NormalTok{) }\OperatorTok{%>%}
-\StringTok{    }\KeywordTok{comp_bayes_lm}\NormalTok{(}\DataTypeTok{prior_param =} \OtherTok{NULL}\NormalTok{)}
-\KeywordTok{autoplot}\NormalTok{(fit_mod_can_pos, }\DataTypeTok{type =} \StringTok{"competition"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{center}\includegraphics[width=1\linewidth]{Figures/can-pos-comp-1} \end{center}
-
-Unsurprisingly we see that canopy and canopy emergent competitors
-generally have negative effects on their neighbors, while shrubs and
-understory competitors have neutral or even positive effects.
-
 \bibliographystyle{agsm}

### 2021-09-14 ecology_and_evolution/paper.Rmd
increased dpi of all images (except one)

@@
   fig.show = "hold",
+  dpi = 600,
   # Etc:
@@
 
-```{r can-pos-comp-plot, out.width="100%", fig.cap="Posterior distributions of all competition parameters.", echo = FALSE}
+```{r can-pos-comp-plot, out.width="100%", fig.cap="Posterior distributions of all competition parameters.", echo = FALSE, fig.width = 16 / 2.3}
 autoplot(fit_mod_can_pos, type = "competition")

### 2021-09-14 ecology_and_evolution/simulation_results/replicate_fig6.R

@@ -216,4 +216,4 @@ cv_plot
 
 filename %>%
   str_c(".pdf") %>%
-  ggsave(plot = cv_plot, width = 16 / 2, height = 9 / 2)
+  ggsave(plot = cv_plot, width = 16 / 2, height = 9 / 2, dpi = 600)

### 2021-10-01 README.Rmd
Update README

@@
 Additionally, these models can test whether the species identity of competitors matters using a permutation test-style shuffling of competitor identity (under the null hypothesis) and subsequently evaluating if model performance changes.
-See Allen and Kim (2020) [A permutation test and spatial cross-validation approach to assess models of interspecific competition between trees](https://doi.org/10.1371/journal.pone.0229930) for a full description.
+See Kim et al.
+(2021) [The `forestecology` R package for fitting and assessing neighborhood models of the effect of interspecific competition on the growth of trees](http://doi.org/10.1002/ece3.8129) for a full description.
 

### 2021-10-01 README.Rmd
adding

@@
 
+[![DOI](https://zenodo.org/badge/144327964.svg)](https://zenodo.org/badge/latestdoi/144327964)
 [![Build Status](https://github.com/rudeboybert/forestecology/workflows/R-CMD-check/badge.svg)](https://github.com/rudeboybert/forestecology/actions)

### 2021-10-02 paper/paper.Rmd
Move ecology and evolution paper back to paper folder

@@
 The colors indicate whether spatial cross-validation was used or not.", echo = FALSE}
-include_graphics("simulation_results/2021-03-03_scbi_49_shuffles.pdf")
+include_graphics("simulation_results/2021-03-03_scbi_49_shuffles.tiff")
 ```

### 2021-10-02 README.Rmd
rebuild stuff

@@
 Additionally, these models can test whether the species identity of competitors matters using a permutation test-style shuffling of competitor identity (under the null hypothesis) and subsequently evaluating if model performance changes.
+
 See Kim et al.
-(2021) [The `forestecology` R package for fitting and assessing neighborhood models of the effect of interspecific competition on the growth of trees](http://doi.org/10.1002/ece3.8129) for a full description.
+(2021) [The `forestecology` R package for fitting and assessing neighborhood models of the effect of interspecific competition on the growth of trees](http://doi.org/10.1002/ece3.8129) for a full description; the source code for this paper can be found in the `paper/` folder.
 
